### 2018-10-11 BF_2D_notebook_01.Rmd
added BF_2D  notebook

---
title: "Basis functions approximations for reduced-rank Gaussian process regression in high-dimensinal space"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(viridis)
library(rstan)
library(rgl)
library(RColorBrewer)
library(brms)
library(rstanarm)
library(geoR)
library(lattice)
library(aqfig)
library(fields)
library(colorRamps)
```

<!-- Gaussian Processes are flexible non-parametric stochastic models for $D$-dimensional functions. However, a Gaussian process model suffer from observational dimensionality $(N)$, where the inversion of the covariance matrix requires of $O(N^3)$ computational expenses.  -->

<!-- This notebook deals with the implementation of an approximate eigendecomposition of the covariance function of a D-dimensional Gaussian process. The method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods. This results in a reduced-rank approximation for the covariance function (Solin and Särkkä, 2018). Comparison of this approximated framework with the full Gaussian process and a splines-based model is carried out.  -->

### D-dimensional Gaussian process model

(Formulation of the D-dimensional GP model is missing)

### Approximate eigendecomposition of the covariance function of a D-dimensional Gaussian process

(Formulation of the D-dimensional Basis function aproximation model is missing)

### Application to simulated data

Simulating data from the following 2D function $f(x_1,x_2)$ and contaminate it with Gaussian noise $e \sim N(0,\sigma^2)$.

$$
f(x_1,x_2) = \frac{1}{100} \cdot (x_1^3 - x_2^3) + \frac{1}{10} \cdot x_1 \cdot x_2
$$
$$
y(x_1,x_2) = f(x_1,x_2) + e
$$

```{r echo=TRUE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# color palettes
col1 <- brewer.pal(11,"BrBG")[11:1] 
col2 <- cm.colors(11, alpha = 0.5)
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(15), colorRampPalette(c("white", "red"), bias=10)(15)[2:15])

# number of simulated observations and true function
nobs <- 120  
f <- function(x1,x2) 1/100*(x1^3 - x2^3) + 1/10*x1*x2

# grid input values
x_grid <- pred_grid(coords= c(-5,5), y.coords= c(-5,5), by= 0.3)

# breaks for colors
par(mfrow=c(1,2))
brk <- seq(min(f(x_grid[,1],x_grid[,2])),max(f(x_grid[,1],x_grid[,2])), length.out= 12)

# true function surface evaluated in the grid input values
xx <- unique(x_grid[,1])
image(xx, xx, matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)), col=col1, breaks= brk, main="true function surface f and simulated locations", xlab="x1", ylab="x2", cex.main=0.7)

# training input values
set.seed(10)
x1 <- runif(nobs,-5,5)  
x2 <- runif(nobs,-5,5)

# simulated training observations
y <-  f(x1,x2) + rnorm(length(x1), 0, 0.3)

points(x1, x2, pch=1, cex=0.7)
plot(f(x1, x2), y, col=4, cex=0.6, main="true function values f vs noisy simulations y", xlab="f", ylab="y", cex.main=0.7); abline(a=0, b=1, col="gray")

# noisy observations in the grid input locations
y_grid <- f(x_grid[,1],x_grid[,2]) + rnorm(length(dim(x_grid)[1]), 0, 0.2)

# total (training and grid) input values
x <- rbind(cbind(x1,x2),as.matrix(x_grid))

# number of training, grid and total observations
N1 <- length(x1)
N_grid <- dim(x_grid)[1]
N <- N1 + N_grid
```

### 2D Gaussian process model

Standata.

```{r}
D <- 2  # number of dimensions

standata_gp12<- list(D= D, x= cbind(x1,x2), y= y, N1= N1, N_grid= N_grid, x_grid= x_grid, y_grid= y_grid)
```

Stancode.

(The code of the D-dimensional GP model is missing)

Load and store the outcomes of the GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_gp12.rData")

gp12_f <- summary(aj_gp12, pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_y <- summary(aj_gp12, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_f_grid <- summary(aj_gp12, pars = c("f_grid"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_y_grid <- summary(aj_gp12, pars = c("y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_elpd_grid <- summary(aj_gp12, pars= c("log_y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary= 4)$summary
gp12_time <- sum(get_elapsed_time(aj_gp12)[2])/(aj_gp12@sim$iter - aj_gp12@sim$warmup)
```

Posteriors plots.

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(0, 0, 2, 1.5))

# training
plot(f(x1,x2), gp12_f[,1], col=4, cex=0.6, xlab="f", ylab="gp_f", main="true values f vs posterior means gp_f", mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(a=0, b=1, col="gray")
plot(f(x1,x2), f(x1,x2) - gp12_f[,1], xlab="f", ylab="f - gp_f", main="posterior residuals (f - gp_f)", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(h=0, col="gray")

title(main= "GP model", outer= TRUE)

# predicting
plot(f(x_grid[,1],x_grid[,2]), gp12_f_grid[,1], col=4, cex=0.6, xlab="f", ylab="gp_f", main="true values f vs posterior means gp_f", mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(a=0, b=1, col="gray")
plot(f(x_grid[,1],x_grid[,2]), f(x_grid[,1],x_grid[,2]) - gp12_f_grid[,1], xlab="f", ylab="f - gp12_f", main="posterior residuals (f - gp_f)", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(h=0, col="gray")
plot(f(x_grid[,1],x_grid[,2]), gp12_elpd_grid[,1], col=2, cex=0.8, xlab="f", ylab="ELPD", mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(h=mean(gp12_elpd_grid[,1]), col="red", lty=2)

# histogram residuals
gp12_error_grid <- matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)) - gp12_f_grid[,1]

hist(gp12_error_grid, mgp= c(1.7, 0.5, 0), xlab="f - gp_f", ylab="", main="posterior residuals (f - gp_f)", cex.main=0.9)

#images
image(xx, xx, matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)), col=col1, breaks= brk, main="true surface f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.9)

image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col= col1, breaks= brk , main="posterior mean surface gp_f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.9)

brk1 <- seq(-max(abs(gp12_error_grid)),max(abs(gp12_error_grid)), length.out= 30)

image.plot(unique(x_grid[,1]), unique(x_grid[,1]), gp12_error_grid, col= col3, breaks= brk1, main="posterior residuals surface", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), legend.width=1.2, cex.main=0.9)
points(x1, x2, pch=1, cex=0.6)
```

### 2D Basis functions model

Standata.

```{r}
M <- c(2,4,6,10,15,20,30)  #number of basis functions
D <- 2  #number of dimensions

indices <- list()
for(j in 1:length(M)){
	indices[[j]] <- matrix(NA, M[j]^D, D)
	mm=0;
	for (m1 in 1:M[j]){
		for (m2 in 1:M[j]){
				mm = mm+1
				indices[[j]][mm,] = c(m1, m2)
		}
	}
}

standat_bf12 <- list()
for(j in 1:length(M)){  #number of basis functions
  	standat_bf12[[j]] <- list(D= D, M= M[j], M_nD= M[j]^D, L= c(5/2*max(x1),5/2*max(x2)), x= rbind(cbind(x1,x2),as.matrix(x_grid)), y= c(y,y_grid), N1= N1, indices= indices[[j]], N_grid= N_grid)
}
```

Stancode.

(The code of the D-dimensional Basis function aproximation model is missing)

Load and store the outcomes of the different Basis functions models as function of the number of basis functions.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_bf12.rData")

bf12_f <- list()
bf12_y <- list()
bf12_elpd <- list()
bf12_time <- list()
for(j in 1:length(M)){  #number of basis functions
	bf12_f[[j]] <- summary(aj_bf12[[j]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_y[[j]] <- summary(aj_bf12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_elpd[[j]] <- summary(aj_bf12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_time[[j]] <- sum(get_elapsed_time(aj_bf12[[j]])[2])/(aj_bf12[[j]]@sim$iter - aj_bf12[[j]]@sim$warmup)
}
```

Posteriors plots as function of the number of basis functions.

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
for(j in 1:length(M)){
  par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(gp12_f[,1], bf12_f[[j]][1:N1,1], col=4, cex=0.6, xlab="gp_f ", ylab="bf_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f[,1], gp12_f[,1] - bf12_f[[j]][1:N1,1], xlab="gp_f ", ylab="gp_f - bf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - bf_f)", cex.main=0.8); abline(h=0, col="gray")

	title(main= paste("BF model with M =",M[j]), outer= TRUE)

	# predicting
	plot(gp12_f_grid[,1], bf12_f[[j]][-(1:N1),1], col=4, cex=0.6, xlab="gp_f ", ylab="bf_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f_grid[,1], gp12_f_grid[,1] - bf12_f[[j]][-(1:N1),1], xlab="gp_f ", ylab="gp_f - bf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - bf_f)", cex.main=0.8); abline(h=0, col="gray")
	plot(gp12_f_grid[,1], bf12_elpd[[j]][-(1:N1),1], col=2, cex=0.8, xlab="gp_f", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(bf12_elpd[[j]][-(1:N1),1]), col="red", lty=2)

	# histogram residuals
	bf12_error <- matrix(gp12_f_grid[,1],length(xx),length(xx)) - bf12_f[[j]][-(1:N1),1]

	hist(bf12_error, mgp= c(1.7, 0.5, 0), xlab="gp_f - bf_f", ylab="", main="posterior residuals (gp_f - bf_f)", cex.main=0.8)

	# images
	image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, main="GP posterior means surface gp_f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.8)

	image(xx, xx, matrix(bf12_f[[j]][-(1:N1),1],length(xx),length(xx)), col= col1, breaks= brk, main="BF posterior means surface bf_f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.8)

	brk1 <- seq(-max(abs(bf12_error)),max(abs(bf12_error)), length.out= 30)

	image.plot(xx, xx, bf12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="posterior residuals surface", cex.main=0.8)
	points(x1, x2, pch=1, cex=0.5)
}
```

### 2D Additive Gaussian process model

Standata.

```{r}
standat_AGP <- list(D= 2, x= cbind(x1,x2), y= y, N1= N1, N_grid= N_grid, x_grid= x_grid, y_grid= y_grid)
```

Stancode.

(The code of the D-dimensional additive GP model is missing)

Load and store the outcomes of the additive GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_gp1gp2.rData")

gp1gp2_f <- summary(aj_gp1gp2, pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_y <- summary(aj_gp1gp2, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_f_grid <- summary(aj_gp1gp2, pars = c("f_grid"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_y_grid <- summary(aj_gp1gp2, pars = c("y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_elpd_grid <- summary(aj_gp1gp2, pars= c("log_y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary= 4)$summary
gp1gp2_time <- sum(get_elapsed_time(aj_gp1gp2)[2])/(aj_gp1gp2@sim$iter - aj_gp1gp2@sim$warmup)
```

Posterior plots.

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(0, 0, 1.5, 1.5))

# training
plot(gp12_f[,1], gp1gp2_f[1:N1,1], col=4, cex=0.6, xlab="gp_f ", ylab="agp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs AGP posterior means agp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
plot(gp12_f[,1], gp12_f[,1] - gp1gp2_f[1:N1,1], xlab="gp12_f ", ylab="gp12_f - agp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - agp_f)", cex.main=0.8); abline(h=0, col="gray")

title(main= paste("Additive GP model"), outer= TRUE)

# predicting
plot(gp12_f_grid[,1], gp1gp2_f_grid[,1], col=4, cex=0.6, xlab="gp_f ", ylab="agp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs AGP posterior means agp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
plot(gp12_f_grid[,1], gp12_f_grid[,1] - gp1gp2_f_grid[,1], xlab="gp_f ", ylab="gp_f - agp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - agp_f)", cex.main=0.8); abline(h=0, col="gray")
plot(gp12_f_grid[,1], gp1gp2_elpd_grid[,1], col=2, cex=0.8, xlab="gp_f ", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(gp1gp2_elpd_grid[,1]), col="red", lty=2)

# histogram residuals
gp1gp2_error <- matrix(gp12_f_grid[,1],length(xx),length(xx)) - gp1gp2_f_grid[,1]

hist(gp1gp2_error, mgp= c(1.7, 0.5, 0), xlab="gp_f - agp_f", ylab="", main="posterior residuals (gp_f - agp_f)", cex.main=0.8)

# images
image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="GP posterior means surface gp_f", cex.main=0.8)

image(xx, xx, matrix(gp1gp2_f_grid[,1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="AGP posterior means surface agp_f", cex.main=0.8)

brk1 <- seq(-max(abs(gp1gp2_error)),max(abs(gp1gp2_error)), length.out= 30)

image.plot(xx, xx, gp1gp2_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="posterior residuals surface", cex.main=0.8)
points(x1, x2, pch=1, cex=0.5)
```

### 2D Additive Basis functions model

Standata.

```{r}
M <- c(2,4,6,10,15,20,30,40)  # number of basis functions
D <- 2  # number of dimensions

standat_bf1bf2 <- list()
for(j in 1:length(M)){
	standat_bf1bf2[[j]] <- list(D= D, M= M[j], L= c(5/2*max(x1), 5/2*max(x2)), x= rbind(cbind(x1,x2),as.matrix(x_grid)), y= c(y,y_grid), N1= N1, N_grid= N_grid)
}
```

Stancode.

(The code of the D-dimensional additive Basis function aproximation model is missing)

Load and store the outcomes of the different additive BF models as function of the number of basis functions.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_bf1bf2.rData")

bf1bf2_f <- list()
bf1bf2_y <- list()
bf1bf2_elpd <- list()
bf1bf2_time <- list()
for(j in 1:length(M)){  # number of Basis functions
	bf1bf2_f[[j]] <- summary(aj_bf1bf2[[j]], pars = c("F"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf1bf2_y[[j]] <- summary(aj_bf1bf2[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf1bf2_elpd[[j]] <- summary(aj_bf1bf2[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf1bf2_time[[j]] <- sum(get_elapsed_time(aj_bf1bf2[[j]])[2])/(aj_bf1bf2[[j]]@sim$iter - aj_bf1bf2[[j]]@sim$warmup)
}
```

Posterior plots as function of the number of basis functions.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
for(j in 1:length(M)){
  par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(gp1gp2_f[,1], bf1bf2_f[[j]][1:N1,1], col=4, cex=0.6, xlab="agp_f ", ylab="abf_f", mgp= c(1.7, 0.5, 0), main="AGP posterior means agp_f vs ABF posterior means abf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp1gp2_f[,1], gp1gp2_f[,1] - bf1bf2_f[[j]][1:N1,1], xlab="agp_f ", ylab="agp_f - abf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (agp_f - abf_f)", cex.main=0.8); abline(h=0, col="gray")

	title(main= paste("Additive BF model with M =",M[j]), outer= TRUE)

	# predicting
	plot(gp1gp2_f_grid[,1], bf1bf2_f[[j]][-(1:N1),1], col=4, cex=0.6, xlab="agp_f ", ylab="abf_f", mgp= c(1.7, 0.5, 0), main="AGP posterior means agp_f vs ABF posterior means abf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp1gp2_f_grid[,1], gp1gp2_f_grid[,1] - bf1bf2_f[[j]][-(1:N1),1], xlab="agp_f ", ylab="agp_f - abf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (agp_f - abf_f)", cex.main=0.8); abline(h=0, col="gray")
	plot(gp1gp2_f_grid[,1], bf1bf2_elpd[[j]][-(1:N1),1], col=2, cex=0.8, xlab="agp_f ", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(bf1bf2_elpd[[j]][,1]), col="red", lty=2)

	# histogram residuals
	bf12_error <- matrix(gp1gp2_f_grid[,1],length(xx),length(xx)) - bf1bf2_f[[j]][-(1:N1),1]

	hist(bf12_error, mgp= c(1.7, 0.5, 0), xlab="agp_f - abf_f", ylab="", main="posterior residuals (agp_f - abf_f)", cex.main=0.8)

	#images
	image(xx, xx, matrix(gp1gp2_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="AGP posterior means surface agp_f", cex.main=0.8)

	image(xx, xx, matrix(bf1bf2_f[[j]][-(1:N1),1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="ABF posterior means surface abf_f", cex.main=0.8)

	brk1 <- seq(-max(abs(bf12_error)),max(abs(bf12_error)), length.out= 30)

	image.plot(xx, xx, bf12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="posterior residuals surface", cex.main=0.8)
	points(x1, x2, pch=1, cex=0.6)
}
```

### 2D Splines-based model

Stancode and standata as a function of the number of knots.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
kn <- c(4,5,6,7,10)  # number of knots

stancode_sp12 <- list()
standata_sp12 <- list()
for(j in 1:length(kn)){  # number of knots

  # Stancode
	stancode_sp12[[j]] <- make_stancode(y ~ t2(x1, x2, k= kn[j]), data = data.frame(x1= c(x1,x_grid[,1]), x2= c(x2,x_grid[,2]), y= c(y, y_grid)))

	# Standata
	standata_sp12[[j]] <- make_standata(y ~ t2(x1, x2, k= kn[j]), data.frame(x1= c(x1,x_grid[,1]), x2= c(x2,x_grid[,2]), y= c(y, y_grid)))

	standata_sp12[[j]]$N1 <- N1
	standata_sp12[[j]]$N_grid <- N_grid
}
```

Load and store the outcomes of the splines-based model as a function of the number of knots.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_sp12.rData")

sp12_f <- list()
sp12_y <- list()
sp12_elpd <- list()
sp12_time <- list()
for(j in 1:length(kn)){ # number of knots
	sp12_f[[j]] <- summary(aj_sp12[[j]], pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_y[[j]] <- summary(aj_sp12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_elpd[[j]] <- summary(aj_sp12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_time[[j]] <- sum(get_elapsed_time(aj_sp12[[j]])[2])/(aj_sp12[[j]]@sim$iter - aj_sp12[[j]]@sim$warmup)
}
```

Posterior plots of the Splines-based model versus the True function.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots  True function vs SP
for(j in 1:length(kn)){
  par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(f(x1,x2), sp12_f[[j]][1:N1,1], col=4, cex=0.6, xlab="f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="true function f vs SP posterior means sp_f", cex.main=0.9); abline(a=0, b=1, col="gray")
	plot(f(x1,x2), f(x1,x2) - sp12_f[[j]][1:N1,1], xlab="f", ylab="f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (f - sp_f)", cex.main=0.9); abline(h=0, col="gray")

	title(main= paste("True function f vs SP model with knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3]), outer= TRUE)

	# predicting
	plot(f(x_grid[,1],x_grid[,2]), sp12_f[[j]][(N1+1):N,1], col=4, cex=0.6, xlab="f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="true function f vs SP posterior means sp_f", cex.main=0.9); abline(a=0, b=1, col="gray")
	plot(f(x_grid[,1],x_grid[,2]), f(x_grid[,1],x_grid[,2]) - sp12_f[[j]][(N1+1):N,1], xlab="f", ylab="f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (f - sp_f)", cex.main=0.9); abline(h=0, col="gray")
	plot(f(x_grid[,1],x_grid[,2]), sp12_elpd[[j]][(N1+1):N,1], col=2, cex=0.8, xlab="f", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(sp12_elpd[[j]][(N1+1):N,1]), col="red", lty=2)

	# histogram residuals
	sp12_error <- matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)) - sp12_f[[j]][(N1+1):N,1]

	hist(sp12_error, mgp= c(1.7, 0.5, 0), xlab="f - sp_f", ylab="", main="posterior residuals (f - sp_f)", cex.main=0.9)

	# images
	image(xx, xx, matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="True function surface f", cex.main=0.9)

	image(xx, xx, matrix(sp12_f[[j]][(N1+1):N,1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="SP posterior means surface sp_f", cex.main=0.9)

	brk1 <- seq(-max(abs(sp12_error)),max(abs(sp12_error)), length.out= 30)

	image.plot(xx, xx, sp12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), legend.width=1.2, main="posterior residuals surface", cex.main=0.9)
	points(x1, x2, pch=1, cex=0.5)
}
```

Posterior plots of the Splines-based model versus the Gaussian process model.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots GP vs SP
for(j in 1:length(kn)){
	par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(gp12_f[,1], sp12_f[[j]][1:N1,1], col=4, cex=0.6, xlab="gp_f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs SP posterior means sp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f[,1], gp12_f[,1] - sp12_f[[j]][1:N1,1], xlab="gp_f", ylab="gp_f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - sp_f)", cex.main=0.8); abline(h=0, col="gray")

	title(main= paste("GP vs SP model with knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3]), outer= TRUE)

	# predicting
	plot(gp12_f_grid[,1], sp12_f[[j]][(N1+1):N,1], col=4, cex=0.6, xlab="gp_f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs SP posterior means sp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f_grid[,1], gp12_f_grid[,1] - sp12_f[[j]][(N1+1):N,1], xlab="gp_f", ylab="gp_f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - sp_f)", cex.main=0.8); abline(h=0, col="gray")
	plot(gp12_f_grid[,1], sp12_elpd[[j]][(N1+1):N,1], col=2, cex=0.8, xlab="gp_f", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(sp12_elpd[[j]][(N1+1):N,1]), col="red", lty=2)

	# histogram residuals
	sp12_error <- matrix(gp12_f_grid[,1],table(x_grid[,2])[1],table(x_grid[,2])[1]) - sp12_f[[j]][(N1+1):N,1]

	hist(sp12_error, mgp= c(1.7, 0.5, 0), xlab="gp_f - sp_f", ylab="", main="posterior residuals (gp_f - sp_f)", cex.main=0.8)

	# images
	image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="GP posterior means surface gp_f", cex.main=0.8)

	image(xx, xx, matrix(sp12_f[[j]][(N1+1):N,1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="SP posterior means surface sp_f", cex.main=0.8)

	brk1 <- seq(-max(abs(sp12_error)),max(abs(sp12_error)), length.out= 30)

	image.plot(xx, xx, sp12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), legend.width=1.2, main="posterior residuals surface", cex.main=0.8)
	points(x1, x2, pch=1, cex=0.5)
}
```

### Computation time

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(M[1:8], c(unlist(bf1bf2_time)[1:8],NA,NA,NA)[1:8], type="b", col=5, ylim=c(0,1.5), xlab="M ; Knots", ylab="seconds per iteration")
lines(M[1:8], c(unlist(bf12_time)[1:7],NA), type="b", col=4)
lines(c(4,6,8,10,16), unlist(sp12_time)[1:5], type="b", col=6)
abline(h= c(gp1gp2_time+0.01, gp12_time), col=3:2)
legend("topright", legend=c("GP  2D","AGP  2D","BF  2D","ABF  2D","SP  2D"), col=2:6, lty=1, cex=0.7)
```

### Application to the Diabetes dataset

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# reference Notebook
# https://rawgit.com/avehtari/modelselection_tutorial/master/diabetes.html

diabetes <- read.csv("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Diabetes/diabetes.csv", header = TRUE)
summary(diabetes)
```

```{r fig.height=3, fig.width=4, message=FALSE, warning=FALSE, paged.print=FALSE}
# removing those observation rows with 0 in any of the variables
for (i in 2:6) {
      diabetes <- diabetes[-which(diabetes[, i] == 0), ]
}
# scale the covariates for easier comparison of coefficient posteriors
for (i in 1:8) {
      diabetes[i] <- as.vector(scale(diabetes[i]))
}

# modify the data column names slightly for easier typing
names(diabetes)[7] <- "dpf"
names(diabetes) <- tolower(names(diabetes))

# bi-plots
p <- ggplot(diabetes, aes(age, glucose))
p + geom_point(aes(colour = outcome), size = 2)

p <- ggplot(diabetes, aes(pregnancies, glucose))
p + geom_point(aes(colour = outcome), size = 2)
```

#### Preparing data for 2-Dimensions

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
### data 2D
set.seed(115)

# indices of test data
ind2 <- sample(1:dim(diabetes)[1], 15)
N2 <- length(ind2)

# indices of training data
ind1 <- setdiff(1:dim(diabetes)[1], ind2)
N1 <- length(ind1)

# grid of predicting data
pred <- pred_grid(coords= c(min(diabetes$pregnancies),max(diabetes$pregnancies)+0.5), y.coords = c(min(diabetes$glucose),max(diabetes$glucose)+0.5), by= 0.5)

# total, training, test and predicting input variables
x1 <- c(diabetes$pregnancies, pred[,1])
x2 <- c(diabetes$glucose, pred[,2])

# response variable
y <- c(diabetes$outcome, rep(0,dim(pred)[1]))

# indices predicting grid
ind_grid <-  setdiff(1:length(x1), 1:(N1+N2))
N_grid <- length(ind_grid)

length(x1)==N1+N2+N_grid
```

### 2D Gaussian process model

Standata y stancode.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Standata
standata_gp12 <- make_standata(y ~ gp(x1,x2, scale= FALSE) - 1 , data = data.frame(x1, x2, y), family= binomial(link = "logit"))

# Stancode
# stancode <- make_stancode(y ~ gp(x1,x2, scale= FALSE) - 1 , data = data.frame(x1, x2, y), family= binomial(link = "logit")) #, save_model= "C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/stancode_gp12")

standata_gp12$N1 <- N1
standata_gp12$N2 <- N2
standata_gp12$ind1 <- ind1
standata_gp12$ind2 <- ind2
```

Load and store the outcomes of the GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_gp12.rData")

gp12_f <- summary(aj_gp12, pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_y <- summary(aj_gp12, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_finv <- summary(aj_gp12, pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_elpd <- summary(aj_gp12, pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_time <- sum(get_elapsed_time(aj_gp12)[2])/(aj_gp12@sim$iter - aj_gp12@sim$warmup)
```

Confusion matrix on training data.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
print(table(as.factor(as.numeric(gp12_finv[ind1,1]>0.5)), y[ind1]))
```

Confusion matrix on test data.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
print(table(as.factor(as.numeric(gp12_finv[ind2,1]>0.5)), y[ind2]))
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
par(mfrow=c(2,2))

hist(as.numeric(gp12_finv[ind1,1]>0.5), main="GP posterior means", xlab="gp_f", cex.main=0.7)
hist(y[ind1], main="observed data", xlab="y", cex.main=0.7)

plot(gp12_finv[ind2,1], gp12_elpd[,1], col=4, cex=0.6, xlab="gp_f", ylab="ELPD"); abline(a=0, b=1, col="gray")

n_x <- length(unique(pred[,1]))
n_y <- length(unique(pred[,2]))
z <- matrix(as.vector(gp12_finv[ind_grid,1]), n_x, n_y)
filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, zlim= range(z, finite= TRUE), levels= pretty(range(z, finite = TRUE), n = 10), col= cm.colors(length(pretty(range(z, finite = TRUE), n = 10)) - 1, alpha=0.5), xlab="x1",
plot.axes = {axis(1, round(pred[,1],2)); axis(2, round(pred[,2],2)); points(x=diabetes$pregnancies, y=diabetes$glucose, pch=16, cex=0.7, col=cm.colors(2)[cut(y, breaks=2)])},
plot.title=title(main= "GP posterior mean", xlab="x1", ylab="x2")
)
```

### 2D Basis function model

Preparing data for 2D Basis function approach.

```{r}
D <- 2  # number of dimensions
M <- c(2,4,6,10,20,30,40,60)  # number of basis functions

indices <- list()
for(j in 1:length(M)){
	indices[[j]] <- matrix(NA, M[j]^D, D)
	mm=0;
	for (m1 in 1:M[j]){
		for (m2 in 1:M[j]){
			# for (m3 in 1:M[j]){
				mm = mm+1
				indices[[j]][mm,] = c(m1, m2) #, m3)
			# }
		}
	}
}

# Standata
standata_bf12 <- list()
for(j in 1:length(M)){  # number of basis functions
  standata_bf12[[j]] <- standata_gp12
	standata_bf12[[j]]$D <- 2
	standata_bf12[[j]]$M <- M[j]
	standata_bf12[[j]]$M_nD <- M[j]^D
	standata_bf12[[j]]$indices <- indices[[j]]
	standata_bf12[[j]]$L <- c(5/2*max(x1),5/2*max(x2))
}
```

Stancode.

(The code of the D-dimensional Basis function aproximation model is missing)

Load and store the outcomes of the different Basis function models as function of the number of basis functions.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_bf12.rData")

bf12_f <- list()
bf12_finv <- list()
bf12_y <- list()
bf12_elpd <- list()
bf12_time <- list()
for(j in 1:length(aj_bf12)){
	bf12_f[[j]] <- summary(aj_bf12[[j]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_finv[[j]] <- summary(aj_bf12[[j]], pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_y[[j]] <- summary(aj_bf12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_elpd[[j]] <- summary(aj_bf12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_time[[j]] <- sum(get_elapsed_time(aj_bf12[[j]])[2])/(aj_bf12[[j]]@sim$iter - aj_bf12[[j]]@sim$warmup)
}
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(10), colorRampPalette(c("white", "red"), bias=10)(10)[2:10])

for(j in 1:length(aj_bf12)){
	par(mfrow=c(2,2), oma=c(1, 0, 1.5, 1.5))

  # scatter  plots of training data
	plot(gp12_finv[ind1,1], bf12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="bf_f", main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.7); abline(a=0, b=1, col="gray")

	plot(gp12_finv[ind1,1], gp12_finv[ind1,1] - bf12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="gp_f - bf_f", main="posterior residuals (gp_f - bf_f)", cex.main=0.7); abline(h=0, col="gray")

	# histograms
	hist(as.numeric(bf12_finv[[j]][ind1,1]>0.5), main="BF posterior means", xlab="bf_f", cex.main=0.7)
	hist(y[ind1], main="observed data", xlab="y", cex.main=0.7)

	title(main= paste("GP vs BF model with M =",M[j]), outer= TRUE)

	# Confussion matrices
	# print(paste("Confussion matrix of BF model with M =",M[j]))
	# print("training data")
	# print(table(as.factor(as.numeric(bf12_finv[[j]][ind1,1]>0.5)), y[ind1]))
	# print("test data")
	# print(table(as.factor(as.numeric(bf12_finv[[j]][ind2,1]>0.5)), y[ind2]))

	# Posterior means
	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(bf12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, zlim= range(z, finite= TRUE), levels= pretty(range(z, finite = TRUE), n = 10), col= cm.colors(length(pretty(range(z, finite = TRUE), n = 10)) - 1, alpha=0.5), plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))
	points(x=diabetes$pregnancies, y=diabetes$glucose, pch=16, cex=0.7, col=cm.colors(2)[cut(y, breaks=2)] )},	plot.title=title(main = paste("bf(x1;x2;M=",M[[j]],sep="",")")) )

	# GP vs BF posterior means
	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(gp12_finv[ind_grid,1]) - as.vector(bf12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, levels= seq(-max(abs(z)),max(abs(z)), length.out= 20), col= col3, plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))},
	plot.title=title(main= paste("gp(x1,x2) - bf(x1;x2;M=",M[[j]],sep="",")")) )
}
```

## 2D Splines-based model

Standata and stancode as a function of the number of knots.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
kn <- c(2,3,4,5,8)  # number of knots

stancode <- list()
standata_sp12 <- list()
for(j in 1:length(kn)){

  # Stancode
	stancode[[j]] <- make_stancode(y | trials(n) ~ t2(x1, x2, k= 2 + kn[j]), data = data.frame(x1= x1, x2= x2, y= y, n= rep(1,N1+N2+N_grid)), family = binomial("logit")) #, save_model= "C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/stancode_gp12")

	# Standata
	standata_sp12[[j]] <- make_standata(y | trials(n) ~ t2(x1, x2, k= 2 + kn[j]), data = data.frame(x1= x1, x2= x2, y= y, n= rep(1,N1+N2+N_grid)), family= binomial("logit"))

	standata_sp12[[j]]$N1 <- N1
	standata_sp12[[j]]$N2 <- N2
	standata_sp12[[j]]$ind1 <- ind1
	standata_sp12[[j]]$ind2 <- ind2
}

N <- standata_sp12[[1]]$N
```

Load and store de outcomes of the different Splines models as function of the number of knots.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_sp12.rData")

sp12_f <- list()
sp12_finv <- list()
sp12_y <- list()
sp12_elpd <- list()
sp12_time <- list()
for(j in 1:length(kn)){ # number of knots
	sp12_f[[j]] <- summary(aj_sp12[[j]], pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_finv[[j]] <- summary(aj_sp12[[j]], pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_y[[j]] <- summary(aj_sp12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_elpd[[j]] <- summary(aj_sp12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_time[[j]] <- sum(get_elapsed_time(aj_sp12[[j]])[2])/(aj_sp12[[j]]@sim$iter - aj_sp12[[j]]@sim$warmup)
}
```

```{r fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(10), colorRampPalette(c("white", "red"), bias=10)(10)[2:10])

for(j in 1:length(kn)){  # number of knots
	par(mfrow=c(2,2), oma=c(1, 0, 1.5, 1.5))

	plot(gp12_finv[ind1,1], sp12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="sp_f", main="GP posterior means gp_f vs SP posterior means sp_f", cex.main= 0.7); abline(a=0, b=1, col="gray")

	plot(gp12_finv[ind1,1], gp12_finv[ind1,1] - sp12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="gp_f - sp_f", main="posterior residuals (gp_f - sp_f)", cex.main= 0.7); abline(h=0, col="gray")

	hist(as.numeric(sp12_finv[[j]][ind1,1]>0.5), main="SP posterior means", xlab="sp_f", cex.main=0.7)
	hist(y[ind1], main="observed data", xlab="y", cex.main=0.7)

	title(main= paste("GP vs SP model with knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3]), outer= TRUE)

	# print(paste("Confussion matrix of the Splines model with Knots =",kn[j]))
	# print("training data")
	# print(table(as.factor(as.numeric(sp12_y[[j]][ind1,1]>0.5)), y[ind1]))
	# print("test data")
	# print(table(as.factor(as.numeric(sp12_y[[j]][ind2,1]>0.5)), y[ind2]))

	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(sp12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, zlim= range(z, finite= TRUE), levels= pretty(range(z, finite = TRUE), n = 10), col= cm.colors(length(pretty(range(z, finite = TRUE), n = 10)) - 1, alpha=0.5), plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))
	points(x=diabetes$pregnancies, y=diabetes$glucose, pch=16, cex=0.7, col=cm.colors(2)[cut(y, breaks=2)] )},	plot.title=title(main = paste("sp(x1;x2;Knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3],")")) )

	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(gp12_finv[ind_grid,1]) - as.vector(sp12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, levels= seq(-max(abs(z)),max(abs(z)), length.out= 20), col= col3, plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))},
	plot.title=title(main= paste("gp(x1,x2) - sp(x1;x2;Knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3],")")) )
}
```

### Computation time

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(M[1:8], c(unlist(bf12_time)[1:5],NA,NA,NA)[1:8], type="b", col=4, ylim=c(0,gp12_time+1), xlab="M ; Knots", ylab="seconds per iteration")
lines(kn[1:5], unlist(sp12_time)[1:5], type="b", col=6)
abline(h= gp12_time, col=2)
legend("bottomright", legend=c("GP  2D","BF  2D","SP  2D"), col=c(2,4,6), lty=1, cex=0.7)
```

#### Preparing data for 3-Dimensions

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
### data 3D
set.seed(115)
ind2 <- sample(1:dim(diabetes)[1], 15)
N2 <- length(ind2)

ind1 <- setdiff(1:dim(diabetes)[1], ind2)
N1 <- length(ind1)

x1 <- diabetes$pregnancies
x2 <- diabetes$glucose
x3 <- diabetes$age
y <- diabetes$outcome

length(x1)==N1+N2
```

### 3D Gaussian process model

Standata y stancode.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Standata
standata_gp123 <- make_standata(y ~ gp(x1,x2,x3, scale= FALSE) - 1 , data = data.frame(x1, x2, x3, y), family= binomial(link = "logit"))

# Stan code
# stancode <- make_stancode(y ~ gp(x1,x2,x3, scale= FALSE) - 1 , data = data.frame(x1, x2, x3, y), family= binomial(link = "logit")) #, save_model= "C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/stancode_gp123")

standata_gp123$N1 <- N1
standata_gp123$N2 <- N2
standata_gp123$ind1 <- ind1
standata_gp123$ind2 <- ind2
```

Load and store the outcomes of the GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_gp123.rData")

gp123_f <- summary(aj_gp123, pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_y <- summary(aj_gp123, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_finv <- summary(aj_gp123, pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_elpd <- summary(aj_gp123, pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_time <- sum(get_elapsed_time(aj_gp123)[2])/(aj_gp123@sim$iter - aj_gp123@sim$warmup)
```

Confussion matrix of the GP model for training data.

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
table(as.factor(as.numeric(gp123_y[,1]>0.5)), y)
```

Confussion matrix of the GP model for test data.

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
table(as.factor(as.numeric(gp123_y[ind2,1]>0.5)), y[ind2])
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
par(mfrow=c(2,2))

hist(as.numeric(gp123_y[,1]>0.5), main="GP posterior means", xlab="gp_y", cex.main=0.7)
hist(y, main="observed data", xlab="y", cex.main=0.7)

plot(gp123_y[ind2,1], gp123_elpd[,1], col=4, cex=0.6, xlab="gp_y", ylab="ELPD"); abline(a=0, b=1, col="gray")
```

### 3D Basis function model

Standata.

```{r}
D <- 3  # number of dimensions
M <- c(2,4,6,10,20,30,40,60)  # number of basis functions

indices <- list()
for(j in 1:length(M)){
	indices[[j]] <- matrix(NA, M[j]^D, D)
	mm=0;
	for (m1 in 1:M[j]){
		for (m2 in 1:M[j]){
			for (m3 in 1:M[j]){
				mm = mm+1
				indices[[j]][mm,] = c(m1, m2, m3)
			}
		}
	}
}

standata_bf123 <- list()
for(j in 1:length(M)){
  standata_bf123[[j]] <- standata_gp123
	standata_bf123[[j]]$D <- D
	standata_bf123[[j]]$M <- M[j]
	standata_bf123[[j]]$M_nD <- M[j]^D
	standata_bf123[[j]]$indices <- indices[[j]]
	standata_bf123[[j]]$L <- c(5/2*max(x1),5/2*max(x2),5/2*max(x3))
}
```

Stancode.

(The code of the D-dimensional Basis fucntion aproximation model is missing2)

Load and store the outcomes of the Splines-based model model as function of the number of knots.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_bf123.rData")

bf123_f <- list()
bf123_finv <- list()
bf123_y <- list()
bf123_elpd <- list()
bf123_time <- list()
for(j in 1:length(aj_bf123)){
	bf123_f[[j]] <- summary(aj_bf123[[j]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_finv[[j]] <- summary(aj_bf123[[j]], pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_y[[j]] <- summary(aj_bf123[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_elpd[[j]] <- summary(aj_bf123[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_time[[j]] <- sum(get_elapsed_time(aj_bf123[[j]])[2])/(aj_bf123[[j]]@sim$iter - aj_bf123[[j]]@sim$warmup)
}
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
for(j in 1:length(aj_bf123)){
	par(mfrow=c(2,2), oma=c(1, 0, 1.5, 1.5))

	plot(gp123_finv[,1], bf123_finv[[j]][,1], col=4, cex=0.6, xlab="gp_f", ylab="bf_f", main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.7); abline(a=0, b=1, col="gray")

	plot(gp123_finv[,1], gp123_finv[,1] - bf123_finv[[j]][,1], col=4, cex=0.6, xlab="gp_f", ylab="gp_f - bf_f", main="posterior residuals (gp_f - bf_f)", cex.main=0.7); abline(h=0, col="gray")

	hist(as.numeric(bf123_y[[j]][,1]>0.5), main="BF posterior means", xlab="bf_y", cex.main=0.7)
	hist(y, main="observed data", xlab="y", cex.main=0.7)

	title(main= paste("GP vs BF model with M =",M[j]), outer= TRUE)

	print(paste("Confussion matrix of the Basis function model with M =",M[j]))
	print("training data")
	print(table(as.factor(as.numeric(bf123_y[[j]][,1]>0.5)), y))
  print("test data")
	print(table(as.factor(as.numeric(bf123_y[[j]][ind2,1]>0.5)), y[ind2]))
}
```

### Computational time

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(M[1:6], c(unlist(bf123_time)[1:5],NA,NA,NA,NA)[1:6], type="b", col=4, ylim=c(0,gp123_time+0.5), ylab="seconds per iteration", xlab="M")
abline(h= c(gp123_time+0.01), col=5)
legend("topright", legend=c("GP  3D","BF  3D"), col=5:4, lty=1, cex=0.7)
```

### 2019-02-19 Paper/manuscript/template/interacttfssample.tex
paper

% interacttfssample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\begin{document}

\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{Taylor \& Francis \LaTeX\ template for authors (\textsf{Interact} layout + reference style~S)}

\author{
\name{A.~N. Author\textsuperscript{a}\thanks{CONTACT A.~N. Author. Email: latex.helpdesk@tandf.co.uk} and John Smith\textsuperscript{b}}
\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK; \textsuperscript{b}Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany}
}

\maketitle

\begin{abstract}
This template is for authors who are preparing a manuscript for a Taylor \& Francis journal using the \LaTeX\ document preparation system and the \texttt{interact} class file, which is available via selected journals' home pages on the Taylor \& Francis website.
\end{abstract}

\begin{keywords}
Sections; lists; figures; tables; mathematics; fonts; references; appendices
\end{keywords}

\section{Introduction}

In order to assist authors in the process of preparing a manuscript for a journal, the Taylor \& Francis `\textsf{Interact}' layout style has been implemented as a \LaTeXe\ class file based on the \texttt{article} document class. A \textsc{Bib}\TeX\ bibliography style file and a sample bibliography are also provided in order to assist with the formatting of your references.

Commands that differ from or are provided in addition to standard \LaTeXe\ are described in this document, which is \emph{not} a substitute for a \LaTeXe\ tutorial.

The \texttt{interacttfssample.tex} file can be used as a template for a manuscript by cutting, pasting, inserting and deleting text as appropriate, using the preamble and the \LaTeX\ environments provided (e.g.\ \verb"\begin{abstract}", \verb"\begin{keywords}").

\subsection{The \textsf{Interact} class file}\label{class}

The \texttt{interact} class file preserves the standard \LaTeXe\ interface such that any document that can be produced using \texttt{article.cls} can also be produced with minimal alteration using the \texttt{interact} class file as described in this document.

If your article is accepted for publication it will be typeset as the journal requires in Minion Pro and/or Myriad Pro. Since most authors will not have these fonts installed, the page make-up is liable to alter slightly with the change of font. Also, the \texttt{interact} class file produces only single-column format, which is preferred for peer review and will be converted to two-column format by the typesetter if necessary during preparation of the proofs. Please therefore do not try to match the typeset format exactly, but use the standard \LaTeX\ fonts instead and ignore details such as slightly long lines of text or figures/tables not appearing in exact synchronization with their citations in the text: these details will be dealt with by the typesetter. Similarly, it is unnecessary to spend time addressing warnings in the log file -- if your .tex file compiles to produce a PDF document that correctly shows how you wish your paper to appear, such warnings will not prevent your source files being imported into the typesetter's program.

\subsection{Submission of manuscripts prepared using \emph{\LaTeX}}

Manuscripts for possible publication should be submitted to the Editors for review as directed in the journal's Instructions for Authors, and in accordance with any technical instructions provided in the journal's ScholarOne Manuscripts or Editorial Manager site. Your \LaTeX\ source file(s), the class file and any graphics files will be required in addition to the final PDF version when final, revised versions of accepted manuscripts are submitted.

Please ensure that any author-defined macros used in your article are gathered together in the preamble of your .tex file, i.e.\ before the \verb"\begin{document}" command. Note that if serious problems are encountered in the coding of a document (missing author-defined macros, for example), the typesetter may resort to rekeying it.

\section{Using the \texttt{interact} class file}

For convenience, simply copy the \texttt{interact.cls} file into the same directory as your manuscript files (you do not need to install it in your \TeX\ distribution). In order to use the \texttt{interact} document class, replace the command \verb"\documentclass{article}" at the beginning of your document with the command \verb"\documentclass{interact}".

The following document-class options should \emph{not} be used with the \texttt{interact} class file:
\begin{itemize}
  \item \texttt{10pt}, \texttt{11pt}, \texttt{12pt} -- unavailable;
  \item \texttt{oneside}, \texttt{twoside} -- not necessary, \texttt{oneside} is the default;
  \item \texttt{leqno}, \texttt{titlepage} -- should not be used;
  \item \texttt{twocolumn} -- should not be used (see Subsection~\ref{class});
  \item \texttt{onecolumn} -- not necessary as it is the default style.
\end{itemize}
To prepare a manuscript for a journal that is printed in A4 (two column) format, use the \verb"largeformat" document-class option provided by \texttt{interact.cls}; otherwise the class file produces pages sized for B5 (single column) format by default. The \texttt{geometry} package should not be used to make any further adjustments to the page dimensions.

%If your manuscript has supplementary content you can use the \verb"interact" class file to prepare all or part of it using the \verb"suppldata" document-class option, which will suppress the `article history' date. This option \emph{must not} be used on any primary content. Note that authors are solely responsible for the preparation of all supplemental material.

\section{Additional features of the \texttt{interact} class file}

\subsection{Title, authors' names and affiliations, abstracts and article types}

The title should be generated at the beginning of your article using the \verb"\maketitle" command.
In the final version the author name(s) and affiliation(s) must be followed immediately by \verb"\maketitle" as shown below in order for them to be displayed in your PDF document.
To prepare an anonymous version for double-blind peer review, you can put the \verb"\maketitle" between the \verb"\title" and the \verb"\author" in order to hide the author name(s) and affiliation(s) temporarily.
Next you should include the abstract if your article has one, enclosed within an \texttt{abstract} environment.
The \verb"\articletype" command is also provided as an \emph{optional} element which should \emph{only} be included if your article actually needs it.
For example, the titles for this document begin as follows:
\begin{verbatim}
\articletype{ARTICLE TEMPLATE}

\title{Taylor \& Francis \LaTeX\ template for authors (\textsf{Interact}
layout + reference style S)}

\author{
\name{A.~N. Author\textsuperscript{a}\thanks{CONTACT A.~N. Author.
Email: latex.helpdesk@tandf.co.uk} and John Smith\textsuperscript{b}}
\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton
Park, Abingdon, UK; \textsuperscript{b}Institut f\"{u}r Informatik,
Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany} }

\maketitle

\begin{abstract}
This template is for authors who are preparing a manuscript for a
Taylor \& Francis journal using the \LaTeX\ document preparation system
and the \texttt{interact} class file, which is available via selected
journals' home pages on the Taylor \& Francis website.
\end{abstract}
\end{verbatim}

An additional abstract in another language (preceded by a translation of the article title) may be included within the \verb"abstract" environment if required.

A graphical abstract may also be included if required. Within the \verb"abstract" environment you can include the code
\begin{verbatim}
\\\resizebox{25pc}{!}{\includegraphics{abstract.eps}}
\end{verbatim}
where the graphical abstract is to appear, where \verb"abstract.eps" is the name of the file containing the graphic (note that \verb"25pc" is the recommended maximum width, expressed in pica, for the graphical abstract in your manuscript).

\subsection{Abbreviations}

A list of abbreviations may be included if required, enclosed within an \texttt{abbreviations} environment, i.e.\ \verb"\begin{abbreviations}"\ldots\verb"\end{abbreviations}", immediately following the \verb"abstract" environment.

\subsection{Keywords}

A list of keywords may be included if required, enclosed within a \texttt{keywords} environment, i.e.\ \verb"\begin{keywords}"\ldots\verb"\end{keywords}". Additional keywords in other languages (preceded by a translation of the word `keywords') may also be included within the \verb"keywords" environment if required.

\subsection{Subject classification codes}

AMS, JEL or PACS classification codes may be included if required. The \texttt{interact} class file provides an \texttt{amscode} environment, i.e.\ \verb"\begin{amscode}"\ldots\verb"\end{amscode}", a \texttt{jelcode} environment, i.e.\ \verb"\begin{jelcode}"\ldots\verb"\end{jelcode}", and a \texttt{pacscode} environment, i.e.\ \verb"\begin{pacscode}"\ldots\verb"\end{pacscode}" to assist with this.

\subsection{Additional footnotes to the title or authors' names}

The \verb"\thanks" command may be used to create additional footnotes to the title or authors' names if required. Footnote symbols for this purpose should be used in the order
$^\ast$~(coded as \verb"$^\ast$"), $\dagger$~(\verb"$\dagger$"), $\ddagger$~(\verb"$\ddagger$"), $\S$~(\verb"$\S$"), $\P$~(\verb"$\P$"), $\|$~(\verb"$\|$"),
$\dagger\dagger$~(\verb"$\dagger\dagger$"), $\ddagger\ddagger$~(\verb"$\ddagger\ddagger$"), $\S\S$~(\verb"$\S\S$"), $\P\P$~(\verb"$\P\P$").

Note that any \verb"footnote"s to the main text will automatically be assigned the superscript symbols 1, 2, 3, etc. by the class file.\footnote{If preferred, the \texttt{endnotes} package may be used to set the notes at the end of your text, before the bibliography. The symbols will be changed to match the style of the journal if necessary by the typesetter.}

\section{Some guidelines for using the standard features of \LaTeX}

\subsection{Sections}

The \textsf{Interact} layout style allows for five levels of section heading, all of which are provided in the \texttt{interact} class file using the standard \LaTeX\ commands \verb"\section", \verb"\subsection", \verb"\subsubsection", \verb"\paragraph" and \verb"\subparagraph". Numbering will be automatically generated for all these headings by default.

\subsection{Lists}

Numbered lists are produced using the \texttt{enumerate} environment, which will number each list item with arabic numerals by default. For example,
\begin{enumerate}
  \item first item
  \item second item
  \item third item
\end{enumerate}
was produced by
\begin{verbatim}
\begin{enumerate}
  \item first item
  \item second item
  \item third item
\end{enumerate}
\end{verbatim}
Alternative numbering styles can be achieved by inserting an optional argument in square brackets to each \verb"item", e.g.\ \verb"\item[(i)] first item"\, to create a list numbered with roman numerals at level one.

Bulleted lists are produced using the \texttt{itemize} environment. For example,
\begin{itemize}
  \item First bulleted item
  \item Second bulleted item
  \item Third bulleted item
\end{itemize}
was produced by
\begin{verbatim}
\begin{itemize}
  \item First bulleted item
  \item Second bulleted item
  \item Third bulleted item
\end{itemize}
\end{verbatim}

\subsection{Figures}

The \texttt{interact} class file will deal with positioning your figures in the same way as standard \LaTeX. It should not normally be necessary to use the optional \texttt{[htb]} location specifiers of the \texttt{figure} environment in your manuscript; you may, however, find the \verb"[p]" placement option or the \verb"endfloat" package useful if a journal insists on the need to separate figures from the text.

Figure captions appear below the figures themselves, therefore the \verb"\caption" command should appear after the body of the figure. For example, Figure~\ref{sample-figure} with caption and sub-captions is produced using the following commands:
\begin{verbatim}
\begin{figure}
\centering
\subfloat[An example of an individual figure sub-caption.]{%
\resizebox*{5cm}{!}{\includegraphics{graph1.eps}}}\hspace{5pt}
\subfloat[A slightly shorter sub-caption.]{%
\resizebox*{5cm}{!}{\includegraphics{graph2.eps}}}
\caption{Example of a two-part figure with individual sub-captions
 showing that captions are flush left and justified if greater
 than one line of text.} \label{sample-figure}
\end{figure}
\end{verbatim}
\begin{figure}
\centering
\subfloat[An example of an individual figure sub-caption.]{%
\resizebox*{5cm}{!}{\includegraphics{graph1.eps}}}\hspace{5pt}
\subfloat[A slightly shorter sub-caption.]{%
\resizebox*{5cm}{!}{\includegraphics{graph2.eps}}}
\caption{Example of a two-part figure with individual sub-captions
 showing that captions are flush left and justified if greater
 than one line of text.} \label{sample-figure}
\end{figure}

To ensure that figures are correctly numbered automatically, the \verb"\label" command should be included just after the \verb"\caption" command, or in its argument.

The \verb"\subfloat" command requires \verb"subfig.sty", which is called in the preamble of the \texttt{interacttfssample.tex} file (to allow your choice of an alternative package if preferred) and included in the \textsf{Interact} \LaTeX\ bundle for convenience. Please supply any additional figure macros used with your article in the preamble of your .tex file.

The source files of any figures will be required when the final, revised version of a manuscript is submitted. Authors should ensure that these are suitable (in terms of lettering size, etc.) for the reductions they envisage.

The \texttt{epstopdf} package can be used to incorporate encapsulated PostScript (.eps) illustrations when using PDF\LaTeX, etc. Please provide the original .eps source files rather than the generated PDF images of those illustrations for production purposes.

\subsection{Tables}

The \texttt{interact} class file will deal with positioning your tables in the same way as standard \LaTeX. It should not normally be necessary to use the optional \texttt{[htb]} location specifiers of the \texttt{table} environment in your manuscript; you may, however, find the \verb"[p]" placement option or the \verb"endfloat" package useful if a journal insists on the need to separate tables from the text.

The \texttt{tabular} environment can be used as shown to create tables with single horizontal rules at the head, foot and elsewhere as appropriate. The captions appear above the tables in the \textsf{Interact} style, therefore the \verb"\tbl" command should be used before the body of the table. For example, Table~\ref{sample-table} is produced using the following commands:
\begin{table}
\tbl{Example of a table showing that its caption is as wide as
 the table itself and justified.}
{\begin{tabular}{lcccccc} \toprule
 & \multicolumn{2}{l}{Type} \\ \cmidrule{2-7}
 Class & One & Two & Three & Four & Five & Six \\ \midrule
 Alpha\textsuperscript{a} & A1 & A2 & A3 & A4 & A5 & A6 \\
 Beta & B2 & B2 & B3 & B4 & B5 & B6 \\
 Gamma & C2 & C2 & C3 & C4 & C5 & C6 \\ \bottomrule
\end{tabular}}
\tabnote{\textsuperscript{a}This footnote shows how to include
 footnotes to a table if required.}
\label{sample-table}
\end{table}
\begin{verbatim}
\begin{table}
\tbl{Example of a table showing that its caption is as wide as
 the table itself and justified.}
{\begin{tabular}{lcccccc} \toprule
 & \multicolumn{2}{l}{Type} \\ \cmidrule{2-7}
 Class & One & Two & Three & Four & Five & Six \\ \midrule
 Alpha\textsuperscript{a} & A1 & A2 & A3 & A4 & A5 & A6 \\
 Beta & B2 & B2 & B3 & B4 & B5 & B6 \\
 Gamma & C2 & C2 & C3 & C4 & C5 & C6 \\ \bottomrule
\end{tabular}}
\tabnote{\textsuperscript{a}This footnote shows how to include
 footnotes to a table if required.}
\label{sample-table}
\end{table}
\end{verbatim}

To ensure that tables are correctly numbered automatically, the \verb"\label" command should be included just before \verb"\end{table}".

The \verb"\toprule", \verb"\midrule", \verb"\bottomrule" and \verb"\cmidrule" commands are those used by \verb"booktabs.sty", which is called by the \texttt{interact} class file and included in the \textsf{Interact} \LaTeX\ bundle for convenience. Tables produced using the standard commands of the \texttt{tabular} environment are also compatible with the \texttt{interact} class file.

\subsection{Landscape pages}

If a figure or table is too wide to fit the page it will need to be rotated, along with its caption, through 90$^{\circ}$ anticlockwise. Landscape figures and tables can be produced using the \verb"rotating" package, which is called by the \texttt{interact} class file. The following commands (for example) can be used to produce such pages.
\begin{verbatim}
\setcounter{figure}{1}
\begin{sidewaysfigure}
\centerline{\epsfbox{figname.eps}}
\caption{Example landscape figure caption.}
\label{landfig}
\end{sidewaysfigure}
\end{verbatim}
\begin{verbatim}
\setcounter{table}{1}
\begin{sidewaystable}
 \tbl{Example landscape table caption.}
  {\begin{tabular}{@{}llllcll}
    .
    .
    .
  \end{tabular}}\label{landtab}
\end{sidewaystable}
\end{verbatim}
Before any such float environment, use the \verb"\setcounter" command as above to fix the numbering of the caption (the value of the counter being the number given to the preceding figure or table). Subsequent captions will then be automatically renumbered accordingly. The \verb"\epsfbox" command requires \verb"epsfig.sty", which is called by the \texttt{interact} class file and is also included in the \textsf{Interact} \LaTeX\ bundle for convenience.

Please note that if the \verb"endfloat" package is used, one or both of the commands
\begin{verbatim}
\DeclareDelayedFloatFlavor{sidewaysfigure}{figure}
\DeclareDelayedFloatFlavor{sidewaystable}{table}
\end{verbatim}
will need to be included in the preamble of your .tex file, after the \verb"endfloat" package is loaded, in order to process any landscape figures and/or tables correctly.

\subsection{Theorem-like structures}

A predefined \verb"proof" environment is provided by the \texttt{amsthm} package (which is called by the \texttt{interact} class file), as follows:
\begin{proof}
More recent algorithms for solving the semidefinite programming relaxation are particularly efficient, because they explore the structure of the MAX-CUT problem.
\end{proof}
\noindent This was produced by simply typing:
\begin{verbatim}
\begin{proof}
More recent algorithms for solving the semidefinite programming
relaxation are particularly efficient, because they explore the
structure of the MAX-CUT problem.
\end{proof}
\end{verbatim}
Other theorem-like environments (theorem, definition, remark, etc.) need to be defined as required, e.g.\ using \verb"\newtheorem{theorem}{Theorem}" in the preamble of your .tex file (see the preamble of \verb"interacttfssample.tex" for more examples). You can define the numbering scheme for these structures however suits your article best. Please note that the format of the text in these environments may be changed if necessary to match the style of individual journals by the typesetter during preparation of the proofs.

\subsection{Mathematics}

\subsubsection{Displayed mathematics}

The \texttt{interact} class file will set displayed mathematical formulas centred on the page without equation numbers if you use the \texttt{displaymath} environment or the equivalent \verb"\[...\]" construction. For example, the equation
\[
 \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
\]
was typeset using the commands
\begin{verbatim}
\[
 \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
\]
\end{verbatim}

For those of your equations that you wish to be automatically numbered sequentially throughout the text for future reference, use the \texttt{equation} environment, e.g.
\begin{equation}
 \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
\end{equation}
was typeset using the commands
\begin{verbatim}
\begin{equation}
 \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
\end{equation}
\end{verbatim}

Part numbers for sets of equations may be generated using the \texttt{subequations} environment, e.g.
\begin{subequations} \label{subeqnexample}
\begin{equation}
     \varepsilon \rho w_{tt}(s,t) = N[w_{s}(s,t),w_{st}(s,t)]_{s},
     \label{subeqnparta}
\end{equation}
\begin{equation}
     w_{tt}(1,t)+N[w_{s}(1,t),w_{st}(1,t)] = 0,   \label{subeqnpartb}
\end{equation}
\end{subequations}
which was typeset using the commands
\begin{verbatim}
\begin{subequations} \label{subeqnexample}
\begin{equation}
     \varepsilon \rho w_{tt}(s,t) = N[w_{s}(s,t),w_{st}(s,t)]_{s},
     \label{subeqnparta}
\end{equation}
\begin{equation}
     w_{tt}(1,t)+N[w_{s}(1,t),w_{st}(1,t)] = 0,   \label{subeqnpartb}
\end{equation}
\end{subequations}
\end{verbatim}
This is made possible by the \texttt{amsmath} package, which is called by the class file. If you put a \verb"\label" just after the \verb"\begin{subequations}" command, references can be made to the collection of equations, i.e.\ `(\ref{subeqnexample})' in the example above. Or, as the example also shows, you can label and refer to each equation individually -- i.e.\ `(\ref{subeqnparta})' and `(\ref{subeqnpartb})'.

Displayed mathematics should be given end-of-line punctuation appropriate to the running text sentence of which it forms a part, if required.

\subsubsection{Math fonts}

\paragraph{Superscripts and subscripts}
Superscripts and subscripts will automatically come out in the correct size in a math environment (i.e.\ enclosed within \verb"\(...\)" or \verb"$...$" commands in running text, or within \verb"\[...\]" or the \texttt{equation} environment for displayed equations). Sub/superscripts that are physical variables should be italic, whereas those that are labels should be roman (e.g.\ $C_p$, $T_\mathrm{eff}$). If the subscripts or superscripts need to be other than italic, they must be coded individually.

\paragraph{Upright Greek characters and the upright partial derivative sign}
Upright lowercase Greek characters can be obtained by inserting the letter `u' in the control code for the character, e.g.\ \verb"\umu" and \verb"\upi" produce $\umu$ (used, for example, in the symbol for the unit microns -- $\umu\mathrm{m}$) and $\upi$ (the ratio of the circumference of a circle to its diameter). Similarly, the control code for the upright partial derivative $\upartial$ is \verb"\upartial". Bold lowercase as well as uppercase Greek characters can be obtained by \verb"{\bm \gamma}", for example, which gives ${\bm \gamma}$, and \verb"{\bm \Gamma}", which gives ${\bm \Gamma}$.

\section*{Acknowledgement(s)}

An unnumbered section, e.g.\ \verb"\section*{Acknowledgements}", may be used for thanks, etc.\ if required and included \emph{in the non-anonymous version} before any Notes or References.

\section*{Disclosure statement}

An unnumbered section, e.g.\ \verb"\section*{Disclosure statement}", may be used to declare any potential conflict of interest and included \emph{in the non-anonymous version} before any Notes or References, after any Acknowledgements and before any Funding information.

\section*{Funding}

An unnumbered section, e.g.\ \verb"\section*{Funding}", may be used for grant details, etc.\ if required and included \emph{in the non-anonymous version} before any Notes or References.

\section*{Notes on contributor(s)}

An unnumbered section, e.g.\ \verb"\section*{Notes on contributors}", may be included \emph{in the non-anonymous version} if required. A photograph may be added if requested.

\section*{Nomenclature/Notation}

An unnumbered section, e.g.\ \verb"\section*{Nomenclature}" (or \verb"\section*{Notation}"), may be included if required, before any Notes or References.

\section*{Notes}

An unnumbered `Notes' section may be included before the References (if using the \verb"endnotes" package, use the command \verb"\theendnotes" where the notes are to appear, instead of creating a \verb"\section*").

\section{References}

\subsection{References cited in the text}

References should be cited in the text by numbers in square brackets based on the order in which they appear in an alphabetical list of references at the end of the document (not the order of citation), so the first reference cited in the text might be [23]. For example, these may take the forms [32], [5,\,6,\,14], [21--55] (\emph{not} [21]--[55]). For further details on this reference style, see the Instructions for Authors on the Taylor \& Francis website.

Each bibliographic entry has a key, which is assigned by the author and is used to refer to that entry in the text. In this document, the key \verb"Ali95" in the citation form \verb"\cite{Ali95}" produces `\cite{Ali95}', and the keys \verb"Bow76" and \verb"BGP02" in the citation form \verb"\cite{Bow76,BGP02}" produce `\cite{Bow76,BGP02}'. The citation for a range of bibliographic entries (e.g. `\cite{GMW81,Hor96,Con96,Har97,FGK03,Fle80,Ste98,Ell98,Str97,Coo03,Kih01,Hol03,Hai01,Hag03,Hov03,Mul00,GHGsoft,Pow00}') will automatically be produced by \verb"\cite{GMW81,Hor96,Con96,Har97,FGK03,Fle80,Ste98,Ell98,Str97,Coo03,Kih01," \verb"Hol03,Hai01,Hag03,Hov03,Mul00,GHGsoft,Pow00}".
Optional notes may be included at the beginning and/or end of a citation by the use of square brackets, e.g. \verb"\cite[cf.][]{Tay03}" produces `\cite[cf.][]{Tay03}', and \verb"\cite[see][and references therein]{Wei95}" produces `\cite[see][and references therein]{Wei95}'.

\subsection{The list of references}

To produce the list of references, the bibliographic data about each reference item should be listed in \texttt{thebibliography} environment in alphabetical order. References with the same author or group of authors are further sorted chronologically, beginning with the earliest. The following list shows some sample references prepared in Taylor \& Francis' Reference Style S.

\begin{thebibliography}{99}

\bibitem{Ali95}
S. Alinhac, \emph{Blowup for Nonlinear Hyperbolic Equations}, Progress in
  Nonlinear Differential Equations and Applications Vol.~17, Birkh{\"a}user,
  Boston, MA, 1995.

\bibitem{Bow76}
F.E. Bowder (ed.), \emph{Nonlinear Operators and Nonlinear Equations of
  Evolution in Banach Spaces}, Proceedings of Symposia in Pure Mathematics
  Vol. 18, Part~2, American Mathematical Society, Providence, RI, 1976.

\bibitem{BGP02}
J. Burckhardt, M. Gunzburger, and J. Peterson, \emph{Insensitive functionals,
  inconsistent gradients, spurious minima, and regularized functionals in flow
  optimization problems}, Int. J. Comput. Fluid Dyn. 16 (2002), pp. 171--185.

\bibitem{Con96}
A.R. Conn and P.L. Toint, \emph{An algorithm using quadratic interpolation}, in
  \emph{Nonlinear Optimization and Applications}, G. Di~Pillo and F. Giannessi,
  eds., Kluwer Academic/Plenum Publishers, New York, 1996, pp. 27--47.

\bibitem{Coo03}
I.D. Coope and C.J. Price, \emph{Positive bases in numerical optimization},
  Comput. Optim. Appl. 21 (2003), pp. 169--175.

\bibitem{Ell98}
J.S. Ellenberg, \emph{Hilbert modular forms and the Galois representations
  associated to Hilbert-Blumenthal abelian varieties}, Ph.D. diss., Harvard
  University, 1998.

\bibitem{Fle80}
R. Fletcher, \emph{Practical Methods of Optimization}, 2nd ed., Vol.~2, Wiley
  and Sons, New York, 1980.

\bibitem{FGK03}
R. Fourer, D.M. Gay, and B.W. Kernighan, \emph{AMPL: A Modeling Language for
  Mathematical Programming}, 2nd ed., Thomson/Brooks/Cole, Pacific Grove, CA,
  2003.

\bibitem{GMW81}
P.E. Gill, W. Murray, and M.H. Wright, \emph{Practical Optimization}, Academic
  Press, London, 1981.

\bibitem{GHGsoft}
T.G. Golda, P.D. Hough, and G. Gay, \emph{APPSPACK (Asynchronous parallel
  pattern search package)}. Software available at
  http://software.sandia.gov/appspack.

\bibitem{Hag03}
J. Haglund, \emph{Conjectured statistics for the {$q,t$}-Catalan numbers},
  preprint (2003), to appear in Adv. Math. Available at
  http://www.math.upenn.edu/$\sim$jhaglund.

\bibitem{Hai01}
M. Haiman, \emph{Hilbert schemes, polygraphs, and the Macdonald positivity
  conjecture}, J. Amer. Math. Soc. 14 (2001), pp. 941--1006. Available at
  http://www.math.berkeley.edu/$\sim$mhaiman. MR~2002c:14008

\bibitem{Har97}
W.E. Hart, \emph{A stationary point convergence theory of evolutionary
  algorithms}, in \emph{Foundations of Genetic Algorithms 4}, R.K. Belew and
  M.D. Vose, eds., Morgan Kaufmann, San Francisco, 1997, pp. 127--134.

\bibitem{Hol03}
J. Holt, \emph{Multiple bumping of components of deformation spaces of
  hyperbolic 3-manifolds}, Amer. J. Math. 125 (2003), pp. 691--736. Available at
  http://muse.jhu.edu/journals/american\_journal\_of\_mathematics/v125/125.4holt.pdf.

\bibitem{Hor96}
U. Hornung (ed.), \emph{Homogenization and Porous Media}, Springer, Berlin,
  1996.

\bibitem{Hov03}
P. Hovland, \emph{Automatic differentiation and its role in simulation-based
  optimization}, IMA Workshop, Minneapolis, MN, 2003.

\bibitem{Kih01}
S. Kihara, \emph{On an elliptic curve over {$Q(t)$} of rank $\geq$14}, Proc.
  Japan Acad. Ser. A Math. Sci. 77 (2001), pp. 50--51. MR~2002a:11057

\bibitem{Mul00}
\emph{MultiSimplex 2.0}. Grabitech Solutions AB, Sundvall, Sweden, 2000;
  software available at http://www.multisimplex.com.

\bibitem{Pow00}
M.J.D. Powell, \emph{On the Lagrange functions of quadratic models that are
  defined by interpolation}, Tech. Rep. DAMTP 2000/NA10, Department of
  Applied Mathematics and Theoretical Physics, University of Cambridge,
  Cambridge, UK, 2000.

\bibitem{Ste98}
G.W. Stewart, \emph{Matrix Algorithms. Volume 1: Basic Decompositions}, SIAM,
  Philadelphia, 1998.

\bibitem{Str97}
N.P. Strickland, \emph{Finite subgroups of formal groups}, J. Pure Appl.
  Algebra 121 (1997), pp. 161--208.

\bibitem{Tay03}
R. Taylor, \emph{On the meromorphic continuation of degree two L-functions},
  preprint (2003). Available at http://abel.math.harvard.edu/$\sim$rtaylor/.

\bibitem{Wei95}
A. Weil, \emph{Basic Number Theory}, Springer-Verlag, Berlin, 1995.

\end{thebibliography}
\bigskip
\noindent This was produced by typing:
\medskip
\begin{verbatim}
\begin{thebibliography}{99}

\bibitem{Ali95}
S. Alinhac, \emph{Blowup for Nonlinear Hyperbolic Equations}, Progress
 in Nonlinear Differential Equations and Applications Vol.~17,
 Birkh{\"a}user, Boston, MA, 1995.

\bibitem{Bow76}
F.E. Bowder (ed.), \emph{Nonlinear Operators and Nonlinear Equations
 of Evolution in Banach Spaces}, Proceedings of Symposia in Pure
 Mathematics Vol. 18, Part~2, American Mathematical Society,
 Providence, RI, 1976.

\bibitem{BGP02}
J. Burckhardt, M. Gunzburger, and J. Peterson, \emph{Insensitive
 functionals, inconsistent gradients, spurious minima, and regularized
 functionals in flow optimization problems}, Int. J. Comput. Fluid Dyn.
 16 (2002), pp. 171--185.

\bibitem{Con96}
A.R. Conn and P.L. Toint, \emph{An algorithm using quadratic
 interpolation}, in \emph{Nonlinear Optimization and Applications},
 G. Di~Pillo and F. Giannessi, eds., Kluwer Academic/Plenum Publishers,
 New York, 1996, pp. 27--47.

\bibitem{Coo03}
I.D. Coope and C.J. Price, \emph{Positive bases in numerical
 optimization}, Comput. Optim. Appl. 21 (2003), pp. 169--175.

\bibitem{Ell98}
J.S. Ellenberg, \emph{Hilbert modular forms and the Galois
 representations associated to Hilbert-Blumenthal abelian varieties},
 Ph.D. diss., Harvard University, 1998.

\bibitem{Fle80}
R. Fletcher, \emph{Practical Methods of Optimization}, 2nd ed., Vol.~2,
 Wiley and Sons, New York, 1980.

\bibitem{FGK03}
R. Fourer, D.M. Gay, and B.W. Kernighan, \emph{AMPL: A Modeling
 Language for Mathematical Programming}, 2nd ed., Thomson/Brooks/Cole,
 Pacific Grove, CA, 2003.

\bibitem{GMW81}
P.E. Gill, W. Murray, and M.H. Wright, \emph{Practical Optimization},
 Academic Press, London, 1981.

\bibitem{GHGsoft}
T.G. Golda, P.D. Hough, and G. Gay, \emph{APPSPACK (Asynchronous
 parallel pattern search package)}. Software available at
 http://software.sandia.gov/appspack.

\bibitem{Hag03}
J. Haglund, \emph{Conjectured statistics for the {$q,t$}-Catalan
 numbers}, preprint (2003), to appear in Adv. Math. Available at
 http://www.math.upenn.edu/$\sim$jhaglund.

\bibitem{Hai01}
M. Haiman, \emph{Hilbert schemes, polygraphs, and the Macdonald
 positivity conjecture}, J. Amer. Math. Soc. 14 (2001), pp. 941--1006.
 Available at http://www.math.berkeley.edu/$\sim$mhaiman.
 MR~2002c:14008

\bibitem{Har97}
W.E. Hart, \emph{A stationary point convergence theory of evolutionary
 algorithms}, in \emph{Foundations of Genetic Algorithms 4}, R.K. Belew
 and M.D. Vose, eds., Morgan Kaufmann, San Francisco, 1997,
 pp. 127--134.

\bibitem{Hol03}
J. Holt, \emph{Multiple bumping of components of deformation spaces of
 hyperbolic 3-manifolds}, Amer. J. Math. 125 (2003), pp. 691--736.
 Available at
 http://muse.jhu.edu/journals/american\_journal\_of\_mathematics/v125/125.4holt.pdf.

\bibitem{Hor96}
U. Hornung (ed.), \emph{Homogenization and Porous Media}, Springer,
 Berlin, 1996.

\bibitem{Hov03}
P. Hovland, \emph{Automatic differentiation and its role in simulation-
 based optimization}, IMA Workshop, Minneapolis, MN, 2003.

\bibitem{Kih01}
S. Kihara, \emph{On an elliptic curve over {$Q(t)$} of rank $\geq$14},
 Proc. Japan Acad. Ser. A Math. Sci. 77 (2001), pp. 50--51.
 MR~2002a:11057

\bibitem{Mul00}
\emph{MultiSimplex 2.0}. Grabitech Solutions AB, Sundvall, Sweden,
 2000; software available at http://www.multisimplex.com.

\bibitem{Pow00}
M.J.D. Powell, \emph{On the Lagrange functions of quadratic models that
 are defined by interpolation}, Tech. Rep. DAMTP 2000/NA10, Department
 of Applied Mathematics and Theoretical Physics, University of
 Cambridge, Cambridge, UK, 2000.

\bibitem{Ste98}
G.W. Stewart, \emph{Matrix Algorithms. Volume 1: Basic Decompositions},
 SIAM, Philadelphia, 1998.

\bibitem{Str97}
N.P. Strickland, \emph{Finite subgroups of formal groups}, J. Pure
 Appl. Algebra 121 (1997), pp. 161--208.
\bibitem{Tay03}
R. Taylor, \emph{On the meromorphic continuation of degree two
 L-functions}, preprint (2003). Available at
 http://abel.math.harvard.edu/$\sim$rtaylor/.

\bibitem{Wei95}
A. Weil, \emph{Basic Number Theory}, Springer-Verlag, Berlin, 1995.

\end{thebibliography}
\end{verbatim}
\bigskip
\noindent Each entry takes the form:
\begin{verbatim}
\bibitem{key} Bibliography entry
\end{verbatim}
where `\texttt{key}' is the tag that is to be used as an argument for the \verb"\cite{}" commands in the text of the article and `\texttt{Bibliography entry}' is the material that is to appear in the list of references, suitably formatted. The commands
\begin{verbatim}
\usepackage[numbers,sort&compress]{natbib}
\bibpunct[, ]{[}{]}{,}{n}{,}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
\end{verbatim}
need to be included in the preamble of your .tex file in order to generate the citations and bibliography as described above.

Instead of typing the bibliography by hand, you may prefer to create the list of references using a \textsc{Bib}\TeX\ database. In order to do this, the \texttt{tfs.bst} file needs to be in your working folder or an appropriate directory, and the lines
\begin{verbatim}
\bibliographystyle{tfs}
\bibliography{interacttfssample}
\end{verbatim}
included where the list of references is to appear, where \texttt{tfs.bst} is the name of the \textsc{Bib}\TeX\ bibliography style file for Taylor \& Francis' Reference Style S and \texttt{interacttfssample.bib} is the bibliographic database included with the \textsf{Interact}-TFS \LaTeX\ bundle (to be replaced with the name of your own .bib file). \LaTeX/\textsc{Bib}\TeX\ will extract from your .bib file only those references that are cited in your .tex file and list them in the References section.

Please include a copy of your .bib file and/or the final generated .bbl file among your source files if your .tex file does not contain a reference list in a \texttt{thebibliography} environment.

\section{Appendices}

Any appendices should be placed after the list of references, beginning with the command \verb"\appendix" followed by the command \verb"\section" for each appendix title, e.g.
\begin{verbatim}
\appendix
\section{This is the title of the first appendix}
\section{This is the title of the second appendix}
\end{verbatim}
produces:\medskip

\noindent\textbf{Appendix A. This is the title of the first appendix}\medskip

\noindent\textbf{Appendix B. This is the title of the second appendix}\medskip

\noindent Subsections, equations, figures, tables, etc.\ within appendices will then be automatically numbered as appropriate. Some theorem-like environments may need to have their counters reset manually (e.g.\ if they are not numbered within sections in the main text). You can achieve this by using \verb"\numberwithin{remark}{section}" (for example) just after the \verb"\appendix" command.

Note that if the \verb"endfloat" package is used on a document containing any appendices, the \verb"\processdelayedfloats" command must be included immediately before the \verb"\appendix" command in order to ensure that the floats belonging to the main body of the text are numbered as such.

%\processdelayedfloats %%% See above for an explanation of why this command might be needed here.

\appendix

\section{Troubleshooting}

Authors may occasionally encounter problems with the preparation of a manuscript using \LaTeX. The appropriate action to take will depend on the nature of the problem:
\begin{enumerate}
\item[(i)] If the problem is with \LaTeX\ itself, rather than with the actual macros, please consult an appropriate \LaTeXe\ manual for initial advice. If the solution cannot be found, or if you suspect that the problem does lie with the macros, then please contact Taylor \& Francis for assistance (\texttt{latex.helpdesk@tandf.co.uk}), clearly stating the title of the journal to which you are submitting.
\item[(ii)] Problems with page make-up (e.g.\ occasional overlong lines of text; figures or tables appearing out of order): please do not try to fix these using `hard' page make-up commands -- the typesetter will deal with such problems. (You may, if you wish, draw attention to particular problems when submitting the final version of your manuscript.)
\item[(iii)] If a required font is not available on your system, allow \TeX\ to substitute the font and specify which font is required in a covering letter accompanying your files.
\end{enumerate}

\section{Obtaining the template and class file}

\subsection{Via the Taylor \& Francis website}

This article template and the \texttt{interact} class file may be obtained via the `Instructions for Authors' pages of selected Taylor \& Francis journals.

Please note that the class file calls up the open-source \LaTeX\ packages booktabs.sty, epsfig.sty and rotating.sty, which will, for convenience, unpack with the downloaded template and class file. The template calls for natbib.sty and subfig.sty, which are also supplied for convenience.

\subsection{Via e-mail}

This article template, the \texttt{interact} class file and the associated open-source \LaTeX\ packages are also available via e-mail. Requests should be addressed to \texttt{latex.helpdesk@tandf.co.uk}, clearly stating for which journal you require the template and class file.

\end{document}

### 2019-02-22 Paper/manuscript/manuscript_05.tex
restructuring

% interacttfssample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.

%\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables

%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty  %

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}

\graphicspath{{./images/}}

\bibpunct[, ]{[}{]}{,}{}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\begin{document}

%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{Hilbert space methods to approximate Gaussian processes using Stan}

%\author{
%\name{Gabriel Riutort-Mayol, Michael Riis Anderseen, Aki Vehtari, Paul-Christian Burkner}
%\affil{\textsuperscript{a}Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain; }
%}

\maketitle

\begin{abstract}

\end{abstract}

\begin{keywords}
Gaussian processes; Low-rank Gaussian processes; Hilbert Space methods; Sparse Gaussian processes
\end{keywords}

\tableofcontents

%Contents
\newpage

\section{Introduction}\label{sec:bf_intro}

Gaussian processes (GPs) are natural and flexible non-parametric models to define probability distributions over multi-dimensional functions \citep{rasmussen2006gaussian,neal1997monte}. This enables to apply probabilistic inference directly to functions. The defining feature of a GP model is that the joint distribution of function values is a multivariate normal distribution which is characterized by a mean function and a covariance function. The covariance function encodes our prior assumptions about the function, such as continuity, smoothness, periodicity and scale properties of the GP, and defines a correlation structure which characterize the correlations between function values at different inputs. A GP model with a covariance function allows for non-linear effects on the function values and can handle interactions between input variables implicitly. Several types of covariance functions with different behaviors can be used, from stationary (e.g. exponentiated quadratic, Matern, periodic, quasyperiodic...) to non-stationary (e.g. dot product, neuronal network,...). Combinations, sum and product, of different covariance function can also be used in order to obtain more modelling flexibility on real complex problems. GP models are often referred to as non-parametric because the number  of parameters is not fixed, but rather spanned by the number of data points, which allows to adapt model complexity to the data. GP models are powerful non-parametric prior models for regression and classification \citep{rasmussen2006gaussian}, and they can also be used as non-linear components in additive models. GP models is of broad interest across machine learning and statistics, with applications in the fields of spatial epidemiology \citep{diggle2013statistical,carlin2014hierarchical}, robotics and control \citep{deisenroth2015gaussian}, signal processing \citep{sarkka2013spatiotemporal}, Bayesian optimization and probabilistic numerics, \citep{roberts2010bayesian,briol2015probabilistic,hennig2015probabilistic}, and many others.

One of the main limitations of GPs is that computational demands and memory requirements scale as $O(n^3)$ and $O(n^2)$ respectively, effectively limiting a direct implementation to a number of training samples $n$ at most a few thousand cases.   This problem becomes especially important when using sampling methods for full Bayesian inference, when in each sampling step we need to invert the Gram matrix, usually through Cholesky factorization.

In this paper we want to make full Bayesian inference over Gaussian process models accessible to a broader audience by making them fast and easy to apply. We implement novel method recently  developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models. This method decomposes the GP kernel into basis functions which represent the GP as a linear model and makes it faster to fit. The set of Laplace eigenfunctions considered in this paper forms such a basis and we can compute them analytically. Furthermore, they are independent of the choice of kernel (including the hyperparameters), which means that we do not need to autodiff through the kernel and the Cholesky. In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method. The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions, since it is known that there exist a correlation between the latent values and the hyperparameters which makes inference more difficult especially when the dimension of latent values is high.

There already exist several approximate methods that alleviate these computational demands. Sparse GPs, which are approximate methods to fit GP models, are based on building a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables'. They do reduce the dimension of the posterior, reducing the storage requirements to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$. However it still requires Cholesky factorizations. Furthermore convergence to the full GP model is not straight forward, and reducing the amount of data while maintaining the posterior densities is not a simple procedure. Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods. More recently Bui et. al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time. \cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI).

Other global approach make use of the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}. It is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a specific choice of covariance function, but they do not generally have correspondence with GP models, and they do not have the flexibility of choosing different covariance functions as model structure. The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving,gal2015improving}. Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat´ern class of kernels. Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels.

The approach we propose in this paper is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models. This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods. This results in a reduced-rank approximation for the covariance function.

While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.

There are several GP specific software, which have efficient
computation for restricted set of GP based models allowing scaling to
bigger data sets. In this paper we do not focus on the fastest
possible inference for a specific restricted models, but instead we
are interested in how GPs can be easily used as modular components in
probabilistic programming frameworks.

As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology. Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function. The performance and accuracy of the method are directly related with these factors. The time of computation is mainly dependent on the number of basis functions. Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects. We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the functions.  We make recomendations for the values of these factors based on the recognized relations among them. We provide useful graphs of these relations that will help the users to improve performance and save time of computation. We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.

The rest of the paper is structered as follows. Section 2 describe  the proposed model based on the Hilbert space methods for low-rank approximation to GP. In Section 3 we analyze the accuracy of the method and its dependency on existing relations among the number of basis functions, the box size factor and the roughness or smoothness of the process.

\vspace{3mm}
\section{Method}\label{sec:bf_method}

\subsection{Gaussian process model}

A GP model can be seen as a probability model over an infinite collection of random varibales. However, formally, the defining feature of a GP model is that the joint distribution of function values at any finite number of input values $\mathbf{x}$ is a multivariate normal distribution, 
%
\begin{equation}
f(\mathbf{x}) \sim \mathcal{N}(\mathbf{m},\mathbf{K})
\end{equation}

\noindent with $\mathbf{m}$ the mean vector and $\mathbf{K}$ the covariance matrix. The element $\mathbf{m}_i$ is given by the mean function $m(\mathbf{x}_i)$ and the element $\mathbf{K}_{ij}$ is given by the covariance function $k(x_i,x_j|\theta)$ which is a function of the inputs values $\mathbf{x}$ and the parameters $\theta$. The mean and covariance functions represent how the random functions behave on average and hoe the different points in the input space co-vary with respect to each other:

 In general the mean function $m(\mathbf{x})$ can be any function of the inputs, although this is usually set to zero for simplifying notation.

\vspace{0.5cm}
\noindent ... \\
... \\
...

\vspace{0.5cm}
A stationary covariance function is a function of $x-x'$, so we can write $k(x,x') = k(x-x')$. Isotropic covariance functions are those that are function of the distance between observations, $k(x,x') = k(|x-x'|) = k(r)$. The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which are
%
\begin{eqnarray}
k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r/\ell^2), \nonumber \\
k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
\end{eqnarray}

\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the lengt-scale and magnitud, respectively, of the kernel. The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function. These covariance functions can be easily generalized to the multidimensional case considering multidimensional inputs $\mathbf{r}$ and the multidimensional length-scale $\boldsymbol{\ell}$.

The spectral density functions associated with the univariate Mat\'ern covariance functions written above are
%
\begin{eqnarray}
S_{\nu=\infty}(w)&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 w^2), \nonumber \\
S_{\nu=\frac{1}{2}}(w)&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \nonumber \\
S_{\nu=\frac{3}{2}}(w)&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \nonumber \\
S_{\nu=\frac{5}{2}}(w)&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3}, \nonumber 
\end{eqnarray}

\noindent where variable $w$ is in the frequency domain.

\subsection{Hilbert space-based approximate Gaussian process model}

The approximate Gaussian process method, developed by \cite{solin2018hilbert} and implemented in this paper, lay on the basis of considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators. Then, the pseudo-differential operator is approximated with Hilbert space methods on compact subsets $\Omega \subset {\rm I\!R}^D$, and with some boundary condition.

First, we focus on the unidimensional case of the input space, such as $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value.

\vspace{0.2cm}
The approximate method leads to the approximation of the covariance function between two input values $\{x,x'\} \in \{-L,L\}$ as 
%
\begin{equation}\label{approxcov}
k(x,x') \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
\end{equation}

\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ and $\theta$ the set of hyperparameters of $k$; a stationary covariance function can be equivalently represented in terms of the spectral density \citep{rasmussen2006gaussian}, then the spectral density is a function of the hyperparameters $\theta$. $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator. Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
%
\begin{eqnarray}\label{eigenproblem}
-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
\end{eqnarray}

\noindent The eigenvalues $\lambda_j>0$ are real and positive, because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in eq. (\ref{eigenproblem}) are sinusoidal functions,
%
\begin{eqnarray}
\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
\end{eqnarray}

If we truncate the sum in (\ref{approxcov}) up to $J$ ($j=1,\cdots,J$), the approximate covariance function can be represented as
%
\begin{equation}
k(x,x') \approx \boldsymbol{\phi}(x)^\intercal \Delta \boldsymbol{\phi}(x'), \nonumber
\end{equation}

\noindent where $\boldsymbol{\phi}(x)=\{\phi_j(x)\}_{j=1}^{J} \in {\rm I\!R}^{J}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{J\times J}$ the diagonal matrix of espectral densities $S_{\theta}(\sqrt{\lambda_j})$, 
%
\begin{eqnarray}
\Delta &=&  \begin{bmatrix}
    S_{\theta}(\sqrt{\lambda_1}) & & \\
    & \ddots & \nonumber \\
    & & S_{\theta}(\sqrt{\lambda_J}) \\
  \end{bmatrix}.
\end{eqnarray}

Thus, the Gram matrix $\text{K}$ of the covariance function $k$ for a collection of observations $i=1,\cdots,N$ and collection of input values $\{x^i\}_{i=1}^{N} \in {\rm I\!R}^{N}$ can be represented as follows
%
\begin{equation}
\text{K}= \Phi \Delta \Phi^\intercal, \nonumber
\end{equation}

\noindent where $\Phi \in {\rm I\!R}^{N\times J}$ is the matrix of eigenfunctions $\phi_j(x^i)$,
%
\begin{eqnarray}
\Phi &=&  \left[ {\begin{array}{ccc}
   \phi_1(x^1) & \cdots & \phi_J(x^1)  \\
    \vdots &\ddots & \vdots  \nonumber \\ 
    \phi_1(x^N) & \cdots & \phi_J(x^N) \\
  \end{array} } \right].
\end{eqnarray}

\noindent Now, the Gaussian process prior for the function $f$ can be re-defined as
%
\begin{equation}
\mathbf{f} \sim \text{GP}(0,\Phi \Delta \Phi^\intercal). \nonumber
\end{equation}

\noindent This equivalently leads to a linear representation of the function $f$,
%
\begin{equation}\label{approxf}
f(x) \approx \sum_{j} \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
\end{equation}

\noindent where $\beta_j \sim \text{N}(0,1)$ is a random variable with zero mean and unitary variance. That is, the function $f$ can be approximated with a finite basis function expansion (eigenfunctions $\phi_j$ of the Laplace operator), scaled by the squared root of the spectral density as a function of the eigenvalues $\lambda_j$. The eigenfunctions $\phi_j$ do not depend on the kernel hyperparameters $\theta$, the only dependence on $\theta$ is through the spectral density $S_{\theta}$. The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and the spectral density goes rapidly to zero for bounded covariance functions. Therefore, it is expected a good approximation in eq. (\ref{approxf}) for a finite number, $J$, of terms in the series as long as the inputs values $x^i$ are not near the boundaries of the domain $[-L,L]$, where the Laplacian was taken to be zero.

The computational cost of this approximate model scales $O(n\cdot J+J)$, where $n$ is the number of observations and $J$ the number of basis functions.

\subsection{Generalization to multidimensional input space}

We are going to generalize the results from the previous section to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions. We assume the same number $J$ of basis functions terms in the series expansion for every input dimension, however it is straight forward to generalize this to different number of term series $J_d$ for every dimension.

\vspace{0.2cm}
In a $D$-dimensional input space the total number of eigenfunctions and eigenvalues in the approximation corresponds to the total number of $D$-tuples over $J$, which is $J^D$. Let $\mathbb{S}\in {\rm I\!R}^{J^D \times D}$ be the set of $D$-tuples elements. Each new eigenfunction $\phi^{\ast}_j$ correspond to the product of the univariate eigenfunctions whose indices corresponds to the the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$. And each new eigenvalue $\lambda^{\ast}_j$ is a $D$-vector whose elements are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$. Let $j=\{1,\cdots,J^D\}$,
%
\begin{eqnarray}
\phi^{\ast}_j(\mathbf{x}) &=& \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right)  \nonumber \\
%
\lambda^{\ast}_j &=& \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  \nonumber
\end{eqnarray}

As an example we show the matrix $\mathbb{S}$ for a $three$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2,\mathbf{x}_3\}$ ($D=3$) and $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively. The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$. The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
%
\begin{eqnarray}
\mathbb{S}=
\left[ {\begin{array}{ccc}
1 & 1 & 1 \nonumber \\
1 & 1 & 2 \\
1 & 1 & 3 \\
1 & 2 & 1 \\
1 & 2 & 2 \\
1 & 2 & 3 \\
2 & 1 & 1 \\
2 & 1 & 2 \\
2 & 1 & 3 \\
2 & 2 & 1 \\
2 & 2 & 2 \\
2 & 2 & 3 
\end{array} } \right]
\end{eqnarray}

The approximate covariance function is represented as
%
\begin{equation}\label{approxcov_multi}
k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{J^D} S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right) \phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
\end{equation}

\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function. The $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
%
\begin{eqnarray}
S^{\ast}_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \text{exp}\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
%
S^{\ast}_{\nu}(\mathbf{w})&=& \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
\end{eqnarray}

\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain. We can represent the approximate series expansion of the function $f$ in the multidimensional input space as,
%
\begin{equation}\label{approxf}
f(\mathbf{x}) \approx \sum_{j} \left( S^{\ast}_{\theta}(\sqrt{\boldsymbol{\lambda^{\ast}_j}})\right)^{1/2} \phi^{\ast}_j(\mathbf{x}) \beta_j, 
\end{equation}

\noindent where $\beta_j \sim \text{N}(0,1)$ is a random variable with zero mean and unitary variance. The computational cost in a multidimensional setting scales $O((n+1)\cdot \prod_{d=1}^D J_d )$, where $n$ is the number of observations and $J_d$ is the number of basis functions considered in the dimension $d$.

\subsection{Learning hyperparameters and model inference}

\vspace{2mm}
$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.

\vspace{2mm}
$\bullet$ The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.

\vspace{2mm}
$\bullet$ The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands. The weights depend on the hyperparameters, then they have to be computed in every step of the HMC sampling method.

\vspace{2mm}
$\bullet$ The linear model is computed with complexity $O(nm)$, computed in every step of the HMC sampling method.

\vspace{2mm}
$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method. In addition, the computation of the automatic differentation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$. After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer. (Arno's paper, pag. 7)

\vspace{2mm}
$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions. especially when sampling methods for inference are used. GP prior as latent functions is needed in generalized models.

In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:

\vspace{2mm}
$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation. This operation has to be computed at every step of the HMC or optimizer.

\vspace{2mm}
$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional. It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitud parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

\vspace{2mm}
$\bullet$ In addition, the computation of the automatic differentation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ In a regular GP model the automatic differentation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$. This operation has to be computed at every step of the HMC.

\vspace{2mm}
$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.

\section{The accuracy of the approximation}

The performance and accuracy of the method are directly related with the number of basis functions and the box size, and, at the same time, these two factors depend on the smoothness or roughness of the process. The time of computation is mainly dependent on the number of basis functions. Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.

\vspace{2mm}
$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the function.

\vspace{2mm}
$\bullet$ We make recomendations for the values of these factors based on the recognized relations among them. We provide useful graphs of these relations that will help the users to improve performance and save time of computation.

\vspace{2mm}
$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.

\subsection{Dependency on the number of basis functions and the boundary condition}

The approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, e.g. in a 1-$D$ input space $x\in \Omega=[-L,L]$:
%
\begin{equation}\label{approxcov}
k(\tau) \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
\end{equation}

\noindent where $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$. The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are 
%
\begin{eqnarray}
\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
\end{eqnarray}

The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity $J=1,\dots,j,\dots,\infty$. The number of basis functions can be truncated that the difference between the covariance function and the approximation be less than a threshold $\epsilon$,
%
\begin{eqnarray}\label{diff_covs}
\int k(\tau)d(\tau) - \int \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
\end{eqnarray}

The finite number of basis functions in the approximation to satisfy eq. (\ref{diff_covs}) will depend on the smoothness or roughness of the function, that is, on the lengthscale $\ell$. The approximation will also depends on the box size $L$ which will affect the performance specially near the boundaries.

To illustrate the effects of these two factors, number of basis functions $J$ and box size $L$, on the performance of the approximation, let's consider a simulated dataset of 75 noisy observations from a GP prior with lengthscale $\ell=0.3$ and magnitud $\alpha=1$. The additive noise added to the GP prior simulations is $\sigma=0.2$. A regular GP model and different approximate GP models with different choices for the number of basis function $J$ and box size $L$ are fitted on this dataset.

For a fixed $L$, let's suppose it is set to a right choice, Figure \ref{image2_J} shows the effect of the number of basis functions in approaching the posterior mean and the covariance function. The number of basis functions affects the non-linearity or roughness (or smoothness) of the posteriors. Fewer basis functions implies smoother covariance functions and consequently smoother posteriors. The more roughness the process is, the more basis functions are needed.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.4, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
\caption{(left) Posterior distributions for different number of basis functions $J$. (right) Covariance function for different number of basis functions $J$.}
  \label{fig1_Post_J}
\end{figure}

Similarly, for a fixed number of basis functions $J$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of the box size $L$ in approaching the posterior mean and the covariance function. The box size mainly affects the approximation near the boundaries of the function, and affects the covariances at long distances.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig2_Cov_L.pdf}}
\caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
  \label{fig2_Post_L}
\end{figure}

Therefore, the approximation to the covariance function depends on the number of basis functions $J$ used in the series expansion and the value for boundary condition or box size $L$. Additionally, there exit a relation among the number of basis functions, the box size and the lengthscale or roughness of the process on the performance of the approximation. Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in order to achieve a good approximation to the covariance function of the regular GP using eq. (\ref{diff_covs}). On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data. The countour lines gather the box size which is also normalized by the half-range of the data, to which we refer as the boundary factor $c$, where $L=c \cdot \text{range}(x)/2$. And the Y-axis gathers the number of basis functions. Thus, this figure give us the minimum number of basis functions needed to achieve a good approximation in terms of satisfying eq. (\ref{diff_covs}), for a certain GP process with lengthscale $\ell$ and given a fixed boundary factor $c$. Alternatively, this figure could also be read as the maximum boundary factor $c$ that we should use given a number of basis functions, for certain GP process with certain lengthscale. And also this figure could also be read as the maximum roughness in a process that can be fitted properly given a number of basis functions and a box size. Finally, this figure also gives us the minimum value for the boundary factor $c$ which is suitable for certain roughness or number of basis functions. This figure basically collects the behaviour of the approximate GP model in function of the lengthscale, number of basis functions and boundary condition:

$\bullet$ As the lengthscale or smoothness of the process increase, the boundary factor and the number of basis functions needed for the approach decrease.

$\bullet$ There is a minimum boundary factor to achieve a good approach in function of the lengthscale (The contour lines have an end in function of the lengthscale).

$\bullet$ The lower boundary factor, the fewer basis functions, and the lower lengthscales suitable to be fitted.

$\bullet$ This plot serves as a diagnosis tool in the sense that if the estimated lengthscale is lower than that minimun generated by the figure for the chosen number of basis functions and boundary factor means that we would need to increase the number of basis functions or decrease the boundary factor.

$\bullet$ On the other hand, if the lengthscale is bigger than that generated by the figure then it means that the fit should be good.

Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $J$ and the boundary factor $c$ to approach a certain process characterized by its lengthscale. This Figure will allow us to optimize computations which depends on the number of basis functions. Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit. This figure also provides a diagnosis tool of the fit, in the sense that if the estimated lengthscale is lower than the minimum lengthscale able to fit with the actual number of basis functions and box size, indicates that the number of basis functions needs to be increased or the box size decreased.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
\caption{Relation between the minumum number of basis functions $J$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{\frac{\text{range}(x)}{2}}$), and the boundary condition factor $c$ ($c = \frac{L}{\frac{\text{range}(x)}{2}}$).}
  \label{fig3_lscale_vs_J_vs_c_part1}
\end{figure}

The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior. Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either the regular GP model fit and the approximate GP model fit. Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors distributions and the covariance functions of the approximate GP model and the regular GP model, obtained after fitting the data. Figure \ref{fig5_MSE_vs_J} shows the standarized root mean square error both models in function of the number of basis functions and the boundary factor. Figure \ref{fig6_lscale_vs_J} shows the evolution of the GP hyperparameters, lengthscale and magnitud, in function of the number of basis functions and boundary factor. Following Figure \ref{fig5_MSE_vs_J} it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5. The choice of 10 basis functions and a boundary factor of 1.5 could be accurate enough. This same conclusion can be roughly seen in the right-side plots in Figure \ref{image11_Posteriors_example-I}. This conclusions also agree with Figure \ref{fig3_lscale_vs_J_vs_c_part1} from which we can deduce that for a process with a normalized lengthscale of 0.3, 10 basis functions and a normalized box size (boundary factor) of 1.5 would be enough to get a good approximation following eq. (\ref{diff_covs}).

From Figure \ref{fig5_MSE_vs_J} it can also be drawn the behaviour in terms of performance of the approximate GP model in function of the number of basis functions and the boundary factor. Two main behaviours can be deduced from Figure \ref{fig5_MSE_vs_J}, which can also be easily seen from Figure \ref{fig3_lscale_vs_J_vs_c_part1}:
\\ 
$\bullet$ as the boundary factor increases, more basis functions are needed,
\\
$\bullet$ as fewer basis functions are used, the boundary factor must decrease.

Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and magnitud parameter as function of the number of basis functions and the boundary factor. The same conclusion drawn before can be drawn form this figure where with 15 basis functions and a boundary factor of 1.5 the estimated lengthscale of the approximate GP model approach the estimeted lengthscale of the regular GP model. From this figure it can be drawn that exists a minumum value for the boundary factor under this never a good approach will be achieved. This conclusion also appears in Figure \ref{fig3_lscale_vs_J_vs_c_part1} the contour lines of the boundary factor have an end at certain lengthscales, which means that exist a minimum boundary factor depending on the lengthscale of the process.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part1.pdf}}
\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
\caption{Posterior distributions and covariance functions of the proposed approximate GP models and the exact GP model in function of the number of basis functions $J$ and the boundary factor $c$.}
  \label{fig4_Post_&_Cov_part1}
\end{figure}

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
\caption{Standarized root mean square error (SRMSE) of the proposed approximate GP models and the exact GP model. (left) SRMSE against the number of basis functions $J$ for different values of the boundary factor $c$. (right) SRMSE against the boundary factor $c$ for different values of the number of basis functions $J$. }
  \label{fig5_MSE_vs_J}
\end{figure}

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig6_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
\caption{(left) Estimated lengthscales against the number of basis functions $J$ for different values of the boundary factor $c$. (right) Estimated GP magnitud the number of basis functions $J$ for different values of the boundary factor $c$. }
  \label{fig6_lscale_vs_J}
\end{figure}

\subsection{Comparative between true lengthscales and estimated lengthscales}

In this section we fit the regular GP model and the approximate GP model on different datasets drawn from GP prior models with a squared exponential covariance function with different lengthscales. The approximate GP model is fitted using different number of basis functions.

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{fig7_varing_lscale.pdf}
\caption{Posterior means of the exact and approximate GP models fitted over different datasets with different lengthscales.}
  \label{fig7_varing_lscale}
\end{figure}

Figure \ref{fig8_Tlscale_vs_Elscale} compares the true lengthscales to the estimated lengthscales for the exact GP and the approximate GP. The black line defines the minimun lengthscale that the approximate GP model can fit for each case.

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{fig8_Tlscale_vs_Elscale.pdf}
\caption{True lengthscales versus estimated lscales of the different realizations of datasets using different number of basis functions. }
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

\vspace{3mm}
\section{Univariate illustrative examples}\label{sec:gp_examples1D}

\subsection{Simulated data}

A simulated one-dimensional dataset, with 250 data points drawn from a Gaussian process prior with a Matern covariance function with 3/2 degrees of freedom. The hyperparameters of this GP prior with a Matern kernel, marginal variance and lengthscale are set to 1 and 0.15, respectively. The additive Gaussian noise to the GP model is set to 0.2. We compare the modelling performance over this simulated process of our proposed approach with the performance of a regular GP and a Splines-based model. For our basis function approach we use $J=80$ basis functions and a boundary factor $c=1.2$. For the Splines-based model we use 80 knots or basis in the model.

Figure \ref{fig9_Posteriors_exI} shows the posteriors distributions of the proposed model and the regular GP model, fitted over the simulated dataset. Posteriors distributions of the proposed approximation GP model, of the exact GP model, and the Splines model are plotted jointly with the true GP prior from the data were drawn. The right and left extremes of the simulated data correspond to out-of-sample data or test data which have not been taking part of training the models. Over this test data we can evaluate the predictive performance of every method.

\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{fig9_Posteriors_exI.pdf}
\caption{Posterior distributions of the proposed approximated GP model, the exact GP model, and the Splines model.}
  \label{fig9_Posteriors_exI}
\end{figure}

For assessing the performance of different methods we use the standardized mean squared error (SMSE) over the true function for interpolation and extrapolation data.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.350]{fig10_MSE_exI_inter.pdf}}
\subfigure{\includegraphics[scale=0.350]{fig10_MSE_exI_extra.pdf}}
\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
  \label{fig10_MSE_exI_inter}
\end{figure}

\subsection{Case I: Gay data}\label{sec:bf_caseII}
\subsection{Case II: Birthday data}\label{sec:bf_caseIII}

\vspace{3mm}
\section{Multivariate illustrative examples}\label{sec:gp_examplesMulti}
\subsection{Simulated data}\label{sec:bf_toyexampleMulti}
\subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
\subsection{Case IV: Leukemia data}\label{sec:bf_caseV}
\subsection{Case V: Land use spatio-temporal classification task}\label{sec:bf_caseVII}

\vspace{3mm}
\appendix

\section{Related work}

The GP prior entails an $O(n^3)$ complexity that is computationally intractable for many practical problems, and this problem especially becomes severe when we want to conduct inference using sampling methods. To overcome this scaling problem several schemes have been proposed. One approach is to partition the data set into separate groups \citep{snelson2007local, urtasun2008sparse} and performing local inference in each partition. Other global approach is to build a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables' \citep{quinonero2005unifying,bui2017unifying}. Other global approach make use of basis functions to approximate the covariance function. In \cite{snelson2007local} the authors conduct an approach that combines the idea of local and global approaches.

The literature contains many parametric models that approximate Gaussian process behaviours; for example \cite{bui2014tree} included  tree-structures in the approximation for extra scalability, and \cite{moore2015gaussian} combined local Gaussian
processes with Gaussian random fields.

\subsection{Inducing points methods}

The approach based on inducing points employs a small set of pseudo data points to summarise the actual data. The storage requirements are reduced to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$. Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods. From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modiﬁcations to the so-called Nystr\"om method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum. These methods are basically based on choosing a set of $m$ inducing inputs $x_u$ and scaling the corresponding eigendecomposition of their corresponding covariance matrix $K_{u,u}$ to match that of the actual covariance.

This scheme was originally introduced to the GP context by \cite{williams2001using}. As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model. However, several methods modifying the inducing point approach are widely used. The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix. The SoR method is based on a degenarete prior which produces unresoanble predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).

The Deterministic Training Conditional (DTC) method \citep{ro2001sparse,seeger2003fast}) retains the true covariance for the training data, but uses the approximate cross-covariances between training and test data, which reverse the problem of nonsensical predictive uncertainties. However, since the covariances for training and test cases are computed differently, this method results not to actually be a Gaussian process. This method was presented as Projected Latent Variables (PLV) in \cite{seeger2003fast} and Projected Process Approximation (PPA) in \cite{rasmussen2006gaussian}.

The Variational Approximation (VAR) \citep{titsias2009variational} suggests a variational approach which provides an objective function for optimizing the selection of inducing points. This basically modifies the DTC method by an additional trace term in the likelihood that comes from the variational bound.  \cite{hensman2013gaussian} extended this idea by introducing additional variational parameters to enable stochastic variational inference \citep{hoffman2013stochastic}, achieving a more computationally scalable bound which allows GPs to be fitted to millions of data.

The Fully Independent (Training) Conditional (FIC) \citep{quinonero2005unifying} method originally introduced as Sparse Pseudo-Input GP by \cite{snelson2006sparse} is also based on the Nystr\"om approximation, where they allow the pseudo-point input locations to be optimised by maximising the new model's marginal likelihood whose covariance is parameterized by the locations of an active set not constrained to be a subset of the training and test data.

More recently Bui et. al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time. The new framework is built on standard methods for approximate inference (variational-free-inference, EP and Power EP methods).

In practice, the inducing points-based sparse approximation methods works reasonable well in cases where the field is relatively smooth. \cite{vanhatalo2010approximate}
propose the use of compactly supported covariance function in conjunction with sparse approximations to model both short and long range correlations.

\cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI). This framework improves the scalability and accuracy of fast kernel approximations through kernel interpolation, and naturally combines the advantages of inducing point and structure exploiting for scalability (such as Kronecker \citep{saatcci2012scalable} or Toeplitz \citep{cunningham2008fast}) approaches.

The number of inducing points or their locations are crucial in order to capture the correlation structure. For a discussion on the effects of the inducing points, see \cite{vanhatalo2010approximate}. This behavior applies to all the methods from the Nystr\''om family.

This kind of 'projected process' approximation has also been discussed by e.g. \cite{banerjee2008gaussian}.

\subsection{Basis function methods}

The spectral analysis and series expansions of Gaussian processes has a long history. A classical result (see, e.g, \cite{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}, and references therein) is that the covariance function can be approximated with a finite truncation of Mercer series and the approximation is guaranteed to converge to the exact covariance function when the number of terms is increased.

Another related classical connection is to the works in the relationship of spline interpolation and Gaussian process priors \citep{wahba1978improper,kimeldorf1970correspondence,wahba1990spline}. In particular, it is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a specific choice of covariance function. The relationship of the spline regularization with Laplace operators then leads to series expansion representations that are closely related to the approximations considered here.

Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels. The approximate kernel has a finite basis function expansion.

The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse}, where the spectral representation of the covariance function is used. This model is a stationary sparse GP that can approximate any desired stationary full GP. However, as argued by the authors, this option does not converge to the full GP and can suffer from overfitting to the training data. \citep{gal2015improving} sought to improve the model by integrating out, rather than optimizing the frequencies. Gal and Turner derived a variational approximation that made use of a tractable integral over the frequency space. The result is an  algorithm that suffers less overfitting than the Sparse Spectrum GP, yet remains flexible.

While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.

Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat´ern class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure. They combine the variational methodology with Fourier based approximations.

In spatial statistics similar approaches are called low-rank models \citep{diggle2007springer}. The low rank models assume that the Gaussian field is a linear combination of $m$ basis functions. The type of an approximation depends on the basis functions used. Familiar examples include spectral representation \citep{diggle2007springer,paciorek2007computational,paciorek2007bayesian} and splines \citep{wood2003thin}.

Recent Splines models can reproduce the Matern family of covariance functions, however our approach can reproduce basically all of the stationary covariance functions.

\section{Contributions of the method}

This work is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models. This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods. This results in a reduced-rank approximation for the covariance function. This method has some nice features:

\vspace{2mm}
$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.

\vspace{2mm}
$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method. In addition, the computation of the automatic differentation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$. After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer. (Arno's paper, pag. 7)

\vspace{2mm}
$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions. especially when sampling methods for inference are used. GP prior as latent functions is needed in generalized models.

In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:

\vspace{2mm}
$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation. This operation has to be computed at every step of the HMC or optimizer.

\vspace{2mm}
$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional. It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitud parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

\section{Contributions of our work}

As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology. Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function. The performance and accuracy of the method are directly related with the number of basis functions and the box size. At the same time, successful values for these two factors depend on the smoothness or roughness of the process to be modeled. The time of computation is mainly dependent on the number of basis functions. Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.

\vspace{2mm}
$\bullet$ Firstly, clear summarized formulae of the method for the univariate and multivariate cases is presented.

\vspace{2mm}
$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the functions.

\vspace{2mm}
$\bullet$ We make recomendations for the values of these factors based on the recognized relations among them. We provide useful graphs of these relations that will help the users to improve performance and save time of computation.

\vspace{2mm}
$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.

\vspace{2mm}
$\bullet$ We describe the generalization of the method to the multidimensional case.

\vspace{2mm}
$\bullet$ We implement the approach in a fully probabilistic framework and for the Stan programming probabilistic software.

\vspace{2mm}
$\bullet$ We show several illustrative examples, simulate and real datasets, of the performance of the model, and accompanied by their Stan codes.

%\vspace{3mm}
%A summary of the formulae of the method for the univariate and multivariate cases is presented. The number of basis functions used in the approach and the value for the boundary condition of the model are two factors of main importance in practical applications. The number of basis functions is directly related to the accuracy of the approximation but also to the computation demands. Also exist a relation among the number of basis functions, the specified values for the boundary condition, and the performance of the approach.  Finally, the approach is applied in several illustrative study cases and accompanied by their Stan codes.

\section{Spectral densities of stationary covariance functions}

The covariance function of a stationary process, that is function of $\boldsymbol{\tau}=\mathbf{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).

\vspace{0.2cm}
\textit{(Bochner’s theorem) A complex-valued function $k$ on ${\rm I\!R}^D$ is the covariance function of a weakly stationary mean square continuous complex valued random process on ${\rm I\!R}^D$ if and only if it can be represented as}
%
\begin{equation}
k(\boldsymbol{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \mathbf{s} \cdot \mathbf{\tau}} d\mu(\boldsymbol{\tau}), \nonumber 
\end{equation}

\textit{where $\mu$ is a positive finite measure.}

\vspace{0.2cm}
If the measure $\mu$ has a density, it is known as the spectral density $S(\omega)$ of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the Wiener-Khintchine theorem. It gives the following relations:

\begin{eqnarray}
k(\boldsymbol{\tau})&=& \int S(\mathbf{s}) e^{2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber \\
%
S(\mathbf{s})&=& \int k(\boldsymbol{\tau}) e^{-2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber
\end{eqnarray}

\section{Approximate the covariance function using Hilbert space methods}

Associated to each covariance function $k(\mathbf{x},\mathbf{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
%
\begin{equation}
\mathcal{K} f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}'.
\end{equation}

Assuming that the espectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
%
\begin{equation}
S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
\end{equation}

If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the covariance function $k$,

\begin{equation}
-\nabla^2 f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}',
\end{equation}

\noindent then the covariance function can be represented as

\begin{equation}
k(\mathbf{x},\mathbf{x}')= \sum_j \lambda_j \phi_j(\mathbf{x}) \phi_j(\mathbf{x}'),
\end{equation}

\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator. Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
%
\begin{eqnarray}
-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
\end{eqnarray}

a series expansion of eigenvalues and eigenfunctions

\section{Example of generalization to the multivariate case}

Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension. The number of new multidimensional eigenfunctions $\phi^{\ast}_j$ and eigenvalues $\lambda^{\ast}_j$ is $J^D=3^2=9$ ($j=\{1,\cdots,J^D\}$). The matrix $\mathbb{S}\in {\rm I\!R}^{9 \times 2}$ is
%
\begin{eqnarray}
\mathbb{S}=
\left[ {\begin{array}{cc}
1 & 1 \nonumber \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3
\end{array} } \right]
\end{eqnarray}

\noindent and the multidimensional eigenfunctions and eigenvalues
%
\begin{eqnarray}
\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} \nonumber \\
%
\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber
\end{eqnarray}

Now, we show another example where different number of eigenfunctions and eigenvalues are used for every dimension. We consider a three-dimensional ($D=3$) input space, and sets of $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively. The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$. The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
%
\begin{eqnarray}
\mathbb{S}=
\left[ {\begin{array}{ccc}
1 & 1 & 1 \nonumber \\
1 & 1 & 2 \\
1 & 1 & 3 \\
1 & 2 & 1 \\
1 & 2 & 2 \\
1 & 2 & 3 \\
2 & 1 & 1 \\
2 & 1 & 2 \\
2 & 1 & 3 \\
2 & 2 & 1 \\
2 & 2 & 2 \\
2 & 2 & 3 
\end{array} } \right]
\end{eqnarray}

\noindent and the multidimensional eigenfunctions and eigenvalues
%
\begin{eqnarray}
\phi^{\ast}_1= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_1= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_2= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_2= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_3= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_3= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_4= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_4= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_5= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_5= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_6= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_6= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_7= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_7= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_8= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_8= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_9= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_9= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_{10}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{10}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_{11}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{11}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_{12}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{12}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber
\end{eqnarray}

\vspace{2mm}
\section*{Acknowledgment}

\bibliographystyle{biom}
\bibliography{references}

\end{document}

###########
#######
Next we show a little example of how to construct the new set of eigenfunctions and eigenvalues with input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ in a two-dimensional input space ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension. As a result we obtain sets of $J^D=3^2=9$ new eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$. The eigenfunctions $\phi^{\ast}$ are the product combination of the unidimensional eigenfunctions $\phi$, and eigenvalues $\lambda^{\ast}$ is the $D$-vector of unidimensional eigenvalues $\lambda$:

\begin{eqnarray}
\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber
\end{eqnarray}

### 2019-02-23 Paper/manuscript/manuscript_05.tex
little change. number of equation
number of equation
more contents
small grammar changes 
adding contents in section 3.2

@@
 \begin{eqnarray}\label{eigenproblem}
--\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
+\begin{split}
+-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \\ 
 \phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
+\end{split}
 \end{eqnarray}
@@
 %
-\begin{eqnarray}
-\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
-\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
+\begin{eqnarray}\label{bf&lambda}
+\begin{split}
+\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \\
+\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right).
+\end{split}
 \end{eqnarray}
@@
 We are going to generalize the results from the previous section to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions.
-We assume the same number $J$ of basis functions terms in the series expansion for every input dimension, however it is straight forward to generalize this to different number of term series $J_d$ for every dimension.
 
 \vspace{0.2cm}
-In a $D$-dimensional input space the total number of eigenfunctions and eigenvalues in the approximation corresponds to the total number of $D$-tuples over $J$, which is $J^D$.
-Let $\mathbb{S}\in {\rm I\!R}^{J^D \times D}$ be the set of $D$-tuples elements.
+In a $D$-dimensional input space the total number of eigenfunctions and eigenvalues in the approximation corresponds to the total number of $D$-tuples over $J$, which is $\prod_{d=1}^{D} J_d$, where $J_d$ is the number of basis function for the dimension $d$.
+Let $\mathbb{S}\in {\rm I\!R}^{\prod_{d=1}^{D} J_d \times D}$ be the set of $D$-tuples elements.
 Each new eigenfunction $\phi^{\ast}_j$ correspond to the product of the univariate eigenfunctions whose indices corresponds to the the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
@@
 %
-\begin{eqnarray}
-\phi^{\ast}_j(\mathbf{x}) &=& \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right)  \nonumber \\
-%
-\lambda^{\ast}_j &=& \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  \nonumber
-\end{eqnarray}
-
-As an example we show the matrix $\mathbb{S}$ for a $three$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2,\mathbf{x}_3\}$ ($D=3$) and $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively.
+\begin{eqnarray} \label{bf&lambda_multi}
+\begin{split}
+\phi^{\ast}_j(\mathbf{x}) &=& \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right) \\
+%
+\lambda^{\ast}_j &=& \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  
+\end{split}
+\end{eqnarray}
+
+As an example we show the matrix $\mathbb{S}$ for a $three$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2,\mathbf{x}_3\}$ ($D=3$) and $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimension, respectively.
 The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$.
@@
 %
-\begin{equation}\label{approxcov}
+\begin{equation}\label{diffcov}
 k(\tau) \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
@@
 
-\noindent where $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$. The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are 
-%
-\begin{eqnarray}
-\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
-\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
-\end{eqnarray}
+\noindent where $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) in a unidimensional case and those in equation (\ref{bf&lambda_multi}) in a multidimensional case.
 
 The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity $J=1,\dots,j,\dots,\infty$.
-The number of basis functions can be truncated that the difference between the covariance function and the approximation be less than a threshold $\epsilon$,
+The number of basis functions can be truncated that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
 %
@@
 The finite number of basis functions in the approximation to satisfy eq.
-(\ref{diff_covs}) will depend on the smoothness or roughness of the function, that is, on the lengthscale $\ell$.
+(\ref{diff_covs}) will depend on the smoothness or roughness of the function, that is, on the lengthscale $\ell$ of the GP process.
 The approximation will also depends on the box size $L$ which will affect the performance specially near the boundaries.
@@
 
-For a fixed $L$, let's suppose it is set to a right choice, Figure \ref{image2_J} shows the effect of the number of basis functions in approaching the posterior mean and the covariance function.
+For a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of the number of basis functions in approaching the posterior mean and the covariance function.
 The number of basis functions affects the non-linearity or roughness (or smoothness) of the posteriors.
@@
 (\ref{diff_covs}), for a certain GP process with lengthscale $\ell$ and given a fixed boundary factor $c$.
-Alternatively, this figure could also be read as the maximum boundary factor $c$ that we should use given a number of basis functions, for certain GP process with certain lengthscale.
+Alternatively, this figure could also be read as the minimum boundary factor $c$ that we should use given a number of basis functions, for certain GP process with certain lengthscale.
 And also this figure could also be read as the maximum roughness in a process that can be fitted properly given a number of basis functions and a box size.
-Finally, this figure also gives us the minimum value for the boundary factor $c$ which is suitable for certain roughness or number of basis functions.
 This figure basically collects the behaviour of the approximate GP model in function of the lengthscale, number of basis functions and boundary condition:
@@
 The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
-Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either the regular GP model fit and the approximate GP model fit.
+Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
 Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors distributions and the covariance functions of the approximate GP model and the regular GP model, obtained after fitting the data.
-Figure \ref{fig5_MSE_vs_J} shows the standarized root mean square error both models in function of the number of basis functions and the boundary factor.
+Figure \ref{fig5_MSE_vs_J} shows the standarized root mean square error of both models in function of the number of basis functions and the boundary factor.
 Figure \ref{fig6_lscale_vs_J} shows the evolution of the GP hyperparameters, lengthscale and magnitud, in function of the number of basis functions and boundary factor.
@@
 The choice of 10 basis functions and a boundary factor of 1.5 could be accurate enough.
-This same conclusion can be roughly seen in the right-side plots in Figure \ref{image11_Posteriors_example-I}.
+This same conclusion can be roughly seen in the right-side plots in Figure \ref{fig9_Posteriors_exI}.
 This conclusions also agree with Figure \ref{fig3_lscale_vs_J_vs_c_part1} from which we can deduce that for a process with a normalized lengthscale of 0.3, 10 basis functions and a normalized box size (boundary factor) of 1.5 would be enough to get a good approximation following eq.
@@
 Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and magnitud parameter as function of the number of basis functions and the boundary factor.
-The same conclusion drawn before can be drawn form this figure where with 15 basis functions and a boundary factor of 1.5 the estimated lengthscale of the approximate GP model approach the estimeted lengthscale of the regular GP model.
-From this figure it can be drawn that exists a minumum value for the boundary factor under this never a good approach will be achieved.
+The same conclusion drawn before can be drawn form this figure where with 15 basis functions and a boundary factor of 1.5 the estimated lengthscale of the approximate GP model approach the estimated lengthscale of the regular GP model.
+From this figure, it can be drawn that exists a minumum value for the boundary factor under this a good approach will never be achieved.
 This conclusion also appears in Figure \ref{fig3_lscale_vs_J_vs_c_part1} the contour lines of the boundary factor have an end at certain lengthscales, which means that exist a minimum boundary factor depending on the lengthscale of the process.
@@
 
-In this section we fit the regular GP model and the approximate GP model on different datasets drawn from GP prior models with a squared exponential covariance function with different lengthscales.
+In this section we fit the regular GP model and the approximate GP model on different datasets drawn from a GP prior model with a squared exponential covariance function with different lengthscales for every dataset.
 The approximate GP model is fitted using different number of basis functions.
+Figure \ref{fig7_varing_lscale} shows the posterior means of the exact GP model and the approximate GP model with different number of basis functions for every dataset with different wiggly effects (lengthscale).
 
@@
 
-Figure \ref{fig8_Tlscale_vs_Elscale} compares the true lengthscales to the estimated lengthscales for the exact GP and the approximate GP.
-The black line defines the minimun lengthscale that the approximate GP model can fit for each case.
+Figure \ref{fig8_Tlscale_vs_Elscale} compares the lengthscales of the true generating process to the estimated lengthscales for the different datasets.
+Different comparison are shown using different number of basis functions in the approximate GP model.
+Let's assume that the box size parameter was properly set in every case.
+The dashed black line represents the minimum lengthscales that can be properly fitted by the approximate GP model according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+Figure \ref{fig9_MSE_varing_lscale} shows the SMSE of the approximate GP model for the different datasets with different wiggly effects and the number of basis functions used in the approximation.
+
+From Figure \ref{fig8_Tlscale_vs_Elscale} we can recognize when the lengthscales between the regular GP model and the approximate GP model match each other, which should be an indicator of a good approximation.
+Figure \ref{fig9_MSE_varing_lscale} indicates when the MSE is near to zero which also indicates a good approximation.
+We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for very low lengthscales and very long lengthscales.
+Figure \ref{fig8_Tlscale_vs_Elscale} shows a correpondence between very low lengthscales when the SMSE is large, conversly there is no correspondence between very large lengthscale and the SMSE is close to zero.
+These effects can make sense because, in the case of long lengthscales these have larger variabilities obtaining  similar posteriors, and in the case of low lengthscale these have lower variabilities obtaining diferent posteriors.
+
+The dashed black line represents the minimum lengthscales that can be properly fitted by the approximate GP model according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{fig9_MSE_varing_lscale.pdf}
+\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and the number of basis functions.}
+  \label{fig9_MSE_varing_lscale}
+\end{figure}
+
 \vspace{3mm}
-\section{Univariate illustrative examples}\label{sec:gp_examples1D}
+\section{Univariate examples}\label{sec:gp_examples1D}
 
@@
 A simulated one-dimensional dataset, with 250 data points drawn from a Gaussian process prior with a Matern covariance function with 3/2 degrees of freedom.
-The hyperparameters of this GP prior with a Matern kernel, marginal variance and lengthscale are set to 1 and 0.15, respectively.
-The additive Gaussian noise to the GP model is set to 0.2.
+The hyperparameters of this GP prior with a Matern kernel, marginal variance and lengthscale, are set to 1 and 0.15, respectively.
+The additive Gaussian noise to these drawn points is set to 0.2.
 We compare the modelling performance over this simulated process of our proposed approach with the performance of a regular GP and a Splines-based model.
@@
 
-Figure \ref{fig9_Posteriors_exI} shows the posteriors distributions of the proposed model and the regular GP model, fitted over the simulated dataset.
-Posteriors distributions of the proposed approximation GP model, of the exact GP model, and the Splines model are plotted jointly with the true GP prior from the data were drawn.
-The right and left extremes of the simulated data correspond to out-of-sample data or test data which have not been taking part of training the models.
-Over this test data we can evaluate the predictive performance of every method.
+Figure \ref{fig9_Posteriors_exI} shows the posteriors distributions of the proposed model and the regular GP model, fitted over the data.
+Posteriors distributions of the proposed approximate GP model, of the exact GP model, and the Splines model are plotted jointly with the true GP prior from the data were drawn.
+The right and left extremes of the data, as well as three small regions around the middle of the data, correspond to out-of-sample or test data which have not been taking part of training the models.
 
@@
 \centering
-\includegraphics[scale=0.35]{fig9_Posteriors_exI.pdf}
+\includegraphics[scale=0.35]{fig10_Posteriors_exI.pdf}
 \caption{Posterior distributions of the proposed approximated GP model, the exact GP model, and the Splines model.}
-  \label{fig9_Posteriors_exI}
+  \label{fig10_Posteriors_exI}
 \end{figure}
 
-For assessing the performance of different methods we use the standardized mean squared error (SMSE) over the true function for interpolation and extrapolation data.
+Figure \ref{fig11_MSE_exI_inter} shows the  standardized mean squared error (SMSE) over the true function for interpolation and extrapolation data.
+As it can be seen the splines model does not extrapolate data properly.
+For interpolation both models show similar performance although the approximate GP model seems to be a bit better.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.350]{fig10_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.350]{fig10_MSE_exI_extra.pdf}}
+\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
 \caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
-  \label{fig10_MSE_exI_inter}
+  \label{fig11_MSE_exI_inter}
 \end{figure}
@@
 \vspace{3mm}
-\section{Multivariate illustrative examples}\label{sec:gp_examplesMulti}
+\section{Multivariate examples}\label{sec:gp_examplesMulti}
 \subsection{Simulated data}\label{sec:bf_toyexampleMulti}

### 2019-02-25 Paper/manuscript/manuscript_05.tex
Update section 2

@@
 \usepackage{soul}
-
+\usepackage{xcolor}
 \graphicspath{{./images/}}
@@
 
-\subsection{Gaussian process model}
+\subsection{Gaussian process regression}
+
+A Gaussian process (GP) is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, i.e.
+$\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
+Gaussian processes have the defining property that the marginal distribution of every finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
+\\
+\\
+In this work, Gaussian processes will take the role of prior distributions over function spaces for non-parametric regression in a Bayesian setting. 
+Consider a data set $\mathcal{D} = \left\lbrace \mathbf{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is a noisy observation of an unknown function $f$ at $\mathbf{x}_n \in \mathbb{R}^D$ , i.e.
+$y_n = f(\mathbf{x}_n) + \epsilon_n$.
+Our goal is to predict the value of the function $f^*=f(\mathbf{x}^*)$ evaluated at a new input point $\mathbf{x}^* \in \mathbb{R}^D$.
+That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.
+\\
+\\
+We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R} \rightarrow \mathbb{R}$ and $k: \mathbb{R} \times \mathbb{R} \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
+\begin{align}
+ 	\mu(\mathbf{x}) &= \mathbb{E}\!\left[f(\mathbf{x})\right],\\ 
+ 	k(\mathbf{x}, \mathbf{x}') &= \mathbb{E}\!\left[\left( f(\mathbf{x}) - \mu(\mathbf{x}) \right)\left( f(\mathbf{x}') - \mu(\mathbf{x}') \right)\right].
+ \end{align} 
+The mean and covariance functions completely characterize the distribution of $f$. 
+\\
+\\
+{\color{red} $\bullet$ Introduce the joint model}
+\begin{align}
+	p(\mathbf{y}, \mathbf{f}, f^*) = p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)
+\end{align}
+{\color{red} $\bullet$  argue that marginalizing and conditioning yields the predictive distribution of interest}
+\begin{align}
+	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}, \mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})}
+\end{align}
+{\color{red} $\bullet$ Give equations for predictive mean and variance for $f^*$ assuming Gaussian likelihood}
+\\
+\\
+{\color{red} $\bullet$ Briefly standard computational challenges with GPs: cubic scaling and only closed form solution for Gaussian likelihoods.
+This should motivate the use of basis function expansions in a probabilistic programming setting such that with get both speed and modelling flexibility}
+\\
+\\
 

### 2019-02-25 Paper/manuscript/manuscript_05.tex
improve introduction

--- 2019-02-22 Paper/manuscript/manuscript_05.tex
@@
 
-Gaussian processes (GPs) are natural and flexible non-parametric models to define probability distributions over multi-dimensional functions \citep{rasmussen2006gaussian,neal1997monte}.
-This enables to apply probabilistic inference directly to functions.
-The defining feature of a GP model is that the joint distribution of function values is a multivariate normal distribution which is characterized by a mean function and a covariance function.
-The covariance function encodes our prior assumptions about the function, such as continuity, smoothness, periodicity and scale properties of the GP, and defines a correlation structure which characterize the correlations between function values at different inputs.
-A GP model with a covariance function allows for non-linear effects on the function values and can handle interactions between input variables implicitly.
-Several types of covariance functions with different behaviors can be used, from stationary (e.g.
-exponentiated quadratic, Matern, periodic, quasyperiodic...) to non-stationary (e.g.
-dot product, neuronal network,...).
-Combinations, sum and product, of different covariance function can also be used in order to obtain more modelling flexibility on real complex problems.
-GP models are often referred to as non-parametric because the number  of parameters is not fixed, but rather spanned by the number of data points, which allows to adapt model complexity to the data.
-GP models are powerful non-parametric prior models for regression and classification \citep{rasmussen2006gaussian}, and they can also be used as non-linear components in additive models.
-GP models is of broad interest across machine learning and statistics, with applications in the fields of spatial epidemiology \citep{diggle2013statistical,carlin2014hierarchical}, robotics and control \citep{deisenroth2015gaussian}, signal processing \citep{sarkka2013spatiotemporal}, Bayesian optimization and probabilistic numerics, \citep{roberts2010bayesian,briol2015probabilistic,hennig2015probabilistic}, and many others.
-
-One of the main limitations of GPs is that computational demands and memory requirements scale as $O(n^3)$ and $O(n^2)$ respectively, effectively limiting a direct implementation to a number of training samples $n$ at most a few thousand cases.
-This problem becomes especially important when using sampling methods for full Bayesian inference, when in each sampling step we need to invert the Gram matrix, usually through Cholesky factorization.
-
-In this paper we want to make full Bayesian inference over Gaussian process models accessible to a broader audience by making them fast and easy to apply.
-We implement novel method recently  developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models.
-This method decomposes the GP kernel into basis functions which represent the GP as a linear model and makes it faster to fit.
-The set of Laplace eigenfunctions considered in this paper forms such a basis and we can compute them analytically.
-Furthermore, they are independent of the choice of kernel (including the hyperparameters), which means that we do not need to autodiff through the kernel and the Cholesky.
-In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions, since it is known that there exist a correlation between the latent values and the hyperparameters which makes inference more difficult especially when the dimension of latent values is high.
-
-There already exist several approximate methods that alleviate these computational demands.
-Sparse GPs, which are approximate methods to fit GP models, are based on building a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables'.
-They do reduce the dimension of the posterior, reducing the storage requirements to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$.
-However it still requires Cholesky factorizations.
-Furthermore convergence to the full GP model is not straight forward, and reducing the amount of data while maintaining the posterior densities is not a simple procedure.
-Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods.
-More recently Bui et.
-al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time.
-\cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI).
-
+Gaussian processes (GPs) are natural (Paul: what does "natural" imply and as compared to what?) and flexible non-parametric models to define probability distributions over multi-dimensional functions \citep{rasmussen2006gaussian,neal1997monte}.
+This allows applying probabilistic inference to estimate functional relationships between one or more covariates and a single response variable.
+The defining feature of GPs is that the function values are assumed to jointly follow a multivariate normal distribution, which is characterized primarily by a covariance function, sometimes also referred to as covariance kernel.
+The covariance function encodes our prior assumptions about the functional relationship, such as continuity, smoothness, periodicity and scale properties.
+GPs not only allow for non-linear effects but can also implicitly handle interactions between covariates.
+Several types of covariance functions with varying properties can be used, and they may even be combined for further increased flexibility.
+GPs are often referred to as non-parametric models because the number of parameters is not fixed, but rather increases with the number of data points, which allows to adapt model complexity to the data.
+As such, the term 'non-parametric' does not imply having no parameters but actually having a lot of them.
+Due to their generality and flexibility, GPs are of broad interest across machine learning and statistics \citep{rasmussen2006gaussian,neal1997monte}.
+Among others, they find application in the fields of spatial epidemiology \citep{diggle2013statistical,carlin2014hierarchical}, robotics and control \citep{deisenroth2015gaussian}, signal processing \citep{sarkka2013spatiotemporal}, as well as Bayesian optimization and probabilistic numerics \citep{roberts2010bayesian,briol2015probabilistic,hennig2015probabilistic}.
+
+One of the main limitations of exact GPs is that computational demands and memory requirements scale as $O(n^3)$ and $O(n^2)$, respectively, with the number $n$ of observations in the data used to fit the model.
+This effectively limits their application to rather small data sets of a few thousand observations at most.
+This problem becomes especially severe when performing full Bayesian inference via sampling methods, where in each sampling step we need to invert the Gram matrix of the covariance function, usually through Cholesky factorization.
+To alleviate these computational demands, several approximate methods have been proposed, which we will briefly review in the following.
+
+Sparse GPs are based on building low-rank approximations of the covariance matrix of the complete data by means of 'inducing variables' (Paul: if you introduce this term, you may need to explain it.
+Alternatively, can we just get rid of it?).
+They reduce the dimension of the posterior distribution thereby reducing the memory requirements to $O(nm)$ and computational complexity to $O(nm^2)$ for some $m < n$.
+This approach still requires Cholesky factorization although now based on a smaller covariance matrix.
+While computational complexity is reduced considerably, ensuring convergence to the exact GP model is not necessarily straightforward (Paul: this may need a citation).
+As such, reducing the amount of data while maintaining the correct posterior distribution is not a simple procedure.
+A unifying view on sparse GPs based on approximate generative methods is provided in \cite{quinonero2005unifying}, while a general review can be found in \cite{rasmussen2006gaussian}.
+More recent developments in the context of sparse GPs include points-based sparse approximation methods \citep{wilson2015kernel}, which have been further developed by Bui et.
+al (2017) (Paul: cite this study correctly in latex) to perform approximations at inference time rather than at modeling time.
+
+(Paul: This whole paragraph is unclear to me.
+Are all of these approaches approximations via splines, or are the splines just special cases of a more general method?
+In any case, I think the paragraph needs substantial revision).
 Other global approach make use of the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
@@
 
-The approach we propose in this paper is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models.
-This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods.
-This results in a reduced-rank approximation for the covariance function.
-
-While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
+Yet another approach to approximate GPs -- on which we will focus in the present paper -- was proposed by \cite{solin2018hilbert} who suggested to use reduced-rank approximations of the covariance kernel.
+This method decomposes the kernel into basis functions which represent the GP as a linear model and thus makes its estimation considerably faster.
+In principal, we are free in the choice of basis functions, but the Laplace eigenfunctions were found to be particularily appealing \citep{solin2018hilbert}.
+Not only can they be computed analytically, but they are also independent of the particular choice of the covariance kernel including the hyperparameters.
+The latter comes with several advantages when performing inference via algorithms that require differentiation of the posterior density (Paul: cite something here).
+In a fully Bayesian framework based on sampling methods, the proposed approximate GP model has a computational complexity of just $O(nm+m)$, with $m<<n$ in every sampling step.
+The posterior distribution of the reduced-rank GP approximation is of dimension $m$ and thus considerably smaller than in the case of exact GPs.
+This not only lowers the memory requirements but also reduces the correlations between latent values and the corresponding hyperparameters of the GP, which can help to further improve sampling efficiency.
+
+While \cite{solin2018hilbert} fully developed the mathematical theory behind reduced-rank approximations of GPs, they did not put much effort in analyzing the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space (Paul: I replace 'box size' because, at that point, readers will likely have no idea what that means), or properties of the true functional relationship between covariates and response variable.
+However, for practical implementations of approximate GPs, the latter factors are critical and thus require further investigation.
+In the present paper, we will fill this gap by providing practical recommendations for the choice of these factors based on the recognized relations among them.
+What is more, we provide intuitive visualizations of these relations that will help users to improve performance and save computation time, while at the same time maintaining close approximations of exact GPs.
+
+Although there are several GP specific software packages available to date (Paul: cite those), each provide efficient implementations only for a restricted range of GP based models.
+In this paper we do not focus on the fastest possible inference for a specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks such as Stan (cite).
+To this end, we want to make full Bayesian inference over Gaussian process models accessible to a broader audience by making them easy, fast and save to apply.
+
+The remainder of the paper is structered as follows.
+In Section 2, we introduce GPs and their reduced rank approximations proposed by \cite{solin2018hilbert}.
+In Section 3, we analyze the accuracy of these approximations under several conditions using analytical and numerical methods.
+Several case studies in which we fit
+exact and approximate GPs to real and simulated data are provided in Section 4. We end with a discussion in Section 5.
+
+(Paul: The content and the purpose of the following sentences is not fully clear to me.
+Do we need them in the Introduction or can they be removed, or put into the next section?) While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
 Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.
-
-There are several GP specific software, which have efficient
-computation for restricted set of GP based models allowing scaling to
-bigger data sets. In this paper we do not focus on the fastest
-possible inference for a specific restricted models, but instead we
-are interested in how GPs can be easily used as modular components in
-probabilistic programming frameworks.
-
-As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology.
-Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function.
-The performance and accuracy of the method are directly related with these factors.
-The time of computation is mainly dependent on the number of basis functions.
-Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.
-We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the functions.
-We make recomendations for the values of these factors based on the recognized relations among them.
-We provide useful graphs of these relations that will help the users to improve performance and save time of computation.
-We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
-
-The rest of the paper is structered as follows.
-Section 2 describe  the proposed model based on the Hilbert space methods for low-rank approximation to GP.
-In Section 3 we analyze the accuracy of the method and its dependency on existing relations among the number of basis functions, the box size factor and the roughness or smoothness of the process.
 
@@
 \begin{eqnarray}
-\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
-\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
+\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \\
+\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right).
 \end{eqnarray}

### 2019-02-26 Paper/manuscript/manuscript_05.tex
update GP section following Michael comments

--- 2019-02-25 Paper/manuscript/manuscript_05.tex
@@
 \\
-We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R} \rightarrow \mathbb{R}$ and $k: \mathbb{R} \times \mathbb{R} \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
-\begin{align}
+We assume a Gaussian process prior for the collection of function values $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N $, $\mathbf{f} \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
+%
+\begin{align*}
  	\mu(\mathbf{x}) &= \mathbb{E}\!\left[f(\mathbf{x})\right],\\ 
  	k(\mathbf{x}, \mathbf{x}') &= \mathbb{E}\!\left[\left( f(\mathbf{x}) - \mu(\mathbf{x}) \right)\left( f(\mathbf{x}') - \mu(\mathbf{x}') \right)\right].
- \end{align} 
-The mean and covariance functions completely characterize the distribution of $f$. 
+\end{align*}
+
+The mean and covariance functions represent how the random functions behave on average and how the different points in the input space co-vary with respect to each other.
+The mean and covariance functions completely characterize the multivariate Gaussian for $\mathbf{f}$,
+\begin{align*}
+\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu},K),
+\end{align*}
+
+\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ is given by the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
+%
+\begin{eqnarray*}
+p(\mathbf{f},\mathbf{f}')=\mathcal{N} \left( \left[ \begin{array}{cc}
+\mathbf{f} \\ 
+\mathbf{f}'
+\end{array} \right] | 0,\left[ \begin{array}{cc}
+K_{\mathbf{f},\mathbf{f}} & K_{\mathbf{f},f^*} \\ 
+K_{f^*,\mathbf{f}} & K_{f^*,f^*}
+\end{array} \right] \right),
+\end{eqnarray*}
+
+\noindent where $K_{\mathbf{f},f^*}$ defines the covariances between $\mathbf{f}$ and $f^*$, and $K_{f^*,f^*}$ defines the variance of $f^*$.
+By using the conditioning properties of a multivariate Gaussian, we can compute analytically the predictive distribution for $f^*$  given $\mathbf{f}$,
+\begin{align*}
+	p(f^*| \mathbf{f}) = \mathcal{N}(f^*| K_{f^*,\mathbf{f}} K_{\mathbf{f},\mathbf{f}}^{-1} \mathbf{f},  K_{f^*,f^*}-K_{f^*,\mathbf{f}} K_{\mathbf{f},\mathbf{f}}^{-1} K_{\mathbf{f},f^*}  )
+\end{align*}
+
+The joint distribution of observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
+\begin{align*}
+	p(\mathbf{y}, \mathbf{f}, f^*) = p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)
+\end{align*}
+
+Integrating the joint distribution  $p(\mathbf{y}, \mathbf{f}, f^*)$ over the uncertainty of $\mathbf{f}$ and then conditioning on the observations $\mathbf{y}$ following the Bayes's rule yields the posterior predictive distribution for $f^*$ given the observations $\mathbf{y}$,
+%
+\begin{align*}
+	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})}
+\end{align*}
+
+If the observational model $p(\mathbf{y}|\mathbf{f})$ is Gaussian, the integrals over $\mathbf{f}$ can be computed analytically yielding,
+\begin{align*}
+	p(f^*| \mathbf{y}) = \mathcal{N}(f^*| K_{f^*,\mathbf{f}} (K_{\mathbf{f},\mathbf{f}} + \sigma^2I)^{-1} \mathbf{f},  K_{f^*,f^*}-K_{f^*,\mathbf{f}} (K_{\mathbf{f},\mathbf{f}} + \sigma^2I)^{-1} K_{\mathbf{f},f^*}  )
+\end{align*}
+
+In most modelling problems, however, the observational model is not Gaussian (e.g.
+binomial or multinomial classification, Poisson  regression, and many other modelling problems) and the posterior predictive distribution $p(f^*| \mathbf{y})$ is analytically untractable and can not be obtained in a closed-form expression.
+Approximate inference have to be used to tackle the intractable inference, tipically based on sampling methods such as MCMC \citep{brooks_2011} or analytical Gaussian approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational methods \citep{gibbs2000variational,csato2000efficient}.
+Furthermore, Gaussian processes in a direct implementation scales cubically in the number of sample points, $N$, which effectively limits the modelling inference for medium and large datasets.
+In the introduction section in this paper several methods tipically used for dealing with the computational complexity has been commented.
 \\
 \\
-{\color{red} $\bullet$ Introduce the joint model}
-\begin{align}
-	p(\mathbf{y}, \mathbf{f}, f^*) = p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)
-\end{align}
-{\color{red} $\bullet$  argue that marginalizing and conditioning yields the predictive distribution of interest}
-\begin{align}
-	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}, \mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})}
-\end{align}
-{\color{red} $\bullet$ Give equations for predictive mean and variance for $f^*$ assuming Gaussian likelihood}
+In this paper we proposed an approximate model for Gaussian processes based on basis functions expansion which basically turns the GP prior model into a linear model.
+This linear GP model representation makes inference considerably faster and can be used as latent function in non-Gaussian observational models allowing modelling flexibility.
 \\
 \\
-{\color{red} $\bullet$ Briefly standard computational challenges with GPs: cubic scaling and only closed form solution for Gaussian likelihoods.
-This should motivate the use of basis function expansions in a probabilistic programming setting such that with get both speed and modelling flexibility}
-\\
-\\
-
-A GP model can be seen as a probability model over an infinite collection of random varibales.
-However, formally, the defining feature of a GP model is that the joint distribution of function values at any finite number of input values $\mathbf{x}$ is a multivariate normal distribution,
-%
-\begin{equation}
-f(\mathbf{x}) \sim \mathcal{N}(\mathbf{m},\mathbf{K})
-\end{equation}
-
-\noindent with $\mathbf{m}$ the mean vector and $\mathbf{K}$ the covariance matrix.
-The element $\mathbf{m}_i$ is given by the mean function $m(\mathbf{x}_i)$ and the element $\mathbf{K}_{ij}$ is given by the covariance function $k(x_i,x_j|\theta)$ which is a function of the inputs values $\mathbf{x}$ and the parameters $\theta$.
-The mean and covariance functions represent how the random functions behave on average and hoe the different points in the input space co-vary with respect to each other:
-
- In general the mean function $m(\mathbf{x})$ can be any function of the inputs, although this is usually set to zero for simplifying notation.
-
-\vspace{0.5cm}
-\noindent ... \\
-... \\
-...
-
-\vspace{0.5cm}
-A stationary covariance function is a function of $x-x'$, so we can write $k(x,x') = k(x-x')$.
-Isotropic covariance functions are those that are function of the distance between observations, $k(x,x') = k(|x-x'|) = k(r)$.
+The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
+A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, so we can write $k(\mathbf{x},\mathbf{x}') = k(\mathbf{x}-\mathbf{x}')$, which means that the covariance is invariant to translations.
+Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariacne is both translation and rotation invariant.
 The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which are
@@
 \begin{eqnarray}
-k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r/\ell^2), \nonumber \\
-k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
-k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
-k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
+k_{\nu=\infty}(\mathbf{r})&=&\sigma^2 \text{exp}(-\frac{1}{2} \mathbf{r}/\ell^2), \nonumber \\
+k_{\nu=\frac{1}{2}}(\mathbf{r})&=&\sigma^2 exp(-\mathbf{r}/\ell), \nonumber \\
+k_{\nu=\frac{3}{2}}(\mathbf{r})&=&\sigma^2(1+\sqrt{3}\mathbf{r}/\ell \text{exp}(-\sqrt{3}\mathbf{r}/\ell), \nonumber \\
+k_{\nu=\frac{5}{2}}(\mathbf{r})&=&\sigma^2(1+\sqrt{5}\mathbf{r}/\ell+\frac{5}{3}\mathbf{r}^2/\ell^2) \text{exp}(-\sqrt{5}\mathbf{r}/\ell), \nonumber 
 \end{eqnarray}
@@
 The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function.
-These covariance functions can be easily generalized to the multidimensional case considering multidimensional inputs $\mathbf{r}$ and the multidimensional length-scale $\boldsymbol{\ell}$.
-
-The spectral density functions associated with the univariate Mat\'ern covariance functions written above are
-%
-\begin{eqnarray}
-S_{\nu=\infty}(w)&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 w^2), \nonumber \\
-S_{\nu=\frac{1}{2}}(w)&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \nonumber \\
-S_{\nu=\frac{3}{2}}(w)&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \nonumber \\
-S_{\nu=\frac{5}{2}}(w)&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3}, \nonumber 
-\end{eqnarray}
-
-\noindent where variable $w$ is in the frequency domain.
+These covariance functions can be easily generalized to the use of multidimensional length-scale $\boldsymbol{\ell}$, which basically turns the isotropic covariance function into non-isotropic.
+
+Stationary covariance functions can be represented in terms of their spectral densities \citep{Rasmussen_2006}.
+The spectral density functions associated with the Mat\'ern covariance functions written above are
+%
+\begin{eqnarray}
+S_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 \mathbf{w}^2), \nonumber \\
+S_{\nu=\frac{1}{2}}(\mathbf{w})&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \nonumber \\
+S_{\nu=\frac{3}{2}}(\mathbf{w})&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \nonumber \\
+S_{\nu=\frac{5}{2}}(\mathbf{w})&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3}, \nonumber 
+\end{eqnarray}
+
+\noindent where variable $\mathbf{w}$ is in the frequency domain.
 

### 2019-02-26 Paper/manuscript/manuscript_05.tex
review Paul changes

@@
 
-Sparse GPs are based on building low-rank approximations of the covariance matrix of the complete data by means of 'inducing variables' (Paul: if you introduce this term, you may need to explain it.
-Alternatively, can we just get rid of it?).
+Sparse GPs are based on building low-rank approximations of the covariance matrix of the complete data by means of selecting a subset $m$ of training points on which to base computation.
 They reduce the dimension of the posterior distribution thereby reducing the memory requirements to $O(nm)$ and computational complexity to $O(nm^2)$ for some $m < n$.
@@
 A unifying view on sparse GPs based on approximate generative methods is provided in \cite{quinonero2005unifying}, while a general review can be found in \cite{rasmussen2006gaussian}.
-More recent developments in the context of sparse GPs include points-based sparse approximation methods \citep{wilson2015kernel}, which have been further developed by Bui et.
-al (2017) (Paul: cite this study correctly in latex) to perform approximations at inference time rather than at modeling time.
+(Gabi: Could you choose between one of these two set of sentences in different colors?) \textcolor{blue}{More recent developments in the context of sparse GPs include points-based sparse approximation methods \citep{wilson2015kernel}, which have been further developed by \citep{bui2017unifying} to perform approximations at inference time rather than at modeling time.} \textcolor{red} {New developments in the context of sparse GPs include a structured kernel interpolation (SKI) method \citep{wilson2015kernel}.
+More recently \citep{bui2017unifying} revisit the inducing points-based sparse approximation methods to perform approximations at inference time rather than at modeling time.}
 
@@
 The posterior distribution of the reduced-rank GP approximation is of dimension $m$ and thus considerably smaller than in the case of exact GPs.
-This not only lowers the memory requirements but also reduces the correlations between latent values and the corresponding hyperparameters of the GP, which can help to further improve sampling efficiency.
-
-While \cite{solin2018hilbert} fully developed the mathematical theory behind reduced-rank approximations of GPs, they did not put much effort in analyzing the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space (Paul: I replace 'box size' because, at that point, readers will likely have no idea what that means), or properties of the true functional relationship between covariates and response variable.
+This not only lowers the memory requirements and computational complexity but also reduces the correlations between latent values and the corresponding hyperparameters of the GP, which can help to further improve sampling efficiency.
+
+While \cite{solin2018hilbert} fully developed the mathematical theory behind reduced-rank approximations of GPs, they did not put much effort in analyzing the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
 However, for practical implementations of approximate GPs, the latter factors are critical and thus require further investigation.
@@
 
-Although there are several GP specific software packages available to date (Paul: cite those), each provide efficient implementations only for a restricted range of GP based models.
-In this paper we do not focus on the fastest possible inference for a specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks such as Stan (cite).
+Although there are several GP specific software packages available to date (Paul: cite those.
+Gabi: This full paragraph comes from Aki.
+I think he refers to some Splines methods and maybe others approaches as well), each provide efficient implementations only for a restricted range of GP based models.
+In this paper we do not focus on the fastest possible inference for a specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
 To this end, we want to make full Bayesian inference over Gaussian process models accessible to a broader audience by making them easy, fast and save to apply.
@@
 (Paul: The content and the purpose of the following sentences is not fully clear to me.
-Do we need them in the Introduction or can they be removed, or put into the next section?) While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
+Do we need them in the Introduction or can they be removed, or put into the next section?
+Gabi: In this paragraph I tried to state two differences I found in the literature between our method and the Sparse Spectrum GP and the Splines method.
+But I not know where locate them.) While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
 Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.
@@
 \\
-We assume a Gaussian process prior for the collection of function values $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N $, $\mathbf{f} \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
+We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
 %
@@
 
-The mean and covariance functions represent how the random functions behave on average and how the different points in the input space co-vary with respect to each other.
-The mean and covariance functions completely characterize the multivariate Gaussian for $\mathbf{f}$,
+The mean and covariance functions completely characterize the Gaussian process prior, and represent how the random functions behave on average and how the different points in the input space co-vary with respect to each other, respectively.
+The Gaussian process prior can be interpreted such as any finite number of function values $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is multivariate Gaussian,
 \begin{align*}
@@
 
-\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ is given by the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ are formed by evalating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
 The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
@@
 %
-\begin{eqnarray} \label{bf&lambda_multi}
+\begin{align} \label{bf&lambda_multi}
 \begin{split}
-\phi^{\ast}_j(\mathbf{x}) &=& \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right) \\
-%
-\lambda^{\ast}_j &=& \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  
+\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right) \\
+%
+\lambda^{\ast}_j &= \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  
 \end{split}
-\end{eqnarray}
+\end{align}
 

### 2019-02-27 Paper/manuscript/manuscript_05.tex
improve section 3 and add some case studies

@@
 %\author{
-%\name{Gabriel Riutort-Mayol, Michael Riis Anderseen, Aki Vehtari, Paul-Christian Burkner}
+%\name{}
 %\affil{\textsuperscript{a}Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain; }
@@
 $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
-Gaussian processes have the defining property that the marginal distribution of every finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
-\\
-\\
+Gaussian processes have the defining property that the marginal distribution of every finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.\\
+
 In this work, Gaussian processes will take the role of prior distributions over function spaces for non-parametric regression in a Bayesian setting. 
@@
 Our goal is to predict the value of the function $f^*=f(\mathbf{x}^*)$ evaluated at a new input point $\mathbf{x}^* \in \mathbb{R}^D$.
-That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.
-\\
-\\
+That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
+
 We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
@@
 \\
-\\
+
 In this paper we proposed an approximate model for Gaussian processes based on basis functions expansion which basically turns the GP prior model into a linear model.
-This linear GP model representation makes inference considerably faster and can be used as latent function in non-Gaussian observational models allowing modelling flexibility.
-\\
-\\
+This linear GP model representation makes inference considerably faster and can be used as latent function in non-Gaussian observational models allowing modelling flexibility.\\
+
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
-A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, so we can write $k(\mathbf{x},\mathbf{x}') = k(\mathbf{x}-\mathbf{x}')$, which means that the covariance is invariant to translations.
+A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(\mathbf{x}-\mathbf{x}')$, which means that the covariance is invariant to translations.
 Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariacne is both translation and rotation invariant.
@@
 
-Stationary covariance functions can be represented in terms of their spectral densities \citep{Rasmussen_2006}.
+Stationary covariance functions can be represented in terms of their spectral densities \citep{rasmussen2006gaussian}.
+In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure (\textit{Bochner's theorem}, see, e.g.
+\cite{akhiezer1993theory}).
+If this measure has a density, it is known as the spectral density of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the \textit{Wiener-Khintchine theorem} \citep{rasmussen2006gaussian}.
 The spectral density functions associated with the Mat\'ern covariance functions written above are
@@
 
-The performance and accuracy of the method are directly related with the number of basis functions and the box size, and, at the same time, these two factors depend on the smoothness or roughness of the process.
-The time of computation is mainly dependent on the number of basis functions.
-Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.
-
-\vspace{2mm}
-$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the function.
-
-\vspace{2mm}
-$\bullet$ We make recomendations for the values of these factors based on the recognized relations among them.
-We provide useful graphs of these relations that will help the users to improve performance and save time of computation.
-
-\vspace{2mm}
-$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
+Implementation and performance of the approximate GP model depends on the number of basis functions and box size settings, which are also related each other.
+The time of computation in the approximation mainly depends on the number of basis functions.
+At the same time, the appropriate values for the number of basis functions and box size depend on the non-linear functional effects of the function to be learned.
+In GP functions these non-linear functional effects are characterized by the lengthscale of the covariance function.
+
+In this section we analyze the rule and relationships of the number of basis functions and the box size on approximation and their effects on the learned functions.
+
+We also analyze these two factors in relation to the lengthscale of the functions to be learned.
+We provide a useful graph that retains how these factors, number of basis fucntions, box size and lengthscale, relate each other in relation to the performance of the model.
+This graph basically provides the recommended values for these factors in order to get a close approximation.
+This graph also allows us to diagnose an actual fit.
+
+In this section we also make a comparison between regular GP model and approximate GP model fitted for different dataset with different "wigglyness".
+The comparison is made through comparing their estimated lengthscales.
 
@@
 \begin{equation}\label{diffcov}
-k(\tau) \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
+k(\tau) \approx \sum_{j} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
 \end{equation}
@@
 \noindent where $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) in a unidimensional case and those in equation (\ref{bf&lambda_multi}) in a multidimensional case.
-
-The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity $J=1,\dots,j,\dots,\infty$.
-The number of basis functions can be truncated that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) for the unidimensional case and those in equation (\ref{bf&lambda_multi}) for the multidimensional case.
+
+The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
+The number of basis functions can be truncated at some finite positive value $J$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
 %
 \begin{eqnarray}\label{diff_covs}
-\int k(\tau)d(\tau) - \int \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
-\end{eqnarray}
-
-The finite number of basis functions in the approximation to satisfy eq.
-(\ref{diff_covs}) will depend on the smoothness or roughness of the function, that is, on the lengthscale $\ell$ of the GP process.
-The approximation will also depends on the box size $L$ which will affect the performance specially near the boundaries.
-
-To illustrate the effects of these two factors, number of basis functions $J$ and box size $L$, on the performance of the approximation, let's consider a simulated dataset of 75 noisy observations from a GP prior with lengthscale $\ell=0.3$ and magnitud $\alpha=1$.
-The additive noise added to the GP prior simulations is $\sigma=0.2$.
-A regular GP model and different approximate GP models with different choices for the number of basis function $J$ and box size $L$ are fitted on this dataset.
-
-For a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of the number of basis functions in approaching the posterior mean and the covariance function.
-The number of basis functions affects the non-linearity or roughness (or smoothness) of the posteriors.
+\int k(\tau)d(\tau) - \int \sum_{j}^J S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
+\end{eqnarray}
+
+The finite number $J$ of basis functions in the approximation needed to satisfy eq.
+(\ref{diff_covs}) will depend on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
+The approximation will also depends on the box size $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
+The box size will also have an effect on the number of basis functions needed in the approximation.
+Note that we refer to the box size as the value $L$ of the Dirichelt boundary condition in eq.
+\ref{eigenproblem}.
+The box size $L$ is set as an extension of the desired predicting input domain $\Psi$.
+Considering a zero-mean input domain $\Psi=[-S,S] \in \mathbb{R}$, the box size is defined as,
+%
+\begin{equation}
+L=c \cdot S
+\end{equation}
+
+\noindent where $S$ represents the half-range of the input space, and $c$ ($c \ge 1$) is the proportional extension factor.
+From now on we are going to refer to this proportional factor $c$ as the boundary factor of the approximation.
+The boundary factor can also be regarded as the box size $L$ normalized by the half-range of the input space $S$.\\
+
+For a naive illustration of the effects of these two factors, number of basis functions $J$ and box size $L$, on the performance of the approximation, a set of noisy observations are simulated from a GP model with lengthscale $\ell=0.3$ and magnitud $\alpha=1$ for input vales within the input domain with half-range $S=1$.
+Approximate GP models with different number of basis functions and box sizes $L$ ($L=c\cdot S$, with $c\ge 1$) are fitted to this data.
+The lengthscale and magnitud parameters are considered fix to the true values of the generative GP model.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} ilustrate the individual effects of the number of basis functions and the box size, respectively, on the approximation of the covariance function.
+On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function.
+It can be seen how the number of basis functions affects the non-linear effects of the posteriors.
 Fewer basis functions implies smoother covariance functions and consequently smoother posteriors.
-The more roughness the process is, the more basis functions are needed.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.4, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.4, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
+The more "wigglyness" of the function, the more basis functions are needed.
+On the other hand, and similarly, for a fixed number of basis functions $J$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the box size $L$ in approaching the posterior mean and the covariance function.
+It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances.
+\\
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
 \caption{(left) Posterior distributions for different number of basis functions $J$. (right) Covariance function for different number of basis functions $J$.}
@@
 
-Similarly, for a fixed number of basis functions $J$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of the box size $L$ in approaching the posterior mean and the covariance function.
-The box size mainly affects the approximation near the boundaries of the function, and affects the covariances at long distances.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.4]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig2_Cov_L.pdf}}
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.43]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.43]{fig2_Cov_L.pdf}}
 \caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
@@
 
-Therefore, the approximation to the covariance function depends on the number of basis functions $J$ used in the series expansion and the value for boundary condition or box size $L$.
-Additionally, there exit a relation among the number of basis functions, the box size and the lengthscale or roughness of the process on the performance of the approximation.
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in order to achieve a good approximation to the covariance function of the regular GP using eq.
-(\ref{diff_covs}).
-On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data.
-The countour lines gather the box size which is also normalized by the half-range of the data, to which we refer as the boundary factor $c$, where $L=c \cdot \text{range}(x)/2$.
-And the Y-axis gathers the number of basis functions.
-Thus, this figure give us the minimum number of basis functions needed to achieve a good approximation in terms of satisfying eq.
-(\ref{diff_covs}), for a certain GP process with lengthscale $\ell$ and given a fixed boundary factor $c$.
-Alternatively, this figure could also be read as the minimum boundary factor $c$ that we should use given a number of basis functions, for certain GP process with certain lengthscale.
-And also this figure could also be read as the maximum roughness in a process that can be fitted properly given a number of basis functions and a box size.
-This figure basically collects the behaviour of the approximate GP model in function of the lengthscale, number of basis functions and boundary condition:
-
-$\bullet$ As the lengthscale or smoothness of the process increase, the boundary factor and the number of basis functions needed for the approach decrease.
-
-$\bullet$ There is a minimum boundary factor to achieve a good approach in function of the lengthscale (The contour lines have an end in function of the lengthscale).
-
-$\bullet$ The lower boundary factor, the fewer basis functions, and the lower lengthscales suitable to be fitted.
-
-$\bullet$ This plot serves as a diagnosis tool in the sense that if the estimated lengthscale is lower than that minimun generated by the figure for the chosen number of basis functions and boundary factor means that we would need to increase the number of basis functions or decrease the boundary factor.
-
-$\bullet$ On the other hand, if the lengthscale is bigger than that generated by the figure then it means that the fit should be good.
-
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $J$ and the boundary factor $c$ to approach a certain process characterized by its lengthscale.
-This Figure will allow us to optimize computations which depends on the number of basis functions.
-Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit.
-This figure also provides a diagnosis tool of the fit, in the sense that if the estimated lengthscale is lower than the minimum lengthscale able to fit with the actual number of basis functions and box size, indicates that the number of basis functions needs to be increased or the box size decreased.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
-\caption{Relation between the minumum number of basis functions $J$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{\frac{\text{range}(x)}{2}}$), and the boundary condition factor $c$ ($c = \frac{L}{\frac{\text{range}(x)}{2}}$).}
-  \label{fig3_lscale_vs_J_vs_c_part1}
-\end{figure}
-
-The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
+In the previous Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} we focused on illustrating the individual effects of the number of basis functions and the box size.
+Furthermore, for a clearer illustration of these effects, the lengthscale and magnitud parameters in the approximate GP model were fixed to their true values.
+Now we are going to focus on analyzing the interaction effects between these two factors, number of basis functions and the boundary factor, on the performance of the approximation.
+The lengthscale and magnitud parameters will not be fixed rather they will be estimated in both regular and approximate GP models.
+%The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
 Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
-Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors distributions and the covariance functions of the approximate GP model and the regular GP model, obtained after fitting the data.
-Figure \ref{fig5_MSE_vs_J} shows the standarized root mean square error of both models in function of the number of basis functions and the boundary factor.
-Figure \ref{fig6_lscale_vs_J} shows the evolution of the GP hyperparameters, lengthscale and magnitud, in function of the number of basis functions and boundary factor.
-Following Figure \ref{fig5_MSE_vs_J} it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5.
-The choice of 10 basis functions and a boundary factor of 1.5 could be accurate enough.
-This same conclusion can be roughly seen in the right-side plots in Figure \ref{fig9_Posteriors_exI}.
-This conclusions also agree with Figure \ref{fig3_lscale_vs_J_vs_c_part1} from which we can deduce that for a process with a normalized lengthscale of 0.3, 10 basis functions and a normalized box size (boundary factor) of 1.5 would be enough to get a good approximation following eq.
-(\ref{diff_covs}).
-
-From Figure \ref{fig5_MSE_vs_J} it can also be drawn the behaviour in terms of performance of the approximate GP model in function of the number of basis functions and the boundary factor.
-Two main behaviours can be deduced from Figure \ref{fig5_MSE_vs_J}, which can also be easily seen from Figure \ref{fig3_lscale_vs_J_vs_c_part1}:
+Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors means and the covariance functions obtained after fitting the data, for different number of basis functions and boundary factors.
+Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model over the regular GP model as a function of the number of basis functions and the boundary factors.
+Figure \ref{fig6_lscale_vs_J} shows the lengthscale and magnitud parameters for the regular GP model and for the approximate models as a function of the number of basis functions and boundary factors.
+Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5.
+The choice of 10 basis functions and a boundary factor of 1.5 could be an accurate enough choice.
+This same conclusion can be also be intuitively seen in the posterior means and covariance functions plots in Figure \ref{fig4_Post_&_Cov_part1}.
 \\
+
+From Figure \ref{fig5_MSE_vs_J}, a behaviour in terms of performance of the approximate GP model as a function of the number of basis functions and the boundary factor, can be deduced: \\
 $\bullet$ as the boundary factor increases, more basis functions are needed,
 \\
-$\bullet$ as fewer basis functions are used, the boundary factor must decrease.
-
-Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and magnitud parameter as function of the number of basis functions and the boundary factor.
-The same conclusion drawn before can be drawn form this figure where with 15 basis functions and a boundary factor of 1.5 the estimated lengthscale of the approximate GP model approach the estimated lengthscale of the regular GP model.
-From this figure, it can be drawn that exists a minumum value for the boundary factor under this a good approach will never be achieved.
-This conclusion also appears in Figure \ref{fig3_lscale_vs_J_vs_c_part1} the contour lines of the boundary factor have an end at certain lengthscales, which means that exist a minimum boundary factor depending on the lengthscale of the process.
-
-\begin{figure}[H]
+$\bullet$ as fewer basis functions are used, the boundary factor must decrease. \\
+
+Similar conclusions can be stated from results of Figure \ref{fig6_lscale_vs_J}, This Figure shows the estimated lengthscale and magnitud parameter as function of the number of basis functions and the boundary factor.
+In addition to the above conclusions, it can be easily seen that exists a minimum value for the boundary factor under which a close approximation will never be achieved.
+\\
+
+\begin{figure}
 \centering
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
@@
 
-\subsection{Comparative between true lengthscales and estimated lengthscales}
-
-In this section we fit the regular GP model and the approximate GP model on different datasets drawn from a GP prior model with a squared exponential covariance function with different lengthscales for every dataset.
-The approximate GP model is fitted using different number of basis functions.
-Figure \ref{fig7_varing_lscale} shows the posterior means of the exact GP model and the approximate GP model with different number of basis functions for every dataset with different wiggly effects (lengthscale).
+Additionally, there exit a relation of the number of basis functions $J$ and the boundary factor $c$ with the lengthscale $\ell$ of the function.
+Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in relation to the performance of the approximation evaluated using eq.
+(\ref{diff_covs}).
+On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data $\ell/S$.
+The countour lines gather the boundary factor $c$.
+And the Y-axis represents the number of basis functions.
+Thus, this figure give us, for a certain GP function with lengthscale $\ell$ and given a fixed boundary factor $c$, the minimum number of basis functions needed to achieve a close approximation in terms of satisfying eq.
+(\ref{diff_covs}).
+Alternatively, this figure could also be read as the minimum boundary factor $c$ that we should use given a number of basis functions, for certain GP function with certain lengthscale.
+And also this figure could also be read as the minimum functional lengthscales that can be closely approximated given a number of basis functions and a boundary factor.
+This figure basically collects the behaviour of the approximate GP model as a function of the lengthscale, number of basis functions and boundary factor:
+
+$\bullet$ As the lengthscale of the function increases, the boundary factor and the number of basis functions needed for the approximation decrease.
+
+$\bullet$ There is a minimum boundary factor under which a close approximation never is going to be achieved (The contour lines have an end in function of the lengthscale).
+This statement match with one of the behaviour recognized in Figure \ref{fig6_lscale_vs_J} where were a minimum value for the boundary factor under which a close approximation will never be achieved.
+
+$\bullet$ The lower the boundary factor, the fewer basis functions and the smaller lengthscales that are suitable to be fitted.
+
+$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimun that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
+On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.\\
+
+Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $J$ and the boundary factor $c$ to approximate certain process characterized by its lengthscale.
+This Figure will allow us to optimize computations which depends on the number of basis functions.
+Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit.
+This figure also provides a diagnosis tool of the fit by comparing a actual estimated lengthscale to that the figure bring.
 
@@
 \centering
-\includegraphics[scale=0.4]{fig7_varing_lscale.pdf}
+\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
+\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
+\caption{Relation between the minumum number of basis functions $J$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary condition factor $c$ ($c = \frac{L}{S}$).}
+  \label{fig3_lscale_vs_J_vs_c_part1}
+\end{figure}
+
+If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J} where 10 basis functions and a boundary factor of 1.5 would be enough to closely approximate a function with normalized lengthscale eaual to 0.3, we can recognize that these conclusions also matches the conclusion brought form Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+
+\subsection{Comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model}
+
+Finally, we make a comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model in several datasets with different "wigglyness" (or lengthscales).
+From this analysis we can recognize when an approximation is not close enough to the regular GP when the lengthscales are not similar.
+We contrast these results from these lengthscale similarities between the regular and the approximate GP model with the information brought from the Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+
+We fit the regular GP model and the approximate GP model on different datasets.
+These datasets are drawn from GP prior models with a squared exponential covariance function with different lengthscales for every dataset.
+The approximate GP models is fitted using different number of basis functions.
+The boubdary condition factor is set en ivery fit to a right and optinal choice.
+Figure \ref{fig7_varing_lscale} shows the posterior means of the exact GP model and the approximate GP model with different number of basis functions for every dataset with different wiggly effects (different lengthscale).
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
 \caption{Posterior means of the exact and approximate GP models fitted over different datasets with different lengthscales.}
@@
 
-Figure \ref{fig8_Tlscale_vs_Elscale} compares the lengthscales of the true generating process to the estimated lengthscales for the different datasets.
-Different comparison are shown using different number of basis functions in the approximate GP model.
+The comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model in several datasets is shown in Figure \ref{fig8_Tlscale_vs_Elscale}.
+The X-axes of the plots represent the true lengthscales of the generative functions for every dataset scenario, and in the Y-axes are represented the estimated lengthscales for the regular GP and the approximate GP models.
+Different comparison are made using different number of basis functions in the approximations.
 Let's assume that the box size parameter was properly set in every case.
+The dashed black line represents the minimum lengthscales that can be closely approximate, given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE of the approximate GP model over the regular GP model for the different datasets with different lengthscale scenarios and the number of basis functions used in the approximation.
+
+From Figure \ref{fig8_Tlscale_vs_Elscale} we can recognize similarities and dissimilarities between the estimation for   the lengthscales of the regular GP model and the approximations, which indicates the performance of the model.
+Figure \ref{fig9_MSE_varing_lscale} with the SRMSE also indicates the performance of the model.
+We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for very small lengthscales and very long lengthscales.
+Figure \ref{fig8_Tlscale_vs_Elscale} shows a correpondence between very low lengthscales when the SRMSE is large, conversely there is no correspondence between very large lengthscale and the SRMSE is close to zero.
+These effects can make sense because, in the case of long lengthscales these have larger variabilities obtaining  similar posteriors, and in the case of low lengthscale these have lower variabilities obtaining diferent posteriors.
+
 The dashed black line represents the minimum lengthscales that can be properly fitted by the approximate GP model according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-Figure \ref{fig9_MSE_varing_lscale} shows the SMSE of the approximate GP model for the different datasets with different wiggly effects and the number of basis functions used in the approximation.
-
-From Figure \ref{fig8_Tlscale_vs_Elscale} we can recognize when the lengthscales between the regular GP model and the approximate GP model match each other, which should be an indicator of a good approximation.
-Figure \ref{fig9_MSE_varing_lscale} indicates when the MSE is near to zero which also indicates a good approximation.
-We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for very low lengthscales and very long lengthscales.
-Figure \ref{fig8_Tlscale_vs_Elscale} shows a correpondence between very low lengthscales when the SMSE is large, conversly there is no correspondence between very large lengthscale and the SMSE is close to zero.
-These effects can make sense because, in the case of long lengthscales these have larger variabilities obtaining  similar posteriors, and in the case of low lengthscale these have lower variabilities obtaining diferent posteriors.
-
-The dashed black line represents the minimum lengthscales that can be properly fitted by the approximate GP model according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.4]{fig8_Tlscale_vs_Elscale.pdf}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
 \caption{True lengthscales versus estimated lscales of the different realizations of datasets using different number of basis functions. }
@@
 \centering
-\includegraphics[scale=0.5]{fig9_MSE_varing_lscale.pdf}
+\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
 \caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and the number of basis functions.}
@@
 
-A simulated one-dimensional dataset, with 250 data points drawn from a Gaussian process prior with a Matern covariance function with 3/2 degrees of freedom.
-The hyperparameters of this GP prior with a Matern kernel, marginal variance and lengthscale, are set to 1 and 0.15, respectively.
-The additive Gaussian noise to these drawn points is set to 0.2.
-We compare the modelling performance over this simulated process of our proposed approach with the performance of a regular GP and a Splines-based model.
+A simulated one-dimensional dataset, with 250 observations $f(\mathbf{x}_i)$ ($i=1,\cdots,250$) drawn from a Gaussian process prior with a Matern covariance function with 3/2 degrees of freedom, with inputs $\mathbf{x} \in \{-1,1\} \subset {\rm I\!R}$.
+The hyperparameters of this GP prior with a Matern kernel, marginal variance $\alpha$ and lengthscale $\ell$, are set to 1 and 0.15, respectively.
+The additive Gaussian noise $\sigma$ to these drawn points is set to 0.2.
+The latent Gaussian process model is as follows,
+%
+\begin{eqnarray*}
+\mathbf{y} &=& f(\mathbf{x}) + \boldsymbol{\epsilon} \\
+\boldsymbol{\epsilon} &\sim& \mathcal{N}(0, \sigma^2  \mathbf{I}) \\
+f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x}, \theta)),
+\end{eqnarray*}
+
+\noindent where $\mathbf{I}$ is the identity matrix and $k$ the Matern covariance function in eq. (4). More efficient when using HMC than this latent GP model would be the marginalized model,
+%
+\begin{equation}\label{eq:marginalized_gp}
+\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 \mathbf{I} )
+\end{equation}
+
+\noindent where the element $K_{ij}$ is the covariance function $k(\mathbf{x}_i, \mathbf{x}_j, \theta$) evaluated at the inputs $\mathbf{x}_i$ and $\mathbf{x}_j$.
+
+In the approximate GP model, the covariance function $k$ is approximated as in eq.
+(\ref{approxcov}) with the spectral density in eq.
+(8).
+The latent function $f(\mathbf{x)}$ is approximated as in eq.
+(\ref{approxf}).
+
+We compare the modelling performance over this simulated process of our proposed approach with the performance of a regular GP following the marginalized version in eq.
+(\ref{eq:marginalized_gp}) and a Splines-based model.
 For our basis function approach we use $J=80$ basis functions and a boundary factor $c=1.2$.
@@
 
-\subsection{Case I: Gay data}\label{sec:bf_caseII}
+\subsection{Gay data}\label{sec:gay_data}
+This data set relates the proportion of support for same-sex marriage to the age.
+This proportions correspond to binomial observations per age group $i$.
+The data consist of $i=1,\cdots,74$ binomial observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group.
+
+The observational model for the observed number of people $y_i$ supporting same-age marriage per age group $i$ is a binomial model with the total number of people per age group $n_i$ and the probability per age group $p_i$ as model parameters.
+The total number of people per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ per age group $i$ or mean number of support people per age group $\bar{y}=p_i \cdot n_i$.
+%
+\begin{equation*}
+y_i \sim \text{Binomial}(p_i, n_i)
+\end{equation*}
+
+The probability $p_i$ is related through the logit function to a latent Gaussian process prior distribution with the age values $x_i$ as inputs,
+%
+\begin{eqnarray*} \label{eq:gpprior_gay}
+p_i &=& \text{logit}(f(x_i)) \nonumber \\
+f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x}, \theta))
+\end{eqnarray*}
+
+\noindent with a squared exponetial covariance function $k$ in eq. (2).
+
+In the approximate GP model the covariance function is approximated as in eq.
+(\ref{approxcov}) with spectral density in eq.
+(6),  and the latent function $f(\mathbf{x})$ as in eq.
+(\ref{approxf}).
+As comparison a splines model is also fitted.
+Figure \ref{fig12_Posteriors_gaydata} shows the posterior distributions of the three models.
+In the case of the approximate GP model the posterior shown in the figure corresponds to 20 basis functions and a boundary factor of 1.5.
+For the spline posterior shown in the figure, it correspond to a 20 knots.
+Two small sets of observations have been used as test data.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
+\caption{Posterior distributions of the proposed approximated GP model, the exact GP model, and the Splines model.}
+  \label{fig12_Posteriors_gaydata}
+\end{figure}
+
+Figure \ref{fig13_MSE_train_BF_gaydata} shows the MSE against the regular GP model of the approximate GP model for the different number of basis functions and boundary factor.
+The expected patterns of the approximation in function of the number of basis function and boundary factor are recognized.
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
+\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
+  \label{fig13_MSE_train_BF_gaydata}
+\end{figure}
+
+Figure \ref{fig14_MSE_train_gaydata} shows the MSE of the different models against the actual data, for training set of data and the test set of data.
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
+\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
+  \label{fig14_MSE_train_gaydata}
+\end{figure}
+
 \subsection{Case II: Birthday data}\label{sec:bf_caseIII}

### 2019-03-03 Paper/manuscript/manuscript_05.tex
improve sections 2.2 and 2.3

@@
 A unifying view on sparse GPs based on approximate generative methods is provided in \cite{quinonero2005unifying}, while a general review can be found in \cite{rasmussen2006gaussian}.
-(Gabi: Could you choose between one of these two set of sentences in different colors?) \textcolor{blue}{More recent developments in the context of sparse GPs include points-based sparse approximation methods \citep{wilson2015kernel}, which have been further developed by \citep{bui2017unifying} to perform approximations at inference time rather than at modeling time.} \textcolor{red} {New developments in the context of sparse GPs include a structured kernel interpolation (SKI) method \citep{wilson2015kernel}.
-More recently \citep{bui2017unifying} revisit the inducing points-based sparse approximation methods to perform approximations at inference time rather than at modeling time.}
+New developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
+More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.}
 
@@
 
-\subsection{Hilbert space-based approximate Gaussian process model}
-
-The approximate Gaussian process method, developed by \cite{solin2018hilbert} and implemented in this paper, lay on the basis of considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
-Then, the pseudo-differential operator is approximated with Hilbert space methods on compact subsets $\Omega \subset {\rm I\!R}^D$, and with some boundary condition.
-
-First, we focus on the unidimensional case of the input space, such as $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value.
-
-\vspace{0.2cm}
-The approximate method leads to the approximation of the covariance function between two input values $\{x,x'\} \in \{-L,L\}$ as 
+\subsection{Hilbert space approximate Gaussian process model}
+
+The approximate Gaussian process method, developed by \cite{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
+Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to some boundary condition.
+For brevity, we will refer to these approximate Gaussian processes as HSGPs (Paul: introducing such an abbreviation perhaps shortens what we are going to write in the next sections).
+Below, we will present the main results around HSGPs relevant for practical application.
+More details and mathematical proofs are provided in \cite{solin2018hilbert}.
+
+We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition.
+As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuarcy of HSGPs.
+We will come back to this issue in Section 3.
+
+Within $\Omega$, we can write any stationary covariance function with input values $\{x,x'\} \in \Omega$ as 
 %
 \begin{equation}\label{approxcov}
-k(x,x') \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
-\end{equation}
-
-\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ and $\theta$ the set of hyperparameters of $k$; a stationary covariance function can be equivalently represented in terms of the spectral density \citep{rasmussen2006gaussian}, then the spectral density is a function of the hyperparameters $\theta$.
-$\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
-Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
+k(x,x') = \sum_{j=1}^\infty S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
+\end{equation}
+
+\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section 2.1) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
+The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+Namely, they satisfy the following eigenvalue problem in $\Omega$ when applying the Dirichlet boundary condition (other boundary conditions could be used as well):
 %
@@
 \begin{split}
--\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \\ 
-\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
+-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x \in \Omega \\ 
+\phi_j(x)&=&0, \hspace{2cm} x \notin \Omega.
 \end{split}
@@
 
-\noindent The eigenvalues $\lambda_j>0$ are real and positive, because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in eq.
-(\ref{eigenproblem}) are sinusoidal functions,
+\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in Equation (\ref{eigenproblem}) are sinusoidal functions.
+Independently of the covariance function, they can be computed as
 %
@@
 
-If we truncate the sum in (\ref{approxcov}) up to $J$ ($j=1,\cdots,J$), the approximate covariance function can be represented as
+If we truncate the sum in (\ref{approxcov}) to the first $m$ terms, the approximate covariance function becomes
 %
 \begin{equation}
-k(x,x') \approx \boldsymbol{\phi}(x)^\intercal \Delta \boldsymbol{\phi}(x'), \nonumber
-\end{equation}
-
-\noindent where $\boldsymbol{\phi}(x)=\{\phi_j(x)\}_{j=1}^{J} \in {\rm I\!R}^{J}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{J\times J}$ the diagonal matrix of espectral densities $S_{\theta}(\sqrt{\lambda_j})$,
+k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \boldsymbol{\phi}(x)^\intercal \Delta \boldsymbol{\phi}(x'), \nonumber
+\end{equation}
+
+\noindent where $\boldsymbol{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is the diagonal matrix of the spectral densities $S_{\theta}(\sqrt{\lambda_j})$:
 %
@@
     & \ddots & \nonumber \\
-    & & S_{\theta}(\sqrt{\lambda_J}) \\
+    & & S_{\theta}(\sqrt{\lambda_m}) \\
   \end{bmatrix}.
@@
 
-Thus, the Gram matrix $\text{K}$ of the covariance function $k$ for a collection of observations $i=1,\cdots,N$ and collection of input values $\{x^i\}_{i=1}^{N} \in {\rm I\!R}^{N}$ can be represented as follows
+Thus, the Gram matrix $\text{K}$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
 %
@@
 
-\noindent where $\Phi \in {\rm I\!R}^{N\times J}$ is the matrix of eigenfunctions $\phi_j(x^i)$,
+\noindent where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x^i)$:
 %
@@
 \Phi &=&  \left[ {\begin{array}{ccc}
-   \phi_1(x^1) & \cdots & \phi_J(x^1)  \\
+   \phi_1(x_1) & \cdots & \phi_m(x_1)  \\
     \vdots &\ddots & \vdots  \nonumber \\ 
-    \phi_1(x^N) & \cdots & \phi_J(x^N) \\
+    \phi_1(x_n) & \cdots & \phi_m(x_n) \\
   \end{array} } \right].
@@
 
-\noindent Now, the Gaussian process prior for the function $f$ can be re-defined as
+\noindent As a result, the GP model for the function $f$ can be re-defined as
 %
@@
 
-\noindent This equivalently leads to a linear representation of the function $f$,
+\noindent This equivalently leads to a linear representation of $f$ via
 %
 \begin{equation}\label{approxf}
-f(x) \approx \sum_{j} \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
-\end{equation}
-
-\noindent where $\beta_j \sim \text{N}(0,1)$ is a random variable with zero mean and unitary variance.
-That is, the function $f$ can be approximated with a finite basis function expansion (eigenfunctions $\phi_j$ of the Laplace operator), scaled by the squared root of the spectral density as a function of the eigenvalues $\lambda_j$.
-The eigenfunctions $\phi_j$ do not depend on the kernel hyperparameters $\theta$, the only dependence on $\theta$ is through the spectral density $S_{\theta}$.
-The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and the spectral density goes rapidly to zero for bounded covariance functions.
-Therefore, it is expected a good approximation in eq.
-(\ref{approxf}) for a finite number, $J$, of terms in the series as long as the inputs values $x^i$ are not near the boundaries of the domain $[-L,L]$, where the Laplacian was taken to be zero.
-
-The computational cost of this approximate model scales $O(n\cdot J+J)$, where $n$ is the number of observations and $J$ the number of basis functions.
-
-\subsection{Generalization to multidimensional input space}
-
-We are going to generalize the results from the previous section to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions.
-
-\vspace{0.2cm}
-In a $D$-dimensional input space the total number of eigenfunctions and eigenvalues in the approximation corresponds to the total number of $D$-tuples over $J$, which is $\prod_{d=1}^{D} J_d$, where $J_d$ is the number of basis function for the dimension $d$.
-Let $\mathbb{S}\in {\rm I\!R}^{\prod_{d=1}^{D} J_d \times D}$ be the set of $D$-tuples elements.
-Each new eigenfunction $\phi^{\ast}_j$ correspond to the product of the univariate eigenfunctions whose indices corresponds to the the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
-And each new eigenvalue $\lambda^{\ast}_j$ is a $D$-vector whose elements are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
-Let $j=\{1,\cdots,J^D\}$,
-%
-\begin{align} \label{bf&lambda_multi}
-\begin{split}
-\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right) \\
-%
-\lambda^{\ast}_j &= \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  
-\end{split}
-\end{align}
-
-As an example we show the matrix $\mathbb{S}$ for a $three$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2,\mathbf{x}_3\}$ ($D=3$) and $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimension, respectively.
-The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$.
-The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
-%
+f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
+\end{equation}
+
+\noindent where $\beta_j \sim \text{Normal}(0,1)$ are normal random variables with zero mean and unitary variance.
+That is, the function $f$ can be approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
+A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the covariance hyperparameters $\theta$.
+Instead, the only dependence on $\theta$ is through the spectral density $S_{\theta}$.
+The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions.
+Therefore, Equation (\ref{approxf}) can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
+The computational cost of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
+
+\subsection{Generalization to multidimensional GPs}
+
+We will now show how to generalize the results from the previous section to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions.
+In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is possible combinations of univariate eigenfunctions over all dimensions.
+The number $D$-tuples is given by $m^{\ast} = \prod_{d=1}^{D} m_d$, where $m_d$ is the number of basis function for the dimension $d$.
+Let $\mathbb{S}\in {\rm I\!N}^{m^{\ast} \times D}$ be the matrix of all those $D$-tuples.
+For example, suppose we have $D=3$ covariates and use $m_{1}=2$, $m_{2}=2$ and $m_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimension, respectively.
+Then, the number of multivariate eigenfunctions and eigenvalues is $m^{\ast} = m_{1} \cdot m_{2} \cdot m_{3} = 12$ and the matrix $\mathbb{S}\in {\rm I\!N}^{12 \times 3}$ is given by
+
 \begin{eqnarray}
@@
 
-The approximate covariance function is represented as
+Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
+And each multivariate eigenvalue $\lambda^{\ast}_j$ is a $D$-vector whose elements are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
+Thus, for $\mathbf{x}^D \in \Omega$ and $j=1,\ldots,m$, we have
+%
+\begin{align} \label{bf&lambda_multi}
+\begin{split}
+\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \\
+%
+\lambda^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. 
+\end{split}
+\end{align}
+
+\noindent The approximate covariance function is then represented as
 %
 \begin{equation}\label{approxcov_multi}
-k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{J^D} S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right) \phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
+k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{m^{\ast}} 
+S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right)
+\phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
 \end{equation}
@@
 \noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function.
-The $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
+For example, the $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
 %
@@
 
-\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain. We can represent the approximate series expansion of the function $f$ in the multidimensional input space as,
+\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain (Paul: Could it make sense to move these specific spectral density functions to Section 2.1?).
+We can now write the approximate series expansion of the multivariate function $f$ as,
 %
 \begin{equation}\label{approxf}
-f(\mathbf{x}) \approx \sum_{j} \left( S^{\ast}_{\theta}(\sqrt{\boldsymbol{\lambda^{\ast}_j}})\right)^{1/2} \phi^{\ast}_j(\mathbf{x}) \beta_j, 
-\end{equation}
-
-\noindent where $\beta_j \sim \text{N}(0,1)$ is a random variable with zero mean and unitary variance.
-The computational cost in a multidimensional setting scales $O((n+1)\cdot \prod_{d=1}^D J_d )$, where $n$ is the number of observations and $J_d$ is the number of basis functions considered in the dimension $d$.
+f(\mathbf{x}) \approx \sum_{j=1}^{m^{\ast}} 
+\left( S^{\ast}_{\theta} \left(\sqrt{\lambda^{\ast}_j} \right)\right)^{1/2}
+\phi^{\ast}_j(\mathbf{x}) \beta_j, 
+\end{equation}
+
+\noindent where, again, $\beta_j \sim \text{Normal}(0,1)$ are normal random variables with zero mean and unitary variance.
+The computational cost of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
+Although this still implies linear scaling in $n$, the approximation becomes inefficient more quickly than in the univariate case.
+This is because $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and so increases rapidly with increased dimensionality.
 

### 2019-03-04 Paper/manuscript/manuscript_05.tex
section 3 ready to get feedback

@@
 New developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
-More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.}
+More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.
 
@@
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
-A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(\mathbf{x}-\mathbf{x}')$, which means that the covariance is invariant to translations.
+A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(|\mathbf{x}-\mathbf{x}'|)$, which means that the covariance is invariant to translations.
 Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariacne is both translation and rotation invariant.
-The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which are
+The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which, in the unidimensional case (i.e.
+$\mathbf{x}=x\in \mathbb{R}$, so $\mathbf{r}=r\in \mathbb{R}$), are
 %
 \begin{eqnarray}
-k_{\nu=\infty}(\mathbf{r})&=&\sigma^2 \text{exp}(-\frac{1}{2} \mathbf{r}/\ell^2), \nonumber \\
-k_{\nu=\frac{1}{2}}(\mathbf{r})&=&\sigma^2 exp(-\mathbf{r}/\ell), \nonumber \\
-k_{\nu=\frac{3}{2}}(\mathbf{r})&=&\sigma^2(1+\sqrt{3}\mathbf{r}/\ell \text{exp}(-\sqrt{3}\mathbf{r}/\ell), \nonumber \\
-k_{\nu=\frac{5}{2}}(\mathbf{r})&=&\sigma^2(1+\sqrt{5}\mathbf{r}/\ell+\frac{5}{3}\mathbf{r}^2/\ell^2) \text{exp}(-\sqrt{5}\mathbf{r}/\ell), \nonumber 
+k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r^2/\ell^2), \nonumber \\
+k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
+k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
+k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
 \end{eqnarray}
@@
 The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function.
-These covariance functions can be easily generalized to the use of multidimensional length-scale $\boldsymbol{\ell}$, which basically turns the isotropic covariance function into non-isotropic.
+These covariance functions can be easily generalized to the multidimensional case, with multidimensional inputs $\mathbf{x}\in \mathbb{R}^D$ and multidimensional length-scale $\boldsymbol{\ell}\in \mathbb{R}^D$ (Gabi: We should derive the multidimensional form of the previous kernels, but I have now doubts how is the multidimensional form for the Matern kernels with $\nu$=1/2, $\nu$=3/2 and $\nu$=5/2).
+Using a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.
 
@@
 \begin{eqnarray}
-S_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 \mathbf{w}^2), \nonumber \\
-S_{\nu=\frac{1}{2}}(\mathbf{w})&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \nonumber \\
-S_{\nu=\frac{3}{2}}(\mathbf{w})&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \nonumber \\
-S_{\nu=\frac{5}{2}}(\mathbf{w})&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3}, \nonumber 
-\end{eqnarray}
-
-\noindent where variable $\mathbf{w}$ is in the frequency domain.
+S_{\nu=\infty}(w)&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 w^2), \\
+S_{\nu=\frac{1}{2}}(w)&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \\
+S_{\nu=\frac{3}{2}}(w)&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \\
+S_{\nu=\frac{5}{2}}(w)&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3},  
+\end{eqnarray}
+
+\noindent where variable $w$ is in the frequency domain. The $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
+%
+\begin{eqnarray}
+S^{\ast}_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \text{exp}\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
+%
+S^{\ast}_{\nu}(\mathbf{w})&=& \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
+\end{eqnarray}
+
+\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain.
 
@@
 
-\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function.
-For example, the $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
-%
-\begin{eqnarray}
-S^{\ast}_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \text{exp}\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
-%
-S^{\ast}_{\nu}(\mathbf{w})&=& \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
-\end{eqnarray}
-
-\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain (Paul: Could it make sense to move these specific spectral density functions to Section 2.1?).
+\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section 2.1).
 We can now write the approximate series expansion of the multivariate function $f$ as,
@@
 
-Implementation and performance of the approximate GP model depends on the number of basis functions and box size settings, which are also related each other.
+Implementation and performance of the HSGP model depends on the number of basis functions and boundary settings, which are, at the same time, related  to each other.
 The time of computation in the approximation mainly depends on the number of basis functions.
-At the same time, the appropriate values for the number of basis functions and box size depend on the non-linear functional effects of the function to be learned.
+Furthermore, the appropriate values for the number of basis functions and boundary depend on the non-linear functional effects of the function to be learned.
 In GP functions these non-linear functional effects are characterized by the lengthscale of the covariance function.
 
-In this section we analyze the rule and relationships of the number of basis functions and the box size on approximation and their effects on the learned functions.
-
+In this section we analyze the rule and relationships of the number of basis functions and boundary on the approximation and their effects on the learned functions. 
 We also analyze these two factors in relation to the lengthscale of the functions to be learned.
-We provide a useful graph that retains how these factors, number of basis fucntions, box size and lengthscale, relate each other in relation to the performance of the model.
+We provide a useful graph showing how these factors, number of basis functions, boundary and lengthscale, relate to each other in relation to the performance of the model.
 This graph basically provides the recommended values for these factors in order to get a close approximation.
-This graph also allows us to diagnose an actual fit.
-
-In this section we also make a comparison between regular GP model and approximate GP model fitted for different dataset with different "wigglyness".
+This graph also allows us to diagnose an actual fit by checking whether the value of the estimated lengthscale does not keep to those that the Figure brings as suitable ones.
+
+In this section we also make a comparison between regular GP and HSGP models, fitted for different dataset with different "wigglyness".
 The comparison is made through comparing their estimated lengthscales.
@@
 The approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, e.g.
-in a 1-$D$ input space $x\in \Omega=[-L,L]$:
+in a 1-$D$ input space $x\in \Omega=[-L,L]\subset \mathbb{R}$:
 %
 \begin{equation}\label{diffcov}
-k(\tau) \approx \sum_{j} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
-\end{equation}
-
-\noindent where $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
+k(\tau) = \sum_{j}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
+\end{equation}
+
+\noindent where $L$ is the boundary, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
 The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) for the unidimensional case and those in equation (\ref{bf&lambda_multi}) for the multidimensional case.
@@
 The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
-The number of basis functions can be truncated at some finite positive value $J$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
+The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
 %
 \begin{eqnarray}\label{diff_covs}
-\int k(\tau)d(\tau) - \int \sum_{j}^J S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
-\end{eqnarray}
-
-The finite number $J$ of basis functions in the approximation needed to satisfy eq.
+\int k(\tau)d(\tau) - \int \sum_{j}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
+\end{eqnarray}
+
+The finite number $m$ of basis functions in the approximation needed to satisfy eq.
 (\ref{diff_covs}) will depend on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
-The approximation will also depends on the box size $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
-The box size will also have an effect on the number of basis functions needed in the approximation.
-Note that we refer to the box size as the value $L$ of the Dirichelt boundary condition in eq.
-\ref{eigenproblem}.
-The box size $L$ is set as an extension of the desired predicting input domain $\Psi$.
-Considering a zero-mean input domain $\Psi=[-S,S] \in \mathbb{R}$, the box size is defined as,
+The approximation will also depends on boundary $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
+The boundary will also have an effect on the number of basis functions needed in the approximation.
+The boundary $L$ is set as an extension of the desired predicting input domain $\Psi$.
+Considering a zero-mean input domain $\Psi=[-S,S] \in \mathbb{R}$, the boundary is defined as
 %
 \begin{equation}
-L=c \cdot S
+L=c \cdot S,
 \end{equation}
@@
 From now on we are going to refer to this proportional factor $c$ as the boundary factor of the approximation.
-The boundary factor can also be regarded as the box size $L$ normalized by the half-range of the input space $S$.\\
-
-For a naive illustration of the effects of these two factors, number of basis functions $J$ and box size $L$, on the performance of the approximation, a set of noisy observations are simulated from a GP model with lengthscale $\ell=0.3$ and magnitud $\alpha=1$ for input vales within the input domain with half-range $S=1$.
-Approximate GP models with different number of basis functions and box sizes $L$ ($L=c\cdot S$, with $c\ge 1$) are fitted to this data.
-The lengthscale and magnitud parameters are considered fix to the true values of the generative GP model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} ilustrate the individual effects of the number of basis functions and the box size, respectively, on the approximation of the covariance function.
+The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
+
+For a naive illustration of the effects of these two factors, number of basis functions $m$ and boundary $L$, on the performance of the approximation, a set of noisy observations are drawn from a GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, and input vales within the zero-mean input domain with half-range $S=1$.
+HSGP models with different number of basis functions and boundaries are fitted to this data.
+The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} ilustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function.
 On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function.
@@
 The more "wigglyness" of the function, the more basis functions are needed.
-On the other hand, and similarly, for a fixed number of basis functions $J$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the box size $L$ in approaching the posterior mean and the covariance function.
+On the other hand, and similarly, for a fixed number of basis functions $m$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the boundary $L$ in approaching the posterior mean and the covariance function.
 It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances.
@@
 \subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
-\caption{(left) Posterior distributions for different number of basis functions $J$. (right) Covariance function for different number of basis functions $J$.}
+\caption{(left) Posterior distributions for different number of basis functions $m$. (right) Covariance function for different number of basis functions $m$.}
   \label{fig1_Post_J}
@@
 
-In the previous Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} we focused on illustrating the individual effects of the number of basis functions and the box size.
-Furthermore, for a clearer illustration of these effects, the lengthscale and magnitud parameters in the approximate GP model were fixed to their true values.
+In the previous Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} we focused on illustrating the individual effects of the number of basis functions and the boundary.
+Furthermore, for a clearer illustration of these effects, the lengthscale and marginal variance parameters in the HSGP model were fixed to their true values.
 Now we are going to focus on analyzing the interaction effects between these two factors, number of basis functions and the boundary factor, on the performance of the approximation.
-The lengthscale and magnitud parameters will not be fixed rather they will be estimated in both regular and approximate GP models.
+The lengthscale and marginal variance will not be fixed rather they will be estimated either in regular and HSGP models.
 %The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
 Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
-Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors means and the covariance functions obtained after fitting the data, for different number of basis functions and boundary factors.
-Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model over the regular GP model as a function of the number of basis functions and the boundary factors.
-Figure \ref{fig6_lscale_vs_J} shows the lengthscale and magnitud parameters for the regular GP model and for the approximate models as a function of the number of basis functions and boundary factors.
+Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors means and the covariance functions obtained after fitting the data, for different number of basis functions and boundary factor.
+Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model, as a function of the number of basis functions and the boundary factor.
+Figure \ref{fig6_lscale_vs_J} shows the lengthscale and marginal variance for the regular GP model and for the HSGP models as a function of the number of basis functions and boundary factor.
 Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5.
 The choice of 10 basis functions and a boundary factor of 1.5 could be an accurate enough choice.
-This same conclusion can be also be intuitively seen in the posterior means and covariance functions plots in Figure \ref{fig4_Post_&_Cov_part1}.
-\\
-
-From Figure \ref{fig5_MSE_vs_J}, a behaviour in terms of performance of the approximate GP model as a function of the number of basis functions and the boundary factor, can be deduced: \\
+This same conclusion can be also be intuitively seen in the posterior mean and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}.
+
+From Figure \ref{fig5_MSE_vs_J}, a behaviour in terms of performance of the HSGP model as a function of the number of basis functions and the boundary factor, can be deduced: \\
 $\bullet$ as the boundary factor increases, more basis functions are needed,
 \\
-$\bullet$ as fewer basis functions are used, the boundary factor must decrease. \\
-
-Similar conclusions can be stated from results of Figure \ref{fig6_lscale_vs_J}, This Figure shows the estimated lengthscale and magnitud parameter as function of the number of basis functions and the boundary factor.
-In addition to the above conclusions, it can be easily seen that exists a minimum value for the boundary factor under which a close approximation will never be achieved.
+$\bullet$ as fewer basis functions are used, the boundary factor must decrease.
+
+Similar conclusions can be stated from results of Figure \ref{fig6_lscale_vs_J}.
+This Figure shows the estimated lengthscale and marginal variance as function of the number of basis functions and boundary factor.
+In addition to the above conclusions, in Figure \ref{fig6_lscale_vs_J} it can be easily seen that exists a minimum value for the boundary factor under which a close approximation will never be achieved.
 \\
@@
 \subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
-\caption{Posterior distributions and covariance functions of the proposed approximate GP models and the exact GP model in function of the number of basis functions $J$ and the boundary factor $c$.} 
+\caption{Posterior distributions and covariance functions of the regular GP model and the proposed HSGP model as a function of the number of basis functions $m$ and the boundary factor $c$.} 
   \label{fig4_Post_&_Cov_part1}
@@
 \subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
-\caption{Standarized root mean square error (SRMSE) of the proposed approximate GP models and the exact GP model.
-(left) SRMSE against the number of basis functions $J$ for different values of the boundary factor $c$.
-(right) SRMSE against the boundary factor $c$ for different values of the number of basis functions $J$.
+\caption{Standarized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model.
+(left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$.
+(right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$.
 }
@@
 \subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
-\caption{(left) Estimated lengthscales against the number of basis functions $J$ for different values of the boundary factor $c$.
-(right) Estimated GP magnitud the number of basis functions $J$ for different values of the boundary factor $c$.
-}
+\caption{Estimated lengthscale and marginal variance parameters of the proposed HSGP models and the regular GP model.
+(left) Estimated lengthscales vs the number of basis functions $m$ and for different values of the boundary factor $c$.
+(right) Estimated marginal variance vs the number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig6_lscale_vs_J}
@@
 
-Additionally, there exit a relation of the number of basis functions $J$ and the boundary factor $c$ with the lengthscale $\ell$ of the function.
+Additionally, there exit a relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the function.
 Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in relation to the performance of the approximation evaluated using eq.
@@
 And the Y-axis represents the number of basis functions.
-Thus, this figure give us, for a certain GP function with lengthscale $\ell$ and given a fixed boundary factor $c$, the minimum number of basis functions needed to achieve a close approximation in terms of satisfying eq.
+Thus, this figure give us, for a certain GP function with lengthscale $\ell$ and given a boundary factor $c$, the minimum number of basis functions needed to achieve a close approximation in terms of satisfying eq.
 (\ref{diff_covs}).
@@
 And also this figure could also be read as the minimum functional lengthscales that can be closely approximated given a number of basis functions and a boundary factor.
-This figure basically collects the behaviour of the approximate GP model as a function of the lengthscale, number of basis functions and boundary factor:
+This figure basically collects the behaviour of the HSGP model as a function of the lengthscale, number of basis functions and boundary factor:
 
@@
 $\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimun that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
-On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.\\
-
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $J$ and the boundary factor $c$ to approximate certain process characterized by its lengthscale.
+On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.
+
+Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $m$ and the boundary factor $c$ to approximate certain process characterized by its lengthscale.
 This Figure will allow us to optimize computations which depends on the number of basis functions.
@@
 \subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
-\caption{Relation between the minumum number of basis functions $J$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary condition factor $c$ ($c = \frac{L}{S}$).}
+\caption{Relation between the minumum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
   \label{fig3_lscale_vs_J_vs_c_part1}
@@
 
-If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J} where 10 basis functions and a boundary factor of 1.5 would be enough to closely approximate a function with normalized lengthscale eaual to 0.3, we can recognize that these conclusions also matches the conclusion brought form Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J} where 10 basis functions and a boundary factor of 1.5 would be enough to closely approximate a function with normalized lengthscale equal to 0.3, we can recognize that these conclusions also matches the conclusion brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
 
@@
 
-Finally, we make a comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model in several datasets with different "wigglyness" (or lengthscales).
-From this analysis we can recognize when an approximation is not close enough to the regular GP when the lengthscales are not similar.
-We contrast these results from these lengthscale similarities between the regular and the approximate GP model with the information brought from the Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-We fit the regular GP model and the approximate GP model on different datasets.
-These datasets are drawn from GP prior models with a squared exponential covariance function with different lengthscales for every dataset.
-The approximate GP models is fitted using different number of basis functions.
-The boubdary condition factor is set en ivery fit to a right and optinal choice.
-Figure \ref{fig7_varing_lscale} shows the posterior means of the exact GP model and the approximate GP model with different number of basis functions for every dataset with different wiggly effects (different lengthscale).
+We make a comparative analysis between the lengthscales estimated by the regular GP model and the approximate GP model in several datasets with different "wigglyness" (or lengthscales).
+Basically, comparing the lengthscales we can recognize if an approximation is close enough to the regular GP.
+We contrast these results with the information brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+
+For this analysis, we fit the regular GP model and the approximate GP model on different datasets.
+These datasets consists of noisy draws from a GP prior model with a squared exponential covariance function with different lengthscale.
+Different number of basis functions are used to fit the approximate GP model.
+The boundary factor is set in every case to a right and optimal choice.
+Figure \ref{fig7_varing_lscale} shows the posterior means of the regular GP model and the approximate GP model with different number of basis functions, for every dataset.
 
@@
 \includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
-\caption{Posterior means of the exact and approximate GP models fitted over different datasets with different lengthscales.}
+\caption{Posterior means of the regular GP and approximate GP models fitted over different datasets with different lengthscales.}
   \label{fig7_varing_lscale}
@@
 
-The comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model in several datasets is shown in Figure \ref{fig8_Tlscale_vs_Elscale}.
-The X-axes of the plots represent the true lengthscales of the generative functions for every dataset scenario, and in the Y-axes are represented the estimated lengthscales for the regular GP and the approximate GP models.
-Different comparison are made using different number of basis functions in the approximations.
-Let's assume that the box size parameter was properly set in every case.
-The dashed black line represents the minimum lengthscales that can be closely approximate, given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE of the approximate GP model over the regular GP model for the different datasets with different lengthscale scenarios and the number of basis functions used in the approximation.
-
-From Figure \ref{fig8_Tlscale_vs_Elscale} we can recognize similarities and dissimilarities between the estimation for   the lengthscales of the regular GP model and the approximations, which indicates the performance of the model.
-Figure \ref{fig9_MSE_varing_lscale} with the SRMSE also indicates the performance of the model.
-We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for very small lengthscales and very long lengthscales.
-Figure \ref{fig8_Tlscale_vs_Elscale} shows a correpondence between very low lengthscales when the SRMSE is large, conversely there is no correspondence between very large lengthscale and the SRMSE is close to zero.
-These effects can make sense because, in the case of long lengthscales these have larger variabilities obtaining  similar posteriors, and in the case of low lengthscale these have lower variabilities obtaining diferent posteriors.
-
-The dashed black line represents the minimum lengthscales that can be properly fitted by the approximate GP model according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+The comparative analysis between the lengthscales, estimated by a regular GP model and an approximate GP model in the different datasets with different lengthscales, is shown in Figure \ref{fig8_Tlscale_vs_Elscale}.
+In the X-axes of the plots, the true lengthscales of the data-generative functions for every dataset scenario are represented.
+In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented.
+Different comparisons are made using different number of basis functions in the approximate GP model.
+Let's assume that the boundary factor was properly set in every case.
+The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE of the approximate GP model, computed against the regular GP model, as a function of the lengthscale and number of basis functions.
+
+We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales.
+Very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high SRMSE.
+On the other side, big changes on long lengthscales, which might clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small SRMSE.
 
@@
 \includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
-\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and the number of basis functions.}
+\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and number of basis functions.}
   \label{fig9_MSE_varing_lscale}
@@
 
-A simulated one-dimensional dataset, with 250 observations $f(\mathbf{x}_i)$ ($i=1,\cdots,250$) drawn from a Gaussian process prior with a Matern covariance function with 3/2 degrees of freedom, with inputs $\mathbf{x} \in \{-1,1\} \subset {\rm I\!R}$.
-The hyperparameters of this GP prior with a Matern kernel, marginal variance $\alpha$ and lengthscale $\ell$, are set to 1 and 0.15, respectively.
-The additive Gaussian noise $\sigma$ to these drawn points is set to 0.2.
-The latent Gaussian process model is as follows,
+This example consist of a simulated dataset with 250 draws values from a Gaussian process prior with a Matern covariance function (a Matern with 3/2 degrees of freedom) and inputs $\{\mathbf{x}_i\}_{i=1}^{250} \in \{-1,1\} \subset {\rm I\!R}$.
+The hyperparameters, marginal variance $\alpha$ and lengthscale $\ell$, of this GP prior were set to 1 and 0.15, respectively.
+The Gaussian noise $\sigma$ added to these draws was set to 0.2.
+
+The regular GP model for fitting this simulated dataset $\mathbf{y}$ can be written as follows,
+%
+\begin{eqnarray}\label{eq:latentgp_simudata1}
+\begin{split}
+\mathbf{y} &= f(\mathbf{x}) + \boldsymbol{\epsilon} \\
+\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  \mathbf{I}) \\
+f(\mathbf{x}) &\sim \mathcal{GP}(0, k(\mathbf{x}, \mathbf{x}', \theta)),
+\end{split}
+\end{eqnarray}
+
+\noindent The previous formulation corresponds to the latent form where $f(\mathbf{x})$ represents the functional model underlying the noisy data and modeled by a GP prior with a Matern covariance function $k$ (4).
+$\mathbf{I}$ represents the identity matrix.
+Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where the element $K_{ij}$ is formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j,\theta)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+
+A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using sampling methods such as HMC, would be the marginalized form of the latent GP model in (\ref{eq:latentgp_simudata1}),
+%
+\begin{equation*}\label{eq:marginalizedgp_simudata1}
+\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 \mathbf{I} )
+\end{equation*}
+
+\noindent where the function values $f(\mathbf{x})$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing time of computation and improving the sampling and the effective number of samples.
+
+In the HSGP model, the covariance function $k$ is approximated as (\ref{approxcov}),
+%
+\begin{equation*}
+k(x,x') \approx \sum_{j}^{m}S(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
+\end{equation*}
+
+\noindent with the spectral density $S$ (3),
 %
 \begin{eqnarray*}
-\mathbf{y} &=& f(\mathbf{x}) + \boldsymbol{\epsilon} \\
-\boldsymbol{\epsilon} &\sim& \mathcal{N}(0, \sigma^2  \mathbf{I}) \\
-f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x}, \theta)),
+S(w)&=& 4\alpha^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2},  
 \end{eqnarray*}
 
-\noindent where $\mathbf{I}$ is the identity matrix and $k$ the Matern covariance function in eq. (4). More efficient when using HMC than this latent GP model would be the marginalized model,
-%
-\begin{equation}\label{eq:marginalized_gp}
-\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 \mathbf{I} )
-\end{equation}
-
-\noindent where the element $K_{ij}$ is the covariance function $k(\mathbf{x}_i, \mathbf{x}_j, \theta$) evaluated at the inputs $\mathbf{x}_i$ and $\mathbf{x}_j$.
-
-In the approximate GP model, the covariance function $k$ is approximated as in eq.
-(\ref{approxcov}) with the spectral density in eq.
-(8).
-The latent function $f(\mathbf{x)}$ is approximated as in eq.
-(\ref{approxf}).
-
-We compare the modelling performance over this simulated process of our proposed approach with the performance of a regular GP following the marginalized version in eq.
-(\ref{eq:marginalized_gp}) and a Splines-based model.
-For our basis function approach we use $J=80$ basis functions and a boundary factor $c=1.2$.
-For the Splines-based model we use 80 knots or basis in the model.
-
-Figure \ref{fig9_Posteriors_exI} shows the posteriors distributions of the proposed model and the regular GP model, fitted over the data.
-Posteriors distributions of the proposed approximate GP model, of the exact GP model, and the Splines model are plotted jointly with the true GP prior from the data were drawn.
-The right and left extremes of the data, as well as three small regions around the middle of the data, correspond to out-of-sample or test data which have not been taking part of training the models.
+\noindent and the latent function values $f(\mathbf{x)}$ are approximated as (\ref{approxf}),
+%
+\begin{equation*}
+f(x) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
+\end{equation*}
+
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splinesn approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
+A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
+
+For the HSGP model we use $m=80$ basis functions and a boundary factor $c=1.2$. For the splines model we use 80 knots or basis in the model.
+
+Figure \ref{fig10_Posteriors_exI} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
+The right and left extremes and three small regions around the middle of the plot correspond to out-of-sample or test data which have not been taking part on training the models.
+The test data have been plotted in Figure \ref{fig10_Posteriors_exI} with crosses and the training data with circles.
+The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
 
@@
 \includegraphics[scale=0.35]{fig10_Posteriors_exI.pdf}
-\caption{Posterior distributions of the proposed approximated GP model, the exact GP model, and the Splines model.}
+\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
   \label{fig10_Posteriors_exI}
@@
 
-Figure \ref{fig11_MSE_exI_inter} shows the  standardized mean squared error (SMSE) over the true function for interpolation and extrapolation data.
-As it can be seen the splines model does not extrapolate data properly.
-For interpolation both models show similar performance although the approximate GP model seems to be a bit better.
+Figure \ref{fig11_MSE_exI_inter} shows the  standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
+For this, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been formulated and fitted on this data.
+The SRMSE is computed against the true function.
+Model interpolation has been assessed either using the test data for interpolating and the training data.
+From Figures \ref{fig10_Posteriors_exI} and \ref{fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model, for interpolating and extrapolating data.
+However, the splines model does not extrapolate data properly.
+For interpolation both models show roughly similar performance.
 
@@
 This data set relates the proportion of support for same-sex marriage to the age.
-This proportions correspond to binomial observations per age group $i$.
-The data consist of $i=1,\cdots,74$ binomial observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group.
-
-The observational model for the observed number of people $y_i$ supporting same-age marriage per age group $i$ is a binomial model with the total number of people per age group $n_i$ and the probability per age group $p_i$ as model parameters.
-The total number of people per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ per age group $i$ or mean number of support people per age group $\bar{y}=p_i \cdot n_i$.
+The data consists of 74 binomial observations per age group $i$ ($i=1,\cdots,74$), with the observation of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group.
+
+The observational model for the observed number of people $y_i$ supporting same-age marriage per age group $i$ is a binomial model with the total number of people $n_i$ per age group and the probability $p_i$ per age group as model parameters,
 %
 \begin{equation*}
-y_i \sim \text{Binomial}(p_i, n_i)
+y_i \sim \text{Binomial}(p_i, n_i).
 \end{equation*}
 
-The probability $p_i$ is related through the logit function to a latent Gaussian process prior distribution with the age values $x_i$ as inputs,
+\noindent The total number of people per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ per age group $i$ or mean number of support people per age group $i$.
+The probability $p_i$ is related to the Gaussian process latent values $f(\mathbf{x})$ with age values $\mathbf{x}=\{x_i\}_{i=1}^{74}$ as inputs, through the logit function,
 %
@@
 p_i &=& \text{logit}(f(x_i)) \nonumber \\
-f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x}, \theta))
+f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x},\mathbf{x}', \theta)
 \end{eqnarray*}
 
-\noindent with a squared exponetial covariance function $k$ in eq. (2).
-
-In the approximate GP model the covariance function is approximated as in eq.
-(\ref{approxcov}) with spectral density in eq.
+\noindent with a squared exponetial covariance function $k$ (2).
+
+In the HSGP model the covariance function is approximated as in eq.
+(\ref{approxcov}) with the square exponential spectral density in eq.
 (6), and the latent function $f(\mathbf{x})$ as in eq.
 (\ref{approxf}).
-As comparison a splines model is also fitted.
-Figure \ref{fig12_Posteriors_gaydata} shows the posterior distributions of the three models.
-In the case of the approximate GP model the posterior shown in the figure corresponds to 20 basis functions and a boundary factor of 1.5.
+
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
+A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
+
+For the HSGP model we use $m=20$ basis functions and a boundary factor $c=1.5$. For the splines model we use 20 knots or basis in the model.
+
+Figure \ref{fig12_Posteriors_gaydata} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
+The HSGP model posterior shown in the figure corresponds to 20 basis functions and a boundary factor of 1.5.
 For the spline posterior shown in the figure, it correspond to a 20 knots.
-Two small sets of observations have been used as test data.
+Two small sets of observations have been used as test data, plotted as crosses in the Figure.
 
@@
 \includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
-\caption{Posterior distributions of the proposed approximated GP model, the exact GP model, and the Splines model.}
+\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
   \label{fig12_Posteriors_gaydata}
@@
 
-Figure \ref{fig13_MSE_train_BF_gaydata} shows the MSE against the regular GP model of the approximate GP model for the different number of basis functions and boundary factor.
-The expected patterns of the approximation in function of the number of basis function and boundary factor are recognized.
+For the HSGP model, different models with different number of basis functions and boundary factors have been used.
+The root mean square errors (RMSE) for every one of these models have been computed and plotted as a function of the number of basis functions and the boundary factor in Figure \ref{fig13_MSE_train_BF_gaydata}, for training (left) and test (right) data.
+In this case, the RMSE has been computed against the regular GP model.
+The expected patterns of the approximation in function of the number of basis function and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
 
@@
 
-Figure \ref{fig14_MSE_train_gaydata} shows the MSE of the different models against the actual data, for training set of data and the test set of data.
+Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the different models, regular GP, HSGP and splines, against the actual data, for training and test data.
+Again we can see how the splines models does not extrapolate data properly.
 

### 2019-03-04 Paper/manuscript/manuscript_05.tex
spell check

@@
 This method decomposes the kernel into basis functions which represent the GP as a linear model and thus makes its estimation considerably faster.
-In principal, we are free in the choice of basis functions, but the Laplace eigenfunctions were found to be particularily appealing \citep{solin2018hilbert}.
+In principal, we are free in the choice of basis functions, but the Laplace eigenfunctions were found to be particularly appealing \citep{solin2018hilbert}.
 Not only can they be computed analytically, but they are also independent of the particular choice of the covariance kernel including the hyperparameters.
@@
 
-The remainder of the paper is structered as follows.
+The remainder of the paper is structured as follows.
 In Section 2, we introduce GPs and their reduced rank approximations proposed by \cite{solin2018hilbert}.
@@
 
-\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ are formed by evalating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ are formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
 The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
@@
 binomial or multinomial classification, Poisson  regression, and many other modelling problems) and the posterior predictive distribution $p(f^*| \mathbf{y})$ is analytically untractable and can not be obtained in a closed-form expression.
-Approximate inference have to be used to tackle the intractable inference, tipically based on sampling methods such as MCMC \citep{brooks_2011} or analytical Gaussian approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational methods \citep{gibbs2000variational,csato2000efficient}.
+Approximate inference have to be used to tackle the intractable inference, typically based on sampling methods such as MCMC \citep{brooks_2011} or analytical Gaussian approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational methods \citep{gibbs2000variational,csato2000efficient}.
 Furthermore, Gaussian processes in a direct implementation scales cubically in the number of sample points, $N$, which effectively limits the modelling inference for medium and large datasets.
-In the introduction section in this paper several methods tipically used for dealing with the computational complexity has been commented.
+In the introduction section in this paper several methods typically used for dealing with the computational complexity has been commented.
 \\
@@
 A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(|\mathbf{x}-\mathbf{x}'|)$, which means that the covariance is invariant to translations.
-Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariacne is both translation and rotation invariant.
+Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariance is both translation and rotation invariant.
 The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which, in the unidimensional case (i.e.
@@
 
-\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the lengt-scale and magnitud, respectively, of the kernel.
+\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the length-scale and magnitude, respectively, of the kernel.
 The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function.
@@
 We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition.
-As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuarcy of HSGPs.
+As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs.
 We will come back to this issue in Section 3.
@@
 $\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
+In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
 
@@
 \vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.
+$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
 
@@
 $\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitud parameter.
+However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
 In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
@@
 \vspace{2mm}
-$\bullet$ In addition, the computation of the automatic differentation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ In a regular GP model the automatic differentation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
+$\bullet$ In addition, the computation of the automatic differentiation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.
+
+\vspace{2mm}
+$\bullet$ In a regular GP model the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
 This operation has to be computed at every step of the HMC.
@@
 \vspace{2mm}
-$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.
+$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
 
@@
 The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} ilustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function.
 On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function.
@@
 \subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
-\caption{Standarized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model.
+\caption{Standardized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model.
 (left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$.
@@
 
-$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimun that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
+$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimum that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
 On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.
@@
 \subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
-\caption{Relation between the minumum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
+\caption{Relation between the minimum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
   \label{fig3_lscale_vs_J_vs_c_part1}
@@
 \includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
-\caption{True lengthscales versus estimated lscales of the different realizations of datasets using different number of basis functions. }
+\caption{True lengthscales versus estimated lengthscales of the different realizations of datasets using different number of basis functions. }
   \label{fig8_Tlscale_vs_Elscale}
@@
 
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splinesn approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
 A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
@@
 \subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
-\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
+\caption{Standardized mean square error (SMSE) of the different methods against the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
   \label{fig11_MSE_exI_inter}
@@
 
-\noindent with a squared exponetial covariance function $k$ (2).
+\noindent with a squared exponential covariance function $k$ (2).
 
@@
 \subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
+\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
   \label{fig13_MSE_train_BF_gaydata}
@@
 \subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
+\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
   \label{fig14_MSE_train_gaydata}
@@
 The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix.
-The SoR method is based on a degenarete prior which produces unresoanble predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).
+The SoR method is based on a degenerate prior which produces unreasonable predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).
 
@@
 \vspace{2mm}
-$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the functions.
-
-\vspace{2mm}
-$\bullet$ We make recomendations for the values of these factors based on the recognized relations among them.
+$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthscale of the functions.
+
+\vspace{2mm}
+$\bullet$ We make recommendations for the values of these factors based on the recognized relations among them.
 We provide useful graphs of these relations that will help the users to improve performance and save time of computation.
@@
 
-Assuming that the espectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
+Assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
 %

### 2019-03-04 Paper/manuscript/manuscript_05.tex
edited intro

@@
 
-Gaussian processes (GPs) are natural (Paul: what does "natural" imply and as compared to what?) and flexible non-parametric models to define probability distributions over multi-dimensional functions \citep{rasmussen2006gaussian,neal1997monte}.
-This allows applying probabilistic inference to estimate functional relationships between one or more covariates and a single response variable.
-The defining feature of GPs is that the function values are assumed to jointly follow a multivariate normal distribution, which is characterized primarily by a covariance function, sometimes also referred to as covariance kernel.
+Gaussian processes (GPs) are flexible define probability distributions over multi-dimensional non-linear functions \citep{rasmussen2006gaussian,neal1997monte}.
+GPs are defind by a mean and covariance function and any finite set of function values are jointly distributed as a multivariate normal.
 The covariance function encodes our prior assumptions about the functional relationship, such as continuity, smoothness, periodicity and scale properties.
 GPs not only allow for non-linear effects but can also implicitly handle interactions between covariates.
-Several types of covariance functions with varying properties can be used, and they may even be combined for further increased flexibility.
-GPs are often referred to as non-parametric models because the number of parameters is not fixed, but rather increases with the number of data points, which allows to adapt model complexity to the data.
-As such, the term 'non-parametric' does not imply having no parameters but actually having a lot of them.
+Different types of covariance functions can be used, and they may be combined for further increased flexibility.
 Due to their generality and flexibility, GPs are of broad interest across machine learning and statistics \citep{rasmussen2006gaussian,neal1997monte}.
@@
 
-One of the main limitations of exact GPs is that computational demands and memory requirements scale as $O(n^3)$ and $O(n^2)$, respectively, with the number $n$ of observations in the data used to fit the model.
-This effectively limits their application to rather small data sets of a few thousand observations at most.
-This problem becomes especially severe when performing full Bayesian inference via sampling methods, where in each sampling step we need to invert the Gram matrix of the covariance function, usually through Cholesky factorization.
+Given $n$ observations in the data, the computational cost and memory requirements of exact GP implementation in general scales as $O(n^3)$ and $O(n^2)$, respectively.
+This would limit their application to rather small data sets of a few tens of thousands observations at most.
+The problem becomes especially severe when performing full Bayesian inference via sampling methods, where in each sampling step we have $O(n^3)$ computational cost when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
 To alleviate these computational demands, several approximate methods have been proposed, which we will briefly review in the following.
 
-Sparse GPs are based on building low-rank approximations of the covariance matrix of the complete data by means of selecting a subset $m$ of training points on which to base computation.
-They reduce the dimension of the posterior distribution thereby reducing the memory requirements to $O(nm)$ and computational complexity to $O(nm^2)$ for some $m < n$.
-This approach still requires Cholesky factorization although now based on a smaller covariance matrix.
-While computational complexity is reduced considerably, ensuring convergence to the exact GP model is not necessarily straightforward (Paul: this may need a citation).
-As such, reducing the amount of data while maintaining the correct posterior distribution is not a simple procedure.
-A unifying view on sparse GPs based on approximate generative methods is provided in \cite{quinonero2005unifying}, while a general review can be found in \cite{rasmussen2006gaussian}.
-New developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
+Sparse GPs are based on low-rank approximations of the covariance matrix.
+The low-rank approximation with $m \ll n$ {\it inducing points} can be presented with reduced memory cost $O(nm)$ and the corresponding computational complexity is $O(nm^2)$.
+% While computational complexity is reduced considerably, ensuring convergence to the exact GP model is not necessarily straightforward (Paul: this may need a citation).
+A unifying view on sparse GPs based on approximate generative methods
+is provided in \cite{quinonero2005unifying}, while a general review
+can be found in \cite{rasmussen2006gaussian}.
+% More recent developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
 More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.
 
-(Paul: This whole paragraph is unclear to me.
+% (Paul: This whole paragraph is unclear to me.
 Are all of these approaches approximations via splines, or are the splines just special cases of a more general method?
 In any case, I think the paragraph needs substantial revision).
-Other global approach make use of the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
-It is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a specific choice of covariance function, but they do not generally have correspondence with GP models, and they do not have the flexibility of choosing different covariance functions as model structure.
-The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving,gal2015improving}.
+Alternative low-rank approximation is based on forming a basis function approximation with $m \ll$.
+The basis functions are presented explicitly, but could also be used to form a low rank covariance matrix approximation.
+Common basis function  approximations are based on the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
+Certain spline smoothing basis functions are equivalent to GPs with certain covariance function \citep{wahba1990spline}, but not all spline models are GPs (sometimes splines are not marginalization consistent)-
+%
+The sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving,gal2015improving}.
 Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat´ern class of kernels.
-Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels.
-
-Yet another approach to approximate GPs -- on which we will focus in the present paper -- was proposed by \cite{solin2018hilbert} who suggested to use reduced-rank approximations of the covariance kernel.
-This method decomposes the kernel into basis functions which represent the GP as a linear model and thus makes its estimation considerably faster.
-In principal, we are free in the choice of basis functions, but the Laplace eigenfunctions were found to be particularly appealing \citep{solin2018hilbert}.
-Not only can they be computed analytically, but they are also independent of the particular choice of the covariance kernel including the hyperparameters.
-The latter comes with several advantages when performing inference via algorithms that require differentiation of the posterior density (Paul: cite something here).
-In a fully Bayesian framework based on sampling methods, the proposed approximate GP model has a computational complexity of just $O(nm+m)$, with $m<<n$ in every sampling step.
-The posterior distribution of the reduced-rank GP approximation is of dimension $m$ and thus considerably smaller than in the case of exact GPs.
-This not only lowers the memory requirements and computational complexity but also reduces the correlations between latent values and the corresponding hyperparameters of the GP, which can help to further improve sampling efficiency.
-
-While \cite{solin2018hilbert} fully developed the mathematical theory behind reduced-rank approximations of GPs, they did not put much effort in analyzing the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
-However, for practical implementations of approximate GPs, the latter factors are critical and thus require further investigation.
-In the present paper, we will fill this gap by providing practical recommendations for the choice of these factors based on the recognized relations among them.
-What is more, we provide intuitive visualizations of these relations that will help users to improve performance and save computation time, while at the same time maintaining close approximations of exact GPs.
-
-Although there are several GP specific software packages available to date (Paul: cite those.
-Gabi: This full paragraph comes from Aki.
-I think he refers to some Splines methods and maybe others approaches as well), each provide efficient implementations only for a restricted range of GP based models.
-In this paper we do not focus on the fastest possible inference for a specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
-To this end, we want to make full Bayesian inference over Gaussian process models accessible to a broader audience by making them easy, fast and save to apply.
+Random fourier features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels.
+
+In this paper we focus on Laplace eigenfunction basis function approximation for stationary covariance functions proposed by \citet{solin2018hilbert}.
+Basis function approaches are computationally as linear models, which is attractive property in modular probabilistic programming models where there is a big benefit id approximation specific computation is simple.
+The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance kernel including the hyperparameters.
+While the pre-computation cost of the basis functions is $O(m^2n)$ the computation cost when only the covariance function parameters change is $O(mn+m)$ (or $O(m^3)$ if marginalization can be used).
+This is a big advantage for iterative algorithms such as Markov chain Monte Carlo (MCMC).
+The further advantage is the reduced memory cost in automatic differentation used in modern probabilistic programming frameworks, as the automatic differentation memory cost scales as the computational cost and not as the usual memory cost for the posterior density computation.
+The basis function approach provides also easy way to use non-centered parameterization reducing posterior dependency between parameters presenting the function and the covariance function parameters, which helps to further improve MCMC efficiency.
+
+While \citet{solin2018hilbert} fully developed the mathematical theory behind the specific reduced-rank approximation, for practical implementations of approximate GPs in probabilistic programming framework further work is needed.
+In this paper we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
+We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.
+
+Although there are several GP specific software packages available to date (GPML, GPstuff, GPy, GPflow), each provide efficient implementations only for a restricted range of GP based models.
+In this paper we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
 

### 2019-03-04 Paper/manuscript/manuscript_05.tex
fix inpt encoding problems + microtype

@@
 % v1.05 - August 2017
-
 \documentclass[]{interact}
+\usepackage[utf8]{inputenc}
+\usepackage{times}
+\usepackage{microtype}
 
@@
 The sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving,gal2015improving}.
-Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat´ern class of kernels.
+Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels.
 Random fourier features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels.
@@
 Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods.
-From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modiﬁcations to the so-called Nystr\"om method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum.
+From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modfications to the so-called Nystr{\"o}m method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum.
 These methods are basically based on choosing a set of $m$ inducing inputs $x_u$ and scaling the corresponding eigendecomposition of their corresponding covariance matrix $K_{u,u}$ to match that of the actual covariance.
@@
 
-Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat´ern class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure.
+Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure.
 They combine the variational methodology with Fourier based approximations.

### 2019-03-05 Paper/manuscript/manuscript_05.tex
More changes to sec 2.1

@@
 
-In this work, Gaussian processes will take the role of prior distributions over function spaces for non-parametric regression in a Bayesian setting. 
+In this work, Gaussian processes will take the role of a prior distribution over function spaces for non-parametric regression in a Bayesian setting. 
 Consider a data set $\mathcal{D} = \left\lbrace \mathbf{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is a noisy observation of an unknown function $f$ at $\mathbf{x}_n \in \mathbb{R}^D$ , i.e.
@@
 
-The mean and covariance functions completely characterize the Gaussian process prior, and represent how the random functions behave on average and how the different points in the input space co-vary with respect to each other, respectively.
-The Gaussian process prior can be interpreted such as any finite number of function values $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is multivariate Gaussian,
+The mean and covariance functions completely characterize the Gaussian process prior, and control the a priori behavior of the function $f$.
+Let $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\mathbf{f}$ is a multivariate Gaussian distribution
+%
 \begin{align*}
-\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu},K),
+\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu}, \mathbf{K}),
 \end{align*}
 
-\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ are formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\mathbf{K}$ the covariance matrix, where the element $K_{i,j}$ are formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+The covariance function $k(\textbf{x}, \textbf{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
 The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
@@
 \begin{eqnarray*}
-p(\mathbf{f},\mathbf{f}')=\mathcal{N} \left( \left[ \begin{array}{cc}
+p(\mathbf{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
 \mathbf{f} \\ 
-\mathbf{f}'
+f^*
 \end{array} \right] | 0,\left[ \begin{array}{cc}
-K_{\mathbf{f},\mathbf{f}} & K_{\mathbf{f},f^*} \\ 
-K_{f^*,\mathbf{f}} & K_{f^*,f^*}
+\mathbf{K}_{\mathbf{f},\mathbf{f}} & \mathbf{k}_{\mathbf{f},f^*} \\ 
+\mathbf{k}_{f^*,\mathbf{f}} & k_{f^*,f^*}
 \end{array} \right] \right),
@@
 
-\noindent where $K_{\mathbf{f},f^*}$ defines the covariances between $\mathbf{f}$ and $f^*$, and $K_{f^*,f^*}$ defines the variance of $f^*$.
-By using the conditioning properties of a multivariate Gaussian, we can compute analytically the predictive distribution for $f^*$  given $\mathbf{f}$,
-\begin{align*}
-	p(f^*| \mathbf{f}) = \mathcal{N}(f^*| K_{f^*,\mathbf{f}} K_{\mathbf{f},\mathbf{f}}^{-1} \mathbf{f},  K_{f^*,f^*}-K_{f^*,\mathbf{f}} K_{\mathbf{f},\mathbf{f}}^{-1} K_{\mathbf{f},f^*}  )
-\end{align*}
-
-The joint distribution of observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
+\noindent where $\mathbf{k}_{\mathbf{f},f^*}$ is the covariance between $\mathbf{f}$ and $f^*$, and $k_{f^*,f^*}$ is the prior variance of $f^*$.
+%By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\mathbf{f}$ analytically,
+% \begin{align*}
+	% p(f^*| \mathbf{f}) = \mathcal{N}(f^*| \mathbf{k}_{f^*,\mathbf{f}} \mathbf{K}_{\mathbf{f},\mathbf{f}}^{-1} \mathbf{f},  k_{f^*,f^*}-\mathbf{k}_{f^*,\mathbf{f}} \mathbf{K}_{\mathbf{f},\mathbf{f}}^{-1} \mathbf{k}_{\mathbf{f},f^*}  )
+% \end{align*}
+The joint distribution of the observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
 \begin{align*}
@@
 
-Integrating the joint distribution  $p(\mathbf{y}, \mathbf{f}, f^*)$ over the uncertainty of $\mathbf{f}$ and then conditioning on the observations $\mathbf{y}$ following the Bayes's rule yields the posterior predictive distribution for $f^*$ given the observations $\mathbf{y}$,
-%
+By marginalizing over $\mathbf{f}$ and conditioning on vector of observations $\mathbf{y}$, we obtain the posterior distribution of interest
+%
+\begin{align}\label{eq:posterior}
+	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})},
+\end{align}
+where $p(\mathbf{y})$ is the marginal likelihood and is given by
+\begin{align}\label{eq:evidence}
+	p(\textbf{y}) = \int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*) \text{d}\bm{f}\text{d}f^*.
+\end{align}
+%
+If the observation model $p(\mathbf{y}|\mathbf{f})$ is Gaussian, both integrals in eq.
+\eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters.
+For example, an isotropic Gaussian likelihood yields the following closed-form solution
 \begin{align*}
-	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})}
+	p(f^*| \mathbf{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\mathbf{f}} (\textbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{f},  k_{f^*,f^*}-\mathbf{k}_{f^*,\mathbf{f}} (\mathbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{k}_{\mathbf{f},f^*}  ),
 \end{align*}
-
-If the observational model $p(\mathbf{y}|\mathbf{f})$ is Gaussian, the integrals over $\mathbf{f}$ can be computed analytically yielding,
-\begin{align*}
-	p(f^*| \mathbf{y}) = \mathcal{N}(f^*| K_{f^*,\mathbf{f}} (K_{\mathbf{f},\mathbf{f}} + \sigma^2I)^{-1} \mathbf{f},  K_{f^*,f^*}-K_{f^*,\mathbf{f}} (K_{\mathbf{f},\mathbf{f}} + \sigma^2I)^{-1} K_{\mathbf{f},f^*}  )
-\end{align*}
-
-In most modelling problems, however, the observational model is not Gaussian (e.g.
-binomial or multinomial classification, Poisson  regression, and many other modelling problems) and the posterior predictive distribution $p(f^*| \mathbf{y})$ is analytically untractable and can not be obtained in a closed-form expression.
-Approximate inference have to be used to tackle the intractable inference, typically based on sampling methods such as MCMC \citep{brooks_2011} or analytical Gaussian approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational methods \citep{gibbs2000variational,csato2000efficient}.
-Furthermore, Gaussian processes in a direct implementation scales cubically in the number of sample points, $N$, which effectively limits the modelling inference for medium and large datasets.
-In the introduction section in this paper several methods typically used for dealing with the computational complexity has been commented.
+where $\sigma^2$ is the noise variance.
+
+While Gaussian processes provide an elegant framework for Bayesian non-parametric regression, there are two problems that often arise in practice. 
+in many modelling problems, the observation model is non-Gaussian (e.g.
+Binomial or Poisson regression) and therefore, the posterior distribution $p(f^*| \mathbf{y})$ is analytically intractable and cannot be obtained in closed-form.
+Furthermore, it is often of interest to impose a prior distribution on one or more hyperparameters and then marginalize over the uncertainty.
+This renders the posterior distribution intractable even in the case of a Gaussian likelihood.
+Approximate inference have to be used to tackle the intractable inference problem, typically based on sampling methods such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011} or distributional approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient}.
+Secondly, due to the matrix inverse in eq.~\eqref{eq:posterior} a direct implementation of Gaussian processes scales cubicly in the number of data points, $\mathcal{O}\left(N^3\right)$,  which can be prohibitively slow for many problems of interest.
 \\
-
-In this paper we proposed an approximate model for Gaussian processes based on basis functions expansion which basically turns the GP prior model into a linear model.
-This linear GP model representation makes inference considerably faster and can be used as latent function in non-Gaussian observational models allowing modelling flexibility.\\
-
+\\
+In the following section, we propose an approximate framework for fast and accurate inference for Gaussian processes.
+Using a basis function expansion, we approximate the Gaussian process with a linear model.
+This representation has three main advantages: 1) it makes inference considerably faster due to the linear structure, 2) it is simple to implement, which makes it easy to use Gaussian processes as building blocks in more complicated models and can be used as latent function in non-Gaussian observational models allowing modelling flexibility, 3) it can be made arbitrary accurate and the trade-off between computational complexity and approximation accuracy can easily be controlled.
+\\
+\\
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.

### 2019-03-05 Paper/manuscript/manuscript_05.tex
edits in section 2

@@
 $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
-Gaussian processes have the defining property that the marginal distribution of every finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.\\
+Gaussian processes have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.\\
 
@@
 
-\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\mathbf{K}$ the covariance matrix, where the element $K_{i,j}$ are formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\mathbf{K}$ the covariance matrix, where $K_{i,j}=k(\mathbf{x}_i,\mathbf{x}_j)$.
 The covariance function $k(\textbf{x}, \textbf{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
@@
 %
-\begin{eqnarray*}
+\begin{align*}
 p(\mathbf{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
@@
 f^*
-\end{array} \right] | 0,\left[ \begin{array}{cc}
+\end{array} \right] \,\middle|\, \mathbf{0},\left[ \begin{array}{cc}
 \mathbf{K}_{\mathbf{f},\mathbf{f}} & \mathbf{k}_{\mathbf{f},f^*} \\ 
@@
 \end{array} \right] \right),
-\end{eqnarray*}
+\end{align*}
 
@@
 % \end{align*}
+
+{\color{red} Aki: the following until ``While Gaussian'' seems too basic to be in the paper:}
+
 The joint distribution of the observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
@@
 %
-\begin{eqnarray}
-k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r^2/\ell^2), \nonumber \\
-k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
-k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
-k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
-\end{eqnarray}
+\begin{align}
+k_{\nu=\infty}(r)&=\sigma^2 \exp(-\frac{1}{2} r^2/\ell^2), \nonumber \\
+k_{\nu=\frac{1}{2}}(r)&=\sigma^2 \exp(-r/\ell), \nonumber \\
+k_{\nu=\frac{3}{2}}(r)&=\sigma^2(1+\sqrt{3}r/\ell) \exp(-\sqrt{3}r/\ell), \nonumber \\
+k_{\nu=\frac{5}{2}}(r)&=\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \exp(-\sqrt{5}r/\ell), \nonumber 
+\end{align}
 
@@
 %
-\begin{eqnarray}
-S_{\nu=\infty}(w)&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 w^2), \\
-S_{\nu=\frac{1}{2}}(w)&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \\
-S_{\nu=\frac{3}{2}}(w)&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \\
-S_{\nu=\frac{5}{2}}(w)&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3},  
-\end{eqnarray}
+\begin{align}
+S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right), \\
+S_{\nu=\frac{1}{2}}(w)&= 2\sigma^2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1}, \\
+S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}, \\
+S_{\nu=\frac{5}{2}}(w)&= \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}\left(\frac{5}{\ell^2} + \omega^2\right)^{-3},  
+\end{align}
 
@@
 %
-\begin{eqnarray}
-S^{\ast}_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \text{exp}\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
-%
-S^{\ast}_{\nu}(\mathbf{w})&=& \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
-\end{eqnarray}
+\begin{align}
+S^{\ast}_{\nu=\infty}(\mathbf{w})&= \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \exp\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
+%
+S^{\ast}_{\nu}(\mathbf{w})&= \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
+\end{align}
 
@@
 
-\noindent As a result, the GP model for the function $f$ can be re-defined as
+\noindent As a result, the model for $f$ can be written as
 %
 \begin{equation}
-\mathbf{f} \sim \text{GP}(0,\Phi \Delta \Phi^\intercal). \nonumber
+\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu},\Phi \Delta \Phi^\intercal). \nonumber
 \end{equation}
@@
 \begin{equation}\label{approxf}
-f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
-\end{equation}
-
-\noindent where $\beta_j \sim \text{Normal}(0,1)$ are normal random variables with zero mean and unitary variance.
-That is, the function $f$ can be approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
+f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j,
+\end{equation}
+
+\noindent where $\beta_j \sim \text{Normal}(0,1)$.
+Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
 A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the covariance hyperparameters $\theta$.
@@
 
+The parameterization in \ref{approxf} is naturally in the non-centered
+parameterization form with independent prior distribution on
+$\beta_j$, which makes the posterior inference easier.
+
 \subsection{Generalization to multidimensional GPs}
 
-We will now show how to generalize the results from the previous section to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions.
+The results from the previous section can be generalizes to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions.
 In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is possible combinations of univariate eigenfunctions over all dimensions.
@@
 Let $\mathbb{S}\in {\rm I\!N}^{m^{\ast} \times D}$ be the matrix of all those $D$-tuples.
-For example, suppose we have $D=3$ covariates and use $m_{1}=2$, $m_{2}=2$ and $m_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimension, respectively.
+For example, suppose we have $D=3$ dimensions and use $m_{1}=2$, $m_{2}=2$ and $m_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimension, respectively.
 Then, the number of multivariate eigenfunctions and eigenvalues is $m^{\ast} = m_{1} \cdot m_{2} \cdot m_{3} = 12$ and the matrix $\mathbb{S}\in {\rm I\!N}^{12 \times 3}$ is given by
-
+\begin{align}\small
+\mathbb{S}=
+\left[ {\begin{array}{ccc}
+1 & 1 & 1 \nonumber \\
+1 & 1 & 2 \\
+1 & 1 & 3 \\
+1 & 2 & 1 \\
+1 & 2 & 2 \\
+1 & 2 & 3 \\
+2 & 1 & 1 \\
+2 & 1 & 2 \\
+2 & 1 & 3 \\
+2 & 2 & 1 \\
+2 & 2 & 2 \\
+2 & 2 & 3 
+\end{array} } \right]
+\end{align}
+
+Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\lambda^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
+Thus, for $\mathbf{x}^D \in \Omega$ and $j=1,\ldots,m$, we have
+%
+\begin{align} \label{bf&lambda_multi}
+\begin{split}
+\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \\
+%
+\lambda^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. 
+\end{split}
+\end{align}
+
+\noindent The approximate covariance function is then represented as
+%
+\begin{equation}\label{approxcov_multi}
+k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{m^{\ast}} 
+S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right)
+\phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
+\end{equation}
+
+\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section 2.1).
+We can now write the approximate series expansion of the multivariate function $f$ as,
+%
+\begin{equation}\label{approxf}
+f(\mathbf{x}) \approx \sum_{j=1}^{m^{\ast}} 
+\left( S^{\ast}_{\theta} \left(\sqrt{\lambda^{\ast}_j} \right)\right)^{1/2}
+\phi^{\ast}_j(\mathbf{x}) \beta_j, 
+\end{equation}
+
+\noindent where, again, $\beta_j \sim \text{Normal}(0,1)$.
+The computational cost of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
+Although this still implies linear scaling in $n$, the approximation is more costly than in the univariate case, as $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and grows exponentially with respect to the number of  dimensions.
+
+\subsection{Learning hyperparameters and model inference}
+
+\vspace{2mm}
+$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a linear model.
+
+\vspace{2mm}
+$\bullet$ The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.
+
+\vspace{2mm}
+$\bullet$ The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands.
+The weights depend on the hyperparameters, then they have to be computed in every step of the HMC sampling method.
+
+\vspace{2mm}
+$\bullet$ The linear model is computed with complexity $O(nm)$, computed in every step of the HMC sampling method.
+
+\vspace{2mm}
+$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
+In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
+
+\vspace{2mm}
+$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
+After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
+(Arno's paper, pag.
+7)
+
+\vspace{2mm}
+$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
+especially when sampling methods for inference are used.
+GP prior as latent functions is needed in generalized models.
+
+In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
+
+\vspace{2mm}
+$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
+This operation has to be computed at every step of the HMC or optimizer.
+
+\vspace{2mm}
+$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
+It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
+
+\vspace{2mm}
+$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
+
+\vspace{2mm}
+$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
+However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
+In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
+
+\vspace{2mm}
+$\bullet$ In addition, the computation of the automatic differentiation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.
+
+\vspace{2mm}
+$\bullet$ In a regular GP model the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
+This operation has to be computed at every step of the HMC.
+
+\vspace{2mm}
+$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
+
+\vspace{2mm}
+$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
+
+\section{The accuracy of the approximation}
+
+Implementation and performance of the HSGP model depends on the number of basis functions and boundary settings, which are, at the same time, related  to each other.
+The time of computation in the approximation mainly depends on the number of basis functions.
+Furthermore, the appropriate values for the number of basis functions and boundary depend on the non-linear functional effects of the function to be learned.
+In GP functions these non-linear functional effects are characterized by the lengthscale of the covariance function.
+
+In this section we analyze the rule and relationships of the number of basis functions and boundary on the approximation and their effects on the learned functions. 
+We also analyze these two factors in relation to the lengthscale of the functions to be learned.
+We provide a useful graph showing how these factors, number of basis functions, boundary and lengthscale, relate to each other in relation to the performance of the model.
+This graph basically provides the recommended values for these factors in order to get a close approximation.
+This graph also allows us to diagnose an actual fit by checking whether the value of the estimated lengthscale does not keep to those that the Figure brings as suitable ones.
+
+In this section we also make a comparison between regular GP and HSGP models, fitted for different dataset with different "wigglyness".
+The comparison is made through comparing their estimated lengthscales.
+
+\subsection{Dependency on the number of basis functions and the boundary condition}
+
+The approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, e.g.
+in a 1-$D$ input space $x\in \Omega=[-L,L]\subset \mathbb{R}$:
+%
+\begin{equation}\label{diffcov}
+k(\tau) = \sum_{j}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
+\end{equation}
+
+\noindent where $L$ is the boundary, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) for the unidimensional case and those in equation (\ref{bf&lambda_multi}) for the multidimensional case.
+
+The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
+The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
+%
+\begin{eqnarray}\label{diff_covs}
+\int k(\tau)d(\tau) - \int \sum_{j}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
+\end{eqnarray}
+
+The finite number $m$ of basis functions in the approximation needed to satisfy eq.
+(\ref{diff_covs}) will depend on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
+The approximation will also depends on boundary $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
+The boundary will also have an effect on the number of basis functions needed in the approximation.
+The boundary $L$ is set as an extension of the desired predicting input domain $\Psi$.
+Considering a zero-mean input domain $\Psi=[-S,S] \in \mathbb{R}$, the boundary is defined as
+%
+\begin{equation}
+L=c \cdot S,
+\end{equation}
+
+\noindent where $S$ represents the half-range of the input space, and $c$ ($c \ge 1$) is the proportional extension factor.
+From now on we are going to refer to this proportional factor $c$ as the boundary factor of the approximation.
+The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
+
+For a naive illustration of the effects of these two factors, number of basis functions $m$ and boundary $L$, on the performance of the approximation, a set of noisy observations are drawn from a GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, and input vales within the zero-mean input domain with half-range $S=1$.
+HSGP models with different number of basis functions and boundaries are fitted to this data.
+The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function.
+On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function.
+It can be seen how the number of basis functions affects the non-linear effects of the posteriors.
+Fewer basis functions implies smoother covariance functions and consequently smoother posteriors.
+The more "wigglyness" of the function, the more basis functions are needed.
+On the other hand, and similarly, for a fixed number of basis functions $m$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the boundary $L$ in approaching the posterior mean and the covariance function.
+It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances.
+\\
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
+\caption{(left) Posterior distributions for different number of basis functions $m$. (right) Covariance function for different number of basis functions $m$.}
+  \label{fig1_Post_J}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.43]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.43]{fig2_Cov_L.pdf}}
+\caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
+  \label{fig2_Post_L}
+\end{figure}
+
+In the previous Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} we focused on illustrating the individual effects of the number of basis functions and the boundary.
+Furthermore, for a clearer illustration of these effects, the lengthscale and marginal variance parameters in the HSGP model were fixed to their true values.
+Now we are going to focus on analyzing the interaction effects between these two factors, number of basis functions and the boundary factor, on the performance of the approximation.
+The lengthscale and marginal variance will not be fixed rather they will be estimated either in regular and HSGP models.
+%The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
+Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
+Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors means and the covariance functions obtained after fitting the data, for different number of basis functions and boundary factor.
+Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model, as a function of the number of basis functions and the boundary factor.
+Figure \ref{fig6_lscale_vs_J} shows the lengthscale and marginal variance for the regular GP model and for the HSGP models as a function of the number of basis functions and boundary factor.
+Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5.
+The choice of 10 basis functions and a boundary factor of 1.5 could be an accurate enough choice.
+This same conclusion can be also be intuitively seen in the posterior mean and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}.
+
+From Figure \ref{fig5_MSE_vs_J}, a behaviour in terms of performance of the HSGP model as a function of the number of basis functions and the boundary factor, can be deduced: \\
+$\bullet$ as the boundary factor increases, more basis functions are needed,
+\\
+$\bullet$ as fewer basis functions are used, the boundary factor must decrease.
+
+Similar conclusions can be stated from results of Figure \ref{fig6_lscale_vs_J}.
+This Figure shows the estimated lengthscale and marginal variance as function of the number of basis functions and boundary factor.
+In addition to the above conclusions, in Figure \ref{fig6_lscale_vs_J} it can be easily seen that exists a minimum value for the boundary factor under which a close approximation will never be achieved.
+\\
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part1.pdf}}
+\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
+\caption{Posterior distributions and covariance functions of the regular GP model and the proposed HSGP model as a function of the number of basis functions $m$ and the boundary factor $c$.} 
+  \label{fig4_Post_&_Cov_part1}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
+\caption{Standardized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model.
+(left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$.
+(right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$.
+}
+  \label{fig5_MSE_vs_J}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.4]{fig6_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
+\caption{Estimated lengthscale and marginal variance parameters of the proposed HSGP models and the regular GP model.
+(left) Estimated lengthscales vs the number of basis functions $m$ and for different values of the boundary factor $c$.
+(right) Estimated marginal variance vs the number of basis functions $m$ and for different values of the boundary factor $c$.}
+  \label{fig6_lscale_vs_J}
+\end{figure}
+
+Additionally, there exit a relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the function.
+Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in relation to the performance of the approximation evaluated using eq.
+(\ref{diff_covs}).
+On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data $\ell/S$.
+The countour lines gather the boundary factor $c$.
+And the Y-axis represents the number of basis functions.
+Thus, this figure give us, for a certain GP function with lengthscale $\ell$ and given a boundary factor $c$, the minimum number of basis functions needed to achieve a close approximation in terms of satisfying eq.
+(\ref{diff_covs}).
+Alternatively, this figure could also be read as the minimum boundary factor $c$ that we should use given a number of basis functions, for certain GP function with certain lengthscale.
+And also this figure could also be read as the minimum functional lengthscales that can be closely approximated given a number of basis functions and a boundary factor.
+This figure basically collects the behaviour of the HSGP model as a function of the lengthscale, number of basis functions and boundary factor:
+
+$\bullet$ As the lengthscale of the function increases, the boundary factor and the number of basis functions needed for the approximation decrease.
+
+$\bullet$ There is a minimum boundary factor under which a close approximation never is going to be achieved (The contour lines have an end in function of the lengthscale).
+This statement match with one of the behaviour recognized in Figure \ref{fig6_lscale_vs_J} where were a minimum value for the boundary factor under which a close approximation will never be achieved.
+
+$\bullet$ The lower the boundary factor, the fewer basis functions and the smaller lengthscales that are suitable to be fitted.
+
+$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimum that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
+On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.
+
+Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $m$ and the boundary factor $c$ to approximate certain process characterized by its lengthscale.
+This Figure will allow us to optimize computations which depends on the number of basis functions.
+Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit.
+This figure also provides a diagnosis tool of the fit by comparing a actual estimated lengthscale to that the figure bring.
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
+\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
+\caption{Relation between the minimum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
+  \label{fig3_lscale_vs_J_vs_c_part1}
+\end{figure}
+
+If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J} where 10 basis functions and a boundary factor of 1.5 would be enough to closely approximate a function with normalized lengthscale equal to 0.3, we can recognize that these conclusions also matches the conclusion brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+
+\subsection{Comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model}
+
+We make a comparative analysis between the lengthscales estimated by the regular GP model and the approximate GP model in several datasets with different "wigglyness" (or lengthscales).
+Basically, comparing the lengthscales we can recognize if an approximation is close enough to the regular GP.
+We contrast these results with the information brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+
+For this analysis, we fit the regular GP model and the approximate GP model on different datasets.
+These datasets consists of noisy draws from a GP prior model with a squared exponential covariance function with different lengthscale.
+Different number of basis functions are used to fit the approximate GP model.
+The boundary factor is set in every case to a right and optimal choice.
+Figure \ref{fig7_varing_lscale} shows the posterior means of the regular GP model and the approximate GP model with different number of basis functions, for every dataset.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
+\caption{Posterior means of the regular GP and approximate GP models fitted over different datasets with different lengthscales.}
+  \label{fig7_varing_lscale}
+\end{figure}
+
+The comparative analysis between the lengthscales, estimated by a regular GP model and an approximate GP model in the different datasets with different lengthscales, is shown in Figure \ref{fig8_Tlscale_vs_Elscale}.
+In the X-axes of the plots, the true lengthscales of the data-generative functions for every dataset scenario are represented.
+In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented.
+Different comparisons are made using different number of basis functions in the approximate GP model.
+Let's assume that the boundary factor was properly set in every case.
+The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE of the approximate GP model, computed against the regular GP model, as a function of the lengthscale and number of basis functions.
+
+We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales.
+Very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high SRMSE.
+On the other side, big changes on long lengthscales, which might clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small SRMSE.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
+\caption{True lengthscales versus estimated lengthscales of the different realizations of datasets using different number of basis functions. }
+  \label{fig8_Tlscale_vs_Elscale}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
+\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and number of basis functions.}
+  \label{fig9_MSE_varing_lscale}
+\end{figure}
+
+\vspace{3mm}
+\section{Univariate examples}\label{sec:gp_examples1D}
+
+\subsection{Simulated data}
+
+This example consist of a simulated dataset with 250 draws values from a Gaussian process prior with a Matern covariance function (a Matern with 3/2 degrees of freedom) and inputs $\{\mathbf{x}_i\}_{i=1}^{250} \in \{-1,1\} \subset {\rm I\!R}$.
+The hyperparameters, marginal variance $\alpha$ and lengthscale $\ell$, of this GP prior were set to 1 and 0.15, respectively.
+The Gaussian noise $\sigma$ added to these draws was set to 0.2.
+
+The regular GP model for fitting this simulated dataset $\mathbf{y}$ can be written as follows,
+%
+\begin{eqnarray}\label{eq:latentgp_simudata1}
+\begin{split}
+\mathbf{y} &= f(\mathbf{x}) + \boldsymbol{\epsilon} \\
+\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  \mathbf{I}) \\
+f(\mathbf{x}) &\sim \mathcal{GP}(0, k(\mathbf{x}, \mathbf{x}', \theta)),
+\end{split}
+\end{eqnarray}
+
+\noindent The previous formulation corresponds to the latent form where $f(\mathbf{x})$ represents the functional model underlying the noisy data and modeled by a GP prior with a Matern covariance function $k$ (4).
+$\mathbf{I}$ represents the identity matrix.
+Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where the element $K_{ij}$ is formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j,\theta)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
+
+A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using sampling methods such as HMC, would be the marginalized form of the latent GP model in (\ref{eq:latentgp_simudata1}),
+%
+\begin{equation*}\label{eq:marginalizedgp_simudata1}
+\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 \mathbf{I} )
+\end{equation*}
+
+\noindent where the function values $f(\mathbf{x})$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing time of computation and improving the sampling and the effective number of samples.
+
+In the HSGP model, the covariance function $k$ is approximated as (\ref{approxcov}),
+%
+\begin{equation*}
+k(x,x') \approx \sum_{j}^{m}S(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
+\end{equation*}
+
+\noindent with the spectral density $S$ (3),
+%
+\begin{eqnarray*}
+S(w)&=& 4\alpha^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2},  
+\end{eqnarray*}
+
+\noindent and the latent function values $f(\mathbf{x)}$ are approximated as (\ref{approxf}),
+%
+\begin{equation*}
+f(x) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
+\end{equation*}
+
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
+A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
+
+For the HSGP model we use $m=80$ basis functions and a boundary factor $c=1.2$. For the splines model we use 80 knots or basis in the model.
+
+Figure \ref{fig10_Posteriors_exI} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
+The right and left extremes and three small regions around the middle of the plot correspond to out-of-sample or test data which have not been taking part on training the models.
+The test data have been plotted in Figure \ref{fig10_Posteriors_exI} with crosses and the training data with circles.
+The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.35]{fig10_Posteriors_exI.pdf}
+\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
+  \label{fig10_Posteriors_exI}
+\end{figure}
+
+Figure \ref{fig11_MSE_exI_inter} shows the  standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
+For this, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been formulated and fitted on this data.
+The SRMSE is computed against the true function.
+Model interpolation has been assessed either using the test data for interpolating and the training data.
+From Figures \ref{fig10_Posteriors_exI} and \ref{fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model, for interpolating and extrapolating data.
+However, the splines model does not extrapolate data properly.
+For interpolation both models show roughly similar performance.
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
+\caption{Standardized mean square error (SMSE) of the different methods against the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
+  \label{fig11_MSE_exI_inter}
+\end{figure}
+
+\subsection{Gay data}\label{sec:gay_data}
+This data set relates the proportion of support for same-sex marriage to the age.
+The data consists of 74 binomial observations per age group $i$ ($i=1,\cdots,74$), with the observation of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group.
+
+The observational model for the observed number of people $y_i$ supporting same-age marriage per age group $i$ is a binomial model with the total number of people $n_i$ per age group and the probability $p_i$ per age group as model parameters,
+%
+\begin{equation*}
+y_i \sim \text{Binomial}(p_i, n_i).
+\end{equation*}
+
+\noindent The total number of people per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ per age group $i$ or mean number of support people per age group $i$.
+The probability $p_i$ is related to the Gaussian process latent values $f(\mathbf{x})$ with age values $\mathbf{x}=\{x_i\}_{i=1}^{74}$ as inputs, through the logit function,
+%
+\begin{eqnarray*} \label{eq:gpprior_gay}
+p_i &=& \text{logit}(f(x_i)) \nonumber \\
+f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x},\mathbf{x}', \theta)
+\end{eqnarray*}
+
+\noindent with a squared exponential covariance function $k$ (2).
+
+In the HSGP model the covariance function is approximated as in eq.
+(\ref{approxcov}) with the square exponential spectral density in eq.
+(6), and the latent function $f(\mathbf{x})$ as in eq.
+(\ref{approxf}).
+
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
+A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
+
+For the HSGP model we use $m=20$ basis functions and a boundary factor $c=1.5$. For the splines model we use 20 knots or basis in the model.
+
+Figure \ref{fig12_Posteriors_gaydata} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
+The HSGP model posterior shown in the figure corresponds to 20 basis functions and a boundary factor of 1.5.
+For the spline posterior shown in the figure, it correspond to a 20 knots.
+Two small sets of observations have been used as test data, plotted as crosses in the Figure.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
+\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
+  \label{fig12_Posteriors_gaydata}
+\end{figure}
+
+For the HSGP model, different models with different number of basis functions and boundary factors have been used.
+The root mean square errors (RMSE) for every one of these models have been computed and plotted as a function of the number of basis functions and the boundary factor in Figure \ref{fig13_MSE_train_BF_gaydata}, for training (left) and test (right) data.
+In this case, the RMSE has been computed against the regular GP model.
+The expected patterns of the approximation in function of the number of basis function and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
+\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
+  \label{fig13_MSE_train_BF_gaydata}
+\end{figure}
+
+Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the different models, regular GP, HSGP and splines, against the actual data, for training and test data.
+Again we can see how the splines models does not extrapolate data properly.
+
+\begin{figure}[H]
+\centering
+\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
+\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
+  \label{fig14_MSE_train_gaydata}
+\end{figure}
+
+\subsection{Case II: Birthday data}\label{sec:bf_caseIII}
+
+\vspace{3mm}
+\section{Multivariate examples}\label{sec:gp_examplesMulti}
+\subsection{Simulated data}\label{sec:bf_toyexampleMulti}
+\subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
+\subsection{Case IV: Leukemia data}\label{sec:bf_caseV}
+\subsection{Case V: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
+
+\vspace{3mm}
+\appendix
+
+\section{Related work}
+
+The GP prior entails an $O(n^3)$ complexity that is computationally intractable for many practical problems, and this problem especially becomes severe when we want to conduct inference using sampling methods.
+To overcome this scaling problem several schemes have been proposed.
+One approach is to partition the data set into separate groups \citep{snelson2007local, urtasun2008sparse} and performing local inference in each partition.
+Other global approach is to build a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables' \citep{quinonero2005unifying,bui2017unifying}.
+Other global approach make use of basis functions to approximate the covariance function.
+In \cite{snelson2007local} the authors conduct an approach that combines the idea of local and global approaches.
+
+The literature contains many parametric models that approximate Gaussian process behaviours; for example \cite{bui2014tree} included  tree-structures in the approximation for extra scalability, and \cite{moore2015gaussian} combined local Gaussian
+processes with Gaussian random fields.
+
+\subsection{Inducing points methods}
+
+The approach based on inducing points employs a small set of pseudo data points to summarise the actual data.
+The storage requirements are reduced to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$.
+Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods.
+From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modfications to the so-called Nystr{\"o}m method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum.
+These methods are basically based on choosing a set of $m$ inducing inputs $x_u$ and scaling the corresponding eigendecomposition of their corresponding covariance matrix $K_{u,u}$ to match that of the actual covariance.
+
+This scheme was originally introduced to the GP context by \cite{williams2001using}.
+As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model.
+However, several methods modifying the inducing point approach are widely used.
+The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix.
+The SoR method is based on a degenerate prior which produces unreasonable predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).
+
+The Deterministic Training Conditional (DTC) method \citep{ro2001sparse,seeger2003fast}) retains the true covariance for the training data, but uses the approximate cross-covariances between training and test data, which reverse the problem of nonsensical predictive uncertainties.
+However, since the covariances for training and test cases are computed differently, this method results not to actually be a Gaussian process.
+This method was presented as Projected Latent Variables (PLV) in \cite{seeger2003fast} and Projected Process Approximation (PPA) in \cite{rasmussen2006gaussian}.
+
+The Variational Approximation (VAR) \citep{titsias2009variational} suggests a variational approach which provides an objective function for optimizing the selection of inducing points.
+This basically modifies the DTC method by an additional trace term in the likelihood that comes from the variational bound.
+\cite{hensman2013gaussian} extended this idea by introducing additional variational parameters to enable stochastic variational inference \citep{hoffman2013stochastic}, achieving a more computationally scalable bound which allows GPs to be fitted to millions of data.
+
+The Fully Independent (Training) Conditional (FIC) \citep{quinonero2005unifying} method originally introduced as Sparse Pseudo-Input GP by \cite{snelson2006sparse} is also based on the Nystr\"om approximation, where they allow the pseudo-point input locations to be optimised by maximising the new model's marginal likelihood whose covariance is parameterized by the locations of an active set not constrained to be a subset of the training and test data.
+
+More recently Bui et.
+al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time.
+The new framework is built on standard methods for approximate inference (variational-free-inference, EP and Power EP methods).
+
+In practice, the inducing points-based sparse approximation methods works reasonable well in cases where the field is relatively smooth. \cite{vanhatalo2010approximate}
+propose the use of compactly supported covariance function in conjunction with sparse approximations to model both short and long range correlations.
+
+\cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI).
+This framework improves the scalability and accuracy of fast kernel approximations through kernel interpolation, and naturally combines the advantages of inducing point and structure exploiting for scalability (such as Kronecker \citep{saatcci2012scalable} or Toeplitz \citep{cunningham2008fast}) approaches.
+
+The number of inducing points or their locations are crucial in order to capture the correlation structure.
+For a discussion on the effects of the inducing points, see \cite{vanhatalo2010approximate}.
+This behavior applies to all the methods from the Nystr\''om family.
+
+This kind of 'projected process' approximation has also been discussed by e.g. \cite{banerjee2008gaussian}.
+
+\subsection{Basis function methods}
+
+The spectral analysis and series expansions of Gaussian processes has a long history.
+A classical result (see, e.g, \cite{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}, and references therein) is that the covariance function can be approximated with a finite truncation of Mercer series and the approximation is guaranteed to converge to the exact covariance function when the number of terms is increased.
+
+Another related classical connection is to the works in the relationship of spline interpolation and Gaussian process priors \citep{wahba1978improper,kimeldorf1970correspondence,wahba1990spline}.
+In particular, it is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a specific choice of covariance function.
+The relationship of the spline regularization with Laplace operators then leads to series expansion representations that are closely related to the approximations considered here.
+
+Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels. The approximate kernel has a finite basis function expansion.
+
+The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse}, where the spectral representation of the covariance function is used.
+This model is a stationary sparse GP that can approximate any desired stationary full GP.
+However, as argued by the authors, this option does not converge to the full GP and can suffer from overfitting to the training data.
+\citep{gal2015improving} sought to improve the model by integrating out, rather than optimizing the frequencies.
+Gal and Turner derived a variational approximation that made use of a tractable integral over the frequency space.
+The result is an  algorithm that suffers less overfitting than the Sparse Spectrum GP, yet remains flexible.
+
+While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
+
+Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure.
+They combine the variational methodology with Fourier based approximations.
+
+In spatial statistics similar approaches are called low-rank models \citep{diggle2007springer}.
+The low rank models assume that the Gaussian field is a linear combination of $m$ basis functions.
+The type of an approximation depends on the basis functions used.
+Familiar examples include spectral representation \citep{diggle2007springer,paciorek2007computational,paciorek2007bayesian} and splines \citep{wood2003thin}.
+
+Recent Splines models can reproduce the Matern family of covariance functions, however our approach can reproduce basically all of the stationary covariance functions.
+
+\section{Contributions of the method}
+
+This work is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models.
+This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods.
+This results in a reduced-rank approximation for the covariance function.
+This method has some nice features:
+
+\vspace{2mm}
+$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.
+
+\vspace{2mm}
+$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
+In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
+
+\vspace{2mm}
+$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
+After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
+(Arno's paper, pag.
+7)
+
+\vspace{2mm}
+$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
+especially when sampling methods for inference are used.
+GP prior as latent functions is needed in generalized models.
+
+In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
+
+\vspace{2mm}
+$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
+This operation has to be computed at every step of the HMC or optimizer.
+
+\vspace{2mm}
+$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
+It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
+
+\vspace{2mm}
+$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
+
+\vspace{2mm}
+$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
+However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
+In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
+
+\section{Contributions of our work}
+
+As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology.
+Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function.
+The performance and accuracy of the method are directly related with the number of basis functions and the box size.
+At the same time, successful values for these two factors depend on the smoothness or roughness of the process to be modeled.
+The time of computation is mainly dependent on the number of basis functions.
+Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.
+
+\vspace{2mm}
+$\bullet$ Firstly, clear summarized formulae of the method for the univariate and multivariate cases is presented.
+
+\vspace{2mm}
+$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthscale of the functions.
+
+\vspace{2mm}
+$\bullet$ We make recommendations for the values of these factors based on the recognized relations among them.
+We provide useful graphs of these relations that will help the users to improve performance and save time of computation.
+
+\vspace{2mm}
+$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
+
+\vspace{2mm}
+$\bullet$ We describe the generalization of the method to the multidimensional case.
+
+\vspace{2mm}
+$\bullet$ We implement the approach in a fully probabilistic framework and for the Stan programming probabilistic software.
+
+\vspace{2mm}
+$\bullet$ We show several illustrative examples, simulate and real datasets, of the performance of the model, and accompanied by their Stan codes.
+
+%\vspace{3mm}
+%A summary of the formulae of the method for the univariate and multivariate cases is presented.
+The number of basis functions used in the approach and the value for the boundary condition of the model are two factors of main importance in practical applications.
+The number of basis functions is directly related to the accuracy of the approximation but also to the computation demands.
+Also exist a relation among the number of basis functions, the specified values for the boundary condition, and the performance of the approach.
+Finally, the approach is applied in several illustrative study cases and accompanied by their Stan codes.
+
+\section{Spectral densities of stationary covariance functions}
+
+The covariance function of a stationary process, that is function of $\boldsymbol{\tau}=\mathbf{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).
+
+\vspace{0.2cm}
+\textit{(Bochner’s theorem) A complex-valued function $k$ on ${\rm I\!R}^D$ is the covariance function of a weakly stationary mean square continuous complex valued random process on ${\rm I\!R}^D$ if and only if it can be represented as}
+%
+\begin{equation}
+k(\boldsymbol{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \mathbf{s} \cdot \mathbf{\tau}} d\mu(\boldsymbol{\tau}), \nonumber 
+\end{equation}
+
+\textit{where $\mu$ is a positive finite measure.}
+
+\vspace{0.2cm}
+If the measure $\mu$ has a density, it is known as the spectral density $S(\omega)$ of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the Wiener-Khintchine theorem.
+It gives the following relations:
+
+\begin{eqnarray}
+k(\boldsymbol{\tau})&=& \int S(\mathbf{s}) e^{2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber \\
+%
+S(\mathbf{s})&=& \int k(\boldsymbol{\tau}) e^{-2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber
+\end{eqnarray}
+
+\section{Approximate the covariance function using Hilbert space methods}
+
+Associated to each covariance function $k(\mathbf{x},\mathbf{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
+%
+\begin{equation}
+\mathcal{K} f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}'.
+\end{equation}
+
+Assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
+%
+\begin{equation}
+S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
+\end{equation}
+
+If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the covariance function $k$,
+
+\begin{equation}
+-\nabla^2 f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}',
+\end{equation}
+
+\noindent then the covariance function can be represented as
+
+\begin{equation}
+k(\mathbf{x},\mathbf{x}')= \sum_j \lambda_j \phi_j(\mathbf{x}) \phi_j(\mathbf{x}'),
+\end{equation}
+
+\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
+%
+\begin{eqnarray}
+-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
+\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
+\end{eqnarray}
+
+a series expansion of eigenvalues and eigenfunctions
+
+\section{Example of generalization to the multivariate case}
+
+Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
+The number of new multidimensional eigenfunctions $\phi^{\ast}_j$ and eigenvalues $\lambda^{\ast}_j$ is $J^D=3^2=9$ ($j=\{1,\cdots,J^D\}$).
+The matrix $\mathbb{S}\in {\rm I\!R}^{9 \times 2}$ is
+%
+\begin{eqnarray}
+\mathbb{S}=
+\left[ {\begin{array}{cc}
+1 & 1 \nonumber \\
+1 & 2 \\
+1 & 3 \\
+2 & 1 \\
+2 & 2 \\
+2 & 3 \\
+3 & 1 \\
+3 & 2 \\
+3 & 3
+\end{array} } \right]
+\end{eqnarray}
+
+\noindent and the multidimensional eigenfunctions and eigenvalues
+%
+\begin{eqnarray}
+\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} \nonumber \\
+%
+\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
+%
+\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber
+\end{eqnarray}
+
+Now, we show another example where different number of eigenfunctions and eigenvalues are used for every dimension.
+We consider a three-dimensional ($D=3$) input space, and sets of $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively.
+The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$.
+The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
+%
 \begin{eqnarray}
@@
 
-Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
-And each multivariate eigenvalue $\lambda^{\ast}_j$ is a $D$-vector whose elements are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
-Thus, for $\mathbf{x}^D \in \Omega$ and $j=1,\ldots,m$, we have
-%
-\begin{align} \label{bf&lambda_multi}
-\begin{split}
-\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \\
-%
-\lambda^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. 
-\end{split}
-\end{align}
-
-\noindent The approximate covariance function is then represented as
-%
-\begin{equation}\label{approxcov_multi}
-k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{m^{\ast}} 
-S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right)
-\phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
-\end{equation}
-
-\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section 2.1).
-We can now write the approximate series expansion of the multivariate function $f$ as,
-%
-\begin{equation}\label{approxf}
-f(\mathbf{x}) \approx \sum_{j=1}^{m^{\ast}} 
-\left( S^{\ast}_{\theta} \left(\sqrt{\lambda^{\ast}_j} \right)\right)^{1/2}
-\phi^{\ast}_j(\mathbf{x}) \beta_j, 
-\end{equation}
-
-\noindent where, again, $\beta_j \sim \text{Normal}(0,1)$ are normal random variables with zero mean and unitary variance.
-The computational cost of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
-Although this still implies linear scaling in $n$, the approximation becomes inefficient more quickly than in the univariate case.
-This is because $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and so increases rapidly with increased dimensionality.
-
-\subsection{Learning hyperparameters and model inference}
-
-\vspace{2mm}
-$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.
-
-\vspace{2mm}
-$\bullet$ The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.
-
-\vspace{2mm}
-$\bullet$ The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands.
-The weights depend on the hyperparameters, then they have to be computed in every step of the HMC sampling method.
-
-\vspace{2mm}
-$\bullet$ The linear model is computed with complexity $O(nm)$, computed in every step of the HMC sampling method.
-
-\vspace{2mm}
-$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
-After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
-(Arno's paper, pag.
-7)
-
-\vspace{2mm}
-$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
-especially when sampling methods for inference are used.
-GP prior as latent functions is needed in generalized models.
-
-In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
-
-\vspace{2mm}
-$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
-This operation has to be computed at every step of the HMC or optimizer.
-
-\vspace{2mm}
-$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
-It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
-
-\vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
-In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
-
-\vspace{2mm}
-$\bullet$ In addition, the computation of the automatic differentiation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ In a regular GP model the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
-This operation has to be computed at every step of the HMC.
-
-\vspace{2mm}
-$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-
-\section{The accuracy of the approximation}
-
-Implementation and performance of the HSGP model depends on the number of basis functions and boundary settings, which are, at the same time, related  to each other.
-The time of computation in the approximation mainly depends on the number of basis functions.
-Furthermore, the appropriate values for the number of basis functions and boundary depend on the non-linear functional effects of the function to be learned.
-In GP functions these non-linear functional effects are characterized by the lengthscale of the covariance function.
-
-In this section we analyze the rule and relationships of the number of basis functions and boundary on the approximation and their effects on the learned functions. 
-We also analyze these two factors in relation to the lengthscale of the functions to be learned.
-We provide a useful graph showing how these factors, number of basis functions, boundary and lengthscale, relate to each other in relation to the performance of the model.
-This graph basically provides the recommended values for these factors in order to get a close approximation.
-This graph also allows us to diagnose an actual fit by checking whether the value of the estimated lengthscale does not keep to those that the Figure brings as suitable ones.
-
-In this section we also make a comparison between regular GP and HSGP models, fitted for different dataset with different "wigglyness".
-The comparison is made through comparing their estimated lengthscales.
-
-\subsection{Dependency on the number of basis functions and the boundary condition}
-
-The approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, e.g.
-in a 1-$D$ input space $x\in \Omega=[-L,L]\subset \mathbb{R}$:
-%
-\begin{equation}\label{diffcov}
-k(\tau) = \sum_{j}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
-\end{equation}
-
-\noindent where $L$ is the boundary, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) for the unidimensional case and those in equation (\ref{bf&lambda_multi}) for the multidimensional case.
-
-The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
-The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
-%
-\begin{eqnarray}\label{diff_covs}
-\int k(\tau)d(\tau) - \int \sum_{j}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
-\end{eqnarray}
-
-The finite number $m$ of basis functions in the approximation needed to satisfy eq.
-(\ref{diff_covs}) will depend on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
-The approximation will also depends on boundary $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
-The boundary will also have an effect on the number of basis functions needed in the approximation.
-The boundary $L$ is set as an extension of the desired predicting input domain $\Psi$.
-Considering a zero-mean input domain $\Psi=[-S,S] \in \mathbb{R}$, the boundary is defined as
-%
-\begin{equation}
-L=c \cdot S,
-\end{equation}
-
-\noindent where $S$ represents the half-range of the input space, and $c$ ($c \ge 1$) is the proportional extension factor.
-From now on we are going to refer to this proportional factor $c$ as the boundary factor of the approximation.
-The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
-
-For a naive illustration of the effects of these two factors, number of basis functions $m$ and boundary $L$, on the performance of the approximation, a set of noisy observations are drawn from a GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, and input vales within the zero-mean input domain with half-range $S=1$.
-HSGP models with different number of basis functions and boundaries are fitted to this data.
-The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function.
-On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function.
-It can be seen how the number of basis functions affects the non-linear effects of the posteriors.
-Fewer basis functions implies smoother covariance functions and consequently smoother posteriors.
-The more "wigglyness" of the function, the more basis functions are needed.
-On the other hand, and similarly, for a fixed number of basis functions $m$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the boundary $L$ in approaching the posterior mean and the covariance function.
-It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances.
-\\
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
-\caption{(left) Posterior distributions for different number of basis functions $m$. (right) Covariance function for different number of basis functions $m$.}
-  \label{fig1_Post_J}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.43]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.43]{fig2_Cov_L.pdf}}
-\caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
-  \label{fig2_Post_L}
-\end{figure}
-
-In the previous Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} we focused on illustrating the individual effects of the number of basis functions and the boundary.
-Furthermore, for a clearer illustration of these effects, the lengthscale and marginal variance parameters in the HSGP model were fixed to their true values.
-Now we are going to focus on analyzing the interaction effects between these two factors, number of basis functions and the boundary factor, on the performance of the approximation.
-The lengthscale and marginal variance will not be fixed rather they will be estimated either in regular and HSGP models.
-%The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
-Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
-Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors means and the covariance functions obtained after fitting the data, for different number of basis functions and boundary factor.
-Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model, as a function of the number of basis functions and the boundary factor.
-Figure \ref{fig6_lscale_vs_J} shows the lengthscale and marginal variance for the regular GP model and for the HSGP models as a function of the number of basis functions and boundary factor.
-Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5.
-The choice of 10 basis functions and a boundary factor of 1.5 could be an accurate enough choice.
-This same conclusion can be also be intuitively seen in the posterior mean and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}.
-
-From Figure \ref{fig5_MSE_vs_J}, a behaviour in terms of performance of the HSGP model as a function of the number of basis functions and the boundary factor, can be deduced: \\
-$\bullet$ as the boundary factor increases, more basis functions are needed,
-\\
-$\bullet$ as fewer basis functions are used, the boundary factor must decrease.
-
-Similar conclusions can be stated from results of Figure \ref{fig6_lscale_vs_J}.
-This Figure shows the estimated lengthscale and marginal variance as function of the number of basis functions and boundary factor.
-In addition to the above conclusions, in Figure \ref{fig6_lscale_vs_J} it can be easily seen that exists a minimum value for the boundary factor under which a close approximation will never be achieved.
-\\
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part1.pdf}}
-\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
-\caption{Posterior distributions and covariance functions of the regular GP model and the proposed HSGP model as a function of the number of basis functions $m$ and the boundary factor $c$.} 
-  \label{fig4_Post_&_Cov_part1}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
-\caption{Standardized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model.
-(left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$.
-(right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$.
-}
-  \label{fig5_MSE_vs_J}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.4]{fig6_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
-\caption{Estimated lengthscale and marginal variance parameters of the proposed HSGP models and the regular GP model.
-(left) Estimated lengthscales vs the number of basis functions $m$ and for different values of the boundary factor $c$.
-(right) Estimated marginal variance vs the number of basis functions $m$ and for different values of the boundary factor $c$.}
-  \label{fig6_lscale_vs_J}
-\end{figure}
-
-Additionally, there exit a relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the function.
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in relation to the performance of the approximation evaluated using eq.
-(\ref{diff_covs}).
-On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data $\ell/S$.
-The countour lines gather the boundary factor $c$.
-And the Y-axis represents the number of basis functions.
-Thus, this figure give us, for a certain GP function with lengthscale $\ell$ and given a boundary factor $c$, the minimum number of basis functions needed to achieve a close approximation in terms of satisfying eq.
-(\ref{diff_covs}).
-Alternatively, this figure could also be read as the minimum boundary factor $c$ that we should use given a number of basis functions, for certain GP function with certain lengthscale.
-And also this figure could also be read as the minimum functional lengthscales that can be closely approximated given a number of basis functions and a boundary factor.
-This figure basically collects the behaviour of the HSGP model as a function of the lengthscale, number of basis functions and boundary factor:
-
-$\bullet$ As the lengthscale of the function increases, the boundary factor and the number of basis functions needed for the approximation decrease.
-
-$\bullet$ There is a minimum boundary factor under which a close approximation never is going to be achieved (The contour lines have an end in function of the lengthscale).
-This statement match with one of the behaviour recognized in Figure \ref{fig6_lscale_vs_J} where were a minimum value for the boundary factor under which a close approximation will never be achieved.
-
-$\bullet$ The lower the boundary factor, the fewer basis functions and the smaller lengthscales that are suitable to be fitted.
-
-$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimum that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
-On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.
-
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $m$ and the boundary factor $c$ to approximate certain process characterized by its lengthscale.
-This Figure will allow us to optimize computations which depends on the number of basis functions.
-Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit.
-This figure also provides a diagnosis tool of the fit by comparing a actual estimated lengthscale to that the figure bring.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
-\caption{Relation between the minimum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
-  \label{fig3_lscale_vs_J_vs_c_part1}
-\end{figure}
-
-If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J} where 10 basis functions and a boundary factor of 1.5 would be enough to closely approximate a function with normalized lengthscale equal to 0.3, we can recognize that these conclusions also matches the conclusion brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-\subsection{Comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model}
-
-We make a comparative analysis between the lengthscales estimated by the regular GP model and the approximate GP model in several datasets with different "wigglyness" (or lengthscales).
-Basically, comparing the lengthscales we can recognize if an approximation is close enough to the regular GP.
-We contrast these results with the information brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-For this analysis, we fit the regular GP model and the approximate GP model on different datasets.
-These datasets consists of noisy draws from a GP prior model with a squared exponential covariance function with different lengthscale.
-Different number of basis functions are used to fit the approximate GP model.
-The boundary factor is set in every case to a right and optimal choice.
-Figure \ref{fig7_varing_lscale} shows the posterior means of the regular GP model and the approximate GP model with different number of basis functions, for every dataset.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
-\caption{Posterior means of the regular GP and approximate GP models fitted over different datasets with different lengthscales.}
-  \label{fig7_varing_lscale}
-\end{figure}
-
-The comparative analysis between the lengthscales, estimated by a regular GP model and an approximate GP model in the different datasets with different lengthscales, is shown in Figure \ref{fig8_Tlscale_vs_Elscale}.
-In the X-axes of the plots, the true lengthscales of the data-generative functions for every dataset scenario are represented.
-In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented.
-Different comparisons are made using different number of basis functions in the approximate GP model.
-Let's assume that the boundary factor was properly set in every case.
-The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE of the approximate GP model, computed against the regular GP model, as a function of the lengthscale and number of basis functions.
-
-We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales.
-Very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high SRMSE.
-On the other side, big changes on long lengthscales, which might clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small SRMSE.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
-\caption{True lengthscales versus estimated lengthscales of the different realizations of datasets using different number of basis functions. }
-  \label{fig8_Tlscale_vs_Elscale}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
-\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and number of basis functions.}
-  \label{fig9_MSE_varing_lscale}
-\end{figure}
-
-\vspace{3mm}
-\section{Univariate examples}\label{sec:gp_examples1D}
-
-\subsection{Simulated data}
-
-This example consist of a simulated dataset with 250 draws values from a Gaussian process prior with a Matern covariance function (a Matern with 3/2 degrees of freedom) and inputs $\{\mathbf{x}_i\}_{i=1}^{250} \in \{-1,1\} \subset {\rm I\!R}$.
-The hyperparameters, marginal variance $\alpha$ and lengthscale $\ell$, of this GP prior were set to 1 and 0.15, respectively.
-The Gaussian noise $\sigma$ added to these draws was set to 0.2.
-
-The regular GP model for fitting this simulated dataset $\mathbf{y}$ can be written as follows,
-%
-\begin{eqnarray}\label{eq:latentgp_simudata1}
-\begin{split}
-\mathbf{y} &= f(\mathbf{x}) + \boldsymbol{\epsilon} \\
-\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  \mathbf{I}) \\
-f(\mathbf{x}) &\sim \mathcal{GP}(0, k(\mathbf{x}, \mathbf{x}', \theta)),
-\end{split}
-\end{eqnarray}
-
-\noindent The previous formulation corresponds to the latent form where $f(\mathbf{x})$ represents the functional model underlying the noisy data and modeled by a GP prior with a Matern covariance function $k$ (4).
-$\mathbf{I}$ represents the identity matrix.
-Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where the element $K_{ij}$ is formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j,\theta)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
-
-A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using sampling methods such as HMC, would be the marginalized form of the latent GP model in (\ref{eq:latentgp_simudata1}),
-%
-\begin{equation*}\label{eq:marginalizedgp_simudata1}
-\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 \mathbf{I} )
-\end{equation*}
-
-\noindent where the function values $f(\mathbf{x})$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing time of computation and improving the sampling and the effective number of samples.
-
-In the HSGP model, the covariance function $k$ is approximated as (\ref{approxcov}),
-%
-\begin{equation*}
-k(x,x') \approx \sum_{j}^{m}S(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
-\end{equation*}
-
-\noindent with the spectral density $S$ (3),
-%
-\begin{eqnarray*}
-S(w)&=& 4\alpha^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2},  
-\end{eqnarray*}
-
-\noindent and the latent function values $f(\mathbf{x)}$ are approximated as (\ref{approxf}),
-%
-\begin{equation*}
-f(x) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
-\end{equation*}
-
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
-
-For the HSGP model we use $m=80$ basis functions and a boundary factor $c=1.2$. For the splines model we use 80 knots or basis in the model.
-
-Figure \ref{fig10_Posteriors_exI} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
-The right and left extremes and three small regions around the middle of the plot correspond to out-of-sample or test data which have not been taking part on training the models.
-The test data have been plotted in Figure \ref{fig10_Posteriors_exI} with crosses and the training data with circles.
-The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.35]{fig10_Posteriors_exI.pdf}
-\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
-  \label{fig10_Posteriors_exI}
-\end{figure}
-
-Figure \ref{fig11_MSE_exI_inter} shows the  standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
-For this, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been formulated and fitted on this data.
-The SRMSE is computed against the true function.
-Model interpolation has been assessed either using the test data for interpolating and the training data.
-From Figures \ref{fig10_Posteriors_exI} and \ref{fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model, for interpolating and extrapolating data.
-However, the splines model does not extrapolate data properly.
-For interpolation both models show roughly similar performance.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
-\caption{Standardized mean square error (SMSE) of the different methods against the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
-  \label{fig11_MSE_exI_inter}
-\end{figure}
-
-\subsection{Gay data}\label{sec:gay_data}
-This data set relates the proportion of support for same-sex marriage to the age.
-The data consists of 74 binomial observations per age group $i$ ($i=1,\cdots,74$), with the observation of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group.
-
-The observational model for the observed number of people $y_i$ supporting same-age marriage per age group $i$ is a binomial model with the total number of people $n_i$ per age group and the probability $p_i$ per age group as model parameters,
-%
-\begin{equation*}
-y_i \sim \text{Binomial}(p_i, n_i).
-\end{equation*}
-
-\noindent The total number of people per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ per age group $i$ or mean number of support people per age group $i$.
-The probability $p_i$ is related to the Gaussian process latent values $f(\mathbf{x})$ with age values $\mathbf{x}=\{x_i\}_{i=1}^{74}$ as inputs, through the logit function,
-%
-\begin{eqnarray*} \label{eq:gpprior_gay}
-p_i &=& \text{logit}(f(x_i)) \nonumber \\
-f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x},\mathbf{x}', \theta)
-\end{eqnarray*}
-
-\noindent with a squared exponential covariance function $k$ (2).
-
-In the HSGP model the covariance function is approximated as in eq.
-(\ref{approxcov}) with the square exponential spectral density in eq.
-(6), and the latent function $f(\mathbf{x})$ as in eq.
-(\ref{approxf}).
-
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
-
-For the HSGP model we use $m=20$ basis functions and a boundary factor $c=1.5$. For the splines model we use 20 knots or basis in the model.
-
-Figure \ref{fig12_Posteriors_gaydata} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
-The HSGP model posterior shown in the figure corresponds to 20 basis functions and a boundary factor of 1.5.
-For the spline posterior shown in the figure, it correspond to a 20 knots.
-Two small sets of observations have been used as test data, plotted as crosses in the Figure.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
-\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
-  \label{fig12_Posteriors_gaydata}
-\end{figure}
-
-For the HSGP model, different models with different number of basis functions and boundary factors have been used.
-The root mean square errors (RMSE) for every one of these models have been computed and plotted as a function of the number of basis functions and the boundary factor in Figure \ref{fig13_MSE_train_BF_gaydata}, for training (left) and test (right) data.
-In this case, the RMSE has been computed against the regular GP model.
-The expected patterns of the approximation in function of the number of basis function and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
-  \label{fig13_MSE_train_BF_gaydata}
-\end{figure}
-
-Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the different models, regular GP, HSGP and splines, against the actual data, for training and test data.
-Again we can see how the splines models does not extrapolate data properly.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
-  \label{fig14_MSE_train_gaydata}
-\end{figure}
-
-\subsection{Case II: Birthday data}\label{sec:bf_caseIII}
-
-\vspace{3mm}
-\section{Multivariate examples}\label{sec:gp_examplesMulti}
-\subsection{Simulated data}\label{sec:bf_toyexampleMulti}
-\subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
-\subsection{Case IV: Leukemia data}\label{sec:bf_caseV}
-\subsection{Case V: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
-
-\vspace{3mm}
-\appendix
-
-\section{Related work}
-
-The GP prior entails an $O(n^3)$ complexity that is computationally intractable for many practical problems, and this problem especially becomes severe when we want to conduct inference using sampling methods.
-To overcome this scaling problem several schemes have been proposed.
-One approach is to partition the data set into separate groups \citep{snelson2007local, urtasun2008sparse} and performing local inference in each partition.
-Other global approach is to build a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables' \citep{quinonero2005unifying,bui2017unifying}.
-Other global approach make use of basis functions to approximate the covariance function.
-In \cite{snelson2007local} the authors conduct an approach that combines the idea of local and global approaches.
-
-The literature contains many parametric models that approximate Gaussian process behaviours; for example \cite{bui2014tree} included  tree-structures in the approximation for extra scalability, and \cite{moore2015gaussian} combined local Gaussian
-processes with Gaussian random fields.
-
-\subsection{Inducing points methods}
-
-The approach based on inducing points employs a small set of pseudo data points to summarise the actual data.
-The storage requirements are reduced to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$.
-Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods.
-From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modfications to the so-called Nystr{\"o}m method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum.
-These methods are basically based on choosing a set of $m$ inducing inputs $x_u$ and scaling the corresponding eigendecomposition of their corresponding covariance matrix $K_{u,u}$ to match that of the actual covariance.
-
-This scheme was originally introduced to the GP context by \cite{williams2001using}.
-As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model.
-However, several methods modifying the inducing point approach are widely used.
-The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix.
-The SoR method is based on a degenerate prior which produces unreasonable predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).
-
-The Deterministic Training Conditional (DTC) method \citep{ro2001sparse,seeger2003fast}) retains the true covariance for the training data, but uses the approximate cross-covariances between training and test data, which reverse the problem of nonsensical predictive uncertainties.
-However, since the covariances for training and test cases are computed differently, this method results not to actually be a Gaussian process.
-This method was presented as Projected Latent Variables (PLV) in \cite{seeger2003fast} and Projected Process Approximation (PPA) in \cite{rasmussen2006gaussian}.
-
-The Variational Approximation (VAR) \citep{titsias2009variational} suggests a variational approach which provides an objective function for optimizing the selection of inducing points.
-This basically modifies the DTC method by an additional trace term in the likelihood that comes from the variational bound.
-\cite{hensman2013gaussian} extended this idea by introducing additional variational parameters to enable stochastic variational inference \citep{hoffman2013stochastic}, achieving a more computationally scalable bound which allows GPs to be fitted to millions of data.
-
-The Fully Independent (Training) Conditional (FIC) \citep{quinonero2005unifying} method originally introduced as Sparse Pseudo-Input GP by \cite{snelson2006sparse} is also based on the Nystr\"om approximation, where they allow the pseudo-point input locations to be optimised by maximising the new model's marginal likelihood whose covariance is parameterized by the locations of an active set not constrained to be a subset of the training and test data.
-
-More recently Bui et.
-al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time.
-The new framework is built on standard methods for approximate inference (variational-free-inference, EP and Power EP methods).
-
-In practice, the inducing points-based sparse approximation methods works reasonable well in cases where the field is relatively smooth. \cite{vanhatalo2010approximate}
-propose the use of compactly supported covariance function in conjunction with sparse approximations to model both short and long range correlations.
-
-\cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI).
-This framework improves the scalability and accuracy of fast kernel approximations through kernel interpolation, and naturally combines the advantages of inducing point and structure exploiting for scalability (such as Kronecker \citep{saatcci2012scalable} or Toeplitz \citep{cunningham2008fast}) approaches.
-
-The number of inducing points or their locations are crucial in order to capture the correlation structure.
-For a discussion on the effects of the inducing points, see \cite{vanhatalo2010approximate}.
-This behavior applies to all the methods from the Nystr\''om family.
-
-This kind of 'projected process' approximation has also been discussed by e.g. \cite{banerjee2008gaussian}.
-
-\subsection{Basis function methods}
-
-The spectral analysis and series expansions of Gaussian processes has a long history.
-A classical result (see, e.g, \cite{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}, and references therein) is that the covariance function can be approximated with a finite truncation of Mercer series and the approximation is guaranteed to converge to the exact covariance function when the number of terms is increased.
-
-Another related classical connection is to the works in the relationship of spline interpolation and Gaussian process priors \citep{wahba1978improper,kimeldorf1970correspondence,wahba1990spline}.
-In particular, it is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a specific choice of covariance function.
-The relationship of the spline regularization with Laplace operators then leads to series expansion representations that are closely related to the approximations considered here.
-
-Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels. The approximate kernel has a finite basis function expansion.
-
-The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse}, where the spectral representation of the covariance function is used.
-This model is a stationary sparse GP that can approximate any desired stationary full GP.
-However, as argued by the authors, this option does not converge to the full GP and can suffer from overfitting to the training data.
-\citep{gal2015improving} sought to improve the model by integrating out, rather than optimizing the frequencies.
-Gal and Turner derived a variational approximation that made use of a tractable integral over the frequency space.
-The result is an  algorithm that suffers less overfitting than the Sparse Spectrum GP, yet remains flexible.
-
-While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
-
-Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure.
-They combine the variational methodology with Fourier based approximations.
-
-In spatial statistics similar approaches are called low-rank models \citep{diggle2007springer}.
-The low rank models assume that the Gaussian field is a linear combination of $m$ basis functions.
-The type of an approximation depends on the basis functions used.
-Familiar examples include spectral representation \citep{diggle2007springer,paciorek2007computational,paciorek2007bayesian} and splines \citep{wood2003thin}.
-
-Recent Splines models can reproduce the Matern family of covariance functions, however our approach can reproduce basically all of the stationary covariance functions.
-
-\section{Contributions of the method}
-
-This work is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models.
-This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods.
-This results in a reduced-rank approximation for the covariance function.
-This method has some nice features:
-
-\vspace{2mm}
-$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.
-
-\vspace{2mm}
-$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
-After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
-(Arno's paper, pag.
-7)
-
-\vspace{2mm}
-$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
-especially when sampling methods for inference are used.
-GP prior as latent functions is needed in generalized models.
-
-In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
-
-\vspace{2mm}
-$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
-This operation has to be computed at every step of the HMC or optimizer.
-
-\vspace{2mm}
-$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
-It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
-
-\vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
-In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
-
-\section{Contributions of our work}
-
-As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology.
-Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function.
-The performance and accuracy of the method are directly related with the number of basis functions and the box size.
-At the same time, successful values for these two factors depend on the smoothness or roughness of the process to be modeled.
-The time of computation is mainly dependent on the number of basis functions.
-Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.
-
-\vspace{2mm}
-$\bullet$ Firstly, clear summarized formulae of the method for the univariate and multivariate cases is presented.
-
-\vspace{2mm}
-$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthscale of the functions.
-
-\vspace{2mm}
-$\bullet$ We make recommendations for the values of these factors based on the recognized relations among them.
-We provide useful graphs of these relations that will help the users to improve performance and save time of computation.
-
-\vspace{2mm}
-$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
-
-\vspace{2mm}
-$\bullet$ We describe the generalization of the method to the multidimensional case.
-
-\vspace{2mm}
-$\bullet$ We implement the approach in a fully probabilistic framework and for the Stan programming probabilistic software.
-
-\vspace{2mm}
-$\bullet$ We show several illustrative examples, simulate and real datasets, of the performance of the model, and accompanied by their Stan codes.
-
-%\vspace{3mm}
-%A summary of the formulae of the method for the univariate and multivariate cases is presented.
-The number of basis functions used in the approach and the value for the boundary condition of the model are two factors of main importance in practical applications.
-The number of basis functions is directly related to the accuracy of the approximation but also to the computation demands.
-Also exist a relation among the number of basis functions, the specified values for the boundary condition, and the performance of the approach.
-Finally, the approach is applied in several illustrative study cases and accompanied by their Stan codes.
-
-\section{Spectral densities of stationary covariance functions}
-
-The covariance function of a stationary process, that is function of $\boldsymbol{\tau}=\mathbf{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).
-
-\vspace{0.2cm}
-\textit{(Bochner’s theorem) A complex-valued function $k$ on ${\rm I\!R}^D$ is the covariance function of a weakly stationary mean square continuous complex valued random process on ${\rm I\!R}^D$ if and only if it can be represented as}
-%
-\begin{equation}
-k(\boldsymbol{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \mathbf{s} \cdot \mathbf{\tau}} d\mu(\boldsymbol{\tau}), \nonumber 
-\end{equation}
-
-\textit{where $\mu$ is a positive finite measure.}
-
-\vspace{0.2cm}
-If the measure $\mu$ has a density, it is known as the spectral density $S(\omega)$ of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the Wiener-Khintchine theorem.
-It gives the following relations:
-
-\begin{eqnarray}
-k(\boldsymbol{\tau})&=& \int S(\mathbf{s}) e^{2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber \\
-%
-S(\mathbf{s})&=& \int k(\boldsymbol{\tau}) e^{-2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber
-\end{eqnarray}
-
-\section{Approximate the covariance function using Hilbert space methods}
-
-Associated to each covariance function $k(\mathbf{x},\mathbf{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
-%
-\begin{equation}
-\mathcal{K} f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}'.
-\end{equation}
-
-Assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
-%
-\begin{equation}
-S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
-\end{equation}
-
-If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the covariance function $k$,
-
-\begin{equation}
--\nabla^2 f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}',
-\end{equation}
-
-\noindent then the covariance function can be represented as
-
-\begin{equation}
-k(\mathbf{x},\mathbf{x}')= \sum_j \lambda_j \phi_j(\mathbf{x}) \phi_j(\mathbf{x}'),
-\end{equation}
-
-\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
-Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
-%
-\begin{eqnarray}
--\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
-\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
-\end{eqnarray}
-
-a series expansion of eigenvalues and eigenfunctions
-
-\section{Example of generalization to the multivariate case}
-
-Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
-The number of new multidimensional eigenfunctions $\phi^{\ast}_j$ and eigenvalues $\lambda^{\ast}_j$ is $J^D=3^2=9$ ($j=\{1,\cdots,J^D\}$).
-The matrix $\mathbb{S}\in {\rm I\!R}^{9 \times 2}$ is
-%
-\begin{eqnarray}
-\mathbb{S}=
-\left[ {\begin{array}{cc}
-1 & 1 \nonumber \\
-1 & 2 \\
-1 & 3 \\
-2 & 1 \\
-2 & 2 \\
-2 & 3 \\
-3 & 1 \\
-3 & 2 \\
-3 & 3
-\end{array} } \right]
-\end{eqnarray}
-
-\noindent and the multidimensional eigenfunctions and eigenvalues
-%
-\begin{eqnarray}
-\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} \nonumber \\
-%
-\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber
-\end{eqnarray}
-
-Now, we show another example where different number of eigenfunctions and eigenvalues are used for every dimension.
-We consider a three-dimensional ($D=3$) input space, and sets of $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively.
-The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$.
-The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
-%
-\begin{eqnarray}
-\mathbb{S}=
-\left[ {\begin{array}{ccc}
-1 & 1 & 1 \nonumber \\
-1 & 1 & 2 \\
-1 & 1 & 3 \\
-1 & 2 & 1 \\
-1 & 2 & 2 \\
-1 & 2 & 3 \\
-2 & 1 & 1 \\
-2 & 1 & 2 \\
-2 & 1 & 3 \\
-2 & 2 & 1 \\
-2 & 2 & 2 \\
-2 & 2 & 3 
-\end{array} } \right]
-\end{eqnarray}
-
 \noindent and the multidimensional eigenfunctions and eigenvalues

### 2019-03-05 Paper/manuscript/manuscript_05.tex
some edits in the beginning of section 3

@@
 The time of computation in the approximation mainly depends on the number of basis functions.
-Furthermore, the appropriate values for the number of basis functions and boundary depend on the non-linear functional effects of the function to be learned.
-In GP functions these non-linear functional effects are characterized by the lengthscale of the covariance function.
-
-In this section we analyze the rule and relationships of the number of basis functions and boundary on the approximation and their effects on the learned functions. 
-We also analyze these two factors in relation to the lengthscale of the functions to be learned.
-We provide a useful graph showing how these factors, number of basis functions, boundary and lengthscale, relate to each other in relation to the performance of the model.
-This graph basically provides the recommended values for these factors in order to get a close approximation.
+Furthermore, the appropriate values for the number of basis functions and boundary depend on the non-linear functional effects of the function characterized by the lengthscale of the covariance function.
+
+In this section we analyze the effects of the number of basis functions and boundary on the approximation accuracy.
+We present recommendations how they should be chosen and diagnostics to check the accuracy of the approximation after the inference.
+We illustrate graphically how the number of basis functions, boundary and lengthscale relate to each other and affect the accuracy of the approximation.
+% This graph basically provides the recommended values for these factors in order to get a close approximation.
 This graph also allows us to diagnose an actual fit by checking whether the value of the estimated lengthscale does not keep to those that the Figure brings as suitable ones.
 
-In this section we also make a comparison between regular GP and HSGP models, fitted for different dataset with different "wigglyness".
-The comparison is made through comparing their estimated lengthscales.
+We compare regular GP and HSGP fitted for different dataset with different "wigglyness".
+%The comparison is made through comparing their estimated lengthscales.
 
@@
 
-The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
+The approximation to the covariance function becomes exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
 The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
@@
 The finite number $m$ of basis functions in the approximation needed to satisfy eq.
-(\ref{diff_covs}) will depend on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
-The approximation will also depends on boundary $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
+(\ref{diff_covs}) depends on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
+The approximation also depends on boundary $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
 The boundary will also have an effect on the number of basis functions needed in the approximation.

### 2019-03-06 Paper/manuscript/manuscript_05.tex
more cleaning of the introduction

@@
 
-Gaussian processes (GPs) are flexible define probability distributions over multi-dimensional non-linear functions \citep{rasmussen2006gaussian,neal1997monte}.
-GPs are defind by a mean and covariance function and any finite set of function values are jointly distributed as a multivariate normal.
+Gaussian processes (GPs) are flexible statistical models for specifying probability distributions over multi-dimensional non-linear functions \citep{rasmussen2006gaussian,neal1997monte}.
+Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate normal.
+GPs themselves are defined by two functions, a mean and a covariance function, with the latter being arguably the more important one.
 The covariance function encodes our prior assumptions about the functional relationship, such as continuity, smoothness, periodicity and scale properties.
@@
 
-Given $n$ observations in the data, the computational cost and memory requirements of exact GP implementation in general scales as $O(n^3)$ and $O(n^2)$, respectively.
+Given $n$ observations in the data, the computational comlexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
 This would limit their application to rather small data sets of a few tens of thousands observations at most.
-The problem becomes especially severe when performing full Bayesian inference via sampling methods, where in each sampling step we have $O(n^3)$ computational cost when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
+The problem becomes especially severe when performing full Bayesian inference via sampling methods, where in each sampling step we need $O(n^3)$ computations when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
 To alleviate these computational demands, several approximate methods have been proposed, which we will briefly review in the following.
@@
 Sparse GPs are based on low-rank approximations of the covariance matrix.
-The low-rank approximation with $m \ll n$ {\it inducing points} can be presented with reduced memory cost $O(nm)$ and the corresponding computational complexity is $O(nm^2)$.
-% While computational complexity is reduced considerably, ensuring convergence to the exact GP model is not necessarily straightforward (Paul: this may need a citation).
+The low-rank approximation with $m \ll n$ {\it inducing points} implies reduced memory requirements of $O(nm)$ and corresponding computational complexity of $O(nm^2)$.
 A unifying view on sparse GPs based on approximate generative methods
@@
 
-% (Paul: This whole paragraph is unclear to me.
-Are all of these approaches approximations via splines, or are the splines just special cases of a more general method?
-In any case, I think the paragraph needs substantial revision).
-Alternative low-rank approximation is based on forming a basis function approximation with $m \ll$.
-The basis functions are presented explicitly, but could also be used to form a low rank covariance matrix approximation.
-Common basis function  approximations are based on the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
-Certain spline smoothing basis functions are equivalent to GPs with certain covariance function \citep{wahba1990spline}, but not all spline models are GPs (sometimes splines are not marginalization consistent)-
-%
-The sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving,gal2015improving}.
-Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels.
-Random fourier features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels.
-
-In this paper we focus on Laplace eigenfunction basis function approximation for stationary covariance functions proposed by \citet{solin2018hilbert}.
-Basis function approaches are computationally as linear models, which is attractive property in modular probabilistic programming models where there is a big benefit id approximation specific computation is simple.
+An alternative class of low-rank approximations is based on forming a basis function approximation with $m \ll n$ basis functions.
+The basis functions are usually presented explicitly, but can also be used to form a low rank covariance matrix approximation.
+Common basis function approximations rest on the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
+Sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving}.
+Recently, \cite{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels.
+Another related method for approximating kernels relies on random fourier features \citep{rahimi2008random,rahimi2009weighted}.
+Further, certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline}, but not all spline models are GPs (e.g., sometimes splines are not marginalization consistent).
+
+\textcolor{red}{Paul: Can we clarify how our discussed method relates to the former paragraph,
+which introduces all sorts of basis function approximations. I.e. is our method
+just one out of many or what makes it special? Perhaps we can somehow use the information provided
+in: While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
+Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.}
+
+In this paper we focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
+Basis function approaches behave computationally like linear models, which is an attractive property in modular probabilistic programming models where there is a big benefit id approximation specific computation is simple (Paul: What do you mean with the the last part of the sentence?).
 The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance kernel including the hyperparameters.
-While the pre-computation cost of the basis functions is $O(m^2n)$ the computation cost when only the covariance function parameters change is $O(mn+m)$ (or $O(m^3)$ if marginalization can be used).
-This is a big advantage for iterative algorithms such as Markov chain Monte Carlo (MCMC).
-The further advantage is the reduced memory cost in automatic differentation used in modern probabilistic programming frameworks, as the automatic differentation memory cost scales as the computational cost and not as the usual memory cost for the posterior density computation.
-The basis function approach provides also easy way to use non-centered parameterization reducing posterior dependency between parameters presenting the function and the covariance function parameters, which helps to further improve MCMC efficiency.
-
-While \citet{solin2018hilbert} fully developed the mathematical theory behind the specific reduced-rank approximation, for practical implementations of approximate GPs in probabilistic programming framework further work is needed.
-In this paper we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
+While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost when only changing the covariance function parameters is $O(mn+m)$ (or even $O(m^3)$ if marginalization can be used).
+This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
+Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks (cite!).
+This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation.
+The discussed basis function approach also provides an easy way to apply the non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which helps to further improve MCMC efficiency.
+
+While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
+In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
 We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.
 
-Although there are several GP specific software packages available to date (GPML, GPstuff, GPy, GPflow), each provide efficient implementations only for a restricted range of GP based models.
-In this paper we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
+Although there are several GP specific software packages available to date (GPML, GPstuff, GPy, GPflow, cite them!), each provide efficient implementations only for a restricted range of GP based models.
+In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
 
@@
 exact and approximate GPs to real and simulated data are provided in Section 4. We end with a discussion in Section 5.
-
-(Paul: The content and the purpose of the following sentences is not fully clear to me.
-Do we need them in the Introduction or can they be removed, or put into the next section?
-Gabi: In this paragraph I tried to state two differences I found in the literature between our method and the Sparse Spectrum GP and the Splines method.
-But I not know where locate them.) While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
-Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.
 

### 2019-03-07 Paper/manuscript/manuscript_05.tex
add case studies 4.1 and 4.2. Improve cov. functions in section2

@@
 where $p(\mathbf{y})$ is the marginal likelihood and is given by
+%
 \begin{align}\label{eq:evidence}
@@
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
-A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(|\mathbf{x}-\mathbf{x}'|)$, which means that the covariance is invariant to translations.
-Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariance is both translation and rotation invariant.
+A stationary covariance function is a function of $\bm{\tau}=\mathbf{x}-\mathbf{x}' \in \mathbb{R}^D$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
+Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(r), r\in \mathbb{R}$, which means that the covariance is both translation and rotation invariant.
 The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which, in the unidimensional case (i.e.
-$\mathbf{x}=x\in \mathbb{R}$, so $\mathbf{r}=r\in \mathbb{R}$), are
-%
-\begin{align}
-k_{\nu=\infty}(r)&=\sigma^2 \exp(-\frac{1}{2} r^2/\ell^2), \nonumber \\
-k_{\nu=\frac{1}{2}}(r)&=\sigma^2 \exp(-r/\ell), \nonumber \\
-k_{\nu=\frac{3}{2}}(r)&=\sigma^2(1+\sqrt{3}r/\ell) \exp(-\sqrt{3}r/\ell), \nonumber \\
-k_{\nu=\frac{5}{2}}(r)&=\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \exp(-\sqrt{5}r/\ell), \nonumber 
-\end{align}
+$\mathbf{x}=x\in \mathbb{R}$, so $\mathbf{r}=r\in \mathbb{R}$, and $\ell\in \mathbb{R}$ ), are
+%
+\begin{align*}
+k_{\nu=\infty}(r)&=\sigma^2 \exp(-\frac{1}{2} r^2/\ell^2), \\
+k_{\nu=\frac{1}{2}}(r)&=\sigma^2 \exp(-r/\ell),  \\
+k_{\nu=\frac{3}{2}}(r)&=\sigma^2(1+\sqrt{3}r/\ell) \exp(-\sqrt{3}r/\ell), 
+%k_{\nu=\frac{5}{2}}(r)&=\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \exp(-\sqrt{5}r/\ell), 
+\end{align*}
 
@@
 The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function.
-These covariance functions can be easily generalized to the multidimensional case, with multidimensional inputs $\mathbf{x}\in \mathbb{R}^D$ and multidimensional length-scale $\boldsymbol{\ell}\in \mathbb{R}^D$ (Gabi: We should derive the multidimensional form of the previous kernels, but I have now doubts how is the multidimensional form for the Matern kernels with $\nu$=1/2, $\nu$=3/2 and $\nu$=5/2).
-Using a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.
+These covariance functions can be easily generalized to a $D$-dimensional case, by using the property of separability of multidimensional kernels.
+Let $\mathbf{x}\in \mathbb{R}^D$, $\mathbf{\bm{\tau}}=\{\tau_1,\tau_2\}\in \mathbb{R}^D$ and $\boldsymbol{\ell}=\{\ell_1,\ell_2\}\in \mathbb{R}^D$,
+%
+\begin{align*}
+k_{\nu=\infty}(\bm{\tau})&=\exp(-\frac{1}{2} \tau_1^2/\ell_1^2) \cdot \exp(-\frac{1}{2} \tau_2^2/\ell_2^2), \\
+k_{\nu=\frac{1}{2}}(\bm{\tau})&=\exp(-\tau_1/\ell_1) \cdot \exp(-\tau_2/\ell_2), \\
+k_{\nu=\frac{3}{2}}(\bm{\tau})&=(1+\sqrt{3}\tau_1/\ell_1) \exp(-\sqrt{3}\tau_1/\ell_1) \cdot (1+\sqrt{3}\tau_2/\ell_2) \exp(-\sqrt{3}\tau_2/\ell_2) \\
+\end{align*}
+
+\noindent Using a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.
 
@@
 If this measure has a density, it is known as the spectral density of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the \textit{Wiener-Khintchine theorem} \citep{rasmussen2006gaussian}.
-The spectral density functions associated with the Mat\'ern covariance functions written above are
-%
-\begin{align}
+In the unidimensional case the spectral density functions associated with the Mat\'ern covariance functions written above are
+%
+\begin{align}\label{eq:specdens}
 S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right), \\
@@
 S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}, \\
-S_{\nu=\frac{5}{2}}(w)&= \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}\left(\frac{5}{\ell^2} + \omega^2\right)^{-3},  
+%S_{\nu=\frac{5}{2}}(w)&= \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}\left(\frac{5}{\ell^2} + \omega^2\right)^{-3},  
 \end{align}
 
-\noindent where variable $w$ is in the frequency domain. The $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
-%
-\begin{align}
-S^{\ast}_{\nu=\infty}(\mathbf{w})&= \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \exp\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
-%
-S^{\ast}_{\nu}(\mathbf{w})&= \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
+\noindent where variable $w\in \mathbb{R}$ is in the frequency domain, and $\ell\in \mathbb{R}$.
+By using the separability property of the spectral densities functions, the $D$-dimensional form of these spectral density functions, where $\mathbf{w}=\{w_1,w_2\}$ and $\mathbf{\ell}=\{\ell_1,\ell_2\}$ are
+%
+\begin{align}\label{eq:specdens_multi}
+S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \omega_2^2\right) \\
+S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell_1}\left(\frac{1}{\ell_1^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell_2}\left(\frac{1}{\ell_2^2} + \omega_2^2\right)^{-1} \\
+S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell_1}^{3}\left(\frac{3}{\ell_1^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell_2}^{3}\left(\frac{3}{\ell_2^2} + \omega_2^2\right)^{-2},
 \end{align}
 
-\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain.
+\noindent which can also be derived from the following expressions \citep{rasmussen2006gaussian}
+%
+\begin{align*}
+S^{\ast}_{\nu=\infty}(\mathbf{\omega})&= \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \exp\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{\omega}_d^2\right), \\
+%
+S^{\ast}_{\nu}(\mathbf{\omega})&= \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{\omega}^2 \right)^{(\nu+D/2)}.
+\end{align*}
 
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:boundary}
 L=c \cdot S,
@@
 
-This example consist of a simulated dataset with 250 draws values from a Gaussian process prior with a Matern covariance function (a Matern with 3/2 degrees of freedom) and inputs $\{\mathbf{x}_i\}_{i=1}^{250} \in \{-1,1\} \subset {\rm I\!R}$.
-The hyperparameters, marginal variance $\alpha$ and lengthscale $\ell$, of this GP prior were set to 1 and 0.15, respectively.
-The Gaussian noise $\sigma$ added to these draws was set to 0.2.
+This example consists of a simulated dataset with 250 draws of a Gaussian process prior with a Mat{\'e}rn covariance function with 3/2 degrees of freedom and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, and inputs values $\{\mathbf{x}_i\}_{i=1}^{250} \in \{-1,1\} \subset {\rm I\!R}$.
+To form the final dataset $\mathbf{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
 
@@
 %
-\begin{eqnarray}\label{eq:latentgp_simudata1}
+\begin{eqnarray*}\label{eq:latentgp_simudata1}
 \begin{split}
@@
 \end{split}
-\end{eqnarray}
-
-\noindent The previous formulation corresponds to the latent form where $f(\mathbf{x})$ represents the functional model underlying the noisy data and modeled by a GP prior with a Matern covariance function $k$ (4).
-$\mathbf{I}$ represents the identity matrix.
-Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where the element $K_{ij}$ is formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j,\theta)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$.
-
-A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using sampling methods such as HMC, would be the marginalized form of the latent GP model in (\ref{eq:latentgp_simudata1}),
+\end{eqnarray*}
+
+\noindent where $\mathbf{I}$ represents the identity matrix.
+The previous formulation corresponds to the latent form of a GP model, where $f(\mathbf{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$ (\ref{covfun}).
+Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(\mathbf{x}_i,\mathbf{x}_j,\theta)$.
+
+A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using sampling methods such as HMC, would be its marginalized form,
 %
@@
 
-In the HSGP model, the covariance function $k$ is approximated as (\ref{approxcov}),
+In the HSGP model, the latent function values $f(x)$ are approximated as (\ref{approxf}),
 %
 \begin{equation*}
-k(x,x') \approx \sum_{j}^{m}S(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
+f(x) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j,
 \end{equation*}
 
-\noindent with the spectral density $S$ (3),
+\noindent with the spectral density $S$ (\ref{specdens}) as a function of $\sqrt{\lambda_j}$,
 %
 \begin{eqnarray*}
-S(w)&=& 4\alpha^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2},  
+S(\sqrt{\lambda_j})&=& 4\alpha^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \lambda_j)^{-2}, 
 \end{eqnarray*}
 
-\noindent and the latent function values $f(\mathbf{x)}$ are approximated as (\ref{approxf}),
-%
-\begin{equation*}
-f(x) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
-\end{equation*}
+\noindent and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ (\ref{bf&lambda})
+%
+\begin{eqnarray*}
+\begin{split}
+\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \\
+\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right).
+\end{split}
+\end{eqnarray*}
+
+\noindent $L$ is the boundary and $m$ the number of basis functions.
+The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale parameters of the approximate covariance function, respectively.
 
 In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
-
-For the HSGP model we use $m=80$ basis functions and a boundary factor $c=1.2$. For the splines model we use 80 knots or basis in the model.
-
-Figure \ref{fig10_Posteriors_exI} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
-The right and left extremes and three small regions around the middle of the plot correspond to out-of-sample or test data which have not been taking part on training the models.
-The test data have been plotted in Figure \ref{fig10_Posteriors_exI} with crosses and the training data with circles.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
+
+Figure \ref{fig10_Posteriors_exI} shows the posteriors distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$ (\ref{eq:boundary})), and the splines model with 80 knots.
+The true data-generative function and the noisy observations are also plotted in the Figure.
+The right and left extremes and three small regions in the middle of the plot correspond to out-of-sample or test data which have not been taking part on training the models.
+The test data have been plotted as crosses and the training data as circles.
 The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
@@
 
+In order to assess performance of the models as a function the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
 Figure \ref{fig11_MSE_exI_inter} shows the  standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
-For this, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been formulated and fitted on this data.
 The SRMSE is computed against the true function.
@@
 \subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
-\caption{Standardized mean square error (SMSE) of the different methods against the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
+\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
   \label{fig11_MSE_exI_inter}
@@
 This data set relates the proportion of support for same-sex marriage to the age.
-The data consists of 74 binomial observations per age group $i$ ($i=1,\cdots,74$), with the observation of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group.
-
-The observational model for the observed number of people $y_i$ supporting same-age marriage per age group $i$ is a binomial model with the total number of people $n_i$ per age group and the probability $p_i$ per age group as model parameters,
+The data consists of 74 observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group $i$ ($i=1,\cdots,74$).
+The observational model is a binomial model with parameters population $n_i$ and probability of supporting same-sex marriage $p_i$ per age group $i$,
 %
@@
 
-\noindent The total number of people per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ per age group $i$ or mean number of support people per age group $i$.
-The probability $p_i$ is related to the Gaussian process latent values $f(\mathbf{x})$ with age values $\mathbf{x}=\{x_i\}_{i=1}^{74}$ as inputs, through the logit function,
+\noindent The population per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ or mean number of support people per age group.
+The probabilities $\mathbf{p}=\{p_i\}_{i=1}^{74}$ are modeled by a GP latent function $f(\mathbf{x})$ with a squared exponential covariance function $k$ (2) and age input values $\mathbf{x}=\{x_i\}_{i=1}^{74}$, with the {\it logit} function as a link function,
 %
@@
 p_i &=& \text{logit}(f(x_i)) \nonumber \\
-f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x},\mathbf{x}', \theta)
+f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x},\mathbf{x}', \theta).
 \end{eqnarray*}
 
-\noindent with a squared exponential covariance function $k$ (2).
-
-In the HSGP model the covariance function is approximated as in eq.
-(\ref{approxcov}) with the square exponential spectral density in eq.
-(6), and the latent function $f(\mathbf{x})$ as in eq.
-(\ref{approxf}).
+In the HSGP model the latent function $f(\mathbf{x})$ is approximated as in eq.
+(\ref{approxf}), with the square exponential spectral density as in eq.
+(\ref{specdens}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in eq.
+(\ref{bf&lambda}).
 
@@
 
-For the HSGP model we use $m=20$ basis functions and a boundary factor $c=1.5$. For the splines model we use 20 knots or basis in the model.
-
-Figure \ref{fig12_Posteriors_gaydata} shows the posteriors distributions of the three models, the regular GP, the HSGP and the splines, jointly with the true data-generative function and the noisy observations.
-The HSGP model posterior shown in the figure corresponds to 20 basis functions and a boundary factor of 1.5.
-For the spline posterior shown in the figure, it correspond to a 20 knots.
-Two small sets of observations have been used as test data, plotted as crosses in the Figure.
+Figure \ref{fig12_Posteriors_gaydata} shows the posteriors distributions of the three models, the regular GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the splines model with 20 knots.
+The true data-generative function and the noisy observations are also plotted in the Figure.
+Two small sets of observations have been used as test data, and plotted as crosses in the Figure.
 
@@
 
-For the HSGP model, different models with different number of basis functions and boundary factors have been used.
+For the HSGP model, different models with different number of basis functions and boundary factors have been fitted.
 The root mean square errors (RMSE) for every one of these models have been computed and plotted as a function of the number of basis functions and the boundary factor in Figure \ref{fig13_MSE_train_BF_gaydata}, for training (left) and test (right) data.
 In this case, the RMSE has been computed against the regular GP model.
-The expected patterns of the approximation in function of the number of basis function and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
+The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
 
@@
 \subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
+\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
   \label{fig13_MSE_train_BF_gaydata}
@@
 Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the different models, regular GP, HSGP and splines, against the actual data, for training and test data.
-Again we can see how the splines models does not extrapolate data properly.
+We can see how the splines models does not extrapolate data properly.
 
@@
 \subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
+\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
   \label{fig14_MSE_train_gaydata}
@@
 
+\section{Multidimensional generalization of covariance functions and spectral densities}
+
+\begin{itemize}
+\item 
+$x, x' \in \mathbb{R}$,\\
+$\ell \in \mathbb{R}$, \\
+$\tau=x-x'\in \mathbb{R}$, \\
+$r=||x-x'||\in \mathbb{R}$, \\
+$\tau=r$ \\
+$\omega\in \mathbb{R}$
+
+%
+\begin{align*}
+k_{\nu=\infty}(\tau)&=\exp(-\frac{1}{2} \tau^2/\ell^2) \\
+k_{\nu=\frac{1}{2}}(\tau)&=\exp(-\tau/\ell) \\
+k_{\nu=\frac{3}{2}}(\tau)&=(1+\sqrt{3}\tau/\ell) \exp(-\sqrt{3}\tau/\ell)  \\
+\\
+S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right) \\
+S_{\nu=\frac{1}{2}}(w)&= 2\sigma^2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1} \\
+S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}
+\end{align*}
+
+\item 
+$\mathbf{x}, \mathbf{x}' \in \mathbb{R}^2$, \\
+$\ell \in \mathbb{R}$, \\
+$\bm{\tau}=\mathbf{x}- \mathbf{x}'=\{\tau_1,\tau_2 \}\in \mathbb{R}^2$, \\
+$\bm{\omega}=\{\omega_1,\omega_2 \}\in \mathbb{R}^2$
+
+%
+\begin{align*}
+k_{\nu=\infty}(\bm{\tau})&=\exp(-\frac{1}{2} \tau_1^2/\ell^2) \cdot \exp(-\frac{1}{2} \tau_2^2/\ell^2), \\
+k_{\nu=\frac{1}{2}}(\bm{\tau})&=\exp(-\tau_1/\ell) \cdot \exp(-\tau_2/\ell), \\
+k_{\nu=\frac{3}{2}}(\bm{\tau})&=(1+\sqrt{3}\tau_1/\ell) \exp(-\sqrt{3}\tau_1/\ell) \cdot (1+\sqrt{3}\tau_2/\ell) \exp(-\sqrt{3}\tau_2/\ell),  \\
+\\
+S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 \omega_2^2\right) \\
+S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega_2^2\right)^{-1} \\
+S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega_2^2\right)^{-2}
+\end{align*}
+
+\item
+$\mathbf{x}, \mathbf{x}' \in \mathbb{R}^2$,\\ 
+$\ell \in \mathbb{R}$, \\
+$r=||\mathbf{x}-\mathbf{x}'||=\sqrt{r_1^2+r_2^2}\in \mathbb{R}$, \\
+$\omega=\sqrt{\omega_1^2+\omega_2^2}\in \mathbb{R}$
+
+%
+\begin{align*}
+k_{\nu=\infty}(r)&=\exp(-\frac{1}{2} r^2/\ell^2), \\
+k_{\nu=\frac{1}{2}}(r)&=\exp(-r/\ell), \\
+k_{\nu=\frac{3}{2}}(r)&=(1+\sqrt{3}r/\ell) \exp(-\sqrt{3}r/\ell),    \\
+\\
+S_{\nu=\infty}(w)&=  \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right) \\
+S_{\nu=\frac{1}{2}}(w)&= 2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1} \\
+S_{\nu=\frac{3}{2}}(w)&= 4 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}
+\end{align*}
+
+\item 
+$\mathbf{x}, \mathbf{x}' \in \mathbb{R}^2$, \\
+$\bm{\ell}=\{\ell_1,\ell_2\} \in \mathbb{R}^2$, \\
+$\bm{\tau}=\mathbf{x}- \mathbf{x}'=\{\tau_1,\tau_2\}\in \mathbb{R}^2$, \\
+$\bm{\omega}=\{\omega_1,\omega_2\}\in \mathbb{R}^2$
+
+%
+\begin{align*}
+k_{\nu=\infty}(\bm{\tau})&=\exp(-\frac{1}{2} \tau_1^2/\ell_1^2) \cdot \exp(-\frac{1}{2} \tau_2^2/\ell_2^2), \\
+k_{\nu=\frac{1}{2}}(\bm{\tau})&=\exp(-\tau_1/\ell_1) \cdot \exp(-\tau_2/\ell_2), \\
+k_{\nu=\frac{3}{2}}(\bm{\tau})&=(1+\sqrt{3}\tau_1/\ell_1) \exp(-\sqrt{3}\tau_1/\ell_1) \cdot (1+\sqrt{3}\tau_2/\ell_2) \exp(-\sqrt{3}\tau_2/\ell_2),  \\
+\\
+S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \omega_2^2\right) \\
+S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell_1}\left(\frac{1}{\ell_1^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell_2}\left(\frac{1}{\ell_2^2} + \omega_2^2\right)^{-1} \\
+S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell_1}^{3}\left(\frac{3}{\ell_1^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell_2}^{3}\left(\frac{3}{\ell_2^2} + \omega_2^2\right)^{-2}
+\end{align*}
+
+\end{itemize}
+
+\newpage
 \vspace{2mm}

### 2019-03-10 Paper/manuscript/manuscript_05.tex
improve Section 3

@@
 
-Implementation and performance of the HSGP model depends on the number of basis functions and boundary settings, which are, at the same time, related  to each other.
-The time of computation in the approximation mainly depends on the number of basis functions.
-Furthermore, the appropriate values for the number of basis functions and boundary depend on the non-linear functional effects of the function characterized by the lengthscale of the covariance function.
-
-In this section we analyze the effects of the number of basis functions and boundary on the approximation accuracy.
-We present recommendations how they should be chosen and diagnostics to check the accuracy of the approximation after the inference.
-We illustrate graphically how the number of basis functions, boundary and lengthscale relate to each other and affect the accuracy of the approximation.
+The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
+Furthermore, appropriate values for these factors will depend on the non-linearity of the estimated function, which is in turn characterized by the lengthscale of the covariance function.
+In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
+We present recommendations on how they should be chosen and diagnostics to check the accuracy of the obtained approximation.
+
 % This graph basically provides the recommended values for these factors in order to get a close approximation.
@@
 
-We compare regular GP and HSGP fitted for different dataset with different "wigglyness".
+%We compare regular GP and HSGP fitted for different dataset with different "wigglyness".
 %The comparison is made through comparing their estimated lengthscales.
@@
 
-The approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, e.g.
-in a 1-$D$ input space $x\in \Omega=[-L,L]\subset \mathbb{R}$:
+As explained in Section 2, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset \mathbb{R}$:
 %
 \begin{equation}\label{diffcov}
-k(\tau) = \sum_{j}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
+k(\tau) = \sum_{j=1}^{\infty} S_{\theta} \left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0),  \nonumber
 \end{equation}
 
-\noindent where $L$ is the boundary, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are those in equation (\ref{bf&lambda}) for the unidimensional case and those in equation (\ref{bf&lambda_multi}) for the multidimensional case.
-
-The approximation to the covariance function becomes exact in the limit when the number of basis functions approach infinity, $j=1,\cdots,\infty$.
-The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions be less than a threshold $\epsilon$,
+\noindent where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equation (\ref{bf&lambda}) for the unidimensional case and in Equation (\ref{bf&lambda_multi}) for the multidimensional case.
+The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
 %
 \begin{eqnarray}\label{diff_covs}
-\int k(\tau)d(\tau) - \int \sum_{j}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
+\int k(\tau) \, d\tau - 
+\int \sum_{j=1}^m S_{\theta}\left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0) \, d\tau < \varepsilon.
 \end{eqnarray}
 
-The finite number $m$ of basis functions in the approximation needed to satisfy eq.
-(\ref{diff_covs}) depends on the non-linear effects of the function to be learned, that is, on its lengthscale $\ell$.
-The approximation also depends on boundary $L$ (equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})) which will affect the performance specially near the boundaries.
-The boundary will also have an effect on the number of basis functions needed in the approximation.
-The boundary $L$ is set as an extension of the desired predicting input domain $\Psi$.
-Considering a zero-mean input domain $\Psi=[-S,S] \in \mathbb{R}$, the boundary is defined as
+The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+The approximation also depends on the boundary $L$ (see Equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})), which will affect its accuracy especially near the boundaries.
+As we wil see later on, $L$ will also influence the number of basis functions required in the approximation.
+In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi$.
+Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \mathbb{R}$.
+We now define $L$ as
 %
@@
 
-\noindent where $S$ represents the half-range of the input space, and $c$ ($c \ge 1$) is the proportional extension factor.
-From now on we are going to refer to this proportional factor $c$ as the boundary factor of the approximation.
+\noindent where $S$ (for $S > 0$) represents the half-range of the input space, and $c$ (for $c \geq 1$) is the proportional extension factor.
+In the following, we will to refer to $c$ as the boundary factor of the approximation.
 The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
 
-For a naive illustration of the effects of these two factors, number of basis functions $m$ and boundary $L$, on the performance of the approximation, a set of noisy observations are drawn from a GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, and input vales within the zero-mean input domain with half-range $S=1$.
-HSGP models with different number of basis functions and boundaries are fitted to this data.
-The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function.
-On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function.
-It can be seen how the number of basis functions affects the non-linear effects of the posteriors.
-Fewer basis functions implies smoother covariance functions and consequently smoother posteriors.
-The more "wigglyness" of the function, the more basis functions are needed.
-On the other hand, and similarly, for a fixed number of basis functions $m$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the boundary $L$ in approaching the posterior mean and the covariance function.
-It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances.
-\\
+We start with an illustration on how the number of basis functions $m$ and boundary factor $c$ influences the accuracy of the HSGP approximations, separately.
+For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ using input values from the zero-mean input domain with half-range $S=1$.
+Several HSGP models with varying $m$ and $L$ are fitted to this data.
+In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictions of the estimated function and on the covariance function itself.
+For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the non-linearity of the estimated function: Fewer basis functions imply smoother covariance functions (Paul: when would you consider a covariance function as smooth?
+Also, I think the main point here is that at long distances the implied covariance function becomes negative which can't be in an exact GP model) and consequently more linear functional forms.
+The higher the "wigglyness" of the function to be estimated, the more basis functions will be required (Paul: do we see this in Figure \ref{fig1_Post_J} already?).
+For $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.
 
@@
 
-In the previous Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} we focused on illustrating the individual effects of the number of basis functions and the boundary.
-Furthermore, for a clearer illustration of these effects, the lengthscale and marginal variance parameters in the HSGP model were fixed to their true values.
-Now we are going to focus on analyzing the interaction effects between these two factors, number of basis functions and the boundary factor, on the performance of the approximation.
-The lengthscale and marginal variance will not be fixed rather they will be estimated either in regular and HSGP models.
+Next, we will focus on analyzing the interaction effects between these $m$ and $c$ on the performance of the approximation.
+The lengthscale and marginal variance will no longer be fixed but rather estimated in both regular GP and HSGP models.
 %The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
 Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
-Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors means and the covariance functions obtained after fitting the data, for different number of basis functions and boundary factor.
-Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model, as a function of the number of basis functions and the boundary factor.
-Figure \ref{fig6_lscale_vs_J} shows the lengthscale and marginal variance for the regular GP model and for the HSGP models as a function of the number of basis functions and boundary factor.
-Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5.
-The choice of 10 basis functions and a boundary factor of 1.5 could be an accurate enough choice.
-This same conclusion can be also be intuitively seen in the posterior mean and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}.
-
-From Figure \ref{fig5_MSE_vs_J}, a behaviour in terms of performance of the HSGP model as a function of the number of basis functions and the boundary factor, can be deduced: \\
-$\bullet$ as the boundary factor increases, more basis functions are needed,
-\\
-$\bullet$ as fewer basis functions are used, the boundary factor must decrease.
-
-Similar conclusions can be stated from results of Figure \ref{fig6_lscale_vs_J}.
-This Figure shows the estimated lengthscale and marginal variance as function of the number of basis functions and boundary factor.
-In addition to the above conclusions, in Figure \ref{fig6_lscale_vs_J} it can be easily seen that exists a minimum value for the boundary factor under which a close approximation will never be achieved.
-\\
+Figure \ref{fig4_Post_&_Cov_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$.
+Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model.
+Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models.
+Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Further, the choice of $m = 10$ and $c = 1.5$ could still be an accurate enough choice.
+We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}.
+From these results, some general conclusions may be drawn:
+
+\begin{itemize}
+\item As $c$ increases, $m$ has to increase as well (and vice versa).
+\item There exists a minimum $c$ under which a close approximation will never be achieved regardless of $m$.
+\end{itemize}
 
@@
 
-Additionally, there exit a relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the function.
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in relation to the performance of the approximation evaluated using eq.
-(\ref{diff_covs}).
+Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
+Figure \ref{fig3_lscale_vs_J_vs_c_part1} depicts how these three factors interact with each other in relation to the performance of the HSGP approximation (Paul: Did you say how you choose $\varepsilon$?).
 On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data $\ell/S$.
 The countour lines gather the boundary factor $c$.
-And the Y-axis represents the number of basis functions.
-Thus, this figure give us, for a certain GP function with lengthscale $\ell$ and given a boundary factor $c$, the minimum number of basis functions needed to achieve a close approximation in terms of satisfying eq.
-(\ref{diff_covs}).
-Alternatively, this figure could also be read as the minimum boundary factor $c$ that we should use given a number of basis functions, for certain GP function with certain lengthscale.
-And also this figure could also be read as the minimum functional lengthscales that can be closely approximated given a number of basis functions and a boundary factor.
-This figure basically collects the behaviour of the HSGP model as a function of the lengthscale, number of basis functions and boundary factor:
-
-$\bullet$ As the lengthscale of the function increases, the boundary factor and the number of basis functions needed for the approximation decrease.
-
-$\bullet$ There is a minimum boundary factor under which a close approximation never is going to be achieved (The contour lines have an end in function of the lengthscale).
-This statement match with one of the behaviour recognized in Figure \ref{fig6_lscale_vs_J} where were a minimum value for the boundary factor under which a close approximation will never be achieved.
-
-$\bullet$ The lower the boundary factor, the fewer basis functions and the smaller lengthscales that are suitable to be fitted.
-
-$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimum that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased.
-On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.
-
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $m$ and the boundary factor $c$ to approximate certain process characterized by its lengthscale.
-This Figure will allow us to optimize computations which depends on the number of basis functions.
-Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit.
-This figure also provides a diagnosis tool of the fit by comparing a actual estimated lengthscale to that the figure bring.
+And the Y-axis represents the number of basis functions (Paul: This needs to be put in the figure caption not in the text).
+More precisely, for a given GP function with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig3_lscale_vs_J_vs_c_part1} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
+Alternatively, this figure could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
+Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$.
+We obtain the following main conclusions:
+
+\begin{itemize}
+\item As $\ell$ increases, $c$ and $m$ required for a close enough approximation decrease.
+\item There is a minimum $c$ under which a close approximation is never going to be achieved (The contour lines have an end in function of the lengthscale; Paul: what do you mean?).
+\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve a close approximation.
+\end{itemize}
+
+For users of HSGP models, Figure \ref{fig3_lscale_vs_J_vs_c_part1} also severs as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ is exceeds the corresponding line provided in Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+If $\hat{\ell}$ exceeds the line, the approximation should be close enough.
+If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased (Paul: decreasing $c$ seems dangerous to me if predictions for out of sample data are desired.
+Shell we mention this?).
+We may also use this diagnostic in a iterative procedure.
+Starting from some initial guess of $\ell$, we can fit an HSGP model, check the approximation accuracy, and, if not accurate enough, repeat the process while replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.
 
@@
 
-If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J} where 10 basis functions and a boundary factor of 1.5 would be enough to closely approximate a function with normalized lengthscale equal to 0.3, we can recognize that these conclusions also matches the conclusion brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-\subsection{Comparative analysis between the lengthscales estimated by a regular GP model and approximate GP model}
-
-We make a comparative analysis between the lengthscales estimated by the regular GP model and the approximate GP model in several datasets with different "wigglyness" (or lengthscales).
-Basically, comparing the lengthscales we can recognize if an approximation is close enough to the regular GP.
-We contrast these results with the information brought from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-For this analysis, we fit the regular GP model and the approximate GP model on different datasets.
-These datasets consists of noisy draws from a GP prior model with a squared exponential covariance function with different lengthscale.
-Different number of basis functions are used to fit the approximate GP model.
-The boundary factor is set in every case to a right and optimal choice.
-Figure \ref{fig7_varing_lscale} shows the posterior means of the regular GP model and the approximate GP model with different number of basis functions, for every dataset.
+If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
+
+\subsection{Comparing lengthscale estimates}
+
+Another way to look at the approximation accuracy of HSGPs is to compare the lengthscale estimates
+obtained from the regular GP and HSGP models. As the lenghscale $\ell$ controls the "wigglyness" of
+the functional relationship, it strongly influences the difficulty of obtaining accurate
+inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
+we can expect the HSGP approximation to be accurate as well.
+
+For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3?) and varying lengthscale values.
+Different values of the number of basis functions $m$ are used when estimating the HSGP models while the boundary factor $c$ is fixed to some reasonable value.
+Figure \ref{fig7_varing_lscale} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets.
 
@@
 
-The comparative analysis between the lengthscales, estimated by a regular GP model and an approximate GP model in the different datasets with different lengthscales, is shown in Figure \ref{fig8_Tlscale_vs_Elscale}.
+The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
 In the X-axes of the plots, the true lengthscales of the data-generative functions for every dataset scenario are represented.
-In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented.
-Different comparisons are made using different number of basis functions in the approximate GP model.
-Let's assume that the boundary factor was properly set in every case.
-The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE of the approximate GP model, computed against the regular GP model, as a function of the lengthscale and number of basis functions.
+In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented (Paul: This needs to go into the Figure captions).
+The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1} (This also needs to go into the Figure caption, but may be worth keeping in the text as well).
+Further, Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE (Paul: you did not define that anywhere before) of the HSGP against the regular GP model as a function of the lengthscale and number of basis functions.
 
@@
 On the other side, big changes on long lengthscales, which might clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small SRMSE.
+
+(Paul: I don't get the last paragraph. Can you rewrite this a little bit yourself and also merge it with the former paragraph? I am happy to look at it again afterwards.)
 

### 2019-03-12 Paper/manuscript/manuscript_05.tex
some edits in intro

@@
 Gaussian processes (GPs) are flexible statistical models for specifying probability distributions over multi-dimensional non-linear functions \citep{rasmussen2006gaussian,neal1997monte}.
-Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate normal.
-GPs themselves are defined by two functions, a mean and a covariance function, with the latter being arguably the more important one.
+Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate Gaussian.
+GPs are defined by a mean and a covariance function.
 The covariance function encodes our prior assumptions about the functional relationship, such as continuity, smoothness, periodicity and scale properties.
 GPs not only allow for non-linear effects but can also implicitly handle interactions between covariates.
-Different types of covariance functions can be used, and they may be combined for further increased flexibility.
+Different types of covariance functions can be combined for further increased flexibility.
 Due to their generality and flexibility, GPs are of broad interest across machine learning and statistics \citep{rasmussen2006gaussian,neal1997monte}.
@@
 
-Given $n$ observations in the data, the computational comlexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
+Given $n$ observations in the data, the computational complexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
 This would limit their application to rather small data sets of a few tens of thousands observations at most.
-The problem becomes especially severe when performing full Bayesian inference via sampling methods, where in each sampling step we need $O(n^3)$ computations when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
-To alleviate these computational demands, several approximate methods have been proposed, which we will briefly review in the following.
+The problem becomes more severe when performing full Bayesian inference via sampling methods, where in each sampling step we need $O(n^3)$ computations when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
+To alleviate these computational demands, several approximate methods have been proposed.
 
@@
 can be found in \cite{rasmussen2006gaussian}.
+\citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for accurate approximation.
 % More recent developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
@@
 Another related method for approximating kernels relies on random fourier features \citep{rahimi2008random,rahimi2009weighted}.
-Further, certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline}, but not all spline models are GPs (e.g., sometimes splines are not marginalization consistent).
+Further, certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}.
+% , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).
 
@@
 In this paper we focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
-Basis function approaches behave computationally like linear models, which is an attractive property in modular probabilistic programming models where there is a big benefit id approximation specific computation is simple (Paul: What do you mean with the the last part of the sentence?).
+Basis function approaches behave computationally like linear models, which is an attractive property in modular probabilistic programming models where there is a big benefit if approximation specific computation is simple (Paul: What do you mean with the the last part of the sentence?).
 The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance kernel including the hyperparameters.
-While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost when only changing the covariance function parameters is $O(mn+m)$ (or even $O(m^3)$ if marginalization can be used).
+While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost when only changing the covariance function parameters is $O(mn+m)$.% (or even $O(m^3)$ if marginalization can be used).
 This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
@@
 This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation.
-The discussed basis function approach also provides an easy way to apply the non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which helps to further improve MCMC efficiency.
+The basis function approach also provides an easy way to apply the non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
 

### 2019-03-12 Paper/manuscript/manuscript_05.tex
edits in section 2

@@
 
+\textcolor{red}{Aki: I moved this from section 2, needs to be combined with other intro}
+We propose an approximate framework for fast and accurate inference
+for Gaussian processes. Using a basis function expansion, we
+approximate the Gaussian process with a linear model. This
+representation has three main advantages: 1) it makes inference
+considerably faster due to the linear structure, 2) it is simple to
+implement, which makes it easy to use Gaussian processes as building
+blocks in more complicated models and can be used as latent function
+in non-Gaussian observational models allowing modelling flexibility,
+3) it can be made arbitrary accurate and the trade-off between
+computational complexity and approximation accuracy can easily be
+controlled.
+
 While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
@@
 
-\subsection{Gaussian process regression}
+\subsection{Gaussian process as a prior}
 
@@
 $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
-Gaussian processes have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.\\
-
-In this work, Gaussian processes will take the role of a prior distribution over function spaces for non-parametric regression in a Bayesian setting. 
-Consider a data set $\mathcal{D} = \left\lbrace \mathbf{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is a noisy observation of an unknown function $f$ at $\mathbf{x}_n \in \mathbb{R}^D$ , i.e.
-$y_n = f(\mathbf{x}_n) + \epsilon_n$.
-Our goal is to predict the value of the function $f^*=f(\mathbf{x}^*)$ evaluated at a new input point $\mathbf{x}^* \in \mathbb{R}^D$.
-That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
+Gaussian processes have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
+
+In this work, Gaussian processes will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. 
+Consider a data set $\mathcal{D} = \left\lbrace \mathbf{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\mathbf{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with Gaussian process prior.
+This generalizes trivially to more complex models depending on several unknown functions, for example such as $p(y_n|f(\mathbf{x}_n),g(\mathbf{x}_n))$ or multilevel models.
+Our goal is to obtain posterior distribution for the value of the function $f^*=f(\mathbf{x}^*)$  evaluated at a new input point $\mathbf{x}^* \in \mathbb{R}^D$.
+% That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
 
@@
 Let $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\mathbf{f}$ is a multivariate Gaussian distribution
-%
-\begin{align*}
-\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu}, \mathbf{K}),
-\end{align*}
-
-\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\mathbf{K}$ the covariance matrix, where $K_{i,j}=k(\mathbf{x}_i,\mathbf{x}_j)$.
+$\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu}, \mathbf{K})$,
+ where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\mathbf{K}$ the covariance matrix, where $K_{i,j}=k(\mathbf{x}_i,\mathbf{x}_j)$.
 The covariance function $k(\textbf{x}, \textbf{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
-The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
+The joint distribution of $\mathbf{f}$ and a new $f^*$ is also a multivariate Gaussian as,
 %
@@
 
-{\color{red} Aki: the following until ``While Gaussian'' seems too basic to be in the paper:}
-
-The joint distribution of the observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
-\begin{align*}
-	p(\mathbf{y}, \mathbf{f}, f^*) = p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)
-\end{align*}
-
-By marginalizing over $\mathbf{f}$ and conditioning on vector of observations $\mathbf{y}$, we obtain the posterior distribution of interest
-%
-\begin{align}\label{eq:posterior}
-	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})},
-\end{align}
-where $p(\mathbf{y})$ is the marginal likelihood and is given by
-%
-\begin{align}\label{eq:evidence}
-	p(\textbf{y}) = \int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*) \text{d}\bm{f}\text{d}f^*.
-\end{align}
-%
-If the observation model $p(\mathbf{y}|\mathbf{f})$ is Gaussian, both integrals in eq.
+% {\color{red} Aki: the following until ``While Gaussian'' seems too basic to be in the paper:}
+
+% The joint distribution of the observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
+% \begin{align*}
+% 	p(\mathbf{y}, \mathbf{f}, f^*) = p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)
+% \end{align*}
+
+% By marginalizing over $\mathbf{f}$ and conditioning on vector of observations $\mathbf{y}$, we obtain the posterior distribution of interest
+% %
+% \begin{align}\label{eq:posterior}
+% 	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})},
+% \end{align}
+% where $p(\mathbf{y})$ is the marginal likelihood and is given by
+% %
+% \begin{align}\label{eq:evidence}
+% 	p(\textbf{y}) = \int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*) \text{d}\bm{f}\text{d}f^*.
+% \end{align}
+% %
+% If the observation model $p(\mathbf{y}|\mathbf{f})$ is Gaussian, both integrals in eq.
 \eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters.
 For example, an isotropic Gaussian likelihood yields the following closed-form solution
-\begin{align*}
-	p(f^*| \mathbf{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\mathbf{f}} (\textbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{f},  k_{f^*,f^*}-\mathbf{k}_{f^*,\mathbf{f}} (\mathbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{k}_{\mathbf{f},f^*}  ),
-\end{align*}
-where $\sigma^2$ is the noise variance.
-
-While Gaussian processes provide an elegant framework for Bayesian non-parametric regression, there are two problems that often arise in practice. 
-in many modelling problems, the observation model is non-Gaussian (e.g.
-Binomial or Poisson regression) and therefore, the posterior distribution $p(f^*| \mathbf{y})$ is analytically intractable and cannot be obtained in closed-form.
-Furthermore, it is often of interest to impose a prior distribution on one or more hyperparameters and then marginalize over the uncertainty.
-This renders the posterior distribution intractable even in the case of a Gaussian likelihood.
-Approximate inference have to be used to tackle the intractable inference problem, typically based on sampling methods such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011} or distributional approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient}.
-Secondly, due to the matrix inverse in eq.~\eqref{eq:posterior} a direct implementation of Gaussian processes scales cubicly in the number of data points, $\mathcal{O}\left(N^3\right)$,  which can be prohibitively slow for many problems of interest.
-\\
-\\
-In the following section, we propose an approximate framework for fast and accurate inference for Gaussian processes.
-Using a basis function expansion, we approximate the Gaussian process with a linear model.
-This representation has three main advantages: 1) it makes inference considerably faster due to the linear structure, 2) it is simple to implement, which makes it easy to use Gaussian processes as building blocks in more complicated models and can be used as latent function in non-Gaussian observational models allowing modelling flexibility, 3) it can be made arbitrary accurate and the trade-off between computational complexity and approximation accuracy can easily be controlled.
-\\
-\\
+% \begin{align*}
+% 	p(f^*| \mathbf{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\mathbf{f}} (\textbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{f},  k_{f^*,f^*}-\mathbf{k}_{f^*,\mathbf{f}} (\mathbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{k}_{\mathbf{f},f^*}  ),
+% \end{align*}
+% where $\sigma^2$ is the noise variance.
+
+If $p(y_n|f(\mathbf{x}_n),\phi)=N(y_n|f(\mathbf{x}_n),\sigma)$ then $\mathbf{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GP and $O(nm^2)$ for sparse GP).
+If $p(y_n|f(\mathbf{x}_n),\phi)=N(y_n|f(\mathbf{x}_n),g(\mathbf{x}_n))$ or $p(y_n|f(\mathbf{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
+%
+Furthermore, if a prior distribution is imposed on $\phi$ and $\theta$ to form a joint posterior for $\phi$, $\theta$ and $\mathbf{f}$,
+approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011} Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} need to be used.
+%
+In this paper we focus on use of MCMC for integrating over the joint
+posterior. MCMC is not usually the fastest approach, but allows
+accurate inference for general models in probabilistic programming
+setting. We consider the computational costs of GPs specifically from
+this point of view.
+
+\subsection{Covariance function}
+
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.

### 2019-03-21 Paper/manuscript/manuscript_05.tex
improve section 3 and figures

--- 2019-03-10 Paper/manuscript/manuscript_05.tex
@@
 
-% This graph basically provides the recommended values for these factors in order to get a close approximation.
-This graph also allows us to diagnose an actual fit by checking whether the value of the estimated lengthscale does not keep to those that the Figure brings as suitable ones.
-
-%We compare regular GP and HSGP fitted for different dataset with different "wigglyness".
-%The comparison is made through comparing their estimated lengthscales.
-
 \subsection{Dependency on the number of basis functions and the boundary condition}
@@
 The approximation also depends on the boundary $L$ (see Equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})), which will affect its accuracy especially near the boundaries.
-As we wil see later on, $L$ will also influence the number of basis functions required in the approximation.
+As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
 In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi$.
@@
 Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictions of the estimated function and on the covariance function itself.
-For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the non-linearity of the estimated function: Fewer basis functions imply smoother covariance functions (Paul: when would you consider a covariance function as smooth?
-Also, I think the main point here is that at long distances the implied covariance function becomes negative which can't be in an exact GP model) and consequently more linear functional forms.
-The higher the "wigglyness" of the function to be estimated, the more basis functions will be required (Paul: do we see this in Figure \ref{fig1_Post_J} already?).
-For $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.
+For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation and the non-linearity of the estimated function, in the sense that fewer basis functions implies inaccurate and smoother effects covariance functions (covariance functions with larger lengthscales) and consequently more linear functional forms, and conversely.
+%:Fewer basis functions imply smoother covariance functions (Paul: when would you consider a covariance function as smooth?
+Also, I think the main point here is that at long distances the implied covariance function becomes negative which can't be in an exact GP model.) and consequently more linear functional forms.
+Then, the higher the "wigglyness" of the function to be estimated, the more basis functions will be required. 
+%(Paul: do we see this in Figure \ref{fig1_Post_J} already?). 
+On the other hand, for $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
-\caption{(left) Posterior distributions for different number of basis functions $m$. (right) Covariance function for different number of basis functions $m$.}
+\subfigure{\includegraphics[scale=0.47, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.47, trim = 25mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
+\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different number of basis functions $m$.
+(right) Covariance functions of both regular GP model and HSGP model for different number of basis functions $m$.}
   \label{fig1_Post_J}
@@
 \centering
-\subfigure{\includegraphics[scale=0.43]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.43]{fig2_Cov_L.pdf}}
-\caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
+\subfigure{\includegraphics[scale=0.47, trim = 0mm 0mm 0mm 10mm, clip]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.47, trim = 25mm 0mm 0mm 10mm, clip]{fig2_Cov_L.pdf}}
+\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$.
+(right) Covariance functions of both regular GP model and HSGP model for different boundary factors $c$.}
   \label{fig2_Post_L}
@@
 The lengthscale and marginal variance will no longer be fixed but rather estimated in both regular GP and HSGP models.
-%The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior.
-Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit.
 Figure \ref{fig4_Post_&_Cov_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$.
-Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model.
+Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP model.
 Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models.
@@
 \centering
-\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part1.pdf}}
-\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
-\caption{Posterior distributions and covariance functions of the regular GP model and the proposed HSGP model as a function of the number of basis functions $m$ and the boundary factor $c$.} 
+\subfigure{\includegraphics[scale=0.41, trim = 0mm 3mm 0mm 0mm, clip]{fig4_Post_&_Cov_part1.pdf}}
+\subfigure{\includegraphics[scale=0.41, trim = 0mm 3mm 0mm 3mm, clip]{fig4_Post_&_Cov_part2.pdf}}
+\subfigure{\includegraphics[scale=0.41, trim = 0mm 0mm 0mm 3mm, clip]{fig4_Post_&_Cov_part3.pdf}}
+\caption{Posterior predictive mean functions and covariance functions of both regular GP model and HSGP model for different number of basis functions $m$ and values of the boundary factor $c$.} 
   \label{fig4_Post_&_Cov_part1}
@@
 \centering
-\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
-\caption{Standardized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model.
-(left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$.
-(right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$.
+\subfigure{\includegraphics[scale=0.44]{fig5_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.44]{fig5_MSE_vs_c.pdf}}
+\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
+(left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$.
+(right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$.
 }
@@
 \centering
-\subfigure{\includegraphics[scale=0.4]{fig6_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
-\caption{Estimated lengthscale and marginal variance parameters of the proposed HSGP models and the regular GP model.
-(left) Estimated lengthscales vs the number of basis functions $m$ and for different values of the boundary factor $c$.
-(right) Estimated marginal variance vs the number of basis functions $m$ and for different values of the boundary factor $c$.}
+\subfigure{\includegraphics[scale=0.48, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.48, trim = 31mm 0mm 0mm 0mm, clip]{fig6_magnitud_vs_J.pdf}}
+\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig6_lscale_vs_J}
@@
 Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figure \ref{fig3_lscale_vs_J_vs_c_part1} depicts how these three factors interact with each other in relation to the performance of the HSGP approximation (Paul: Did you say how you choose $\varepsilon$?).
-On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data $\ell/S$.
-The countour lines gather the boundary factor $c$.
-And the Y-axis represents the number of basis functions (Paul: This needs to be put in the figure caption not in the text).
-More precisely, for a given GP function with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig3_lscale_vs_J_vs_c_part1} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
-Alternatively, this figure could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
+Figure \ref{fig3_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to
+% the performance of the HSGP approximation.
+a close approximation of the HSGP model, in the case of a GP with square exponential covariance function and a single input dimension.
+(Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?).
+More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig3_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
+We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
+%
+\begin{eqnarray*}
+ \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
+\end{eqnarray*}
+
+\noindent Alternatively, this figure could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
 Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$.
@@
 \item As $\ell$ increases, $c$ and $m$ required for a close enough approximation decrease.
-\item There is a minimum $c$ under which a close approximation is never going to be achieved (The contour lines have an end in function of the lengthscale; Paul: what do you mean?).
 \item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve a close approximation.
+\item For a given $\ell$ there exist a minimum $c$ under which a close approximation is never going to be achieved regardless of $m$.
+This fact can be appreciated in the Figure as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 \end{itemize}
 
-For users of HSGP models, Figure \ref{fig3_lscale_vs_J_vs_c_part1} also severs as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ is exceeds the corresponding line provided in Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-If $\hat{\ell}$ exceeds the line, the approximation should be close enough.
-If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased (Paul: decreasing $c$ seems dangerous to me if predictions for out of sample data are desired.
-Shell we mention this?).
-We may also use this diagnostic in a iterative procedure.
-Starting from some initial guess of $\ell$, we can fit an HSGP model, check the approximation accuracy, and, if not accurate enough, repeat the process while replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
-\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
-\caption{Relation between the minimum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
-  \label{fig3_lscale_vs_J_vs_c_part1}
-\end{figure}
-
-If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.
-
-\subsection{Comparing lengthscale estimates}
-
-Another way to look at the approximation accuracy of HSGPs is to compare the lengthscale estimates
-obtained from the regular GP and HSGP models. As the lenghscale $\ell$ controls the "wigglyness" of
+As stated above, Figure \ref{fig3_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$.
+This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
+As the lenghscale $\ell$ controls the "wigglyness" of
 the functional relationship, it strongly influences the difficulty of obtaining accurate
@@
 we can expect the HSGP approximation to be accurate as well.
-
-For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3?) and varying lengthscale values.
-Different values of the number of basis functions $m$ are used when estimating the HSGP models while the boundary factor $c$ is fixed to some reasonable value.
+Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig3_lscale_vs_J_vs_c}.
+If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
+If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased (Paul: decreasing $c$ seems dangerous to me if predictions for out of sample data are desired.
+Shell we mention this?
+Gabi: We assume that predicting points are already within $S$.
+Then, a valid $c$ is more related to the lengthscale than to the predicting points.
+In fact, valid $c$s are restricted in function of the lengthscale: A given lengthscale has a minimum $c$ under which a close approximation will never be achieved regarless $m$ (We stated this in line 502 and also in line 462).
+For example, a given $c$ can not work for certain lengthscale but for a larger one).
+We may also use this diagnostic in a iterative procedure.
+Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig3_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
+Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
+%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.44]{fig3_lscale_vs_J_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.44, trim = 0mm 0mm 0mm 0mm, clip]{fig3_lscale_vs_J_vs_c_zoomin.pdf}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$).
+The right-side plot is a zoom in of the left-side plot.}
+  \label{fig3_lscale_vs_J_vs_c}
+\end{figure}
+
+If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig3_lscale_vs_J_vs_c}.
+
+(Gabi: The following paragraph is just an idea to extend the discussion to the multidimensional case)
+
+Figure \ref{fig3_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending of three variables, $m$, $c$ and $\ell$.
+An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending of four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
+However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig3_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
+
+\subsection{Comparing lengthscale estimates}
+
+In this example, we make a comparison of the lengthscale estimates
+obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig3_lscale_vs_J_vs_c}.
+
+%Another way to look at the approximation accuracy of HSGPs is to compare the lengthscale estimates
+%obtained from the regular GP and HSGP models. As the lenghscale $\ell$ controls the "wigglyness" of
+%the functional relationship, it strongly influences the difficulty of obtaining accurate
+%inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
+%we can expect the HSGP approximation to be accurate as well.
+
+For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3?
+Gabi: I've mentioned it in line 491) and varying lengthscale values.
+Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case.
+
 Figure \ref{fig7_varing_lscale} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets.
+The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
+Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models against the regular GP models as a function of the lengthscale and number of basis functions $m$.
 
@@
 \includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
-\caption{Posterior means of the regular GP and approximate GP models fitted over different datasets with different lengthscales.}
+\caption{Posterior predictive means of both regular GP and HSGP models, fitted over different datasets simulated from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generative function ($True function$).}
   \label{fig7_varing_lscale}
 \end{figure}
-
-The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
-In the X-axes of the plots, the true lengthscales of the data-generative functions for every dataset scenario are represented.
-In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented (Paul: This needs to go into the Figure captions).
-The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1} (This also needs to go into the Figure caption, but may be worth keeping in the text as well).
-Further, Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE (Paul: you did not define that anywhere before) of the HSGP against the regular GP model as a function of the lengthscale and number of basis functions.
-
-We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales.
-Very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high SRMSE.
-On the other side, big changes on long lengthscales, which might clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small SRMSE.
-
-(Paul: I don't get the last paragraph. Can you rewrite this a little bit yourself and also merge it with the former paragraph? I am happy to look at it again afterwards.)
 
@@
 \includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
-\caption{True lengthscales versus estimated lengthscales of the different realizations of datasets using different number of basis functions. }
+\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario.
+The Y-axes represent the lengthscales estimates of GP and HSGP models.
+The use of different $m$ in the HSGP models is represented in separated plots.
+The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
   \label{fig8_Tlscale_vs_Elscale}
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
 \includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
-\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and number of basis functions.}
+\caption{RMSE of the HSGP models for the different datasets with different wiggly effects (lengthscales) and number of basis functions $m$.}
   \label{fig9_MSE_varing_lscale}
 \end{figure}
+
+We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales.
+We observe small RMSE for large lengthscales, however we appreciate mismatch between large lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
+This is because big changes on large lengthscales, which might be clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small RMSE.
+On the other side, we observe high RMSE for small lengthscales, however we appreciate match between small lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
+This is because very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high RMSE.
+
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig3_lscale_vs_J_vs_c}.
+It can be appreciated a complete agreement between those recommended minimum lengthscales that can be properly fitted in every case following Figure \ref{fig3_lscale_vs_J_vs_c} and the real approximate accuracy measured by the RMSE in Figure \ref{fig9_MSE_varing_lscale}.
 

### 2019-03-21 Paper/manuscript/manuscript_05.tex

@@
 This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation.
-The discussed basis function approach also provides an easy way to apply the non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which helps to further improve MCMC efficiency.
+The basis function approach also provides an easy way to apply the non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
+
+\textcolor{red}{Aki: I moved this from section 2, needs to be combined with other intro}
+We propose an approximate framework for fast and accurate inference
+for Gaussian processes. Using a basis function expansion, we
+approximate the Gaussian process with a linear model. This
+representation has three main advantages: 1) it makes inference
+considerably faster due to the linear structure, 2) it is simple to
+implement, which makes it easy to use Gaussian processes as building
+blocks in more complicated models and can be used as latent function
+in non-Gaussian observational models allowing modelling flexibility,
+3) it can be made arbitrary accurate and the trade-off between
+computational complexity and approximation accuracy can easily be
+controlled.
 

### 2019-04-10 Paper/manuscript/manuscript_05.tex
two minor fixes
add 2D simulate study case and improve some figures

@@
 Given $n$ observations in the data, the computational complexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
-This would limit their application to rather small data sets of a few tens of thousands observations at most.
+This limit their application to rather small data sets of a few tens of thousands observations at most.
 The problem becomes more severe when performing full Bayesian inference via sampling methods, where in each sampling step we need $O(n^3)$ computations when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
@@
 %
-\begin{align}\label{eq:specdens}
-S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right), \\
-S_{\nu=\frac{1}{2}}(w)&= 2\sigma^2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1}, \\
-S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}, \\
-%S_{\nu=\frac{5}{2}}(w)&= \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}\left(\frac{5}{\ell^2} + \omega^2\right)^{-3},  
+\begin{align}
+S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right), \label{eq:specdens_inf}  \\
+S_{\nu=\frac{1}{2}}(w)&= 2\sigma^2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1}, \label{eq:specdens_12}  \\
+S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}, \label{eq:specdens_32} \\
+%S_{\nu=\frac{5}{2}}(w)&= \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}\left(\frac{5}{\ell^2} + \omega^2\right)^{-3}, \label{eq:specdens_52} 
 \end{align}
@@
 %
-\begin{align}\label{eq:specdens_multi}
-S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \omega_2^2\right) \\
-S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell_1}\left(\frac{1}{\ell_1^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell_2}\left(\frac{1}{\ell_2^2} + \omega_2^2\right)^{-1} \\
-S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell_1}^{3}\left(\frac{3}{\ell_1^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell_2}^{3}\left(\frac{3}{\ell_2^2} + \omega_2^2\right)^{-2},
+\begin{align}
+S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \omega_2^2\right) \label{eq:specdens_multi_inf} \\
+S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell_1}\left(\frac{1}{\ell_1^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell_2}\left(\frac{1}{\ell_2^2} + \omega_2^2\right)^{-1} \label{eq:specdens_multi_12} \\
+S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell_1}^{3}\left(\frac{3}{\ell_1^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell_2}^{3}\left(\frac{3}{\ell_2^2} + \omega_2^2\right)^{-2}, \label{eq:specdens_multi_32}
 \end{align}
@@
 Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to some boundary condition.
-For brevity, we will refer to these approximate Gaussian processes as HSGPs (Paul: introducing such an abbreviation perhaps shortens what we are going to write in the next sections).
+For brevity, we will refer to these approximate Gaussian processes as HSGPs.
 Below, we will present the main results around HSGPs relevant for practical application.
@@
 
-We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition.
+We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in [-L,L] \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition.
 As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs.
@@
 %
-\begin{eqnarray}\label{bf&lambda}
-\begin{split}
-\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \\
-\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right).
-\end{split}
-\end{eqnarray}
+\begin{align}
+\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \label{eigenvalue}\\
+\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \label{eigenfunction}
+\end{align}
 
@@
 
-\noindent where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x^i)$:
+\noindent where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$:
 %
@@
 In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is possible combinations of univariate eigenfunctions over all dimensions.
-The number $D$-tuples is given by $m^{\ast} = \prod_{d=1}^{D} m_d$, where $m_d$ is the number of basis function for the dimension $d$.
+The number $D$-tuples is given by
+%
+\begin{align} \label{m_multi}
+m^{\ast} = \prod_{d=1}^{D} m_d,
+\end{align}
+
+\noindent where $m_d$ is the number of basis function for the dimension $d$.
 Let $\mathbb{S}\in {\rm I\!N}^{m^{\ast} \times D}$ be the matrix of all those $D$-tuples.
@@
 
-Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\lambda^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$.
-Thus, for $\mathbf{x}^D \in \Omega$ and $j=1,\ldots,m$, we have
-%
-\begin{align} \label{bf&lambda_multi}
-\begin{split}
-\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \\
-%
-\lambda^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. 
-\end{split}
+Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\boldsymbol{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\boldsymbol{\cdotp}}$.
+Thus, for $\mathbf{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have
+%
+\begin{align}
+\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenvalue_multi} \\
+%
+\boldsymbol{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenfunction_multi}
 \end{align}
@@
 k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{m^{\ast}} 
-S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right)
+S^{\ast}_{\theta}\left(\sqrt{\boldsymbol{\lambda}^{\ast}_j}\right)
 \phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
@@
 %
-\begin{equation}\label{approxf}
+\begin{equation}\label{approxf_multi}
 f(\mathbf{x}) \approx \sum_{j=1}^{m^{\ast}} 
-\left( S^{\ast}_{\theta} \left(\sqrt{\lambda^{\ast}_j} \right)\right)^{1/2}
+\left( S^{\ast}_{\theta} \left(\sqrt{\boldsymbol{\lambda}^{\ast}_j} \right)\right)^{1/2}
 \phi^{\ast}_j(\mathbf{x}) \beta_j, 
@@
 \noindent where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equation (\ref{bf&lambda}) for the unidimensional case and in Equation (\ref{bf&lambda_multi}) for the multidimensional case.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equation (\ref{eigenvalue} and \ref{eigenfunction}) for the unidimensional case and in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}) for the multidimensional case.
 The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
@@
 The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
-The approximation also depends on the boundary $L$ (see Equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})), which will affect its accuracy especially near the boundaries.
+The approximation also depends on the boundary $L$ (see Equations (\ref{eigenvalue}), (\ref{eigenfunction}), (\ref{eigenvalue}) and (\ref{eigenfunction}), which will affect its accuracy especially near the boundaries.
 As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
@@
 \centering
-\subfigure{\includegraphics[scale=0.47, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.47, trim = 25mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig1_legend.pdf}}
 \caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different number of basis functions $m$.
@@
 \centering
-\subfigure{\includegraphics[scale=0.47, trim = 0mm 0mm 0mm 10mm, clip]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.47, trim = 25mm 0mm 0mm 10mm, clip]{fig2_Cov_L.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig2_legend.pdf}}
 \caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$.
@@
 \centering
-\subfigure{\includegraphics[scale=0.41, trim = 0mm 3mm 0mm 0mm, clip]{fig4_Post_&_Cov_part1.pdf}}
-\subfigure{\includegraphics[scale=0.41, trim = 0mm 3mm 0mm 3mm, clip]{fig4_Post_&_Cov_part2.pdf}}
-\subfigure{\includegraphics[scale=0.41, trim = 0mm 0mm 0mm 3mm, clip]{fig4_Post_&_Cov_part3.pdf}}
-\caption{Posterior predictive mean functions and covariance functions of both regular GP model and HSGP model for different number of basis functions $m$ and values of the boundary factor $c$.} 
-  \label{fig4_Post_&_Cov_part1}
+\subfigure{\includegraphics[scale=0.30, trim = 10mm 95mm 10mm 30mm, clip]{fig4_legend.pdf}}\\
+\vspace{-1mm}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part1.pdf}}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part1.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part2.pdf}}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part2.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part3.pdf}}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part3.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part4.pdf}}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part4.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part5.pdf}}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part5.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 10mm 10mm 20mm, clip]{fig4_Post_part6.pdf}}
+\subfigure{\includegraphics[scale=0.25, trim = 0mm 10mm 10mm 20mm, clip]{fig4_Cov_part6.pdf}}\\
+\vspace{-3mm}
+\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$.
+(right) Covariance functions of both regular GP model and HSGP model for different boundary factors $c$.}
+  \label{fig4_Post_part1}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[scale=0.44]{fig5_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.44]{fig5_MSE_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_MSE_vs_c.pdf}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
@@
 \centering
-\subfigure{\includegraphics[scale=0.48, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.48, trim = 31mm 0mm 0mm 0mm, clip]{fig6_magnitud_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_magnitud_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig6_legend.pdf}}
 \caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
@@
 
-\vspace{3mm}
-\section{Univariate examples}\label{sec:gp_examples1D}
+\section{Univariate examples}
 
@@
 
-This example consists of a simulated dataset with 250 draws of a Gaussian process prior with a Mat{\'e}rn covariance function with 3/2 degrees of freedom and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, and inputs values $\{\mathbf{x}_i\}_{i=1}^{250} \in \{-1,1\} \subset {\rm I\!R}$.
-To form the final dataset $\mathbf{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
+This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws of a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\mathbf{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
+To form the final noisy dataset $\mathbf{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
 
@@
 \mathbf{y} &= f(\mathbf{x}) + \boldsymbol{\epsilon} \\
-\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  \mathbf{I}) \\
+\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
 f(\mathbf{x}) &\sim \mathcal{GP}(0, k(\mathbf{x}, \mathbf{x}', \theta)),
@@
 
-\noindent where $\mathbf{I}$ represents the identity matrix.
-The previous formulation corresponds to the latent form of a GP model, where $f(\mathbf{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$ (\ref{covfun}).
-Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(\mathbf{x}_i,\mathbf{x}_j,\theta)$.
+\noindent where $I$ represents the identity matrix.
+The previous formulation corresponds to the latent form of a GP model, where $f(\mathbf{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$.
+Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.
 
@@
 \begin{equation*}\label{eq:marginalizedgp_simudata1}
-\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 \mathbf{I} )
+\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 I )
 \end{equation*}
@@
 
-In the HSGP model, the latent function values $f(x)$ are approximated as (\ref{approxf}),
+In the HSGP model, the latent function values $f(\mathbf{x})$ are approximated as (\ref{approxf}),
 %
 \begin{equation*}
-f(x) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j,
+f(\mathbf{x}) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(\mathbf{x}) \beta_j,
 \end{equation*}
 
-\noindent with the spectral density $S$ (\ref{specdens}) as a function of $\sqrt{\lambda_j}$,
+\noindent with the spectral density $S$ (\ref{eq:specdens_32}) as a function of $\sqrt{\lambda_j}$,
 %
@@
 
-\noindent and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ (\ref{bf&lambda})
+\noindent and eigenvalues $\lambda_j$ (\ref{eigenvalue}) and eigenfunctions $\phi_j$ (\ref{eigenfunction})
 %
@@
 \lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \\
-\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right).
+\phi_j(\mathbf{x})&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(\mathbf{x}+L)\right).
 \end{split}
@@
 
-\subsection{Gay data}\label{sec:gay_data}
+\subsection{Gay data}
 This data set relates the proportion of support for same-sex marriage to the age.
@@
 (\ref{approxf}), with the square exponential spectral density as in eq.
-(\ref{specdens}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in eq.
-(\ref{bf&lambda}).
+(\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$ (\ref{eigenvalue}) and eigenfunctions $\phi_j$ as in eq.
+(\ref{eigenfunction}).
 
@@
 
-\subsection{Case II: Birthday data}\label{sec:bf_caseIII}
-
-\vspace{3mm}
-\section{Multivariate examples}\label{sec:gp_examplesMulti}
-\subsection{Simulated data}\label{sec:bf_toyexampleMulti}
+\subsection{Case II: Birthday data}
+
+\section{Multivariate examples}
+\subsection{Simulated data}
+
+This example consists of a simulated dataset with 120 ($i=1,\dots,n=120$) single draws from a Gaussian process prior with two ($D=2$) input dimensions.
+A square exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$ for dimension 1 and $\ell_2=0.35$ for dimension 2, is used.
+The corresponding matrix of input values is $X=\{\mathbf{x}_i\}_{i=1}^{n=120} \in {\rm I\!R}^{n\times 2}$ with $\mathbf{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$.
+Gaussian noise $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\mathbf{y} \in {\rm I\!R}^n$.
+
+The regular GP model with 2 input dimensions, over the outcome variable $\mathbf{y}$ and $n \times 2$ matrix of inputs $X$, can be written as follows,
+%
+\begin{eqnarray*}\label{eq:latentgp_simudata2}
+\begin{split}
+\mathbf{y} &= f(X) + \boldsymbol{\epsilon} \\
+\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
+f(X) &\sim \mathcal{GP}(0, k(X, X', \theta)),
+\end{split}
+\end{eqnarray*}
+
+\noindent where $I$ represents the identity matrix and $k$ the square exponential covariance function with hyperarameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
+The marginalized form, by integrating out the latent values $f(X)$, of the previous latent GP model results
+%
+\begin{equation*}\label{eq:marginalizedgp_simudata2}
+\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
+\end{equation*}
+
+\noindent with covariance matrix $K$, and its element $K_{rs}=k(\mathbf{x}_r,\mathbf{x}_s,\theta)$ is the evaluation of the covariance function $k$ at points $r$ and $s$.
+
+In the HSGP model with 2 input dimensions, the latent function $f$, evaluated at input vector $\mathbf{x} \in {\rm I\!R}^2$, is approximated as (\ref{approxf_multi}),
+%
+\begin{equation*}
+f(\mathbf{x}) \approx \sum_{j}^m \left( S(\sqrt{\boldsymbol{\lambda}_j})\right)^{1/2} \phi_j(\mathbf{x}) \beta_j,
+\end{equation*}
+
+\noindent where $\boldsymbol{\lambda}_j$ is the 2-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+%
+\begin{align*}
+\boldsymbol{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D=2} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D=2},
+\end{align*}
+
+\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}), as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+%
+\begin{align*}
+\phi_j(\mathbf{x}) &= \prod_{d=1}^{D=2} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D=2} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
+\end{align*}
+
+\noindent where $x_d$ is the input value corresponding to dimension $d$.
+$S$ is the spectral density (\ref{eq:specdens_multi_inf}), as a function of $\sqrt{\boldsymbol{\lambda}_j}= \left(\sqrt{\lambda_{\mathbb{S}_{j1}}},\sqrt{\lambda_{\mathbb{S}_{j2}}} \right)$, of the 2-dimensional square exponential covariance function,
+%
+\begin{eqnarray*}
+S(\sqrt{\boldsymbol{\lambda}_j})&=& \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \sqrt{\lambda_{\mathbb{S}_{j1}}}^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \sqrt{\lambda_{\mathbb{S}_{j2}}}^2\right) ,
+\end{eqnarray*}
+
+\noindent In the previous equations, $j$ denotes the index for the $m= m_1 \cdot m_2$ multivariate basis functions (\ref{m_multi}).
+$m_1$ and $m_2$ are the number of basis functions considered for dimension 1 and 2, respectively.
+$\mathbb{S}$ is the matrix of $2$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the two ($D$=2) dimensions.
+$L_d$ is the boundary for the dimension $d$.
+The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$, $\ell_1$ and $\ell_2$ are the marginal variance and lengthscale parameters for dimensions 1 and 2, respectively, of the approximate covariance function.
+
+In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
+
+Figure \ref{fig15_Posteriors_exII} shows the data-generative GP function, from where the dataset was drawn and the mean posterior functions of the three models fitted over the dataset: the regular GP, the HSGP, and the splines models.
+The noisy observations are also plotted in the plots as circles.
+For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
+A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ are used.
+For the splines model, 40 knots in each dimension were used.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}}
+\subfigure{\includegraphics[scale=0.5, trim = 0mm 5mm 0mm 155mm, clip]{fig15_colorbar_exII.pdf}}
+\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
+  \label{fig15_Posteriors_exII}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.36, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.36, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.36, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.34, trim = 150mm 20mm 10mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
+\caption{Mean error between the data-generative function and the GP (a), HSGP (b) and splines (c) models.}
+  \label{fig16_errors_exII}
+\end{figure}
+
+In order to assess performance of the models as a function the number of basis functions and knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
+In all models, the same number of basis functions and knots for each dimension were used.
+Figure \ref{fig11_MSE_exI_inter} shows the root mean squared error (RMSE) as a function of the number of univariate basis functions and knots.
+The RMSE is computed against the true function.
+From Figures \ref{fig16_errors_exII} and \ref{fig17_RMSE_exII}, it can be seen a close approximation of the HSGP model to the regular GP model.
+However, the performance of the splines model is significantly worse.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
+\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation.}
+  \label{fig17_RMSE_exII}
+\end{figure}
+
 \subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
@@
 
+\begin{enumerate}
+\item Exquare exponential covariance function
+
+	\begin{align*}
+	k(\mathbf{r})= \mathrm{exp}(-\frac{1}{2} \frac{|\mathbf{r}|^2}{\ell^2})
+	\end{align*}
+
+	\noindent $\mathbf{x}, \mathbf{x}' \in \mathbb{R}^D$\\
+	$\ell \in \mathbb{R}$\\
+
+	\begin{enumerate}
+	\item
+	$r=||\mathbf{x}-\mathbf{x}'||$\\	
+	$r^2=||\mathbf{x}-\mathbf{x}'||^2= \sqrt{(\mathbf{x}-\mathbf{x}')(\mathbf{x}-\mathbf{x}')}^2= (\mathbf{x}-\mathbf{x}')(\mathbf{x}-\mathbf{x}')= (x_1-x_1')^2+\cdots+(x_D-x_D')^2$\\
+	\item
+	$\boldsymbol{\tau}=(\mathbf{x}-\mathbf{x}')$\\
+	$\tau^2=(\mathbf{x}-\mathbf{x}')^2= (\mathbf{x}-\mathbf{x}')(\mathbf{x}-\mathbf{x}')=(x_1-x_1')^2+\cdots+(x_D-x_D')^2$\\
+
+	\end{enumerate}
+
+	So, $r^2=\tau^2$.
+
+	\begin{align*}
+	k(||\mathbf{x}-\mathbf{x}'||) &= \mathrm{exp}\left(-\frac{1}{2} \frac{(\mathbf{x}-\mathbf{x}')\dot (\mathbf{x}-\mathbf{x}')}{\ell^2} \right) \\
+	%
+	&= \mathrm{exp}\left(-\frac{1}{2} \frac{\sum_{i=1}^{D}(x_i-x_i')^2}{\ell^2} \right) \\
+	%
+	&=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell^2} \right)
+	\end{align*}
+
+	If we define $z_i=\frac{(x_i-x_i')}{\ell}$, the $D$-dimensional variable $\mathbf{x}-\mathbf{x}'$ is scaled by $\ell$. Let $\mathbf{z}=\{z_i\}_{i=1}^D$,
+
+	\begin{align*}
+	k(\mathbf{z}) &=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}z_i^2 \right)\\
+	%
+	&=\mathrm{exp}\left(-\frac{1}{2} \mathbf{z}^2 \right)
+	\end{align*}
+
+	We can notice that each component $(x_i-x_i')$ can be scaled differently such that $z_i=\frac{(x_i-x_i')}{\ell_i}$, and the kernel is still valid.
+Then, we can redefine $\mathbf{z}$ such as $z_i=\frac{(x_i-x_i')}{\ell_i}$,
+
+	\begin{align*}
+	k(\mathbf{z}) &=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}z_i^2 \right)\\
+	%
+	&=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2} \right),
+	\end{align*}
+
+	\noindent and now we are using a different lengthscale for each dimension. In this way we can generalize to using a different lengthscale for each dimension,\\
+
+	\noindent $\mathbf{x}, \mathbf{x}' \in \mathbb{R}^D$\\
+	$\boldsymbol{\ell} \in \mathbb{R}^D$\\
+
+	\begin{align*}
+	k(||\mathbf{x}-\mathbf{x}'||) &= \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2} \right)
+	\end{align*}
+
+\end{enumerate}
+
+\section{Multidimensional generalization of covariance functions and spectral densities (old version)}
+
 \begin{itemize}

### 2019-04-18 Paper/manuscript/manuscript_05.tex
improve section 2.2

@@
 \usepackage{microtype}
+\usepackage{lscape}
 
@@
 \newtheorem{notation}{Notation}
+
+\newenvironment{changemargin}[2]{%
+\begin{list}{}{%
+\setlength{\topsep}{0pt}%
+\setlength{\leftmargin}{#1}%
+\setlength{\rightmargin}{#2}%
+\setlength{\listparindent}{\parindent}%
+\setlength{\itemindent}{\parindent}%
+\setlength{\parsep}{\parskip}%
+}%
+\item[]}{\end{list}}
 
@@
 In this work, Gaussian processes will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. 
-Consider a data set $\mathcal{D} = \left\lbrace \mathbf{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\mathbf{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with Gaussian process prior.
-This generalizes trivially to more complex models depending on several unknown functions, for example such as $p(y_n|f(\mathbf{x}_n),g(\mathbf{x}_n))$ or multilevel models.
-Our goal is to obtain posterior distribution for the value of the function $f^*=f(\mathbf{x}^*)$  evaluated at a new input point $\mathbf{x}^* \in \mathbb{R}^D$.
+Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with Gaussian process prior.
+This generalizes trivially to more complex models depending on several unknown functions, for example such as $p(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
+Our goal is to obtain posterior distribution for the value of the function $f^*=f(\bm{x}^*)$  evaluated at a new input point $\bm{x}^* \in \mathbb{R}^D$.
 % That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
 
-We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
+We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
 %
 \begin{align*}
- 	\mu(\mathbf{x}) &= \mathbb{E}\!\left[f(\mathbf{x})\right],\\ 
- 	k(\mathbf{x}, \mathbf{x}') &= \mathbb{E}\!\left[\left( f(\mathbf{x}) - \mu(\mathbf{x}) \right)\left( f(\mathbf{x}') - \mu(\mathbf{x}') \right)\right].
+ 	\mu(\bm{x}) &= \mathbb{E}\!\left[f(\bm{x})\right],\\ 
+ 	k(\bm{x}, \bm{x}') &= \mathbb{E}\!\left[\left( f(\bm{x}) - \mu(\bm{x}) \right)\left( f(\bm{x}') - \mu(\bm{x}') \right)\right].
 \end{align*}
@@
 The mean and covariance functions completely characterize the Gaussian process prior, and control the a priori behavior of the function $f$.
-Let $\mathbf{f}=\left\lbrace f(\mathbf{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\mathbf{f}$ is a multivariate Gaussian distribution
-$\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu}, \mathbf{K})$,
- where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\mathbf{K}$ the covariance matrix, where $K_{i,j}=k(\mathbf{x}_i,\mathbf{x}_j)$.
-The covariance function $k(\textbf{x}, \textbf{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
-The joint distribution of $\mathbf{f}$ and a new $f^*$ is also a multivariate Gaussian as,
+Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution
+$\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$,
+ where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
+The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
+The joint distribution of $\bm{f}$ and a new $f^*$ is also a multivariate Gaussian as,
 %
 \begin{align*}
-p(\mathbf{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
-\mathbf{f} \\ 
+p(\bm{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
+\bm{f} \\ 
 f^*
-\end{array} \right] \,\middle|\, \mathbf{0},\left[ \begin{array}{cc}
-\mathbf{K}_{\mathbf{f},\mathbf{f}} & \mathbf{k}_{\mathbf{f},f^*} \\ 
-\mathbf{k}_{f^*,\mathbf{f}} & k_{f^*,f^*}
+\end{array} \right] \,\middle|\, \bm{0},\left[ \begin{array}{cc}
+\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},f^*} \\ 
+\bm{k}_{f^*,\bm{f}} & k_{f^*,f^*}
 \end{array} \right] \right),
@@
 
-\noindent where $\mathbf{k}_{\mathbf{f},f^*}$ is the covariance between $\mathbf{f}$ and $f^*$, and $k_{f^*,f^*}$ is the prior variance of $f^*$.
-%By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\mathbf{f}$ analytically,
+\noindent where $\bm{k}_{\bm{f},f^*}$ is the covariance between $\bm{f}$ and $f^*$, and $k_{f^*,f^*}$ is the prior variance of $f^*$.
+%By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\bm{f}$ analytically,
 % \begin{align*}
-	% p(f^*| \mathbf{f}) = \mathcal{N}(f^*| \mathbf{k}_{f^*,\mathbf{f}} \mathbf{K}_{\mathbf{f},\mathbf{f}}^{-1} \mathbf{f},  k_{f^*,f^*}-\mathbf{k}_{f^*,\mathbf{f}} \mathbf{K}_{\mathbf{f},\mathbf{f}}^{-1} \mathbf{k}_{\mathbf{f},f^*}  )
+	% p(f^*| \bm{f}) = \mathcal{N}(f^*| \bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{k}_{\bm{f},f^*}  )
 % \end{align*}
@@
 
-% The joint distribution of the observations $\mathbf{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\mathbf{f}$ and $f^*$, $p(\mathbf{y}, \mathbf{f}, f^*)$, is the product of the conditional distribution for $\mathbf{y}$ given $\mathbf{f}$ and the joint distribution for $\mathbf{f}$ and $f^*$,
+% The joint distribution of the observations $\bm{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\bm{f}$ and $f^*$, $p(\bm{y}, \bm{f}, f^*)$, is the product of the conditional distribution for $\bm{y}$ given $\bm{f}$ and the joint distribution for $\bm{f}$ and $f^*$,
 % \begin{align*}
-% 	p(\mathbf{y}, \mathbf{f}, f^*) = p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)
+% 	p(\bm{y}, \bm{f}, f^*) = p(\bm{y}|\bm{f})p(\bm{f}, f^*)
 % \end{align*}
 
-% By marginalizing over $\mathbf{f}$ and conditioning on vector of observations $\mathbf{y}$, we obtain the posterior distribution of interest
+% By marginalizing over $\bm{f}$ and conditioning on vector of observations $\bm{y}$, we obtain the posterior distribution of interest
 % %
 % \begin{align}\label{eq:posterior}
-% 	p(f^*|\mathbf{y}) = \frac{\int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*)\text{d}\mathbf{f}}{p(\mathbf{y})},
+% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\text{d}\bm{f}}{p(\bm{y})},
 % \end{align}
-% where $p(\mathbf{y})$ is the marginal likelihood and is given by
+% where $p(\bm{y})$ is the marginal likelihood and is given by
 % %
 % \begin{align}\label{eq:evidence}
-% 	p(\textbf{y}) = \int p(\mathbf{y}|\mathbf{f})p(\mathbf{f}, f^*) \text{d}\bm{f}\text{d}f^*.
+% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \text{d}\bm{f}\text{d}f^*.
 % \end{align}
 % %
-% If the observation model $p(\mathbf{y}|\mathbf{f})$ is Gaussian, both integrals in eq.
+% If the observation model $p(\bm{y}|\bm{f})$ is Gaussian, both integrals in eq.
 \eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters.
@@
 % \begin{align*}
-% 	p(f^*| \mathbf{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\mathbf{f}} (\textbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{f},  k_{f^*,f^*}-\mathbf{k}_{f^*,\mathbf{f}} (\mathbf{K}_{\mathbf{f},\mathbf{f}} + \sigma^2 \mathbf{I})^{-1} \mathbf{k}_{\mathbf{f},f^*}  ),
+% 	p(f^*| \bm{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
 % \end{align*}
@@
 
-If $p(y_n|f(\mathbf{x}_n),\phi)=N(y_n|f(\mathbf{x}_n),\sigma)$ then $\mathbf{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GP and $O(nm^2)$ for sparse GP).
-If $p(y_n|f(\mathbf{x}_n),\phi)=N(y_n|f(\mathbf{x}_n),g(\mathbf{x}_n))$ or $p(y_n|f(\mathbf{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
-%
-Furthermore, if a prior distribution is imposed on $\phi$ and $\theta$ to form a joint posterior for $\phi$, $\theta$ and $\mathbf{f}$,
+If $p(y_n|f(\bm{x}_n),\phi)=N(y_n|f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GP and $O(nm^2)$ for sparse GP).
+If $p(y_n|f(\bm{x}_n),\phi)=N(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
+%
+Furthermore, if a prior distribution is imposed on $\phi$ and $\theta$ to form a joint posterior for $\phi$, $\theta$ and $\bm{f}$,
 approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011} Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} need to be used.
@@
 
-\subsection{Covariance function}
+\subsection{Covariance function and spectral density}
 
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
-A stationary covariance function is a function of $\bm{\tau}=\mathbf{x}-\mathbf{x}' \in \mathbb{R}^D$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
-Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(r), r\in \mathbb{R}$, which means that the covariance is both translation and rotation invariant.
-The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which, in the unidimensional case (i.e.
-$\mathbf{x}=x\in \mathbb{R}$, so $\mathbf{r}=r\in \mathbb{R}$, and $\ell\in \mathbb{R}$ ), are
+A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
+Isotropic covariance functions are those that are function of the distance between observations, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant.
+The most commonly used distance between observations is the norm L2 $(|\bm{x}-\bm{x}'|_{L2})$, also known as Euclidean distance, although other types of distances can be considered.
+
+The Mat\'ern class of isotropic covariance functions is given by, 
+%which, in the unidimensional case (i.e. $\bm{x}=x\in \mathbb{R}$, so $\bm{r}=r\in \mathbb{R}$, and $\ell\in \mathbb{R}$ ), are
 %
 \begin{align*}
-k_{\nu=\infty}(r)&=\sigma^2 \exp(-\frac{1}{2} r^2/\ell^2),  \\
-k_{\nu=\frac{1}{2}}(r)&=\sigma^2 \exp(-r/\ell),  \\
-k_{\nu=\frac{3}{2}}(r)&=\sigma^2(1+\sqrt{3}r/\ell) \exp(-\sqrt{3}r/\ell), 
-%k_{\nu=\frac{5}{2}}(r)&=\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \exp(-\sqrt{5}r/\ell), 
+k_{\nu}(r)&=\frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\nu}K_{\nu}\left(\frac{\sqrt{2\nu}r}{\ell}\right),
 \end{align*}
 
-\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the length-scale and magnitude, respectively, of the kernel.
-The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function.
-These covariance functions can be easily generalized to a $D$-dimensional case, by using the property of separability of multidimensional kernels.
-Let $\mathbf{x}\in \mathbb{R}^D$, $\mathbf{\bm{\tau}}=\{\tau_1,\tau_2\}\in \mathbb{R}^D$ and $\boldsymbol{\ell}=\{\ell_1,\ell_2\}\in \mathbb{R}^D$,
+\noindent where $\nu$ is the order the kernel, $K_{\nu}$ the modified Bessel function, and the $\ell$ and $\sigma$ are the length-scale and magnitude, respectively, of the kernel.
+The particular case where $\nu=\infty$ and $\nu=3/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
 %
 \begin{align*}
-k_{\nu=\infty}(\bm{\tau})&=\exp(-\frac{1}{2} \tau_1^2/\ell_1^2) \cdot \exp(-\frac{1}{2} \tau_2^2/\ell_2^2), \\
-k_{\nu=\frac{1}{2}}(\bm{\tau})&=\exp(-\tau_1/\ell_1) \cdot \exp(-\tau_2/\ell_2), \\
-k_{\nu=\frac{3}{2}}(\bm{\tau})&=(1+\sqrt{3}\tau_1/\ell_1) \exp(-\sqrt{3}\tau_1/\ell_1) \cdot (1+\sqrt{3}\tau_2/\ell_2) \exp(-\sqrt{3}\tau_2/\ell_2) \\
+k_{\infty}(r)&=\sigma^2 \exp\bigg(-\frac{1}{2} \frac{r^2}{\ell^2}\bigg),  \\
+k_{\frac{3}{2}}(r)&=\sigma^2\bigg(1+\frac{\sqrt{3}r}{\ell}\bigg) \exp\bigg(-\frac{\sqrt{3}r}{\ell}\bigg). 
 \end{align*}
 
-\noindent Using a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.
+\noindent The former is commonly known as squared exponential (exponentiated quadratic) covariance function.
+Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form
+%
+\begin{align*}
+k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\exp\bigg(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\bigg),\\
+k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\Bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg)\exp\Bigg(- \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg). \\
+\end{align*}
+
+\noindent Notice that the previous expressions have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
+The use of a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.
 
@@
 If this measure has a density, it is known as the spectral density of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the \textit{Wiener-Khintchine theorem} \citep{rasmussen2006gaussian}.
-In the unidimensional case the spectral density functions associated with the Mat\'ern covariance functions written above are
+The spectral density functions associated with the Mat\'ern class of covariance functions is given by
+%
+\begin{align*}
+S_{\nu}(\omega)&= \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
+\end{align*}
+
+\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ is a distance in the frequency domain, and $\ell$ and $\sigma$ are the lengthscale and magnitud, respectively, of the kernel.
+The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form
 %
 \begin{align}
-S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right), \label{eq:specdens_inf}  \\
-S_{\nu=\frac{1}{2}}(w)&= 2\sigma^2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1}, \label{eq:specdens_12}  \\
-S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}, \label{eq:specdens_32} \\
-%S_{\nu=\frac{5}{2}}(w)&= \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}\left(\frac{5}{\ell^2} + \omega^2\right)^{-3}, \label{eq:specdens_52} 
+S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq:specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq:specdens_32} 
 \end{align}
 
-\noindent where variable $w\in \mathbb{R}$ is in the frequency domain, and $\ell\in \mathbb{R}$.
-By using the separability property of the spectral densities functions, the $D$-dimensional form of these spectral density functions, where $\mathbf{w}=\{w_1,w_2\}$ and $\mathbf{\ell}=\{\ell_1,\ell_2\}$ are
-%
-\begin{align}
-S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \omega_2^2\right) \label{eq:specdens_multi_inf} \\
-S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell_1}\left(\frac{1}{\ell_1^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell_2}\left(\frac{1}{\ell_2^2} + \omega_2^2\right)^{-1} \label{eq:specdens_multi_12} \\
-S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell_1}^{3}\left(\frac{3}{\ell_1^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell_2}^{3}\left(\frac{3}{\ell_2^2} + \omega_2^2\right)^{-2}, \label{eq:specdens_multi_32}
-\end{align}
-
-\noindent which can also be derived from the following expressions \citep{rasmussen2006gaussian}
+\noindent Particularizing to an input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{D=3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{D=3}$, the spectral densities written above take the form
 %
 \begin{align*}
-S^{\ast}_{\nu=\infty}(\mathbf{\omega})&= \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \exp\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{\omega}_d^2\right), \\
-%
-S^{\ast}_{\nu}(\mathbf{\omega})&= \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{\omega}^2 \right)^{(\nu+D/2)}.
+S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^{D=3}  \prod_{i=1}^{D=3} \ell_i  \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D=3} \ell_i^2 s_i^2 \right),   \\
+S_{\frac{3}{2}}(\omega)&= \sigma^2 32\pi\sqrt{3}^3\prod_{i=1}^{D=3}\ell_i\left(3+\sum_{i=1}^{D=3}\ell_i^2 s_i^2 \right)^{-3}.
 \end{align*}
@@
 \begin{equation}
-k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \boldsymbol{\phi}(x)^\intercal \Delta \boldsymbol{\phi}(x'), \nonumber
+k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \bm{\phi}(x'), \nonumber
 \end{equation}
 
-\noindent where $\boldsymbol{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is the diagonal matrix of the spectral densities $S_{\theta}(\sqrt{\lambda_j})$:
+\noindent where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is the diagonal matrix of the spectral densities $S_{\theta}(\sqrt{\lambda_j})$:
 %
@@
 \begin{equation}
-\mathbf{f} \sim \mathcal{N}(\boldsymbol{\mu},\Phi \Delta \Phi^\intercal). \nonumber
+\bm{f} \sim \mathcal{N}(\bm{\mu},\Phi \Delta \Phi^\intercal). \nonumber
 \end{equation}
@@
 
-Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\boldsymbol{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\boldsymbol{\cdotp}}$.
-Thus, for $\mathbf{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have
+Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
+Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have
 %
 \begin{align}
-\phi^{\ast}_j(\mathbf{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenvalue_multi} \\
-%
-\boldsymbol{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenfunction_multi}
+\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenvalue_multi} \\
+%
+\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenfunction_multi}
 \end{align}
@@
 \begin{equation}\label{approxcov_multi}
-k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{m^{\ast}} 
-S^{\ast}_{\theta}\left(\sqrt{\boldsymbol{\lambda}^{\ast}_j}\right)
-\phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
+k(\bm{x},\bm{x}') \approx \sum_{j=1}^{m^{\ast}} 
+S^{\ast}_{\theta}\left(\sqrt{\bm{\lambda}^{\ast}_j}\right)
+\phi^{\ast}_j(\bm{x}) \phi^{\ast}_j(\bm{x}'),
 \end{equation}
@@
 \begin{equation}\label{approxf_multi}
-f(\mathbf{x}) \approx \sum_{j=1}^{m^{\ast}} 
-\left( S^{\ast}_{\theta} \left(\sqrt{\boldsymbol{\lambda}^{\ast}_j} \right)\right)^{1/2}
-\phi^{\ast}_j(\mathbf{x}) \beta_j, 
+f(\bm{x}) \approx \sum_{j=1}^{m^{\ast}} 
+\left( S^{\ast}_{\theta} \left(\sqrt{\bm{\lambda}^{\ast}_j} \right)\right)^{1/2}
+\phi^{\ast}_j(\bm{x}) \beta_j, 
 \end{equation}
@@
 \centering
-\subfigure{\includegraphics[scale=0.44]{fig3_lscale_vs_J_vs_c.pdf}}
-\subfigure{\includegraphics[scale=0.44, trim = 0mm 0mm 0mm 0mm, clip]{fig3_lscale_vs_J_vs_c_zoomin.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig3_lscale_vs_J_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig3_lscale_vs_J_vs_c_zoomin.pdf}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$).
@@
 \centering
-\includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
+\subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
+\vspace{-3mm}
 \caption{Posterior predictive means of both regular GP and HSGP models, fitted over different datasets simulated from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generative function ($True function$).}
-  \label{fig7_varing_lscale}
+  \label{fig7_posterior_varing_lscale_part1}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
+\begin{table}
+\begin{flushleft}
+\begin{tabular}{ c c c c}
+\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 50mm, clip]{fig8_legend.pdf}}\\
+\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
+\end{tabular}
+\end{flushleft}
 \caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario.
@@
   \label{fig8_Tlscale_vs_Elscale}
-\end{figure}
+\end{table}
 
@@
 
-This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws of a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\mathbf{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
-To form the final noisy dataset $\mathbf{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
-
-The regular GP model for fitting this simulated dataset $\mathbf{y}$ can be written as follows,
+This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws of a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
+To form the final noisy dataset $\bm{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
+
+The regular GP model for fitting this simulated dataset $\bm{y}$ can be written as follows,
 %
@@
 \begin{split}
-\mathbf{y} &= f(\mathbf{x}) + \boldsymbol{\epsilon} \\
-\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
-f(\mathbf{x}) &\sim \mathcal{GP}(0, k(\mathbf{x}, \mathbf{x}', \theta)),
+\bm{y} &= f(\bm{x}) + \bm{\epsilon} \\
+\bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
+f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x}, \bm{x}', \theta)),
 \end{split}
@@
 \noindent where $I$ represents the identity matrix.
-The previous formulation corresponds to the latent form of a GP model, where $f(\mathbf{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$.
-Saying that the function values $f(\mathbf{x})$ follows a GP model is equivalent to say that $f(\mathbf{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.
+The previous formulation corresponds to the latent form of a GP model, where $f(\bm{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$.
+Saying that the function values $f(\bm{x})$ follows a GP model is equivalent to say that $f(\bm{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.
 
@@
 \begin{equation*}\label{eq:marginalizedgp_simudata1}
-\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 I )
+\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I )
 \end{equation*}
 
-\noindent where the function values $f(\mathbf{x})$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing time of computation and improving the sampling and the effective number of samples.
-
-In the HSGP model, the latent function values $f(\mathbf{x})$ are approximated as (\ref{approxf}),
+\noindent where the function values $f(\bm{x})$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing time of computation and improving the sampling and the effective number of samples.
+
+In the HSGP model, the latent function values $f(\bm{x})$ are approximated as (\ref{approxf}),
 %
 \begin{equation*}
-f(\mathbf{x}) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(\mathbf{x}) \beta_j,
+f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
 \end{equation*}
@@
 \begin{eqnarray*}
-S(\sqrt{\lambda_j})&=& 4\alpha^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \lambda_j)^{-2}, 
+S(\sqrt{\lambda_j})&=& \alpha^2 \frac{4\sqrt{3}^{3}}{\ell^3}(\frac{3}{\ell^2} + \lambda_j)^{-2}, 
 \end{eqnarray*}
@@
 \lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \\
-\phi_j(\mathbf{x})&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(\mathbf{x}+L)\right).
+\phi_j(\bm{x})&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(\bm{x}+L)\right).
 \end{split}
@@
 \noindent The population per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ or mean number of support people per age group.
-The probabilities $\mathbf{p}=\{p_i\}_{i=1}^{74}$ are modeled by a GP latent function $f(\mathbf{x})$ with a squared exponential covariance function $k$ (2) and age input values $\mathbf{x}=\{x_i\}_{i=1}^{74}$, with the {\it logit} function as a link function,
+The probabilities $\bm{p}=\{p_i\}_{i=1}^{74}$ are modeled by a GP function $f(\bm{x})$ with a squared exponential covariance function $k$ and age input values $\bm{x}=\{x_i\}_{i=1}^{74}$, with the {\it logit} function as a link function,
 %
@@
 p_i &=& \text{logit}(f(x_i)) \nonumber \\
-f(\mathbf{x}) &\sim& \mathcal{GP}(0, k(\mathbf{x},\mathbf{x}', \theta).
-\end{eqnarray*}
-
-In the HSGP model the latent function $f(\mathbf{x})$ is approximated as in eq.
-(\ref{approxf}), with the square exponential spectral density as in eq.
-(\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$ (\ref{eigenvalue}) and eigenfunctions $\phi_j$ as in eq.
-(\ref{eigenfunction}).
+f(\bm{x}) &\sim& \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta).
+\end{eqnarray*}
+
+In the HSGP model the function $f(\bm{x})$ is approximated as in equation (\ref{approxf}), with the square exponential spectral density as in equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
 
@@
 
-\subsection{Case II: Birthday data}
+\subsection{Birthday data}
 
@@
 A square exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$ for dimension 1 and $\ell_2=0.35$ for dimension 2, is used.
-The corresponding matrix of input values is $X=\{\mathbf{x}_i\}_{i=1}^{n=120} \in {\rm I\!R}^{n\times 2}$ with $\mathbf{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$.
-Gaussian noise $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\mathbf{y} \in {\rm I\!R}^n$.
-
-The regular GP model with 2 input dimensions, over the outcome variable $\mathbf{y}$ and $n \times 2$ matrix of inputs $X$, can be written as follows,
+The corresponding matrix of input values is $X=\{\bm{x}_i\}_{i=1}^{n=120} \in {\rm I\!R}^{n\times 2}$ with $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$.
+Gaussian noise $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^n$.
+
+The regular GP model with 2 input dimensions, over the outcome variable $\bm{y}$ and $n \times 2$ matrix of inputs $X$, can be written as follows,
 %
@@
 \begin{split}
-\mathbf{y} &= f(X) + \boldsymbol{\epsilon} \\
-\boldsymbol{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
+\bm{y} &= f(X) + \bm{\epsilon} \\
+\bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
 f(X) &\sim \mathcal{GP}(0, k(X, X', \theta)),
@@
 \begin{equation*}\label{eq:marginalizedgp_simudata2}
-\mathbf{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
+\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
 \end{equation*}
 
-\noindent with covariance matrix $K$, and its element $K_{rs}=k(\mathbf{x}_r,\mathbf{x}_s,\theta)$ is the evaluation of the covariance function $k$ at points $r$ and $s$.
-
-In the HSGP model with 2 input dimensions, the latent function $f$, evaluated at input vector $\mathbf{x} \in {\rm I\!R}^2$, is approximated as (\ref{approxf_multi}),
+\noindent with covariance matrix $K$, and its element $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$ is the evaluation of the covariance function $k$ at points $r$ and $s$.
+
+In the HSGP model with 2 input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^2$, is approximated as (\ref{approxf_multi}),
 %
 \begin{equation*}
-f(\mathbf{x}) \approx \sum_{j}^m \left( S(\sqrt{\boldsymbol{\lambda}_j})\right)^{1/2} \phi_j(\mathbf{x}) \beta_j,
+f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\bm{\lambda}_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
 \end{equation*}
 
-\noindent where $\boldsymbol{\lambda}_j$ is the 2-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent where $\bm{\lambda}_j$ is the 2-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
 \begin{align*}
-\boldsymbol{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D=2} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D=2},
+\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D=2} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D=2},
 \end{align*}
@@
 \begin{align*}
-\phi_j(\mathbf{x}) &= \prod_{d=1}^{D=2} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D=2} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
+\phi_j(\bm{x}) &= \prod_{d=1}^{D=2} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D=2} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
 \end{align*}
@@
 \noindent where $x_d$ is the input value corresponding to dimension $d$.
-$S$ is the spectral density (\ref{eq:specdens_multi_inf}), as a function of $\sqrt{\boldsymbol{\lambda}_j}= \left(\sqrt{\lambda_{\mathbb{S}_{j1}}},\sqrt{\lambda_{\mathbb{S}_{j2}}} \right)$, of the 2-dimensional square exponential covariance function,
-%
-\begin{eqnarray*}
-S(\sqrt{\boldsymbol{\lambda}_j})&=& \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \sqrt{\lambda_{\mathbb{S}_{j1}}}^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \sqrt{\lambda_{\mathbb{S}_{j2}}}^2\right) ,
-\end{eqnarray*}
-
-\noindent In the previous equations, $j$ denotes the index for the $m= m_1 \cdot m_2$ multivariate basis functions (\ref{m_multi}).
-$m_1$ and $m_2$ are the number of basis functions considered for dimension 1 and 2, respectively.
+$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left(\sqrt{\lambda_{\mathbb{S}_{j1}}},\sqrt{\lambda_{\mathbb{S}_{j2}}} \right)$, of the 2-dimensional square exponential covariance function,
+%
+\begin{eqnarray*}
+S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \ell_1 \ell_2  \mathrm{exp}\Bigg(-\frac{1}{2} \bigg( \ell_1^2 \sqrt{\lambda_{\mathbb{S}_{j1}}}^2 + \ell_2^2 \sqrt{\lambda_{\mathbb{S}_{j2}}}^2 \bigg) \Bigg). 
+\end{eqnarray*}
+
+\noindent In the previous equations, $j$ denotes the index for the $m= m_1 \cdot m_2$ multivariate basis functions (\ref{m_multi}), where $m_1$ and $m_2$ are the number of basis functions considered for dimension 1 and 2, respectively.
 $\mathbb{S}$ is the matrix of $2$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the two ($D$=2) dimensions.
@@
 
-The covariance function of a stationary process, that is function of $\boldsymbol{\tau}=\mathbf{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).
+The covariance function of a stationary process, that is function of $\bm{\tau}=\bm{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).
 
@@
 \begin{equation}
-k(\boldsymbol{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \mathbf{s} \cdot \mathbf{\tau}} d\mu(\boldsymbol{\tau}), \nonumber 
+k(\bm{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \bm{s} \cdot \bm{\tau}} d\mu(\bm{\tau}), \nonumber 
 \end{equation}
@@
 \begin{eqnarray}
-k(\boldsymbol{\tau})&=& \int S(\mathbf{s}) e^{2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber \\
-%
-S(\mathbf{s})&=& \int k(\boldsymbol{\tau}) e^{-2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber
+k(\bm{\tau})&=& \int S(\bm{s}) e^{2\pi i \bm{s} \cdot \bm{\tau}} d\bm{s}  \nonumber \\
+%
+S(\bm{s})&=& \int k(\bm{\tau}) e^{-2\pi i \bm{s} \cdot \bm{\tau}} d\bm{s}  \nonumber
 \end{eqnarray}
@@
 
-Associated to each covariance function $k(\mathbf{x},\mathbf{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
+Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
 %
 \begin{equation}
-\mathcal{K} f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}'.
+\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
 \end{equation}
@@
 \begin{equation}
-S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
+S(\bm{w})=a_0+a_1\bm{w}^2+a_2(\bm{w}^2)^2+a_1(\bm{w}^2)^3+\cdots
 \end{equation}
@@
 \begin{equation}
--\nabla^2 f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}',
+-\nabla^2 f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}',
 \end{equation}
@@
 \begin{equation}
-k(\mathbf{x},\mathbf{x}')= \sum_j \lambda_j \phi_j(\mathbf{x}) \phi_j(\mathbf{x}'),
+k(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
 \end{equation}
@@
 
-Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
+Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\bm{x}=\{\bm{x}_1,\bm{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
 The number of new multidimensional eigenfunctions $\phi^{\ast}_j$ and eigenvalues $\lambda^{\ast}_j$ is $J^D=3^2=9$ ($j=\{1,\cdots,J^D\}$).
@@
 \begin{eqnarray}
-\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} \nonumber \\
-%
-\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
-%
-\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber
+\phi^{\ast}_1(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_1 = \{\lambda_{1}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_2(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_2 = \{\lambda_{1}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} \nonumber \\
+%
+\phi^{\ast}_3(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_3 = \{\lambda_{1}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_4(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_4 = \{\lambda_{2}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_5(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_5 = \{\lambda_{2}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_6(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_6 = \{\lambda_{2}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_7(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_7 = \{\lambda_{3}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_8(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_8 = \{\lambda_{3}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\}  \nonumber \\
+%
+\phi^{\ast}_9(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2)  \hspace{2cm}
+%
+\bm{\lambda}^{\ast}_9 = \{\lambda_{3}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\}  \nonumber
 \end{eqnarray}
@@
 \begin{eqnarray}
-\phi^{\ast}_1= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_1= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_2= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_2= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_3= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_3= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_4= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_4= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_5= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_5= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_6= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_6= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_7= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_7= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_8= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_8= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_9= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_9= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_{10}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{10}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_{11}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{11}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_{12}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{12}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber
+\phi^{\ast}_1= \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_1= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_2= \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_2= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_3= \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_3= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_4= \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_4= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_5= \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_5= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_6= \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_6= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_7= \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_7= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_8= \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_8= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_9= \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_9= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_{10}= \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_{10}= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_{11}= \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_{11}= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
+%
+\phi^{\ast}_{12}= \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_{12}= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber
 \end{eqnarray}
 
+\newpage
+\begin{landscape}
+
 \section{Multidimensional generalization of covariance functions and spectral densities}
 
-\begin{enumerate}
-\item Exquare exponential covariance function
-
-	\begin{align*}
-	k(\mathbf{r})= \mathrm{exp}(-\frac{1}{2} \frac{|\mathbf{r}|^2}{\ell^2})
-	\end{align*}
-
-	\noindent $\mathbf{x}, \mathbf{x}' \in \mathbb{R}^D$\\
-	$\ell \in \mathbb{R}$\\
-
-	\begin{enumerate}
-	\item
-	$r=||\mathbf{x}-\mathbf{x}'||$\\	
-	$r^2=||\mathbf{x}-\mathbf{x}'||^2= \sqrt{(\mathbf{x}-\mathbf{x}')(\mathbf{x}-\mathbf{x}')}^2= (\mathbf{x}-\mathbf{x}')(\mathbf{x}-\mathbf{x}')= (x_1-x_1')^2+\cdots+(x_D-x_D')^2$\\
-	\item
-	$\boldsymbol{\tau}=(\mathbf{x}-\mathbf{x}')$\\
-	$\tau^2=(\mathbf{x}-\mathbf{x}')^2= (\mathbf{x}-\mathbf{x}')(\mathbf{x}-\mathbf{x}')=(x_1-x_1')^2+\cdots+(x_D-x_D')^2$\\
-
-	\end{enumerate}
-
-	So, $r^2=\tau^2$.
-
-	\begin{align*}
-	k(||\mathbf{x}-\mathbf{x}'||) &= \mathrm{exp}\left(-\frac{1}{2} \frac{(\mathbf{x}-\mathbf{x}')\dot (\mathbf{x}-\mathbf{x}')}{\ell^2} \right) \\
-	%
-	&= \mathrm{exp}\left(-\frac{1}{2} \frac{\sum_{i=1}^{D}(x_i-x_i')^2}{\ell^2} \right) \\
-	%
-	&=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell^2} \right)
-	\end{align*}
-
-	If we define $z_i=\frac{(x_i-x_i')}{\ell}$, the $D$-dimensional variable $\mathbf{x}-\mathbf{x}'$ is scaled by $\ell$. Let $\mathbf{z}=\{z_i\}_{i=1}^D$,
-
-	\begin{align*}
-	k(\mathbf{z}) &=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}z_i^2 \right)\\
-	%
-	&=\mathrm{exp}\left(-\frac{1}{2} \mathbf{z}^2 \right)
-	\end{align*}
-
-	We can notice that each component $(x_i-x_i')$ can be scaled differently such that $z_i=\frac{(x_i-x_i')}{\ell_i}$, and the kernel is still valid.
-Then, we can redefine $\mathbf{z}$ such as $z_i=\frac{(x_i-x_i')}{\ell_i}$,
-
-	\begin{align*}
-	k(\mathbf{z}) &=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}z_i^2 \right)\\
-	%
-	&=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2} \right),
-	\end{align*}
-
-	\noindent and now we are using a different lengthscale for each dimension. In this way we can generalize to using a different lengthscale for each dimension,\\
-
-	\noindent $\mathbf{x}, \mathbf{x}' \in \mathbb{R}^D$\\
-	$\boldsymbol{\ell} \in \mathbb{R}^D$\\
-
-	\begin{align*}
-	k(||\mathbf{x}-\mathbf{x}'||) &= \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2} \right)
-	\end{align*}
-
-\end{enumerate}
-
-\section{Multidimensional generalization of covariance functions and spectral densities (old version)}
-
-\begin{itemize}
-\item 
-$x, x' \in \mathbb{R}$,\\
-$\ell \in \mathbb{R}$, \\
-$\tau=x-x'\in \mathbb{R}$, \\
-$r=||x-x'||\in \mathbb{R}$, \\
-$\tau=r$ \\
-$\omega\in \mathbb{R}$
-
-%
-\begin{align*}
-k_{\nu=\infty}(\tau)&=\exp(-\frac{1}{2} \tau^2/\ell^2) \\
-k_{\nu=\frac{1}{2}}(\tau)&=\exp(-\tau/\ell) \\
-k_{\nu=\frac{3}{2}}(\tau)&=(1+\sqrt{3}\tau/\ell) \exp(-\sqrt{3}\tau/\ell)  \\
-\\
-S_{\nu=\infty}(w)&= \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right) \\
-S_{\nu=\frac{1}{2}}(w)&= 2\sigma^2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1} \\
-S_{\nu=\frac{3}{2}}(w)&= 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}
-\end{align*}
-
-\item 
-$\mathbf{x}, \mathbf{x}' \in \mathbb{R}^2$, \\
-$\ell \in \mathbb{R}$, \\
-$\bm{\tau}=\mathbf{x}- \mathbf{x}'=\{\tau_1,\tau_2 \}\in \mathbb{R}^2$, \\
-$\bm{\omega}=\{\omega_1,\omega_2 \}\in \mathbb{R}^2$
-
-%
-\begin{align*}
-k_{\nu=\infty}(\bm{\tau})&=\exp(-\frac{1}{2} \tau_1^2/\ell^2) \cdot \exp(-\frac{1}{2} \tau_2^2/\ell^2), \\
-k_{\nu=\frac{1}{2}}(\bm{\tau})&=\exp(-\tau_1/\ell) \cdot \exp(-\tau_2/\ell), \\
-k_{\nu=\frac{3}{2}}(\bm{\tau})&=(1+\sqrt{3}\tau_1/\ell) \exp(-\sqrt{3}\tau_1/\ell) \cdot (1+\sqrt{3}\tau_2/\ell) \exp(-\sqrt{3}\tau_2/\ell),  \\
-\\
-S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 \omega_2^2\right) \\
-S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega_2^2\right)^{-1} \\
-S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega_2^2\right)^{-2}
-\end{align*}
-
-\item
-$\mathbf{x}, \mathbf{x}' \in \mathbb{R}^2$,\\ 
-$\ell \in \mathbb{R}$, \\
-$r=||\mathbf{x}-\mathbf{x}'||=\sqrt{r_1^2+r_2^2}\in \mathbb{R}$, \\
-$\omega=\sqrt{\omega_1^2+\omega_2^2}\in \mathbb{R}$
-
-%
-\begin{align*}
-k_{\nu=\infty}(r)&=\exp(-\frac{1}{2} r^2/\ell^2), \\
-k_{\nu=\frac{1}{2}}(r)&=\exp(-r/\ell), \\
-k_{\nu=\frac{3}{2}}(r)&=(1+\sqrt{3}r/\ell) \exp(-\sqrt{3}r/\ell),    \\
-\\
-S_{\nu=\infty}(w)&=  \sqrt{2\pi} \cdot \ell \cdot \exp\left(-0.5 \ell^2 w^2\right) \\
-S_{\nu=\frac{1}{2}}(w)&= 2 \frac{1}{\ell}\left(\frac{1}{\ell^2} + \omega^2\right)^{-1} \\
-S_{\nu=\frac{3}{2}}(w)&= 4 \frac{\sqrt{3}}{\ell}^{3}\left(\frac{3}{\ell^2} + \omega^2\right)^{-2}
-\end{align*}
-
-\item 
-$\mathbf{x}, \mathbf{x}' \in \mathbb{R}^2$, \\
-$\bm{\ell}=\{\ell_1,\ell_2\} \in \mathbb{R}^2$, \\
-$\bm{\tau}=\mathbf{x}- \mathbf{x}'=\{\tau_1,\tau_2\}\in \mathbb{R}^2$, \\
-$\bm{\omega}=\{\omega_1,\omega_2\}\in \mathbb{R}^2$
-
-%
-\begin{align*}
-k_{\nu=\infty}(\bm{\tau})&=\exp(-\frac{1}{2} \tau_1^2/\ell_1^2) \cdot \exp(-\frac{1}{2} \tau_2^2/\ell_2^2), \\
-k_{\nu=\frac{1}{2}}(\bm{\tau})&=\exp(-\tau_1/\ell_1) \cdot \exp(-\tau_2/\ell_2), \\
-k_{\nu=\frac{3}{2}}(\bm{\tau})&=(1+\sqrt{3}\tau_1/\ell_1) \exp(-\sqrt{3}\tau_1/\ell_1) \cdot (1+\sqrt{3}\tau_2/\ell_2) \exp(-\sqrt{3}\tau_2/\ell_2),  \\
-\\
-S_{\nu=\infty}(\bm{\omega})&= \sqrt{2\pi} \cdot \ell_1 \cdot \exp\left(-0.5 \ell_1^2 \omega_1^2\right) \cdot  \sqrt{2\pi} \cdot \ell_2 \cdot \exp\left(-0.5 \ell_2^2 \omega_2^2\right) \\
-S_{\nu=\frac{1}{2}}(\bm{\omega})&= 2 \frac{1}{\ell_1}\left(\frac{1}{\ell_1^2} + \omega_1^2\right)^{-1} \cdot 2 \frac{1}{\ell_2}\left(\frac{1}{\ell_2^2} + \omega_2^2\right)^{-1} \\
-S_{\nu=\frac{3}{2}}(\bm{\omega})&= 4 \frac{\sqrt{3}}{\ell_1}^{3}\left(\frac{3}{\ell_1^2} + \omega_1^2\right)^{-2} \cdot 4 \frac{\sqrt{3}}{\ell_2}^{3}\left(\frac{3}{\ell_2^2} + \omega_2^2\right)^{-2}
-\end{align*}
-
-\end{itemize}
+\subsection{Square Exponential covariance function ($k$) \textbf and spectral density ($S$)}
+\vspace{-2mm}
+\subsubsection{Using norm-L2 (Euclidean distance)}
+\vspace{-10mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-20mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L2} =|\bm{x}-\bm{x}'|_{L2} = \sqrt{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')} = \sqrt{(x_1-x_1')^2+(x_2-x_2')^2}  = \sqrt{r_1^2+r_2^2} \in \mathbb{R}; \hspace{2mm} k(\tau_{L2},\ell) = \mathrm{exp}\left(-\frac{1}{2} \frac{\tau^2}{\ell^2} \right)\\
+&& \omega_{L2}= \sqrt{s_1^2+s_2^2} \in \mathbb{R}; \hspace{2mm} S(\omega_{L2},\ell) = \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \omega_{L2}^2 \right)
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+       \hline
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		k(\tau_{L2},\ell) &=& k(|\bm{x}-\bm{x}'|_{L2},\ell)\\
+		 &=& \mathrm{exp}\left(-\frac{1}{2} \frac{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')}{\ell^2} \right)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{\sum_{i=1}^{2}(x_i-x_i')^2}{\ell^2} \right) \\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{\small
+         \begin{eqnarray*}
+		S(\omega_{L2},\ell) &=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 (s_1^2 + s_2^2) \right) \\
+		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell^2 s_i^2) \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-SEPARABLE:} \\
+		\\
+		&&k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
+		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		&&S(\omega_{L2},\bm{\ell})= S(s_1,\ell_1)S(s_2,\ell_2)
+         \end{eqnarray*}
+       }\\
+        \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		k(\tau_{L2},\bm{\ell}) &=& k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{\small
+       \begin{eqnarray*}
+		S(\omega_{L2},\bm{\ell}) &=& \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
+	   \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+
+       } \\
+
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&&= \mathrm{exp}\left(-\frac{1}{2} \frac{r_1^2}{\ell_1^2} \right) \mathrm{exp}\left(-\frac{1}{2} \frac{r_2^2}{\ell_2^2} \right) \\
+		&&=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{\small
+         \begin{eqnarray*}
+         &&S(s_1,\ell_1)S(s_2,\ell_2) \\
+		&&= \sqrt{2\pi} \cdot \ell_1 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_1^2 s_1^2 \right) \\
+		 &&\times  \sqrt{2\pi} \cdot \ell_2 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_2^2 s_2^2 \right)\\
+		&&= \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
+         \end{eqnarray*}         
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+
+       } \\
+
+       \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\newpage
+\subsubsection{Using norm-L1 }
+\vspace{-10mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-20mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L1} =|\bm{x}-\bm{x}'|_{L1} = |x_1-x_1'|+|x_2-x_2'| = r_1+r_2 \in \mathbb{R}; \hspace{2mm} k(\tau_{L1},\ell) = \mathrm{exp}\left(-\frac{1}{2} \frac{\tau_{L1}^2}{\ell^2} \right)\\
+&& \omega_{L1}= s_1+s_2 \in \mathbb{R}; \hspace{2mm} S(\omega_{L1},\ell) = \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \omega_{L1}^2 \right)
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+       \hline
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		k(\tau_{L1},\ell) &=& k(|\bm{x}-\bm{x}'|_{L1},\ell)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{(r_1 + r_2)(r_1 + r_2)}{\ell^2} \right)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{(r_1 + r_2)}{\ell}\frac{(r_1 + r_2)}{\ell} \right)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \left(\frac{r_1}{\ell} + \frac{r_2}{\ell}\right)\left(\frac{r_1}{\ell} + \frac{r_2}{\ell}\right) \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		S(\omega_{L1},\ell) &=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} \ell^2 (s_1 + s_2)\\
+		&\cdot & (s_1 + s_2) \Big) \\
+		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} \ell (s_1 + s_2)\\
+		&\cdot & \ell(s_1 + s_2) \Big) \\
+		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} (\ell s_1 + \ell s_2)\\
+		&\cdot &(\ell s_1 + \ell s_2) \Big)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-NO SEPARABLE:} \\
+		\\
+		&&k(|\bm{x}-\bm{x}'|_{L1},\bm{\ell})\\
+		 &&\neq k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		&&S(\omega_{L1},\bm{\ell})\neq S(s_1,\ell_1)S(s_2,\ell_2)
+         \end{eqnarray*}
+       }\\
+       \vspace{-10mm}\\
+        \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		k(\tau_{L1},\bm{\ell}) &=& k(|\bm{x}-\bm{x}'|_{L1},\bm{\ell})\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \left(\frac{r_1}{\ell_1} + \frac{r_2}{\ell_2}\right)\left(\frac{r_1}{\ell_1} + \frac{r_2}{\ell_2}\right) \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		S(\omega_{L1},\bm{\ell}) &=& \sqrt{2\pi}^D \cdot \ell_1 \ell_2 \cdot \mathrm{exp}\Big(-\frac{1}{2} (\ell_1 s_1 + \ell_2 s_2)\\
+		&\cdot &(\ell_1 s_1 + \ell_2 s_2) \Big) \\
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.2cm}|}{
+
+       } \\  
+       \vspace{-15mm}\\
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+        $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&&= \mathrm{exp}\left(-\frac{1}{2} \frac{r_1^2}{\ell_1^2} \right) \mathrm{exp}\left(-\frac{1}{2} \frac{r_2^2}{\ell_2^2} \right)\\
+		&&= \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{2}\frac{r_i^2}{\ell_i^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{
+       \begin{eqnarray*}
+		&&S(s_1,\ell_1)S(s_2,\ell_2) \\
+		&&= \sqrt{2\pi} \cdot \ell_1 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_1^2 s_1^2 \right) \\
+		&& \times \sqrt{2\pi} \cdot \ell_2 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_2^2 s_2^2 \right)\\
+		&&= \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
+		\end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.2cm}|}{
+
+       } \\          
+       \hline\\   
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\newpage	
+\subsubsection{Using vector difference of inputs}
+\vspace{-5mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-10mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\bm{\tau} =\bm{x}-\bm{x}' = (x_1-x_1',x_2-x_2') = (r_1,r_2) \in \mathbb{R}^2; \hspace{2mm} k(\bm{\tau},\ell) = \mathrm{exp}\left(-\frac{1}{2} \frac{\bm{\tau}^2}{\ell^2} \right)\\
+&& \bm{\omega}= (s_1,s_2) \in \mathbb{R}^2; \hspace{2mm} S(\bm{\omega},\ell) = \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \bm{\omega}^2 \right)
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+
+       \hline
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{
+         \begin{eqnarray*}
+		k(\bm{\tau},\ell) &=& k(\bm{x}-\bm{x}',\ell)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{\bm{\tau}^\top \bm{\tau}}{\ell^2} \right)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{(r_1, r_2)^\top (r_1, r_2)}{\ell^2} \right)\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		S(\bm{\omega},\ell) &=& S(\bm{x}-\bm{x}',\ell)\\
+		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \bm{\omega}^\top\bm{\omega} \right) \\
+		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} \ell^2 (s_1, s_2)^\top\\
+		&\cdot& (s_1, s_2) \Big) \\
+		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\ell^2 s_i^2 \right) \\
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-SEPARABLE:} \\
+		\\
+		&&k(\bm{x}-\bm{x}',\bm{\ell})\\
+		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		&&S(\bm{\omega},\bm{\ell})= S(s_1,\ell_1)S(s_2,\ell_2)
+         \end{eqnarray*}
+       } \\
+       \vspace{-15mm}\\
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{7.2cm}|}{
+         \begin{eqnarray*}
+		k(\bm{\tau},\bm{\ell}) &=& k(\bm{x}-\bm{x}',\bm{\ell})\\
+		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		S(\bm{\omega},\ell) &=& \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{
+
+       } \\
+       \vspace{-10mm}\\
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&&= \mathrm{exp}\left(-\frac{1}{2} \frac{r_1^2}{\ell_1^2} \right) \mathrm{exp}\left(-\frac{1}{2} \frac{r_2^2}{\ell_2^2} \right) \\
+		&&= \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{
+         \begin{eqnarray*}
+		&&S(s_1,\ell_1)S(s_2,\ell_2) \\
+		&&= \sqrt{2\pi} \cdot \ell_1 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_1^2 s_1^2 \right) \\
+		&& \times \sqrt{2\pi} \cdot \ell_2 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_2^2 s_2^2 \right)\\
+		&&= \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{
+
+       } \\
+
+       \hline
+
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\newpage	
+\subsection{Matern($\nu=1/2$) covariance function ($k$) \textbf and spectral density ($S$)}
+\vspace{-2mm}
+\subsubsection{Using norm-L2 (Euclidean distance)}
+\vspace{-8mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-20mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L2} =|\bm{x}-\bm{x}'|_{L2} = \sqrt{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')} = \sqrt{(x_1-x_1')^2+(x_2-x_2')^2}  = \sqrt{r_1^2+r_2^2} \in \mathbb{R}; \hspace{2mm} k(\tau_{L2},\ell) = \mathrm{exp}\left(- \frac{\tau_{L2}}{\ell} \right)\\
+&& \omega_{L2}= \sqrt{s_1^2+s_2^2} \in \mathbb{R}; \hspace{2mm} S_{\nu}(\omega_{L2},\ell) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\omega_{L2}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{1/2}(\omega_{L2},\ell) = \frac{2^D\pi^{\frac{D}{2}}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{D+1}{2}}
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+       \hline
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		k(\tau_{L2},\ell) &=& k(|\bm{x}-\bm{x}'|_{L2},\ell)\\
+		 &=& \mathrm{exp}\left(- \frac{\sqrt{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')}}{\ell} \right)\\
+		&=& \mathrm{exp}\left(- \sqrt{\frac{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')}{\ell^2}} \right)\\
+		&=& \mathrm{exp}\left(- \sqrt{\frac{\sum_{i=1}^{2}(x_i-x_i')^2}{\ell^2}} \right) \\
+		&=& \mathrm{exp}\left(-\sqrt{ \sum_{i=1}^{2}\frac{r_i^2}{\ell^2}} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{\small
+         \begin{eqnarray*}
+		D &=& 2\\
+		S_{1/2}(\omega_{L2},\ell) &=& \frac{2\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2\omega_{L2}^2\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2s_{1}^2+\ell^2s_{2}^2\right)^{-\frac{3}{2}} \\
+		\\
+		D &=& 3\\
+		S_{1/2}(\omega_{L2},\ell) &=& \frac{8\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-2} \\
+		&=& 8\pi\ell^3\left(1+\ell^2\omega_{L2}^2\right)^{-2} \\
+		&=& 8\pi\ell^3\left(1+\ell^2s_{1}^2+\ell^2s_{2}^2+\ell^2s_{3}^2\right)^{-2} 
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-NO SEPARABLE:} \\
+		\\
+		&&k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
+		 &&\neq k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		&&S_{1/2}(\omega_{L2},\ell)\neq S_{1/2}(s_1,\ell)S_{1/2}(s_2,\ell)
+         \end{eqnarray*}
+       }\\
+       \vspace{-10mm}\\
+        \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		k(\tau_{L2},\bm{\ell}) &=& k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
+		&=& \mathrm{exp}\left(-\sqrt{\sum_{i=1}^{2}\frac{r_i^2}{\ell_i^2}} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{\small
+         \begin{eqnarray*}
+		D &=& 2\\
+		S_{1/2}(\omega_{L2},\bm{\ell}) &=& 2\pi\ell_1\ell_2 \left( 1+ \sum_{i=1}^{D}\ell_i^2 s_i^2 \right)^{-\frac{3}{2}}\\
+		S_{1/2}(\omega_{L2},\bm{\ell}) &=& 2\pi\prod_{i=1}^{D}\ell_i \left( 1+ \sum_{i=1}^{D}\ell_i^2 s_i^2 \right)^{-\frac{3}{2}}
+         \end{eqnarray*}
+        }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+
+       } \\  
+       \vspace{-8mm}\\
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&=& \mathrm{exp}\left(- \frac{r_1}{\ell_1} \right) \mathrm{exp}\left(- \frac{r_2}{\ell_2} \right) \\
+		&=& \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell_i} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{8.2cm}|}{\small
+         \begin{eqnarray*}
+         &&S_{1/2}(s_1,\ell_1)S_{1/2}(s_2,\ell_2) \\
+		&=& \frac{2}{\ell_1}\left(\frac{1}{\ell_1^2}+s_1^2 \right)^{-1} \cdot \frac{2}{\ell_2}\left(\frac{1}{\ell_2^2}+s_2^2 \right)^{-1} \\
+		%
+		&=& 4\ell_1\ell_2\left(1+\ell_1^2s_1^2 \right)^{-1} \left(1+\ell_2^2s_2^2 \right)^{-1} \\
+         \end{eqnarray*}        
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+
+       } \\
+
+       \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\newpage
+\subsubsection{Using norm-L1}
+\vspace{-5mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-20mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L1} =|\bm{x}-\bm{x}'|_{L1} = |x_1-x_1'|+|x_2-x_2'| = r_1+r_2 \in \mathbb{R}; \hspace{2mm} k(\tau_{L1},\ell) = \mathrm{exp}\left(- \frac{\tau_{L1}}{\ell} \right)\\
+&& \omega_{L1}= s_1+s_2 \in \mathbb{R}; \hspace{2mm} S_{\nu}(\omega_{L2},\ell) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\omega_{L2}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{1/2}(\omega_{L2},\ell) = \frac{2^D\pi^{\frac{D}{2}}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{D+1}{2}}
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+       \hline
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+		k(\tau_{L1},\ell) &=& k(|\bm{x}-\bm{x}'|_{L1},\ell)\\
+		 &=& \mathrm{exp}\left(- \frac{\tau}{\ell} \right) \\
+		&=& \mathrm{exp}\left(- \frac{r_1 + r_2}{\ell} \right)\\
+		&=& \mathrm{exp}\left(- \sum_{i=1}^{D}\frac{r_i}{\ell} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{9.7cm}|}{\small
+         \begin{eqnarray*}
+		D &=& 2\\
+		S_{1/2}(\omega_{L1},\ell) &=& \frac{2\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L1}^2 \right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2\omega_{L1}^2\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2(s_{1}+s_{2})(s_{1}+s_{2})\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+(\ell s_{1}+\ell s_{2})(\ell s_{1}+\ell s_{2})\right)^{-\frac{3}{2}} \\
+		\\
+		D &=& 3\\
+		S_{1/2}(\omega_{L1},\ell) &=& \frac{8\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L1}^2 \right)^{-2} \\
+		&=& 8\pi\ell^3\left(1+\ell^2\omega_{L1}^2\right)^{-2} \\
+		&=& 8\pi\ell^3\left(1+\ell^2(s_{1}+s_{2}+s_{3})(s_{1}+s_{2}+s_{3})\right)^{-2} \\
+		&=& 8\pi\ell^3\left(1+(\ell s_{1}+\ell s_{2}+\ell s_{3})(\ell s_{1}+\ell s_{2}+\ell s_{3})\right)^{-2} 
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-SEPARABLE:} \\
+		\\
+		&&k(|\bm{x}-\bm{x}'|_{L1},\bm{\ell})\\
+		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		 \\
+		 &&\text{IT SHOULD BE SEPARABLE}\\
+		 &&\text{IN THE SPECTRAL DENSITY} \\
+		 &&\text{AS WELL?}\\
+		&&S_{1/2}(\omega_{L1},\ell) \text{ should be equal to }\\
+		 &&S_{1/2}(s_1,\ell)S_{1/2}(s_2,\ell)
+         \end{eqnarray*}
+       }\\
+       \vspace{-10mm}\\
+        \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+		k(\tau_{L1},\bm{\ell}) &=& \mathrm{exp}\left(- \sum_{i=1}^{D}\frac{r_i}{\ell_i} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{9.7cm}|}{\small
+         \begin{eqnarray*}
+         D &=& 2\\
+		S_{1/2}(\omega_{L1},\bm{\ell}) &=& 2\pi\ell_1\ell_2 \left(1+(\ell_1 s_{1}+\ell_2 s_{2})(\ell_1 s_{1}+\ell_2 s_{2})\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell_1\ell_2 \left( 1+ (\ell_1 s_{1}+\ell_2 s_{2})(\ell_1 s_{1}+\ell_2 s_{2}) \right)^{-\frac{3}{2}}
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.7cm}|}{\small
+
+       } \\
+
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&&= \mathrm{exp}\left(- \frac{r_1}{\ell_1} \right) \mathrm{exp}\left(- \frac{r_2}{\ell_2} \right) \\
+		&&= \mathrm{exp}\left(- \sum_{i=1}^{D}\frac{r_i}{\ell_i} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{9.7cm}|}{\small
+         \begin{eqnarray*}
+         &&S_{1/2}(s_1,\ell_1)S_{1/2}(s_2,\ell_2) \\
+		&=& \frac{2}{\ell_1}\left(\frac{1}{\ell_1^2}+s_1^2 \right)^{-1} \cdot \frac{2}{\ell_2}\left(\frac{1}{\ell_2^2}+s_2^2 \right)^{-1} \\
+		%
+		&=& 4\ell_1\ell_2\left(1+\ell_1^2s_1^2 \right)^{-1} \left(1+\ell_2^2s_2^2 \right)^{-1} \\
+         \end{eqnarray*}      
+       }
+
+       & \multicolumn{1}{|p{6.7cm}|}{\small
+
+       } \\
+
+       \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\newpage
+\subsubsection{Using  the vector difference of inputs}
+\vspace{-5mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-20mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\bm{\tau} =\bm{x}-\bm{x}' = (x_1-x_1',x_2-x_2') = (r_1,r_2) \in \mathbb{R}^2; \hspace{2mm} k(\bm{\tau},\ell) = \mathrm{exp}\left(- \frac{\bm{\tau}}{\ell} \right)\\
+&& \bm{\omega}= (s_1,s_2) \in \mathbb{R}^2; \hspace{2mm}S_{\nu}(\bm{\omega},\ell) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\bm{\omega}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{1/2}(\bm{\omega},\ell) = \frac{2^D\pi^{\frac{D}{2}}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^2 \right)^{-\frac{D+1}{2}}
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+       \hline
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+		k(\bm{\tau},\ell) &=& k(\bm{x}-\bm{x}',\ell)\\
+		 &=& \mathrm{exp}\left(- \frac{\bm{\tau}}{\ell} \right) \\
+		&=& \mathrm{exp}\left(- \frac{(r_1, r_2)}{\ell} \right)\\
+		&&\text{(using dot product?)}\\
+		&=& \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell} \right)
+		\end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{9.7cm}|}{\small
+         \begin{eqnarray*}
+		D &=& 2\\
+		S_{1/2}(\bm{\omega},\ell) &=& \frac{2\pi}{\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^2 \right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2\bm{\omega}^2\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2(s_{1},s_{2})^\top(s_{1},s_{2})\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\ell^2(s_{1}^2+s_{2}^2)\right)^{-\frac{3}{2}} \\
+		&=& 2\pi\ell^2\left(1+\sum_{i=1}^{D}\ell^2s_{i}^2\right)^{-\frac{3}{2}} \\
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-SEPARABLE:} \\
+		&&k(\bm{x}-\bm{x}',\bm{\ell})\\
+		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		 &&\text{IT SHOULD BE SEPARABLE}\\
+		 &&\text{IN THE SPECTRAL DENSITY} \\
+		 &&\text{AS WELL?}\\
+		&&S_{1/2}(\bm{\omega},\ell) \text{ should be equal to }\\
+		 &&S_{1/2}(s_1,\ell)S_{1/2}(s_2,\ell)
+         \end{eqnarray*}
+       }\\
+       \vspace{-10mm}\\
+        \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+		k(\bm{\tau},\bm{\ell}) &=& \mathrm{exp}\left(- \frac{\bm{\tau}}{\bm{\ell}} \right) \\
+		&=& \mathrm{exp}\left(- \frac{(r_1, r_2)}{(\ell_1,\ell_2)} \right)\\
+		&&\text{(using dot product?)}\\
+		&=& \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell_i} \right) 
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{9.7cm}|}{\small
+         \begin{eqnarray*}
+		S_{1/2}(\bm{\omega},\bm{\ell}) &=& 2\pi\ell_1\ell_2 \left(1+\sum_{i=1}^{D}\ell_i^2s_{i}^2\right)^{-\frac{3}{2}}
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{5.7cm}|}{\small
+
+       } \\
+
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{5.7cm}|}{\small
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&&= \mathrm{exp}\left(- \frac{r_1}{\ell_1} \right) \mathrm{exp}\left(- \frac{r_2}{\ell_2} \right) \\
+		&&= \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell_i} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{9.7cm}|}{\small
+         \begin{eqnarray*}
+         &&S_{1/2}(s_1,\ell_1)S_{1/2}(s_2,\ell_2) \\
+		&=& \frac{2}{\ell_1}\left(\frac{1}{\ell_1^2}+s_1^2 \right)^{-1} \cdot \frac{2}{\ell_2}\left(\frac{1}{\ell_2^2}+s_2^2 \right)^{-1} \\
+		%
+		&=& 4\ell_1\ell_2\left(1+\ell_1^2s_1^2 \right)^{-1} \left(1+\ell_2^2s_2^2 \right)^{-1} \\
+         \end{eqnarray*}      
+       }
+
+       & \multicolumn{1}{|p{6.7cm}|}{\small
+
+       } \\
+
+       \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\newpage	
+\subsection{Matern($\nu=3/2$) covariance function ($k$) \textbf and spectral density ($S$)}
+\vspace{-2mm}
+\subsubsection{Using norm-L2 (Euclidean distance)}
+\vspace{-8mm}
+
+\begin{table}[H]
+\small
+\begin{eqnarray*}
+\vspace{-20mm}
+&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L2} =|\bm{x}-\bm{x}'|_{L2} = \sqrt{(\bm{x}-\bm{x}')(\bm{x}-\bm{x}')} = \sqrt{(x_1-x_1')^2+(x_2-x_2')^2}  = \sqrt{r_1^2+r_2^2} \in \mathbb{R}; \hspace{2mm} k(\tau_{L2},\ell) = \left(1+\frac{\sqrt{3}\tau}{\ell}\right) \mathrm{exp}\left(- \frac{\sqrt{3}\tau}{\ell} \right)\\
+&& \omega_{L2}= \sqrt{s_1^2+s_2^2} \in \mathbb{R}; \hspace{2mm} S_{\nu}(\omega_{L2}) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\omega_{L2}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{3/2}(\omega_{L2}) = \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{D+3}{2}}
+\end{eqnarray*}
+\normalsize
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}
+       \hline
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\ell \in \mathbb{R}$
+       }
+
+        & \multicolumn{1}{|p{8.2cm}|}{\small
+         \begin{eqnarray*}
+		&&k(\tau_{L2},\ell) = k(|\bm{x}-\bm{x}'|_{L2},\ell)\\
+		&&= \left(1+\frac{\sqrt{3}\sqrt{\sum_{i=1}^{2}r_i^2}}{\ell} \right)\mathrm{exp}\left(-\frac{\sqrt{3}\sqrt{\sum_{i=1}^{2}r_i^2}}{\ell} \right)\\
+		&&= \left(1+\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell^2}} \right)\mathrm{exp}\left(-\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell^2}} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{7.5cm}|}{\small
+         \begin{eqnarray*}
+		D &=& 2\\
+		\\
+		S_{3/2}(\omega_{L2},\ell) &=& \frac{6\pi\sqrt{3}^3}{\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{5}{2}} \\
+		&=& 6\pi\sqrt{3}^3\ell^2\left(3+\ell^2\omega^2 \right)^{-\frac{5}{2}} \\
+		&=& 6\pi\sqrt{3}^3\ell^2\left(3+\ell^2s_1^2+\ell^2s_2^2 \right)^{-\frac{5}{2}} \\
+		\\
+		D &=& 3\\
+		S_{3/2}(\omega_{L2},\ell) &=& 32\pi\sqrt{3}^3\ell^3\left(3+\ell^2\omega^2 \right)^{-3}
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+         \begin{eqnarray*}
+        &&\text{-ISOTROPIC}\\
+        \\
+		&&\text{-NO SEPARABLE:} \\
+		\\
+		&&k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
+		 &&\neq k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
+		 \\
+		&&S_{3/2}(\omega_{L2},\ell)\neq S_{3/2}(s_1,\ell)S_{1/2}(s_2,\ell)
+         \end{eqnarray*} 
+       }\\
+       \vspace{-10mm}\\
+        \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+       }
+
+         & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		&&k(\tau_{L2},\bm{\ell}) = k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
+		&&= \left(1+\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell_i^2}} \right)\mathrm{exp}\left(-\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell_i^2}} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		S_{3/2}(\omega_{L2},\bm{\ell}) &=& 6\pi\sqrt{3}^3\ell_1\ell_2\left(3+\ell_1^2s_1^2+\ell_2^2s_2^2 \right)^{-\frac{5}{2}} \\
+		&=& 6\pi\sqrt{3}^3\prod_{i=1}^{D}\ell_i\left(3+\sum_{i=1}^{D}\ell_i^2s_i^2 \right)^{-\frac{5}{2}} \\
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+
+       } \\  
+       \vspace{-10mm}\\
+       \cline{1-3}
+
+       \multicolumn{1}{|p{1.5cm}|}{
+       \vspace{1mm}
+       $\bm{\ell} \in \mathbb{R}^2$
+
+       {Separable kernel} 
+       }
+
+        & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
+		&&= \left(1+\sqrt{3}\frac{r_1}{\ell_1}\right)\mathrm{exp}\left(-\sqrt{3}\frac{r_1}{\ell_1} \right)\\
+		&&\cdot \left(1+\sqrt{3}\frac{r_2}{\ell_2}\right)\mathrm{exp}\left(-\sqrt{3}\frac{r_2}{\ell_2} \right) \\
+		&&= \left(1+\sqrt{\frac{3r_1^2}{\ell_1^2}}\right)\left(1+\sqrt{\frac{3r_2^2}{\ell_2^2}}\right)\mathrm{exp}\left(-\sum_{i=1}^{2}\sqrt{\frac{3r_i^2}{\ell_i^2}} \right) \\
+		&&= \left(1+\sum_{i=1}^{2}\sqrt{\frac{3r_i^2}{\ell_i^2}}+\frac{3r_1^2r_2^2}{\ell_1^2\ell_2^2}\right)\mathrm{exp}\left(-\sum_{i=1}^{2}\sqrt{\frac{3r_i^2}{\ell_i^2}} \right)
+         \end{eqnarray*}
+       }
+
+       & \multicolumn{1}{|p{7.2cm}|}{\small
+         \begin{eqnarray*}
+         &&S_{3/2}(s_1,\ell_1)S_{3/2}(s_2,\ell_2) \\
+		&=& \frac{4\sqrt{3}^3}{\ell_1^3}\left(\frac{3}{\ell_1^2}+s_1^2 \right)^{-3} \cdot \frac{4\sqrt{3}^3}{\ell_2^3}\left(\frac{3}{\ell_2^2}+s_2^2 \right)^{-3}\\ 
+		&=& 4^2\sqrt{3}^6\bm{\ell}^\top\bm{\ell}\left(3+\ell_1^2s_1^2 \right)^{-2} \left(3+\ell_2^2s_2^2 \right)^{-2}\\        
+		\end{eqnarray*}        
+       }
+
+       & \multicolumn{1}{|p{6.2cm}|}{\small
+
+       } \\
+
+       \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\end{landscape}
 
@@
 #######
-Next we show a little example of how to construct the new set of eigenfunctions and eigenvalues with input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ in a two-dimensional input space ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
+Next we show a little example of how to construct the new set of eigenfunctions and eigenvalues with input vector $\bm{x}=\{\bm{x}_1,\bm{x}_2\}$ in a two-dimensional input space ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
 As a result we obtain sets of $J^D=3^2=9$ new eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$.
@@
 \begin{eqnarray}
-\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
-%
-\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber
+\phi^{\ast}_1(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_1 = \{\lambda_{1}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_2(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_2 = \{\lambda_{1}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_3(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_3 = \{\lambda_{1}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_4(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_4 = \{\lambda_{2}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_5(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_5 = \{\lambda_{2}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_6(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_6 = \{\lambda_{2}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_7(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_7 = \{\lambda_{3}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_8(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_8 = \{\lambda_{3}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
+%
+\phi^{\ast}_9(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_2+L_2)\right) \nonumber \\
+%
+\bm{\lambda}^{\ast}_9 = \{\lambda_{3}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber
 \end{eqnarray}

### 2019-04-20 Paper/manuscript/manuscript_05.tex
improved figures

@@
 \begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.30, trim = 10mm 95mm 10mm 30mm, clip]{fig4_legend.pdf}}\\
-\vspace{-1mm}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part1.pdf}}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part1.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part2.pdf}}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part2.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part3.pdf}}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part3.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part4.pdf}}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part4.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Post_part5.pdf}}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig4_Cov_part5.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 10mm 10mm 20mm, clip]{fig4_Post_part6.pdf}}
-\subfigure{\includegraphics[scale=0.25, trim = 0mm 10mm 10mm 20mm, clip]{fig4_Cov_part6.pdf}}\\
-\vspace{-3mm}
-\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$.
-(right) Covariance functions of both regular GP model and HSGP model for different boundary factors $c$.}
-  \label{fig4_Post_part1}
+\begin{center}
+\begin{tabular}{ c c c}
+\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\
+\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
+\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
+\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
+\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
+\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
+\end{tabular}
+\end{center}
+\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario.
+The Y-axes represent the lengthscales estimates of GP and HSGP models.
+The use of different $m$ in the HSGP models is represented in separated plots.
+The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
+  \label{fig3_Post_part1}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_MSE_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
@@
 }
-  \label{fig5_MSE_vs_J}
+  \label{fig4_MSE_vs_J}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_magnitud_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig6_legend.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig5_legend.pdf}}
 \caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
-  \label{fig6_lscale_vs_J}
+  \label{fig5_lscale_vs_J}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig3_lscale_vs_J_vs_c.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig3_lscale_vs_J_vs_c_zoomin.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$).
 The right-side plot is a zoom in of the left-side plot.}
-  \label{fig3_lscale_vs_J_vs_c}
+  \label{fig6_lscale_vs_J_vs_c}
 \end{figure}
@@
 
-\begin{table}
+\begin{figure}
 \begin{flushleft}
 \begin{tabular}{ c c c c}
-\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 50mm, clip]{fig8_legend.pdf}}\\
+\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 45mm, clip]{fig8_legend.pdf}}\\
 \includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
@@
 The use of different $m$ in the HSGP models is represented in separated plots.
-The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
+The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig6_lscale_vs_J_vs_c}.}
   \label{fig8_Tlscale_vs_Elscale}
-\end{table}
+\end{figure}
 
@@
 \centering
-\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
+\includegraphics[scale=0.35]{fig9_MSE_varing_lscale.pdf}
 \caption{RMSE of the HSGP models for the different datasets with different wiggly effects (lengthscales) and number of basis functions $m$.}
@@
 
-\subsection{Simulated data}
+\subsection{Study case I: Simulated data}
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
-\includegraphics[scale=0.35]{fig10_Posteriors_exI.pdf}
+\includegraphics[width=\textwidth]{fig10_Posteriors_exI.pdf}
 \caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
-\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
-\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_extra.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 60mm 120mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
+\caption{Standarized root mean square error of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation.}
   \label{fig11_MSE_exI_inter}
@@
 
-\subsection{Gay data}
+\subsection{Study case II: Gay data}
 This data set relates the proportion of support for same-sex marriage to the age.
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
-\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_train_BF_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_pred_BF_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig13_legend_gaydata.pdf}}
+\caption{Root mean square error (RMSE) of the different methods against the true function. (left) RMSE for training data. (right) RMSE for test data.}
   \label{fig13_MSE_train_BF_gaydata}
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
-\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 50mm 120mm 0mm, clip]{fig14_legend_gaydata.pdf}}
+\caption{Mean square error (MSE) of the different methods againts the actual data. (left) MSE for training data. (right) MSE for test data.}
   \label{fig14_MSE_train_gaydata}
@@
 
-\subsection{Birthday data}
-
+%\begin{figure}[H]
+%\centering
+%\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
+%\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
+%\caption{Mean square error (MSE) of the different methods againts the actual data. (left) MSE for training data. (right) MSE for test data.}
+%  \label{fig14_MSE_train_gaydata}
+%\end{figure}
+
+\subsection{Study case III: Birthday data}
+
+\vspace{5mm}
 \section{Multivariate examples}
-\subsection{Simulated data}
+\subsection{Study case IV: Simulated data}
 
@@
 \begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}}
-\subfigure{\includegraphics[scale=0.5, trim = 0mm 5mm 0mm 155mm, clip]{fig15_colorbar_exII.pdf}}
+\begin{center}
+\begin{tabular}{ c c c}
+\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 10mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\
+\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
+\end{tabular}
+\end{center}
 \caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
@@
 
+%\begin{figure}
+%\centering
+%\subfigure{\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf}}
+%\subfigure{\includegraphics[scale=0.40, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf}}
+%\subfigure{\includegraphics[scale=0.40, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}}
+%\subfigure{\includegraphics[scale=0.40, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}}
+%\subfigure{\includegraphics[scale=0.5, trim = 0mm 5mm 0mm 155mm, clip]{fig15_colorbar_exII.pdf}}
+%\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
+%  \label{fig15_Posteriors_exII}
+%\end{figure}
+
 \begin{figure}
 \centering
-\subfigure{\includegraphics[scale=0.36, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.36, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.36, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.34, trim = 150mm 20mm 10mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 5mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
 \caption{Mean error between the data-generative function and the GP (a), HSGP (b) and splines (c) models.}
@@
 
-\subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
-\subsection{Case IV: Leukemia data}\label{sec:bf_caseV}
-\subsection{Case V: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
+\subsection{Study case V: Diabetes data}\label{sec:bf_caseV}
+The next example presents an epidemiological study of the diabetes diseases.
+The study aims to relate the probability of suffering from diabetes to some risk factors.
+The data set consists of 392 individuals ($i=1,\cdots,n=392$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed.
+The matrix $X=\{\bm{x}_i\}_{i=1}^{n=392} \in {\rm I\!R}^{n\times 3}$,  with $\bm{x}_i=\{x_{i1},x_{i2},x_{i3}\} \in {\rm I\!R}^{3}$, contains the risk factor values, Glucose ($x_{i1}$), Pregnancy ($x_{i2}$) and Age ($x_{i3}$), per individual $i$,.
+The observational model is a bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
+%
+\begin{equation*}
+y_i \sim \text{Bernoulli}(p_i).
+\end{equation*}
+
+\noindent The goal is to estimate the 
+ probability $p_i$ as a function of the risk factors, which function $f(X)$ is modeled as a Gaussian process with a squared exponential covariance function $k$ depending of the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
+%
+\begin{eqnarray*} \label{eq:gpprior_gay}
+p_i &=& \text{logit}(f(\bm{x}_i)) \nonumber \\
+f(X) &\sim& \mathcal{GP}(0, k(X,X, \theta).
+\end{eqnarray*}
+
+\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale of the GP process.
+Notice that a scalar lengthscale is considered only in the multivariate covariance function.
+
+In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{approxf_multi}),
+%
+\begin{equation*}
+f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\bm{\lambda}_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
+\end{equation*}
+
+\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+%
+\begin{align*}
+\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D},
+\end{align*}
+
+\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}), as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+%
+\begin{align*}
+\phi_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
+\end{align*}
+
+\noindent where $x_d$ is the input value corresponding to dimension $d$.
+$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left\{ \sqrt{\lambda_{\mathbb{S}_{jd}}} \right\}_{d=1}^{D}$, of the $D$-dimensional square exponential covariance function with a scalar lengthscale $\ell$,
+%
+\begin{eqnarray*}
+S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \ell^D  \mathrm{exp}\bigg(-\frac{1}{2} \ell^2 \sum_{d=1}^D  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg). 
+\end{eqnarray*}
+
+\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$.
+$\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the $D$ dimensions.
+$L_d$ is the boundary for the dimension $d$.
+The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale, respectively, of the approximate multivariate covariance function.
+
+In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
+
+In order to analyze the increase of computational time  of the HSGP model with higher dimensions $D$, we perform the study twice, first using two risk factors only, Glucose and Pregnancy, and second using the three risk factors, Glucose, Pregnancy and Age.
+With the HSGP approach in a multivariate input space, choosing the optimal number of basis functions per dimension can be crucial to avoid excessive computational times.
+
+%Figure \ref{fig15_Posteriors_exII} shows the data-generative GP function, from where the dataset was drawn and the mean posterior functions of the three models fitted over the dataset: the regular GP, the HSGP, and the splines models.
+The noisy observations are also plotted in the plots as circles.
+For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
+A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ are used.
+For the splines model, 40 knots in each dimension were used.
+
+\subsection{Study case VI: Leukemia data}\label{sec:bf_caseVI}
+\subsection{Study case VII: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
 

### 2019-04-26 Paper/manuscript/manuscript_05.tex
add Diabetes dataset

@@
 
-%\begin{figure}
-%\centering
-%\subfigure{\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf}}
-%\subfigure{\includegraphics[scale=0.40, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf}}
-%\subfigure{\includegraphics[scale=0.40, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}}
-%\subfigure{\includegraphics[scale=0.40, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}}
-%\subfigure{\includegraphics[scale=0.5, trim = 0mm 5mm 0mm 155mm, clip]{fig15_colorbar_exII.pdf}}
-%\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
-%  \label{fig15_Posteriors_exII}
-%\end{figure}
-
 \begin{figure}
@@
 
-In order to assess performance of the models as a function the number of basis functions and knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
+In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
 In all models, the same number of basis functions and knots for each dimension were used.
-Figure \ref{fig11_MSE_exI_inter} shows the root mean squared error (RMSE) as a function of the number of univariate basis functions and knots.
+Figure \ref{fig17_RMSE_exII}(left) shows the root mean squared error (RMSE) as a function of the boundary factor, the number of univariate basis functions and knots.
 The RMSE is computed against the true function.
-From Figures \ref{fig16_errors_exII} and \ref{fig17_RMSE_exII}, it can be seen a close approximation of the HSGP model to the regular GP model.
+From Figures \ref{fig16_errors_exII} and \ref{fig17_RMSE_exII}(left), it can be seen a close approximation of the HSGP model to the regular GP model.
 However, the performance of the splines model is significantly worse.
+Figure \ref{fig17_RMSE_exII}(right) shows the times of computation of the different models as a function of the boundary factor, number of basis functions and knots.
+From both plots in Figure \ref{fig17_RMSE_exII} it can be noticed that choosing the optimal boundary factor allows for less number of basis functions and less computational time.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
-\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation.}
+\caption{Standarized mean square error (SMSE) of the different methods against the true function (left), and time of computation of the different models (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
   \label{fig17_RMSE_exII}
@@
 The study aims to relate the probability of suffering from diabetes to some risk factors.
-The data set consists of 392 individuals ($i=1,\cdots,n=392$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed.
-The matrix $X=\{\bm{x}_i\}_{i=1}^{n=392} \in {\rm I\!R}^{n\times 3}$,  with $\bm{x}_i=\{x_{i1},x_{i2},x_{i3}\} \in {\rm I\!R}^{3}$, contains the risk factor values, Glucose ($x_{i1}$), Pregnancy ($x_{i2}$) and Age ($x_{i3}$), per individual $i$,.
+The data contains 392 individuals ($i=1,\cdots,n=392$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed.
+The matrix $X=\{\bm{x}_i\}_{i=1}^{n=392} \in {\rm I\!R}^{n\times 4}$,  with $\bm{x}_i=\{x_{i1},x_{i2},x_{i3},x_{i4}\} \in {\rm I\!R}^{4}$, contains the risk factors, \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$.
 The observational model is a bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
@@
 \noindent The goal is to estimate the 
- probability $p_i$ as a function of the risk factors, which function $f(X)$ is modeled as a Gaussian process with a squared exponential covariance function $k$ depending of the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
+ probability $p_i$ as a function of the risk factors, which function $f(X)$ is modeled as a Gaussian process with a squared exponential covariance function $k$ depending on the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
 %
@@
 
-\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale of the GP process.
-Notice that a scalar lengthscale is considered only in the multivariate covariance function.
+\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+Notice that a scalar lengthscale is considered in the multivariate covariance function.
 
@@
 
-\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) which elements are the univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
@@
 
-\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}), as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}) as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
@@
 
-In order to analyze the increase of computational time  of the HSGP model with higher dimensions $D$, we perform the study twice, first using two risk factors only, Glucose and Pregnancy, and second using the three risk factors, Glucose, Pregnancy and Age.
-With the HSGP approach in a multivariate input space, choosing the optimal number of basis functions per dimension can be crucial to avoid excessive computational times.
-
-%Figure \ref{fig15_Posteriors_exII} shows the data-generative GP function, from where the dataset was drawn and the mean posterior functions of the three models fitted over the dataset: the regular GP, the HSGP, and the splines models.
-The noisy observations are also plotted in the plots as circles.
-For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
-A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ are used.
-For the splines model, 40 knots in each dimension were used.
+Figure \ref{fig18_gpfun_diabetes} shows the mean posterior probability ($p_i$) function of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset with the 2 input dimensions ($D=2$) \textit{Glucose} and \textit{Pregnancy}.
+The binary observations $y_i$ are also plotted in the plots as colored points.
+For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions.
+A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used.
+For the splines model, 20 knots in each dimension were used.
+%
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 149mm 18mm 5mm 20mm, clip]{fig18_legend_diabetes.pdf}}
+\caption{(a) Posterior mean function of the GP model. (b) Posterior mean function of the HSGP model. (c) Posterior mean function of the splines (SP) model.}
+  \label{fig18_gpfun_diabetes}
+\end{figure}
+
+In order to assess performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factor for the HSGP model and different number of knots for the splines model have been fitted.
+In all models, the same boundary factor, number of basis functions and knots for each dimension were used.
+Figure \ref{fig19_ELPD_diabetes} shows the expected log predictive density (ELPD) as a function of the boundary factor, the number of univariate basis functions and knots.
+The ELPD is computed over the actual observations.
+Basically, with slightly differences, all models show similar performances, due to the process is very smooth with a estimated lengthscale $\ell=4.51$ very large.
+A slight pattern of improvement of performance can be noticed as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales.
+%
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig19_ELPD_diabetes.pdf}}
+\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$ and knots.}
+  \label{fig19_ELPD_diabetes}
+\end{figure}
+
+Figure \ref{fig20_time2D_diabetes} shows the times of computation in the logarithm scale of the different models fitted over the dataset with 2 input dimensions (left), 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor, number of univariate basis functions and knots.
+We can notice the significant increase of computational time with higher dimensions for the HSGP and splines models.
+This fact shows that choosing optimal values for the number of basis functions and bounfary factor is essential to avoid excessive time of computation.
+It is interesting to be noticed that considering more than 10 knots per dimension in the splines model with 3D is not allowed for an amount of 392 observations.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 0mm 13mm 0mm, clip]{fig20_time2D_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time3D_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time4D_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig20_legend_diabetes.pdf}}
+\caption{Time of computation in the log scale of the different models fitted over the dataset with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+  \label{fig20_time2D_diabetes}
+\end{figure}
 

### 2019-04-27 Paper/manuscript/manuscript_05.tex
fix a couple of figure references

@@
 The lengthscale and marginal variance will no longer be fixed but rather estimated in both regular GP and HSGP models.
-Figure \ref{fig4_Post_&_Cov_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$.
-Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP model.
-Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models.
-Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Figure \ref{fig3_Post_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$.
+Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP model.
+Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models.
+Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
 Further, the choice of $m = 10$ and $c = 1.5$ could still be an accurate enough choice.
-We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}.
+We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig3_Post_part1}.
 From these results, some general conclusions may be drawn:
@@
 Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figure \ref{fig3_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to
+Figure \ref{fig6_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to
 % the performance of the HSGP approximation.
@@
 (Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?).
-More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig3_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
+More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
 We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
@@
 
-As stated above, Figure \ref{fig3_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$.
+As stated above, Figure \ref{fig6_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 we can expect the HSGP approximation to be accurate as well.
-Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig3_lscale_vs_J_vs_c}.
+Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c}.
 If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
@@
 We may also use this diagnostic in a iterative procedure.
-Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig3_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
+Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
 Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
@@
 
-If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig3_lscale_vs_J_vs_c}.
+If we look back to the conclusions drawn from Figures \ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig6_lscale_vs_J_vs_c}.
 
@@
 
-Figure \ref{fig3_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending of three variables, $m$, $c$ and $\ell$.
+Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending of three variables, $m$, $c$ and $\ell$.
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending of four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
-However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig3_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
+However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
 
@@
 In this example, we make a comparison of the lengthscale estimates
-obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig3_lscale_vs_J_vs_c}.
+obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
 
@@
 
-Figure \ref{fig7_varing_lscale} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets.
+Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets.
 The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
@@
 \vspace{-3mm}
-\caption{Posterior predictive means of both regular GP and HSGP models, fitted over different datasets simulated from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generative function ($True function$).}
+\caption{Posterior predictive means of both regular GP and HSGP models, fitted over different datasets simulated from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generative function (\textit{True function}).}
   \label{fig7_posterior_varing_lscale_part1}
@@
 
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig3_lscale_vs_J_vs_c}.
-It can be appreciated a complete agreement between those recommended minimum lengthscales that can be properly fitted in every case following Figure \ref{fig3_lscale_vs_J_vs_c} and the real approximate accuracy measured by the RMSE in Figure \ref{fig9_MSE_varing_lscale}.
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig6_lscale_vs_J_vs_c}.
+It can be appreciated a complete agreement between those recommended minimum lengthscales that can be properly fitted in every case following Figure \ref{fig6_lscale_vs_J_vs_c} and the real approximate accuracy measured by the RMSE in Figure \ref{fig9_MSE_varing_lscale}.
 

### 2019-04-29 Paper/manuscript/manuscript_05.tex
fix some mistakes and make some improvements. Ready for feedback

@@
 
-The parameterization in \ref{approxf} is naturally in the non-centered
+The parameterization in (\ref{approxf}) is naturally in the non-centered
 parameterization form with independent prior distribution on
@@
 In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is possible combinations of univariate eigenfunctions over all dimensions.
-The number $D$-tuples is given by
+The number of $D$-tuples is given by
 %
@@
 
+(Gabi: The following paragraph is a first idea to introduce that the recommendations and diagnostics depends on the kernel considered.
+And that, at the moment, we have built these recommendations only for the square exponential kernel.
+I know that what I wrote in not enough and should be a bit more elaborate; your suggestions where and how elaborate it would be useful.)
+
+Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
+In this work we built these relationships for the square exponential covariance function.
+For other kernels, the relationships will be slightly different.
+
 \subsection{Dependency on the number of basis functions and the boundary condition}
@@
 \noindent where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equation (\ref{eigenvalue} and \ref{eigenfunction}) for the unidimensional case and in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}) for the multidimensional case.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}) for the unidimensional case and in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}) for the multidimensional case.
 The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
@@
 The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
-The approximation also depends on the boundary $L$ (see Equations (\ref{eigenvalue}), (\ref{eigenfunction}), (\ref{eigenvalue}) and (\ref{eigenfunction}), which will affect its accuracy especially near the boundaries.
+The approximation also depends on the boundary $L$ (see Equations (\ref{eigenvalue}), (\ref{eigenfunction}), (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
 As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
@@
 We start with an illustration on how the number of basis functions $m$ and boundary factor $c$ influences the accuracy of the HSGP approximations, separately.
-For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ using input values from the zero-mean input domain with half-range $S=1$.
+For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
 Several HSGP models with varying $m$ and $L$ are fitted to this data.
@@
 \subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig1_legend.pdf}}
-\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different number of basis functions $m$.
-(right) Covariance functions of both regular GP model and HSGP model for different number of basis functions $m$.}
+\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
   \label{fig1_Post_J}
@@
 \subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig2_legend.pdf}}
-\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$.
-(right) Covariance functions of both regular GP model and HSGP model for different boundary factors $c$.}
+\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
   \label{fig2_Post_L}
@@
 \end{center}
-\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario.
-The Y-axes represent the lengthscales estimates of GP and HSGP models.
-The use of different $m$ in the HSGP models is represented in separated plots.
-The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
+\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig3_Post_part1}
@@
 \subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$).
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
 The right-side plot is a zoom in of the left-side plot.}
@@
 
-Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending of three variables, $m$, $c$ and $\ell$.
-An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending of four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
+Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending on three variables, $m$, $c$ and $\ell$.
+An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
 However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
+
+(Gabi: The following paragraph is another idea to introduce that the diagnosis plot in Figure \ref{fig6_lscale_vs_J_vs_c} depends on the kernel considered.
+At the moment we have built this figure only for the square exponential kernel.)
+
+Ultimately, Figure \ref{fig6_lscale_vs_J_vs_c} showing the relationships among the number of basis functions, the boundary factor and the lengthscale of the function depends on the particular choice of covariance function.
+In this work we have built these relationships only for the square exponential covariance function.
+For other kernels, the Figure will be slightly different.
 
@@
 For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3?
-Gabi: I've mentioned it in line 491) and varying lengthscale values.
+Gabi: I did some attempt about that in previous subsection, but I am not sure whether succesfully.) and varying lengthscale values.
 Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case.
 
-Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets.
-The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
-Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models against the regular GP models as a function of the lengthscale and number of basis functions $m$.
+Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of both regular GP and HSGP models fitted to those datasets.
+The lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
+As commented previously, an accurate estimate of the lengthscale can be a good indicator of a close approximation of the HSGP model to the regular GP model.
+Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP models as a function of the lengthscale and number of basis functions $m$.
 
@@
 \subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
-\vspace{-3mm}
+\vspace{-2mm}
 \subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
@@
 \vspace{-3mm}
-\caption{Posterior predictive means of both regular GP and HSGP models, fitted over different datasets simulated from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generative function (\textit{True function}).}
+\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
   \label{fig7_posterior_varing_lscale_part1}
@@
 \end{flushleft}
-\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario.
-The Y-axes represent the lengthscales estimates of GP and HSGP models.
-The use of different $m$ in the HSGP models is represented in separated plots.
-The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig6_lscale_vs_J_vs_c}.}
+\caption{Data-generating functional lengthscales (X-axis), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates from the regular GP and HSGP models (Y-axis).
+95\% confident intervals of the lengthscale estimates are plotted as dot lines.
+The different plots represent the use of different number of basis functions $m$ in the HSGP model.
+The dashed black line represents the recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c} that can be closely approximated by the HSGP model in every case.}
   \label{fig8_Tlscale_vs_Elscale}
@@
 \includegraphics[scale=0.35]{fig9_MSE_varing_lscale.pdf}
-\caption{RMSE of the HSGP models for the different datasets with different wiggly effects (lengthscales) and number of basis functions $m$.}
+\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects (lengthscales).}
   \label{fig9_MSE_varing_lscale}
@@
 
-We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales.
-We observe small RMSE for large lengthscales, however we appreciate mismatch between large lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
-This is because big changes on large lengthscales, which might be clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small RMSE.
-On the other side, we observe high RMSE for small lengthscales, however we appreciate match between small lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
-This is because very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high RMSE.
+Comparing the information of models approximation provided by Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale}, it can be appreciated an agreement between both figures, excepting for small lengthscales and long lengthscales.
+We observe small RMSE for large lengthscales in Figure \ref{fig9_MSE_varing_lscale}, however there exist mismatch between large lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
+This is because big changes on large lengthscales, which might be clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior predictions and therefore a small RMSE.
+On the other side, we observe high RMSE for small lengthscales in Figure \ref{fig9_MSE_varing_lscale}, however there exist match between small lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
+This is because very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors predictions and therefore high RMSE.
 
@@
 
-\noindent $L$ is the boundary and $m$ the number of basis functions.
-The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale parameters of the approximate covariance function, respectively.
+\noindent In the previous equations, $L$ is the boundary and $m$ the number of basis functions.
+The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale parameters, respectively, of the approximate covariance function.
 
@@
 
-Figure \ref{fig10_Posteriors_exI} shows the posteriors distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$ (\ref{eq:boundary})), and the splines model with 80 knots.
-The true data-generative function and the noisy observations are also plotted in the Figure.
-The right and left extremes and three small regions in the middle of the plot correspond to out-of-sample or test data which have not been taking part on training the models.
-The test data have been plotted as crosses and the training data as circles.
+Figure \ref{fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$ (\ref{eq:boundary})), and the splines model with 80 knots.
+The true data-generative function and the noisy observations are also plotted in the Figure; the sample observations are plotted as circles and the out-of-sample or test data, which have not been taking part on training the models, are plotted as crosses.
 The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
@@
 \includegraphics[width=\textwidth]{fig10_Posteriors_exI.pdf}
-\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
+\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dot lines.}
   \label{fig10_Posteriors_exI}
@@
 Figure \ref{fig11_MSE_exI_inter} shows the standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
-The SRMSE is computed against the true function.
-Model interpolation has been assessed either using the test data for interpolating and the training data.
+The SRMSE is computed against the data-generating function.
 From Figures \ref{fig10_Posteriors_exI} and \ref{fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model, for interpolating and extrapolating data.
 However, the splines model does not extrapolate data properly.
-For interpolation both models show roughly similar performance.
+Both models show roughly similar interpolating performance.
+
+Figure \ref{fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the splines model.
+The computational time is represented in the Y-axis of the figure, which is in a logarithmic scale.
+The HSGP model is on average roughly 400 times faster than the regular GP, in the particular case of applying over this dataset.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 22mm 60mm 120mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
-\caption{Standarized root mean square error of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation.}
+\caption{Standarized root mean square error (SRMSE) of the different methods against the data-generating function.
+(left) SRMSE for interpolation.
+(right) SRMSE for extrapolation.
+The Standard deviation of the mean of the SRMSE are plotted as dot lines.}
   \label{fig11_MSE_exI_inter}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_time_exI.pdf}}
+\caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
+  \label{fig11_time_exI}
 \end{figure}
@@
 
-In the HSGP model the function $f(\bm{x})$ is approximated as in equation (\ref{approxf}), with the square exponential spectral density as in equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
+In the HSGP model the function $f(\bm{x})$ is approximated as in Equation (\ref{approxf}), with the square exponential spectral density as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
 
@@
 
-Figure \ref{fig12_Posteriors_gaydata} shows the posteriors distributions of the three models, the regular GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the splines model with 20 knots.
-The true data-generative function and the noisy observations are also plotted in the Figure.
-Two small sets of observations have been used as test data, and plotted as crosses in the Figure.
+Figure \ref{fig12_Posteriors_gaydata} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the splines model with 20 knots.
+Sample observations are plotted as circles in the figure, and the out-of-sample observations, which have been used for testing, are plotted as crosses.
 
@@
 \includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
-\caption{Posterior distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
+\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
   \label{fig12_Posteriors_gaydata}
@@
 
-For the HSGP model, different models with different number of basis functions and boundary factors have been fitted.
-The root mean square errors (RMSE) for every one of these models have been computed and plotted as a function of the number of basis functions and the boundary factor in Figure \ref{fig13_MSE_train_BF_gaydata}, for training (left) and test (right) data.
-In this case, the RMSE has been computed against the regular GP model.
+For the HSGP model, different models with different number of basis functions and boundary factor have been fitted.
+The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model and plotted as a function of the number of basis functions $m$ and the boundary factor $c$ in Figure \ref{fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
 The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
@@
 \subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_pred_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig13_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the different methods against the true function. (left) RMSE for training data. (right) RMSE for test data.}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 50mm 120mm 0mm, clip]{fig13_legend_gaydata.pdf}}
+\caption{Root mean square error (RMSE) of the HSGP model, computed against the regular GP model, as a function of the number of basis functions $m$ and boundary factor $c$.
+(left) RMSE for sample data.
+(right) RMSE for test data.}
   \label{fig13_MSE_train_BF_gaydata}
@@
 
-Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the different models, regular GP, HSGP and splines, against the actual data, for training and test data.
+Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models, computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model.
 We can see how the splines models does not extrapolate data properly.
+Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration, as a function of the number of basis functions $m$ for the HSGP model and knots for the splines model.
+The computational times is represented in the Y-axis which is in a logarithmic scale.
+The HSGP model is on average roughly 10 times faster than the regular GP, in this particular case.
 
@@
 \subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 50mm 120mm 0mm, clip]{fig14_legend_gaydata.pdf}}
-\caption{Mean square error (MSE) of the different methods againts the actual data. (left) MSE for training data. (right) MSE for test data.}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 42mm 120mm 0mm, clip]{fig14_legend_gaydata.pdf}}
+\caption{Root mean square error (RMSE) of the different methods, regular GP, HSGP and splines models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots for the splines model.
+(left) RMSE for sample data.
+(right) RMSE for test data.}
   \label{fig14_MSE_train_gaydata}
@@
 
-%\begin{figure}[H]
-%\centering
-%\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
-%\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
-%\caption{Mean square error (MSE) of the different methods againts the actual data. (left) MSE for training data. (right) MSE for test data.}
-%  \label{fig14_MSE_train_gaydata}
-%\end{figure}
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_1_time_gaydata.pdf}}
+\caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
+  \label{fig14_1_time_gaydata}
+\end{figure}
 
@@
 
-The regular GP model with 2 input dimensions, over the outcome variable $\bm{y}$ and $n \times 2$ matrix of inputs $X$, can be written as follows,
+The regular GP model over the outcome variable $\bm{y}$ and $n \times 2$ matrix of inputs $X$, can be written as follows,
 %
@@
 
-In the HSGP model with 2 input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^2$, is approximated as (\ref{approxf_multi}),
+In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{approxf_multi}),
 %
@@
 
-\noindent where $\bm{\lambda}_j$ is the 2-vector (\ref{eigenvalue_multi}) with elements that are univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) which elements are the univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
 \begin{align*}
-\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D=2} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D=2},
+\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D},
 \end{align*}
 
-\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}), as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}) as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
 \begin{align*}
-\phi_j(\bm{x}) &= \prod_{d=1}^{D=2} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D=2} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
+\phi_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
 \end{align*}
@@
 \noindent where $x_d$ is the input value corresponding to dimension $d$.
-$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left(\sqrt{\lambda_{\mathbb{S}_{j1}}},\sqrt{\lambda_{\mathbb{S}_{j2}}} \right)$, of the 2-dimensional square exponential covariance function,
-%
-\begin{eqnarray*}
-S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \ell_1 \ell_2  \mathrm{exp}\Bigg(-\frac{1}{2} \bigg( \ell_1^2 \sqrt{\lambda_{\mathbb{S}_{j1}}}^2 + \ell_2^2 \sqrt{\lambda_{\mathbb{S}_{j2}}}^2 \bigg) \Bigg). 
-\end{eqnarray*}
-
-\noindent In the previous equations, $j$ denotes the index for the $m= m_1 \cdot m_2$ multivariate basis functions (\ref{m_multi}), where $m_1$ and $m_2$ are the number of basis functions considered for dimension 1 and 2, respectively.
-$\mathbb{S}$ is the matrix of $2$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the two ($D$=2) dimensions.
+$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left\{ \sqrt{\lambda_{\mathbb{S}_{jd}}} \right\}_{d=1}^{D}$, of the $D$-dimensional square exponential covariance function,
+%
+\begin{eqnarray*}
+S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \prod_{d=1}^D \ell_d  \mathrm{exp}\bigg(-\frac{1}{2} \sum_{d=1}^D \ell_d^2  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg). 
+\end{eqnarray*}
+
+\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$.
+$\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the $D$ dimensions.
 $L_d$ is the boundary for the dimension $d$.
-The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$, $\ell_1$ and $\ell_2$ are the marginal variance and lengthscale parameters for dimensions 1 and 2, respectively, of the approximate covariance function.
+The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell_d$ are the marginal variance and lengthscale of dimension $d$, respectively, of the approximate multivariate covariance function.
 
@@
 
-Figure \ref{fig15_Posteriors_exII} shows the data-generative GP function, from where the dataset was drawn and the mean posterior functions of the three models fitted over the dataset: the regular GP, the HSGP, and the splines models.
-The noisy observations are also plotted in the plots as circles.
+Figure \ref{fig15_Posteriors_exII} shows the data-generating GP function from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset.
+Sample observations are also plotted in the plots as circles.
 For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
-A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ are used.
+A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ were used.
 For the splines model, 40 knots in each dimension were used.
@@
 \end{center}
-\caption{(a) True data-generative function. (b) Posterior mean function of the GP model. (c) Posterior mean function of the HSGP model. (d) Posterior mean function of the splines (SP) model.}
+\caption{(a) Data-generating function.
+(b) Mean posterior predictive function of the GP model.
+(c) Mean posterior predictive function of the HSGP model.
+(d) Mean posterior predictive function of the splines (SP) model.}
   \label{fig15_Posteriors_exII}
@@
 \subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 5mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
-\caption{Mean error between the data-generative function and the GP (a), HSGP (b) and splines (c) models.}
+\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models.}
   \label{fig16_errors_exII}
@@
 In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
-In all models, the same number of basis functions and knots for each dimension were used.
-Figure \ref{fig17_RMSE_exII}(left) shows the root mean squared error (RMSE) as a function of the boundary factor, the number of univariate basis functions and knots.
-The RMSE is computed against the true function.
+In all models, the same number of basis functions and knots per dimension were used.
+Figure \ref{fig17_RMSE_exII}(left) shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, the number of univariate basis functions $m$ and knots.
 From Figures \ref{fig16_errors_exII} and \ref{fig17_RMSE_exII}(left), it can be seen a close approximation of the HSGP model to the regular GP model.
 However, the performance of the splines model is significantly worse.
-Figure \ref{fig17_RMSE_exII}(right) shows the times of computation of the different models as a function of the boundary factor, number of basis functions and knots.
-From both plots in Figure \ref{fig17_RMSE_exII} it can be noticed that choosing the optimal boundary factor allows for less number of basis functions and less computational time.
+Figure \ref{fig17_RMSE_exII}(right) shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots.
+Figure \ref{fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computational time.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
-\caption{Standarized mean square error (SMSE) of the different methods against the true function (left), and time of computation of the different models (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+\caption{Root mean square error (RMSE) (left) and computational time in seconds per iteration (right) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
   \label{fig17_RMSE_exII}
@@
 
-In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{approxf_multi}),
-%
-\begin{equation*}
-f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\bm{\lambda}_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
-\end{equation*}
-
-\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) which elements are the univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
-%
-\begin{align*}
-\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D},
-\end{align*}
-
-\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}) as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
-%
-\begin{align*}
-\phi_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
-\end{align*}
-
-\noindent where $x_d$ is the input value corresponding to dimension $d$.
-$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left\{ \sqrt{\lambda_{\mathbb{S}_{jd}}} \right\}_{d=1}^{D}$, of the $D$-dimensional square exponential covariance function with a scalar lengthscale $\ell$,
-%
-\begin{eqnarray*}
-S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \ell^D  \mathrm{exp}\bigg(-\frac{1}{2} \ell^2 \sum_{d=1}^D  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg). 
-\end{eqnarray*}
-
-\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$.
-$\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the $D$ dimensions.
-$L_d$ is the boundary for the dimension $d$.
-The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale, respectively, of the approximate multivariate covariance function.
+In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as in equation (\ref{approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq:specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$, and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
 
@@
 
-Figure \ref{fig18_gpfun_diabetes} shows the mean posterior probability ($p_i$) function of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset with the 2 input dimensions ($D=2$) \textit{Glucose} and \textit{Pregnancy}.
+Figure \ref{fig18_gpfun_diabetes} shows the mean posterior predictions of probabilities ($p_i$) of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset with the 2 input dimensions ($D=2$) \textit{Glucose} and \textit{Pregnancy}.
 The binary observations $y_i$ are also plotted in the plots as colored points.
@@
 A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used.
-For the splines model, 20 knots in each dimension were used.
+For the splines model, 20 knots per dimension were used.
 %
@@
 \subfigure{\includegraphics[scale=0.31, trim = 149mm 18mm 5mm 20mm, clip]{fig18_legend_diabetes.pdf}}
-\caption{(a) Posterior mean function of the GP model. (b) Posterior mean function of the HSGP model. (c) Posterior mean function of the splines (SP) model.}
+\caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the splines (SP) model.}
   \label{fig18_gpfun_diabetes}
@@
 In order to assess performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factor for the HSGP model and different number of knots for the splines model have been fitted.
-In all models, the same boundary factor, number of basis functions and knots for each dimension were used.
-Figure \ref{fig19_ELPD_diabetes} shows the expected log predictive density (ELPD) as a function of the boundary factor, the number of univariate basis functions and knots.
+In all models, the same boundary factor, number of basis functions and knots per dimension were used.
+Figure \ref{fig19_ELPD_diabetes} shows the expected log predictive density (ELPD) as a function of the boundary factor $c$, the number of univariate basis functions $m$ and knots.
 The ELPD is computed over the actual observations.
-Basically, with slightly differences, all models show similar performances, due to the process is very smooth with a estimated lengthscale $\ell=4.51$ very large.
-A slight pattern of improvement of performance can be noticed as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales.
+Basically, with slightly differences, all models show similar performances, due to the fact that the process is very smooth with a relatively very large lengthscale estimate $\ell=4.51$.
+Even tough, a slightly pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (\ref{fig6_lscale_vs_J_vs_c}).
 %
@@
 
-Figure \ref{fig20_time2D_diabetes} shows the times of computation in the logarithm scale of the different models fitted over the dataset with 2 input dimensions (left), 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor, number of univariate basis functions and knots.
-We can notice the significant increase of computational time with higher dimensions for the HSGP and splines models.
-This fact shows that choosing optimal values for the number of basis functions and bounfary factor is essential to avoid excessive time of computation.
+Figure \ref{fig20_time2D_diabetes} shows the computational times in the logarithm scale of the different models, regular GP, HSGP and splines, fitted over the dataset with 2 input dimensions (left), 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots.
+We can appreciate the significant increase of computational time with higher dimensions for the HSGP and splines models.
+This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid excessive computational time, specially in high input dimensionality.
 It is interesting to be noticed that considering more than 10 knots per dimension in the splines model with 3D is not allowed for an amount of 392 observations.
+Similarly, just the computation of the input data for the splines model with 4D input dimension is computationally very expensive.
 

### 2019-04-29 Paper/manuscript/manuscript_05.tex
some improvements to Section 3

@@
 Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictions of the estimated function and on the covariance function itself.
-For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation and the non-linearity of the estimated function, in the sense that fewer basis functions implies inaccurate and smoother effects covariance functions (covariance functions with larger lengthscales) and consequently more linear functional forms, and conversely.
-%:Fewer basis functions imply smoother covariance functions (Paul: when would you consider a covariance function as smooth?
-Also, I think the main point here is that at long distances the implied covariance function becomes negative which can't be in an exact GP model.) and consequently more linear functional forms.
-Then, the higher the "wigglyness" of the function to be estimated, the more basis functions will be required. 
-%(Paul: do we see this in Figure \ref{fig1_Post_J} already?). 
-On the other hand, for $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.
+For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently more linear functional forms.
+The higher the "wigglyness" of the function to be estimated, the more basis functions will be required.
+If $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.
 
@@
 \item As $c$ increases, $m$ has to increase as well (and vice versa).
-\item There exists a minimum $c$ under which a close approximation will never be achieved regardless of $m$.
+\item There exists a minimum $c$ below which a close approximation will never be achieved regardless of $m$.
 \end{itemize}
@@
 If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
-If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased (Paul: decreasing $c$ seems dangerous to me if predictions for out of sample data are desired.
-Shell we mention this?
-Gabi: We assume that predicting points are already within $S$.
-Then, a valid $c$ is more related to the lengthscale than to the predicting points.
-In fact, valid $c$s are restricted in function of the lengthscale: A given lengthscale has a minimum $c$ under which a close approximation will never be achieved regarless $m$ (We stated this in line 502 and also in line 462).
-For example, a given $c$ can not work for certain lengthscale but for a larger one).
-We may also use this diagnostic in a iterative procedure.
+If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased We may also use this diagnostic in a iterative procedure.
 Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
@@
 
-For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3?
-Gabi: I did some attempt about that in previous subsection, but I am not sure whether succesfully.) and varying lengthscale values.
+For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying lengthscale values.
 Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case.
@@
 The lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
-As commented previously, an accurate estimate of the lengthscale can be a good indicator of a close approximation of the HSGP model to the regular GP model.
-Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP models as a function of the lengthscale and number of basis functions $m$.
+As noted previously, an accurate estimate of the lengthscale can be a good indicator of a close approximation of the HSGP model to the regular GP model.
+Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP models, as a function of the lengthscale and number of basis functions.
 
@@
 
-Comparing the information of models approximation provided by Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale}, it can be appreciated an agreement between both figures, excepting for small lengthscales and long lengthscales.
+Comparing the accuracy of the lengthscale in Figure 
+\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure \ref{fig9_MSE_varing_lscale},
+we see that they agree closely with each other for medium lengthscales.
+That is, a good estimation of the lengthscale implies a
+small RMSE. This is no longer true for very small or large lengthscales.
+In small lengthscales, even very small inaccuracies 
+may have a strong influence on the posteriors predictions and thus on the RMSE.
+In large lengthscales, larger inaccuracies change the posterior predictions
+only little and may thus not yield large RMSEs.
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the 
+minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}. 
+We observe that whenever the estimated lengthscale exceeds the minimially
+estimable lengthscale, the RMSE of the posterior predictions is small
+(see Figure \ref{fig9_MSE_varing_lscale}).
+Conversely, when the estimated lengthscale is smaller than the minimally
+estimable one, the RMSE becomes very large.
+
+(Paul: Figure 8 and 9 are not rendered correctly in my PDF viewer after 
+compiling the tex file, which has something to do with the font of the facet 
+and legend labels I believe. Latex also issues warnings about this.)
+
+% Comparing the information of models approximation provided by Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale}, it can be appreciated an agreement between both figures, excepting for small lengthscales and long lengthscales.
 We observe small RMSE for large lengthscales in Figure \ref{fig9_MSE_varing_lscale}, however there exist mismatch between large lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
@@
 
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig6_lscale_vs_J_vs_c}.
+%The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig6_lscale_vs_J_vs_c}.
 It can be appreciated a complete agreement between those recommended minimum lengthscales that can be properly fitted in every case following Figure \ref{fig6_lscale_vs_J_vs_c} and the real approximate accuracy measured by the RMSE in Figure \ref{fig9_MSE_varing_lscale}.

### 2019-09-07 Paper/manuscript/manuscript_06.tex
improved figures and added all study cases except birthday dataset

--- 2019-04-29 Paper/manuscript/manuscript_05.tex
@@
 \usepackage{subfigure}
-\usepackage{multirow}
+\usepackage{multirow,slashbox}
+\usepackage{array}
 \usepackage{float}
@@
 
-Although there are several GP specific software packages available to date (GPML, GPstuff, GPy, GPflow, cite them!), each provide efficient implementations only for a restricted range of GP based models.
+Although there are several GP specific software packages available to date (GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP based models.
 In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
@@
 $\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$,
- where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
+ where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$, which needs be positive semidefinite.
 The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
@@
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
+A covariance function needs be symmetric and positive semidefinite.
 A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
@@
 \begin{align}
-\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenvalue_multi} \\
-%
-\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenfunction_multi}
+\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenfunction_multi} \\
+%
+\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenvalue_multi}
 \end{align}
@@
 
-(Gabi: The following paragraph is a first idea to introduce that the recommendations and diagnostics depends on the kernel considered.
-And that, at the moment, we have built these recommendations only for the square exponential kernel.
-I know that what I wrote in not enough and should be a bit more elaborate; your suggestions where and how elaborate it would be useful.)
-
 Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
-In this work we built these relationships for the square exponential covariance function.
-For other kernels, the relationships will be slightly different.
+In this work we built these relationships for the square exponential covariance function and Matern ($\nu=3/2$) covariance function.
+For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig1_legend.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 5mm 0mm 5mm 0mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 122mm 10mm, clip]{fig1_legend.pdf}}
 \caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig2_legend.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 5mm 0mm 5mm 0mm, clip]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 120mm 0mm, clip]{fig2_legend.pdf}}
 \caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
@@
 \begin{center}
-\begin{tabular}{ c c c}
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
+\begin{tabular}{| c | c | c | c |}
+\hline
+c  = 1.05 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part1.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part1.pdf} & \multirow{36}{1.3cm}{\includegraphics[scale=0.35, trim = 28mm 40mm 100mm 30mm, clip]{fig3_legend.pdf}}\\
+\cline{1-3}
+c  = 1.1 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part2.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part2.pdf} &\\
+\cline{1-3}
+c  = 1.2 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part3.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part3.pdf} &\\
+\cline{1-3}
+c  = 1.5 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part4.pdf} & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part4.pdf} & \\
+\cline{1-3}
+c  = 2 & \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part5.pdf} &  \includegraphics[scale=0.23, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Cov_part5.pdf} &\\
+\cline{1-3}
+c  = 2.5 & \includegraphics[scale=0.23, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Post_part6.pdf} & \includegraphics[scale=0.23, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Cov_part6.pdf} &\\
+\hline
 \end{tabular}
@@
 
+%\begin{figure}
+%\begin{tabular}{ c c c c c}
+%\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} &  \hspace{-0mm}
+%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\ 
+%\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} & \hspace{-0mm}
+%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.35, trim = 25mm 63mm 120mm 50mm, clip]{fig3_legend.pdf}\\
+%\includegraphics[scale=0.20, trim = 10mm 5mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} & \hspace{-0mm}
+%\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} 
+%\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} & 
+%\end{tabular}
+%\caption{Mean posterior predictive functions (left-first and right-first columns) and covariance functions (left-second and right-second columns) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
+%  \label{fig3_Post_part1}
+%\end{figure}
+
 \begin{figure}
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 0mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig5_legend.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 28mm 119mm 0mm, clip]{fig5_legend.pdf}}
 \caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
@@
 Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figure \ref{fig6_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to
 % the performance of the HSGP approximation.
-a close approximation of the HSGP model, in the case of a GP with square exponential covariance function and a single input dimension.
-(Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?).
+a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension.
 More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
+Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern ($\nu=3/2$) covariance function.
 We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
@@
 
-As stated above, Figure \ref{fig6_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$.
+As stated above, Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
-If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased We may also use this diagnostic in a iterative procedure.
+If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
+We may also use this diagnostic in a iterative procedure.
 Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
 \hspace{3mm}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
@@
 
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_Matern.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\frac{3}{2}$) covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+  \label{fig6_lscale_vs_J_vs_c_Matern}
+\end{figure}
+
 If we look back to the conclusions drawn from Figures \ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig6_lscale_vs_J_vs_c}.
@@
 
-Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending on three variables, $m$, $c$ and $\ell$.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, and result in a surface depending on three variables, $m$, $c$ and $\ell$.
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
 However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
-
-(Gabi: The following paragraph is another idea to introduce that the diagnosis plot in Figure \ref{fig6_lscale_vs_J_vs_c} depends on the kernel considered.
-At the moment we have built this figure only for the square exponential kernel.)
-
-Ultimately, Figure \ref{fig6_lscale_vs_J_vs_c} showing the relationships among the number of basis functions, the boundary factor and the lengthscale of the function depends on the particular choice of covariance function.
-In this work we have built these relationships only for the square exponential covariance function.
-For other kernels, the Figure will be slightly different.
+\begin{center}
+\end{center}
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
-\vspace{-2mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
+\subfigure{\includegraphics[scale=0.4, trim = 10mm 100mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
+\vspace{-3mm}
+\subfigure{\includegraphics[scale=0.31, trim = 10mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
 \vspace{-3mm}
@@
 \begin{flushleft}
-\begin{tabular}{ c c c c}
-\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 45mm, clip]{fig8_legend.pdf}}\\
-\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
+\begin{tabular}{ccc}
+\multicolumn{3}{c}{ \includegraphics[scale=0.40, trim = 25mm 115mm 90mm 45mm, clip]{fig8_legend.pdf}}\\
+\includegraphics[scale=0.30, trim = 0mm 24mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.30, trim = 25mm 24mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 24mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \\
+\includegraphics[scale=0.30, trim = 0mm 0mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 0mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 0mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf}
 \end{tabular}
@@
 \centering
-\includegraphics[scale=0.35]{fig9_MSE_varing_lscale.pdf}
+\includegraphics[scale=0.38]{fig9_MSE_varing_lscale.pdf}
 \caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects (lengthscales).}
@@
 
-This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws of a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
+This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
 To form the final noisy dataset $\bm{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
@@
 \noindent where $I$ represents the identity matrix.
-The previous formulation corresponds to the latent form of a GP model, where $f(\bm{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$.
+The previous formulation corresponds to the latent form of a GP model, where $f(\bm{x})$ represents the underlying functional model to the noisy data and is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$.
 Saying that the function values $f(\bm{x})$ follows a GP model is equivalent to say that $f(\bm{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.
@@
 The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
+The posteriors of the three models, regular GP, HSGP, and Splines, are pretty similar within the interpolation input space.
+However, when extrapolating the splines solution clearly differs from the regular GP and HSGP models as well as the actual observations.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_extra.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 60mm 120mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
+\subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_extra.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 55mm 110mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
 \caption{Standarized root mean square error (SRMSE) of the different methods against the data-generating function.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_time_exI.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig11_time_exI.pdf}}
 \caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_train_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_pred_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 50mm 120mm 0mm, clip]{fig13_legend_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig13_MSE_train_BF_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_pred_BF_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 119mm 0mm, clip]{fig13_legend_gaydata.pdf}}
 \caption{Root mean square error (RMSE) of the HSGP model, computed against the regular GP model, as a function of the number of basis functions $m$ and boundary factor $c$.
@@
 We can see how the splines models does not extrapolate data properly.
+
 Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration, as a function of the number of basis functions $m$ for the HSGP model and knots for the splines model.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 42mm 120mm 0mm, clip]{fig14_legend_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 25mm 30mm 123mm 0mm, clip]{fig14_legend_gaydata.pdf}}
 \caption{Root mean square error (RMSE) of the different methods, regular GP, HSGP and splines models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots for the splines model.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_1_time_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig14_1_time_gaydata.pdf}}
 \caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
@@
 \begin{tabular}{ c c c}
-\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 10mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\
-\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
+\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 0mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\
+\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
 \end{tabular}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 5mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 39.2mm 20mm, clip]{fig16_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 19mm 23mm 39.2mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 19mm 23mm 39.2mm 20mm, clip]{fig16_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 0mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
 \caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig17_RMSE_exII.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 30mm 123mm 0mm, clip]{fig17_legend_exII.pdf}}
 \caption{Root mean square error (RMSE) (left) and computational time in seconds per iteration (right) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 149mm 18mm 5mm 20mm, clip]{fig18_legend_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 12mm 23mm 35mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 35mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 38mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 150mm 13mm 2mm 15mm, clip]{fig18_legend_diabetes.pdf}}
 \caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the splines (SP) model.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig19_ELPD_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig19_ELPD_diabetes.pdf}}
 \caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$ and knots.}
@@
 
+%\begin{figure}
+%\centering
+%\subfigure{\includegraphics[scale=0.29, trim = 10mm 0mm 13mm 0mm, clip]{fig20_time2D_diabetes.pdf}}
+%\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time3D_diabetes.pdf}}
+%\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time4D_diabetes.pdf}}
+%\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig20_legend_diabetes.pdf}}
+%\caption{Time of computation in the log scale of the different models fitted over the dataset with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+%  \label{fig20_time2D_diabetes}
+%\end{figure}
+
 \begin{figure}
 \centering
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 0mm 13mm 0mm, clip]{fig20_time2D_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time3D_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time4D_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig20_legend_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{fig20_time2D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 30.2mm 0mm 10mm 0mm, clip]{fig20_time3D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 30.2mm 0mm 10mm 0mm, clip]{fig20_time4D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 25mm 53mm 123mm 10mm, clip]{fig20_legend_diabetes_2.pdf}}
 \caption{Time of computation in the log scale of the different models fitted over the dataset with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
@@
 \subsection{Study case VI: Leukemia data}\label{sec:bf_caseVI}
+The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
+The data set consist in survival times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for 1043 cases ($i=1,\cdots,n=1043$).
+Some 16\% of cases were censored.
+Predictors are \textit{age} ($\bm{x}_1$), \textit{sex} ($\bm{x}_2$), \textit{white blood cell} (WBC) ($\bm{x}_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($\bm{x}_4$) which is a measure of deprivation for district of residence.
+We denote the matrix $X=\{\bm{x}_1,\bm{x}_2,\bm{x}_3,\bm{x}_4\} \in {\rm I\!R}^{n\times 4}$ which contains the predictors.
+
+As the WBC measurements were strictly positive and highly skewed, we fit the model to its logarithm.
+Continuous predictors were normalized to have zero mean and unit standard deviation.
+Survival time was normalized to have zero mean for the logarithm of time.
+We assume a Gaussian observation model for the logarithm of the observed survival time, $t'_i=log(t_i)$, with a function of the predictors, $f(X)$, as the location parameter, and $\sigma$ as the Gaussian noise:
+%
+\begin{equation*}
+p(t'_i)= \mathcal{N}(t'_i|f(X_i),\sigma^2)
+\end{equation*}
+
+As we do not have a model for the censoring process, we do not have a full observation model, and the observational model for the censored data $t_i$ is assumed to be the complementary cumulative normal probability distribution:
+%
+\begin{equation*}
+p(y_i > t'_i)= \int_{t'_i}^{\infty} \mathcal{N}(y_i|f(X_i),\sigma^2) dy_i=  1 - \Phi \left( \frac{y_i-f(X_i)}{\sigma} \right),
+\end{equation*}
+
+\noindent where $y_i$ denotes the logarithm of the uncensored time.
+
+The latent function $f(X)$ is modeled as a Gaussian process, centered in a linear model of the predictors $X$, and with a squared exponential covariance function $k$ depending on predictors and hyperparameters $\theta=\{\alpha,\ell\}$,
+%
+\begin{eqnarray*} 
+f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta)).
+\end{eqnarray*}
+
+\noindent where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+Notice that a scalar lengthscale is considered in the multivariate covariance function.
+
+%The marginalized forms, by integrating out the latent values $f(X)$, of the previous previous observational models, observed and censored, result
+%%
+%\begin{align*}
+%p(t'_i)&= \int_{-\infty}^{\infty} \mathcal{N}(t'_i|c + \bm{\beta}X, K_{ii} + \sigma^2 )dt'_i,\\
+%p(y_i > t'_i)&= \int_{t'_i}^{\infty} \mathcal{N}(y_i|c + \bm{\beta}X, K_{ii} + \sigma^2) dy_i=  1 - \Phi \left( \frac{y_i-(c + \bm{\beta}X)}{K_{ii} + \sigma} \right),
+%\end{align*}
+%
+%\noindent with covariance matrix $K$, and its element $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$ is the evaluation of the covariance function $k$ at points $r$ and $s$.
+
+Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable (1 for female and 2 for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
+The relative contribution of a GP function given one of the levels of the predictor (\ref{eq:gpml2}) to a general mean GP function (\ref{eq:gpml1}) is defined.
+For the other level of the predictor, the GP function effects are set to zero.
+This multilevel construction is depicted as following:
+%
+\begin{eqnarray} 
+&f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta_1)) \label{eq:gpml1} \\
+&g(X|\bm{x}_2=2) \sim \mathcal{GP}(0, k(X,X, \theta_2|\bm{x}_2=2)) \label{eq:gpml2} \\ 
+&f(X|\bm{x}_2=2) = f(X|\bm{x}_2=2) + g(X|\bm{x}_2=2) \nonumber
+\end{eqnarray}
+
+\noindent In the previous equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of the GP function restricted to the male sex ($\bm{x}_2=2$).
+
+Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x}|x_2=2)$ are approximated as in equation (\ref{approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq:specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
+
+%Figure \ref{fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
+The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
+
+Figure \ref{fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
+These posterior estimates correspond to the HSPG model with $m=10$ basis functions and $c=3$ boundary factor.
+The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.70, trim = 0mm 0mm 0mm 0mm, clip]{fig21_posteriors_leukemia_01.pdf}
+\caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
+The thick line in each graph is the posterior mean estimated using a HSGP model, and the thin lines represent pointwise 95\% intervals.}
+  \label{fig21_posteriors_leukemia}
+\end{figure}
+
+Figure \ref{fig22_rmse_leukemia} shows the root mean square error (RMSE) and the time of computation in seconds per iteration as a function of the number of univariate basis functions $m$ and boundary factor $c$.
+As the functions are smooth, few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig22_rmse_leukemia}-right); Small boundary factors are not allowed when large lengthscales as can be seen in \ref{fig6_lscale_vs_J_vs_c}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig22_rmse_leukemia}-rleft).
+%
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig22_rmse_leukemia.pdf}} 
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig23_time_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 120mm 0mm, clip]{fig24_legend_leukemia.pdf}}
+\caption{Root mean square error (RMSE) and time of computation (in seconds per iteration) as a function of the number of basis functions $m$ and boundary factor $c$.}
+  \label{fig22_rmse_leukemia}
+\end{figure}
+
+%\subsection{Study case VII: Land use spatio-temporal classification task (first form)}\label{sec:bf_caseVII}
+%The next example presents an spatio-temporal classification task in land-use of parcels between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits.
+A set of 200 parcels with known class are used for training the model and a set of 800 parcels in which to predict their unknown classes.
+The data is recorded in a time series of 5 years (2006,2008,2010,2012,2015) within the period.
+The class of each parcel $i$ and time $t$ is stored by a categorical variable $y_{it}$ representing the 5 different possible classes ($k=1,\cdots,K=5$): $k=1$, adult independent citrus fruits; $k=2$, aligned citrus fruits; $k=3$, irregular citrus fruits; $k=4$, abandoned citrus fruits; $k=5$, young citrus fruits.
+%
+%A bunch of 52 characteristic variables were measured as predictors for every parcel and time.
+These variables were computed from satellite color images and cadastral map, and they concern spectral intensities and empirical semivariogram of the pixels within a parcel, as well as descriptive statistics of the shape of the parcels.
+%
+%Due to the fact that 52 input variables are too many for a multivariate HSGP model as it scales $O(n\cdot m^D)$ with $m$ the number of basis functions in the HSGP model and $D$ the number of input variables, the multivariate HSGP model will be formulated as an additive HSGP model.
+%
+%In order to avoid high-order components in the additive HSGP model as much as possible, since it is known multivariate HSGP models to be computationally demanding, a Principal Component Analysis is carried out over the characteristic variables.
+The resulting principal components are linearly uncorrelated each other and will be used jointly with the time variable as predictors in the classifying first-order additive HSGP model.
+Let's denote the matrix $X=\{\bm{x}_{it}\}$ which contains the predictors, where $\bm{x}_{it} \in {\rm I\!R}^{D}$, $D=53$ (52 principal components + time) is the $D$-vector of characteristic variables of the parcel $i$ and time $t$.
+The use of a first-order additive model is justified due to the input variables correspond to principal components which are linearly uncorrelated each other.
+%
+%The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{i}=\{p_{ik}\}_{k=1}^{K=5}$ of belonging to class $k$ per parcel $i$,
+%%
+%\begin{equation*}
+%y_i \sim \text{multinomial}(\bm{p}_i).
+%\end{equation*}
+%
+%\noindent The goal is to estimate the vector of probabilities $\bm{p}_i$ as a function of the input values, which is a multivariate function $f(\bm{x}_i):\bm{x_i}\in {\rm I\!R}^{D} \rightarrow \{f_k(\bm{x}_i)\}_{k=1}^K \in {\rm I\!R}^{K}$, related to the vector of probabilities $\bm{p}_i$ through the 'softmax' link function.
+%%
+%\begin{eqnarray*}
+%\bm{p}_i = \text{softmax}(f(\bm{x}_i)).
+%\end{eqnarray*}
+%
+%Each individual function $f_k(\bm{x})$ is modeled as a first-order additive model, 
+%%
+%\begin{eqnarray*}
+%f_k(\bm{x}) \sim \sum_{d=1}^{D} g_d(x^d)
+%%\mathcal{HSGP}(x^d, S(\bm{\lambda}), \theta_d),
+%\end{eqnarray*}
+%
+%\noindent where each individual component $g_d(x^d)$ is modeled as a unidimensional HSGP model,
+%%
+%\begin{eqnarray*}
+%g_d(x^d) \sim \mathcal{HSGP}(x^d, S, \theta_d),
+%\end{eqnarray*}
+%
+%In the HSGP model, the function $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
+The vector of hyperparameters $\theta_d=\{\alpha_d,\ell_d\}$ contains the marginal variance $\alpha_d$ and lengthscale $\ell_d$ of the HSGP process.
+%
+%For each individual component $g_d(x^d)$ modeled as a HSGP model, $m$=15 basis functions and a boundary factor $c$=2.5 were used.
+All the input variables were previously standarized with the exception of the time variable.
+%
+%After the adjustment of the model, the lengthscales estimates $\{\hat{\ell}_d\}_{d=1}^{D=14}$ normalized by half of the input range ($\frac{2\cdot \hat{\ell}_d}{|x^d_{max}-x^d_{min}|}$) are all bigger than the minimun lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as function of $m$, $c$.
+Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling the data.
+%
+%Table \ref{tab1} shows the confusion classification matrix after fit the model following a $Q$-fold cross-validation procedure over the training data. Every fold was composed of $Q$=10 observations.
+%
+%\begin{table}
+%\begin{center}
+%\begin{tabular}{|c|*{6}{c|}}\hline
+%\backslashbox{Estimate}{True} & \multicolumn{1}{p{1.5cm}|}{\centering caind \\(k=1)} &  \multicolumn{1}{p{1.5cm}|}{\centering calign \\(k=2)} & \multicolumn{1}{p{1.5cm}|}{\centering irr \\(k=3)} & \multicolumn{1}{p{1.5cm}|}{\centering aband \\(k=4)} & \multicolumn{1}{p{1.5cm}|}{\centering young \\(k=5)} & \multicolumn{1}{p{1.5cm}|}{\centering } \\
+%\hline \multicolumn{1}{|p{3cm}|}{ \centering caind (k=1)} &90&39&14&3&11&42\%\\
+%\hline \multicolumn{1}{|p{3cm}|}{\centering calign (k=2)} &46&301&8&2&3&16\%\\
+%\hline \multicolumn{1}{|p{3cm}|}{\centering irr (k=3)}  &8&4&59&4&1&19\%\\
+%\hline \multicolumn{1}{|p{3cm}|}{\centering aband (k=4)}  &5&2&6&342&5&5\%\\
+%\hline \multicolumn{1}{|p{3cm}|}{\centering young (k=5)}  &8&2&1&0&38&19\%\\
+%\hline \multicolumn{1}{|p{3cm}|}{\centering }  &42\%&13\%&32\%&2\%&34\%&$\textbf{17\%}$\\
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Confusion matrix after the Q-fold cross-validation procedure over the training data.}
+%  \label{tab1}
+%\end{table}
+
 \subsection{Study case VII: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
+The next example presents an spatio-temporal classification task in land-use of parcels between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits.
+A set of 200 parcels with known class are used for training the model and a set of 800 parcels in which to predict their unknown classes.
+The data is recorded in a time series of 5 years (2006,2008,2010,2012,2015) within the period.
+The class of each parcel $i$ and time $t$ is stored by a categorical variable $y_{it}$ representing the 5 different possible classes ($k=1,\cdots,K=5$): $k=1$, adult independent citrus fruits; $k=2$, aligned citrus fruits; $k=3$, irregular citrus fruits; $k=4$, abandoned citrus fruits; $k=5$, young citrus fruits.
+
+A bunch of 52 characteristic variables were available for every parcel and time.
+These variables were computed from satellite color images and cadastral map, and they concern spectral intensities and empirical semivariogram of the pixels within a parcel, as well as descriptive statistics of the shape of the parcels.
+
+%In order to reduce the number of input variables and extract the most informative information contained among all of them, we computed a Principal Component Analysis on these variables and keep the 13 first principal components which will be used jointly with the time variable as predictors in the classifying GP model.
+Let's denote the matrix $X=\{x_{it,j}\}$ which contains the predictors.
+
+%The trained GP model is applied over another set of 800 parcels in the time series ($n'=800\cdot5=4000$) with unknown class and input values $X'=\{\bm{x}'_i\}_{i'=1}^{n'=800} \in {\rm I\!R}^{n\times 14}$,  with $\bm{x}'_i=\{x'_{ij}\}_{j=1}^{14} \in {\rm I\!R}^{14}$ with the aim to classify these parcels in space and time.
+
+Due to the fact that 52 input variables are too many for a multivariate HSGP model as it scales $O(n\cdot m^D)$ with $m$ the number of basis functions in the HSGP model and $D$ the number of input variables, the multivariate HSGP model will be formulated as an additive HSGP model.
+%The use of a first-order additive model is justified due to the input variables correspond to principal components which are linearly uncorrelated each other.
+
+As it is known, computational demands of a multivariate HSGP model component increases quickly with the number of dimensions.
+Therefore, with the aim of avoiding as much as possible interaction components in the additive HSGP model, a Principal Component Analysis is carried out over the characteristic variables.
+The resulting principal components are linearly uncorrelated each other, so interaction effects between principal components are irrelevant or null.
+
+The principal components will be used jointly with the time variable as predictors in the classifying additive HSGP model.
+Let's denote the matrix $X=\{\bm{x}_{it}\}$ which contains the predictors, where $\bm{x}_{it} \in {\rm I\!R}^{D}$, $D=53$ (52 principal components + time) is the $D$-vector of characteristic variables of the parcel $i$ and time $t$.
+%The classifying additive HSGP model will be composed of first-order additive components with the 53 predictors and second-order additive components corresponding to the interaction components among the 52 principal components and the time variable.
+
+The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{it}=\{p_{it,k}\}_{k=1}^{K=5}$ of belonging to class $k$ per parcel $i$ and time $t$,
+%
+\begin{equation*}
+y_{it} \sim \text{multinomial}(\bm{p}_{it}).
+\end{equation*}
+
+\noindent The goal is to estimate the vector of probabilities $\bm{p}_{it}$ as a function of the predictors, which is a multivariate function $f(\bm{x}_{it}):\bm{x}_{it}\in {\rm I\!R}^{D} \rightarrow \{f_k(\bm{x}_{it})\}_{k=1}^K \in {\rm I\!R}^{K}$, which is related to the vector of probabilities $\bm{p}_{it}$ through the 'softmax' link function.
+%
+\begin{eqnarray*}
+\bm{p}_{it} = \text{softmax}(f(\bm{x}_{it})).
+\end{eqnarray*}
+
+Each individual function $f_k(\bm{x})$ is modeled as a partial second-order additive model in the form: 
+%
+\begin{eqnarray} \label{eq1_sec54}
+f_k(\bm{x}) = \sum_{d=1}^{D} g_d(x^d) + \sum_{d=1}^{D-1} g_{d+D}(x^d,x^D),
+%\mathcal{HSGP}(x^d, S(\bm{\lambda}), \theta_d),
+\end{eqnarray}
+
+\noindent where $x^D$ represents the time variable. The first-order components $g_d(x^d)$, for $d=1,\dots,D$, in Equation (\ref{eq1_sec54}), are modeled as unidimensional HSGP models
+%
+\begin{eqnarray*}
+g_d(x^d) \sim \mathcal{HSGP}(x^d, S, \theta_d).
+\end{eqnarray*}
+
+\noindent In the HSGP model, the first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
+
+The second-order components $g_{d+D}(x^d,x^D)$, for $d=1,\dots,D-1$, in Equation (\ref{eq1_sec54}), are modeled as two-dimensional HSGP models
+%
+\begin{eqnarray*}
+g_{d+D}(x^d,x^D) \sim \mathcal{HSGP}(x^d,x^D, S, \theta_{d+D}).
+\end{eqnarray*}
+
+\noindent In the HSGP model, the second-order components $g_{d+D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf_multi}), with the two-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
+
+The vectors of hyperparameters $\{\theta_d=[\alpha_d,\ell_d]\}_{d=1}^{2D-1}$ contains the marginal variance $\alpha_d$ and lengthscale $\ell_d$ of the d$th$ HSGP model component.
+
+For the first-order HSGP components $\{g_d(x^d)\}_{d=1}^D$, $m$=15 basis functions and a boundary factor $c$=2.5 were used.
+For the second-order HSGP components $\{g_{d+D}(x^d,x^D)\}_{d=1}^{D-1}$, $m_1$=15 and $m_2$=15 basis functions for each dimension, respectively, were used, which lead to a total of 225 multivariate basis functions.
+A boundary factor for each dimension $c_1$=2.5 and $c_2$=2.5 were used.
+All the input variables were previously standarized.
+
+In the case of the first-order components, the lengthscales estimates $\{\hat{\ell}_d\}_{d=1}^D$ normalized by half of the input range $\left(\frac{2\cdot \hat{\ell}_d}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimun lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as function of $m$, $c$.
+Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.
+
+For the second-order components we have not built a similar figure to that Figure \ref{fig6_lscale_vs_J_vs_c} but for the two-dimensional case. 
+%The figure for the two-dimensional case would depend on the ralation between the values of both lengthscales. In the particular case of this stufy case, the ratio would be one.
+However, we can approximately analyze the lengthscale estimates of the second-order HSGP components analyzing each dimension separately as unidimensional HSGP models.
+
+Table \ref{tab1} shows the confusion classification matrix after fitting the model following a $Q$-fold cross-validation procedure over the training data.
+Every fold was composed of $Q$=10 observations.
+As can be seen, there exist a high misclassification rate between classes $k=1$ and $k=2$, between classes $k=1$ and $k=3$, and between classes $k=1$ and $k=5$.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|*{6}{c|}}\hline
+\backslashbox{Estimate}{True} & \multicolumn{1}{p{1.5cm}|}{\centering k = 1} &  \multicolumn{1}{p{1.5cm}|}{\centering k = 2} & \multicolumn{1}{p{1.5cm}|}{\centering k = 3} & \multicolumn{1}{p{1.5cm}|}{\centering k = 4} & \multicolumn{1}{p{1.5cm}|}{\centering k = 5} & \multicolumn{1}{p{1.5cm}|}{\centering } \\
+\hline \multicolumn{1}{|p{3cm}|}{ \centering k = 1} &90&39&14&3&11&42\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering k = 2} &46&301&8&2&3&16\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering k = 3}  &8&4&59&4&1&19\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering k = 4}  &5&2&6&342&5&5\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering k = 5}  &8&2&1&0&38&19\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering }  &42\%&13\%&32\%&2\%&34\%&$\textbf{17\%}$\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Confusion matrix after the Q-fold cross-validation procedure over the training data.}
+  \label{tab1}
+\end{table}
+
+%Figure \ref{fig21_classifying_map} shows the image map of predicted classes for the whole set of parcels of the territory.
+%
+%\begin{figure}
+%\centering
+%\includegraphics[scale=0.35, trim = 0mm 0mm 0mm 0mm, clip]{recorte_newparcels3.png}
+%\caption{Image classification map with the predicted classes for  the whole set, training and predicting, classes.}
+%  \label{fig21_classifying_map}
+%\end{figure}
+
+%\subsubsection{Markov Chain structure underlying the time evolution of classes}
+%The underlying Markov chain distribution if fitted to the sequence of classes along time.
+This provides the probabilities transition matrix between classes (Figure \ref{fig25_transmatrix_landuses}), which is useful to check the smooth and monotonic correlation assumption of using a square exponential covariance function for the \textit{time} predictor in a Gaussian process.
+There is significant probability
+%%
+%\begin{figure}
+%\centering
+%\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig26_histlag_landuses.pdf}} \hspace{2mm}
+%\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig25_transmatrix_landuses.pdf}} \hspace{2mm}
+%\caption{Probabilities transition matrix between classes}
+%  \label{fig25_transmatrix_landuses}
+%\end{figure}
 

### 2019-09-07 Paper/manuscript/manuscript_03.tex

% interacttfssample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.

%\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables

%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty  %

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}

\graphicspath{{./images/}}

\bibpunct[, ]{[}{]}{,}{}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\begin{document}

%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{Hilbert space methods to approximate Gaussian processes using Stan}

%\author{
%\name{Gabriel Riutort-Mayol, Michael Riis Anderseen, Aki Vehtari, Paul-Christian Burkner}
%\affil{\textsuperscript{a}Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain; }
%}

\maketitle

\begin{abstract}

\end{abstract}

\begin{keywords}
Gaussian processes; Low-rank Gaussian processes; Hilbert Space methods; Sparse Gaussian processes
\end{keywords}

\tableofcontents

%Contents
\newpage

\section{Introduction}\label{sec:bf_intro}

\subsection{Motivation}

Efficient computation in Gaussian process (GP) models is of broad interest across machine learning and statistics, with applications in the fields of spatial epidemiology \citep{diggle2013statistical,carlin2014hierarchical}, robotics and control \citep{deisenroth2015gaussian}, signal processing \citep{sarkka2013spatiotemporal}, Bayesian optimization and probabilistic numerics, \citep{roberts2010bayesian,briol2015probabilistic,hennig2015probabilistic} , and many others.

One of the main limitations of Gaussian processes (GPs) \citep{rasmussen2006gaussian} is that computational demands and memory requirements scale as $O(n^3)$ and $O(n^2)$ respectively, effectively limiting a direct implementation. This limits the applicability of GPs to a number of training samples $n$ at most a few thousand cases. The computational requirements arise because in solving the GP regression problem we need to invert the $n \times n$ Gram matrix $\mathbf{K}$ of the covariance function $k(\mathbf{x},\mathbf{x}')$, which is an $O(n^3)$ operation in general. This problem becomes especially important when using sampling methods for full Bayesian inference, when in each sampling step we need to invert the Gram matrix, usually through Cholesky factorization. Then, approximate methods that will alleviate these computational demands are needed. Sparse GPs, which are approximate methods to fit GP models, do reduce the dimension of the posterior, but it still requires Cholesky factorizations. 
%Sparse GPs alleviate the computational demands for small and medium amounts of data but, due to 
They are basically based on reducing the amount of data used to fit the model, and convergence to the full GP model is not straight forward. Furthermore, reducing the amount of data while maintaining the posterior densities is not a simple procedure.

We want to make full Bayesian inference over Gaussian process models accessible to a broader audience by making them fast and easy to apply. If we can find a method to decompose the GP kernel into basis functions, we can represent the GP as a linear model, which makes it faster to fit. The set of Laplace eigenfunctions considered in this paper forms such a basis and we can compute them analytically. Furthermore, they are independent of the choice of kernel (including the hyperparameters), which means that we do not need to autodiff through the kernel and the Cholesky. We are going to achieve this by making it possible to use GPs in a probabilistic programming language like Stan.

Well-known splines models are a sort of basis functions expansion models but they do not generally have correspondence with GP models, and they do not have the flexibility of choosing different covariance functions as model structure.

\subsection{Contributions of the method}

This work is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models. This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods. This results in a reduced-rank approximation for the covariance function. This method has some nice features:

\vspace{2mm}
$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.

\vspace{2mm}
$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method. In addition, the computation of the automatic differentation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$. After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer. (Arno's paper, pag. 7)

\vspace{2mm}
$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions. especially when sampling methods for inference are used. GP prior as latent functions is needed in generalized models.

In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:

\vspace{2mm}
$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation. This operation has to be computed at every step of the HMC or optimizer.

\vspace{2mm}
$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional. It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitud parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

\subsection{Contributions of our work}

As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology. Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function. The performance and accuracy of the method are directly related with the number of basis functions and the box size. At the same time, successful values for these two factors depend on the smoothness or roughness of the process to be modeled. The time of computation is mainly dependent on the number of basis functions. Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.

\vspace{2mm}
$\bullet$ Firstly, clear summarized formulae of the method for the univariate and multivariate cases is presented.

\vspace{2mm}
$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the functions.

\vspace{2mm}
$\bullet$ We make recomendations for the values of these factors based on the recognized relations among them. We provide useful graphs of these relations that will help the users to improve performance and save time of computation.

\vspace{2mm}
$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.

\vspace{2mm}
$\bullet$ We describe the generalization of the method to the multidimensional case.

\vspace{2mm}
$\bullet$ We implement the approach in a fully probabilistic framework and for the Stan programming probabilistic software.

\vspace{2mm}
$\bullet$ We show several illustrative examples, simulate and real datasets, of the performance of the model, and accompanied by their Stan codes.

%\vspace{3mm}
%A summary of the formulae of the method for the univariate and multivariate cases is presented. The number of basis functions used in the approach and the value for the boundary condition of the model are two factors of main importance in practical applications. The number of basis functions is directly related to the accuracy of the approximation but also to the computation demands. Also exist a relation among the number of basis functions, the specified values for the boundary condition, and the performance of the approach.  Finally, the approach is applied in several illustrative study cases and accompanied by their Stan codes.

\subsection{Related work}

The GP prior entails an $O(n^3)$ complexity that is computationally intractable for many practical problems, and this problem especially becomes severe when we want to conduct inference using sampling methods. To overcome this scaling problem several schemes have been proposed. One approach is to partition the data set into separate groups \citep{snelson2007local, urtasun2008sparse} and performing local inference in each partition. Other global approach is to build a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables' \citep{quinonero2005unifying,bui2017unifying}. Other global approach make use of basis functions to approximate the covariance function. In \cite{snelson2007local} the authors conduct an approach that combines the idea of local and global approaches.

The literature contains many parametric models that approximate Gaussian process behaviours; for example \cite{bui2014tree} included  tree-structures in the approximation for extra scalability, and \cite{moore2015gaussian} combined local Gaussian
processes with Gaussian random fields.

\subsubsection{Inducing points methods}

The approach based on inducing points employs a small set of pseudo data points to summarise the actual data. The storage requirements are reduced to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$. Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods. From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modiﬁcations to the so-called Nystr\"om method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum. These methods are basically based on choosing a set of $m$ inducing inputs $x_u$ and scaling the corresponding eigendecomposition of their corresponding covariance matrix $K_{u,u}$ to match that of the actual covariance.

This scheme was originally introduced to the GP context by \cite{williams2001using}. As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model. However, several methods modifying the inducing point approach are widely used. The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix. The SoR method is based on a degenarete prior which produces unresoanble predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).

The Deterministic Training Conditional (DTC) method \citep{ro2001sparse,seeger2003fast}) retains the true covariance for the training data, but uses the approximate cross-covariances between training and test data, which reverse the problem of nonsensical predictive uncertainties. However, since the covariances for training and test cases are computed differently, this method results not to actually be a Gaussian process. This method was presented as Projected Latent Variables (PLV) in \cite{seeger2003fast} and Projected Process Approximation (PPA) in \cite{rasmussen2006gaussian}.

The Variational Approximation (VAR) \citep{titsias2009variational} suggests a variational approach which provides an objective function for optimizing the selection of inducing points. This basically modifies the DTC method by an additional trace term in the likelihood that comes from the variational bound.  \cite{hensman2013gaussian} extended this idea by introducing additional variational parameters to enable stochastic variational inference \citep{hoffman2013stochastic}, achieving a more computationally scalable bound which allows GPs to be fitted to millions of data.

The Fully Independent (Training) Conditional (FIC) \citep{quinonero2005unifying} method originally introduced as Sparse Pseudo-Input GP by \cite{snelson2006sparse} is also based on the Nystr\"om approximation, where they allow the pseudo-point input locations to be optimised by maximising the new model's marginal likelihood whose covariance is parameterized by the locations of an active set not constrained to be a subset of the training and test data.

More recently Bui et. al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time. The new framework is built on standard methods for approximate inference (variational-free-inference, EP and Power EP methods).

In practice, the inducing points-based sparse approximation methods works reasonable well in cases where the field is relatively smooth. \cite{vanhatalo2010approximate}
propose the use of compactly supported covariance function in conjunction with sparse approximations to model both short and long range correlations.

\cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI). This framework improves the scalability and accuracy of fast kernel approximations through kernel interpolation, and naturally combines the advantages of inducing point and structure exploiting for scalability (such as Kronecker \citep{saatcci2012scalable} or Toeplitz \citep{cunningham2008fast}) approaches.

The number of inducing points or their locations are crucial in order to capture the correlation structure. For a discussion on the effects of the inducing points, see \cite{vanhatalo2010approximate}. This behavior applies to all the methods from the Nystr\''om family.

This kind of 'projected process' approximation has also been discussed by e.g. \cite{banerjee2008gaussian}.

\subsubsection{Basis function methods}

The spectral analysis and series expansions of Gaussian processes has a long history. A classical result (see, e.g, \cite{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}, and references therein) is that the covariance function can be approximated with a finite truncation of Mercer series and the approximation is guaranteed to converge to the exact covariance function when the number of terms is increased.

Another related classical connection is to the works in the relationship of spline interpolation and Gaussian process priors \citep{wahba1978improper,kimeldorf1970correspondence,wahba1990spline}. In particular, it is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a speciﬁc choice of covariance function. The relationship of the spline regularization with Laplace operators then leads to series expansion representations that are closely related to the approximations considered here.

Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels. The approximate kernel has a finite basis function expansion.

The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse}, where the spectral representation of the covariance function is used. This model is a stationary sparse GP that can approximate any desired stationary full GP. However, as argued by the authors, this option does not converge to the full GP and can suffer from overfitting to the training data. \citep{gal2015improving} sought to improve the model by integrating out, rather than optimizing the frequencies.   Gal and Turner derived a variational approximation that made use of a tractable integral over the frequency space. The result is an  algorithm that suffers less overfitting than the Sparse Spectrum GP, yet remains flexible.

While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.

Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat´ern class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure. They combine the variational methodology with Fourier based approximations.

In spatial statistics similar approaches are called low-rank models \citep{diggle2007springer}. The low rank models assume that the Gaussian field is a linear combination of $m$ basis functions. The type of an approximation depends on the basis functions used. Familiar examples include spectral representation \citep{diggle2007springer,paciorek2007computational,paciorek2007bayesian} and splines \citep{wood2003thin}.

Recent Splines models can reproduce the Matern family of covariance functions, however our approach can reproduce basically all of the stationary covariance functions.

\vspace{3mm}
\section{Gaussian process prior model}

Gaussian processes (GP) are natural and flexible non-parametric models to define probability distributions over multi-dimensional functions \citep{rasmussen2006gaussian;neal1997monte}. This enables to apply probabilistic inference directly to functions. As a continuous function is a uncontable infite collection of values, a GP model can be seen as a probability model over an infinite collection of random varibales. However, formally, the defining feature of a GP model is that the joint distribution of function values at any finite number of input values is a multivariate normal distribution.

\vspace{0.5cm}
\noindent ... \\
... \\
...

\vspace{0.5cm}
A stationary covariance function is a function of $x-x'$, so we can write $k(x,x') = k(x-x')$. Isotropic covariance functions are those that are function of the distance between observations, $k(x,x') = k(|x-x'|) = k(r)$. The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which are
%
\begin{eqnarray}
k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r/\ell^2), \nonumber \\
k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
\end{eqnarray}

\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the lengt-scale and magnitud, respectively, of the kernel. The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function. These covariance functions can be easily generalized to the multidimensional case considering multidimensional inputs $\mathbf{r}$ and the multidimensional length-scale $\boldsymbol{\ell}$.

The spectral density functions associated with the univariate Mat\'ern covariance functions written above are
%
\begin{eqnarray}
S_{\nu=\infty}(w)&=& \sigma^2 \sqrt{2\pi} \cdot \ell \cdot \text{exp}(-0.5 \ell^2 w^2), \nonumber \\
S_{\nu=\frac{1}{2}}(w)&=& 2\sigma^2 \frac{1}{\ell}(\frac{1}{\ell^2} + \omega^2)^{-1}, \nonumber \\
S_{\nu=\frac{3}{2}}(w)&=& 4\sigma^2 \frac{\sqrt{3}}{\ell}^{3}(\frac{3}{\ell^2} + \omega^2)^{-2}, \nonumber \\
S_{\nu=\frac{5}{2}}(w)&=& \frac{16}{3}\sigma^2 \frac{\sqrt{3}}{\ell}^{5}(\frac{5}{\ell^2} + \omega^2)^{-3}, \nonumber 
\end{eqnarray}

\noindent where variable $w$ is in the frequency domain.

\vspace{3mm}
\section{Hilbert space-based approximate Gaussian process prior model}\label{sec:bf_method}

\subsection{Method}

The approximate Gaussian process method, developed by \cite{solin2018hilbert} and implemented in this paper, lay on the basis of considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators. Then, the pseudo-differential operator is approximated with Hilbert space methods on compact subsets $\Omega \subset {\rm I\!R}^D$, and with some boundary condition.

First, we focus on the unidimensional case of the input space, such as $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value.

\vspace{0.2cm}
The approximate method leads to the approximation of the covariance function between two input values $\{x,x'\} \in \{-L,L\}$ as 
%
\begin{equation}\label{approxcov}
k(x,x') \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
\end{equation}

\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ and $\theta$ the set of hyperparameters of $k$; a stationary covariance function can be equivalently represented in terms of the spectral density \citep{rasmussen2006gaussian}, then the spectral density is a function of the hyperparameters $\theta$. $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator. Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
%
\begin{eqnarray}\label{eigenproblem}
-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
\end{eqnarray}

\noindent The eigenvalues $\lambda_j>0$ are real and positive, because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in eq. (\ref{eigenproblem}) are sinusoidal functions,
%
\begin{eqnarray}
\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
\end{eqnarray}

If we truncate the sum in (\ref{approxcov}) up to $J$ ($j=1,\cdots,J$), the approximate covariance function can be represented as
%
\begin{equation}
k(x,x') \approx \boldsymbol{\phi}(x)^\intercal \Delta \boldsymbol{\phi}(x'), \nonumber
\end{equation}

\noindent where $\boldsymbol{\phi}(x)=\{\phi_j(x)\}_{j=1}^{J} \in {\rm I\!R}^{J}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{J\times J}$ the diagonal matrix of espectral densities $S_{\theta}(\sqrt{\lambda_j})$, 
%
\begin{eqnarray}
\Delta &=&  \begin{bmatrix}
    S_{\theta}(\sqrt{\lambda_1}) & & \\
    & \ddots & \nonumber \\
    & & S_{\theta}(\sqrt{\lambda_J}) \\
  \end{bmatrix}.
\end{eqnarray}

Thus, the Gram matrix $\text{K}$ of the covariance function $k$ for a collection of observations $i=1,\cdots,N$ and collection of input values $\{x^i\}_{i=1}^{N} \in {\rm I\!R}^{N}$ can be represented as follows
%
\begin{equation}
\text{K}= \Phi \Delta \Phi^\intercal, \nonumber
\end{equation}

\noindent where $\Phi \in {\rm I\!R}^{N\times J}$ is the matrix of eigenfunctions $\phi_j(x^i)$,
%
\begin{eqnarray}
\Phi &=&  \left[ {\begin{array}{ccc}
   \phi_1(x^1) & \cdots & \phi_J(x^1)  \\
    \vdots &\ddots & \vdots  \nonumber \\ 
    \phi_1(x^N) & \cdots & \phi_J(x^N) \\
  \end{array} } \right].
\end{eqnarray}

\noindent Now, the Gaussian process prior for the function $f$ can be re-defined as
%
\begin{equation}
\mathbf{f} \sim \text{GP}(0,\Phi \Delta \Phi^\intercal). \nonumber
\end{equation}

\noindent This equivalently leads to a linear representation of the function $f$,
%
\begin{equation}\label{approxf}
f(x) \approx \sum_{j} \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j, \nonumber
\end{equation}

\noindent where $\beta_j \sim \text{N}(0,1)$ is a random variable with zero mean and unitary variance. That is, the function $f$ can be approximated with a finite basis function expansion (eigenfunctions $\phi_j$ of the Laplace operator), scaled by the squared root of the spectral density as a function of the eigenvalues $\lambda_j$. The eigenfunctions $\phi_j$ do not depend on the kernel hyperparameters $\theta$, the only dependence on $\theta$ is through the spectral density $S_{\theta}$. The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and the spectral density goes rapidly to zero for bounded covariance functions. Therefore, it is expected a good approximation in eq. (\ref{approxf}) for a finite number, $J$, of terms in the series as long as the inputs values $x^i$ are not near the boundaries of the domain $[-L,L]$, where the Laplacian was taken to be zero.

The computational cost of this approximate model scales $O(n\cdot J+J)$, where $n$ is the number of observations and $J$ the number of basis functions.

\subsection{Generalization to multidimensional input space}

We are going to generalize the results from the previous section to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions. We assume the same number $J$ of basis functions terms in the series expansion for every input dimension, however it is straight forward to generalize this to different number of term series $J_d$ for every dimension.

\vspace{0.2cm}
In a $D$-dimensional input space the total number of eigenfunctions and eigenvalues in the approximation corresponds to the total number of $D$-tuples over $J$, which is $J^D$. Let $\mathbb{S}\in {\rm I\!R}^{J^D \times D}$ be the set of $D$-tuples elements. Each new eigenfunction $\phi^{\ast}_j$ correspond to the product of the univariate eigenfunctions whose indices corresponds to the the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$. And each new eigenvalue $\lambda^{\ast}_j$ is a $D$-vector whose elements are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$. Let $j=\{1,\cdots,J^D\}$,
%
\begin{eqnarray}
\phi^{\ast}_j(\mathbf{x}) &=& \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(\mathbf{x}_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(\mathbf{x}_d+L_d)\right)  \nonumber \\
%
\lambda^{\ast}_j &=& \{\lambda_{\mathbb{S}_{jd}} \}_{d=1}^D =  \{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \}_{d=1}^D  \nonumber
\end{eqnarray}

As an example we show the matrix $\mathbb{S}$ for a $three$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2,\mathbf{x}_3\}$ ($D=3$) and $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively. The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$. The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
%
\begin{eqnarray}
\mathbb{S}=
\left[ {\begin{array}{ccc}
1 & 1 & 1 \nonumber \\
1 & 1 & 2 \\
1 & 1 & 3 \\
1 & 2 & 1 \\
1 & 2 & 2 \\
1 & 2 & 3 \\
2 & 1 & 1 \\
2 & 1 & 2 \\
2 & 1 & 3 \\
2 & 2 & 1 \\
2 & 2 & 2 \\
2 & 2 & 3 
\end{array} } \right]
\end{eqnarray}

The approximate covariance function is represented as
%
\begin{equation}\label{approxcov_multi}
k(\mathbf{x},\mathbf{x}') \approx \sum_{j=1}^{J^D} S^{\ast}_{\theta}\left(\sqrt{\lambda^{\ast}_j}\right) \phi^{\ast}_j(\mathbf{x}) \phi^{\ast}_j(\mathbf{x}'),
\end{equation}

\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function. The $D$-dimensional spectral density functions associated with the Mat\'ern covariance functions considered in the previous section are
%
\begin{eqnarray}
S^{\ast}_{\nu=\infty}(\mathbf{w})&=& \sigma^2 \sqrt{2\pi}^D \prod_{d=1}^D l_d \cdot \text{exp}\left(-0.5 \sum_{d=1}^D l_d^2 \mathbf{w}_d^2\right), \nonumber \\
%
S^{\ast}_{\nu}(\mathbf{w})&=& \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\mathbf{w}^2 \right)^{(\nu+D/2)}, \nonumber
\end{eqnarray}

\noindent where the $D$-vector $\mathbf{w}$ is in the frequency domain. We can represent the approximate series expansion of the function $f$ in the multidimensional input space as,
%
\begin{equation}\label{approxf}
f(\mathbf{x}) \approx \sum_{j} \left( S^{\ast}_{\theta}(\sqrt{\boldsymbol{\lambda^{\ast}_j}})\right)^{1/2} \phi^{\ast}_j(\mathbf{x}) \beta_j, 
\end{equation}

\noindent where $\beta_j \sim \text{N}(0,1)$ is a random variable with zero mean and unitary variance. The computational cost in a multidimensional setting scales $O((n+1)\cdot \prod_{d=1}^D J_d )$, where $n$ is the number of observations and $J_d$ is the number of basis functions considered in the dimension $d$.

\subsection{Details in computational complexity}

\vspace{2mm}
$\bullet$ The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.

\vspace{2mm}
$\bullet$ The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands. The weights depend on the hyperparameters, then they have to be computed in every step of the HMC sampling method.

\vspace{2mm}
$\bullet$ The linear model is computed with complexity $O(nm)$, computed in every step of the HMC sampling method.

\vspace{2mm}
$\bullet$ Then, the operations in this whole linear model will scale $O(nm)$ and computed in every step of the HMC sampling method.

\vspace{2mm}
$\bullet$ In addition, the computation of the automatic differentation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ In a regular GP model the automatic differentation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$. This operation has to be computed at every step of the HMC.

\vspace{2mm}
$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.

\subsection{Dependency on the number of basis functions and the boundary condition}

The approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, e.g. in a 1-$D$ input space $x\in \Omega=[-L,L]$:
%
\begin{equation}\label{approxcov}
k(\tau) \approx \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0),  \nonumber
\end{equation}

\noindent where $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$. The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are 
%
\begin{eqnarray}
\lambda_j&=&\left(\frac{j\pi}{2L}\right)^2, \nonumber \\
\phi_j(x)&=&\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \nonumber
\end{eqnarray}

The approximation to the covariance function will become exact in the limit when the number of basis functions approach infinity $J=1,\dots,j,\dots,\infty$. The number of basis functions can be truncated that the difference between the covariance function and the approximation be less than a threshold $\epsilon$,
%
\begin{eqnarray}\label{diff_covs}
\int k(\tau)d(\tau) - \int \sum_{j}S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) d(\tau) < \epsilon.
\end{eqnarray}

The finite number of basis functions in the approximation to satisfy eq. (\ref{diff_covs}) will depend on the smoothness or roughness of the function, that is, on the lengthscale $\ell$. The approximation will also depends on the box size $L$ which will affect the performance specially near the boundaries.

To illustrate the effects of these two factors, number of basis functions $J$ and box size $L$, on the performance of the approximation, let's consider a simulated dataset of 75 noisy observations from a GP prior with lengthscale $\ell=0.3$ and magnitud $\alpha=1$. The additive noise added to the GP prior simulations is $\sigma=0.2$. A regular GP model and different approximate GP models with different choices for the number of basis function $J$ and box size $L$ are fitted on this dataset.

For a fixed $L$, let's suppose it is set to a right choice, Figure \ref{image2_J} shows the effect of the number of basis functions in approaching the posterior mean and the covariance function. The number of basis functions affects the non-linearity or roughness (or smoothness) of the posteriors. Fewer basis functions implies smoother covariance functions and consequently smoother posteriors. The more roughness the process is, the more basis functions are needed.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.4, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
\caption{(left) Posterior distributions for different number of basis functions $J$. (right) Covariance function for different number of basis functions $J$.}
  \label{fig1_Post_J}
\end{figure}

Similarly, for a fixed number of basis functions $J$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of the box size $L$ in approaching the posterior mean and the covariance function. The box size mainly affects the approximation near the boundaries of the function, and affects the covariances at long distances.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig2_Cov_L.pdf}}
\caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
  \label{fig2_Post_L}
\end{figure}

Therefore, the approximation to the covariance function depends on the number of basis functions $J$ used in the series expansion and the value for boundary condition or box size $L$. Additionally, there exit a relation among the number of basis functions, the box size and the lengthscale or roughness of the process on the performance of the approximation. Figure \ref{fig3_lscale_vs_J_vs_c_part1} collects how these three factors relate each other in order to achieve a good approximation to the covariance function of the regular GP using eq. (\ref{diff_covs}). On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data. The countour lines gather the box size which is also normalized by the half-range of the data, to which we refer as the boundary factor $c$, where $L=c \cdot \text{range}(x)/2$. And the Y-axis gathers the number of basis functions. Thus, this figure give us the minimum number of basis functions needed to achieve a good approximation in terms of satisfying eq. (\ref{diff_covs}), for a certain GP process with lengthscale $\ell$ and given a fixed boundary factor $c$. Alternatively, this figure could also be read as the maximum boundary factor $c$ that we should use given a number of basis functions, for certain GP process with certain lengthscale. And also this figure could also be read as the maximum roughness in a process that can be fitted properly given a number of basis functions and a box size. Finally, this figure also gives us the minimum value for the boundary factor $c$ which is suitable for certain roughness or number of basis functions. This figure basically collects the behaviour of the approximate GP model in function of the lengthscale, number of basis functions and boundary condition:

$\bullet$ As the lengthscale or smoothness of the process increase, the boundary factor and the number of basis functions needed for the approach decrease.

$\bullet$ There is a minimum boundary factor to achieve a good approach in function of the lengthscale (The contour lines have an end in function of the lengthscale).

$\bullet$ The lower boundary factor, the fewer basis functions, and the lower lengthscales suitable to be fitted.

$\bullet$ This plot serves as a diagnosis tool in the sense that if the estimated lengthscale is lower than that minimun generated by the figure for the chosen number of basis functions and boundary factor means that we would need to increase the number of basis functions or decrease the boundary factor.

$\bullet$ On the other hand, if the lengthscale is bigger than that generated by the figure then it means that the fit should be good.

Figure \ref{fig3_lscale_vs_J_vs_c_part1} is useful for the user to know the suitable values for the number of basis functions $J$ and the boundary factor $c$ to approach a certain process characterized by its lengthscale. This Figure will allow us to optimize computations which depends on the number of basis functions. Starting from some guess of the lengthscale of the process and knowing the range of the input data which we are interested in, the user can do a few iterative adjustments to obtain an optimal fit. This figure also provides a diagnosis tool of the fit, in the sense that if the estimated lengthscale is lower than the minimum lengthscale able to fit with the actual number of basis functions and box size, indicates that the number of basis functions needs to be increased or the box size decreased.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
\caption{Relation between the minumum number of basis functions $J$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{\frac{\text{range}(x)}{2}}$), and the boundary condition factor $c$ ($c = \frac{L}{\frac{\text{range}(x)}{2}}$).}
  \label{fig3_lscale_vs_J_vs_c_part1}
\end{figure}

The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior. Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either the regular GP model fit and the approximate GP model fit. Figure \ref{fig4_Post_&_Cov_part1} shows the posteriors distributions and the covariance functions of the approximate GP model and the regular GP model, obtained after fitting the data. Figure \ref{fig5_MSE_vs_J} shows the standarized root mean square error both models in function of the number of basis functions and the boundary factor. Figure \ref{fig6_lscale_vs_J} shows the evolution of the GP hyperparameters, lengthscale and magnitud, in function of the number of basis functions and boundary factor. Following Figure \ref{fig5_MSE_vs_J} it can be drawn that the optimal choice in terms of precision and computations would be 15 basis functions and a boundary factor between 1.5 and 2.5. The choice of 10 basis functions and a boundary factor of 1.5 could be accurate enough. This same conclusion can be roughly seen in the right-side plots in Figure \ref{image11_Posteriors_example-I}. This conclusions also agree with Figure \ref{fig3_lscale_vs_J_vs_c_part1} from which we can deduce that for a process with a normalized lengthscale of 0.3, 10 basis functions and a normalized box size (boundary factor) of 1.5 would be enough to get a good approximation following eq. (\ref{diff_covs}).

From Figure \ref{fig5_MSE_vs_J} it can also be drawn the behaviour in terms of performance of the approximate GP model in function of the number of basis functions and the boundary factor. Two main behaviours can be deduced from Figure \ref{fig5_MSE_vs_J}, which can also be easily seen from Figure \ref{fig3_lscale_vs_J_vs_c_part1}:
\\ 
$\bullet$ as the boundary factor increases, more basis functions are needed,
\\
$\bullet$ as fewer basis functions are used, the boundary factor must decrease.

Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and magnitud parameter as function of the number of basis functions and the boundary factor. The same conclusion drawn before can be drawn form this figure where with 15 basis functions and a boundary factor of 1.5 the estimated lengthscale of the approximate GP model approach the estimeted lengthscale of the regular GP model. From this figure it can be drawn that exists a minumum value for the boundary factor under this never a good approach will be achieved. This conclusion also appears in Figure \ref{fig3_lscale_vs_J_vs_c_part1} the contour lines of the boundary factor have an end at certain lengthscales, which means that exist a minimum boundary factor depending on the lengthscale of the process.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part1.pdf}}
\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
\caption{Posterior distributions and covariance functions of the proposed approximate GP models and the exact GP model in function of the number of basis functions $J$ and the boundary factor $c$.}
  \label{fig4_Post_&_Cov_part1}
\end{figure}

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
\caption{Standarized root mean square error (SRMSE) of the proposed approximate GP models and the exact GP model. (left) SRMSE against the number of basis functions $J$ for different values of the boundary factor $c$. (right) SRMSE against the boundary factor $c$ for different values of the number of basis functions $J$. }
  \label{fig5_MSE_vs_J}
\end{figure}

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig6_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
\caption{(left) Estimated lengthscales against the number of basis functions $J$ for different values of the boundary factor $c$. (right) Estimated GP magnitud the number of basis functions $J$ for different values of the boundary factor $c$. }
  \label{fig6_lscale_vs_J}
\end{figure}

\subsection{Comparative between true lengthscales and estimated lengthscales}

We fit the exact and the approximate GP model on different datasets drawn from a squared exponential GP model with different lengthscales. We fit the approximate GP model using different number of basis functions.

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{fig7_varing_lscale.pdf}
\caption{Posterior means of the exact and approximate GP models fitted over different datasets with different lengthscales.}
  \label{fig7_varing_lscale}
\end{figure}

Figure \ref{fig8_Tlscale_vs_Elscale} compares the true lengthscales to the estimated lengthscales for the exact GP and the approximate GP. The black line defines the minimun lengthscale that the approximate GP model can fit for each case.

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{fig8_Tlscale_vs_Elscale.pdf}
\caption{True lengthscales versus estimated lscales of the different realizations of datasets using different number of basis functions. }
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

\vspace{3mm}
\section{Univariate illustrative examples}\label{sec:gp_examples1D}
\subsection{Simulated data}

A simulated one-dimensional dataset, with 250 data points drawn from a Gaussian process prior with a Matern covariance function with 3/2 degrees of freedom. The hyperparameters of this GP prior with a Matern kernel, marginal variance and lengthscale are set to 1 and 0.15, respectively. The additive Gaussian noise to the GP model is set to 0.2. We compare the modelling performance over this simulated process of our proposed approach with the performance of a regular GP and a Splines-based model. For our basis function approach we use $J=80$ basis functions and a boundary factor $c=1.2$. For the Splines-based model we use 80 knots or basis in the model.

Figure \ref{fig9_Posteriors_exI} shows the posteriors distributions of the proposed model and the regular GP model, fitted over the simulated dataset. Posteriors distributions of the proposed approximation GP model, of the exact GP model, and the Splines model are plotted jointly with the true GP prior from the data were drawn. The right and left extremes of the simulated data correspond to out-of-sample data or test data which have not been taking part of training the models. Over this test data we can evaluate the predictive performance of every method.

\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{fig9_Posteriors_exI.pdf}
\caption{Posterior distributions of the proposed approximated GP model, the exact GP model, and the Splines model.}
  \label{fig9_Posteriors_exI}
\end{figure}

For assessing the performance of different methods we use the standardized mean squared error (SMSE) over the true function for interpolation and extrapolation data.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.350]{fig10_MSE_exI_inter.pdf}}
\subfigure{\includegraphics[scale=0.350]{fig10_MSE_exI_extra.pdf}}
\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
  \label{fig10_MSE_exI_inter}
\end{figure}

\subsection{Case I: Gay data}\label{sec:bf_caseII}
\subsection{Case II: Birthday data}\label{sec:bf_caseIII}

\vspace{3mm}
\section{Multivariate illustrative examples}\label{sec:gp_examplesMulti}
\subsection{Simulated data}\label{sec:bf_toyexampleMulti}
\subsection{Case III: Diabetes data}\label{sec:bf_caseIV}
\subsection{Case IV: Leukemia data}\label{sec:bf_caseV}

\vspace{3mm}
\section{Novel application cases}\label{sec:bf_novelcases}
\subsection{Case V: Land use spatio-temporal classification task}\label{sec:bf_caseVII}

\vspace{3mm}
\appendix

\section{Spectral densities of stationary covariance functions}

The covariance function of a stationary process, that is function of $\boldsymbol{\tau}=\mathbf{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).

\vspace{0.2cm}
\textit{(Bochner’s theorem) A complex-valued function $k$ on ${\rm I\!R}^D$ is the covariance function of a weakly stationary mean square continuous complex valued random process on ${\rm I\!R}^D$ if and only if it can be represented as}
%
\begin{equation}
k(\boldsymbol{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \mathbf{s} \cdot \mathbf{\tau}} d\mu(\boldsymbol{\tau}), \nonumber 
\end{equation}

\textit{where $\mu$ is a positive finite measure.}

\vspace{0.2cm}
If the measure $\mu$ has a density, it is known as the spectral density $S(\omega)$ of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the Wiener-Khintchine theorem. It gives the following relations:

\begin{eqnarray}
k(\boldsymbol{\tau})&=& \int S(\mathbf{s}) e^{2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber \\
%
S(\mathbf{s})&=& \int k(\boldsymbol{\tau}) e^{-2\pi i \mathbf{s} \cdot \boldsymbol{\tau}} d\mathbf{s}  \nonumber
\end{eqnarray}

\section{Approximate the covariance function using Hilbert space methods}

Associated to each covariance function $k(\mathbf{x},\mathbf{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
%
\begin{equation}
\mathcal{K} f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}'.
\end{equation}

Assuming that the espectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
%
\begin{equation}
S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
\end{equation}

If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the covariance function $k$,

\begin{equation}
-\nabla^2 f(\mathbf{x}) = \int k(\mathbf{x},\mathbf{x}') f(\mathbf{x}') d\mathbf{x}',
\end{equation}

\noindent then the covariance function can be represented as

\begin{equation}
k(\mathbf{x},\mathbf{x}')= \sum_j \lambda_j \phi_j(\mathbf{x}) \phi_j(\mathbf{x}'),
\end{equation}

\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator. Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
%
\begin{eqnarray}
-\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
\end{eqnarray}

a series expansion of eigenvalues and eigenfunctions

\section{Example of generalization to the multivariate case}

Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension. The number of new multidimensional eigenfunctions $\phi^{\ast}_j$ and eigenvalues $\lambda^{\ast}_j$ is $J^D=3^2=9$ ($j=\{1,\cdots,J^D\}$). The matrix $\mathbb{S}\in {\rm I\!R}^{9 \times 2}$ is
%
\begin{eqnarray}
\mathbb{S}=
\left[ {\begin{array}{cc}
1 & 1 \nonumber \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3
\end{array} } \right]
\end{eqnarray}

\noindent and the multidimensional eigenfunctions and eigenvalues
%
\begin{eqnarray}
\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} \nonumber \\
%
\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\}  \nonumber \\
%
\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2)  \hspace{2cm}
%
\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\}  \nonumber
\end{eqnarray}

Now, we show another example where different number of eigenfunctions and eigenvalues are used for every dimension. We consider a three-dimensional ($D=3$) input space, and sets of $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively. The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$. The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
%
\begin{eqnarray}
\mathbb{S}=
\left[ {\begin{array}{ccc}
1 & 1 & 1 \nonumber \\
1 & 1 & 2 \\
1 & 1 & 3 \\
1 & 2 & 1 \\
1 & 2 & 2 \\
1 & 2 & 3 \\
2 & 1 & 1 \\
2 & 1 & 2 \\
2 & 1 & 3 \\
2 & 2 & 1 \\
2 & 2 & 2 \\
2 & 2 & 3 
\end{array} } \right]
\end{eqnarray}

\noindent and the multidimensional eigenfunctions and eigenvalues
%
\begin{eqnarray}
\phi^{\ast}_1= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_1= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_2= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_2= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_3= \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_3= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_4= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_4= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_5= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_5= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_6= \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_6= \{\lambda_1(\mathbf{x}_1), \lambda_1(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_7= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_7= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_8= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_8= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_9= \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_9= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_{10}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{1}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{10}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_1(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_{11}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{2}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{11}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_2(\mathbf{x}_3)\} \nonumber \\
%
\phi^{\ast}_{12}= \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) \cdot \phi_{3}(\mathbf{x}_3) \hspace{1.5cm} \boldsymbol{\lambda}^{\ast}_{12}= \{\lambda_2(\mathbf{x}_1), \lambda_2(\mathbf{x}_2), \lambda_3(\mathbf{x}_3)\} \nonumber
\end{eqnarray}

\vspace{2mm}
\section*{Acknowledgment}

\bibliographystyle{biom}
\bibliography{references}

\end{document}

###########
#######
Next we show a little example of how to construct the new set of eigenfunctions and eigenvalues with input vector $\mathbf{x}=\{\mathbf{x}_1,\mathbf{x}_2\}$ in a two-dimensional input space ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension. As a result we obtain sets of $J^D=3^2=9$ new eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$. The eigenfunctions $\phi^{\ast}$ are the product combination of the unidimensional eigenfunctions $\phi$, and eigenvalues $\lambda^{\ast}$ is the $D$-vector of unidimensional eigenvalues $\lambda$:

\begin{eqnarray}
\phi^{\ast}_1(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_1 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_2(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_2 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_3(\mathbf{x}) = \phi_{1}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_3 = \{\lambda_{1}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_4(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_4 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_5(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_5 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_6(\mathbf{x}) = \phi_{2}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_6 = \{\lambda_{2}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_7(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{1}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_7 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{1}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_8(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{2}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_8 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{2}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
%
\phi^{\ast}_9(\mathbf{x}) = \phi_{3}(\mathbf{x}_1) \cdot \phi_{3}(\mathbf{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\mathbf{x}_2+L_2)\right) \nonumber \\
%
\boldsymbol{\lambda}^{\ast}_9 = \{\lambda_{3}(\mathbf{x}_1), \lambda_{3}(\mathbf{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber
\end{eqnarray}

### 2019-09-07 Paper/manuscript/manuscript_05.tex

@@
 \usepackage{subfigure}
-\usepackage{multirow}
+\usepackage{multirow,slashbox}
 \usepackage{float}
@@
 Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
-In this work we built these relationships for the square exponential covariance function.
-For other kernels, the relationships will be slightly different.
+In this work we built these relationships for the square exponential covariance function and Matern ($\nu=3/2$) covariance function.
+For other kernels, the relationships will be slightly different, in function mainly of the smoothness or wigglyness of the kernel effects.
 
@@
 
-\begin{figure}
-\begin{center}
-\begin{tabular}{ c c c}
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
+%\begin{figure}
+%\begin{center}
+%\begin{tabular}{ c c c}
+%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\
+%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
+%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
+%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
+%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
+%\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
+%\end{tabular}
+%\end{center}
+%\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
+%  \label{fig3_Post_part1}
+%\end{figure}
+
+\begin{figure}
+\begin{tabular}{ c c c c c}
+\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} &  \hspace{-0mm}
+\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\ 
+\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} & \hspace{-0mm}
+\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.35, trim = 25mm 63mm 120mm 50mm, clip]{fig3_legend.pdf}\\
+\includegraphics[scale=0.20, trim = 10mm 5mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} & \hspace{-0mm}
+\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} 
+\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} & 
 \end{tabular}
-\end{center}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
+\caption{Mean posterior predictive functions (left-first and right-first columns) and covariance functions (left-second and right-second columns) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig3_Post_part1}
@@
 
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_Matern.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\frac{3}{2}$) covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+  \label{fig6_lscale_vs_J_vs_c_Matern}
+\end{figure}
+
 If we look back to the conclusions drawn from Figures \ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig6_lscale_vs_J_vs_c}.
@@
 
-Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending on three variables, $m$, $c$ and $\ell$.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, and result in a surface depending on three variables, $m$, $c$ and $\ell$.
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
 However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
-
-(Gabi: The following paragraph is another idea to introduce that the diagnosis plot in Figure \ref{fig6_lscale_vs_J_vs_c} depends on the kernel considered.
-At the moment we have built this figure only for the square exponential kernel.)
-
-Ultimately, Figure \ref{fig6_lscale_vs_J_vs_c} showing the relationships among the number of basis functions, the boundary factor and the lengthscale of the function depends on the particular choice of covariance function.
-In this work we have built these relationships only for the square exponential covariance function.
-For other kernels, the Figure will be slightly different.
+\begin{center}
+•
+\end{center}
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
-\vspace{-2mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
+\subfigure{\includegraphics[scale=0.4, trim = 10mm 100mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
+\vspace{-2.5mm}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 7mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 7mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 10mm 7mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
 \vspace{-3mm}
@@
 \begin{flushleft}
-\begin{tabular}{ c c c c}
-\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 45mm, clip]{fig8_legend.pdf}}\\
-\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
+\begin{tabular}{ccc}
+\multicolumn{3}{c}{ \includegraphics[scale=0.35, trim = 10mm 115mm 80mm 45mm, clip]{fig8_legend.pdf}}\\
+\includegraphics[scale=0.28, trim = 0mm 26mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.28, trim = 20mm 26mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.28, trim = 20mm 26mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \\
+\includegraphics[scale=0.28, trim = 0mm 5mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.28, trim = 20mm 5mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.28, trim = 20mm 5mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf}
 \end{tabular}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}} \hspace{2mm}
 \subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
@@
 
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.27, trim = 0mm 0mm 5mm 0mm, clip]{fig20_time2D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.27, trim = 21mm 0mm 5mm 0mm, clip]{fig20_time3D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.27, trim = 21mm 0mm 5mm 0mm, clip]{fig20_time4D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 27mm 65mm 130mm 20mm, clip]{fig20_legend_diabetes_2.pdf}}
+\caption{Time of computation in the log scale of the different models fitted over the dataset with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+  \label{fig20_time2D_diabetes}
+\end{figure}
+
 \subsection{Study case VI: Leukemia data}\label{sec:bf_caseVI}
+The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
+The data set consist in survival times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for 1043 cases ($i=1,\cdots,n=1043$).
+Some 16\% of cases were censored.
+Predictors are \textit{age} ($\bm{x}_1$), \textit{sex} ($\bm{x}_2$), \textit{white blood cell} (WBC) ($\bm{x}_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($\bm{x}_4$) which is a measure of deprivation for district of residence.
+We denote the matrix $X=\{\bm{x}_1,\bm{x}_2,\bm{x}_3,\bm{x}_4\} \in {\rm I\!R}^{n\times 4}$ which contains the predictors.
+
+As the WBC measurements were strictly positive and highly skewed, we fit the model to its logarithm.
+Continuous predictors were normalized to have zero mean and unit standard deviation.
+Survival time was normalized to have zero mean for the logarithm of time.
+We assume a Gaussian observation model for the logarithm of the observed survival time, $t'_i=log(t_i)$, with a function of the predictors, $f(X)$, as the location parameter, and $\sigma$ as the Gaussian noise:
+%
+\begin{equation*}
+p(t'_i)= \mathcal{N}(t'_i|f(X_i),\sigma^2)
+\end{equation*}
+
+As we do not have a model for the censoring process, we do not have a full observation model, and the observational model for the censored data $t_i$ is assumed to be the complementary cumulative normal probability distribution:
+%
+\begin{equation*}
+p(y_i > t'_i)= \int_{t'_i}^{\infty} \mathcal{N}(y_i|f(X_i),\sigma^2) dy_i=  1 - \Phi \left( \frac{y_i-f(X_i)}{\sigma} \right),
+\end{equation*}
+
+\noindent where $y_i$ denotes the logarithm of the uncensored time.
+
+The latent function $f(X)$ is modeled as a Gaussian process, centered in a linear model of the predictors $X$, and with a squared exponential covariance function $k$ depending on predictors and hyperparameters $\theta=\{\alpha,\ell\}$,
+%
+\begin{eqnarray*} 
+f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta)).
+\end{eqnarray*}
+
+\noindent where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+Notice that a scalar lengthscale is considered in the multivariate covariance function.
+
+%The marginalized forms, by integrating out the latent values $f(X)$, of the previous previous observational models, observed and censored, result
+%%
+%\begin{align*}
+%p(t'_i)&= \int_{-\infty}^{\infty} \mathcal{N}(t'_i|c + \bm{\beta}X, K_{ii} + \sigma^2 )dt'_i,\\
+%p(y_i > t'_i)&= \int_{t'_i}^{\infty} \mathcal{N}(y_i|c + \bm{\beta}X, K_{ii} + \sigma^2) dy_i=  1 - \Phi \left( \frac{y_i-(c + \bm{\beta}X)}{K_{ii} + \sigma} \right),
+%\end{align*}
+%
+%\noindent with covariance matrix $K$, and its element $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$ is the evaluation of the covariance function $k$ at points $r$ and $s$.
+
+Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable (1 for female and 2 for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
+The relative contribution of a GP function given one of the levels of the predictor (\ref{eq:gpml2}) to a general mean GP function (\ref{eq:gpml1}) is defined.
+For the other level of the predictor, the GP function effects are set to zero.
+This multilevel construction is depicted as following:
+%
+\begin{eqnarray} 
+&f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta_1)) \label{eq:gpml1} \\
+&g(X|\bm{x}_2=2) \sim \mathcal{GP}(0, k(X,X, \theta_2|\bm{x}_2=2)) \label{eq:gpml2} \\ 
+&f(X|\bm{x}_2=2) = f(X|\bm{x}_2=2) + g(X|\bm{x}_2=2) \nonumber
+\end{eqnarray}
+
+\noindent In the previous equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of the GP function restricted to the male sex ($\bm{x}_2=2$).
+
+Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x}|x_2=2)$ are approximated as in equation (\ref{approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq:specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
+
+%Figure \ref{fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
+The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
+
+Figure \ref{fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
+These posterior estimates correspond to the HSPG model with $m=10$ basis functions and $c=3$ boundary factor.
+The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.70, trim = 0mm 0mm 0mm 0mm, clip]{fig21_posteriors_leukemia_01.pdf}
+\caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
+The thick line in each graph is the posterior mean estimated using a HSGP model, and the thin lines represent pointwise 95\% intervals.}
+  \label{fig21_posteriors_leukemia}
+\end{figure}
+
+Figure \ref{fig22_rmse_leukemia} shows the root mean square error (RMSE) and the time of computation in seconds per iteration as a function of the number of univariate basis functions $m$ and boundary factor $c$.
+As the functions are smooth, few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig22_rmse_leukemia}-right); Small boundary factors are not allowed when large lengthscales as can be seen in \ref{fig6_lscale_vs_J_vs_c}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig22_rmse_leukemia}-rleft).
+%
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig22_rmse_leukemia.pdf}} \hspace{2mm}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig23_time_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 22mm 65mm 120mm 0mm, clip]{fig24_legend_leukemia.pdf}}
+\caption{Root mean square error (RMSE) as a function of the number of basis functions $m$ and boundary factor $c$.}
+  \label{fig22_rmse_leukemia}
+\end{figure}
+
 \subsection{Study case VII: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
+The next example presents an spatio-temporal classification task in land-use of parcels between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits.
+A set of 200 parcels with known class are used for training the model and a set of 800 parcels in which to predict their unknown classes.
+The data is recorded in a time series of 5 years (2006,2008,2010,2012,2015) within the period.
+The class of each parcel $i$ and time $t$ is stored by a categorical variable $y_{it}$ representing the 5 different possible classes ($k=1,\cdots,K=5$): $k=1$, abandoned citrus fruits; $k=2$, aligned citrus fruits; $k=3$, irregular citrus fruits; $k=4$, scrub; $k=5$, young citrus fruits.
+
+A bunch of 52 characteristic variables were measured as predictors for every parcel and time.
+These variables were computed from satellite color images and cadastral map, and they concern spectral intensities and empirical semivariogram of the pixels within a parcel, as well as descriptive statistics of the shape of the parcels.
+
+In order to reduce the number of input variables and extract the most informative information contained among all of them, we computed a Principal Component Analysis on these variables and keep the 13 first principal components which will be used jointly with the time variable as predictors in the classifying GP model.
+Let's denote the matrix $X=\{x_{it,j}\}$ which contains the predictors.
+
+The underlying Markov chain distribution if fitted to the sequence of classes along time.
+This provides the probabilities transition matrix between classes (Figure \ref{fig25_transmatrix_landuses}), which is useful to check the smooth and monotonic correlation assumption of using a square exponential covariance function for the \textit{time} predictor in a Gaussian process.
+There is significant probability
+%
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig26_histlag_landuses.pdf}} \hspace{2mm}
+\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig25_transmatrix_landuses.pdf}} \hspace{2mm}
+\caption{Probabilities transition matrix between classes}
+  \label{fig25_transmatrix_landuses}
+\end{figure}
+
+The trained GP model is applied over another set of 800 parcels in the time series ($n'=800\cdot5=4000$) with unknown class and input values $X'=\{\bm{x}'_i\}_{i'=1}^{n'=800} \in {\rm I\!R}^{n\times 14}$,  with $\bm{x}'_i=\{x'_{ij}\}_{j=1}^{14} \in {\rm I\!R}^{14}$ with the aim to classify these parcels in space and time.
+
+Due to the fact that 14 (13 principal components + time) input variables are too many for a multivariate HSGP model as it scales $O(n\cdot m^D)$ with $m$ the number of basis functions in the HSGP model and $D$ the number of input variables, the multivariate HSGP model will be formulated as a first-order additive HSGP model.
+The use of a first-order additive model is justified due to the input variables correspond to principal components which are linearly uncorrelated each other..
+
+The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{i}=\{p_{ik}\}_{k=1}^{K=5}$ of belonging to class $k$ per parcel $i$,
+%
+\begin{equation*}
+y_i \sim \text{multinomial}(\bm{p}_i).
+\end{equation*}
+
+\noindent The goal is to estimate the vector of probabilities $\bm{p}_i$ as a function of the input values, which is a multivariate function $f(\bm{x}_i):\bm{x_i}\in {\rm I\!R}^{D} \rightarrow \{f_k(\bm{x}_i)\}_{k=1}^K \in {\rm I\!R}^{K}$, related to the vector of probabilities $\bm{p}_i$ through the 'softmax' link function.
+%
+\begin{eqnarray*}
+\bm{p}_i = \text{softmax}(f(\bm{x}_i)).
+\end{eqnarray*}
+
+Each individual function $f_k(\bm{x})$ is modeled as a first-order additive model, 
+%
+\begin{eqnarray*}
+f_k(\bm{x}) \sim \sum_{d=1}^{D} g_d(x^d)
+%\mathcal{HSGP}(x^d, S(\bm{\lambda}), \theta_d),
+\end{eqnarray*}
+
+\noindent where each individual component $g_d(x^d)$ is modeled as a unidimensional HSGP model,
+%
+\begin{eqnarray*}
+g_d(x^d) \sim \mathcal{HSGP}(x^d, S, \theta_d),
+\end{eqnarray*}
+
+In the HSGP model, the function $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
+The vector of hyperparameters $\theta_d=\{\alpha_d,\ell_d\}$ contains the marginal variance $\alpha_d$ and lengthscale $\ell_d$ of the HSGP process.
+
+For each individual component $g_d(x^d)$ modeled as a HSGP model, $m$=15 basis functions and a boundary factor $c$=2.5 were used.
+All the input variables were previously standarized with the exception of the time variable.
+
+After the adjustment of the model, the lengthscales estimates $\{\hat{\ell}_d\}_{d=1}^{D=14}$ normalized by half of the input range ($\frac{2\cdot \hat{\ell}_d}{|x^d_{max}-x^d_{min}|}$) are all bigger than the minimun lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as function of $m$, $c$.
+Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling the data.
+
+Table \ref{tab1} shows the confusion classification matrix after fit the model following a $Q$-fold cross-validation procedure over the training data. Every fold was composed of $Q$=10 observations.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|*{6}{c|}}\hline
+\backslashbox{Estimate}{True} & \multicolumn{1}{p{1.5cm}|}{\centering caind \\(k=1)} &  \multicolumn{1}{p{1.5cm}|}{\centering calign \\(k=2)} & \multicolumn{1}{p{1.5cm}|}{\centering irr \\(k=3)} & \multicolumn{1}{p{1.5cm}|}{\centering aband \\(k=4)} & \multicolumn{1}{p{1.5cm}|}{\centering young \\(k=5)} & \multicolumn{1}{p{1.5cm}|}{\centering } \\
+\hline \multicolumn{1}{|p{3cm}|}{ \centering caind (k=1)} &90&39&14&3&11&42\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering calign (k=2)} &46&301&8&2&3&16\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering irr (k=3)}  &8&4&59&4&1&19\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering aband (k=4)}  &5&2&6&342&5&5\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering young (k=5)}  &8&2&1&0&38&19\%\\
+\hline \multicolumn{1}{|p{3cm}|}{\centering }  &42\%&13\%&32\%&2\%&34\%&$\textbf{17\%}$\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Confusion matrix after the Q-fold cross-validation procedure over the training data.}
+  \label{tab1}
+\end{table}
+
+Figure \ref{fig21_classifying_map} shows the image map of predicted classes for the whole set of parcels of the territory.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.35, trim = 0mm 0mm 0mm 0mm, clip]{recorte_newparcels3.png}
+\caption{Image classification map with the predicted classes for  the whole set, training and predicting, classes.}
+  \label{fig21_classifying_map}
+\end{figure}
 

### 2020-03-14 Paper/manuscript_v1/manuscript_06.tex
update in manuscript_v2 and online appendice

--- 2019-09-07 Paper/manuscript/manuscript_06.tex
@@
 
+%% Add a new subsection
+\usepackage{titlesec}
+\usepackage{hyperref}
+
+\titleclass{\subsubsubsection}{straight}[\subsection]
+
+\newcounter{subsubsubsection}[subsubsection]
+\renewcommand\thesubsubsubsection{\thesubsubsection.\arabic{subsubsubsection}}
+\renewcommand\theparagraph{\thesubsubsubsection.\arabic{paragraph}} % optional; useful if paragraphs are to be numbered
+
+\titleformat{\subsubsubsection}
+  {\normalfont\normalsize\bfseries}{\thesubsubsubsection}{1em}{}
+\titlespacing*{\subsubsubsection}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+
+\makeatletter
+\renewcommand\paragraph{\@startsection{paragraph}{5}{\z@}%
+  {3.25ex \@plus1ex \@minus.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries}}
+\renewcommand\subparagraph{\@startsection{subparagraph}{6}{\parindent}%
+  {3.25ex \@plus1ex \@minus .2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries}}
+\def\toclevel@subsubsubsection{4}
+\def\toclevel@paragraph{5}
+\def\toclevel@paragraph{6}
+\def\l@subsubsubsection{\@dottedtocline{4}{7em}{4em}}
+\def\l@paragraph{\@dottedtocline{5}{10em}{5em}}
+\def\l@subparagraph{\@dottedtocline{6}{14em}{6em}}
+\makeatother
+
+\setcounter{secnumdepth}{4}
+\setcounter{tocdepth}{4}
+%%
+
 \begin{document}
@@
 
-\section{Introduction}\label{sec:bf_intro}
+\section{Introduction}\label{sec_bf_intro}
 
@@
 This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
-Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks (cite!).
+Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
 This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation.
@@
 
-While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
+While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
 In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
@@
 \vspace{3mm}
-\section{Method}\label{sec:bf_method}
+\section{Method}\label{sec_bf_method}
 
@@
 $\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$,
- where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$, which needs be positive semidefinite.
+ where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$, which needs be symmetric and positive semi-definite.
 The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
@@
 The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
-A covariance function needs be symmetric and positive semidefinite.
+A covariance function needs be symmetric and positive semi-definite.
 A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
@@
 \begin{align}
-S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq:specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq:specdens_32} 
+S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
@@
 More details and mathematical proofs are provided in \cite{solin2018hilbert}.
+Our starting point for presenting the method is the main result obtained by \cite{solin2018hilbert} of the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
+The mathematical details of this approximation have been briefly presented in the Appendix \ref{app_details_formulas_approx_covfun} of this work.
 
@@
 %
-\begin{equation}\label{approxcov}
+\begin{equation}\label{eq_approxcov}
 k(x,x') = \sum_{j=1}^\infty S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
@@
 \noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section 2.1) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
-The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
 Namely, they satisfy the following eigenvalue problem in $\Omega$ when applying the Dirichlet boundary condition (other boundary conditions could be used as well):
 %
-\begin{eqnarray}\label{eigenproblem}
+\begin{eqnarray}\label{eq_eigenproblem}
 \begin{split}
@@
 
-\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in Equation (\ref{eigenproblem}) are sinusoidal functions.
+\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in Equation (\ref{eq_eigenproblem}) are sinusoidal functions.
 Independently of the covariance function, they can be computed as
@@
 \begin{align}
-\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \label{eigenvalue}\\
-\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \label{eigenfunction}
+\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \label{eq_eigenvalue}\\
+\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \label{eq_eigenfunction}
 \end{align}
 
-If we truncate the sum in (\ref{approxcov}) to the first $m$ terms, the approximate covariance function becomes
+If we truncate the sum in (\ref{eq_approxcov}) to the first $m$ terms, the approximate covariance function becomes
 %
@@
 %
-\begin{equation}\label{approxf}
+\begin{equation}\label{eq_approxf}
 f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j,
@@
 The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions.
-Therefore, Equation (\ref{approxf}) can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
+Therefore, Equation (\ref{eq_approxf}) can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
 The computational cost of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
 
-The parameterization in (\ref{approxf}) is naturally in the non-centered
+The parameterization in (\ref{eq_approxf}) is naturally in the non-centered
 parameterization form with independent prior distribution on
@@
 %
-\begin{align} \label{m_multi}
+\begin{align} \label{eq_m_multi}
 m^{\ast} = \prod_{d=1}^{D} m_d,
@@
 \begin{align}
-\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenfunction_multi} \\
-%
-\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenvalue_multi}
+\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eq_eigenvalue_multi} \\
+%
+\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eq_eigenfunction_multi}
 \end{align}
@@
 %
-\begin{equation}\label{approxcov_multi}
+\begin{equation}\label{eq_approxcov_multi}
 k(\bm{x},\bm{x}') \approx \sum_{j=1}^{m^{\ast}} 
@@
 %
-\begin{equation}\label{approxf_multi}
+\begin{equation}\label{eq_approxf_multi}
 f(\bm{x}) \approx \sum_{j=1}^{m^{\ast}} 
@@
 
-\subsection{Learning hyperparameters and model inference}
-
-\vspace{2mm}
-$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a linear model.
-
-\vspace{2mm}
-$\bullet$ The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.
-
-\vspace{2mm}
-$\bullet$ The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands.
-The weights depend on the hyperparameters, then they have to be computed in every step of the HMC sampling method.
-
-\vspace{2mm}
-$\bullet$ The linear model is computed with complexity $O(nm)$, computed in every step of the HMC sampling method.
-
-\vspace{2mm}
-$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
-After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
-(Arno's paper, pag.
-7)
-
-\vspace{2mm}
-$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
-especially when sampling methods for inference are used.
-GP prior as latent functions is needed in generalized models.
-
-In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
-
-\vspace{2mm}
-$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
-This operation has to be computed at every step of the HMC or optimizer.
-
-\vspace{2mm}
-$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
-It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
-
-\vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
-In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
-
-\vspace{2mm}
-$\bullet$ In addition, the computation of the automatic differentiation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ In a regular GP model the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
-This operation has to be computed at every step of the HMC.
-
-\vspace{2mm}
-$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-
 \section{The accuracy of the approximation}
@@
 
-\subsection{Dependency on the number of basis functions and the boundary condition}
+\subsection{Dependency on the number of basis functions and the boundary condition} \label{sec_dependency}
 
@@
 %
-\begin{equation}\label{diffcov}
-k(\tau) = \sum_{j=1}^{\infty} S_{\theta} \left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0),  \nonumber
-\end{equation}
+\begin{equation*}
+k(\tau) = \sum_{j=1}^{\infty} S_{\theta} \left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0), 
+\end{equation*}
 
 \noindent where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}) for the unidimensional case and in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}) for the multidimensional case.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in Equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
 The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
 %
-\begin{eqnarray}\label{diff_covs}
+\begin{eqnarray}\label{eq_diff_covs}
 \int k(\tau) \, d\tau - 
@@
 
-The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
-The approximation also depends on the boundary $L$ (see Equations (\ref{eigenvalue}), (\ref{eigenfunction}), (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
+The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{eq_diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+The approximation also depends on the boundary $L$ (see Equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
 As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
@@
 %
-\begin{equation}\label{eq:boundary}
+\begin{equation}\label{eq_boundary}
 L=c \cdot S,
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 5mm 0mm 5mm 0mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig1_Post_J.pdf}}
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 5mm 0mm 5mm 0mm, clip]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig2_Post_L.pdf}}
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
@@
 a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
+More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{eq_diff_covs}).
 Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern ($\nu=3/2$) covariance function.
-We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
+We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function,
 %
@@
 
-\noindent Alternatively, this figure could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
+\noindent Alternatively, these figures could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
 Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$.
@@
 
-(Gabi: The following paragraph is just an idea to extend the discussion to the multidimensional case)
-
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, and result in a surface depending on three variables, $m$, $c$ and $\ell$.
-An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
-However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
-\begin{center}
-\end{center}
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
+An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be graphically represented.
+However, as an approximation we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} to analyze individually the different dimensions of a multidimensional GP model.
+
+(Gabi: Perhaps, in the two-dimensional case, whether the approximation is close enough might depend only on the ratio between the variables $\ell_1$ and $\ell_2$.)
 
@@
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
+\subfigure{\includegraphics[scale=0.30, trim = 1mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
+\subfigure{\includegraphics[scale=0.30, trim = 1mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 10mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
+\subfigure{\includegraphics[scale=0.30, trim = 1mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
 \vspace{-3mm}
-\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
+\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
   \label{fig7_posterior_varing_lscale_part1}
@@
 \begin{tabular}{ccc}
-\multicolumn{3}{c}{ \includegraphics[scale=0.40, trim = 25mm 115mm 90mm 45mm, clip]{fig8_legend.pdf}}\\
+\multicolumn{3}{c}{ \includegraphics[scale=0.40, trim = 25mm 117mm 90mm 45mm, clip]{fig8_legend.pdf}}\\
 \includegraphics[scale=0.30, trim = 0mm 24mm 7mm 10mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
@@
 \end{flushleft}
-\caption{Data-generating functional lengthscales (X-axis), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates from the regular GP and HSGP models (Y-axis).
+\caption{Data-generating functional lengthscales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates $(\hat{\ell})$ from the regular GP and HSGP models.
 95\% confident intervals of the lengthscale estimates are plotted as dot lines.
@@
 \includegraphics[scale=0.38]{fig9_MSE_varing_lscale.pdf}
-\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects (lengthscales).}
+\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects ($\ell$).}
   \label{fig9_MSE_varing_lscale}
@@
 %
-\begin{eqnarray*}\label{eq:latentgp_simudata1}
+\begin{eqnarray*}\label{eq_latentgp_simudata1}
 \begin{split}
@@
 %
-\begin{equation*}\label{eq:marginalizedgp_simudata1}
-\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I )
+\begin{equation*}\label{eq_marginalizedgp_simudata1}
+\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
 \end{equation*}
@@
 
-In the HSGP model, the latent function values $f(\bm{x})$ are approximated as (\ref{approxf}),
+In the HSGP model, the latent function values $f(\bm{x})$ are approximated as (\ref{eq_approxf}),
 %
@@
 
-\noindent with the spectral density $S$ (\ref{eq:specdens_32}) as a function of $\sqrt{\lambda_j}$,
+\noindent with the spectral density $S$ (\ref{eq_specdens_32}) as a function of $\sqrt{\lambda_j}$,
 %
@@
 
-\noindent and eigenvalues $\lambda_j$ (\ref{eigenvalue}) and eigenfunctions $\phi_j$ (\ref{eigenfunction})
+\noindent and eigenvalues $\lambda_j$ (\ref{eq_eigenvalue}) and eigenfunctions $\phi_j$ (\ref{eq_eigenfunction})
 %
@@
 
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
-
-Figure \ref{fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$ (\ref{eq:boundary})), and the splines model with 80 knots.
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
+
+Figure \ref{fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$ (\ref{eq_boundary})), and the splines model with 80 knots.
 The true data-generative function and the noisy observations are also plotted in the Figure; the sample observations are plotted as circles and the out-of-sample or test data, which have not been taking part on training the models, are plotted as crosses.
@@
 \includegraphics[width=\textwidth]{fig10_Posteriors_exI.pdf}
-\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dot lines.}
+\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dashed lines.}
   \label{fig10_Posteriors_exI}
@@
 
-In order to assess performance of the models as a function the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
+In order to assess performance of the models as a function the number of basis functions and number of knots, different models with different number of basis functions, for the HSGP model, and different number of knots, for the splines model, have been fitted.
 Figure \ref{fig11_MSE_exI_inter} shows the standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
@@
 \subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_extra.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 55mm 110mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 29mm 55mm 110mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
 \caption{Standarized root mean square error (SRMSE) of the different methods against the data-generating function.
@@
 (right) SRMSE for extrapolation.
-The Standard deviation of the mean of the SRMSE are plotted as dot lines.}
+The standard deviation of the mean of the SRMSE is plotted as dashed lines.}
   \label{fig11_MSE_exI_inter}
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig11_time_exI.pdf}}
-\caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
+\caption{Computational time (Y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the Splines model.
+The Y-axis is in a logarithmic scale.
+The standard deviation of the computational time is plotted as dashed lines.}
   \label{fig11_time_exI}
@@
 %
-\begin{eqnarray*} \label{eq:gpprior_gay}
+\begin{eqnarray*} \label{eq_gpprior_gay}
 p_i &=& \text{logit}(f(x_i)) \nonumber \\
@@
 
-In the HSGP model the function $f(\bm{x})$ is approximated as in Equation (\ref{approxf}), with the square exponential spectral density as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
-
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
+In the HSGP model the function $f(\bm{x})$ is approximated as in Equation (\ref{eq_approxf}), with the square exponential spectral density as in Equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+A Bayesian approach is used to fit this splines model using the R-package \textit{brms} \citep{burkner2017brms}.
 
@@
 \includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
-\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
+\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dashed lines.}
   \label{fig12_Posteriors_gaydata}
@@
 For the HSGP model, different models with different number of basis functions and boundary factor have been fitted.
-The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model and plotted as a function of the number of basis functions $m$ and the boundary factor $c$ in Figure \ref{fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
+The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
 The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
@@
 (left) RMSE for sample data.
-(right) RMSE for test data.}
+(right) RMSE for out-of-sample data.}
   \label{fig13_MSE_train_BF_gaydata}
@@
 
-Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration, as a function of the number of basis functions $m$ for the HSGP model and knots for the splines model.
+Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the splines model.
 The computational times is represented in the Y-axis which is in a logarithmic scale.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 25mm 30mm 123mm 0mm, clip]{fig14_legend_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 5mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 27mm 35mm 112mm 0mm, clip]{fig14_legend_gaydata.pdf}}
 \caption{Root mean square error (RMSE) of the different methods, regular GP, HSGP and splines models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots for the splines model.
 (left) RMSE for sample data.
-(right) RMSE for test data.}
+(right) RMSE for out-of-sample data.}
   \label{fig14_MSE_train_gaydata}
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig14_1_time_gaydata.pdf}}
-\caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
+\caption{Computational time (Y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$ and knots.
+The Y-axis is in a logarithmic scale.}
   \label{fig14_1_time_gaydata}
@@
 
+\newpage
 \subsection{Study case III: Birthday data}
-
+This example is an analysis of patterns in birthday frequencies in a dataset containing records of all births in the United States on each day during the years 1969–1988.
+The model decomposes the number of births along all the period in longer-term trend effects, patterns during the year, day-of-week effects, and special days effects.
+The special days effects cover patterns such as possible fewer births on Halloween, christmas or new year, and excess of births on Valentine’s Day or the days after christmas (due, presumably, to choices involved in scheduled deliveries, along with decisions of whether to induce a birth for health reasons).
+This analysis was originally addressed in \cite{gelman2013bayesian}.
+The total number of days within the period is 7305 ($t=1,\dots,T=7305$), then a regular GP model is unfeasible to be fitted on this dataset as we know inference scales $O(T^3)$ in covariance matrix inversion.
+Therefore, an approximate approach has to be used to fit a GP model on this data.
+We will use the HSGP model developed in this work, as well as a low-rank GP model with a periodic covariance function (Appendix \ref{sec_cov_periodic}) based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
+
+Let's denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
+%
+\begin{equation*}
+y_{t} \sim \mathcal{N}(\mu(t),\sigma^2).
+\end{equation*}
+
+\noindent The mean function $\mu(t)$ will be defined as an additive model in the form: 
+%
+\begin{eqnarray} \label{eq_mean_brithday}
+\mu(t) = f_1(t) + f_2(t) + f_3(t) + f_4(t),
+\end{eqnarray}
+
+\noindent where $f_1(t)$ represents the long-term trends modeled by a GP with squared exponential covariance function,
+%
+\begin{eqnarray*}
+f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\bigg(-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\bigg), 
+\end{eqnarray*}
+
+\noindent $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic square exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
+%
+\begin{eqnarray*}
+f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\bigg(-\frac{2\text{sin}^2(\pi(t-t')/365.25}{\ell_2^2}\bigg), 
+\end{eqnarray*}
+
+\noindent $f_3(t)$ represents the weekly smooth pattern using a periodic square exponential covariance function (with period 7 of length of the week) in a GP model,
+%
+\begin{eqnarray*}
+f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\bigg(-\frac{2\text{sin}^2(\pi(t-t')/7}{\ell_3^2}\bigg), 
+\end{eqnarray*}
+
+\noindent and $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{piironen2017sparsity}:
+%
+\begin{eqnarray*}
+f_4(t) \sim \mathcal{N}(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
+\end{eqnarray*}
+
+\noindent A horse-shoe prior allows for sparse distributed effects.
+Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
+Different levels of sparsity can be accommodated by changing the value of $\tau$: with large $\tau$ all the variables have very diffuse priors with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
+%More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.
+
+GP priors have been defined over the components $f_1(t)$, $f_2(t)$ and $f_3(t)$.
+Then, low-rank representation of the GP priors have to be used in the modeling and inference.
+The component $f_1(t)$ will be approximated using the HSGP model.
+Thus, the function values $f_1(t)$ are approximated as in Equation (\ref{eq_approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+
+The year effects $f_2(t)$ and week effects $f_3(t)$, as they use a periodic covariance function, they do no fit under the framework of the HSGP approximation covered in this work.
+However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators \citep{solin2014explicit}, which approximation has been briefly covered in Appendix \ref{sec_cov_periodic}.
+Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in Equation \ref{eq_f_period}, with variance coefficients $\tilde{q}_j^2$ as in Equation \ref{eq_q_2}.
+
+For the component $f_1(t)$, $m$=30 basis functions and a boundary factor $c$=1.5 were used.
+The lengthscale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimun lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as function of $m$ and $c$.
+Which means that the used number of basis functions and boundary factor are suitable values for modeling accurately the input effects.
+
+For the components $f_2(t)$ and $f_3(t)$, $J$=10 cosine terms were used.
+Similarly to above, the  lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimun lengthscale reported by Figure \ref{fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are accurate enough.
+
+Figure \ref{fig27_posteriors_birthday} shows the posterior mens of the long-term trend $f_1(t)$ and year patterns $f_2(t)$ for the whole period, jointly with the observed data.
+Figure \ref{fig27_posteriors_oneyear_birthday} show the process for one year (1972) only.
+In this figure, the special days effects $f_4(t)$ in the year can be clearly represented.
+The posterior means of the the function $\mu(t)$ (\ref{eq_mean_brithday}), and the components $f_1(t)$ (long-term trend) and $f_2(t)$ (year pattern) are also plotted.
+Figure \ref{fig27_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3(t)$ can be clearly represented.
+The mean of the the function $\mu(t)$ (\ref{eq_mean_brithday}) and components $f_1(t)$ (long-term trend), $f_2(t)$ (year pattern) and $f_4(t)$ (special-days effects) are also plotted.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig27_posteriors_birthday.pdf}}
+\caption{Posterior means of the function $\mu(\cdot)$ for the whole series. The long-term trend ($f_1(\cdot)$) and year effects ($f_2(\cdot)$) are also plotted. }
+  \label{fig27_posteriors_birthday}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig27_posteriors_oneyear_birthday.pdf}}
+\caption{Posterior means of the function $\mu(\cdot)$ for the year 1972 of the series.
+The special days effects pattern ($f_4(\cdot)$) in the year is also represented, as well as the long-term trend ($f_1(\cdot)$) and year effects pattern ($f_2(\cdot)$).
+}
+  \label{fig27_posteriors_oneyear_birthday}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig27_posteriors_onemonth_birthday.pdf}}
+\caption{Posterior means of the function $\mu(\cdot)$ for the month of January of 1972.
+The week effects pattern ($f_3(\cdot)$) in the month is also represented, as well as the long-term trend ($f_1(\cdot)$), year effects pattern ($f_2(\cdot)$) and special days effects pattern ($f_4(\cdot)$).
+}
+  \label{fig27_posteriors_onemonth_birthday}
+\end{figure}
+
+\newpage
 \vspace{5mm}
@@
 This example consists of a simulated dataset with 120 ($i=1,\dots,n=120$) single draws from a Gaussian process prior with two ($D=2$) input dimensions.
-A square exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$ for dimension 1 and $\ell_2=0.35$ for dimension 2, is used.
+A square exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$, for dimension 1, and $\ell_2=0.35$, for dimension 2, is used.
 The corresponding matrix of input values is $X=\{\bm{x}_i\}_{i=1}^{n=120} \in {\rm I\!R}^{n\times 2}$ with $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$.
@@
 %
-\begin{eqnarray*}\label{eq:latentgp_simudata2}
+\begin{eqnarray*}\label{eq_latentgp_simudata2}
 \begin{split}
@@
 %
-\begin{equation*}\label{eq:marginalizedgp_simudata2}
+\begin{equation*}\label{eq_marginalizedgp_simudata2}
 \bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
@@
 
-In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{approxf_multi}),
+In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{eq_approxf_multi}),
 %
@@
 
-\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) which elements are the univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent where $S$ is the spectral density (\ref{eq_specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}$, of the $D$-dimensional square exponential covariance function,
+%
+\begin{eqnarray*}
+S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \prod_{d=1}^D \ell_d  \mathrm{exp}\bigg(-\frac{1}{2} \sum_{d=1}^D \ell_d^2  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg),
+\end{eqnarray*}
+
+\noindent and $\bm{\lambda}_j$ is the $D$-vector (\ref{eq_eigenvalue_multi}) which elements are the univariate eigenvalues (\ref{eq_eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
@@
 
-\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}) as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eq_eigenfunction_multi}) as the product of univariate eigenfunctions (\ref{eq_eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
 %
@@
 \noindent where $x_d$ is the input value corresponding to dimension $d$.
-$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left\{ \sqrt{\lambda_{\mathbb{S}_{jd}}} \right\}_{d=1}^{D}$, of the $D$-dimensional square exponential covariance function,
-%
-\begin{eqnarray*}
-S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \prod_{d=1}^D \ell_d  \mathrm{exp}\bigg(-\frac{1}{2} \sum_{d=1}^D \ell_d^2  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg). 
-\end{eqnarray*}
-
-\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$.
-$\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the $D$ dimensions.
+
+\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{eq_m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$.
+$\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eq_eigenvalue}) over the $D$ dimensions.
 $L_d$ is the boundary for the dimension $d$.
@@
 
-In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
-
-Figure \ref{fig15_Posteriors_exII} shows the data-generating GP function from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset.
+In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
+
+Figure \ref{fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset.
 Sample observations are also plotted in the plots as circles.
@@
 
+Figure \ref{fig16_errors_exII} shows the difference functions between the data-generating function and the GP, HSGP and splines models, respectively.
+
 \begin{figure}
@@
 \begin{tabular}{ c c c}
-\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 0mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\
-\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
+\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 18mm 0mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\
+\includegraphics[scale=0.35, trim = 0mm 18mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 18mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
 \end{tabular}
@@
 (c) Mean posterior predictive function of the HSGP model.
-(d) Mean posterior predictive function of the splines (SP) model.}
+(d) Mean posterior predictive function of the splines model.
+Sample points are plotted as circles}
   \label{fig15_Posteriors_exII}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 39.2mm 20mm, clip]{fig16_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 19mm 23mm 39.2mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 19mm 23mm 39.2mm 20mm, clip]{fig16_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 0mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
-\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models.}
+\subfigure{\includegraphics[scale=0.35, trim = 1mm 18mm 39.2mm 20mm, clip]{fig16_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 19mm 18mm 39.2mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 19mm 18mm 39.2mm 20mm, clip]{fig16_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.33, trim = 150mm 13mm 0mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
+\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models. Sample points are plotted as circles.}
   \label{fig16_errors_exII}
@@
 
-In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
+In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions, for the HSGP model, and different number of knots, for the splines model, have been fitted.
 In all models, the same number of basis functions and knots per dimension were used.
-Figure \ref{fig17_RMSE_exII}(left) shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, the number of univariate basis functions $m$ and knots.
+Figure \ref{fig17_RMSE_exII}(left) shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, and the number of univariate basis functions $m$, for the HSGP model, and knots, for the splines model.
 From Figures \ref{fig16_errors_exII} and \ref{fig17_RMSE_exII}(left), it can be seen a close approximation of the HSGP model to the regular GP model.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig17_RMSE_exII.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 30mm 123mm 0mm, clip]{fig17_legend_exII.pdf}}
-\caption{Root mean square error (RMSE) (left) and computational time in seconds per iteration (right) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+\subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 5mm 0mm, clip]{fig17_RMSE_exII.pdf}}
+\subfigure{\includegraphics[scale=0.37, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 113mm 0mm, clip]{fig17_legend_exII.pdf}}
+\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
   \label{fig17_RMSE_exII}
@@
 
-\subsection{Study case V: Diabetes data}\label{sec:bf_caseV}
+\subsection{Study case V: Diabetes data}\label{sec_bf_caseV}
 The next example presents an epidemiological study of the diabetes diseases.
@@
 %
-\begin{eqnarray*} \label{eq:gpprior_gay}
+\begin{eqnarray*} \label{eq_gpprior_gay}
 p_i &=& \text{logit}(f(\bm{x}_i)) \nonumber \\
@@
 
-In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as in equation (\ref{approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq:specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$, and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
+In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$, and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 12mm 23mm 35mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 35mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 38mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 5mm 23mm 39mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 39mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 39mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
 \subfigure{\includegraphics[scale=0.31, trim = 150mm 13mm 2mm 15mm, clip]{fig18_legend_diabetes.pdf}}
-\caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the splines (SP) model.}
+\caption{(a) Mean posterior predictive functions of the GP model.
+(b) Mean posterior predictive functions of the HSGP model.
+(c) Mean posterior predictive functions of the splines (SP) model.
+Samples observations of suffering (red points) and not suffering (blue points) form the disease are plotted.}
   \label{fig18_gpfun_diabetes}
@@
 In all models, the same boundary factor, number of basis functions and knots per dimension were used.
-Figure \ref{fig19_ELPD_diabetes} shows the expected log predictive density (ELPD) as a function of the boundary factor $c$, the number of univariate basis functions $m$ and knots.
+Figure \ref{fig19_ELPD_diabetes} shows the expected log predictive density (ELPD) as a function of the boundary factor $c$ and the number of univariate basis functions $m$, for the HSGP model, and knots, for the splines model.
 The ELPD is computed over the actual observations.
 Basically, with slightly differences, all models show similar performances, due to the fact that the process is very smooth with a relatively very large lengthscale estimate $\ell=4.51$.
-Even tough, a slightly pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (\ref{fig6_lscale_vs_J_vs_c}).
+Even tough, a slightly pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (Figure \ref{fig6_lscale_vs_J_vs_c}).
 %
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{fig19_ELPD_diabetes.pdf}}
-\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$ and knots.}
+\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$, for the HSGP model, and knots, for the splines model.}
   \label{fig19_ELPD_diabetes}
@@
 
-Figure \ref{fig20_time2D_diabetes} shows the computational times in the logarithm scale of the different models, regular GP, HSGP and splines, fitted over the dataset with 2 input dimensions (left), 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots.
+Figure \ref{fig20_time2D_diabetes} shows the computational times of the different models, regular GP, HSGP and splines, fitted over the dataset, with 2 input dimensions (left), 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$ and number of univariate basis functions $m$ and knots.
 We can appreciate the significant increase of computational time with higher dimensions for the HSGP and splines models.
@@
 \centering
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{fig20_time2D_diabetes_2.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 30.2mm 0mm 10mm 0mm, clip]{fig20_time3D_diabetes_2.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 30.2mm 0mm 10mm 0mm, clip]{fig20_time4D_diabetes_2.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 25mm 53mm 123mm 10mm, clip]{fig20_legend_diabetes_2.pdf}}
-\caption{Time of computation in the log scale of the different models fitted over the dataset with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+\subfigure{\includegraphics[scale=0.28, trim = 1mm 0mm 10mm 0mm, clip]{fig20_time2D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.28, trim = 30.2mm 0mm 10mm 0mm, clip]{fig20_time3D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.28, trim = 30.2mm 0mm 10mm 0mm, clip]{fig20_time4D_diabetes_2.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 60mm 112mm 10mm, clip]{fig20_legend_diabetes_2.pdf}}
+\caption{Time of computation in seconds per iteration (iteration of the HMC sampling method) of the different models fitted over the dataset, with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$ and number of basis functions $m$, for the HSGP model, and knots, for the splines model.}
   \label{fig20_time2D_diabetes}
@@
 
-\subsection{Study case VI: Leukemia data}\label{sec:bf_caseVI}
+\subsection{Study case VI: Leukemia data}\label{sec_bf_caseVI}
 The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
@@
 
-The latent function $f(X)$ is modeled as a Gaussian process, centered in a linear model of the predictors $X$, and with a squared exponential covariance function $k$ depending on predictors and hyperparameters $\theta=\{\alpha,\ell\}$,
-%
-\begin{eqnarray*} 
-f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta)).
+The latent function $f(X)$ is modeled as a Gaussian process, centered in a linear model of the predictors $X$, and with a squared exponential covariance function $k$ depending on predictors $X$ and hyperparameters $\theta=\{\alpha,\ell\}$,
+%
+\begin{eqnarray*} 
+f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta)),
 \end{eqnarray*}
@@
 Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable (1 for female and 2 for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
-The relative contribution of a GP function given one of the levels of the predictor (\ref{eq:gpml2}) to a general mean GP function (\ref{eq:gpml1}) is defined.
+The relative contribution of a GP function given one of the levels of the predictor (\ref{eq_gpml2}) to a general mean GP function (\ref{eq_gpml1}) is defined.
 For the other level of the predictor, the GP function effects are set to zero.
@@
 \begin{eqnarray} 
-&f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta_1)) \label{eq:gpml1} \\
-&g(X|\bm{x}_2=2) \sim \mathcal{GP}(0, k(X,X, \theta_2|\bm{x}_2=2)) \label{eq:gpml2} \\ 
+&f(X) \sim \mathcal{GP}(c + \bm{\beta}X, k(X,X, \theta_1)) \label{eq_gpml1} \\
+&g(X|\bm{x}_2=2) \sim \mathcal{GP}(0, k(X,X, \theta_2|\bm{x}_2=2)) \label{eq_gpml2} \\ 
 &f(X|\bm{x}_2=2) = f(X|\bm{x}_2=2) + g(X|\bm{x}_2=2) \nonumber
@@
 
-Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x}|x_2=2)$ are approximated as in equation (\ref{approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq:specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
+Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x}|x_2=2)$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.
 
@@
 Figure \ref{fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
-These posterior estimates correspond to the HSPG model with $m=10$ basis functions and $c=3$ boundary factor.
+These posterior estimates correspond to the HSGP model with $m=10$ basis functions and $c=3$ boundary factor.
 The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
@@
 \centering
-\includegraphics[scale=0.70, trim = 0mm 0mm 0mm 0mm, clip]{fig21_posteriors_leukemia_01.pdf}
+\includegraphics[scale=0.70, trim = 0mm 0mm 0mm 0mm, clip]{fig21_posteriors_leukemia_02.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
@@
 
-Figure \ref{fig22_rmse_leukemia} shows the root mean square error (RMSE) and the time of computation in seconds per iteration as a function of the number of univariate basis functions $m$ and boundary factor $c$.
-As the functions are smooth, few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig22_rmse_leukemia}-right); Small boundary factors are not allowed when large lengthscales as can be seen in \ref{fig6_lscale_vs_J_vs_c}.
-Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig22_rmse_leukemia}-rleft).
+Figure \ref{fig22_rmse_leukemia} shows the root mean square error (RMSE) computed against the regular GP, and the time of computation as a function of the number of univariate basis functions $m$ and boundary factor $c$.
+As the functions are smooth, few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig22_rmse_leukemia}-right); Small boundary factors are not allowed when large lengthscales as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig22_rmse_leukemia}-left).
 %
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 120mm 0mm, clip]{fig24_legend_leukemia.pdf}}
-\caption{Root mean square error (RMSE) and time of computation (in seconds per iteration) as a function of the number of basis functions $m$ and boundary factor $c$.}
+\caption{Root mean square error (RMSE) computed against the GP and time of computation in seconds per iteration (iteration of the HMC sampling method) as a function of the number of basis functions $m$ and boundary factor $c$.}
   \label{fig22_rmse_leukemia}
@@
 
-%\subsection{Study case VII: Land use spatio-temporal classification task (first form)}\label{sec:bf_caseVII}
-%The next example presents an spatio-temporal classification task in land-use of parcels between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits.
-A set of 200 parcels with known class are used for training the model and a set of 800 parcels in which to predict their unknown classes.
-The data is recorded in a time series of 5 years (2006,2008,2010,2012,2015) within the period.
-The class of each parcel $i$ and time $t$ is stored by a categorical variable $y_{it}$ representing the 5 different possible classes ($k=1,\cdots,K=5$): $k=1$, adult independent citrus fruits; $k=2$, aligned citrus fruits; $k=3$, irregular citrus fruits; $k=4$, abandoned citrus fruits; $k=5$, young citrus fruits.
-%
-%A bunch of 52 characteristic variables were measured as predictors for every parcel and time.
-These variables were computed from satellite color images and cadastral map, and they concern spectral intensities and empirical semivariogram of the pixels within a parcel, as well as descriptive statistics of the shape of the parcels.
-%
-%Due to the fact that 52 input variables are too many for a multivariate HSGP model as it scales $O(n\cdot m^D)$ with $m$ the number of basis functions in the HSGP model and $D$ the number of input variables, the multivariate HSGP model will be formulated as an additive HSGP model.
-%
-%In order to avoid high-order components in the additive HSGP model as much as possible, since it is known multivariate HSGP models to be computationally demanding, a Principal Component Analysis is carried out over the characteristic variables.
-The resulting principal components are linearly uncorrelated each other and will be used jointly with the time variable as predictors in the classifying first-order additive HSGP model.
-Let's denote the matrix $X=\{\bm{x}_{it}\}$ which contains the predictors, where $\bm{x}_{it} \in {\rm I\!R}^{D}$, $D=53$ (52 principal components + time) is the $D$-vector of characteristic variables of the parcel $i$ and time $t$.
-The use of a first-order additive model is justified due to the input variables correspond to principal components which are linearly uncorrelated each other.
-%
-%The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{i}=\{p_{ik}\}_{k=1}^{K=5}$ of belonging to class $k$ per parcel $i$,
-%%
-%\begin{equation*}
-%y_i \sim \text{multinomial}(\bm{p}_i).
-%\end{equation*}
-%
-%\noindent The goal is to estimate the vector of probabilities $\bm{p}_i$ as a function of the input values, which is a multivariate function $f(\bm{x}_i):\bm{x_i}\in {\rm I\!R}^{D} \rightarrow \{f_k(\bm{x}_i)\}_{k=1}^K \in {\rm I\!R}^{K}$, related to the vector of probabilities $\bm{p}_i$ through the 'softmax' link function.
-%%
-%\begin{eqnarray*}
-%\bm{p}_i = \text{softmax}(f(\bm{x}_i)).
-%\end{eqnarray*}
-%
-%Each individual function $f_k(\bm{x})$ is modeled as a first-order additive model, 
-%%
-%\begin{eqnarray*}
-%f_k(\bm{x}) \sim \sum_{d=1}^{D} g_d(x^d)
-%%\mathcal{HSGP}(x^d, S(\bm{\lambda}), \theta_d),
-%\end{eqnarray*}
-%
-%\noindent where each individual component $g_d(x^d)$ is modeled as a unidimensional HSGP model,
-%%
-%\begin{eqnarray*}
-%g_d(x^d) \sim \mathcal{HSGP}(x^d, S, \theta_d),
-%\end{eqnarray*}
-%
-%In the HSGP model, the function $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
-The vector of hyperparameters $\theta_d=\{\alpha_d,\ell_d\}$ contains the marginal variance $\alpha_d$ and lengthscale $\ell_d$ of the HSGP process.
-%
-%For each individual component $g_d(x^d)$ modeled as a HSGP model, $m$=15 basis functions and a boundary factor $c$=2.5 were used.
-All the input variables were previously standarized with the exception of the time variable.
-%
-%After the adjustment of the model, the lengthscales estimates $\{\hat{\ell}_d\}_{d=1}^{D=14}$ normalized by half of the input range ($\frac{2\cdot \hat{\ell}_d}{|x^d_{max}-x^d_{min}|}$) are all bigger than the minimun lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as function of $m$, $c$.
-Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling the data.
-%
-%Table \ref{tab1} shows the confusion classification matrix after fit the model following a $Q$-fold cross-validation procedure over the training data. Every fold was composed of $Q$=10 observations.
-%
-%\begin{table}
-%\begin{center}
-%\begin{tabular}{|c|*{6}{c|}}\hline
-%\backslashbox{Estimate}{True} & \multicolumn{1}{p{1.5cm}|}{\centering caind \\(k=1)} &  \multicolumn{1}{p{1.5cm}|}{\centering calign \\(k=2)} & \multicolumn{1}{p{1.5cm}|}{\centering irr \\(k=3)} & \multicolumn{1}{p{1.5cm}|}{\centering aband \\(k=4)} & \multicolumn{1}{p{1.5cm}|}{\centering young \\(k=5)} & \multicolumn{1}{p{1.5cm}|}{\centering } \\
-%\hline \multicolumn{1}{|p{3cm}|}{ \centering caind (k=1)} &90&39&14&3&11&42\%\\
-%\hline \multicolumn{1}{|p{3cm}|}{\centering calign (k=2)} &46&301&8&2&3&16\%\\
-%\hline \multicolumn{1}{|p{3cm}|}{\centering irr (k=3)}  &8&4&59&4&1&19\%\\
-%\hline \multicolumn{1}{|p{3cm}|}{\centering aband (k=4)}  &5&2&6&342&5&5\%\\
-%\hline \multicolumn{1}{|p{3cm}|}{\centering young (k=5)}  &8&2&1&0&38&19\%\\
-%\hline \multicolumn{1}{|p{3cm}|}{\centering }  &42\%&13\%&32\%&2\%&34\%&$\textbf{17\%}$\\
-%\hline
-%\end{tabular}
-%\end{center}
-%\caption{Confusion matrix after the Q-fold cross-validation procedure over the training data.}
-%  \label{tab1}
-%\end{table}
-
-\subsection{Study case VII: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
+\subsection{Study case VII: Land use spatio-temporal classification task}\label{sec_bf_caseVII}
 The next example presents an spatio-temporal classification task in land-use of parcels between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits.
@@
 %
-\begin{eqnarray} \label{eq1_sec54}
+\begin{eqnarray} \label{eq_f_parcels}
 f_k(\bm{x}) = \sum_{d=1}^{D} g_d(x^d) + \sum_{d=1}^{D-1} g_{d+D}(x^d,x^D),
@@
 
-\noindent where $x^D$ represents the time variable. The first-order components $g_d(x^d)$, for $d=1,\dots,D$, in Equation (\ref{eq1_sec54}), are modeled as unidimensional HSGP models
+\noindent where $x^D$ represents the time variable. The first-order components $g_d(x^d)$, for $d=1,\dots,D$, in Equation (\ref{eq_f_parcels}), are modeled as unidimensional HSGP models
 %
@@
 
-\noindent In the HSGP model, the first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
-
-The second-order components $g_{d+D}(x^d,x^D)$, for $d=1,\dots,D-1$, in Equation (\ref{eq1_sec54}), are modeled as two-dimensional HSGP models
+\noindent In the HSGP model, the first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in Equation (\ref{eq_approxf}), with the square exponential spectral density $S$ as in Equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+
+The second-order components $g_{d+D}(x^d,x^D)$, for $d=1,\dots,D-1$, in Equation (\ref{eq_f_parcels}), are modeled as two-dimensional HSGP models
 %
@@
 
-\noindent In the HSGP model, the second-order components $g_{d+D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in Equation (\ref{approxf_multi}), with the two-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in Equation (\ref{eq:specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
+\noindent In the HSGP model, the second-order components $g_{d+D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in Equation (\ref{eq_approxf_multi}), with the two-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in Equation (\ref{eq_specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in Equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.
 
@@
 
-Table \ref{tab1} shows the confusion classification matrix after fitting the model following a $Q$-fold cross-validation procedure over the training data.
+Table \ref{tab_parcels} shows the confusion classification matrix after fitting the model following a $Q$-fold cross-validation procedure over the training data.
 Every fold was composed of $Q$=10 observations.
@@
 \caption{Confusion matrix after the Q-fold cross-validation procedure over the training data.}
-  \label{tab1}
+  \label{tab_parcels}
 \end{table}
@@
 
-\section{Related work}
+\section{Approximate the covariance function using Hilbert space methods} \label{sec_details_formulas_approx_covfun}
+
+In this section we briefly present the mathematical details of the approximation of a stationary covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
+This statement is basically an extract of the work \cite{solin2018hilbert}, where the authors fully develop the mathematical theory behind the Hilbert Space approximation for Gaussian processes.
+
+Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
+%
+\begin{equation*}
+\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
+\end{equation*}
+
+From the Bochner’s and Wiener-Khintchine theorem, the spectral density of a stationary covariance function $k(\bm{x},\bm{x}') = k(||\bm{r}||), \bm{r}=k(\bm{x}-\bm{x}')$, is the Fourier transform of the covariance operator $\mathcal{K}$
+
+\begin{eqnarray*}
+S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \cdot \bm{\tau}} = \mathcal{F}(\mathcal{K}) \nonumber
+\end{eqnarray*}
+
+\noindent where $\bm{w}$ is in the frequency domain. Thus, the spectral density $S(\bm{w})$ gives also the approximate eigenvalues of the operator $\mathcal{K}$.
+
+In the isotropic case $S(\bm{w}) = S(||\bm{w}||)$ and assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
+%
+\begin{equation}\label{eq_S}
+S(||\bm{w}||)=a_0+a_1||\bm{w}||^2+a_2(||\bm{w}||^2)^2+a_3(||\bm{w}||^2)^3+\cdots
+\end{equation}
+
+\noindent Recall that the Fourier transform of the Laplace operator $\nabla^2$ is $-||\bm{w}||$, thus the Fourier transform of $S(||\bm{w}||)$ is
+%
+\begin{equation}\label{eq_K}
+\mathcal{K}=a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+a_3(-\nabla^2)^3+\cdots,
+\end{equation}
+
+\noindent defining a pseudo-differential operator as a series of Laplace operators.
+
+If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the formal kernel $l$,
+%
+\begin{equation*}
+-\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}',
+\end{equation*}
+
+\noindent then the formal kernel can be represented as 
+%
+\begin{equation*}
+l(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{equation*}
+
+\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
+%
+\begin{eqnarray*}
+-\nabla^2 \phi_j(\bm{x})&=&\lambda \phi_j(\bm{x}), \hspace{1cm}  x\in \Omega \nonumber \\ 
+\phi_j(\bm{x})&=&0, \hspace{2cm} x\notin \Omega.
+\end{eqnarray*}
+
+\noindent Because $-\nabla^2$ is a positive definite Hermitian operator, the set of eigenfunctions $\phi_j(\cdot)$ are orthonormal with respect to the inner product
+%
+\begin{eqnarray*}
+<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) d(\bm{x})
+\end{eqnarray*}
+
+\noindent that is,
+%
+\begin{eqnarray*}
+\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) d(\bm{x}) = \delta_{ij},
+\end{eqnarray*}
+
+\noindent and all the eigenvalues $\lambda_j$ are real and positive.
+
+Due to normality of the basis of the representation of the formal kernel $l(\bm{x},\bm{x}')$, its formal powers $s=1,2,\dots$ can be write as
+%
+\begin{eqnarray}\label{eq_formalkernel}
+l(\bm{x},\bm{x}')^s= \sum_j \lambda_j^s \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{eqnarray}
+
+\noindent which are again to be interpreted to mean that
+%
+\begin{equation*}
+(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
+\end{equation*}
+
+\noindent This implies that we also have
+%
+\begin{equation*}
+[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}') d\bm{x}'.
+\end{equation*}
+
+Then, looking at Equations (\ref{eq_K}) and (\ref{eq_formalkernel}), it can be concluded 
+%
+\begin{eqnarray}\label{eq_k}
+k(\bm{x},\bm{x}')= \sum_j [a_0+a_1\lambda_j^1+a_2\lambda_j^2+\cdots] \phi_j(\bm{x}) \phi_j(\bm{x}').
+\end{eqnarray}
+
+\noindent By letting $||\bm{w}||^2=\lambda_j$ the spectral density in Equation (\ref{eq_S}) becomes
+%
+\begin{equation*}
+S(\sqrt{\lambda_j})=a_0+a_1\lambda_j+a_2\lambda_j^2+a_3\lambda_j^3+\cdots,
+\end{equation*}
+
+\noindent and substituting in Equation (\ref{eq_k}) then leads to the final searched approximation
+%
+\begin{eqnarray}\label{eq_k}
+k(\bm{x},\bm{x}')= \sum_j S(\sqrt{\lambda_j}) \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{eqnarray}
+
+\noindent where $S(\cdot)$ is the spectral density of the covariance function, $\lambda_j$ is the $j$th eigenvalue and $\phi_j(\cdot)$ the eigenfunction of the Laplace operator in a given domain.
+
+\section{Low-rank GP model with a periodic square exponential covariance function} \label{sec_cov_periodic}
+
+The periodic square exponential covariance function takes the form
+%
+\begin{eqnarray} \label{eq_cov_periodic}
+k(\bm{\tau})= \sigma^2 \exp\left(-\frac{2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)}{\ell^2}\right),
+\end{eqnarray}
+
+\noindent where $\sigma^2$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
+A GP model with a periodic covariance function does no fit under the framework of the HSGP approximation covered in this paper.
+However, it do has a state space representation.
+Next, we present the results from \cite{solin2014explicit}, where the authors obtain an approximate linear representation of a periodic covariance function based on expanding a periodic covariance functions into a series of stochastic resonators.
+
+In \cite{solin2014explicit}, the authors come to a cosine series expansion for the periodic covariance function (\ref{eq_cov_periodic}) as follows,
+% 
+\begin{eqnarray} \label{eq_cov_periodic_taylor_approx}
+k(\tau)= \sigma^2 \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 \tau),
+\end{eqnarray}
+
+\noindent which comes basically from a Taylor series representation of the periodic covariance function (\ref{eq_cov_periodic}). The coefficients $\tilde{q}_j^2$ of the previous expression are
+%
+\begin{eqnarray} \label{eq_q}
+\tilde{q}_j^2= \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
+\end{eqnarray}
+
+\noindent where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
+%
+\begin{eqnarray} \label{eq_q0}
+\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
+\end{eqnarray}
+
+\noindent Note that the covariance in Equation (\ref{eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
+As \cite{solin2014explicit} argue, this approximation (\ref{eq_cov_periodic_taylor_approx}) converge to Equation (\ref{eq_cov_periodic}) when $J \rightarrow \infty$.
+
+If the requirement of a valid covariance function is relaxed and only an optimal series approximation is required, taking the limit $J \rightarrow \infty$ in the sub-sums (\ref{eq_q}) gives the following variance coefficients
+%
+\begin{eqnarray} \label{eq_q_2}
+\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)}, 
+\end{eqnarray}
+
+\noindent for $j=1,2,\cdots,J$, and $\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)}$, where the $\text{I}_{\alpha}(z)$ is the modified Bessel function \citep{handbook1970m} of the first order of $\alpha$.
+Note that the terms in Equations (\ref{eq_q}) and (\ref{eq_q0}) are underneath bounded approximations for any $J$.
+
+In order to assess the accuracy of this representation as a function of the number of cosine terms $J$ considered in the approximation, an empirical evaluation is carried out in a similar way than that in Section \ref{sec_dependency} of this paper.
+Thus, Figure \ref{fig28_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic square exponential kernel as a function of the lengthscale of the kernel.
+We have considered an approximation to be a close enough in terms of satisfying Equation (\ref{eq_diff_covs}) with $\varepsilon=0.5\%$.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig28_m_lscale_periodic_appendix.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig28_m_lscale_periodic_appendix_zoom.pdf}}
+\caption{Relation among the minimum number of terms $J$ in the approximation and the lengthscale ($\ell$) of the periodic square exponential covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+  \label{fig28_m_lscale_periodic}
+\end{figure}
+
+The function values of a GP model with this state space representation of the periodic exponential covariance function can be easily derived: Considering the identity
+\begin{equation*}
+\text{cos}(j\omega_0 (x-x'))=\text{cos}(j\omega_0 x) \text{cos}(j\omega_0 x') + \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x'),
+\end{equation*}
+
+\noindent the covariance $k(\tau)$ (\ref{eq_cov_periodic_taylor_approx}) can be re-writting as
+%
+\begin{eqnarray} \label{eq_cov_periodic_taylor_approx_2}
+k(x,x')= \sigma^2 \left( \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 x)  \text{cos}(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x') \right).
+\end{eqnarray}
+
+\noindent where $\tau=x-x'$.
+With this approximation (\ref{eq_cov_periodic_taylor_approx_2}) for the periodic square exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \mathcal{GP}(0,k(x,x')$ equivalently leads to a linear representation of $f(\cdot)$ via
+%
+\begin{equation} \label{eq_f_period}
+f(x) \approx \sigma \left( \sum_{j=0}^J  \tilde{q}_j \text{cos}(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \text{sin}(j\omega_0 x) \beta_{J+1+j} \right),
+\end{equation}
+
+\noindent where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\cdots,2J+1$.
+The cosine $\text{cos}(j\omega_0 x)$ and sinus $\text{sin}(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.
+The only dependence on the hyperparameter $\ell$ is through the coefficients $\tilde{q}_j$, which are $J$-dimensional.
+The computational cost of this approximation scales as $O\big(n(2J+1) + (2J+1)\big)$, where $n$ is the number of observations and $J$ the number of cosine terms.
+
+The parameterization in (\ref{eq_f_period}) is naturally in the non-centered
+parameterization form with independent prior distribution on
+$\beta_j$, which makes the posterior inference easier.
+
+\section{ADDITIONAL MATERIAL -- NOT TO BE INCLUDED IN THE PAPER}
+
+\subsection{State space representation of a periodic and stationary covariance function} \label{sec_cov_periodic_otra}
+
+Let consider $k(x,x')$, with $x,x' \in {\rm I\!R}$, being a stationary and periodic covariance function
+%
+\begin{eqnarray}\label{eq_periodic_k}
+k(x,x')=k(x-x')=k(\tau).
+\end{eqnarray}
+
+\noindent Being $k(\cdot)$ also symmetric $k(\bm{\tau})=k(-\bm{\tau})$, it can be expanded into a convergence Fourier series
+%
+\begin{eqnarray}\label{eq_periodic_k_fourier}
+k(\tau)= \sum_{j=0}^{\infty} q_j^2\text{cos}(j\omega_0 \tau),
+\end{eqnarray}
+
+\noindent where $\omega_0$ is the angular frequency defining the periodicity.
+The question now remains how to determine the coefficients $q_j^2$ from a given periodic and stationary covariance function $k(\cdot)$ (\ref{eq_periodic_k}).
+One way to determine the coefficients is via projection to the cosine basis
+%
+\begin{eqnarray} \label{eq_exact_q}
+q_j^2= \frac{\omega_0}{\pi} \int_{\frac{-\pi}{\omega_0}}^{\frac{\pi}{\omega_0}} k(\tau)\text{cos}(j\omega_0 \tau)\text{d}\tau.
+\end{eqnarray}
+
+\noindent However, these coefficients can also be determined in a different way.
+Next, we present the results from \cite{solin2014explicit}, where the authors convert the square exponential periodic covariance function into a state space form.
+They lead to an expression for $k(\bm{\tau})$ equivalent to Equation (\ref{eq_periodic_k_fourier}), and approximations for the coefficients $q_j^2$.
+
+\paragraph{State space representation of a square exponential periodic covariance function}
+
+The square exponential periodic covariance function takes the form
+%
+\begin{eqnarray*}
+k(\bm{\tau})= \sigma^2 \exp\left(-\frac{2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)}{\ell^2}\right),
+\end{eqnarray*}
+
+\noindent where $\sigma^2$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
+
+Considering the identity $2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)=1-\text{cos}^j(\omega_0\bm{\tau})$, the above covariance function can be written as follows
+%
+\begin{eqnarray*}
+k(\tau)= \sigma^2 \exp\left(-\frac{1}{\ell^2}\right) \exp\left(1-\text{cos}(\omega_0\tau)\right),
+\end{eqnarray*}
+
+\noindent and taking into account the Taylor expansion of $\exp(1-\text{cos}(\omega_0\bm{\tau}))$:
+%
+\begin{eqnarray*}
+k(\tau)= \sigma^2 \exp\left(-\frac{1}{\ell^2}\right) \sum_{j=0}^{\infty} \frac{1}{j!} \text{cos}^j(\omega_0\tau).
+\end{eqnarray*}
+
+Now considering the series to be truncated at $J$, and recall that the powers of cosine can be expressed as sums of cosines with multiplied angles, and applying reductions formulas, lead to an expression of the form given by (\ref{eq_periodic_k_fourier}):
+%
+\begin{eqnarray} \label{eq_k_taylor}
+k(\tau)= \sigma^2 \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 \tau),
+\end{eqnarray}
+
+\noindent where the coefficients $\tilde{q}_j^2$ are
+%
+\begin{eqnarray} \label{eq_app_q}
+\tilde{q}_j^2= \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
+\end{eqnarray}
+
+\noindent where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
+%
+\begin{eqnarray*} 
+\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
+\end{eqnarray*}
+
+If the requirement of a valid covariance function is relaxed and only an optimal series approximation is required, taking the limit $J \rightarrow \infty$ in the sub-sums (\ref{eq_app_q}) gives the following variances coefficients
+%
+\begin{eqnarray} \label{eq_q_2_otra}
+\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)}, 
+\end{eqnarray}
+
+\noindent for $j=1,2,\cdots,J$, and $\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)}$, where the $\text{I}_{\alpha}(z)$ is the modified Bessel function \citep{handbook1970m} of the first order of $\alpha$.
+This is also the solution to the corresponding integral in (\ref{eq_exact_q}).
+Note that the terms in (\ref{eq_app_q}) are bounded from above such that $\tilde{q}^2_j < q^2_j$ for any $J$.
+
+Section 3.4 in \cite{solin2014explicit} gives a discussion about the error of the approximation in (\ref{eq_k_taylor}).
+However, in addition, we made an empirical evaluation of the accuracy of this approximation, in a similar way that was carried out in Section \ref{sec_dependency} of this paper.
+Thus, Figure \ref{fig28_m_lscale_periodic_otra} shows shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic square exponential kernel  as a function of the lengthscale of the kernel.
+We have considered an approximation to be a close enough in terms of satisfying Equation (\ref{eq_diff_covs}) with $\varepsilon=0.5\%$.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig28_m_lscale_periodic_appendix.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{fig28_m_lscale_periodic_appendix_zoom.pdf}}
+\caption{Relation among the minimum number of terms $J$ in the approximation and the lengthscale ($\ell$) of the periodic square exponential covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+  \label{fig28_m_lscale_periodic_otra}
+\end{figure}
+
+\paragraph{Low-rank GP model with a periodic square exponential covariance function}
+
+Considering the identity
+\begin{equation*}
+\text{cos}(j\omega_0 (x-x'))=\text{cos}(j\omega_0 x) \text{cos}(j\omega_0 x') + \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x'),
+\end{equation*}
+
+\noindent the covariance $k(\tau)$ (\ref{eq_k_taylor}), with $\tau=x-x'$, can be re-writting as
+%
+\begin{eqnarray} \label{eq_k_taylor_2}
+k(x,x')= \sigma^2 \left( \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 x)  \text{cos}(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x') \right).
+\end{eqnarray}
+
+With this approximation (\ref{eq_k_taylor_2}) for the periodic square exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \mathcal{GP}(0,k(x,x')$ equivalently leads to a linear representation of $f(\cdot)$ via
+%
+\begin{equation} \label{eq_f_approx_period}
+f(x) \approx \sigma \left( \sum_{j=0}^J  \tilde{q}_j \text{cos}(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \text{sin}(j\omega_0 x) \beta_{J+1+j} \right),
+\end{equation}
+
+\noindent where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\cdots,2J+1$.
+The cosine $\text{cos}(j\omega_0 x)$ and sinus $\text{sin}(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.
+The only dependence on the hyperparameter $\ell$ is through the coefficients $\tilde{q}_j$, which are $J$-dimensional.
+The computational cost of this approximation scales as $O\big(n(2J+1) + (2J+1)\big)$, where $n$ is the number of observations and $J$ the number of cosine terms.
+
+The parameterization in (\ref{eq_f_approx_period}) is naturally in the non-centered
+parameterization form with independent prior distribution on
+$\beta_j$, which makes the posterior inference easier.
+
+\subsection{Related work}
 
@@
 
-\subsection{Inducing points methods}
+\subsubsection{Inducing points methods}
 
@@
 
-\subsection{Basis function methods}
+\subsubsection{Basis function methods}
 
@@
 
-\section{Contributions of the method}
-
-This work is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models.
-This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods.
-This results in a reduced-rank approximation for the covariance function.
-This method has some nice features:
+\subsection{Brief details about computational demands and model inference}
+
+This paper focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
+This method has an attractive computational cost as this basically turns the regular GP model into a linear model.
 
 \vspace{2mm}
-$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.
-
-\vspace{2mm}
-$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
+\begin{itemize}
+	\item The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.
+
+	\item In learning the covariance function hyperparameters the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the optimizer. This demand includes:
+
+	\begin{itemize}
+		\item The linear model is computed with complexity $O(nm)$, computed in every step of the optimizer.
+
+		\item The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands.
+The weights depend on the hyperparameters, then they have to be computed in every step of the optimizer.
+	\end{itemize}
+
+	\item The computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?
+which is related to the computational complexity, an operation that must be computed in every step of the optimizer.
+
+	\item Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
 After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
-(Arno's paper, pag.
-7)
-
-\vspace{2mm}
-$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
-especially when sampling methods for inference are used.
+
+	\item The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions, especially when sampling methods for inference are used.
 GP prior as latent functions is needed in generalized models.
 
-In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
-
-\vspace{2mm}
-$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
-This operation has to be computed at every step of the HMC or optimizer.
-
-\vspace{2mm}
-$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
+	\item In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
+	\begin{itemize}
+		\item In regular GPs, the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
+These two operations have to be computed at every step of the HMC or optimizer.
+
+		\item In regular GPs, the parameter posterior distributions is $N$-dimensional.
 It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
 
-\vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
+		\item In conventional sparse GP approximations, based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
+
+		\item The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
 However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
 In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
-
-\section{Contributions of our work}
+	\end{itemize}
+\end{itemize}
+
+\subsection{Contributions of the work}
 
@@
 
-\section{Spectral densities of stationary covariance functions}
+\subsection{Spectral densities of stationary covariance functions}
 
@@
 
-\section{Approximate the covariance function using Hilbert space methods}
-
-Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
-%
-\begin{equation}
-\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
-\end{equation}
-
-Assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
-%
-\begin{equation}
-S(\bm{w})=a_0+a_1\bm{w}^2+a_2(\bm{w}^2)^2+a_1(\bm{w}^2)^3+\cdots
-\end{equation}
-
-If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the covariance function $k$,
-
-\begin{equation}
--\nabla^2 f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}',
-\end{equation}
-
-\noindent then the covariance function can be represented as
-
-\begin{equation}
-k(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
-\end{equation}
-
-\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
-Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
-%
-\begin{eqnarray}
--\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
-\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
-\end{eqnarray}
-
-a series expansion of eigenvalues and eigenfunctions
-
-\section{Example of generalization to the multivariate case}
+\subsection{Example of generalization to the multivariate case}
 
@@
 
-\section{Multidimensional generalization of covariance functions and spectral densities}
-
-\subsection{Square Exponential covariance function ($k$) \textbf and spectral density ($S$)}
+\subsection{Multidimensional generalization of covariance functions and spectral densities}
+
+\subsubsection{Square Exponential covariance function ($k$) \textbf and spectral density ($S$)}
 \vspace{-2mm}
-\subsubsection{Using norm-L2 (Euclidean distance)}
+\subsubsubsection{Using norm-L2 (Euclidean distance)}
 \vspace{-10mm}
@@
 \newpage
-\subsubsection{Using norm-L1 }
+\subsubsubsection{Using norm-L1 }
 \vspace{-10mm}
@@
 \newpage	
-\subsubsection{Using vector difference of inputs}
+\subsubsubsection{Using vector difference of inputs}
 \vspace{-5mm}
@@
 \newpage	
-\subsection{Matern($\nu=1/2$) covariance function ($k$) \textbf and spectral density ($S$)}
+\subsubsection{Matern($\nu=1/2$) covariance function ($k$) \textbf and spectral density ($S$)}
 \vspace{-2mm}
-\subsubsection{Using norm-L2 (Euclidean distance)}
+\subsubsubsection{Using norm-L2 (Euclidean distance)}
 \vspace{-8mm}
@@
 \newpage
-\subsubsection{Using norm-L1}
+\subsubsubsection{Using norm-L1}
 \vspace{-5mm}
@@
 \newpage
-\subsubsection{Using  the vector difference of inputs}
+\subsubsubsection{Using  the vector difference of inputs}
 \vspace{-5mm}
@@
 \newpage	
-\subsection{Matern($\nu=3/2$) covariance function ($k$) \textbf and spectral density ($S$)}
+\subsubsection{Matern($\nu=3/2$) covariance function ($k$) \textbf and spectral density ($S$)}
 \vspace{-2mm}
-\subsubsection{Using norm-L2 (Euclidean distance)}
+\subsubsubsection{Using norm-L2 (Euclidean distance)}
 \vspace{-8mm}

### 2020-03-14 Paper/manuscript_v2/manuscript_07_01.tex
update in manuscript_v2 and online appendice

--- 2019-04-29 Paper/manuscript/manuscript_05.tex
@@
-% interacttfssample.tex
-% v1.05 - August 2017
-\documentclass[]{interact}
+\documentclass[onecolumn,a4paper,11pt]{article}
+\voffset -1cm
+\hoffset -1cm
+\textheight 22.3cm
+\textwidth 15.0cm
+
+\usepackage{blindtext}
+\usepackage{titlesec}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{microtype}
+
 \usepackage{lscape}
 
-\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
-
-%\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
-
-%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
-
-%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
-%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
-%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled
-
-\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty  %
-
+\setcounter{secnumdepth}{5}
+\setcounter{tocdepth}{5}
+
+\usepackage{epstopdf}
+
+\usepackage[numbers,sort&compress]{natbib}
 \usepackage{subfigure}
@@
 
-\bibpunct[, ]{[}{]}{,}{}{,}{,}% Citation support using natbib.sty
-\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
-
-\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
-\newtheorem{notation}{Notation}
-
-\newenvironment{changemargin}[2]{%
-\begin{list}{}{%
-\setlength{\topsep}{0pt}%
-\setlength{\leftmargin}{#1}%
-\setlength{\rightmargin}{#2}%
-\setlength{\listparindent}{\parindent}%
-\setlength{\itemindent}{\parindent}%
-\setlength{\parsep}{\parskip}%
-}%
-\item[]}{\end{list}}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{mathtools}
+\usepackage{amsthm}
+\usepackage{bm}
+
+\usepackage[titletoc]{appendix}
+
+\providecommand{\keywords}[1]
+{
+  \small	
+  \textbf{\textit{Keywords---}} #1
+}
+
+\bibpunct[, ]{[}{]}{,}{}{,}{,}
+\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
+
+\usepackage{listings}
+\usepackage{slashbox}
+\usepackage{caption}
+\captionsetup{font=footnotesize}
+
+\bibliographystyle{authordate1}
+
+\title{Hilbert space approximate Gaussian process using Stan}
 
@@
 
-%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate
-
-\title{Hilbert space methods to approximate Gaussian processes using Stan}
-
-%\author{
-%\name{}
-%\affil{\textsuperscript{a}Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain; }
-%}
-
 \maketitle
@@
 \begin{abstract}
-
+Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large.
+In this paper, we focus on a novel approach for low-rank approximate Gaussian processes, based on the basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
+The main contribution of this paper is the realization of a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, desired prediction space and non-linearity of the function to be learned.
+Intuitive visualizations and useful recommendations for the values of these factors which help users to improve performance and computation are provided.
+A diagnosis procedure of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is also proposed.
+Furthermore, we show the simplicity of the method, with an attractive computational complexity due to its linear structure, which makes it easier to be used as modular components in probabilistic programming frameworks.
+Several illustrative examples of the performance and applicability of the method in the Stan programming software are presented, and their Stan model codes are also provided.
 \end{abstract}
 
-\begin{keywords}
-Gaussian processes; Low-rank Gaussian processes; Hilbert Space methods; Sparse Gaussian processes
-\end{keywords}
+\keywords{Gaussian process; Low-rank Gaussian process; Hilbert space methods; Sparse Gaussian process.}
 
@@
 
-%Contents
-\newpage
-
-\section{Introduction}\label{sec:bf_intro}
+\section{Introduction}\label{sec_bf_intro}
 
@@
 The covariance function encodes our prior assumptions about the functional relationship, such as continuity, smoothness, periodicity and scale properties.
-GPs not only allow for non-linear effects but can also implicitly handle interactions between covariates.
+GPs not only allow for non-linear effects but can also implicitly handle interactions between input variables (covariates).
 Different types of covariance functions can be combined for further increased flexibility.
@@
 
+The key element of a GP is the covariance function that defines the dependence structure between function values at different inputs.
+However, the covariance function is also a computational issue because of the need of inverting its Gram matrix to optimize the hyperparameters.
 Given $n$ observations in the data, the computational complexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
@@
 The basis functions are usually presented explicitly, but can also be used to form a low rank covariance matrix approximation.
-Common basis function approximations rest on the spectral analysis and series expansions of Gaussian processes \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
+Common basis function approximations rest on the spectral analysis and series expansions of GPs \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
 Sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving}.
-Recently, \cite{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels.
+Recently, \cite{hensman2017variational} presented a variational Fourier feature approximation for GPs that was derived for the Mat{\'e}rn class of kernels.
 Another related method for approximating kernels relies on random fourier features \citep{rahimi2008random,rahimi2009weighted}.
@@
 
-\textcolor{red}{Paul: Can we clarify how our discussed method relates to the former paragraph,
+{\color{blue} Paul: Can we clarify how our discussed method relates to the former paragraph,
 which introduces all sorts of basis function approximations. I.e. is our method
@@
 
-In this paper we focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
-Basis function approaches behave computationally like linear models, which is an attractive property in modular probabilistic programming models where there is a big benefit if approximation specific computation is simple (Paul: What do you mean with the the last part of the sentence?).
+Furthermore, a recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation (see, e.g.
+\cite{grenander1981abstract}) can be found in \cite{JSSv090i10}.
+
+In this paper we propose an approximate framework for fast and accurate inference for GPs.
+We focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
+Using basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.
+A linear model structure makes GPs easier to implement as building blocks in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if approximation specific computation is simple.
+Furthermore, a linear representation of a GP makes it easier to be used as latent function
+in non-Gaussian observational models allowing modelling flexibility.
+The basis function approximation via Laplace eigenfunctions can be made arbitrary accurate and the trade-off between computational complexity and approximation accuracy can easily be controlled.
+
 The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance kernel including the hyperparameters.
-While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost when only changing the covariance function parameters is $O(mn+m)$.% (or even $O(m^3)$ if marginalization can be used).
+While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost of learning the covariance function parameters is $O(mn+m)$ in every step of the optimizer.
 This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
-Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks (cite!).
+Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
 This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation.
@@
 
-\textcolor{red}{Aki: I moved this from section 2, needs to be combined with other intro}
-We propose an approximate framework for fast and accurate inference
-for Gaussian processes. Using a basis function expansion, we
-approximate the Gaussian process with a linear model. This
-representation has three main advantages: 1) it makes inference
-considerably faster due to the linear structure, 2) it is simple to
-implement, which makes it easy to use Gaussian processes as building
-blocks in more complicated models and can be used as latent function
-in non-Gaussian observational models allowing modelling flexibility,
-3) it can be made arbitrary accurate and the trade-off between
-computational complexity and approximation accuracy can easily be
-controlled.
-
-While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks such as Stan \citep{carpenter2017stan}.
+%\textcolor{red}{Aki: I moved this from section 2, needs to be combined with other intro}
+%We propose an approximate framework for fast and accurate inference
+%for GPs. Using a basis function expansion, we
+%approximate the GPs with a linear model. This
+%representation has three main advantages: 1) it makes inference
+%considerably faster due to the linear structure, 2) it is simple to
+%implement, which makes it easy to use GPs as building
+%blocks in more complicated models and can be used as latent function
+%in non-Gaussian observational models allowing modelling flexibility,
+%3) it can be made arbitrary accurate and the trade-off between
+%computational complexity and approximation accuracy can easily be
+%controlled.
+
+While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
+In this paper, the relations going on among the key factors of the method such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the performance and accuracy of the method.
+Practical recommendations for the values of the key factors, based on useful graphs that encode the recognized relationships, are given.
+Which will help users to chose valid and optimized values for these factors, improving computational performance and saving time of computation.
+A diagnosis of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is proposed.
+
+{\color{blue} (Gabi: This paragraph is the old version of the former paragraph) While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
 In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
-We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.
-
-Although there are several GP specific software packages available to date (GPML, GPstuff, GPy, GPflow, cite them!), each provide efficient implementations only for a restricted range of GP based models.
+We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.}
+
+The approach is implemented in a fully probabilistic framework and for the Stan programming probabilistic software.
+This work has served as a basis for the subsequent implementation of the method in the \textit{brms} library \citep{burkner2017brms} of the R software \citep{R2014R}.
+Several illustrative examples, simulated and real datasets, of the performance and applicability of the model, and accompanied by their Stan model codes, are developed.
+
+Although there are several GP specific software packages available to date (GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP based models.
 In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
@@
 The remainder of the paper is structured as follows.
-In Section 2, we introduce GPs and their reduced rank approximations proposed by \cite{solin2018hilbert}.
-In Section 3, we analyze the accuracy of these approximations under several conditions using analytical and numerical methods.
+In Section \ref{ch4_gp}, we introduce GPs with the covariance and spectral density functions.
+In Section \ref{ch5_sec_method}, the reduced rank approximation to GPs proposed by \cite{solin2018hilbert} is described.
+In Section \ref{ch5_sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
 Several case studies in which we fit
-exact and approximate GPs to real and simulated data are provided in Section 4. We end with a discussion in Section 5.
-
-\vspace{3mm}
-\section{Method}\label{sec:bf_method}
-
-\subsection{Gaussian process as a prior}
-
-A Gaussian process (GP) is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, i.e.
+exact and approximate GPs to real and simulated data are provided in Sections \ref{ch5_sec_univariate_cases} and \ref{ch5_sec_multivariate_cases}.
+We end with a discussion in Section \ref{ch5_sec_conclusion}.
+Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind this Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
+Additional content as online complementary material with more case studies illustrating the performance and applicability of the method can be found through the following link:
+
+\section{Gaussian process as a prior}\label{ch4_gp}
+
+A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, i.e.
 $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
-Gaussian processes have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
-
-In this work, Gaussian processes will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. 
-Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with Gaussian process prior.
+GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
+
+In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. 
+Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
 This generalizes trivially to more complex models depending on several unknown functions, for example such as $p(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
-Our goal is to obtain posterior distribution for the value of the function $f^*=f(\bm{x}^*)$  evaluated at a new input point $\bm{x}^* \in \mathbb{R}^D$.
+Our goal is to obtain posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
 % That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
 
-We assume a Gaussian process prior for $f \sim \mathcal{GP}(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: \mathbb{R}^D \rightarrow \mathbb{R}$ and $k: \mathbb{R}^D \times \mathbb{R}^D \rightarrow \mathbb{R}$ are the mean and covariance functions, respectively,
+We assume a GP prior for $f \sim \mathcal{GP}(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \rightarrow {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \rightarrow {\rm I\!R}$ are the mean and covariance functions, respectively,
 %
@@
 
-The mean and covariance functions completely characterize the Gaussian process prior, and control the a priori behavior of the function $f$.
+The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$.
 Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution
@@
 The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
-The joint distribution of $\bm{f}$ and a new $f^*$ is also a multivariate Gaussian as,
+The joint distribution of $\bm{f}$ and a new $\tilde{f}$ is also a multivariate Gaussian as,
 %
 \begin{align*}
-p(\bm{f}, f^*)=\mathcal{N} \left( \left[ \begin{array}{cc}
+p(\bm{f}, \tilde{f})=\mathcal{N} \left( \left[ \begin{array}{cc}
 \bm{f} \\ 
@@
 \end{array} \right] \,\middle|\, \bm{0},\left[ \begin{array}{cc}
-\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},f^*} \\ 
-\bm{k}_{f^*,\bm{f}} & k_{f^*,f^*}
+\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},\tilde{f}} \\ 
+\bm{k}_{\tilde{f},\bm{f}} & k_{\tilde{f},\tilde{f}}
 \end{array} \right] \right),
@@
 
-\noindent where $\bm{k}_{\bm{f},f^*}$ is the covariance between $\bm{f}$ and $f^*$, and $k_{f^*,f^*}$ is the prior variance of $f^*$.
+\noindent where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$.
 %By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\bm{f}$ analytically,
@@
 If $p(y_n|f(\bm{x}_n),\phi)=N(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
-%
-Furthermore, if a prior distribution is imposed on $\phi$ and $\theta$ to form a joint posterior for $\phi$, $\theta$ and $\bm{f}$,
-approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011} Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} need to be used.
-%
+Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011}, Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} need to be used.
 In this paper we focus on use of MCMC for integrating over the joint
@@
 
-\subsection{Covariance function and spectral density}
-
-The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
+\subsection{Covariance function and spectral density}\label{ch4_sec_cov}
+
+The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
+A covariance function needs be symmetric and positive semi-definite \citep{rasmussen2006gaussian}.
 A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
@@
 The Mat\'ern class of isotropic covariance functions is given by, 
-%which, in the unidimensional case (i.e. $\bm{x}=x\in \mathbb{R}$, so $\bm{r}=r\in \mathbb{R}$, and $\ell\in \mathbb{R}$ ), are
 %
 \begin{align*}
-k_{\nu}(r)&=\frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\nu}K_{\nu}\left(\frac{\sqrt{2\nu}r}{\ell}\right),
+k_{\nu}(r)&=\sigma^2 \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\nu}K_{\nu}\left(\frac{\sqrt{2\nu}r}{\ell}\right),
 \end{align*}
 
-\noindent where $\nu$ is the order the kernel, $K_{\nu}$ the modified Bessel function, and the $\ell$ and $\sigma$ are the length-scale and magnitude, respectively, of the kernel.
+\noindent where $\nu$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell$ and $\sigma$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
 The particular case where $\nu=\infty$ and $\nu=3/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
@@
 \noindent The former is commonly known as squared exponential (exponentiated quadratic) covariance function.
-Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form
+Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form:
 %
 \begin{align*}
-k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\exp\bigg(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\bigg),\\
-k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\Bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg)\exp\Bigg(- \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg). \\
+k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\sigma^2 \exp\bigg(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\bigg),\\
+k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\sigma^2 \Bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg)\exp\Bigg(- \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg). \\
 \end{align*}
@@
 \begin{align*}
-S_{\nu}(\omega)&= \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
+S_{\nu}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
 \end{align*}
 
-\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ is a distance in the frequency domain, and $\ell$ and $\sigma$ are the lengthscale and magnitud, respectively, of the kernel.
-The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form
+\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ is a distance in the frequency domain, and $\ell$ and $\sigma$ are the lengthscale and magnitud (marginal variance), respectively, of the kernel.
+The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form:
 %
 \begin{align}
-S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq:specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq:specdens_32} 
+S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
 
-\noindent Particularizing to an input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{D=3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{D=3}$, the spectral densities written above take the form
+\noindent Particularizing to an input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form:
 %
 \begin{align*}
-S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^{D=3}  \prod_{i=1}^{D=3} \ell_i  \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D=3} \ell_i^2 s_i^2 \right),   \\
-S_{\frac{3}{2}}(\omega)&= \sigma^2 32\pi\sqrt{3}^3\prod_{i=1}^{D=3}\ell_i\left(3+\sum_{i=1}^{D=3}\ell_i^2 s_i^2 \right)^{-3}.
+S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
+S_{\frac{3}{2}}(\omega)&= \sigma^2 32\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
 \end{align*}
 
-\subsection{Hilbert space approximate Gaussian process model}
-
-The approximate Gaussian process method, developed by \cite{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
+\section{Hilbert space approximate Gaussian process model}\label{ch5_sec_method}
+
+The approximate GP method, developed by \cite{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
 Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to some boundary condition.
@@
 More details and mathematical proofs are provided in \cite{solin2018hilbert}.
+Our starting point for presenting the method is the main result obtained by \cite{solin2018hilbert} of the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
+The mathematical details of this approximation have been briefly presented in the Appendix \ref{ch5_app_approx_covfun} of this paper.
 
@@
 As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs.
-We will come back to this issue in Section 3.
+We will come back to this issue in Section \ref{ch5_sec_accuracy}.
 
@@
 
-\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section 2.1) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
-The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
+The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
 Namely, they satisfy the following eigenvalue problem in $\Omega$ when applying the Dirichlet boundary condition (other boundary conditions could be used as well):
 %
-\begin{eqnarray}\label{eigenproblem}
+\begin{align}\label{eq_eigenproblem}
 \begin{split}
--\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x \in \Omega \\ 
-\phi_j(x)&=&0, \hspace{2cm} x \notin \Omega.
+-\nabla^2 \phi_j(x)&=\lambda \phi_j(x), \hspace{1cm}  x \in \Omega \\ 
+\phi_j(x)&= 0, \hspace{1.85cm}   x \notin \Omega.
 \end{split}
-\end{eqnarray}
-
-\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in Equation (\ref{eigenproblem}) are sinusoidal functions.
+\end{align}
+
+\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in equation (\ref{eq_eigenproblem}) are sinusoidal functions.
 Independently of the covariance function, they can be computed as
@@
 A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the covariance hyperparameters $\theta$.
-Instead, the only dependence on $\theta$ is through the spectral density $S_{\theta}$.
+Instead, the only dependence of the model on $\theta$ is through the spectral density $S_{\theta}$.
 The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions.
-Therefore, Equation (\ref{approxf}) can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
+Therefore, equation (\ref{eq_approxf}) can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
 The computational cost of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
 
-The parameterization in (\ref{approxf}) is naturally in the non-centered
-parameterization form with independent prior distribution on
-$\beta_j$, which makes the posterior inference easier.
-
-\subsection{Generalization to multidimensional GPs}
+The parameterization in equation (\ref{eq_approxf}) is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which makes the posterior inference easier.
+Furthermore, all dependencies on the covariance kernel and the hyperparameters is through the prior distribution of the regression weights $\beta_j$.
+The parameter posterior distribution $p(\bm{\beta}|\bm{y})$ is $m$-dimensional, where $m$ is much smaller than the number of observations $n$.
+Therefore, the parameter space is greatly reduced and this makes inference faster, especially when sampling methods are used.
+
+\subsection{Generalization to multidimensional GPs} \label{ch5_sec_method_multi}
 
@@
 2 & 2 & 3 
-\end{array} } \right]
+\end{array} } \right].
 \end{align}
@@
 Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
-Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have
+Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have:
 %
 \begin{align}
-\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right) \label{eigenvalue_multi} \\
-%
-\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D. \label{eigenfunction_multi}
+\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D, \label{eq_eigenvalue_multi} \\
+%
+\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right).
+\label{eq_eigenfunction_multi}
 \end{align}
@@
 
-\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section 2.1).
-We can now write the approximate series expansion of the multivariate function $f$ as,
-%
-\begin{equation}\label{approxf_multi}
+\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}).
+We can now write the approximate series expansion of the multivariate function $f$ as
+%
+\begin{equation}\label{eq_approxf_multi}
 f(\bm{x}) \approx \sum_{j=1}^{m^{\ast}} 
@@
 \noindent where, again, $\beta_j \sim \text{Normal}(0,1)$.
-The computational cost of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
+The computational cost, in learning the covariance function hyperparameters, of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
 Although this still implies linear scaling in $n$, the approximation is more costly than in the univariate case, as $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and grows exponentially with respect to the number of  dimensions.
 
-\subsection{Learning hyperparameters and model inference}
-
-\vspace{2mm}
-$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a linear model.
-
-\vspace{2mm}
-$\bullet$ The design matrix of the proposed linear model, which is composed of a basis of Laplace eigenfunctions, can be computed analytically and does not depend on the hyperparameters of the model, then it has to be computed only once with $O(n+m)$ computational demands.
-
-\vspace{2mm}
-$\bullet$ The weights associated to the basis functions in this linear model is a $m$-dimensional vector ($m$ is the number of basis functions) and their computation is an operation with $O(m)$ computational demands.
-The weights depend on the hyperparameters, then they have to be computed in every step of the HMC sampling method.
-
-\vspace{2mm}
-$\bullet$ The linear model is computed with complexity $O(nm)$, computed in every step of the HMC sampling method.
-
-\vspace{2mm}
-$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
-After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
-(Arno's paper, pag.
-7)
-
-\vspace{2mm}
-$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
-especially when sampling methods for inference are used.
-GP prior as latent functions is needed in generalized models.
-
-In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
-
-\vspace{2mm}
-$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
-This operation has to be computed at every step of the HMC or optimizer.
-
-\vspace{2mm}
-$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
-It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
-
-\vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
-In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
-
-\vspace{2mm}
-$\bullet$ In addition, the computation of the automatic differentiation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ In a regular GP model the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$.
-This operation has to be computed at every step of the HMC.
-
-\vspace{2mm}
-$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-
-\section{The accuracy of the approximation}
+\section{The accuracy of the approximation}\label{ch5_sec_accuracy}
 
@@
 
-(Gabi: The following paragraph is a first idea to introduce that the recommendations and diagnostics depends on the kernel considered.
-And that, at the moment, we have built these recommendations only for the square exponential kernel.
-I know that what I wrote in not enough and should be a bit more elaborate; your suggestions where and how elaborate it would be useful.)
-
 Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
-In this work we built these relationships for the square exponential covariance function.
-For other kernels, the relationships will be slightly different.
-
-\subsection{Dependency on the number of basis functions and the boundary condition}
-
-As explained in Section 2, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset \mathbb{R}$:
-%
-\begin{equation}\label{diffcov}
-k(\tau) = \sum_{j=1}^{\infty} S_{\theta} \left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0),  \nonumber
-\end{equation}
+In this work we built these relationships for the square exponential covariance function and Matern ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix \ref{ch5_sec_periodic}.
+For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects.
+
+\subsection{Dependency on the number of basis functions and the boundary condition} \label{ch5_subsec_dependency}
+
+As explained in Section \ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$:
+%
+\begin{equation*}
+k(\tau) = \sum_{j=1}^{\infty} S_{\theta} \left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0), 
+\end{equation*}
 
 \noindent where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}) for the unidimensional case and in Equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}) for the multidimensional case.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
 The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
 %
-\begin{eqnarray}\label{diff_covs}
+\begin{eqnarray}\label{eq_diff_covs}
 \int k(\tau) \, d\tau - 
@@
 
-The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
-The approximation also depends on the boundary $L$ (see Equations (\ref{eigenvalue}), (\ref{eigenfunction}), (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
+The finite number $m$ of basis functions in the approximation needed to satisfy equation (\ref{eq_diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
 As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
 In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi$.
-Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \mathbb{R}$.
+Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \Omega$.
 We now define $L$ as
 %
-\begin{equation}\label{eq:boundary}
+\begin{equation}\label{eq_boundary}
 L=c \cdot S,
@@
 \noindent where $S$ (for $S > 0$) represents the half-range of the input space, and $c$ (for $c \geq 1$) is the proportional extension factor.
-In the following, we will to refer to $c$ as the boundary factor of the approximation.
+In the following, we will refer to $c$ as the boundary factor of the approximation.
 The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
@@
 We start with an illustration on how the number of basis functions $m$ and boundary factor $c$ influences the accuracy of the HSGP approximations, separately.
-For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
+For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ of the kernel function, using input values from the zero-mean input domain with half-range $S=1$.
 Several HSGP models with varying $m$ and $L$ are fitted to this data.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig1_Cov_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig1_legend.pdf}}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 122mm 10mm, clip]{ch5_fig1_legend.pdf}}
+\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.
+Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
   \label{fig1_Post_J}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig2_Cov_L.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig2_legend.pdf}}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 120mm 0mm, clip]{ch5_fig2_legend.pdf}}
+\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.
+Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
   \label{fig2_Post_L}
@@
 Next, we will focus on analyzing the interaction effects between these $m$ and $c$ on the performance of the approximation.
-The lengthscale and marginal variance will no longer be fixed but rather estimated in both regular GP and HSGP models.
+The lengthscale and marginal variance of the kernel will no longer be fixed but rather estimated in both regular GP and HSGP models.
 Figure \ref{fig3_Post_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$.
@@
 \begin{center}
-\begin{tabular}{ c c c}
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
-\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
-\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
+\begin{tabular}{| c | c | c | c |}
+\hline
+c  = 1.05 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & \multirow{32}{1.3cm}{\includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\
+\cline{1-3}
+c  = 1.1 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
+\cline{1-3}
+c  = 1.2 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
+\cline{1-3}
+c  = 1.5 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & \\
+\cline{1-3}
+c  = 2 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} &  \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
+\cline{1-3}
+c  = 2.5 & \includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf} & \includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
+\hline
 \end{tabular}
 \end{center}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
+\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
+Notice that the dashed red line of the regular GP can hardly be seen in the some of the plots because it is under the blue lines.}
   \label{fig3_Post_part1}
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig4_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig4_MSE_vs_c.pdf}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
@@
 \centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig5_legend.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig5_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig5_magnitud_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 28mm 119mm 0mm, clip]{ch5_fig5_legend.pdf}}
 \caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
@@
 Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figure \ref{fig6_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to
-% the performance of the HSGP approximation.
-a close approximation of the HSGP model, in the case of a GP with square exponential covariance function and a single input dimension.
-(Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?).
-More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}).
-We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern ($\nu$=3/2) covariance function, respectively, and a single input dimension.
+More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying equation (\ref{eq_diff_covs}).
+Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern ($\nu$=3/2) covariance function.
+We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function:
 %
@@
 If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
-If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased We may also use this diagnostic in a iterative procedure.
+If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
+We may also use this diagnostic in a iterative procedure.
 Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
 Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
-%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c.pdf}}
 \hspace{3mm}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
@@
 
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_Matern.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\nu$=3/2) covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+  \label{fig6_lscale_vs_J_vs_c_Matern}
+\end{figure}
+
 If we look back to the conclusions drawn from Figures \ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig6_lscale_vs_J_vs_c}.
 
-(Gabi: The following paragraph is just an idea to extend the discussion to the multidimensional case)
-
-Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending on three variables, $m$, $c$ and $\ell$.
-An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented.
-However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
-
-(Gabi: The following paragraph is another idea to introduce that the diagnosis plot in Figure \ref{fig6_lscale_vs_J_vs_c} depends on the kernel considered.
-At the moment we have built this figure only for the square exponential kernel.)
-
-Ultimately, Figure \ref{fig6_lscale_vs_J_vs_c} showing the relationships among the number of basis functions, the boundary factor and the lengthscale of the function depends on the particular choice of covariance function.
-In this work we have built these relationships only for the square exponential covariance function.
-For other kernels, the Figure will be slightly different.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
+An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be graphically represented.
+More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wiggliness in every dimensions.
+For instance, in the two-dimensional case it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
+Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
+However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} to check the accuracy by analyze individually the different dimensions of a multidimensional GP model.
 
@@
 obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
-
-%Another way to look at the approximation accuracy of HSGPs is to compare the lengthscale estimates
-%obtained from the regular GP and HSGP models. As the lenghscale $\ell$ controls the "wigglyness" of
-%the functional relationship, it strongly influences the difficulty of obtaining accurate
-%inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
-%we can expect the HSGP approximation to be accurate as well.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
-\vspace{-2mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
+\subfigure{\includegraphics[scale=0.4, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig7_legend.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
+\subfigure{\includegraphics[scale=0.30, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part3.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
+\subfigure{\includegraphics[scale=0.30, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part6.pdf}}\\
 \vspace{-3mm}
-\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
+\subfigure{\includegraphics[scale=0.30, trim = 1mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
+\vspace{-3mm}
+\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).
+Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
   \label{fig7_posterior_varing_lscale_part1}
@@
 \begin{flushleft}
-\begin{tabular}{ c c c c}
-\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 45mm, clip]{fig8_legend.pdf}}\\
-\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
+\begin{tabular}{ccc}
+\multicolumn{3}{c}{ \includegraphics[scale=0.40, trim = 25mm 117mm 90mm 45mm, clip]{ch5_fig8_legend.pdf}}\\
+\includegraphics[scale=0.30, trim = 0mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.30, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part3.pdf} \\
+\includegraphics[scale=0.30, trim = 0mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part6.pdf}
 \end{tabular}
 \end{flushleft}
-\caption{Data-generating functional lengthscales (X-axis), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates from the regular GP and HSGP models (Y-axis).
+\caption{Data-generating functional lengthscales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates $(\hat{\ell})$ from the regular GP and HSGP models.
 95\% confident intervals of the lengthscale estimates are plotted as dot lines.
@@
 \centering
-\includegraphics[scale=0.35]{fig9_MSE_varing_lscale.pdf}
-\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects (lengthscales).}
+\includegraphics[scale=0.38]{ch5_fig9_MSE_varing_lscale.pdf}
+\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects ($\ell$).}
   \label{fig9_MSE_varing_lscale}
@@
 
-(Paul: Figure 8 and 9 are not rendered correctly in my PDF viewer after 
-compiling the tex file, which has something to do with the font of the facet 
-and legend labels I believe. Latex also issues warnings about this.)
-
-% Comparing the information of models approximation provided by Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale}, it can be appreciated an agreement between both figures, excepting for small lengthscales and long lengthscales.
-We observe small RMSE for large lengthscales in Figure \ref{fig9_MSE_varing_lscale}, however there exist mismatch between large lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
-This is because big changes on large lengthscales, which might be clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior predictions and therefore a small RMSE.
-On the other side, we observe high RMSE for small lengthscales in Figure \ref{fig9_MSE_varing_lscale}, however there exist match between small lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}.
-This is because very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors predictions and therefore high RMSE.
-
-%The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig6_lscale_vs_J_vs_c}.
-It can be appreciated a complete agreement between those recommended minimum lengthscales that can be properly fitted in every case following Figure \ref{fig6_lscale_vs_J_vs_c} and the real approximate accuracy measured by the RMSE in Figure \ref{fig9_MSE_varing_lscale}.
-
-\section{Univariate examples}
-
-\subsection{Study case I: Simulated data}
-
-This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws of a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
+\section{Univariate case studies}\label{ch5_sec_univariate_cases}
+
+\subsection{Simulated data}\label{ch5_sec_univariate_simu}
+
+This example consists of a simulated dataset with $n=250$ ($i=1,\dots,n$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=(x_1\dots,x_n)$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
 To form the final noisy dataset $\bm{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
@@
 %
-\begin{eqnarray*}\label{eq:latentgp_simudata1}
+\begin{eqnarray*}\label{ch5_eq_latentgp_simudata1}
 \begin{split}
-\bm{y} &= f(\bm{x}) + \bm{\epsilon} \\
+\bm{y} &= \bm{f} + \bm{\epsilon} \\
 \bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
-f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x}, \bm{x}', \theta)),
+f(x) &\sim \mathcal{GP}(0, k(x, x', \theta)),
 \end{split}
@@
 
-\noindent where $I$ represents the identity matrix.
-The previous formulation corresponds to the latent form of a GP model, where $f(\bm{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$.
-Saying that the function values $f(\bm{x})$ follows a GP model is equivalent to say that $f(\bm{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.
+\noindent where $I$ represents the identity matrix and $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function values to the noisy data.
+The previous formulation corresponds to the latent form of a GP model.
+The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with a Mat{\'e}rn($\nu$=3/2) covariance function $k$.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
 
@@
 %
-\begin{equation*}\label{eq:marginalizedgp_simudata1}
-\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I )
-\end{equation*}
-
-\noindent where the function values $f(\bm{x})$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing time of computation and improving the sampling and the effective number of samples.
-
-In the HSGP model, the latent function values $f(\bm{x})$ are approximated as (\ref{approxf}),
-%
-\begin{equation*}
-f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\lambda_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
-\end{equation*}
-
-\noindent with the spectral density $S$ (\ref{eq:specdens_32}) as a function of $\sqrt{\lambda_j}$,
-%
-\begin{eqnarray*}
-S(\sqrt{\lambda_j})&=& \alpha^2 \frac{4\sqrt{3}^{3}}{\ell^3}(\frac{3}{\ell^2} + \lambda_j)^{-2}, 
-\end{eqnarray*}
-
-\noindent and eigenvalues $\lambda_j$ (\ref{eigenvalue}) and eigenfunctions $\phi_j$ (\ref{eigenfunction})
-%
-\begin{eqnarray*}
-\begin{split}
-\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \\
-\phi_j(\bm{x})&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(\bm{x}+L)\right).
-\end{split}
-\end{eqnarray*}
-
-\noindent In the previous equations, $L$ is the boundary and $m$ the number of basis functions.
-The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell$ are the marginal variance and lengthscale parameters, respectively, of the approximate covariance function.
-
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
-
-Figure \ref{fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$ (\ref{eq:boundary})), and the splines model with 80 knots.
-The true data-generative function and the noisy observations are also plotted in the Figure; the sample observations are plotted as circles and the out-of-sample or test data, which have not been taking part on training the models, are plotted as crosses.
-The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{fig10_Posteriors_exI.pdf}
-\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dot lines.}
-  \label{fig10_Posteriors_exI}
-\end{figure}
-
-In order to assess performance of the models as a function the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
-Figure \ref{fig11_MSE_exI_inter} shows the standardized root mean squared error (SRMSE), for interpolation and extrapolating data, as a function of the number of basis functions and knots.
-The SRMSE is computed against the data-generating function.
-From Figures \ref{fig10_Posteriors_exI} and \ref{fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model, for interpolating and extrapolating data.
-However, the splines model does not extrapolate data properly.
-Both models show roughly similar interpolating performance.
-
-Figure \ref{fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the splines model.
-The computational time is represented in the Y-axis of the figure, which is in a logarithmic scale.
-The HSGP model is on average roughly 400 times faster than the regular GP, in the particular case of applying over this dataset.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_MSE_exI_extra.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 60mm 120mm 0mm, clip]{fig11_MSE_exI_legend.pdf}}
-\caption{Standarized root mean square error (SRMSE) of the different methods against the data-generating function.
-(left) SRMSE for interpolation.
-(right) SRMSE for extrapolation.
-The Standard deviation of the mean of the SRMSE are plotted as dot lines.}
-  \label{fig11_MSE_exI_inter}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig11_time_exI.pdf}}
-\caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
-  \label{fig11_time_exI}
-\end{figure}
-
-\subsection{Study case II: Gay data}
-This data set relates the proportion of support for same-sex marriage to the age.
-The data consists of 74 observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group $i$ ($i=1,\cdots,74$).
-The observational model is a binomial model with parameters population $n_i$ and probability of supporting same-sex marriage $p_i$ per age group $i$,
-%
-\begin{equation*}
-y_i \sim \text{Binomial}(p_i, n_i).
-\end{equation*}
-
-\noindent The population per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ or mean number of support people per age group.
-The probabilities $\bm{p}=\{p_i\}_{i=1}^{74}$ are modeled by a GP function $f(\bm{x})$ with a squared exponential covariance function $k$ and age input values $\bm{x}=\{x_i\}_{i=1}^{74}$, with the {\it logit} function as a link function,
-%
-\begin{eqnarray*} \label{eq:gpprior_gay}
-p_i &=& \text{logit}(f(x_i)) \nonumber \\
-f(\bm{x}) &\sim& \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta).
-\end{eqnarray*}
-
-In the HSGP model the function $f(\bm{x})$ is approximated as in Equation (\ref{approxf}), with the square exponential spectral density as in Equation (\ref{eq:specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in Equations (\ref{eigenvalue}) and (\ref{eigenfunction}).
-
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.
-
-Figure \ref{fig12_Posteriors_gaydata} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the splines model with 20 knots.
-Sample observations are plotted as circles in the figure, and the out-of-sample observations, which have been used for testing, are plotted as crosses.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
-\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
-  \label{fig12_Posteriors_gaydata}
-\end{figure}
-
-For the HSGP model, different models with different number of basis functions and boundary factor have been fitted.
-The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model and plotted as a function of the number of basis functions $m$ and the boundary factor $c$ in Figure \ref{fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
-The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_train_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig13_MSE_pred_BF_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 50mm 120mm 0mm, clip]{fig13_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the HSGP model, computed against the regular GP model, as a function of the number of basis functions $m$ and boundary factor $c$.
-(left) RMSE for sample data.
-(right) RMSE for test data.}
-  \label{fig13_MSE_train_BF_gaydata}
-\end{figure}
-
-Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models, computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model.
-We can see how the splines models does not extrapolate data properly.
-Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration, as a function of the number of basis functions $m$ for the HSGP model and knots for the splines model.
-The computational times is represented in the Y-axis which is in a logarithmic scale.
-The HSGP model is on average roughly 10 times faster than the regular GP, in this particular case.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_train_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_MSE_pred_gaydata.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 42mm 120mm 0mm, clip]{fig14_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the different methods, regular GP, HSGP and splines models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots for the splines model.
-(left) RMSE for sample data.
-(right) RMSE for test data.}
-  \label{fig14_MSE_train_gaydata}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig14_1_time_gaydata.pdf}}
-\caption{Computational time (Y-axis), in seconds per iteration, as a function of the number of basis functions $m$ and knots. The Y-axis is in a logarithmic scale.}
-  \label{fig14_1_time_gaydata}
-\end{figure}
-
-\subsection{Study case III: Birthday data}
-
-\vspace{5mm}
-\section{Multivariate examples}
-\subsection{Study case IV: Simulated data}
-
-This example consists of a simulated dataset with 120 ($i=1,\dots,n=120$) single draws from a Gaussian process prior with two ($D=2$) input dimensions.
-A square exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$ for dimension 1 and $\ell_2=0.35$ for dimension 2, is used.
-The corresponding matrix of input values is $X=\{\bm{x}_i\}_{i=1}^{n=120} \in {\rm I\!R}^{n\times 2}$ with $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$.
-Gaussian noise $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^n$.
-
-The regular GP model over the outcome variable $\bm{y}$ and $n \times 2$ matrix of inputs $X$, can be written as follows,
-%
-\begin{eqnarray*}\label{eq:latentgp_simudata2}
-\begin{split}
-\bm{y} &= f(X) + \bm{\epsilon} \\
-\bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
-f(X) &\sim \mathcal{GP}(0, k(X, X', \theta)),
-\end{split}
-\end{eqnarray*}
-
-\noindent where $I$ represents the identity matrix and $k$ the square exponential covariance function with hyperarameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
-The marginalized form, by integrating out the latent values $f(X)$, of the previous latent GP model results
-%
-\begin{equation*}\label{eq:marginalizedgp_simudata2}
+\begin{equation*}\label{ch5_eq_marginalizedgp_simudata1}
 \bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
@@
 
-\noindent with covariance matrix $K$, and its element $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$ is the evaluation of the covariance function $k$ at points $r$ and $s$.
-
-In the HSGP model with $D$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as (\ref{approxf_multi}),
+\noindent where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
+
+In the HSGP model, the latent function values $f(x)$ are approximated as in equation (\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in equation (\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
+
+In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
+
+Figure \ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
+The true data-generative function and the noisy observations are also plotted.
+The sample observations are plotted as circles and the out-of-sample or test data, which have not been taking part on training the models, are plotted as crosses.
+The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
+The posteriors of the three models, regular GP, HSGP and spline, are pretty similar within the interpolation input space.
+However, when extrapolating the spline model solution clearly differs from the regular GP and HSGP models as well as the actual observations.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{ch5_fig10_Posteriors_exI.pdf}
+\caption{Posterior predictive means of the proposed HSGP model, the regular GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
+  \label{ch5_fig10_Posteriors_exI}
+\end{figure}
+
+In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted.
+Figure \ref{ch5_fig11_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots.
+The SRMSE is computed against the data-generating model.
+From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model for interpolating and extrapolating data.
+However, the spline model does not extrapolate data properly.
+Both models show roughly similar interpolating performance.
+
+Figure \ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+The computational time is represented in the y-axis of the figure, which is in a logarithmic scale.
+The HSGP model is on average roughly 400 times faster than the regular GP, in the particular case of applying over this dataset.
+
+The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found through the following link:
+%
+\begin{lstlisting}[breaklines]
+https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-I_1D-Simulated-data
+\end{lstlisting}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig11_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig11_MSE_exI_extra.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 29mm 55mm 110mm 0mm, clip]{ch5_fig11_MSE_exI_legend.pdf}}
+\caption{Standardized root mean square error (SRMSE) of the different methods against the data-generating function.
+SRMSE for interpolation (left) and SRMSE for extrapolation (right).
+The standard deviation of the mean of the SRMSE is plotted as dashed lines.}
+  \label{ch5_fig11_MSE_exI_inter}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig11_time_exI.pdf}}
+\caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+The y-axis is on a logarithmic scale.
+The standard deviation of the computational time is plotted as dashed lines.}
+  \label{ch5_fig11_time_exI}
+\end{figure}
+
+\subsection{Birthday data}\label{ch5_sec_birthday}
+This example is an analysis of patterns in birthday frequencies in a dataset containing records of all births in the United States on each day during the period 1969–1988.
+The model decomposes the number of births along all the period in longer-term trend effects, patterns during the year, day-of-week effects, and special days effects.
+The special days effects cover patterns such as possible fewer births on Halloween, Christmas or new year, and excess of births on Valentine’s Day or the days after Christmas (due, presumably, to choices involved in scheduled deliveries, along with decisions of whether to induce a birth for health reasons).
+This analysis was originally addressed in \cite{gelman2013bayesian}.
+The total number of days within the period is $T=7305$ ($t=1,\dots,T$), then a regular GP model is unfeasible to be fitted on this dataset as we know inference scales $O(T^3)$ in covariance matrix inversion.
+Therefore, an approximate approach has to be used to fit a GP model on this data.
+We will use the HSGP model developed in this Chapter, as well as the low-rank GP model with a periodic covariance function introduced in Section \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
+
+Let's denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
 %
 \begin{equation*}
-f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\bm{\lambda}_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j,
+y_{t} \sim \mathcal{N}(\mu(t),\sigma^2).
 \end{equation*}
 
-\noindent where $\bm{\lambda}_j$ is the $D$-vector (\ref{eigenvalue_multi}) which elements are the univariate eigenvalues (\ref{eigenvalue}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
+\noindent The mean function $\mu(t)$ will be defined as an additive model in the form: 
+%
+\begin{eqnarray} \label{ch5_eq_mean_brithday}
+\mu(t) = f_1(t) + f_2(t) + f_3(t) + f_4(t).
+\end{eqnarray}
+
+The component $f_1(t)$ represents the long-term trends modeled by a GP with squared exponential covariance function,
+%
+\begin{eqnarray*}
+f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\bigg(-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\bigg), 
+\end{eqnarray*}
+
+\noindent which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
+
+The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
+%
+\begin{eqnarray*}
+f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\bigg(-\frac{2\text{sin}^2(\pi(t-t')/365.25}{\ell_2^2}\bigg), 
+\end{eqnarray*}
+
+\noindent which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.
+
+The component $f_3(t)$ represents the weekly smooth pattern using a periodic squared exponential covariance function (with period 7 of length of the week) in a GP model,
+%
+\begin{eqnarray*}
+f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\bigg(-\frac{2\text{sin}^2(\pi(t-t')/7}{\ell_3^2}\bigg), 
+\end{eqnarray*}
+
+\noindent which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.
+
+The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{piironen2017sparsity}:
+%
+\begin{eqnarray*}
+f_4(t) \sim \mathcal{N}(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
+\end{eqnarray*}
+
+\noindent A horse-shoe prior allows for sparse distributed effects.
+Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
+Different levels of sparsity can be accommodated by changing the value of $\tau$: with large $\tau$ all the variables have very diffuse priors with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
+%More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.
+
+GP priors have been defined over the components $f_1(t)$, $f_2(t)$ and $f_3(t)$.
+Then, low-rank representations of the GP priors have to be used in the modeling and inference.
+The component $f_1(t)$ will be approximated using the HSGP model.
+Thus, the function values $f_1(t)$ are approximated as in equation (\ref{eq_approxf}), with the squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+
+The year effects $f_2(t)$ and week effects $f_3(t)$, as they use a periodic covariance function, they do no fit under the main framework of the HSGP approximation covered in this chapter.
+However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix \ref{ch5_sec_periodic}).
+Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in equation (\ref{ch5_eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in equation (\ref{ch5_eq_q_2}).
+
+For the component $f_1(t)$, $m=30$ basis functions and a boundary factor $c=1.5$ were used.
+The lengthscale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
+Which means that the used number of basis functions and boundary factor are suitable values for modeling accurately the input effects.
+
+For the components $f_2(t)$ and $f_3(t)$, $J=10$ cosine terms were used.
+The  lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are accurate enough.
+
+Figure \ref{ch5_fig27_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and year patterns $f_2(t)$ for the whole period, jointly with the observed data.
+Figure \ref{ch5_fig27_posteriors_oneyear_birthday} show the process for one year (1972) only.
+In this figure, the special days effects $f_4(t)$ in the year can be clearly represented.
+The posterior means of the the function $\mu(t)$ and the components $f_1(t)$ (long-term trend) and $f_2(t)$ (year pattern) are also plotted in this Figure \ref{ch5_fig27_posteriors_oneyear_birthday}.
+Figure \ref{ch5_fig27_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3(t)$ can be clearly represented.
+The mean of the the function $\mu(t)$ and components $f_1(t)$ (long-term trend), $f_2(t)$ (year pattern) and $f_4(t)$ (special-days effects) are also plotted in this Figure \ref{ch5_fig27_posteriors_onemonth_birthday}.
+
+The Stan model code for the approximate GP model of this case study can be found in:
+%
+\begin{lstlisting}[breaklines]
+https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-II_Birthday-data
+\end{lstlisting}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig27_posteriors_birthday.pdf}}
+\caption{Posterior means of the long-term trend ($f_1(\cdot)$) and year effects pattern ($f_2(\cdot)$) for the whole series. }
+  \label{ch5_fig27_posteriors_birthday}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig27_posteriors_oneyear_birthday.pdf}}
+\caption{Posterior means of the function $\mu(\cdot)$ for the year 1972 of the series.
+The special days effects pattern ($f_4(\cdot)$) in the year is also represented, as well as the long-term trend ($f_1(\cdot)$) and year effects pattern ($f_2(\cdot)$).
+}
+  \label{ch5_fig27_posteriors_oneyear_birthday}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig27_posteriors_onemonth_birthday.pdf}}
+\caption{Posterior means of the function $\mu(\cdot)$ for the month of January of 1972.
+The week effects pattern ($f_3(\cdot)$) in the month is also represented, as well as the long-term trend ($f_1(\cdot)$), year effects pattern ($f_2(\cdot)$) and special days effects pattern ($f_4(\cdot)$).
+}
+  \label{ch5_fig27_posteriors_onemonth_birthday}
+\end{figure}
+
+\section{Multivariate case studies}\label{ch5_sec_multivariate_cases}
+
+\subsection{Leukemia data}\label{ch5_sec_bf_caseVI}
+The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
+The data set consist in survival times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for $n=1043$ cases ($i=1,\dots,n$).
+Some 16\% of cases were censored.
+Predictors are \textit{age} ($\bm{x}_1$), \textit{sex} ($\bm{x}_2$), \textit{white blood cell} (WBC) ($\bm{x}_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($\bm{x}_4$) which is a measure of deprivation for district of residence.
+We denote the matrix $X=[\bm{x}_1\; \bm{x}_2 \; \bm{x}_3 \; \bm{x}_4]^\top \in {\rm I\!R}^{n\times 4}$ which contains the predictors.
+
+As the WBC measurements were strictly positive and highly skewed, we fit the model to its logarithm.
+Continuous predictors were normalized to have zero mean and unit standard deviation.
+%Survival time was normalized to have zero mean for the logarithm of time.
+We assume a log Gaussian observation model for the observed survival time, $t_i$, with a function of the predictors, $f(X_i)$, as the location parameter, and $\sigma$ as the Gaussian noise: 
+%
+\begin{equation*}
+p(t_i)= \mathrm{LogNormal}(t_i|f(X_i),\sigma^2)
+\end{equation*}
+
+\noindent with $X_i=\{x_{i1},x_{i2},x_{i3},x_{i4}\} \in {\rm I\!R}^4$ and $f:{\rm I\!R} \to {\rm I\!R}$.
+
+As we do not have a model for the censoring process, we do not have a full observation model, and the observational model for the censored data $t_i$ is assumed to be the complementary cumulative normal probability distribution:
+%
+\begin{equation*}
+p(y_i > t_i)= \int_{t_i}^{\infty} \mathrm{LogNormal}(y_i|f(X_i),\sigma^2) dy_i=  1 - \Phi \left( \frac{\mathrm{log}(y_i)-f(X_i)}{\sigma} \right),
+\end{equation*}
+
+\noindent where $y_i$ denotes the uncensored time.
+
+The latent function $f(\cdot)$ is modeled as a Gaussian process, centered in a linear model of the predictors $X$, and with a squared exponential covariance function $k$ depending on predictors $X$ and hyperparameters $\theta=(\alpha,\ell)$,
+%
+\begin{eqnarray*} 
+f(\bm{x}) \sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta)),
+\end{eqnarray*}
+
+\noindent where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ are multivariate Gaussian distributed with mean function $\mu(\cdot)$ and covariance matrix $K$, where $\mu(\bm{x}_i)=c + \bm{\beta}\bm{x}_i$ and $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,n$.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+Notice that a scalar lengthscale is considered in the multivariate covariance function.
+
+Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable (1 for female and 2 for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
+The relative contribution of a GP function given one of the levels of the predictor (equation \ref{ch5_eq_gpml2}) to a general mean GP function (equation \ref{ch5_eq_gpml1}) is defined.
+For the other level of the predictor, the GP function effects are set to zero.
+This multilevel construction is depicted as follows:
+%
+\begin{eqnarray} 
+&f(\bm{x}) \sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
+&g(\bm{x}) \sim \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2=2)) \label{ch5_eq_gpml2} \\ 
+&f(\bm{x}|\bm{x}_2=2) = f(X|\bm{x}_2=2) + g(\bm{x}) \nonumber
+\end{eqnarray}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.70, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia_02.pdf}
+\caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
+The thick line in each graph is the posterior mean estimated using a HSGP model, and the thin lines represent pointwise 95\% credible intervals.}
+  \label{ch5_fig21_posteriors_leukemia}
+\end{figure}
+
+\noindent In the previous equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of the GP function restricted to the male sex ($\bm{x}_2=2$).
+
+Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+
+Figure \ref{ch5_fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
+These posterior estimates correspond to the HSGP model with $m=10$ basis functions and $c=3$ boundary factor.
+The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
+
+Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) and time of computation as function of the number of univariate basis functions $m$ and boundary factor $c$.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).
+%
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig22_elpd_leukemia.pdf}} 
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig23_time_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 120mm 0mm, clip]{ch5_fig24_legend_leukemia.pdf}}
+\caption{Expected log predictive density (ELPD) and time of computation in seconds per iteration (iteration of the HMC sampling method) as a function of the number of basis functions $m$ and boundary factor $c$.}
+  \label{ch5_fig22_elpd_leukemia}
+\end{figure}
+
+The Stan model codes for the exact GP and the approximate GP models of this case study can be found in:
+%
+\begin{lstlisting}[breaklines]
+https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-VII_Leukemia-data
+\end{lstlisting}
+
+\section{Conclusion}\label{ch5_sec_conclusion}
+The GP model entails a complexity that is computationally intractable for many applications, and this problem especially becomes severe when we want to perform inference using sampling methods.
+In this paper, a novel approach for a low-rank representation of GPs, originally and theoretically developed by \cite{solin2018hilbert}, has been implemented and analyzed.
+The method is based on a basis function approximation via Laplace eigenfunctions.
+The method has an attractive computational cost as this basically turns the regular GP model into a linear model, which is also an attractive property in modular probabilistic programming models.
+
+The formulae of the method and how it applies for all the stationary covariance functions as long as they can be represented in terms of their spectral densities has been clearly presented.
+The dominating cost per log density evaluation (during sampling) is $O(nm+m)$, which is a big benefit in comparison to $O(n^3)$ of a regular GP model.
+The design matrix is independent of hyperparameters and therefore only needs to be constructed once, at cost $O(nm)$.
+All dependencies on the kernel and the hyperparameters are through the prior distribution of the regression weights.
+The parameter posterior distribution is $m$-dimensional, where $m$ is much smaller than the number of observations $n$, which is greatly reduced in comparison to regular GPs and this makes inference faster, especially when sampling methods are used.
+The drawbacks of the method are the boundary conditions and scaling with respect to the number of dimensions, i.e.
+the number of basis functions $m$ scales exponentially with the number of dimensions.
+
+The main contribution of this paper has been the realization of a deep analysis and diagnosis of the performance and accuracy of the approximation in relation to the key factors of the method, the number of basis functions, the boundary condition of the Laplace eigenfunctions and the non-linearity of the function to be learned.
+Recommendations for the values of these key factors based on the recognized relations among them have been provided, as well as useful graphs of these relations that will help the users to improve performance and save time of computation, which also serve as a powerful diagnosis tool whether the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
+
+Furthermore, the focus has been put on showing how GPs can be easily used as modular components in probabilistic programming frameworks (i.e.
+Stan, WinBUGS and others) and can be used as latent functions in non-Gaussian observational models.
+Several illustrative examples, simulated and real datasets, of the performance of the model, where we demonstrated the applicability and the implementation of the methodology, the reduction of the computation and the improvement in sampling efficiency have been carried out.
+
+The main drawback of this approach is that its computational complexity scales exponentially with the number of dimensions.
+Hence, in practice, input dimensionalities larger than 3 start to be too computationally demanding.
+In these cases, the proposed HSGP model can be used as low-dimensional components in an additive modeling scheme.
+
+Future research will focus on constructing analytical models for the relationships between the key factors of the number of basis functions, the boundary factor and the lengthscale of the function, depicted in Figures \ref{fig6_lscale_vs_J_vs_c}, \ref{fig6_lscale_vs_J_vs_c_Matern} and \ref{ch5_fig28_m_lscale_periodic}, on which ultimately depend the performance of the approximation.
+This analytical models can be useful to automatize the diagnosis of the performance of the approximation.
+These relationships have been obtained, in the present study, for the unidimensional case only.
+So, another future research line will be focused on analyzing these relationships in the multidimensional case, building useful graphs or analytical models that encode these relationships in multidimensional cases.
+
+\appendix
+
+\numberwithin{equation}{section}
+\numberwithin{figure}{section}
+
+\section{Approximation of the covariance function using Hilbert space methods} \label{ch5_app_approx_covfun}
+
+In this section, we briefly present a summary of the mathematical details of the approximation of a stationary covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
+This statement is basically an extract of the work \cite{solin2018hilbert}, where the authors fully develop the mathematical theory behind the Hilbert Space approximation for stationary covariance functions.
+
+Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ over a function $f(\bm{x})$ as follows:
+%
+\begin{equation*}
+\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
+\end{equation*}
+
+From the Bochner’s and Wiener-Khintchine theorem, the spectral density of a stationary covariance function $k(\bm{x},\bm{x}') = k(\bm{\tau})$, $\bm{\tau}=(\bm{x}-\bm{x}')$, is the Fourier transform of the covariance function,
+%
+\begin{eqnarray*}
+S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \cdot \bm{\tau}}, \nonumber
+\end{eqnarray*}
+
+\noindent where $\bm{w}$ is in the frequency domain.
+The operator $\mathcal{K}$ will be translation invariant if the covariance function is stationary.
+This allows for a Fourier representation of the operator $\mathcal{K}$ as a transfer function which is the spectral density of the Gaussian process.
+%When the covariance function is homogeneous, the corresponding operator $\mathcal{K}$ will be translation invariant thus allowing for Fourier-representation as a transfer function.
+This transfer function is the spectral density of the Gaussian process
+%%
+%\begin{eqnarray*}
+%S(\bm{w}) = \mathcal{F}(\mathcal{K}). \nonumber
+%\end{eqnarray*}
+%
+%\noindent 
+Thus, the spectral density $S(\bm{w})$ also gives the approximate eigenvalues of the operator $\mathcal{K}$.
+
+In the isotropic case $S(\bm{w}) = S(||\bm{w}||)$ and assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
+%
+\begin{equation}\label{ch5_eq_S}
+S(||\bm{w}||)=a_0+a_1||\bm{w}||^2+a_2(||\bm{w}||^2)^2+a_3(||\bm{w}||^2)^3+\cdots.
+\end{equation}
+
+\noindent The Fourier transform of the Laplace operator $\nabla^2$ is $-||\bm{w}||$, thus the Fourier transform of $S(||\bm{w}||)$ is
+%
+\begin{equation}\label{ch5_eq_K}
+\mathcal{K}=a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+a_3(-\nabla^2)^3+\cdots,
+\end{equation}
+
+\noindent defining a pseudo-differential operator as a series of Laplace operators.
+
+If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the formal kernel $l$,
+%
+\begin{equation*}
+-\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}',
+\end{equation*}
+
+\noindent then the formal kernel can be represented as 
+%
+\begin{equation*}
+l(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{equation*}
+
+\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
 %
 \begin{align*}
-\bm{\lambda}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^{D} = \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^{D},
+-\nabla^2 \phi_j(\bm{x})&=\lambda \phi_j(\bm{x}), \hspace{1cm}  x\in \Omega \nonumber \\ 
+\phi_j(\bm{x})&=0, \hspace{1.85cm} x\notin \Omega.
 \end{align*}
 
-\noindent and $\phi_j$ is the multivariate eigenfunction (\ref{eigenfunction_multi}) as the product of univariate eigenfunctions (\ref{eigenfunction}) whose indices correspond to the elements of the $D$-tupple $\mathbb{S}_{j\cdotp}$,
-%
-\begin{align*}
-\phi_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}} = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right),
-\end{align*}
-
-\noindent where $x_d$ is the input value corresponding to dimension $d$.
-$S$ is the spectral density (\ref{eq:specdens_inf}), as a function of $\sqrt{\bm{\lambda}_j}= \left\{ \sqrt{\lambda_{\mathbb{S}_{jd}}} \right\}_{d=1}^{D}$, of the $D$-dimensional square exponential covariance function,
+\noindent Because $-\nabla^2$ is a positive definite Hermitian operator, the set of eigenfunctions $\phi_j(\cdot)$ are orthonormal with respect to the inner product
 %
 \begin{eqnarray*}
-S(\sqrt{\bm{\lambda}_j})= \sigma^2 2\pi \prod_{d=1}^D \ell_d  \mathrm{exp}\bigg(-\frac{1}{2} \sum_{d=1}^D \ell_d^2  \sqrt{\lambda_{\mathbb{S}_{jd}}}^2 \bigg). 
+<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) d(\bm{x})
 \end{eqnarray*}
 
-\noindent In the previous equations, $j$ denotes the index for the $m= \prod_{d=1}^{D} m_d$ multivariate basis functions (\ref{m_multi}), where $m_d$ is the number of basis functions considered for dimension $d$.
-$\mathbb{S}$ is the matrix of $D$-tuples, with rows being the indices of every possible combinations of univariate eigenvalues (\ref{eigenvalue}) over the $D$ dimensions.
-$L_d$ is the boundary for the dimension $d$.
-The parameters $\beta_j$ are $\mathcal{N}(0,1)$ distributed, and $\alpha$ and $\ell_d$ are the marginal variance and lengthscale of dimension $d$, respectively, of the approximate multivariate covariance function.
-
-In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
-
-Figure \ref{fig15_Posteriors_exII} shows the data-generating GP function from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset.
-Sample observations are also plotted in the plots as circles.
-For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
-A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ were used.
-For the splines model, 40 knots in each dimension were used.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{ c c c}
-\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 10mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\
-\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
-\end{tabular}
-\end{center}
-\caption{(a) Data-generating function.
-(b) Mean posterior predictive function of the GP model.
-(c) Mean posterior predictive function of the HSGP model.
-(d) Mean posterior predictive function of the splines (SP) model.}
-  \label{fig15_Posteriors_exII}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 5mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
-\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models.}
-  \label{fig16_errors_exII}
-\end{figure}
-
-In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions for the HSGP model and different number of knots for the splines model have been fitted.
-In all models, the same number of basis functions and knots per dimension were used.
-Figure \ref{fig17_RMSE_exII}(left) shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, the number of univariate basis functions $m$ and knots.
-From Figures \ref{fig16_errors_exII} and \ref{fig17_RMSE_exII}(left), it can be seen a close approximation of the HSGP model to the regular GP model.
-However, the performance of the splines model is significantly worse.
-Figure \ref{fig17_RMSE_exII}(right) shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots.
-Figure \ref{fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computational time.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
-\caption{Root mean square error (RMSE) (left) and computational time in seconds per iteration (right) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
-  \label{fig17_RMSE_exII}
-\end{figure}
-
-\subsection{Study case V: Diabetes data}\label{sec:bf_caseV}
-The next example presents an epidemiological study of the diabetes diseases.
-The study aims to relate the probability of suffering from diabetes to some risk factors.
-The data contains 392 individuals ($i=1,\cdots,n=392$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed.
-The matrix $X=\{\bm{x}_i\}_{i=1}^{n=392} \in {\rm I\!R}^{n\times 4}$,  with $\bm{x}_i=\{x_{i1},x_{i2},x_{i3},x_{i4}\} \in {\rm I\!R}^{4}$, contains the risk factors, \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$.
-The observational model is a bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
+\noindent that is,
+%
+\begin{eqnarray*}
+\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) d(\bm{x}) = \delta_{ij},
+\end{eqnarray*}
+
+\noindent and all the eigenvalues $\lambda_j$ are real and positive.
+
+Due to normality of the basis of the representation of the formal kernel $l(\bm{x},\bm{x}')$, its formal powers $s=1,2,\dots$ can be write as
+%
+\begin{eqnarray}\label{ch5_eq_formalkernel}
+l(\bm{x},\bm{x}')^s= \sum_j \lambda_j^s \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{eqnarray}
+
+\noindent which are again to be interpreted to mean that
 %
 \begin{equation*}
-y_i \sim \text{Bernoulli}(p_i).
+(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
 \end{equation*}
 
-\noindent The goal is to estimate the 
- probability $p_i$ as a function of the risk factors, which function $f(X)$ is modeled as a Gaussian process with a squared exponential covariance function $k$ depending on the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
-%
-\begin{eqnarray*} \label{eq:gpprior_gay}
-p_i &=& \text{logit}(f(\bm{x}_i)) \nonumber \\
-f(X) &\sim& \mathcal{GP}(0, k(X,X, \theta).
-\end{eqnarray*}
-
-\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-Notice that a scalar lengthscale is considered in the multivariate covariance function.
-
-In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$, evaluated at input vector $\bm{x} \in {\rm I\!R}^D$, is approximated as in equation (\ref{approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) square exponential spectral density $S$ as in equation (\ref{eq:specdens_inf}), and the $D$-vector of eigenvalues $\bm{\lambda}_j$, and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eigenvalue_multi}) and (\ref{eigenfunction_multi}), respectively.
-
-In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional intergrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
-
-Figure \ref{fig18_gpfun_diabetes} shows the mean posterior predictions of probabilities ($p_i$) of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset with the 2 input dimensions ($D=2$) \textit{Glucose} and \textit{Pregnancy}.
-The binary observations $y_i$ are also plotted in the plots as colored points.
-For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions.
-A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used.
-For the splines model, 20 knots per dimension were used.
-%
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 149mm 18mm 5mm 20mm, clip]{fig18_legend_diabetes.pdf}}
-\caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the splines (SP) model.}
-  \label{fig18_gpfun_diabetes}
-\end{figure}
-
-In order to assess performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factor for the HSGP model and different number of knots for the splines model have been fitted.
-In all models, the same boundary factor, number of basis functions and knots per dimension were used.
-Figure \ref{fig19_ELPD_diabetes} shows the expected log predictive density (ELPD) as a function of the boundary factor $c$, the number of univariate basis functions $m$ and knots.
-The ELPD is computed over the actual observations.
-Basically, with slightly differences, all models show similar performances, due to the fact that the process is very smooth with a relatively very large lengthscale estimate $\ell=4.51$.
-Even tough, a slightly pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (\ref{fig6_lscale_vs_J_vs_c}).
-%
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig19_ELPD_diabetes.pdf}}
-\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$ and knots.}
-  \label{fig19_ELPD_diabetes}
-\end{figure}
-
-Figure \ref{fig20_time2D_diabetes} shows the computational times in the logarithm scale of the different models, regular GP, HSGP and splines, fitted over the dataset with 2 input dimensions (left), 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots.
-We can appreciate the significant increase of computational time with higher dimensions for the HSGP and splines models.
-This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid excessive computational time, specially in high input dimensionality.
-It is interesting to be noticed that considering more than 10 knots per dimension in the splines model with 3D is not allowed for an amount of 392 observations.
-Similarly, just the computation of the input data for the splines model with 4D input dimension is computationally very expensive.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.29, trim = 10mm 0mm 13mm 0mm, clip]{fig20_time2D_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time3D_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 25mm 0mm 13mm 0mm, clip]{fig20_time4D_diabetes.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 22mm 55mm 120mm 0mm, clip]{fig20_legend_diabetes.pdf}}
-\caption{Time of computation in the log scale of the different models fitted over the dataset with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
-  \label{fig20_time2D_diabetes}
-\end{figure}
-
-\subsection{Study case VI: Leukemia data}\label{sec:bf_caseVI}
-\subsection{Study case VII: Land use spatio-temporal classification task}\label{sec:bf_caseVII}
-
-\vspace{3mm}
-\appendix
-
-\section{Related work}
-
-The GP prior entails an $O(n^3)$ complexity that is computationally intractable for many practical problems, and this problem especially becomes severe when we want to conduct inference using sampling methods.
-To overcome this scaling problem several schemes have been proposed.
-One approach is to partition the data set into separate groups \citep{snelson2007local, urtasun2008sparse} and performing local inference in each partition.
-Other global approach is to build a low-rank approximation to the covariance matrix of the complete data based around 'inducing variables' \citep{quinonero2005unifying,bui2017unifying}.
-Other global approach make use of basis functions to approximate the covariance function.
-In \cite{snelson2007local} the authors conduct an approach that combines the idea of local and global approaches.
-
-The literature contains many parametric models that approximate Gaussian process behaviours; for example \cite{bui2014tree} included  tree-structures in the approximation for extra scalability, and \cite{moore2015gaussian} combined local Gaussian
-processes with Gaussian random fields.
-
-\subsection{Inducing points methods}
-
-The approach based on inducing points employs a small set of pseudo data points to summarise the actual data.
-The storage requirements are reduced to $O(nm)$ and complexity to $O(nm^2)$, where $m < n$.
-Some of these methods have been reviewed in \cite{rasmussen2006gaussian}, and \cite{quinonero2005unifying} provide a unifying view of these methods based on approximate generative methods.
-From a spectral point of view, several of these methods (e.g., SOR, DTC, VAR, FIC) can be interpreted as modfications to the so-called Nystr{\"o}m method (see \cite{arthur1979baker} and \cite{williams2001using}), a scheme for approximating the eigenspectrum.
-These methods are basically based on choosing a set of $m$ inducing inputs $x_u$ and scaling the corresponding eigendecomposition of their corresponding covariance matrix $K_{u,u}$ to match that of the actual covariance.
-
-This scheme was originally introduced to the GP context by \cite{williams2001using}.
-As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model.
-However, several methods modifying the inducing point approach are widely used.
-The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix.
-The SoR method is based on a degenerate prior which produces unreasonable predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).
-
-The Deterministic Training Conditional (DTC) method \citep{ro2001sparse,seeger2003fast}) retains the true covariance for the training data, but uses the approximate cross-covariances between training and test data, which reverse the problem of nonsensical predictive uncertainties.
-However, since the covariances for training and test cases are computed differently, this method results not to actually be a Gaussian process.
-This method was presented as Projected Latent Variables (PLV) in \cite{seeger2003fast} and Projected Process Approximation (PPA) in \cite{rasmussen2006gaussian}.
-
-The Variational Approximation (VAR) \citep{titsias2009variational} suggests a variational approach which provides an objective function for optimizing the selection of inducing points.
-This basically modifies the DTC method by an additional trace term in the likelihood that comes from the variational bound.
-\cite{hensman2013gaussian} extended this idea by introducing additional variational parameters to enable stochastic variational inference \citep{hoffman2013stochastic}, achieving a more computationally scalable bound which allows GPs to be fitted to millions of data.
-
-The Fully Independent (Training) Conditional (FIC) \citep{quinonero2005unifying} method originally introduced as Sparse Pseudo-Input GP by \cite{snelson2006sparse} is also based on the Nystr\"om approximation, where they allow the pseudo-point input locations to be optimised by maximising the new model's marginal likelihood whose covariance is parameterized by the locations of an active set not constrained to be a subset of the training and test data.
-
-More recently Bui et.
-al (2017) revisit the inducing points-based sparse approximation methods, in which all the necessary approximation is performed at inference time, rather than at the modelling time.
-The new framework is built on standard methods for approximate inference (variational-free-inference, EP and Power EP methods).
-
-In practice, the inducing points-based sparse approximation methods works reasonable well in cases where the field is relatively smooth. \cite{vanhatalo2010approximate}
-propose the use of compactly supported covariance function in conjunction with sparse approximations to model both short and long range correlations.
-
-\cite{wilson2015kernel} introduce a new unifying framework for inducing point methods, called structured kernel interpolation (SKI).
-This framework improves the scalability and accuracy of fast kernel approximations through kernel interpolation, and naturally combines the advantages of inducing point and structure exploiting for scalability (such as Kronecker \citep{saatcci2012scalable} or Toeplitz \citep{cunningham2008fast}) approaches.
-
-The number of inducing points or their locations are crucial in order to capture the correlation structure.
-For a discussion on the effects of the inducing points, see \cite{vanhatalo2010approximate}.
-This behavior applies to all the methods from the Nystr\''om family.
-
-This kind of 'projected process' approximation has also been discussed by e.g. \cite{banerjee2008gaussian}.
-
-\subsection{Basis function methods}
-
-The spectral analysis and series expansions of Gaussian processes has a long history.
-A classical result (see, e.g, \cite{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}, and references therein) is that the covariance function can be approximated with a finite truncation of Mercer series and the approximation is guaranteed to converge to the exact covariance function when the number of terms is increased.
-
-Another related classical connection is to the works in the relationship of spline interpolation and Gaussian process priors \citep{wahba1978improper,kimeldorf1970correspondence,wahba1990spline}.
-In particular, it is well-known (see, e.g., \cite{wahba1990spline}) that spline smoothing can be seen as Gaussian process regression with a specific choice of covariance function.
-The relationship of the spline regularization with Laplace operators then leads to series expansion representations that are closely related to the approximations considered here.
-
-Random Fourier Features \citep{rahimi2008random,rahimi2009weighted} is a method for approximating kernels. The approximate kernel has a finite basis function expansion.
-
-The Sparse Spectrum GP is based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse}, where the spectral representation of the covariance function is used.
-This model is a stationary sparse GP that can approximate any desired stationary full GP.
-However, as argued by the authors, this option does not converge to the full GP and can suffer from overfitting to the training data.
-\citep{gal2015improving} sought to improve the model by integrating out, rather than optimizing the frequencies.
-Gal and Turner derived a variational approximation that made use of a tractable integral over the frequency space.
-The result is an  algorithm that suffers less overfitting than the Sparse Spectrum GP, yet remains flexible.
-
-While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
-
-Recently \citep{hensman2017variational} presented a variational Fourier feature approximation for Gaussian processes that was derived for the Mat{\'e}rn class of kernels, where the approximation structure is set up by a low-rank plus diagonal structure.
-They combine the variational methodology with Fourier based approximations.
-
-In spatial statistics similar approaches are called low-rank models \citep{diggle2007springer}.
-The low rank models assume that the Gaussian field is a linear combination of $m$ basis functions.
-The type of an approximation depends on the basis functions used.
-Familiar examples include spectral representation \citep{diggle2007springer,paciorek2007computational,paciorek2007bayesian} and splines \citep{wood2003thin}.
-
-Recent Splines models can reproduce the Matern family of covariance functions, however our approach can reproduce basically all of the stationary covariance functions.
-
-\section{Contributions of the method}
-
-This work is based on the novel method developed by \cite{solin2018hilbert} for reduced-rank approximations of GP models.
-This method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods.
-This results in a reduced-rank approximation for the covariance function.
-This method has some nice features:
-
-\vspace{2mm}
-$\bullet$ It has an attractive computational cost as this basically turns the regular GP model into a lineal model.
-
-\vspace{2mm}
-$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method.
-In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.
-
-\vspace{2mm}
-$\bullet$ Using maximizing marginal likelihood methods, the proposed model has a overall complexity of $O(nm^2)$.
-After this, evaluating the marginal likelihood and marginal likelihood gradients is an $O(m^3)$ operation in every step of the optimizer.
-(Arno's paper, pag.
-7)
-
-\vspace{2mm}
-$\bullet$ The parameter posterior distribution in this approximate GP model is $m$-dimensional ($m<<n$) which helps the use of GP priors as latent functions.
-especially when sampling methods for inference are used.
-GP prior as latent functions is needed in generalized models.
-
-In regular GPs and other approximate GP models and Splines models these features do not have so nice properties:
-
-\vspace{2mm}
-$\bullet$ In a regular GPs, the main computational complexity comes from the inversion of the covariance matrix which is in general a $O(n^3)$ operation.
-This operation has to be computed at every step of the HMC or optimizer.
-
-\vspace{2mm}
-$\bullet$ In regular GPs, the parameter posterior distributions is $N$-dimensional.
-It is known that when $N$ is of medium or large size there is high correlation between the $N$-dimensional latent function and the hyperparameters of the GP prior.
-
-\vspace{2mm}
-$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.
-
-\vspace{2mm}
-$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach.
-However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter.
-In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.
-
-\section{Contributions of our work}
-
-As said above the proposed method was already developed by \cite{solin2018hilbert} where they fully develop, describe and generalize the methodology.
-Though, they do not put much effort in describing and analyzing the relation among the key factors of the box size (or boundary condition), the number of basis functions, and the smoothness or roughness of the function.
-The performance and accuracy of the method are directly related with the number of basis functions and the box size.
-At the same time, successful values for these two factors depend on the smoothness or roughness of the process to be modeled.
-The time of computation is mainly dependent on the number of basis functions.
-Our main contributions to this recently developed methodology for low-rank GP model by \cite{solin2018hilbert} goes around these aspects.
-
-\vspace{2mm}
-$\bullet$ Firstly, clear summarized formulae of the method for the univariate and multivariate cases is presented.
-
-\vspace{2mm}
-$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthscale of the functions.
-
-\vspace{2mm}
-$\bullet$ We make recommendations for the values of these factors based on the recognized relations among them.
-We provide useful graphs of these relations that will help the users to improve performance and save time of computation.
-
-\vspace{2mm}
-$\bullet$ We also diagnose if the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
-
-\vspace{2mm}
-$\bullet$ We describe the generalization of the method to the multidimensional case.
-
-\vspace{2mm}
-$\bullet$ We implement the approach in a fully probabilistic framework and for the Stan programming probabilistic software.
-
-\vspace{2mm}
-$\bullet$ We show several illustrative examples, simulate and real datasets, of the performance of the model, and accompanied by their Stan codes.
-
-%\vspace{3mm}
-%A summary of the formulae of the method for the univariate and multivariate cases is presented.
-The number of basis functions used in the approach and the value for the boundary condition of the model are two factors of main importance in practical applications.
-The number of basis functions is directly related to the accuracy of the approximation but also to the computation demands.
-Also exist a relation among the number of basis functions, the specified values for the boundary condition, and the performance of the approach.
-Finally, the approach is applied in several illustrative study cases and accompanied by their Stan codes.
-
-\section{Spectral densities of stationary covariance functions}
-
-The covariance function of a stationary process, that is function of $\bm{\tau}=\bm{x-x'}$ can be represented as the Fourier transform of a positive finite measure ($Bochner's theorem$).
-
-\vspace{0.2cm}
-\textit{(Bochner’s theorem) A complex-valued function $k$ on ${\rm I\!R}^D$ is the covariance function of a weakly stationary mean square continuous complex valued random process on ${\rm I\!R}^D$ if and only if it can be represented as}
-%
-\begin{equation}
-k(\bm{\tau})= \int_{{\rm I\!R}^D} e^{2\pi i \bm{s} \cdot \bm{\tau}} d\mu(\bm{\tau}), \nonumber 
+\noindent This implies that we also have
+%
+\begin{equation*}
+[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}') d\bm{x}'.
+\end{equation*}
+
+Then, looking at equations (\ref{ch5_eq_K}) and (\ref{ch5_eq_formalkernel}), it can be concluded 
+%
+\begin{eqnarray}\label{ch5_eq_k}
+k(\bm{x},\bm{x}')= \sum_j [a_0+a_1\lambda_j^1+a_2\lambda_j^2+\cdots] \phi_j(\bm{x}) \phi_j(\bm{x}').
+\end{eqnarray}
+
+\noindent By letting $||\bm{w}||^2=\lambda_j$ the spectral density in Equation (\ref{ch5_eq_S}) becomes
+%
+\begin{equation*}
+S(\sqrt{\lambda_j})=a_0+a_1\lambda_j+a_2\lambda_j^2+a_3\lambda_j^3+\cdots,
+\end{equation*}
+
+\noindent and substituting in equation (\ref{ch5_eq_k}) then leads to the final searched approximation
+%
+\begin{eqnarray}\label{ch5_eq_k_2}
+k(\bm{x},\bm{x}')= \sum_j S(\sqrt{\lambda_j}) \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{eqnarray}
+
+\noindent where $S(\cdot)$ is the spectral density of the covariance function, $\lambda_j$ is the $j$th eigenvalue and $\phi_j(\cdot)$ the eigenfunction of the Laplace operator in a given domain.
+
+\section{Low-rank Gaussian process with a periodic covariance function}\label{ch5_sec_periodic}
+
+A GP model with a periodic covariance function does no fit under the framework of the HSGP approximation covered so far in this study.
+However, it do has a low-rank representation.
+In this section, we first give a brief presentation of the results from \cite{solin2014explicit}, where the authors obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
+Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic square exponential covariance function.
+
+The periodic squared exponential covariance function takes the form
+%
+\begin{eqnarray} \label{ch5_eq_cov_periodic}
+k(\bm{\tau})= \sigma^2 \exp\Big(-\frac{2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)}{\ell^2}\Big),
+\end{eqnarray}
+
+\noindent where $\sigma^2$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
+
+In \cite{solin2014explicit}, the authors come to a cosine series expansion for the periodic covariance function (\ref{ch5_eq_cov_periodic}) as follows,
+% 
+\begin{eqnarray} \label{ch5_eq_cov_periodic_taylor_approx}
+k(\tau)= \sigma^2 \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 \tau),
+\end{eqnarray}
+
+\noindent which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ of the previous expression are
+%
+\begin{eqnarray} \label{ch5_eq_q}
+\tilde{q}_j^2= \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
+\end{eqnarray}
+
+\noindent where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
+%
+\begin{eqnarray} \label{ch5_eq_q0}
+\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
+\end{eqnarray}
+
+\noindent Note that the covariance in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
+As \cite{solin2014explicit} argue, this approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
+
+%If the requirement of a valid covariance function is relaxed and only an optimal series approximation is required, taking the limit $J \rightarrow \infty$ in the sub-sums in equation (\ref{ch5_eq_q}) gives the following variance coefficients
+An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}), and thus leading to the following variance coefficients:
+%
+\begin{equation}\label{ch5_eq_q_2}
+\begin{split}
+\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)}, \\
+\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)},
+\end{split}
 \end{equation}
 
-\textit{where $\mu$ is a positive finite measure.}
-
-\vspace{0.2cm}
-If the measure $\mu$ has a density, it is known as the spectral density $S(\omega)$ of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the Wiener-Khintchine theorem.
-It gives the following relations:
-
-\begin{eqnarray}
-k(\bm{\tau})&=& \int S(\bm{s}) e^{2\pi i \bm{s} \cdot \bm{\tau}} d\bm{s}  \nonumber \\
-%
-S(\bm{s})&=& \int k(\bm{\tau}) e^{-2\pi i \bm{s} \cdot \bm{\tau}} d\bm{s}  \nonumber
+\noindent for $j=1,2,\cdots,J$, and where the $\text{I}_{\alpha}(z)$ is the modified Bessel function \citep{handbook1970m} of the first order of $\alpha$.
+This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required \citep{solin2014explicit}.
+A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function can be found in \cite{solin2014explicit}.
+
+%Note that the terms in equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}) are underneath bounded approximations for any $J$.
+
+In order to assess the accuracy of this representation as a function of the number of cosine terms $J$ considered in the approximation, an empirical evaluation is carried out in a similar way than that in Section \ref{ch5_sec_accuracy} of this work.
+Thus, Figure \ref{ch5_fig28_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the lengthscale of the kernel.
+We have considered an approximation to be close enough in terms of satisfying equation (\ref{ch5_eq_diff_covs}) with $\varepsilon=0.5\%$.
+Notice that since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix.pdf}}
+\hspace{3mm}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix_zoom.pdf}}
+\caption{Relation among the minimum number of terms $J$ in the approximation and the lengthscale ($\ell$) of the periodic squared exponential covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+  \label{ch5_fig28_m_lscale_periodic}
+\end{figure}
+
+The function values of a GP model with this low-rank representation of the periodic exponential covariance function can be easily derived. Considering the identity
+\begin{equation*}
+\text{cos}(j\omega_0 (x-x'))=\text{cos}(j\omega_0 x) \text{cos}(j\omega_0 x') + \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x'),
+\end{equation*}
+
+\noindent the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be re-writting as
+%
+\begin{eqnarray} \label{ch5_eq_cov_periodic_taylor_approx_2}
+k(x,x')= \sigma^2 \Big( \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 x)  \text{cos}(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x') \Big).
 \end{eqnarray}
 
-\section{Approximate the covariance function using Hilbert space methods}
-
-Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ as follows:
-%
-\begin{equation}
-\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
+\noindent where $\tau=x-x'$.
+With this approximation for the periodic squared exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \mathcal{GP}(0,k(x,x')$ equivalently leads to a linear representation of $f(\cdot)$ via
+%
+\begin{equation} \label{ch5_eq_f_period}
+f(x) \approx \sigma \Big( \sum_{j=0}^J  \tilde{q}_j \text{cos}(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \text{sin}(j\omega_0 x) \beta_{J+1+j} \Big),
 \end{equation}
 
-Assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
-%
-\begin{equation}
-S(\bm{w})=a_0+a_1\bm{w}^2+a_2(\bm{w}^2)^2+a_1(\bm{w}^2)^3+\cdots
-\end{equation}
-
-If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the covariance function $k$,
-
-\begin{equation}
--\nabla^2 f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}',
-\end{equation}
-
-\noindent then the covariance function can be represented as
-
-\begin{equation}
-k(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
-\end{equation}
-
-\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
-Namely, they satisfy the following eigenvalue problem in the compact subset $x \in \{-L,L\}$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
-%
-\begin{eqnarray}
--\nabla^2 \phi_j(x)&=&\lambda \phi_j(x), \hspace{1cm}  x\in \{-L,L\} \nonumber \\ 
-\phi_j(x)&=&0, \hspace{2cm} x\notin \{-L,L\}.
-\end{eqnarray}
-
-a series expansion of eigenvalues and eigenfunctions
-
-\section{Example of generalization to the multivariate case}
-
-Next, as an example we show the matrix $\mathbb{S}$ and eigenfunctions and eigenvalues for a $two$-dimensional input vector $\bm{x}=\{\bm{x}_1,\bm{x}_2\}$ ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
-The number of new multidimensional eigenfunctions $\phi^{\ast}_j$ and eigenvalues $\lambda^{\ast}_j$ is $J^D=3^2=9$ ($j=\{1,\cdots,J^D\}$).
-The matrix $\mathbb{S}\in {\rm I\!R}^{9 \times 2}$ is
-%
-\begin{eqnarray}
-\mathbb{S}=
-\left[ {\begin{array}{cc}
-1 & 1 \nonumber \\
-1 & 2 \\
-1 & 3 \\
-2 & 1 \\
-2 & 2 \\
-2 & 3 \\
-3 & 1 \\
-3 & 2 \\
-3 & 3
-\end{array} } \right]
-\end{eqnarray}
-
-\noindent and the multidimensional eigenfunctions and eigenvalues
-%
-\begin{eqnarray}
-\phi^{\ast}_1(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_1 = \{\lambda_{1}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_2(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_2 = \{\lambda_{1}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} \nonumber \\
-%
-\phi^{\ast}_3(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_3 = \{\lambda_{1}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_4(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_4 = \{\lambda_{2}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_5(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_5 = \{\lambda_{2}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_6(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_6 = \{\lambda_{2}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_7(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_7 = \{\lambda_{3}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_8(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_8 = \{\lambda_{3}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\}  \nonumber \\
-%
-\phi^{\ast}_9(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2)  \hspace{2cm}
-%
-\bm{\lambda}^{\ast}_9 = \{\lambda_{3}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\}  \nonumber
-\end{eqnarray}
-
-Now, we show another example where different number of eigenfunctions and eigenvalues are used for every dimension.
-We consider a three-dimensional ($D=3$) input space, and sets of $J_{1}=2$, $J_{2}=2$ and $J_{3}=3$ eigenfunctions and eigenvalues for the first, second and third dimensions, respectively.
-The number of new multidimensional eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$ is $J_{1}\cdot J_{2}\cdot J_{3}=2\cdot 2\cdot 3=12$.
-The matrix $\mathbb{S}\in {\rm I\!R}^{12 \times 3}$ is
-%
-\begin{eqnarray}
-\mathbb{S}=
-\left[ {\begin{array}{ccc}
-1 & 1 & 1 \nonumber \\
-1 & 1 & 2 \\
-1 & 1 & 3 \\
-1 & 2 & 1 \\
-1 & 2 & 2 \\
-1 & 2 & 3 \\
-2 & 1 & 1 \\
-2 & 1 & 2 \\
-2 & 1 & 3 \\
-2 & 2 & 1 \\
-2 & 2 & 2 \\
-2 & 2 & 3 
-\end{array} } \right]
-\end{eqnarray}
-
-\noindent and the multidimensional eigenfunctions and eigenvalues
-%
-\begin{eqnarray}
-\phi^{\ast}_1= \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_1= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_2= \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_2= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_3= \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_3= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_4= \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_4= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_5= \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_5= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_6= \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_6= \{\lambda_1(\bm{x}_1), \lambda_1(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_7= \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_7= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_8= \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_8= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_9= \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_9= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_{10}= \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{1}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_{10}= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_1(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_{11}= \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{2}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_{11}= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_2(\bm{x}_3)\} \nonumber \\
-%
-\phi^{\ast}_{12}= \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) \cdot \phi_{3}(\bm{x}_3) \hspace{1.5cm} \bm{\lambda}^{\ast}_{12}= \{\lambda_2(\bm{x}_1), \lambda_2(\bm{x}_2), \lambda_3(\bm{x}_3)\} \nonumber
-\end{eqnarray}
-
-\newpage
-\begin{landscape}
-
-\section{Multidimensional generalization of covariance functions and spectral densities}
-
-\subsection{Square Exponential covariance function ($k$) \textbf and spectral density ($S$)}
-\vspace{-2mm}
-\subsubsection{Using norm-L2 (Euclidean distance)}
-\vspace{-10mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-20mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L2} =|\bm{x}-\bm{x}'|_{L2} = \sqrt{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')} = \sqrt{(x_1-x_1')^2+(x_2-x_2')^2}  = \sqrt{r_1^2+r_2^2} \in \mathbb{R}; \hspace{2mm} k(\tau_{L2},\ell) = \mathrm{exp}\left(-\frac{1}{2} \frac{\tau^2}{\ell^2} \right)\\
-&& \omega_{L2}= \sqrt{s_1^2+s_2^2} \in \mathbb{R}; \hspace{2mm} S(\omega_{L2},\ell) = \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \omega_{L2}^2 \right)
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-       \hline
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		k(\tau_{L2},\ell) &=& k(|\bm{x}-\bm{x}'|_{L2},\ell)\\
-		 &=& \mathrm{exp}\left(-\frac{1}{2} \frac{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')}{\ell^2} \right)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{\sum_{i=1}^{2}(x_i-x_i')^2}{\ell^2} \right) \\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{\small
-         \begin{eqnarray*}
-		S(\omega_{L2},\ell) &=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 (s_1^2 + s_2^2) \right) \\
-		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell^2 s_i^2) \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-SEPARABLE:} \\
-		\\
-		&&k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
-		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		&&S(\omega_{L2},\bm{\ell})= S(s_1,\ell_1)S(s_2,\ell_2)
-         \end{eqnarray*}
-       }\\
-        \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		k(\tau_{L2},\bm{\ell}) &=& k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{\small
-       \begin{eqnarray*}
-		S(\omega_{L2},\bm{\ell}) &=& \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
-	   \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-
-       } \\
-
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&&= \mathrm{exp}\left(-\frac{1}{2} \frac{r_1^2}{\ell_1^2} \right) \mathrm{exp}\left(-\frac{1}{2} \frac{r_2^2}{\ell_2^2} \right) \\
-		&&=\mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{\small
-         \begin{eqnarray*}
-         &&S(s_1,\ell_1)S(s_2,\ell_2) \\
-		&&= \sqrt{2\pi} \cdot \ell_1 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_1^2 s_1^2 \right) \\
-		 &&\times  \sqrt{2\pi} \cdot \ell_2 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_2^2 s_2^2 \right)\\
-		&&= \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
-         \end{eqnarray*}         
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-
-       } \\
-
-       \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\newpage
-\subsubsection{Using norm-L1 }
-\vspace{-10mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-20mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L1} =|\bm{x}-\bm{x}'|_{L1} = |x_1-x_1'|+|x_2-x_2'| = r_1+r_2 \in \mathbb{R}; \hspace{2mm} k(\tau_{L1},\ell) = \mathrm{exp}\left(-\frac{1}{2} \frac{\tau_{L1}^2}{\ell^2} \right)\\
-&& \omega_{L1}= s_1+s_2 \in \mathbb{R}; \hspace{2mm} S(\omega_{L1},\ell) = \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \omega_{L1}^2 \right)
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-       \hline
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		k(\tau_{L1},\ell) &=& k(|\bm{x}-\bm{x}'|_{L1},\ell)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{(r_1 + r_2)(r_1 + r_2)}{\ell^2} \right)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{(r_1 + r_2)}{\ell}\frac{(r_1 + r_2)}{\ell} \right)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \left(\frac{r_1}{\ell} + \frac{r_2}{\ell}\right)\left(\frac{r_1}{\ell} + \frac{r_2}{\ell}\right) \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		S(\omega_{L1},\ell) &=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} \ell^2 (s_1 + s_2)\\
-		&\cdot & (s_1 + s_2) \Big) \\
-		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} \ell (s_1 + s_2)\\
-		&\cdot & \ell(s_1 + s_2) \Big) \\
-		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} (\ell s_1 + \ell s_2)\\
-		&\cdot &(\ell s_1 + \ell s_2) \Big)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-NO SEPARABLE:} \\
-		\\
-		&&k(|\bm{x}-\bm{x}'|_{L1},\bm{\ell})\\
-		 &&\neq k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		&&S(\omega_{L1},\bm{\ell})\neq S(s_1,\ell_1)S(s_2,\ell_2)
-         \end{eqnarray*}
-       }\\
-       \vspace{-10mm}\\
-        \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		k(\tau_{L1},\bm{\ell}) &=& k(|\bm{x}-\bm{x}'|_{L1},\bm{\ell})\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \left(\frac{r_1}{\ell_1} + \frac{r_2}{\ell_2}\right)\left(\frac{r_1}{\ell_1} + \frac{r_2}{\ell_2}\right) \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		S(\omega_{L1},\bm{\ell}) &=& \sqrt{2\pi}^D \cdot \ell_1 \ell_2 \cdot \mathrm{exp}\Big(-\frac{1}{2} (\ell_1 s_1 + \ell_2 s_2)\\
-		&\cdot &(\ell_1 s_1 + \ell_2 s_2) \Big) \\
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.2cm}|}{
-
-       } \\  
-       \vspace{-15mm}\\
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-        $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&&= \mathrm{exp}\left(-\frac{1}{2} \frac{r_1^2}{\ell_1^2} \right) \mathrm{exp}\left(-\frac{1}{2} \frac{r_2^2}{\ell_2^2} \right)\\
-		&&= \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{2}\frac{r_i^2}{\ell_i^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{
-       \begin{eqnarray*}
-		&&S(s_1,\ell_1)S(s_2,\ell_2) \\
-		&&= \sqrt{2\pi} \cdot \ell_1 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_1^2 s_1^2 \right) \\
-		&& \times \sqrt{2\pi} \cdot \ell_2 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_2^2 s_2^2 \right)\\
-		&&= \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
-		\end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.2cm}|}{
-
-       } \\          
-       \hline\\   
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\newpage	
-\subsubsection{Using vector difference of inputs}
-\vspace{-5mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-10mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\bm{\tau} =\bm{x}-\bm{x}' = (x_1-x_1',x_2-x_2') = (r_1,r_2) \in \mathbb{R}^2; \hspace{2mm} k(\bm{\tau},\ell) = \mathrm{exp}\left(-\frac{1}{2} \frac{\bm{\tau}^2}{\ell^2} \right)\\
-&& \bm{\omega}= (s_1,s_2) \in \mathbb{R}^2; \hspace{2mm} S(\bm{\omega},\ell) = \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \bm{\omega}^2 \right)
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-
-       \hline
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{
-         \begin{eqnarray*}
-		k(\bm{\tau},\ell) &=& k(\bm{x}-\bm{x}',\ell)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{\bm{\tau}^\top \bm{\tau}}{\ell^2} \right)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \frac{(r_1, r_2)^\top (r_1, r_2)}{\ell^2} \right)\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		S(\bm{\omega},\ell) &=& S(\bm{x}-\bm{x}',\ell)\\
-		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \ell^2 \bm{\omega}^\top\bm{\omega} \right) \\
-		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\Big(-\frac{1}{2} \ell^2 (s_1, s_2)^\top\\
-		&\cdot& (s_1, s_2) \Big) \\
-		&=& \sqrt{2\pi}^D \cdot \ell^D \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\ell^2 s_i^2 \right) \\
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-SEPARABLE:} \\
-		\\
-		&&k(\bm{x}-\bm{x}',\bm{\ell})\\
-		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		&&S(\bm{\omega},\bm{\ell})= S(s_1,\ell_1)S(s_2,\ell_2)
-         \end{eqnarray*}
-       } \\
-       \vspace{-15mm}\\
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{7.2cm}|}{
-         \begin{eqnarray*}
-		k(\bm{\tau},\bm{\ell}) &=& k(\bm{x}-\bm{x}',\bm{\ell})\\
-		&=& \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		S(\bm{\omega},\ell) &=& \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{
-
-       } \\
-       \vspace{-10mm}\\
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&&= \mathrm{exp}\left(-\frac{1}{2} \frac{r_1^2}{\ell_1^2} \right) \mathrm{exp}\left(-\frac{1}{2} \frac{r_2^2}{\ell_2^2} \right) \\
-		&&= \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{D}\frac{r_i^2}{\ell_i^2} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{
-         \begin{eqnarray*}
-		&&S(s_1,\ell_1)S(s_2,\ell_2) \\
-		&&= \sqrt{2\pi} \cdot \ell_1 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_1^2 s_1^2 \right) \\
-		&& \times \sqrt{2\pi} \cdot \ell_2 \cdot \mathrm{exp}\left(-\frac{1}{2} \ell_2^2 s_2^2 \right)\\
-		&&= \sqrt{2\pi}^D \cdot \prod_{i=1}^D \ell_i \cdot \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^D \ell_i^2 s_i^2 \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{
-
-       } \\
-
-       \hline
-
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\newpage	
-\subsection{Matern($\nu=1/2$) covariance function ($k$) \textbf and spectral density ($S$)}
-\vspace{-2mm}
-\subsubsection{Using norm-L2 (Euclidean distance)}
-\vspace{-8mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-20mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L2} =|\bm{x}-\bm{x}'|_{L2} = \sqrt{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')} = \sqrt{(x_1-x_1')^2+(x_2-x_2')^2}  = \sqrt{r_1^2+r_2^2} \in \mathbb{R}; \hspace{2mm} k(\tau_{L2},\ell) = \mathrm{exp}\left(- \frac{\tau_{L2}}{\ell} \right)\\
-&& \omega_{L2}= \sqrt{s_1^2+s_2^2} \in \mathbb{R}; \hspace{2mm} S_{\nu}(\omega_{L2},\ell) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\omega_{L2}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{1/2}(\omega_{L2},\ell) = \frac{2^D\pi^{\frac{D}{2}}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{D+1}{2}}
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-       \hline
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		k(\tau_{L2},\ell) &=& k(|\bm{x}-\bm{x}'|_{L2},\ell)\\
-		 &=& \mathrm{exp}\left(- \frac{\sqrt{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')}}{\ell} \right)\\
-		&=& \mathrm{exp}\left(- \sqrt{\frac{(\bm{x}-\bm{x}')^\top(\bm{x}-\bm{x}')}{\ell^2}} \right)\\
-		&=& \mathrm{exp}\left(- \sqrt{\frac{\sum_{i=1}^{2}(x_i-x_i')^2}{\ell^2}} \right) \\
-		&=& \mathrm{exp}\left(-\sqrt{ \sum_{i=1}^{2}\frac{r_i^2}{\ell^2}} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{\small
-         \begin{eqnarray*}
-		D &=& 2\\
-		S_{1/2}(\omega_{L2},\ell) &=& \frac{2\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2\omega_{L2}^2\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2s_{1}^2+\ell^2s_{2}^2\right)^{-\frac{3}{2}} \\
-		\\
-		D &=& 3\\
-		S_{1/2}(\omega_{L2},\ell) &=& \frac{8\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-2} \\
-		&=& 8\pi\ell^3\left(1+\ell^2\omega_{L2}^2\right)^{-2} \\
-		&=& 8\pi\ell^3\left(1+\ell^2s_{1}^2+\ell^2s_{2}^2+\ell^2s_{3}^2\right)^{-2} 
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-NO SEPARABLE:} \\
-		\\
-		&&k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
-		 &&\neq k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		&&S_{1/2}(\omega_{L2},\ell)\neq S_{1/2}(s_1,\ell)S_{1/2}(s_2,\ell)
-         \end{eqnarray*}
-       }\\
-       \vspace{-10mm}\\
-        \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		k(\tau_{L2},\bm{\ell}) &=& k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
-		&=& \mathrm{exp}\left(-\sqrt{\sum_{i=1}^{2}\frac{r_i^2}{\ell_i^2}} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{\small
-         \begin{eqnarray*}
-		D &=& 2\\
-		S_{1/2}(\omega_{L2},\bm{\ell}) &=& 2\pi\ell_1\ell_2 \left( 1+ \sum_{i=1}^{D}\ell_i^2 s_i^2 \right)^{-\frac{3}{2}}\\
-		S_{1/2}(\omega_{L2},\bm{\ell}) &=& 2\pi\prod_{i=1}^{D}\ell_i \left( 1+ \sum_{i=1}^{D}\ell_i^2 s_i^2 \right)^{-\frac{3}{2}}
-         \end{eqnarray*}
-        }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-
-       } \\  
-       \vspace{-8mm}\\
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&=& \mathrm{exp}\left(- \frac{r_1}{\ell_1} \right) \mathrm{exp}\left(- \frac{r_2}{\ell_2} \right) \\
-		&=& \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell_i} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{8.2cm}|}{\small
-         \begin{eqnarray*}
-         &&S_{1/2}(s_1,\ell_1)S_{1/2}(s_2,\ell_2) \\
-		&=& \frac{2}{\ell_1}\left(\frac{1}{\ell_1^2}+s_1^2 \right)^{-1} \cdot \frac{2}{\ell_2}\left(\frac{1}{\ell_2^2}+s_2^2 \right)^{-1} \\
-		%
-		&=& 4\ell_1\ell_2\left(1+\ell_1^2s_1^2 \right)^{-1} \left(1+\ell_2^2s_2^2 \right)^{-1} \\
-         \end{eqnarray*}        
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-
-       } \\
-
-       \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\newpage
-\subsubsection{Using norm-L1}
-\vspace{-5mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-20mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L1} =|\bm{x}-\bm{x}'|_{L1} = |x_1-x_1'|+|x_2-x_2'| = r_1+r_2 \in \mathbb{R}; \hspace{2mm} k(\tau_{L1},\ell) = \mathrm{exp}\left(- \frac{\tau_{L1}}{\ell} \right)\\
-&& \omega_{L1}= s_1+s_2 \in \mathbb{R}; \hspace{2mm} S_{\nu}(\omega_{L2},\ell) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\omega_{L2}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{1/2}(\omega_{L2},\ell) = \frac{2^D\pi^{\frac{D}{2}}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{D+1}{2}}
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-       \hline
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-		k(\tau_{L1},\ell) &=& k(|\bm{x}-\bm{x}'|_{L1},\ell)\\
-		 &=& \mathrm{exp}\left(- \frac{\tau}{\ell} \right) \\
-		&=& \mathrm{exp}\left(- \frac{r_1 + r_2}{\ell} \right)\\
-		&=& \mathrm{exp}\left(- \sum_{i=1}^{D}\frac{r_i}{\ell} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{9.7cm}|}{\small
-         \begin{eqnarray*}
-		D &=& 2\\
-		S_{1/2}(\omega_{L1},\ell) &=& \frac{2\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L1}^2 \right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2\omega_{L1}^2\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2(s_{1}+s_{2})(s_{1}+s_{2})\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+(\ell s_{1}+\ell s_{2})(\ell s_{1}+\ell s_{2})\right)^{-\frac{3}{2}} \\
-		\\
-		D &=& 3\\
-		S_{1/2}(\omega_{L1},\ell) &=& \frac{8\pi}{\ell}\left(\frac{1}{\ell^2}+\omega_{L1}^2 \right)^{-2} \\
-		&=& 8\pi\ell^3\left(1+\ell^2\omega_{L1}^2\right)^{-2} \\
-		&=& 8\pi\ell^3\left(1+\ell^2(s_{1}+s_{2}+s_{3})(s_{1}+s_{2}+s_{3})\right)^{-2} \\
-		&=& 8\pi\ell^3\left(1+(\ell s_{1}+\ell s_{2}+\ell s_{3})(\ell s_{1}+\ell s_{2}+\ell s_{3})\right)^{-2} 
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-SEPARABLE:} \\
-		\\
-		&&k(|\bm{x}-\bm{x}'|_{L1},\bm{\ell})\\
-		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		 \\
-		 &&\text{IT SHOULD BE SEPARABLE}\\
-		 &&\text{IN THE SPECTRAL DENSITY} \\
-		 &&\text{AS WELL?}\\
-		&&S_{1/2}(\omega_{L1},\ell) \text{ should be equal to }\\
-		 &&S_{1/2}(s_1,\ell)S_{1/2}(s_2,\ell)
-         \end{eqnarray*}
-       }\\
-       \vspace{-10mm}\\
-        \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-		k(\tau_{L1},\bm{\ell}) &=& \mathrm{exp}\left(- \sum_{i=1}^{D}\frac{r_i}{\ell_i} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{9.7cm}|}{\small
-         \begin{eqnarray*}
-         D &=& 2\\
-		S_{1/2}(\omega_{L1},\bm{\ell}) &=& 2\pi\ell_1\ell_2 \left(1+(\ell_1 s_{1}+\ell_2 s_{2})(\ell_1 s_{1}+\ell_2 s_{2})\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell_1\ell_2 \left( 1+ (\ell_1 s_{1}+\ell_2 s_{2})(\ell_1 s_{1}+\ell_2 s_{2}) \right)^{-\frac{3}{2}}
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.7cm}|}{\small
-
-       } \\
-
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&&= \mathrm{exp}\left(- \frac{r_1}{\ell_1} \right) \mathrm{exp}\left(- \frac{r_2}{\ell_2} \right) \\
-		&&= \mathrm{exp}\left(- \sum_{i=1}^{D}\frac{r_i}{\ell_i} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{9.7cm}|}{\small
-         \begin{eqnarray*}
-         &&S_{1/2}(s_1,\ell_1)S_{1/2}(s_2,\ell_2) \\
-		&=& \frac{2}{\ell_1}\left(\frac{1}{\ell_1^2}+s_1^2 \right)^{-1} \cdot \frac{2}{\ell_2}\left(\frac{1}{\ell_2^2}+s_2^2 \right)^{-1} \\
-		%
-		&=& 4\ell_1\ell_2\left(1+\ell_1^2s_1^2 \right)^{-1} \left(1+\ell_2^2s_2^2 \right)^{-1} \\
-         \end{eqnarray*}      
-       }
-
-       & \multicolumn{1}{|p{6.7cm}|}{\small
-
-       } \\
-
-       \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\newpage
-\subsubsection{Using  the vector difference of inputs}
-\vspace{-5mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-20mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\bm{\tau} =\bm{x}-\bm{x}' = (x_1-x_1',x_2-x_2') = (r_1,r_2) \in \mathbb{R}^2; \hspace{2mm} k(\bm{\tau},\ell) = \mathrm{exp}\left(- \frac{\bm{\tau}}{\ell} \right)\\
-&& \bm{\omega}= (s_1,s_2) \in \mathbb{R}^2; \hspace{2mm}S_{\nu}(\bm{\omega},\ell) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\bm{\omega}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{1/2}(\bm{\omega},\ell) = \frac{2^D\pi^{\frac{D}{2}}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^2 \right)^{-\frac{D+1}{2}}
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-       \hline
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-		k(\bm{\tau},\ell) &=& k(\bm{x}-\bm{x}',\ell)\\
-		 &=& \mathrm{exp}\left(- \frac{\bm{\tau}}{\ell} \right) \\
-		&=& \mathrm{exp}\left(- \frac{(r_1, r_2)}{\ell} \right)\\
-		&&\text{(using dot product?)}\\
-		&=& \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell} \right)
-		\end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{9.7cm}|}{\small
-         \begin{eqnarray*}
-		D &=& 2\\
-		S_{1/2}(\bm{\omega},\ell) &=& \frac{2\pi}{\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^2 \right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2\bm{\omega}^2\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2(s_{1},s_{2})^\top(s_{1},s_{2})\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\ell^2(s_{1}^2+s_{2}^2)\right)^{-\frac{3}{2}} \\
-		&=& 2\pi\ell^2\left(1+\sum_{i=1}^{D}\ell^2s_{i}^2\right)^{-\frac{3}{2}} \\
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-SEPARABLE:} \\
-		&&k(\bm{x}-\bm{x}',\bm{\ell})\\
-		 &&= k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		 &&\text{IT SHOULD BE SEPARABLE}\\
-		 &&\text{IN THE SPECTRAL DENSITY} \\
-		 &&\text{AS WELL?}\\
-		&&S_{1/2}(\bm{\omega},\ell) \text{ should be equal to }\\
-		 &&S_{1/2}(s_1,\ell)S_{1/2}(s_2,\ell)
-         \end{eqnarray*}
-       }\\
-       \vspace{-10mm}\\
-        \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-		k(\bm{\tau},\bm{\ell}) &=& \mathrm{exp}\left(- \frac{\bm{\tau}}{\bm{\ell}} \right) \\
-		&=& \mathrm{exp}\left(- \frac{(r_1, r_2)}{(\ell_1,\ell_2)} \right)\\
-		&&\text{(using dot product?)}\\
-		&=& \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell_i} \right) 
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{9.7cm}|}{\small
-         \begin{eqnarray*}
-		S_{1/2}(\bm{\omega},\bm{\ell}) &=& 2\pi\ell_1\ell_2 \left(1+\sum_{i=1}^{D}\ell_i^2s_{i}^2\right)^{-\frac{3}{2}}
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{5.7cm}|}{\small
-
-       } \\
-
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{5.7cm}|}{\small
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&&= \mathrm{exp}\left(- \frac{r_1}{\ell_1} \right) \mathrm{exp}\left(- \frac{r_2}{\ell_2} \right) \\
-		&&= \mathrm{exp}\left(- \sum_{i=1}^{2}\frac{r_i}{\ell_i} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{9.7cm}|}{\small
-         \begin{eqnarray*}
-         &&S_{1/2}(s_1,\ell_1)S_{1/2}(s_2,\ell_2) \\
-		&=& \frac{2}{\ell_1}\left(\frac{1}{\ell_1^2}+s_1^2 \right)^{-1} \cdot \frac{2}{\ell_2}\left(\frac{1}{\ell_2^2}+s_2^2 \right)^{-1} \\
-		%
-		&=& 4\ell_1\ell_2\left(1+\ell_1^2s_1^2 \right)^{-1} \left(1+\ell_2^2s_2^2 \right)^{-1} \\
-         \end{eqnarray*}      
-       }
-
-       & \multicolumn{1}{|p{6.7cm}|}{\small
-
-       } \\
-
-       \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\newpage	
-\subsection{Matern($\nu=3/2$) covariance function ($k$) \textbf and spectral density ($S$)}
-\vspace{-2mm}
-\subsubsection{Using norm-L2 (Euclidean distance)}
-\vspace{-8mm}
-
-\begin{table}[H]
-\small
-\begin{eqnarray*}
-\vspace{-20mm}
-&&\bm{x}, \bm{x}' \in \mathbb{R}^{D=2}; \hspace{2mm}\tau_{L2} =|\bm{x}-\bm{x}'|_{L2} = \sqrt{(\bm{x}-\bm{x}')(\bm{x}-\bm{x}')} = \sqrt{(x_1-x_1')^2+(x_2-x_2')^2}  = \sqrt{r_1^2+r_2^2} \in \mathbb{R}; \hspace{2mm} k(\tau_{L2},\ell) = \left(1+\frac{\sqrt{3}\tau}{\ell}\right) \mathrm{exp}\left(- \frac{\sqrt{3}\tau}{\ell} \right)\\
-&& \omega_{L2}= \sqrt{s_1^2+s_2^2} \in \mathbb{R}; \hspace{2mm} S_{\nu}(\omega_{L2}) = \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+\omega_{L2}^2 \right)^{-(\nu+D/2)}; \hspace{2mm} S_{3/2}(\omega_{L2}) = \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega_{L2}^2 \right)^{-\frac{D+3}{2}}
-\end{eqnarray*}
-\normalsize
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-       \hline
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\ell \in \mathbb{R}$
-       }
-
-        & \multicolumn{1}{|p{8.2cm}|}{\small
-         \begin{eqnarray*}
-		&&k(\tau_{L2},\ell) = k(|\bm{x}-\bm{x}'|_{L2},\ell)\\
-		&&= \left(1+\frac{\sqrt{3}\sqrt{\sum_{i=1}^{2}r_i^2}}{\ell} \right)\mathrm{exp}\left(-\frac{\sqrt{3}\sqrt{\sum_{i=1}^{2}r_i^2}}{\ell} \right)\\
-		&&= \left(1+\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell^2}} \right)\mathrm{exp}\left(-\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell^2}} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{7.5cm}|}{\small
-         \begin{eqnarray*}
-		D &=& 2\\
-		\\
-		S_{3/2}(\omega_{L2},\ell) &=& \frac{6\pi\sqrt{3}^3}{\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{5}{2}} \\
-		&=& 6\pi\sqrt{3}^3\ell^2\left(3+\ell^2\omega^2 \right)^{-\frac{5}{2}} \\
-		&=& 6\pi\sqrt{3}^3\ell^2\left(3+\ell^2s_1^2+\ell^2s_2^2 \right)^{-\frac{5}{2}} \\
-		\\
-		D &=& 3\\
-		S_{3/2}(\omega_{L2},\ell) &=& 32\pi\sqrt{3}^3\ell^3\left(3+\ell^2\omega^2 \right)^{-3}
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-         \begin{eqnarray*}
-        &&\text{-ISOTROPIC}\\
-        \\
-		&&\text{-NO SEPARABLE:} \\
-		\\
-		&&k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
-		 &&\neq k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2)\\
-		 \\
-		&&S_{3/2}(\omega_{L2},\ell)\neq S_{3/2}(s_1,\ell)S_{1/2}(s_2,\ell)
-         \end{eqnarray*} 
-       }\\
-       \vspace{-10mm}\\
-        \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-       }
-
-         & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		&&k(\tau_{L2},\bm{\ell}) = k(|\bm{x}-\bm{x}'|_{L2},\bm{\ell})\\
-		&&= \left(1+\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell_i^2}} \right)\mathrm{exp}\left(-\sqrt{\frac{\sum_{i=1}^{2}3r_i^2}{\ell_i^2}} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		S_{3/2}(\omega_{L2},\bm{\ell}) &=& 6\pi\sqrt{3}^3\ell_1\ell_2\left(3+\ell_1^2s_1^2+\ell_2^2s_2^2 \right)^{-\frac{5}{2}} \\
-		&=& 6\pi\sqrt{3}^3\prod_{i=1}^{D}\ell_i\left(3+\sum_{i=1}^{D}\ell_i^2s_i^2 \right)^{-\frac{5}{2}} \\
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-
-       } \\  
-       \vspace{-10mm}\\
-       \cline{1-3}
-
-       \multicolumn{1}{|p{1.5cm}|}{
-       \vspace{1mm}
-       $\bm{\ell} \in \mathbb{R}^2$
-
-       {Separable kernel} 
-       }
-
-        & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-		&&k(|x_1-x_1'|,\ell_1)k(|x_2-x_2'|,\ell_2) \\
-		&&= \left(1+\sqrt{3}\frac{r_1}{\ell_1}\right)\mathrm{exp}\left(-\sqrt{3}\frac{r_1}{\ell_1} \right)\\
-		&&\cdot \left(1+\sqrt{3}\frac{r_2}{\ell_2}\right)\mathrm{exp}\left(-\sqrt{3}\frac{r_2}{\ell_2} \right) \\
-		&&= \left(1+\sqrt{\frac{3r_1^2}{\ell_1^2}}\right)\left(1+\sqrt{\frac{3r_2^2}{\ell_2^2}}\right)\mathrm{exp}\left(-\sum_{i=1}^{2}\sqrt{\frac{3r_i^2}{\ell_i^2}} \right) \\
-		&&= \left(1+\sum_{i=1}^{2}\sqrt{\frac{3r_i^2}{\ell_i^2}}+\frac{3r_1^2r_2^2}{\ell_1^2\ell_2^2}\right)\mathrm{exp}\left(-\sum_{i=1}^{2}\sqrt{\frac{3r_i^2}{\ell_i^2}} \right)
-         \end{eqnarray*}
-       }
-
-       & \multicolumn{1}{|p{7.2cm}|}{\small
-         \begin{eqnarray*}
-         &&S_{3/2}(s_1,\ell_1)S_{3/2}(s_2,\ell_2) \\
-		&=& \frac{4\sqrt{3}^3}{\ell_1^3}\left(\frac{3}{\ell_1^2}+s_1^2 \right)^{-3} \cdot \frac{4\sqrt{3}^3}{\ell_2^3}\left(\frac{3}{\ell_2^2}+s_2^2 \right)^{-3}\\ 
-		&=& 4^2\sqrt{3}^6\bm{\ell}^\top\bm{\ell}\left(3+\ell_1^2s_1^2 \right)^{-2} \left(3+\ell_2^2s_2^2 \right)^{-2}\\        
-		\end{eqnarray*}        
-       }
-
-       & \multicolumn{1}{|p{6.2cm}|}{\small
-
-       } \\
-
-       \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\end{landscape}
-
-\newpage
-\vspace{2mm}
+\noindent where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\dots,2J+1$.
+The cosine $\text{cos}(j\omega_0 x)$ and sinus $\text{sin}(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.
+The only dependence on the hyperparameter $\ell$ is through the coefficients $\tilde{q}_j$, which are $J$-dimensional.
+The computational cost of this approximation scales as $O\big(n(2J+1) + (2J+1)\big)$, where $n$ is the number of observations and $J$ the number of cosine terms.
+
+The parameterization in equation (\ref{ch5_eq_f_period}) is naturally in the non-centered
+parameterization form with independent prior distribution on
+$\beta_j$, which makes the posterior inference easier.
+
 \section*{Acknowledgment}
 
-\bibliographystyle{biom}
 \bibliography{references}
@@
 \end{document}
-
-###########
-#######
-Next we show a little example of how to construct the new set of eigenfunctions and eigenvalues with input vector $\bm{x}=\{\bm{x}_1,\bm{x}_2\}$ in a two-dimensional input space ($D=2$) and three eigenfunctions and eigenvalues $(J=3)$ for every dimension.
-As a result we obtain sets of $J^D=3^2=9$ new eigenfunctions $\phi^{\ast}$ and eigenvalues $\lambda^{\ast}$.
-The eigenfunctions $\phi^{\ast}$ are the product combination of the unidimensional eigenfunctions $\phi$, and eigenvalues $\lambda^{\ast}$ is the $D$-vector of unidimensional eigenvalues $\lambda$:
-
-\begin{eqnarray}
-\phi^{\ast}_1(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_1 = \{\lambda_{1}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_2(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_2 = \{\lambda_{1}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_3(\bm{x}) = \phi_{1}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_3 = \{\lambda_{1}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\} &=& \{ \left(\tfrac{1\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_4(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_4 = \{\lambda_{2}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_5(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_5 = \{\lambda_{2}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_6(\bm{x}) = \phi_{2}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_6 = \{\lambda_{2}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\} &=& \{ \left(\tfrac{2\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_7(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{1}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_1}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_7 = \{\lambda_{3}(\bm{x}_1), \lambda_{1}(\bm{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{1\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_8(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{2}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_2}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_8 = \{\lambda_{3}(\bm{x}_1), \lambda_{2}(\bm{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{2\pi}{2L_2}\right)^2 \} \nonumber \\
-%
-\phi^{\ast}_9(\bm{x}) = \phi_{3}(\bm{x}_1) \cdot \phi_{3}(\bm{x}_2) &=& \sqrt{\tfrac{1}{L_1}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_1+L_1)\right) \cdot \sqrt{\tfrac{1}{L_2}} \text{sin}\left(\sqrt{\lambda_3}(\bm{x}_2+L_2)\right) \nonumber \\
-%
-\bm{\lambda}^{\ast}_9 = \{\lambda_{3}(\bm{x}_1), \lambda_{3}(\bm{x}_2)\} &=& \{ \left(\tfrac{3\pi}{2L_1}\right)^2, \left(\tfrac{3\pi}{2L_2}\right)^2 \} \nonumber
-\end{eqnarray}

### 2020-03-14 Paper/online_appendices/online_appendices.tex
update in manuscript_v2 and online appendice

\documentclass[onecolumn,a4paper,11pt]{article}
\voffset -1cm
\hoffset -1cm
\textheight 22.3cm
\textwidth 15.0cm

\usepackage{blindtext}
\usepackage{titlesec}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage{microtype}

\usepackage{lscape}

\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}

\usepackage{epstopdf}

\usepackage[numbers,sort&compress]{natbib}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{bm}

\usepackage[titletoc]{appendix}

\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\bibpunct[, ]{[}{]}{,}{}{,}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}

\usepackage{listings}
\usepackage{slashbox}
\usepackage{caption}
\captionsetup{font=footnotesize}

\bibliographystyle{authordate1}

\title{Online appendices}

\begin{document}

\maketitle

\tableofcontents

\appendix

\section{Univariate case studies}

\subsection{Same-sex marriage data}\label{ch5_sec_studycaseII}
This data set relates the proportion of support for same-sex marriage to the age. The data consists of 74 observations of the amount of people $y_i$ supporting same-sex marriage from a population $n_i$ per age group $i$ ($i=1,\dots,74$). The observational model is a binomial model with parameters population $n_i$ and probability of supporting same-sex marriage $p_i$ per age group $i$,
%
\begin{equation*}
y_i \sim \text{Binomial}(p_i, n_i).
\end{equation*}

\noindent The population per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ or mean number of support people per age group. Probabilities $\bm{p}=(p_1,\dots,p_{74})$ are modeled by a GP function $f:{\rm I\!R} \to {\rm I\!R}$ with a squared exponential covariance function $k$, as a function of age input values $\bm{x}=(x_1,\dots,x_{74})$, and through the {\it logit} function as a link function,
%
\begin{align*} \label{ch5_eq_gpprior_gay}
p_i &= \text{logit}(f(x_i)) \nonumber \\
f(x) &\sim \mathcal{GP}(0, k(x,x', \theta)).
\end{align*}

\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,74$. In the HSGP model, the function $f(x)$ is approximated as in equation (\ref{eq_approxf}), with the squared exponential spectral density as in equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).

In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}. A Bayesian approach is used to fit this splines model using the R-package \textit{brms} \citep{burkner2017brms}.

Figure \ref{ch5_fig12_Posteriors_gaydata} shows the posterior mean predictive distributions of the three models, the regular GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the splines model with 20 knots. Sample observations are plotted as circles in the figure, and the out-of-sample observations, which have been used for testing, are plotted as crosses.

\begin{figure}
\centering
\includegraphics[scale=0.50]{ch5_fig12_Posteriors_gaydata.pdf}
\caption{Posterior mean predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dashed lines.}
  \label{ch5_fig12_Posteriors_gaydata}
\end{figure}

For the HSGP model, different models with different number of basis functions and boundary factor have been fitted. The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data. The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.

Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models, computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model. We can see how the splines models do not extrapolate data properly.

Figure \ref{ch5_fig14_1_time_gaydata} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the splines model. The computational times is represented in the y-axis which is on a logarithmic scale. The HSGP model is on average roughly 10 times faster than the regular GP, in this particular case.

The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found in:
%
\begin{lstlisting}[breaklines]
https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-V_Same-sex-marriage
\end{lstlisting}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig13_MSE_train_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig13_MSE_pred_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 119mm 0mm, clip]{ch5_fig13_legend_gaydata.pdf}}
\caption{Root mean square error (RMSE) of the HSGP model, computed against the regular GP model, as a function of the number of basis functions $m$ and boundary factor $c$. RMSE for sample data (left) and RMSE for out-of-sample data (right).}
  \label{ch5_fig13_MSE_train_BF_gaydata}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig14_MSE_train_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig14_MSE_pred_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 27mm 35mm 112mm 0mm, clip]{ch5_fig14_legend_gaydata.pdf}}
\caption{Root mean square error (RMSE) of the different methods, regular GP, HSGP and splines models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots for the splines model. RMSE for sample data (left) and RMSE for out-of-sample data (right).}
  \label{ch5_fig14_MSE_train_gaydata}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig14_1_time_gaydata.pdf}}
\caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$ and knots. The y-axis is in a logarithmic scale.}
  \label{ch5_fig14_1_time_gaydata}
\end{figure}

\section{Multivariate case studies}

\subsection{2D Simulated data}

This example consists of a simulated dataset with $n=120$ ($i=1,\dots,n$) single draws from a Gaussian process prior with two input dimensions ($D=2$). A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$, for the first dimension, and $\ell_2=0.35$, for the second dimension, is used. The corresponding matrix of input values is $X=[\bm{x}_1 \; \bm{x}_2 \; \cdots \; \bm{x}_{120}]^\top \in {\rm I\!R}^{120\times 2}$ with $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$. Gaussian noise $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^{120}$.

The regular GP model over the outcome variable $\bm{y}$ and matrix of inputs $X \in {\rm I\!R}^{120 \times 2}$, can be written as follows,
%
\begin{eqnarray*}\label{ch5_eq_latentgp_simudata2}
\begin{split}
\bm{y} &= \bm{f} + \bm{\epsilon} \\
\bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x}, \bm{x}', \theta)),
\end{split}
\end{eqnarray*}

\noindent where $I$ represents the identity matrix, $\bm{e}$ the noise term, and $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^{120}$ represents the underlying function values to the noisy observations $\bm{y}$. The function $f:{\rm I\!R}^2 \to {\rm I\!R}$ is a GP prior with a multivariate squared exponential covariance function $k$. Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,120$. The covariance function $k$ depends on the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell_1,\ell_2\}$. The hyperparameters $\alpha$, $\ell_1$ and $\ell_2$ represent the marginal variance and lengthscales for first and second dimensions, respectively, of the GP process.

The marginalized form, by integrating out the latent values $\bm{f}$, of the previous latent GP model results:
%
\begin{equation*}\label{ch5_eq_marginalizedgp_simudata2}
\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ).
\end{equation*}

In the HSGP model with $2$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^2$, is approximated as in equation (\ref{eq_approxf_multi}), with the $2$-dimensional squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $2$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.

In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2015package}. A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.

Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset. Sample observations are also plotted in the plots as circles. For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions. A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ were used. For the splines model, 40 knots in each dimension were used.

Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the data-generating function and the GP, HSGP and splines models, respectively.

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.40, trim = 0mm 33mm 35mm 20mm, clip]{ch5_fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.40, trim = 19mm 33mm 35mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 18mm 0mm 10mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}}\\ 
\includegraphics[scale=0.40, trim = 0mm 18mm 35mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.40, trim = 19mm 18mm 35mm 20mm, clip]{ch5_fig15_spfun_exII.pdf}
\end{tabular}
\end{center}
\caption{(a) Data-generating function. (b) Mean posterior predictive function of the GP model. (c) Mean posterior predictive function of the HSGP model. (d) Mean posterior predictive function of the splines model. Sample points are plotted as circles}
  \label{ch5_fig15_Posteriors_exII}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorHSGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorSP_exII.pdf}}
\subfigure{\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models. Sample points are plotted as circles.}
  \label{ch5_fig16_errors_exII}
\end{figure}

In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions, for the HSGP model, and different number of knots, for the splines model, have been fitted. In all models, the same number of basis functions and knots per dimension were used. Figure \ref{ch5_fig17_RMSE_exII}-(left) shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, and the number of univariate basis functions $m$, for the HSGP model, and knots, for the splines model. From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-(left), it can be seen a close approximation of the HSGP model to the regular GP model. However, the performance of the splines model is significantly worse. Figure \ref{ch5_fig17_RMSE_exII}-(right) shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots. Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computational time.

The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found in:
%
\begin{lstlisting}[breaklines]
https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-VI_Simulated-data_2D
\end{lstlisting}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 113mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
  \label{ch5_fig17_RMSE_exII}
\end{figure}

\subsection{Diabetes data}\label{ch5_sec_bf_caseV}
The next example presents an epidemiological study of diabetes disease. The study aims to relate the probability of suffering from diabetes to some risk factors. The data contains $n=392$ individuals ($i=1,\dots,n$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed. The matrix $X=[\bm{x}_1\;\bm{x}_2\;\cdots\;\bm{x}_n]^\top \in {\rm I\!R}^{n\times 4}$,  with $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$, contains the risk factors, \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$. The observational model is a Bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
%
\begin{equation*}
y_i \sim \text{Bernoulli}(p_i).
\end{equation*}

\noindent The goal is to estimate the probability $p_i$ as a function of the risk factors, which function $f(\cdot):{\rm I\!R}^{4} \to {\rm I\!R}$ is modeled as a Gaussian process with a multivariate squared exponential covariance function $k$ depending on the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
%
\begin{eqnarray*} \label{ch5_eq_gpprior_gay}
p_i &=& \text{logit}(f(\bm{x}_i)) \nonumber \\
f(\bm{x}) &\sim& \mathcal{GP}(0, k(\bm{x},\bm{x}, \theta).
\end{eqnarray*}

\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^n$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$. The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process. Notice that a scalar lengthscale is considered in the multivariate covariance function.

In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$ evaluated at input vector $\bm{x} \in {\rm I\!R}^D$ is approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}) and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.

In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional spline-based model is also fitted using a cubic spline basis penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized} and implemented in the R-package \textit{mgcv}. A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.

Figure \ref{ch5_fig18_gpfun_diabetes} shows the mean posterior predictions of probabilities ($p_i$) of the three models, the regular GP, the HSGP and the spline, fitted over the dataset with the 2 input dimensions \textit{Glucose} and \textit{Pregnancy} ($D=2$). The binary observations $y_i$ are also plotted in the plots as colored points. For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions. A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used. For the spline model, 20 knots per dimension were used.
%
\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.37, trim = 5mm 23mm 39mm 20mm, clip]{ch5_fig18_gpfun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.37, trim = 20mm 23mm 39mm 20mm, clip]{ch5_fig18_bffun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.37, trim = 20mm 23mm 39mm 20mm, clip]{ch5_fig18_spfun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.30, trim = 150mm 4mm 2mm 15mm, clip]{ch5_fig18_legend_diabetes.pdf}}
\caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the spline (SP) model. Samples observations of suffering (red points) and not suffering (blue points) form the disease are plotted.}
  \label{ch5_fig18_gpfun_diabetes}
\end{figure}

In order to assess the performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factor for the HSGP model and different number of knots for the spline model have been fitted. In all models, the same boundary factor, number of basis functions and knots per dimension were used. Figure \ref{ch5_fig19_ELPD_diabetes} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) as a function of the boundary factor $c$ and the number of univariate basis functions $m$, for the HSGP model, and knots, for the spline model. The ELPD is computed over the actual observations by cross-validation. Basically, with slightly differences, all models show similar performances, due to the fact that the process is very smooth with a relatively very large lengthscale estimate $\ell=4.51$. Even though, a slight pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (Figure \ref{fig6_lscale_vs_J_vs_c}).
%
\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig19_ELPD_diabetes.pdf}}
\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$, for the HSGP model, and knots, for the spline model.}
  \label{ch5_fig19_ELPD_diabetes}
\end{figure}

Figure \ref{ch5_fig20_time2D_diabetes} shows the computational times of the different models, regular GP, HSGP and spline, fitted over the dataset, with 2 input dimensions, 3 input dimensions and 4 input dimensions, as a function of the boundary factor $c$ and number of univariate basis functions $m$ and knots. We can appreciate the significant increase of computational time with higher dimensions for the HSGP and spline models. This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid a excessive computational time, especially in high input dimensionality. It is interesting to be noticed that considering more than 10 knots per dimension in the spline model with 3D is not allowed for an amount of 392 observations. Similarly, just the computation of the input data for the spline model in a 4D input space is computationally very expensive.

The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found in:
%
\begin{lstlisting}[breaklines]
https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-III_Diabetes-data
\end{lstlisting}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 10mm 0mm, clip]{ch5_fig20_time2D_diabetes_2.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 30.2mm 0mm 10mm 0mm, clip]{ch5_fig20_time3D_diabetes_2.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 30.2mm 0mm 10mm 0mm, clip]{ch5_fig20_time4D_diabetes_2.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 28mm 50mm 112mm 10mm, clip]{ch5_fig20_legend_diabetes_2.pdf}}
\caption{Time of computation in seconds per iteration (iteration of the HMC sampling method) of the different models fitted over the dataset, with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$ and number of basis functions $m$, for the HSGP model, and knots, for the spline model.}
  \label{ch5_fig20_time2D_diabetes}
\end{figure}

\subsection{Spatio-temporal land-use classification}\label{ch5_sec_application_caseII}

The next example presents an spatio-temporal classification in land-use of plots between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits. A sampling set consists of $N=200$ plots with known class. The data is recorded in a time series of $T=5$ years (2006, 2008, 2010, 2012, 2015) within the period. The class of each parcel $i$ and time $t$ is stored by a categorical variable $y_{it}$ representing the $K=5$ different possible classes ($k=1,\dots,K$): $k=1$, adult independent citrus fruits; $k=2$, aligned citrus fruits; $k=3$, irregular citrus fruits; $k=4$, abandoned citrus fruits; $k=5$, young citrus fruits.

A bunch of 52 characteristic variables was available for every parcel and time. These variables were computed from satellite color images and cadastral map by using the software FETEX for automatic descriptive feature extraction from image-objects \citep{ruiz2011feature}. These variables concern spectral intensities and empirical semivariogram of the pixels within a plot, as well as descriptive statistics of the shape of the plots.

Due to the fact that 52 input variables are too high-dimensional for a multivariate HSGP model, which computational cost scales as $O(n m^D+m^D)$, with $m$ the number of basis functions and $D$ the number of input variables, the multivariate HSGP model will be formulated as an additive HSGP model.

As it is known, the computational demand of a multivariate HSGP model component increases quickly with the number of dimensions, so we should avoid high-dimensional HSGP components in the additive model. Original input variables are highly correlated, which would imply the use of high-order interaction components in the additive model to achieve accurate model performance. Therefore, instead of using the original variables as inputs, we use their principal components, which are expect to be linearly uncorrelated. Using the principal components as inputs helps, in principle, not to have to use as many high-order interaction components in the additive model.

The principal components (PCs) will be used jointly with the time variable as inputs in the classifying additive HSGP model. Let's denote the matrix $X=[\bm{x}_{11}\, \cdots \, \bm{x}_{it} \, \cdots \, \bm{x}_{NT}]^\top \in {\rm I\!R}^{NT\times D}$, which contains the input vectors $\bm{x}_{it} \in {\rm I\!R}^D$, $D=53$ ($52 \text{ PCs plus time}$) for the spatio-temporal observations (plots $i=1,\dots,200$, and times $t=1,\dots,5$).

The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{it}=(p_{it,1},\dots,p_{it,K})$, where $p_{it,k}$ is the probability of belonging to class $k$ per parcel $i$ and time $t$,
%
\begin{equation*}
y_{it} \sim \text{multinomial}(\bm{p}_{it}).
\end{equation*}

\noindent The goal is to estimate the vector of probabilities $\bm{p}_{it}$ as a function of the predictors, which is a multivariate function $f(\bm{x}_{it}):{\rm I\!R}^{D} \rightarrow {\rm I\!R}^{K}$,
$$
f(\bm{x}_{it})=\big(f_1(\bm{x}_{it}),\dots,f_K(\bm{x}_{it}) \big),
$$

\noindent which is related to the vector of probabilities $\bm{p}_{it}$ through the 'softmax' link function (equation \ref{ch2_eq_softmax} in chapter \ref{ch2_bayes}):
%
\begin{eqnarray*}
\bm{p}_{it} = \text{softmax}(f(\bm{x}_{it})).
\end{eqnarray*}

Each individual function $f_k(\bm{x})$ is modeled as a first-order additive model plus the second-order additive effects between time input variable ($x^D$) and all the other inputs as follows: 
%
\begin{eqnarray} \label{ch5_eq_f_parcels}
f_k(\bm{x}) = \sum_{d=1}^{D} g_d(x^d) + \sum_{d=1}^{D-1} h_{d,D}(x^d,x^D).
%\mathcal{HSGP}(x^d, S(\bm{\lambda}), \theta_d),
\end{eqnarray}

\noindent The first-order components $\{g_d(x^d)\}_{d=1}^D$ in equation (\ref{ch5_eq_f_parcels}) are modeled as unidimensional HSGP models:
%
\begin{eqnarray*}
g_d(x^d) \sim \mathcal{HSGP}(x^d, S, \theta_{d,1}).
\end{eqnarray*}

\noindent In the HSGP model, a first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in equation (\ref{eq_approxf}) with the squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}) and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.

The second-order components $\{h_{d,D}(x^d,x^D)\}_{d=1}^{D-1}$ in equation (\ref{ch5_eq_f_parcels}) are modeled as two-dimensional HSGP models:
%
\begin{eqnarray*}
h_{d,D}(x^d,x^D) \sim \mathcal{HSGP}(x^d,x^D, S, \theta_{d,D}).
\end{eqnarray*}

\noindent In the HSGP model, a second-order component $h_{d,D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in equation (\ref{eq_approxf_multi}) with the two-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}) and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.

The vector of hyperparameters $\theta_{d,1}=(\alpha_{d,1},\ell_{d,1})$ contains the marginal variance $\alpha_{d,1}$ and lengthscale $\ell_{d,1}$ of the $g_d(x^d)$ model component. And, the vector of hyperparameters $\theta_{d,D}=(\alpha_{d,D},\ell_{d,D})$ contains the marginal variance $\alpha_{d,D}$ and lengthscale $\ell_{d,D}$ of the $h_{d,D}(x^d,x^D)$ model component.

For the first-order components $g_d(x^d)$, $m=15$ basis functions and a boundary factor $c=2.5$ were used. For the second-order components $h_{d,D}(x^d,x^D)$, $m_1=15$ and $m_2=15$ basis functions for each dimension, respectively, were used, which lead to a total of 225 multivariate basis functions. A boundary factor for each dimension $c_1=2.5$ and $c_2=2.5$ were used. All the input variables were previously standardized.

In the case of the first-order components, the normalized lengthscale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$, $c$. Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.

For the second-order components, the relationships between the number of basis functions, the boundary factor and the lengthscale is not available for the multivariate case. However, we can approximately analyze the lengthscale estimates of the second-order HSGP components analyzing each dimension separately as unidimensional HSGP models.

Table \ref{ch5_tab_parcels} shows the confusion matrix after fitting the model following a $Q$-fold cross-validation procedure, with $Q=100$, over the training data. Thus, every fold contains $10$ observations. The confusion matrix evaluates the rate of misclassification per class. Columns represent the true classes and rows represent the estimated classes. The values within the matrix correspond to the number of items that fall into every cell. 
The marginals of the columns (true classes) represent the percentage of misclassified items in relation to the 'truth', commonly known as the \textit{omission error}. And the marginals of the rows (estimated classes) represent the percentage of misclassified items in relation to the estimates (classifier), commonly known as the \textit{commission error}. The percentage in the down right cell of the matrix is the overall mean misclassification rate. As can be seen, there exist a high misclassification rate between classes $k=1$ and $k=2$, between classes $k=1$ and $k=3$, and between classes $k=1$ and $k=5$.

\begin{table}
\begin{center}
\begin{tabular}{|c|*{6}{c|}}\hline
\backslashbox{\small Estimate}{\small True} & \multicolumn{1}{p{1cm}|}{\centering k = 1} &  \multicolumn{1}{p{1cm}|}{\centering k = 2} & \multicolumn{1}{p{1.5cm}|}{\centering k = 3} & \multicolumn{1}{p{1cm}|}{\centering k = 4} & \multicolumn{1}{p{1cm}|}{\centering k = 5} & \multicolumn{1}{p{1cm}|}{\centering } \\ 
\hline \multicolumn{1}{|p{3cm}|}{ \centering k = 1} &90&39&14&3&11&42\%\\
\hline \multicolumn{1}{|p{3cm}|}{\centering k = 2} &46&301&8&2&3&16\%\\
\hline \multicolumn{1}{|p{3cm}|}{\centering k = 3}  &8&4&59&4&1&19\%\\
\hline \multicolumn{1}{|p{3cm}|}{\centering k = 4}  &5&2&6&342&5&5\%\\
\hline \multicolumn{1}{|p{3cm}|}{\centering k = 5}  &8&2&1&0&38&19\%\\
\hline \multicolumn{1}{|p{3cm}|}{\centering }  &42\%&13\%&32\%&2\%&34\%&$\textbf{17\%}$\\
\hline
\end{tabular}
\end{center}
\caption{Confusion matrix after the Q-fold cross-validation procedure over the training data.}
  \label{ch5_tab_parcels}
\end{table}

The Stan model code for the approximate GP model of this case study can be found in:
%
\begin{lstlisting}[breaklines]
https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-IV_Land-use-classification
\end{lstlisting}

\bibliography{references}

\end{document}

### 2020-03-14 Paper/manuscript_v2/manuscript_07_01.tex
update manuscript

@@
 A diagnosis procedure of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is also proposed.
-Furthermore, we show the simplicity of the method, with an attractive computational complexity due to its linear structure, which makes it easier to be used as modular components in probabilistic programming frameworks.
+We show the simplicity of the method, with an attractive computational complexity due to its linear structure, which makes it easier to be used as modular components in probabilistic programming frameworks.
 Several illustrative examples of the performance and applicability of the method in the Stan programming software are presented, and their Stan model codes are also provided.
@@
 In this paper, the relations going on among the key factors of the method such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the performance and accuracy of the method.
-Practical recommendations for the values of the key factors, based on useful graphs that encode the recognized relationships, are given.
-Which will help users to chose valid and optimized values for these factors, improving computational performance and saving time of computation.
+Practical recommendations for the values of the key factors based on useful graphs that encode the recognized relationships are given.
+Which will help users to chose valid and optimized values for these factors, improving performance and saving time of computation.
 A diagnosis of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is proposed.
 
-{\color{blue} (Gabi: This paragraph is the old version of the former paragraph) While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
-In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
-We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.}
+%In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
+We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.
 
@@
 exact and approximate GPs to real and simulated data are provided in Sections \ref{ch5_sec_univariate_cases} and \ref{ch5_sec_multivariate_cases}.
-We end with a discussion in Section \ref{ch5_sec_conclusion}.
+A brief conclusion of the work is made in Section \ref{ch5_sec_conclusion}.
 Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind this Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
-Additional content as online complementary material with more case studies illustrating the performance and applicability of the method can be found through the following link:
+
+Online complementary material with more case studies illustrating the performance and applicability of the method can be found through the following link:
+%
+\begin{lstlisting}[breaklines]
+https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/online_appendices
+\end{lstlisting}
 
@@
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
-(left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$.
-(right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$.
+RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$ (left).
+RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$ (right).
 }
@@
 \centering
-\subfigure{\includegraphics[scale=0.4, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig7_legend.pdf}}\\
+\subfigure{\includegraphics[scale=0.42, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig7_legend.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.30, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part3.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part3.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.30, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part6.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part6.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.30, trim = 1mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 1mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
 \vspace{-3mm}
-\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).
+\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
 Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
@@
 \begin{tabular}{ccc}
-\multicolumn{3}{c}{ \includegraphics[scale=0.40, trim = 25mm 117mm 90mm 45mm, clip]{ch5_fig8_legend.pdf}}\\
-\includegraphics[scale=0.30, trim = 0mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
-\includegraphics[scale=0.30, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part3.pdf} \\
-\includegraphics[scale=0.30, trim = 0mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.30, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part6.pdf}
+\multicolumn{3}{c}{ \includegraphics[scale=0.42, trim = 25mm 117mm 90mm 45mm, clip]{ch5_fig8_legend.pdf}}\\
+\includegraphics[scale=0.31, trim = 0mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part3.pdf} \\
+\includegraphics[scale=0.31, trim = 0mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part6.pdf}
 \end{tabular}

### 2020-03-16 Paper/manuscript_v2/manuscript_07_01.tex
add 'Bayesian' to title and abstract

@@
 
-\title{Hilbert space approximate Gaussian process using Stan}
+\title{Hilbert space approximate Bayesian Gaussian processes using Stan}
 
@@
 \begin{abstract}
-Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large.
-In this paper, we focus on a novel approach for low-rank approximate Gaussian processes, based on the basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
+Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov-Chain Monte-Carlo.
+In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on the basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
 The main contribution of this paper is the realization of a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, desired prediction space and non-linearity of the function to be learned.
@@
 We show the simplicity of the method, with an attractive computational complexity due to its linear structure, which makes it easier to be used as modular components in probabilistic programming frameworks.
-Several illustrative examples of the performance and applicability of the method in the Stan programming software are presented, and their Stan model codes are also provided.
+Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.
 \end{abstract}
 
-\keywords{Gaussian process; Low-rank Gaussian process; Hilbert space methods; Sparse Gaussian process.}
+\keywords{Gaussian process; Low-rank Gaussian process; Hilbert space methods; Sparse Gaussian process; Bayesian statistics; Stan.}
 

### 2020-03-16 Paper/manuscript_v2/manuscript_07_01.tex
minor edits to the intro

@@
 \usepackage{bm}
+\usepackage{url}
 
@@
 \title{Hilbert space approximate Bayesian Gaussian processes using Stan}
+
+\author{TODO}
+
+\date{}
 
@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 The key element of a GP is the covariance function that defines the dependence structure between function values at different inputs.
-However, the covariance function is also a computational issue because of the need of inverting its Gram matrix to optimize the hyperparameters.
+However, evaluating the covariance function comes with a computational issue because of the need of inverting its Gram matrix to optimize the hyperparameters.
 Given $n$ observations in the data, the computational complexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
@@
 Further, certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}.
-% , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).
-
-{\color{blue} Paul: Can we clarify how our discussed method relates to the former paragraph,
-which introduces all sorts of basis function approximations. I.e. is our method
-just one out of many or what makes it special? Perhaps we can somehow use the information provided
-in: While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
-Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.}
-
 Furthermore, a recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation (see, e.g.
 \cite{grenander1981abstract}) can be found in \cite{JSSv090i10}.
-
-In this paper we propose an approximate framework for fast and accurate inference for GPs.
+% , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).
+
+%{\color{blue} Paul: Can we clarify how our discussed method relates to the former paragraph,
+%which introduces all sorts of basis function approximations. I.e. is our method
+%just one out of many or what makes it special? Perhaps we can somehow use the information provided
+%in: While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
+Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.}
+
+In this paper we propose and evaluate an new framework for fast and accurate inference for fully Bayesian GPs using Basis function approximations.
 We focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
@@
 Furthermore, a linear representation of a GP makes it easier to be used as latent function
-in non-Gaussian observational models allowing modelling flexibility.
+in non-Gaussian observational models allowing for more modelling flexibility.
 The basis function approximation via Laplace eigenfunctions can be made arbitrary accurate and the trade-off between computational complexity and approximation accuracy can easily be controlled.
@@
 The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance kernel including the hyperparameters.
-While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost of learning the covariance function parameters is $O(mn+m)$ in every step of the optimizer.
+While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost of learning the covariance function parameters is $O(mn+m)$ in every step of the optimizer or sampler.
 This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
@@
 While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
-In this paper, the relations going on among the key factors of the method such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the performance and accuracy of the method.
-Practical recommendations for the values of the key factors based on useful graphs that encode the recognized relationships are given.
-Which will help users to chose valid and optimized values for these factors, improving performance and saving time of computation.
+In this paper, the interactions among the key factors of the method such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the performance and accuracy of the method.
+Practical recommendations are given for the values of the key factors based on intuitive graphical summaries that encode the recognized relationships.
+Our recommendations will help users to chose valid and optimized values for these factors, improving performance, and saving time of computation.
 A diagnosis of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is proposed.
@@
 
-The approach is implemented in a fully probabilistic framework and for the Stan programming probabilistic software.
-This work has served as a basis for the subsequent implementation of the method in the \textit{brms} library \citep{burkner2017brms} of the R software \citep{R2014R}.
-Several illustrative examples, simulated and real datasets, of the performance and applicability of the model, and accompanied by their Stan model codes, are developed.
-
-Although there are several GP specific software packages available to date (GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP based models.
+We have implemented the approach in the probabilistic programming language Stan \citep{carpenter2017stan} as well as subsequently in the \textit{brms} package \citep{burkner2017brms} of the R software \citep{R2019R}.
+Several illustrative examples of the performance and applicability of the method are shown using both simulated and real datasets.
+All examples are accompanied by the corresponding Stan and R code.
+
+Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, or GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP-based models.
 In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
@@
 Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind this Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
-
-Online complementary material with more case studies illustrating the performance and applicability of the method can be found through the following link:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/online_appendices
-\end{lstlisting}
+Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \texttt{Paper/online\_appendices}.
 

### 2020-03-18 Paper/manuscript_v2/manuscript_07_01.tex
more edits across the paper

@@
 Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
-This generalizes trivially to more complex models depending on several unknown functions, for example such as $p(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
+This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
 Our goal is to obtain posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
@@
  where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
+In the following, we focus on zero-mean Gaussian processes, that is set $\mu(\bm{x}) = 0$.
 The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation.
@@
 
-If $p(y_n|f(\bm{x}_n),\phi)=N(y_n|f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GP and $O(nm^2)$ for sparse GP).
-If $p(y_n|f(\bm{x}_n),\phi)=N(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
+If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
+If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
 Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011}, Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} need to be used.
-In this paper we focus on use of MCMC for integrating over the joint
+In this paper, we focus on the use of MCMC for integrating over the joint
 posterior. MCMC is not usually the fastest approach, but allows
-accurate inference for general models in probabilistic programming
-setting. We consider the computational costs of GPs specifically from
+accurate inference and uncertainty estimates for general models in probabilistic programming
+settings. We consider the computational costs of GPs specifically from
 this point of view.
@@
 \begin{align*}
-k_{\nu}(r)&=\sigma^2 \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\nu}K_{\nu}\left(\frac{\sqrt{2\nu}r}{\ell}\right),
+k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\nu}K_{\nu}\left(\frac{\sqrt{2\nu}r}{\ell}\right),
 \end{align*}
 
-\noindent where $\nu$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell$ and $\sigma$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
+\noindent where $\nu$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell$ and $\alpha$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
 The particular case where $\nu=\infty$ and $\nu=3/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
@@
 \begin{align*}
-k_{\infty}(r)&=\sigma^2 \exp\bigg(-\frac{1}{2} \frac{r^2}{\ell^2}\bigg),  \\
-k_{\frac{3}{2}}(r)&=\sigma^2\bigg(1+\frac{\sqrt{3}r}{\ell}\bigg) \exp\bigg(-\frac{\sqrt{3}r}{\ell}\bigg). 
+k_{\infty}(r)&=\alpha \exp\bigg(-\frac{1}{2} \frac{r^2}{\ell^2}\bigg),  \\
+k_{\frac{3}{2}}(r)&=\alpha\bigg(1+\frac{\sqrt{3}r}{\ell}\bigg) \exp\bigg(-\frac{\sqrt{3}r}{\ell}\bigg). 
 \end{align*}
@@
 \begin{align*}
-k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\sigma^2 \exp\bigg(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\bigg),\\
-k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\sigma^2 \Bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg)\exp\Bigg(- \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg). \\
+k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp\bigg(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\bigg),\\
+k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \Bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg)\exp\Bigg(- \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg). \\
 \end{align*}
@@
 \noindent Notice that the previous expressions have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
-The use of a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.
+The use of a multidimensional length-scale basically turns the isotropic covariance function into a non-isotropic one.
+That is, for a non-isotropic covariance function, the smoothness may vary across different input dimensions.
 
@@
 \begin{align*}
-S_{\nu}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
+S_{\nu}(\omega)&= \alpha \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
 \end{align*}
 
-\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ is a distance in the frequency domain, and $\ell$ and $\sigma$ are the lengthscale and magnitud (marginal variance), respectively, of the kernel.
+\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ is a distance in the frequency domain, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
 The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form:
@@
 \begin{align}
-S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \sigma^2 \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
+S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \alpha \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
@@
 \begin{align*}
-S_{\infty}(\omega)&= \sigma^2 \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
-S_{\frac{3}{2}}(\omega)&= \sigma^2 32\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
+S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
+S_{\frac{3}{2}}(\omega)&= \alpha 32\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
 \end{align*}
@@
 Our starting point for presenting the method is the main result obtained by \cite{solin2018hilbert} of the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
-The mathematical details of this approximation have been briefly presented in the Appendix \ref{ch5_app_approx_covfun} of this paper.
+The mathematical details of this approximation are briefly presented in Appendix \ref{ch5_app_approx_covfun}.
 
@@
 
-Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
+Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
 In this work we built these relationships for the square exponential covariance function and Matern ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix \ref{ch5_sec_periodic}.
@@
 \item For a given $\ell$ there exist a minimum $c$ under which a close approximation is never going to be achieved regardless of $m$.
-This fact can be appreciated in the Figure as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+This fact can be seen in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 \end{itemize}
@@
 we can expect the HSGP approximation to be accurate as well.
-Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c}.
+Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
-If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
+If, however, $\hat{\ell}$ does not exceed recommended minimum lengthscale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
 We may also use this diagnostic in a iterative procedure.
-Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$.
-Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
+Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
+As mentioned earlier, $c$ cannot be decreased as much as desired because it is restricted by the lengthscale, so increasing $m$ may usually the preferred approach.
 
@@
 Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
-An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be graphically represented.
+An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
 More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wiggliness in every dimensions.
-For instance, in the two-dimensional case it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
+For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
@@
 obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
-
 For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying lengthscale values.
@@
 \caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
-Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
+Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.}
   \label{fig7_posterior_varing_lscale_part1}
@@
 
-\section{Univariate case studies}\label{ch5_sec_univariate_cases}
-
-\subsection{Simulated data}\label{ch5_sec_univariate_simu}
+%\section{Univariate case studies}\label{ch5_sec_univariate_cases}
+\section{Case studies}\label{ch5_sec_cases}
+
+In this section, we will present several simulated and real case studies in which we apply the developed HSGP models. More case studies are presented in the online supplemental materials.
+
+\subsection{Simulated data for a univariate function}\label{ch5_sec_univariate_simu}
 
 This example consists of a simulated dataset with $n=250$ ($i=1,\dots,n$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=(x_1\dots,x_n)$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
-To form the final noisy dataset $\bm{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.
+To form the final noisy dataset $\bm{y}$, Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws.
 
@@
 In the HSGP model, the latent function values $f(x)$ are approximated as in equation (\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in equation (\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
-
 In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
@@
 Figure \ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The computational time is represented in the y-axis of the figure, which is in a logarithmic scale.
-The HSGP model is on average roughly 400 times faster than the regular GP, in the particular case of applying over this dataset.
-
-The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found through the following link:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-I_1D-Simulated-data
-\end{lstlisting}
+The HSGP model is on average roughly 400 times faster than the regular GP, in this particular application.
+
+The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at \url{https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-I_1D-Simulated-data}.
+%
 
@@
 Therefore, an approximate approach has to be used to fit a GP model on this data.
-We will use the HSGP model developed in this Chapter, as well as the low-rank GP model with a periodic covariance function introduced in Section \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
+We will use the HSGP method as well as the low-rank GP model with a periodic covariance function introduced in Appendix \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
 
@@
 The lengthscale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
-Which means that the used number of basis functions and boundary factor are suitable values for modeling accurately the input effects.
+This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found in:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-II_Birthday-data
-\end{lstlisting}
+The Stan model code for the approximate GP model of this case study can be found at \url{https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-II_Birthday-data}.
 
@@
 
-\section{Multivariate case studies}\label{ch5_sec_multivariate_cases}
+%\section{Multivariate case studies}\label{ch5_sec_multivariate_cases}
 
@@
 The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
-The data set consist in survival times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for $n=1043$ cases ($i=1,\dots,n$).
-Some 16\% of cases were censored.
+The data set consists of survival times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for $n=1043$ cases ($i=1,\dots,n$).
+About 16\% of cases were censored.
 Predictors are \textit{age} ($\bm{x}_1$), \textit{sex} ($\bm{x}_2$), \textit{white blood cell} (WBC) ($\bm{x}_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($\bm{x}_4$) which is a measure of deprivation for district of residence.
@@
 %Survival time was normalized to have zero mean for the logarithm of time.
-We assume a log Gaussian observation model for the observed survival time, $t_i$, with a function of the predictors, $f(X_i)$, as the location parameter, and $\sigma$ as the Gaussian noise: 
+We assume a log-normal observation model for the observed survival time, $t_i$, with a function of the predictors, $f(X_i)$, as the location parameter, and $\sigma$ as the Gaussian noise: 
 %
@@
 
-The latent function $f(\cdot)$ is modeled as a Gaussian process, centered in a linear model of the predictors $X$, and with a squared exponential covariance function $k$ depending on predictors $X$ and hyperparameters $\theta=(\alpha,\ell)$,
+The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$ depending on predictors $x$ and hyperparameters $\theta=(\alpha,\ell)$,
 %
@@
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-Notice that a scalar lengthscale is considered in the multivariate covariance function.
+Notice that a scalar lengthscale is considered in the multivariate covariance function assuming an isotropic function.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 120mm 0mm, clip]{ch5_fig24_legend_leukemia.pdf}}
-\caption{Expected log predictive density (ELPD) and time of computation in seconds per iteration (iteration of the HMC sampling method) as a function of the number of basis functions $m$ and boundary factor $c$.}
+\caption{Expected log predictive density (ELPD; left) and time of computation in seconds per iteration (iteration of the HMC sampling method; right) as a function of the number of basis functions $m$ and boundary factor $c$.}
   \label{ch5_fig22_elpd_leukemia}
@@
 
-The Stan model codes for the exact GP and the approximate GP models of this case study can be found in:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-VII_Leukemia-data
-\end{lstlisting}
+The Stan model code for the exact GP and the approximate GP models of this case study can be found at \url{https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-VII_Leukemia-data}.
+%
 
 \section{Conclusion}\label{ch5_sec_conclusion}
-The GP model entails a complexity that is computationally intractable for many applications, and this problem especially becomes severe when we want to perform inference using sampling methods.
-In this paper, a novel approach for a low-rank representation of GPs, originally and theoretically developed by \cite{solin2018hilbert}, has been implemented and analyzed.
+Modeling unknown functions using exact GPs is computationally intractable for a lot of applications.
+This problem becomes especially severe when performing full Bayesian inference using sampling-based methods.
+In this paper, a novel approach for a low-rank representation of stationary GPs, originally proposed by \cite{solin2018hilbert}, has been implemented and analyzed in detail.
 The method is based on a basis function approximation via Laplace eigenfunctions.
-The method has an attractive computational cost as this basically turns the regular GP model into a linear model, which is also an attractive property in modular probabilistic programming models.
-
-The formulae of the method and how it applies for all the stationary covariance functions as long as they can be represented in terms of their spectral densities has been clearly presented.
+The method has an attractive computational cost as it effectively approximates GPs by linear models, which is also an attractive property in modular probabilistic programming programming frameworks.
 The dominating cost per log density evaluation (during sampling) is $O(nm+m)$, which is a big benefit in comparison to $O(n^3)$ of a regular GP model.
-The design matrix is independent of hyperparameters and therefore only needs to be constructed once, at cost $O(nm)$.
+The obtained design matrix is independent of hyperparameters and therefore only needs to be constructed once, at cost $O(nm)$.
 All dependencies on the kernel and the hyperparameters are through the prior distribution of the regression weights.
-The parameter posterior distribution is $m$-dimensional, where $m$ is much smaller than the number of observations $n$, which is greatly reduced in comparison to regular GPs and this makes inference faster, especially when sampling methods are used.
-The drawbacks of the method are the boundary conditions and scaling with respect to the number of dimensions, i.e.
-the number of basis functions $m$ scales exponentially with the number of dimensions.
-
-The main contribution of this paper has been the realization of a deep analysis and diagnosis of the performance and accuracy of the approximation in relation to the key factors of the method, the number of basis functions, the boundary condition of the Laplace eigenfunctions and the non-linearity of the function to be learned.
-Recommendations for the values of these key factors based on the recognized relations among them have been provided, as well as useful graphs of these relations that will help the users to improve performance and save time of computation, which also serve as a powerful diagnosis tool whether the chosen values for the number of basis functions and the box size are adequate to fit to the actual data.
-
-Furthermore, the focus has been put on showing how GPs can be easily used as modular components in probabilistic programming frameworks (i.e.
-Stan, WinBUGS and others) and can be used as latent functions in non-Gaussian observational models.
-Several illustrative examples, simulated and real datasets, of the performance of the model, where we demonstrated the applicability and the implementation of the methodology, the reduction of the computation and the improvement in sampling efficiency have been carried out.
-
-The main drawback of this approach is that its computational complexity scales exponentially with the number of dimensions.
-Hence, in practice, input dimensionalities larger than 3 start to be too computationally demanding.
-In these cases, the proposed HSGP model can be used as low-dimensional components in an additive modeling scheme.
-
-Future research will focus on constructing analytical models for the relationships between the key factors of the number of basis functions, the boundary factor and the lengthscale of the function, depicted in Figures \ref{fig6_lscale_vs_J_vs_c}, \ref{fig6_lscale_vs_J_vs_c_Matern} and \ref{ch5_fig28_m_lscale_periodic}, on which ultimately depend the performance of the approximation.
+The parameters' posterior distribution is $m$-dimensional, where $m$ is usually much smaller than the number of observations $n$.
+
+%In terms of scaling, one main drawback of the method is the scaling with respect to the number of input dimensions, as the number of basis functions $m$ scales exponentially with the number of dimensions.
+
+The main contribution of this paper is an in-depth analysis and diagnosis of the performance and accuracy of the approximation in relation to the key factors of the method, that is, the number of basis functions, the boundary condition of the Laplace eigenfunctions, and the non-linearity of the function to be learned.
+Recommendations for the values of these key factors based on the recognized relations among them have been provided along with illustrations of these relations.
+These illustrations will not only help users to improve performance and save computation time, but also serve as a powerful diagnosis tool whether the chosen values for the number of basis functions and the boundary condition are adequate to fit to the data at hand with sufficient accuracy.
+
+The developed approximate GPs can be easily applied as modular components in probabilistic programming frameworks such as Stan in both Gaussian and non-Gaussian observational models.
+Using several simulated and real datasets, we have demonstrated the practical applicability and improved sampling effiency, as compared to regular GPs, of the developed method.
+The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
+Hence, in practice, input dimensionalities larger than 3 start to be quite computationally demanding.
+In these cases, the proposed approxiamate GP method may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
+
+The obtained functional relationships between the key factors influencing the approximation not only help users to visually assess the accuracy of the method
+but can also serve an automatic diagnostic tool, if appropriately implemented.
+In this paper, we primarily studied the functional relationships for univariate inputs.
+Accordingly, investigating the functional relationships more thoroughly for multivariate inputs remains a topic for future research.
+
+%Future research will focus on constructing analytical models for the relationships between the key factors of the number of basis functions, the boundary factor and the lengthscale of the function, depicted in Figures \ref{fig6_lscale_vs_J_vs_c}, \ref{fig6_lscale_vs_J_vs_c_Matern} and \ref{ch5_fig28_m_lscale_periodic}, on which ultimately depend the performance of the approximation.
 This analytical models can be useful to automatize the diagnosis of the performance of the approximation.

### 2020-03-18 Paper/manuscript_v2/manuscript_07_01.tex
minor edits to the appendices

@@
 
-Due to normality of the basis of the representation of the formal kernel $l(\bm{x},\bm{x}')$, its formal powers $s=1,2,\dots$ can be write as
+Due to normality of the basis of the representation of the formal kernel $l(\bm{x},\bm{x}')$, its formal powers $s=1,2,\dots$ can be written as
 %
@@
 
-A GP model with a periodic covariance function does no fit under the framework of the HSGP approximation covered so far in this study.
+A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study.
 However, it do has a low-rank representation.
@@
 \begin{eqnarray} \label{ch5_eq_cov_periodic}
-k(\bm{\tau})= \sigma^2 \exp\Big(-\frac{2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)}{\ell^2}\Big),
+k(\bm{\tau})= \alpha \exp\Big(-\frac{2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)}{\ell^2}\Big),
 \end{eqnarray}
 
-\noindent where $\sigma^2$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
+\noindent where $\alpha$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
 
@@
 \begin{eqnarray} \label{ch5_eq_cov_periodic_taylor_approx}
-k(\tau)= \sigma^2 \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 \tau),
+k(\tau)= \alpha \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 \tau),
 \end{eqnarray}
@@
 
-\noindent for $j=1,2,\cdots,J$, and where the $\text{I}_{\alpha}(z)$ is the modified Bessel function \citep{handbook1970m} of the first order of $\alpha$.
+\noindent for $j=1,2,\cdots,J$, and where the $\text{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
 This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required \citep{solin2014explicit}.
@@
 Thus, Figure \ref{ch5_fig28_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the lengthscale of the kernel.
-We have considered an approximation to be close enough in terms of satisfying equation (\ref{ch5_eq_diff_covs}) with $\varepsilon=0.5\%$.
+We have considered an approximation to be close enough in terms of satisfying equation (\ref{eq_diff_covs}) with $\varepsilon=0.5\%$.
 Notice that since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
@@
 The computational cost of this approximation scales as $O\big(n(2J+1) + (2J+1)\big)$, where $n$ is the number of observations and $J$ the number of cosine terms.
-
-The parameterization in equation (\ref{ch5_eq_f_period}) is naturally in the non-centered
-parameterization form with independent prior distribution on
-$\beta_j$, which makes the posterior inference easier.
+The parameterization in equation (\ref{ch5_eq_f_period}) is naturally in the non-centered form with independent prior distributions on
+$\beta_j$, which makes posterior inference easier.
 

### 2020-03-21 Paper/manuscript_v2/manuscript_07_01.tex
updated manuscript_v2, online material and added case studies codes

@@
 Several illustrative examples of the performance and applicability of the method are shown using both simulated and real datasets.
-All examples are accompanied by the corresponding Stan and R code.
+All examples are accompanied by the corresponding Stan %and R
+code.
 
@@
 Several case studies in which we fit
-exact and approximate GPs to real and simulated data are provided in Sections \ref{ch5_sec_univariate_cases} and \ref{ch5_sec_multivariate_cases}.
+exact and approximate GPs to real and simulated data are provided in Section \ref{ch5_sec_cases}.
 A brief conclusion of the work is made in Section \ref{ch5_sec_conclusion}.
 Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind this Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
-Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \texttt{Paper/online\_appendices}.
+Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
 
@@
 \begin{align*}
-k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\nu}K_{\nu}\left(\frac{\sqrt{2\nu}r}{\ell}\right),
+k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\Big(\frac{\sqrt{2\nu}r}{\ell}\Big)^{\!\nu} \! K_{\nu} \Big(\frac{\sqrt{2\nu}r}{\ell}\Big),
 \end{align*}
@@
 \begin{align*}
-k_{\infty}(r)&=\alpha \exp\bigg(-\frac{1}{2} \frac{r^2}{\ell^2}\bigg),  \\
-k_{\frac{3}{2}}(r)&=\alpha\bigg(1+\frac{\sqrt{3}r}{\ell}\bigg) \exp\bigg(-\frac{\sqrt{3}r}{\ell}\bigg). 
+k_{\infty}(r)&=\alpha \exp\!\! \big(\! -\frac{1}{2} \frac{r^2}{\ell^2}\big),  \\
+k_{\frac{3}{2}}(r)&=\alpha\Big(1+\frac{\sqrt{3}r}{\ell}\Big) \! \exp\!\! \big(\! -\frac{\sqrt{3}r}{\ell}\big). 
 \end{align*}
@@
 \begin{align*}
-k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp\bigg(-\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\bigg),\\
-k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \Bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg)\exp\Bigg(- \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\Bigg). \\
+k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp\!\!\Big(\! -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\Big),\\
+k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\bigg) \! \exp\!\!\Big(\! - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\Big). \\
 \end{align*}
@@
 \begin{align*}
-S_{\nu}(\omega)&= \alpha \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
+S_{\nu}(\omega)&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, l^{2\nu}}\Big(\frac{2\nu}{l^2}+4\pi^2\omega^2 \Big)^{\!(\nu+D/2)}
 \end{align*}
@@
 \begin{align}
-S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^D  \ell^D  \exp\left(-0.5 \ell^2 \omega^2\right), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \alpha \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
+S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^D  \ell^D \exp(-0.5 \ell^2 \omega^2), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\,\ell^3}\Big(\frac{3}{\ell^2}+\omega^2 \Big)^{\!-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
@@
 \begin{align*}
-S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \mathrm{exp}\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
-S_{\frac{3}{2}}(\omega)&= \alpha 32\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
+S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \exp\!\!\big(\!-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \big),   \\
+S_{\frac{3}{2}}(\omega)&= \alpha \, 32\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\big(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \big)^{\!-3}.
 \end{align*}
@@
 \begin{align}
-\lambda_j&=\left(\frac{j\pi}{2L}\right)^2, \label{eq_eigenvalue}\\
-\phi_j(x)&=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right). \label{eq_eigenfunction}
+\lambda_j&=\Big(\frac{j\pi}{2L}\Big)^{\!2}, \label{eq_eigenvalue}\\
+\phi_j(x)&=\sqrt{\frac{1}{L}}\, \sin\!\!\big(\sqrt{\lambda_j}(x+L)\big). \label{eq_eigenfunction}
 \end{align}
@@
 \begin{equation}
-k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \bm{\phi}(x'), \nonumber
+k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \, \bm{\phi}(x'), \nonumber
 \end{equation}
@@
 %
-\begin{eqnarray}
-\Delta &=&  \begin{bmatrix}
+\begin{align}
+\Delta =  \begin{bmatrix}
     S_{\theta}(\sqrt{\lambda_1}) & & \\
@@
   \end{bmatrix}.
-\end{eqnarray}
+\end{align}
 
@@
 %
-\begin{eqnarray}
-\Phi &=&  \left[ {\begin{array}{ccc}
+\begin{align}
+\Phi =  \left[ {\begin{array}{ccc}
    \phi_1(x_1) & \cdots & \phi_m(x_1)  \\
@@
   \end{array} } \right].
-\end{eqnarray}
+\end{align}
 
@@
 \begin{equation}\label{eq_approxf}
-f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j,
+f(x) \approx \sum_{j}^m \big( S_{\theta}(\sqrt{\lambda_j})\big)^{\! \frac{1}{2}} \phi_j(x) \beta_j,
 \end{equation}
@@
 \begin{align}
-\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D, \label{eq_eigenvalue_multi} \\
-%
-\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right).
-\label{eq_eigenfunction_multi}
+\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \Big\{ \Big(\frac{\pi \mathbb{S}_{jd}}{2L_d}\Big)^{\!2} \Big\}_{d=1}^D, \label{eq_eigenvalue_multi} \\
+%
+\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \sin\!\!\big(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\big). \label{eq_eigenfunction_multi}
 \end{align}
@@
 k(\bm{x},\bm{x}') \approx \sum_{j=1}^{m^{\ast}} 
-S^{\ast}_{\theta}\left(\sqrt{\bm{\lambda}^{\ast}_j}\right)
+S^{\ast}_{\theta}\big(\sqrt{\bm{\lambda}^{\ast}_j}\big)
 \phi^{\ast}_j(\bm{x}) \phi^{\ast}_j(\bm{x}'),
@@
 f(\bm{x}) \approx \sum_{j=1}^{m^{\ast}} 
-\left( S^{\ast}_{\theta} \left(\sqrt{\bm{\lambda}^{\ast}_j} \right)\right)^{1/2}
-\phi^{\ast}_j(\bm{x}) \beta_j, 
+\big( S^{\ast}_{\theta} \big(\sqrt{\bm{\lambda}^{\ast}_j} \big)\big)^{\! \frac{1}{2}} \phi^{\ast}_j(\bm{x}) \beta_j, 
 \end{equation}
@@
 \begin{equation*}
-k(\tau) = \sum_{j=1}^{\infty} S_{\theta} \left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0), 
+k(\tau) = \sum_{j=1}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
 \end{equation*}
@@
 %
-\begin{eqnarray}\label{eq_diff_covs}
-\int k(\tau) \, d\tau - 
-\int \sum_{j=1}^m S_{\theta}\left(\sqrt{\lambda_j} \right) \phi_j(\tau) \phi_j(0) \, d\tau < \varepsilon.
-\end{eqnarray}
+\begin{equation}\label{eq_diff_covs}
+\Big( \int k(\tau) \mathrm{d}\tau - 
+\int \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) \mathrm{d}\tau \Big) < \varepsilon.
+\end{equation}
 
@@
 Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern ($\nu$=3/2) covariance function, respectively, and a single input dimension.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
 More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying equation (\ref{eq_diff_covs}).
-Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern ($\nu$=3/2) covariance function.
+Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function.
 We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function:
 %
-\begin{eqnarray*}
- \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
-\end{eqnarray*}
+\begin{equation*}
+ \frac{\varepsilon}{\int k(\tau) \mathrm{d}\tau} < 0.01.
+\end{equation*}
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
 The right-side plot is a zoom in of the left-side plot.}
@@
 \subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\nu$=3/2) covariance function.
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\nu$=3/2) covariance function.
 The right-side plot is a zoom in of the left-side plot.}
@@
 
-%\section{Univariate case studies}\label{ch5_sec_univariate_cases}
 \section{Case studies}\label{ch5_sec_cases}
@@
 %
-\begin{eqnarray*}\label{ch5_eq_latentgp_simudata1}
+\begin{align*}\label{ch5_eq_latentgp_simudata1}
 \begin{split}
@@
 \end{split}
-\end{eqnarray*}
+\end{align*}
 
@@
 Figure \ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The HSGP model is on average roughly 400 times faster than the regular GP, in this particular application.
-
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at \url{https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-I_1D-Simulated-data}.
+{\color{blue}The HSGP model is on average roughly 400 times faster than the regular GP and 10 times faster than the spline model, for this particular application with a univariate input space.
+The increase of computation as function of the number of basis functions in a univariate input space is relatively slight, as can be seen in the figure.}
+
+The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}.
 %
@@
 \end{figure}
+
+{\color{blue}In the online supplemental material, some other case studies are presented.
+From those examples, it can be seen how computation demand of the HSGP model increases severely with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see equation (\ref{eq_m_multi})).
+Even though in a bivariate input space the computation demand increases significantly with $D$, the HSGP model works significantly better that the regular GP for most of the non-linear $2D$ functions (even highly wiggly functions) (see Figures B.3-right and C.3 in the online material).
+However, HSGPs are slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$) in the approximation (see Figure C.3 in the online material).
+It can be appreciated in all the cases how choosing the optimal boundary factor in the HSGP approximation allows for less number of basis functions (see Figures A.3, B.3-left and C.2 in the online material) and therefore less computational time, which can makes the difference in multivariate input spaces.
+
+%This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid a excessive computational time, especially in multivariate input spaces.
+
+Roughly similar or even worse consideration is found for splines where serious difficulties have been encountered in building the spline model from $D=3$ and \textit{knots} $> 10$, or even for $D=2$ and \textit{knots} $> 40$ (see Figures B.3-right and C.3 in the online material).}
 
@@
 %
-\begin{eqnarray} \label{ch5_eq_mean_brithday}
+\begin{equation} \label{ch5_eq_mean_brithday}
 \mu(t) = f_1(t) + f_2(t) + f_3(t) + f_4(t).
-\end{eqnarray}
+\end{equation}
 
@@
 %
-\begin{eqnarray*}
-f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\bigg(-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\bigg), 
-\end{eqnarray*}
+\begin{equation*}
+f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\Big(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\Big), 
+\end{equation*}
 
@@
 %
-\begin{eqnarray*}
-f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\bigg(-\frac{2\text{sin}^2(\pi(t-t')/365.25}{\ell_2^2}\bigg), 
-\end{eqnarray*}
+\begin{equation*}
+f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\Big(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\Big), 
+\end{equation*}
 
@@
 %
-\begin{eqnarray*}
-f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\bigg(-\frac{2\text{sin}^2(\pi(t-t')/7}{\ell_3^2}\bigg), 
-\end{eqnarray*}
+\begin{equation*}
+f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\Big(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\Big), 
+\end{equation*}
 
@@
 %
-\begin{eqnarray*}
+\begin{equation*}
 f_4(t) \sim \mathcal{N}(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
-\end{eqnarray*}
+\end{equation*}
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found at \url{https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-II_Birthday-data}.
+The Stan model code for the approximate GP model of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}.
 
@@
 
-%\section{Multivariate case studies}\label{ch5_sec_multivariate_cases}
-
 \subsection{Leukemia data}\label{ch5_sec_bf_caseVI}
@@
 Predictors are \textit{age} ($\bm{x}_1$), \textit{sex} ($\bm{x}_2$), \textit{white blood cell} (WBC) ($\bm{x}_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($\bm{x}_4$) which is a measure of deprivation for district of residence.
-We denote the matrix $X=[\bm{x}_1\; \bm{x}_2 \; \bm{x}_3 \; \bm{x}_4]^\top \in {\rm I\!R}^{n\times 4}$ which contains the predictors.
+We denote the matrix $X=[\bm{x}_1\,\, \bm{x}_2\,\, \bm{x}_3\,\, \bm{x}_4] \in {\rm I\!R}^{n\times 4}$ which contains the column-vectors $\bm{x}_d \in {\rm I\!R}^n$ of predictors ($d=1,\dots,4$).
 
@@
 \begin{equation*}
-p(y_i > t_i)= \int_{t_i}^{\infty} \mathrm{LogNormal}(y_i|f(X_i),\sigma^2) dy_i=  1 - \Phi \left( \frac{\mathrm{log}(y_i)-f(X_i)}{\sigma} \right),
+p(y_i > t_i)= \int_{t_i}^{\infty} \mathrm{LogNormal}(y_i|f(X_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \Big( \frac{\mathrm{log}(y_i)-f(X_i)}{\sigma} \Big),
 \end{equation*}
@@
 %
-\begin{eqnarray*} 
+\begin{equation*} 
 f(\bm{x}) \sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta)),
-\end{eqnarray*}
+\end{equation*}
 
@@
 %
-\begin{eqnarray} 
+\begin{align} 
 &f(\bm{x}) \sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
-&g(\bm{x}) \sim \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2=2)) \label{ch5_eq_gpml2} \\ 
-&f(\bm{x}|\bm{x}_2=2) = f(X|\bm{x}_2=2) + g(\bm{x}) \nonumber
-\end{eqnarray}
+&g(\bm{x}) \sim \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
+&f(\bm{x}|\bm{x}_2\!=\!2) = f(X|\bm{x}_2\!=\!2) + g(\bm{x}) \nonumber
+\end{align}
 
@@
 
-Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) and time of computation as function of the number of univariate basis functions $m$ and boundary factor $c$.
+Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
 As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).
+{\color{blue} With a moderate number of univariate basis functions ($m=15$) the HSGP model already becomes slower than the exact GP model, in this specific application with $3$ input variables.}
 %
@@
 
-The Stan model code for the exact GP and the approximate GP models of this case study can be found at \url{https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-VII_Leukemia-data}.
+The Stan model code for the exact GP and the approximate GP models of this case study can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}.
 %
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-Hence, in practice, input dimensionalities larger than 3 start to be quite computationally demanding.
-In these cases, the proposed approxiamate GP method may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
+{\color{red}Hence, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time, especially in multivariate input spaces.
+However, in practice, input dimensionalities larger than 3 start to be quite computationally demanding even for moderate wiggly functions.} In these cases, the proposed approxiamate GP method may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
 
@@
 \begin{equation*}
-\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
+\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}'.
 \end{equation*}
@@
 %
-\begin{eqnarray*}
-S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \cdot \bm{\tau}}, \nonumber
-\end{eqnarray*}
+\begin{equation*}
+S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \mathrm{d}\bm{\tau}, \nonumber
+\end{equation*}
 
@@
 \begin{equation*}
--\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}',
+-\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}',
 \end{equation*}
@@
 %
-\begin{eqnarray*}
-<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) d(\bm{x})
-\end{eqnarray*}
+\begin{align*}
+<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) \mathrm{d}(\bm{x})
+\end{align*}
 
@@
 %
-\begin{eqnarray*}
-\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) d(\bm{x}) = \delta_{ij},
-\end{eqnarray*}
+\begin{align*}
+\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) \mathrm{d}(\bm{x}) = \delta_{ij},
+\end{align*}
 
@@
 \begin{equation*}
-(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') d\bm{x}'.
+(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}'.
 \end{equation*}
@@
 \begin{equation*}
-[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}') d\bm{x}'.
+[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \mathrm{d}\bm{x}'.
 \end{equation*}
@@
 %
-\begin{eqnarray}\label{ch5_eq_k}
+\begin{equation}\label{ch5_eq_k}
 k(\bm{x},\bm{x}')= \sum_j [a_0+a_1\lambda_j^1+a_2\lambda_j^2+\cdots] \phi_j(\bm{x}) \phi_j(\bm{x}').
-\end{eqnarray}
+\end{equation}
 
@@
 %
-\begin{eqnarray}\label{ch5_eq_k_2}
+\begin{equation}\label{ch5_eq_k_2}
 k(\bm{x},\bm{x}')= \sum_j S(\sqrt{\lambda_j}) \phi_j(\bm{x}) \phi_j(\bm{x}'),
-\end{eqnarray}
+\end{equation}
 
@@
 %
-\begin{eqnarray} \label{ch5_eq_cov_periodic}
-k(\bm{\tau})= \alpha \exp\Big(-\frac{2\text{sin}^2\left(\omega_0\frac{\bm{\tau}}{2}\right)}{\ell^2}\Big),
-\end{eqnarray}
+\begin{equation} \label{ch5_eq_cov_periodic}
+k(\bm{\tau})= \alpha \exp\!\!\Big(\! -\frac{2\sin^{\!2}\!(\omega_0\frac{\bm{\tau}}{2})}{\ell^2}\Big),
+\end{equation}
 
@@
 % 
-\begin{eqnarray} \label{ch5_eq_cov_periodic_taylor_approx}
-k(\tau)= \alpha \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 \tau),
-\end{eqnarray}
+\begin{equation} \label{ch5_eq_cov_periodic_taylor_approx}
+k(\tau)= \alpha \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 \tau),
+\end{equation}
 
@@
 %
-\begin{eqnarray} \label{ch5_eq_q}
-\tilde{q}_j^2= \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
-\end{eqnarray}
+\begin{equation} \label{ch5_eq_q}
+\tilde{q}_j^2= \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
+\end{equation}
 
@@
 %
-\begin{eqnarray} \label{ch5_eq_q0}
-\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp\left(\frac{1}{\ell^2}\right)} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
-\end{eqnarray}
+\begin{equation} \label{ch5_eq_q0}
+\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
+\end{equation}
 
@@
 \begin{split}
-\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)}, \\
-\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp\left(\frac{1}{\ell^2}\right)},
+\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
+\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
 \end{split}
@@
 The function values of a GP model with this low-rank representation of the periodic exponential covariance function can be easily derived. Considering the identity
-\begin{equation*}
-\text{cos}(j\omega_0 (x-x'))=\text{cos}(j\omega_0 x) \text{cos}(j\omega_0 x') + \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x'),
+%
+\begin{equation*}
+\cos(j\omega_0 (x-x'))=\cos(j\omega_0 x) \cos(j\omega_0 x') + \sin(j\omega_0 x) \sin(j\omega_0 x'),
 \end{equation*}
@@
 %
-\begin{eqnarray} \label{ch5_eq_cov_periodic_taylor_approx_2}
-k(x,x')= \sigma^2 \Big( \sum_{j=0}^{J} \tilde{q}_j^2 \text{cos}(j\omega_0 x)  \text{cos}(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \text{sin}(j\omega_0 x) \text{sin}(j\omega_0 x') \Big).
-\end{eqnarray}
+\begin{equation} \label{ch5_eq_cov_periodic_taylor_approx_2}
+k(x,x')= \sigma^2 \Big( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \Big).
+\end{equation}
 
@@
 \begin{equation} \label{ch5_eq_f_period}
-f(x) \approx \sigma \Big( \sum_{j=0}^J  \tilde{q}_j \text{cos}(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \text{sin}(j\omega_0 x) \beta_{J+1+j} \Big),
+f(x) \approx \sigma \Big( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \Big),
 \end{equation}
@@
 \noindent where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\dots,2J+1$.
-The cosine $\text{cos}(j\omega_0 x)$ and sinus $\text{sin}(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.
+The cosine $\cos(j\omega_0 x)$ and sinus $\sin(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.
 The only dependence on the hyperparameter $\ell$ is through the coefficients $\tilde{q}_j$, which are $J$-dimensional.

### 2020-03-21 Paper/online_supplemental_material/online_material.tex
updated manuscript_v2, online material and added case studies codes

--- 2020-03-14 Paper/online_appendices/online_appendices.tex
@@
 
+\usepackage{chngcntr}
+
 \bibliographystyle{authordate1}
 
-\title{Online appendices}
+\title{Online supplemental material of the article:\\ \vspace{7mm} 'Hilbert space approximate Bayesian Gaussian processes using Stan'\\ \vspace{7mm}
+Case Studies}
+
+\author{TODO}
+
+\date{}
 
@@
 
-\section{Univariate case studies}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+\counterwithin{equation}{section}
 
-\subsection{Same-sex marriage data}\label{ch5_sec_studycaseII}
+\section{Same-sex marriage case study}\label{ch5_sec_studycaseII}
 This data set relates the proportion of support for same-sex marriage to the age.
@@
 \begin{align*} \label{ch5_eq_gpprior_gay}
-p_i &= \text{logit}(f(x_i)) \nonumber \\
+p_i &= \mathrm{logit}(f(x_i)) \nonumber \\
 f(x) &\sim \mathcal{GP}(0, k(x,x', \theta)).
@@
 \noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,74$.
-In the HSGP model, the function $f(x)$ is approximated as in equation (\ref{eq_approxf}), with the squared exponential spectral density as in equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+In the HSGP model, the function $f(x)$ is approximated as in equation (7) %(\ref{eq_approxf})
+, with the squared exponential spectral density as in equation (1)%(\ref{eq_specdens_inf})
+, and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (5) %(\ref{eq_eigenvalue}) 
+and (6)%(\ref{eq_eigenfunction})
+.
 
@@
 The computational times is represented in the y-axis which is on a logarithmic scale.
-The HSGP model is on average roughly 10 times faster than the regular GP, in this particular case.
+The HSGP model is on average roughly 15 times faster than the regular GP and 5 times faster than the spline model, for this particular case and univariate input space.
+{\color{blue}The increase of computation as function of the number of basis functions in a univariate input space is relatively slight, as can be seen in the figure}.
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found in:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-V_Same-sex-marriage
-\end{lstlisting}
+The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}.
+
+%\begin{lstlisting}[breaklines]
+%https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data
+%\end{lstlisting}
 
@@
 
-\section{Multivariate case studies}
-
-\subsection{2D Simulated data}
+\section{Case study of simulated data for a bivariate function}
 
@@
 
-In the HSGP model with $2$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^2$, is approximated as in equation (\ref{eq_approxf_multi}), with the $2$-dimensional squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $2$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+In the HSGP model with $2$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^2$, is approximated as in equation (12)% (\ref{eq_approxf_multi})
+, with the $2$-dimensional squared exponential spectral density $S$ as in equation (1)% (\ref{eq_specdens_inf})
+, and the multivariate eigenfunctions $\phi_j$ and the $2$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (10) %(\ref{eq_eigenfunction_multi})
+ and (9)% (\ref{eq_eigenvalue_multi})
+ , respectively.
 
@@
 In all models, the same number of basis functions and knots per dimension were used.
-Figure \ref{ch5_fig17_RMSE_exII}-(left) shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, and the number of univariate basis functions $m$, for the HSGP model, and knots, for the splines model.
-From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-(left), it can be seen a close approximation of the HSGP model to the regular GP model.
+Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, and the number of univariate basis functions $m$, for the HSGP model, and knots, for the splines model.
+From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen a close approximation of the HSGP model to the regular GP model.
 However, the performance of the splines model is significantly worse.
-Figure \ref{ch5_fig17_RMSE_exII}-(right) shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots.
+Figure \ref{ch5_fig17_RMSE_exII}-right shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots.
 Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computational time.
+{\color{blue}Even though in a bivariate input space the computation demand increases significantly with the number of dimensions or knots, either the HSGP or spline models work significantly better than regular GP, even for highly wiggly functions where a high number of basis functions or knots are required in the approximations.
+However, severe difficulties have been found in building the spline model with $50$ knots for this bivariate function}.
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found in:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-VI_Simulated-data_2D
-\end{lstlisting}
+The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}.
 
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 113mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
 \caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
@@
 
-\subsection{Diabetes data}\label{ch5_sec_bf_caseV}
+\section{Diabetes case study}\label{ch5_sec_bf_caseV}
 The next example presents an epidemiological study of diabetes disease.
@@
 %
-\begin{eqnarray*} \label{ch5_eq_gpprior_gay}
-p_i &=& \text{logit}(f(\bm{x}_i)) \nonumber \\
-f(\bm{x}) &\sim& \mathcal{GP}(0, k(\bm{x},\bm{x}, \theta).
-\end{eqnarray*}
+\begin{align*} \label{ch5_eq_gpprior_gay}
+p_i &= \mathrm{logit}(f(\bm{x}_i)) \nonumber \\
+f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x},\bm{x}, \theta).
+\end{align*}
 
@@
 
-In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$ evaluated at input vector $\bm{x} \in {\rm I\!R}^D$ is approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}) and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.
+In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$ evaluated at input vector $\bm{x} \in {\rm I\!R}^D$ is approximated as in equation (12)% (\ref{eq_approxf_multi})
+, with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (1) %(\ref{eq_specdens_inf}) 
+and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (9) % (\ref{eq_eigenvalue_multi}) 
+and (10)% (\ref{eq_eigenfunction_multi})
+, respectively.
 
@@
 Basically, with slightly differences, all models show similar performances, due to the fact that the process is very smooth with a relatively very large lengthscale estimate $\ell=4.51$.
-Even though, a slight pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (Figure \ref{fig6_lscale_vs_J_vs_c}).
+Even though, a slight pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (Figure 6% \ref{fig6_lscale_vs_J_vs_c}
+).
 %
@@
 We can appreciate the significant increase of computational time with higher dimensions for the HSGP and spline models.
-This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid a excessive computational time, especially in high input dimensionality.
+In this sense, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure 6% \ref{fig6_lscale_vs_J_vs_c}
+, can be essential to avoid a excessive computational time, especially in multivariate input spaces.
 It is interesting to be noticed that considering more than 10 knots per dimension in the spline model with 3D is not allowed for an amount of 392 observations.
@@
 
-The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found in:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-III_Diabetes-data
-\end{lstlisting}
+The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}.
 
@@
 \subfigure{\includegraphics[scale=0.35, trim = 28mm 50mm 112mm 10mm, clip]{ch5_fig20_legend_diabetes_2.pdf}}
-\caption{Time of computation in seconds per iteration (iteration of the HMC sampling method) of the different models fitted over the dataset, with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$ and number of basis functions $m$, for the HSGP model, and knots, for the spline model.}
+\caption{Time of computation in seconds per iteration (iteration of the HMC sampling method) of the different models fitted over the dataset, with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$ and number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+The y-axis is on a logarithmic scale.}
   \label{ch5_fig20_time2D_diabetes}
@@
 
-\subsection{Spatio-temporal land-use classification}\label{ch5_sec_application_caseII}
+\section{Spatio-temporal land-use classification case study}\label{ch5_sec_application_caseII}
 
@@
 
-\noindent which is related to the vector of probabilities $\bm{p}_{it}$ through the 'softmax' link function (equation \ref{ch2_eq_softmax} in chapter \ref{ch2_bayes}):
+\noindent which is related to the vector of probabilities $\bm{p}_{it}$ through the 'softmax' link function:
 %
-\begin{eqnarray*}
-\bm{p}_{it} = \text{softmax}(f(\bm{x}_{it})).
-\end{eqnarray*}
+\begin{equation*}
+\bm{p}_{it} = \mathrm{softmax}(f(\bm{x}_{it})).
+\end{equation*}
 
@@
 %
-\begin{eqnarray} \label{ch5_eq_f_parcels}
+\begin{equation} \label{ch5_eq_f_parcels}
 f_k(\bm{x}) = \sum_{d=1}^{D} g_d(x^d) + \sum_{d=1}^{D-1} h_{d,D}(x^d,x^D).
 %\mathcal{HSGP}(x^d, S(\bm{\lambda}), \theta_d),
-\end{eqnarray}
+\end{equation}
 
@@
 %
-\begin{eqnarray*}
+\begin{equation*}
 g_d(x^d) \sim \mathcal{HSGP}(x^d, S, \theta_{d,1}).
-\end{eqnarray*}
+\end{equation*}
 
-\noindent In the HSGP model, a first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in equation (\ref{eq_approxf}) with the squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}) and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
+\noindent In the HSGP model, a first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in equation (7) % (\ref{eq_approxf}) 
+with the squared exponential spectral density $S$ as in equation (1) % (\ref{eq_specdens_inf})
+ and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (5) % (\ref{eq_eigenvalue})
+and (6)% (\ref{eq_eigenfunction})
+  , respectively.
 
@@
 %
-\begin{eqnarray*}
+\begin{equation*}
 h_{d,D}(x^d,x^D) \sim \mathcal{HSGP}(x^d,x^D, S, \theta_{d,D}).
-\end{eqnarray*}
+\end{equation*}
 
-\noindent In the HSGP model, a second-order component $h_{d,D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in equation (\ref{eq_approxf_multi}) with the two-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}) and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}), respectively.
+\noindent In the HSGP model, a second-order component $h_{d,D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in equation (12)% (\ref{eq_approxf_multi})
+ with the two-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (1) % (\ref{eq_specdens_inf}) 
+ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (9) % (\ref{eq_eigenvalue_multi}) 
+ and (10)% (\ref{eq_eigenfunction_multi})
+ , respectively.
 
@@
 
-In the case of the first-order components, the normalized lengthscale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$, $c$.
-Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.
+In the case of the first-order components, the normalized lengthscale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum lengthscale reported by Figure (6) % \ref{fig6_lscale_vs_J_vs_c}
+as a function of $m$, $c$. Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found in:
-%
-\begin{lstlisting}[breaklines]
-https://github.com/gabriuma/Doctoral_thesis/tree/master/Case-study-IV_Land-use-classification
-\end{lstlisting}
+The Stan model code for the approximate GP model of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}.
 

### 2020-03-21 Paper/manuscript_v2/manuscript_07_01.tex
little update

@@
 About 16\% of cases were censored.
-Predictors are \textit{age} ($\bm{x}_1$), \textit{sex} ($\bm{x}_2$), \textit{white blood cell} (WBC) ($\bm{x}_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($\bm{x}_4$) which is a measure of deprivation for district of residence.
-We denote the matrix $X=[\bm{x}_1\,\, \bm{x}_2\,\, \bm{x}_3\,\, \bm{x}_4] \in {\rm I\!R}^{n\times 4}$ which contains the column-vectors $\bm{x}_d \in {\rm I\!R}^n$ of predictors ($d=1,\dots,4$).
+Predictors are \textit{age} ($x_1$), \textit{sex} ($x_2$), \textit{white blood cell} (WBC) ($x_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($x_4$) which is a measure of deprivation for district of residence.
+We denote $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$ as the vector of predictor values for observation $i$.
+
+%We denote the matrix $X=[\bm{x}_1\, \cdots \bm{x}_i\, \cdots \bm{x}_n] \in {\rm I\!R}^{n\times 4}$ where $\bm{x}_i=(x_1,x_2,x_3,x_4) \in {\rm I\!R}^{4}$ the vector of predictor values for observations $i$
 
@@
 %Survival time was normalized to have zero mean for the logarithm of time.
-We assume a log-normal observation model for the observed survival time, $t_i$, with a function of the predictors, $f(X_i)$, as the location parameter, and $\sigma$ as the Gaussian noise: 
-%
-\begin{equation*}
-p(t_i)= \mathrm{LogNormal}(t_i|f(X_i),\sigma^2)
-\end{equation*}
-
-\noindent with $X_i=\{x_{i1},x_{i2},x_{i3},x_{i4}\} \in {\rm I\!R}^4$ and $f:{\rm I\!R} \to {\rm I\!R}$.
+We assume a log-normal observation model for the observed survival time, $t_i$, with a function of the predictors, $f(\bm{x}_i):{\rm I\!R}^4 \to {\rm I\!R}$, as the location parameter, and $\sigma$ as the Gaussian noise:
+%
+\begin{equation*}
+p(t_i)= \mathrm{LogNormal}(t_i|f(\bm{x}_i),\sigma^2).
+\end{equation*}
 
@@
 \begin{equation*}
-p(y_i > t_i)= \int_{t_i}^{\infty} \mathrm{LogNormal}(y_i|f(X_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \Big( \frac{\mathrm{log}(y_i)-f(X_i)}{\sigma} \Big),
+p(y_i > t_i)= \int_{t_i}^{\infty} \mathrm{LogNormal}(y_i|f(\bm{x}_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \Big( \frac{\mathrm{log}(y_i)-f(\bm{x}_i)}{\sigma} \Big),
 \end{equation*}
@@
 \begin{align} 
-&f(\bm{x}) \sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
-&g(\bm{x}) \sim \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
-&f(\bm{x}|\bm{x}_2\!=\!2) = f(X|\bm{x}_2\!=\!2) + g(\bm{x}) \nonumber
+f(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
+g(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
+f(\bm{x}|\bm{x}_2\!=\!2) &= f(\bm{x}|\bm{x}_2\!=\!2) + g(\bm{x}). \nonumber
 \end{align}

### 2020-03-22 Paper/manuscript_v2/manuscript_07_01.tex
add preliminary author list and minor cleaning

@@
 
-\author{TODO}
-
-\date{}
+\author{Gabriel R. Mayol$^{1*}$, Paul-Christian Bürkner$^{2}$, Michael R. Anderson$^3$, Aki Vehtari$^{2}$}
+
+\date{
+$^1$ TODO\break
+$^2$ Department of Computer Science, Aalto University, Finland\break
+$^3$ TODO\break
+$^*$ Corresponding author, Email: gabriuma@gmail.com
+}
 
@@
 Figure \ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-{\color{blue}The HSGP model is on average roughly 400 times faster than the regular GP and 10 times faster than the spline model, for this particular application with a univariate input space.
-The increase of computation as function of the number of basis functions in a univariate input space is relatively slight, as can be seen in the figure.}
+The HSGP model is on average roughly 400 times faster than the regular GP and 10 times faster than the spline model, for this particular application with a univariate input space.
+Also, the increase in computation time as a function of the number of basis functions in a univariate input space is relatively slight.
 
@@
 
-{\color{blue}In the online supplemental material, some other case studies are presented.
-From those examples, it can be seen how computation demand of the HSGP model increases severely with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see equation (\ref{eq_m_multi})).
-Even though in a bivariate input space the computation demand increases significantly with $D$, the HSGP model works significantly better that the regular GP for most of the non-linear $2D$ functions (even highly wiggly functions) (see Figures B.3-right and C.3 in the online material).
-However, HSGPs are slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$) in the approximation (see Figure C.3 in the online material).
-It can be appreciated in all the cases how choosing the optimal boundary factor in the HSGP approximation allows for less number of basis functions (see Figures A.3, B.3-left and C.2 in the online material) and therefore less computational time, which can makes the difference in multivariate input spaces.
+In the online supplemental material, some other case studies are presented.
+From those examples, it can be seen how computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see equation (\ref{eq_m_multi})).
+Even though, in a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the regular GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
+However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure C.3 in the online material).
+In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticably (see Figures A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
 
@@
 
-Roughly similar or even worse consideration is found for splines where serious difficulties have been encountered in building the spline model from $D=3$ and \textit{knots} $> 10$, or even for $D=2$ and \textit{knots} $> 40$ (see Figures B.3-right and C.3 in the online material).}
+Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and \textit{knots} $> 10$, or even for $D=2$ and \textit{knots} $> 40$ (see Figures B.3-right and C.3 in the online material).
 
@@
 g(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
-f(\bm{x}|\bm{x}_2\!=\!2) &= f(\bm{x}|\bm{x}_2\!=\!2) + g(\bm{x}). \nonumber
+f(\bm{x}|\bm{x}_2\!=\!2) &= f(\bm{x}|\bm{x}_2\!=\!1) + g(\bm{x}). \nonumber
 \end{align}
@@
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).
-{\color{blue} With a moderate number of univariate basis functions ($m=15$) the HSGP model already becomes slower than the exact GP model, in this specific application with $3$ input variables.}
+With a moderate number of univariate basis functions ($m=15$), the HSGP model becomes slower than the exact GP model, in this specific application with $3$ input variables, as the total number of multivariate basis functions becomes $15^3 = 3375$ and is therefore quite high.
 %
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-{\color{red}Hence, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time, especially in multivariate input spaces.
-However, in practice, input dimensionalities larger than 3 start to be quite computationally demanding even for moderate wiggly functions.} In these cases, the proposed approxiamate GP method may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
+Hence, choosing optimal values for the number of basis functions and the boundary factor, using the
+ recommendations and diagnostics provided in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time especially in multivariate input spaces.
+However, in practice, input dimensionalities larger than 3 start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
+In these high dimensional cases, the proposed approxiamte GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
 

### 2020-03-23 Paper/manuscript_v2/manuscript_07_01.tex
Leukemia's multilevel formulae proposal

@@
 
-Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable (1 for female and 2 for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
-The relative contribution of a GP function given one of the levels of the predictor (equation \ref{ch5_eq_gpml2}) to a general mean GP function (equation \ref{ch5_eq_gpml1}) is defined.
-For the other level of the predictor, the GP function effects are set to zero.
+Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable ('1' for female and '2' for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
+The relative contribution of a GP function given one of the levels of the predictor (equation (\ref{ch5_eq_gpml2})) to a general mean GP function ($\mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1))$ in  equation (\ref{ch5_eq_gpml1})) is defined.
+For the other level of the predictor, the GP function effects are set to zero (equation (\ref{ch5_eq_gpml2})).
 This multilevel construction is depicted as follows:
@@
 \begin{align} 
-f(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
-g(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2|\bm{x}_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
-f(\bm{x}|\bm{x}_2\!=\!2) &= f(\bm{x}|\bm{x}_2\!=\!1) + g(\bm{x}). \nonumber
+f(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) + g(\bm{x}), \label{ch5_eq_gpml1}
 \end{align}
+
+\noindent with
+%
+\begin{align} \label{ch5_eq_gpml2}
+g(\bm{x}) =
+\left\{
+	\begin{array}{ll}
+		0 & \;\; \mbox{if } \;\; x_2\!=\!\text{'}1\text{'}  \\
+		\mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2))  & \;\; \mbox{if } \;\; x_2\!=\!\text{'}2\text{'} 
+	\end{array}.
+\right.
+\end{align}
+%
+%
+
+%\begin{align} 
+%f(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
+%g(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2 \,|\, x_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
+%f(\bm{x} \, |\, x_2\!=\!2) &= f(\bm{x} \,|\, x_2\!=\!1) + g(\bm{x}). \nonumber
+%\end{align}
 
@@
 
-\noindent In the previous equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of the GP function restricted to the male sex ($\bm{x}_2=2$).
+\noindent In the previous equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($\bm{x}_2=\text{'}2\text{'}$).
 

### 2020-03-24 Paper/manuscript_v2/manuscript_07_01.tex
improve dummy coding explanation

@@
 
-Due to the predictor \textit{sex} ($\bm{x}_2$) is a categorical variable ('1' for female and '2' for male), we can outline a multilevel model for the GP function, in a similar way like categorical effects are treated in linear models.
-The relative contribution of a GP function given one of the levels of the predictor (equation (\ref{ch5_eq_gpml2})) to a general mean GP function ($\mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1))$ in  equation (\ref{ch5_eq_gpml1})) is defined.
-For the other level of the predictor, the GP function effects are set to zero (equation (\ref{ch5_eq_gpml2})).
-This multilevel construction is depicted as follows:
+Due to the predictor \textit{sex} ($\bm{x}_2$) being a categorical variable ('1' for female and '2' for male), we apply dummy/treatment coding for the GP functions, in a similar way such coding is applied in linear models.
+A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise.
+More formally, this construction reads as follows:
 %
@@
 		0 & \;\; \mbox{if } \;\; x_2\!=\!\text{'}1\text{'}  \\
-		\mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2))  & \;\; \mbox{if } \;\; x_2\!=\!\text{'}2\text{'} 
-	\end{array}.
+		\mathcal{GP}(0, k(\bm{x},\bm{x}', \theta_2))  & \;\; \mbox{if } \;\; x_2\!=\!\text{'}2\text{'} 
+	\end{array}
 \right.
@@
 
-\noindent In the previous equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($\bm{x}_2=\text{'}2\text{'}$).
+\noindent In the former equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($\bm{x}_2=\text{'}2\text{'}$).
 

### 2020-03-24 Paper/manuscript_v2/manuscript_07_01.tex
author's list

@@
 
-\author{Gabriel R. Mayol$^{1*}$, Paul-Christian Bürkner$^{2}$, Michael R. Anderson$^3$, Aki Vehtari$^{2}$}
-
-\date{
-$^1$ TODO\break
-$^2$ Department of Computer Science, Aalto University, Finland\break
-$^3$ TODO\break
+\author{Gabriel Riutort-Mayol$^{1*}$, Michael R. Andersen$^2$, Paul-Christian Bürkner$^{3}$, Aki Vehtari$^{2}$}
+
+\date{ \small
+$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain. \break
+$^2$ TODO\break
+$^3$ Department of Computer Science, Aalto University, Finland\break
 $^*$ Corresponding author, Email: gabriuma@gmail.com
@@
 
-Due to the predictor \textit{sex} ($\bm{x}_2$) being a categorical variable ('1' for female and '2' for male), we apply dummy/treatment coding for the GP functions, in a similar way such coding is applied in linear models.
+Due to the predictor \textit{sex} ($x_2$) being a categorical variable ('1' for female and '2' for male), we apply dummy/treatment coding for the GP functions, in a similar way such coding is applied in linear models.
 A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise.
@@
 
-\noindent In the former equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($\bm{x}_2=\text{'}2\text{'}$).
+\noindent In the former equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($x_2=\text{'}2\text{'}$).
 

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
Update affiliation

@@
 $^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain. \break
-$^2$ TODO\break
+$^2$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
 $^3$ Department of Computer Science, Aalto University, Finland\break

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to abstract

@@
 Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov-Chain Monte-Carlo.
-In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on the basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
-The main contribution of this paper is the realization of a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, desired prediction space and non-linearity of the function to be learned.
+In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
+The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, desired prediction space and smoothness of the function to be learned.
 Intuitive visualizations and useful recommendations for the values of these factors which help users to improve performance and computation are provided.
-A diagnosis procedure of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is also proposed.
-We show the simplicity of the method, with an attractive computational complexity due to its linear structure, which makes it easier to be used as modular components in probabilistic programming frameworks.
+A diagnosis procedure for whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is also proposed.
+The proposed method is simple and exhibits an attractive computational complexity due to its linear structure and it is easy to implmenent in probabilistic programming frameworks.
 Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to introduction and added a few to dos

@@
 \bibliographystyle{authordate1}
+% TODO macro
+\newcommand{\todo}[1]{\textcolor{blue}{\textbf{[#1]}}}
 
@@
 Due to their generality and flexibility, GPs are of broad interest across machine learning and statistics \citep{rasmussen2006gaussian,neal1997monte}.
-Among others, they find application in the fields of spatial epidemiology \citep{diggle2013statistical,carlin2014hierarchical}, robotics and control \citep{deisenroth2015gaussian}, signal processing \citep{sarkka2013spatiotemporal}, as well as Bayesian optimization and probabilistic numerics \citep{roberts2010bayesian,briol2015probabilistic,hennig2015probabilistic}.
+Among others, they find application in the fields of spatial epidemiology \citep{diggle2013statistical,carlin2014hierarchical}, robotics and control \citep{deisenroth2015gaussian}, signal processing \citep{sarkka2013spatiotemporal}, neuroimaging \citep{andersen2017} as well as Bayesian optimization and probabilistic numerics \citep{roberts2010bayesian,briol2015probabilistic,hennig2015probabilistic}.
 
 The key element of a GP is the covariance function that defines the dependence structure between function values at different inputs.
-However, evaluating the covariance function comes with a computational issue because of the need of inverting its Gram matrix to optimize the hyperparameters.
-Given $n$ observations in the data, the computational complexity and memory requirements of exact GP implementation in general scale as $O(n^3)$ and $O(n^2)$, respectively.
-This limit their application to rather small data sets of a few tens of thousands observations at most.
+However, evaluating the covariance function \todo{evaluating the covariance function is not a problem, but computing the posterior is} comes with a computational issue because of the need of inverting its Gram matrix to optimize the hyperparameters.
+Given $n$ observations in the data, the computational complexity and memory requirements of computing the posterior distribution for a GP in general scale as $O(n^3)$ and $O(n^2)$, respectively.
+This limits their application to rather small data sets of a few tens of thousands observations at most.
 The problem becomes more severe when performing full Bayesian inference via sampling methods, where in each sampling step we need $O(n^3)$ computations when inverting the Gram matrix of the covariance function, usually through Cholesky factorization.
@@
 can be found in \cite{rasmussen2006gaussian}.
-\citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for accurate approximation.
+\citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for an accurate approximation.
 % More recent developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
@@
 Recently, \cite{hensman2017variational} presented a variational Fourier feature approximation for GPs that was derived for the Mat{\'e}rn class of kernels.
-Another related method for approximating kernels relies on random fourier features \citep{rahimi2008random,rahimi2009weighted}.
+Another related method for approximating kernels relies on random Fourier features \citep{rahimi2008random,rahimi2009weighted}.
 Further, certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}.
@@
 
-In this paper we propose and evaluate an new framework for fast and accurate inference for fully Bayesian GPs using Basis function approximations.
+In this paper we propose and evaluate a new framework for fast and accurate inference for fully Bayesian GPs using basis function approximations.
 We focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
-Using basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.
-A linear model structure makes GPs easier to implement as building blocks in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if approximation specific computation is simple.
+Using a basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.
+The linear model structure makes GPs easy to implement as a building block in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if the approximation specific computation is simple.
 Furthermore, a linear representation of a GP makes it easier to be used as latent function
@@
 
-The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance kernel including the hyperparameters.
+The Laplace eigenfunctions can be computed analytically and they are independent of the particular choice of the covariance function including the hyperparameters.
 While the pre-computation cost of the basis functions is $O(m^2n)$, the computational cost of learning the covariance function parameters is $O(mn+m)$ in every step of the optimizer or sampler.
@@
 Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
-This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation.
-The basis function approach also provides an easy way to apply the non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
+This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation I don't understand the sentence about autodiff and computational complexity}.
+The basis function approach also provides an easy way to apply a non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
 
@@
 Practical recommendations are given for the values of the key factors based on intuitive graphical summaries that encode the recognized relationships.
-Our recommendations will help users to chose valid and optimized values for these factors, improving performance, and saving time of computation.
+Our recommendations will help users to choose valid and optimized values for these factors, improving performance, and saving time of computation.
 A diagnosis of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is proposed.
@@
 
-Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, or GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP-based models.
+Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, and GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP-based models.
 In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
@@
 The remainder of the paper is structured as follows.
-In Section \ref{ch4_gp}, we introduce GPs with the covariance and spectral density functions.
+In Section \ref{ch4_gp}, we introduce GPs, covariance functions and spectral density functions.
 In Section \ref{ch5_sec_method}, the reduced rank approximation to GPs proposed by \cite{solin2018hilbert} is described.

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
Added another todo in introduciton

@@
 Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
-This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation I don't understand the sentence about autodiff and computational complexity}.
+This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation \todo{I don't understand the sentence about autodiff and computational complexity}.
 The basis function approach also provides an easy way to apply a non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits and added to dos to Section 2

@@
 In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. 
-Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters $f$ and $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
+Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
 This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
-Our goal is to obtain posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
+Our goal is to obtain the posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
 % That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
@@
 If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
-If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have closed form solution.
-Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011}, Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} need to be used.
+If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),g(\bm{x}_n))$ \todo{Is something missing here?
+The function g is missing on the left hand side.
+} or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed form solution.
+Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011}, Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} are required.
 In this paper, we focus on the use of MCMC for integrating over the joint
-posterior. MCMC is not usually the fastest approach, but allows
+posterior. MCMC is usually not the fastest approach, but it is flexible and allows
 accurate inference and uncertainty estimates for general models in probabilistic programming
@@
 
-\subsection{Covariance function and spectral density}\label{ch4_sec_cov}
-
-The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs.
+\subsection{Covariance functions and spectral density}\label{ch4_sec_cov}
+
+The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and defines a correlation structure which characterizes the correlations between function values at different inputs.
 A covariance function needs be symmetric and positive semi-definite \citep{rasmussen2006gaussian}.
 A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
-Isotropic covariance functions are those that are function of the distance between observations, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant.
-The most commonly used distance between observations is the norm L2 $(|\bm{x}-\bm{x}'|_{L2})$, also known as Euclidean distance, although other types of distances can be considered.
+Isotropic covariance functions are only depends on the input points through the norm of the difference, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant.
+The most commonly used distance between observations is the L2-norm $(|\bm{x}-\bm{x}'|_{L2})$, also known as Euclidean distance, although other types of distances can be considered.
 
@@
 
-\noindent where $\nu$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell$ and $\alpha$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
+\noindent where $\nu > 0$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell > 0$ and $\alpha > 0$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
 The particular case where $\nu=\infty$ and $\nu=3/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
@@
 
-\noindent The former is commonly known as squared exponential (exponentiated quadratic) covariance function.
-Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form:
+\noindent The former is commonly known as the squared exponential (exponentiated quadratic) covariance function.
+Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form
 %
@@
 \noindent Notice that the previous expressions have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
-The use of a multidimensional length-scale basically turns the isotropic covariance function into a non-isotropic one.
+Using individual length-scales for each dimension turns the isotropic covariance function into a non-isotropic covariance function.
 That is, for a non-isotropic covariance function, the smoothness may vary across different input dimensions.
@@
 
-\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ is a distance in the frequency domain, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
-The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form:
+\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ denotes the frequency, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
+The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form
 %
@@
 \end{align}
-
-\noindent Particularizing to an input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form:
+\todo{Is it only purpose that only some equations have numbers?}
+\noindent For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
 %

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits and added to dos in Section 3

@@
 \usepackage{url}
+\usepackage{parskip}
 
@@
 
-The approximate GP method, developed by \cite{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a homogeneous (stationary) covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
-Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to some boundary condition.
+The approximate GP method, developed by \cite{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a stationary covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
+Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to boundary conditions.
 For brevity, we will refer to these approximate Gaussian processes as HSGPs.
-Below, we will present the main results around HSGPs relevant for practical application.
+Below, we will present the main results around HSGPs relevant for practical applications.
 More details and mathematical proofs are provided in \cite{solin2018hilbert}.
@@
 
-We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in [-L,L] \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition.
+We begin by focusing on the case of a unidimensional input space (i.e.
+on GPs with just a single covariate) such that $\Omega \in [-L,L] \subset {\rm I\!R}$, where $L$ is some positive real number to which we also refer as boundary condition.
 As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs.
@@
 The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
-Namely, they satisfy the following eigenvalue problem in $\Omega$ when applying the Dirichlet boundary condition (other boundary conditions could be used as well):
+Namely, they satisfy the following eigenvalue problem in $\Omega$ when applying the Dirichlet boundary condition (other boundary conditions could be used as well)
 %
@@
 
-\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalues problem in equation (\ref{eq_eigenproblem}) are sinusoidal functions.
-Independently of the covariance function, they can be computed as
+\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in equation (\ref{eq_eigenproblem}) are sinusoidal functions.
+The solution to the eigenvalue problem is independent of  the specific choice of covariance function and is given by
 %
@@
 
-If we truncate the sum in (\ref{eq_approxcov}) to the first $m$ terms, the approximate covariance function becomes
+If we truncate the sum in eq. \eqref{eq_approxcov} to the first $m$ terms, the approximate covariance function becomes
 %
@@
 
-\noindent where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is the diagonal matrix of the spectral densities $S_{\theta}(\sqrt{\lambda_j})$:
+\noindent where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is a diagonal matrix of the spectral density evaluated at the square root of the eigenvalues, i.e.
+$S_{\theta}(\sqrt{\lambda_j})$,
 %
@@
 
-\noindent where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$:
+\noindent where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$
 %
@@
 Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
-A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the covariance hyperparameters $\theta$.
+A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the hyperparameters of the covariance function $\theta$.
 Instead, the only dependence of the model on $\theta$ is through the spectral density $S_{\theta}$.
 The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions.
-Therefore, equation (\ref{eq_approxf}) can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
+Therefore, eq.
+\eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
 The computational cost of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
 
-The parameterization in equation (\ref{eq_approxf}) is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which makes the posterior inference easier.
-Furthermore, all dependencies on the covariance kernel and the hyperparameters is through the prior distribution of the regression weights $\beta_j$.
-The parameter posterior distribution $p(\bm{\beta}|\bm{y})$ is $m$-dimensional, where $m$ is much smaller than the number of observations $n$.
+The parameterization in eq.
+\eqref{eq_approxf} is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which makes the posterior inference easier.
+Furthermore, all dependencies on the covariance function and the hyperparameters is through the prior distribution of the regression weights $\beta_j$.
+The posterior distribution of the parameters $p(\bm{\beta}|\bm{y})$ is a distribution over a $m$-dimensional space, where $m$ is much smaller than the number of observations $n$.
 Therefore, the parameter space is greatly reduced and this makes inference faster, especially when sampling methods are used.
@@
 
-The results from the previous section can be generalizes to a multidimensional input space with compact regular domain $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions.
+The results from the previous section can be generalizes to a multidimensional input space with compact support, $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions. 
 In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is possible combinations of univariate eigenfunctions over all dimensions.
@@
 Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
-Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have:
+Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have
 %
@@
 \noindent where, again, $\beta_j \sim \text{Normal}(0,1)$.
-The computational cost, in learning the covariance function hyperparameters, of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
+The computational cost, in learning the covariance function hyperparameters, of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions \todo{I dont think the first of part of this sentence is correct.
+Do you mean the cost of evaluating the joint distribution in a single iteration rather than "the cost of learnaing the covariance function hyperparameters?}.
 Although this still implies linear scaling in $n$, the approximation is more costly than in the univariate case, as $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and grows exponentially with respect to the number of  dimensions.

### 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
fix last Michael's comments

@@
 
-\author{Gabriel Riutort-Mayol$^{1*}$, Michael R. Andersen$^2$, Paul-Christian Bürkner$^{3}$, Aki Vehtari$^{2}$}
+\author{Gabriel Riutort-Mayol$^{1*}$, Michael R. Andersen$^2$, Paul-Christian Bürkner$^{3}$, Aki Vehtari$^{3}$}
 
 \date{ \small
-$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain. \break
+$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain \break
 $^2$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
-$^3$ Department of Computer Science, Aalto University, Finland\break
+\break
+$^3$ Department of Computer Science, Aalto University, Finland
+\break
 $^*$ Corresponding author, Email: gabriuma@gmail.com
@@
 The key element of a GP is the covariance function that defines the dependence structure between function values at different inputs.
-However, evaluating the covariance function \todo{evaluating the covariance function is not a problem, but computing the posterior is} comes with a computational issue because of the need of inverting its Gram matrix to optimize the hyperparameters.
+However, computing the posterior distribution of a GP comes with a computational issue because of the need of inverting the covariance matrix.
 Given $n$ observations in the data, the computational complexity and memory requirements of computing the posterior distribution for a GP in general scale as $O(n^3)$ and $O(n^2)$, respectively.
@@
 If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
-If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),g(\bm{x}_n))$ \todo{Is something missing here?
-The function g is missing on the left hand side.
-} or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed form solution.
+If $p(y_n|f(\bm{x}_n),g(\bm{x}_n))=\mathcal{N}(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed form solution.
 Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011}, Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} are required.
@@
 \end{align}
-\todo{Is it only purpose that only some equations have numbers?}
+\todo{Is it only purpose that only some equations have numbers? Res: Only text referenced equations have number (Gabi). }
 \noindent For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
@@
 \eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
-The computational cost of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
+The computational cost of evaluating the posterior distribution of univariate HSGPs scales as $O(nm + m)$ in a single iteration, where $n$ is the number of observations and $m$ the number of basis functions.
 
@@
 \noindent where, again, $\beta_j \sim \text{Normal}(0,1)$.
-The computational cost, in learning the covariance function hyperparameters, of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions \todo{I dont think the first of part of this sentence is correct.
-Do you mean the cost of evaluating the joint distribution in a single iteration rather than "the cost of learnaing the covariance function hyperparameters?}.
+The computational cost of evaluating the posterior distribution  of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$ in a single iteration, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
 Although this still implies linear scaling in $n$, the approximation is more costly than in the univariate case, as $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and grows exponentially with respect to the number of  dimensions.

### 2020-03-26 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits and to dos in sec 4

--- 2020-03-25 Paper/manuscript_v2/manuscript_07_01.tex
@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the non-linearity of the estimated function, which is in turn characterized by the lengthscale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the estimated function, which is in turn characterized by the lengthscale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 
-Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function.
-In this work we built these relationships for the square exponential covariance function and Matern ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix \ref{ch5_sec_periodic}.
+Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the lengthscale of the function, which depend on the particular choice of the kernel function.
+In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix \ref{ch5_sec_periodic}.
 For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects.
+\todo{The last sentence is not very clear.
+}
 
@@
 
-As explained in Section \ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$:
+As explained in Section \ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
 %
@@
 The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
-The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
+The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$
 %
@@
 
-The finite number $m$ of basis functions in the approximation needed to satisfy equation (\ref{eq_diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+\todo{I think the equation above is wrong, it should compared the areas under the curves and not the difference between the exact and approximate covariance function.
+It should the total variation difference between the true kernel and the approximate kernel, i.e.
+something like
+\begin{align*}
+\int|k(\tau) -S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) \mathrm{d}\tau |\mathrm{d}\tau 
+\end{align*}.
+How did you compute it in the code?
+}
+
+The specific numbe of basis functions  $m$ in the approximation needed to satisfy equation (\ref{eq_diff_covs}) depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
 The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
 As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
-In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi$.
+In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi = \max_{i}|x_i|$.
 Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \Omega$.
@@
 
-\noindent where $S$ (for $S > 0$) represents the half-range of the input space, and $c$ (for $c \geq 1$) is the proportional extension factor.
+\noindent where $S$ (for $S > 0$) represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
 In the following, we will refer to $c$ as the boundary factor of the approximation.
@@
 
-We start with an illustration on how the number of basis functions $m$ and boundary factor $c$ influences the accuracy of the HSGP approximations, separately.
-For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ of the kernel function, using input values from the zero-mean input domain with half-range $S=1$.
+We start by illustrating how the number of basis functions $m$ and boundary factor $c$ influence the accuracy of the HSGP approximations individually.
+For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ of the covariance function \todo{which covarariance function?}, using input values from the zero-mean input domain with half-range $S=1$.
 Several HSGP models with varying $m$ and $L$ are fitted to this data.
@@
 The higher the "wigglyness" of the function to be estimated, the more basis functions will be required.
-If $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.
+If $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries.
 

### 2020-03-26 Paper/manuscript_v2/manuscript_07_01.tex
More editing and to dos for sec 4

@@
 
-Next, we will focus on analyzing the interaction effects between these $m$ and $c$ on the performance of the approximation.
-The lengthscale and marginal variance of the kernel will no longer be fixed but rather estimated in both regular GP and HSGP models.
-Figure \ref{fig3_Post_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$.
-Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP model.
+Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
+The lengthscale and marginal variance of the covariance function will no longer be fixed but instead estimated in both regular GP and HSGP models.
+\todo{Specify how the hyperparameters are handled} Figure \ref{fig3_Post_part1} shows the posterior predictive mean of the function and the covariance function obtained after fitting the data for varying $m$ and $c$.
+Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the regular GP model.
 Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models.
 Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
-Further, the choice of $m = 10$ and $c = 1.5$ could still be an accurate enough choice.
+Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application.
 We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig3_Post_part1}.
@@
 \begin{itemize}
-\item As $c$ increases, $m$ has to increase as well (and vice versa).
-\item There exists a minimum $c$ below which a close approximation will never be achieved regardless of $m$.
+\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq. \eqref{eq_eigenvalue}, where $L$ appears in the denominator	.
+\item There exists a minimum $c$ below which a accurate approximation will never be achieved regardless of the number of basis functions $m$.
 \end{itemize}
@@
 
-Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying equation (\ref{eq_diff_covs}).
+Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
+More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying equation (\ref{eq_diff_covs}).
 Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function.
-We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function:
+We considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function
+\todo{1) I don't fully understand the statement in the sentence about.
+Which density are you referring to?
+2) I also think the is a problem with the equation below.
+The integral $\int k(\tau) \mathrm{d}\tau$ is just a constant (for fixed hyperparameters).
+Furthermore, we need to use different symbols to represent the exact covariance function and the approximate covariance function, you could use $k$ and $\tilde{k}$ for example.
+}
 %
@@
 As the lenghscale $\ell$ controls the "wigglyness" of
-the functional relationship, it strongly influences the difficulty of obtaining accurate
+the function, it strongly influences the difficulty of obtaining accurate
 inference about the function from the data. Basically, if the lengthscale estimate is accurate, 

### 2020-03-31 Paper/manuscript_v2/manuscript_07_01.tex

@@
 \usepackage[titletoc]{appendix}
+\usepackage{booktabs}
+\usepackage{boldline}
+\usepackage{colortbl}
 
@@
 \newcommand{\todo}[1]{\textcolor{blue}{\textbf{[#1]}}}
+\newcommand{\resp}[1]{\textcolor{red}{\textbf{[#1]}}}
 
@@
 
-\author{Gabriel Riutort-Mayol$^{1*}$, Michael R. Andersen$^2$, Paul-Christian Bürkner$^{3}$, Aki Vehtari$^{2}$}
+\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R. Andersen$^{3}$, Aki Vehtari$^{2}$}
 
 \date{ \small
-$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain. \break
-$^2$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
-$^3$ Department of Computer Science, Aalto University, Finland\break
+$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
+\break
+$^2$ Department of Computer Science, Aalto University, Finland
+\break
+$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
+\break
 $^*$ Corresponding author, Email: gabriuma@gmail.com
@@
 A diagnosis procedure for whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is also proposed.
-The proposed method is simple and exhibits an attractive computational complexity due to its linear structure and it is easy to implmenent in probabilistic programming frameworks.
+The proposed method is simple and exhibits an attractive computational complexity due to its linear structure and it is easy to implement in probabilistic programming frameworks.
 Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.
@@
 Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
-This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation \todo{I don't understand the sentence about autodiff and computational complexity}.
+This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation \todo{I don't understand the sentence about autodiff and computational complexity} \resp{Res: This sentence scapes me.
+I think this was written by Aki}.
 The basis function approach also provides an easy way to apply a non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
@@
 \end{align}
-\todo{Is it only purpose that only some equations have numbers?}
+
 \noindent For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
@@
 In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix \ref{ch5_sec_periodic}.
-For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects.
-\todo{The last sentence is not very clear.
-}
+For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects on the posterior function.
 
@@
 The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
-The number of basis functions can be truncated at some finite positive value $m$ such that the difference between the densities of the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$
+The number of basis functions can be truncated at some finite positive value $m$ such that the total variation difference between the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
 %
 \begin{equation}\label{eq_diff_covs}
-\Big( \int k(\tau) \mathrm{d}\tau - 
-\int \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) \mathrm{d}\tau \Big) < \varepsilon.
-\end{equation}
-
-\todo{I think the equation above is wrong, it should compared the areas under the curves and not the difference between the exact and approximate covariance function.
-It should the total variation difference between the true kernel and the approximate kernel, i.e.
-something like
-\begin{align*}
-\int|k(\tau) -S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0) \mathrm{d}\tau |\mathrm{d}\tau 
-\end{align*}.
-How did you compute it in the code?
-}
+ \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \mathrm{d}\tau  < \varepsilon.
+\end{equation}
 
@@
 We start by illustrating how the number of basis functions $m$ and boundary factor $c$ influence the accuracy of the HSGP approximations individually.
-For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ of the covariance function \todo{which covarariance function?}, using input values from the zero-mean input domain with half-range $S=1$.
+For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
 Several HSGP models with varying $m$ and $L$ are fitted to this data.
 In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictions of the estimated function and on the covariance function itself.
-For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently more linear functional forms.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function and on the covariance function itself.
+For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation, either in mean or uncertainty, and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently more linear functional forms.
 The higher the "wigglyness" of the function to be estimated, the more basis functions will be required.
-If $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 122mm 10mm, clip]{ch5_fig1_legend.pdf}}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.
-Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
+If $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, although along the whole domain in the uncertainty.
+The approximation shows to be slighter sensitive in the uncertainty than in mean.
+
+\begin{figure}
+\centering
+%\subfigure{\includegraphics[scale=0.38, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig1_legend_01.pdf}}\\
+%\vspace{-9mm}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
+%\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig1_legend.pdf}}
+\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
   \label{fig1_Post_J}
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 27mm 32mm 120mm 0mm, clip]{ch5_fig2_legend.pdf}}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.
-Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is under the blue lines.}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 10mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
+%\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig2_legend.pdf}}
+\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
   \label{fig2_Post_L}
@@
 Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The lengthscale and marginal variance of the covariance function will no longer be fixed but instead estimated in both regular GP and HSGP models.
-\todo{Specify how the hyperparameters are handled} Figure \ref{fig3_Post_part1} shows the posterior predictive mean of the function and the covariance function obtained after fitting the data for varying $m$ and $c$.
+The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using HMC methods, for both regular GP and HSGP models.
+Figure \ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function and the covariance function obtained after fitting the data for varying $m$ and $c$.
 Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the regular GP model.
@@
 \begin{figure}
-\begin{center}
-\begin{tabular}{| c | c | c | c |}
-\hline
-c  = 1.05 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & \multirow{32}{1.3cm}{\includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\
-\cline{1-3}
-c  = 1.1 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
-\cline{1-3}
-c  = 1.2 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
-\cline{1-3}
-c  = 1.5 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & \\
-\cline{1-3}
-c  = 2 & \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} &  \includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
-\cline{1-3}
-c  = 2.5 & \includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf} & \includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
-\hline
+\centering
+\begin{tabular}{ c c c }
+\arrayrulecolor{darkgray}\hline
+c = 1.05 &
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf}
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part1.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & 
+\multirow{32}{1cm}{ \includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\ 
+\arrayrulecolor{lightgray}\cline{1-2}
+c = 1.1 &
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part2.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
+\cline{1-2}
+c = 1.2 &
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part3.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
+\cline{1-2}
+c = 1.5 &
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part4.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & \\
+\cline{1-2}
+c = 2 &
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part5.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
+\cline{1-2}
+c = 2.5 &
+\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf}
+\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Sigma_part6.pdf}
+\includegraphics[scale=0.215, trim = 0mm 4mm 5mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
+\arrayrulecolor{darkgray}\hline
 \end{tabular}
-\end{center}
-\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
+\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
 Notice that the dashed red line of the regular GP can hardly be seen in the some of the plots because it is under the blue lines.}
@@
 Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying equation (\ref{eq_diff_covs}).
+More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying equation (\ref{eq_diff_covs}).
 Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function.
-We considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function
-\todo{1) I don't fully understand the statement in the sentence about.
-Which density are you referring to?
-2) I also think the is a problem with the equation below.
-The integral $\int k(\tau) \mathrm{d}\tau$ is just a constant (for fixed hyperparameters).
-Furthermore, we need to use different symbols to represent the exact covariance function and the approximate covariance function, you could use $k$ and $\tilde{k}$ for example.
-}
+We considered an approximation to be a close enough when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$:
 %

### 2020-04-01 Paper/manuscript_v2/manuscript_07_01.tex
fix a few references

@@
 In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function and on the covariance function itself.
+Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
 For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation, either in mean or uncertainty, and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently more linear functional forms.
@@
 The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using HMC methods, for both regular GP and HSGP models.
-Figure \ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function and the covariance function obtained after fitting the data for varying $m$ and $c$.
+Figure \ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the data for varying $m$ and $c$.
 Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the regular GP model.
@@
 In the HSGP model, the latent function values $f(x)$ are approximated as in equation (\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in equation (\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
-In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.

### 2020-04-09 Paper/manuscript_v2/manuscript_07_01.tex
spell check

@@
 \begin{abstract}
-Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov-Chain Monte-Carlo.
+Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
 In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
@@
 This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
-Another advantage is the reduced memory requirements of automatic differentation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
-This is because the memory requirements of automatic differentation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation \todo{I don't understand the sentence about autodiff and computational complexity} \resp{Res: This sentence scapes me.
+Another advantage is the reduced memory requirements of automatic differentiation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
+This is because the memory requirements of automatic differentiation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation \todo{I don't understand the sentence about autodiff and computational complexity} \resp{Res: This sentence escapes me.
 I think this was written by Aki}.
@@
 
-The specific numbe of basis functions  $m$ in the approximation needed to satisfy equation (\ref{eq_diff_covs}) depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+The specific number of basis functions  $m$ in the approximation needed to satisfy equation (\ref{eq_diff_covs}) depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
 The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
@@
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-As the lenghscale $\ell$ controls the "wigglyness" of
+As the lengthscale $\ell$ controls the "wigglyness" of
 the function, it strongly influences the difficulty of obtaining accurate
@@
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
-More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wiggliness in every dimensions.
+More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions.
 For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
@@
 minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}. 
-We observe that whenever the estimated lengthscale exceeds the minimially
+We observe that whenever the estimated lengthscale exceeds the minimally
 estimable lengthscale, the RMSE of the posterior predictions is small
@@
 However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure C.3 in the online material).
-In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticably (see Figures A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
+In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
 
@@
 The developed approximate GPs can be easily applied as modular components in probabilistic programming frameworks such as Stan in both Gaussian and non-Gaussian observational models.
-Using several simulated and real datasets, we have demonstrated the practical applicability and improved sampling effiency, as compared to regular GPs, of the developed method.
+Using several simulated and real datasets, we have demonstrated the practical applicability and improved sampling efficiency, as compared to regular GPs, of the developed method.
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
@@
 However, in practice, input dimensionalities larger than 3 start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
-In these high dimensional cases, the proposed approxiamte GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
+In these high dimensional cases, the proposed approximate GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
 
@@
 
-\noindent the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be re-writting as
+\noindent the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be re-writing as
 %

### 2020-04-09 Paper/manuscript_v2/manuscript_07_01.tex
edited abstract

@@
 In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
-The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, desired prediction space and smoothness of the function to be learned.
-Intuitive visualizations and useful recommendations for the values of these factors which help users to improve performance and computation are provided.
-A diagnosis procedure for whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is also proposed.
-The proposed method is simple and exhibits an attractive computational complexity due to its linear structure and it is easy to implement in probabilistic programming frameworks.
+The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, support of the prediction space, and smoothness of the latent function.
+We provide intuitive visualizations and recommendations for choosing the values of these factors, which make it easier for users to improve approximation accuracy and computational performance.
+We also propose a diagnostics for checking that the number of basis functions and the support of the prediction space are adequate given the data.
+The proposed approach is simple and exhibits an attractive computational complexity due to its linear structure, and it is easy to implement in probabilistic programming frameworks.
 Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.

### 2020-04-09 Paper/manuscript_v2/manuscript_07_01.tex
fix cites

@@
 
-\usepackage[numbers,sort&compress]{natbib}
+\usepackage[sort&compress]{natbib}
 \usepackage{subfigure}
@@
 
-\bibpunct[, ]{[}{]}{,}{}{,}{,}
+% \bibpunct[, ]{[}{]}{,}{}{,}{,}
 \renewcommand\bibfont{\fontsize{10}{12}\selectfont}
@@
 A unifying view on sparse GPs based on approximate generative methods
-is provided in \cite{quinonero2005unifying}, while a general review
-can be found in \cite{rasmussen2006gaussian}.
+is provided by \citet{quinonero2005unifying}, while a general review
+is provided by \citet{rasmussen2006gaussian}.
 \citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for an accurate approximation.
@@
 Sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving}.
-Recently, \cite{hensman2017variational} presented a variational Fourier feature approximation for GPs that was derived for the Mat{\'e}rn class of kernels.
+Recently, \citet{hensman2017variational} presented a variational Fourier feature approximation for GPs that was derived for the Mat{\'e}rn class of kernels.
 Another related method for approximating kernels relies on random Fourier features \citep{rahimi2008random,rahimi2009weighted}.
-Further, certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}.
-Furthermore, a recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation (see, e.g.
-\cite{grenander1981abstract}) can be found in \cite{JSSv090i10}.
+Certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}.
+A recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.][]{grenander1981abstract} is presented by \citet{JSSv090i10}.
 % , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).
@@
 In Section \ref{ch4_gp}, we introduce GPs, covariance functions and spectral density functions.
-In Section \ref{ch5_sec_method}, the reduced rank approximation to GPs proposed by \cite{solin2018hilbert} is described.
+In Section \ref{ch5_sec_method}, the reduced rank approximation to GPs proposed by \citet{solin2018hilbert} is described.
 In Section \ref{ch5_sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
@@
 Stationary covariance functions can be represented in terms of their spectral densities \citep{rasmussen2006gaussian}.
-In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure (\textit{Bochner's theorem}, see, e.g.
-\cite{akhiezer1993theory}).
+In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure \citep[\textit{Bochner's theorem}, see, e.g.
+][]{akhiezer1993theory}.
 If this measure has a density, it is known as the spectral density of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the \textit{Wiener-Khintchine theorem} \citep{rasmussen2006gaussian}.
@@
 
-The approximate GP method, developed by \cite{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a stationary covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
+The approximate GP method, developed by \citet{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a stationary covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
 Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to boundary conditions.
@@
 Below, we will present the main results around HSGPs relevant for practical applications.
-More details and mathematical proofs are provided in \cite{solin2018hilbert}.
-Our starting point for presenting the method is the main result obtained by \cite{solin2018hilbert} of the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
+More details and mathematical proofs are provided by \citet{solin2018hilbert}.
+Our starting point for presenting the method is the main result obtained by \citet{solin2018hilbert} of the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
 The mathematical details of this approximation are briefly presented in Appendix \ref{ch5_app_approx_covfun}.
@@
 In the HSGP model, the latent function values $f(x)$ are approximated as in equation (\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in equation (\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
-In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
@@
 The special days effects cover patterns such as possible fewer births on Halloween, Christmas or new year, and excess of births on Valentine’s Day or the days after Christmas (due, presumably, to choices involved in scheduled deliveries, along with decisions of whether to induce a birth for health reasons).
-This analysis was originally addressed in \cite{gelman2013bayesian}.
+This analysis was originally presented by \citet{gelman2013bayesian}.
 The total number of days within the period is $T=7305$ ($t=1,\dots,T$), then a regular GP model is unfeasible to be fitted on this dataset as we know inference scales $O(T^3)$ in covariance matrix inversion.
@@
 
-Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
+Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; see][]{vehtari_2012} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
 As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
@@
 This problem becomes especially severe when performing full Bayesian inference using sampling-based methods.
-In this paper, a novel approach for a low-rank representation of stationary GPs, originally proposed by \cite{solin2018hilbert}, has been implemented and analyzed in detail.
+In this paper, a novel approach for a low-rank representation of stationary GPs, originally proposed by \citet{solin2018hilbert}, has been implemented and analyzed in detail.
 The method is based on a basis function approximation via Laplace eigenfunctions.
@@
 In this section, we briefly present a summary of the mathematical details of the approximation of a stationary covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
-This statement is basically an extract of the work \cite{solin2018hilbert}, where the authors fully develop the mathematical theory behind the Hilbert Space approximation for stationary covariance functions.
+This statement is based on the work by \citet{solin2018hilbert}, who developed the mathematical theory behind the Hilbert Space approximation for stationary covariance functions.
 
@@
 However, it do has a low-rank representation.
-In this section, we first give a brief presentation of the results from \cite{solin2014explicit}, where the authors obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
+In this section, we first give a brief presentation of the results by \citep{solin2014explicit}, whi obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
 Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic square exponential covariance function.
@@
 
-In \cite{solin2014explicit}, the authors come to a cosine series expansion for the periodic covariance function (\ref{ch5_eq_cov_periodic}) as follows,
+\citet{solin2014explicit} come to a cosine series expansion for the periodic covariance function (\ref{ch5_eq_cov_periodic}) as follows,
 % 
@@
 \noindent Note that the covariance in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
-As \cite{solin2014explicit} argue, this approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
+As \citet{solin2014explicit} argue, this approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
 
@@
 This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required \citep{solin2014explicit}.
-A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function can be found in \cite{solin2014explicit}.
+A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.
 

### 2020-04-09 Paper/manuscript_v2/manuscript_07_01.tex
edits until mid 5.3

@@
 \usepackage{blindtext}
-\usepackage{titlesec}
+%\usepackage{titlesec}
 \usepackage[utf8]{inputenc}
@@
 
-\usepackage[sort&compress]{natbib}
+\usepackage{natbib}
+\bibliographystyle{plainnat}
+\bibpunct{(}{)}{;}{a}{,}{,}
 \usepackage{subfigure}
@@
 \usepackage{colortbl}
+\DeclareMathOperator{\LogNormal}{LogNormal}
 
@@
 
-\bibliographystyle{authordate1}
+%\bibliographystyle{authordate1}
 % TODO macro
@@
 In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
-The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, support of the prediction space, and smoothness of the latent function.
+The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, domain of the prediction space, and smoothness of the latent function.
 We provide intuitive visualizations and recommendations for choosing the values of these factors, which make it easier for users to improve approximation accuracy and computational performance.
-We also propose a diagnostics for checking that the number of basis functions and the support of the prediction space are adequate given the data.
+We also propose a diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
 The proposed approach is simple and exhibits an attractive computational complexity due to its linear structure, and it is easy to implement in probabilistic programming frameworks.
@@
 
-In this paper we propose and evaluate a new framework for fast and accurate inference for fully Bayesian GPs using basis function approximations.
-We focus on the basis function approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
+In this paper, we focus on a new framework for fast and accurate inference for fully Bayesian GPs using basis function approximations based on approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
 Using a basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.
@@
 This is a big advantage in terms of speed for iterative algorithms such as Markov chain Monte Carlo (MCMC).
-Another advantage is the reduced memory requirements of automatic differentiation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan}, WinBUGS \citep{lunn2000winbugs} and others.
-This is because the memory requirements of automatic differentiation rather scale with the computational complexity instead of with the usual memory requirements for the posterior density computation \todo{I don't understand the sentence about autodiff and computational complexity} \resp{Res: This sentence escapes me.
-I think this was written by Aki}.
+Another advantage is the reduced memory requirements of automatic differentiation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan} and others.
+This is because the memory requirements of automatic differentiation scale with the size of the autodiff expression tree which in direct implementations is simpler for basis function than covariance matrix based approach.
+\todo{Better?}.
 The basis function approach also provides an easy way to apply a non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
@@
 While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
-In this paper, the interactions among the key factors of the method such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the performance and accuracy of the method.
+In this paper, the interactions among the key factors of the method such as the number of basis functions, domain of the prediction space, and properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the computational performance and accuracy of the method.
 Practical recommendations are given for the values of the key factors based on intuitive graphical summaries that encode the recognized relationships.
-Our recommendations will help users to choose valid and optimized values for these factors, improving performance, and saving time of computation.
-A diagnosis of whether the chosen values for the number of basis functions and the desired prediction space are adequate to fit to the actual data is proposed.
+Our recommendations will help users to choose valid and optimized values for these factors, improving computational performance without sacrificing modeling accuracy.
+We propose also diagnostics to indicate whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
 
@@
 A brief conclusion of the work is made in Section \ref{ch5_sec_conclusion}.
-Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind this Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
+Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
 Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
@@
 
-A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, i.e.
-$\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
+A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
 GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
@@
 In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. 
-Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n|f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
-This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
+Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
+This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
 Our goal is to obtain the posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
-% That is, we want to obtain the predictive distribution $p(f^*|\mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
+% That is, we want to obtain the predictive distribution $p(f^* \mid \mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
 
@@
 \end{align*} 
-
-\noindent where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$.
+where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$.
 %By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\bm{f}$ analytically,
@@
 
-If $p(y_n|f(\bm{x}_n),\phi)=\mathcal{N}(y_n|f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
-If $p(y_n|f(\bm{x}_n),g(\bm{x}_n))=\mathcal{N}(y_n|f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n|f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed form solution.
-Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo (MCMC) \citep{brooks_2011}, Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational Bayes methods \citep{gibbs2000variational,csato2000efficient} are required.
+If $p(y_n \mid f(\bm{x}_n),\phi)=\mathcal{N}(y_n \mid f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
+If $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))=\mathcal{N}(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n \mid f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed form solution.
+Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo \citep[MCMC; ][]{brooks_2011}, Laplace approximation \citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational inference methods \citep{gibbs2000variational,csato2000efficient} are required.
 In this paper, we focus on the use of MCMC for integrating over the joint
@@
 The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and defines a correlation structure which characterizes the correlations between function values at different inputs.
-A covariance function needs be symmetric and positive semi-definite \citep{rasmussen2006gaussian}.
+A covariance function needs to be symmetric and positive semi-definite \citep{rasmussen2006gaussian}.
 A stationary covariance function is a function of $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^D$, such that it can be written $k(\bm{x},\bm{x}') = k(\bm{\tau})$, which means that the covariance is invariant to translations.
-Isotropic covariance functions are only depends on the input points through the norm of the difference, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant.
+Isotropic covariance functions depend only on the input points through the norm of the difference, $k(\bm{x},\bm{x}') = k(|\bm{x}-\bm{x}'|) = k(r), r\in {\rm I\!R}$, which means that the covariance is both translation and rotation invariant.
 The most commonly used distance between observations is the L2-norm $(|\bm{x}-\bm{x}'|_{L2})$, also known as Euclidean distance, although other types of distances can be considered.
@@
 \begin{align*}
-k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\Big(\frac{\sqrt{2\nu}r}{\ell}\Big)^{\!\nu} \! K_{\nu} \Big(\frac{\sqrt{2\nu}r}{\ell}\Big),
+k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\!\nu} \! K_{\nu} \left(\frac{\sqrt{2\nu}r}{\ell}\right),
 \end{align*}
-
-\noindent where $\nu > 0$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell > 0$ and $\alpha > 0$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
+where $\nu > 0$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell > 0$ and $\alpha > 0$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
 The particular case where $\nu=\infty$ and $\nu=3/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
@@
 \begin{align*}
-k_{\infty}(r)&=\alpha \exp\!\! \big(\! -\frac{1}{2} \frac{r^2}{\ell^2}\big),  \\
-k_{\frac{3}{2}}(r)&=\alpha\Big(1+\frac{\sqrt{3}r}{\ell}\Big) \! \exp\!\! \big(\! -\frac{\sqrt{3}r}{\ell}\big). 
+k_{\infty}(r)&=\alpha \exp \left( -\frac{1}{2} \frac{r^2}{\ell^2}\right),  \\
+k_{\frac{3}{2}}(r)&=\alpha\left(1+\frac{\sqrt{3}r}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}r}{\ell}\right). 
 \end{align*}
-
-\noindent The former is commonly known as the squared exponential (exponentiated quadratic) covariance function.
+The former is commonly known as the squared exponential or exponentiated quadratic covariance function.
 Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form
@@
 \begin{align*}
-k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp\!\!\Big(\! -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\Big),\\
-k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \bigg(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\bigg) \! \exp\!\!\Big(\! - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\Big). \\
+k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
+k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right) \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right). \\
 \end{align*}
-
-\noindent Notice that the previous expressions have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
+The previous expressions can be easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
 Using individual length-scales for each dimension turns the isotropic covariance function into a non-isotropic covariance function.
@@
 Stationary covariance functions can be represented in terms of their spectral densities \citep{rasmussen2006gaussian}.
-In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure \citep[\textit{Bochner's theorem}, see, e.g.
+In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure \citep[\textit{Bochner's theorem}; see, e.g.
 ][]{akhiezer1993theory}.
 If this measure has a density, it is known as the spectral density of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the \textit{Wiener-Khintchine theorem} \citep{rasmussen2006gaussian}.
-The spectral density functions associated with the Mat\'ern class of covariance functions is given by
+The spectral density functions associated with the Mat\'ern class of covariance functions are given by
 %
 \begin{align*}
-S_{\nu}(\omega)&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, l^{2\nu}}\Big(\frac{2\nu}{l^2}+4\pi^2\omega^2 \Big)^{\!(\nu+D/2)}
+S_{\nu}(\omega)&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
 \end{align*}
-
-\noindent in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ denotes the frequency, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
+in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ denotes the frequency, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
 The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form
@@
 S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^D  \ell^D \exp(-0.5 \ell^2 \omega^2), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\,\ell^3}\Big(\frac{3}{\ell^2}+\omega^2 \Big)^{\!-\frac{D+3}{2}}. \label{eq_specdens_32} 
+S_{\frac{3}{2}}(\omega)&= \alpha  \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
-
-\noindent For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
+For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
 %
 \begin{align*}
-S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \exp\!\!\big(\!-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \big),   \\
-S_{\frac{3}{2}}(\omega)&= \alpha \, 32\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\big(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \big)^{\!-3}.
+S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
+S_{\frac{3}{2}}(\omega)&= \alpha \, 32\,\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
 \end{align*}
@@
 
-The approximate GP method, developed by \citet{solin2018hilbert} and implemented in this paper, is based on considering the covariance operator of a stationary covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
-Then, the pseudo-differential operator is approximated with Hilbert space methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to boundary conditions.
+The approximate GP method, developed by \citet{solin2018hilbert} and further analysed in this paper, is based on considering the covariance operator of a stationary covariance function as a pseudo-differential operator constructed as a series of Laplace operators.
+Then, the pseudo-differential operator is approximated with \emph{Hilbert space} methods on a compact subset $\Omega \subset {\rm I\!R}^D$ subject to boundary conditions.
 For brevity, we will refer to these approximate Gaussian processes as HSGPs.
@@
 More details and mathematical proofs are provided by \citet{solin2018hilbert}.
-Our starting point for presenting the method is the main result obtained by \citet{solin2018hilbert} of the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
+Our starting point for presenting the method is the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
 The mathematical details of this approximation are briefly presented in Appendix \ref{ch5_app_approx_covfun}.
+
+\subsection{Unidimensional GPs} \label{ch5_sec_method_uni}
 
@@
 \end{equation} 
-
-\noindent where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
+where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
 The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
@@
 \end{align} 
-
-\noindent The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in equation (\ref{eq_eigenproblem}) are sinusoidal functions.
+The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in equation (\ref{eq_eigenproblem}) are sinusoidal functions.
 The solution to the eigenvalue problem is independent of  the specific choice of covariance function and is given by
@@
 \begin{align}
-\lambda_j&=\Big(\frac{j\pi}{2L}\Big)^{\!2}, \label{eq_eigenvalue}\\
-\phi_j(x)&=\sqrt{\frac{1}{L}}\, \sin\!\!\big(\sqrt{\lambda_j}(x+L)\big). \label{eq_eigenfunction}
+\lambda_j&=\left(\frac{j\pi}{2L}\right)^{\!2}, \label{eq_eigenvalue}\\
+\phi_j(x)&=\sqrt{\frac{1}{L}}\, \sin\!\!\left(\sqrt{\lambda_j}(x+L)\right). \label{eq_eigenfunction}
 \end{align}
@@
 \begin{equation}
-k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \, \bm{\phi}(x'), \nonumber
-\end{equation}
-
-\noindent where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is a diagonal matrix of the spectral density evaluated at the square root of the eigenvalues, i.e.
-$S_{\theta}(\sqrt{\lambda_j})$,
+k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \bm{\phi}(x'), \nonumber
+\end{equation}
+where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is a diagonal matrix of the spectral density evaluated at the square root of the eigenvalues, that is, $S_{\theta}(\sqrt{\lambda_j})$,
 %
@@
 \end{equation}
-
-\noindent where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$
+where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$
 %
@@
 \end{align}
-
-\noindent As a result, the model for $f$ can be written as
+As a result, the model for $f$ can be written as
 %
@@
 \end{equation}
-
-\noindent This equivalently leads to a linear representation of $f$ via
+This equivalently leads to a linear representation of $f$ via
 %
 \begin{equation}\label{eq_approxf}
-f(x) \approx \sum_{j}^m \big( S_{\theta}(\sqrt{\lambda_j})\big)^{\! \frac{1}{2}} \phi_j(x) \beta_j,
-\end{equation}
-
-\noindent where $\beta_j \sim \text{Normal}(0,1)$.
+f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
+\end{equation}
+where $\beta_j \sim \text{Normal}(0,1)$.
 Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
@@
 \eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
-The computational cost of evaluating the posterior distribution of univariate HSGPs scales as $O(nm + m)$ in a single iteration, where $n$ is the number of observations and $m$ the number of basis functions.
+The computational cost of evaluating the log posterior density of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
 
 The parameterization in eq.
-\eqref{eq_approxf} is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which makes the posterior inference easier.
+\eqref{eq_approxf} is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which can make the posterior inference easier \citep[see, e.g.][]{Betancourt+Girolami:2019}.
 Furthermore, all dependencies on the covariance function and the hyperparameters is through the prior distribution of the regression weights $\beta_j$.
@@
 
-The results from the previous section can be generalizes to a multidimensional input space with compact support, $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions. 
-In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is possible combinations of univariate eigenfunctions over all dimensions.
+The results from the previous section can be generalized to a multidimensional input space with compact support, $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions. 
+In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is, possible combinations of univariate eigenfunctions over all dimensions.
 The number of $D$-tuples is given by
@@
 \end{align}
-
-\noindent where $m_d$ is the number of basis function for the dimension $d$.
+where $m_d$ is the number of basis function for the dimension $d$.
 Let $\mathbb{S}\in {\rm I\!N}^{m^{\ast} \times D}$ be the matrix of all those $D$-tuples.
@@
 \begin{align}
-\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \Big\{ \Big(\frac{\pi \mathbb{S}_{jd}}{2L_d}\Big)^{\!2} \Big\}_{d=1}^D, \label{eq_eigenvalue_multi} \\
-%
-\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \sin\!\!\big(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\big). \label{eq_eigenfunction_multi}
+\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\frac{\pi \mathbb{S}_{jd}}{2L_d}\right)^{\!2} \right\}_{d=1}^D, \label{eq_eigenvalue_multi} \\
+%
+\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \sin\!\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right). \label{eq_eigenfunction_multi}
 \end{align}
-
-\noindent The approximate covariance function is then represented as
+The approximate covariance function is then represented as
 %
@@
 k(\bm{x},\bm{x}') \approx \sum_{j=1}^{m^{\ast}} 
-S^{\ast}_{\theta}\big(\sqrt{\bm{\lambda}^{\ast}_j}\big)
+S^{\ast}_{\theta}\left(\sqrt{\bm{\lambda}^{\ast}_j}\right)
 \phi^{\ast}_j(\bm{x}) \phi^{\ast}_j(\bm{x}'),
 \end{equation}
-
-\noindent where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}).
+where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}).
 We can now write the approximate series expansion of the multivariate function $f$ as
@@
 f(\bm{x}) \approx \sum_{j=1}^{m^{\ast}} 
-\big( S^{\ast}_{\theta} \big(\sqrt{\bm{\lambda}^{\ast}_j} \big)\big)^{\! \frac{1}{2}} \phi^{\ast}_j(\bm{x}) \beta_j, 
-\end{equation}
-
-\noindent where, again, $\beta_j \sim \text{Normal}(0,1)$.
-The computational cost of evaluating the posterior distribution  of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$ in a single iteration, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
+\left( S^{\ast}_{\theta} \left(\sqrt{\bm{\lambda}^{\ast}_j} \right)\right)^{\! \frac{1}{2}} \phi^{\ast}_j(\bm{x}) \beta_j, 
+\end{equation}
+where, again, $\beta_j \sim \text{Normal}(0,1)$.
+The computational cost of evaluating the log posterior density of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
 Although this still implies linear scaling in $n$, the approximation is more costly than in the univariate case, as $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and grows exponentially with respect to the number of  dimensions.
@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the estimated function, which is in turn characterized by the lengthscale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the estimated function, which is in turn is characterized by the lengthscale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 \end{equation*} 
-
-\noindent where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
+where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
 The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
@@
 
-The specific number of basis functions  $m$ in the approximation needed to satisfy equation (\ref{eq_diff_covs}) depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+The specific number of basis functions  $m$ needed to satisfy equation (\ref{eq_diff_covs}) depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
 The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
@@
 \end{equation} 
-
-\noindent where $S$ (for $S > 0$) represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
+where $S$ (for $S > 0$) represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
 In the following, we will refer to $c$ as the boundary factor of the approximation.
@@
 Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
-For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation, either in mean or uncertainty, and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently more linear functional forms.
-The higher the "wigglyness" of the function to be estimated, the more basis functions will be required.
-If $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, although along the whole domain in the uncertainty.
-The approximation shows to be slighter sensitive in the uncertainty than in mean.
+For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation, either in mean or uncertainty, and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently less wiggly functional forms.
+The higher the wigglyness of the function to be estimated, the more basis functions will be required.
+If $m$ is fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, although along the whole domain in the uncertainty.
+The approximation error for the variance tends to be bigger for the variance than for the mean.
 
@@
 Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using HMC methods, for both regular GP and HSGP models.
+The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}, for both regular GP and HSGP models.
 Figure \ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the data for varying $m$ and $c$.
@@
 \caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
-Notice that the dashed red line of the regular GP can hardly be seen in the some of the plots because it is under the blue lines.}
+The dashed red line of the regular GP can hardly be seen in the some of the plots because it is under the blue lines.}
   \label{fig3_Post_part1}
@@
 
-\noindent Alternatively, these figures could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
+Alternatively, these figures could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
 Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$.
@@
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-As the lengthscale $\ell$ controls the "wigglyness" of
+As the lengthscale $\ell$ controls the wigglyness of
 the function, it strongly influences the difficulty of obtaining accurate
@@
 we can expect the HSGP approximation to be accurate as well.
-Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
-If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough.
+
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
+If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation is assumed to be good.
 If, however, $\hat{\ell}$ does not exceed recommended minimum lengthscale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
-We may also use this diagnostic in a iterative procedure.
-Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
-As mentioned earlier, $c$ cannot be decreased as much as desired because it is restricted by the lengthscale, so increasing $m$ may usually the preferred approach.
+We may also use this diagnostic in an iterative procedure by
+starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if
+the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
+As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the lengthscale.
+Thus, increasing $m$ may usually the preferred approach.
 
@@
 \caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
-Notice that the dashed red line of the regular GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.}
+The dashed red line of the regular GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.}
   \label{fig7_posterior_varing_lscale_part1}
@@
 \end{align*}
-
-\noindent where $I$ represents the identity matrix and $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function values to the noisy data.
+where $I$ represents the identity matrix and $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function values to the noisy data.
 The previous formulation corresponds to the latent form of a GP model.
@@
 
-A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using sampling methods such as HMC, would be its marginalized form,
+A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using MCMC sampling methods, would be its marginalized form,
 %
@@
 \end{equation*}
-
-\noindent where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
+where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
 
@@
 
-Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and \textit{knots} $> 10$, or even for $D=2$ and \textit{knots} $> 40$ (see Figures B.3-right and C.3 in the online material).
+Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures B.3-right and C.3 in the online material).
 
@@
 The special days effects cover patterns such as possible fewer births on Halloween, Christmas or new year, and excess of births on Valentine’s Day or the days after Christmas (due, presumably, to choices involved in scheduled deliveries, along with decisions of whether to induce a birth for health reasons).
-This analysis was originally presented by \citet{gelman2013bayesian}.
-The total number of days within the period is $T=7305$ ($t=1,\dots,T$), then a regular GP model is unfeasible to be fitted on this dataset as we know inference scales $O(T^3)$ in covariance matrix inversion.
-Therefore, an approximate approach has to be used to fit a GP model on this data.
-We will use the HSGP method as well as the low-rank GP model with a periodic covariance function introduced in Appendix \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
+\citet{gelman2013bayesian} presented an analysis using regular GP and maximum a posteriori inference.
+As the total number of days within the period is $T=7305$ ($t=1,\dots,T$), a full Bayesian inference with MCMC for a regular GP model is memory and time consuming.
+We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
 
@@
 \end{equation*}
-
-\noindent The mean function $\mu(t)$ will be defined as an additive model in the form: 
+The mean function $\mu(t)$ will be defined as an additive model in the form: 
 %
@@
 \begin{equation*}
-f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\Big(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\Big), 
-\end{equation*}
-
-\noindent which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
+f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
+\end{equation*}
+which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
 
@@
 \begin{equation*}
-f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\Big(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\Big), 
-\end{equation*}
-
-\noindent which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.
+f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+\end{equation*}
+% which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.
 
@@
 \begin{equation*}
-f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\Big(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\Big), 
-\end{equation*}
-
-\noindent which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.
-
-The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{piironen2017sparsity}:
+f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
+\end{equation*}
+% which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.
+
+The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
 %
@@
 \end{equation*}
-
-\noindent A horse-shoe prior allows for sparse distributed effects.
+A horse-shoe prior allows for sparse distributed effects.
 Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
@@
 
-GP priors have been defined over the components $f_1(t)$, $f_2(t)$ and $f_3(t)$.
-Then, low-rank representations of the GP priors have to be used in the modeling and inference.
-The component $f_1(t)$ will be approximated using the HSGP model.
-Thus, the function values $f_1(t)$ are approximated as in equation (\ref{eq_approxf}), with the squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$  and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
-
-The year effects $f_2(t)$ and week effects $f_3(t)$, as they use a periodic covariance function, they do no fit under the main framework of the HSGP approximation covered in this chapter.
+The component $f_1(t)$ will be approximated using the HSGP
+model and the function values $f_1(t)$ are approximated as in
+equation (\ref{eq_approxf}), with the squared exponential spectral
+density $S$ as in equation (\ref{eq_specdens_inf}), and eigenvalues
+$\lambda_j$ and eigenfunctions $\phi_j$ as in equations
+(\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+We use $m=30$ basis functions and a boundary factor $c=1.5$.
+The lengthscale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
+This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
+
+The year effects $f_2(t)$ and week effects $f_3(t)$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper.
 However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix \ref{ch5_sec_periodic}).
 Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in equation (\ref{ch5_eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in equation (\ref{ch5_eq_q_2}).
-
-For the component $f_1(t)$, $m=30$ basis functions and a boundary factor $c=1.5$ were used.
-The lengthscale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
-This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
-
-For the components $f_2(t)$ and $f_3(t)$, $J=10$ cosine terms were used.
-The  lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are accurate enough.
-
-Figure \ref{ch5_fig27_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and year patterns $f_2(t)$ for the whole period, jointly with the observed data.
-Figure \ref{ch5_fig27_posteriors_oneyear_birthday} show the process for one year (1972) only.
+We use $J=10$ cosine terms.
+The  lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
+
+Figure \ref{ch5_fig27_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and yearly pattern $f_2(t)$ for the whole period, jointly with the observed data.
+Figure \ref{ch5_fig27_posteriors_oneyear_birthday} shows the model for one year (1972) only.
 In this figure, the special days effects $f_4(t)$ in the year can be clearly represented.
@@
 
-As the WBC measurements were strictly positive and highly skewed, we fit the model to its logarithm.
+As the WBC measurements were strictly positive and highly skewed, logarithm transformation is used for it.
 Continuous predictors were normalized to have zero mean and unit standard deviation.
@@
 \begin{equation*}
-p(t_i)= \mathrm{LogNormal}(t_i|f(\bm{x}_i),\sigma^2).
-\end{equation*}
-
-As we do not have a model for the censoring process, we do not have a full observation model, and the observational model for the censored data $t_i$ is assumed to be the complementary cumulative normal probability distribution:
-%
-\begin{equation*}
-p(y_i > t_i)= \int_{t_i}^{\infty} \mathrm{LogNormal}(y_i|f(\bm{x}_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \Big( \frac{\mathrm{log}(y_i)-f(\bm{x}_i)}{\sigma} \Big),
-\end{equation*}
-
-\noindent where $y_i$ denotes the uncensored time.
+p(t_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
+\end{equation*}
+
+As we do not have a model for the censoring process, we do not have a full observation model, and the observational model for the censored data $t_i$ is assumed to be the complementary cumulative log-normal probability distribution:
+%
+\begin{equation*}
+p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \left( \frac{\mathrm{log}(y_i)-f(\bm{x}_i)}{\sigma} \right),
+\end{equation*}
+where $y_i$ denotes the uncensored time.
 
@@
 \end{equation*}
-
-\noindent where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
+where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
 Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ are multivariate Gaussian distributed with mean function $\mu(\cdot)$ and covariance matrix $K$, where $\mu(\bm{x}_i)=c + \bm{\beta}\bm{x}_i$ and $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,n$.
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-Notice that a scalar lengthscale is considered in the multivariate covariance function assuming an isotropic function.
+A scalar lengthscale is used in the multivariate covariance function assuming an isotropic function.
 
@@
 \end{align}
-
-\noindent with
+with
 %
@@
 \end{figure}
-
-\noindent In the former equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($x_2=\text{'}2\text{'}$).
+In the former equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($x_2=\text{'}2\text{'}$).
 
@@
 \end{equation*}
-
-\noindent where $\bm{w}$ is in the frequency domain.
+where $\bm{w}$ is in the frequency domain.
 The operator $\mathcal{K}$ will be translation invariant if the covariance function is stationary.
@@
 \end{equation}
-
-\noindent The Fourier transform of the Laplace operator $\nabla^2$ is $-||\bm{w}||$, thus the Fourier transform of $S(||\bm{w}||)$ is
+The Fourier transform of the Laplace operator $\nabla^2$ is $-||\bm{w}||$, thus the Fourier transform of $S(||\bm{w}||)$ is
 %
@@
 \end{equation}
-
-\noindent defining a pseudo-differential operator as a series of Laplace operators.
+defining a pseudo-differential operator as a series of Laplace operators.
 
@@
 \end{equation*} 
-
-\noindent then the formal kernel can be represented as 
+then the formal kernel can be represented as 
 %
@@
 \end{equation*}
-
-\noindent where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
+where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
 Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
@@
 \end{align*}  
-
-\noindent Because $-\nabla^2$ is a positive definite Hermitian operator, the set of eigenfunctions $\phi_j(\cdot)$ are orthonormal with respect to the inner product
+Because $-\nabla^2$ is a positive definite Hermitian operator, the set of eigenfunctions $\phi_j(\cdot)$ are orthonormal with respect to the inner product
 %
@@
 \end{align*} 
-
-\noindent that is,
+that is,
 %
@@
 \end{align*} 
-
-\noindent and all the eigenvalues $\lambda_j$ are real and positive.
+and all the eigenvalues $\lambda_j$ are real and positive.
 
@@
 \end{eqnarray} 
-
-\noindent which are again to be interpreted to mean that
+which are again to be interpreted to mean that
 %
@@
 \end{equation*} 
-
-\noindent This implies that we also have
+This implies that we also have
 %
@@
 \end{equation} 
-
-\noindent By letting $||\bm{w}||^2=\lambda_j$ the spectral density in Equation (\ref{ch5_eq_S}) becomes
+By letting $||\bm{w}||^2=\lambda_j$ the spectral density in Equation (\ref{ch5_eq_S}) becomes
 %
@@
 \end{equation*}
-
-\noindent and substituting in equation (\ref{ch5_eq_k}) then leads to the final searched approximation
+and substituting in equation (\ref{ch5_eq_k}) then leads to the final searched approximation
 %
@@
 \end{equation} 
-
-\noindent where $S(\cdot)$ is the spectral density of the covariance function, $\lambda_j$ is the $j$th eigenvalue and $\phi_j(\cdot)$ the eigenfunction of the Laplace operator in a given domain.
+where $S(\cdot)$ is the spectral density of the covariance function, $\lambda_j$ is the $j$th eigenvalue and $\phi_j(\cdot)$ the eigenfunction of the Laplace operator in a given domain.
 
@@
 \end{equation}
-
-\noindent where $\alpha$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
+where $\alpha$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
 
@@
 \end{equation}
-
-\noindent which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ of the previous expression are
+which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ of the previous expression are
 %
@@
 \end{equation}
-
-\noindent where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
+where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
 %
@@
 \end{equation}
-
-\noindent Note that the covariance in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
+The covariance in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
 As \citet{solin2014explicit} argue, this approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
@@
 \end{equation} 
-
-\noindent for $j=1,2,\cdots,J$, and where the $\text{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
+for $j=1,2,\cdots,J$, and where the $\text{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
 This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required \citep{solin2014explicit}.
@@
 We have considered an approximation to be close enough in terms of satisfying equation (\ref{eq_diff_covs}) with $\varepsilon=0.5\%$.
-Notice that since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
+Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
 
@@
 \end{equation*}
-
-\noindent the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be re-writing as
+the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be re-writing as
 %
@@
 \end{equation}
-
-\noindent where $\tau=x-x'$.
+where $\tau=x-x'$.
 With this approximation for the periodic squared exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \mathcal{GP}(0,k(x,x')$ equivalently leads to a linear representation of $f(\cdot)$ via
@@
 \end{equation}
-
-\noindent where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\dots,2J+1$.
+where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\dots,2J+1$.
 The cosine $\cos(j\omega_0 x)$ and sinus $\sin(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.

### 2020-04-10 Paper/manuscript_v2/manuscript_07_01.tex
edits from 5.3 to end

@@
 \DeclareMathOperator{\LogNormal}{LogNormal}
+\DeclareMathOperator{\GP}{\mathcal{GP}}
 
@@
 
-We assume a GP prior for $f \sim \mathcal{GP}(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \rightarrow {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \rightarrow {\rm I\!R}$ are the mean and covariance functions, respectively,
+We assume a GP prior for $f \sim \GP(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \rightarrow {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \rightarrow {\rm I\!R}$ are the mean and covariance functions, respectively,
 %
@@
 %\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig1_legend.pdf}}
-\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
+\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
   \label{fig1_Post_J}
@@
 %\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig2_legend.pdf}}
-\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
+\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
   \label{fig2_Post_L}
@@
 Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}, for both regular GP and HSGP models.
+The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}, for both exact GP and HSGP models.
 Figure \ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the data for varying $m$ and $c$.
-Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the regular GP model.
-Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models.
+Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
+Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the exact GP model and the HSGP models.
 Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
@@
 \end{tabular}
-\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
-The dashed red line of the regular GP can hardly be seen in the some of the plots because it is under the blue lines.}
+\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
+The dashed red line of the exact GP can hardly be seen in the some of the plots because it is under the blue lines.}
   \label{fig3_Post_part1}
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig4_MSE_vs_c.pdf}}
-\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model.
+\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the exact GP model.
 RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$ (left).
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 28mm 119mm 0mm, clip]{ch5_fig5_legend.pdf}}
-\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
+\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both exact GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig5_lscale_vs_J}
@@
 In this example, we make a comparison of the lengthscale estimates
-obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
+obtained from the exact GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
 For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying lengthscale values.
@@
 
-Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of both regular GP and HSGP models fitted to those datasets.
-The lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
-As noted previously, an accurate estimate of the lengthscale can be a good indicator of a close approximation of the HSGP model to the regular GP model.
-Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP models, as a function of the lengthscale and number of basis functions.
+Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of both exact GP and HSGP models fitted to those datasets.
+The lengthscale estimates as obtained by exact GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
+As noted previously, an accurate estimate of the lengthscale can be a good indicator of a close approximation of the HSGP model to the exact GP model.
+Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the exact GP models, as a function of the lengthscale and number of basis functions.
 
@@
 \vspace{-3mm}
-\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
-The dashed red line of the regular GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.}
+\caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
+The dashed red line of the exact GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.}
   \label{fig7_posterior_varing_lscale_part1}
@@
 \end{flushleft}
-\caption{Data-generating functional lengthscales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates $(\hat{\ell})$ from the regular GP and HSGP models.
+\caption{Data-generating functional lengthscales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates $(\hat{\ell})$ from the exact GP and HSGP models.
 95\% confident intervals of the lengthscale estimates are plotted as dot lines.
@@
 
-The regular GP model for fitting this simulated dataset $\bm{y}$ can be written as follows,
+The exact GP model for fitting this simulated dataset $\bm{y}$ can be written as follows,
 %
@@
 \bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
-f(x) &\sim \mathcal{GP}(0, k(x, x', \theta)),
+f(x) &\sim \GP(0, k(x, x', \theta)),
 \end{split}
@@
 In the HSGP model, the latent function values $f(x)$ are approximated as in equation (\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in equation (\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
-In order to do model comparison, in addition to the regular GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
 
-Figure \ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the regular GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
+Figure \ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
 The true data-generative function and the noisy observations are also plotted.
@@
 The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
-The posteriors of the three models, regular GP, HSGP and spline, are pretty similar within the interpolation input space.
-However, when extrapolating the spline model solution clearly differs from the regular GP and HSGP models as well as the actual observations.
+The posteriors of the three models, exact GP, HSGP and spline, are pretty similar within the interpolation input space.
+However, when extrapolating the spline model solution clearly differs from the exact GP and HSGP models as well as the actual observations.
 
@@
 \includegraphics[width=\textwidth]{ch5_fig10_Posteriors_exI.pdf}
-\caption{Posterior predictive means of the proposed HSGP model, the regular GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
+\caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
   \label{ch5_fig10_Posteriors_exI}
@@
 The SRMSE is computed against the data-generating model.
-From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the regular GP model for interpolating and extrapolating data.
+From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the exact GP model for interpolating and extrapolating data.
 However, the spline model does not extrapolate data properly.
@@
 Figure \ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The HSGP model is on average roughly 400 times faster than the regular GP and 10 times faster than the spline model, for this particular application with a univariate input space.
+The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model, for this particular application with a univariate input space.
 Also, the increase in computation time as a function of the number of basis functions in a univariate input space is relatively slight.
@@
 From those examples, it can be seen how computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see equation (\ref{eq_m_multi})).
-Even though, in a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the regular GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
+Even though, in a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
 However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure C.3 in the online material).
@@
 The special days effects cover patterns such as possible fewer births on Halloween, Christmas or new year, and excess of births on Valentine’s Day or the days after Christmas (due, presumably, to choices involved in scheduled deliveries, along with decisions of whether to induce a birth for health reasons).
-\citet{gelman2013bayesian} presented an analysis using regular GP and maximum a posteriori inference.
-As the total number of days within the period is $T=7305$ ($t=1,\dots,T$), a full Bayesian inference with MCMC for a regular GP model is memory and time consuming.
+\citet{gelman2013bayesian} presented an analysis using exact GP and maximum a posteriori inference.
+As the total number of days within the period is $T=7305$ ($t=1,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
 We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
@@
 \begin{equation*}
-f_1(t) \sim \mathcal{GP}(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
+f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
 \end{equation*}
@@
 \begin{equation*}
-f_2(t) \sim \mathcal{GP}(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
 \end{equation*}
@@
 \begin{equation*}
-f_3(t) \sim \mathcal{GP}(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
+f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
 \end{equation*}
@@
 The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
-The data set consists of survival times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for $n=1043$ cases ($i=1,\dots,n$).
+The data set consists of survival and censoring times $t_i$ and censoring indicator $z_i$ (0 for observed and 1 for censored) for $n=1043$ cases ($i=1,\dots,n$).
 About 16\% of cases were censored.
@@
 
-As the WBC measurements were strictly positive and highly skewed, logarithm transformation is used for it.
+As the WBC predictor values were strictly positive and highly skewed, logarithm transformation is used for it.
 Continuous predictors were normalized to have zero mean and unit standard deviation.
@@
 
-As we do not have a model for the censoring process, we do not have a full observation model, and the observational model for the censored data $t_i$ is assumed to be the complementary cumulative log-normal probability distribution:
+We do not have a full observation model, as we do not have a model for the censoring process.
+We use the complementary cumulative log-normal probability distribution for the censored data conditionally on the censoring time $t_i$:
 %
@@
 \end{equation*}
-where $y_i$ denotes the uncensored time.
+where $y_i>t_i$ denotes the unobserved survival time.
 
@@
 \begin{equation*} 
-f(\bm{x}) \sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta)),
+f(\bm{x}) \sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta)),
 \end{equation*}
 where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ are multivariate Gaussian distributed with mean function $\mu(\cdot)$ and covariance matrix $K$, where $\mu(\bm{x}_i)=c + \bm{\beta}\bm{x}_i$ and $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,n$.
+% Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ are multivariate Gaussian distributed with mean function $\mu(\cdot)$ and covariance matrix $K$, where $\mu(\bm{x}_i)=c + \bm{\beta}\bm{x}_i$ and $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,n$.
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
@@
 
-Due to the predictor \textit{sex} ($x_2$) being a categorical variable ('1' for female and '2' for male), we apply dummy/treatment coding for the GP functions, in a similar way such coding is applied in linear models.
-A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise.
-More formally, this construction reads as follows:
+Due to the predictor \textit{sex} ($x_2$) being a categorical variable ('1' for female and '2' for male), we apply indicator variable coding for the GP functions, in a similar way such coding is applied in linear models \citep{Gelman+Hill+Vehtari:2020:ROS}.
+A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise:
 %
 \begin{align} 
-f(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) + g(\bm{x}), \label{ch5_eq_gpml1}
+f(\bm{x}) &\sim \GP\left(c + \bm{\beta}\bm{x},\, k(\bm{x},\bm{x}', \theta_1)\right) + g(\bm{x}), \label{ch5_eq_gpml1}
 \end{align}
@@
 	\begin{array}{ll}
-		0 & \;\; \mbox{if } \;\; x_2\!=\!\text{'}1\text{'}  \\
-		\mathcal{GP}(0, k(\bm{x},\bm{x}', \theta_2))  & \;\; \mbox{if } \;\; x_2\!=\!\text{'}2\text{'} 
+		0 & \;\; \mbox{if } \;\; x_2=1  \\
+		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right)  & \;\; \mbox{if } \;\; x_2=2 \,.
 	\end{array}
@@
 \end{align}
-%
-%
 
 %\begin{align} 
-%f(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
-%g(\bm{x}) &\sim \mathcal{GP}(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2 \,|\, x_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
+%f(\bm{x}) &\sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
+%g(\bm{x}) &\sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2 \,|\, x_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
 %f(\bm{x} \, |\, x_2\!=\!2) &= f(\bm{x} \,|\, x_2\!=\!1) + g(\bm{x}). \nonumber
@@
 \end{figure}
-In the former equations, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a specific GP function restricted to the male sex ($x_2=\text{'}2\text{'}$).
+Above, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a GP function specific to the male sex ($x_2=2$).
 
@@
 
-Figure \ref{ch5_fig21_posteriors_leukemia} shows estimated conditional comparison of each predictor with all others fixed to their mean values.
+Figure \ref{ch5_fig21_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
 These posterior estimates correspond to the HSGP model with $m=10$ basis functions and $c=3$ boundary factor.
-The model has found smooth non-linear patterns and the right bottom subplot also shows that the conditional comparison associated with WBC has an interaction with TDI.
-
-Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; see][]{vehtari_2012} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
+There are clear  non-linear patterns and the right bottom subplot also shows that the conditional function associated with WBC has an interaction with TDI.
+
+Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
 As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
@@
 The method has an attractive computational cost as it effectively approximates GPs by linear models, which is also an attractive property in modular probabilistic programming programming frameworks.
-The dominating cost per log density evaluation (during sampling) is $O(nm+m)$, which is a big benefit in comparison to $O(n^3)$ of a regular GP model.
+The dominating cost per log density evaluation (during sampling) is $O(nm+m)$, which is a big benefit in comparison to $O(n^3)$ of a exact GP model.
 The obtained design matrix is independent of hyperparameters and therefore only needs to be constructed once, at cost $O(nm)$.
@@
 
-The developed approximate GPs can be easily applied as modular components in probabilistic programming frameworks such as Stan in both Gaussian and non-Gaussian observational models.
-Using several simulated and real datasets, we have demonstrated the practical applicability and improved sampling efficiency, as compared to regular GPs, of the developed method.
+The developed approximate GPs can be easily applied as modular components in probabilistic programming frameworks such as Stan in both Gaussian and non-Gaussian observation models.
+Using several simulated and real datasets, we have demonstrated the practical applicability and improved sampling efficiency, as compared to exact GPs, of the developed method.
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
@@
 
-From the Bochner’s and Wiener-Khintchine theorem, the spectral density of a stationary covariance function $k(\bm{x},\bm{x}') = k(\bm{\tau})$, $\bm{\tau}=(\bm{x}-\bm{x}')$, is the Fourier transform of the covariance function,
+From the Bochner’s and Wiener-Khintchine theorems, the spectral density of a stationary covariance function $k(\bm{x},\bm{x}') = k(\bm{\tau})$, $\bm{\tau}=(\bm{x}-\bm{x}')$, is the Fourier transform of the covariance function,
 %
@@
 where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator.
-Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (another boundary condition could be used as well):
+Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (other boundary conditions could be used as well):
 %
@@
 \end{equation*}
-and substituting in equation (\ref{ch5_eq_k}) then leads to the final searched approximation
+and substituting in equation (\ref{ch5_eq_k}) then leads to the final form
 %
@@
 
-A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study.
-However, it do has a low-rank representation.
-In this section, we first give a brief presentation of the results by \citep{solin2014explicit}, whi obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
+A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study, but it has also a low-rank representation.
+In this section, we first give a brief presentation of the results by \citet{solin2014explicit}, who obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
 Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic square exponential covariance function.
@@
 \begin{equation} \label{ch5_eq_cov_periodic}
-k(\bm{\tau})= \alpha \exp\!\!\Big(\! -\frac{2\sin^{\!2}\!(\omega_0\frac{\bm{\tau}}{2})}{\ell^2}\Big),
+k(\bm{\tau})= \alpha \exp\left( -\frac{2\sin^{2}(\omega_0\frac{\bm{\tau}}{2})}{\ell^2}\right),
 \end{equation}
@@
 
-\citet{solin2014explicit} come to a cosine series expansion for the periodic covariance function (\ref{ch5_eq_cov_periodic}) as follows,
+\citet{solin2014explicit} derive a cosine series expansion for the periodic covariance function (\ref{ch5_eq_cov_periodic}) as follows,
 % 
@@
 \end{equation}
-which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ of the previous expression are
+which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ 
 %
@@
 The covariance in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
-As \citet{solin2014explicit} argue, this approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
+This approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
 
@@
 for $j=1,2,\cdots,J$, and where the $\text{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
-This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required \citep{solin2014explicit}.
+This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required.
 A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.
@@
 \end{equation*}
-the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be re-writing as
+the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be written as
 %
 \begin{equation} \label{ch5_eq_cov_periodic_taylor_approx_2}
-k(x,x')= \sigma^2 \Big( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \Big).
-\end{equation}
-where $\tau=x-x'$.
-With this approximation for the periodic squared exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \mathcal{GP}(0,k(x,x')$ equivalently leads to a linear representation of $f(\cdot)$ via
+k(x,x')= \sigma^2 \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
+\end{equation}
+With this approximation for the periodic squared exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \GP\left(0,k(x,x')\right)$ equivalently leads to a linear representation of $f(\cdot)$ via
 %
 \begin{equation} \label{ch5_eq_f_period}
-f(x) \approx \sigma \Big( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \Big),
+f(x) \approx \sigma \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
 \end{equation}

### 2020-04-10 Paper/manuscript_v2/manuscript_07_01.tex
Abstract and Sec. 1

@@
 \usepackage{listings}
-\usepackage{slashbox}
+%\usepackage{slashbox}
 \usepackage{caption}
 \captionsetup{font=footnotesize}
+
+\usepackage[colorlinks=true]{hyperref}
 
@@
 \begin{abstract}
-Gaussian processes are powerful non-parametric probabilistic models for stochastic functions, however they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
-In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis functions approximation via Laplace eigenfunctions for stationary covariance functions.
+Gaussian processes are powerful non-parametric probabilistic models for stochastic functions.
+However they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
+In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions.
 The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, domain of the prediction space, and smoothness of the latent function.
 We provide intuitive visualizations and recommendations for choosing the values of these factors, which make it easier for users to improve approximation accuracy and computational performance.
-We also propose a diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
+We also propose diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
 The proposed approach is simple and exhibits an attractive computational complexity due to its linear structure, and it is easy to implement in probabilistic programming frameworks.
@@
 The low-rank approximation with $m \ll n$ {\it inducing points} implies reduced memory requirements of $O(nm)$ and corresponding computational complexity of $O(nm^2)$.
-A unifying view on sparse GPs based on approximate generative methods
-is provided by \citet{quinonero2005unifying}, while a general review
-is provided by \citet{rasmussen2006gaussian}.
+A unifying view on sparse GPs based on approximate generative methods is provided by \citet{quinonero2005unifying}, while a general review is provided by \citet{rasmussen2006gaussian}.
 \citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for an accurate approximation.
@@
 More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.
-
 An alternative class of low-rank approximations is based on forming a basis function approximation with $m \ll n$ basis functions.
-The basis functions are usually presented explicitly, but can also be used to form a low rank covariance matrix approximation.
+The basis functions are usually presented explicitly, but can also be used to form a low-rank covariance matrix approximation.
 Common basis function approximations rest on the spectral analysis and series expansions of GPs \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}.
@@
 Certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}.
-A recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.][]{grenander1981abstract} is presented by \citet{JSSv090i10}.
+Recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.,][]{grenander1981abstract} is presented by \citet{JSSv090i10}.
 % , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).
@@
 The linear model structure makes GPs easy to implement as a building block in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if the approximation specific computation is simple.
-Furthermore, a linear representation of a GP makes it easier to be used as latent function
-in non-Gaussian observational models allowing for more modelling flexibility.
+Furthermore, a linear representation of a GP makes it easier to be used as latent function in non-Gaussian observational models allowing for more modelling flexibility.
 The basis function approximation via Laplace eigenfunctions can be made arbitrary accurate and the trade-off between computational complexity and approximation accuracy can easily be controlled.
@@
 This is because the memory requirements of automatic differentiation scale with the size of the autodiff expression tree which in direct implementations is simpler for basis function than covariance matrix based approach.
-\todo{Better?}.
 The basis function approach also provides an easy way to apply a non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
@@
 Our recommendations will help users to choose valid and optimized values for these factors, improving computational performance without sacrificing modeling accuracy.
-We propose also diagnostics to indicate whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
+We also propose  diagnostics to indicate whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
 
@@
 Several illustrative examples of the performance and applicability of the method are shown using both simulated and real datasets.
-All examples are accompanied by the corresponding Stan %and R
-code.
-
-Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, and GPflow \citep{GPflow2017}), each provide efficient implementations only for a restricted range of GP-based models.
+All examples are accompanied by the corresponding Stan code.
+Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, and GPflow \citep{GPflow2017}, each provide efficient implementations only for a restricted range of GP-based models.
 In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
@@
 The remainder of the paper is structured as follows.
-In Section \ref{ch4_gp}, we introduce GPs, covariance functions and spectral density functions.
-In Section \ref{ch5_sec_method}, the reduced rank approximation to GPs proposed by \citet{solin2018hilbert} is described.
-In Section \ref{ch5_sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
-Several case studies in which we fit
-exact and approximate GPs to real and simulated data are provided in Section \ref{ch5_sec_cases}.
-A brief conclusion of the work is made in Section \ref{ch5_sec_conclusion}.
-Appendix \ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
+In Section~\ref{ch4_gp}, we introduce GPs, covariance functions and spectral density functions.
+In Section~\ref{ch5_sec_method}, the reduced-rank approximation to GPs proposed by \citet{solin2018hilbert} is described.
+In Section~\ref{ch5_sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
+Several case studies in which we fit exact and approximate GPs to real and simulated data are provided in Section~\ref{ch5_sec_cases}.
+A brief conclusion of the work is made in Section~\ref{ch5_sec_conclusion}.
+Appendix~\ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
 Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.

### 2020-04-10 Paper/online_supplemental_material/online_material.tex
update online material

@@
 \begin{tabular}{|c|*{6}{c|}}\hline
-\backslashbox{\small Estimate}{\small True} & \multicolumn{1}{p{1cm}|}{\centering k = 1} &  \multicolumn{1}{p{1cm}|}{\centering k = 2} & \multicolumn{1}{p{1.5cm}|}{\centering k = 3} & \multicolumn{1}{p{1cm}|}{\centering k = 4} & \multicolumn{1}{p{1cm}|}{\centering k = 5} & \multicolumn{1}{p{1cm}|}{\centering } \\
+\backslashbox{\small Estimate}{\small True} & \multicolumn{1}{p{1cm}|}{\centering k = 1} &  \multicolumn{1}{p{1cm}|}{\centering k = 2} & \multicolumn{1}{p{1cm}|}{\centering k = 3} & \multicolumn{1}{p{1cm}|}{\centering k = 4} & \multicolumn{1}{p{1cm}|}{\centering k = 5} & \multicolumn{1}{p{1cm}|}{\centering } \\
 \hline \multicolumn{1}{|p{3cm}|}{ \centering k = 1} &90&39&14&3&11&42\%\\

### 2020-04-10 Paper/manuscript_v2/manuscript_07_01.tex
add Arno to author list
Sec. 2

@@
 
-We assume a GP prior for $f \sim \GP(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \rightarrow {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \rightarrow {\rm I\!R}$ are the mean and covariance functions, respectively,
+We assume a GP prior for $f \sim \GP(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \to {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \to {\rm I\!R}$ are the mean and covariance functions, respectively,
 %
@@
 The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$.
-Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution
-$\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$,
- where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
+Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution $\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$, where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
 In the following, we focus on zero-mean Gaussian processes, that is set $\mu(\bm{x}) = 0$.
@@
 \end{align*}
+%
 where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$.
@@
 If $p(y_n \mid f(\bm{x}_n),\phi)=\mathcal{N}(y_n \mid f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
-If $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))=\mathcal{N}(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n \mid f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed form solution.
+If $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))=\mathcal{N}(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n \mid f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed-form solution.
 Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo \citep[MCMC; ][]{brooks_2011}, Laplace approximation \citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational inference methods \citep{gibbs2000variational,csato2000efficient} are required.
-In this paper, we focus on the use of MCMC for integrating over the joint
-posterior. MCMC is usually not the fastest approach, but it is flexible and allows
-accurate inference and uncertainty estimates for general models in probabilistic programming
-settings. We consider the computational costs of GPs specifically from
-this point of view.
+In this paper, we focus on the use of MCMC for integrating over the joint posterior.
+MCMC is usually not the fastest approach, but it is flexible and allows accurate inference and uncertainty estimates for general models in probabilistic programming settings.
+We consider the computational costs of GPs specifically from this point of view.
 
@@
 \end{align*}
+%
 where $\nu > 0$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell > 0$ and $\alpha > 0$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
@@
 \end{align*}
+%
 The former is commonly known as the squared exponential or exponentiated quadratic covariance function.
@@
 k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
-k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right) \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right). \\
+k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right) \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right).
 \end{align*}
+%
 The previous expressions can be easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
@@
 
-Stationary covariance functions can be represented in terms of their spectral densities \citep{rasmussen2006gaussian}.
-In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure \citep[\textit{Bochner's theorem}; see, e.g.
-][]{akhiezer1993theory}.
+Stationary covariance functions can be represented in terms of their spectral densities \citep[see, e.g.,][]{rasmussen2006gaussian}.
+In this sense, the covariance function of a stationary process can be represented as the Fourier transform of a positive finite measure \citep[\textit{Bochner's theorem}; see, e.g., ][]{akhiezer1993theory}.
 If this measure has a density, it is known as the spectral density of the covariance function, and the covariance function and the spectral density are Fourier duals, known as the \textit{Wiener-Khintchine theorem} \citep{rasmussen2006gaussian}.
@@
 \end{align*}
+%
 in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ denotes the frequency, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
-The particular cases where $\nu=\infty$ and $\nu=3/2$ take the form
+The particular cases, where $\nu=\infty$ and $\nu=3/2$, take the form
 %
 \begin{align}
-S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^D  \ell^D \exp(-0.5 \ell^2 \omega^2), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \alpha  \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2})\sqrt{3}^3}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
+S_{\infty}(\omega)&= \alpha (\sqrt{2\pi})^D  \ell^D \exp(-\ell^2 \omega^2 / 2), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \alpha  \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
+%
 For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
@@
 \begin{align*}
-S_{\infty}(\omega)&= \alpha \sqrt{2\pi}^{3}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
-S_{\frac{3}{2}}(\omega)&= \alpha \, 32\,\pi\sqrt{3}^3\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
+S_{\infty}(\omega)&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
+S_{\frac{3}{2}}(\omega)&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
 \end{align*}

### 2020-04-10 Paper/manuscript_v2/manuscript_07_01.tex
hyperref colors

@@
 
-\usepackage[colorlinks=true]{hyperref}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
 
@@
 
-\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R. Andersen$^{3}$, Aki Vehtari$^{2}$}
+\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R. Andersen$^{3}$,\\
+ Aki Vehtari$^{2}$, Arno Solin$^{2}$}
 
@@
 \caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
-The dashed red line of the exact GP can hardly be seen in the some of the plots because it is under the blue lines.}
+%The dashed red line of the exact GP can hardly be seen in the some of the plots because it is under the blue lines.
+}
   \label{fig3_Post_part1}
@@
 \caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
-The dashed red line of the exact GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.}
+%The dashed red line of the exact GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.
+}
   \label{fig7_posterior_varing_lscale_part1}

### 2020-04-11 Paper/manuscript_v2/manuscript_07_01.tex
Minor

@@
 \captionsetup{font=footnotesize}
-
 \usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
@@
 
-\title{Hilbert space approximate Bayesian Gaussian processes using Stan}
-
-\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R. Andersen$^{3}$,\\
- Aki Vehtari$^{2}$, Arno Solin$^{2}$}
+\title{Hilbert space approximate Bayesian \\ Gaussian processes using Stan}
+
+\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R.\ Andersen$^{3}$,\\
+  Arno Solin$^{2}$, Aki Vehtari$^{2}$}
 
@@
 
-In this paper, we focus on a new framework for fast and accurate inference for fully Bayesian GPs using basis function approximations based on approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
+In this paper, we focus on a recent framework for fast and accurate inference for fully Bayesian GPs using basis function approximations based on approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
 Using a basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.

### 2020-04-11 Paper/manuscript_v2/manuscript_07_01.tex
Edits to Sec. 3

@@
 \begin{align}
-S_{\infty}(\omega)&= \alpha (\sqrt{2\pi})^D  \ell^D \exp(-\ell^2 \omega^2 / 2), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \alpha  \frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
+S_{\infty}(\omega)&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\ell^2 \omega^2 / 2), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\omega)&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
 \end{align}
@@
 Our starting point for presenting the method is the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
-The mathematical details of this approximation are briefly presented in Appendix \ref{ch5_app_approx_covfun}.
+The mathematical details of this approximation are briefly presented in Appendix~\ref{ch5_app_approx_covfun}.
 
@@
 
-We begin by focusing on the case of a unidimensional input space (i.e.
-on GPs with just a single covariate) such that $\Omega \in [-L,L] \subset {\rm I\!R}$, where $L$ is some positive real number to which we also refer as boundary condition.
+We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in [-L,L] \subset {\rm I\!R}$, where $L$ is some positive real number to which we also refer as boundary condition.
 As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs.
-We will come back to this issue in Section \ref{ch5_sec_accuracy}.
+We will come back to this issue in Section~\ref{ch5_sec_accuracy}.
 
@@
 \end{equation} 
+%
 where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
@@
 \end{align} 
-The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in equation (\ref{eq_eigenproblem}) are sinusoidal functions.
+%
+The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in equation \eqref{eq_eigenproblem} are sinusoidal functions.
 The solution to the eigenvalue problem is independent of  the specific choice of covariance function and is given by
@@
 
-If we truncate the sum in eq. \eqref{eq_approxcov} to the first $m$ terms, the approximate covariance function becomes
+If we truncate the sum in eq.~\eqref{eq_approxcov} to the first $m$ terms, the approximate covariance function becomes
 %
@@
 \end{equation}
+%
 where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is a diagonal matrix of the spectral density evaluated at the square root of the eigenvalues, that is, $S_{\theta}(\sqrt{\lambda_j})$,
@@
 
-Thus, the Gram matrix $\text{K}$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
+Thus, the Gram matrix $\mathrm{K}$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
 %
 \begin{equation}
-\text{K}= \Phi \Delta \Phi^\intercal, \nonumber
+\mathrm{K} = \Phi \Delta \Phi^\intercal, \nonumber
 \end{equation}
@@
 \end{equation}
+%
 where $\beta_j \sim \text{Normal}(0,1)$.
@@
 The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions.
-Therefore, eq.
-\eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
+Therefore, eq.~\eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
 The computational cost of evaluating the log posterior density of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.
 
-The parameterization in eq.
-\eqref{eq_approxf} is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which can make the posterior inference easier \citep[see, e.g.][]{Betancourt+Girolami:2019}.
+The parameterization in eq.~\eqref{eq_approxf} is naturally in the non-centered parameterization form with independent prior distribution on $\beta_j$, which can make the posterior inference easier \citep[see, e.g., ][]{Betancourt+Girolami:2019}.
 Furthermore, all dependencies on the covariance function and the hyperparameters is through the prior distribution of the regression weights $\beta_j$.

### 2020-04-11 Paper/manuscript_v2/manuscript_07_01.tex
More edits

@@
 \end{align}
+%
 where $m_d$ is the number of basis function for the dimension $d$.
@@
 Then, the number of multivariate eigenfunctions and eigenvalues is $m^{\ast} = m_{1} \cdot m_{2} \cdot m_{3} = 12$ and the matrix $\mathbb{S}\in {\rm I\!N}^{12 \times 3}$ is given by
+%
 \begin{align}\small
@@
 Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
-Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,\ldots,m^{\ast}$, we have
+Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,2,\ldots,m^{\ast}$, we have
 %
@@
 \end{align}
+%
 The approximate covariance function is then represented as
@@
 \end{equation}
+%
 where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}).
@@
 \end{equation}
+%
 where, again, $\beta_j \sim \text{Normal}(0,1)$.
@@
 Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the lengthscale of the function, which depend on the particular choice of the kernel function.
-In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix \ref{ch5_sec_periodic}.
+In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
 For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects on the posterior function.
@@
 
-As explained in Section \ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
+As explained in Section~\ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
 %
@@
 \end{equation*}
+%
 where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
@@
 \end{equation} 
+%
 where $S$ (for $S > 0$) represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
@@
 In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
-Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
+Figures~\ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
 For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation, either in mean or uncertainty, and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently less wiggly functional forms.
 The higher the wigglyness of the function to be estimated, the more basis functions will be required.
-If $m$ is fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, although along the whole domain in the uncertainty.
+If $m$ is fixed to a large enough value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, although along the whole domain in the uncertainty.
 The approximation error for the variance tends to be bigger for the variance than for the mean.
@@
 The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}, for both exact GP and HSGP models.
-Figure \ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the data for varying $m$ and $c$.
+Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the data for varying $m$ and $c$.
 Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
@@
 Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application.
-We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig3_Post_part1}.
+We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig3_Post_part1}.
 From these results, some general conclusions may be drawn:
@@
 \begin{itemize}
-\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq. \eqref{eq_eigenvalue}, where $L$ appears in the denominator	.
+\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
 \item There exists a minimum $c$ below which a accurate approximation will never be achieved regardless of the number of basis functions $m$.
@@
 \begin{equation*}
- \frac{\varepsilon}{\int k(\tau) \mathrm{d}\tau} < 0.01.
-\end{equation*}
-
+ \frac{\varepsilon}{\int k(\tau) \,\mathrm{d}\tau} < 0.01.
+\end{equation*}
+%
 Alternatively, these figures could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
@@
 \item For a given $\ell$ there exist a minimum $c$ under which a close approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 \end{itemize}
 
-As stated above, Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$.
+As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-As the lengthscale $\ell$ controls the wigglyness of
-the function, it strongly influences the difficulty of obtaining accurate
-inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
-we can expect the HSGP approximation to be accurate as well.
-
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
+As the lengthscale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data.
+Basically, if the lengthscale estimate is accurate, we can expect the HSGP approximation to be accurate as well.
+
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation is assumed to be good.
 If, however, $\hat{\ell}$ does not exceed recommended minimum lengthscale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
-We may also use this diagnostic in an iterative procedure by
-starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if
-the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
+We may also use this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
 As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the lengthscale.
@@
 
-If we look back to the conclusions drawn from Figures \ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig6_lscale_vs_J_vs_c}.
-
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_lscale_vs_J_vs_c}.
+
+Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
@@
 
-In the HSGP model, the latent function values $f(x)$ are approximated as in equation (\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in equation (\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
+In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in eq.
+~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
 In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
@@
 
-Figure \ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
+Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
 The true data-generative function and the noisy observations are also plotted.
@@
 
-Figure \ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
 The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model, for this particular application with a univariate input space.
@@
 In the online supplemental material, some other case studies are presented.
-From those examples, it can be seen how computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see equation (\ref{eq_m_multi})).
+From those examples, it can be seen how computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
 Even though, in a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
-However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure C.3 in the online material).
-In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
+However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
+In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
 
@@
 
-Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures B.3-right and C.3 in the online material).
+Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
 
 \subsection{Birthday data}\label{ch5_sec_birthday}
+%
 This example is an analysis of patterns in birthday frequencies in a dataset containing records of all births in the United States on each day during the period 1969–1988.
@@
 As the total number of days within the period is $T=7305$ ($t=1,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
-We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix \ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
-
-Let's denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
+We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix~\ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
+
+Let us denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
 %
@@
 \end{equation*}
+%
 The mean function $\mu(t)$ will be defined as an additive model in the form: 
@@
 \end{equation*}
+%
 which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
@@
 \end{equation*}
+%
 A horse-shoe prior allows for sparse distributed effects.
@@
 
-The component $f_1(t)$ will be approximated using the HSGP
-model and the function values $f_1(t)$ are approximated as in
-equation (\ref{eq_approxf}), with the squared exponential spectral
-density $S$ as in equation (\ref{eq_specdens_inf}), and eigenvalues
-$\lambda_j$ and eigenfunctions $\phi_j$ as in equations
-(\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+The component $f_1(t)$ will be approximated using the HSGP model and the function values $f_1(t)$ are approximated as in equation (\ref{eq_approxf}), with the squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
 We use $m=30$ basis functions and a boundary factor $c=1.5$.
@@
 The year effects $f_2(t)$ and week effects $f_3(t)$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper.
-However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix \ref{ch5_sec_periodic}).
+However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{ch5_sec_periodic}).
 Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in equation (\ref{ch5_eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in equation (\ref{ch5_eq_q_2}).
@@
 
-Figure \ref{ch5_fig27_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and yearly pattern $f_2(t)$ for the whole period, jointly with the observed data.
+Figure~\ref{ch5_fig27_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and yearly pattern $f_2(t)$ for the whole period, jointly with the observed data.
 Figure \ref{ch5_fig27_posteriors_oneyear_birthday} shows the model for one year (1972) only.
@@
 \subsection{Leukemia data}\label{ch5_sec_bf_caseVI}
+
 The next example presents a survival analysis in acute myeloid leukemia (AML) in adults, with data recorded between 1982 and 1998 in the North West Leukemia Register in the United Kingdom.
@@
 \end{equation*}
+%
 where $y_i>t_i$ denotes the unobserved survival time.
@@
 \section{Conclusion}\label{ch5_sec_conclusion}
+
 Modeling unknown functions using exact GPs is computationally intractable for a lot of applications.
 This problem becomes especially severe when performing full Bayesian inference using sampling-based methods.
-In this paper, a novel approach for a low-rank representation of stationary GPs, originally proposed by \citet{solin2018hilbert}, has been implemented and analyzed in detail.
+In this paper, a recent approach for a low-rank representation of stationary GPs, originally proposed by \citet{solin2018hilbert}, has been implemented and analyzed in detail.
 The method is based on a basis function approximation via Laplace eigenfunctions.
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-Hence, choosing optimal values for the number of basis functions and the boundary factor, using the
- recommendations and diagnostics provided in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time especially in multivariate input spaces.
-However, in practice, input dimensionalities larger than 3 start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
+Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time especially in multivariate input spaces.
+However, in practice, input dimensionalities larger than three start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
 In these high dimensional cases, the proposed approximate GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.

### 2020-04-13 Paper/manuscript_v2/manuscript_07_01.tex
update Fig. 15 and title
Changed Normal to N
Notation

@@
 
-\title{Hilbert space approximate Bayesian \\ Gaussian processes using Stan}
+\title{Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming}
 
@@
 %
-where $\beta_j \sim \text{Normal}(0,1)$.
+where $\beta_j \sim \mathcal{N}(0,1)$.
 Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
@@
 %
-where, again, $\beta_j \sim \text{Normal}(0,1)$.
+where, again, $\beta_j \sim \mathcal{N}(0,1)$.
 The computational cost of evaluating the log posterior density of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
@@
 
-Comparing the accuracy of the lengthscale in Figure 
-\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure \ref{fig9_MSE_varing_lscale},
-we see that they agree closely with each other for medium lengthscales.
-That is, a good estimation of the lengthscale implies a
-small RMSE. This is no longer true for very small or large lengthscales.
-In small lengthscales, even very small inaccuracies 
-may have a strong influence on the posteriors predictions and thus on the RMSE.
-In large lengthscales, larger inaccuracies change the posterior predictions
-only little and may thus not yield large RMSEs.
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the 
-minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}. 
-We observe that whenever the estimated lengthscale exceeds the minimally
-estimable lengthscale, the RMSE of the posterior predictions is small
-(see Figure \ref{fig9_MSE_varing_lscale}).
-Conversely, when the estimated lengthscale is smaller than the minimally
-estimable one, the RMSE becomes very large.
+Comparing the accuracy of the lengthscale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig9_MSE_varing_lscale}, we see that they agree closely with each other for medium lengthscales.
+That is, a good estimation of the lengthscale implies a small RMSE.
+This is no longer true for very small or large lengthscales.
+In small lengthscales, even very small inaccuracies  may have a strong influence on the posteriors predictions and thus on the RMSE.
+In large lengthscales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}.
+We observe that whenever the estimated lengthscale exceeds the minimally estimable lengthscale, the RMSE of the posterior predictions is small
+(see Figure \ref{fig9_MSE_varing_lscale}). Conversely, when the estimated lengthscale is smaller than the minimally estimable one, the RMSE becomes very large.
 
@@
 
-This example consists of a simulated dataset with $n=250$ ($i=1,\dots,n$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=(x_1\dots,x_n)$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
+This example consists of a simulated dataset with $n=250$ ($i=1,2,\dots,n$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
 To form the final noisy dataset $\bm{y}$, Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws.
@@
 \end{align*}
+%
 where $I$ represents the identity matrix and $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function values to the noisy data.
@@
 \end{equation*}
+%
 where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
 
-In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in eq.
-~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
+In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
 In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
@@
 \begin{equation*}
-f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\mathrm{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
 \end{equation*}
@@
 \end{equation}
-where $\beta_j \sim \text{Normal}(0,1)$, with $j=1,\dots,2J+1$.
+where $\beta_j \sim \mathcal{N}(0,1)$, with $j=1,\dots,2J+1$.
 The cosine $\cos(j\omega_0 x)$ and sinus $\sin(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.

### 2020-04-13 Paper/manuscript_v2/manuscript_07_01.tex
Notation

@@
 % \begin{align}\label{eq:posterior}
-% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\text{d}\bm{f}}{p(\bm{y})},
+% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\mathrm{d}\bm{f}}{p(\bm{y})},
 % \end{align}
@@
 % \begin{align}\label{eq:evidence}
-% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \text{d}\bm{f}\text{d}f^*.
+% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \mathrm{d}\bm{f}\mathrm{d}f^*.
 % \end{align}
@@
 
-Thus, the Gram matrix $\mathrm{K}$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
+Thus, the Gram matrix $K$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
 %
 \begin{equation}
-\mathrm{K} = \Phi \Delta \Phi^\intercal, \nonumber
-\end{equation}
+  K = \Phi \Delta \Phi^\intercal, \nonumber
+\end{equation}
+%
 where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$
@@
 \begin{equation}\label{eq_diff_covs}
- \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \mathrm{d}\tau  < \varepsilon.
+ \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \,\mathrm{d}\tau  < \varepsilon.
 \end{equation}
@@
 \begin{equation*}
-f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\mathrm{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
 \end{equation*}
@@
 \begin{equation*}
-f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
+f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
 \end{equation*}
@@
 \begin{equation*}
-p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \left( \frac{\mathrm{log}(y_i)-f(\bm{x}_i)}{\sigma} \right),
+p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
 \end{equation*}
@@
 \begin{equation*}
-\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}'.
+\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') \,\mathrm{d}\bm{x}'.
 \end{equation*}
@@
 \begin{equation*}
-S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \mathrm{d}\bm{\tau}, \nonumber
+S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \,\mathrm{d}\bm{\tau}, \nonumber
 \end{equation*}
@@
 \begin{equation*}
--\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}',
+-\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') \,\mathrm{d}\bm{x}',
 \end{equation*} 
@@
 \begin{align*}
-<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) \mathrm{d}(\bm{x})
+<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) \,\mathrm{d}(\bm{x})
 \end{align*} 
@@
 \begin{align*}
-\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) \mathrm{d}(\bm{x}) = \delta_{ij},
+\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) \,\mathrm{d}(\bm{x}) = \delta_{ij},
 \end{align*} 
@@
 \begin{equation*}
-(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}'.
+(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') \,\mathrm{d}\bm{x}'.
 \end{equation*} 
@@
 \begin{equation*}
-[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \mathrm{d}\bm{x}'.
+[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \,\mathrm{d}\bm{x}'.
 \end{equation*}
@@
 \begin{split}
-\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
-\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
+\tilde{q}_j^2= \frac{2\mathrm{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
+\tilde{q}_0^2= \frac{\mathrm{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
 \end{split}
 \end{equation} 
-for $j=1,2,\cdots,J$, and where the $\text{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
+for $j=1,2,\cdots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
 This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required.

### 2020-04-13 Paper/manuscript_v2/manuscript_07_01.tex
Changed notation again following Aki

@@
 \DeclareMathOperator{\GP}{\mathcal{GP}}
+\DeclareMathOperator{\Normal}{Normal}
 
@@
 The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$.
-Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution $\bm{f} \sim \mathcal{N}(\bm{\mu}, \bm{K})$, where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
+Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution $\bm{f} \sim \Normal(\bm{\mu}, \bm{K})$, where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$.
 In the following, we focus on zero-mean Gaussian processes, that is set $\mu(\bm{x}) = 0$.
@@
 \begin{align*}
-p(\bm{f}, \tilde{f})=\mathcal{N} \left( \left[ \begin{array}{cc}
+p(\bm{f}, \tilde{f})=\Normal \left( \left[ \begin{array}{cc}
 \bm{f} \\ 
@@
 % \begin{align*}
-	% p(f^*| \bm{f}) = \mathcal{N}(f^*| \bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{k}_{\bm{f},f^*}  )
+	% p(f^*| \bm{f}) = \Normal(f^*| \bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{k}_{\bm{f},f^*}  )
 % \end{align*}
@@
 % \begin{align*}
-% 	p(f^*| \bm{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
+% 	p(f^*| \bm{y}) = \Normal(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
 % \end{align*}
@@
 
-If $p(y_n \mid f(\bm{x}_n),\phi)=\mathcal{N}(y_n \mid f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
-If $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))=\mathcal{N}(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n \mid f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed-form solution.
+If $p(y_n \mid f(\bm{x}_n),\phi)=\Normal(y_n \mid f(\bm{x}_n),\sigma)$ then $\bm{f}$ can be integrated out analytically (with a computational cost of $O(n^3)$ for exact GPs and $O(nm^2)$ for sparse GPs).
+If $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))=\Normal(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or $p(y_n \mid f(\bm{x}_n),\phi)$ is non-Gaussian, the marginalization does not have a closed-form solution.
 Furthermore, if a prior distribution is imposed on $\phi$ and $\bm{\theta}$ to form a joint posterior for $\phi$, $\bm{\theta}$ and $\bm{f}$, approximate inference such as Markov chain Monte Carlo \citep[MCMC; ][]{brooks_2011}, Laplace approximation \citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational inference methods \citep{gibbs2000variational,csato2000efficient} are required.
@@
 \begin{equation}
-  K = \Phi \Delta \Phi^\intercal, \nonumber
+ \bm{K} = \Phi \Delta \Phi^\intercal, \nonumber
 \end{equation}
@@
 \begin{equation}
-\bm{f} \sim \mathcal{N}(\bm{\mu},\Phi \Delta \Phi^\intercal). \nonumber
+\bm{f} \sim \Normal(\bm{\mu},\Phi \Delta \Phi^\intercal). \nonumber
 \end{equation}
@@
 %
-where $\beta_j \sim \mathcal{N}(0,1)$.
+where $\beta_j \sim \Normal(0,1)$.
 Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values.
@@
 %
-where, again, $\beta_j \sim \mathcal{N}(0,1)$.
+where, again, $\beta_j \sim \Normal(0,1)$.
 The computational cost of evaluating the log posterior density of multivariate HSGPs scales as $O(n m^{\ast} + m^{\ast})$, where $n$ is the number of observations and $m^{\ast}$ is the number of multivariate basis functions.
@@
 \bm{y} &= \bm{f} + \bm{\epsilon} \\
-\bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
+\bm{\epsilon} &\sim \Normal(0, \sigma^2  I) \\
 f(x) &\sim \GP(0, k(x, x', \theta)),
@@
 The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with a Mat{\'e}rn($\nu$=3/2) covariance function $k$.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
 
@@
 \begin{equation*}\label{ch5_eq_marginalizedgp_simudata1}
-\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ),
+\bm{y} \sim \Normal(0,\bm{K}+ \sigma^2 \bm{I} ),
 \end{equation*}
@@
 \begin{equation*}
-y_{t} \sim \mathcal{N}(\mu(t),\sigma^2).
+y_{t} \sim \Normal(\mu(t),\sigma^2).
 \end{equation*}
@@
 %
-which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
+which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $\bm{K}_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
 
@@
 \begin{equation*}
-f_4(t) \sim \mathcal{N}(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
+f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
 \end{equation*}
@@
 \end{equation}
-where $\beta_j \sim \mathcal{N}(0,1)$, with $j=1,\dots,2J+1$.
+where $\beta_j \sim \Normal(0,1)$, with $j=1,\dots,2J+1$.
 The cosine $\cos(j\omega_0 x)$ and sinus $\sin(j\omega_0 x)$ terms do not depend on the covariance hyperparameters $\ell$.

### 2020-04-15 Paper/manuscript_v2/manuscript_07_01.tex
Fixes in bib
fix variance in equations B6 and B7

@@
 \captionsetup{font=footnotesize}
-\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
+%\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
 
@@
 Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying equation (\ref{eq_diff_covs}).
+More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying eq.
+(\ref{eq_diff_covs}).
 Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function.
-We considered an approximation to be a close enough when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$:
+We considered an approximation to be a close enough when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
+(\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$:
 %
@@
 \begin{equation} \label{ch5_eq_cov_periodic_taylor_approx_2}
-k(x,x')= \sigma^2 \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
+k(x,x')= \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
 \end{equation}
@@
 \begin{equation} \label{ch5_eq_f_period}
-f(x) \approx \sigma \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
+f(x) \approx \alpha^{1/2} \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
 \end{equation}
@@
 
-\section*{Acknowledgment}
+%\section*{Acknowledgment}
 

### 2020-04-16 Paper/manuscript_v2/manuscript_07_01.tex
update 2nd paragraph of case study I

@@
 \captionsetup{font=footnotesize}
-%\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
 
@@
 A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
-GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
+GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_N) \right\rbrace$, is a multivariate Gaussian distribution.
 
@@
 %
-where $I$ represents the identity matrix and $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function values to the noisy data.
+where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $I$ representing the identity matrix.
 The previous formulation corresponds to the latent form of a GP model.
-The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with a Mat{\'e}rn($\nu$=3/2) covariance function $k$.
+The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with a Mat{\'e}rn($\nu$=3/2) covariance function $k(x_i,x_j,\theta)$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
 Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
@@
 
-The Stan model code for the approximate GP model of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}.
+The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}.
 
@@
 
-The Stan model code for the exact GP and the approximate GP models of this case study can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}.
+The Stan model code for the exact GP and the approximate GP models of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}.
 %

### 2020-04-16 Paper/online_supplemental_material/online_material.tex
edits to online material

@@
 
-\usepackage[numbers,sort&compress]{natbib}
+\usepackage{natbib}
+\bibliographystyle{plainnat}
+\bibpunct{(}{)}{;}{a}{,}{,}
 \usepackage{subfigure}
@@
 \usepackage{url}
+\usepackage{parskip}
 
 \usepackage[titletoc]{appendix}
+\usepackage{booktabs}
+\usepackage{boldline}
+\usepackage{colortbl}
+\DeclareMathOperator{\LogNormal}{LogNormal}
+\DeclareMathOperator{\GP}{\mathcal{GP}}
+\DeclareMathOperator{\Normal}{Normal}
 
@@
 \captionsetup{font=footnotesize}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
 
-\usepackage{chngcntr}
+%\bibliographystyle{authordate1}
+% TODO macro
+\newcommand{\todo}[1]{\textcolor{blue}{\textbf{[#1]}}}
+\newcommand{\resp}[1]{\textcolor{red}{\textbf{[#1]}}}
 
-\bibliographystyle{authordate1}
+\title{Online supplemental material of the article:\\ \vspace{7mm} 'Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming'\\ \vspace{7mm}
+\normalsize by}
 
-\title{Online supplemental material of the article:\\ \vspace{7mm} 'Hilbert space approximate Bayesian Gaussian processes using Stan'\\ \vspace{7mm}
-Case Studies}
-
-\author{TODO}
+\author{Gabriel Riutort-Mayol and Paul-Christian Bürkner and Michael R.\ Andersen and\\
+  Arno Solin and Aki Vehtari}
 
@@
 \maketitle
+
+This document consists of four case studies of application of the article 'Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming'.
 
@@
 \begin{equation*}
-y_i \sim \text{Binomial}(p_i, n_i).
+y_i \sim \mathrm{Binomial}(p_i, n_i).
 \end{equation*}
@@
 \noindent The population per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ or mean number of support people per age group.
-Probabilities $\bm{p}=(p_1,\dots,p_{74})$ are modeled by a GP function $f:{\rm I\!R} \to {\rm I\!R}$ with a squared exponential covariance function $k$, as a function of age input values $\bm{x}=(x_1,\dots,x_{74})$, and through the {\it logit} function as a link function,
+Probabilities $\bm{p}=(p_1,\dots,p_{74})$ are modeled by a Gaussian process (GP) function $f:{\rm I\!R} \to {\rm I\!R}$ with a squared exponential covariance function $k$, as a function of age input values $\bm{x}=(x_1,\dots,x_{74})$, and through the {\it logit} function as a link function,
 %
@@
 
-\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,74$.
-In the HSGP model, the function $f(x)$ is approximated as in equation (7) %(\ref{eq_approxf})
+\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,74$.
+The covariance function $k$ depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+
+In the HSGP model, the function $f(x)$ is approximated as in equation (7)%(\ref{eq_approxf})
 , with the squared exponential spectral density as in equation (1)%(\ref{eq_specdens_inf})
@@
 
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this splines model using the R-package \textit{brms} \citep{burkner2017brms}.
@@
 For the HSGP model, different models with different number of basis functions and boundary factor have been fitted.
-The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
-The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
+The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}-left, for sample data, and in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}-right, for test data.
+As the boundary factor increases, more basis functions are needed for a accurate approximation, as it was expected.
 
-Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models, computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model.
-We can see how the splines models do not extrapolate data properly.
+Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model.
+We can see how the splines models do not extrapolate data properly (Figure \ref{ch5_fig14_MSE_train_gaydata}-right).
 
@@
 The HSGP model is on average roughly 15 times faster than the regular GP and 5 times faster than the spline model, for this particular case and univariate input space.
-{\color{blue}The increase of computation as function of the number of basis functions in a univariate input space is relatively slight, as can be seen in the figure}.
+Computation time increases relatively slight with the number of basis functions in a univariate input space, as can be seen in the figure.
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}.
+The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}}.
 
@@
 
-This example consists of a simulated dataset with $n=120$ ($i=1,\dots,n$) single draws from a Gaussian process prior with two input dimensions ($D=2$).
-A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$, for the first dimension, and $\ell_2=0.35$, for the second dimension, is used.
-The corresponding matrix of input values is $X=[\bm{x}_1 \; \bm{x}_2 \; \cdots \; \bm{x}_{120}]^\top \in {\rm I\!R}^{120\times 2}$ with $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$.
-Gaussian noise $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^{120}$.
+This example consists of a simulated dataset with $n=120$ ($i=1,\dots,n$) single draws from a GP prior as function of input values $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$ in a bivariate input space ($D=2$).
+A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$, for the first input dimension, and $\ell_2=0.35$, for the second input dimension, is used for the GP prior.
+Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^{120}$.
 
-The regular GP model over the outcome variable $\bm{y}$ and matrix of inputs $X \in {\rm I\!R}^{120 \times 2}$, can be written as follows,
+The regular GP model over the outcome variable $\bm{y}$ can be written as follows,
 %
@@
 \bm{y} &= \bm{f} + \bm{\epsilon} \\
-\bm{\epsilon} &\sim \mathcal{N}(0, \sigma^2  I) \\
+\bm{\epsilon} &\sim \Normal(0, \sigma^2  I) \\
 f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x}, \bm{x}', \theta)),
@@
 
-\noindent where $I$ represents the identity matrix, $\bm{e}$ the noise term, and $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^{120}$ represents the underlying function values to the noisy observations $\bm{y}$.
-The function $f:{\rm I\!R}^2 \to {\rm I\!R}$ is a GP prior with a multivariate squared exponential covariance function $k$.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $K$, where $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,120$.
-The covariance function $k$ depends on the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
-The hyperparameters $\alpha$, $\ell_1$ and $\ell_2$ represent the marginal variance and lengthscales for first and second dimensions, respectively, of the GP process.
+\noindent where $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^{120}$ represents the underlying function at the input values $\bm{x_i}\in {\rm I\!R}^{2}$, $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, and $I$ represents the identity matrix.
+The function $f:{\rm I\!R}^2 \to {\rm I\!R}$ is a GP prior with a multivariate squared exponential covariance function $k(\bm{x}_i,\bm{x}_j,\theta)$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
+The hyperparameters $\alpha$, $\ell_1$ and $\ell_2$ represent the marginal variance and lengthscales for first and second input dimensions, respectively, of the GP process.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(\bm{x}_i,\bm{x}_j,\theta)$, with $i,j=1,\dots,120$.
 
@@
 \begin{equation*}\label{ch5_eq_marginalizedgp_simudata2}
-\bm{y} \sim \mathcal{N}(0, K + \sigma^2 I ).
+\bm{y} \sim \Normal(0, \bm{K} + \sigma^2 I ).
 \end{equation*}
 
-In the HSGP model with $2$ input dimensions, the latent function $f$, evaluated at input vector $\bm{x} \in {\rm I\!R}^2$, is approximated as in equation (12)% (\ref{eq_approxf_multi})
+In the HSGP model with $2$ input dimensions, the latent function $f(\bm{x})$ is approximated as in equation (12)% (\ref{eq_approxf_multi})
 , with the $2$-dimensional squared exponential spectral density $S$ as in equation (1)% (\ref{eq_specdens_inf})
-, and the multivariate eigenfunctions $\phi_j$ and the $2$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (10) %(\ref{eq_eigenfunction_multi})
- and (9)% (\ref{eq_eigenvalue_multi})
+, and the $2$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (9) %(\ref{eq_eigenfunction_multi})
+ and (10)% (\ref{eq_eigenvalue_multi})
  , respectively.
 
-In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2015package}.
+In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
@@
 Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset.
-Sample observations are also plotted in the plots as circles.
+Sample observations are also plotted as circles.
 For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
@@
 However, the performance of the splines model is significantly worse.
+
 Figure \ref{ch5_fig17_RMSE_exII}-right shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots.
-Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computational time.
-{\color{blue}Even though in a bivariate input space the computation demand increases significantly with the number of dimensions or knots, either the HSGP or spline models work significantly better than regular GP, even for highly wiggly functions where a high number of basis functions or knots are required in the approximations.
-However, severe difficulties have been found in building the spline model with $50$ knots for this bivariate function}.
+Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computation time.
+Even though in a bivariate input space the computation demand increases significantly with the number of dimensions or knots, either the HSGP or spline models work significantly better than regular GP, even for highly wiggly functions that require a high number of basis functions or knots for a accurate approximation.
+However, serious difficulties with cumputation time were encountered in building the spline model with $50$ knots for this bivariate function.
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}.
+The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}.
 
@@
 The data contains $n=392$ individuals ($i=1,\dots,n$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed.
-The matrix $X=[\bm{x}_1\;\bm{x}_2\;\cdots\;\bm{x}_n]^\top \in {\rm I\!R}^{n\times 4}$,  with $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$, contains the risk factors, \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$.
+The input vector $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$, in a 4D input space ($D=4$), contains the risk factors, \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$.
 The observational model is a Bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
@@
 \begin{equation*}
-y_i \sim \text{Bernoulli}(p_i).
+y_i \sim \mathrm{Bernoulli}(p_i).
 \end{equation*}
 
-\noindent The goal is to estimate the probability $p_i$ as a function of the risk factors, which function $f(\cdot):{\rm I\!R}^{4} \to {\rm I\!R}$ is modeled as a Gaussian process with a multivariate squared exponential covariance function $k$ depending on the matrix $X$ of risk factors and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
+\noindent The goal is to estimate the probability $p_i$ as a function of the risk factors, which function $f(\cdot):{\rm I\!R}^{4} \to {\rm I\!R}$ is modeled as a GP with a multivariate squared exponential covariance function $k$ depending on the risk factors $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
 %
@@
 p_i &= \mathrm{logit}(f(\bm{x}_i)) \nonumber \\
-f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x},\bm{x}, \theta).
+f(\bm{x}) &\sim \mathcal{GP}(0, k(\bm{x},\bm{x}', \theta).
 \end{align*}
 
-\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^n$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
+\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^n$ are multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-Notice that a scalar lengthscale is considered in the multivariate covariance function.
+A scalar lengthscale has been considered in the multivariate covariance function $k$.
 
@@
 
-In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional spline-based model is also fitted using a cubic spline basis penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized} and implemented in the R-package \textit{mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms}.
+In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional spline-based model is also fitted using a cubic spline basis penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
 
@@
 
-Figure \ref{ch5_fig20_time2D_diabetes} shows the computational times of the different models, regular GP, HSGP and spline, fitted over the dataset, with 2 input dimensions, 3 input dimensions and 4 input dimensions, as a function of the boundary factor $c$ and number of univariate basis functions $m$ and knots.
-We can appreciate the significant increase of computational time with higher dimensions for the HSGP and spline models.
-In this sense, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure 6% \ref{fig6_lscale_vs_J_vs_c}
-, can be essential to avoid a excessive computational time, especially in multivariate input spaces.
-It is interesting to be noticed that considering more than 10 knots per dimension in the spline model with 3D is not allowed for an amount of 392 observations.
-Similarly, just the computation of the input data for the spline model in a 4D input space is computationally very expensive.
+Figure \ref{ch5_fig20_time2D_diabetes} shows the computational times of the different models, regular GP, HSGP and spline, fitted over the dataset with 2, 3 and 4 input dimensions, as a function of the boundary factor $c$ and number of univariate basis functions $m$ and knots.
+The time of omputation increases significantly with higher dimensions for the HSGP and spline models.
+In this sense, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure 6 % \ref{fig6_lscale_vs_J_vs_c}
+in the main document, can be essential to avoid a excessive computation time, especially in multivariate input spaces.
+Serious difficulties with computation time were encountered in building spline models in a 4D input space or even in a 3D input space with more than 10 knots.
 
-The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}.
+The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}}.
 
@@
 The next example presents an spatio-temporal classification in land-use of plots between 2006 and 2015 in a part of the territory of Valencia in Spain dedicated to growing citrus fruits.
-A sampling set consists of $N=200$ plots with known class.
+A sampling set consists of $n=200$ plots with known class.
 The data is recorded in a time series of $T=5$ years (2006, 2008, 2010, 2012, 2015) within the period.
@@
 The principal components (PCs) will be used jointly with the time variable as inputs in the classifying additive HSGP model.
-Let's denote the matrix $X=[\bm{x}_{11}\, \cdots \, \bm{x}_{it} \, \cdots \, \bm{x}_{NT}]^\top \in {\rm I\!R}^{NT\times D}$, which contains the input vectors $\bm{x}_{it} \in {\rm I\!R}^D$, $D=53$ ($52 \text{ PCs plus time}$) for the spatio-temporal observations (plots $i=1,\dots,200$, and times $t=1,\dots,5$).
+Let's denote the matrix $\bm{X}=[\bm{x}_{11}\, \cdots \, \bm{x}_{it} \, \cdots \, \bm{x}_{nT}]^\top \in {\rm I\!R}^{nT\times D}$, which contains the input vectors $\bm{x}_{it} \in {\rm I\!R}^D$, $D=53$ ($52 \text{ PCs plus time}$) for the spatio-temporal observations (plots $i=1,\dots,200$, and times $t=1,\dots,5$).
 
-The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{it}=(p_{it,1},\dots,p_{it,K})$, where $p_{it,k}$ is the probability of belonging to class $k$ per parcel $i$ and time $t$,
+The observational model is a multinomial model with parameters the vector of probabilities $\bm{p}_{it}=(p_{it,1},p_{it,2},\dots,p_{it,K})$, where $p_{it,k}$ is the probability of belonging to class $k$ per parcel $i$ and time $t$,
 %
 \begin{equation*}
-y_{it} \sim \text{multinomial}(\bm{p}_{it}).
+y_{it} \sim \mathrm{Multinomial}(\bm{p}_{it}).
 \end{equation*}
@@
 In the case of the first-order components, the normalized lengthscale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum lengthscale reported by Figure (6) % \ref{fig6_lscale_vs_J_vs_c}
-as a function of $m$, $c$. Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.
+as a function of $m$ and $c$. Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}.
+The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}.
 

### 2020-04-17 Paper/online_supplemental_material/online_material.tex
a few more edits to online material

@@
 \usepackage{xcolor}
-\graphicspath{{./images/}}
+\graphicspath{{../manuscript_v2/images/}}
 
@@
 The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}-left, for sample data, and in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}-right, for test data.
-As the boundary factor increases, more basis functions are needed for a accurate approximation, as it was expected.
+As the boundary factor increases, more basis functions are needed for a accurate approximation.
 
@@
 
-Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset.
+Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, regular GP, HSGP, and splines, fitted over the dataset.
 Sample observations are also plotted as circles.
@@
 \caption{(a) Data-generating function.
-(b) Mean posterior predictive function of the GP model.
-(c) Mean posterior predictive function of the HSGP model.
-(d) Mean posterior predictive function of the splines model.
+(b) Posterior mean predictive function of the GP model.
+(c) Posterior mean predictive function of the HSGP model.
+(d) Posterior mean predictive function of the splines model.
 Sample points are plotted as circles}
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
-\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models. Sample points are plotted as circles.}
+\caption{Error between the data-generating mean function and the posterior mean of the GP (a), HSGP (b) and spline (c) models. Sample points are plotted as circles.}
   \label{ch5_fig16_errors_exII}
@@
 
-Figure \ref{ch5_fig18_gpfun_diabetes} shows the mean posterior predictions of probabilities ($p_i$) of the three models, the regular GP, the HSGP and the spline, fitted over the dataset with the 2 input dimensions \textit{Glucose} and \textit{Pregnancy} ($D=2$).
+Figure \ref{ch5_fig18_gpfun_diabetes} shows the posterior mean predictions of probabilities ($p_i$) of the three models, regular GP, HSGP and spline, fitted over the dataset with the 2 input dimensions \textit{Glucose} and \textit{Pregnancy} ($D=2$).
 The binary observations $y_i$ are also plotted in the plots as colored points.
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 4mm 2mm 15mm, clip]{ch5_fig18_legend_diabetes.pdf}}
-\caption{(a) Mean posterior predictive functions of the GP model.
-(b) Mean posterior predictive functions of the HSGP model.
-(c) Mean posterior predictive functions of the spline (SP) model.
+\caption{(a) Posterior mean predictive functions of the GP model.
+(b) Posterior mean predictive functions of the HSGP model.
+(c) Posterior mean predictive functions of the spline (SP) model.
 Samples observations of suffering (red points) and not suffering (blue points) form the disease are plotted.}
@@
 The time of omputation increases significantly with higher dimensions for the HSGP and spline models.
-In this sense, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure 6 % \ref{fig6_lscale_vs_J_vs_c}
-in the main document, can be essential to avoid a excessive computation time, especially in multivariate input spaces.
+In this sense, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure 6% \ref{fig6_lscale_vs_J_vs_c}
+, can be essential to avoid a excessive computation time, especially in multivariate input spaces.
 Serious difficulties with computation time were encountered in building spline models in a 4D input space or even in a 3D input space with more than 10 knots.
@@
 
-\bibliography{references}
+\bibliography{../manuscript_v2/references}
 

### 2020-04-17 Paper/online_supplemental_material/online_material.tex
few edits in captions in online material

@@
 \includegraphics[scale=0.50]{ch5_fig12_Posteriors_gaydata.pdf}
-\caption{Posterior mean predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model. 95\% credible intervals are plotted as dashed lines.}
+\caption{Posterior predictive means of the proposed HSGP model, the regular GP model, and the Splines model. 95\% confident intervals are plotted as dashed lines.}
   \label{ch5_fig12_Posteriors_gaydata}
@@
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}}.
+The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}}\,.
 
@@
 \end{center}
-\caption{(a) Data-generating function.
-(b) Posterior mean predictive function of the GP model.
-(c) Posterior mean predictive function of the HSGP model.
-(d) Posterior mean predictive function of the splines model.
-Sample points are plotted as circles}
+\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline (d) models. Sample points are plotted as circles}
   \label{ch5_fig15_Posteriors_exII}
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
-\caption{Error between the data-generating mean function and the posterior mean of the GP (a), HSGP (b) and spline (c) models. Sample points are plotted as circles.}
+\caption{Error between the data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline (c) models. Sample points are plotted as circles.}
   \label{ch5_fig16_errors_exII}
@@
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}.
+The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
 
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 4mm 2mm 15mm, clip]{ch5_fig18_legend_diabetes.pdf}}
-\caption{(a) Posterior mean predictive functions of the GP model.
-(b) Posterior mean predictive functions of the HSGP model.
-(c) Posterior mean predictive functions of the spline (SP) model.
-Samples observations of suffering (red points) and not suffering (blue points) form the disease are plotted.}
+\caption{Posterior predictive mean functions of the GP (a), HSGP (b) and spline (c) models.
+Samples observations of suffering (red points) and not suffering (blue points) from the disease are plotted.}
   \label{ch5_fig18_gpfun_diabetes}
@@
 
-The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}}.
+The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}}\,.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}.
+The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}\,.
 

### 2020-04-17 Paper/manuscript_v2/manuscript_07_01.tex
edit kernel magnitude notations in case study II

@@
 
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}.
+The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
 %
@@
 \begin{equation*}
-f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
-\end{equation*}
-%
-which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $\bm{K}_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
+f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \alpha_1 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
+\end{equation*}
+%
+which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
+$\ell_1$ and $\alpha_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component.
 
@@
 \begin{equation*}
-f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \alpha_2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
 \end{equation*}
@@
 \begin{equation*}
-f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
+f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \alpha_3 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
 \end{equation*}
@@
 
-The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}.
+The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}\,.
 
@@
 
-The Stan model code for the exact GP and the approximate GP models of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}.
+The Stan model code for the exact GP and the approximate GP models of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
 %

### 2020-04-18 Paper/manuscript_v2/manuscript_07_01.tex
little edits

@@
 Appendix~\ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
-Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
+Online supplemental material with more case studies illustrating the performance and applicability of the method can be found online at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
 
@@
 The previous formulation corresponds to the latent form of a GP model.
-The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with a Mat{\'e}rn($\nu$=3/2) covariance function $k(x_i,x_j,\theta)$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
+The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
@@
 which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
-$\ell_1$ and $\alpha_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component.
+$\alpha_1$ and $\ell_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}\,.
+The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}\,.
 
@@
 
-The Stan model code for the exact GP and the approximate GP models of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
+The Stan model code for the exact GP and the approximate GP models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
 %

### 2020-04-18 Paper/online_supplemental_material/online_material.tex
little edits

@@
 \noindent The population per age group $n_i$ is a known quantity and the goal is to estimate the same-sex support probability $p_i$ or mean number of support people per age group.
-Probabilities $\bm{p}=(p_1,\dots,p_{74})$ are modeled by a Gaussian process (GP) function $f:{\rm I\!R} \to {\rm I\!R}$ with a squared exponential covariance function $k$, as a function of age input values $\bm{x}=(x_1,\dots,x_{74})$, and through the {\it logit} function as a link function,
+Probabilities $\bm{p}=(p_1,\dots,p_{74})$ are modeled by a Gaussian process (GP) function $f:{\rm I\!R} \to {\rm I\!R}$ with a squared exponential covariance function $k$, as a function of age input values $\bm{x}=(x_1,\dots,x_{74})$, and through the {\it logit} function as a link function:
 %
@@
 \noindent where $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^{120}$ represents the underlying function at the input values $\bm{x_i}\in {\rm I\!R}^{2}$, $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, and $I$ represents the identity matrix.
-The function $f:{\rm I\!R}^2 \to {\rm I\!R}$ is a GP prior with a multivariate squared exponential covariance function $k(\bm{x}_i,\bm{x}_j,\theta)$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
+The function $f:{\rm I\!R}^2 \to {\rm I\!R}$ is a GP prior with a multivariate squared exponential covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
 The hyperparameters $\alpha$, $\ell_1$ and $\ell_2$ represent the marginal variance and lengthscales for first and second input dimensions, respectively, of the GP process.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(\bm{x}_i,\bm{x}_j,\theta)$, with $i,j=1,\dots,120$.
 
@@
 \begin{equation*}\label{ch5_eq_marginalizedgp_simudata2}
-\bm{y} \sim \Normal(0, \bm{K} + \sigma^2 I ).
+\bm{y} \sim \Normal(0, \bm{K} + \sigma^2 I ),
 \end{equation*}
+%
+where $\bm{K} \in {\rm I\!R}^{n\times n}$ if the covariance matrix with an element $K_{ij}=k(\bm{x}_i,\bm{x}_j,\theta)$, and $\sigma^2$ is the noise variance.
 
@@
 
-\noindent Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}=\{f(\bm{x}_i)\}_{i=1}^n$ are multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
-The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
 A scalar lengthscale has been considered in the multivariate covariance function $k$.
@@
 
-The Stan model code for the approximate GP model of this case study can be found at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}\,.
+The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}\,.
 

### 2020-04-20 Paper/manuscript_v2/manuscript_07_01.tex
More edits

@@
 \end{align*} 
-
+%
 The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$.
@@
 \bm{y} &= \bm{f} + \bm{\epsilon} \\
-\bm{\epsilon} &\sim \Normal(0, \sigma^2  I) \\
+\bm{\epsilon} &\sim \Normal(\bm{0}, \sigma^2 \bm{I}) \\
 f(x) &\sim \GP(0, k(x, x', \theta)),
@@
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,2,\dots,n$.
 
@@
 \begin{equation*}\label{ch5_eq_marginalizedgp_simudata1}
-\bm{y} \sim \Normal(0,\bm{K}+ \sigma^2 \bm{I} ),
+\bm{y} \sim \Normal(\bm{0},\bm{K}+ \sigma^2 \bm{I} ),
 \end{equation*}
@@
 \citet{gelman2013bayesian} presented an analysis using exact GP and maximum a posteriori inference.
-As the total number of days within the period is $T=7305$ ($t=1,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
+As the total number of days within the period is $T=7305$ ($t=1,2,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
 We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix~\ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
@@
 $\alpha_1$ and $\ell_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component.
-
 The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
@@
 where $y_i>t_i$ denotes the unobserved survival time.
-
 The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$ depending on predictors $x$ and hyperparameters $\theta=(\alpha,\ell)$,
@@
 
-Due to the predictor \textit{sex} ($x_2$) being a categorical variable ('1' for female and '2' for male), we apply indicator variable coding for the GP functions, in a similar way such coding is applied in linear models \citep{Gelman+Hill+Vehtari:2020:ROS}.
+Due to the predictor \textit{sex} ($x_2$) being a categorical variable (`1' for female and `2' for male), we apply indicator variable coding for the GP functions, in a similar way such coding is applied in linear models \citep{Gelman+Hill+Vehtari:2020:ROS}.
 A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise:
@@
 	\begin{array}{ll}
-		0 & \;\; \mbox{if } \;\; x_2=1  \\
-		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right)  & \;\; \mbox{if } \;\; x_2=2 \,.
+		0, & \;\; \mbox{if } \;\; x_2=1,  \\
+		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right),  & \;\; \mbox{if } \;\; x_2=2 \,.
 	\end{array}
@@
 There are clear  non-linear patterns and the right bottom subplot also shows that the conditional function associated with WBC has an interaction with TDI.
-
-Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
+Figure~\ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
 As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.

### 2020-04-21 Paper/manuscript_v2/manuscript_07_01.tex
Edit to sec 5.1

@@
 
-This example consists of a simulated dataset with $n=250$ ($i=1,2,\dots,n$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
-To form the final noisy dataset $\bm{y}$, Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws.
+In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
+The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and lengthscale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
+$\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
 
@@
 %
-where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $I$ representing the identity matrix.
-The previous formulation corresponds to the latent form of a GP model.
+where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $\bm{I}$ representing the identity matrix.
 The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
@@
 Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
-The true data-generative function and the noisy observations are also plotted.
-The sample observations are plotted as circles and the out-of-sample or test data, which have not been taking part on training the models, are plotted as crosses.
+The true data-generating function and the noisy observations are also plotted.
+The sample observations are plotted as circles and the out-of-sample or test data are plotted as crosses.
 The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
-The posteriors of the three models, exact GP, HSGP and spline, are pretty similar within the interpolation input space.
+The posteriors of the three models, exact GP, HSGP and spline, are similar in the interpolation regions of the input space.
 However, when extrapolating the spline model solution clearly differs from the exact GP and HSGP models as well as the actual observations.
@@
 The SRMSE is computed against the data-generating model.
-From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the exact GP model for interpolating and extrapolating data.
+From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it is seen that the HSGP method yield a good approximation of the exact GP model for both interpolation and extrapolation.
 However, the spline model does not extrapolate data properly.
@@
 Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model, for this particular application with a univariate input space.
-Also, the increase in computation time as a function of the number of basis functions in a univariate input space is relatively slight.
+The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
+Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
 
@@
 
-In the online supplemental material, some other case studies are presented.
-From those examples, it can be seen how computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-Even though, in a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
-However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
+In the online supplemental material, additional case studies are presented.
+From those examples, it can be seen how the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
+Even though, for a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
+For moderate sized datasets, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
+However, the HSGP method will be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
 In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.

### 2020-04-21 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to sec 5.2

@@
 
-Let us denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
+Let $y_t$ denote the number of births on the $t$'th day. The observation model is a normal distribution with mean function $\mu(t)$ and noise variance $\sigma^2$,
 %
@@
 %
-The mean function $\mu(t)$ will be defined as an additive model in the form: 
+The mean function $\mu(t)$ will be defined as an additive model in the form
 %
@@
 Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
-Different levels of sparsity can be accommodated by changing the value of $\tau$: with large $\tau$ all the variables have very diffuse priors with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
+Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
 %More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.
@@
 We use $J=10$ cosine terms.
-The lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
+The lengthscale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
 

### 2020-04-21 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to sec 5.3

@@
 
-As the WBC predictor values were strictly positive and highly skewed, logarithm transformation is used for it.
+As the WBC predictor values were strictly positive and highly skewed, a logarithm transformation is used.
 Continuous predictors were normalized to have zero mean and unit standard deviation.
@@
 \begin{equation*}
-  p(t_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
+  p(t_i|f_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
 \end{equation*}
@@
 \begin{equation*}
-p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
+p(y_i > t_i|f)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
 \end{equation*}
@@
 Figure~\ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).

### 2020-04-21 Paper/manuscript_v2/manuscript_07_01.tex
update eq 16 of sec 5.3 and related text around

@@
 \begin{equation*}
-  p(t_i|f_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
+  p(t_i \mid f_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
 \end{equation*}
@@
 \begin{equation*}
-p(y_i > t_i|f)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
+p(y_i > t_i \mid f)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
 \end{equation*}
@@
 where $y_i>t_i$ denotes the unobserved survival time.
-The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$ depending on predictors $x$ and hyperparameters $\theta=(\alpha,\ell)$,
-%
-\begin{equation*} 
-f(\bm{x}) \sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta)),
-\end{equation*}
-where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
-% Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ are multivariate Gaussian distributed with mean function $\mu(\cdot)$ and covariance matrix $K$, where $\mu(\bm{x}_i)=c + \bm{\beta}\bm{x}_i$ and $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,n$.
-The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-A scalar lengthscale is used in the multivariate covariance function assuming an isotropic function.
-
+The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$.
 Due to the predictor \textit{sex} ($x_2$) being a categorical variable (`1' for female and `2' for male), we apply indicator variable coding for the GP functions, in a similar way such coding is applied in linear models \citep{Gelman+Hill+Vehtari:2020:ROS}.
-A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise:
-%
-\begin{align} 
-f(\bm{x}) &\sim \GP\left(c + \bm{\beta}\bm{x},\, k(\bm{x},\bm{x}', \theta_1)\right) + g(\bm{x}), \label{ch5_eq_gpml1}
-\end{align}
-with
-%
-\begin{align} \label{ch5_eq_gpml2}
-g(\bm{x}) =
-\left\{
-	\begin{array}{ll}
-		0, & \;\; \mbox{if } \;\; x_2=1,  \\
-		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right),  & \;\; \mbox{if } \;\; x_2=2 \,.
-	\end{array}
-\right. 
-\end{align}
-
-%\begin{align} 
-%f(\bm{x}) &\sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
-%g(\bm{x}) &\sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2 \,|\, x_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
-%f(\bm{x} \, |\, x_2\!=\!2) &= f(\bm{x} \,|\, x_2\!=\!1) + g(\bm{x}). \nonumber
+The latent function $f(\bm{x})$, besides of being centered around a linear model, is composed of a general mean GP function, $h(\bm{x})$, defined for all observations, plus a second GP function, $g(\bm{x})$, that only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise:
+%
+\begin{align*}
+h(\bm{x}) &\sim \GP\left(\bm{0}, \, k(\bm{x},\bm{x}', \theta_0) \right),  \\[1pt] 
+%
+g(\bm{x}) &\sim \GP\left(\bm{0}, \, k(\bm{x},\bm{x}', \theta_1) \mid x_2=2 \right),\\[1pt] 
+%
+f(\bm{x}) &= c + \bm{\beta} \bm{x} + h(\bm{x}) + g(\bm{x}). 
+\end{align*}
+%
+Above, $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
+$\theta_0$ contains the hyperparameters $\alpha_0$ and $\ell_0$ which are the marginal variance and lengthscale of the general mean GP function, and $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of a GP function specific to the male sex ($x_2=2$).
+Scalar lengthscales, $l_0$ and $l_1$, are used in both multivariate covariance functions, assuming isotropic functions.
+
+%with
+%%
+%\begin{align} \label{ch5_eq_gpml2}
+%g(\bm{x}) =
+%\left\{
+%	\begin{array}{ll}
+%		0, & \;\; \mbox{if } \;\; x_2=1,  \\
+%		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right),  & \;\; \mbox{if } \;\; x_2=2 \,.
+%	\end{array}
+%\right. 
 %\end{align}
@@
 \end{figure}
-Above, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a GP function specific to the male sex ($x_2=2$).
-
-Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+
+Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
 

### 2020-04-22 Paper/manuscript_v2/manuscript_07_01.tex
fix statement about minimun valid c in pag 11
add details of sampling and testing data in sec 5.1

@@
 \begin{itemize}
-\item As $\ell$ increases, $c$ and $m$ required for a close enough approximation decrease.
+\item As $\ell$ increases, $m$ required for a close enough approximation decrease.
 \item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve a close approximation.
@@
 This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+A $\ell$ increases, the minimum valid $c$ also increases.
 \end{itemize}
@@
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
-
-The exact GP model for fitting this simulated dataset $\bm{y}$ can be written as follows,
+From this synthetic dataset, $155$ data points are used as sample data, $45$ as interpolation out-of-sample test data, and $50$ as extrapolation out-of-sample test data.
+
+The exact GP model for fitting and predicting this simulated dataset $\bm{y}$ can be written as follows,
 %

### 2020-04-22 Paper/manuscript_v2/manuscript_07_01.tex
add details of the hyperparameters' prior distributions of sec 5.1

@@
 
+The joint posterior parameter distributions are estimated from the sampling data using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for lengthscale $\ell$, either in exact GP or HSGP models.
+
 Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.

### 2020-04-22 Paper/manuscript_v2/manuscript_07_01.tex
added some acks

--- 2020-04-21 Paper/manuscript_v2/manuscript_07_01.tex
@@
 Appendix~\ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
-Online supplemental material with more case studies illustrating the performance and applicability of the method can be found online at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
+Online supplemental material with more case studies illustrating the performance and applicability of the method can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
 
@@
 A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
-GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_N) \right\rbrace$, is a multivariate Gaussian distribution.
+GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_K) \right\rbrace$, is a multivariate Gaussian distribution.
 
@@
 \end{align*}
-%
+
 The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$.
@@
 Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying eq.
-(\ref{eq_diff_covs}).
+More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying equation (\ref{eq_diff_covs}).
 Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function.
-We considered an approximation to be a close enough when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
-(\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$:
+We considered an approximation to be a close enough when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$:
 %
@@
 
-In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
-The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and lengthscale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
-$\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
+This example consists of a simulated dataset with $n=250$ ($i=1,2,\dots,n$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1] \subset {\rm I\!R}$.
+To form the final noisy dataset $\bm{y}$, Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws.
 
@@
 \bm{y} &= \bm{f} + \bm{\epsilon} \\
-\bm{\epsilon} &\sim \Normal(\bm{0}, \sigma^2 \bm{I}) \\
+\bm{\epsilon} &\sim \Normal(0, \sigma^2  I) \\
 f(x) &\sim \GP(0, k(x, x', \theta)),
@@
 %
-where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $\bm{I}$ representing the identity matrix.
-The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
-The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,2,\dots,n$.
+where $I$ represents the identity matrix and $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function values to the noisy data.
+The previous formulation corresponds to the latent form of a GP model.
+The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with a Mat{\'e}rn($\nu$=3/2) covariance function $k$.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,\dots,n$.
 
@@
 \begin{equation*}\label{ch5_eq_marginalizedgp_simudata1}
-\bm{y} \sim \Normal(\bm{0},\bm{K}+ \sigma^2 \bm{I} ),
+\bm{y} \sim \Normal(0,\bm{K}+ \sigma^2 \bm{I} ),
 \end{equation*}
@@
 Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
-The true data-generating function and the noisy observations are also plotted.
-The sample observations are plotted as circles and the out-of-sample or test data are plotted as crosses.
+The true data-generative function and the noisy observations are also plotted.
+The sample observations are plotted as circles and the out-of-sample or test data, which have not been taking part on training the models, are plotted as crosses.
 The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
-The posteriors of the three models, exact GP, HSGP and spline, are similar in the interpolation regions of the input space.
+The posteriors of the three models, exact GP, HSGP and spline, are pretty similar within the interpolation input space.
 However, when extrapolating the spline model solution clearly differs from the exact GP and HSGP models as well as the actual observations.
@@
 The SRMSE is computed against the data-generating model.
-From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it is seen that the HSGP method yield a good approximation of the exact GP model for both interpolation and extrapolation.
+From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it can be seen a close approximation of the HSGP model to the exact GP model for interpolating and extrapolating data.
 However, the spline model does not extrapolate data properly.
@@
 Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
-Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
-
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
+The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model, for this particular application with a univariate input space.
+Also, the increase in computation time as a function of the number of basis functions in a univariate input space is relatively slight.
+
+The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}.
 %
@@
 
-In the online supplemental material, additional case studies are presented.
-From those examples, it can be seen how the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-Even though, for a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
-For moderate sized datasets, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
-However, the HSGP method will be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
+In the online supplemental material, some other case studies are presented.
+From those examples, it can be seen how computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
+Even though, in a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
+However, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
 In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
@@
 \citet{gelman2013bayesian} presented an analysis using exact GP and maximum a posteriori inference.
-As the total number of days within the period is $T=7305$ ($t=1,2,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
+As the total number of days within the period is $T=7305$ ($t=1,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
 We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix~\ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
 
-Let $y_t$ denote the number of births on the $t$'th day. The observation model is a normal distribution with mean function $\mu(t)$ and noise variance $\sigma^2$,
+Let us denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
 %
@@
 %
-The mean function $\mu(t)$ will be defined as an additive model in the form
+The mean function $\mu(t)$ will be defined as an additive model in the form: 
 %
@@
 \begin{equation*}
-f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \alpha_1 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
-\end{equation*}
-%
-which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
-$\alpha_1$ and $\ell_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component.
+f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \sigma_1^2 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
+\end{equation*}
+%
+which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $\bm{K}_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
+
 The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
@@
 \begin{equation*}
-f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \alpha_2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
 \end{equation*}
@@
 \begin{equation*}
-f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \alpha_3 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
+f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
 \end{equation*}
@@
 Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
-Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
+Different levels of sparsity can be accommodated by changing the value of $\tau$: with large $\tau$ all the variables have very diffuse priors with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
 %More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.
@@
 We use $J=10$ cosine terms.
-The lengthscale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
+The lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}\,.
+The Stan model code for the approximate GP model of this case study can be found at \small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}.
 
@@
 
-As the WBC predictor values were strictly positive and highly skewed, a logarithm transformation is used.
+As the WBC predictor values were strictly positive and highly skewed, logarithm transformation is used for it.
 Continuous predictors were normalized to have zero mean and unit standard deviation.
@@
 \begin{equation*}
-  p(t_i \mid f_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
+  p(t_i)= \LogNormal(t_i \mid f(\bm{x}_i),\sigma^2).
 \end{equation*}
@@
 \begin{equation*}
-p(y_i > t_i \mid f)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
+p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
 \end{equation*}
@@
 where $y_i>t_i$ denotes the unobserved survival time.
-The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$.
-Due to the predictor \textit{sex} ($x_2$) being a categorical variable (`1' for female and `2' for male), we apply indicator variable coding for the GP functions, in a similar way such coding is applied in linear models \citep{Gelman+Hill+Vehtari:2020:ROS}.
-The latent function $f(\bm{x})$, besides of being centered around a linear model, is composed of a general mean GP function, $h(\bm{x})$, defined for all observations, plus a second GP function, $g(\bm{x})$, that only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise:
-%
-\begin{align*}
-h(\bm{x}) &\sim \GP\left(\bm{0}, \, k(\bm{x},\bm{x}', \theta_0) \right),  \\[1pt] 
-%
-g(\bm{x}) &\sim \GP\left(\bm{0}, \, k(\bm{x},\bm{x}', \theta_1) \mid x_2=2 \right),\\[1pt] 
-%
-f(\bm{x}) &= c + \bm{\beta} \bm{x} + h(\bm{x}) + g(\bm{x}). 
-\end{align*}
-%
-Above, $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
-$\theta_0$ contains the hyperparameters $\alpha_0$ and $\ell_0$ which are the marginal variance and lengthscale of the general mean GP function, and $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of a GP function specific to the male sex ($x_2=2$).
-Scalar lengthscales, $l_0$ and $l_1$, are used in both multivariate covariance functions, assuming isotropic functions.
-
-%with
-%%
-%\begin{align} \label{ch5_eq_gpml2}
-%g(\bm{x}) =
-%\left\{
-%	\begin{array}{ll}
-%		0, & \;\; \mbox{if } \;\; x_2=1,  \\
-%		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right),  & \;\; \mbox{if } \;\; x_2=2 \,.
-%	\end{array}
-%\right. 
+
+The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$ depending on predictors $x$ and hyperparameters $\theta=(\alpha,\ell)$,
+%
+\begin{equation*} 
+f(\bm{x}) \sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta)),
+\end{equation*}
+where $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
+% Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ are multivariate Gaussian distributed with mean function $\mu(\cdot)$ and covariance matrix $K$, where $\mu(\bm{x}_i)=c + \bm{\beta}\bm{x}_i$ and $K_{rs}=k(\bm{x}_r,\bm{x}_s,\theta)$, with $r,s=1,\dots,n$.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+A scalar lengthscale is used in the multivariate covariance function assuming an isotropic function.
+
+Due to the predictor \textit{sex} ($x_2$) being a categorical variable ('1' for female and '2' for male), we apply indicator variable coding for the GP functions, in a similar way such coding is applied in linear models \citep{Gelman+Hill+Vehtari:2020:ROS}.
+A general mean GP function is defined for all observations, while a second GP function only applies to one of the predictor levels ('male' in this case) and is set to zero otherwise:
+%
+\begin{align} 
+f(\bm{x}) &\sim \GP\left(c + \bm{\beta}\bm{x},\, k(\bm{x},\bm{x}', \theta_1)\right) + g(\bm{x}), \label{ch5_eq_gpml1}
+\end{align}
+with
+%
+\begin{align} \label{ch5_eq_gpml2}
+g(\bm{x}) =
+\left\{
+	\begin{array}{ll}
+		0 & \;\; \mbox{if } \;\; x_2=1  \\
+		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right)  & \;\; \mbox{if } \;\; x_2=2 \,.
+	\end{array}
+\right. 
+\end{align}
+
+%\begin{align} 
+%f(\bm{x}) &\sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_1)) \label{ch5_eq_gpml1} \\
+%g(\bm{x}) &\sim \GP(c + \bm{\beta}\bm{x}, k(\bm{x},\bm{x}', \theta_2 \,|\, x_2\!=\!2)) \label{ch5_eq_gpml2} \\ 
+%f(\bm{x} \, |\, x_2\!=\!2) &= f(\bm{x} \,|\, x_2\!=\!1) + g(\bm{x}). \nonumber
 %\end{align}
@@
 \end{figure}
-
-Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+Above, $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale of the general mean GP function, and $\theta_2$ contains the hyperparameters $\alpha_2$ and $\ell_2$ which are the marginal variance and lengthscale, respectively, of a GP function specific to the male sex ($x_2=2$).
+
+Using the HSGP approximation, the functions $f(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
 
@@
 There are clear  non-linear patterns and the right bottom subplot also shows that the conditional function associated with WBC has an interaction with TDI.
-Figure~\ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+
+Figure \ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not allowed when large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).
@@
 
-The Stan model code for the exact GP and the approximate GP models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
+The Stan model code for the exact GP and the approximate GP models of this case study can be found at \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}.
 %
@@
 \begin{equation} \label{ch5_eq_cov_periodic_taylor_approx_2}
-k(x,x')= \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
+k(x,x')= \sigma^2 \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
 \end{equation}
@@
 \begin{equation} \label{ch5_eq_f_period}
-f(x) \approx \alpha^{1/2} \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
+f(x) \approx \sigma \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
 \end{equation}
@@
 
-%\section*{Acknowledgment}
+\section*{Acknowledgements}
+
+We thank Academy of Finland (grants 298742 and 313122), Finnish Center
+for Artificial Intelligence, and Technology Industries of Finland
+Centennial Foundation (grant 70007503; Artificial Intelligence for
+Research and Development) for partial support of this research. We
+also acknowledge the computational resources provided by the Aalto
+Science-IT project.
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2020-04-22 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to sec 5.1

@@
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
-From this synthetic dataset, $155$ data points are used as sample data, $45$ as interpolation out-of-sample test data, and $50$ as extrapolation out-of-sample test data.
+We split the dataset into three parts: $155$ data points are used for fitting the model (training set), $45$ data ponts are used for the interpolation test set, and the remaining $50$ data points are used for the extrapolation test set.
 
@@
 
-The joint posterior parameter distributions are estimated from the sampling data using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
-A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for lengthscale $\ell$, either in exact GP or HSGP models.
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for lengthscale $\ell$.
+We use the same prior distributions for the exact GP model as for the HSGP models.
 

### 2020-04-22 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to sec 5.3
Added ack

@@
 %
-g(\bm{x}) &\sim \GP\left(\bm{0}, \, k(\bm{x},\bm{x}', \theta_1) \mid x_2=2 \right),\\[1pt] 
-%
-f(\bm{x}) &= c + \bm{\beta} \bm{x} + h(\bm{x}) + g(\bm{x}). 
+g(\bm{x}) &\sim \GP\left(\bm{0}, \, k(\bm{x},\bm{x}', \theta_1)\right),\\[1pt] 
+%
+f(\bm{x}) &= c + \bm{\beta} \bm{x} + h(\bm{x}) + \mathbb{I}\!\left[x_2=2\right]g(\bm{x}),
 \end{align*}
-%
+where $\mathbb{I}\left[\cdot\right]$ is an indicator function.
 Above, $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
@@
 
-We thank Academy of Finland (grants 298742 and 313122), Finnish Center
-for Artificial Intelligence, and Technology Industries of Finland
-Centennial Foundation (grant 70007503; Artificial Intelligence for
-Research and Development) for partial support of this research. We
-also acknowledge the computational resources provided by the Aalto
-Science-IT project.
+We thank Academy of Finland (grants 298742, 308640, and 313122), Finnish Center for Artificial Intelligence, and Technology Industries of Finland Centennial Foundation (grant 70007503; Artificial Intelligence for Research and Development) for partial support of this research.
+We also acknowledge the computational resources provided by the Aalto Science-IT project.
 

### 2020-04-23 Paper/manuscript_v2/manuscript_07_01.tex
layout the paper

@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centering
@@
 %\vspace{-9mm}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
 %\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig1_legend.pdf}}
+\vspace{-1mm}
 \caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 10mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 0mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
 %\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig2_legend.pdf}}
+\vspace{-1mm}
 \caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig4_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig4_MSE_vs_c.pdf}}
+\begin{figure}[!t]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{ch5_fig4_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig4_MSE_vs_c.pdf}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the exact GP model.
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig5_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig5_magnitud_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 28mm 119mm 0mm, clip]{ch5_fig5_legend.pdf}}
+\begin{figure}[!h]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig5_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig5_magnitud_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 119mm 0mm, clip]{ch5_fig5_legend.pdf}}
 \caption{Estimated lengthscale (left) and marginal variance (right) parameters of both exact GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
@@
 
-\begin{figure}
+Comparing the accuracy of the lengthscale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig9_MSE_varing_lscale}, we see that they agree closely with each other for medium lengthscales.
+That is, a good estimation of the lengthscale implies a small RMSE.
+This is no longer true for very small or large lengthscales.
+In small lengthscales, even very small inaccuracies  may have a strong influence on the posteriors predictions and thus on the RMSE.
+In large lengthscales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}.
+We observe that whenever the estimated lengthscale exceeds the minimally estimable lengthscale, the RMSE of the posterior predictions is small
+(see Figure \ref{fig9_MSE_varing_lscale}). Conversely, when the estimated lengthscale is smaller than the minimally estimable one, the RMSE becomes very large.
+
+\begin{figure}[H]
 \centering
@@
 \subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
-\vspace{-3mm}
 \caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
@@
 \includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part3.pdf} \\
-\includegraphics[scale=0.31, trim = 0mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 0mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part6.pdf}
+\includegraphics[scale=0.31, trim = 0mm 2mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part6.pdf}
 \end{tabular}
@@
 \centering
-\includegraphics[scale=0.38]{ch5_fig9_MSE_varing_lscale.pdf}
+\includegraphics[scale=0.38, trim = 0mm 9mm 0mm 0mm, clip]{ch5_fig9_MSE_varing_lscale.pdf}
 \caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects ($\ell$).}
@@
 
-Comparing the accuracy of the lengthscale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig9_MSE_varing_lscale}, we see that they agree closely with each other for medium lengthscales.
-That is, a good estimation of the lengthscale implies a small RMSE.
-This is no longer true for very small or large lengthscales.
-In small lengthscales, even very small inaccuracies  may have a strong influence on the posteriors predictions and thus on the RMSE.
-In large lengthscales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}.
-We observe that whenever the estimated lengthscale exceeds the minimally estimable lengthscale, the RMSE of the posterior predictions is small
-(see Figure \ref{fig9_MSE_varing_lscale}). Conversely, when the estimated lengthscale is smaller than the minimally estimable one, the RMSE becomes very large.
-
+\newpage
 \section{Case studies}\label{ch5_sec_cases}
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig11_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig11_MSE_exI_extra.pdf}}
+\begin{figure}[!t]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 5mm 0mm, clip]{ch5_fig11_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig11_MSE_exI_extra.pdf}}
 \subfigure{\includegraphics[scale=0.40, trim = 29mm 55mm 110mm 0mm, clip]{ch5_fig11_MSE_exI_legend.pdf}}
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig11_time_exI.pdf}}
+\begin{figure}[!h]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig11_time_exI.pdf}}
 \caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
@@
 
-\begin{figure}
+\begin{figure}[!h]
 \centering
@@
 
+\newpage 
 \subsection{Leukemia data}\label{ch5_sec_bf_caseVI}
@@
 \centering
-\includegraphics[scale=0.70, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia_02.pdf}
+\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia_02.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig22_elpd_leukemia.pdf}} 
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig23_time_leukemia.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 120mm 0mm, clip]{ch5_fig24_legend_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{ch5_fig22_elpd_leukemia.pdf}} 
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig23_time_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 50mm 120mm 0mm, clip]{ch5_fig24_legend_leukemia.pdf}}
 \caption{Expected log predictive density (ELPD; left) and time of computation in seconds per iteration (iteration of the HMC sampling method; right) as a function of the number of basis functions $m$ and boundary factor $c$.}
@@
 \centering
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix.pdf}}
 \hspace{3mm}
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix_zoom.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix_zoom.pdf}}
 \caption{Relation among the minimum number of terms $J$ in the approximation and the lengthscale ($\ell$) of the periodic squared exponential covariance function.
@@
 \begin{equation} \label{ch5_eq_cov_periodic_taylor_approx_2}
-k(x,x')= \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
+k(x,x')\approx \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
 \end{equation}

### 2020-04-24 Paper/manuscript_v2/manuscript_07_02.tex
remove unnecessary figures and commented text. Update file name

--- 2020-04-23 Paper/manuscript_v2/manuscript_07_01.tex
@@
 \usepackage{listings}
-%\usepackage{slashbox}
 \usepackage{caption}
 \captionsetup{font=footnotesize}
-\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
-
-%\bibliographystyle{authordate1}
-% TODO macro
-\newcommand{\todo}[1]{\textcolor{blue}{\textbf{[#1]}}}
-\newcommand{\resp}[1]{\textcolor{red}{\textbf{[#1]}}}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,breaklinks=true]{hyperref}
 
@@
 
-%\tableofcontents
-
 \section{Introduction}\label{sec_bf_intro}
@@
 \citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for an accurate approximation.
-% More recent developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}.
-More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.
 An alternative class of low-rank approximations is based on forming a basis function approximation with $m \ll n$ basis functions.
@@
 Recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.,][]{grenander1981abstract} is presented by \citet{JSSv090i10}.
-% , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).
-
-%{\color{blue} Paul: Can we clarify how our discussed method relates to the former paragraph,
-%which introduces all sorts of basis function approximations. I.e. is our method
-%just one out of many or what makes it special? Perhaps we can somehow use the information provided
-%in: While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank.
-Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.}
 
@@
 
-%\textcolor{red}{Aki: I moved this from section 2, needs to be combined with other intro}
-%We propose an approximate framework for fast and accurate inference
-%for GPs. Using a basis function expansion, we
-%approximate the GPs with a linear model. This
-%representation has three main advantages: 1) it makes inference
-%considerably faster due to the linear structure, 2) it is simple to
-%implement, which makes it easy to use GPs as building
-%blocks in more complicated models and can be used as latent function
-%in non-Gaussian observational models allowing modelling flexibility,
-%3) it can be made arbitrary accurate and the trade-off between
-%computational complexity and approximation accuracy can easily be
-%controlled.
-
 While \citet{solin2018hilbert} have fully developed the mathematical theory behind this specific approximation of GPs, further work is needed for its practical implementation in probabilistic programming frameworks.
@@
 We also propose  diagnostics to indicate whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
-
-%In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable.
-We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.
 
@@
 Our goal is to obtain the posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
-% That is, we want to obtain the predictive distribution $p(f^* \mid \mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\
 
@@
 where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$.
-%By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\bm{f}$ analytically,
-% \begin{align*}
-	% p(f^*| \bm{f}) = \Normal(f^*| \bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{k}_{\bm{f},f^*}  )
-% \end{align*}
-
-% {\color{red} Aki: the following until ``While Gaussian'' seems too basic to be in the paper:}
-
-% The joint distribution of the observations $\bm{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\bm{f}$ and $f^*$, $p(\bm{y}, \bm{f}, f^*)$, is the product of the conditional distribution for $\bm{y}$ given $\bm{f}$ and the joint distribution for $\bm{f}$ and $f^*$,
-% \begin{align*}
-% 	p(\bm{y}, \bm{f}, f^*) = p(\bm{y}|\bm{f})p(\bm{f}, f^*)
-% \end{align*}
-
-% By marginalizing over $\bm{f}$ and conditioning on vector of observations $\bm{y}$, we obtain the posterior distribution of interest
-% %
-% \begin{align}\label{eq:posterior}
-% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\mathrm{d}\bm{f}}{p(\bm{y})},
-% \end{align}
-% where $p(\bm{y})$ is the marginal likelihood and is given by
-% %
-% \begin{align}\label{eq:evidence}
-% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \mathrm{d}\bm{f}\mathrm{d}f^*.
-% \end{align}
-% %
-% If the observation model $p(\bm{y}|\bm{f})$ is Gaussian, both integrals in eq.
-\eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters.
-For example, an isotropic Gaussian likelihood yields the following closed-form solution
-% \begin{align*}
-% 	p(f^*| \bm{y}) = \Normal(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
-% \end{align*}
-% where $\sigma^2$ is the noise variance.
 
@@
 \centering
-%\subfigure{\includegraphics[scale=0.38, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig1_legend_01.pdf}}\\
-%\vspace{-9mm}
 \subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
@@
 \subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
-%\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig1_legend.pdf}}
 \vspace{-1mm}
@@
 \subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
-%\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig2_legend.pdf}}
 \vspace{-1mm}
@@
 \end{tabular}
-\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.
-%The dashed red line of the exact GP can hardly be seen in the some of the plots because it is under the blue lines.
-}
+\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig3_Post_part1}
@@
 \caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
-%The dashed red line of the exact GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.
 }
@@
 
-%This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid a excessive computational time, especially in multivariate input spaces.
-
 Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
@@
 \end{equation*}
-% which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.
 
@@
 \end{equation*}
-% which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.
 
@@
 Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
-%More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.
 
@@
 
-%We denote the matrix $X=[\bm{x}_1\, \cdots \bm{x}_i\, \cdots \bm{x}_n] \in {\rm I\!R}^{n\times 4}$ where $\bm{x}_i=(x_1,x_2,x_3,x_4) \in {\rm I\!R}^{4}$ the vector of predictor values for observations $i$
-
-As the WBC predictor values were strictly positive and highly skewed, a logarithm transformation is used.
-Continuous predictors were normalized to have zero mean and unit standard deviation.
-%Survival time was normalized to have zero mean for the logarithm of time.
+As the WBC predictor values were strictly positive and highly skewed, a logarithm transformation is used. Continuous predictors were normalized to have zero mean and unit standard deviation.  
 We assume a log-normal observation model for the observed survival time, $t_i$, with a function of the predictors, $f(\bm{x}_i):{\rm I\!R}^4 \to {\rm I\!R}$, as the location parameter, and $\sigma$ as the Gaussian noise:
@@
 \end{align*}
+%
 where $\mathbb{I}\left[\cdot\right]$ is an indicator function.
@@
 Scalar lengthscales, $l_0$ and $l_1$, are used in both multivariate covariance functions, assuming isotropic functions.
-
-%with
-%%
-%\begin{align} \label{ch5_eq_gpml2}
-%g(\bm{x}) =
-%\left\{
-%	\begin{array}{ll}
-%		0, & \;\; \mbox{if } \;\; x_2=1,  \\
-%		\GP\left(0, k(\bm{x},\bm{x}', \theta_2)\right),  & \;\; \mbox{if } \;\; x_2=2 \,.
-%	\end{array}
-%\right. 
-%\end{align}
-
+%
 \begin{figure}
 \centering
-\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia_02.pdf}
+\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
@@
 
-%In terms of scaling, one main drawback of the method is the scaling with respect to the number of input dimensions, as the number of basis functions $m$ scales exponentially with the number of dimensions.
-
 The main contribution of this paper is an in-depth analysis and diagnosis of the performance and accuracy of the approximation in relation to the key factors of the method, that is, the number of basis functions, the boundary condition of the Laplace eigenfunctions, and the non-linearity of the function to be learned.
@@
 
-%Future research will focus on constructing analytical models for the relationships between the key factors of the number of basis functions, the boundary factor and the lengthscale of the function, depicted in Figures \ref{fig6_lscale_vs_J_vs_c}, \ref{fig6_lscale_vs_J_vs_c_Matern} and \ref{ch5_fig28_m_lscale_periodic}, on which ultimately depend the performance of the approximation.
-This analytical models can be useful to automatize the diagnosis of the performance of the approximation.
-These relationships have been obtained, in the present study, for the unidimensional case only.
-So, another future research line will be focused on analyzing these relationships in the multidimensional case, building useful graphs or analytical models that encode these relationships in multidimensional cases.
-
 \appendix
@@
 This allows for a Fourier representation of the operator $\mathcal{K}$ as a transfer function which is the spectral density of the Gaussian process.
-%When the covariance function is homogeneous, the corresponding operator $\mathcal{K}$ will be translation invariant thus allowing for Fourier-representation as a transfer function.
-This transfer function is the spectral density of the Gaussian process
-%%
-%\begin{eqnarray*}
-%S(\bm{w}) = \mathcal{F}(\mathcal{K}). \nonumber
-%\end{eqnarray*}
-%
-%\noindent 
 Thus, the spectral density $S(\bm{w})$ also gives the approximate eigenvalues of the operator $\mathcal{K}$.
@@
 
-%If the requirement of a valid covariance function is relaxed and only an optimal series approximation is required, taking the limit $J \rightarrow \infty$ in the sub-sums in equation (\ref{ch5_eq_q}) gives the following variance coefficients
 An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}), and thus leading to the following variance coefficients:
@@
 A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.
-
-%Note that the terms in equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}) are underneath bounded approximations for any $J$.
 
@@
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2020-04-28 Paper/manuscript_v2/manuscript_07_02.tex
the order of the panels doesn't match the ordering in the caption; posterior sigma and posterior Cov are reversed relative to the caption

@@
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
+\caption{Mean posterior predictive functions (left), posterior standard deviations (center) and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
   \label{fig1_Post_J}
@@
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
+\caption{Mean posterior predictive functions (left), posterior standard deviations (center) and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
   \label{fig2_Post_L}
@@
 \end{tabular}
-\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
+\caption{Posterior mean predictive functions (left), posterior standard deviations (center) and covariance functions (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig3_Post_part1}

### 2020-05-11 Paper/manuscript_v2/manuscript_07_02.tex
Fix captions in fig 1&2

--- 2020-04-24 Paper/manuscript_v2/manuscript_07_02.tex
@@
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
+\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
   \label{fig1_Post_J}
@@
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
+\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
   \label{fig2_Post_L}

### 2020-05-11 Paper/manuscript_v2/manuscript_07_02.tex
Minor edits

@@
 
-The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and defines a correlation structure which characterizes the correlations between function values at different inputs.
+The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and characterizes the correlations between function values at different locations in the input space.
 A covariance function needs to be symmetric and positive semi-definite \citep{rasmussen2006gaussian}.
@@
 
-Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
+Each multivariate eigenfunction $\phi^{\ast}_j:\Omega \rightarrow {\rm I\!R}$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
 Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,2,\ldots,m^{\ast}$, we have
@@
 In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
-For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects on the posterior function.
+For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
 
@@
 \begin{equation} \label{ch5_eq_cov_periodic}
-k(\bm{\tau})= \alpha \exp\left( -\frac{2\sin^{2}(\omega_0\frac{\bm{\tau}}{2})}{\ell^2}\right),
+k(\tau)= \alpha \exp\left( -\frac{2\sin^{2}(\omega_0\frac{\tau}{2})}{\ell^2}\right),
 \end{equation}

### 2020-05-13 Paper/manuscript_v2/manuscript_07_02.tex
More minor edits

@@
 Gaussian processes are powerful non-parametric probabilistic models for stochastic functions.
-However they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
+However, they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
 In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions.
@@
 Gaussian processes (GPs) are flexible statistical models for specifying probability distributions over multi-dimensional non-linear functions \citep{rasmussen2006gaussian,neal1997monte}.
-Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate Gaussian.
+Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate Gaussian distribution.
 GPs are defined by a mean and a covariance function.
@@
 The key element of a GP is the covariance function that defines the dependence structure between function values at different inputs.
-However, computing the posterior distribution of a GP comes with a computational issue because of the need of inverting the covariance matrix.
+However, computing the posterior distribution of a GP comes with a computational issue because of the need to invert the covariance matrix.
 Given $n$ observations in the data, the computational complexity and memory requirements of computing the posterior distribution for a GP in general scale as $O(n^3)$ and $O(n^2)$, respectively.
@@
 Using a basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.
-The linear model structure makes GPs easy to implement as a building block in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if the approximation specific computation is simple.
+The linear model structure makes GPs easy to implement as building blocks in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if the approximation specific computation is simple.
 Furthermore, a linear representation of a GP makes it easier to be used as latent function in non-Gaussian observational models allowing for more modelling flexibility.
@@
 Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, and GPflow \citep{GPflow2017}, each provide efficient implementations only for a restricted range of GP-based models.
-In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
+In this paper, we do not focus on the fastest possible inference for a small set of specific GP models, but instead we are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
 
@@
 
-A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
+A GP is a stochastic process which defines the distribution of a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
 GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_N) \right\rbrace$, is a multivariate Gaussian distribution.
@@
 In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting.
-Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
+Consider a data set $\mathcal{D} = \left\lbrace (\bm{x}_n, y_n) \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with a GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
 This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
@@
 Below, we will present the main results around HSGPs relevant for practical applications.
-More details and mathematical proofs are provided by \citet{solin2018hilbert}.
+More details on the theoretical background are provided by \citet{solin2018hilbert}.
 Our starting point for presenting the method is the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
@@
 
-Within $\Omega$, we can write any stationary covariance function with input values $\{x,x'\} \in \Omega$ as
+Within $\Omega$, we can write any stationary covariance function with input values $x,x' \in \Omega$ as
 %
@@
 %
-where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
+where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ is the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
 The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
@@
 \begin{split}
--\nabla^2 \phi_j(x)&=\lambda \phi_j(x), \hspace{1cm}  x \in \Omega \\ 
+-\nabla^2 \phi_j(x)&=\lambda_j \phi_j(x), \hspace{1cm}  x \in \Omega \\ 
 \phi_j(x)&= 0, \hspace{1.85cm}   x \notin \Omega.
@@
 
-Thus, the Gram matrix $K$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
+Thus, the Gram matrix $\bm{K}$ for the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n}$ can be represented as
 %
@@
 \begin{equation}\label{eq_approxf}
-f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
+f(x) \approx \sum_{j=1}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
 \end{equation}

### 2020-05-15 Paper/manuscript_v2/manuscript_07_02.tex
Edit to sec 4

@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the estimated function, which is in turn is characterized by the lengthscale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the function to be estimated, which is in turn is characterized by the lengthscale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the lengthscale of the function, which depend on the particular choice of the kernel function.
-In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
+In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu$=3/2) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
 For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
@@
 
-The specific number of basis functions $m$ needed to satisfy equation (\ref{eq_diff_covs}) depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP.
+This inequality can be satisfied for arbritrary small $\epsilon$ provided that $L$ and $m$ are sufficiently large \cite[Theorem 1 and 4]{solin2018hilbert}. 
+The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the G.
 The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
 As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
-In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi = \max_{i}|x_i|$.
-Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \Omega$.
+
+In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ will include the smallest interval that contains all the inputs points $x_i$.
+Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.
+Let $S = \max_{i}|x_i|$, then it follows that $x_i \in \left[-S, S\right]$ for all $i$.
 We now define $L$ as
-%
 \begin{equation}\label{eq_boundary}
@@
 %
-where $S$ (for $S > 0$) represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
+where $S > 0$ represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
 In the following, we will refer to $c$ as the boundary factor of the approximation.
@@
 For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
-Several HSGP models with varying $m$ and $L$ are fitted to this data.
+Several HSGP models with varying $m$ and $c$ are fitted to this data.
 In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
 Figures~\ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
-For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation, either in mean or uncertainty, and the non-linearity of the estimated function, in the sense that fewer basis functions inaccurately imply larger lengthscales and consequently less wiggly functional forms.
-The higher the wigglyness of the function to be estimated, the more basis functions will be required.
-If $m$ is fixed to a large enough value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, although along the whole domain in the uncertainty.
+For $c$ fixed to a sufficiently large value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation for both the posterior mean or uncertainty.
+It is seen that if the number of basis functions $m$ is too small, the estimated function tend to be overly smooth because the necessary high frequency components are missing.
+In general, the higher the degree of wigglyness of the function to be estimated, the larger number of  basis functions will be required.
+If $m$ is fixed to a sufficiently large value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, but along the whole domain in the standard deviation.
 The approximation error for the variance tends to be bigger for the variance than for the mean.
@@
 Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The lengthscale and marginal variance of the covariance function will no longer be fixed but instead their posterior marginal distributions estimated using dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}, for both exact GP and HSGP models.
-Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the data for varying $m$ and $c$.
+The lengthscale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} for both the exact GP and the HSGP models.
+Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
 Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
 Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the exact GP model and the HSGP models.
-Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
 Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application.
@@
 \item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
-\item There exists a minimum $c$ below which a accurate approximation will never be achieved regardless of the number of basis functions $m$.
+\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
 \end{itemize}

### 2020-05-18 Paper/manuscript_v2/manuscript_07_02.tex
More edits to sec 4

@@
 Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying eq.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension.
+More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
-Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function.
-We considered an approximation to be a close enough when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
-(\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$:
-%
-\begin{equation*}
- \frac{\varepsilon}{\int k(\tau) \,\mathrm{d}\tau} < 0.01.
-\end{equation*}
-%
-Alternatively, these figures could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
-Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$.
+Similarly,  Figure \ref{fig6_lscale_vs_J_vs_c_Matern} shows the corresponding plot for the  Mat\'ern ($\nu=3/2$) covariance function.
+We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
+(\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
+%
+\begin{equation*}
+ \int | k(\tau) - \tilde{k}_m(\tau)|  \,\mathrm{d}\tau  < 0.01 \int k(\tau) \,\mathrm{d}\tau,
+\end{equation*}
+where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions.
+Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
+Of course, we may also read it as providing the minimum $\ell$ that can be approximated with high accuracy given $m$ and $c$.
 We obtain the following main conclusions:
@@
 \begin{itemize}
-\item As $\ell$ increases, $m$ required for a close enough approximation decrease.
-\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve a close approximation.
-\item For a given $\ell$ there exist a minimum $c$ under which a close approximation is never going to be achieved regardless of $m$.
+\item As $\ell$ increases, $m$ required for an accurate approximation decrease.
+\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
+\item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
 This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
@@
 
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_lscale_vs_J_vs_c}.
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_lscale_vs_J_vs_c}.
 

### 2020-06-17 Paper/Stancon_slides/HSGP_Stancon.tex
Gabriel's slides for Stancon

\documentclass[8pt]{beamer} %serif
\usefonttheme{structurebold} %structuresmallcapsserif
%\setbeameroption{show notes}

\mode<presentation>
{
\usetheme{CambridgeUS}
\usecolortheme{seagull}
}

\setbeamersize{text margin left=5mm,text margin right=8mm}

% Uncomment if want to show notes
% \setbeameroption{show notes}

\usepackage{blindtext}
%\usepackage{titlesec}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{epic,epsfig}
\usepackage{subfigure,float}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{babel}
\usepackage{afterpage}
\usepackage{url}
\urlstyle{same}
\usepackage{amsbsy}
\usepackage{eucal}
\usepackage{rotating}
\usepackage{listings}
\usepackage{lstbayes}
\usepackage{setspace}
%\usepackage{enumitem}

\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{bibentry}

\usepackage{bm}
\usepackage{graphics}
\usepackage{xcolor}
\usepackage{dsfont}
\usepackage{enumerate}
\usepackage{graphicx} % Allows including images
\usepackage{tabulary,booktabs,tabularx}
\usepackage{ragged2e}
\usepackage{tikz}
\usetikzlibrary{bayesnet}
\usepackage{multirow}
\usepackage{changepage}
\usepackage{makecell}
\usepackage{ulem}
\usepackage{tcolorbox}
\usepackage{pbox}
\usepackage{colortbl}
\usepackage{array}
\usepackage{multirow}
\usepackage{diagbox}
\usepackage{stackengine}
\usepackage{xfrac}
\usepackage{nicefrac}
\usepackage{cancel}
%\usepackage{hyperref}

 \definecolor{hutblue}{rgb}{0,0.2549,0.6784}
 \definecolor{midnightblue}{rgb}{0.0977,0.0977,0.4375}
 \definecolor{hutsilver}{rgb}{0.4863,0.4784,0.4784}
 \definecolor{lightgray}{rgb}{0.95,0.95,0.95}
 \definecolor{section}{rgb}{0,0.2549,0.6784}
 \definecolor{list1}{rgb}{0,0.2549,0.6784}
 \definecolor{navyblue}{rgb}{0,0,0.55}
 \definecolor{lightblue}{rgb}{0.91,0.91,0.97}
\renewcommand{\emph}[1]{\textcolor{navyblue}{#1}}

\graphicspath{{./images/}}

\parindent=0pt
\parskip=8pt
\tolerance=9000
\abovedisplayshortskip=0pt

\setbeamertemplate{navigation symbols}{}
%\setbeamertemplate{headline}[default]{}
%\setbeamertemplate{headline}[text line]{\insertsection}
\setbeamertemplate{footline}[frame number]

\setbeamercolor{frametitle}{fg=black,bg=white}
\setbeamercolor{title}{bg=lightblue}
\setbeamercolor{section in head/foot}{fg=black,bg=lightblue}
\setbeamerfont{section in head/foot}{size=\scriptsize}
\setbeamerfont{subsection in head/foot}{size=\scriptsize}%,series=\bfseries
\setbeamercolor{subsection in head/foot}{fg=black,bg=lightblue}

\def\o{{\mathbf o}}
\def\t{{\mathbf \theta}}
\def\w{{\mathbf w}}
\def\x{{\mathbf x}}
\def\y{{\mathbf y}}
\def\z{{\mathbf z}}
\def\eff{\mathrm{eff}}

\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\Sd}{Sd}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\Gammad}{Gamma}
\DeclareMathOperator{\Invgamma}{Inv-gamma}
\DeclareMathOperator{\Bin}{Bin}
\DeclareMathOperator{\Negbin}{Neg-bin}
\DeclareMathOperator{\Poisson}{Poisson}
\DeclareMathOperator{\Beta}{Beta}
\DeclareMathOperator{\logit}{logit}
\DeclareMathOperator{\Normal}{Normal}
\DeclareMathOperator{\GP}{GP}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\BF}{BF}
\DeclareMathOperator{\Invchi2}{Inv-\chi^2}
\DeclareMathOperator{\NInvchi2}{N-Inv-\chi^2}
\DeclareMathOperator{\InvWishart}{Inv-Wishart}
\DeclareMathOperator{\tr}{tr}
% \DeclareMathOperator{\Pr}{Pr}
\def\euro{{\footnotesize \EUR\, }}
\DeclareMathOperator{\rep}{\mathrm{rep}}
\DeclareMathOperator*{\argmax}{argmax}

%\renewcommand\labelitemi{\tiny$\bullet$}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{\large \color{black} List of contents}
    \tableofcontents[currentsection]
  \end{frame}
}

\makeatletter
    \newenvironment{withoutheadline}{
        \setbeamertemplate{headline}[default]
        \def\beamer@entrycode{\vspace*{-\headheight}}
    }{}
\makeatother

\newtheorem*{thm}{Theorem}
\newtheorem*{remark}{Remark}

\newcommand{\tabitem}{%
  \usebeamertemplate{itemize item}\hspace*{\labelsep}}

\date{}

\begin{document}

%----------------------------------------------------------------
\begin{frame}
\setcounter{tocdepth}{2}
\frametitle{\large \color{black} List of contents}
\tableofcontents
\end{frame}
%----------------------------------------------------------------

\section{Introduction}

\subsection*{Motivation}
%----------------------------------------------------------------
\begin{frame}[t]
\frametitle{\normalsize \color{black} Motivation}
\setstretch{1.1}

\vspace{-4mm}
\begin{itemize}\setlength\itemsep{2mm}
\item {\color{navyblue} GPs} are flexible {\color{navyblue} non-parametric probabilistic models} for stochastic non-linear functions.

\begin{itemize}\setlength\itemsep{0.5mm}
\item Non-parametric means that the model complexity adapts to the data.
\item In parametric models the model complexity is pre-specified by the functional form.

\only<2>{$N=15$} \only<3>{$N=18$} \only<4>{$N=19$} \only<5>{$N=20$} \only<6>{$N=22$} \visible<7->{$N=23$} \\ 
\only<1-2>{\includegraphics[scale=0.40]{motivation_fig1_1.pdf}}
\only<3>{\includegraphics[scale=0.40, trim= 0mm 0mm 0mm 0mm, clip]{motivation_fig1_2.pdf}}
\only<4>{\includegraphics[scale=0.40, trim= 0mm 0mm 0mm 0mm, clip]{motivation_fig1_3.pdf}}
\only<5>{\includegraphics[scale=0.40, trim= 0mm 0mm 0mm 0mm, clip]{motivation_fig1_4.pdf}}
\only<6>{\includegraphics[scale=0.40, trim= 0mm 0mm 0mm 0mm, clip]{motivation_fig1_5.pdf}}
\visible<7->{\includegraphics[scale=0.40, trim= 0mm 0mm 0mm 0mm, clip]{motivation_fig1_6.pdf}}

\end{itemize}

\item<8-> Many applications:

\begin{center}
\begin{tabular}{m{3cm} m{3cm} m{3cm}}
\visible<8->{Regression} & \visible<8->{Classification} & \\ 
%
\visible<8->{\includegraphics[scale=0.35,trim=90mm 0mm 0mm 8mm, clip]{motivation_fig2.png}} & \visible<8->{\includegraphics[scale=0.22,trim=120mm 0mm 230mm 12mm, clip]{motivation_fig3.pdf}} & \visible<8->{\pbox{5cm}{Density estimation\\[1mm]
Dimensionality reduction\\[1mm]
Solving differential equation\\[1mm]
...}}
\end{tabular}
\end{center}

\end{itemize}
\end{frame}
%----------------------------------------------------------------

%----------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item The key element of a GP is the {\color{navyblue} covariance function} as it encodes the covariance structure of function values.
%\item Covariance function is the main drawback of GPs in a direct implementation.

\item The covariance function generates the {\color{navyblue} covariance matrix} of the collection of function values.

\item Computing the posterior distribution of a GP {\color{navyblue} needs to invert the covariance matrix} $\to$\\ 
\begin{center}
{$\to$ \color{red} $O(n^3)$}
\end{center}

\item This {\color{navyblue} limits their application} to rather small data sets (a few tens of thousands observations).

\item The problem {\color{navyblue} becomes more severe} with full Bayesian inference via sampling methods.

\end{itemize}
\end{frame}
%----------------------------------------------------------------

%----------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{3mm}
\item We focus on a low-rank representation of GPs using a {\color{navyblue} basis function approximation based on Hilbert space} {\color{darkgray} \citep{solin2018hilbert}}.

\item A basis function approximation to GPs has a linear functional structure:\\[2mm]
	\begin{itemize}\setlength\itemsep{1mm}
	\item Inference considerably faster.
	\item Improves sampling efficiency.
	\item Big benefit to be implemented in modular probabilistic programming frameworks. 
	\item Easier to be used as latent functions in non-Gaussian observational models. 
	\end{itemize}
\end{itemize}
\end{frame}
%----------------------------------------------------------------

\subsection*{Contribution}
%----------------------------------------------------------------
\begin{frame}
\frametitle{\normalsize Contribution}
\setstretch{1.1}

{\color{darkgray} \cite{solin2018hilbert} fully developed the mathematical theory behind this {\color{darkgray} low-rank GP}}. \\[4mm]

{\small \textit{\color{darkgray} Our contribution}:}\vspace{-2mm}
\begin{itemize}\setlength\itemsep{3mm}
\item A detailed analysis for its practical implementation in probabilistic programming.

\item The performance of the method depends on some interrelated key factors:\\[2mm]
	\begin{itemize}\setlength\itemsep{1mm}
	\normalsize
	\item {\color{navyblue} Discover the relationships} among these key factors in relation to accuracy and computational performance.
	\item {\color{navyblue} Practical recommendations} based on these relationships. %to improve performance and save computation time
	\item {\color{navyblue} Diagnosis} of the accuracy of the approximation.
	\end{itemize}

\item Implementation of the method in a fully probabilistic programming framework such as Stan.
\end{itemize}
\end{frame}
%----------------------------------------------------------------

\section{Gaussian process as a prior}

\subsection*{Gaussian process prior model}
%-----------------------------------------------------------------
\begin{frame}
\frametitle{\normalsize Gaussian process as a prior distribution for functions}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2.5mm}
\item GP is a {\color{navyblue} stochastic process} which defines the distribution over a collection of random variables, $\left\lbrace  f(\bm{x}): \bm{x} \in \mathcal{X}\right\rbrace$ of a function $f:{\rm I\!R}^D \to {\rm I\!R}$ and some input domain $\mathcal{X} \subset {\rm I\!R}^D$.
%
\item Common notation:
%
\begin{align*}
f \sim \GP(m(\bm{x}), k(\bm{x}, \bm{x}'))
\end{align*}
%
\item The {\color{navyblue}mean} $m: {\rm I\!R}^D \to {\rm I\!R}$ and {\color{navyblue}covariance function}\, $k: {\rm I\!R}^D \times {\rm I\!R}^D \to {\rm I\!R}$ completely characterize the GP prior and control the a priori behavior of\, $f$.

\item Defining property: 
\vspace{-1mm}
%
\begin{align*}
\left\lbrace  f(\bm{x}_1), f(\bm{x}_2), \hdots, f(\bm{x}_N) \right\rbrace &\sim \Normal(\bm{\mu}, K),\\[1mm]
\mu_i &= m(\bm{x}_i),\\
\hspace{27mm} K_{i,j} &=\cov( f(\bm{x}_i),f(\bm{x}_j))=k(\bm{x}_i,\bm{x}_j).
\end{align*}
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item Let \, $\bm{f}= \left\lbrace f(\bm{x}_1), f(\bm{x}_2), \hdots, f(\bm{x}_N) \right\rbrace$

\item Let \, $\bm{f}_1$ and \, $\bm{f}_2$ a partitioning of \, $\bm{f}= \bm{f}_1 \cup \bm{f}_2$:

\item<2-> Multivariate normal is {\color{navyblue} closed under conditioning}:
\end{itemize}

\vspace{-7mm}
\visible<2->{
\begin{align*}
\bm{f}_2 \mid \bm{f}_1 &\sim \Normal(\bm{\mu}, K),\\[1mm]
\bm{\mu} &= K_{2,1} K_{1,1}^{-1}(\bm{f}_1- \bm{\mu}_1) + \bm{\mu}_2,\\
K &= K_{2,2}-K_{2,1} K_{1,1}^{-1} K_{1,2}.
\end{align*}
}

\vspace{-10mm}
\begin{adjustwidth}{-1.6em}{-2em}
\begin{tabular}{ c c c }
\visible<3->{\scriptsize $\bm{f}\sim \Normal(\bm{\mu}, K)$} & \visible<4->{\scriptsize $\bm{f}_2 \mid \bm{f}_1$} & \visible<5->{\scriptsize $p(\bm{f}_2 \mid \bm{f}_1)$} \\
%
\visible<3->{\includegraphics[width=4cm, trim = 7mm 12mm 1mm 0mm, clip]{GP_fig1.pdf}} & \visible<4->{\hspace{-3.5mm} \includegraphics[width=4cm, trim = 8mm 13mm 1mm 0mm, clip]{GP_fig2.pdf}} & \visible<5->{\hspace{-3.5mm} \includegraphics[width=4cm, trim = 9mm 13mm 2mm 0mm, clip]{GP_fig3.pdf}} \\[-2mm]
%
\visible<3->{\small $x$} & \visible<4->{\small $x$} & \visible<5->{\small $x$}
\end{tabular}
\end{adjustwidth}

\vspace{2mm}
\visible<5->{\scriptsize *These images belong to Nicolas Durrande, \textit{Course presentation in Gaussian Process Summer School 2019}, University of Sheffield.}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]\frametitle{\normalsize Latent Gaussian process model}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item GP functions can be combined with noisy observations $\bm{y}$
\item and used as {\color{navyblue} latent functions} for observational models:
%
\vspace{2mm}
\begin{align*}
&{\Large p(y_i \mid f(\bm{x}_i),\phi)}\\[3mm]
&f:{\rm I\!R}^D \to {\rm I\!R}\; \text{is a latent function with a GP prior}
\end{align*}

\end{itemize}
\end{frame}
%---------------------------------------------------------------

\subsection*{Covariance function}
%---------------------------------------------------------------
\begin{frame}
\frametitle{\normalsize Covariance function}
\setstretch{1.1}

\vspace{-2mm}
\begin{adjustwidth}{-2.5em}{-0.5em}
\begin{itemize}\setlength\itemsep{2mm}
\item[] A covariance function\, $k : \mathcal{X} \times \mathcal{X} \to {\rm I\!R}$ maps a pair of inputs $x_1, x_2 \in \mathcal{X}$, from some input space $\mathcal{X} \in {\rm I\!R}^D$, to the real line ${\rm I\!R}$.\\[1mm]

	\begin{itemize}\setlength\itemsep{3mm}
	\item Encodes our {\color{navyblue} prior assumptions} about the variation of the function.

	\item Defines a {\color{navyblue} correlation structure} of function values.

	\item Generates the {\color{navyblue} covariance matrix} $K \in {\rm I\!R}^{N\times N}$, with\, $K_{ij}= \cov(f(\bm{x}_i),f(\bm{x}_j))= k(\bm{x}_i,\bm{x}_j)$.

	\item Covariance functions must be symmetric and Positive (Semi) Definite, such that,
	\begin{align*}
	\text{(Symmetric)}\hspace{5mm} &K = K^\intercal\\
	\text{(PSD)}\hspace{5mm} &\forall \bm{x} \neq 0 : \;\; \bm{x}^\intercal 		K \bm{x} \geq 0
	\end{align*}

	\item A covariance function is said to be {\color{navyblue} stationary} (invariant to translations) if \\[-1mm]% if it is only a function of difference of the input values,
	$$
	k(\bm{x},\bm{x}')= {\color{navyblue} k(\bm{x}-\bm{x}')}= k(\bm{\tau}), \;\; \bm{\tau} \in {\rm I\!R}^D
	$$

	\item A covariance functions is said to be {\color{navyblue} isotropic} (translation and rotation invariant) if\\[-1mm] % if it only depends on the input points through the norm of the difference, 
	$$
	k(\bm{x},\bm{x}') = {\color{navyblue} k(|\bm{x}-\bm{x}'|)} = k(r), \;\; r\in {\rm I\!R}
	$$

	\vspace{1mm}
	\hspace{20mm} \tiny$\bullet$ \scriptsize  {\color{navyblue} L2-norm $(|\bm{x}-\bm{x}'|_{L2})$}, although other types of distances can be considered.
	\end{itemize}
\end{itemize}
\end{adjustwidth}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

The {\color{navyblue} Mat\'ern class} of isotropic covariance functions is given by:

\vspace{-5mm}
\begin{align*}
k_{\nu}(r)=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\!\nu} \! K_{\nu} \left(\frac{\sqrt{2\nu}r}{\ell}\right)
\end{align*}

\vspace{-2mm}
\begin{itemize}\setlength\itemsep{1mm}
\item<1->$\nu > 0$\, is order of the kernel\; $\Rightarrow$\; {\color{navyblue} Samples paths are ($\nu-1$) times differentiable}\\[2mm]
\end{itemize}

\begin{adjustwidth}{-0.5cm}{-1cm}
\begin{tabular}{c c c c}
\small \visible<1->{$\nu=\infty$} & \small \visible<1->{$\nu=5/2$} & \small \visible<1->{$\nu=3/2$} & \small \visible<1->{$\nu=1/2$}\\[-1mm]
%
\scriptsize \visible<1->{Squared exponential} & \scriptsize \visible<1->{Matern52} & \scriptsize \visible<1->{Matern32} & \scriptsize \visible<1->{Matern12}\\[-1mm]
%
\visible<1->{\includegraphics[width=2.8cm, trim = 0mm 0mm 0mm 5mm, clip]{covfun_fig1.pdf}} & \visible<1->{\includegraphics[width=2.8cm, trim = 0mm 0mm 0mm 5mm, clip]{covfun_fig2.pdf}} & \visible<1->{\includegraphics[width=2.8cm, trim = 0mm 0mm 0mm 5mm, clip]{covfun_fig3.pdf}} & \visible<1->{\includegraphics[width=2.8cm, trim = 0mm 0mm 0mm 5mm, clip]{covfun_fig4.pdf}}
\end{tabular}
\end{adjustwidth}

 {\scriptsize *These images belong to Nicolas Durrande, \textit{Course presentation in Gaussian Process Summer School 2019}, University of Sheffield.}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}
\item $\ell > 0$,\: {\color{navyblue} length-scale} controls the {\color{navyblue} smoothness/wigglyness} of the function samples.\\[1mm]

\begin{itemize}\setlength\itemsep{0mm}
\item Using\, $\bm{\ell}=(\ell_1,\dots,\ell_D)\in {\rm I\!R}^D \rightarrow$  smoothness may vary across different input dimensions.
\end{itemize}

\vspace{1mm}
\begin{adjustwidth}{-3em}{-1em}

{ \color{black} \hspace{17mm} \centering $\ell=0.1$ \hspace{30mm} $\ell=0.3$ \hspace{30mm} $\ell=1.0$} \\[-4pt]
{\scriptsize \color{black} Squared exponential} \\
\includegraphics[width=12.5cm, trim = 0mm 3mm 0mm 5mm, clip]{covfun_fig5.pdf}

{\scriptsize \color{black} Mattern32}\\[1mm]
\includegraphics[width=12.5cm, trim = 0mm 0mm 0mm 5mm, clip]{covfun_fig6.pdf}

\end{adjustwidth}

\item $\alpha > 0$,\: {\color{navyblue} magnitude} or marginal variance of the process.
\end{itemize}
\end{frame}
%---------------------------------------------------------------

\subsection*{Computation requirements}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Computation and learning hyperparameters}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item Optimize kernel hyperparameters $\theta$ from data.

\item Ideally, specify priors $p(\theta)$ and compute the joint posterior:
 %
 \begin{align*}
	p(\bm{f},\theta \mid \bm{y}) = \frac{p(\bm{y}\mid \bm{f})\, p(\bm{f}\mid\theta) p(\theta)}{p(\bm{y})}
 \end{align*}

but marginal variance\, $p(\bm{y}) = \int p(\bm{y}\mid \bm{f})\, p(\bm{f}\mid\theta) p(\theta)\mathrm{d}\theta$\, is almost always intractable.

\vspace{5mm}
\item Using {\color{navyblue} Sampling methods} (MCMC), generates a sample from the unnormalized posterior, %as the posterior is proportional to the likelihood and priors:
%
\begin{align*}
p(\bm{f},\theta\mid y) \propto p(y\mid\bm{f})p(\bm{f}\mid\theta)p(\theta)
\end{align*}

\begin{itemize}\setlength\itemsep{2mm}
\item Evaluating\, $p(\bm{f} \mid\theta) \sim \Normal(\bm{\mu}, K(\theta))$\,  in every sampling step is a\, {\color{navyblue} $O(N^3)$ operation}.\\[2mm]

\vspace{-1mm}
\item MCMC {\color{navyblue} is not the fastest} approach, but allows {\color{navyblue} accurate inference}.
\item It is commonly referred as {\color{navyblue}full Bayesian inference}.
\end{itemize}
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item {\color{navyblue} Maximum a posteriori} (MAP) optimizes the posterior distribution over the parameter space.
 %
 \begin{gather*}
	p(\theta \mid \bm{y}) = \frac{\int p(\bm{y}\mid \bm{f})\, p(\bm{f}\mid \theta)\, p(\theta) \operatorname{d}\bm{f}}{p(\bm{y})}  = \frac{p(\bm{y}\mid \theta)\, p(\theta)}{p(\bm{y})} \propto p(\bm{y}\mid \theta)\, p(\theta)\\[3mm]
%
 \hat{\theta} = \argmax_\theta \left( \ln \left( p(\bm{y}\mid \theta)\, p(\theta)\right) \right) = \argmax_\theta \left( \ln p(\bm{y}\mid \theta) + \ln p(\theta) \right)
 \end{gather*}

\item {\color{navyblue} Maximum likelihood} (ML) methods optimize the likelihood function over the parameter space.\\[1mm]
 ML $\Leftrightarrow$ MAP with $p(\theta)\propto 1$.
 %
 \begin{gather*}
 \hat{\theta} = \argmax_\theta \left( \ln p(\bm{y}\mid \theta) \right)
 \end{gather*}

\item e.g. with Gaussian observational model, to optimize $p(\bm{y}\mid \theta)$ wrt. $\theta$ using gradient based methods,
%
\begin{align*}
\nabla_\theta \ln p(\bm{y}\mid \theta)= \nabla_\theta\; -\frac{N}{2}\ln(2\pi)-\frac{1}	{2}\ln|\sigma^2I+K|-\frac{1}{2}\bm{y}^\intercal(\sigma^2I+K)^{-1}\bm{y},
\end{align*}
%
the {\color{navyblue} covariance matrix needs to be inverted} which is an $O(N^3)$ operation.

\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

{\color{navyblue} MCMC}, {\color{navyblue} MAP} and {\color{navyblue} ML} need to invert the covariance matrix with cost $O(N^3)$.\\[0.7cm]

$O(N^3)$:
	\begin{itemize}\setlength\itemsep{2mm}
	\item[] $N < 1000 \to$ Fine
	\item[] $N < 10000 \to$ Slow, but possible
	\item[] $N > 10000 \to$ Prohibitively slow
	\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\textsc{\small Other approximate methods for Bayesian inference:}
\begin{itemize}\setlength\itemsep{2mm}
\item {\color{navyblue} Laplace approximation}
\item {\color{navyblue} Expectation propagation}
\item {\color{navyblue} Variational Bayes}
\end{itemize}
\end{frame}
%---------------------------------------------------------------

\subsection*{Related low-rank GP methods}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Related low-rank approximate GP methods}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item {\color{navyblue} Sparse GPs} are based on low-rank approximations of the covariance matrix with $m \ll n$ inducing points, and computational complexity of $O(nm^2 + m^3)$ \\[1mm]
	%
	\begin{itemize}\setlength\itemsep{2mm}
		\item Inducing points-based sparse approximation
		\item Structured kernel interpolation method
	\end{itemize}

\item {\color{navyblue} Basis function approximations} with $m \ll n$ basis functions and computational complexity of $O(nm+m)$.\\[1mm]
	%
	\begin{itemize}\setlength\itemsep{2mm}
	\item Spectral analysis and series expansions of GPs
	\item Sparse spectrum GPs
	\item Variational Fourier feature approximation for GPs
	\end{itemize}

\item ...

\end{itemize}

\centering
\begin{tcolorbox}[boxrule=0.5pt,colframe=blue!20, colback=blue!5]
A Hilbert space method to approximate GPs is a {\color{navyblue} basis function approximation via Laplace eigenfunctions} for stationary covariance functions.
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

\subsection*{Spectral density function}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Spectral density function}
\setstretch{1.1}

\vspace{-2mm}
Stationary covariance functions can be represented in terms of their {\color{navyblue} spectral density functions}.

\vspace{2mm}
\begin{thm}[Bochner]
Any {\color{navyblue} stationary covariance function}\, $k:{\rm I\!R}^{D} \to {\rm I\!R}$\, and its {\color{navyblue} spectral density}\, $S:{\rm I\!R}^{D} \to {\rm I\!R}$\, are Fourier duals.
\vspace{-1mm}
\begin{align*}
k(\bm{\tau})=& \int S(\bm{\omega}) e^{2\pi i \bm{\omega} \cdot \bm{\tau}} d\bm{\omega}, \hspace{10mm} \text{(Inverse Fourier Transform)} \nonumber \\
%
S(\bm{\omega})=& \int k(\bm{\tau}) e^{-2\pi i \bm{\omega} \cdot \bm{\tau}} d\bm{\tau}. \hspace{10mm} \text{(Fourier Transform)}  \nonumber
\end{align*}

where\, $\bm{\tau}=\bm{x}-\bm{x}' \in {\rm I\!R}^{D}$, and\, $\bm{\omega} \in {\rm I\!R}^{D}$\, is a frequency.
\end{thm}

\vspace{2mm}
\begin{adjustwidth}{-0.2em}{-0.2em}
\begin{itemize}\setlength\itemsep{2mm}
\item Given a {\color{navyblue} kernel} $k(\bm{\tau})$, the {\color{navyblue} frequencies} that it considers can be obtained by solving the FT.

\item Given a {\color{navyblue} spectral density} $S(\bm{\omega})$, its {\color{navyblue} similarity function} can be obtained by solving the IFT.

\item We can custom spectral densities (frequencies) and get new kernels ({\color{navyblue} kernel learning}).
\end{itemize}
\end{adjustwidth}

\end{frame}
%---------------------------------------------------------------

\section[Hilbert space approximate GP model (HSGP)]{Hilbert space approximate Gaussian process model (HSGP)}

\subsection*{HSGP model}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Hilbert space approximate GP model (HSGP)}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}

\item The covariance function is represented as a {\color{navyblue} infinite series expansion} of eigenfunctions $\phi_j(x)$ scaled by the spectral density\, $S_{\theta}(\sqrt{\lambda_j})$ {\color{darkgray} \citep{solin2018hilbert}}:
%
\begin{align*}
k(x,x') = \sum_{j=1}^\infty S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'), \hspace{10mm} x,x' \in [-L,L]
\end{align*}

\item We choose the {\color{navyblue} eigenvalues} $\lambda_j$ and {\color{navyblue} eigenfunctions} $\phi_j(x)$ of the Laplacian operator $-\nabla^2$:
%
\begin{align*}
-\nabla^2 \phi_j(x)&=\lambda_j \phi_j(x), \hspace{0.8cm}  x \in [-L,L] \\ 
\hspace{6mm}\phi_j(x)&=0, \hspace{1.6cm} x \notin [-L,L]
\end{align*}

given by:\\[-5mm]
%
\begin{gather*}
\lambda_j=\left(\frac{j\pi}{2L}\right)^2 \hspace{10mm}
\phi_j(x)=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right)
\end{gather*}

\item $L > 0$ represents the {\color{navyblue} boundaries} where this representation is valid.
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\vspace{-1mm}
\begin{itemize}\setlength\itemsep{1mm}
\item<1-> To make this representation useful we can truncated the sum to the first $m$ terms:
%
\vspace{-1mm}
\begin{align*}
k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'), \hspace{10mm} x,x' \in [-L,L]
\end{align*}
\end{itemize}

\vspace{-10mm}
\begin{adjustwidth}{-2em}{-1em}
\begin{tabular}{c c}
 \only<2-6>{$\phi_j(x)$} & \visible<5->{$S_{\theta}(\sqrt{\lambda_j})$} \\
 %
 \only<2>{\hspace{2mm}\includegraphics[width=6cm, trim = 0mm 0mm 0mm 0mm, clip]{HSGP_fig1_1.pdf}}
\only<3>{\hspace{1mm}\includegraphics[width=6cm, trim = 0mm 0mm 0mm 0mm, clip]{HSGP_fig1_2.pdf}}
\only<4-6>{\includegraphics[width=6cm, trim = 0mm 0mm 0mm 0mm, clip]{HSGP_fig1_3.pdf}} & \visible<5->{\includegraphics[width=6.1cm, trim = 0mm 0mm 0mm 0mm, clip]{HSGP_fig2.pdf}}\\[-2mm]
\end{tabular}
\end{adjustwidth}

\begin{itemize}\setlength\itemsep{1mm}
\item<6-> $S_{\theta}(\sqrt{\lambda_j})$ decays rapidly to zero.
\item<6-> This rate of dacay depends on the lenghtscale of $k$ or smoothness of the function to be learned.
\item<6-> A finite number $m$ of terms for an accurate approximation.
\item<6-> As long as $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}

\item This representation\, $k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x')$\, leads to the\\ {\color{navyblue} linear representation} of the function\, $f$:
%
\begin{align*}
f(x) \approx \sum_{j}^m \phi_j(x) \beta_j, \hspace{10mm} \beta_j \sim \Normal(0,S_{\theta}(\sqrt{\lambda_j})).
\end{align*}
\end{itemize}

%\vspace{10mm}
%\begin{itemize}\setlength\itemsep{1mm}
%\small
%\item Several common GP approximations lead to approximations of similar form (i.e. sparse GPs etc).
%\item All of them require a Cholesky factorization or eigendecomposition of the kernel matrix.
%\item However, Hilbert space methods give analytical expressions for both eigenfunctions and variance of the weights $\beta_j$.
%\end{itemize}

\begin{tcolorbox}[colframe=blue!20, colback=white, title=\small Properties of the method, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
\setstretch{1.1}
\begin{itemize}\setlength\itemsep{1.5mm}
\item[+] The basis functions $\phi_j(x)$ do not depend on the hyperparameters, so they just need to be computed once with cost $O(nm)$.

\item[+]  All dependencies on the hyperparameters is through the variance of the weights $\beta_j$.

%\item[+]  In every step of the sampler the computational complexity is that of a linear model $O(nm)$.

%\item[+] $f(x)$ is naturally in the {\color{navyblue} non-centered parameterization form} with independent prior distribution on $\beta_j$ which makes posterior inference easier. %\citep[see, e.g., ][]{Betancourt+Girolami:2019}

\item[+]  The parameter posterior\, $p(\bm{\beta}|\bm{y})$ is\, $m$-dimensional ($m\ll n$), which makes posterior inference easier.

\item[+] Trade-off between accuracy and computation.

\item[-] Scales badly with the number of input dimensions $D$.

\item[-] In practice,\, $D>3$\, starts to be too computationally demanding, even for smooth functions.
\end{itemize}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

\subsection*{Aim of the study}
%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{tcolorbox}[colframe=blue!20, colback=white, title=\small Aim of the study, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
\setstretch{1.1}

Determine the minimum $m$ ({\color{navyblue} basis functions}) and valid $L$ ({\color{navyblue} boundary}) to obtain accurate approximations, as a function of the {\color{navyblue} lengthscale}.

\begin{enumerate}
\item How many basis functions $m$ are needed for a good approximation?

\item Which is the effect of the boundary box $L$ on the approximation?

\item Which are valid values for the boundary box $L$?

\item How $m$ and $L$ relate to each other and to the wigglyness of the function to be learned?
\end{enumerate}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

\subsection*{Generalization to multidimensional input space}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Generalization to multidimensional input space}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item Regular and compact domain\, $\Omega=[-L_1,L_1] \times \dots \times [-L_D,L_D] \subset {\rm I\!R}^D$

\item The total number of eigenfunctions and eigenvalues is\, $m^{\ast} = \prod_{d=1}^{D} m_d$ %that is possible combinations of univariate eigenfunctions over all dimensions

\vspace{4mm}
\begin{columns}
\column{0.3\textwidth}
\centering
 {\small e.g. $D=2$:} \\[1mm]
 \includegraphics[width=4cm, trim = 0mm 0mm 0mm 20mm, clip]{HSGP_fig3.pdf}

\column{0.5\textwidth}
\begin{align*}\small
\mathbb{S}^\intercal=
\left[ {\begin{array}{cccccccccccc}
1 & 1 & 1 & 2 & 2 & 2 & 3 & 3 & 3 & 4 & 4 & 4  \\
1 & 2 & 3 & 1 & 2 & 3 & 1 & 2 & 3 & 1 & 2 & 3  
\end{array} } \right] \in {\rm I\!R}^{2\times 12}
\end{align*} 
\end{columns}

\vspace{-2mm}
\item Thus, for\, $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$\, and\, $j=1,\ldots,m^{\ast}$, we have:
%
\begin{align*}
\bm{\lambda}^{\ast}_j &= \left\{ \lambda_{\mathbb{S}_{jd}} \right\}_{d=1}^D =  \left\{ \left(\tfrac{\pi \mathbb{S}_{jd}}{2L_d}\right)^2 \right\}_{d=1}^D \\[1mm]
%
\phi^{\ast}_j(\bm{x}) &= \prod_{d=1}^{D} \phi_{\mathbb{S}_{jd}}(x_d) = \prod_{d=1}^{D} \sqrt{\frac{1}{L_d}} \text{sin}\left(\sqrt{\lambda_{\mathbb{S}_{jd}}}(x_d+L_d)\right)
\end{align*}

\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{4mm}
\item And the approximate covariance function\, $k(\bm{x},\bm{x}',\theta)$\, and\, $f(\bm{x})$:
%
\begin{align*}
k(\bm{x},\bm{x}') &\approx \sum_{j=1}^{m^{\ast}} 
S^{\ast}_{\theta}\left(\sqrt{\bm{\lambda}^{\ast}_j}\right)
\phi^{\ast}_j(\bm{x}) \phi^{\ast}_j(\bm{x}') \\
%
f(\bm{x}) &\approx \sum_{j=1}^{m^{\ast}} 
\left( S^{\ast}_{\theta} \left(\sqrt{\bm{\lambda}^{\ast}_j} \right)\right)^{1/2}
\phi^{\ast}_j(\bm{x}) \beta_j
\end{align*}

where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function.

\vspace{2mm}
\item The number of multivariate basis functions,\, $m^{\ast} = \prod_{d=1}^{D} m_d$,\, {\color{navyblue} grows exponentially} with respect to $D$.\\[-2mm]
 \[O(nm^*+m^*)\]

\item In practice,\, $D>3$\, starts to be too computationally demanding, even for smooth functions.

\end{itemize}
\end{frame}
%---------------------------------------------------------------

\section{Performance analysis of HSGP}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Performance analysis of HSGP}
\setstretch{1.1}

The {\color{navyblue} performance}, {\color{navyblue} accuracy} and {\color{navyblue} speed} of the method mainly depends on:
%
\begin{center}
\begin{tabular}{ l l l}
 \pbox{5cm}{Number of basis functions ($m$) \\ Boundary box ($L$)} & $\Leftrightarrow$ & Smoothness ($\ell$)
\end{tabular}
\end{center}

\begin{tcolorbox}[colframe=blue!20, colback=white, title=Objectives, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
\setstretch{1.1}
\begin{enumerate}\setlength\itemsep{2mm}
\item Analysis of the effects of\, $m$\, and\, $L$ on the approximation.
\item A model gathering the relationships among\, $m$, $L$\, and\, $\ell$ in terms of an accurate approximation \\

 \centering $\Downarrow$
 \\[2mm]
 \begin{center}
 \begin{tabular}{@{}l@{}}
 \tabitem make useful and valuable {\color{navyblue} recomendations} \\
 \tabitem make {\color{navyblue} diagnosis} of the performance \\
 \tabitem save time of computation 
 \end{tabular}
 \end{center}

\end{enumerate}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

\subsection*{Effects of $m$ and $L$}
%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{enumerate}
\item<1-> \textbf{Effects of}\, $m$\, \textbf{and}\, $L$\, \textbf{on the approximation}

\vspace{3mm}
 \pbox{6cm}{Boundary box\, $\Omega \in [-L,L]$ \\ Desired input domain\, $\Phi \in [-S,S]$}\;\;  $\to$\;\;  $L=c \cdot S$, \hspace{3mm} $c \geqslant 1$, \hspace{1mm} $c \equiv$ {\color{navyblue} Boundary factor}

\vspace{3mm}
\begin{itemize}\setlength\itemsep{1mm}
\item<2-> \normalsize Effect of\, $m$\; ($c$ is fixed)

\begin{adjustwidth}{-4.7em}{-3em}
\centering
{\scriptsize \hspace{5mm} Covariance \hspace{30mm} \visible<3->{Mean function} \hspace{27mm} \visible<4->{Standard deviation}}\\[-4mm]
%
\includegraphics[scale=0.23, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig1_Cov_J.pdf}
\visible<3->{\includegraphics[scale=0.23, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
\visible<4->{\includegraphics[scale=0.23, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
\end{adjustwidth}

\normalsize 
\item<3-> $m$\, affects the smoothness of the posterior function:\; The lower\, $m$,\, the smoother the posterior.
\item<3-> The higher the wigglyness of the function to be learned, the higher\, $m$\, is required.
\item<4-> $m$\, affects the uncertainty as well (obviously).
\item<4-> The approximation is more sensitive in the uncertainty than in mean.
\end{itemize}
\end{enumerate}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item \normalsize Effect of\, $c$\; ($m$ is fixed)\\[2mm]

\begin{adjustwidth}{-2.7em}{-3em}
\begin{center}
{\scriptsize \hspace{5mm} Covariance \hspace{30mm} \visible<1->{Mean function} \hspace{27mm} \visible<2->{Standard deviation}}\\[-4mm]
%
\includegraphics[scale=0.23, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig2_Cov_L.pdf}
\visible<1->{\includegraphics[scale=0.23, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
\visible<2->{\includegraphics[scale=0.23, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
\end{center}
\end{adjustwidth}

\vspace{2mm}
\begin{itemize}\setlength\itemsep{2mm}
\normalsize
\item<1-> $c$\, affects the mean mainly near the boundaries.
\item<2-> $c$\, affects the uncertainty along the whole domain and the error is greater where there is less data. 
\end{itemize}
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item \normalsize Effect of\, $m$\, and\, $c$\\[3mm]
\begin{adjustwidth}{-4.3em}{-3em}
%
\centering
\begin{tabular}{ c c c }
\arrayrulecolor{darkgray}\hline
c = 1.05 &
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf}
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part1.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & 
\multirow{10}{1cm}{ \includegraphics[scale=0.25, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\ 
\arrayrulecolor{gray!50}\hline
c = 1.1 &
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part2.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
\hline
c = 1.2 &
\includegraphics[scale=0.17, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} 
\includegraphics[scale=0.17, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Sigma_part3.pdf} 
\includegraphics[scale=0.17, trim = 0mm 4mm 5mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
\hline
\end{tabular}
\end{adjustwidth}
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item \normalsize Effect of\, $m$\, and\, $c$\\[3mm]
\begin{adjustwidth}{-4.3em}{-3em}
%
\centering
\begin{tabular}{ c c c }
\arrayrulecolor{gray!50}\hline
c = 1.5 &
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part4.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & 
\multirow{10}{1cm}{ \includegraphics[scale=0.25, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\
\hline
c = 2 &
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part5.pdf} 
\includegraphics[scale=0.17, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
\hline
c = 2.5 &
\includegraphics[scale=0.17, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf}
\includegraphics[scale=0.17, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Sigma_part6.pdf}
\includegraphics[scale=0.17, trim = 0mm 4mm 5mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
\arrayrulecolor{darkgray}\hline
\end{tabular}
\end{adjustwidth}
\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\vspace{-3mm}
\begin{adjustwidth}{-0.5em}{-2em}
\only<1>{\includegraphics[scale=0.34, trim = 0mm 8mm 0mm 0mm, clip]{ch5_fig4_MSE_vs_J.pdf} }
\only<2>{\includegraphics[scale=0.34, trim = 0mm 8mm 0mm 0mm, clip]{performance_fig1.pdf} }
\only<3>{\includegraphics[scale=0.34, trim = 0mm 8mm 0mm 0mm, clip]{performance_fig2.pdf} }
\only<4-5>{\includegraphics[scale=0.34, trim = 0mm 8mm 0mm 0mm, clip]{performance_fig3.pdf} }
%
\visible<5->{\includegraphics[scale=0.35, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig5_lscale_vs_J.pdf}}
\end{adjustwidth}

\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small Main conclusions}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, right=1mm, left=1mm, width=0.99\textwidth]
\setstretch{1.1}
\begin{itemize}
\item \; As\, $c$\, increases,\, $m$ has to increase as well (and vice versa). 
$$
\small
\text{As}\;\, L=c\cdot S,\hspace{3mm} \text{if}\; c\uparrow,\; \text{then}\; \lambda_j=\left(\frac{j\pi}{2L}\right)^{\!2}\downarrow,\, \text{and therefore}\;\, S_{\theta}(\sqrt{\lambda_j})\downarrow \; \text{and}\; m\; \downarrow
$$ 
\item \; There exists a minimum valid\, $c$,\, regardless\, $m$. % below which a accurate approximation will never be achieved.
\end{itemize}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

\subsection*{Model of relationships among $m$, $L$ and $\ell$}
%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{enumerate}\setcounter{enumi}{1}
\item \textbf{Model of relationships among}\, $m$,\, $c$\, \textbf{and}\, $\ell$ \\[2mm]

 \begin{itemize}
 \item make useful and valuable {\color{navyblue} recomendations} 
 \item make {\color{navyblue} diagnosis} of the performance 
 \item save time of computation 
 \end{itemize}
\end{enumerate}

\vspace{3mm}
\begin{columns}
\column{0.2\textwidth}

\column{0.4\textwidth}
\hspace{5mm} \centering \scriptsize Squared exponential kernel\\
\includegraphics[scale=0.25, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}

\column{0.6\textwidth}
\begin{itemize}
\scriptsize
\item[] $m$\, and $c$\, govern the approximation.\\
\item[] $\ell$\, controls the wigglyness of the function.
\end{itemize}

\begin{columns}
\column{5mm}
 \includegraphics[scale=0.25, trim = 80mm 110mm 210mm 40mm, clip]{performance_fig4.pdf}
\column{40mm}
 \pbox{4cm}{\scriptsize in terms of an accurate approximation}
\column{5mm}
\end{columns}
\end{columns}

\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize Criteria of an accurate approximation as a function of $m$}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.8\textwidth]
\setstretch{1.1}
\scriptsize
{\small Relative difference among approximate and exact covariance:}
%
\begin{align*}
 \int \frac{ | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|}{k(\tau)} \,\mathrm{d}\tau < 0.01
\end{align*}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\vspace{3mm}

\begin{columns}
\column{0.4\textwidth}
\hspace{7mm} \centering \scriptsize Squared exponential\\
\includegraphics[scale=0.29, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}

\column{0.5\textwidth}
\hspace{7mm} \centering \scriptsize Mattern($\nu$=3/2) kernel\\
\includegraphics[scale=0.29, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}
\end{columns}

\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small This model says...}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
\setstretch{1.1}
\begin{itemize}\setlength\itemsep{1mm}
\item For a given\, $m$\, and\, $c$,\, the minimum\, $\ell$\, that can be accurately approximated.
\item For a given\, $\ell$\, and\, $c$,\, the minimum\, $m$\, required for an accurate approximation.
\item For a given\, $\ell$\, and\, $m$,\, the highest\, $c$\, allowed for an accurate approximation.
\item The minimum valid\, $c$\, for a given\, $\ell$.
\end{itemize}
\end{tcolorbox}

\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small Insights from this model of relationships}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]

\begin{itemize}\setlength\itemsep{1mm}
\item Allows to  choose the combination ($m,c$) that accurately approximate\, $\ell$\, and save computation time.\\[-5mm]
$$
O(nm+m)
$$

\item Serves as a {\color{navyblue} diagnosis tool} of the approximation:

\begin{remark}
\setstretch{1.1}
As the lengthscale is the parameter that ultimately characterizes the non-linearity of the posterior functions, we assume we can look at the {\color{navyblue} estimate lengthscale} to diagnose the approximation.
\end{remark}

\vspace{2mm}
\item Given $m$ and $c$

\item If\, $\hat{\ell}$\, {\color{navyblue} is below} the minimum\, $\ell$\, that can be accurately approximated (for those given\, $m$\, and\, $c$), then the approximation {\color{navyblue} is not sufficiently accurate}, and\, $m$\, must be increased or\, $c$\, decreased.

\item If\, $\hat{\ell}$\, {\color{navyblue} is above} the minimum\, $\ell$, then the approximation should be {\color{navyblue} sufficiently accurate}.

\item This can be done iteratively.

%\item $c$\, can not be decreased as much as desired because it is restricted to\, $\ell$.
\end{itemize}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\textsc{\small Example of diagnosis}\\

\vspace{-4mm}
\centering
\includegraphics[scale=0.25, trim = 0mm 8mm 0mm 0mm, clip]{ch5_fig4_MSE_vs_J.pdf}

\vspace{2mm}
\begin{columns}
\column{0.5\textwidth}
\begin{itemize}[<+->]\setlength\itemsep{4mm}
\small
\item[] $c=1.05,1.1,1.2$\; $\nexists$
\item[] \pbox{1cm}{$c=2.5$\\ $m=10$} \; $\to$ \; $\hat{\ell}=0.255$
\item[] \pbox{1cm}{$c=2$\\ $m=10$} \; $\to$ \; $\hat{\ell}=0.267$ 
\item[] \pbox{1cm}{$c=2$\\ $m=15$} \; $\to$ \; $\hat{\ell}=0.31$
\item[] \pbox{1cm}{$c=1.5$\\ $m=10$} \; $\to$ \; $\hat{\ell}=0.29$
%\item[] \pbox{1cm}{$c=1.5$\\ $m=15$} \; $\to$ \; $\hat{\ell}=0.31$
\end{itemize}

\column{0.5\textwidth}
\includegraphics[scale=0.28, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}
\end{columns}
\end{frame}
%---------------------------------------------------------------

\subsection*{Future research}
%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\textsc{\small Future research on this model of relationships}\\[5mm]

\begin{itemize}\setlength\itemsep{3mm}
\item Investigate and obtain this model of relationships for $D$-dimensional input spaces.
\item Analytical models  would be useful to automatize diagnosis.
\end{itemize}
\end{frame}
%---------------------------------------------------------------

\section{Case studies}

\subsection*{Univariate synthetic case study}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Univariate synthetic case study}
\setstretch{1.1}
\scriptsize

\vspace{0mm}
\begin{columns}
\scriptsize
\column{0.05\textwidth}

\column{0.4\textwidth}
$x \in [-1,1] \subset {\rm I\!R}$\\[3mm]
GP Mattern($\nu$=3/2) and\, $\ell=0.15$ and\, $\alpha=1$ + $\bm{\epsilon} \sim \Normal(0, 0.2^2)$:\\[2mm]
\pbox{7cm}{
155 obs. for training\\
45 obs. for testing interpolation\\
50 obs. for testing extrapolation}\\[2mm]

\column{0.6\textwidth}
\begin{itemize}\setlength\itemsep{2mm}
\item<2->
Observation model:\;\; \pbox{6cm}{$\bm{y} = \bm{f} + \bm{\epsilon}$ \\
$\bm{\epsilon} \sim \Normal(\bm{0}, \sigma^2 \bm{I})$}

\item<3-> Exact GP prior for\, $\bm{f}$:\;\;
\pbox{3cm}{$f(x) \sim \GP(0, k_{\nicefrac{3}{2}}(x, x'))$}

\item<4-> HSGP prior for\, $\bm{f}$:\;\;
\pbox{6cm}{$f(x) \approx \sum_{j=1}^m \left( S_{\nicefrac{3}{2}}(\sqrt{\lambda_j})\right)^{1/2} \phi_j(x) \beta_j$ \\
$\beta_j \sim \Normal(0,1)$\\
$m=60$\\
$c=1.5$}

\item<5-> Thin plate regression spline for\, $\bm{f}$ using the R-package \textit{mgcv}.
\end{itemize}
\end{columns}

\vspace{-4mm}
\centering
\includegraphics[scale=0.22, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig10_Posteriors_exI.pdf}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{center}
\begin{tabular}{c c c }
\hspace{5mm} \scriptsize Interpolation error & \hspace{4mm} \scriptsize Extrapolation error & \\[-3mm]
%
\includegraphics[scale=0.24, trim = 0mm 0mm 10mm 10mm, clip]{ch5_fig11_MSE_exI_inter.pdf} & \hspace{-4mm} \includegraphics[scale=0.24, trim = 0mm 0mm 10mm 10mm, clip]{ch5_fig11_MSE_exI_extra.pdf} & \hspace{-4mm}
\raisebox{10mm}{\includegraphics[scale=0.33, trim = 29mm 68mm 110mm 30mm, clip]{ch5_fig11_MSE_exI_legend.pdf}}
\end{tabular}
\end{center}

\begin{columns}
\column{0.28\textwidth}
 \centering \hspace{12mm}  \scriptsize Computation time\\[-3mm]
\centering \includegraphics[scale=0.24, trim = 0mm 0mm 10mm 5mm, clip]{ch5_fig11_time_exI.pdf}

\column{0.7\textwidth}
 \begin{itemize}\setlength\itemsep{1.5mm}
 \small
\item The HSGP is roughly 400 times faster than the exact GP
\item The HSGP is 10 times faster than the spline model 
\item In 1D the computation time of HSGP increases slowly as a function of $m$.
 \end{itemize}

\end{columns}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\textsc{\small Diagnosis}

\begin{columns}
\column{0.4\textwidth}

\vspace{3mm}
\scriptsize Interpolation error\\[-4mm]
\hspace{50mm} \only<1-3>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig1.pdf}}
\only<4>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig2.pdf}}
\only<5-7>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig1.pdf}}
\only<8>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig3.pdf}}
\only<9-11>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig1.pdf}}
\only<12>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig4.pdf}}
\only<13-15>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig1.pdf}}
\only<16>{\includegraphics[scale=0.20, trim = 0mm 0mm 10mm 10mm, clip]{caseI_fig5.pdf}}
\\[4mm]

\begin{enumerate}\setlength\itemsep{3mm}
\small
\item \pbox{1cm}{$m=20$\\ $c=1.5$} \: $\to$\, $\hat{\ell}=0.09$\\[4mm]
\item<5-> \pbox{1cm}{$m=30$\\ $c=1.5$} \: $\to$\, $\hat{\ell}=0.12$\\[4mm]
\item<9-> \pbox{1cm}{$m=40$\\ $c=1.5$} \: $\to$\, $\hat{\ell}=0.14$\\[4mm]
\item<13-> \pbox{1cm}{$m=60$\\ $c=1.5$} \: $\to$\, $\hat{\ell}=0.14$\\[4mm]
\end{enumerate}

\column{0.6\textwidth}
\hspace{9mm} \scriptsize Mattern($\nu$=3/2) kernel\\
\only<1>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
%
\only<2>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_1.pdf}}
%
\only<3-4>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_1_1.pdf}}
%
\only<5>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
%
\only<6>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_2.pdf}}
%
\only<7-8>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_2_1.pdf}}
%
\only<9>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
%
\only<10>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_3.pdf}}
%
\only<11-12>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_3_1.pdf}}
%
\only<13>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
%
\only<14>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_4.pdf}}
%
\only<15-16>{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 10mm, clip]{caseI_fig6_4_1.pdf}}

\end{columns}
\end{frame}
%---------------------------------------------------------------

\subsection*{Bivariate synthetic case study}
%---------------------------------------------------------------
\begin{frame}\frametitle{Bivariate synthetic case study}
\setstretch{1.1}

\vspace{-1mm}
\begin{itemize}\setlength\itemsep{1mm}
\item[]
\begin{itemize}\setlength\itemsep{1mm}
\item[] $\bm{x} \in [-1,1] \times [-1,1] \subset {\rm I\!R}^2$
\item[] 120 obs. from a GP with SE kernel and $\ell_1=0.1$ and $\ell_2=0.35$.
\end{itemize}
\end{itemize}

\vspace{-3mm}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.27, trim = 0mm 33mm 35mm 20mm, clip]{ch5_fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.27, trim = 19mm 33mm 35mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.27, trim = 150mm 18mm 0mm 0mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}}\\ 
%
\includegraphics[scale=0.27, trim = 0mm 18mm 35mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.27, trim = 19mm 18mm 35mm 20mm, clip]{ch5_fig15_spfun_exII.pdf}
\end{tabular}
\end{center}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\vspace{-3mm}
\begin{center}
\begin{tabular}{ c c c}
\small Error & \small Computation time & \\[-5mm]
%
\includegraphics[scale=0.29, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf} & \hspace{-3mm}
\includegraphics[scale=0.29, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf} & \hspace{-4.5mm}
\raisebox{8mm}{\includegraphics[scale=0.38, trim = 28mm 53mm 112mm 20mm, clip]{ch5_fig17_legend_exII.pdf}}
\end{tabular}
\end{center}

\vspace{-4mm}
\begin{itemize}\setlength\itemsep{3mm}
\small
\item Choosing the optimal $c$ allows for lower (univariate) $m$ and then less computation time.

\item In 2D computation increases significantly with $m$ or $knots$,

\item however HSGP and spline still work significantly better than regular GP, even for highly wiggly functions.

%\item<5-> Serious difficulties with computation time were encountered in building the spline model with $50$ knots for this bivariate function.
\end{itemize}
\end{frame}
%---------------------------------------------------------------

\subsection*{Diabetes disease case study}
%---------------------------------------------------------------
\begin{frame}[t]\frametitle{\normalsize Epidemiological study of diabetes disease}
\setstretch{1.1}

\vspace{-4mm}

\begin{itemize}\setlength\itemsep{1mm}
\small
\item[]  $392$ observations $\to$ $y_i=\{1,0\}$, suffering or not suffering from diabetes.
\item[] $\bm{x}_i \in {\rm I\!R}^4$, \hspace{1mm} \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$).
\end{itemize}

\begin{center}
\begin{tabular}{c c c}
\small
\pbox{4cm}{
$y_i \sim \mathrm{Bernoulli}(p_i)$ \\[2mm]
$p_i = \mathrm{logit}(f(\bm{x}_i))$}
 & $\to$ & \pbox{10cm}{\scriptsize Exact GP for \, $\bm{f}$:\;
$f(x) \sim \GP(0, k(x, x'))$ \\[2mm]
%
\scriptsize HSGP for \, $\bm{f}$:\;
$f(\bm{x}) \approx \sum_{j}^m \left( S(\sqrt{\bm{\lambda}_j})\right)^{1/2} \phi_j(\bm{x}) \beta_j$\\[2mm]
\scriptsize Thin plate spline
}
\end{tabular}
\end{center}

\vspace{5mm}
{\small $\bm{x} \in {\rm I\!R}^2$}\\[-20mm]
%
\begin{center}
\includegraphics[scale=0.27, trim = 5mm 23mm 39mm 20mm, clip]{ch5_fig18_gpfun_diabetes.pdf}
\includegraphics[scale=0.27, trim = 20mm 23mm 39mm 20mm, clip]{ch5_fig18_bffun_diabetes.pdf}
\includegraphics[scale=0.27, trim = 20mm 23mm 39mm 20mm, clip]{ch5_fig18_spfun_diabetes.pdf}
\includegraphics[scale=0.22, trim = 150mm 4mm 2mm 15mm, clip]{ch5_fig18_legend_diabetes.pdf}
\end{center}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\begin{center}
\centering  \small Computation time \\[-1mm]
%
\includegraphics[scale=0.20, trim = 1mm 0mm 10mm 0mm, clip]{ch5_fig20_time2D_diabetes_2.pdf}
\includegraphics[scale=0.20, trim = 30.2mm 0mm 10mm 0mm, clip]{ch5_fig20_time3D_diabetes_2.pdf}
\includegraphics[scale=0.20, trim = 30.2mm 0mm 10mm 0mm, clip]{ch5_fig20_time4D_diabetes_2.pdf}
\hspace{-1mm} \includegraphics[scale=0.28, trim = 28mm 60mm 112mm 10mm, clip]{ch5_fig20_legend_diabetes_2.pdf}
\end{center}

\vspace{-4mm}
\begin{itemize}\setlength\itemsep{1mm}
\small
\item Computation time increases significantly with higher $D$. 
\item Choosing optimal values for $c$ can be essential to avoid a excessive computation time. 
\end{itemize}

\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small On time of computation}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, right=1mm, left=1mm, width=0.99\textwidth]
\setstretch{1.1}
\begin{itemize}\setlength\itemsep{2mm}
\small
\item HSGP $\left(O(nm^D)\right)$ is computationally faster than exact GP $\left(O(n^3)\right)$ for larger datasets ($n$):

{\centering $O(nm^D) \ll O(n^3)$, as long as $D$ is not too large.} % (approx. $D\leq 4$).

\item HSGP is significantly faster than exact GP for $D=2$, even for highly wiggly functions.

\item If $n$ is low (i.e. $n\approx 400$), HSGP is slower than exact GP for $D>3$, even for smooth functions, or even for $D=3$ and moderate wiggly functions.

\item However, if $n$ increases, HSGP does not significantly increase as much as exact GP does.

%\item Choosing the optimal $c$\, reduces the number of required basis functions.
\end{itemize}
\end{tcolorbox}
\end{frame}
%---------------------------------------------------------------

\subsection*{Spatio-temporal land-use classification}
%---------------------------------------------------------------
\begin{frame}[t]\frametitle{Spatio-temporal land-use classification}
\setstretch{1.1}

\begin{columns}
\column{0.6\textwidth}
\begin{itemize}\setlength\itemsep{1mm}
\small
\item[] Plots dedicated to growing citrus fruits.
\item[] $n=200$ plots with known class.

\item[] $T=5$ years (2006, 2008, 2010, 2012, 2015). \\[5mm]

\item<1->[] $K=5$ classes:\\[2mm]
 \hspace{2mm} \pbox{6cm}{$k=1$, \; adult independent\\ $k=2$, \; aligned\\ $k=3$, \; irregular\\ $k=4$, \; abandoned\\ $k=5$, \; young}

\vspace{5mm}
\item<1->[] $\bm{x}\in {\rm I\!R}^{53}$  characteristic variables:\\[2mm] 
\hspace{2mm} \pbox{6cm}{ Spectral intensities \\ Empirical semivariogram of the pixels within a plot \\ Shape-descriptive statistics of the plots \\ Time}

%\item<3->[] $\mathrm{PCA}(\bm{x})$
\end{itemize}

\column{0.5\textwidth}
\includegraphics[scale=0.15, trim = 1mm 0mm 10mm 0mm, clip]{parcelas.pdf}
\end{columns}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\textsc{\small Model}:

\vspace{-8mm}
\begin{align*}
\visible<1->{y_{it} &\sim \mathrm{multinomial}(\bm{p}_{it}), \hspace{5mm} \bm{p}_{it}=\{p_k\}_{k=1}^5}  \\[2mm]
%
\visible<1->{\bm{p}_{it} &= \mathrm{softmax}(f(\bm{x}_{it})), \hspace{5mm} f(\bm{x}_{it})=\{f_k(\bm{x}_{it})\}_{k=1}^5} \\[2mm]
%
\visible<1->{f_k(\bm{x}) &= \sum_{d=1}^{D} g_d(x^d) + \sum_{d=1}^{D-1} h_{d,D}(x^d,x^D)} \\[2mm]
%
\visible<1->{g_d(x^d) &\sim \mathrm{HSGP}(x^d, S, \theta_{d,1})\\[2mm]
h_{d,D}(x^d,x^D) &\sim \mathrm{HSGP}(x^d,x^D, S, \theta_{d,D})}
\end{align*}

\textsc{\small Diagnosis}:\\[-7mm]

\centering
{\includegraphics[scale=0.22, trim = 0mm 0mm 5mm 10mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\textsc{10-fold cross-validation confusion matrix}

\begin{center}
\arrayrulecolor{gray!50}
\begin{tabular}{|c|*{6}{c|}}
\hline
\diagbox[linecolor=gray!50,width=2cm]{\small Estimate}{\small True} & \multicolumn{1}{p{1cm}|}{\centering \scriptsize independent} &  \multicolumn{1}{p{1cm}|}{\centering \scriptsize aligned} & \multicolumn{1}{p{1cm}|}{\centering \scriptsize irregular} & \multicolumn{1}{p{1cm}|}{\centering \scriptsize abandoned} & \multicolumn{1}{p{1cm}|}{\centering \scriptsize young} & \multicolumn{1}{p{1cm}|}{\centering } \\ 
\hline \multicolumn{1}{|p{1.5cm}|}{ \centering \scriptsize independent} &90&39&14&3&11&42\%\\
\hline \multicolumn{1}{|p{1.5cm}|}{\centering \scriptsize aligned} &46&301&8&2&3&16\%\\
\hline \multicolumn{1}{|p{1.5cm}|}{\centering \scriptsize irregular}  &8&4&59&4&1&19\%\\
\hline \multicolumn{1}{|p{1.5cm}|}{\centering \scriptsize abandoned}  &5&2&6&342&5&5\%\\
\hline \multicolumn{1}{|p{1.5cm}|}{\centering \scriptsize young}  &8&2&1&0&38&19\%\\
\hline \multicolumn{1}{|p{1.5cm}|}{\centering }  &42\%&13\%&32\%&2\%&34\%&$\textbf{17\%}$\\
\hline
\end{tabular}
\end{center}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}[t]
\setstretch{1.1}

\textsc{Further modeling considerations}

\begin{itemize}\setlength\itemsep{2mm}
\item SE kernel for the time dimension might be too simple for this case study

\item because land-uses are expected to switch in time in different ways.
\end{itemize}

\vspace{5mm}

First-order transition probability %$\Rightarrow$ class-change probability in one interval of time.

\begin{columns}
\column{0.3\textwidth}
\vspace{1mm}
\centering \includegraphics[scale=0.25, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig25_transmatrix_landuses.pdf}

\column{0.65\textwidth}
\begin{itemize}\setlength\itemsep{2mm}
\small
%\item There is a significant probability of some of the classes to change in time.

%\item Suggests a Markov chain distribution for the time dimension.

\item[] Further research:

\begin{itemize}\setlength\itemsep{1mm}
\item A Markov chain model for the transition probabilities of classes in time.

\item Multivariate GP prior to relate the transition probabilities among plots (space).
\end{itemize}
\end{itemize}
\end{columns}
\end{frame}
%---------------------------------------------------------------

\section{Conclusions}
%---------------------------------------------------------------
\begin{frame}\frametitle{\normalsize Conclusions}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}
\item Exact GP is computationally intractable for a lot of applications.

\item With special limitations when using full Bayesian inference.

\item We focused on a low-rank representation of stationary GPs.%, based on a basis function approximation via Laplace eigenfunctions.

\item This low-rank GP has a linear functional structure.

\begin{itemize}\setlength\itemsep{1mm}
\item Inference faster and improve sampling efficiency.

\item Attractive for modular probabilistic programming frameworks.
\end{itemize}

\item Recommendations for the values of the key factors of the method.\\[1mm]

\begin{itemize}\setlength\itemsep{1mm}
\item Help improving performance and save computation time.
\item Powerful diagnosis tool.
\end{itemize}

\end{itemize}
\end{frame}
%---------------------------------------------------------------

%---------------------------------------------------------------
\begin{frame}
\setstretch{1.1}

\begin{itemize}\setlength\itemsep{2mm}

\item The method is implemented in the Stan probabilistic programming framework,

\item ..in both Gaussian and non-Gaussian observation models.

\item We demonstrated the practical applicability and the improvement in sampling efficiency.

\item The main drawback is that computational complexity scales exponentially with $D$.

\item Choosing optimal values for $m$ and $c$ is essential to avoid a excessive computational time, especially in multivariate input spaces.

%\item In practice, $D>3$ starts to be quite computationally demanding even for moderate wiggly functions.

%\item In these high dimensional cases, the method may still be used for low-dimensional components in an additive modeling scheme.

%\item Investigating the relationships among the key factors of the method for multivariate input spaces remains a topic for future research. 
\end{itemize}
\end{frame}
%---------------------------------------------------------------

\nobibliography{references}
\end{document}

### 2020-06-17 Paper/manuscript_v2/manuscript_07_02.tex
some update to manuscript

--- 2020-05-13 Paper/manuscript_v2/manuscript_07_02.tex
@@
 Gaussian processes are powerful non-parametric probabilistic models for stochastic functions.
-However, they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
+However they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
 In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions.
@@
 Gaussian processes (GPs) are flexible statistical models for specifying probability distributions over multi-dimensional non-linear functions \citep{rasmussen2006gaussian,neal1997monte}.
-Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate Gaussian distribution.
+Their name stems from the fact that any finite set of function values is jointly distributed as a multivariate Gaussian.
 GPs are defined by a mean and a covariance function.
@@
 The key element of a GP is the covariance function that defines the dependence structure between function values at different inputs.
-However, computing the posterior distribution of a GP comes with a computational issue because of the need to invert the covariance matrix.
+However, computing the posterior distribution of a GP comes with a computational issue because of the need of inverting the covariance matrix.
 Given $n$ observations in the data, the computational complexity and memory requirements of computing the posterior distribution for a GP in general scale as $O(n^3)$ and $O(n^2)$, respectively.
@@
 Using a basis function expansion, a GP is approximated with a linear model which makes inference considerably faster.
-The linear model structure makes GPs easy to implement as building blocks in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if the approximation specific computation is simple.
+The linear model structure makes GPs easy to implement as a building block in more complicated models in modular probabilistic programming frameworks, where there is a big benefit if the approximation specific computation is simple.
 Furthermore, a linear representation of a GP makes it easier to be used as latent function in non-Gaussian observational models allowing for more modelling flexibility.
@@
 Although there are several GP specific software packages available to date, for example, GPML \citep{rasmussen2010gpml},  GPstuff \citep{vanhatalo2013gpstuff}, GPy \citep{gpy2014}, and GPflow \citep{GPflow2017}, each provide efficient implementations only for a restricted range of GP-based models.
-In this paper, we do not focus on the fastest possible inference for a small set of specific GP models, but instead we are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
+In this paper, we do not focus on the fastest possible inference for some specific GP models, but instead are interested in how GPs can be easily used as modular components in probabilistic programming frameworks.
 
@@
 
-A GP is a stochastic process which defines the distribution of a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
+A GP is a stochastic process which defines the distribution over a collection of random variables indexed by a continuous variable, that is, $\left\lbrace f(t): t \in \mathcal{T}\right\rbrace$ for some index set $\mathcal{T}$.
 GPs have the defining property that the marginal distribution of any finite subset of random variables, $\left\lbrace f(t_1), f(t_2), \hdots, f(t_N) \right\rbrace$, is a multivariate Gaussian distribution.
@@
 In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting.
-Consider a data set $\mathcal{D} = \left\lbrace (\bm{x}_n, y_n) \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with a GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
+Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$.
 This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models.
@@
 
-The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and characterizes the correlations between function values at different locations in the input space.
+The covariance function is the crucial ingredient in a GP as it encodes our prior assumptions about the function, and defines a correlation structure which characterizes the correlations between function values at different inputs.
 A covariance function needs to be symmetric and positive semi-definite \citep{rasmussen2006gaussian}.
@@
 Below, we will present the main results around HSGPs relevant for practical applications.
-More details on the theoretical background are provided by \citet{solin2018hilbert}.
+More details and mathematical proofs are provided by \citet{solin2018hilbert}.
 Our starting point for presenting the method is the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
@@
 
-Within $\Omega$, we can write any stationary covariance function with input values $x,x' \in \Omega$ as
+Within $\Omega$, we can write any stationary covariance function with input values $\{x,x'\} \in \Omega$ as
 %
@@
 %
-where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ is the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
+where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
 The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
@@
 
-Thus, the Gram matrix $\bm{K}$ for the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n}$ can be represented as
+Thus, the Gram matrix $K$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
 %
@@
 \begin{equation}\label{eq_approxf}
-f(x) \approx \sum_{j=1}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
+f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
 \end{equation}
@@
 
-The results from the previous section can be generalized to a multidimensional input space with compact support, $\Omega=[-L_1,L_1] \times \dots \times [-L_d,L_d]$ and Dirichlet boundary conditions. 
+The results from the previous section can be generalized to a multidimensional input space with compact support, $\Omega=[-L_1,L_1] \times \dots \times [-L_D,L_D]$ and Dirichlet boundary conditions. 
 In a $D$-dimensional input space, the total number of eigenfunctions and eigenvalues in the approximation is equal to the number of $D$-tuples, that is, possible combinations of univariate eigenfunctions over all dimensions.
@@
 
-Each multivariate eigenfunction $\phi^{\ast}_j:\Omega \rightarrow {\rm I\!R}$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
+Each multivariate eigenfunction $\phi^{\ast}_j$ corresponds to the product of the univariate eigenfunctions whose indices corresponds to the elements of the $D$-tuple $\mathbb{S}_{j\cdotp}$, and each multivariate eigenvalue $\bm{\lambda}^{\ast}_j$ is a $D$-vector with elements that are the univariate eigenvalues whose indices correspond to the elements of the $D$-tuple $\mathbb{S}_{j\bm{\cdotp}}$.
 Thus, for $\bm{x}=\{x_d\}_{d=1}^D \in \Omega$ and $j=1,2,\ldots,m^{\ast}$, we have
@@
 In this work we investigate these relationships for the square exponential covariance function and Mat{\'e}rn ($\nu$=3/2) covariance function in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
-For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
+For other kernels, the relationships will be slightly different, in function of mainly the smoothness or wigglyness of the kernel effects on the posterior function.
 
@@
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
+\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
   \label{fig1_Post_J}
@@
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
+\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
   \label{fig2_Post_L}
@@
 \begin{equation} \label{ch5_eq_cov_periodic}
-k(\tau)= \alpha \exp\left( -\frac{2\sin^{2}(\omega_0\frac{\tau}{2})}{\ell^2}\right),
+k(\bm{\tau})= \alpha \exp\left( -\frac{2\sin^{2}(\omega_0\frac{\bm{\tau}}{2})}{\ell^2}\right),
 \end{equation}

### 2020-09-06 Paper/manuscript_v2/manuscript_07_02.tex
Updates in sections 2.1 and 3.2

@@
 \begin{align*}
-k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\!\nu} \! K_{\nu} \left(\frac{\sqrt{2\nu}r}{\ell}\right),
+k_{\nu}(\bm{\tau})&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}\bm{\tau}}{\ell}\right)^{\!\nu} \! K_{\nu} \left(\frac{\sqrt{2\nu}\bm{\tau}}{\ell}\right),
 \end{align*}
@@
 \begin{align*}
-k_{\infty}(r)&=\alpha \exp \left( -\frac{1}{2} \frac{r^2}{\ell^2}\right),  \\
-k_{\frac{3}{2}}(r)&=\alpha\left(1+\frac{\sqrt{3}r}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}r}{\ell}\right). 
+k_{\infty}(\bm{\tau})&=\alpha \exp \left( -\frac{1}{2} \frac{\bm{\tau}^2}{\ell^2}\right),  \\
+k_{\frac{3}{2}}(\bm{\tau})&=\alpha\left(1+\frac{\sqrt{3}\bm{\tau}}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}\bm{\tau}}{\ell}\right). 
 \end{align*}
@@
 \begin{align*}
-k_{\infty}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
-k_{\frac{3}{2}}(|\bm{x}-\bm{x}'|_{L2})&=\alpha \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right) \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right).
+k_{\infty}(r) &= \alpha \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
+k_{\frac{3}{2}}(r) &= \alpha \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right) \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right).
 \end{align*}
 %
-The previous expressions can be easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
+Notice that the previous expressions, $k_{\infty}(r)$ and $k_{\frac{3}{2}}(r)$, have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
 Using individual length-scales for each dimension turns the isotropic covariance function into a non-isotropic covariance function.
@@
 \begin{align*}
-S_{\nu}(\omega)&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, l^{2\nu}}\left(\frac{2\nu}{l^2}+4\pi^2\omega^2 \right)^{(\nu+D/2)}
+S_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^2 \right)^{(\nu+D/2)}
 \end{align*}
 %
-in $D$ dimensions, where variable $\omega\in {\rm I\!R}$ denotes the frequency, and $\ell$ and $\alpha$ are the lengthscale and magnitude (marginal variance), respectively, of the kernel.
+in $D$ dimensions, where vector $\bm{\omega}\in {\rm I\!R^D}$ is in the frequency domain, and $\ell$ and $\alpha$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
 The particular cases, where $\nu=\infty$ and $\nu=3/2$, take the form
@@
 \begin{align}
-S_{\infty}(\omega)&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\ell^2 \omega^2 / 2), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\omega)&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\omega^2 \right)^{-\frac{D+3}{2}}. \label{eq_specdens_32} 
+S_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\ell^2 \bm{\omega}^2 / 2), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^2 \right)^{-\frac{D+3}{2}}.
+\label{eq_specdens_32}
 \end{align}
 %
-For input dimension $D=3$ and Euclidean distance $\omega=\sqrt{\sum_{i=1}^{3}s_i^2}$, and considering a multidimensional lengthscale $\bm{\ell} \in {\rm I\!R}^{3}$, the spectral densities written above take the form
+For instance, with input dimensionality $D=3$ and $\bm{\omega}=(\omega_1,\omega_2,\omega_3)$, the spectral densities written above take the form
 %
 \begin{align*}
-S_{\infty}(\omega)&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 s_i^2 \right),   \\
-S_{\frac{3}{2}}(\omega)&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 s_i^2 \right)^{-3}.
+S_{\infty}(\bm{\omega})&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 \omega_i^2 \right),   \\
+S_{\frac{3}{2}}(\bm{\omega})&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-3},
 \end{align*}
+%
+where individual length-scales $\ell_i$ for each frequency dimension $\omega_i$ have been used.
 
@@
 %
-where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}).
+where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}) as a function of $\sqrt{\bm{\lambda}^{\ast}_j}$ that denotes the element-wise square root of the vector $\bm{\lambda}^{\ast}_j$.
 We can now write the approximate series expansion of the multivariate function $f$ as

### 2020-09-06 Paper/manuscript_v2/manuscript_07_02.tex
Unify 'lengthscale' word

@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the function to be estimated, which is in turn is characterized by the lengthscale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the function to be estimated, which is in turn is characterized by the length-scale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 
-Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the lengthscale of the function, which depend on the particular choice of the kernel function.
+Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the length-scale of the function, which depend on the particular choice of the kernel function.
 In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu$=3/2) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
@@
 This inequality can be satisfied for arbritrary small $\epsilon$ provided that $L$ and $m$ are sufficiently large \cite[Theorem 1 and 4]{solin2018hilbert}. 
-The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the G.
+The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its length-scale $\ell$, which constitutes a hyperparameter of the G.
 The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
@@
 We start by illustrating how the number of basis functions $m$ and boundary factor $c$ influence the accuracy of the HSGP approximations individually.
-For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
+For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of length-scale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
 Several HSGP models with varying $m$ and $c$ are fitted to this data.
-In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
+In this example, the length-scale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
 Figures~\ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
@@
 Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The lengthscale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} for both the exact GP and the HSGP models.
+The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} for both the exact GP and the HSGP models.
 Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
 Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
-Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the exact GP model and the HSGP models.
+Figure \ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
 Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 119mm 0mm, clip]{ch5_fig5_legend.pdf}}
-\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both exact GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
+\caption{Estimated length-scale (left) and marginal variance (right) parameters of both exact GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig5_lscale_vs_J}
@@
 
-Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function.
+Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
 Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a squared exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+More precisely, for a given GP model (with a squared exponential covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
@@
 
-As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$.
+As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-As the lengthscale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data.
-Basically, if the lengthscale estimate is accurate, we can expect the HSGP approximation to be accurate as well.
-
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
-If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation is assumed to be good.
-If, however, $\hat{\ell}$ does not exceed recommended minimum lengthscale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
-We may also use this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
-As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the lengthscale.
+As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data.
+Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.
+
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
+If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
+If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
+We may also use this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if the estimated $\hat{\ell}$ is below the minimum length-scale, repeat the process while increasing $m$ or decreasing $c$.
+As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the length-scale.
 Thus, increasing $m$ may usually the preferred approach.
@@
 \subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
 The right-side plot is a zoom in of the left-side plot.}
@@
 \subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\nu$=3/2) covariance function.
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\nu$=3/2) covariance function.
 The right-side plot is a zoom in of the left-side plot.}
@@
 
-\subsection{Comparing lengthscale estimates}
-
-In this example, we make a comparison of the lengthscale estimates
-obtained from the exact GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
-For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying lengthscale values.
+\subsection{Comparing length-scale estimates}
+
+In this example, we make a comparison of the length-scale estimates
+obtained from the exact GP and HSGP models. We also have a look at those recommended minimum length-scales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
+For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying length-scale values.
 Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case.
@@
 Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of both exact GP and HSGP models fitted to those datasets.
-The lengthscale estimates as obtained by exact GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
-As noted previously, an accurate estimate of the lengthscale can be a good indicator of a close approximation of the HSGP model to the exact GP model.
-Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the exact GP models, as a function of the lengthscale and number of basis functions.
-
-Comparing the accuracy of the lengthscale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig9_MSE_varing_lscale}, we see that they agree closely with each other for medium lengthscales.
-That is, a good estimation of the lengthscale implies a small RMSE.
-This is no longer true for very small or large lengthscales.
-In small lengthscales, even very small inaccuracies  may have a strong influence on the posteriors predictions and thus on the RMSE.
-In large lengthscales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum lengthscale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}.
-We observe that whenever the estimated lengthscale exceeds the minimally estimable lengthscale, the RMSE of the posterior predictions is small
-(see Figure \ref{fig9_MSE_varing_lscale}). Conversely, when the estimated lengthscale is smaller than the minimally estimable one, the RMSE becomes very large.
+The length-scale estimates as obtained by exact GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
+As noted previously, an accurate estimate of the length-scale can be a good indicator of a close approximation of the HSGP model to the exact GP model.
+Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the exact GP models, as a function of the length-scale and number of basis functions.
+
+Comparing the accuracy of the length-scale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig9_MSE_varing_lscale}, we see that they agree closely with each other for medium length-scales.
+That is, a good estimation of the length-scale implies a small RMSE.
+This is no longer true for very small or large length-scales.
+In small length-scales, even very small inaccuracies  may have a strong influence on the posteriors predictions and thus on the RMSE.
+In large length-scales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum length-scale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}.
+We observe that whenever the estimated length-scale exceeds the minimally estimable length-scale, the RMSE of the posterior predictions is small
+(see Figure \ref{fig9_MSE_varing_lscale}). Conversely, when the estimated length-scale is smaller than the minimally estimable one, the RMSE becomes very large.
 
@@
 \subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
-\caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
+\caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic length-scales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
 }
@@
 \end{flushleft}
-\caption{Data-generating functional lengthscales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates $(\hat{\ell})$ from the exact GP and HSGP models.
-95\% confident intervals of the lengthscale estimates are plotted as dot lines.
+\caption{Data-generating functional length-scales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding length-scale estimates $(\hat{\ell})$ from the exact GP and HSGP models.
+95\% confident intervals of the length-scale estimates are plotted as dot lines.
 The different plots represent the use of different number of basis functions $m$ in the HSGP model.
-The dashed black line represents the recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c} that can be closely approximated by the HSGP model in every case.}
+The dashed black line represents the recommended minimum length-scales provided by Figure \ref{fig6_lscale_vs_J_vs_c} that can be closely approximated by the HSGP model in every case.}
   \label{fig8_Tlscale_vs_Elscale}
@@
 In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
-The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and lengthscale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
+The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and length-scale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
@@
 The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
-The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and length-scale, respectively, of the GP process.
 Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,2,\dots,n$.
@@
 The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
-A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for lengthscale $\ell$.
+A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for length-scale $\ell$.
 We use the same prior distributions for the exact GP model as for the HSGP models.
@@
 which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$.
-$\alpha_1$ and $\ell_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component.
+$\alpha_1$ and $\ell_1$ represent the marginal variance and length-scale, respectively, of this GP prior component.
 The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
@@
 We use $m=30$ basis functions and a boundary factor $c=1.5$.
-The lengthscale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum lengthscale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
+The length-scale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum length-scale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
 This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
@@
 We use $J=10$ cosine terms.
-The lengthscale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
+The length-scale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum length-scale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
 
@@
 Above, $c$ and $\bm{\beta}$ are the intercept and vector of coefficients, respectively, of the linear model.
-$\theta_0$ contains the hyperparameters $\alpha_0$ and $\ell_0$ which are the marginal variance and lengthscale of the general mean GP function, and $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and lengthscale, respectively, of a GP function specific to the male sex ($x_2=2$).
-Scalar lengthscales, $l_0$ and $l_1$, are used in both multivariate covariance functions, assuming isotropic functions.
+$\theta_0$ contains the hyperparameters $\alpha_0$ and $\ell_0$ which are the marginal variance and length-scale of the general mean GP function, and $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and length-scale, respectively, of a GP function specific to the male sex ($x_2=2$).
+Scalar length-scales, $l_0$ and $l_1$, are used in both multivariate covariance functions, assuming isotropic functions.
 %
@@
 
-Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
 
@@
 Figure~\ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large lengthscales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).
@@
 \end{equation}
-where $\alpha$ is the magnitude scale of the covariance, $\ell$ is the characteristic lengthscale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
+where $\alpha$ is the magnitude scale of the covariance, $\ell$ is the characteristic length-scale of the covariance, and $\omega_0$ is the angular frequency defining the periodicity.
 
@@
 In order to assess the accuracy of this representation as a function of the number of cosine terms $J$ considered in the approximation, an empirical evaluation is carried out in a similar way than that in Section \ref{ch5_sec_accuracy} of this work.
-Thus, Figure \ref{ch5_fig28_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the lengthscale of the kernel.
+Thus, Figure \ref{ch5_fig28_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale of the kernel.
 We have considered an approximation to be close enough in terms of satisfying equation (\ref{eq_diff_covs}) with $\varepsilon=0.5\%$.
@@
 \subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix_zoom.pdf}}
-\caption{Relation among the minimum number of terms $J$ in the approximation and the lengthscale ($\ell$) of the periodic squared exponential covariance function.
+\caption{Relation among the minimum number of terms $J$ in the approximation and the length-scale ($\ell$) of the periodic squared exponential covariance function.
 The right-side plot is a zoom in of the left-side plot.}

### 2020-09-07 Paper/manuscript_v2/manuscript_07_02.tex
add label for the y-axis in figure 17
little corrections to section 2.1

@@
 \begin{align*}
-k_{\nu}(\bm{\tau})&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}\bm{\tau}}{\ell}\right)^{\!\nu} \! K_{\nu} \left(\frac{\sqrt{2\nu}\bm{\tau}}{\ell}\right),
+k_{\nu}(r)&=\alpha \, \frac{2^{1-\nu}}{\Gamma(\nu)}\left(\frac{\sqrt{2\nu}r}{\ell}\right)^{\!\nu} \! K_{\nu} \left(\frac{\sqrt{2\nu}r}{\ell}\right),
 \end{align*}
@@
 \begin{align*}
-k_{\infty}(\bm{\tau})&=\alpha \exp \left( -\frac{1}{2} \frac{\bm{\tau}^2}{\ell^2}\right),  \\
-k_{\frac{3}{2}}(\bm{\tau})&=\alpha\left(1+\frac{\sqrt{3}\bm{\tau}}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}\bm{\tau}}{\ell}\right). 
+k_{\infty}(r)&=\alpha \exp \left( -\frac{1}{2} \frac{r^2}{\ell^2}\right),  \\
+k_{\frac{3}{2}}(r)&=\alpha\left(1+\frac{\sqrt{3}r}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}r}{\ell}\right). 
 \end{align*}
@@
 \begin{align*}
-S_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^2 \right)^{(\nu+D/2)}
+S_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^\intercal \bm{\omega} \right)^{(\nu+D/2)}
 \end{align*}
@@
 \begin{align}
-S_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\ell^2 \bm{\omega}^2 / 2), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^2 \right)^{-\frac{D+3}{2}}.
+S_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\frac{1}{2}\ell^2 \bm{\omega}^\intercal \bm{\omega}), \label{eq_specdens_inf}  \\
+S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+3}{2}}.
 \label{eq_specdens_32}
@@
 %
-For instance, with input dimensionality $D=3$ and $\bm{\omega}=(\omega_1,\omega_2,\omega_3)$, the spectral densities written above take the form
+For instance, with input dimensionality $D=3$ and $\bm{\omega}=(\omega_1,\omega_2,\omega_3)^\intercal$, the spectral densities written above take the form
 %
@@
 \centering
-\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia.pdf}
+\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia_01.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.

### 2020-09-07 Paper/manuscript_v2/manuscript_07_02.tex
add diagnosis example for case study I

@@
 \usepackage{multirow}
+\usepackage{multicol}
 \usepackage{float}
@@
 \usepackage{parskip}
+\usepackage{pbox}
 
@@
 
-Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
+Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=60$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
 The true data-generating function and the noisy observations are also plotted.
@@
 
-Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
-Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
-
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
-%
+In practice, as the high computational demands in some cases of exact GPs prevents their use, to diagnose the approximation as a function of $m$, as proposed in section \ref{ch5_subsec_dependency} by using the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is a need.
+The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$, for a given $m$ and $c$, with that minimum length-scale $\ell^*$ brought by the model of relationships and that can be accurately approximated for those given $m$ and $c$.
+Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
+Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis for this case study.
+In a first iteration we choose $m=20$ and $c=1.5$, which is not a feasible combination of values as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$.
+In a second iteration we increase up to $m=30$, the HSGP estimates a length-scale as $\hat{\ell}=0.12$, and the approachable minimum length-scale that the model of relationships brings is $\ell^*=0.19$.
+As $\hat{\ell} < \ell^*$, the approximation should not be accurate enough (see the higher interpolation error of HSGP that GP in Figure \ref{ch5_fig11_MSE_exI_inter}-left), and $m$ needs to be increased.
+In a third iteration we increase up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships brings a approachable minimum length-scale of $\ell^*=0.15$.
+As $\hat{\ell}$ is still less than $\ell^*$ but both are very closed, the approximation should be almost accurate enough (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the HSGP error is approaching that of the exact GP.
+In a fourth iteration we increase up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum length-scale brought by the model of relationships is $\ell^*=0.11$, then the approximation should be accurate (The error is the same than the exact GP in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
 
@@
 
+\begin{figure}
+\begin{tabular}{c c c c}
+{\Large 1}\, \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\,  {\footnotesize$\hat{\ell}=0.09$} & {\Large 2}\, \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.12$} & {\Large 3}\, \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} & {\Large 4}\, \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
+\includegraphics[scale=0.21, trim = 0mm 0mm 4mm 10mm, clip]{caseI_fig6_1_1_vposter.pdf} &
+\includegraphics[scale=0.21, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_2_1_vposter.pdf} &
+\includegraphics[scale=0.21, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_3_1_vposter.pdf} &
+\includegraphics[scale=0.21, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_4_1_vposter.pdf} \\[1mm]
+	& $\hat{\ell} < \ell^*$ & $\hat{\ell} < \ell^*$ & $\hat{\ell} > \ell^*$
+\end{tabular}
+\caption{Diagnosis.}
+	\label{ch5_fig12_diagnosis}
+\end{figure}
+
+Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
+Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
+
 \begin{figure}[!h]
@@
 \end{figure}
+
+The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
 

### 2020-09-08 Paper/manuscript_v2/manuscript_07_02.tex
add diagnosis example for case study I (2)

--- 2020-09-07 Paper/manuscript_v2/manuscript_07_02.tex
@@
 
-Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
-Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
-
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
-%
+{\color{black!75}
+In practice, as the high computational demands of exact GPs prevents their use to asses the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{ch5_subsec_dependency} is recommended.
+%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}.
+Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
+Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis applied for the this case study.
+In a first iteration, the HSGP is launched with $m=20$ and $c=1.5$, which is not a valid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
+In a second iteration, we increase $m$ up to $m=30$ and the HSGP estimates a length-scale of $\hat{\ell}=0.12$; and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
+As $\hat{\ell} < \ell^*$, the approximation should not be sufficiently accurate (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
+In a third iteration, we increase up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$ and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
+As $\hat{\ell}$ is still lower than $\ell^*$, the approximation should not be sufficiently accurate, although, since $\hat{\ell}$ is very closed to $\ell^*$, the approximation might be close to the GP (see errors in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
+In a fourth iteration, we increase up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, then the approximation should be sufficiently accurate.}
 
@@
 
+\begin{figure}
+\begin{tabular}{cccc}
+\hspace{-6mm} {\Large 1}\, \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$}\!
+\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\,  {\footnotesize$\hat{\ell}=0.09$} & \hspace{-5mm} {\Large 2}\, \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\!
+\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.12$} & \hspace{-3mm} {\Large 3}\, \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\!
+\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} & \hspace{-3mm} {\Large 4}\, \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\!
+\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
+	& $\hat{\ell} < \ell^*$ & $\hat{\ell} < \ell^*$ & $\hat{\ell} > \ell^*$\\[1mm]
+\hspace{-3mm} \includegraphics[scale=0.22, trim = 0mm 0mm 4mm 10mm, clip]{caseI_fig6_1_1_vposter.pdf} & \hspace{-3mm}
+\includegraphics[scale=0.22, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_2_1_vposter.pdf} & \hspace{-3mm}
+\includegraphics[scale=0.22, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_3_1_vposter.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.22, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_4_1_vposter.pdf} \\[1mm]
+\end{tabular}
+\caption{Diagnosis of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.}
+	\label{ch5_fig12_diagnosis}
+\end{figure}
+
+Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
+Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
+
 \begin{figure}[!h]

### 2020-09-09 Paper/manuscript_v2/manuscript_07_02.tex
little edit to diagnosis example in case I

@@
 Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
-Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis applied for the this case study.
+Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis applied for this case study.
 In a first iteration, the HSGP is launched with $m=20$ and $c=1.5$, which is not a valid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
@@
 As $\hat{\ell} < \ell^*$, the approximation should not be sufficiently accurate (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
-In a third iteration, we increase up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$ and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
+In a third iteration, we increase $m$ up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$ and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
 As $\hat{\ell}$ is still lower than $\ell^*$, the approximation should not be sufficiently accurate, although, since $\hat{\ell}$ is very closed to $\ell^*$, the approximation might be close to the GP (see errors in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
-In a fourth iteration, we increase up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, then the approximation should be sufficiently accurate.}
+In a fourth iteration, we increase $m$ up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, as $\hat{\ell} > \ell^*$ then the approximation should be sufficiently accurate.}
 
@@
 \begin{tabular}{cccc}
-\hspace{-6mm} {\Large 1}\, \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$}\!
-\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\,  {\footnotesize$\hat{\ell}=0.09$} & \hspace{-5mm} {\Large 2}\, \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\!
+\hspace{-25mm} {\Large 1}\, \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} & \hspace{-5mm} {\Large 2}\, \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\!
 \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.12$} & \hspace{-3mm} {\Large 3}\, \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\!

### 2020-09-09 Paper/manuscript_v2/manuscript_07_02.tex
minor edits to the new paragraph

@@
 Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis applied for this case study.
-In a first iteration, the HSGP is launched with $m=20$ and $c=1.5$, which is not a valid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
-In a second iteration, we increase $m$ up to $m=30$ and the HSGP estimates a length-scale of $\hat{\ell}=0.12$; and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
-As $\hat{\ell} < \ell^*$, the approximation should not be sufficiently accurate (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
-In a third iteration, we increase $m$ up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$ and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
-As $\hat{\ell}$ is still lower than $\ell^*$, the approximation should not be sufficiently accurate, although, since $\hat{\ell}$ is very closed to $\ell^*$, the approximation might be close to the GP (see errors in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
-In a fourth iteration, we increase $m$ up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, as $\hat{\ell} > \ell^*$ then the approximation should be sufficiently accurate.}
+In the first iteration, the HSGP is run with $m=20$ and $c=1.5$, which is not a valid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
+In the second iteration, we increase $m$ up to $m=30$ and the HSGP estimates a length-scale of $\hat{\ell}=0.12$; and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
+As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
+In the third iteration, we increase $m$ up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$ and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
+As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
+In the fourth iteration, we increase $m$ up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, and as $\hat{\ell} > \ell^*$ then approximation is diagnosed to be accurate.}
 

### 2020-09-10 Paper/manuscript_v2/manuscript_07_02.tex
minor update to new paragraph

@@
 
-{\color{black!75}
-In practice, as the high computational demands of exact GPs prevents their use to asses the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{ch5_subsec_dependency} is recommended.
+In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{ch5_subsec_dependency} is recommended.
 %The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}.
@@
 Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis applied for this case study.
-In the first iteration, the HSGP is run with $m=20$ and $c=1.5$, which is not a valid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
-In the second iteration, we increase $m$ up to $m=30$ and the HSGP estimates a length-scale of $\hat{\ell}=0.12$; and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
+In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
+In the second iteration, $m$ is increased up to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
 As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
-In the third iteration, we increase $m$ up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$ and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
+In the third iteration, $m$ is increased up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
 As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
-In the fourth iteration, we increase $m$ up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, and as $\hat{\ell} > \ell^*$ then approximation is diagnosed to be accurate.}
+In the fourth iteration, $m$ is increased up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, and as $\hat{\ell} > \ell^*$ then approximation is diagnosed to be accurate.
 
@@
 \end{tabular}
-\caption{Diagnosis of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.}
+\caption{Iterative diagnosis of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.
+The iterative steps go from the left side to the right side of the figure.}
 	\label{ch5_fig12_diagnosis}

### 2020-09-10 Paper/Manuscript_Stats&Comp/manuscript_ST&CO.tex
add folder with Stats&Comp manuscript files

--- 2020-09-10 Paper/manuscript_v2/manuscript_07_02.tex
@@
-\documentclass[onecolumn,a4paper,11pt]{article}
-\voffset -1cm
-\hoffset -1cm
-\textheight 22.3cm
-\textwidth 15.0cm
-
+%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% This is a general template file for the LaTeX package SVJour3
+% for Springer journals.          Springer Heidelberg 2010/09/16
+%
+% Copy it to a new file with a new name and use it as the basis
+% for your article. Delete % signs as needed.
+%
+% This template includes a few options for different layouts and
+% content for various journals. Please consult a previous issue of
+% your journal as needed.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% First comes an example EPS file -- just ignore it and
+% proceed on the \documentclass line
+% your LaTeX will extract the file if required
+\begin{filecontents*}{example.eps}
+%!PS-Adobe-3.0 EPSF-3.0
+%%BoundingBox: 19 19 221 221
+%%CreationDate: Mon Sep 29 1997
+%%Creator: programmed by hand (JK)
+%%EndComments
+gsave
+newpath
+  20 20 moveto
+  20 220 lineto
+  220 220 lineto
+  220 20 lineto
+closepath
+2 setlinewidth
+gsave
+  .4 setgray fill
+grestore
+stroke
+grestore
+\end{filecontents*}
+%
+\RequirePackage{fix-cm}
+%
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{graphicx}
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
 \usepackage{blindtext}
@@
 \usepackage{microtype}
-
 \usepackage{lscape}
-
-\setcounter{secnumdepth}{5}
-\setcounter{tocdepth}{5}
-
 \usepackage{epstopdf}
-
 \usepackage{natbib}
-\bibliographystyle{plainnat}
-\bibpunct{(}{)}{;}{a}{,}{,}
 \usepackage{subfigure}
@@
 \usepackage{xcolor}
-\graphicspath{{./images/}}
-
 \usepackage{amssymb}
@@
 \usepackage{mathtools}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 \usepackage{bm}
@@
 \usepackage{pbox}
-
 \usepackage[titletoc]{appendix}
@@
 \usepackage{colortbl}
+\usepackage{listings}
+\usepackage{caption}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,breaklinks=true]{hyperref}
+
+% please place your own definitions here and don't use \def but
+\setcounter{secnumdepth}{5}
+\setcounter{tocdepth}{5}
+\graphicspath{{./images/}}
 \DeclareMathOperator{\LogNormal}{LogNormal}
@@
 \DeclareMathOperator{\Normal}{Normal}
-
-\providecommand{\keywords}[1]
-{
-  \small	
-  \textbf{\textit{Keywords---}} #1
+\captionsetup{font=footnotesize}
+
+%
+% Insert the name of "your journal" with
+% \journalname{myjournal}
+%
+\begin{document}
+
+\title{Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming%\thanks{Grants or other notes
+%about the article that should go on the front page should be
+%placed here. General acknowledgments should be placed at the end of the article.}
 }
-
-% \bibpunct[, ]{[}{]}{,}{}{,}{,}
-\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
-
-\usepackage{listings}
-\usepackage{caption}
-\captionsetup{font=footnotesize}
-\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,breaklinks=true]{hyperref}
-
-\title{Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming}
-
-\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R.\ Andersen$^{3}$,\\
-  Arno Solin$^{2}$, Aki Vehtari$^{2}$}
-
-\date{ \small
-$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
-\break
-$^2$ Department of Computer Science, Aalto University, Finland
-\break
-$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
-\break
-$^*$ Corresponding author, Email: gabriuma@gmail.com
+%\subtitle{Do you have a subtitle?\\ If so, write it here}
+
+%\titlerunning{Short form of title}        % if too long for running head
+
+\author{Gabriel Riutort-Mayol \and Paul-Christian Bürkner \and Michael R.\ Andersen \and Arno Solin \and Aki Vehtari
 }
 
-\begin{document}
+%\authorrunning{Short form of author list} % if too long for running head
+
+\institute{Gabriel Riutort-Mayol \at
+              Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain\\
+              \email{gabriuma@gmail.com}           %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+           \and
+           Paul-Christian Bürkner \at
+              Department of Computer Science, Aalto University, Finland
+           \and
+           Michael R.\ Andersen \at
+              Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
+           \and
+           Arno Solin \at
+              Department of Computer Science, Aalto University, Finland
+           \and
+           Aki Vehtari \at
+              Department of Computer Science, Aalto University, Finland
+}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
@@
 Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.
+\keywords{Gaussian process \and Low-rank Gaussian process \and Hilbert space methods \and Sparse Gaussian process \and Bayesian statistics \and Stan}
+% \PACS{PACS code1 \and PACS code2 \and more}
+% \subclass{MSC code1 \and MSC code2 \and more}
 \end{abstract}
-
-\keywords{Gaussian process; Low-rank Gaussian process; Hilbert space methods; Sparse Gaussian process; Bayesian statistics; Stan.}
 
@@
 In Section~\ref{ch4_gp}, we introduce GPs, covariance functions and spectral density functions.
-In Section~\ref{ch5_sec_method}, the reduced-rank approximation to GPs proposed by \citet{solin2018hilbert} is described.
-In Section~\ref{ch5_sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
-Several case studies in which we fit exact and approximate GPs to real and simulated data are provided in Section~\ref{ch5_sec_cases}.
-A brief conclusion of the work is made in Section~\ref{ch5_sec_conclusion}.
-Appendix~\ref{ch5_app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{ch5_sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
+In Section~\ref{sec_method}, the reduced-rank approximation to GPs proposed by \citet{solin2018hilbert} is described.
+In Section~\ref{sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
+Several case studies in which we fit exact and approximate GPs to real and simulated data are provided in Section~\ref{sec_cases}.
+A brief conclusion of the work is made in Section~\ref{sec_conclusion}.
+Appendix~\ref{app_approx_covfun} includes a brief presentation of the mathematical details behind the Hilbert space approximation of a stationary covariance function, and Appendix \ref{sec_periodic} presents a low-rank representation of a GP for the particular case of a periodic covariance function.
 Online supplemental material with more case studies illustrating the performance and applicability of the method can be found online at \url{https://github.com/gabriuma/basis_functions_approach_to_GP} in the subfolder \url{Paper/online_supplemental_material}.
@@
 \begin{align*}
-k_{\infty}(r) &= \alpha \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
-k_{\frac{3}{2}}(r) &= \alpha \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right) \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right).
+k_{\infty}(r) = \alpha & \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
+k_{\frac{3}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right)\\
+& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right).
 \end{align*}
@@
 
-\section{Hilbert space approximate Gaussian process model}\label{ch5_sec_method}
+\section{Hilbert space approximate Gaussian process model}\label{sec_method}
 
@@
 Our starting point for presenting the method is the definition of the covariance function as a series expansion of eigenvalues and eigenfunctions of the Laplacian operator.
-The mathematical details of this approximation are briefly presented in Appendix~\ref{ch5_app_approx_covfun}.
-
-\subsection{Unidimensional GPs} \label{ch5_sec_method_uni}
+The mathematical details of this approximation are briefly presented in Appendix~\ref{app_approx_covfun}.
+
+\subsection{Unidimensional GPs} \label{sec_method_uni}
 
@@
 As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs.
-We will come back to this issue in Section~\ref{ch5_sec_accuracy}.
+We will come back to this issue in Section~\ref{sec_accuracy}.
 
@@
 %
-The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in equation \eqref{eq_eigenproblem} are sinusoidal functions.
+The eigenvalues $\lambda_j>0$ are real and positive because the Laplacian is a positive definite Hermitian operator, and the eigenfunctions $\phi_j$ for the eigenvalue problem in eq.~\eqref{eq_eigenproblem} are sinusoidal functions.
 The solution to the eigenvalue problem is independent of  the specific choice of covariance function and is given by
@@
 
-\subsection{Generalization to multidimensional GPs} \label{ch5_sec_method_multi}
+\subsection{Generalization to multidimensional GPs} \label{sec_method_multi}
 
@@
 %
+%\begin{align}\small
+%\mathbb{S}=
+%\left[ {\begin{array}{ccc}
+%1 & 1 & 1 \nonumber \\
+%1 & 1 & 2 \\
+%1 & 1 & 3 \\
+%1 & 2 & 1 \\
+%1 & 2 & 2 \\
+%1 & 2 & 3 \\
+%2 & 1 & 1 \\
+%2 & 1 & 2 \\
+%2 & 1 & 3 \\
+%2 & 2 & 1 \\
+%2 & 2 & 2 \\
+%2 & 2 & 3 
+%\end{array} } \right].
+%\end{align} 
+%
 \begin{align}\small
 \mathbb{S}=
-\left[ {\begin{array}{ccc}
-1 & 1 & 1 \nonumber \\
-1 & 1 & 2 \\
-1 & 1 & 3 \\
-1 & 2 & 1 \\
-1 & 2 & 2 \\
-1 & 2 & 3 \\
-2 & 1 & 1 \\
-2 & 1 & 2 \\
-2 & 1 & 3 \\
-2 & 2 & 1 \\
-2 & 2 & 2 \\
-2 & 2 & 3 
-\end{array} } \right].
+\left[ {\begin{array}{cccccccccccc}
+1 & 1 & 1 & 1 & 1 & 1 & 2 & 2 & 2 & 2 & 2 & 2 \\
+1 & 1 & 1 & 2 & 2 & 2 & 1 & 1 & 1 & 2 & 2 & 2 \\
+1 & 2 & 3 & 1 & 2 & 3 & 1 & 2 & 3 & 1 & 2 & 3
+\end{array} } \right]^\intercal.
 \end{align}
@@
 
-\section{The accuracy of the approximation}\label{ch5_sec_accuracy}
+\section{The accuracy of the approximation}\label{sec_accuracy}
 
@@
 Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the length-scale of the function, which depend on the particular choice of the kernel function.
-In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu$=3/2) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{ch5_sec_periodic}.
+In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu$=3/2) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
 For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
 
-\subsection{Dependency on the number of basis functions and the boundary condition} \label{ch5_subsec_dependency}
-
-As explained in Section~\ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
+\subsection{Dependency on the number of basis functions and the boundary condition} \label{subsec_dependency}
+
+As explained in Section~\ref{sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
 %
@@
 This inequality can be satisfied for arbritrary small $\epsilon$ provided that $L$ and $m$ are sufficiently large \cite[Theorem 1 and 4]{solin2018hilbert}. 
-The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its length-scale $\ell$, which constitutes a hyperparameter of the G.
+The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its length-scale $\ell$, which constitutes a hyperparameter of the GP.
 The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
@@
 
-\begin{figure}[!t]
-\centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
+\begin{figure*}[!t]
+\centering
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig1_Sigma_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Cov_J.pdf}}
 \vspace{-1mm}
@@
   \label{fig1_Post_J}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig2_Sigma_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Cov_L.pdf}}
 \vspace{-1mm}
@@
   \label{fig2_Post_L}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
@@
 c = 1.05 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf}
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part1.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & 
-\multirow{32}{1cm}{ \includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\ 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part1.pdf}
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part1.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part1.pdf} & 
+\multirow{43}{*}{ \includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{fig3_legend.pdf}}\\ 
 \arrayrulecolor{lightgray}\cline{1-2}
 c = 1.1 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part2.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part2.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part2.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part2.pdf} &\\
 \cline{1-2}
 c = 1.2 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part3.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part3.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part3.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part3.pdf} &\\
 \cline{1-2}
 c = 1.5 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part4.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & \\
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part4.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part4.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part4.pdf} & \\
 \cline{1-2}
 c = 2 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part5.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part5.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part5.pdf} 
+\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part5.pdf} & \\
 \cline{1-2}
 c = 2.5 &
-\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf}
-\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Sigma_part6.pdf}
-\includegraphics[scale=0.215, trim = 0mm 4mm 5mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
+\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Post_part6.pdf}
+\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Sigma_part6.pdf}
+\includegraphics[scale=0.215, trim = 0mm 4mm 5mm 14mm, clip]{fig3_Cov_part6.pdf} &\\
 \arrayrulecolor{darkgray}\hline
@@
   \label{fig3_Post_part1}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{ch5_fig4_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig4_MSE_vs_c.pdf}}
+\end{figure*}
+
+\begin{figure*}[!t]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the exact GP model.
@@
   \label{fig4_MSE_vs_J}
-\end{figure}
-
-\begin{figure}[!h]
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig5_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig5_magnitud_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 119mm 0mm, clip]{ch5_fig5_legend.pdf}}
+\end{figure*}
+
+\begin{figure*}[!h]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
 \caption{Estimated length-scale (left) and marginal variance (right) parameters of both exact GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig5_lscale_vs_J}
-\end{figure}
+\end{figure*}
 
 Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension.
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu$=3/2) covariance function, respectively, and a single input dimension.
 More precisely, for a given GP model (with a squared exponential covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
-Similarly,  Figure \ref{fig6_lscale_vs_J_vs_c_Matern} shows the corresponding plot for the  Mat\'ern ($\nu=3/2$) covariance function.
+Similarly,  Figure \ref{fig7_lscale_vs_J_vs_c_Matern} shows the corresponding plot for the  Mat\'ern ($\nu$=3/2) covariance function.
 We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
@@
 \item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 A $\ell$ increases, the minimum valid $c$ also increases.
@@
 
-As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig7_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c.pdf}}
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
 \hspace{3mm}
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
@@
   \label{fig6_lscale_vs_J_vs_c}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_Matern.pdf}}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig7_lscale_vs_J_vs_c_Matern.pdf}}
 \hspace{3mm}
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\nu$=3/2) covariance function.
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig7_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\'ern ($\nu$=3/2) covariance function.
 The right-side plot is a zoom in of the left-side plot.}
-  \label{fig6_lscale_vs_J_vs_c_Matern}
-\end{figure}
+  \label{fig7_lscale_vs_J_vs_c_Matern}
+\end{figure*}
 
@@
 
-Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
+Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
@@
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} to check the accuracy by analyze individually the different dimensions of a multidimensional GP model.
+However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} to check the accuracy by analyze individually the different dimensions of a multidimensional GP model.
 
@@
 
-Figure \ref{fig7_posterior_varing_lscale_part1} shows the posterior predictions of both exact GP and HSGP models fitted to those datasets.
+Figure \ref{fig8_posterior_varing_lscale_part1} shows the posterior predictions of both exact GP and HSGP models fitted to those datasets.
 The length-scale estimates as obtained by exact GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
 As noted previously, an accurate estimate of the length-scale can be a good indicator of a close approximation of the HSGP model to the exact GP model.
-Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the exact GP models, as a function of the length-scale and number of basis functions.
-
-Comparing the accuracy of the length-scale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig9_MSE_varing_lscale}, we see that they agree closely with each other for medium length-scales.
+Further, Figure \ref{fig10_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the exact GP models, as a function of the length-scale and number of basis functions.
+
+Comparing the accuracy of the length-scale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig10_MSE_varing_lscale}, we see that they agree closely with each other for medium length-scales.
 That is, a good estimation of the length-scale implies a small RMSE.
@@
 We observe that whenever the estimated length-scale exceeds the minimally estimable length-scale, the RMSE of the posterior predictions is small
-(see Figure \ref{fig9_MSE_varing_lscale}). Conversely, when the estimated length-scale is smaller than the minimally estimable one, the RMSE becomes very large.
-
-\begin{figure}[H]
-\centering
-\subfigure{\includegraphics[scale=0.42, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig7_legend.pdf}}\\
+(see Figure \ref{fig10_MSE_varing_lscale}). Conversely, when the estimated length-scale is smaller than the minimally estimable one, the RMSE becomes very large.
+
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.42, trim = 10mm 100mm 5mm 30mm, clip]{fig8_legend.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part3.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part1.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part2.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part3.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part6.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part4.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part5.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part6.pdf}}\\
 \vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 1mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
+\subfigure{\includegraphics[scale=0.31, trim = 1mm 5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part7.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part8.pdf}}
+\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part9.pdf}}\\
 \caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic length-scales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
 }
-  \label{fig7_posterior_varing_lscale_part1}
-\end{figure}
-
-\begin{figure}
-\begin{flushleft}
+  \label{fig8_posterior_varing_lscale_part1}
+\end{figure*}
+
+\begin{figure*}
+%\begin{flushleft}
+\centering
 \begin{tabular}{ccc}
-\multicolumn{3}{c}{ \includegraphics[scale=0.42, trim = 25mm 117mm 90mm 45mm, clip]{ch5_fig8_legend.pdf}}\\
-\includegraphics[scale=0.31, trim = 0mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
-\includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part3.pdf} \\
-\includegraphics[scale=0.31, trim = 0mm 2mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{ch5_fig8_Tlscale_vs_Elscale_part6.pdf}
+\multicolumn{3}{c}{ \includegraphics[scale=0.42, trim = 25mm 117mm 90mm 45mm, clip]{fig9_legend.pdf}}\\
+\includegraphics[scale=0.31, trim = 0mm 24mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part3.pdf} \\
+\includegraphics[scale=0.31, trim = 0mm 2mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part6.pdf}
 \end{tabular}
-\end{flushleft}
-\caption{Data-generating functional length-scales ($\ell$), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding length-scale estimates $(\hat{\ell})$ from the exact GP and HSGP models.
+%\end{flushleft}
+\caption{Data-generating functional length-scales ($\ell$), of the various datasets illustrated in Figure \ref{fig8_posterior_varing_lscale_part1}, versus the corresponding length-scale estimates $(\hat{\ell})$ from the exact GP and HSGP models.
 95\% confident intervals of the length-scale estimates are plotted as dot lines.
@@
   \label{fig8_Tlscale_vs_Elscale}
+\end{figure*}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.38, trim = 0mm 9mm 0mm 0mm, clip]{fig10_MSE_varing_lscale.pdf}
+\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects ($\ell$).}
+  \label{fig10_MSE_varing_lscale}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics[scale=0.38, trim = 0mm 9mm 0mm 0mm, clip]{ch5_fig9_MSE_varing_lscale.pdf}
-\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects ($\ell$).}
-  \label{fig9_MSE_varing_lscale}
-\end{figure}
-
-\newpage
-\section{Case studies}\label{ch5_sec_cases}
+\section{Case studies}\label{sec_cases}
 
@@
 
-\subsection{Simulated data for a univariate function}\label{ch5_sec_univariate_simu}
+\subsection{Simulated data for a univariate function}\label{sec_univariate_simu}
 
 In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
-The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and length-scale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
+The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn ($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and length-scale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
@@
 %
-\begin{align*}\label{ch5_eq_latentgp_simudata1}
+\begin{align*}\label{eq_latentgp_simudata1}
 \begin{split}
@@
 where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $\bm{I}$ representing the identity matrix.
-The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
+The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn ($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
 The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and length-scale, respectively, of the GP process.
@@
 %
-\begin{equation*}\label{ch5_eq_marginalizedgp_simudata1}
+\begin{equation*}\label{eq_marginalizedgp_simudata1}
 \bm{y} \sim \Normal(\bm{0},\bm{K}+ \sigma^2 \bm{I} ),
@@
 
-In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn($\nu$=3/2) spectral density $S$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
+In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn ($\nu$=3/2) spectral density $S$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
 In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
@@
 
-The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
-A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for length-scale $\ell$.
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.\break A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for length-scale $\ell$.
 We use the same prior distributions for the exact GP model as for the HSGP models.
 
-Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=60$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots.
+Figure~\ref{fig11_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=60$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~(\ref{eq_boundary})), and the spline model with 80 knots.
 The true data-generating function and the noisy observations are also plotted.
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{ch5_fig10_Posteriors_exI.pdf}
+\begin{figure*}
+\centering
+\includegraphics[width=0.85\textwidth]{fig11_Posteriors_exI.pdf}
 \caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
-  \label{ch5_fig10_Posteriors_exI}
-\end{figure}
+  \label{fig11_Posteriors_exI}
+\end{figure*}
 
 In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted.
-Figure \ref{ch5_fig11_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots.
+Figure \ref{fig12_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots.
 The SRMSE is computed against the data-generating model.
-From Figures \ref{ch5_fig10_Posteriors_exI} and \ref{ch5_fig11_MSE_exI_inter}, it is seen that the HSGP method yield a good approximation of the exact GP model for both interpolation and extrapolation.
+From Figures \ref{fig11_Posteriors_exI} and \ref{fig12_MSE_exI_inter}, it is seen that the HSGP method yield a good approximation of the exact GP model for both interpolation and extrapolation.
 However, the spline model does not extrapolate data properly.
@@
 
-In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{ch5_subsec_dependency} is recommended.
-%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}.
+In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.
+%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}.
 Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
-Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis applied for this case study.
-In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{ch5_fig12_diagnosis}.
-In the second iteration, $m$ is increased up to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
-As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
+Figure \ref{fig13_diagnosis} depicts the diagnosis applied for this case study.
+In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnosis}.
+In the second iteration, $m$ is increased up to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig7_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
+As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{fig12_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
 In the third iteration, $m$ is increased up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
-As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{ch5_fig11_MSE_exI_inter}-left).
+As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{fig12_MSE_exI_inter}-left).
 In the fourth iteration, $m$ is increased up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, and as $\hat{\ell} > \ell^*$ then approximation is diagnosed to be accurate.
 
-\begin{figure}[!t]
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 5mm 0mm, clip]{ch5_fig11_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig11_MSE_exI_extra.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 29mm 55mm 110mm 0mm, clip]{ch5_fig11_MSE_exI_legend.pdf}}
+\begin{figure*}[!t]
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_inter.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_extra.pdf}}
 \caption{Standardized root mean square error (SRMSE) of the different methods against the data-generating function.
@@
 The standard deviation of the mean of the SRMSE is plotted as dashed lines.}
-  \label{ch5_fig11_MSE_exI_inter}
-\end{figure}
-
-\begin{figure}
-\begin{tabular}{cccc}
-\hspace{-25mm} {\Large 1}\, \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} & \hspace{-5mm} {\Large 2}\, \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\!
-\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.12$} & \hspace{-3mm} {\Large 3}\, \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\!
-\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} & \hspace{-3mm} {\Large 4}\, \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\!
-\pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
-	& $\hat{\ell} < \ell^*$ & $\hat{\ell} < \ell^*$ & $\hat{\ell} > \ell^*$\\[1mm]
-\hspace{-3mm} \includegraphics[scale=0.22, trim = 0mm 0mm 4mm 10mm, clip]{caseI_fig6_1_1_vposter.pdf} & \hspace{-3mm}
-\includegraphics[scale=0.22, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_2_1_vposter.pdf} & \hspace{-3mm}
-\includegraphics[scale=0.22, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_3_1_vposter.pdf} & \hspace{-4mm}
-\includegraphics[scale=0.22, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_4_1_vposter.pdf} \\[1mm]
+  \label{fig12_MSE_exI_inter}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\begin{tabular}{llll}
+{\Large 1}\;\;\; \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} &  {\Large 2}\;\;\; \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.12$} &  {\Large 3}\;\;\; \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} &  {\Large 4}\;\;\; \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
+	& \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} > \ell^*$\\[1mm]
+\hspace{-3mm} \includegraphics[scale=0.26, trim = 0mm 0mm 4mm 10mm, clip]{fig13_diag_iter1_exI.pdf} & \hspace{-3mm}
+\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter2_exI.pdf} & \hspace{-3mm}
+\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter3_exI.pdf} & \hspace{-4mm}
+\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter4_exI.pdf} \\[1mm]
 \end{tabular}
@@
 The iterative steps go from the left side to the right side of the figure.}
-	\label{ch5_fig12_diagnosis}
-\end{figure}
-
-Figure~\ref{ch5_fig11_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+	\label{fig13_diagnosis}
+\end{figure*}
+
+Figure~\ref{fig14_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
 The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig11_time_exI.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 10mm 0mm, clip]{fig14_time_exI.pdf}}
 \caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
@@
 The standard deviation of the computational time is plotted as dashed lines.}
-  \label{ch5_fig11_time_exI}
+  \label{fig14_time_exI}
 \end{figure}
@@
 
-\subsection{Birthday data}\label{ch5_sec_birthday}
+\subsection{Birthday data}\label{sec_birthday}
 %
@@
 As the total number of days within the period is $T=7305$ ($t=1,2,\dots,T$), a full Bayesian inference with MCMC for a exact GP model is memory and time consuming.
-We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix~\ref{ch5_sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
+We will use the HSGP method as well as the low-rank GP model with a periodic covariance function described in Appendix~\ref{sec_periodic} which is based on expanding the periodic covariance function into a series of stochastic resonators \citep{solin2014explicit}.
 
@@
 %
-\begin{equation} \label{ch5_eq_mean_brithday}
+\begin{equation} \label{eq_mean_brithday}
 \mu(t) = f_1(t) + f_2(t) + f_3(t) + f_4(t).
@@
 %
-\begin{equation*}
-f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \alpha_2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
-\end{equation*}
+\begin{align*}
+&f_2(t) \sim \GP(0,k_2), \\
+&k_2(t,t') = \alpha_2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
+\end{align*}
 
@@
 %
-\begin{equation*}
-f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \alpha_3 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
-\end{equation*}
+\begin{align*}
+&f_3(t) \sim \GP(0,k_3), \\
+&k_3(t,t') = \alpha_3 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
+\end{align*}
 
@@
 
-The component $f_1(t)$ will be approximated using the HSGP model and the function values $f_1(t)$ are approximated as in equation (\ref{eq_approxf}), with the squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+The component $f_1(t)$ will be approximated using the HSGP model and the function values $f_1(t)$ are approximated as in eq.~(\ref{eq_approxf}), with the squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
 We use $m=30$ basis functions and a boundary factor $c=1.5$.
@@
 The year effects $f_2(t)$ and week effects $f_3(t)$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper.
-However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{ch5_sec_periodic}).
-Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in equation (\ref{ch5_eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in equation (\ref{ch5_eq_q_2}).
+However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{sec_periodic}).
+Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in eq.~(\ref{eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in eq.~(\ref{eq_q_2}).
 We use $J=10$ cosine terms.
-The length-scale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum length-scale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
-
-Figure~\ref{ch5_fig27_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and yearly pattern $f_2(t)$ for the whole period, jointly with the observed data.
-Figure \ref{ch5_fig27_posteriors_oneyear_birthday} shows the model for one year (1972) only.
+The length-scale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum length-scale reported by Figure \ref{figB1_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
+
+Figure~\ref{fig15_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and yearly pattern $f_2(t)$ for the whole period, jointly with the observed data.
+Figure \ref{fig16_posteriors_oneyear_birthday} shows the model for one year (1972) only.
 In this figure, the special days effects $f_4(t)$ in the year can be clearly represented.
-The posterior means of the the function $\mu(t)$ and the components $f_1(t)$ (long-term trend) and $f_2(t)$ (year pattern) are also plotted in this Figure \ref{ch5_fig27_posteriors_oneyear_birthday}.
-Figure \ref{ch5_fig27_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3(t)$ can be clearly represented.
-The mean of the the function $\mu(t)$ and components $f_1(t)$ (long-term trend), $f_2(t)$ (year pattern) and $f_4(t)$ (special-days effects) are also plotted in this Figure \ref{ch5_fig27_posteriors_onemonth_birthday}.
+The posterior means of the the function $\mu(t)$ and the components $f_1(t)$ (long-term trend) and $f_2(t)$ (year pattern) are also plotted in this Figure \ref{fig16_posteriors_oneyear_birthday}.
+Figure \ref{fig17_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3(t)$ can be clearly represented.
+The mean of the the function $\mu(t)$ and components $f_1(t)$ (long-term trend), $f_2(t)$ (year pattern) and $f_4(t)$ (special-days effects) are also plotted in this Figure \ref{fig17_posteriors_onemonth_birthday}.
 
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig27_posteriors_birthday.pdf}}
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[width=0.85\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig15_posteriors_birthday.pdf}}
 \caption{Posterior means of the long-term trend ($f_1(\cdot)$) and year effects pattern ($f_2(\cdot)$) for the whole series. }
-  \label{ch5_fig27_posteriors_birthday}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig27_posteriors_oneyear_birthday.pdf}}
+  \label{fig15_posteriors_birthday}
+\end{figure*}
+
+\begin{figure*}[!t]
+\centering
+\subfigure{\includegraphics[width=0.85\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig16_posteriors_oneyear_birthday.pdf}}
 \caption{Posterior means of the function $\mu(\cdot)$ for the year 1972 of the series.
@@
 }
-  \label{ch5_fig27_posteriors_oneyear_birthday}
-\end{figure}
-
-\begin{figure}[!h]
-\centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{ch5_fig27_posteriors_onemonth_birthday.pdf}}
+  \label{fig16_posteriors_oneyear_birthday}
+\end{figure*}
+
+\begin{figure*}[!h]
+\centering
+\subfigure{\includegraphics[width=0.85\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig17_posteriors_onemonth_birthday.pdf}}
 \caption{Posterior means of the function $\mu(\cdot)$ for the month of January of 1972.
@@
 }
-  \label{ch5_fig27_posteriors_onemonth_birthday}
-\end{figure}
-
-\newpage 
-\subsection{Leukemia data}\label{ch5_sec_bf_caseVI}
+  \label{fig17_posteriors_onemonth_birthday}
+\end{figure*}
+
+\subsection{Leukemia data}\label{sec_bf_caseVI}
 
@@
 %
-\begin{equation*}
-p(y_i > t_i \mid f)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
-\end{equation*}
+\begin{align*}
+p(y_i > t_i \mid f) &= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i \\
+&=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
+\end{align*}
 %
@@
 %
-\begin{figure}
-\centering
-\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{ch5_fig21_posteriors_leukemia_01.pdf}
+\begin{figure*}
+\centering
+\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{fig18_posteriors_leukemia.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
 The thick line in each graph is the posterior mean estimated using a HSGP model, and the thin lines represent pointwise 95\% credible intervals.}
-  \label{ch5_fig21_posteriors_leukemia}
-\end{figure}
-
-Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in equation (\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in equation (\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
-
-Figure \ref{ch5_fig21_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
+  \label{fig18_posteriors_leukemia}
+\end{figure*}
+
+Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in eq.~(\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+
+Figure \ref{fig18_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
 These posterior estimates correspond to the HSGP model with $m=10$ basis functions and $c=3$ boundary factor.
 There are clear  non-linear patterns and the right bottom subplot also shows that the conditional function associated with WBC has an interaction with TDI.
-Figure~\ref{ch5_fig22_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in equation (\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{ch5_fig22_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
-Increasing the boundary factor also significantly increases the time of computation (Figure \ref{ch5_fig22_elpd_leukemia}-right).
+Figure~\ref{fig19_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig19_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig19_elpd_leukemia}-right).
 With a moderate number of univariate basis functions ($m=15$), the HSGP model becomes slower than the exact GP model, in this specific application with $3$ input variables, as the total number of multivariate basis functions becomes $15^3 = 3375$ and is therefore quite high.
 %
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{ch5_fig22_elpd_leukemia.pdf}} 
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 10mm 0mm, clip]{ch5_fig23_time_leukemia.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 50mm 120mm 0mm, clip]{ch5_fig24_legend_leukemia.pdf}}
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig19_elpd_leukemia.pdf}} 
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig19_time_leukemia.pdf}}
 \caption{Expected log predictive density (ELPD; left) and time of computation in seconds per iteration (iteration of the HMC sampling method; right) as a function of the number of basis functions $m$ and boundary factor $c$.}
-  \label{ch5_fig22_elpd_leukemia}
-\end{figure}
+  \label{fig19_elpd_leukemia}
+\end{figure*}
 
@@
 
-\section{Conclusion}\label{ch5_sec_conclusion}
+\section{Conclusion}\label{sec_conclusion}
 
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time especially in multivariate input spaces.
+Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time especially in multivariate input spaces.
 However, in practice, input dimensionalities larger than three start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
-In these high dimensional cases, the proposed approximate GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions.
+In these high dimensional cases, the proposed approximate GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions, as complexity is linear with the number of additive components.
 
@@
 
-\section{Approximation of the covariance function using Hilbert space methods} \label{ch5_app_approx_covfun}
+\section{Approximation of the covariance function using Hilbert space methods} \label{app_approx_covfun}
 
@@
 %
-\begin{equation}\label{ch5_eq_S}
+\begin{equation}\label{eq_S}
 S(||\bm{w}||)=a_0+a_1||\bm{w}||^2+a_2(||\bm{w}||^2)^2+a_3(||\bm{w}||^2)^3+\cdots.
@@
 %
-\begin{equation}\label{ch5_eq_K}
+\begin{equation}\label{eq_K}
 \mathcal{K}=a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+a_3(-\nabla^2)^3+\cdots,
@@
 %
-\begin{eqnarray}\label{ch5_eq_formalkernel}
+\begin{eqnarray}\label{eq_formalkernel}
 l(\bm{x},\bm{x}')^s= \sum_j \lambda_j^s \phi_j(\bm{x}) \phi_j(\bm{x}'),
@@
 %
-\begin{equation*}
-[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \,\mathrm{d}\bm{x}'.
-\end{equation*}
-
-Then, looking at equations (\ref{ch5_eq_K}) and (\ref{ch5_eq_formalkernel}), it can be concluded 
-%
-\begin{equation}\label{ch5_eq_k}
+\begin{align*}
+&[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \\
+& \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \,\mathrm{d}\bm{x}'.
+\end{align*}
+
+Then, looking at equations (\ref{eq_K}) and (\ref{eq_formalkernel}), it can be concluded 
+%
+\begin{equation}\label{eq_k}
 k(\bm{x},\bm{x}')= \sum_j [a_0+a_1\lambda_j^1+a_2\lambda_j^2+\cdots] \phi_j(\bm{x}) \phi_j(\bm{x}').
 \end{equation} 
-By letting $||\bm{w}||^2=\lambda_j$ the spectral density in Equation (\ref{ch5_eq_S}) becomes
+By letting $||\bm{w}||^2=\lambda_j$ the spectral density in eq.~(\ref{eq_S}) becomes
 %
@@
 \end{equation*}
-and substituting in equation (\ref{ch5_eq_k}) then leads to the final form
-%
-\begin{equation}\label{ch5_eq_k_2}
+and substituting in eq.~(\ref{eq_k}) then leads to the final form
+%
+\begin{equation}\label{eq_k_2}
 k(\bm{x},\bm{x}')= \sum_j S(\sqrt{\lambda_j}) \phi_j(\bm{x}) \phi_j(\bm{x}'),
@@
 
-\section{Low-rank Gaussian process with a periodic covariance function}\label{ch5_sec_periodic}
+\section{Low-rank Gaussian process with a periodic covariance function}\label{sec_periodic}
 
@@
 %
-\begin{equation} \label{ch5_eq_cov_periodic}
+\begin{equation} \label{eq_cov_periodic}
 k(\tau)= \alpha \exp\left( -\frac{2\sin^{2}(\omega_0\frac{\tau}{2})}{\ell^2}\right),
@@
 
-\citet{solin2014explicit} derive a cosine series expansion for the periodic covariance function (\ref{ch5_eq_cov_periodic}) as follows,
-% 
-\begin{equation} \label{ch5_eq_cov_periodic_taylor_approx}
+\citet{solin2014explicit} derive a cosine series expansion for the periodic covariance function (\ref{eq_cov_periodic}) as follows,
+% 
+\begin{equation} \label{eq_cov_periodic_taylor_approx}
 k(\tau)= \alpha \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 \tau),
@@
 %
-\begin{equation} \label{ch5_eq_q}
+\begin{equation} \label{eq_q}
 \tilde{q}_j^2= \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
@@
 %
-\begin{equation} \label{ch5_eq_q0}
+\begin{equation} \label{eq_q0}
 \tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
 \end{equation}
-The covariance in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
-This approximation converges to equation (\ref{ch5_eq_cov_periodic}) when $J \rightarrow \infty$.
-
-An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}), and thus leading to the following variance coefficients:
-%
-\begin{equation}\label{ch5_eq_q_2}
+The covariance in eq.~(\ref{eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation.
+This approximation converges to eq.~(\ref{eq_cov_periodic}) when $J \rightarrow \infty$.
+
+An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{eq_q}) and (\ref{eq_q0}), and thus leading to the following variance coefficients:
+%
+\begin{equation}\label{eq_q_2}
 \begin{split}
-\tilde{q}_j^2= \frac{2\mathrm{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
-\tilde{q}_0^2= \frac{\mathrm{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
+\tilde{q}_j^2=& \frac{2\mathrm{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
+\tilde{q}_0^2=& \frac{\mathrm{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
 \end{split}
@@
 
-In order to assess the accuracy of this representation as a function of the number of cosine terms $J$ considered in the approximation, an empirical evaluation is carried out in a similar way than that in Section \ref{ch5_sec_accuracy} of this work.
-Thus, Figure \ref{ch5_fig28_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale of the kernel.
-We have considered an approximation to be close enough in terms of satisfying equation (\ref{eq_diff_covs}) with $\varepsilon=0.5\%$.
+In order to assess the accuracy of this representation as a function of the number of cosine terms $J$ considered in the approximation, an empirical evaluation is carried out in a similar way than that in Section \ref{sec_accuracy} of this work.
+Thus, Figure \ref{figB1_m_lscale_periodic} shows the minimum number of terms $J$ required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale of the kernel.
+We have considered an approximation to be close enough in terms of satisfying eq.~(\ref{eq_diff_covs}) with $\varepsilon=0.005 \int k(\tau) \,\mathrm{d}\tau$ (0.5$\%$ of the total area under the curve of the exact covariance function $k$).
 Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix.pdf}}
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{figB1_m_lscale_periodic_appendix.pdf}}
 \hspace{3mm}
-\subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{ch5_fig28_m_lscale_periodic_appendix_zoom.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 0mm 8mm 5mm 0mm, clip]{figB1_m_lscale_periodic_appendix_zoom.pdf}}
 \caption{Relation among the minimum number of terms $J$ in the approximation and the length-scale ($\ell$) of the periodic squared exponential covariance function.
 The right-side plot is a zoom in of the left-side plot.}
-  \label{ch5_fig28_m_lscale_periodic}
-\end{figure}
+  \label{figB1_m_lscale_periodic}
+\end{figure*}
 
@@
 \end{equation*}
-the covariance $k(\tau)$ in equation (\ref{ch5_eq_cov_periodic_taylor_approx}) can be written as
-%
-\begin{equation} \label{ch5_eq_cov_periodic_taylor_approx_2}
-k(x,x')\approx \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
-\end{equation}
+the covariance $k(\tau)$ in eq.~(\ref{eq_cov_periodic_taylor_approx}) can be written as
+%
+\begin{align} \label{eq_cov_periodic_taylor_approx_2}
+k(x,x')\approx \; & \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') \right. \nonumber\\
+& \left. + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
+\end{align}
 With this approximation for the periodic squared exponential covariance function $k(x,x')$, the approximate GP model $f(x) \sim \GP\left(0,k(x,x')\right)$ equivalently leads to a linear representation of $f(\cdot)$ via
 %
-\begin{equation} \label{ch5_eq_f_period}
-f(x) \approx \alpha^{1/2} \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
-\end{equation}
+\begin{align} \label{eq_f_period}
+f(x) \approx \; & \alpha^{1/2} \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j \right. \nonumber\\
+& \left.+  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
+\end{align}
 where $\beta_j \sim \Normal(0,1)$, with $j=1,\dots,2J+1$.
@@
 The computational cost of this approximation scales as $O\big(n(2J+1) + (2J+1)\big)$, where $n$ is the number of observations and $J$ the number of cosine terms.
-The parameterization in equation (\ref{ch5_eq_f_period}) is naturally in the non-centered form with independent prior distributions on
+The parameterization in eq.~(\ref{eq_f_period}) is naturally in the non-centered form with independent prior distributions on
 $\beta_j$, which makes posterior inference easier.
 
-\section*{Acknowledgements}
-
+\begin{acknowledgements}
 We thank Academy of Finland (grants 298742, 308640, and 313122), Finnish Center for Artificial Intelligence, and Technology Industries of Finland Centennial Foundation (grant 70007503; Artificial Intelligence for Research and Development) for partial support of this research.
 We also acknowledge the computational resources provided by the Aalto Science-IT project.
-
-\bibliography{references}
+\end{acknowledgements}
+
+% Authors must disclose all relationships or interests that 
+% could have direct or potential influence or impart bias on 
+% the work: 
+%
+% \section*{Conflict of interest}
+%
+% The authors declare that they have no conflict of interest.
+
+% BibTeX users please use one of
+\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+\bibliography{references}   % name your BibTeX data base
+
+%% Non-BibTeX users please use
+%\begin{thebibliography}{}
+%%
+%% and use \bibitem to create references. Consult the Instructions
+%% for authors for reference list style.
+%%
+%\bibitem{RefJ}
+%% Format for Journal Reference
+%Author, Article title, Journal, Volume, page numbers (year)
+%% Format for books
+%\bibitem{RefB}
+%Author, Book title, page numbers. Publisher, place (year)
+%% etc
+%\end{thebibliography}
 
 \end{document}
+% end of file template.tex

### 2020-09-10 Paper/StanCon/StanCon_summary_01.tex
add StanCon poster folder

\documentclass[onecolumn,a4paper,11pt]{article}
\voffset -1cm
\hoffset -1cm
\textheight 22.3cm
\textwidth 15.0cm

\usepackage{times}
\usepackage{parskip}

\title{Practical implementation of a Hilbert space reduced-rank approximate Bayesian Gaussian process}

\author{Gabriel Riutort-Mayol$^{1}$, Paul-Christian Bürkner$^2$, Michael R.\ Andersen$^{3}$,\\
  Arno Solin$^{2}$, Aki Vehtari$^{2}$}

\date{\small Summary for StanCon 2020}

\begin{document}

\maketitle

Gaussian processes are powerful non-parametric probabilistic models for stochastic functions. However, when they are implemented using covariance matrices, the computation cost of evaluating the log posterior density is $O(n^3)$ for $n$ observations, which makes them computationally intractable with moderate or large datasets. In this talk, we focus on a novel approach for reduced-rank approximate Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions. This reduced-rank approximation is simple and exhibits an attractive computational complexity due to its linear structure. The Laplace eigenfunctions can be pre-computed analytically with cost $O(m^2n)$ for $m$ basis functions, and the computational cost of evaluating the log posterior density is $O(mn + m)$. An additional advantage is the reduced memory requirements of automatic differentiation methods used in modern probabilistic programming frameworks, such as Stan. The main contribution of this work is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, domain of the prediction space, and wiggliness of the latent function. We provide intuitive visualizations and recommendations for choosing the values of these factors based on the recognized relationships, which make it easier for users to improve approximation accuracy and computational performance. We also propose diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.  Trade-off between accuracy and computation can also be carried out. However, the number of basis function in multi-dimensional input spaces grows exponentially with dimensions, which makes looking at the recommendations and the diagnosis tool proposed in this work essential to avoid excessive computation time in multivariate cases. In high dimensional cases, this method may still be used for low-dimensional components in an additive modeling scheme as complexity is linear with the number of additive components. Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.

---------------------
\vspace{-2mm}
\begin{flushleft}
\scriptsize
$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
\break
$^2$ Department of Computer Science, Aalto University, Finland
\break
$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
\end{flushleft}

%\bibliography{references}

\end{document}

### 2020-09-10 Paper/StanCon/poster/HSGP_StanCon_poster.tex
add StanCon poster folder

\documentclass[portrait,a1,final]{a0poster} % a0poster class sets the paper size
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[SCI]{aaltologo} % RGB, Coated, Uncoated
%\usepackage{times}
%\usepackage{lastpage}
\usepackage[center]{titlesec} % For changing the font on chapters, sections, etc.
\usepackage{amsmath,amsfonts,amssymb,amsbsy}
\usepackage{soul}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{sectsty}
\usepackage{natbib}
\usepackage{subcaption}
\usepackage{url}
\usepackage{bm}
\usepackage[colorlinks,linkcolor=black,citecolor=black, urlcolor=blue,filecolor=blue]{hyperref}
\usepackage[left=0.5cm,right=0.5cm,bottom=1cm,top=1.5cm]{geometry} 
\usepackage{changepage}
\usepackage{amsthm}
\usepackage{tcolorbox}
\usepackage{pbox}
\usepackage{multicol}
\usepackage{nicefrac}
\usepackage{setspace}

\newtheorem*{remark}{Remark}

%%%%%%titleformat
\newcommand{\cajados}[1]{\fcolorbox{aaltoLightGreen!30}{aaltoLightGreen!30}
{\parbox[c][8mm][c]{0.95\textwidth}
{\centering\normalsize\sc\bfseries\sffamily\color{black} \thesection. #1}}}

\titleformat{\section}[block]
{\normalfont\filcenter\color{aaltoGray}}{}{1em}{\cajados}

%\titleformat{\section}{\centering\normalsize\sc\bfseries\sffamily\color{aaltoPurple}}{\textcolor{aaltoPurple}{\thesection}}{1em}{}

\titleformat{\subsection}{\centering\normalsize\sc\bfseries\sffamily\color{aaltoPurple}}{\textcolor{aaltoPurple}{\thesubsection}}{1em}{} % Text size for a1 posters

\renewcommand{\emph}[1]{\sffamily{\textcolor{aaltoPurple}{#1}}}

% Beamer-style math font
%\usepackage{fouriernc}
\usepackage{sfmath}

% redefine the bullet symbol 
\renewcommand\labelitemi{\textcolor{aaltoPurple}{$\blacktriangleright$}\,}

% reduce the space between bullets
\let\tempone\itemize
\let\temptwo\enditemize
\renewenvironment{itemize}{\tempone\addtolength{\itemsep}{-0.3\baselineskip}}{\temptwo}

%%%%%%%%%%%%%%%%%%% for tikz figures %%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{ifthen}
\usepackage{tikz}
\usepackage{pgfplots}  
\usetikzlibrary{positioning} 
% \usepgfplotslibrary{external}
%\tikzexternalize

%\usetikzlibrary{matrix}
%\usetikzlibrary{calc}
\newlength{\figurewidth}
\newlength{\figureheight}

\def\figpdfdir{fig/} % directory for pdf-figures
\def\figtikzdir{tikz/} % directory for tikz-figures

% this is replacement for the \input command used in the figure-environment which
% takes into account whether pdf is forced
\newcommand{\minput}[2][]{
\ifthenelse{\equal{#1}{pdf}}
	{ \includegraphics{\figpdfdir #2} }
	{ \tikzset{external/remake next} \tikzsetnextfilename{#2} \input{\figtikzdir #2} }
}

% for externalization
% \usetikzlibrary{external}
% \tikzexternalize[prefix=\figpdfdir]
% \tikzset{external/system call={lualatex
% 	\tikzexternalcheckshellescape -halt-on-error -interaction=batchmode
% 	-jobname "\image" "\texsource"}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\vc}[1] { #1 } %
\newcommand{\vs}[1] { #1 }
\newcommand{\tp}{\mathsf{T}}
\newcommand{\ti}[1] { \tilde{#1} }
\newcommand{\mc}[1] { \mathcal{#1} } 
\newcommand{\tx}[1] { \text{#1} } 
\newcommand{\given} { \,|\, }
\newcommand{\data} {D}
\newcommand{\subjto}{ \, \tx{s.t.} }
\newcommand{\tr}{^{\text{T}} }
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}

%
\newcommand{\pr}[1]{ \Pr {\left[#1\right]} }
\newcommand{\partiald}[2]{\frac{\partial {#1}}{\partial {#2}}}
\newcommand{\mean}[2][] { \mathrm{E}_{#1} {\left[#2\right]} }
\newcommand{\Var}[1] { \mathrm{Var} {\left[#1\right]} }
\newcommand{\var}[1] { \mathrm{var} {\left(#1\right)} }
\newcommand{\Cov}[1] { \mathrm{Cov} {\left[#1\right]} }
%\newcommand{\cov}[1] { \mathrm{cov} {\left({#1}\right)}}
\newcommand{\Tr}[1] { \mathrm{Tr} {\left[#1\right]} }
\newcommand{\KL}[2] { \mathrm{KL} {\left(#1 \, \| \, #2\right)} }
\newcommand{\x}{\mathbf{x}}
\newcommand{\cx}{\mathbf{c}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\id}{\text{d}}
%
%\newcommand{\Normal}[1] { \mathrm{N} {\left(#1\right)}  } 
\newcommand{\Student}[2][] { t_{#1} {\left(#2\right)} }
\newcommand{\halfStudent}[2][] { t_{#1}^+ {\left(#2\right)} }
\newcommand{\SInvchi}[1] { \mathrm{Scale\text{-}Inv\text{-}}\chi^2 {\left(#1\right)} }
\newcommand{\InvGamma}[1] { \mathrm{Inv\text{-}Gamma} {\left(#1\right)} }
\newcommand{\Cauchy}[1] { \mathrm{C} {\left(#1\right)} }
\newcommand{\halfCauchy}[1] { \mathrm{C}^+ {\left(#1\right)} }
\newcommand{\Beta}[1] {\mathrm{Beta}{\left(#1\right)} }
\newcommand{\Ber}[1] {\mathrm{Ber}{\left(#1\right)} }

%
\newcommand{\NormalNP} { \mathrm{N} } 
\newcommand{\StudentNP}[1][] { t_{#1} }
\newcommand{\halfStudentNP}[1][] { t_{#1}^+ }
\newcommand{\SInvchiNP} { \mathrm{Scale\text{-}Inv\text{-}}\chi^2  }
\newcommand{\InvGammaNP} { \mathrm{Inv\text{-}Gamma} }
\newcommand{\CauchyNP} { \mathrm{C}  }
\newcommand{\halfCauchyNP} { \mathrm{C}^+ }
\newcommand{\BetaNP} {\mathrm{Beta} }
\newcommand{\BerNP} {\mathrm{Ber} }

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\GP}{GP}
%\DeclareMathOperator{\Normal}{Normal}
\DeclareMathOperator{\cov}{cov}

% --- Acronyms
\usepackage{xspace}
\newcommand{\acr}[1]{\textsc{#1}\xspace}
\newcommand{\gp}{\acr{gp}}
\newcommand{\ep}{\acr{ep}}
\newcommand{\gps}{\acr{gp}{\!\,}s }
\newcommand{\abc}{\acr{abc}}
\newcommand{\bo}{\acr{bo}}
% optimization

\definecolor{navyblue}{rgb}{0,0,0.55}
\definecolor{lightblue}{rgb}{0.91,0.91,0.97}

\newcommand{\sectionspace}{0em} % Free space before each section inside a minipage
\newcommand{\figurespace}{0.5em} % Free space around figures inside a minipage (where floats are not allowed)

% Reduce the vertical space taken by the equations
\setlength{\abovedisplayskip}{10pt}
\setlength{\belowdisplayskip}{10pt}
%\setlength{\abovedisplayshortskip}{10pt}
%\setlength{\belowdisplayshortskip}{10pt}

%% Font sizes for a0poster are
%\tiny
%\scriptsize
%\footnotesize
%\small
%\normalsize
%\large
%\Large
%\LARGE
%\huge
%\Huge
%\veryHuge
%\VeryHuge
%\VERYHuge

%% Official colors from aaltologo-package (visual-identity guideline)
% aaltoBlack
% aaltoGray
% aaltoGrayScale (for b&w prints)
% aaltoYellow
% aaltoOrange
% aaltoRed
% aaltoFuchsia
% aaltoPurple
% aaltoBlue
% aaltoTurquoise
% aaltoGreen
% aaltoLightGreen
% You can use \textcolor{<aaltocolor>}{<your text>) to change the colors of text, or

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\thispagestyle{empty} % Removes the page number

%------------ minipage for the logo & title ------------%
\begin{minipage}[t]{0.98\linewidth} 
\vspace{0pt} % A trick to align the parallel minipages on top

\centering

\vspace{-0.015\linewidth}

%-- minipage for the logo at left side --%
\begin{minipage}[t]{0.17\linewidth} % logo
\vspace{1.3cm}

\begin{adjustwidth}{-0.3cm}{-0cm}
\includegraphics[scale=0.35]{\figpdfdir StanCon_logo.png}
\end{adjustwidth}

\end{minipage} % no empty line before the next begin
%-- minipage for title and authors --%
\begin{minipage}[t]{0.76\linewidth}
\vspace{0pt}

\centering
{\renewcommand{\baselinestretch}{0.6} 
\textcolor{aaltoGreen}{\huge{\bfseries{\textsf{Practical implementation of Hilbert space reduced-rank approximate Gaussian processes} }}}
\par}

\vspace{0.015\linewidth}

\normalsize{\textsf{\bfseries{Gabriel Riutort-Mayol$^{1}$ \;\;\; Paul-Christian Bürkner$^2$ \;\;\; Michael Riis Andersen$^{3}$ \;\;\; Arno Solin$^{2}$ \;\;\; Aki Vehtari$^{2}$}}}\\[2mm]

%\setstretch{1.2}
\textcolor{aaltoGray}{\textsf{\bfseries{\small
$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
\\
$^2$ Department of Computer Science, Aalto University, Finland
\\
$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
\\[-2mm]
\footnotesize Contact: gabriuma@gmail.com, aki.vehtari@aalto.fi}}}

\end{minipage}
%-- minipage for the logo at rigth side --%
\begin{minipage}[t]{0.06\linewidth}
\vspace{0.7cm}

%\begin{adjustwidth}{-1cm}{-0cm}
{\centering \includegraphics[scale=0.8]{\figpdfdir aaltologo}}\\[0.5cm]
\includegraphics[scale=0.7]{\figpdfdir marca_UPV_secundaria_negro300.png}
%\end{adjustwidth}
\end{minipage}

\end{minipage}

% Space according to the visual identity guidelines...
\vspace{0.02\linewidth}

% Centering helps in placement
\centering

% Default font for text
\sffamily
%\small % Text size for a1 posters
\footnotesize % Text size for a1 posters

%------------ minipage for the main text ------------%
\begin{minipage}{1\linewidth} 
%------------ minipage for first column ------------%
%\hspace{0.001\textwidth}
\centering
\begin{minipage}[t]{0.48\linewidth}
\vspace{0pt}
\setlength{\parindent}{10mm} % Paragraph indent

\vspace{-0.5cm}
This poster is based on the pre-print publication at \url{https://arxiv.org/abs/2004.11408}

\vspace{-1cm}
\section{Introduction}
\vspace{-0.2cm}

\begin{itemize}
	\setlength{\itemsep}{3mm}
	\item Gaussian processes (GPs) are powerful non-parametric probabilistic models for stochastic functions.
	\item However, when they are implemented using covariance matrices, the computation cost of evaluating the log posterior density is\, $O(n^3)$\, for\, $n$\, observations, which makes them computationally intractable with moderate or large datasets.
	\item We focus on a novel approach for reduced-rank approximate GPs, based on a basis function approximation via Laplace eigenfunctions
for stationary covariance functions {\color{darkgray} \citep{solin2018hilbert}}. We call this reduced-rank GP as {\color{navyblue} Hilbert space approximate Gaussian process} (HSGP).
	\item This reduced-rank approximation is simple and exhibits an attractive computational complexity due to its linear structure\; $\to$\; Improves sampling efficiency and makes inference considerably faster.

%	\begin{itemize}
%		\setlength{\itemsep}{1mm}
%		\item Improves sampling efficiency and makes inference considerably faster
%		\item Big benefit to be implemented in modular probabilistic programming frameworks
%		\item Easier to be used as latent functions in non-Gaussian observational models
%	\end{itemize}
%\item While {\color{darkgray} \cite{solin2018hilbert}} fully developed the mathematical theory behind this low-rank GP, further work is needed for its practical implementation:

\end{itemize}

\vspace{-5mm}
\hspace{1mm}
\centering  %%blue!20; ligthblue
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize \color{black!80} Our contribution:}\; A performance analysis for the practical implementation of this reduce-rank GP (HSGP), colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.98\textwidth]
	\begin{itemize}\setlength\itemsep{2mm}
	\item {\color{navyblue} Discover the relationships} among the key factors of the method such as the number of basis functions, domain of the prediction space, and wiggliness of the function to be learned.
	\item {\color{navyblue} Practical recommendations} for choosing the values of these key factors based on these relationships. %to improve performance and save computation time
	\item {\color{navyblue} Diagnosis} for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
	\end{itemize}
\end{tcolorbox}

%\subsection{Our contributions}
%
%\begin{itemize}
%	\setlength{\itemsep}{4mm}
%	\item A detailed analysis of the performance and implementation of the method in relation to key factors such as the number of basis functions, domain of the prediction space, and wiggliness of the latent function.
%	\item We provide recommendations for choosing the values of these key factors based on the recognized relationships, which make it easier for users to improve approximation accuracy and computational performance. 
%	\item We also propose diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
%\end{itemize}
%
%\subsection{Drawback of the method}
%
%\begin{itemize}
%	\setlength{\itemsep}{4mm}
%	\item The number of basis function in multi-dimensional input spaces grows exponentially with dimensions, which makes looking at the recommendations and the diagnosis tool proposed in this work essential to avoid excessive computation time in multivariate cases. 
%	\item In high dimensional cases, this method may still be used for low-dimensional components in an additive modeling scheme as complexity is linear with the number of additive components.
%\end{itemize}

\vspace{-0.7cm}
\section{Gaussian process as a prior for latent functions}
\vspace{-0.2cm}

\begin{itemize}\setlength\itemsep{3mm}
\item GP is a {\color{navyblue} stochastic process} which defines the distribution over a collection of random variables, $\left\lbrace  f(\bm{x}): \bm{x} \in \mathcal{X}\right\rbrace$ of a function\, $f:{\rm I\!R}^D \to {\rm I\!R}$\, and some input domain\, $\mathcal{X} \subset {\rm I\!R}^D$:
%
\begin{align*}
f \sim \GP(m(\bm{x}), k(\bm{x}, \bm{x}'))
\end{align*}
%
\item The {\color{navyblue}mean}\, $m: {\rm I\!R}^D \to {\rm I\!R}$\, and {\color{navyblue}covariance function}\, $k: {\rm I\!R}^D \times {\rm I\!R}^D \to {\rm I\!R}$\, completely characterize the GP prior and control the a priori behavior of\, $f$\, \cite{rasmussen2006gaussian}.

\item Function values\, $\bm{f}=\left\lbrace  f(\bm{x}_1), f(\bm{x}_2), \hdots, f(\bm{x}_n) \right\rbrace$\, following a GP model has the defining property:
%
\begin{align*}
p(\bm{f}) &= \mathcal{N}(\bm{f} \mid \bm{\mu}, K),\\[1mm]
\mu_i &= m(\bm{x}_i),\\
\hspace{27mm} K_{i,j} &=\cov( f(\bm{x}_i),f(\bm{x}_j))=k(\bm{x}_i,\bm{x}_j).
\end{align*}

%\item GP functions $\bm{f}$ can be used as {\color{navyblue} latent functions} for observational models $p(\bm{y}|\bm{f},\phi)$ for noisy observations $\bm{y}$
%
%\item The joint distribution of observations $\bm{y}$ and latent function values $\bm{f}$ is
%%
%\begin{align*}
%p(\bm{y}, \bm{f}) = p(\bm{y}|\bm{f},\phi)p(\bm{f}) = p(\bm{y}|\bm{f},\phi) \, \mathcal{N}\left(\bm{f}|\bm{\mu}, \bm{K}\right),
%\end{align*}
%where $\phi$ represents the parameters of a specific observation model.
%\end{comment}
%
%\begin{comment}
%\item A covariance function is said to be {\color{navyblue} stationary} (invariant to translations) if 
%%
%\begin{align*}
%	k(\bm{x},\bm{x}')= {\color{navyblue} k(\bm{x}-\bm{x}')}= k(\bm{\tau}), \;\; \bm{\tau} \in {\rm I\!R}^D
%\end{align*}

%\item The most widely used family of covariance functions is the {\color{navyblue} Mat\'ern class} of stationary covariance functions.

%\item Figure:  Function samples from different covariance function from the {\color{navyblue} Mat\'ern class} of stationary covariance functions:
%
%\begin{tabular}{c c c c}
%\scriptsize Squared exponential & \scriptsize Matern52 & \scriptsize Matern32 & \scriptsize Matern12\\[-1mm]
%%
%\includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig1.pdf} & \includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig2.pdf} & \includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig3.pdf} & \includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig4.pdf}
%\end{tabular}
%
% {\scriptsize *These images belong to Nicolas Durrande, \textit{Course presentation in Gaussian Process Summer School 2019}, University of Sheffield.}

\item The covariance function encodes our prior assumptions about the variation of the function and defines the {\color{navyblue} correlation structure} of function values.

\item The main parameter of covariance functions from the Matern class is the {\color{navyblue} length-scale}\, $\ell > 0$, which controls the {\color{navyblue} smoothness/wigglyness} of the function samples.

\item Function samples {\scriptsize(from a squared exponential covariance function)} for different lengthscale\, $\ell$:\:

\hspace{45mm} $\ell=0.1$ \hspace{63mm} $\ell=0.3$ \hspace{64mm} $\ell=0.6$\\
\includegraphics[scale=0.57, trim = 0mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_1.pdf}
\includegraphics[scale=0.57, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_2.pdf}
\includegraphics[scale=0.57, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_3.pdf}

\end{itemize}

\vspace{-0.6cm}
\section{Hilbert space approximate Gaussian process (HSGP)}
\vspace{-0.2cm}

\begin{itemize}\setlength\itemsep{3mm}

\item The covariance function is represented as a {\color{navyblue} infinite series expansion} of eigenfunctions\, $\phi_j(x)$\, scaled by the spectral density\, $S_{\theta}(\sqrt{\lambda_j})$\, {\color{darkgray} \citep{solin2018hilbert}}, and {\color{navyblue} approximated by the first $m$ terms}:
%
\begin{align*}
k(x,x') = \sum_{j=1}^m S(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'), \hspace{10mm} x,x' \in [-L,L]
\end{align*}

\item where $S(\cdot)$ is the {\color{navyblue}spectral density} of the covariance function $k(\cdot,\cdot)$.

\begin{remark}
Any stationary covariance function\, $k:\bm{x} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, and its {\color{navyblue} spectral density}\, $S:\bm{w} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, are Fourier duals, where\, $\bm{w}$\, is a frequency. So, stationary covariance functions can be represented in terms of their spectral density functions.
\end{remark}

\item We choose the {\color{navyblue} eigenvalues}\, $\lambda_j$\, and {\color{navyblue} eigenfunctions}\, $\phi_j(x)$\, of the Laplacian operator $-\nabla^2$:
%
\begin{align*}
\left. \begin{array}{cc}
 -\nabla^2 \phi_j(x) = \lambda_j \phi_j(x), \hspace{0.8cm}  x \in [-L,L] \\
\hspace{6mm} \phi_j(x) = 0, \hspace{2cm} x \notin [-L,L] 
\end{array} \right\} \; \Rightarrow \hspace{5mm} \lambda_j=\left(\frac{j\pi}{2L}\right)^2 \hspace{5mm} \text{and} \hspace{5mm}
\phi_j(x)=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right)
\end{align*}

%\begin{minipage}{0.35\textwidth}
%%
%\begin{align*}
%-\nabla^2 \phi_j(x)&=\lambda_j \phi_j(x), \hspace{0.8cm}  x \in [-L,L] \\ 
%\hspace{6mm}\phi_j(x)&=0, \hspace{1.6cm} x \notin [-L,L]
%\end{align*}
%
%\end{minipage}
%\begin{minipage}{0.65\textwidth}
%%
%\begin{gather*}
%\text{given by:} \hspace{5mm} \lambda_j=\left(\frac{j\pi}{2L}\right)^2 \hspace{5mm} \text{and} \hspace{5mm}
%\phi_j(x)=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right)
%\end{gather*}
%\end{minipage}

\item $L > 0$\, represents the {\color{navyblue} boundaries} where this representation is valid.

\item $S(\sqrt{\lambda_j})$\, is monotonically decreasing with\, $j$, and decays rapidly to zero, and this rate of decay depends on the lenghtscale of\, $k$\, (smoothness of the function to be learned) \; $\rightarrow$ \; This points to {\color{navyblue} a finite number}\, $m$\, {\color{navyblue} of terms for an accurate approximation}, as long as\, $x_i$\, are not too close to the boundaries\, $-L$\, and\, $L$\, of\, $\Omega$.

\item This representation of the covariance function\, $k(x,x')$\, leads to the {\color{navyblue} linear representation} of the function\, $f$:
%
\begin{align*}
f(x) \approx \sum_{j}^m \phi_j(x) \beta_j, \hspace{10mm} \beta_j \sim \mathcal{N}(0,S(\sqrt{\lambda_j})).
\end{align*}
\end{itemize}

\vspace{-0.5cm}
\begin{tcolorbox}[colframe=blue!20, colback=white, title=\scriptsize Properties of the HSGP method, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
\begin{itemize}\setlength\itemsep{2mm}
\item[+] The basis functions\, $\phi_j(x)$\, do not depend on the hyperparameters, so they just need to be computed once with cost\, $O(m^2n)$.

\item[+]  All dependencies on the hyperparameters is through the variance of the weights\, $\beta_j$. The computational cost of evaluating the log posterior density is\, $O(mn+m)$, which contrasts with\, $O(n^3)$\, of exact GPs.

%\item[+] $f(x)$ is naturally in the {\color{navyblue} non-centered parameterization form} with independent prior distribution on $\beta_j$ which makes posterior inference easier. \citep[see, e.g., ][]{Betancourt+Girolami:2019}

\item[+] The parameter posterior\, $p(\bm{\beta}|\bm{y})$\, is\, $m$-dimensional ($m\ll n$), which makes posterior inference easier.

\item[+] Reduced memory requirements of automatic differentiation methods used in Stan  software.

\item[+] Trade-off between accuracy and computation can also be carried out.

\item[-] The number of multivariate basis functions,\, $m^{\ast} = \prod_{d=1}^{D} m_d$,\, grows exponentially with respect to\, $D$.

\item[-] In practice,\, $D>3$\, starts to be too computationally demanding, even for smooth functions.

\end{itemize}
\end{tcolorbox}

\end{minipage}% end minipage first column
\hspace{0.015\linewidth} 
%------------ minipage for second column ------------%
\begin{minipage}[t]{0.48\textwidth}
\vspace{0pt}
\setlength{\parindent}{10mm}

\vspace{0.3cm}
\section{Performance analysis of HSGP}
\vspace{-0.2cm}

\begin{itemize}\setlength\itemsep{2mm}
\item Determine the minimum\, $m$\, ({\color{navyblue} basis functions}) and valid\, $L$\, ({\color{navyblue} boundary}) to obtain accurate approximations, as a function of the {\color{navyblue} lengthscale}\, $\ell$.

%\item How many basis functions $m$ are needed for a good approximation?
%
%\item Which is the effect of the boundary box $L$ on the approximation?
%
%\item Which are valid values for the boundary box $L$?

%\item How $m$ and $L$ relate to each other and to the wigglyness $\ell$ of the function to be learned, in terms of an accurate approximation?
\end{itemize}

\vspace{0.2cm}
\begin{remark}
\pbox{10cm}{Boundary box\, $\Omega \in [-L,L]$ \\ Desired input domain\, $\Phi \in [-S,S]$}\;\;  $\to$\;\;  $L=c \cdot S$, \hspace{3mm} $c \geqslant 1$, \hspace{1mm} $c \equiv$ {\color{navyblue} Boundary factor}
\end{remark}
\vspace{0.2cm}

\begin{enumerate}
\item \fcolorbox{aaltoLightGreen!10}{aaltoLightGreen!10}{\textsc{\textbf{Analysis of the effects of $m$ and $L$ on the approximation}}}

\vspace{-2mm}
\hspace{-1cm}
\begin{minipage}{0.70\textwidth}
{\scriptsize \hspace{20mm} Covariance \hspace{35mm} Mean function \hspace{25mm} Standard deviation}\\[-6mm]
%
\includegraphics[width=0.32\textwidth, trim = 3mm 26mm 0mm 0mm, clip]{\figpdfdir ch5_fig1_Cov_J.pdf}
\includegraphics[width=0.32\textwidth, trim = 2mm 26mm 0mm 0mm, clip]{\figpdfdir ch5_fig1_Post_J.pdf}
\includegraphics[width=0.32\textwidth, trim = 3mm 26mm 0mm 0mm, clip]{\figpdfdir ch5_fig1_Sigma_J.pdf}\\
\includegraphics[width=0.32\textwidth, trim = 3mm 0mm 0mm 17mm, clip]{\figpdfdir ch5_fig2_Cov_L.pdf}
\includegraphics[width=0.32\textwidth, trim = 2mm 0mm 0mm 17mm, clip]{\figpdfdir ch5_fig2_Post_L.pdf}
\includegraphics[width=0.32\textwidth, trim = 3mm 0mm 0mm 17mm, clip]{\figpdfdir ch5_fig2_Sigma_L.pdf}

\end{minipage}
\begin{minipage}{0.29\textwidth}

\includegraphics[width=1\linewidth, trim = 0mm 8mm 5mm 0mm, clip]{\figpdfdir performance_fig3_vposter.pdf}\\
For this example, a optimal choice would be\, $m=10$\, and\, $c=1.5$.\, And a bit more conservative choice would\, $m=15$\, and\, $c=1.5$.\\
\end{minipage}

\vspace{-3mm}
\item \fcolorbox{aaltoLightGreen!10}{aaltoLightGreen!10}{\textsc{\textbf{Model gathering the relationships among $m$, $L$ and $\ell$ in terms of an accurate approximation}}}
\end{enumerate}

\vspace{3mm}
\hspace{-1cm}
\begin{minipage}{0.35\textwidth}
\hspace{0.25\textwidth} \scriptsize Squared exponential kernel\\
% \hspace{0.3\textwidth}  \scriptsize Mattern($\nu$=3/2) kernel\\
\includegraphics[width=0.98\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf} \;\;
%\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}

\end{minipage}
\begin{minipage}{0.65\textwidth}

\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize This model says...}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.975\textwidth]
\begin{itemize}\setlength\itemsep{1mm}
\item For a given\, $m$\, and\, $c$,\, the minimum\, $\ell$\, that can be accurately approximated.
\item For a given\, $\ell$\, and\, $c$,\, the minimum\, $m$\, required for an accurate approximation.
\item For a given\, $\ell$\, and\, $m$,\, the highest\, $c$\, allowed for an accurate approximation.
\item The minimum valid\, $c$\, for a given\, $\ell$.
\end{itemize}
\end{tcolorbox}
\end{minipage}

\vspace{-3mm}
{\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize This model serves as a \textsc{\textbf{diagnosis tool}} of the approximation:}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.98\textwidth]
\begin{remark}
As the lengthscale is the parameter that ultimately characterizes the non-linearity of the posterior functions, we assume we can look at the {\color{navyblue} estimate lengthscale}\, $\hat{\ell}$\, to diagnose the approximation.
\end{remark}

\begin{itemize}\setlength\itemsep{1mm}
\item Given $m$ and $c$

\item If\, $\hat{\ell}$\, {\color{navyblue} is below} the minimum\, $\ell$\, that can be accurately approximated (for those given\, $m$\, and\, $c$), then the approximation {\color{navyblue} is not sufficiently accurate}, and\, $m$\, must be increased or\, $c$\, decreased.

\item If\, $\hat{\ell}$\, {\color{navyblue} is above} the minimum\, $\ell$,\, then the approximation should be {\color{navyblue} sufficiently accurate}.

\item This can be done iteratively.

%\item $c$\, can not be decreased as much as desired because it is restricted to\, $\ell$.
\end{itemize}
\end{tcolorbox}
}

\vspace{-0.7cm}
\section{Univariate synthetic case study}
\vspace{-0.2cm}

%\hspace{-1.1cm}
\begin{minipage}{0.33\textwidth}
\begin{itemize}\setlength\itemsep{4mm}
\item \textit{\scriptsize Samples draws from}:%\\ $x \in [-1,1] \subset {\rm I\!R}$,
\\ GP Mattern($\nu$=3/2, $\ell$=0.15, $\alpha$=1)\\ + $\bm{\epsilon} \sim \mathcal{N}(0, 0.2^2)$

\item \textit{\scriptsize Observation model}: \pbox{6cm}{$\bm{y} = \bm{f} + \bm{\epsilon}$ \\
$\bm{\epsilon} \sim \mathcal{N}(\bm{0}, \sigma^2 \bm{I})$}

\item \pbox{8cm}{\textit{\scriptsize Exact GP prior for $\bm{f}$}:\\ $f(x) \sim \GP(0, k_{\nicefrac{3}{2}}(x, x'))$}

\item \pbox{11cm}{\textit{\scriptsize HSGP prior for $\bm{f}$}:\\ $f(x) \approx \sum_{j=1}^m \phi_j(x) \beta_j$,\\ 
$\beta_j \sim \mathcal{N}(0,S_{\nicefrac{3}{2}}(\sqrt{\lambda_j}))$}
%\\ $m=60$, $c=1.5$}
\end{itemize}

\end{minipage}
\begin{minipage}{0.67\textwidth}
\centering
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir ch5_fig10_Posteriors_exI_vposter.pdf}
\end{minipage}

\vspace{6mm}
\hspace{-7mm} \textsc{\textbf{\small Diagnosis}}:\\[-4mm]

\begin{minipage}[t]{0.49\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 1}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=20$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.09$}

\vspace{0.3cm}
%{\hspace{2.2cm} \scriptsize Error}\\[-3mm]
\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig2_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_1_1_vposter.pdf}

\end{minipage}

\end{minipage}
\begin{minipage}[t]{0.49\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 2}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=30$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.12$}

\vspace{0.3cm}
\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig3_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_2_1_vposter.pdf}

\end{minipage}

\end{minipage}

\vspace{5mm}
\begin{minipage}[t]{0.49\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 3}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=40$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.14$}

\vspace{0.3cm}
\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig4_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_3_1_vposter.pdf}

\end{minipage}

\end{minipage}
\begin{minipage}[t]{0.49\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 4}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=60$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.14$}

\vspace{0.3cm}
\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig5_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  {\tiny Mattern($\nu$=3/2) kernel}\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_4_1_vposter.pdf}

\end{minipage}

\end{minipage}

% bibliography
\vspace*{3mm}
{\centering\textbf{References}}
\renewcommand{\bibsection}{}
{\tiny % % Bibliography text size begins...
\bibliographystyle{unsrt}
\bibliography{references}
} % <- bibliography text size ends

\end{minipage} % End minipage second column
\hspace{0.02\linewidth} 
\end{minipage} % end minipage main text

%%------------ minipage for third column ------------%
%\begin{minipage}[t]{0.48\textwidth} % minipage for the latter two columns
%\setlength{\parindent}{10mm} % Paragraph indent
%
%\end{minipage} % End minipage third column
%\hspace{0.03\linewidth}

%\vspace*{0.03\linewidth} % Increase the bottom margins

\end{document}

### 2020-09-10 Paper/StanCon/poster/HSGP_StanCon_poster_landscape.tex
add StanCon poster folder

--- 2020-09-10 Paper/StanCon/poster/HSGP_StanCon_poster.tex
@@
-\documentclass[portrait,a1,final]{a0poster} % a0poster class sets the paper size
+\documentclass[landscape,a1,final]{a0poster} % a0poster class sets the paper size
 \usepackage[utf8]{inputenc}
@@
 %\usepackage{sectsty}
-\usepackage{natbib}
 \usepackage{subcaption}
@@
 \usepackage{setspace}
+\usepackage{bibentry}
+
+\usepackage{natbib}
+\bibpunct{(}{)}{;}{a}{,}{,}
 
@@
 
-\titleformat{\subsection}{\centering\normalsize\sc\bfseries\sffamily\color{aaltoPurple}}{\textcolor{aaltoPurple}{\thesubsection}}{1em}{} % Text size for a1 posters
+%\newcommand{\caja_sub}[1]{\fcolorbox{aaltoLightGreen!10}{aaltoLightGreen!10}
+%{\parbox[c][8mm][c]{0.95\textwidth}
+%{\centering\normalsize\sc\bfseries\sffamily\color{black} \thesubsection. #1}}}
+%
+%\titleformat{\subsection}[block]
+%{\normalfont\filcenter\color{aaltoGray}}{}{1em}{\caja_sub}
+
+\titleformat{\subsection}{\normalsize\sc\bfseries\sffamily\color{aaltoPurple}}{\textcolor{black}{\thesubsection}}{1em}{}
 
@@
 %------------ minipage for the logo & title ------------%
-\begin{minipage}[t]{0.98\linewidth} 
+\begin{minipage}[t]{1\linewidth} 
 \vspace{0pt} % A trick to align the parallel minipages on top
@@
 %-- minipage for the logo at left side --%
-\begin{minipage}[t]{0.17\linewidth} % logo
-\vspace{1.3cm}
+\begin{minipage}[t]{0.14\linewidth} % logo
+\vspace{0.5cm}
 
-\begin{adjustwidth}{-0.3cm}{-0cm}
+%\begin{adjustwidth}{-1cm}{-0cm}
 \includegraphics[scale=0.35]{\figpdfdir StanCon_logo.png}
-\end{adjustwidth}
+%\end{adjustwidth}
 
@@
 %-- minipage for title and authors --%
-\begin{minipage}[t]{0.76\linewidth}
-\vspace{0pt}
+\begin{minipage}[t]{0.50\linewidth}
+\vspace{1cm}
 
@@
 
-\vspace{0.015\linewidth}
-
-\normalsize{\textsf{\bfseries{Gabriel Riutort-Mayol$^{1}$ \;\;\; Paul-Christian Bürkner$^2$ \;\;\; Michael Riis Andersen$^{3}$ \;\;\; Arno Solin$^{2}$ \;\;\; Aki Vehtari$^{2}$}}}\\[2mm]
-
-%\setstretch{1.2}
-\textcolor{aaltoGray}{\textsf{\bfseries{\small
-$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
-\\
-$^2$ Department of Computer Science, Aalto University, Finland
-\\
-$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
-\\[-2mm]
-\footnotesize Contact: gabriuma@gmail.com, aki.vehtari@aalto.fi}}}
+\vspace{5mm}
+\hspace{-1cm}
+\normalsize{\textsf{\bfseries{Gabriel Riutort-Mayol$^{1}$ \;\; Paul-Christian Bürkner$^2$ \;\; Michael Riis Andersen$^{3}$ \;\; Arno Solin$^{2}$ \;\; Aki Vehtari$^{2}$}}}
 
@@
 %-- minipage for the logo at rigth side --%
-\begin{minipage}[t]{0.06\linewidth}
-\vspace{0.7cm}
+\begin{minipage}[t]{0.22\linewidth}
+
+\vspace{-2mm}
+\textcolor{aaltoGray}{\textsf{\bfseries{ \footnotesize
+\begin{itemize}
+	\setlength{\itemsep}{0mm}
+	\item[] $^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain
+\item[] $^2$ Department of Computer Science, Aalto University, Finland
+\item[] $^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
+\vspace{2mm}
+\item[] {\small Contact: gabriuma@gmail.com,\, aki.vehtari@aalto.fi}
+\end{itemize}
+}}}
+
+\end{minipage}
+%-- minipage for the logo at rigth side --%
+\hspace{15mm}
+\begin{minipage}[t]{0.09\linewidth}
+\vspace{10mm}
 
 %\begin{adjustwidth}{-1cm}{-0cm}
-{\centering \includegraphics[scale=0.8]{\figpdfdir aaltologo}}\\[0.5cm]
-\includegraphics[scale=0.7]{\figpdfdir marca_UPV_secundaria_negro300.png}
+\includegraphics[scale=0.8]{\figpdfdir aaltologo}
+\raisebox{-5mm}{\includegraphics[scale=0.7]{\figpdfdir marca_UPV_secundaria_negro300.png}}
 %\end{adjustwidth}
@@
 
+% Default font for text
+\sffamily
+\normalsize % Text size
+%\small % Text size
+%\footnotesize % Text size
+
+\vspace{1cm}
+This poster is based on the pre-print publication at {\url{https://arxiv.org/abs/2004.11408}}
+
 % Space according to the visual identity guidelines...
-\vspace{0.02\linewidth}
+%\vspace{0.02\linewidth}
 
@@
 \centering
-
-% Default font for text
-\sffamily
-%\small % Text size for a1 posters
-\footnotesize % Text size for a1 posters
 
@@
 \centering
-\begin{minipage}[t]{0.48\linewidth}
+\hspace{3pt}
+\begin{minipage}[t]{0.30\linewidth}
 \vspace{0pt}
@@
 
-\vspace{-0.5cm}
-This poster is based on the pre-print publication at \url{https://arxiv.org/abs/2004.11408}
-
-\vspace{-1cm}
+\vspace{0.8cm}
 \section{Introduction}
@@
 	\setlength{\itemsep}{3mm}
-	\item Gaussian processes (GPs) are powerful non-parametric probabilistic models for stochastic functions.
-	\item However, when they are implemented using covariance matrices, the computation cost of evaluating the log posterior density is\, $O(n^3)$\, for\, $n$\, observations, which makes them computationally intractable with moderate or large datasets.
+%	\item Gaussian processes (GPs) are powerful non-parametric probabilistic models for stochastic functions.
+	\item When Gaussian processes are implemented using covariance matrices, the computation cost of evaluating the log posterior density is\, $O(n^3)$\, for\, $n$\, observations, which makes them computationally intractable with moderate or large datasets.
 	\item We focus on a novel approach for reduced-rank approximate GPs, based on a basis function approximation via Laplace eigenfunctions
-for stationary covariance functions {\color{darkgray} \citep{solin2018hilbert}}. We call this reduced-rank GP as {\color{navyblue} Hilbert space approximate Gaussian process} (HSGP).
-	\item This reduced-rank approximation is simple and exhibits an attractive computational complexity due to its linear structure\; $\to$\; Improves sampling efficiency and makes inference considerably faster.
-
-%	\begin{itemize}
-%		\setlength{\itemsep}{1mm}
-%		\item Improves sampling efficiency and makes inference considerably faster
-%		\item Big benefit to be implemented in modular probabilistic programming frameworks
-%		\item Easier to be used as latent functions in non-Gaussian observational models
-%	\end{itemize}
-%\item While {\color{darkgray} \cite{solin2018hilbert}} fully developed the mathematical theory behind this low-rank GP, further work is needed for its practical implementation:
+for stationary covariance functions \href{https://arxiv.org/abs/1401.5508}{\citep{solin2018hilbert}}.
+We call this reduced-rank GP as {\color{navyblue} Hilbert space approximate Gaussian process} (HSGP).
+	\item This HSGP is simple in complexity due to its linear structure\; $\to$\; Improves sampling efficiency and makes inference considerably faster.
 
@@
 
-\vspace{-5mm}
-\hspace{1mm}
+\vspace{-7mm}
 \centering  %%blue!20; ligthblue
-\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize \color{black!80} Our contribution:}\; A performance analysis for the practical implementation of this reduce-rank GP (HSGP), colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.98\textwidth]
+\hspace{-2mm}
+\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small \color{black!80} Our contribution:} Performance analysis for practical implementation of this HSGP, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=1\textwidth]
 	\begin{itemize}\setlength\itemsep{2mm}
@@
 
-%\subsection{Our contributions}
-%
-%\begin{itemize}
-%	\setlength{\itemsep}{4mm}
-%	\item A detailed analysis of the performance and implementation of the method in relation to key factors such as the number of basis functions, domain of the prediction space, and wiggliness of the latent function.
-%	\item We provide recommendations for choosing the values of these key factors based on the recognized relationships, which make it easier for users to improve approximation accuracy and computational performance.
-%	\item We also propose diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
-%\end{itemize}
-%
-%\subsection{Drawback of the method}
-%
-%\begin{itemize}
-%	\setlength{\itemsep}{4mm}
-%	\item The number of basis function in multi-dimensional input spaces grows exponentially with dimensions, which makes looking at the recommendations and the diagnosis tool proposed in this work essential to avoid excessive computation time in multivariate cases.
-%	\item In high dimensional cases, this method may still be used for low-dimensional components in an additive modeling scheme as complexity is linear with the number of additive components.
-%\end{itemize}
-
 \vspace{-0.7cm}
 \section{Gaussian process as a prior for latent functions}
-\vspace{-0.2cm}
 
 \begin{itemize}\setlength\itemsep{3mm}
-\item GP is a {\color{navyblue} stochastic process} which defines the distribution over a collection of random variables, $\left\lbrace  f(\bm{x}): \bm{x} \in \mathcal{X}\right\rbrace$ of a function\, $f:{\rm I\!R}^D \to {\rm I\!R}$\, and some input domain\, $\mathcal{X} \subset {\rm I\!R}^D$:
-%
-\begin{align*}
-f \sim \GP(m(\bm{x}), k(\bm{x}, \bm{x}'))
-\end{align*}
-%
-\item The {\color{navyblue}mean}\, $m: {\rm I\!R}^D \to {\rm I\!R}$\, and {\color{navyblue}covariance function}\, $k: {\rm I\!R}^D \times {\rm I\!R}^D \to {\rm I\!R}$\, completely characterize the GP prior and control the a priori behavior of\, $f$\, \cite{rasmussen2006gaussian}.
-
-\item Function values\, $\bm{f}=\left\lbrace  f(\bm{x}_1), f(\bm{x}_2), \hdots, f(\bm{x}_n) \right\rbrace$\, following a GP model has the defining property:
-%
+\item Function values\, $\bm{f}=\left\lbrace  f(\bm{x}_1), f(\bm{x}_2), \hdots, f(\bm{x}_n) \right\rbrace$,\, following a GP model has the defining property:
+%\vspace{-4mm}
 \begin{align*}
@@
 
-%\item GP functions $\bm{f}$ can be used as {\color{navyblue} latent functions} for observational models $p(\bm{y}|\bm{f},\phi)$ for noisy observations $\bm{y}$
-%
-%\item The joint distribution of observations $\bm{y}$ and latent function values $\bm{f}$ is
-%%
-%\begin{align*}
-%p(\bm{y}, \bm{f}) = p(\bm{y}|\bm{f},\phi)p(\bm{f}) = p(\bm{y}|\bm{f},\phi) \, \mathcal{N}\left(\bm{f}|\bm{\mu}, \bm{K}\right),
-%\end{align*}
-%where $\phi$ represents the parameters of a specific observation model.
-%\end{comment}
-%
-%\begin{comment}
-%\item A covariance function is said to be {\color{navyblue} stationary} (invariant to translations) if 
-%%
-%\begin{align*}
-%	k(\bm{x},\bm{x}')= {\color{navyblue} k(\bm{x}-\bm{x}')}= k(\bm{\tau}), \;\; \bm{\tau} \in {\rm I\!R}^D
-%\end{align*}
-
-%\item The most widely used family of covariance functions is the {\color{navyblue} Mat\'ern class} of stationary covariance functions.
-
-%\item Figure:  Function samples from different covariance function from the {\color{navyblue} Mat\'ern class} of stationary covariance functions:
-%
-%\begin{tabular}{c c c c}
-%\scriptsize Squared exponential & \scriptsize Matern52 & \scriptsize Matern32 & \scriptsize Matern12\\[-1mm]
-%%
-%\includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig1.pdf} & \includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig2.pdf} & \includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig3.pdf} & \includegraphics[width=0.23\linewidth, trim = 0mm 0mm 0mm 5mm, clip]{\figpdfdir covfun_fig4.pdf}
-%\end{tabular}
-%
-% {\scriptsize *These images belong to Nicolas Durrande, \textit{Course presentation in Gaussian Process Summer School 2019}, University of Sheffield.}
-
-\item The covariance function encodes our prior assumptions about the variation of the function and defines the {\color{navyblue} correlation structure} of function values.
+\item The covariance function encodes our prior assumptions about the variation of the function and defines the {\color{navyblue} correlation structure} of function values \citep{rasmussen2006gaussian}.
 
@@
 
-\item Function samples {\scriptsize(from a squared exponential covariance function)} for different lengthscale\, $\ell$:\:
-
-\hspace{45mm} $\ell=0.1$ \hspace{63mm} $\ell=0.3$ \hspace{64mm} $\ell=0.6$\\
-\includegraphics[scale=0.57, trim = 0mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_1.pdf}
-\includegraphics[scale=0.57, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_2.pdf}
-\includegraphics[scale=0.57, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_3.pdf}
-
+%\item Function samples (from a squared exponential covariance function) for different lengthscale\, $\ell$:\: 
 \end{itemize}
 
-\vspace{-0.6cm}
+\begin{flushleft}
+\hspace{40mm} $\ell=0.1$ \hspace{48mm} $\ell=0.3$ \hspace{47mm} $\ell=0.6$
+\end{flushleft}
+\vspace{-3mm}
+\includegraphics[scale=0.45, trim = 0mm 0mm 7mm 9mm, clip]{\figpdfdir func_samples_1.pdf}
+\includegraphics[scale=0.45, trim = 34mm 0mm 7mm 9mm, clip]{\figpdfdir func_samples_2.pdf}
+\includegraphics[scale=0.45, trim = 34mm 0mm 7mm 9mm, clip]{\figpdfdir func_samples_3.pdf}
+
+\end{minipage}% end minipage first column
+\hspace{0.015\linewidth} 
+%------------ minipage for second column ------------%
+\begin{minipage}[t]{0.31\textwidth}
+\vspace{0pt}
+\setlength{\parindent}{10mm}
+
+\vspace{0.3cm}
 \section{Hilbert space approximate Gaussian process (HSGP)}
@@
 
-\item where $S(\cdot)$ is the {\color{navyblue}spectral density} of the covariance function $k(\cdot,\cdot)$.
+\item where $S(\cdot)$ is the {\color{navyblue}spectral density function} of the covariance function $k(\cdot,\cdot)$.
 
-\begin{remark}
-Any stationary covariance function\, $k:\bm{x} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, and its {\color{navyblue} spectral density}\, $S:\bm{w} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, are Fourier duals, where\, $\bm{w}$\, is a frequency.
+%\begin{remark}
+%Any stationary covariance function\, $k:\bm{x} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, and its {\color{navyblue} spectral density}\, $S:\bm{w} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, are Fourier duals, where\, $\bm{w}$\, is a frequency.
 So, stationary covariance functions can be represented in terms of their spectral density functions.
-\end{remark}
+%\end{remark}
 
-\item We choose the {\color{navyblue} eigenvalues}\, $\lambda_j$\, and {\color{navyblue} eigenfunctions}\, $\phi_j(x)$\, of the Laplacian operator $-\nabla^2$:
+\item We choose the {\color{navyblue} eigenvalues}\, $\lambda_j$\, and {\color{navyblue} eigenfunctions}\, $\phi_j(x)$\, of the Laplacian $-\nabla^2$:
 %
@@
 \hspace{6mm} \phi_j(x) = 0, \hspace{2cm} x \notin [-L,L] 
-\end{array} \right\} \; \Rightarrow \hspace{5mm} \lambda_j=\left(\frac{j\pi}{2L}\right)^2 \hspace{5mm} \text{and} \hspace{5mm}
+\end{array} \right\} \Rightarrow \begin{array}{ll}
+ \lambda_j=\left(\frac{j\pi}{2L}\right)^2 \\
 \phi_j(x)=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right) 
+\end{array}
 \end{align*}
-
-%\begin{minipage}{0.35\textwidth}
-%%
-%\begin{align*}
-%-\nabla^2 \phi_j(x)&=\lambda_j \phi_j(x), \hspace{0.8cm}  x \in [-L,L] \\ 
-%\hspace{6mm}\phi_j(x)&=0, \hspace{1.6cm} x \notin [-L,L]
-%\end{align*}
-%
-%\end{minipage}
-%\begin{minipage}{0.65\textwidth}
-%%
-%\begin{gather*}
-%\text{given by:} \hspace{5mm} \lambda_j=\left(\frac{j\pi}{2L}\right)^2 \hspace{5mm} \text{and} \hspace{5mm}
-%\phi_j(x)=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right)
-%\end{gather*}
-%\end{minipage}
 
@@
 
-\item $S(\sqrt{\lambda_j})$\, is monotonically decreasing with\, $j$, and decays rapidly to zero, and this rate of decay depends on the lenghtscale of\, $k$\, (smoothness of the function to be learned) \; $\rightarrow$ \; This points to {\color{navyblue} a finite number}\, $m$\, {\color{navyblue} of terms for an accurate approximation}, as long as\, $x_i$\, are not too close to the boundaries\, $-L$\, and\, $L$\, of\, $\Omega$.
+%\item $S(\sqrt{\lambda_j})$\, is monotonically decreasing with\, $j$, and decays rapidly to zero, and this rate of decay depends on the lenghtscale of\, $k$\, (smoothness of the function to be learned) \; $\rightarrow$ \; This points to {\color{navyblue} a finite number}\, $m$\, {\color{navyblue} of terms for an accurate approximation}, as long as\, $x_i$\, are not too close to the boundaries\, $-L$\, and\, $L$\, of\, $\Omega$.
 
@@
 \vspace{-0.5cm}
-\begin{tcolorbox}[colframe=blue!20, colback=white, title=\scriptsize Properties of the HSGP method, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
-\begin{itemize}\setlength\itemsep{2mm}
+\begin{tcolorbox}[colframe=blue!20, colback=white, title=\small \color{black!80} Properties of the HSGP model, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
+\begin{itemize}\setlength\itemsep{1mm}
 \item[+] The basis functions\, $\phi_j(x)$\, do not depend on the hyperparameters, so they just need to be computed once with cost\, $O(m^2n)$.
@@
 
-\item[+] Reduced memory requirements of automatic differentiation methods used in Stan  software.
+\item[+] Reduced memory requirements of autodiff methods used in Stan software.
 
-\item[+] Trade-off between accuracy and computation can also be carried out.
+%\item[+] Trade-off between accuracy and computation can also be carried out.
 
 \item[-] The number of multivariate basis functions,\, $m^{\ast} = \prod_{d=1}^{D} m_d$,\, grows exponentially with respect to\, $D$.
+In practice,\, $D>3$\, starts to be too computationally demanding, even for smooth functions.
 
-\item[-] In practice,\, $D>3$\, starts to be too computationally demanding, even for smooth functions.
+\item[+] In high dimensional cases, HSGP may still be used for low-dimensional components in an additive modeling scheme as complexity is linear with the number of additive components.
 
@@
 
-\end{minipage}% end minipage first column
-\hspace{0.015\linewidth} 
-%------------ minipage for second column ------------%
-\begin{minipage}[t]{0.48\textwidth}
-\vspace{0pt}
-\setlength{\parindent}{10mm}
-
-\vspace{0.3cm}
+\vspace{-0.5cm}
 \section{Performance analysis of HSGP}
@@
 \begin{itemize}\setlength\itemsep{2mm}
-\item Determine the minimum\, $m$\, ({\color{navyblue} basis functions}) and valid\, $L$\, ({\color{navyblue} boundary}) to obtain accurate approximations, as a function of the {\color{navyblue} lengthscale}\, $\ell$.
+\item Determine the minimum\, $m$\, ({\color{navyblue} basis functions}) and valid\, $L$\, ({\color{navyblue} boundary}) to obtain accurate approximations, as a function of\, $\ell$\, ({\color{navyblue} lengthscale}).
 
-%\item How many basis functions $m$ are needed for a good approximation?
-%
-%\item Which is the effect of the boundary box $L$ on the approximation?
-%
-%\item Which are valid values for the boundary box $L$?
-
-%\item How $m$ and $L$ relate to each other and to the wigglyness $\ell$ of the function to be learned, in terms of an accurate approximation?
 \end{itemize}
 
-\vspace{0.2cm}
-\begin{remark}
-\pbox{10cm}{Boundary box\, $\Omega \in [-L,L]$ \\ Desired input domain\, $\Phi \in [-S,S]$}\;\;  $\to$\;\;  $L=c \cdot S$, \hspace{3mm} $c \geqslant 1$, \hspace{1mm} $c \equiv$ {\color{navyblue} Boundary factor}
-\end{remark}
-\vspace{0.2cm}
+%% bibliography
+%\rule{\textwidth}{1mm}
+%{\centering\textbf{References}}
+%\renewcommand{\bibsection}{}
+%{\footnotesize 
+\bibliographystyle{plainnat} %unsrt
+\nobibliography{references}
+%}
 
-\begin{enumerate}
-\item \fcolorbox{aaltoLightGreen!10}{aaltoLightGreen!10}{\textsc{\textbf{Analysis of the effects of $m$ and $L$ on the approximation}}}
+\end{minipage}% end minipage second column
+\hspace{0.015\linewidth} 
+%------------ minipage for third column ------------%
+\begin{minipage}[t]{0.345\textwidth}
+\vspace{0pt}
+\setlength{\parindent}{10mm}
 
-\vspace{-2mm}
+\vspace{-0.8cm}
 \hspace{-1cm}
-\begin{minipage}{0.70\textwidth}
-{\scriptsize \hspace{20mm} Covariance \hspace{35mm} Mean function \hspace{25mm} Standard deviation}\\[-6mm]
-%
-\includegraphics[width=0.32\textwidth, trim = 3mm 26mm 0mm 0mm, clip]{\figpdfdir ch5_fig1_Cov_J.pdf}
-\includegraphics[width=0.32\textwidth, trim = 2mm 26mm 0mm 0mm, clip]{\figpdfdir ch5_fig1_Post_J.pdf}
-\includegraphics[width=0.32\textwidth, trim = 3mm 26mm 0mm 0mm, clip]{\figpdfdir ch5_fig1_Sigma_J.pdf}\\
-\includegraphics[width=0.32\textwidth, trim = 3mm 0mm 0mm 17mm, clip]{\figpdfdir ch5_fig2_Cov_L.pdf}
-\includegraphics[width=0.32\textwidth, trim = 2mm 0mm 0mm 17mm, clip]{\figpdfdir ch5_fig2_Post_L.pdf}
-\includegraphics[width=0.32\textwidth, trim = 3mm 0mm 0mm 17mm, clip]{\figpdfdir ch5_fig2_Sigma_L.pdf}
+%\subsection{Model of relationships among $m$, $L$ and $\ell$ in terms of an accurate approximation}
+\textsc{\textbf{4.1 Model of relationships among $m$, $L$ and $\ell$ in terms of an accurate approximation}}
 
-\end{minipage}
-\begin{minipage}{0.29\textwidth}
-
-\includegraphics[width=1\linewidth, trim = 0mm 8mm 5mm 0mm, clip]{\figpdfdir performance_fig3_vposter.pdf}\\
-For this example, a optimal choice would be\, $m=10$\, and\, $c=1.5$.\, And a bit more conservative choice would\, $m=15$\, and\, $c=1.5$.\\
-\end{minipage}
-
-\vspace{-3mm}
-\item \fcolorbox{aaltoLightGreen!10}{aaltoLightGreen!10}{\textsc{\textbf{Model gathering the relationships among $m$, $L$ and $\ell$ in terms of an accurate approximation}}}
-\end{enumerate}
-
-\vspace{3mm}
-\hspace{-1cm}
-\begin{minipage}{0.35\textwidth}
-\hspace{0.25\textwidth} \scriptsize Squared exponential kernel\\
+\vspace{4mm}
+\hspace{-1.3cm}
+\begin{minipage}{0.32\textwidth}
+\hspace{0.22\textwidth} \scriptsize Squared exponential kernel\\
 % \hspace{0.3\textwidth}  \scriptsize Mattern($\nu$=3/2) kernel\\
@@
 \end{minipage}
-\begin{minipage}{0.65\textwidth}
+\begin{minipage}{0.67\textwidth}
 
+\vspace{-1.5cm}
 \centering
-\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize This model says...}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.975\textwidth]
+\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small \color{black!80} This model of relationships says...}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.975\textwidth]
 \begin{itemize}\setlength\itemsep{1mm}
@@
 {\centering
-\begin{tcolorbox}[colframe=blue!20, colback=white, title={\scriptsize This model serves as a \textsc{\textbf{diagnosis tool}} of the approximation:}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.98\textwidth]
+\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small \color{black!80} This model of relationships serves as a\, \textsc{\textbf{\color{black} \normalsize diagnosis tool}}\, of the approximation:}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.98\textwidth]
 \begin{remark}
@@
 
-\vspace{-0.7cm}
+%%%%%%titleformat
+\newcommand{\pepito}[1]{\fcolorbox{aaltoLightGreen!30}{aaltoLightGreen!30}
+{\parbox[c][16mm][c]{0.95\textwidth}
+{\centering\normalsize\sc\bfseries\sffamily\color{black} \thesection. #1}}}
+
+\titleformat{\section}[block]
+{\normalfont\filcenter\color{aaltoGray}}{}{1em}{\pepito}
+%%%%%%
+
+\begin{minipage}{0.35\textwidth}
+
+\vspace{-2cm}
 \section{Univariate synthetic case study}
-\vspace{-0.2cm}
+\vspace{-0.5cm}
+\begin{itemize}\setlength\itemsep{3mm}
+\small
+%\item \textit{\scriptsize Samples draws from}:%\\ $x \in [-1,1] \subset {\rm I\!R}$,
+%\\ GP Mattern($\nu$=3/2, $\ell$=0.15, $\alpha$=1)\\ + $\bm{\epsilon} \sim \mathcal{N}(0, 0.2^2)$
 
-%\hspace{-1.1cm}
-\begin{minipage}{0.33\textwidth}
-\begin{itemize}\setlength\itemsep{4mm}
-\item \textit{\scriptsize Samples draws from}:%\\ $x \in [-1,1] \subset {\rm I\!R}$,
-\\ GP Mattern($\nu$=3/2, $\ell$=0.15, $\alpha$=1)\\ + $\bm{\epsilon} \sim \mathcal{N}(0, 0.2^2)$
+%\item \textit{\scriptsize Observation model}: \pbox{6cm}{$\bm{y} = \bm{f} + \bm{\epsilon}$ \\
+%$\bm{\epsilon} \sim \mathcal{N}(\bm{0}, \sigma^2 \bm{I})$}
 
-\item \textit{\scriptsize Observation model}: \pbox{6cm}{$\bm{y} = \bm{f} + \bm{\epsilon}$ \\
-$\bm{\epsilon} \sim \mathcal{N}(\bm{0}, \sigma^2 \bm{I})$}
+\item \pbox{8cm}{\textit{\footnotesize Exact GP prior for $\bm{f}$}:\\ $f(x) \sim \GP(0, k_{\nicefrac{3}{2}}(x, x'))$}
 
-\item \pbox{8cm}{\textit{\scriptsize Exact GP prior for $\bm{f}$}:\\ $f(x) \sim \GP(0, k_{\nicefrac{3}{2}}(x, x'))$}
-
-\item \pbox{11cm}{\textit{\scriptsize HSGP prior for $\bm{f}$}:\\ $f(x) \approx \sum_{j=1}^m \phi_j(x) \beta_j$,\\ 
+\item \pbox{11cm}{\textit{\footnotesize HSGP prior for $\bm{f}$}:\\ $f(x) \approx \sum_{j=1}^m \phi_j(x) \beta_j$,\\ 
 $\beta_j \sim \mathcal{N}(0,S_{\nicefrac{3}{2}}(\sqrt{\lambda_j}))$}
@@
 \end{minipage}
-\begin{minipage}{0.67\textwidth}
+\begin{minipage}{0.62\textwidth}
 \centering
@@
 
-\vspace{6mm}
-\hspace{-7mm} \textsc{\textbf{\small Diagnosis}}:\\[-4mm]
+\vspace{-0.8cm}
+\hspace{-10mm} \textsc{\textbf{\normalsize Diagnosis}}: {\small(iterative procedure)}\\[-18pt]
 
-\begin{minipage}[t]{0.49\textwidth}
+\hspace{-1cm}
+\begin{minipage}[t]{0.48\textwidth}
 
 \begin{minipage}[t]{0.45\textwidth}
-{\large 1}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=20$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.09$}
+{\large 1}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.09$}
 
@@
 %{\hspace{2.2cm} \scriptsize Error}\\[-3mm]
-\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig2_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig2_vposter.pdf}
 
@@
 \hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
-\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_1_1_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig6_1_1_vposter.pdf}
 
@@
 \end{minipage}
-\begin{minipage}[t]{0.49\textwidth}
+\hspace{0.5cm}
+\begin{minipage}[t]{0.48\textwidth}
 
 \begin{minipage}[t]{0.45\textwidth}
-{\large 2}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=30$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.12$}
+{\large 2}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.12$}
 
 \vspace{0.3cm}
-\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig3_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig3_vposter.pdf}
 
@@
 \vspace{5mm}
-\begin{minipage}[t]{0.49\textwidth}
+\hspace{-1cm}
+\begin{minipage}[t]{0.48\textwidth}
 
 \begin{minipage}[t]{0.45\textwidth}
-{\large 3}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=40$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.14$}
+{\large 3}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.14$}
 
 \vspace{0.3cm}
-\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig4_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig4_vposter.pdf}
 
@@
 \hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
-\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_3_1_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig6_3_1_vposter.pdf}
 
@@
 \end{minipage}
-\begin{minipage}[t]{0.49\textwidth}
+\hspace{0.5cm}
+\begin{minipage}[t]{0.48\textwidth}
 
 \begin{minipage}[t]{0.45\textwidth}
-{\large 4}\hspace{0.7cm} \pbox{1.7cm}{\scriptsize $m=60$\\ $c=1.5$} $\to$\, {\scriptsize$\hat{\ell}=0.14$}
+{\large 4}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.14$}
 
 \vspace{0.3cm}
-\includegraphics[width=0.90\textwidth, trim = 0mm 0mm 10mm 10mm, clip]{\figpdfdir caseI_fig5_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig5_vposter.pdf}
 
@@
 \hspace{0.25\textwidth}  {\tiny Mattern($\nu$=3/2) kernel}\\
-\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_4_1_vposter.pdf}
+\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig6_4_1_vposter.pdf}
 
@@
 
-% bibliography
-\vspace*{3mm}
-{\centering\textbf{References}}
-\renewcommand{\bibsection}{}
-{\tiny % % Bibliography text size begins...
-\bibliographystyle{unsrt}
-\bibliography{references}
-} % <- bibliography text size ends
-
-\end{minipage} % End minipage second column
+\end{minipage} % End minipage third column
 \hspace{0.02\linewidth} 
@@
 
-%%------------ minipage for third column ------------%
-%\begin{minipage}[t]{0.48\textwidth} % minipage for the latter two columns
-%\setlength{\parindent}{10mm} % Paragraph indent
-%
-%\end{minipage} % End minipage third column
-%\hspace{0.03\linewidth}
-
-%\vspace*{0.03\linewidth} % Increase the bottom margins
-
 \end{document}

### 2020-09-10 Paper/StanCon/poster/HSGP_StanCon_poster_landscape_full-content.tex
add StanCon poster folder

--- 2020-09-10 Paper/StanCon/poster/HSGP_StanCon_poster.tex
@@
 %-- minipage for title and authors --%
-\begin{minipage}[t]{0.76\linewidth}
-\vspace{0pt}
+\begin{minipage}[t]{0.50\linewidth}
+\vspace{0.5cm}
 
@@
 
-\vspace{0.015\linewidth}
-
-\normalsize{\textsf{\bfseries{Gabriel Riutort-Mayol$^{1}$ \;\;\; Paul-Christian Bürkner$^2$ \;\;\; Michael Riis Andersen$^{3}$ \;\;\; Arno Solin$^{2}$ \;\;\; Aki Vehtari$^{2}$}}}\\[2mm]
-
-%\setstretch{1.2}
-\textcolor{aaltoGray}{\textsf{\bfseries{\small
-$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
-\\
-$^2$ Department of Computer Science, Aalto University, Finland
-\\
-$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
-\\[-2mm]
-\footnotesize Contact: gabriuma@gmail.com, aki.vehtari@aalto.fi}}}
+\vspace{5mm}
+\normalsize{\textsf{\bfseries{Gabriel Riutort-Mayol$^{1}$ \;\;\; Paul-Christian Bürkner$^2$ \;\;\; Michael Riis Andersen$^{3}$ \;\;\; Arno Solin$^{2}$ \;\;\; Aki Vehtari$^{2}$}}}
 
@@
 %-- minipage for the logo at rigth side --%
-\begin{minipage}[t]{0.06\linewidth}
-\vspace{0.7cm}
+\begin{minipage}[t]{0.22\linewidth}
+
+\vspace{3mm}
+\textcolor{aaltoGray}{\textsf{\bfseries{ \scriptsize
+\begin{itemize}
+	\setlength{\itemsep}{1mm}
+	\item[] $^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain
+\item[] $^2$ Department of Computer Science, Aalto University, Finland
+\item[] $^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
+\item[] Contact: gabriuma@gmail.com, aki.vehtari@aalto.fi
+\end{itemize}
+}}}
+
+\end{minipage}
+%-- minipage for the logo at rigth side --%
+\hspace{15mm}
+\begin{minipage}[t]{0.09\linewidth}
+\vspace{1.5mm}
 
 %\begin{adjustwidth}{-1cm}{-0cm}
-{\centering \includegraphics[scale=0.8]{\figpdfdir aaltologo}}\\[0.5cm]
-\includegraphics[scale=0.7]{\figpdfdir marca_UPV_secundaria_negro300.png}
+\includegraphics[scale=0.8]{\figpdfdir aaltologo}
+\raisebox{-5mm}{\includegraphics[scale=0.7]{\figpdfdir marca_UPV_secundaria_negro300.png}}
 %\end{adjustwidth}
@@
 \sffamily
-%\small % Text size for a1 posters
-\footnotesize % Text size for a1 posters
+\small % Text size
+%\footnotesize % Text size
 
@@
 \centering
-\begin{minipage}[t]{0.48\linewidth}
+\begin{minipage}[t]{0.31\linewidth}
 \vspace{0pt}
@@
 \hspace{45mm} $\ell=0.1$ \hspace{63mm} $\ell=0.3$ \hspace{64mm} $\ell=0.6$\\
-\includegraphics[scale=0.57, trim = 0mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_1.pdf}
-\includegraphics[scale=0.57, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_2.pdf}
-\includegraphics[scale=0.57, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_3.pdf}
+\includegraphics[scale=0.53, trim = 0mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_1.pdf}
+\includegraphics[scale=0.53, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_2.pdf}
+\includegraphics[scale=0.53, trim = 25.5mm 10mm 7mm 18mm, clip]{\figpdfdir func_samples_3.pdf}
 
@@
 
-\vspace{-0.6cm}
+\end{minipage}% end minipage first column
+\hspace{0.015\linewidth} 
+%------------ minipage for second column ------------%
+\begin{minipage}[t]{0.31\textwidth}
+\vspace{0pt}
+\setlength{\parindent}{10mm}
+
+\vspace{0.3cm}
 \section{Hilbert space approximate Gaussian process (HSGP)}
@@
 
-\end{minipage}% end minipage first column
+\end{minipage}% end minipage second column
 \hspace{0.015\linewidth} 
-%------------ minipage for second column ------------%
-\begin{minipage}[t]{0.48\textwidth}
+%------------ minipage for third column ------------%
+\begin{minipage}[t]{0.31\textwidth}
 \vspace{0pt}
@@
 
-\end{minipage} % End minipage second column
+\end{minipage} % End minipage third column
 \hspace{0.02\linewidth} 

### 2020-09-10 Paper/online_supplemental_material/online_material.tex
update supplemental material
updating supplementary material

@@
 \documentclass[onecolumn,a4paper,11pt]{article}
-\voffset -1cm
+\voffset -2cm
 \hoffset -1cm
-\textheight 22.3cm
+\textheight 23.3cm
 \textwidth 15.0cm
@@
 \usepackage{xcolor}
-\graphicspath{{../manuscript_v2/images/}}
+\graphicspath{{./images/}}
 
@@
 \captionsetup{font=footnotesize}
-\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,breaklinks=true]{hyperref}
 
@@
 
-\title{Online supplemental material of the article:\\ \vspace{7mm} 'Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming'\\ \vspace{7mm}
-\normalsize by}
+\title{\begin{flushleft}
+\textsf{\small Statistics and Computing}
+\end{flushleft}  \vspace{0.5cm}  Supplementary material for "Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming"}
 
-\author{Gabriel Riutort-Mayol and Paul-Christian Bürkner and Michael R.\ Andersen and\\
-  Arno Solin and Aki Vehtari}
+\author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R.\ Andersen$^{3}$,\\
+  Arno Solin$^{2}$, Aki Vehtari$^{2}$}
 
-\date{}
+\date{ \small
+$^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain 
+\break
+$^2$ Department of Computer Science, Aalto University, Finland
+\break
+$^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
+\break
+$^*$ Corresponding author, Email: gabriuma@gmail.com
+}
 
@@
 
-This document consists of four case studies of application of the article 'Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming'.
+This supplementary material consists of four case studies of application of the methodology proposed in the article.
 
@@
 The covariance function $k$ depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
-The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
+The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and length-scale, respectively, of the GP process.
 
@@
 
-In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this splines model using the R-package \textit{brms} \citep{burkner2017brms}.
+In order to do model comparison, in addition to the exact GP model and HSGP model, an spline-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
 
-Figure \ref{ch5_fig12_Posteriors_gaydata} shows the posterior mean predictive distributions of the three models, the regular GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the splines model with 20 knots.
+Figure \ref{ch5_fig12_Posteriors_gaydata} shows the posterior mean predictive distributions of the three models, the exact GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the spline-based model with 20 knots.
 Sample observations are plotted as circles in the figure, and the out-of-sample observations, which have been used for testing, are plotted as crosses.
@@
 \includegraphics[scale=0.50]{ch5_fig12_Posteriors_gaydata.pdf}
-\caption{Posterior predictive means of the proposed HSGP model, the regular GP model, and the Splines model. 95\% confident intervals are plotted as dashed lines.}
+\caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the Splines model. 95\% confident intervals are plotted as dashed lines.}
   \label{ch5_fig12_Posteriors_gaydata}
@@
 For the HSGP model, different models with different number of basis functions and boundary factor have been fitted.
-The root mean square errors (RMSE) for every one of these models have been computed against the regular GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}-left, for sample data, and in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}-right, for test data.
-As the boundary factor increases, more basis functions are needed for a accurate approximation.
+The root mean square errors (RMSE) for every one of these models have been computed against the exact GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
+The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
 
-Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model.
-We can see how the splines models do not extrapolate data properly (Figure \ref{ch5_fig14_MSE_train_gaydata}-right).
+Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the exact GP, HSGP and spline-based models computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the spline-based model.
+We can see how the spline-based model does not extrapolate data properly (Figure \ref{ch5_fig14_MSE_train_gaydata}-right).
 
-Figure \ref{ch5_fig14_1_time_gaydata} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the splines model.
+Figure \ref{ch5_fig14_1_time_gaydata} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline-based model.
 The computational times is represented in the y-axis which is on a logarithmic scale.
-The HSGP model is on average roughly 15 times faster than the regular GP and 5 times faster than the spline model, for this particular case and univariate input space.
+The HSGP model is on average roughly 15 times faster than the exact GP and 5 times faster than the spline-based model, for this particular case and univariate input space.
 Computation time increases relatively slight with the number of basis functions in a univariate input space, as can be seen in the figure.
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}}\,.
+The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Same-sex-marriage-data}}\,.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 119mm 0mm, clip]{ch5_fig13_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the HSGP model, computed against the regular GP model, as a function of the number of basis functions $m$ and boundary factor $c$.
+\caption{Root mean square error (RMSE) of the HSGP model, computed against the exact GP model, as a function of the number of basis functions $m$ and boundary factor $c$.
 RMSE for sample data (left) and RMSE for out-of-sample data (right).}
@@
 \subfigure{\includegraphics[scale=0.40, trim = 27mm 35mm 112mm 0mm, clip]{ch5_fig14_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the different methods, regular GP, HSGP and splines models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots for the splines model.
+\caption{Root mean square error (RMSE) of the different methods, exact GP, HSGP and spline-based models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots, for the spline-based model.
 RMSE for sample data (left) and RMSE for out-of-sample data (right).}
@@
 
-This example consists of a simulated dataset with $n=120$ ($i=1,\dots,n$) single draws from a GP prior as function of input values $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$ in a bivariate input space ($D=2$).
-A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and lengthscales $\ell_1=0.10$, for the first input dimension, and $\ell_2=0.35$, for the second input dimension, is used for the GP prior.
+This example consists of a synthetic dataset with $n=120$ ($i=1,\dots,n$) single draws from a GP prior as a function of input values $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$ in a bivariate input space ($D=2$).
+A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, for the first input dimension, and $\ell_2=0.35$, for the second input dimension, is used for the GP prior.
 Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^{120}$.
 
-The regular GP model over the outcome variable $\bm{y}$ can be written as follows,
+The exact GP model over the outcome variable $\bm{y}$ can be written as follows,
 %
@@
 The function $f:{\rm I\!R}^2 \to {\rm I\!R}$ is a GP prior with a multivariate squared exponential covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell_1,\ell_2\}$.
-The hyperparameters $\alpha$, $\ell_1$ and $\ell_2$ represent the marginal variance and lengthscales for first and second input dimensions, respectively, of the GP process.
+The hyperparameters $\alpha$, $\ell_1$ and $\ell_2$ represent the marginal variance and length-scales for first and second input dimensions, respectively, of the GP process.
+Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(\bm{x}_i,\bm{x}_j,\theta)$, with $i,j=1,\dots,120$.
 
@@
 
-In the HSGP model with $2$ input dimensions, the latent function $f(\bm{x})$ is approximated as in equation (12)% (\ref{eq_approxf_multi})
+In the HSGP model with $2$ input dimensions, the latent function $f(\bm{x})$ is approximated as in equation (13)% (\ref{eq_approxf_multi})
 , with the $2$-dimensional squared exponential spectral density $S$ as in equation (1)% (\ref{eq_specdens_inf})
-, and the $2$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (9) %(\ref{eq_eigenfunction_multi})
- and (10)% (\ref{eq_eigenvalue_multi})
+, and the $2$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (10) %(\ref{eq_eigenfunction_multi})
+ and (11)% (\ref{eq_eigenvalue_multi})
  , respectively.
 
-In order to do model comparison, in addition to the regular GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
+In order to do model comparison, in addition to the exact GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
 
-Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, regular GP, HSGP, and splines, fitted over the dataset.
+Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, exact GP, HSGP, and splines, fitted over the dataset.
 Sample observations are also plotted as circles.
-For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension respectively, were used, which lead to a total of 1600 multivariate basis functions.
-A boundary factor for each dimension $c_1=1.5$ and $c_2=1.5$ were used.
-For the splines model, 40 knots in each dimension were used.
+For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension, respectively, were used, which lead to a total of 1600 multivariate basis functions.
+A boundary factor for each dimension, $c_1=1.5$ and $c_2=1.5$, respectively, were used.
+For the spline-based model, 40 knots in each dimension were used.
 
-Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the data-generating function and the GP, HSGP and splines models, respectively.
+Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the data-generating function and the GP, HSGP and spline-based models, respectively.
 
@@
 
-In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions, for the HSGP model, and different number of knots, for the splines model, have been fitted.
-In all models, the same number of basis functions and knots per dimension were used.
-Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$, and the number of univariate basis functions $m$, for the HSGP model, and knots, for the splines model.
-From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen a close approximation of the HSGP model to the regular GP model.
-However, the performance of the splines model is significantly worse.
+In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions, for the HSGP model, and different number of knots, for the spline-based model, have been fitted.
+In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
+Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and the number of univariate basis functions $m$, for the HSGP model, and knots, for the spline-based model.
+From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen an accurate approximation of the HSGP model to the exact GP model.
+However, the performance of the spline-based model is significantly worse.
 
@@
 Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computation time.
-Even though in a bivariate input space the computation demand increases significantly with the number of dimensions or knots, either the HSGP or spline models work significantly better than regular GP, even for highly wiggly functions that require a high number of basis functions or knots for a accurate approximation.
-However, serious difficulties with cumputation time were encountered in building the spline model with $50$ knots for this bivariate function.
+Even though in a bivariate input space the computational demands increase significantly with the number of dimensions or knots, either the HSGP or spline-based models work significantly better than exact GP, even for highly wiggly functions that require a high number of basis functions or knots for an accurate approximation.
+It is interesting to be noticed that just the computation of the input data for the spline model in a 2D input space and considering more than 50 knots per dimension is computationally very expensive.
 
-The Stan model codes for the exact GP, the approximate GP and the splines models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
+The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
-\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods computed against the data-generating function, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
   \label{ch5_fig17_RMSE_exII}
@@
 The data contains $n=392$ individuals ($i=1,\dots,n$) from which the binary variable of suffering ($y_i=1$) or not suffering ($y_i=0$) from diabetes have been observed.
-The input vector $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$, in a 4D input space ($D=4$), contains the risk factors, \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$.
+The input vector $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$, in a 4D input space ($D=4$), contains the risk factors: \textit{Glucose} ($x_{i1}$), \textit{Pregnancy} ($x_{i2}$), \textit{Age} ($x_{i3}$) and \textit{BMI} ($x_{i4}$), per individual $i$.
 The observational model is a Bernoulli model with parameter the probability $p_i$ of suffering from diabetes per observation $i$,
@@
 
-\noindent The goal is to estimate the probability $p_i$ as a function of the risk factors, which function $f(\cdot):{\rm I\!R}^{4} \to {\rm I\!R}$ is modeled as a GP with a multivariate squared exponential covariance function $k$ depending on the risk factors $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
+\noindent The goal is to estimate the probability $p_i$ as a function $f$ of the risk factors.
+Which function $f(\cdot):{\rm I\!R}^{4} \to {\rm I\!R}$ is modeled as a GP with a multivariate squared exponential covariance function $k$ depending on the risk factors $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
 %
@@
 
-\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and lengthscale, respectively, of the GP process.
-A scalar lengthscale has been considered in the multivariate covariance function $k$.
+\noindent The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and length-scale, respectively, of the GP process.
+A scalar length-scale has been considered in the multivariate covariance function $k$.
 
-In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$ evaluated at input vector $\bm{x} \in {\rm I\!R}^D$ is approximated as in equation (12)% (\ref{eq_approxf_multi})
-, with the $D$-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (1) %(\ref{eq_specdens_inf}) 
-and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (9) % (\ref{eq_eigenvalue_multi}) 
-and (10)% (\ref{eq_eigenfunction_multi})
+In the HSGP model with $D$ input dimensions, the function $f(\bm{x})$ evaluated at input vector $\bm{x} \in {\rm I\!R}^D$ is approximated as in equation (13)% (\ref{eq_approxf_multi})
+, with the $D$-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in equation (1) %(\ref{eq_specdens_inf}) 
+and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (10) % (\ref{eq_eigenvalue_multi}) 
+and (11)% (\ref{eq_eigenfunction_multi})
 , respectively.
 
-In order to do model comparison, in addition to the regular GP and HSGP models, a $D$-dimensional spline-based model is also fitted using a cubic spline basis penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
+In order to do model comparison, in addition to the exact GP and HSGP models, a $D$-dimensional spline-based model is also fitted using a cubic spline basis penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
 
-Figure \ref{ch5_fig18_gpfun_diabetes} shows the posterior mean predictions of probabilities ($p_i$) of the three models, regular GP, HSGP and spline, fitted over the dataset with the 2 input dimensions \textit{Glucose} and \textit{Pregnancy} ($D=2$).
+Figure \ref{ch5_fig18_gpfun_diabetes} shows the posterior mean predictions of probabilities ($p_i$) of the three models, exact GP, HSGP and splines, fitted over the dataset with the two input dimensions \textit{Glucose} and \textit{Pregnancy} ($D=2$).
 The binary observations $y_i$ are also plotted in the plots as colored points.
 For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions.
-A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used.
-For the spline model, 20 knots per dimension were used.
+A boundary factor for each dimension, $c_1=4$ and $c_2=4$, respectively, were used.
+For the spline-based model, 20 knots per dimension were used.
 %
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 4mm 2mm 15mm, clip]{ch5_fig18_legend_diabetes.pdf}}
-\caption{Posterior predictive mean functions of the GP (a), HSGP (b) and spline (c) models.
-Samples observations of suffering (red points) and not suffering (blue points) from the disease are plotted.}
+\caption{Posterior predictive mean functions of the GP (a), HSGP (b) and spline (c) models. Sample observations of suffering (red points) and not suffering (blue points) from the disease are plotted.}
   \label{ch5_fig18_gpfun_diabetes}
@@
 
-In order to assess the performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factor for the HSGP model and different number of knots for the spline model have been fitted.
-In all models, the same boundary factor, number of basis functions and knots per dimension were used.
-Figure \ref{ch5_fig19_ELPD_diabetes} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) as a function of the boundary factor $c$ and the number of univariate basis functions $m$, for the HSGP model, and knots, for the spline model.
+In order to assess the performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factors, for the HSGP model, and different number of knots, for the spline model, have been fitted.
+In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
+Figure \ref{ch5_fig19_ELPD_diabetes} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP model, and knots, for the spline-based model.
 The ELPD is computed over the actual observations by cross-validation.
-Basically, with slightly differences, all models show similar performances, due to the fact that the process is very smooth with a relatively very large lengthscale estimate $\ell=4.51$.
-Even though, a slight pattern of performance improvement can be appreciated as the boundary factor $c$ increases, which fact is because small boundary factors are not allowed when large lengthscales (Figure 6% \ref{fig6_lscale_vs_J_vs_c}
+With slightly differences, all models show similar performances due to the fact that the process is very smooth with a relatively very large length-scale estimate $\ell=4.51$.
+Even though, a slight improvement can be appreciated as the boundary factor $c$ increases, because small boundary factors are not allowed when large length-scales (Figure 6% \ref{fig6_lscale_vs_J_vs_c}
 ).
@@
 
-Figure \ref{ch5_fig20_time2D_diabetes} shows the computational times of the different models, regular GP, HSGP and spline, fitted over the dataset with 2, 3 and 4 input dimensions, as a function of the boundary factor $c$ and number of univariate basis functions $m$ and knots.
-The time of omputation increases significantly with higher dimensions for the HSGP and spline models.
-In this sense, choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure 6% \ref{fig6_lscale_vs_J_vs_c}
-, can be essential to avoid a excessive computation time, especially in multivariate input spaces.
-Serious difficulties with computation time were encountered in building spline models in a 4D input space or even in a 3D input space with more than 10 knots.
+Figure \ref{ch5_fig20_time2D_diabetes} shows the computational times of the different models, exact GP, HSGP and splines, fitted over the dataset with 2, 3 and 4 input dimensions, as a function of the boundary factor $c$ and number of univariate basis functions $m$ and knots.
+We can appreciate the significant increase
+of computational time with higher dimensions for the HSGP and spline models.
+This fact reveals that choosing optimal values for the number of basis functions and
+boundary factor, looking at the recommendations and diagnosis provided by Figure 6% \ref{fig6_lscale_vs_J_vs_c}
+, is essential to avoid a excessive computational time, especially in high input
+dimensionality. It is interesting to be noticed that considering the spline model in a 4D input space is not allowed for an amount of 392
+observations. Similarly, just the computation of the input data for the spline model
+in a 3D input space and considering more than 10 knots per dimension is computationally very expensive.
 
-The Stan model codes for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}}\,.
+The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Diabetes-data}}\,.
 
@@
 
-Due to the fact that 52 input variables are too high-dimensional for a multivariate HSGP model, which computational cost scales as $O(n m^D+m^D)$, with $m$ the number of basis functions and $D$ the number of input variables, the multivariate HSGP model will be formulated as an additive HSGP model.
+Due to the fact that 52 input variables are too high-dimensional for a multivariate HSGP model, which computational cost scales as $O(n m^D+m^D)$ for $m$ basis functions and $D$ input variables, the multivariate HSGP model will be formulated as an additive HSGP model.
 
 As it is known, the computational demand of a multivariate HSGP model component increases quickly with the number of dimensions, so we should avoid high-dimensional HSGP components in the additive model.
-Original input variables are highly correlated, which would imply the use of high-order interaction components in the additive model to achieve accurate model performance.
-Therefore, instead of using the original variables as inputs, we use their principal components, which are expect to be linearly uncorrelated.
-Using the principal components as inputs helps, in principle, not to have to use as many high-order interaction components in the additive model.
+Original input variables are highly correlated, which would imply the use of high-order interaction components in the additive model.
+Therefore, instead of using the original variables as inputs, we use their principal components (PCs), which are expect to be linearly uncorrelated.
+%Using the principal components as inputs helps, in principle, not to have to use as many high-order interaction components in the additive model.
 
-The principal components (PCs) will be used jointly with the time variable as inputs in the classifying additive HSGP model.
-Let's denote the matrix $\bm{X}=[\bm{x}_{11}\, \cdots \, \bm{x}_{it} \, \cdots \, \bm{x}_{nT}]^\top \in {\rm I\!R}^{nT\times D}$, which contains the input vectors $\bm{x}_{it} \in {\rm I\!R}^D$, $D=53$ ($52 \text{ PCs plus time}$) for the spatio-temporal observations (plots $i=1,\dots,200$, and times $t=1,\dots,5$).
+The PCs will be used jointly with the time variable as inputs in the classifying additive HSGP model.
+Let's denote the matrix $\bm{X}=[\bm{x}_{11}\, \cdots \, \bm{x}_{it} \, \cdots \, \bm{x}_{nT}]^\top \in {\rm I\!R}^{nT\times D}$, which contains the input vectors $\bm{x}_{it} \in {\rm I\!R}^D$, $D=53$ ($52 \text{ PCs plus time}$), for the spatio-temporal observations (plots $i=1,\dots,200$, and times $t=1,\dots,5$).
 
@@
 $$
-f(\bm{x}_{it})=\big(f_1(\bm{x}_{it}),\dots,f_K(\bm{x}_{it}) \big),
+f(\bm{x}_{it})=\big(f_1(\bm{x}_{it}),\dots,f_K(\bm{x}_{it}) \big).
 $$
 
-\noindent which is related to the vector of probabilities $\bm{p}_{it}$ through the 'softmax' link function:
+\noindent $f(\bm{x}_{it})$ is related to the vector of probabilities $\bm{p}_{it}$ through the 'softmax' link function,
 %
@@
 
-\noindent In the HSGP model, a first-order components $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in equation (7) % (\ref{eq_approxf}) 
+\noindent In the HSGP model, a first-order component $g_d(x^d)$, evaluated at input value $x^d \in {\rm I\!R}$, is approximated as in equation (7), % (\ref{eq_approxf}) 
 with the squared exponential spectral density $S$ as in equation (1) % (\ref{eq_specdens_inf})
@@
 
-\noindent In the HSGP model, a second-order component $h_{d,D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in equation (12)% (\ref{eq_approxf_multi})
- with the two-dimensional (with a scalar lengthscale) squared exponential spectral density $S$ as in equation (1) % (\ref{eq_specdens_inf}) 
- and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (9) % (\ref{eq_eigenvalue_multi}) 
- and (10)% (\ref{eq_eigenfunction_multi})
+\noindent In the HSGP model, a second-order component $h_{d,D}(x^d,x^D)$, evaluated at inputs $x^d \in {\rm I\!R}$ and $x^D \in {\rm I\!R}$, is approximated as in equation (13), % (\ref{eq_approxf_multi})
+ with the two-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in equation (1) % (\ref{eq_specdens_inf}) 
+ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (10) % (\ref{eq_eigenvalue_multi}) 
+ and (11)% (\ref{eq_eigenfunction_multi})
  , respectively.
 
-The vector of hyperparameters $\theta_{d,1}=(\alpha_{d,1},\ell_{d,1})$ contains the marginal variance $\alpha_{d,1}$ and lengthscale $\ell_{d,1}$ of the $g_d(x^d)$ model component.
-And, the vector of hyperparameters $\theta_{d,D}=(\alpha_{d,D},\ell_{d,D})$ contains the marginal variance $\alpha_{d,D}$ and lengthscale $\ell_{d,D}$ of the $h_{d,D}(x^d,x^D)$ model component.
+The vector of hyperparameters $\theta_{d,1}=(\alpha_{d,1},\ell_{d,1})$ contains the marginal variance $\alpha_{d,1}$ and length-scale $\ell_{d,1}$ of the $g_d(x^d)$ model component.
+And, the vector of hyperparameters $\theta_{d,D}=(\alpha_{d,D},\ell_{d,D})$ contains the marginal variance $\alpha_{d,D}$ and length-scale $\ell_{d,D}$ of the $h_{d,D}(x^d,x^D)$ model component.
 
@@
 For the second-order components $h_{d,D}(x^d,x^D)$, $m_1=15$ and $m_2=15$ basis functions for each dimension, respectively, were used, which lead to a total of 225 multivariate basis functions.
-A boundary factor for each dimension $c_1=2.5$ and $c_2=2.5$ were used.
+A boundary factor for each dimension, $c_1=2.5$ and $c_2=2.5$, respectively, were used.
 All the input variables were previously standardized.
 
-In the case of the first-order components, the normalized lengthscale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum lengthscale reported by Figure (6) % \ref{fig6_lscale_vs_J_vs_c}
+In the case of the first-order components, the normalized length-scale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum approachable length-scale reported by Figure (6) % \ref{fig6_lscale_vs_J_vs_c}
 as a function of $m$ and $c$. Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.
 
-For the second-order components, the relationships between the number of basis functions, the boundary factor and the lengthscale is not available for the multivariate case.
-However, we can approximately analyze the lengthscale estimates of the second-order HSGP components analyzing each dimension separately as unidimensional HSGP models.
+For the second-order components, the relationships between the number of basis functions, the boundary factor and the length-scale is not available for the multivariate case.
+However, we can approximately diagnose the length-scale estimates of the second-order HSGP components analyzing each dimension separately as unidimensional HSGP models.
 
@@
 
-The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}\,.
+The Stan model code for the HSGP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Land-use-classification}}\,.
 

### 2020-09-11 Paper/Manuscript_Stats&Comp/manuscript_ST&CO.tex
Minor edits to new paragraph on the diagnostic

@@
 The SRMSE is computed against the data-generating model.
-From Figures \ref{fig11_Posteriors_exI} and \ref{fig12_MSE_exI_inter}, it is seen that the HSGP method yield a good approximation of the exact GP model for both interpolation and extrapolation.
+From Figures \ref{fig11_Posteriors_exI} and \ref{fig12_MSE_exI_inter}, it is seen that the HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation.
 However, the spline model does not extrapolate data properly.
@@
 
-In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.
+%In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.
+
 %The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}.
 Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
-Figure \ref{fig13_diagnosis} depicts the diagnosis applied for this case study.
+Next, we demonstrate how to apply the diagnostic for approximation quality as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency}.
+Figure \ref{fig13_diagnosis} depicts the diagnosis applied to case study.
 In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnosis}.
-In the second iteration, $m$ is increased up to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig7_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
+In the second iteration, $m$ is increased to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig7_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
 As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{fig12_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
-In the third iteration, $m$ is increased up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$.
+In the third iteration, $m$ is increased to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships suggests a minimum length-scale of $\ell^*=0.15$.
 As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{fig12_MSE_exI_inter}-left).
-In the fourth iteration, $m$ is increased up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, and as $\hat{\ell} > \ell^*$ then approximation is diagnosed to be accurate.
+In the fourth iteration, $m$ is increased up $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum length-scale generated by the model of relationships is $\ell^*=0.11$, and the approximation is finally diagnosed to be accurate as $\hat{\ell} > \ell^*$.
 

### 2020-09-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
remane manuscript file and add a footnote

--- 2020-09-10 Paper/Manuscript_Stats&Comp/manuscript_ST&CO.tex
@@
 % your LaTeX will extract the file if required
-\begin{filecontents*}{example.eps}
-%!PS-Adobe-3.0 EPSF-3.0
-%%BoundingBox: 19 19 221 221
-%%CreationDate: Mon Sep 29 1997
-%%Creator: programmed by hand (JK)
-%%EndComments
-gsave
-newpath
-  20 20 moveto
-  20 220 lineto
-  220 220 lineto
-  220 20 lineto
-closepath
-2 setlinewidth
-gsave
-  .4 setgray fill
-grestore
-stroke
-grestore
-\end{filecontents*}
-%
-\RequirePackage{fix-cm}
+%\begin{filecontents*}{example.eps}
+%%!PS-Adobe-3.0 EPSF-3.0
+%%%BoundingBox: 19 19 221 221
+%%%CreationDate: Mon Sep 29 1997
+%%%Creator: programmed by hand (JK)
+%%%EndComments
+%gsave
+%newpath
+%  20 20 moveto
+%  20 220 lineto
+%  220 220 lineto
+%  220 20 lineto
+%closepath
+%2 setlinewidth
+%gsave
+%  .4 setgray fill
+%grestore
+%stroke
+%grestore
+%\end{filecontents*}
+%
+%\RequirePackage{fix-cm}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
            Paul-Christian Bürkner \at
-              Department of Computer Science, Aalto University, Finland
+              Department of Computer Science, University of Stuttgard, Germany \\
+              \emph{Most of work was done while at Aalto University}
            \and

### 2020-09-11 Paper/online_supplemental_material/online_supplemental_material.tex
update and rename online supplemental material

--- 2020-09-10 Paper/online_supplemental_material/online_material.tex
@@
 
-In order to do model comparison, in addition to the exact GP model and HSGP model, an spline-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+In order to do model comparison, in addition to the exact GP and HSGP models, an spline-based model is also fitted using the Thin Plate Regression Splines approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
 
-Figure \ref{ch5_fig12_Posteriors_gaydata} shows the posterior mean predictive distributions of the three models, the exact GP, the HSGP model with $m=20$ basis functions and boundary factor $c=1.5$, and the spline-based model with 20 knots.
-Sample observations are plotted as circles in the figure, and the out-of-sample observations, which have been used for testing, are plotted as crosses.
+Figure \ref{ch5_fig12_Posteriors_gaydata} shows the posterior mean predictive distributions of the three models, the exact GP, the HSGP with $m=20$ basis functions and boundary factor $c=1.5$, and the spline-based model with 20 knots.
+Sample observations are plotted as circles in the figure, and the out-of-sample (test) observations, which have been used for testing, are plotted as crosses.
 
@@
 \includegraphics[scale=0.50]{ch5_fig12_Posteriors_gaydata.pdf}
-\caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the Splines model. 95\% confident intervals are plotted as dashed lines.}
+\caption{Posterior predictive means of the proposed HSGP, the exact GP, and the spline-based models. 95\% confident intervals are plotted as dashed lines.}
   \label{ch5_fig12_Posteriors_gaydata}
@@
 
-For the HSGP model, different models with different number of basis functions and boundary factor have been fitted.
-The root mean square errors (RMSE) for every one of these models have been computed against the exact GP model, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
+For the HSGP, different models with different number of basis functions and boundary factor have been fitted.
+The root mean square errors (RMSE) for every one of these models have been computed against the exact GP, and plotted as a function of the number of basis functions $m$ and boundary factor $c$ in Figure \ref{ch5_fig13_MSE_train_BF_gaydata}, for sample (left) and test (right) data.
 The expected patterns of the approximation as a function of the number of basis functions and boundary factor are recognized: as the boundary factor increases, more basis functions are needed.
 
-Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the exact GP, HSGP and spline-based models computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the spline-based model.
+Figure \ref{ch5_fig14_MSE_train_gaydata} shows the RMSE of the exact GP, HSGP and spline-based models computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$, for the HSGP, and knots for the spline-based model.
 We can see how the spline-based model does not extrapolate data properly (Figure \ref{ch5_fig14_MSE_train_gaydata}-right).
 
-Figure \ref{ch5_fig14_1_time_gaydata} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline-based model.
+Figure \ref{ch5_fig14_1_time_gaydata} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP, and knots, for the spline-based model.
 The computational times is represented in the y-axis which is on a logarithmic scale.
-The HSGP model is on average roughly 15 times faster than the exact GP and 5 times faster than the spline-based model, for this particular case and univariate input space.
-Computation time increases relatively slight with the number of basis functions in a univariate input space, as can be seen in the figure.
+The HSGP is on average roughly 15 times faster than the exact GP and 5 times faster than the spline-based model, for this particular case and univariate input space.
+It can be appreciated that the computational time increases relatively slightly with the number of basis functions in a univariate input space.
 
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 45mm 119mm 0mm, clip]{ch5_fig13_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the HSGP model, computed against the exact GP model, as a function of the number of basis functions $m$ and boundary factor $c$.
-RMSE for sample data (left) and RMSE for out-of-sample data (right).}
+\caption{Root mean square error (RMSE) of the HSGP, computed against the exact GP, as a function of the number of basis functions $m$ and boundary factor $c$.
+RMSE for sample (left) and test (right) data.}
   \label{ch5_fig13_MSE_train_BF_gaydata}
@@
 \subfigure{\includegraphics[scale=0.40, trim = 27mm 35mm 112mm 0mm, clip]{ch5_fig14_legend_gaydata.pdf}}
-\caption{Root mean square error (RMSE) of the different methods, exact GP, HSGP and spline-based models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and knots, for the spline-based model.
-RMSE for sample data (left) and RMSE for out-of-sample data (right).}
+\caption{Root mean square error (RMSE) of the different methods, exact GP, HSGP and spline-based models, computed against the actual data, as a function of the number of basis functions $m$ and boundary factor $c$, for the HSGP, and knots, for the spline-based model.
+RMSE for sample (left) and test (right) data.}
   \label{ch5_fig14_MSE_train_gaydata}
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig14_1_time_gaydata.pdf}}
-\caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$ and knots.
+\caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP, and knots, for the spline-based model.
 The y-axis is in a logarithmic scale.}
@@
 In the HSGP model with $2$ input dimensions, the latent function $f(\bm{x})$ is approximated as in equation (13)% (\ref{eq_approxf_multi})
-, with the $2$-dimensional squared exponential spectral density $S$ as in equation (1)% (\ref{eq_specdens_inf})
+, with the $2$D squared exponential spectral density $S$ as in equation (1)% (\ref{eq_specdens_inf})
 , and the $2$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (10) %(\ref{eq_eigenfunction_multi})
@@
 
-In order to do model comparison, in addition to the exact GP and HSGP models, a two-dimensional splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+In order to do model comparison, in addition to the exact GP and HSGP models, a $2$D splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
@@
 Sample observations are also plotted as circles.
-For the HSGP model, $m_1=40$ and $m_2=40$ basis functions for each dimension, respectively, were used, which lead to a total of 1600 multivariate basis functions.
-A boundary factor for each dimension, $c_1=1.5$ and $c_2=1.5$, respectively, were used.
+For the HSGP, $m_1=40$ and $m_2=40$ basis functions for each dimension, respectively, were used, which lead to a total of 1600 multivariate basis functions.
+A boundary factor for each dimension, $c_1=1.5$ and $c_2=1.5$, were used.
 For the spline-based model, 40 knots in each dimension were used.
@@
 \end{center}
-\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline (d) models. Sample points are plotted as circles}
+\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline-based (d) models. Sample points are plotted as circles}
   \label{ch5_fig15_Posteriors_exII}
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
-\caption{Error between the data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline (c) models. Sample points are plotted as circles.}
+\caption{Error between the data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models. Sample points are plotted as circles.}
   \label{ch5_fig16_errors_exII}
@@
 
-In order to assess performance of the models as a function of the number of basis functions and knots, different models with different number of basis functions, for the HSGP model, and different number of knots, for the spline-based model, have been fitted.
+In order to assess performance of the models as a function of the number of basis functions, boundary factor and knots, different models with different number of basis functions, for the HSGP, and different number of knots, for the spline-based model, have been fitted.
 In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
-Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and the number of univariate basis functions $m$, for the HSGP model, and knots, for the spline-based model.
-From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen an accurate approximation of the HSGP model to the exact GP model.
+Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the spline-based model.
+From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen an accurate approximation of the HSGP to the exact GP.
 However, the performance of the spline-based model is significantly worse.
 
-Figure \ref{ch5_fig17_RMSE_exII}-right shows the computational times of the different models as a function of the boundary factor, number of basis functions and knots.
+Figure \ref{ch5_fig17_RMSE_exII}-right shows the computation times of the different models as a function of the boundary factor, number of basis functions and knots.
 Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computation time.
@@
 \subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
-\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods computed against the data-generating function, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
   \label{ch5_fig17_RMSE_exII}
@@
 
-\noindent The goal is to estimate the probability $p_i$ as a function $f$ of the risk factors.
+\noindent The goal is to estimate the probability $p_i$ as a function ($f$) of the risk factors.
 Which function $f(\cdot):{\rm I\!R}^{4} \to {\rm I\!R}$ is modeled as a GP with a multivariate squared exponential covariance function $k$ depending on the risk factors $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$, and related to the probabilities $p_i$ through the {\it logit} link function,
@@
 The binary observations $y_i$ are also plotted in the plots as colored points.
-For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions.
-A boundary factor for each dimension, $c_1=4$ and $c_2=4$, respectively, were used.
+For the HSGP, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions.
+A boundary factor for each dimension, $c_1=4$ and $c_2=4$, were used.
 For the spline-based model, 20 knots per dimension were used.
@@
 \subfigure{\includegraphics[scale=0.30, trim = 150mm 4mm 2mm 15mm, clip]{ch5_fig18_legend_diabetes.pdf}}
-\caption{Posterior predictive mean functions of the GP (a), HSGP (b) and spline (c) models. Sample observations of suffering (red points) and not suffering (blue points) from the disease are plotted.}
+\caption{Posterior predictive mean functions of the GP (a), HSGP (b) and spline-based (c) models.
+Sample observations of suffering (red points) and not suffering (blue points) from the disease are plotted.}
   \label{ch5_fig18_gpfun_diabetes}
@@
 
-In order to assess the performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factors, for the HSGP model, and different number of knots, for the spline model, have been fitted.
+In order to assess the performance of the models as a function of the boundary factor, the number of basis functions and knots, different models with different number of basis functions and boundary factors, for the HSGP, and different number of knots, for the spline-based model, have been fitted.
 In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
-Figure \ref{ch5_fig19_ELPD_diabetes} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP model, and knots, for the spline-based model.
+Figure \ref{ch5_fig19_ELPD_diabetes} shows the expected log predictive density (ELPD; see \cite{vehtari_2012}) as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the spline-based model.
 The ELPD is computed over the actual observations by cross-validation.
 With slightly differences, all models show similar performances due to the fact that the process is very smooth with a relatively very large length-scale estimate $\ell=4.51$.
-Even though, a slight improvement can be appreciated as the boundary factor $c$ increases, because small boundary factors are not allowed when large length-scales (Figure 6% \ref{fig6_lscale_vs_J_vs_c}
+Even though, a slight improvement can be appreciated with larger boundary factor $c$, since small boundary factors are not allowed when large length-scales (Figure 6% \ref{fig6_lscale_vs_J_vs_c}
 ).
@@
 \subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig19_ELPD_diabetes.pdf}}
-\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$, for the HSGP model, and knots, for the spline model.}
+\caption{Expected log predictive density (ELPD) of the different methods as a function of the boundary factor $c$ and the number of basis functions $m$, for the HSGP model, and knots, for the spline-based model.}
   \label{ch5_fig19_ELPD_diabetes}
@@
 \subfigure{\includegraphics[scale=0.35, trim = 28mm 50mm 112mm 10mm, clip]{ch5_fig20_legend_diabetes_2.pdf}}
-\caption{Time of computation in seconds per iteration (iteration of the HMC sampling method) of the different models fitted over the dataset, with 2 input dimensions (left) and 3 input dimensions (center) and 4 input dimensions (right), as a function of the boundary factor $c$ and number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+\caption{Time of computation in seconds per iteration (iteration of the HMC sampling method) of the different models fitted over the dataset, with 2 (left), 3 (center) and 4 (right) input dimensions, as a function of the boundary factor $c$ and number of basis functions $m$, for the HSGP model, and knots, for the spline model.
 The y-axis is on a logarithmic scale.}
@@
 
-Due to the fact that 52 input variables are too high-dimensional for a multivariate HSGP model, which computational cost scales as $O(n m^D+m^D)$ for $m$ basis functions and $D$ input variables, the multivariate HSGP model will be formulated as an additive HSGP model.
+Due to the fact that 52 input variables are too high-dimensional for a multivariate HSGP model, since computation scales as $O(n m^D+m^D)$ for $m$ basis functions and $D$ input variables, the multivariate HSGP model will be formulated as an additive HSGP model, where computational complexity is linear with the number of additive components.
 
@@
  with the two-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in equation (1) % (\ref{eq_specdens_inf}) 
- and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations (10) % (\ref{eq_eigenvalue_multi}) 
+ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ and multivariate eigenfunctions $\phi_j$ as in equations (10) % (\ref{eq_eigenvalue_multi}) 
  and (11)% (\ref{eq_eigenfunction_multi})
@@
 A boundary factor for each dimension, $c_1=2.5$ and $c_2=2.5$, respectively, were used.
-All the input variables were previously standardized.
+All the input variables were previously standardised.
 
-In the case of the first-order components, the normalized length-scale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum approachable length-scale reported by Figure (6) % \ref{fig6_lscale_vs_J_vs_c}
+In the case of the first-order components, the normalized length-scale estimates $\left(\frac{2\cdot \hat{\ell}_{d,1}}{|x^d_{max}-x^d_{min}|}\right)$ are all bigger than the minimum length-scale reported by Figure (6) % \ref{fig6_lscale_vs_J_vs_c}
 as a function of $m$ and $c$. Which means that the used number of basis functions ($m=15$) and boundary factor ($c=2.5$) are suitable values for modeling accurately the input effects.

### 2020-09-15 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
update manuscript and supplemental material

@@
            Paul-Christian Bürkner \at
-              Department of Computer Science, University of Stuttgard, Germany \\
-              \emph{Most of work was done while at Aalto University}
+              Excellence Cluster for Simulation Technology, University of Stuttgart, Germany \\
+              \emph{Most of work was done while at Dept. of Computer Science at Aalto University}
            \and

### 2021-02-11 Paper/online_supplemental_material/online_supplemental_material.tex
update bib path in suplemental material

@@
 \author{Gabriel Riutort-Mayol$^{1*}$, Paul-Christian Bürkner$^2$, Michael R.\ Andersen$^{3}$,\\
-  Arno Solin$^{2}$, Aki Vehtari$^{2}$}
+  Arno Solin$^{4}$, Aki Vehtari$^{4}$}
 
@@
 \break
-$^2$ Department of Computer Science, Aalto University, Finland
+$^2$ Excellence Cluster for Simulation Technology, University of Stuttgart, Germany
 \break
 $^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
+\break
+$^4$ Department of Computer Science, Aalto University, Finland
 \break
@@
 
-\bibliography{../manuscript_v2/references}
+\bibliography{../Manuscript_Stats&Comp/references}
 

### 2022-01-02 Paper/Manuscript_Stats&Comp/1st_journal_review/Numeric_diagnosis_function/Diagnosis_HSGP_v2.Rmd
add 1st journal review

---
title: 'HSGP model diagnosis tool'
# subtitle: 'Diagnosis tool'
date: "`r format(Sys.Date())`."
output:
  bookdown::html_document2:
    theme: readable
    toc: yes
    toc_depth: 3
    # toc_float:
    #   collapsed: false
    #   smooth_scroll: true
    code_download: yes
    number_sections: true
    df_print: paged # kable #
---

<style type="text/css">
body, td{ font-size: 16px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
h1.title {
  font-size: 32px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 20px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
}
h3 { /* Header 3 */
    font-size: 18px;
  color: Black;
}
#TOC {
  color: DarkBlue;
  font-size: 16px;
}
</style>

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  comment = ''
)
```

The report with the reviewers' comments and our responses and thoughts to them can be found at [link](https://docs.google.com/document/d/1aDvI598Z6PNqRlPWrzsnqrXDYC5k_w9bdV_D_vJRgH4/edit).

- This notebook addresses the task of automatizing diagnosis required by the reviewers in **comments 1), 2), 3)**.

- **Comment 13)** suggests:

1. Build a user guide with the iterative steps to perform the diagnosis using the diagnosis tool we have developed in comments 1), 2) and 3).

2. Propose how much user has to increase $m$ in each iteration of the diagnosis.

- **Comment 4.1)** relates to make the application of the method to periodic kernels more explicit in the main text; move the approximation for periodic kernels and its diagnosis' Figure from the appendix to the main text.

- **Comment 4.2)** is a specific comment about writing the steps the user needs to do in order to make HSGPs work for a new kernel (i.e., we need Fourier transform of the kernel). Kernel combinations that satisfy the requirements can also be used.

- **Comments 5) and 11)** are general comments about the general goal of the work. Maybe these could be the last comments to be addressed.

- **Comments 7), 9), 10), 15) and 16)** are about:

1. Make the diagnosis prominent in all the case studies

2. Make more clear how the dimensionality affects to our approach

3.  Move at least one 2D case study to the main text and investigate whether the diagnosis in multi-D cases can be done by just inspecting each single dimension independently.

- **Comment 8)** is a specific comment about the text-length of the first case study. It can be addressed after addressing comment 7).

- About **Comment 12)** we have already made some discussion in the revision document.

- **Comment 14)** is about restructering some parts of the paper.

- **Comments 17) and 18)** are specific comments about comparing our approach with variational methods and Nystrom. Aki already made some discussion about this in the revision document.

```{r echo=TRUE, results='hold'}  
library(RColorBrewer)
library(reshape2)
library(ggplot2)
library(latex2exp)
library(gridExtra)
```

As said before, this notebook addresses the task of automatizing diagnosis required by the reviewers in **comments 1), 2), 3)**.

# Sampling parameter values

Half input domain $S$

```{r, eval=TRUE, echo=TRUE}
S <- 3
S
```

Number of basis functions ($m$)

```{r, eval=TRUE, echo=TRUE}
m <- c(seq(2,40,1),seq(42,70,2),seq(75,130,5),seq(140,200,10))
m
```

Lenghtscale $l$ normalized by the half input domain $S$

```{r, eval=TRUE, echo=TRUE}
l <- seq(0.01, 8, by=0.01) / S
```

Boundary factor $c$

```{r, eval=TRUE, echo=TRUE}
c <- c(seq(1.125, 3, length.out=16), seq(4, 6, length.out= 3))
c
```

# Squared exponential covariance function: Empirical relationships among $m$, $l$ and $c$

## Loading calculated areas under the kernel curves

Area under the GP kernel curve

```{r, eval=TRUE, echo=TRUE}
load("area_gp.rData")
str(area_gp)
```

Difference (error) between the area under the GP kernel curve and the area under the HSGP kernel curve

```{r, eval=TRUE, echo=TRUE}
load("area_diff.rData")
str(area_diff)
```

### Error ratio

Proportion of the difference of areas (*area_diff*) within the area under the GP kernel curve (*area_gp*)

```{r eval=TRUE}
e_ratio <- area_diff/area_gp
str(e_ratio)
```

## Empirical relationships: Minimun $m$ given $l$ and $c$

Table with the number of basis functions $m$ needed as a function of $l$ and $c$.

```{r eval=TRUE}
obsM <- array(NA,c(length(l),length(c)))
for(i in 1:length(l)){
  for(j in 1:length(c)){
    obsM[i,j] <- if(which.max(e_ratio[i,,j]<0.005)==which.min(e_ratio[i,,j]<0.005))
                    NA 
                  else
                    m[which.max(e_ratio[i,,j]<0.01)]
  }
}
dimnames(obsM)[[2]] <- as.character(c)

obsM <- data.frame(l=l, obsM)
names(obsM) <- c("l", as.character(c))
round(obsM, 3)
```

Plot of the empirical relationships:

```{r empirical, fig.cap="Empirical relationships between m, l and c for the squared exponential kernel. Right side plot is in the log scale for m and l.", echo=FALSE, eval=TRUE, warning=FALSE, fig.height=4, fig.width=10}
obsM_melt <- melt(obsM, id.vars= "l", measure.vars= as.character(c), variable.name = "c", value.name = "m")
obsM_melt <- obsM_melt[obsM_melt$c %in% c(1.25,1.5,2,3,4,5,6) & obsM_melt$l < 8,] # %in% c(1.25,2,3,4)

p1 <- ggplot(obsM_melt, aes(l, m, color = as.factor(c))) +
  geom_line(size=0.8) + 
  theme_classic() + 
  theme(legend.position=c(.8,.7), axis.title.y = element_text(angle = 0, vjust=0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  scale_color_brewer("c", palette="PuBu") +
  labs(y = "m", x = "l")

p2 <- ggplot(obsM_melt, aes(log(l), log(m), color = as.factor(c))) +
  geom_line(size=0.8) +
  theme_classic() + 
  theme(legend.position="none", axis.title.y = element_text(angle = 0, vjust=0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  scale_color_brewer("c", palette="PuBu") +  #palette="OrRd"
  labs(y = "log(m)", x = "log(l)")

grid.arrange(p1, p2, ncol = 2)
```

### Relationship between allowed $l$ depending on $c$ and vice versa

As a function of $c$ there is a maximun $l$ valid for HSGP approach and, the other way around, as a function of $l$ there is a minimum $c$ valid for HSGP approach.

* Data of this relationship:

```{r eval=TRUE, echo=FALSE}
max_l_ind <- apply(!is.na(obsM[,!(names(obsM) %in% c("l","5","6"))]), 2, function(x) max(which(x)))
max_l <- l[max_l_ind]
max_l <- data.frame(c=c[!(c %in% c("5","6"))], l=max_l)
max_l
```

* Functions for this relationship:

```{r eval=TRUE, fig.height=3, fig.width=7, echo=FALSE}
model_1 <- lm(l ~ c, data=max_l)
# summary(model_1)

p1 <- ggplot(max_l, aes(c, l)) +
  geom_point(size=2) + 
  geom_line(data= max_l, aes(c, model_1$fitted.values), col="red", size=1) +
  theme_classic() + 
  theme(axis.title.y = element_text(angle = 0, vjust=0.5), plot.title = element_text(hjust = 0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  labs(y = "l", x = "c", title = TeX(c("$l \\, \\leq \\, -0.67 + 0.69 \\, c")))

model_2 <- lm(c ~ l, data=max_l)
# summary(model_2)

p2 <- ggplot(max_l, aes(l, c)) +
  geom_point(size=2) + 
  geom_line(data= max_l, aes(l, model_2$fitted.values), col="red", size=1) +
  theme_classic() + 
  theme(axis.title.y = element_text(angle = 0, vjust=0.5), plot.title = element_text(hjust = 0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  labs(y = "c", x = "l", title = TeX(c("$c \\, \\geq \\, 0.97 + 1.45 \\, l")))

grid.arrange(p1, p2, ncol = 2)
```

## Adjusted function for the empirical relationships {#sec-model}

Due to the number of basis functions $m$ is an integer variable and $l$ is a continuous variable, the relationship between $m$ and $l$ has a part that looks like stairs (see \@ref(fig:empirical)). In order to fit a function on these relationships it is better not to include this stairs like part in the fitting data set.

```{r echo=FALSE, eval=TRUE, warning=FALSE, fig.height=3, fig.width=5}
# Filtering vector to remove the stairs like part from the fitting data set
ind_stairs <- (obsM_melt$c==1.25 & log(obsM_melt$l)<(-2.4)) | (obsM_melt$c==1.5 & log(obsM_melt$l)<(-2.3)) | (obsM_melt$c==2 & log(obsM_melt$l)<(-2)) | (obsM_melt$c==3 & log(obsM_melt$l)<(-1.7)) | (obsM_melt$c==4 & log(obsM_melt$l)<(-1.5)) | (obsM_melt$c==5 & log(obsM_melt$l)<(-1.3)) | (obsM_melt$c==6 & log(obsM_melt$l)<(-1.2))
```

### Linear function: $\; log(m) \sim a + b \cdot log(l) + c \cdot log(c), \;$ where  $\; c \, \geq \, 0.97 + 1.45 \, l$

Model fit

```{r echo=FALSE, eval=TRUE, warning=FALSE, fig.height=3, fig.width=5, echo=FALSE}
obsM_melt$c <- as.numeric(as.character(obsM_melt$c))

model <- lm(log(m) ~ log(l) + log(c), data=obsM_melt[ind_stairs, ])
summary(model)

par(mgp=c(1.5,0.5,0), mai=c(0.7,0.5,0.3,0.1))
plot(model$model$'log(m)', model$residuals, cex=0.7)
abline(h=0, lty=2, col="gray")
```

Overlay the fitted functions to the empirical relationships

```{r echo=FALSE, eval=TRUE, fig.height=15, fig.width=15, warning=FALSE}
pred <- predict.lm(model, obsM_melt[!is.na(obsM_melt$m),], na.action = na.exclude)
pred <- as.data.frame(pred)

obsM_melt$type <- "Observed relationships"
pred$type <- "Fitted relationships"
model$model$type <- "Fitted relationships"

p1 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(log(l), log(m), group = as.factor(c), color = type)) +
  geom_line(size=1) + 
  geom_line(data=pred, aes(log(obsM_melt[!is.na(obsM_melt$m),]$l), pred, group = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=1, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.95), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=14), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 0.55 - 1.05 \\, log(l) + log(c) \\, \\right| \\, c \\, \\geq \\, 0.97 + 1.45 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=log(l), y=log(m) - pred$pred, color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(0.55 - 1.05 \\, log(l) + log(c)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(l, m, group = as.factor(c), color = type)) +
  geom_line(size=1) + 
  geom_line(data=pred, aes(obsM_melt[!is.na(obsM_melt$m),]$l, exp(pred), group = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=1, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.73,.85), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=14), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 1.73 \\, \\frac{c}{l^{1.05}} \\, \\right| \\, c \\, \\geq \\, 0.97 + 1.45 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=l, y=m - exp(pred$pred), color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\left(\\, 1.73 \\, \\frac{c}{l^{1.05}}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### Approximation of the fitted function by $2 \frac{c}{l}$

```{r echo=FALSE, eval=TRUE, fig.height=15, fig.width=15, warning=FALSE}
pred2 <- pred
pred2$pred = 0.69 - log(obsM_melt[!is.na(obsM_melt$m),]$l) + log(obsM_melt[!is.na(obsM_melt$m),]$c)

p1 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(log(l), log(m), fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(log(obsM_melt[!is.na(obsM_melt$m),]$l), pred, fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.65,.85), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 0.69 - log(l) + log(c) \\, \\right| \\, c \\, \\geq \\, 0.97 + 1.45 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=log(l), y=log(m) - pred2$pred, color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(0.69 - log(l) + log(c)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(l, m, fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(obsM_melt[!is.na(obsM_melt$m),]$l, exp(pred), fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.73,.85), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 2 \\, \\frac{c}{l} \\, \\right| \\, c \\, \\geq \\, 0.97 + 1.45 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=l, y=m - exp(pred2$pred), color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\, \\left(2 \\, \\frac{c}{l}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

# Mattern-$\nu=3/2$ covariance function: Empirical relationship among $m$, $l$ and $c$

## Loading calculated areas under the kernel curves

Area under the GP kernel curve

```{r, eval=TRUE, echo=TRUE}
load("area_gp_Matern.rData")
str(area_gp)
```

Difference (error) between the area under the GP kernel curve and the area under the HSGP kernel curve

```{r, eval=TRUE, echo=TRUE}
load("area_diff_Matern.rData")
str(area_diff)
```

### Error ratio

Proportion of the difference of areas (*area_diff*) within the area under the GP kernel curve (*area_gp*)

```{r eval=TRUE}
e_ratio <- area_diff/area_gp
str(e_ratio)
```

## Empirical relationships: minimun $m$ given $l$ and $c$

Table with the number of basis functions $m$ needed as a function of $l$ and $c$.

```{r eval=TRUE}
obsM <- array(NA,c(length(l),length(c)))
for(i in 1:length(l)){
  for(j in 1:length(c)){
    obsM[i,j] <- if(which.max(e_ratio[i,,j]<0.005)==which.min(e_ratio[i,,j]<0.005))
                    NA 
                  else
                    m[which.max(e_ratio[i,,j]<0.01)]
  }
}
dimnames(obsM)[[2]] <- as.character(c)

obsM <- data.frame(l=l, obsM)
names(obsM) <- c("l", as.character(c))
round(obsM, 3)
```

Plot of the empirical relationships:

```{r empirical-relationships-Mattern32, fig.cap="Empirical relationships between m, l and c for the Mattern-3/2 kernel. Right side plot is in the log scale for m and l.", echo=FALSE, eval=TRUE, warning=FALSE, fig.height=4, fig.width=10}
obsM_melt <- melt(obsM, id.vars= "l", measure.vars= as.character(c), variable.name = "c", value.name = "m")
obsM_melt <- obsM_melt[obsM_melt$c %in% c(1.5,2,2.5,3,4,5,6) & obsM_melt$l < 8,] # %in% c(1.25,2,3,4)

p1 <- ggplot(obsM_melt, aes(l, m, color = as.factor(c))) +
  geom_line(size=0.8) + 
  theme_classic() + 
  theme(legend.position="none", axis.title.y = element_text(angle = 0, vjust=0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  scale_color_brewer("c", palette="PuBu") +
  labs(y = "m", x = "l")

p2 <- ggplot(obsM_melt, aes(log(l), log(m), color = as.factor(c))) +
  geom_line(size=0.8) +
  theme_classic() + 
  theme(legend.position="none", axis.title.y = element_text(angle = 0, vjust=0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  scale_color_brewer("c", palette="PuBu") +  #palette="OrRd"
  labs(y = "log(m)", x = "log(l)") +
  coord_cartesian(ylim = c(0, 5))

grid.arrange(p1, p2, ncol = 2)
```

### Relationship between allowed $l$ depending on $c$ and vice versa

As a function of $c$ there is a maximun $l$ valid for HSGP approach and, the other way around, as a function of $l$ there is a minimum $c$ valid for HSGP approach.

* Data of this relationship:

```{r eval=TRUE, echo=FALSE}
max_l_ind <- apply(!is.na(obsM[,!(names(obsM) %in% c("l","5","6"))]), 2, function(x) max(which(x)))
max_l <- l[max_l_ind]
max_l <- data.frame(c=c[!(c %in% c("5","6"))], l=max_l)
max_l
```

* Functions for this relationship:

```{r eval=TRUE, fig.height=3, fig.width=7, echo=FALSE}
model_1 <- lm(l ~ c, data=max_l)
# summary(model_1)

p1 <- ggplot(max_l, aes(c, l)) +
  geom_point(size=2) + 
  geom_line(data= max_l, aes(c, model_1$fitted.values), col="red", size=1) +
  theme_classic() + 
  theme(axis.title.y = element_text(angle = 0, vjust=0.5), plot.title = element_text(hjust = 0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  labs(y = "l", x = "c", title = TeX(c("$l \\, \\leq \\, -0.49 + 0.50 \\, c")))

model_2 <- lm(c ~ l, data=max_l)
# summary(model_2)

p2 <- ggplot(max_l, aes(l, c)) +
  geom_point(size=2) + 
  geom_line(data= max_l, aes(l, model_2$fitted.values), col="red", size=1) +
  theme_classic() + 
  theme(axis.title.y = element_text(angle = 0, vjust=0.5), plot.title = element_text(hjust = 0.5), axis.title = element_text(size=12), axis.text = element_text(size=10)) +
  labs(y = "c", x = "l", title = TeX(c("$c \\, \\geq \\, 0.98 + 2.01 \\, l")))

grid.arrange(p1, p2, ncol = 2)
```

## Adjusted function for the empirical relationships {#sec-model-Mattern}

Due to the number of basis functions $m$ is an integer variable and $l$ is a continuous variable, the relationship between $m$ and $l$ has a part that looks like stairs (see \@ref(fig:empirical-relationships-Mattern32). In order to fit a function on these relationships it is better not to include this stairs like part in the fitting data set.

```{r echo=FALSE, eval=TRUE, warning=FALSE, fig.height=3, fig.width=5}
# Filtering vector to remove the stairs like part from the fitting data set
ind_stairs <- log(obsM_melt$m)>3.7
```

### Linear function: $log(m) \sim a + b \cdot log(l) + c \cdot log(c), \;$ where  $\; c \, \geq \, 0.98 + 2.01 \, l$

Model fit

```{r echo=FALSE, eval=TRUE, warning=FALSE, fig.height=3, fig.width=5}
obsM_melt$c <- as.numeric(as.character(obsM_melt$c))

model <- lm(log(m) ~ log(l) + log(c), data=obsM_melt[ind_stairs, ])
summary(model)

par(mgp=c(1.5,0.5,0), mai=c(0.7,0.5,0.3,0.1))
plot(model$fitted.values, model$residuals, cex=0.7)
abline(h=0, lty=2, col="gray")
```

Overlay the fitted model to the observed relationships

```{r echo=FALSE, eval=TRUE, fig.height=15, fig.width=15, warning=FALSE}
pred <- predict.lm(model, obsM_melt[!is.na(obsM_melt$m),], na.action = na.exclude)
pred <- as.data.frame(pred)

obsM_melt$type <- "Observed relationships"
pred$type <- "Fitted relationships"

p1 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(log(l), log(m), fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred, aes(log(obsM_melt[!is.na(obsM_melt$m),]$l), pred, fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.95), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 1.20 - 1.08 \\, log(l) + 0.99 \\, log(c) \\, \\right| \\, c \\, \\geq \\, 0.98 + 2.01 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM_melt, aes(x=log(l), y=log(m) - pred$pred, color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(1.20 - 1.08 \\, log(l) + 0.99 \\, log(c)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(l, m, fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred, aes(obsM_melt[!is.na(obsM_melt$m),]$l, exp(pred), fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.6), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 3.32 \\, \\frac{c^{0.99}}{l^{1.08}} \\, \\right| \\, c \\, \\geq \\, 0.98 + 2.01 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=l, y=m - exp(pred$pred), color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\left(\\, 3.32 \\, \\frac{c^{0.99}}{l^{1.08}}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### Approximation of the fitted model by $3.32 \frac{c}{l^{1.08}}$

```{r echo=FALSE, eval=TRUE, fig.height=15, fig.width=15, warning=FALSE}
pred2 <- pred
pred2$pred = 1.204 - 1.084*log(obsM_melt[!is.na(obsM_melt$m),]$l) + log(obsM_melt[!is.na(obsM_melt$m),]$c)

p1 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(log(l), log(m), fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(log(obsM_melt[!is.na(obsM_melt$m),]$l), pred, fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.95), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 1.20 - 1.08 \\, log(l) + \\, log(c) \\, \\right| \\, c \\, \\geq \\, 0.98 + 2.01 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=log(l), y=log(m) - pred2$pred, color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(1.20 - 1.08 \\, log(l) + \\, log(c)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(l, m, fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(obsM_melt[!is.na(obsM_melt$m),]$l, exp(pred), fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.6), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 3.32 \\, \\frac{c}{l^{1.08}} \\, \\right| \\, c \\, \\geq \\, 0.98 + 2.01 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=l, y=m - exp(pred2$pred), color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\left(\\, 3.32 \\, \\frac{c}{l^{1.08}}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### Approximation of the fitted model by $3.32 \frac{c}{l}$

```{r echo=FALSE, eval=TRUE, fig.height=15, fig.width=15, warning=FALSE}
pred2 <- pred
pred2$pred = 1.204 - log(obsM_melt[!is.na(obsM_melt$m),]$l) + log(obsM_melt[!is.na(obsM_melt$m),]$c)

p1 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(log(l), log(m), fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(log(obsM_melt[!is.na(obsM_melt$m),]$l), pred, fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.95), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 1.20 - \\, log(l) + \\, log(c) \\, \\right| \\, c \\, \\geq \\, 0.98 + 2.01 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=log(l), y=log(m) - pred2$pred, color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(1.20 - \\, log(l) + \\, log(c)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(l, m, fill = as.factor(c), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(obsM_melt[!is.na(obsM_melt$m),]$l, exp(pred), fill = as.factor(obsM_melt[!is.na(obsM_melt$m),]$c), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.6), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$\\left. 3.32 \\, \\frac{c}{l} \\, \\right| \\, c \\, \\geq \\, 0.98 + 2.01 \\, l", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM_melt[!is.na(obsM_melt$m),], aes(x=l, y=m - exp(pred2$pred), color = as.factor(c))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  facet_grid(rows=vars(c)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\left(\\, 3.32 \\, \\frac{c}{l}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

# Periodic covariance function: Empirical relationship among $m$ and $l$

## Setting parameters values

Number of basis functions ($m$)

```{r eval=TRUE, message=FALSE, size='small', warn=FALSE}
m <- c(seq(1,40,1),seq(42,70,2),seq(75,130,5),seq(140,200,10))
m
```

Lenghtscale ($l$)

```{r eval=TRUE, message=FALSE, size='small', warn=FALSE}
l <- c(seq(0.04,0.1,0.01),seq(0.12,2,0.02),seq(2.03,4,0.03),seq(4.04,5,0.04))
round(l,3)
```

## Loading calculated areas under the kernel curves

Area under the GP kernel curve

```{r, eval=TRUE, echo=TRUE}
load("area_gp_periodic.rData")
str(area_gp)
```

Difference (error) between the area under the GP kernel curve and the area under the HSGP kernel curve

```{r, eval=TRUE, echo=TRUE}
load("area_diff_periodic.rData")
str(area_diff)
```

### Error ratio

Proportion of the difference of areas (*area_diff*) within the area under the GP kernel curve (*area_gp*)

```{r eval=TRUE}
e_ratio <- area_diff/area_gp
str(e_ratio)
```

## Empirical relationships: minimun $m$ given $l$

```{r eval=TRUE}
obsM <- vector()
for(r in 1:length(l)){
    obsM[r] <- if(which.max(e_ratio[r,]<0.005)==which.min(e_ratio[r,]<0.005)){
                    if(which.max(e_ratio[r,]<0.005)==1){
                      1} 
                    else{
                      NA}}
                else{
                  m[which.max(e_ratio[r,]<0.005)]}
}

obsM <- data.frame(l=l, m=obsM)
round(obsM, 3)
```

Plot of the empirical relationships:

```{r empirical-relationships-periodic, fig.cap="Empirical relationships between m and l for the periodic kernel. Right side plot is in the log scale for m and l.", echo=FALSE, eval=TRUE, warning=FALSE, fig.height=4, fig.width=10}
p1 <- ggplot(obsM, aes(l, m)) +
  geom_line(size=0.8) + 
  theme_classic() + theme(legend.position=c(.8,.7), axis.title.y = element_text(angle = 0, vjust=0.5)) +
  scale_color_brewer(palette="PuBu") +
  labs(y = "m", x = "l") +
  coord_cartesian(ylim = c(0, 100))

p2 <- ggplot(obsM, aes(log(l), log(m))) +
  geom_line(size=0.8) +
  theme_classic() + theme(legend.position="none", axis.title.y = element_text(angle = 0, vjust=0.5)) +
  scale_color_brewer(palette="PuBu") +  #palette="OrRd"
  labs(y = "log(m)", x = "log(l)") +
  coord_cartesian(ylim = c(0, 5))

grid.arrange(p1, p2, ncol = 2)
```

## Adjusted function for the empirical relationships {#sec-model-periodic}

Due to the number of basis functions $m$ is an integer variable and $l$ is a continuous variable, the relationship between $m$ and $l$ has a part that looks like stairs (see \@ref(fig:empirical-relationships-periodic). In order to fit a function on these relationships it is better not to include this stairs like part in the fitting data set.

```{r echo=TRUE, eval=TRUE, warning=FALSE, fig.height=3, fig.width=5}
# Filtering vector to remove the stairs like part from the fitting data set
ind_stairs <- log(obsM$m)>2.3
```

### Linear model: $log(m) \sim a + b \cdot log(l)$

Model fit

```{r echo=FALSE, eval=TRUE, warning=FALSE, fig.height=3, fig.width=5}
model <- lm(log(m) ~ log(l), data=obsM[ind_stairs, ])
summary(model)

par(mgp=c(1.5,0.5,0), mai=c(0.7,0.5,0.3,0.1))
plot(model$fitted.values, model$residuals, cex=0.7)
abline(h=0, lty=2, col="gray")
```

Overlay the fitted model to the observed relationships

```{r echo=FALSE, eval=TRUE, fig.height=10, fig.width=15, warning=FALSE}
pred <- predict.lm(model, obsM, na.action = na.exclude)
pred <- as.data.frame(pred)

obsM$type <- "Observed relationships"
pred$type <- "Fitted relationships"

p1 <- ggplot(obsM, aes(log(l), log(m), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred, aes(log(obsM$l), pred, color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.95), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$1.20 - 1.05 \\, log(l)", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM, aes(x=log(l), y= log(m) - pred$pred)) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(1.20 - 1.05 \\, log(l)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM, aes(l, m, color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred, aes(obsM$l, exp(pred), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.75), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$3.31 \\, \\frac{1}{l^{1.05}}", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM, aes(x=l, y= m - exp(pred$pred))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=12)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\left(\\, 3.31 \\, \\frac{1}{l^{1.05}}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### Approximation of the fitted model by $3.31 \, \frac{1}{l}$

```{r echo=FALSE, eval=TRUE, fig.height=10, fig.width=15, warning=FALSE}
pred2 <- pred
pred2$pred = 1.20 - 1*log(obsM$l)

p1 <- ggplot(obsM, aes(log(l), log(m), color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(log(obsM$l), pred, color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.95), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$1.20 - log(l)", "Empirical relationships (m_{obs})")))) +
  labs(y = "log(m)", x = "log(l)")  +
  coord_cartesian(ylim = c(0, 6))

p2 <- ggplot(obsM, aes(x=log(l), y= log(m) - pred2$pred)) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  labs(x = "log(l)", y = TeX("$log(m_{obs}) \\, - \\, \\left(1.20 - log(l)\\right)$"), title = "Empirical - Fitted")

p3 <- ggplot(obsM, aes(l, m, color = type)) +
  geom_line(size=0.8) + 
  geom_line(data=pred2, aes(obsM$l, exp(pred), color = type), size=0.8, linetype=2) + 
  theme_classic() + 
  theme(legend.position=c(.5,.75), axis.title.y = element_text(angle = 0, vjust=0.5, size=16), axis.title.x = element_text(size=16), axis.text = element_text(size=16), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  scale_color_brewer("", palette="Set1", labels = unname(TeX(c("$3.31 \\, \\frac{1}{l}", "Empirical relationships (m_{obs})")))) +
  labs(y = "m", x = "l")  +
  coord_cartesian(ylim = c(0, 100))

p4 <- ggplot(obsM, aes(x=l, y= m - exp(pred2$pred))) +
  geom_point() +
  theme_bw() +
  theme(legend.position= "none", axis.title = element_text(size=16), axis.text = element_text(size=14)) +
  labs(x = "l", y = TeX(c("$m_{obs} \\, - \\left(3.31 \\, \\frac{1}{l}\\right)")), title = "Empirical - Fitted")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

# On whether there is theoretical evidence on the coefficients of the fitted functions.

As the approximate covariance function is a series expansion of eigenfunctions $\phi_j(x)$ scaled by the spectral density $S_{\ell}(\sqrt{\lambda_j})$:

$$
k(x,x') = \sum_{j=1}^m S_{\ell}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x')
$$

From Michael's work I extract that the rapid convergence of this series expansion mainly depends on how rapidly $S_{\ell}(\sqrt{\lambda_j})$ goes to zero as a function of $j$ \,($m$ is the number of basis functions), and depending on the lenght-scale $\ell$.

- $\sqrt{λ_j} = \frac{j \pi}{2L}$ is monotonically increasing in $j$. The velocity of the increase depends on $L$ (or equivalently on the boundary factor $c$).

- $S(\omega)= \alpha\,(\sqrt{2\pi})  \ell \exp(-\frac{1}{2}\ell^2 \omega^\intercal \omega)$, with $\omega=\sqrt{λ_j}$, goes to zero as $\omega$ increases.

```{r spectral-densities, echo=FALSE, fig.cap="", out.width = '50%', fig.align='center'}
knitr::include_graphics("img/spectral_densities.jpg")
```

- The $j$ component $\,S_{\ell}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = S_{\ell}(\sqrt{\lambda_j}) \sqrt{\frac{1}{L}} sin\left(\sqrt{\lambda_j}(x+L)\right) \sqrt{\frac{1}{L}} sin\left(\sqrt{\lambda_j}(x'+L)\right)$ becomes close to zero when $S_{\ell}(\sqrt{\lambda_j}) \frac{1}{L}$ is close to zero.

I guess a theoretical derivation of the minimum $m$ that makes the last component $S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x')$, with $j=m$, of the series expansion $k(x,x')$ equal to zero mi be quite complex. An empirical approximation to determine that minimum $m$ as a function of $\ell$ and $c$ might be more practical. However, I guess that it would mostly be equivalent to what we have done in our paper by looking at the differences between the areas under the curves of the exact and approximate covariance functions as a function $m$, $\ell$ and $c$.

# Next steps {#sec-next-steps}

1. Build a user guide with the iterative steps to perform the diagnosis.

2. Propose how much user has to increase $m$ in each iteration of the diagnosis.

### 2022-01-02 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
add 1st journal review

@@
 \end{tabular}
-\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
+\caption{Posterior mean predictive functions (left), posterior standard deviations (center) and  covariance functions (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
   \label{fig3_Post_part1}

### 2022-01-12 Paper/Manuscript_Stats&Comp/1st_journal_review/case1D_diagnosis/Rnb_case1D_diagnosis.Rmd
revision update: adding R-notebook with the diagnosis for 1D case study

---
title: '1D simulated case study: Diagnosis of the HSGP approximation'
date: "`r format(Sys.Date())`."
output:
  html_document:
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: true
    code_download: yes
    df_print: kable
---

<style type="text/css">
body, td{ font-size: 14px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
h1.title {
  font-size: 24px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 20px;
  color: Black;
}
h3 { /* Header 3 */
    font-size: 16px;
  color: Black;
}
h4 { /* Header 4 */
    font-size: 14px;
  color: Black;
}
#TOC {
  color: DarkBlue;
  font-size: 16px;
}
</style>

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

## Load packages

```{r message=FALSE, warn=FALSE, eval=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(mgcv)
library(brms)
library(RColorBrewer)
library(ggplot2)
library(loo)
library(latex2exp)
library(gridExtra)
library(reshape2)
library(plyr)
```

## Synthetic dataset

A synthetic dataset $\boldsymbol{y}$ with a true data generating function $f(x)$ which is a Gaussian process (GP) with a Matern-$\nu$=3/2 kernel. Errors $\boldsymbol{\epsilon}$ with noise $\sigma$ are added to $f(x)$ to form the final synthetic noisy dataset $\boldsymbol{y}$.

\begin{align*}
\begin{split}
\boldsymbol{y} &= \boldsymbol{f} + \boldsymbol{\epsilon} \\
\boldsymbol{\epsilon} &\sim \text{Normal}(\boldsymbol{0}, \sigma^2 \boldsymbol{I}) \\
f(x) &\sim \text{GP}(0, k(x, x', \theta)),
\end{split}
\end{align*}

where $k(x, x', \theta)$ is a Matern-$\nu$=3/2 kernel as a function of input values $x$ and parameters $\theta$ (marginal variance of the GP and lengthscale of the kernel)

### Draw the true data generative function $f(x)$ from a GP

#### Input predictive space

Vector of predictive input points `x_pred`, length of `x_pred` (`N_pred`) and vector of indices of `x_pred` (`vv_pred`).

```{r}
x_pred <- seq(-1,1,0.002)
N_pred <- length(x_pred)  
vv_pred <- 1:length(x_pred)
```

Half range of the input domain

```{r message=FALSE, warn=FALSE}
S <- abs((max(x_pred) - min(x_pred))/2)
S
```

#### Draw $f(x)$ using Stan

`param_v` identifies the order $\nu$ of the kernel

```{r}
param_v <- 3
```

Marginal variance `gpscale` and lengthscale `lscale`

```{r}
gpscale <- 1
lscale <- 0.2
```

Random variable `eta` to simulate from the GP

```{r}
seed <- 88646
set.seed(seed)
eta <- rnorm(N_pred, 0, 1)
```

Data to be passed to Stan

```{r}
standata <- list(x= x_pred, 
                  N= N_pred, 
                  param_v= param_v, 
                  gpscale= gpscale, 
                  lscale= lscale, 
                  eta= eta)
```

GP simulating model in Stan code

```{ echo=FALSE}
//saved in 'stancode_SIMU.stan'
functions {
	//GP with a SE kernel
	vector gp_SE(real[] x, real gpscale, real lscale, vector zgp) { 
		int n= size(x);
		matrix[n, n] cov;
		cov = cov_exp_quad(x, gpscale, lscale);
		for (i in 1:n)
			cov[i, i] = cov[i, i] + 1e-12;
		return cholesky_decompose(cov) * zgp;
	}
	//GP with a Matern kernel
	vector gp_mattern(real[] x, real gpscale, real lscale, vector zgp, int v) {
		int n= size(x);	
		matrix[n, n] cov;
		//Mattern 3/2
		if (v==3)
			for (i in 1:n)
				for (j in 1:n)
					cov[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x[i]-x[j])))/lscale) * exp(-sqrt(3)*(fabs(x[i]-x[j]))/lscale);
		//Mattern 5/2
		if (v==5)
			for (i in 1:n)
				for (j in 1:n)
					cov[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x[i]-x[j])))/lscale + (5*(fabs(x[i]-x[j]))^2)/(3*lscale^2)) * exp(-sqrt(5)*(fabs(x[i]-x[j]))/lscale);
		for (i in 1:n)
			cov[i, i] = cov[i, i] + 1e-12;
		return cholesky_decompose(cov) * zgp;
	}
}
data {
	int<lower=1> N;				
	real x[N];					
	int param_v;				//parameter v in the Mattern cov. fun.
	real lscale;
	real gpscale;
	vector[N] eta;
}
generated quantities{
	vector[N] f;
	f = gp_mattern(x, gpscale, lscale, eta, param_v);
}
```

Compiling the simulating model

```{r eval=TRUE}
simuGP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_SIMU.stan")
```

Model sampling

```{r eval=TRUE, results=FALSE}
fit_SIMU <- simuGP_mod$sample(data= standata,
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=1, thin=1, init=0.5,
                                fixed_param=TRUE)
fit_SIMU <- read_stan_csv(fit_SIMU$output_files())
```

Drawn $f(x)$

```{r eval=TRUE, fig.height=2, fig.width=5, echo=FALSE}
f_true <- summary(fit_SIMU, pars = "f", probs = c(0.025, 0.5, 0.975))$summary[,1]

ggplot(data.frame(f=f_true, x=x_pred), aes(x=x, y=f))  +
  geom_line(color = 'black', size=1) +
  theme_classic()
```

### Adding noise to $f(x)$

Five different noisy datasets $\boldsymbol{y}$ over true data generative function $f(x)$

```{r fig.height=2, fig.width=5}
seed <- 6734
set.seed(seed)

sd <- 0.2
y <- list()
y <- f_true + rnorm(N_pred, 0, sd)

ggplot(data.frame(f=f_true, x=x_pred), aes(x=x, y=f))  +
    geom_line(color = 'black', size=1) +
    geom_point(y= y, size=0.7, color='black') +
    theme_classic()
```

## Modeling data

A random sample `x_sample` (n=250) from the input predicting points `x_pred`

```{r fig.height=2, fig.width=5}
seed <- 6734
set.seed(seed)
vv_sample <- sort(sample(vv_pred, 250))
x_sample <- x_pred[vv_sample]
N_sample <- length(x_sample)

ggplot(data.frame(f=f_true[vv_sample], x=x_sample), aes(x=x, y=f))  +
    geom_line(color = 'black', size=1) +
    geom_point(y= y[vv_sample], size=0.9, color='black') +
    theme_classic()
```

### Interpolating data

Select samples from `x_sample` for interpolation

```{r}
seed <- 6734
set.seed(seed)
vv_inter <- vv_sample[which((x_sample>(-0.50)&x_sample<(-0.35))|(x_sample>(0.13)&x_sample<0.30))]
N_inter <- length(vv_inter)
x_inter <- x_pred[vv_inter]
```

### Extrapolating data

Select samples from `x_sample` for extrapolation

```{r}
vv_extra <- vv_sample[which(x_sample<(-0.8)|x_sample>0.82)]
N_extra <- length(vv_extra)
x_extra <- x_pred[vv_extra]
```

### Training data

Select samples from `x_sample` for training

```{r}
vv_train <- setdiff(vv_sample, c(vv_extra, vv_inter))
N_train <- length(vv_train)
x_train <- x_pred[vv_train]
```

### Plot of training (black), interpolation (red) and extrapolation (red) points

```{r fig.height=2, fig.width=5, echo=FALSE}
ggplot(data.frame(f=f_true[vv_sample], x=x_sample), aes(x=x, y=f), size=0.7, color='black')  +
  geom_line() +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= as.vector(y[vv_train]), x=x_train), size=0.7, color='black') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= as.vector(y[vv_inter]), x=x_inter), size=0.7, color='red') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= as.vector(y[vv_extra]), x=x_extra), size=0.7, color='red') +
  theme_classic()
```

## GP model fitting

Initialization of programming objects

```{r eval=TRUE}
fit_GP <- list()
```

### Data

Data to Stan

```{r}
standata_GP <- list(x_pred= x_pred, 
                     y= y, 
                     vv_train= vv_train, 
                     N_pred= N_pred, 
                     N_train= N_train, 
                     param_v= param_v,
                     f_true= f_true) 
```

### Model fitting

Exact GP model in Stan code

```{ echo=FALSE}
//saved in 'stancode_GP.stan'
functions {
	//SE kernel
	vector gp_SE(real[] x, real sdgp, real lscale, vector zgp) { 
		int n= size(x);
		matrix[n, n] cov;
		cov = gp_exp_quad_cov(x, sdgp, lscale);
		for (i in 1:n)
			cov[i, i] = cov[i, i] + 1e-12;
		return cholesky_decompose(cov) * zgp;
	}
	//Mattern kernel
	vector gp_mattern(real[] x, real sdgp, real lscale, vector zgp, int v) {
		int n= size(x);	
		matrix[n, n] cov;
		//Mattern 3/2
		if (v==3)
			for (i in 1:n)
				for (j in 1:n)
					cov[i,j] = sdgp^2 * (1 + (sqrt(3)*(fabs(x[i]-x[j])))/lscale) * exp(-sqrt(3)*(fabs(x[i]-x[j]))/lscale);
		//Mattern 5/2
		if (v==5)
			for (i in 1:n)
				for (j in 1:n)
					cov[i,j] = sdgp^2 * (1 + (sqrt(3)*(fabs(x[i]-x[j])))/lscale + (5*(fabs(x[i]-x[j]))^2)/(3*lscale^2)) * exp(-sqrt(5)*(fabs(x[i]-x[j]))/lscale);
		for (i in 1:n)
			cov[i, i] = cov[i, i] + 1e-12;
		return cholesky_decompose(cov) * zgp;
	}
	//SE kernel for prediction
	vector gp_SE_rng(real[] x2, vector f1, real[] x1, real gpscale, real lscale, real delta) {
	int N1 = rows(f1);
	int N2 = size(x2);
	vector[N2] f2;
	 {
	  matrix[N1, N1] K =   gp_exp_quad_cov(x1, gpscale, lscale) + diag_matrix(rep_vector(delta, N1));
	  matrix[N1, N1] L_K = cholesky_decompose(K);
	  vector[N1] L_K_div_f1 = mdivide_left_tri_low(L_K, f1);
	  vector[N1] K_div_f1 = mdivide_right_tri_low(L_K_div_f1', L_K)';
	  matrix[N1, N2] k_x1_x2 = gp_exp_quad_cov(x1, x2, gpscale, lscale);
	  vector[N2] f2_mu = (k_x1_x2' * K_div_f1);
	  matrix[N1, N2] v_pred = mdivide_left_tri_low(L_K, k_x1_x2);
	  matrix[N2, N2] cov_f2 =   gp_exp_quad_cov(x2, gpscale, lscale) - v_pred' * v_pred + diag_matrix(rep_vector(delta, N2));
	  f2 = multi_normal_rng(f2_mu, cov_f2);
	 }
	return f2;
	}
  //Mattern kernel for prediction
	vector gp_Matern_rng(real[] x2, vector f1, real[] x1, real gpscale, real lscale, real delta, real v) {
	int N1 = rows(f1);
	int N2 = size(x2);
	vector[N2] f2;
	matrix[N1, N1] K;
	matrix[N1, N2] k_x1_x2;
	matrix[N2, N2] cov_f2;
	matrix[N1, N1] L_K;
	vector[N1] L_K_div_f1;
	vector[N1] K_div_f1;
	vector[N2] f2_mu;
	matrix[N1, N2] v_pred;
	{
	  for (i in 1:N1)
  		for (j in i:N1){
  			if (v==3)
  			  K[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x1[i]-x1[j])))/lscale) * exp(-sqrt(3)*(fabs(x1[i]-x1[j]))/lscale);
  			if (v==5)
  			  K[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x1[i]-x1[j])))/lscale + (5*(fabs(x1[i]-x1[j]))^2)/(3*lscale^2)) * exp(-sqrt(5)*(fabs(x1[i]-x1[j]))/lscale);
  		}
	  for (i in 1:N1){
	     K[i,i] = K[i, i] + delta;
	     for (j in (i+1):N1)
			   K[j, i] = K[i, j];
	  }
	  L_K = cholesky_decompose(K);
	  L_K_div_f1 = mdivide_left_tri_low(L_K, f1);
	  K_div_f1 = mdivide_right_tri_low(L_K_div_f1', L_K)';
	  for (i in 1:N1)
  		for (j in 1:N2){
  			if (v==3)
  			  k_x1_x2[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x1[i]-x2[j])))/lscale) * exp(-sqrt(3)*(fabs(x1[i]-x2[j]))/lscale);
  			if (v==5)
  			  k_x1_x2[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x1[i]-x2[j])))/lscale + (5*(fabs(x1[i]-x2[j]))^2)/(3*lscale^2)) * exp(-sqrt(5)*(fabs(x1[i]-x2[j]))/lscale);
  		}
	  f2_mu = (k_x1_x2' * K_div_f1); //'
	  v_pred = mdivide_left_tri_low(L_K, k_x1_x2);
	  for (i in 1:N2)
  		for (j in i:N2){
  			if (v==3)
  			  cov_f2[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x2[i]-x2[j])))/lscale) * exp(-sqrt(3)*(fabs(x2[i]-x2[j]))/lscale);
  			if (v==5)
  			  cov_f2[i,j] = gpscale^2 * (1 + (sqrt(3)*(fabs(x2[i]-x2[j])))/lscale + (5*(fabs(x2[i]-x2[j]))^2)/(3*lscale^2)) * exp(-sqrt(5)*(fabs(x2[i]-x2[j]))/lscale);
  		}
	  for (i in 1:N2) {
	     for (j in (i+1):N2) {
			cov_f2[j, i] = cov_f2[i, j];
		 }
	  }

	  cov_f2 =   cov_f2 - v_pred' * v_pred + diag_matrix(rep_vector(delta, N2)); //'
	  f2 = multi_normal_rng(f2_mu, cov_f2);
	}
	return f2;
	}
}
data {
	int<lower=1> N_train;			
	int<lower=1> N_pred;					
	real x_pred[N_pred];			
	vector[N_pred] y;	
	vector[N_pred] f_true;	
	int param_v;	
	int vv_train[N_train];			
}
parameters {
	real<lower=0> lscale;
	real<lower=0> sigma;
	real<lower=0> gpscale;
	vector[N_train] eta;
}
transformed parameters{
	vector[N_train] f;
	f = gp_mattern(x_pred[vv_train], gpscale, lscale, eta, param_v);
	// f = gp_SE(x[vv_train], gpscale, lscale, eta);	
}
model{
	eta ~ normal(0, 1);
	lscale ~ gamma(2.5, 10); //gamma(3.75, 25); //normal(0,1); //
	sigma ~ gamma(1, 1);
	gpscale ~ gamma(5, 5);
	y[vv_train] ~ normal(f, sigma); 
}
generated quantities{
	vector[N_pred] f_pred;
	vector[N_pred] lpd_f;
	vector[N_pred] lpd_y;
	f_pred = gp_Matern_rng(x_pred, f, x_pred[vv_train], gpscale, lscale, 1e-12, param_v);
	for(i in 1:N_pred){
		lpd_y[i] = normal_lpdf(y[i] | f_pred[i], sigma);
		lpd_f[i] = normal_lpdf(f_true[i] | f_pred[i], sigma);
	}
}
```

Compilation of the exact GP model

```{r eval=TRUE}
GP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_GP.stan")
```

Model fitting

```{r eval=TRUE}
load("fit_GP.rData")
```

```{r eval=TRUE}
# fit_GP <- GP_mod$sample(data= standata_GP,
#                           iter_warmup=50,
#                           iter_sampling=50,
#                           chains=4, thin=1, init=0.5)
# fit_GP <- read_stan_csv(fit_GP$output_files())
```

Summaries of variable estimates

```{r message=FALSE, warn=FALSE, eval=TRUE, echo=FALSE}
param = c("lscale","gpscale","sigma")
print(summary(fit_GP, pars = param, probs = c(0.025, 0.5, 0.975))$summary)
```

Simulation chains for the variables after warmup

```{r message=FALSE, warn=FALSE, fig.height=3, fig.width=15, eval=TRUE, echo=FALSE}
traceplot(fit_GP, pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Posterior predictive distributions

```{r include=TRUE, echo=TRUE, echo=FALSE}
f_pred_GP <- summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
time_GP <- sum(get_elapsed_time(fit_GP)[2])/(fit_GP@sim$iter - fit_GP@sim$warmup)
```

```{r fig.height=4, fig.width=10, include=TRUE, echo=FALSE}
df_pred_GP <- data.frame(as.data.frame(f_pred_GP), x=x_pred)
ggplot(df_pred_GP, aes(x=x, y=mean))  +
  geom_line(size=1, color='red') +
  geom_ribbon(aes(ymin=X2.5., ymax=X97.5., color=NA), fill='red', size=0.7, linetype=2, alpha=0.1) +
  # geom_line(mapping=aes(x=x, y=f), data=data.frame(f=f_true[vv_sample], x=x_sample), size=0.7, color='black') +  //true function
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=0.9, color='black') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=0.9, color='red') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=0.9, color='red') +
  scale_color_brewer(palette="Set1") +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12))
```

## Diagnosis

### Relationships among $m$, $l$ and $c$

Relationship among the number of basis functions $m$, the lengthscale $l$ and the boundary factor $c$ for a **Matern-$\nu=3/2$ kernel**:

$$ m= 3.32 \frac{c}{\left(\frac{l}{S}\right)^{1.08}} \;\; \Leftrightarrow \;\; \frac{l}{S}= 3.04 \, \frac{c^{0.926}}{m^{0.926}}$$

```{r}
m_MAT32 <- function(c,l,S){ ceiling(3.32 * c / (l/S)^1.08) }

l_MAT32 <- function(c,m,S){ round(S * 3.04 * c^0.926 / m^0.926, 3) }
```

with $$c \, \geq \, 0.98 + 2.01 \, \frac{l}{S}$$

as a function of $l$ and $S$, where $S$ is the half range of the input domain,

```{r}
c_vs_l_MAT32 <- function(l,S){
  if(0.98+2.01*l/S < 1.25){
    c = 1.3
  }else{
    if(0.98+2.01*l/S >= 1.25 & 0.98+2.01*l/S < 1.45){
      c = 1.5
    }else{
      if(0.98+2.01*l/S >= 1.45 & 0.98+2.01*l/S < 1.95){
        c = 2
      }else{
        if(0.98+2.01*l/S >= 1.95 & 0.98+2.01*l/S < 2.45){
          c = 2.5
        }else{
          if(0.98+2.01*l/S >= 2.45 & 0.98+2.01*l/S < 2.95){
            c = 3
          }else{
            c = ceiling(0.98+2.01*l/S)
          } } } } }
  c
}
```

or, equivalently, $$c \, \geq \, 31868.07 \, m^{-0.172 \, log(m)^2 \, + \,  1.963 \, log(m) \,  - \, 7.634}$$

as a function of $m$.

```{r}
c_vs_m_MAT32 <- function(m){ 
  if(31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) < 1.25){
    c = 1.3
  }else{
    if(31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) >= 1.25 & 31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) < 1.45){
      c = 1.5
    }else{
      if(31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) >= 1.45 & 31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) < 1.95){
        c = 2
      }else{
        if(31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) >= 1.95 & 31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) < 2.45){
          c = 2.5
        }else{
          if(31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) >= 2.45 & 31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338) < 2.95){
            c = 3
          }else{
            c = ceiling(31868.07*m^(-0.1722*log(m)^2+1.9633*log(m)-7.6338))
          } } } } }
  c
}
```

### User-guide for diagnosis

**Assumption of the diagnosis tool:**

- Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is (always) smaller than the true lengthscale $l$.

**User-guide with the steps to perform diagnosis:**

0. Make a first guess of the lengthscale $l_1$ of the function to be learned.

*Iteration 1*

1. Obtain the minimum valid boundary factor $c_1$ determined by $l_1$ using the functional relationship $c_1 \geq \, 0.98 +  2.01 \, \frac{l_1}{S}$, where $S$ is the half range of the input domain.

2. Obtain the mimimum valid number of basis functions $m_1$ determined by $l_1$ and $c_1$ using the functional relationship $m_1= 3.32 \, \frac{c_1}{\left(\frac{l_1}{S}\right)^{1.08}}$. Notice that $l_1$ can also be read as the minimum lengthscale that can be accurately fitted determined by $m_1$ and $c_1$ $\left(\frac{l_1}{S}= 3.04 \, \frac{(c_1)^{0.926}}{(m_1)^{0.926}}\right)$.

3. Fit the HSGP model and assess residuals: check any residual trends, compute *rmse* (Root mean square error), $R^2$ (Coefficient of variation) and *elpd* (Expected log predicitve density).

4. Check the diagnostic of whether $\, \hat{l_1} \geq l_1$.

    4.1. If the diagnostic is TRUE, the HSGP model approximation must be sufficiently accurate or, at least, very close to be sufficiently accurate. Then user can continue with step 5.

    4.2. If the diagnostic is FALSE, the HSGP model approximation can not be sufficiently accurate. Then user must move to step 10.

*Iteration 2*

*(If the diagnostic in step 4 is TRUE)*

5. Set $\, m_2 = m_1 + 10$.

6. Obtain the minimum valid boundary factor $c_2$ determined by the estimated $\hat{l_1}$ using the functional relationship $c_2 \, \geq \, 0.98 + 2.01 \, \frac{\hat{l_1}}{S}$, where $S$ is the half range of the input domain.

7. Obtain the mimimum lengthscale $l_2$ that can be accurately fitted determined by $m_2$ and $c_2$ using the functional relationship $\frac{l_2}{S}= 3.04 \, \frac{(c_2)^{0.926}}{(m_2)^{0.926}}$.

8. Fit the HSGP model and

    8.1. Check whether $\, \hat{l_2} \geq \hat{l_1}$

    8.2. Check stability in *rmse*, $R^2$ and *elpd* from previous iteration

9. If the two verifications above are TRUE, the HSGP model approximation must be sufficiently accurate, and diagnosis ends here. Otherwise, repeat steps 5-9 and update parameters.

*(If the diagnostic in step 4 is FALSE)*

10. Set $\, l_2 = \hat{l_1}$.

11. Repeat steps 1-4 and update parameters.

### Iteration 1

Initializing programming objects:

```{r message=FALSE, warn=FALSE}
standata <- list()
fit_diagnosis <- list()
l <- vector()
c <- vector()
m <- vector()
l_hat <- vector()
residual <- list()
evaluation <- list()
diagnosis <- list()
rmse <- vector()
eR2 <- vector()
elpd <- vector()
```

Iteration index

```{r message=FALSE, warn=FALSE}
i <- 1
```

#### Setting $m$, $l$ and $c$

1. Making the first guess that the lengthscale $l_1$ might be 0.5

```{r message=FALSE, warn=FALSE}
l[i] <- 0.5
l[i]
```

2. The boundary factor $c_1$ has to fulfill $c_1 \, \geq \,0.98 + 2.01 \, \frac{l_1}{S}$

```{r message=FALSE, warn=FALSE}
c[i] <- c_vs_l_MAT32(l=l[i], S=S)
c[i] 
```

3. The number of basis functions $m_1$ as a function of $l_1$ and $c_1$:  $\; m_1 = 3.32 \, \frac{c_1}{(l_1/S)^{1.08}}$

```{r message=FALSE, warn=FALSE}
m[i] <- m_MAT32(c=c[i], l=l[i], S=S)
m[i]
```

#### Data to Stan

```{r message=FALSE, warn=FALSE}
standata[[i]] <- list(y= y, 
                       f_true= f_true, 
                       x_pred= x_pred, 
                       vv_train= vv_train, 
                       N_pred= N_pred, 
                       N_train= N_train, 
                       param_v= param_v, 
                       c= c[i], 
                       L= c[i]*S, 
                       M= m[i])
str(standata[[i]])
```

#### Model fitting

HSGP model in Stan code

```{ echo=FALSE}
//saved in 'stancode_HSGP.stan'
functions {
	real lambda(real L, int m) {
		real lam;
		lam = ((m*pi())/(2*L))^2;
		return lam;
	}
	real spd_SE(real gpscale, real lscale, real w) {
		real S;
		S = (gpscale^2) * sqrt(2*pi()) * lscale * exp(-0.5*(lscale^2)*(w^2));
		return S;
	}
	real spd_Matern(real gpscale, real lscale, real w, int v) {
		real S;
		//Mattern 3/2
		if (v==3)
			S = 4*gpscale^2 * (sqrt(3)/lscale)^3 * 1/((sqrt(3)/lscale)^2 + w^2)^2;
		//Mattern 5/2
		if (v==5)
			S = 16.0/3*gpscale^2 * (sqrt(5)/lscale)^5 * 1/((sqrt(5)/lscale)^2 + w^2)^3;
		return S;
	}
	vector phi(real L, int m, vector x) {
		vector[rows(x)] fi;
		fi = 1/sqrt(L) * sin(m*pi()/(2*L) * (x+L));
		return fi;
	}
}
data {
	int<lower=1> N_pred;
	int<lower=1> N_train;
	vector[N_pred] x_pred;			
	vector[N_pred] y;
	vector[N_pred] f_true;	
	int param_v;	
	int vv_train[N_train];
	real L;	
	int<lower=1> M;	
}
transformed data{
	//Basis functions for f
	matrix[N_train,M] PHI_train;
	matrix[N_pred,M] PHI_pred;
	for (m in 1:M){
	  PHI_train[,m] = phi(L, m, x_pred[vv_train]);// - mean(phi(L, m, x_pred[vv_train]));
	  PHI_pred[,m] = phi(L, m, x_pred);// - mean(phi(L, m, x_pred));	
	}
}
parameters {
  // real intercept;
	real<lower=0> lscale;
	real<lower=0> noise;
	real<lower=0> gpscale;
	vector[M] beta;
}
transformed parameters{
	vector[N_train] f;
	vector[M] diagSPD;
	vector[M] SPD_beta;
	//Spectral densities for f
	for(m in 1:M){ 
		diagSPD[m] =  sqrt(spd_Matern(gpscale, lscale, sqrt(lambda(L, m)), param_v)); 
		// diagSPD[m] =  sqrt(spd_SE(gpscale, lscale, sqrt(lambda(L, m))));
	}
	SPD_beta = diagSPD .* beta;
	f= PHI_train[,] * SPD_beta;
}
model{
	lscale ~ gamma(2.5, 10); //gamma(3.75, 25); //normal(0,1); //gamma(3, 5); //
	noise ~ gamma(1, 1); //normal(0,1); //
	gpscale ~ gamma(5, 5); //normal(0,1); //
	beta ~ normal(0, 1);
	// intercept ~ normal(0, 2);
	target += normal_lpdf(y[vv_train] | f, noise);
}
generated quantities{
	vector[N_pred] f_pred;
	vector[N_pred] lpd_f;
	vector[N_pred] lpd_y;
	f_pred= PHI_pred[,] * SPD_beta;
	for(i in 1:N_pred){
		lpd_y[i] = normal_lpdf(y[i] | f_pred[i], noise);
		lpd_f[i] = normal_lpdf(f_true[i] | f_pred[i], noise);
	}
}
```

Compiling the HSGP model

```{r message=FALSE, warn=TRUE, eval=TRUE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP.stan")
```

Model fitting

```{r eval=TRUE, message=FALSE, include=TRUE, warn=TRUE, results=FALSE}
fit_diagnosis[[i]] <- hsgp_mod$sample(data= standata[[i]], 
                                iter_warmup=1000, 
                                iter_sampling=1000, 
                                chains=4, thin=10, init=0.5)
fit_diagnosis[[i]] <- rstan::read_stan_csv(fit_diagnosis[[i]]$output_files())
```

Summaries of variable estimates

```{r message=FALSE, warn=FALSE, eval=TRUE, echo=FALSE}
param = c("lscale","gpscale","noise")
summary(fit_diagnosis[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
```

Simulation chains for the variables after warmup

```{r message=FALSE, warn=FALSE, fig.height=3, fig.width=15, eval=TRUE, echo=FALSE}
traceplot(fit_diagnosis[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior distributions

```{r include=TRUE, echo=TRUE, echo=FALSE}
f <- summary(fit_diagnosis[[i]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
```

```{r fig.height=4, fig.width=10,  include=TRUE, echo=TRUE, echo=FALSE}
f_GP <- data.frame(as.data.frame(f_pred_GP), x=x_pred, type="GP")
f_HSGP <- data.frame(as.data.frame(f), x=x_pred, type="HSGP")
f_HSGP_GP <- rbind(f_GP, f_HSGP)

ggplot(f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=type, color=type))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

#### Model evaluation

Residuals `residual`, Bayesian $R^2$ (Coefficient of determination), and Log predictive density `lpd`

```{r fig.height=2, fig.width=10, message=FALSE, warn=FALSE, echo=FALSE}
#Residuals
f <- summary(fit_diagnosis[[i]], pars = c("f"))$summary[,1]
residual[[i]] <- standata[[i]]$y[vv_train] - f

p1 <- ggplot(as.data.frame(residual[[i]]), aes(residual[[i]]))  +
  geom_histogram(color = 'white') +
  labs(x="residual") +
  theme_classic()

#Bayesian $R^2$
noise <- as.matrix(fit_diagnosis[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit_diagnosis[[i]], pars = c("f")), 1, sd)
R2 <- sd_f^2/(sd_f^2 + noise^2)

p2 <- ggplot(as.data.frame(R2), aes(R2))  +
  geom_histogram(color = 'white') +
  theme_classic()

#Log predictive density (lpd)
lpd <- summary(fit_diagnosis[[i]], pars = c("lpd_y"))$summary[,1]

p3 <- ggplot(as.data.frame(lpd), aes(lpd))  +
  geom_histogram(color = 'white') +
  theme_classic()

grid.arrange(p1, p2, p3, ncol=3)
```

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination), and expected log predictive density `elpd`

```{r fig.height=3, fig.width=5, message=FALSE, warn=FALSE, echo=FALSE}
rmse[i] <- round(sqrt(mean(residual[[i]]^2)), 3)
eR2[i] <- round(median(R2), 3)
elpd[i] <- round(median(lpd), 3)

evaluation[[i]] <- data.frame(iter= i,
                                rsme = rmse[i],
                                R2 = eR2[i],
                                elpd = elpd[i]
                                )
evaluation[[i]]
```

#### Diagnosis

1. Estimated lengthscale $\hat{l_1}$

```{r message=FALSE, warn=FALSE}
l_hat[i] <- round(summary(fit_diagnosis[[i]], pars = "lscale")$summary[,1], 3)
l_hat[i]
```

2. Check the diagnostic of whether $\, \hat{l_1} \geq l_1$

```{r message=FALSE, warn=FALSE}
l_hat[i] >= l[i]
```

**Summary table**

```{r fig.height=3, fig.width=5, message=FALSE, warn=FALSE, echo=FALSE}
diagnosis[[i]] <- data.frame(iter= i,
                              GP_func= c("f"),
                              l= l[i],
                              c= c[i],
                              m= m[i],
                              l_hat= l_hat[i],
                              'l_hat > l' = l_hat[i] >= l[i],
                              rsme = rmse[i],
                              R2 = eR2[i],
                              elpd = elpd[i]
                              )
names(diagnosis[[i]]) <- c("iter", "GP_func", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "R2", "elpd")
diagnosis[[i]]
```

### Iteration 2

Iteration index

```{r message=FALSE, warn=FALSE}
i <- 2
```

#### Setting $m$, $l$ and $c$

1. If the diagnostic $\hat{l_1} \geq l_1$ is FALSE:

    1.1. Update $l_2 = \hat{l_1}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  print("Diagnoistic is TRUE")
}else{
  l[i] <- l_hat[i-1]
  l[i]
}
```

1.2. The boundary factor $c_2$ has to fulfill $c_2 \, \geq \, 0.98 + 2.01 \, \frac{l_2}{S}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  print("Diagnoistic is TRUE")
}else{
  c[i] <- c_vs_l_MAT32(l=l[i], S=S)
  c[i]
}
```

1.3. The number of basis functions $m_2$ as a function of $l_2$ and $c_2$:  $\; m_2= 3.32 \, \frac{c_2}{(l_2/S)^{1.08}}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  print("Diagnoistic is TRUE")
}else{
  m[i] <- m_MAT32(c=c[i], l=l[i], S=S)
  m[i]
}
```

2. If the diagnostic $\hat{l_1} \geq l_1$ is TRUE:

    2.1. Update $m_2 = m_1 + 10$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  m[i] <- m[i-1] + 10
  m[i]
}else{
  print("Diagnoistic is FALSE")
}
```

2.2. The boundary factor $c_2$ has to fulfill $c_2 \, \geq \, 31868.07 \, (m_2)^{-0.172 \, log(m_2)^2 \, + \, 1.963 \, log(m_2) \,  - \, 7.634}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  c[i] <- c_vs_m_MAT32(m=m[i])
  c[i]
}else{
  print("Diagnoistic is FALSE")
}
```

2.3. The minimum valid lengthscale $l_2$ determined by $m_2$ and $c_2$:  $\; l_2= S \cdot 3.04 \, \frac{(c_2)^{0.926}}{(m_2)^{0.926}}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  l[i] <- l_MAT32(c=c[i], m=m[i], S=S)
  l[i]
}else{
  print("Diagnoistic is FALSE")
}
```

#### Data to Stan

```{r message=FALSE, warn=FALSE}
standata[[i]] <- list(y= y, 
                       f_true= f_true, 
                       x_pred= x_pred, 
                       vv_train= vv_train, 
                       N_pred= N_pred, 
                       N_train= N_train, 
                       param_v= param_v, 
                       c= c[i], 
                       L= c[i]*S, 
                       M= m[i])
str(standata[[i]])
```

#### Model fitting

Compiling the HSGP model

```{r message=FALSE, warn=TRUE, eval=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP.stan")
```

Model fitting

```{r warn=TRUE, message=FALSE, eval=TRUE, include=TRUE, results=FALSE}
fit_diagnosis[[i]] <- hsgp_mod$sample(data= standata[[i]], 
                                iter_warmup=1000, 
                                iter_sampling=1000, 
                                chains=4, thin=10, init=0.5)
fit_diagnosis[[i]] <- rstan::read_stan_csv(fit_diagnosis[[i]]$output_files())
```

Summaries of variable estimates

```{r message=FALSE, warn=FALSE, eval=TRUE, echo=FALSE}
param = c("lscale","gpscale","noise")
summary(fit_diagnosis[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
```

Simulation chains for the variables after warmup

```{r message=FALSE, warn=FALSE, fig.height=3, fig.width=15, eval=TRUE, echo=FALSE}
traceplot(fit_diagnosis[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior distributions

```{r include=TRUE, echo=TRUE, echo=FALSE}
f <- summary(fit_diagnosis[[i]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
```

```{r fig.height=4, fig.width=10,  include=TRUE, echo=FALSE}
f_GP <- data.frame(as.data.frame(f_pred_GP), x=x_pred, type="GP")
f_HSGP <- data.frame(as.data.frame(f), x=x_pred, type="HSGP")
f_HSGP_GP <- rbind(f_GP, f_HSGP)

ggplot(f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=type, color=type))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

#### Model evaluation

Residuals `residual`, Bayesian $R^2$ (Coefficient of determination), and Log predictive density `lpd`

```{r fig.height=2, fig.width=10, message=FALSE, warn=FALSE, echo=FALSE}
#Residuals
f <- summary(fit_diagnosis[[i]], pars = c("f"))$summary[,1]
residual[[i]] <- standata[[i]]$y[vv_train] - f

p1 <- ggplot(as.data.frame(residual[[i]]), aes(residual[[i]]))  +
  geom_histogram(color = 'white') +
  labs(x="residual") +
  theme_classic()

#Bayesian $R^2$
noise <- as.matrix(fit_diagnosis[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit_diagnosis[[i]], pars = c("f")), 1, sd)
R2 <- sd_f^2/(sd_f^2 + noise^2)

p2 <- ggplot(as.data.frame(R2), aes(R2))  +
  geom_histogram(color = 'white') +
  theme_classic()

#Log predictive density (lpd)
lpd <- summary(fit_diagnosis[[i]], pars = c("lpd_y"))$summary[,1]

p3 <- ggplot(as.data.frame(lpd), aes(lpd))  +
  geom_histogram(color = 'white') +
  theme_classic()

grid.arrange(p1, p2, p3, ncol=3)
```

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination), and expected log predictive density `elpd`

```{r fig.height=3, fig.width=5, message=FALSE, warn=FALSE, echo=FALSE}
rmse[i] <- round(sqrt(mean(residual[[i]]^2)), 3)
eR2[i] <- round(median(R2), 3)
elpd[i] <- round(median(lpd), 3)

evaluation[[i]] <- data.frame(iter= i,
                                rsme = rmse[i],
                                R2 = eR2[i],
                                elpd = elpd[i]
                                )
rbind(evaluation[[i-1]],evaluation[[i]])
```

#### Diagnosis

1. Estimated lengthscale $\hat{l_2}$

```{r message=FALSE, warn=FALSE}
l_hat[i] <- round(summary(fit_diagnosis[[i]], pars = "lscale")$summary[,1], 3)
l_hat[i]
```

2. Check the diagnostic of whether $\, \hat{l_2} \geq l_2$

```{r message=FALSE, warn=FALSE}
l_hat[i] >= l[i]
```

**Summary table**

```{r fig.height=3, fig.width=5, message=FALSE, warn=FALSE, echo=FALSE}
diagnosis[[i]] <- data.frame(iter= i,
                                GP_func= c("f"),
                                l= l[i],
                                c= c[i],
                                m= m[i],
                                l_hat= l_hat[i],
                                'l_hat > l' = l_hat[i] >= l[i],
                                rsme = rmse[i],
                                R2 = eR2[i],
                                elpd = elpd[i]
                                )
names(diagnosis[[i]]) <- c("iter", "GP_func", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "R2", "elpd")

diagnosis_summ <- rbind(diagnosis[[1]],diagnosis[[2]])
diagnosis_summ
```

### Iteration 3

Iteration index

```{r message=FALSE, warn=FALSE}
i <- 3
```

#### Setting $m$, $l$ and $c$

1. If the diagnostic $\hat{l_2} \geq l_2$ is FALSE:

    1.1. Update $l_3 = \hat{l_2}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  print("Diagnoistic is TRUE")
}else{
  l[i] <- l_hat[i-1]
  l[i]
}
```

1.2. The boundary factor $c_3$ has to fulfill $c_3 \, \geq \, 0.98 + 2.01 \, \frac{l_3}{S}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  print("Diagnoistic is TRUE")
}else{
  c[i] <- c_vs_l_MAT32(l=l[i], S=S)
  c[i]
}
```

1.3. The number of basis functions $m_3$ as a function of $l_3$ and $c_3$:  $\; m_3= 3.32 \, \frac{c_3}{(l_3/S)^{1.08}}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  print("Diagnoistic is TRUE")
}else{
  m[i] <- m_MAT32(c=c[i], l=l[i], S=S)
  m[i]
}
```

2. If the diagnostic $\hat{l_2} \geq l_2$ is TRUE:

    2.1. Update $m_3 = m_2 + 10$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  m[i] <- m[i-1] + 10
  m[i]
}else{
  print("Diagnosis is FALSE")
}
```

2.2. The boundary factor $c_3$ has to fulfill $c_3 \, \geq \, 31868.07 \, (m_3)^{-0.172 \, log(m_3)^2 \, + \,  1.963 \, log(m_3) \,  - \, 7.634}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  c[i] <- c_vs_m_MAT32(m=m[i])
  c[i]
}else{
  print("Diagnosis is FALSE")
}
```

2.3. The minimum valid lengthscale $l_3$ determined by $m_3$ and $c_3$:  $\; l_3= S \cdot 3.04 \, \frac{(c_3)^{0.926}}{(m_3)^{0.926}}$

```{r message=FALSE, warn=FALSE}
if(l_hat[i-1] >= l[i-1]){
  l[i] <- l_MAT32(c=c[i], m=m[i], S=S)
  l[i]
}else{
  print("Diagnosis is FALSE")
}
```

#### Data to Stan

```{r message=FALSE, warn=FALSE}
standata[[i]] <- list(y= y, 
                       f_true= f_true, 
                       x_pred= x_pred, 
                       vv_train= vv_train, 
                       N_pred= N_pred, 
                       N_train= N_train, 
                       param_v= param_v, 
                       c= c[i], 
                       L= c[i]*S, 
                       M= m[i])
str(standata[[i]])
```

#### Model fitting

Compiling the HSGP model

```{r message=FALSE, warn=TRUE, eval=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP.stan")
```

Model fitting

```{r warn=TRUE, message=FALSE, eval=TRUE, include=TRUE, results=FALSE}
fit_diagnosis[[i]] <- hsgp_mod$sample(data= standata[[i]], 
                                iter_warmup=1000, 
                                iter_sampling=1000, 
                                chains=4, thin=10, init=0.5)
fit_diagnosis[[i]] <- rstan::read_stan_csv(fit_diagnosis[[i]]$output_files())
```

Summaries of variable estimates

```{r message=FALSE, warn=FALSE, eval=TRUE, echo=FALSE}
param = c("lscale","gpscale","noise")
summary(fit_diagnosis[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
```

Simulation chains for the variables after warmup

```{r message=FALSE, warn=FALSE, fig.height=3, fig.width=15, eval=TRUE, echo=FALSE}
traceplot(fit_diagnosis[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior distributions

```{r include=TRUE, echo=TRUE, echo=FALSE}
f <- summary(fit_diagnosis[[i]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
```

```{r fig.height=4, fig.width=10,  include=TRUE, echo=TRUE, echo=FALSE}
f_GP <- data.frame(as.data.frame(f_pred_GP), x=x_pred, type="GP")
f_HSGP <- data.frame(as.data.frame(f), x=x_pred, type="HSGP")
f_HSGP_GP <- rbind(f_GP, f_HSGP)

ggplot(f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=type, color=type))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

#### Model evaluation

Residuals `residual`, Bayesian $R^2$ (Coefficient of determination), and Log predictive density `lpd`

```{r fig.height=2, fig.width=10, message=FALSE, warn=FALSE, echo=FALSE}
#Residuals
f <- summary(fit_diagnosis[[i]], pars = c("f"))$summary[,1]
residual[[i]] <- standata[[i]]$y[vv_train] - f

p1 <- ggplot(as.data.frame(residual[[i]]), aes(residual[[i]]))  +
  geom_histogram(color = 'white') +
  labs(x="residual") +
  theme_classic()

#Bayesian $R^2$
noise <- as.matrix(fit_diagnosis[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit_diagnosis[[i]], pars = c("f")), 1, sd)
R2 <- sd_f^2/(sd_f^2 + noise^2)

p2 <- ggplot(as.data.frame(R2), aes(R2))  +
  geom_histogram(color = 'white') +
  theme_classic()

#Log predictive density (lpd)
lpd <- summary(fit_diagnosis[[i]], pars = c("lpd_y"))$summary[,1]

p3 <- ggplot(as.data.frame(lpd), aes(lpd))  +
  geom_histogram(color = 'white') +
  theme_classic()

grid.arrange(p1, p2, p3, ncol=3)
```

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination), and expected log predictive density `elpd`

```{r fig.height=3, fig.width=5, message=FALSE, warn=FALSE, echo=FALSE}
rmse[i] <- round(sqrt(mean(residual[[i]]^2)), 3)
eR2[i] <- round(median(R2), 3)
elpd[i] <- round(median(lpd), 3)

evaluation[[i]] <- data.frame(iter= i,
                                rsme = rmse[i],
                                R2 = eR2[i],
                                elpd = elpd[i]
                                )
rbind(evaluation[[i-2]], evaluation[[i-1]],evaluation[[i]])
```

#### Diagnosis

1. Estimated lengthscale $\hat{l_3}$

```{r message=FALSE, warn=FALSE, echo=FALSE}
l_hat[i] <- round(summary(fit_diagnosis[[i]], pars = "lscale")$summary[,1], 2)
l_hat[i]
```

2. Check the diagnostic of whether $\, \hat{l_3} \geq l_3$

```{r message=FALSE, warn=FALSE, echo=FALSE}
l_hat[i] >= l[i]
```

**Summary table**

```{r message=FALSE, warn=FALSE, echo=FALSE}
diagnosis[[i]] <- data.frame(iter= i,
                                GP_func= c("f"),
                                l= l[i],
                                c= c[i],
                                m= m[i],
                                l_hat= l_hat[i],
                                'l_hat > l' = l_hat[i] >= l[i],
                                rsme = rmse[i],
                                R2 = eR2[i],
                                elpd = elpd[i]
                                )
names(diagnosis[[i]]) <- c("iter", "GP_func", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "R2", "elpd")
diagnosis[[i]]

diagnosis_summ <- rbind(diagnosis[[1]],diagnosis[[2]],diagnosis[[3]])
diagnosis_summ
```

### 2022-02-09 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
update manuscript

@@
 \usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,breaklinks=true]{hyperref}
+\usepackage{nicefrac}
 
@@
 
+{\color{blue} Aki: we could cite in the intro https://arxiv.org/abs/2111.01084 as a paper which
+reviews alternative methods}
+
 In this paper, we focus on a recent framework for fast and accurate inference for fully Bayesian GPs using basis function approximations based on approximation via Laplace eigenfunctions for stationary covariance functions proposed by \citet{solin2018hilbert}.
@@
 where $\nu > 0$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell > 0$ and $\alpha > 0$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
-The particular case where $\nu=\infty$ and $\nu=3/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
+The particular case where $\nu=\infty$, $\nu=1/2$, $\nu=3/2$ and $\nu=5/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
 %
@@
 k_{\infty}(r)&=\alpha \exp \left( -\frac{1}{2} \frac{r^2}{\ell^2}\right),  \\
-k_{\frac{3}{2}}(r)&=\alpha\left(1+\frac{\sqrt{3}r}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}r}{\ell}\right). 
+k_{\frac{1}{2}}(r)&=\alpha \exp \left( -\frac{r}{\ell}\right),  \\
+k_{\frac{3}{2}}(r)&=\alpha\left(1+\frac{\sqrt{3}r}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}r}{\ell}\right),   \\
+k_{\frac{5}{2}}(r)&=\alpha\left(1+\frac{\sqrt{5}r}{\ell} + \frac{5r^2}{3\ell^2}\right) \exp\! \left( -\frac{\sqrt{5}r}{\ell}\right). 
 \end{align*}
@@
 k_{\infty}(r) = \alpha & \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
+k_{\frac{1}{2}}(r) = \alpha & \exp \left( -\sum_{i=1}^{D}\frac{(x_i-x_i')}{\ell_i}\right),\\
 k_{\frac{3}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right)\\
-& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right).
-\end{align*}
-%
-Notice that the previous expressions, $k_{\infty}(r)$ and $k_{\frac{3}{2}}(r)$, have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
+& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right),\\
+k_{\frac{5}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}+\frac{5}{3} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\;\right)\\
+& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{5(x_i-x_i')^2}{\ell_i^2}}\;\right).
+\end{align*}
+%
+Notice that the previous expressions, $k_{\infty}(r)$, $k_{\frac{1}{2}}(r)$, $k_{\frac{3}{2}}(r)$ and $k_{\frac{5}{2}}(r)$, have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
 Using individual length-scales for each dimension turns the isotropic covariance function into a non-isotropic covariance function.
@@
 \begin{align*}
-S_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^\intercal \bm{\omega} \right)^{(\nu+D/2)}
+S_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^\intercal \bm{\omega} \right)^{-(\nu+D/2)}
 \end{align*}
@@
 in $D$ dimensions, where vector $\bm{\omega}\in {\rm I\!R^D}$ is in the frequency domain, and $\ell$ and $\alpha$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
-The particular cases, where $\nu=\infty$ and $\nu=3/2$, take the form
+The particular cases, where $\nu=\infty$, $\nu=1/2$, $\nu=3/2$ and $\nu=5/2$, take the form
 %
@@
 S_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\frac{1}{2}\ell^2 \bm{\omega}^\intercal \bm{\omega}), \label{eq_specdens_inf}  \\
-S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+3}{2}}.
-\label{eq_specdens_32}
+S_{\frac{1}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+1}{2}} \label{eq_specdens_12},     \\
+S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+3}{2}} \label{eq_specdens_32},     \\
+S_{\frac{5}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+5}{2}){5}^{5/2}}{\frac{3}{4}\sqrt{\pi}\ell^5}\left(\frac{5}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+5}{2}}.
+\label{eq_specdens_52}
 \end{align}
@@
 S_{\infty}(\bm{\omega})&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 \omega_i^2 \right),   \\
-S_{\frac{3}{2}}(\bm{\omega})&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-3},
+S_{\frac{1}{2}}(\bm{\omega})&= \alpha \, 8\,\pi \prod_{i=1}^{3}\ell_i\left(1+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-2},    \\
+S_{\frac{3}{2}}(\bm{\omega})&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-3},   \\
+S_{\frac{5}{2}}(\bm{\omega})&= \alpha \, \frac{64}{3}\,\pi {5}^{5/2}\prod_{i=1}^{3}\ell_i\left(5+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-4}.
 \end{align*}
@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the function to be estimated, which is in turn is characterized by the length-scale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the function to be estimated, which is in turn characterized by the length-scale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 
-\begin{figure*}[!t]
+\begin{figure*} %[!t]
 \centering
@@
 
-\begin{figure*}[!h]
+\begin{figure*} %[!h]
 \centering
@@
 
-Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu$=3/2) covariance function, respectively, and a single input dimension.
+\subsection{Theoretical evidence of linear proportionality between $m$, $c$ and $l$} \label{subsec_theoretic_linearity}
+
+There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
+In this Section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
+In next Section \ref{subsec_empiric_relations} we conduct an empirical experiments to prove this expected linear proportionality relationship between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
+
+{\color{blue}Aki: We can also look at the mean zero-level upcrossings, which for QE is
+proportional to 1/l which gives as that part, and the effect of c should also
+be quite linear when c is not very close to 1 and m is not very small as it
+linearly makes the interval shorter so that the there are less zero-level
+upcrossings available in the basis functions. You can visualize this using
+{\small \url{https://avehtari.github.io/casestudies/Motorcycle/motorcycle_gpcourse.html}} and
+look how the plot of 6 basis functions change when you set c=1, c=1.5, c=2.25
+(ie $1.5^2$). Instead of looking at mean zero-level crossings, we can look at the
+spectral densities and when they go to zero. This is illustrated in the same
+notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
+explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
+(linear in l) and to explain 80\% of the prior variance we need 2, 4, 8, 16
+basis functions. These spectral densities don't depend on c. When c increaes
+the basis functions are truncated and for example going from c=1.5 to c=2.25,
+basis function 9 has the flexibility basis function 6 had, so 1.5 more basis
+functions are needed. So we have proportionality c/l. }
+
+{\color{red} Gabi: The spectral densities do depend on $c$, see equations (\ref{eq_approxf}) and (\ref{eq_eigenvalue}).}
+
+\subsection{Empirical discovering of the functional form of the relationships between $m$, $c$ and $l$} \label{subsec_empiric_relations}
+
+Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu$=3/2) covariance function, respectively, and a single input dimension.
 More precisely, for a given GP model (with a squared exponential covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
-Similarly,  Figure \ref{fig7_lscale_vs_J_vs_c_Matern} shows the corresponding plot for the  Mat\'ern ($\nu$=3/2) covariance function.
+Similarly,  Figure \ref{fig6_7_relationships} shows the corresponding plot for the  Mat\'ern ($\nu$=3/2) covariance function.
 We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
@@
 \item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
-A $\ell$ increases, the minimum valid $c$ also increases.
+This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+As $\ell$ increases, the minimum valid $c$ also increases.
 \end{itemize}
 
-As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+\subsubsection{Numerical equations}
+
+From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
+Following we show numerical functions governing these relationships for every kernel.
+
+\paragraph*{Squared exponential}:
+%
+\begin{align}\label{eq_m_l_QE}
+&m= 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 1.75 \, \frac{c}{m},
+\end{align}
+%
+with
+%
+\begin{align}\label{eq_c_vs_l_QE}
+&c = c(l) \, \geq \, 3.2 \, \nicefrac{l}{S}         
+\end{align}
+
+\paragraph*{Mat\'ern ($\nu$=3/2)}:
+%
+\begin{align}\label{eq_m_l_mat32}
+&m= 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 3.42 \, \frac{c}{m},
+\end{align}
+%
+with
+%
+\begin{align}\label{eq_c_vs_l_mat32}
+&c \, \geq \, 4.5 \, \nicefrac{l}{S}         
+\end{align}
+
+\paragraph*{Mat\'ern ($\nu$=5/2)}:
+%
+\begin{align}\label{eq_m_l_mat52}
+&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
+\end{align}
+%
+with
+%
+\begin{align}\label{eq_c_vs_l_mat52}
+&c \, \geq \, 4.1 \, \nicefrac{l}{S}         
+\end{align}
+
+Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
+An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
+More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions.
+For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
+Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
+However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+
+\subsection{Diagnosis of the approximation} \label{subsec_diagnosis}
+
+As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig7_lscale_vs_J_vs_c_Matern} (depending on which kernel is used).
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_7_relationships} (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
@@
 
+%\begin{figure*}
+%\centering
+%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
+%\hspace{3mm}
+%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
+%\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+%  \label{fig7_lscale_vs_J_vs_c_Matern}
+%\end{figure*}
+
+%\begin{figure*}
+%\centering
+%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_7_relationships.pdf}}
+%\hspace{3mm}
+%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig7_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
+%\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\'ern ($\nu$=3/2) covariance function.
+The right-side plot is a zoom in of the left-side plot.}
+%  \label{fig6_7_relationships}
+%\end{figure*}
+
 \begin{figure*}
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
-\hspace{3mm}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
-The right-side plot is a zoom in of the left-side plot.}
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_7_relationships.png}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
   \label{fig6_lscale_vs_J_vs_c}
@@
 
-\begin{figure*}
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig7_lscale_vs_J_vs_c_Matern.pdf}}
-\hspace{3mm}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig7_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\'ern ($\nu$=3/2) covariance function.
-The right-side plot is a zoom in of the left-side plot.}
-  \label{fig7_lscale_vs_J_vs_c_Matern}
-\end{figure*}
-
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_lscale_vs_J_vs_c}.
-
-Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
-An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
-More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions.
-For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
-Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern} to check the accuracy by analyze individually the different dimensions of a multidimensional GP model.
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_7_relationships}.
+
+\subsubsection{User guide with the steps to perform diagnosis}
+
+\paragraph*{Assumption of the diagnosis tool}: Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is smaller than the exact GP lengthscale estimate $l$.
+Thus the diagnostic that determines is the approximation is sufficiently accurate is:
+%
+\begin{align}\label{eq_diagnostic}
+&\hat{l} \pm 0.02 \geq l
+\end{align}
+
+\paragraph*{User-guide with the steps to perform diagnosis}:
+
+\begin{enumerate}
+\item[Step 1.] Make a first guess on the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, around 1.
+\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
+\item[Step 2.] Obtain the valid minimum boundary factor $c_1$ determined by $l_1$ by using (\ref{eq_c_vs_l_QE}).
+\item[Step 3.] Obtain the valid mimimum number of basis functions $m_1$ determined by $l_1$ and $c_1$ by using (\ref{eq_m_l_QE}).
+Notice that $l_1$ can also be read as the minimum lengthscale that can be accurately fitted determined by $m_1$ and $c_1$ by using (\ref{eq_m_l_QE}).
+\item[Step 4.] Fit the HSGP model and assess residuals: compute \textit{rmse*} (Root mean square error), $R^2$ (Coefficient of variation) and \textit{elpd} (Expected log predicitve density).
+\item[Step 5.] Check the diagnostic of whether $\, \hat{l_1} \pm 0.02 \geq l_1$ (\ref{eq_diagnostic}).
+	\begin{enumerate}
+    \item If the diagnostic is TRUE, the HSGP model approximation must be close to be sufficiently accurate. Then user can continue with step 6.
+       	\begin{enumerate}
+    	\item[] \hspace{-1cm} \textit{(Next iterations, $i>1$)}
+		\item[Step 6.] Set $\, m_i = m_{i-1} + 10$.
+		\item[Step 7.] Obtain the valid minimum boundary factor $c_i$ by using (\ref{eq_c_vs_l_QE}), with $c(\hat{l}_{i-1})$.
+		\item[Step 8.] Obtain the mimimum lengthscale $l_i$ that can be accurately fitted determined by $m_i$ and $c_i$ by using (\ref{eq_m_l_QE}).
+		\item[Step 9.] Fit the HSGP model and
+			\begin{enumerate}
+        	\item[Step 9.1.] Check whether $\, \hat{l}_i \pm 0.02 \geq l_i$ (\ref{eq_diagnostic}),
+        	\item[Step 9.2.] Check stability in $\hat{l}_i$, \textit{rmse}, $R^2$ and \textit{elpd} relative to previous iteration.
+			\end{enumerate}
+		\item[Step 10.] If the verifications in steps 9.1 and 9.2 are TRUE, the HSGP model approximation should be sufficiently accurate, and diagnosis ends here.
+Otherwise, repeat steps 6-10 and update parameters.
+    	\end{enumerate}
+     \item If the diagnostic is FALSE, the HSGP model approximation can not be sufficiently accurate. Then user has to move to step 11.
+     	\begin{enumerate}
+		\item[Step 11.] Set $\, l_i = \hat{l}_{i-1}$.
+		\item[Step 12.] Repeat steps 2-5 and update parameters.
+		\end{enumerate}
+    \end{enumerate}
+
+\end{enumerate}
 
@@
 In this example, we make a comparison of the length-scale estimates
-obtained from the exact GP and HSGP models. We also have a look at those recommended minimum length-scales provided by Figure \ref{fig6_lscale_vs_J_vs_c}.
+obtained from the exact GP and HSGP models. We also have a look at those recommended minimum length-scales provided by Figure \ref{fig6_7_relationships}.
 For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying length-scale values.
@@
 In large length-scales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum length-scale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_lscale_vs_J_vs_c}.
+The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum length-scale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_7_relationships}.
 We observe that whenever the estimated length-scale exceeds the minimally estimable length-scale, the RMSE of the posterior predictions is small
@@
 The different plots represent the use of different number of basis functions $m$ in the HSGP model.
-The dashed black line represents the recommended minimum length-scales provided by Figure \ref{fig6_lscale_vs_J_vs_c} that can be closely approximated by the HSGP model in every case.}
+The dashed black line represents the recommended minimum length-scales provided by Figure \ref{fig6_7_relationships} that can be closely approximated by the HSGP model in every case.}
   \label{fig8_Tlscale_vs_Elscale}
@@
 
-%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}.
+%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships}.
 Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
@@
 In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnosis}.
-In the second iteration, $m$ is increased to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig7_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$.
+In the second iteration, $m$ is increased to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_7_relationships} is $\ell^*=0.19$.
 As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{fig12_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
@@
 We use $m=30$ basis functions and a boundary factor $c=1.5$.
-The length-scale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum length-scale reported by Figure \ref{fig6_lscale_vs_J_vs_c} as a function of $m$ and $c$.
+The length-scale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum length-scale reported by Figure \ref{fig6_7_relationships} as a function of $m$ and $c$.
 This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
@@
 Figure~\ref{fig19_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig19_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_lscale_vs_J_vs_c}.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig19_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_7_relationships}.
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig19_elpd_leukemia}-right).
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}, is essential to avoid a excessive computational time especially in multivariate input spaces.
+Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships}, is essential to avoid a excessive computational time especially in multivariate input spaces.
 However, in practice, input dimensionalities larger than three start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.

### 2022-02-13 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
little update of main text

@@
 where $\nu > 0$ is the order the kernel, $K_{\nu}$ the modified Bessel function of the second kind, and the $\ell > 0$ and $\alpha > 0$ are the length-scale and magnitude (marginal variance), respectively, of the kernel.
-The particular case where $\nu=\infty$, $\nu=1/2$, $\nu=3/2$ and $\nu=5/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
+The particular case where $\nu=\infty$, $\nu=3/2$ and $\nu=5/2$ are probably the most commonly used kernels \citep{rasmussen2006gaussian},
 %
@@
 k_{\infty}(r)&=\alpha \exp \left( -\frac{1}{2} \frac{r^2}{\ell^2}\right),  \\
-k_{\frac{1}{2}}(r)&=\alpha \exp \left( -\frac{r}{\ell}\right),  \\
+%k_{\frac{1}{2}}(r)&=\alpha \exp \left( -\frac{r}{\ell}\right),  \\
 k_{\frac{3}{2}}(r)&=\alpha\left(1+\frac{\sqrt{3}r}{\ell}\right) \exp\! \left( -\frac{\sqrt{3}r}{\ell}\right),   \\
@@
 The former is commonly known as the squared exponential or exponentiated quadratic covariance function.
-Assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernels written above take the form
-%
-\begin{align*}
-k_{\infty}(r) = \alpha & \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right),\\
-k_{\frac{1}{2}}(r) = \alpha & \exp \left( -\sum_{i=1}^{D}\frac{(x_i-x_i')}{\ell_i}\right),\\
-k_{\frac{3}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right)\\
-& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right),\\
-k_{\frac{5}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}+\frac{5}{3} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\;\right)\\
-& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{5(x_i-x_i')^2}{\ell_i^2}}\;\right).
-\end{align*}
-%
-Notice that the previous expressions, $k_{\infty}(r)$, $k_{\frac{1}{2}}(r)$, $k_{\frac{3}{2}}(r)$ and $k_{\frac{5}{2}}(r)$, have been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
-Using individual length-scales for each dimension turns the isotropic covariance function into a non-isotropic covariance function.
+As an example, assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernel $K_{\nu}$ written above takes the form
+%
+\begin{align*}
+k_{\infty}(r) = \alpha & \exp \left( -\frac{1}{2} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\right).
+%k_{\frac{1}{2}}(r) = \alpha & \exp \left( -\sum_{i=1}^{D}\frac{(x_i-x_i')}{\ell_i}\right),\\
+%k_{\frac{3}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right)\\
+%& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}\;\right),\\
+%k_{\frac{5}{2}}(r) = \alpha & \left(1+\sqrt{\sum_{i=1}^{D}\frac{3(x_i-x_i')^2}{\ell_i^2}}+\frac{5}{3} \sum_{i=1}^{D}\frac{(x_i-x_i')^2}{\ell_i^2}\;\right)\\
+%& \times \exp\! \left( - \sqrt{\sum_{i=1}^{D}\frac{5(x_i-x_i')^2}{\ell_i^2}}\;\right).
+\end{align*}
+%
+Notice that the previous expressions $k_{\infty}(r)$ has been easily generalized to using a multidimensional length-scale $\bm{\ell}\in {\rm I\!R}^D$.
+Using individual length-scales for each dimension turns an isotropic covariance function into a non-isotropic covariance function.
 That is, for a non-isotropic covariance function, the smoothness may vary across different input dimensions.
@@
 S_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\frac{1}{2}\ell^2 \bm{\omega}^\intercal \bm{\omega}), \label{eq_specdens_inf}  \\
-S_{\frac{1}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+1}{2}} \label{eq_specdens_12},     \\
+%S_{\frac{1}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+1}{2}} \label{eq_specdens_12},     \\
 S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+3}{2}} \label{eq_specdens_32},     \\
@@
 S_{\infty}(\bm{\omega})&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 \omega_i^2 \right),   \\
-S_{\frac{1}{2}}(\bm{\omega})&= \alpha \, 8\,\pi \prod_{i=1}^{3}\ell_i\left(1+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-2},    \\
+%S_{\frac{1}{2}}(\bm{\omega})&= \alpha \, 8\,\pi \prod_{i=1}^{3}\ell_i\left(1+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-2},    \\
 S_{\frac{3}{2}}(\bm{\omega})&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-3},   \\
@@
 
+\begin{figure*} %[!t]
+\centering
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Post_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig1_Sigma_J.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Cov_J.pdf}}
+\vspace{-1mm}
+\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
+  \label{fig1_Post_J}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Post_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig2_Sigma_L.pdf}}
+\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Cov_L.pdf}}
+\vspace{-1mm}
+\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
+  \label{fig2_Post_L}
+\end{figure*}
+
 As explained in Section~\ref{sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
@@
 The approximation error for the variance tends to be bigger for the variance than for the mean.
-
-\begin{figure*} %[!t]
-\centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig1_Sigma_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Cov_J.pdf}}
-\vspace{-1mm}
-\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
-  \label{fig1_Post_J}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig2_Sigma_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Cov_L.pdf}}
-\vspace{-1mm}
-\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
-  \label{fig2_Post_L}
-\end{figure*}
-
-Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} for both the exact GP and the HSGP models.
-Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
-Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
-Figure \ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
-Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
-Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application.
-We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig3_Post_part1}.
-From these results, some general conclusions may be drawn:
-
-\begin{itemize}
-\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
-\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
-\end{itemize}
 
@@
 
+Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
+The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} for both the exact GP and the HSGP models.
+Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
+Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
+Figure \ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
+Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application.
+We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig3_Post_part1}.
+From these results, some general conclusions may be drawn:
+
+\begin{itemize}
+\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
+\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
+\end{itemize}
+
 \subsection{Theoretical evidence of linear proportionality between $m$, $c$ and $l$} \label{subsec_theoretic_linearity}
@@
 There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
-In this Section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
-In next Section \ref{subsec_empiric_relations} we conduct an empirical experiments to prove this expected linear proportionality relationship between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
+In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
+In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
 
@@
 
-Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} depict how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function and Mat\'ern ($\nu$=3/2) covariance function, respectively, and a single input dimension.
-More precisely, for a given GP model (with a squared exponential covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+\begin{figure*}
+\centering
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_7_relationships.png}}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
+  \label{fig6_7_relationships}
+\end{figure*}
+
+Figure \ref{fig6_7_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_7_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
-Similarly,  Figure \ref{fig6_7_relationships} shows the corresponding plot for the  Mat\'ern ($\nu$=3/2) covariance function.
 We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
@@
 \item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+This fact can be seen in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 As $\ell$ increases, the minimum valid $c$ also increases.
@@
 From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
-Following we show numerical functions governing these relationships for every kernel.
+Following we show the numerical functions governing these relationships for every kernel.
 
@@
 \begin{align}\label{eq_c_vs_l_QE}
-&c = c(l) \, \geq \, 3.2 \, \nicefrac{l}{S}         
+&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2      
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat32}
-&c \, \geq \, 4.5 \, \nicefrac{l}{S}         
+&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2        
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat52}
-&c \, \geq \, 4.1 \, \nicefrac{l}{S}         
-\end{align}
-
-Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} were build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
+&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
+\end{align}
+
+Notice that we establish a lower bound for $c$ set to 1.2.
+
+Figure~\ref{fig6_7_relationships} was build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
 An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
@@
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships} to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
 
@@
 
-As stated above, Figures~\ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_7_relationships} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, Figure~\ref{fig6_7_relationships} provides the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_7_relationships} (depending on which kernel is used).
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_7_relationships} (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
@@
 
-%\begin{figure*}
-%\centering
-%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c.pdf}}
-%\hspace{3mm}
-%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin.pdf}}
-%\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a square exponential covariance function.
-The right-side plot is a zoom in of the left-side plot.}
-%  \label{fig7_lscale_vs_J_vs_c_Matern}
-%\end{figure*}
-
-%\begin{figure*}
-%\centering
-%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig6_7_relationships.pdf}}
-%\hspace{3mm}
-%\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{fig7_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
-%\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\'ern ($\nu$=3/2) covariance function.
-The right-side plot is a zoom in of the left-side plot.}
-%  \label{fig6_7_relationships}
-%\end{figure*}
-
-\begin{figure*}
-\centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_7_relationships.png}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
-  \label{fig6_lscale_vs_J_vs_c}
-\end{figure*}
-
 If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_7_relationships}.
@@
 
-\paragraph*{Assumption of the diagnosis tool}: Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is smaller than the exact GP lengthscale estimate $l$.
-Thus the diagnostic that determines is the approximation is sufficiently accurate is:
+\paragraph*{Assumption of the diagnosis tool}: Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is smaller than the minimum length-scale $l^*$ determined by $m$ and $c$ in Figure~\ref{fig6_7_relationships}.
+Thus the diagnostic pointing towards an accurate approximation is:
 %
 \begin{align}\label{eq_diagnostic}
-&\hat{l} \pm 0.02 \geq l
+&\hat{l} \pm 0.02 \geq l^*
 \end{align}
@@
 Notice that $l_1$ can also be read as the minimum lengthscale that can be accurately fitted determined by $m_1$ and $c_1$ by using (\ref{eq_m_l_QE}).
-\item[Step 4.] Fit the HSGP model and assess residuals: compute \textit{rmse*} (Root mean square error), $R^2$ (Coefficient of variation) and \textit{elpd} (Expected log predicitve density).
+\item[Step 4.] Fit the HSGP model and assess residuals: compute \textit{rmse} (Root mean square error), $R^2$ (Coefficient of variation) and \textit{elpd} (Expected log predicitve density).
 \item[Step 5.] Check the diagnostic of whether $\, \hat{l_1} \pm 0.02 \geq l_1$ (\ref{eq_diagnostic}).
@@
     \end{enumerate}
-
 \end{enumerate}
 
-\subsection{Comparing length-scale estimates}
-
-In this example, we make a comparison of the length-scale estimates
-obtained from the exact GP and HSGP models. We also have a look at those recommended minimum length-scales provided by Figure \ref{fig6_7_relationships}.
-For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function and varying length-scale values.
-Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case.
-
-Figure \ref{fig8_posterior_varing_lscale_part1} shows the posterior predictions of both exact GP and HSGP models fitted to those datasets.
-The length-scale estimates as obtained by exact GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}.
-As noted previously, an accurate estimate of the length-scale can be a good indicator of a close approximation of the HSGP model to the exact GP model.
-Further, Figure \ref{fig10_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models, computed against the exact GP models, as a function of the length-scale and number of basis functions.
-
-Comparing the accuracy of the length-scale in Figure~\ref{fig8_Tlscale_vs_Elscale} to the RMSE in Figure~\ref{fig10_MSE_varing_lscale}, we see that they agree closely with each other for medium length-scales.
-That is, a good estimation of the length-scale implies a small RMSE.
-This is no longer true for very small or large length-scales.
-In small length-scales, even very small inaccuracies  may have a strong influence on the posteriors predictions and thus on the RMSE.
-In large length-scales, larger inaccuracies change the posterior predictions only little and may thus not yield large RMSEs.
-The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the  minimum length-scale that can be closely approximated under the given condition, according to the results presented in Figure \ref{fig6_7_relationships}.
-We observe that whenever the estimated length-scale exceeds the minimally estimable length-scale, the RMSE of the posterior predictions is small
-(see Figure \ref{fig10_MSE_varing_lscale}). Conversely, when the estimated length-scale is smaller than the minimally estimable one, the RMSE becomes very large.
+\subsection{Performance analysis of the diagnosis}
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.42, trim = 10mm 100mm 5mm 30mm, clip]{fig8_legend.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part1.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part2.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part3.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part4.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part5.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part6.pdf}}\\
-\vspace{-3mm}
-\subfigure{\includegraphics[scale=0.31, trim = 1mm 5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part7.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part8.pdf}}
-\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{fig8_posterior_varing_lscale_part9.pdf}}\\
-\caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic length-scales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}).
-}
-  \label{fig8_posterior_varing_lscale_part1}
-\end{figure*}
-
-\begin{figure*}
-%\begin{flushleft}
-\centering
-\begin{tabular}{ccc}
-\multicolumn{3}{c}{ \includegraphics[scale=0.42, trim = 25mm 117mm 90mm 45mm, clip]{fig9_legend.pdf}}\\
-\includegraphics[scale=0.31, trim = 0mm 24mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
-\includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 24mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part3.pdf} \\
-\includegraphics[scale=0.31, trim = 0mm 2mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.31, trim = 25mm 2mm 7mm 10mm, clip]{fig9_Tlscale_vs_Elscale_part6.pdf}
-\end{tabular}
-%\end{flushleft}
-\caption{Data-generating functional length-scales ($\ell$), of the various datasets illustrated in Figure \ref{fig8_posterior_varing_lscale_part1}, versus the corresponding length-scale estimates $(\hat{\ell})$ from the exact GP and HSGP models.
-95\% confident intervals of the length-scale estimates are plotted as dot lines.
-The different plots represent the use of different number of basis functions $m$ in the HSGP model.
-The dashed black line represents the recommended minimum length-scales provided by Figure \ref{fig6_7_relationships} that can be closely approximated by the HSGP model in every case.}
-  \label{fig8_Tlscale_vs_Elscale}
-\end{figure*}
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_sec45_comparing_lscale.png}}
+\caption{Estimated exact GP and approximate HSGP lengthscales for various datasets with varying smoothness.}
+  \label{fig_sec45_comparing_lscale}
+\end{figure*}
+
+Figure \ref{fig_sec45_comparing_lscale}-left shows a comparison of the length-scale estimates
+obtained from the exact GP and HSGP models with a squared exponential kernel from various noisy datasets drawn from underlying functions with varying smoothness.
+Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using equation (\ref{eq_c_vs_l_QE}).
+Figure \ref{fig_sec45_comparing_lscale}-right shows the RMSE of the HSGP models with the exact GP model as the reference.
+
+Using large number of basis functions in smooth datasets (with length-scales greater than 0.5) causes overestimating the underlying functions slightly and so the length-scale too (Figure \ref{fig_sec45_comparing_lscale}-left).
+Here it comes out a possible weaknesses of the approach that when we want to fit very smooth functions we should try not to use too many basis functions than necessary because we will run into overestimating problems.
+Apart from that, an accurate estimate of the length-scale (Figure \ref{fig_sec45_comparing_lscale}-left) implies a small RMSE (Figure \ref{fig_sec45_comparing_lscale}-right).
 
@@
 \centering
-\includegraphics[scale=0.38, trim = 0mm 9mm 0mm 0mm, clip]{fig10_MSE_varing_lscale.pdf}
-\caption{RMSE of the HSGP models with different number of basis functions $m$, for the various datasets with different wiggly effects ($\ell$).}
-  \label{fig10_MSE_varing_lscale}
+\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_sec45_2_oversmoothing_effect.png}
+\caption{Oversmooth posterior functions when using large number of basis functions for smooth datasets.}
+  \label{fig_sec45_2_oversmoothing_effect}
 \end{figure}
+
+Next we apply the diagnosis procedure over all the datasets of Figure \ref{fig_sec45_comparing_lscale}-left and we show that following our recomendations and the diagnosis we get the optimum solution with minimum requirements.
 

### 2022-02-14 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
update step-by-step user guide

@@
 \subsubsection{Numerical equations}
+\label{num_equations}
 
@@
 
-\subsection{Diagnosis of the approximation} \label{subsec_diagnosis}
+\subsection{Diagnostics of the approximation} \label{subsec_diagnostics}
 
@@
 
-\subsubsection{User guide with the steps to perform diagnosis}
-
-\paragraph*{Assumption of the diagnosis tool}: Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is smaller than the minimum length-scale $l^*$ determined by $m$ and $c$ in Figure~\ref{fig6_7_relationships}.
-Thus the diagnostic pointing towards an accurate approximation is:
-%
-\begin{align}\label{eq_diagnostic}
-&\hat{l} \pm 0.02 \geq l^*
-\end{align}
-
-\paragraph*{User-guide with the steps to perform diagnosis}:
+\subsubsection{A step-by-step user-guide to apply the diagnostics}
+
+Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
+The procedure is split into two phases, Phase A and B, which have to be completed consequtively.
+%In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.
+
+\paragraph*{Phase A:}
 
 \begin{enumerate}
-\item[Step 1.] Make a first guess on the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, around 1.
-\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
-\item[Step 2.] Obtain the valid minimum boundary factor $c_1$ determined by $l_1$ by using (\ref{eq_c_vs_l_QE}).
-\item[Step 3.] Obtain the valid mimimum number of basis functions $m_1$ determined by $l_1$ and $c_1$ by using (\ref{eq_m_l_QE}).
-Notice that $l_1$ can also be read as the minimum lengthscale that can be accurately fitted determined by $m_1$ and $c_1$ by using (\ref{eq_m_l_QE}).
-\item[Step 4.] Fit the HSGP model and assess residuals: compute \textit{rmse} (Root mean square error), $R^2$ (Coefficient of variation) and \textit{elpd} (Expected log predicitve density).
-\item[Step 5.] Check the diagnostic of whether $\, \hat{l_1} \pm 0.02 \geq l_1$ (\ref{eq_diagnostic}).
+\item[Step A1.] Make an initial guess on true the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, about $l_1 \approx 1$. Set the iteration index to $i=1$.
+%\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
+\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $l = l_i$ for the given kernel as per Section \ref{num_equations} 
+\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $l = l_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}. 
+\item[Step A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
+\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.
+
+If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $l_{i+1} = \hat{l}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.
+
+If the diagnostic is TRUE, the HSGP approximation is close to be sufficiently accurate. Continue with Phase B.
+
+\end{enumerate}
+
+\paragraph*{Phase B:}
+
 \begin{enumerate}
-    \item If the diagnostic is TRUE, the HSGP model approximation must be close to be sufficiently accurate. Then user can continue with step 6.
-       	\begin{enumerate}
-    	\item[] \hspace{-1cm} \textit{(Next iterations, $i>1$)}
-		\item[Step 6.] Set $\, m_i = m_{i-1} + 10$.
-		\item[Step 7.] Obtain the valid minimum boundary factor $c_i$ by using (\ref{eq_c_vs_l_QE}), with $c(\hat{l}_{i-1})$.
-		\item[Step 8.] Obtain the mimimum lengthscale $l_i$ that can be accurately fitted determined by $m_i$ and $c_i$ by using (\ref{eq_m_l_QE}).
-		\item[Step 9.] Fit the HSGP model and
-			\begin{enumerate}
-        	\item[Step 9.1.] Check whether $\, \hat{l}_i \pm 0.02 \geq l_i$ (\ref{eq_diagnostic}),
-        	\item[Step 9.2.] Check stability in $\hat{l}_i$, \textit{rmse}, $R^2$ and \textit{elpd} relative to previous iteration.
+
+\item[Step B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD).
+
+\item[Step B2.] Set $m_{i+1} = m_{i} + 10$ and increase the iteration index $i = i + 1$.
+
+\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $l = \hat{l}_{i-1}$ for the given kernel as per Section \ref{num_equations}.
+
+\item[Step B4.] Obtain the mimimum valid lengthscale $l_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}.
+
+\item[Step B5.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
+
+\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.
+
+Check the stability of both $\hat{l}_i$ and the measures of predictive accuracy relative to the previous iteration.
+
+If all the stability checks succeeed, the HSGP approximation of the latest model should be sufficiently accurate and the procedure ends here.
+
+Otherwise, go back to Step B1.
 \end{enumerate}
-		\item[Step 10.] If the verifications in steps 9.1 and 9.2 are TRUE, the HSGP model approximation should be sufficiently accurate, and diagnosis ends here.
-Otherwise, repeat steps 6-10 and update parameters.
-    	\end{enumerate}
-     \item If the diagnostic is FALSE, the HSGP model approximation can not be sufficiently accurate. Then user has to move to step 11.
-     	\begin{enumerate}
-		\item[Step 11.] Set $\, l_i = \hat{l}_{i-1}$.
-		\item[Step 12.] Repeat steps 2-5 and update parameters.
-		\end{enumerate}
-    \end{enumerate}
-\end{enumerate}
-
-\subsection{Performance analysis of the diagnosis}
+
+\subsection{Performance analysis of the diagnostics}
 

### 2022-02-15 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Add paragraph on other covariance functions

@@
 %
-\begin{equation*}
+\begin{equation} \label{eq:rel_total_distance}
  \int | k(\tau) - \tilde{k}_m(\tau)|  \,\mathrm{d}\tau  < 0.01 \int k(\tau) \,\mathrm{d}\tau,
-\end{equation*}
+\end{equation}
 where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions.
@@
 Next we apply the diagnosis procedure over all the datasets of Figure \ref{fig_sec45_comparing_lscale}-left and we show that following our recomendations and the diagnosis we get the optimum solution with minimum requirements.
+
+\subsection{Other covariance functions}
+
+In Section \ref{sec_accuracy}, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
+We specifically focused on the Matérn and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
+The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
+The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_7_relationships} is not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
+\eqref{eq:rel_total_distance}.
+Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
 

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edited 4.2

@@
 
-\subsection{Theoretical evidence of linear proportionality between $m$, $c$ and $l$} \label{subsec_theoretic_linearity}
-
-There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
+\subsection{Near linear proportionality between $m$, $c$, and $l$} \label{subsec_theoretic_linearity}
+
+A priori the terms in the series expansion \eqref{eq_approxf_multi} with very small spectral density are unlikely to contribute to the approximation.
+Given the boundary factor $c$ and the length-scale $\ell$, we can compute cumulative sum of the spectral densities and find out how many basis functions are a priori explaining almost 100\% of the variation.
+Thus, given $c$ and $\ell$, we can estimate a good choice for the number of basis functions $m$ for any covariance function.
+
+When considering exponentiated quadratic and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
+With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$ and with increasing $c$, the $m$ should grow near linearly with $c$.
+%
+This is natural as with decreasing $\ell$, more higher frequency basis functions are needed.
+With increasing $c$, the smaller range of the basis functions are used in the approximation, the expected number of zero up-crossings goes down linearly with $c$, and thus more higher frequency basis functions are needed to compensate this.
+%
+When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$ if $\ell$ or $c$ is changed.
+
+% There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
 In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
-In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
-
-{\color{blue}Aki: We can also look at the mean zero-level upcrossings, which for QE is
-proportional to 1/l which gives as that part, and the effect of c should also
-be quite linear when c is not very close to 1 and m is not very small as it
-linearly makes the interval shorter so that the there are less zero-level
-upcrossings available in the basis functions. You can visualize this using
-{\small \url{https://avehtari.github.io/casestudies/Motorcycle/motorcycle_gpcourse.html}} and
-look how the plot of 6 basis functions change when you set c=1, c=1.5, c=2.25
-(ie $1.5^2$). Instead of looking at mean zero-level crossings, we can look at the
-spectral densities and when they go to zero. This is illustrated in the same
-notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
-explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
-(linear in l) and to explain 80\% of the prior variance we need 2, 4, 8, 16
-basis functions. These spectral densities don't depend on c. When c increaes
-the basis functions are truncated and for example going from c=1.5 to c=2.25,
-basis function 9 has the flexibility basis function 6 had, so 1.5 more basis
-functions are needed. So we have proportionality c/l. }
-
-{\color{red} Gabi: The spectral densities do depend on $c$, see equations (\ref{eq_approxf}) and (\ref{eq_eigenvalue}).}
+% In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
+
+% {\color{blue}Aki: We can also look at the mean zero-level upcrossings, which for QE is
+% proportional to 1/l which gives as that part, and the effect of c should also
+% be quite linear when c is not very close to 1 and m is not very small as it
+% linearly makes the interval shorter so that the there are less zero-level
+% upcrossings available in the basis functions. You can visualize this using
+% {\small \url{https://avehtari.github.io/casestudies/Motorcycle/motorcycle_gpcourse.html}} and
+% look how the plot of 6 basis functions change when you set c=1, c=1.5, c=2.25
+% (ie $1.5^2$). Instead of looking at mean zero-level crossings, we can look at the
+% spectral densities and when they go to zero. This is illustrated in the same
+% notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
+% explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
+% (linear in l) and to explain 80\% of the prior variance we need 2, 4, 8, 16
+% basis functions. These spectral densities don't depend on c. When c increaes
+% the basis functions are truncated and for example going from c=1.5 to c=2.25,
+% basis function 9 has the flexibility basis function 6 had, so 1.5 more basis
+% functions are needed. So we have proportionality c/l. }
+
+% {\color{red} Gabi: The spectral densities do depend on $c$, see equations (\ref{eq_approxf}) and (\ref{eq_eigenvalue}).}
 

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
updates in section 4

--- 2022-02-15 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 Although this still implies linear scaling in $n$, the approximation is more costly than in the univariate case, as $m^{\ast}$ is the product of the number of univariate basis functions over the input dimensions and grows exponentially with respect to the number of  dimensions.
+
+\subsection{Linear representation of a periodic squared exponential covariance function} \label{sec_method_periodic}
+
+A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation.
+In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic square exponential covariance function.
 
@@
 \centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_7_relationships.png}}
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
-  \label{fig6_7_relationships}
-\end{figure*}
-
-Figure \ref{fig6_7_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_7_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+  \label{fig6_relationships}
+\end{figure*}
+
+Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $l$.
+Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
@@
 \item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+This fact can be seen in Figure~\ref{fig6_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 As $\ell$ increases, the minimum valid $c$ also increases.
@@
 
+\subsubsection{Relationships between $J$ and $\ell$ for a periodic squared exponential covariance function} \label{subsec_empiric_relations_periodic}
+
+In Appendix~\ref{sec_periodic}, we analyze the accuracy of the linear representation of a periodic squared exponential covariance function and derive the minimum number of terms $J$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
+Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
+This relationship between $J$ and $\ell$ is gathered in Figure~\ref{figB1_m_lscale_periodic}.
+
 \subsubsection{Numerical equations}
-\label{num_equations}
-
-From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
-Following we show the numerical functions governing these relationships for every kernel.
+\label{sec_num_equations}
+
+From the empirical observations, numerical equations governing the relationships between $m$, $c$ and $\ell$ were estimated, which show an agreement with the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$.
+The estimated numerical equations for each particular covariance function are the following:
 
@@
 \begin{align}\label{eq_m_l_QE}
-&m= 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 1.75 \, \frac{c}{m},
+&m \geq 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 1.75 \, \frac{c}{m},
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_QE}
-&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2      
+&c \geq 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2      
 \end{align}
@@
 \begin{align}\label{eq_m_l_mat32}
-&m= 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 3.42 \, \frac{c}{m},
+&m \geq 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 3.42 \, \frac{c}{m},
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat32}
-&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2        
+&c \geq 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2        
 \end{align}
@@
 \begin{align}\label{eq_m_l_mat52}
-&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
+&m \geq 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 2.65 \, \frac{c}{m},
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat52}
-&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
-\end{align}
-
-Notice that we establish a lower bound for $c$ set to 1.2.
-
-Figure~\ref{fig6_7_relationships} was build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$.
-An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented.
+&c \geq 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2       
+\end{align}
+
+Notice that we establish a lower bound at 1.2 for $c$ in every case.
+
+Figure~\ref{fig6_relationships} and previous equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
+Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$.
 More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions.
-For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented.
+For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$.
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships} to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+However, as an approximation, we can use the unidimensional GP conclusions in Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+
+\subsubsection{Numerical equation for a periodic squared exponential covariance function} \label{sec_num_equation_periodic}
+
+The numerical equation governing the relationships between $J$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\ref{eq_J_l_periodic} which is depicted next:
+%
+\begin{align*}
+&J \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{J}
+\end{align*}
 
@@
 
-As stated above, Figure~\ref{fig6_7_relationships} provides the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) in previous Section~\ref{sec_num_equations} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_7_relationships} (depending on which kernel is used).
+Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided by equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
-If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased.
-We may also use this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if the estimated $\hat{\ell}$ is below the minimum length-scale, repeat the process while increasing $m$ or decreasing $c$.
-As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the length-scale.
-Thus, increasing $m$ may usually the preferred approach.
-
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_7_relationships}.
-
-\subsubsection{A step-by-step user-guide to apply the diagnostics}
+If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated.
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from eq.
+(\ref{eq_m_l_QE}).
+
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, equations in previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to spend minimum computations.
+
+Guidelines for updating $m$ and/or $c$ must be provided:
+
+\begin{itemize}
+	\item $c$ must be big enough for the length-scale $l$ of the function to be learned and,
+	\item $m$ must be big enough as a function of $\ell$ and $c$. 
+\end{itemize}
+
+We recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using equations per Section~\ref{sec_num_equations} , and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$ provided by equations in Section~\ref{sec_num_equations}, update $\ell$ with $\hat{\ell}$ and repeat the process.
+
+A detailed guideline with the steps to perform diagnosis, ensuring that a good approximation is achieved with the minimum computational requirements (i.e.
+minimum number of basis functions $m$) is provided in  next Section~\ref{sec_user_guide}.
+
+\subsubsection{A step-by-step user-guide to apply the diagnostics} \label{sec_user_guide}
 
 Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
-The procedure is split into two phases, Phase A and B, which have to be completed consequtively.
+The procedure is split into two phases, Phase A and B, which have to be completed consecutively.
 %In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.
@@
 \begin{enumerate}
-\item[Step A1.] Make an initial guess on true the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, about $l_1 \approx 1$. Set the iteration index to $i=1$.
-%\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
-\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $l = l_i$ for the given kernel as per Section \ref{num_equations} 
-\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $l = l_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}. 
+
+\item[Step A1.] Make an initial guess on the true length-scale $\ell_1$ of the function to be learned.
+If there is no approximate prior information, we recommend a medium-large length-scale, about $\ell_1 \approx [0.5-1]$.
+Set the iteration index to $i=1$.
+
+\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}
+
+\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
+
 \item[Step A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
-\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.
-
-If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $l_{i+1} = \hat{l}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.
+
+\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i + 0.01 \geq \ell_i$.
+
+If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell_{i+1} = \hat{\ell}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.
 
@@
 
-\item[Step B2.] Set $m_{i+1} = m_{i} + 10$ and increase the iteration index $i = i + 1$.
-
-\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $l = \hat{l}_{i-1}$ for the given kernel as per Section \ref{num_equations}.
-
-\item[Step B4.] Obtain the mimimum valid lengthscale $l_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}.
+\item[Step B2.] Set $m_{i+1} = m_{i} + 5$ and increase the iteration index $i = i + 1$.
+
+\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \hat{\ell}_{i-1}$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[Step B4.] Obtain the mimimum valid lengthscale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
 
@@
 
-\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.
-
-Check the stability of both $\hat{l}_i$ and the measures of predictive accuracy relative to the previous iteration.
+\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i + 0.01 \geq \ell_i$.
+
+Check the stability of both $\hat{\ell}_i$ and the measures of predictive accuracy relative to the previous iteration.
 
@@
 \centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_sec45_comparing_lscale.png}}
-\caption{Estimated exact GP and approximate HSGP lengthscales for various datasets with varying smoothness.}
-  \label{fig_sec45_comparing_lscale}
-\end{figure*}
-
-Figure \ref{fig_sec45_comparing_lscale}-left shows a comparison of the length-scale estimates
-obtained from the exact GP and HSGP models with a squared exponential kernel from various noisy datasets drawn from underlying functions with varying smoothness.
-Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using equation (\ref{eq_c_vs_l_QE}).
-Figure \ref{fig_sec45_comparing_lscale}-right shows the RMSE of the HSGP models with the exact GP model as the reference.
-
-Using large number of basis functions in smooth datasets (with length-scales greater than 0.5) causes overestimating the underlying functions slightly and so the length-scale too (Figure \ref{fig_sec45_comparing_lscale}-left).
-Here it comes out a possible weaknesses of the approach that when we want to fit very smooth functions we should try not to use too many basis functions than necessary because we will run into overestimating problems.
-Apart from that, an accurate estimate of the length-scale (Figure \ref{fig_sec45_comparing_lscale}-left) implies a small RMSE (Figure \ref{fig_sec45_comparing_lscale}-right).
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
+\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP lengthscale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
+  \label{fig7_lscale_comparison}
+\end{figure*}
+
+First of all in this section we illustrate that accurate estimates of the length-scale implies accurate approximations with a small RMSEs.
+Thus, Figure~\ref{fig7_lscale_comparison}-left shows a comparison of the length-scale estimates
+obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness.
+Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.
+(\ref{eq_c_vs_l_QE}).
+Figure \ref{fig7_lscale_comparison}-right shows the RMSE of the HSGP models with the exact GP model as the reference.
+It can be seen that accurate estimates of the length-scale implies small RMSEs.
+
+Furthermore, from Figure~\ref{fig7_lscale_comparison}-left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e.
+true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison}-left).
+Here it comes out a possible weaknesses of the approach which consists in using too many basis functions relative to the true length-scale causes overestimating the posteriors.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_sec45_2_oversmoothing_effect.png}
-\caption{Oversmooth posterior functions when using large number of basis functions for smooth datasets.}
-  \label{fig_sec45_2_oversmoothing_effect}
+\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_oversmoothed_posterios.png}
+\caption{Oversmoothing effect on posterior functions when using too many number of basis functions relative to the wigglyness of the function to be learned.}
+  \label{fig8_oversmoothed_posterios}
 \end{figure}
 
-Next we apply the diagnosis procedure over all the datasets of Figure \ref{fig_sec45_comparing_lscale}-left and we show that following our recomendations and the diagnosis we get the optimum solution with minimum requirements.
+Secondly in this section, Figure~\ref{fig9_diagnosis} shows the iterative steps of applying the diagnosis procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison}-left, showing that by following our recommendations, an optimum solution with minimum computational requirements is achieved.
+Figure~\ref{fig10_diagnosis_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
+Between 2 and 4 iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
+Furthermore, as can be seen in Figure~\ref{fig10_diagnosis_lscale_comparison}, following our diagnostic recommendations avoids overestimation.
+
+As examples, the iterative steps applied to perform diagnosis on two of the data sets in Figure~\ref{fig9_diagnosis} are described next:
+
+\paragraph*{Data set with the exact GP length-scale $\ell_{\text{GP}} = 0.08$:} 
+\begin{itemize}
+\item[] Iteration 1:
+	\begin{enumerate}
+	\item Making the initial guess that the length-scale $\ell_1=0.5$.
+
+	\item Computing the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Computing the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
+
+	\item Fitting the HSGP model and obtain the estimated $\hat{\ell}_1$.
+
+	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives FALSE.
+	\end{enumerate}
+
+\item[] Iteration 2:
+	\begin{enumerate}
+	\item As the diagnostic in previous iteration 1 was FALSE, update $\ell_2$ with $\hat{\ell}_1$.
+
+	\item Computing the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.
+
+	\item Fitting the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+
+	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives FALSE.
+	\end{enumerate}
+
+\item[] Iteration 3:
+	\begin{enumerate}
+	\item As the diagnostic in previous iteration 2 was FALSE, updating $\ell_3$ with $\hat{\ell}_2$ and repeat the process as in iteration 2.
+
+	\item The diagnostic $\hat{\ell}_3 + 0.01 \geq \ell_3$ gives TRUE.
+	\end{enumerate}
+
+\item[] Iteration 4:
+	\begin{enumerate}
+	\item As the diagnostic in step 4 in previous iteration 3 was TRUE, increasing $m_4 = m_3 + 5$.
+
+	\item Computing the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Fitting the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.
+
+	\item The diagnostic $\hat{\ell}_4 + 0.01 \geq \ell_4$ gives TRUE.
+
+	\item As RMSE, $R^2$ and ELPD are stable relative to previous iteration 4 the approximation must be sufficiently accurate.
+	\end{enumerate}
+
+\end{itemize}
+
+\paragraph*{Data set with the exact GP length-scale $\ell_{\text{GP}} = 1.4$:} 
+\begin{itemize}
+\item[] Iteration 1:
+	\begin{enumerate}
+	\item Making the initial guess that the length-scale $\ell_1=1$.
+
+	\item Computing the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Computing the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
+
+	\item Fitting the HSGP model and obtain the estimated $\hat{\ell}_1$.
+
+	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives TRUE.
+	\end{enumerate}
+
+\item[] Iteration 2:
+	\begin{enumerate}
+	\item As the diagnostic in previous iteration 1 was TRUE, increasing $m_2 = m_1 + 5$.
+
+	\item Computing the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Fitting the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+
+	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives TRUE.
+
+	\item As RMSE, $R^2$ and ELPD are stable relative to previous iteration 2 the approximation must be sufficiently accurate.
+	\end{enumerate}
+
+\end{itemize}
+
+It is worth emphasizing that using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential in uni-dimensional cases with very large data sets, and even more in low-dimensional cases ($D \leq 4$; Notice from section~\ref{sec_method_multi} that for $D > 4$ it starts to be unpractical in any case) even with small data sets, since knowing the optimum values for $m$ and $c$ for each individual dimension can reduce significantly the computational requirements in these multi-dimensional cases, otherwise they would be unfeasible to be fitted.
+
+\begin{figure*}
+\centering
+\begin{tabular}{ c c c c c c c c c c | c c}
+\arrayrulecolor{gray}\hline \\[-3mm]
+$\ell_{\text{GP}}$ & iter. & $\ell$ & $c$ & $m$ & $\hat{\ell}$ & $\hat{\ell} + 0.01 \geq \ell$ & RMSE & $R^2$ & ELPD & RMSE* & \\ 
+%\arrayrulecolor{lightgray}\hline \\[-1mm]
+%\multirow{5}{*}{ 0.05} 
+% & 1 & 0.50 & 1.60 & 6 & 0.127 & FALSE & 0.800 & 0.264 & -0.980 & 0.748 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_1.png}}\\
+% & 2 & 0.127 & 1.20 & 17 & 0.061 & FALSE & 0.427 & 0.779 & -0.396 & 0.330 & \\
+% & 3 & 0.061 & 1.20 & 35 & 0.059 & TRUE & 0.264 & 0.908 & 0.008 & 0.063 & \\
+% & 4 & 0.053 & 1.20 & 40 & 0.059 & TRUE & 0.254 & 0.915 & 0.088 & 0.029 & \\
+% \\[1mm]
+\arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{5}{*}{ 0.08} 
+ & 1 & 0.50 & 1.60 & 6 & 0.169 & FALSE & 0.447 & 0.610 & -0.396 & 0.360 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_2.png}}\\
+ & 2 & 0.169 & 1.20 & 13 & 0.068 & FALSE & 0.357 & 0.738 & -0.176 & 0.242 & \\
+ & 3 & 0.068 & 1.20 & 31 & 0.080 & TRUE & 0.242 & 0.871 & 0.204 & 0.013 & \\
+ & 4 & 0.058 & 1.20 & 36 & 0.081 & TRUE & 0.241 & 0.871 & 0.207 & 0.009 & \\
+ \\[1mm]
+ \arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{3}{*}{ 0.13} 
+ & 1 & 0.50 & 1.60 & 6 & 0.250 & FALSE & 0.461 & 0.625 & -0.482 & 0.330 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_3.png}}\\
+ & 2 & 0.250 & 1.20 & 9 & 0.148 & FALSE & 0.352 & 0.780 & -0.135 & 0.153 & \\
+ & 3 & 0.148 & 1.20 & 15 & 0.144 & TRUE & 0.321 & 0.807 & -0.074 & 0.084 & \\
+ & 4 & 0.105 & 1.20 & 20 & 0.117 & TRUE & 0.297 & 0.830 & -0.006 & 0.013 & \\
+ \\[1mm]
+\arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{3}{*}{ 0.25} 
+ & 1 & 0.50 & 1.60 & 6 & 0.237 & FALSE & 0.368 & 0.399 & -0.146 & 0.214 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_4.png}}\\
+ & 2 & 0.237 & 1.20 & 9 & 0.263 & TRUE & 0.294 & 0.603 & -0.085 & 0.024 & \\
+ & 3 & 0.250 & 1.20 & 14 & 0.257 & TRUE & 0.289 & 0.615 & -0.050 & 0.009 & \\
+ \\
+ \\[1mm]
+\arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{3}{*}{ 0.54} 
+ & 1 & 0.50 & 1.60 & 6 & 0.376 & FALSE & 0.272 & 0.858 & 0.034 & 0.013 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_5.png}}\\
+ & 2 & 0.376 & 1.20 & 6 & 0.469 & TRUE & 0.277 & 0.855 & 0.029 & 0.051 & \\
+ & 3 & 0.239 & 1.50 & 11 & 0.525 & TRUE & 0.272 & 0.858 & 0.036 & 0.014 & \\
+ \\
+ \\[1mm]
+\arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{3}{*}{ 0.80} 
+ & 1 & 1.00 & 3.20 & 6 & 0.516 & FALSE & 0.286 & 0.923 & 0.133 & 0.025 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_6.png}}\\
+ & 2 & 0.516 & 1.65 & 6 & 0.975 & TRUE & 0.286 & 0.923 & 0.107 & 0.030 & \\
+ & 3 & 0.496 & 3.12 & 11 & 0.772 & TRUE & 0.283 & 0.924 & 0.109 & 0.005 & \\
+ \\
+ \\[1mm]
+\arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{2}{*}{ 1.40} 
+ & 1 & 1.00 & 3.20 & 6 & 1.023 & TRUE & 0.283 & 0.561 & 0.139 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_7.png}}\\
+ & 2 & 0.521 & 3.27 & 11 & 1.233 & TRUE & 0.281 & 0.566 & 0.136 & 0.003 & \\
+ \\
+ \\
+ \\[1mm]
+ \arrayrulecolor{lightgray}\hline \\[-1mm]
+\multirow{2}{*}{ 2.94} 
+ & 1 & 0.50 & 1.60 & 6 & 1.192 & TRUE & 0.343 & 0.449 & -0.104 & 0.058 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_8.png}}\\
+ & 2 & 0.607 & 3.81 & 11 & 2.577 & TRUE & 0.341 & 0.454 & -0.074 & 0.008 & \\
+ \\
+ \\
+ \\[1mm]
+\arrayrulecolor{gray}\hline
+\end{tabular}
+\caption{Diagnosis on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+  \label{fig9_diagnosis}
+\end{figure*}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_diagnosis_lscale_comparison.png}}
+\caption{Comparison of the true and estimated length-scales in every iteration for the different data sets varying the length-scale.}
+  \label{fig10_diagnosis_lscale_comparison}
+\end{figure}
 
@@
 The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
-The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_7_relationships} is not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
+The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_7_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52} are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
 \eqref{eq:rel_total_distance}.
@@
 $\theta_0$ contains the hyperparameters $\alpha_0$ and $\ell_0$ which are the marginal variance and length-scale of the general mean GP function, and $\theta_1$ contains the hyperparameters $\alpha_1$ and $\ell_1$ which are the marginal variance and length-scale, respectively, of a GP function specific to the male sex ($x_2=2$).
-Scalar length-scales, $l_0$ and $l_1$, are used in both multivariate covariance functions, assuming isotropic functions.
+Scalar length-scales, $\ell_0$ and $\ell_1$, are used in both multivariate covariance functions, assuming isotropic functions.
 %
@@
 Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
+
+From the empirical observations, a numerical equation governing the relationships between $J$ and $\ell$ were estimated in equation \ref{eq_J_l_periodic}, which show linear proportionality between $J$ and $\ell$.
+%
+\begin{align}\label{eq_J_l_periodic}
+&J \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{J},
+\end{align}
 

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edited 4.3.1

--- 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 
-When considering exponentiated quadratic and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
+When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
 With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$ and with increasing $c$, the $m$ should grow near linearly with $c$.
@@
 
-From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
+When $c$ is large enough, there is near
+linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ (see Section~\ref{subsec_theoretic_linearity}).
+To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the needed constant terms.
+We establish a lower bound for $c$ set to 1.2.
+
+% From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
 Following we show the numerical functions governing these relationships for every kernel.
@@
 
+\paragraph*{Mat\'ern ($\nu$=5/2)}:
+%
+\begin{align}\label{eq_m_l_mat52}
+&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
+\end{align}
+%
+with
+%
+\begin{align}\label{eq_c_vs_l_mat52}
+&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
+\end{align}
+
 \paragraph*{Mat\'ern ($\nu$=3/2)}:
@@
 
-\paragraph*{Mat\'ern ($\nu$=5/2)}:
-%
-\begin{align}\label{eq_m_l_mat52}
-&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
-\end{align}
-%
-with
-%
-\begin{align}\label{eq_c_vs_l_mat52}
-&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
-\end{align}
-
-Notice that we establish a lower bound for $c$ set to 1.2.
+These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$).
+Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions.
+If the aim is to find minimal $m$ to speedup the computation, a further refined formula can be obtained for new covariance functions.
 

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
more edits

@@
 \item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
-As $\ell$ increases, the minimum valid $c$ also increases.
+This fact can be seen in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships}
+% TODO FIX the figure reference!
+as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$). As $\ell$ increases, the minimum valid $c$ also increases.
 \end{itemize}
@@
 As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the length-scale.
-Thus, increasing $m$ may usually the preferred approach.
-
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_7_relationships}.
+Thus, increasing $m$ is usually the preferred approach.
+
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also match those obtained from Figure~\ref{fig6_7_relationships}.
 
 \subsubsection{A step-by-step user-guide to apply the diagnostics}
+
+If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the first fitting, it is likely that the result is sufficiently accurate.
+The following iterative approach works as safety approach and allows also finding a smaller sufficient $m$ which can be useful if the user wants to optimized the computational speed for repeated fitting (e.g.
+in cross-validation or simulation based calibration).
 
@@
 \begin{enumerate}
-\item[Step A1.] Make an initial guess on true the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, about $l_1 \approx 1$. Set the iteration index to $i=1$.
+\item[Step A1.] Make an initial guess on the lengthscale $l_1$.
+If there is no useful information available, we recommend to start with a large lengthscale, about $l_1 \approx 1$.
+Set the iteration index to $i=1$.
 %\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
-\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $l = l_i$ for the given kernel as per Section \ref{num_equations} 
+\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $l = l_i$ for the given kernel as per Section \ref{num_equations} .
 \item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $l = l_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}. 

### 2022-02-16 Paper/Case-study_Birthday-data/Rnotebook_birthday.Rmd

---
title: 'Case study: Birthday data'
author: "G. Riutort-Mayol, M.R. Andersen, P.C. Burkner, A. Solin and A. Vehtari"
date: "`r format(Sys.Date())`."
output:
  html_document:
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_download: yes
---

<style type="text/css">
body, td{ font-size: 16px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
</style>

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r message=FALSE, warn=FALSE, eval=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(ggplot2)
library(loo)
library(latex2exp)
```

## Data

```{r message=FALSE, warn=FALSE}
data <- read.csv(file="../births_usa_1969.csv")
str(data)
```

```{r message=TRUE, warn=TRUE, fig.height=5, fig.width=8}
par(mai=c(1.02,1.2,0.3,0.42))

ind <- data$id
births <- data$births
labels_at = aggregate(data, by=list(data$year), FUN=min)$id

plot(ind, births[ind], type="p", pch=20, bg=grey(0.4), cex=0.6, col=grey(0.5), xlab="", ylab="", lwd=1, mgp=c(3.5, 1, 0), frame.plot=TRUE, yaxs="r", cex.axis=1.2, cex.lab=1.2, las=1, xaxt="n", yaxt="n", fg=grey(0.5), family="serif")
axis(1, at = labels_at, c("1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988"), lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.2, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.2, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab="Year", mgp=c(2.2, 1, 0), cex.lab=1.2, las=1)
title(ylab="Births", mgp=c(3.7, 0.7, 0), cex.lab=1.2, las=1)
legend("topleft",inset=c(0.22,0.02),legend=c("Observations"), col=grey(0.4), lty=NA, pch=20, lwd=1, cex=1, xpd=TRUE, bty="n", x.intersp=0.1, text.font=1)
```

## HSGP modeling and fitting

Input vector

```{r message=FALSE, warn=FALSE}
x <- scale(data$id[], center=TRUE, scale=TRUE)
```

Output vector

```{r message=FALSE, warn=FALSE}
y <- scale(data$births[], center=TRUE, scale=TRUE)
```

Standard deviation and mean of the output

```{r message=FALSE, warn=FALSE}
std_y <- attr(y,"scaled:scale")
m_y <- attr(y,"scaled:center")
```

Standard deviation of the input

```{r message=FALSE, warn=FALSE}
std_x <- attr(x,"scaled:scale")
```

Year and week periods

```{r message=FALSE, warn=FALSE}
period_year <- 365.25/std_x
period_week <- 7/std_x
```

`Day of the year2` is a variable that indicates the correspondence between all the days with the days of a leap-year

```{r message=FALSE, warn=FALSE}
day_of_year2 <- data$day_of_year2
```

### HSGP tunnig factors

Number of basis functions $M$ for the smooth trend function $f_1$

```{r message=FALSE, warn=FALSE}
M_f1 <- 40
```

Boundary factor $c$ for $f_1$

```{r message=FALSE, warn=FALSE}
c_f1 <- 1.5
```

Number of basis functions $J$ for year effects function $f_2$

```{r message=FALSE, warn=FALSE}
J_f2 <- 20		
```

Number of basis functions $J$ for the week effects function $f_3$

```{r message=FALSE, warn=FALSE}
J_f3 <- 7
```

### Data to be passed to Stan

```{r message=FALSE, warn=FALSE}
standata <- list(M_f1= M_f1, 
                 c_f1= c_f1, 
                 J_f2= J_f2, 
                 J_f3= J_f3, 
                 x= x[,1], 
                 y= y[,1], 
                 N= length(x), 
                 period_year= period_year, 
                 period_week= period_week,
                 day_of_year2= day_of_year2
)
```

### Stan model

```
//saved in 'stancode_def.stan'
functions {
  vector diagSPD_EQ(real gpscale, real lscale, real L, int M) {
    return sqrt((gpscale^2) * sqrt(2*pi()) * lscale * exp(-0.5*(lscale*pi()/2/L)^2 * linspaced_vector(M, 1, M)^2));
  }
  vector diagSPD_periodic(real gpscale, real lscale, int M) {
    real a = 1/lscale^2;
    int one_to_M[M];
    for (m in 1:M) one_to_M[m] = m;
    vector[M] q = sqrt(gpscale^2 * 2 / exp(a) * to_vector(modified_bessel_first_kind(one_to_M, a)));
    return append_row(q,q);
  }
  matrix PHI_EQ(int N, int M, real L, vector x) {
    return sin(diag_post_multiply(rep_matrix(pi()/(2*L) * (x+L), M), linspaced_vector(M, 1, M)))/sqrt(L);
  }
  matrix PHI_periodic(int N, int M, real w0, vector x) {
    matrix[N,M] mw0x = diag_post_multiply(rep_matrix(w0*x, M), linspaced_vector(M, 1, M));
    return append_col(cos(mw0x), sin(mw0x));
  }
}
data {
    real c_f1;                  //boundary value for function 1
    int<lower=1> M_f1;          //num basis functions for function 1
    int<lower=-1> J_f2;         //num cosine and sinu functions for function 3
    int<lower=-1> J_f3;         //num cosine and sinu functions for function 4
    int<lower=1> N;             //num observations
    vector[N] x;                //input vector
    vector[N] y;                //target vector
    real period_year;           //period of the year
    real period_week;           //period of the week
    int day_of_year2[N];        //day of the year inside a leap-year
}
transformed data {
    real L_f1= c_f1*max(fabs(x));
    //Basis functions for f1, f2 and f3
    matrix[N,M_f1] PHI_f1 = PHI_EQ(N, M_f1, L_f1, x);
    matrix[N,2*J_f2] PHI_f2 = PHI_periodic(N, J_f2, 2*pi()/period_year, x);
    matrix[N,2*J_f3] PHI_f3 = PHI_periodic(N, J_f3, 2*pi()/period_week, x);
}
parameters {
    vector[N] f;
    real intercept;
    //variables for the basis function models
    vector[M_f1] beta_f1;
    vector[2*J_f2] beta_f2;
    vector[2*J_f3] beta_f3;
    //hyperparameters
    vector<lower=0>[3] lscale;
    vector<lower=0>[3] gpscale;
    real<lower=0> noise;
    //
    //t-student prior for special days effects
    vector[366] f4;
    real<lower=0> sigma_f4;
}
transformed parameters{
    vector[N] f1;
    vector[N] f2;
    vector[N] f3;
    {
    vector[M_f1] diagSPD_f1 = diagSPD_EQ(gpscale[1], lscale[1], L_f1, M_f1);
    vector[2*J_f2] diagSPD_f2 = diagSPD_periodic(gpscale[2], lscale[2], J_f2);
    vector[2*J_f3] diagSPD_f3 = diagSPD_periodic(gpscale[3], lscale[3], J_f3);
    //
    vector[M_f1] SPD_beta_f1 = diagSPD_f1 .* beta_f1;
    vector[2*J_f2] SPD_beta_f2 = diagSPD_f2 .* beta_f2;
    vector[2*J_f3] SPD_beta_f3 = diagSPD_f3 .* beta_f3;
    //
    f1 = PHI_f1[,] * SPD_beta_f1;
    f2 = PHI_f2[,] * SPD_beta_f2;
    f3 = PHI_f3[,] * SPD_beta_f3;
    }
}
model{
    intercept ~ normal(0,1);
    beta_f1 ~ normal(0,1);
    beta_f2 ~ normal(0,1);
    beta_f3 ~ normal(0,1);
    //
    lscale ~ normal(0,2);           //GP lengthscales
    gpscale ~ normal(0,10);         //GP magnitudes
    noise ~ normal(0,1);            //model noise
    //
    //t-student prior for special days effects
    f4 ~ student_t(1, 0, sigma_f4);
    sigma_f4 ~ normal(0, 0.1);
    //
    f= intercept + f1 + f2 + f3 + f4[day_of_year2];
    target += normal_lpdf(y | f, noise);
}
generated quantities{
  vector[N] y_rep;
  vector[N] log_lik;
  for(n in 1:N){
    y_rep[n] = normal_rng(f[n], noise);
    log_lik[n] = normal_lpdf(y[n] | f[n], noise);
  }
}
```

### Model sampling

Compiling the model

```{r message=FALSE, warn=TRUE, eval=TRUE}
birthday_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_def.stan")
```

Model sampling (using `cmdstanr` package)

```{r warn=TRUE, message=FALSE, eval=TRUE}
fit <- birthday_mod$sample(data= standata, iter_warmup=200, iter_sampling=200, chains=4, thin=4, init=0.5, adapt_delta=0.9, save_warmup=FALSE)
```

<!-- Save or load model output -->

```{r message=TRUE, warn=TRUE, eval=FALSE, include=FALSE}
# save(fit, file="fit.rData")
# load("fit.rData")
```

<!-- Permanent saving of the output CSV files in the local directory -->

```{r message=TRUE, warn=TRUE, eval=FALSE, include=FALSE}
fit$save_output_files(dir = getwd())
```

Conversion of the model output from 'cmdstanr' package to 'rstan' package. This uses the output CSV files saved at a temporary directory. If you want a permament saving of these files at a local directory use `cmdstanr::fit$save_output_files()`

```{r message=FALSE, warn=FALSE, eval=TRUE}
fit_rstan <- rstan::read_stan_csv(fit$output_files())
```

<!-- Diagnostics of the sampling -->

```{r message=TRUE, warn=TRUE, eval=FALSE, include=FALSE}
check_hmc_diagnostics(fit_rstan)
```

## Model diagnosis

Posterior samples

```{r message=FALSE, warn=FALSE, eval=TRUE}
posterior <- as.array(fit_rstan)
```

Log posterior density

```{r message=FALSE, warn=FALSE, eval=TRUE}
lp <- log_posterior(fit_rstan)
```

Values for parameters of the performance of NUTS sampler

```{r message=FALSE, warn=FALSE, eval=TRUE}
np <- nuts_params(fit_rstan)
```

Draws

```{r message=FALSE, warn=FALSE, eval=TRUE}
draws <- extract(fit_rstan)
```

Summaries of variable estimates

```{r message=FALSE, warn=FALSE, eval=TRUE}
param = c("intercept","lscale[1]","lscale[2]","lscale[3]","gpscale[1]","gpscale[2]","gpscale[3]","noise","sigma_f4")
summary(fit_rstan, pars = param, probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
```

Simulation chains for the variables after warmup

```{r message=FALSE, warn=FALSE, fig.height=10, fig.width=15, eval=TRUE}
traceplot(fit_rstan, pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Bivariate samples plots

```{r message=FALSE, warn=FALSE, fig.height=10, fig.width=15, eval=TRUE}
mcmc_pairs(posterior, np= np, pars= param, off_diag_args= list(size = 1))
```

Autocorrelation plots

```{r message=FALSE, warn=FALSE, fig.height=5, fig.width=8, eval=TRUE}
mcmc_acf(posterior, pars = param, lags = 10)
```

Posterior predictive checks

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ppc_dens_overlay(y[,1], draws$y_rep)
```

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ppc_stat(y[,1], draws$y_rep, stat = "median")
```

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ppc_stat(y[,1], draws$y_rep, stat = "mean")
```

Leave-one-out PIT values using `loo` package

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
loo1 <- loo(fit_rstan, save_psis = TRUE, cores = 2)
psis1 <- loo1$psis_object
lw <- weights(psis1)
ppc_loo_pit_overlay(y[,1], draws$y_rep, lw)
```

Number of effective samples ratios for variables

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ratios <- neff_ratio(fit_rstan, pars= param)
mcmc_neff(ratios, size = 2) + yaxis_text(hjust = 0)
```

Rhat statistics for sampling chains

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
rhats <- rhat(fit_rstan, pars= param)
mcmc_rhat(rhats) + yaxis_text(hjust = 0)
```

Divergence

```{r message=FALSE, warn=FALSE, fig.height=4, fig.width=4, eval=TRUE}
mcmc_nuts_divergence(np, lp)
```

Energy

```{r message=FALSE, warn=FALSE, fig.height=5, fig.width=6, eval=TRUE}
mcmc_nuts_energy(np)
```

## Ploting estimated functions

Extracting estimated function components $f_1$, $f_2$, $f_3$ and $f_4$

```{r message=FALSE, warn=FALSE, eval=TRUE}
f <- summary(fit_rstan, pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f1 <- summary(fit_rstan, pars = c("f1"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f2 <- summary(fit_rstan, pars = c("f2"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f3 <- summary(fit_rstan, pars = c("f3"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f4 <- summary(fit_rstan, pars = c("f4"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
```

#### Plot of only one year, the year 1972

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
data_year <- data[data$year==1972,]
ind <- data_year$id
axis_labels_at <- aggregate(data_year, by=list(data_year$month), FUN=min)$id

plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", pch=21, bg=grey(0.7), cex=0.9, col=grey(0.4), xlab="", ylab="", lwd=1, ylim=c(0.73,1.16), mgp= c(2, 1, 0), frame.plot = TRUE, yaxs="r", cex.axis=1.6, cex.lab=1.6, las=1, xaxt="n", yaxt="n", fg=grey(0.5), family="serif")
axis(1, at=axis_labels_at, labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"), tick=TRUE, lty=1, mgp=c(2, 1, 0), las=1, cex.axis=1.6, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, at=NULL, labels=TRUE, tick=TRUE, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.6, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab ="Month", mgp= c(3, 1, 0), cex.lab=1.6, las=1)
title(ylab ="Proportion of births over the mean", mgp= c(3, 0.7, 0), cex.lab=1.6, las=1)

lines(ind, (f[ind,1]*std_y+m_y)/m_y, col=grey(0.6), lwd=1)	# f
abline(h=1, lty=2)						                              # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=2)			    # f1 smooth trend
lines(ind, (f2[ind,1]*std_y+m_y)/m_y, col=3, lwd=2)	 		    # f3 year effect
lines(ind, (f4[data$day_of_year2,1][ind]*std_y+m_y)/m_y, col=6, lwd=2) 		    # f5 horseshoe

#labels special days
text(ind[data_year$month==1&data_year$day==1]+2, y = 0.84, labels = "New year", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==2&data_year$day==14], y = 1.045, labels = "Valentine's day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==2&data_year$day==29], y = 0.978, labels = "Leap day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==4&data_year$day==1]+2, y = 0.980, labels = "April 1st", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==5&data_year$day==27], y = 0.98, labels = "Memorial day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==7&data_year$day==4], y = 0.86, labels = "Independence day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==9&data_year$day==2], y = 0.97, labels = "Labor day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==10&data_year$day==30]+2, y = 0.985, labels = "Halloween", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==11&data_year$day==25]+5, y = 0.95, labels = "Thanks-giving", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==12&data_year$day==25], y = 0.82, labels = "Christmas", pos =NULL, offset = 0, family="serif", cex=1.3)

legend("topleft", inset=c(0.05,0.01), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)'),TeX('Special days effects ($f_4$)'),TeX('$\\mu$=f_1+f_2+f_3+f_4$')), col=c(grey(0.5),2,3,6,grey(0.5)), lty=c(NA,1,1,1,1), pch=c(20,NA,NA,NA,NA), lwd=c(2,3,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=2, seg.len=1.3)
```

#### Plot of all the years

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
ind <- data$id
labels_at = aggregate(data, by=list(data$year), FUN=min)$id
plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", pch=20, bg=grey(0.4), cex=0.6, col=grey(0.5), xlab="", ylab="", lwd=1, ylim=c(0.7,1.3), mgp=c(3.5, 1, 0), frame.plot=TRUE, yaxs="r", cex.axis=1.6, cex.lab=1.6, las=1, xaxt="n", yaxt="n", fg=grey(0.5), family="serif")

axis(1, at=labels_at, c("1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988"), tick=TRUE, lty=1, mgp=c(3, 1, 0), las=1, cex.axis=1.6, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, at=NULL, labels=TRUE, tick=TRUE, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.6, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab="Year", mgp=c(3, 1, 0), cex.lab=1.6, las=1)
title(ylab="Proportion of births over the mean", mgp=c(3, 0.7, 0), cex.lab=1.6, las=1)

abline(h=1, lty=2)												                # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=2)				# smooth trend
lines(ind, (f2[ind,1]*std_y+m_y)/m_y, col=3, lwd=2)				# year effect

legend("topleft", inset=c(0.22,0.02), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)')), col=c(grey(0.4),2,3,"grey"), lty=c(NA,1,1,1), pch=c(20,NA,NA,NA), lwd=c(1,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=1, seg.len=1.5)
```

#### Plot of only the first month

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
data_month <- data[data$month==1&data$year==1972,]
ind <- data_month$id
axis_labels_at <- aggregate(data_month, by=list(data_month$day), FUN=min)$id

id_week <- data_month$id[data_month$day_of_week==1]

plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", pch=21, bg=grey(0.7), cex=1.2, col=grey(0.4), xlab="", ylab="", lwd=1, ylim=c(0.7,1.2), mgp=c(3.5, 1, 0), frame.plot=TRUE, yaxs="r", cex.axis=1.6, cex.lab=1.6, las=1, xaxt="n", yaxt="n",fg=grey(0.5), family="serif")

axis(1, at=axis_labels_at[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29,31)], labels=as.character(1:31)[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29,31)], tick=TRUE, lty=1, mgp=c(3, 1, 0), las=1, cex.axis=1.6, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(1, at=id_week, labels=rep(c("Monday"),5), tick=TRUE, lty=1, mgp=c(-1, -1.2, 0), las=1, cex.axis=1.3, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, at=NULL, labels=TRUE, tick=TRUE, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.6, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab="Day", mgp=c(3, 1, 0), cex.lab=1.6, las=1)
title(ylab="Proportion of births over the mean", mgp=c(3, 0.7, 0), cex.lab=1.6, las=1)

lines(ind, (f[ind,1]*std_y+m_y)/m_y, col="black", lwd=2)	# f
abline(h=1, lty=2)											                  # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=2)			  # f1 smoth trend
lines(ind, (f2[ind,1]*std_y+m_y)/m_y, col=3, lwd=2)	 		  # f3 year effect
lines(ind, (f3[ind,1]*std_y+m_y)/m_y, col=4, lwd=2)	 		  # f4 week effect
lines(ind, (f4[data$day_of_year2,1][ind]*std_y+m_y)/m_y, col=6, lwd=2) 		  # f5 horseshoe

abline(v=id_week, lty=2, col="grey")

#labels special days
text(ind[data_month$month==1&data_month$day==1]-0.05, y = 0.84, labels = "New year", pos =NULL, offset = 0, family="serif", cex=1.3, xpd=TRUE,)

legend("topleft", inset=c(0.107,0.005), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)'),TeX('Week effects ($f_3$)'),TeX('Special-days effects ($f_4$)'),TeX('$\\mu$=f_1+f_2+f_3+f_4$')), col=c(grey(0.5),2,3,4,6,"black"), lty=c(NA,1,1,1,1,1), pch=c(20,NA,NA,NA,NA,NA), lwd=c(3,3,3,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=2, seg.len=1.3)
```

#### Plot of the first four years

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
data_year <- data[data$year==1969 | data$year==1970 | data$year==1971 | data$year==1972,]
ind <- data_year$id
axis_labels_at <- aggregate(data_year, by=list(data_year$year), FUN=min)$id

plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", lty=1, pch=18, cex=0.4, col="black", xlab="", ylab="Proportion of births over the mean", cex.lab=1.5, cex.axis=1.6, xaxt="n", ylim=c(0.7,1.3))

axis(1, at=axis_labels_at, labels=c("1969","1970","1971","1972"), tick=TRUE, cex.axis=1.6)

lines(ind, (f[ind,1]*std_y+m_y)/m_y, col="grey", lwd=1)		# f
lines(range(ind), c(1,1), lty=2, lwd=2)						        # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=1)			  # f1
lines(ind, (f3[ind,1]*std_y+m_y)/m_y, col=3, lwd=1)	 		  # f3
lines(ind, (f4[data$day_of_year2,1][ind]*std_y+m_y)/m_y, col=6, lwd=1) 		  # f5

legend("topleft", inset=c(0.05,0.005), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)'),TeX('Week effects ($f_3$)'),TeX('Special-days effects ($f_4$)'),TeX('$\\mu$=f_1+f_2+f_3+f_4$')), col=c(grey(0.5),2,3,4,6,"black"), lty=c(NA,1,1,1,1,1), pch=c(20,NA,NA,NA,NA,NA), lwd=c(3,3,3,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=2, seg.len=1.3)
```

### 2022-02-16 Paper/Manuscript_Stats&Comp/1st_journal_review/Diagnosis_analysis/Rnb_Diagnosis_analysis.Rmd

---
title: 'Section 4.2 of the paper'
date: "`r format(Sys.Date())`."
output:
  bookdown::html_document2:
  # html_document:
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: true
    code_download: yes
    number_sections: true
    df_print: kable # paged # 
---

<style type="text/css">
  body, td{ font-size: 14px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
h1.title {
  font-size: 24px;
  color: Black;
}
h1 { /* Header 1 */
    font-size: 22px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
}
h3 { /* Header 3 */
    font-size: 16px;
  color: Black;
}
h4 { /* Header 4 */
    font-size: 14px;
  color: Black;
}
#TOC {
color: DarkBlue;
font-size: 16px;
}
</style>

```{r eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, comment="", eval=TRUE, include=FALSE)
```

```{r message=FALSE, eval=TRUE, warning=FALSE, include=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(latex2exp)
library(reshape2)
SEED <- 4945 #Seed for reproducibility
```

# Relationships among $m$, $l$ and $c$

The relationship among the number of basis functions $m$, the lengthscale $l$, and the boundary factor $c$, for a quadratic exponentiated kernel adjusted on the empirical observations are:

$$ m= 1.73 \, \frac{c}{\left(\frac{l}{S}\right)^{1.05}} \;\; \Leftrightarrow \;\; \frac{l}{S}= 1.68 \, \frac{c^{0.95}}{m^{0.95}}$$

where $S$ is the half range of the input domain. Assuming linear proportionality between $m$ and $c/l$ and between $l$ and $c/m$, the adjusted coefficients of this relationship become:

$$\begin{align*}
m= 1.97 \, \frac{c}{\frac{l}{S}} \;\; \Leftrightarrow \;\; \frac{l}{S}= 1.97 \, \frac{c}{m}        (\#eq:m-l-QE)
\end{align*}$$

```{r eval=TRUE, include=TRUE}
m_QE <- function(c,l,S){ #ceiling(1.73 * c / (l/S)^1.05)
                        ceiling(1.97 * c / (l/S)) }

l_QE <- function(c,m,S){ #round(S * 1.68 * c / m, 3)
                        round(S * 1.97 * c / m, 3) }
```

In the above equations, the boundary factor $c$ has to fulfill \@ref(eq:c-vs-l-QE), as a function of $l$, and \@ref(eq:c-vs-m-QE), as a function of $m$.

$$\begin{align*}
\, c = c(l) \, \geq \, 0.97 + 1.45 \, \frac{l}{S}         (\#eq:c-vs-l-QE)
\end{align*}$$

```{r eval=TRUE, include=TRUE}
c_vs_l_QE <- function(l,S){  #,m,m_max
    c = 0.97+1.45*l/S #+(1*(l/S)^2)*(m-5)/(m_max-5)
    if(c < 1.25){
      c = 1.5
    }else{
      if(c >= 1.25 & c < 1.45){
        c = 1.5
      }else{
        if(c >= 1.45 & c < 1.95){
          c = 2
        }else{
          if(c >= 1.95 & c < 2.45){
            c = 2.5
          }else{
            if(c >= 2.45 & c < 2.95){
              c = 3
            }else{
              c = ceiling(c)
            } } } } }
  c
}
```

$$\begin{align*}
\, c = c(m) \, \geq \, 234.159 \, m^{-0.316 \, \text{log}^2m \, + \,  2.370 \, \text{log}m \,  - \, 6.008}       (\#eq:c-vs-m-QE)  
\end{align*}$$

```{r eval=TRUE, include=TRUE}
c_vs_m_QE <- function(m){
  c = 234.159*m^(-0.316*log(m)^2+2.370*log(m)-6.008)
  if(c < 1.25 | m > 30){
    c = 1.5
  }else{
    if(c >= 1.25 & c < 1.45){
      c = 1.5
    }else{
      if(c >= 1.45 & c < 1.95){
        c = 2
      }else{
        if(c >= 1.95 & c < 2.45){
          c = 2.5
        }else{
          if(c >= 2.45 & c < 2.95){
            c = 3
          }else{
            c = ceiling(c)
          } } } } }
  c
}
```

When $c$ can be determined either by a function of $l$ and $m$, the maximum between both has to be selected:

$$\begin{align*}
\text{max}\left(c(l), c(m)\right)        (\#eq:max-c)
\end{align*}$$

```{r eval=TRUE, include=TRUE}
c_vs_l_m_QE <- function(l,m,S) max(c_vs_l_QE(l,S), c_vs_m_QE(m))  #,m_max
```

# Diagnostic function

The diagnostic to evaluate whether the approximation is sufficiently accurate is:

$$\begin{align*}
\, \hat{l} \pm 0.02 \geq l          (\#eq:diagnostic)
\end{align*}$$

where $\hat{l}$ is the lengthscale estimate in the HSGP approximation and $l$ is the minimum lengthscale determined by the $m$ and $c$ used in the HSGP approximation by using equation \@ref(eq:m-l-QE). Notice the threshold $\,\pm 0.02$ in the diagnostic.

```{r eval=TRUE, include=TRUE}
diagnostic <- function(l,l_hat) l_hat + 0.02 >= l | l_hat - 0.02 >= l
```

# User-guide for diagnosis

**Assumption of the diagnosis tool:**

- Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is smaller than the exact GP lengthscale estimate $l$.

**User-guide with the steps to perform diagnosis:**

0. Make a first guess on the lengthscale $l_1$ of the function to be learned.

*Initial iteration*, $i=1$

1. Obtain the valid minimum boundary factor $c_1$ determined by $l_1$ by using \@ref(eq:c-vs-l-QE).

2. Obtain the valid mimimum number of basis functions $m_1$ determined by $l_1$ and $c_1$ by using \@ref(eq:m-l-QE). Notice that $l_1$ can also be read as the minimum lengthscale that can be accurately fitted determined by $m_1$ and $c_1$ by using \@ref(eq:m-l-QE).

3. Fit the HSGP model and assess residuals: check any residual trends, compute *rmse* (Root mean square error), $R^2$ (Coefficient of variation) and *elpd* (Expected log predicitve density).

4. Check the diagnostic of whether $\, \hat{l_1} \pm 0.02 \geq l_1$ (eq. \@ref(eq:diagnostic)).

    4.1. If the diagnostic is TRUE, the HSGP model approximation must be sufficiently accurate or, at least, very close to be sufficiently accurate. Then user can continue with step 5.

    4.2. If the diagnostic is FALSE, the HSGP model approximation can not be sufficiently accurate. Then user has to move to step 10.

*Next iterations*, $i>1$

- If the diagnostic in step 4 is TRUE

    5. Set $\, m_i = m_{i-1} + 10$.

    6. Obtain the valid minimum boundary factor $c_i$ by using \@ref(eq:max-c), with $c(\hat{l}_{i-1})$ and $c(m_i)$.

    7. Obtain the mimimum lengthscale $l_i$ that can be accurately fitted determined by $m_i$ and $c_i$ by using \@ref(eq:m-l-QE).

    8. Fit the HSGP model and

        8.1. Check whether $\, \hat{l}_i \pm 0.02 \geq l_i$ (eq. \@ref(eq:diagnostic)),

        8.2. Check stability in $\hat{l}_i$, *rmse*, $R^2$ and *elpd* relative to previous iteration.

    9. If the verifications in steps 8.1 and 8.2 are TRUE, the HSGP model approximation should be sufficiently accurate, and diagnosis ends here. Otherwise, repeat steps 5-9 and update parameters.

- If the diagnostic in step 4 is FALSE

    10. Set $\, l_i = \hat{l}_{i-1}$.

    11. Repeat steps 1-4 and update parameters.

# Synthetic datasets

Draw various Gaussian process (GP) functions with different lengthscales. For each one of those GP functions, various noisy datasets are generated.

Synthetic datasets $\boldsymbol{y}$ with a true data generating function $f$ which is a GP with a squared exponential kernel. Errors $\boldsymbol{\epsilon}$ with noise $\sigma$ are added to $f$ to form the final synthetic noisy dataset $\boldsymbol{y}$.

\begin{align*}
\begin{split}
\boldsymbol{y} &= \boldsymbol{f} + \boldsymbol{\epsilon} \\
\boldsymbol{\epsilon} &\sim \text{Normal}(\boldsymbol{0}, \sigma^2 \boldsymbol{I}) \\
f(x) &\sim \text{GP}(0, k(x, x', \theta)),
\end{split}
\end{align*}

where $k(x, x', \theta)$ is a squared exponential kernel as a function of input values $x$ and parameters $\theta$ (marginal variance of $f$ and lengthscale of the kernel).

## Input space

Vector of predictive input points `x_pred`, length of `x_pred` (`N_pred`) and vector of indices of `x_pred` (`vv_pred`).

```{r eval=TRUE}
x_pred <- seq(-1,1,0.001)
N_pred <- length(x_pred)  
vv_pred <- 1:length(x_pred)
```

Half range of the input domain.

```{r eval=TRUE}
S <- abs((max(x_pred) - min(x_pred))/2)
S
```

## Draw the true data generative functions $f$ from GPs varying the lengthscale

Initializing programming objects.

```{r eval=TRUE}
standata_SIMU <- list()
fit_SIMU <- list()
```

Marginal variance `gpscale` and lengthscale `lscale`.

```{r eval=TRUE, include=TRUE}
gpscale <- 1
lscale <- c(0.05,0.1,0.15,0.2,0.4,0.6,0.5,0.55,0.8,1.1)
lscale
```

Random variables `eta` to simulate from the different GPs with different lengthscales.

```{r eval=TRUE}
set.seed(SEED)
SEED_l <- sample(1:9999,length(lscale))

eta <- list()
for(l in 1:length(lscale)){
  set.seed(SEED_l[l])
  eta[[l]] <- rnorm(N_pred, 0, 1)
}
```

Data to be passed to Stan.

```{r eval=TRUE}
for(l in 1:length(lscale))
  standata_SIMU[[l]] <- list(x= x_pred, 
                              N= N_pred, 
                              gpscale= gpscale, 
                              lscale= lscale[l], 
                              eta= eta[[l]])
```

Model sampling.

```{r eval=TRUE}
load("fit_SIMU.rData")
```

```{r eval=TRUE}
# simuGP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_SIMU.stan")
# for(l in 10){
#   fit_SIMU[[l]] <- simuGP_mod$sample(data= standata_SIMU[[l]],
#                                   iter_warmup=100,
#                                   iter_sampling=100,
#                                   chains=1, thin=1, init=0.5,
#                                   fixed_param=TRUE)
#   fit_SIMU[[l]] <- read_stan_csv(fit_SIMU[[l]]$output_files())
# }
```

Drawn true data generative functions $f$ from the different GPs with different lengthscales.

```{r eval=TRUE}
f_true <- list()
for(l in 1:length(lscale))
  f_true[[l]] <- summary(fit_SIMU[[l]], pars = "f", probs = c(0.025, 0.5, 0.975))$summary[,1]
names(f_true) <- lscale
# str(f_true)
```

```{r eval=TRUE, fig.height=7, fig.width=15, include=TRUE, echo=FALSE}
f_true_plt <- as.data.frame(f_true)
f_true_plt$x <- x_pred
f_true_plt <- melt(f_true_plt, id.vars = "x")
levels(f_true_plt$variable) <- as.character(lscale)
f_true_plt$variable <- as.numeric(as.character(f_true_plt$variable))
ggplot(f_true_plt, aes(x=x, y=value))  +
  geom_line(color = 'black', size=1) +
  facet_wrap(~ variable, ncol=5) +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5), axis.title=element_text(size=18), axis.text=element_text(size=16), strip.text = element_text(size = 18)) +
  labs(y="y")
```

## Sample points

```{r eval=TRUE}
vv_sample <- list()
x_sample <- list()
N_sample <- list()
n <- c(200,170,140,110,90,100,100,120,90,100)
for(l in 1:length(lscale)){
	set.seed(SEED_l[l])
	vv_sample[[l]] <- sort(sample(vv_pred, n[l]))
	x_sample[[l]] <- x_pred[vv_sample[[l]]]
	N_sample[[l]] <- length(x_sample[[l]])
}
```

## Adding noise to $f$

Five different noisy datasets $\boldsymbol{y}$ over each one of the true data generative functions $f$.

```{r eval=TRUE, fig.height=5, fig.width=15}
set.seed(SEED)
SEED_l_i <- t(matrix(sample(1:9999,length(lscale)*5), 5, length(lscale)))

noise <- rep(0.3,length(lscale))
y <- list()
for(l in 1:length(lscale)){
  y[[l]] <- list()
  for(i in 1:5){
  	set.seed(SEED_l_i[l,i])
  	y[[l]][[i]] <- as.vector(f_true[[l]][vv_sample[[l]]]) + rnorm(N_sample[[l]], 0, noise[l])
  }
}

names(y) <- lscale  
for(l in 1:length(lscale)){
  names(y[[l]]) <-c("dataset1","dataset2","dataset3","dataset4","dataset5")
}
# str(y)
```

Plot of the noisy datasets.

```{r eval=TRUE, fig.height=20, fig.width=15}
y_plt<-list()
for(l in 1:length(lscale)){
  y_plt[[l]] <- as.data.frame(y[[l]])
  y_plt[[l]]$lscale <- lscale[l]
  y_plt[[l]]$x <- x_sample[[l]]
}
y_plt <- rbind(y_plt[[1]],y_plt[[2]],y_plt[[3]],y_plt[[4]],y_plt[[5]],y_plt[[6]],y_plt[[7]],y_plt[[8]],y_plt[[9]],y_plt[[10]])
y_plt <- melt(y_plt, id.vars=c("lscale","x"), measure.vars=c("dataset1","dataset2","dataset3","dataset4","dataset5"), variable.name = "dataset")

ggplot(y_plt[,], aes(x=x, y=value))  +
  geom_point(color = 'black', size=1) +
  facet_grid(rows=vars(lscale), cols=vars(dataset)) +
  # facet_wrap(~ lscale + dataset, ncol=5) +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5), axis.title=element_text(size=20), axis.text=element_text(size=16), strip.text = element_text(size = 18)) +
  labs(y="y")
```

# GP model fit

Initializing programming objects.

```{r eval=TRUE}
standata_GP <- list()
fit_GP <- list()
```

Data to be passed to Stan.

```{r eval=TRUE}
for(l in 1:length(lscale)){
	standata_GP[[l]] <- list()
	for(i in 1:5){
		standata_GP[[l]][[i]] <- list(x= x_sample[[l]], 
		                              y= y[[l]][[i]], 
		                              N= N_sample[[l]], 
		                              param_v= 99)
	}
}
# str(standata_GP)
```

Model fitting.

```{r eval=TRUE}
load("fit_GP.rData")
```

```{r eval=TRUE}
# gp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_GP_marginalized.stan")
# for(l in 10){
# 	# fit_GP[[l]] <- list()
# 	for(i in 1){
#     fit_GP[[l]][[i]] <- gp_mod$sample(data= standata_GP[[l]][[i]],
#                                     iter_warmup=400,
#                                     iter_sampling=400,
#                                     chains=4, thin=5, init=0.5,
#                                     save_warmup = FALSE)
#     fit_GP[[l]][[i]] <- rstan::read_stan_csv(fit_GP[[l]][[i]]$output_files())
# 	}
# }
```

Summaries of variable estimates.

```{r eval=TRUE}
param = c("lscale","gpscale","noise")
for(l in 1){
	for(i in 1){
	  print(paste("dataset: lengthscale =", lscale[l], "dataset =", i), quote = FALSE)
    print(round(summary(fit_GP[[l]][[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary,4))
	}
}
```

Simulation chains for the variables after warmup.

```{r eval=TRUE, fig.height=3, fig.width=15}
for(l in 1){
	for(i in 1){
	  print(paste("dataset: lengthscale =", lscale[l], "dataset =", i), quote = FALSE)
    print(traceplot(fit_GP[[l]][[i]], pars= param, include= TRUE, unconstrain= FALSE, window= NULL, nrow= NULL, ncol= NULL))
	}
}
```

## Posterior GP mean functions

```{r eval=TRUE}
f_GP <- array(NA,c(N_sample[[1]],length(lscale),5,3))
for(l in 1:length(lscale)){
	for(i in 1:5){
		f_GP[1:N_sample[[l]],l,i,1:3] <- summary(fit_GP[[l]][[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
	}
}
dimnames(f_GP)[[1]] <- 1:N_sample[[1]]
dimnames(f_GP)[[2]] <- lscale
dimnames(f_GP)[[3]] <- 1:5
dimnames(f_GP)[[4]] <- c("mean","X2.5","X97.5")
names(dimnames(f_GP)) <- c("id","lscale","dataset","percentil")
# str(f_GP)
```

```{r eval=TRUE, fig.height=25, fig.width=15, include=TRUE, echo=FALSE}
f_GP_plt <- melt(f_GP, na.rm=TRUE)
f_GP_plt$x <- unlist(x_sample)
f_GP_plt <- dcast(f_GP_plt, id + lscale + dataset + x ~ percentil)

y_plt$dataset <- as.numeric(y_plt$dataset)

ggplot(y_plt[,], aes(x=x, y=value))  +
  geom_point(color = 'black', size=1) +
  geom_line(data=f_GP_plt[,], mapping=aes(x=x, y=mean), size=1, color="red") +
  geom_ribbon(mapping=aes(x=x, y=mean, ymin=X2.5, ymax=X97.5), data=f_GP_plt, alpha=0.2, color=NA, fill="red", inherit.aes=FALSE) +
  # facet_wrap(~ lscale, ncol=3) +
  facet_grid(rows=vars(lscale), cols=vars(dataset)) +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5), axis.title=element_text(size=20), axis.text=element_text(size=16), strip.text = element_text(size = 18)) +
  labs(y="y")
```

## GP-lengthscale estimates

```{r eval=TRUE}
lscale_GP <- array(NA,c(length(lscale),5,3))
for(l in 1:length(lscale)){
	for(i in 1:5){
		lscale_GP[l,i,] <- summary(fit_GP[[l]][[i]], pars = c("lscale"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
	}
}

dimnames(lscale_GP)[[1]] <- lscale
dimnames(lscale_GP)[[2]] <- 1:5
dimnames(lscale_GP)[[3]] <- c("mean","X2.5","X97.5")
names(dimnames(lscale_GP)) <- c("lscale","dataset","percentil")
str(lscale_GP)
```

## rmse

```{r eval=TRUE}
mse_GP <- matrix(NA,length(lscale),5)
for(l in 1:length(lscale)){
	for(i in 1:5){
    mse_GP[l,i] <- sqrt(sum((f_true[[l]][vv_sample[[l]]] - f_GP[1:N_sample[[l]],l,i,1])^2)/N_sample[[l]])
	}
}

dimnames(mse_GP)[[1]] <- lscale
dimnames(mse_GP)[[2]] <- 1:5
names(dimnames(mse_GP)) <- c("lscale","dataset")
str(mse_GP)
```

# HSGP model fit

```{r eval=TRUE}
standata_HSGP <- list()
fit_HSGP <- list()
```

Number of basis functions.

```{r eval=TRUE}
M <- c(5,10,15,20,30,40,60,80,100)
M
```

Boundary factor $c$ as a function of $l$ and $m$, by using equation \@ref(eq:max-c)

```{r eval=TRUE, include=TRUE, echo=FALSE}
c <- matrix(NA, length(lscale), length(M))
for(l in 1:length(lscale)){
  for(m in 1:length(M)){
    c[l,m] <- c_vs_l_m_QE(apply(lscale_GP[,,1], 1, max)[l], M[m], S)
  }
}
dimnames(c)[[1]] <- round(apply(lscale_GP[,,1], 1, max),2)
dimnames(c)[[2]] <- M
names(dimnames(c)) <- c("GP-lengthscale","m")
c

# c[4,6:9] <- c(2,2,2.5,2.5)
# c[5,7:9] <- c(3,3.5,4.5)
# c[6,7:9] <- c(3.5,4,5)
# c
```

Boundary

```{r eval=TRUE, include=TRUE}
L <- c * S
L
```

Data to be passed to Stan.

```{r eval=TRUE}
for(m in 1:length(M)){
	standata_HSGP[[m]] <- list()
	for(l in 1:length(lscale)){
		standata_HSGP[[m]][[l]] <- list()
		for(i in 1:5){
			standata_HSGP[[m]][[l]][[i]] <- list(y= y[[l]][[i]], 
			                                   x= x_sample[[l]], 
			                                   N= N_sample[[l]], 
			                                   L= L[l,m], M= M[m])
		}
	}
}
# str(standata_HSGP)
```

Model fitting.

```{r eval=TRUE}
load("fit_HSGP.rData")
```

```{r eval=TRUE}
# hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP.stan")
# for(m in c(3)){
# 	# fit_HSGP[[m]] <- list()
# 	for(l in c(10)){
# 		# fit_HSGP[[m]][[l]] <- list()
# 		for(i in c(5)){
# 			fit_HSGP[[m]][[l]][[i]] <- hsgp_mod$sample(data= standata_HSGP[[m]][[l]][[i]],
#                                                 iter_warmup=2000,
#                                                 iter_sampling=2000,
#                                                 chains=4, thin=10, init=2) #
#       fit_HSGP[[m]][[l]][[i]] <- rstan::read_stan_csv(fit_HSGP[[m]][[l]][[i]]$output_files())
# 		}
# 	}
# }
```

Summaries of variable estimates.

```{r eval=TRUE}
param = c("lscale","gpscale","noise")
for(m in 1){
	for(l in 1){
		for(i in 1){
		  print(paste("m =", M[m], "; dataset: lengthscale =", lscale[l], "dataset =", i), quote = FALSE)
		  print(summary(fit_HSGP[[m]][[l]][[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary)
		}
	}
}
```

Simulation chains for the variables after warmup.

```{r eval=TRUE, fig.height=3, fig.width=15}
for(m in 1){
	for(l in 1){
		for(i in 1){
		  print(paste("m =", M[m], "; dataset: lengthscale =", lscale[l], "dataset =", i), quote = FALSE)
  	  print(traceplot(fit_HSGP[[m]][[l]][[i]], pars = param, include = TRUE, unconstrain = FALSE, window = NULL, nrow = NULL, ncol = NULL))
		}
	}
}
```

## Posterior HSGP mean functions

```{r eval=TRUE}
f_HSGP <-array(NA,c(N_sample[[1]],length(lscale),length(M),5,3))
for(m in 1:length(M)){
	for(l in 1:length(lscale)){
		for(i in 1:5){
			f_HSGP[1:N_sample[[l]],l,m,i,] <- summary(fit_HSGP[[m]][[l]][[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
		}
	}
}
dimnames(f_HSGP)[[1]] <- 1:N_sample[[1]]
dimnames(f_HSGP)[[2]] <- lscale
dimnames(f_HSGP)[[3]] <- M
dimnames(f_HSGP)[[4]] <- 1:5
dimnames(f_HSGP)[[5]] <- c("mean","X2.5","X97.5")
names(dimnames(f_HSGP)) <- c("id","lscale","m","dataset","percentil")
# str(f_HSGP)
```

```{r eval=TRUE, fig.height=30, fig.width=20, include=TRUE, echo=FALSE}
f_HSGP_plt <- melt(f_HSGP, na.rm=TRUE)
f_HSGP_plt$x <- unlist(x_sample)
f_HSGP_plt <- dcast(f_HSGP_plt, id + lscale + m + dataset + x ~ percentil)
# str(f_HSGP_plt)

ggplot(f_HSGP_plt[,], aes(x=x, y=mean, ymin=X2.5, ymax=X97.5, fill=as.factor(m), color=as.factor(m)))  +
  geom_line(size=1.5) +
  # geom_ribbon(alpha=0.2, color=NA) +
  scale_fill_brewer(palette="Set1") +
  scale_color_brewer(palette="Set1") +
  geom_line(mapping=aes(x=x, y=mean), data=f_GP_plt[,], size=1, color='red', linetype=2, inherit.aes=FALSE) +
  # geom_ribbon(mapping=aes(x=x, y=mean, ymin=X2.5, ymax=X97.5), data=f_GP_plt[,], alpha=0.2, color=NA, fill="blue", inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=value), data=y_plt[,], size=0.7, color='black', inherit.aes=FALSE) +
  labs(y="y", x="x", title="HSGP vs GP") +
  theme_classic() +
  theme(legend.position="top", title=element_text(size=22), axis.title.y=element_text(angle=0, vjust=0.5), axis.title=element_text(size=26), axis.text=element_text(size=22), strip.text = element_text(size = 26), legend.title = element_text(size=24), legend.text = element_text(size=22), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  coord_cartesian(ylim = c(-3,3)) +
  facet_grid(rows=vars(lscale), cols=vars(dataset))
```

## HSGP-lengthscale estimates

```{r eval=TRUE}
lscale_HSGP <- array(NA,c(length(M),length(lscale),5,3))
for(m in 1:length(M)){
	for(l in 1:length(lscale)){
		for(i in 1:5){
			lscale_HSGP[m,l,i,] <- summary(fit_HSGP[[m]][[l]][[i]], pars = c("lscale"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
		}
	}
}

dimnames(lscale_HSGP)[[1]] <- M
dimnames(lscale_HSGP)[[2]] <- lscale
dimnames(lscale_HSGP)[[3]] <- 1:5
dimnames(lscale_HSGP)[[4]] <- c("mean","X2.5","X97.5")
names(dimnames(lscale_HSGP)) <- c("m","lscale","dataset","percentil")
str(lscale_HSGP)
```

## HSGP-marginal variance estimates

```{r eval=TRUE}
gpscale_HSGP <- array(NA,c(length(M),length(lscale),5,3))
for(m in 1:length(M)){
	for(l in 1:length(lscale)){
		for(i in 1:5){
			gpscale_HSGP[m,l,i,] <- summary(fit_HSGP[[m]][[l]][[i]], pars = c("gpscale"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
		}
	}
}

dimnames(gpscale_HSGP)[[1]] <- M
dimnames(gpscale_HSGP)[[2]] <- lscale
dimnames(gpscale_HSGP)[[3]] <- 1:5
dimnames(gpscale_HSGP)[[4]] <- c("mean","X2.5","X97.5")
names(dimnames(gpscale_HSGP)) <- c("m","lscale","dataset","percentil")
str(gpscale_HSGP)
```

## rmse

```{r eval=TRUE}
mse_HSGP <- array(NA,c(length(M),length(lscale),5))
for(m in 1:length(M)){
	for(l in 1:length(lscale)){
		for(i in 1:5){
			mse_HSGP[m,l,i] <- sqrt(sum((f_GP[1:N_sample[[l]],l,i,1] - f_HSGP[1:N_sample[[l]],l,m,i,1])^2)/N_sample[[l]])
		}
	}
}

dimnames(mse_HSGP)[[1]] <- M
dimnames(mse_HSGP)[[2]] <- lscale
dimnames(mse_HSGP)[[3]] <- 1:5
names(dimnames(mse_HSGP)) <- c("m","lscale","dataset")
str(mse_HSGP)
```

# HSGP-lengthscale vs GP-lengthscale and HSGP-rmse vs GP-lengthscale

```{r eval=TRUE, fig.height=15, fig.width=15}
lscale_HSGP_melt <- melt(lscale_HSGP)
lscale_GP_melt <- melt(lscale_GP)

lscale_GP_HSGP_2 <- merge(lscale_HSGP_melt[lscale_HSGP_melt$lscale %in% lscale[c(1:10)],], 
                          lscale_GP_melt[lscale_GP_melt$lscale %in% lscale[c(1:10)],], 
                          by=c("lscale","percentil","dataset"), 
                          suffixes = c(".HSGP",".GP"))

p1 <- ggplot(lscale_GP_HSGP_2[lscale_GP_HSGP_2$percentil=="mean" ,], aes(x=value.GP, y=value.HSGP, color=as.factor(m))) +
  geom_abline(linetype=2) +
  geom_vline(aes(xintercept = value.GP), data=lscale_GP_HSGP_2[lscale_GP_HSGP_2$percentil=="mean" ,], linetype=2, color="grey") +
  geom_line(size=1.2) +
  scale_color_manual(values = RColorBrewer::brewer.pal(9, "Set1")) +
  geom_point(data=lscale_GP_HSGP_2[lscale_GP_HSGP_2$percentil=="mean" ,], mapping=aes(x=value.GP, y=value.HSGP), size=3) +
  labs(y="lscale_HSGP", x="lscale_GP") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  facet_wrap(~ dataset, ncol=2, scales = "free_x") + 
  scale_x_continuous(limits=c(0,4)) 
```

<!-- Mean of all datasets -->

```{r eval=TRUE, fig.height=8, fig.width=15}
lscale_HSGP_melt_aggr <- aggregate(value ~ m + lscale + percentil, data= lscale_HSGP_melt[lscale_HSGP_melt$lscale %in% lscale[c(1:10)],], mean)

lscale_GP_melt_aggr <- aggregate(value ~ lscale + percentil, data= lscale_GP_melt[lscale_GP_melt$lscale %in% lscale[c(1:10)],], mean)

lscale_GP_HSGP <- merge(lscale_HSGP_melt_aggr, 
                        lscale_GP_melt_aggr, 
                        by=c("lscale","percentil"), 
                        suffixes = c(".HSGP",".GP"))

p2 <- ggplot(lscale_GP_HSGP[lscale_GP_HSGP$percentil=="mean",], aes(x=value.GP, y=value.HSGP, fill=as.factor(m), color=as.factor(m))) +
  geom_abline(linetype=2) +
  geom_vline(aes(xintercept = value.GP), data=lscale_GP_HSGP[lscale_GP_HSGP$percentil=="mean",], linetype=2, color="grey") +
  geom_line(size=1.2) +
  scale_color_manual(values = RColorBrewer::brewer.pal(9, "Set1")) +
  # scale_colour_gradient2() +
  geom_point(data=lscale_GP_HSGP[lscale_GP_HSGP$percentil=="mean",], mapping=aes(x=value.GP, y=value.HSGP), size=3) +
  labs(y="lscale_HSGP", x="lscale_GP") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm'))
```

<!-- # HSGP-rmse vs GP-lengthscale -->

```{r eval=TRUE, fig.height=8, fig.width=15}
mse_HSGP_melt <- melt(mse_HSGP)
mse_HSGP_melt2 <- merge(mse_HSGP_melt[mse_HSGP_melt$lscale %in% lscale[c(1:10)],], 
                        lscale_GP_melt[lscale_GP_melt$percentil=="mean" & lscale_GP_melt$lscale %in% lscale[c(1:10)], c("lscale","value","dataset")], 
                        by=c("lscale","dataset"))
names(mse_HSGP_melt2) <- c("lscale","dataset","m","value","lscale_GP")

p3 <- ggplot(mse_HSGP_melt2[,], aes(x=lscale_GP, y=value, fill=as.factor(m), color=as.factor(m))) +
  geom_line(size=1.2) +
  geom_vline(aes(xintercept = lscale_GP), linetype=2, color="grey") +
  geom_hline(yintercept = 0, linetype=2) +
  scale_color_manual(values = RColorBrewer::brewer.pal(9, "Set1")) +
  labs(y="RMSE", x="lscale") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  geom_point(size=3) +
  coord_cartesian(ylim = c(0,0.15)) +
  facet_wrap(~ dataset, ncol=2, scales = "free_x") + 
  scale_x_continuous(limits=c(0,4))
```

Mean of all data realizations for each function $f$.

```{r eval=TRUE, fig.height=8, fig.width=15}
mse_HSGP_melt3 <- merge(mse_HSGP_melt[mse_HSGP_melt$lscale %in% lscale[c(1:10)],], 
                        lscale_GP_melt_aggr[lscale_GP_melt_aggr$percentil=="mean" & lscale_GP_melt_aggr$lscale %in% lscale[c(1:10)], c("lscale","value")], 
                        by="lscale")
names(mse_HSGP_melt3) <- c("lscale","m","dataset","value","lscale_GP")

mse_HSGP_plt <- aggregate(value ~ m + lscale_GP, data = mse_HSGP_melt3, mean)
mse_HSGP_plt$X33 <- mse_HSGP_plt$value - aggregate(value ~ m + lscale_GP, data = mse_HSGP_melt3, sd)$value #/sqrt(5)
mse_HSGP_plt$X67 <- mse_HSGP_plt$value + aggregate(value ~ m + lscale_GP, data = mse_HSGP_melt3, sd)$value #/sqrt(5)

p4 <- ggplot(mse_HSGP_plt[,], aes(x=lscale_GP, y=value, ymin=X33, ymax=X67, fill=as.factor(m), color=as.factor(m))) +
  geom_vline(aes(xintercept = lscale_GP), linetype=2, color="grey") +
  geom_ribbon(alpha=0.1, color=NA) +
  geom_hline(yintercept = 0, linetype=2) +
  geom_line(size=1.2) +
  geom_point(size=3) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(9, "Set1")) +
  # scale_fill_gradient2() +
  scale_color_manual(values = RColorBrewer::brewer.pal(9, "Set1")) +
  # scale_colour_gradient2() +
  labs(y="rmse", x="lscale_GP") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  coord_cartesian(ylim = c(0,0.20))
```

```{r eval=TRUE, fig.height=8, fig.width=15, include=TRUE, echo=FALSE}
grid.arrange(p2 + scale_x_continuous(breaks=c(0.1,0.5,1,2,3), trans="sqrt") + scale_y_continuous(breaks=c(0.1,0.5,1,2,3), trans="sqrt"), 
             p4 + scale_x_continuous(breaks=c(0.1,0.5,1,2,3), trans="sqrt"), 
             ncol=2)
```

HSGP-lengthscale estimates tend to be overestimated in both very large lengthscales and high number of basis functions. Which is also slightly reflected in the accuracy of the approximation with slightly bigger errors. This overestimation of the lengthscale estimate can be reversed by increasing the boundary factor.

<!-- ## Increasing $c$ -->

<!-- Boundary factor $c$ as a function of $l$ and $m$, by using equation \@ref(eq:max-c) -->

<!-- ```{r eval=TRUE, include=TRUE, echo=FALSE} -->
<!-- c <- matrix(NA, length(lscale), length(M)) -->
<!-- for(l in 1:length(lscale)){ -->
<!--   for(m in 1:length(M)){ -->
<!--     c[l,m] <- c_vs_l_m_QE(apply(lscale_GP[,,1], 1, max)[l], M[m], S, max(M)) -->
<!--   } -->
<!-- } -->
<!-- dimnames(c)[[1]] <- round(apply(lscale_GP[,,1], 1, max),2) -->
<!-- dimnames(c)[[2]] <- M -->
<!-- names(dimnames(c)) <- c("GP-lengthscale","m") -->
<!-- c -->

<!-- # c[4,6:9] <- c(2,2,2.5,2.5) -->
<!-- # c[5,7:9] <- c(3,3.5,4.5) -->
<!-- # c[5:10,9] <- c(3,4,5,6,9,12) -->
<!-- # c -->
<!-- ``` -->

<!-- Boundary -->

<!-- ```{r eval=TRUE, include=TRUE} -->
<!-- L <- c * S -->
<!-- L -->
<!-- ``` -->

<!-- Data to be passed to Stan. -->

<!-- ```{r eval=TRUE} -->
<!-- for(m in 1:length(M)){ -->
<!-- 	standata_HSGP[[m]] <- list() -->
<!-- 	for(l in 1:length(lscale)){ -->
<!-- 		standata_HSGP[[m]][[l]] <- list() -->
<!-- 		for(i in 1:5){ -->
<!-- 			standata_HSGP[[m]][[l]][[i]] <- list(y= y[[l]][[i]],  -->
<!-- 			                                   x= x_sample[[l]],  -->
<!-- 			                                   N= N_sample[[l]],  -->
<!-- 			                                   L= L[l,m], M= M[m]) -->
<!-- 		} -->
<!-- 	} -->
<!-- } -->
<!-- # str(standata_HSGP) -->
<!-- ``` -->

<!-- Model fitting. -->

<!-- ```{r eval=TRUE} -->
<!-- load("fit_HSGP_modifying_c.rData") -->
<!-- ``` -->

<!-- ```{r eval=TRUE} -->
<!-- # hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP.stan") -->
<!-- # for(m in 9){ -->
<!-- # 	# fit_HSGP[[m]] <- list() -->
<!-- # 	for(l in c(9)){ -->
<!-- # 		# fit_HSGP[[m]][[l]] <- list() -->
<!-- # 		for(i in c(2)){ -->
<!-- # 			fit_HSGP[[m]][[l]][[i]] <- hsgp_mod$sample(data= standata_HSGP[[m]][[l]][[i]], -->
<!-- #                                                 iter_warmup=2000, -->
<!-- #                                                 iter_sampling=2000, -->
<!-- #                                                 chains=4, thin=10, init=2) # -->
<!-- #       fit_HSGP[[m]][[l]][[i]] <- rstan::read_stan_csv(fit_HSGP[[m]][[l]][[i]]$output_files()) -->
<!-- # 		} -->
<!-- # 	} -->
<!-- # } -->
<!-- # save(fit_HSGP, file="fit_HSGP_modifying_c.rData") -->
<!-- ``` -->

<!-- Simulation chains for the variables after warmup. -->

<!-- ```{r eval=TRUE, fig.height=3, fig.width=15} -->
<!-- for(m in 9){ -->
<!-- 	for(l in 10){ -->
<!-- 		for(i in 1:5){ -->
<!-- 		  print(paste("m =", M[m], "; dataset: lengthscale =", lscale[l], "dataset =", i), quote = FALSE) -->
<!--   	  print(traceplot(fit_HSGP[[m]][[l]][[i]], pars = param, include = TRUE, unconstrain = FALSE, window = NULL, nrow = NULL, ncol = NULL)) -->
<!-- 		} -->
<!-- 	} -->
<!-- } -->
<!-- ``` -->

<!-- ### Posterior HSGP mean functions -->

<!-- ```{r eval=TRUE} -->
<!-- f_HSGP <-array(NA,c(N_sample[[1]],length(lscale),length(M),5,3)) -->
<!-- for(m in 1:length(M)){ -->
<!-- 	for(l in 1:length(lscale)){ -->
<!-- 		for(i in 1:5){ -->
<!-- 			f_HSGP[1:N_sample[[l]],l,m,i,] <- summary(fit_HSGP[[m]][[l]][[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)] -->
<!-- 		} -->
<!-- 	} -->
<!-- } -->
<!-- dimnames(f_HSGP)[[1]] <- 1:N_sample[[1]] -->
<!-- dimnames(f_HSGP)[[2]] <- lscale -->
<!-- dimnames(f_HSGP)[[3]] <- M -->
<!-- dimnames(f_HSGP)[[4]] <- 1:5 -->
<!-- dimnames(f_HSGP)[[5]] <- c("mean","X2.5","X97.5") -->
<!-- names(dimnames(f_HSGP)) <- c("id","lscale","m","dataset","percentil") -->
<!-- # str(f_HSGP) -->
<!-- ``` -->

<!-- ### HSGP-lengthscale estimates -->

<!-- ```{r eval=TRUE} -->
<!-- lscale_HSGP <- array(NA,c(length(M),length(lscale),5,3)) -->
<!-- for(m in 1:length(M)){ -->
<!-- 	for(l in 1:length(lscale)){ -->
<!-- 		for(i in 1:5){ -->
<!-- 			lscale_HSGP[m,l,i,] <- summary(fit_HSGP[[m]][[l]][[i]], pars = c("lscale"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)] -->
<!-- 		} -->
<!-- 	} -->
<!-- } -->

<!-- dimnames(lscale_HSGP)[[1]] <- M -->
<!-- dimnames(lscale_HSGP)[[2]] <- lscale -->
<!-- dimnames(lscale_HSGP)[[3]] <- 1:5 -->
<!-- dimnames(lscale_HSGP)[[4]] <- c("mean","X2.5","X97.5") -->
<!-- names(dimnames(lscale_HSGP)) <- c("m","lscale","dataset","percentil") -->
<!-- str(lscale_HSGP) -->
<!-- ``` -->

<!-- ### rmse -->

<!-- ```{r eval=TRUE} -->
<!-- mse_HSGP <- array(NA,c(length(M),length(lscale),5)) -->
<!-- for(m in 1:length(M)){ -->
<!-- 	for(l in 1:length(lscale)){ -->
<!-- 		for(i in 1:5){ -->
<!-- 			mse_HSGP[m,l,i] <- sqrt(sum((f_GP[1:N_sample[[l]],l,i,1] - f_HSGP[1:N_sample[[l]],l,m,i,1])^2)/N_sample[[l]]) -->
<!-- 		} -->
<!-- 	} -->
<!-- } -->

<!-- dimnames(mse_HSGP)[[1]] <- M -->
<!-- dimnames(mse_HSGP)[[2]] <- lscale -->
<!-- dimnames(mse_HSGP)[[3]] <- 1:5 -->
<!-- names(dimnames(mse_HSGP)) <- c("m","lscale","dataset") -->
<!-- str(mse_HSGP) -->
<!-- ``` -->

<!-- ### HSGP-lengthscale vs GP-lengthscale and HSGP-rmse vs GP-lengthscale -->

<!-- <!-- Mean of all datasets --> -->

<!-- ```{r eval=TRUE, fig.height=8, fig.width=15} -->
<!-- lscale_HSGP_melt <- melt(lscale_HSGP) -->
<!-- lscale_GP_melt <- melt(lscale_GP) -->
<!-- lscale_HSGP_melt_aggr <- aggregate(value ~ m + lscale + percentil, data= lscale_HSGP_melt[lscale_HSGP_melt$lscale %in% lscale[c(1:10)],], mean) -->

<!-- lscale_GP_melt_aggr <- aggregate(value ~ lscale + percentil, data= lscale_GP_melt[lscale_GP_melt$lscale %in% lscale[c(1:10)],], mean) -->

<!-- lscale_GP_HSGP <- merge(lscale_HSGP_melt_aggr,  -->
<!--                         lscale_GP_melt_aggr,  -->
<!--                         by=c("lscale","percentil"),  -->
<!--                         suffixes = c(".HSGP",".GP")) -->

<!-- p2 <- ggplot(lscale_GP_HSGP[lscale_GP_HSGP$percentil=="mean",], aes(x=value.GP, y=value.HSGP, color=as.factor(m))) + -->
<!--   geom_abline(linetype=2) + -->
<!--   geom_vline(aes(xintercept = value.GP), data=lscale_GP_HSGP[lscale_GP_HSGP$percentil=="mean",], linetype=2, color="grey") + -->
<!--   geom_line(size=1.2) + -->
<!--   scale_color_manual(values = RColorBrewer::brewer.pal(9, "Set1")) + #[c(1:4,8:11)] -->
<!--   geom_point(data=lscale_GP_HSGP[lscale_GP_HSGP$percentil=="mean",], mapping=aes(x=value.GP, y=value.HSGP), size=3) + -->
<!--   labs(y="lscale_HSGP", x="lscale_GP") + -->
<!--   theme_classic() + -->
<!--   theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) -->
<!-- ``` -->

<!-- <!-- # HSGP-rmse vs GP-lengthscale --> -->

<!-- Mean of all data realizations for each function $f$. -->

<!-- ```{r eval=TRUE, fig.height=8, fig.width=15} -->
<!-- mse_HSGP_melt <- melt(mse_HSGP) -->
<!-- mse_HSGP_melt3 <- merge(mse_HSGP_melt[mse_HSGP_melt$lscale %in% lscale[c(1:10)],],  -->
<!--                         lscale_GP_melt_aggr[lscale_GP_melt_aggr$percentil=="mean" & lscale_GP_melt_aggr$lscale %in% lscale[c(1:10)], c("lscale","value")],  -->
<!--                         by="lscale") -->
<!-- names(mse_HSGP_melt3) <- c("lscale","m","dataset","value","lscale_GP") -->

<!-- mse_HSGP_plt <- aggregate(value ~ m + lscale_GP, data = mse_HSGP_melt3, mean) -->
<!-- mse_HSGP_plt$X33 <- mse_HSGP_plt$value - aggregate(value ~ m + lscale_GP, data = mse_HSGP_melt3, sd)$value #/sqrt(5) -->
<!-- mse_HSGP_plt$X67 <- mse_HSGP_plt$value + aggregate(value ~ m + lscale_GP, data = mse_HSGP_melt3, sd)$value #/sqrt(5) -->

<!-- p4 <- ggplot(mse_HSGP_plt[,], aes(x=lscale_GP, y=value, ymin=X33, ymax=X67, fill=as.factor(m), color=as.factor(m))) + -->
<!--   geom_line(size=1.2) + -->
<!--   geom_vline(aes(xintercept = lscale_GP), linetype=2, color="grey") + -->
<!--   geom_ribbon(alpha=0.1, color=NA) + -->
<!--   geom_hline(yintercept = 0, linetype=2) + -->
<!--   scale_fill_manual(values = RColorBrewer::brewer.pal(9, "Set1")) + #[c(1:4,8:11)] -->
<!--   scale_color_manual(values = RColorBrewer::brewer.pal(9, "Set1")) + #[c(1:4,8:11)] -->
<!--   labs(y="rmse", x="lscale_GP") + -->
<!--   theme_classic() + -->
<!--   theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) + -->
<!--   geom_point(size=3) + -->
<!--   coord_cartesian(ylim = c(0,0.10)) -->
<!-- ``` -->

<!-- ```{r eval=TRUE, fig.height=8, fig.width=15, include=TRUE, echo=FALSE} -->
<!-- grid.arrange(p2, p4, nrow=1) -->
<!-- ``` -->

# Diagnosis

Initializing programming objects.

```{r eval=TRUE}
standata <- list()
for(i in 1:length(lscale)) standata[[i]] <- list()

fit_diagnosis <- list()
for(i in 1:length(lscale)) fit_diagnosis[[i]] <- list()

diagnosis <- list()
for(i in 1:length(lscale)) diagnosis[[i]] <- list()

niter <- 10
ndata <- 5
l <- array(NA,c(length(lscale),niter,ndata))
dimnames(l)[[1]] <- lscale
dimnames(l)[[2]] <- 1:niter
dimnames(l)[[3]] <- 1:ndata

c <- array(NA,c(length(lscale),niter,ndata))
dimnames(c)[[1]] <- lscale
dimnames(c)[[2]] <- 1:niter
dimnames(c)[[3]] <- 1:ndata

m <- array(NA,c(length(lscale),niter,ndata))
dimnames(m)[[1]] <- lscale
dimnames(m)[[2]] <- 1:niter
dimnames(m)[[3]] <- 1:ndata

l_hat <- array(NA,c(length(lscale),niter,ndata))
dimnames(l_hat)[[1]] <- lscale
dimnames(l_hat)[[2]] <- 1:niter
dimnames(l_hat)[[3]] <- 1:ndata

check <- array(NA,c(length(lscale),niter,ndata))
dimnames(check)[[1]] <- lscale
dimnames(check)[[2]] <- 1:niter
dimnames(check)[[3]] <- 1:ndata

rmse <- array(NA,c(length(lscale),niter,ndata))
dimnames(rmse)[[1]] <- lscale
dimnames(rmse)[[2]] <- 1:niter
dimnames(rmse)[[3]] <- 1:ndata

rmse_gp <- array(NA,c(length(lscale),niter,ndata))
dimnames(rmse_gp)[[1]] <- lscale
dimnames(rmse_gp)[[2]] <- 1:niter
dimnames(rmse_gp)[[3]] <- 1:ndata

eR2 <- array(NA,c(length(lscale),niter,ndata))
dimnames(eR2)[[1]] <- lscale
dimnames(eR2)[[2]] <- 1:niter
dimnames(eR2)[[3]] <- 1:ndata

elpd <- array(NA,c(length(lscale),niter,ndata))
dimnames(elpd)[[1]] <- lscale
dimnames(elpd)[[2]] <- 1:niter
dimnames(elpd)[[3]] <- 1:ndata

f <- array(NA,c(N_sample[[1]],length(lscale),niter,3,ndata))
dimnames(f)[[1]] <- 1:N_sample[[1]]
dimnames(f)[[2]] <- lscale
dimnames(f)[[3]] <- 1:niter
dimnames(f)[[4]] <- c("mean","X2.5","X97.5")
dimnames(f)[[5]] <- 1:ndata
```

## Select data realization

```{r eval=TRUE, include=TRUE, results='hold'}
i_data <- 1
print(paste("Data realization =", i_data), quote = FALSE)
```

## Select function $f$

```{r eval=TRUE, include=TRUE, results='hold'}
r <- 1
print(paste("True data generative function-lengthcale =", round(lscale[r],2), ", GP-Lengthcale =", round(lscale_GP[r,i_data,1],2)), quote = FALSE)
```

## Iteration index

```{r eval=TRUE, include=TRUE, results='hold'}
i <- 1
print(paste("Iteration =", i), quote = FALSE)
```

## Setting $m$, $l$ and $c$

**If it is the first iteration, $i=1$**

```{r eval=TRUE, include=TRUE}
if(i==1){
  print("It is the first iteration", quote = FALSE)
}else{
  print("It is NOT the first iteration", quote = FALSE)
}
```

1. Making the first guess of $m_1$ being equal to 5

```{r eval=TRUE, include=TRUE}
if(i==1){
  m[r,i,i_data] <- 5
  print(paste("m[r,i] =", m[r,i,i_data]), quote = FALSE)
}else{
  print("It is NOT the first iteration", quote = FALSE)
}
```

2. The boundary factor $c_1$ has to fulfill equation \@ref(eq:c-vs-m-QE)

```{r eval=TRUE, include=TRUE}
if(i==1){
  c[r,i,i_data] <- c_vs_m_QE(m=m[r,i,i_data])
  print(paste("c[r,i] =", c[r,i,i_data]), quote = FALSE)
}else{
  print("It is NOT the first iteration", quote = FALSE)
}
```

**If it is the second iteration or further, $i>1$**

```{r eval=TRUE, include=TRUE}
if(i==1){
  print("It is NOT the second iteration or further", quote = FALSE)
}else{
  print("It is the second iteration or further", quote = FALSE)
}
```

1. If the diagnostic $\hat{l}_{i-1} \geq l_{i-1}$ is TRUE:

    2.1. Update $m_i = m_{i-1} + 10$

    2.2. The boundary factor $c_i$ has to fulfill \@ref(eq:max-c) where $c(l)$ is evaluated as a function of $\hat{l}_{i-1}$ and $c(m)$ is obtained as a function of $m_{i}$.

2. If the diagnostic $\hat{l}_{i-1} \geq l_{i-1}$ is FALSE:

    2.1. Update $l_i = \hat{l}_{i-1}$

    2.2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-vs-l-QE)

    2.3. The number of basis functions $m_i$ as a function of $l_i$ and $c_i$ by using \@ref(eq:m-l-QE)

    2.4. Update the boundary factor $c_i$ obtained in step 2.3 by equation \@ref(eq:max-c) where $c(l)$ is evaluated as a function of $l_i$ and $c(m)$ is obtained as a function of $m_i$.

```{r eval=TRUE, include=TRUE, results='hold'}
if(i!=1){
  if(diagnostic(l[r,i-1,i_data], l_hat[r,i-1,i_data])){
    print("Diagnostic is TRUE", quote = FALSE)
    m[r,i,i_data] <- m[r,i-1,i_data] + 10
    c[r,i,i_data] <- c_vs_l_m_QE(l=l_hat[r,i-1,i_data], m=m[r,i,i_data], S=S)
    print(paste("m[r,i] =", m[r,i,i_data]), quote = FALSE) 
    print(paste("c[r,i] =", c[r,i,i_data]), quote = FALSE)
  }else{
    print("Diagnostic is FALSE", quote = FALSE)
    l[r,i,i_data] <- l_hat[r,i-1,i_data]
    c[r,i,i_data] <- c_vs_l_QE(l=l[r,i,i_data], S=S)
    m[r,i,i_data] <- m_QE(c=c[r,i,i_data], l=l[r,i,i_data], S=S)
    c[r,i,i_data] <- c_vs_l_m_QE(l=l[r,i,i_data], m=m[r,i,i_data], S=S)
    print(paste("l[r,i] =", l[r,i,i_data]), quote = FALSE)
    print(paste("c[r,i] =", c[r,i,i_data]), quote = FALSE)
    print(paste("m[r,i] =", m[r,i,i_data]), quote = FALSE)
    print(paste("c[r,i] =", c[r,i,i_data]), quote = FALSE)
  }
}else{
  print("It is NOT the second iteration or further", quote = FALSE)
}
```

## Data to be passed to Stan

```{r eval=TRUE, include=TRUE}
standata[[r]][[i]] <- list(y= y[[r]][[i_data]],
                      x= x_sample[[r]],
                      N= N_sample[[r]],
                      L= c[r,i,i_data]*S,
                      M= m[r,i,i_data])
str(standata[[r]][[i]])
```

## Model fitting

Model fitting.

```{r eval=TRUE, include=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP.stan")
fit_diagnosis[[r]][[i]] <- hsgp_mod$sample(data= standata[[r]][[i]], 
                                iter_warmup=1000, 
                                iter_sampling=1000, 
                                chains=4, thin=10, init=2)
fit_diagnosis[[r]][[i]] <- rstan::read_stan_csv(fit_diagnosis[[r]][[i]]$output_files())
```

Summaries of variable estimates and simulation chains for the variables after warmup.

```{r fig.height=3, fig.width=15, eval=TRUE, include=TRUE, results='hold', collapse=TRUE}
param = c("lscale","gpscale","noise")
print(paste("m =", m[r,i,i_data], "; True data generative function-lengthcale =", round(lscale[r],2), "(r =", r, ")"), quote = FALSE)
summary(fit_diagnosis[[r]][[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit_diagnosis[[r]][[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior mean functions.

```{r fig.height=4, fig.width=10, eval=TRUE, include=TRUE, echo=FALSE}
f[1:N_sample[[r]],r,i,1:3,i_data] <- summary(fit_diagnosis[[r]][[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
dimnames(f)[[1]] <- 1:N_sample[[1]]
dimnames(f)[[2]] <- lscale
dimnames(f)[[3]] <- 1:niter
dimnames(f)[[4]] <- c("mean","X2.5","X97.5")

f_GP_diag <- data.frame(as.data.frame(f_GP[1:N_sample[[r]],r,i_data,1:3]), x=x_sample[[r]], type="GP")
f_HSGP <- data.frame(as.data.frame(f[1:N_sample[[r]],r,i,1:3,i_data]), x=x_sample[[r]], type="HSGP")
f_HSGP_GP <- rbind(f_GP_diag, f_HSGP)

ggplot(f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5, ymax=X97.5, fill=type, color=type))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[[r]][[i_data]], x=x_sample[[r]]), size=1, color='black', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

## Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination), and expected log predictive density `elpd`.

```{r eval=TRUE, include=TRUE}
# rmse
residual <- y[[r]][[i_data]] - f[1:N_sample[[r]],r,i,1,i_data]
rmse[r,i,i_data] <- round(sqrt(mean(residual^2)), 3)

residual_gp <- f_GP[1:N_sample[[r]],r,i_data,1] - f[1:N_sample[[r]],r,i,1,i_data]
rmse_gp[r,i,i_data] <- round(sqrt(mean(residual_gp^2)), 3)

# Coefficient of determination R^2
noise <- as.matrix(fit_diagnosis[[r]][[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit_diagnosis[[r]][[i]], pars = c("f")), 1, sd)
eR2[r,i,i_data] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 3)

# elpd
elpd[r,i,i_data] <- round(median(summary(fit_diagnosis[[r]][[i]], pars = c("lpd"))$summary[,1]), 3)
```

## Diagnosis

1. If it is the first iteration (i.e. $i=1$) or the diagnostic $\, \hat{l}_{i-1} \pm 0.02 \geq l_{i-1}$ is true, the minimum lengthscale $l_{i}$ determined by $m_{i}$ and $c_{i}$ used in the HSGP approximation has to be computed by using \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE}
if(i==1){
  l[r,i,i_data] <- l_QE(c=c[r,i,i_data], m=m[r,i,i_data], S=S)
}else{
  if(diagnostic(l[r,i-1,i_data], l_hat[r,i-1,i_data]))
    l[r,i,i_data] <- l_QE(c=c[r,i,i_data], m=m[r,i,i_data], S=S)
}
print(paste("l[r,i] =", l[r,i,i_data]), quote = FALSE)
```

2. Lengthscale estimate $\hat{l_i}$ in the HSGP approximation

```{r eval=TRUE, include=TRUE}
l_hat[r,i,i_data] <- round(summary(fit_diagnosis[[r]][[i]], pars = "lscale")$summary[,1], 3)
print(paste("l_hat[r,i] =", l_hat[r,i,i_data]), quote = FALSE)
```

3. Check the diagnostic of whether $\, \hat{l_i} \pm 0.02 \geq l_i$

```{r eval=TRUE, include=TRUE}
check[r,i,i_data] <- diagnostic(l[r,i,i_data], l_hat[r,i,i_data])

diagnosis[[r]][[i]] <- data.frame(iter= i,
                              l= l[r,i,i_data],
                              c= c[r,i,i_data],
                              m= m[r,i,i_data],
                              l_hat= l_hat[r,i,i_data],
                              'l_hat > l' = check[r,i,i_data],
                              rsme = rmse[r,i,i_data],
                              rsme_gp = rmse_gp[r,i,i_data],
                              R2 = eR2[r,i,i_data],
                              elpd = elpd[r,i,i_data])
names(diagnosis[[r]][[i]]) <- c("iter", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "rmse_gp", "R2", "elpd")

if(i==1){ 
  diagnosis[[r]][[i]]
}else{
  diagnosis[[r]][[i]] <- rbind(diagnosis[[r]][[i-1]],diagnosis[[r]][[i]])
  diagnosis[[r]][[i]]
}
```

```{r eval=TRUE}
# save(l, file = "l.rData")
# save(c, file = "c.rData")
# save(m, file = "m.rData")
# save(l_hat, file = "l_hat.rData")
# save(check, file = "check.rData")
# save(rmse, file = "rmse.rData")
# save(rmse_gp, file = "rmse_gp.rData")
# save(eR2, file = "eR2.rData")
# save(elpd, file = "elpd.rData")
# save(f, file = "f.rData")
```

# Diagnosis for the different functions $f$

```{r eval=TRUE, include=TRUE}
load("l.rData")
load("c.rData")
load("m.rData")
load("l_hat.rData")
load("check.rData")
load("rmse.rData")
load("rmse_gp.rData")
load("eR2.rData")
load("elpd.rData")
load("f.rData")
```

```{r eval=TRUE, include=TRUE, echo=FALSE}
diagnosis_all <- merge(melt(l, value.name="l", na.rm=TRUE), melt(c, value.name="c", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(m, value.name="m", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(l_hat, value.name="l_hat", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(check, value.name="check", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(rmse, value.name="rmse", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(rmse_gp, value.name="rmse_gp", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(eR2, value.name="eR2", na.rm=TRUE), by=c("Var1","Var2","Var3"))
diagnosis_all <- merge(diagnosis_all, melt(elpd, value.name="elpd", na.rm=TRUE), by=c("Var1","Var2","Var3"))
names(diagnosis_all) <- c("f","iter", "i_data", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "rmse_gp", "R2", "elpd")
diagnosis_all <- merge(diagnosis_all, round(lscale_GP_melt[lscale_GP_melt$percentil=="mean",c("value","lscale","dataset")],3), by.x=c("f","i_data"), by.y=c("lscale","dataset"))
names(diagnosis_all) <- c("f","i_data","iter", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "rmse_gp", "R2", "elpd","lscale_GP")
diagnosis_all <- diagnosis_all[,c("f","i_data","lscale_GP","iter", "l", "c", "m", "l_hat", "l_hat > l", "rmse", "rmse_gp", "R2", "elpd")]
diagnosis_all <- diagnosis_all[with(diagnosis_all, order(f,i_data, iter)), ]
diagnosis_all[diagnosis_all$i_data==1,]
```

```{r eval=TRUE, fig.height=8, fig.width=15, include=TRUE, echo=FALSE}
# diagnosis_all_plt <- aggregate( . ~ f + iter, data= diagnosis_all, mean)
diagnosis_all_plt <- diagnosis_all[diagnosis_all$i_data==1,]

p10 <- ggplot(diagnosis_all_plt, aes(x=lscale_GP, y=l_hat, color=m)) +
  geom_vline(aes(xintercept = lscale_GP), linetype=2, color="grey") +
  geom_abline(linetype=2) +
  scale_colour_gradient2() +
  # scale_colour_continuous(type = "viridis", direction = -1) +
  geom_point(size=4) +
  labs(y="lscale_HSGP", x="lscale_GP") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm'))

p11 <- ggplot(diagnosis_all_plt, aes(x=lscale_GP, y=rmse_gp, color=m)) +
  geom_hline(yintercept = 0, linetype=2) +
  geom_vline(aes(xintercept = lscale_GP), linetype=2, color="grey") +
  scale_colour_gradient2() +
  # scale_colour_continuous(type = "viridis", direction = -1) +
  geom_point(size=4) +
  labs(y="rmse", x="lscale_GP") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=90, vjust=0.5, size=18), legend.position="top", axis.title = element_text(size=18), axis.text = element_text(size=16), legend.title = element_text(size=18), legend.text = element_text(size=16), legend.key.size = unit(1, 'cm'), legend.key.height = unit(1, 'cm'), legend.key.width = unit(1, 'cm')) +
  coord_cartesian(ylim = c(0,0.2))

grid.arrange(p10 + scale_x_continuous(breaks=c(0.1,0.5,1,2,3),trans="sqrt") + scale_y_continuous(breaks=c(0.1,0.5,1,2,3),trans="sqrt"),
             p11 + scale_x_continuous(breaks=c(0.1,0.5,1,2,3),trans="sqrt"),
             ncol=2)
```

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex

--- 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 
+As explained in Section~\ref{sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
+%
+\begin{equation*}
+k(\tau) = \sum_{j=1}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
+\end{equation*}
+%
+where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
+The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
+The number of basis functions can be truncated at some finite positive value $m$ such that the total variation difference between the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
+%
+\begin{equation}\label{eq_diff_covs}
+ \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \,\mathrm{d}\tau  < \varepsilon.
+\end{equation}
+
+This inequality can be satisfied for arbritrary small $\epsilon$ provided that $L$ and $m$ are sufficiently large \cite[Theorem 1 and 4]{solin2018hilbert}. 
+The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its length-scale $\ell$, which constitutes a hyperparameter of the GP.
+The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
+As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
+
+In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ will include the smallest interval that contains all the inputs points $x_i$.
+Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.
+Let $S = \max_{i}|x_i|$, then it follows that $x_i \in \left[-S, S\right]$ for all $i$.
+We now define $L$ as
+\begin{equation}\label{eq_boundary}
+L=c \cdot S,
+\end{equation} 
+%
+where $S > 0$ represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
+In the following, we will refer to $c$ as the boundary factor of the approximation.
+The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
+
 \begin{figure*} %[!t]
@@
 \end{figure*}
-
-As explained in Section~\ref{sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
-%
-\begin{equation*}
-k(\tau) = \sum_{j=1}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
-\end{equation*}
-%
-where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$.
-The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case.
-The number of basis functions can be truncated at some finite positive value $m$ such that the total variation difference between the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
-%
-\begin{equation}\label{eq_diff_covs}
- \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \,\mathrm{d}\tau  < \varepsilon.
-\end{equation}
-
-This inequality can be satisfied for arbritrary small $\epsilon$ provided that $L$ and $m$ are sufficiently large \cite[Theorem 1 and 4]{solin2018hilbert}. 
-The specific number of basis functions $m$ needed depends on the degree of non-linearity of the function to be estimated, that is on its length-scale $\ell$, which constitutes a hyperparameter of the GP.
-The approximation also depends on the boundary $L$ (see equations (\ref{eq_eigenvalue}), (\ref{eq_eigenfunction}), (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi})), which will affect its accuracy especially near the boundaries.
-As we will see later on, $L$ will also influence the number of basis functions required in the approximation.
-
-In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ will include the smallest interval that contains all the inputs points $x_i$.
-Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.
-Let $S = \max_{i}|x_i|$, then it follows that $x_i \in \left[-S, S\right]$ for all $i$.
-We now define $L$ as
-\begin{equation}\label{eq_boundary}
-L=c \cdot S,
-\end{equation} 
-%
-where $S > 0$ represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
-In the following, we will refer to $c$ as the boundary factor of the approximation.
-The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
 
@@
 
-\subsection{Theoretical evidence of linear proportionality between $m$, $c$ and $l$} \label{subsec_theoretic_linearity}
-
-There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
+\subsection{Near linear proportionality between $m$, $c$, and $l$} \label{subsec_theoretic_linearity}
+
+A priori the terms in the series expansion \eqref{eq_approxf_multi} with very small spectral density are unlikely to contribute to the approximation.
+Given the boundary factor $c$ and the length-scale $\ell$, we can compute cumulative sum of the spectral densities and find out how many basis functions are a priori explaining almost 100\% of the variation.
+Thus, given $c$ and $\ell$, we can estimate a good choice for the number of basis functions $m$ for any covariance function.
+
+When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
+With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$ and with increasing $c$, the $m$ should grow near linearly with $c$.
+%
+This is natural as with decreasing $\ell$, more higher frequency basis functions are needed.
+With increasing $c$, the smaller range of the basis functions are used in the approximation, the expected number of zero up-crossings goes down linearly with $c$, and thus more higher frequency basis functions are needed to compensate this.
+%
+When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$ if $\ell$ or $c$ is changed.
+
+% There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
 In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
-In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
-
-{\color{blue}Aki: We can also look at the mean zero-level upcrossings, which for QE is
-proportional to 1/l which gives as that part, and the effect of c should also
-be quite linear when c is not very close to 1 and m is not very small as it
-linearly makes the interval shorter so that the there are less zero-level
-upcrossings available in the basis functions. You can visualize this using
-{\small \url{https://avehtari.github.io/casestudies/Motorcycle/motorcycle_gpcourse.html}} and
-look how the plot of 6 basis functions change when you set c=1, c=1.5, c=2.25
-(ie $1.5^2$). Instead of looking at mean zero-level crossings, we can look at the
-spectral densities and when they go to zero. This is illustrated in the same
-notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
-explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
-(linear in l) and to explain 80\% of the prior variance we need 2, 4, 8, 16
-basis functions. These spectral densities don't depend on c. When c increaes
-the basis functions are truncated and for example going from c=1.5 to c=2.25,
-basis function 9 has the flexibility basis function 6 had, so 1.5 more basis
-functions are needed. So we have proportionality c/l. }
-
-{\color{red} Gabi: The spectral densities do depend on $c$, see equations (\ref{eq_approxf}) and (\ref{eq_eigenvalue}).}
+% In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
+
+% {\color{blue}Aki: We can also look at the mean zero-level upcrossings, which for QE is
+% proportional to 1/l which gives as that part, and the effect of c should also
+% be quite linear when c is not very close to 1 and m is not very small as it
+% linearly makes the interval shorter so that the there are less zero-level
+% upcrossings available in the basis functions. You can visualize this using
+% {\small \url{https://avehtari.github.io/casestudies/Motorcycle/motorcycle_gpcourse.html}} and
+% look how the plot of 6 basis functions change when you set c=1, c=1.5, c=2.25
+% (ie $1.5^2$). Instead of looking at mean zero-level crossings, we can look at the
+% spectral densities and when they go to zero. This is illustrated in the same
+% notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
+% explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
+% (linear in l) and to explain 80\% of the prior variance we need 2, 4, 8, 16
+% basis functions. These spectral densities don't depend on c. When c increaes
+% the basis functions are truncated and for example going from c=1.5 to c=2.25,
+% basis function 9 has the flexibility basis function 6 had, so 1.5 more basis
+% functions are needed. So we have proportionality c/l. }
+
+% {\color{red} Gabi: The spectral densities do depend on $c$, see equations (\ref{eq_approxf}) and (\ref{eq_eigenvalue}).}
 
@@
 \end{equation}
+%
 where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions.
@@
 
-From the empirical observations, numerical equations governing the relationships between $m$, $c$ and $\ell$ were estimated, which show an agreement with the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$.
-The estimated numerical equations for each particular covariance function are the following:
+When $c$ is large enough, there is near
+linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ (see Section~\ref{subsec_theoretic_linearity}).
+To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the needed constant terms.
+We establish a lower bound for $c$ set to 1.2.
+
+% From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
+Following we show the numerical functions governing these relationships for every kernel.
 
@@
 \begin{align}\label{eq_m_l_QE}
-&m \geq 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 1.75 \, \frac{c}{m},
+&m= 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 1.75 \, \frac{c}{m},
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_QE}
-&c \geq 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2      
+&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2      
 \end{align}
 
+\paragraph*{Mat\'ern ($\nu$=5/2)}:
+%
+\begin{align}\label{eq_m_l_mat52}
+&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
+\end{align}
+%
+with
+%
+\begin{align}\label{eq_c_vs_l_mat52}
+&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
+\end{align}
+
 \paragraph*{Mat\'ern ($\nu$=3/2)}:
@@
 \begin{align}\label{eq_m_l_mat32}
-&m \geq 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 3.42 \, \frac{c}{m},
+&m= 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 3.42 \, \frac{c}{m},
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat32}
-&c \geq 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2        
+&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2        
 \end{align}
 
-\paragraph*{Mat\'ern ($\nu$=5/2)}:
-%
-\begin{align}\label{eq_m_l_mat52}
-&m \geq 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 2.65 \, \frac{c}{m},
-\end{align}
-%
-with
-%
-\begin{align}\label{eq_c_vs_l_mat52}
-&c \geq 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2       
-\end{align}
-
-Notice that we establish a lower bound at 1.2 for $c$ in every case.
+These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$).
+Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions.
+If the aim is to find minimal $m$ to speedup the computation, a further refined formula can be obtained for new covariance functions.
 
@@
 If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated.
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from eq.
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also match those obtained from eq.
 (\ref{eq_m_l_QE}).
@@
 
+If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the first fitting, it is likely that the result is sufficiently accurate.
+The following iterative approach works as safety approach and allows also finding a smaller sufficient $m$ which can be useful if the user wants to optimized the computational speed for repeated fitting (e.g.
+in cross-validation or simulation based calibration).
+
 Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
@@
 
-\item[Step A1.] Make an initial guess on the true length-scale $\ell_1$ of the function to be learned.
-If there is no approximate prior information, we recommend a medium-large length-scale, about $\ell_1 \approx [0.5-1]$.
+\item[Step A1.] Make an initial guess on the lengthscale $l_1$.
+If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \approx [0.5-1]$.
 Set the iteration index to $i=1$.
 
-\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}
+\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.
 

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Gabriel's edits

@@
 
-Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, equations in previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to spend minimum computations.
-
-Guidelines for updating $m$ and/or $c$ must be provided:
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, equations in previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to minimizing the computational cost.
+
+Guidelines for updating $m$ and/or $c$ are:
 

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex

@@
 The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
-The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_7_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52} are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
+The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
 \eqref{eq:rel_total_distance}.

### 2022-02-17 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edits

@@
 
-Ultimately, these recommendations are based on the relationships among the number of basis functions, the boundary condition and the length-scale of the function, which depend on the particular choice of the kernel function.
-In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu$=3/2) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
+Ultimately, these recommendations are based on the relationships among the number of basis functions $m$, the boundary condition $c$, and the length-scale $\ell$, which depend on the particular choice of the kernel function.
+In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
 For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
@@
 Figures~\ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
-For $c$ fixed to a sufficiently large value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation for both the posterior mean or uncertainty.
+For a sufficiently large fixed value of $c$, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation for both the posterior mean or uncertainty.
 It is seen that if the number of basis functions $m$ is too small, the estimated function tend to be overly smooth because the necessary high frequency components are missing.
@@
 Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
-The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} for both the exact GP and the HSGP models.
+The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models.
 Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
@@
 
-A priori the terms in the series expansion \eqref{eq_approxf_multi} with very small spectral density are unlikely to contribute to the approximation.
-Given the boundary factor $c$ and the length-scale $\ell$, we can compute cumulative sum of the spectral densities and find out how many basis functions are a priori explaining almost 100\% of the variation.
+A priori, the terms in the series expansion \eqref{eq_approxf_multi} with very small spectral density are unlikely to contribute to the approximation.
+Given the boundary factor $c$ and the length-scale $\ell$, we can compute cumulative sum of the spectral densities and find out how many basis functions are, a priori, explaining almost 100\% of the variation.
 Thus, given $c$ and $\ell$, we can estimate a good choice for the number of basis functions $m$ for any covariance function.
 
-When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation (see TODO ADD REF HERE), the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
-With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$ and with increasing $c$, the $m$ should grow near linearly with $c$.
+When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation, the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
+With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$, and with increasing $c$, the $m$ should grow near linearly with $c$.
 %
 This is natural as with decreasing $\ell$, more higher frequency basis functions are needed.
-With increasing $c$, the smaller range of the basis functions are used in the approximation, the expected number of zero up-crossings goes down linearly with $c$, and thus more higher frequency basis functions are needed to compensate this.
-%
-When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$ if $\ell$ or $c$ is changed.
+With increasing $c$, as a smaller range of the basis functions are used in the approximation, the expected number of zero up-crossings goes down linearly with $c$, and thus more higher frequency basis functions are needed to compensate this.
+%
+When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ is changed.
 
@@
 linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ (see Section~\ref{subsec_theoretic_linearity}).
-To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the needed constant terms.
+To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the practically useful constant terms.
 We establish a lower bound for $c$ set to 1.2.

### 2022-02-17 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edits primarily to section 4

@@
 
-Ultimately, these recommendations are based on the relationships among the number of basis functions $m$, the boundary condition $c$, and the length-scale $\ell$, which depend on the particular choice of the kernel function.
+Ultimately, these recommendations are based on the relationships among the number of basis functions $m$, the boundary condition $L$, and the length-scale $\ell$, which depend on the particular choice of the kernel function.
 In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
@@
 In general, the higher the degree of wigglyness of the function to be estimated, the larger number of  basis functions will be required.
-If $m$ is fixed to a sufficiently large value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation mainly near the boundaries in mean, but along the whole domain in the standard deviation.
-The approximation error for the variance tends to be bigger for the variance than for the mean.
+If $m$ is fixed to a sufficiently large value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation of the mean mainly near the boundaries, while the approximation of the standard deviation is affected across the whole domain.
+What is more, the approximation error tends to be bigger for the standard deviation than for the mean.
 
@@
 
-Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation.
+Next, we analyze how the interaction between $m$ and $c$ affects the quality of the approximation.
 The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models.
@@
 A priori, the terms in the series expansion \eqref{eq_approxf_multi} with very small spectral density are unlikely to contribute to the approximation.
-Given the boundary factor $c$ and the length-scale $\ell$, we can compute cumulative sum of the spectral densities and find out how many basis functions are, a priori, explaining almost 100\% of the variation.
+Given the boundary factor $c$ and the length-scale $\ell$, we can compute the cumulative sum of the spectral densities and find out how many basis functions are, a priori, explaining almost 100\% of the variation.
 Thus, given $c$ and $\ell$, we can estimate a good choice for the number of basis functions $m$ for any covariance function.
 
-When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation, the $m$ first terms needed to explain almost 100\% of the variation has near linear relationship with $\ell$ and $c$.
+When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation, the number of $m$ first terms needed to explain almost 100\% of the variation has a near linear relationship with $\ell$ and $c$.
 With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$, and with increasing $c$, the $m$ should grow near linearly with $c$.
@@
 %
-When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ is changed.
+When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ are changed.
 
@@
 
+TODO: Is $J$ equal to $m$ or is this something else?
+
 In Appendix~\ref{sec_periodic}, we analyze the accuracy of the linear representation of a periodic squared exponential covariance function and derive the minimum number of terms $J$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
@@
 
-When $c$ is large enough, there is near
-linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ (see Section~\ref{subsec_theoretic_linearity}).
+As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near
+linear proportionality between $m$, $\frac{l}{S}$ (lengthscale $l$ relative to the half range of the input domain $S$) and $c$.
 To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the practically useful constant terms.
-We establish a lower bound for $c$ set to 1.2.
+We require a lower bound for $c$ of $c \geq 1.2$ such that the equations below are precise enough for practical application.
 
@@
 
-As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) in previous Section~\ref{sec_num_equations} provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) provide the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
-If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated.
+If, however, $\hat{\ell}$ does not exceed the recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated.
 If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also match those obtained from eq.
@@
 
-Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, equations in previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to minimizing the computational cost.
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to minimizing the computational cost.
 
@@
 \begin{itemize}
-	\item $c$ must be big enough for the length-scale $l$ of the function to be learned and,
-	\item $m$ must be big enough as a function of $\ell$ and $c$. 
+	\item $c$ must be big enough for the given length-scale $l$ of the function to be learned and,
+	\item $m$ must be big enough for the given $\ell$ and $c$. 
 \end{itemize}
 
-We recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using equations per Section~\ref{sec_num_equations} , and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$ provided by equations in Section~\ref{sec_num_equations}, update $\ell$ with $\hat{\ell}$ and repeat the process.
-
-A detailed guideline with the steps to perform diagnosis, ensuring that a good approximation is achieved with the minimum computational requirements (i.e.
+We recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using the equations in Section~\ref{sec_num_equations}; and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$, update $\ell$ with $\hat{\ell}$ and repeat the process.
+
+If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
+The following iterative procedure works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed for repeated fitting (e.g., in cross-validation or simulation based calibration).
+
+%A detailed guideline with the steps to perform diagnostic, ensuring that a good approximation is achieved with the minimum computational requirements (i.e.
 minimum number of basis functions $m$) is provided in  next Section~\ref{sec_user_guide}.
@@
 
-If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the first fitting, it is likely that the result is sufficiently accurate.
-The following iterative approach works as safety approach and allows also finding a smaller sufficient $m$ which can be useful if the user wants to optimized the computational speed for repeated fitting (e.g.
-in cross-validation or simulation based calibration).
-
-Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
+Based on the above proposed diagnostics, we obtain a simple, iterative step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
 The procedure is split into two phases, Phase A and B, which have to be completed consecutively.
@@
 
-\item[Step A1.] Make an initial guess on the lengthscale $l_1$.
-If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \approx [0.5-1]$.
+\item[A1.] Make an initial guess on the lengthscale $l_1$.
+If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \in [0.5, 1]$.
 Set the iteration index to $i=1$.
 
-\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[Step A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
-
-\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i + 0.01 \geq \ell_i$.
-
-If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell_{i+1} = \hat{\ell}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.
+\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
+
+\item[A5.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
+
+If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell_{i+1} = \hat{\ell}_{i}$, increase the iteration index $i = i + 1$, and go back to A2.
 
@@
 
-\item[Step B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD).
-
-\item[Step B2.] Set $m_{i+1} = m_{i} + 5$ and increase the iteration index $i = i + 1$.
-
-\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \hat{\ell}_{i-1}$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[Step B4.] Obtain the mimimum valid lengthscale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[Step B5.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
-
-\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i + 0.01 \geq \ell_i$.
+\item[B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD).
+
+\item[B2.] Set $m_{i+1} = m_{i} + 5$ and increase the iteration index $i = i + 1$.
+
+\item[B3.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \hat{\ell}_{i-1}$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[B4.] Obtain the mimimum valid lengthscale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[B5.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
+
+\item[B6.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
 
@@
 
-Otherwise, go back to Step B1.
+Otherwise, go back to B1.
 \end{enumerate}
@@
 
-First of all in this section we illustrate that accurate estimates of the length-scale implies accurate approximations with a small RMSEs.
-Thus, Figure~\ref{fig7_lscale_comparison}-left shows a comparison of the length-scale estimates
+In this section, we first illustrate that accurate estimates of the length-scale implies accurate approximations via HSGPs.
+Figure~\ref{fig7_lscale_comparison} left shows a comparison of the length-scale estimates
 obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness.
@@
 (\ref{eq_c_vs_l_QE}).
-Figure \ref{fig7_lscale_comparison}-right shows the RMSE of the HSGP models with the exact GP model as the reference.
-It can be seen that accurate estimates of the length-scale implies small RMSEs.
-
-Furthermore, from Figure~\ref{fig7_lscale_comparison}-left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e.
-true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison}-left).
+Figure \ref{fig7_lscale_comparison} right shows the RMSE of the HSGP models with the exact GP model as the reference.
+It can be seen that accurate estimates of the length-scale imply small RMSEs.
+
+Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison} left).
 Here it comes out a possible weaknesses of the approach which consists in using too many basis functions relative to the true length-scale causes overestimating the posteriors.
+
+TODO: Please clarify the last sentence above. I don't get that is ment here.
 
@@
 
-Secondly in this section, Figure~\ref{fig9_diagnosis} shows the iterative steps of applying the diagnosis procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison}-left, showing that by following our recommendations, an optimum solution with minimum computational requirements is achieved.
-Figure~\ref{fig10_diagnosis_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
-Between 2 and 4 iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
-Furthermore, as can be seen in Figure~\ref{fig10_diagnosis_lscale_comparison}, following our diagnostic recommendations avoids overestimation.
-
-As examples, the iterative steps applied to perform diagnosis on two of the data sets in Figure~\ref{fig9_diagnosis} are described next:
-
-\paragraph*{Data set with the exact GP length-scale $\ell_{\text{GP}} = 0.08$:} 
+Figure~\ref{fig9_diagnostic} shows the iterative steps of applying the diagnostic procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison} left.
+It is clearly visible that by following our recommendations, an optimum solution with minimum computational requirements is achieved in these cases.
+Figure~\ref{fig10_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
+Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
+Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids overestimation.
+
+TODO: overestimation of what?
+
+TODO: Personally, I think the two examples of the procedure below should go to an appendix, as they take a lot of space and don't add so much additional value.
+
+As concrete examples, the iterative steps applied to perform diagnostic on two of the data sets in Figure~\ref{fig9_diagnostic} are described next:
+
+\paragraph*{Data set with the squared exponential kernel and true length-scale $\ell_{\text{GP}} = 0.08$:} 
 \begin{itemize}
@@
 	\begin{enumerate}
-	\item Making the initial guess that the length-scale $\ell_1=0.5$.
-
-	\item Computing the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Computing the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
-
-	\item Fitting the HSGP model and obtain the estimated $\hat{\ell}_1$.
+	\item Make the initial guess that the length-scale $\ell_1=0.5$.
+
+	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	TODO: name the value of $c_1$ and all the other quantities in the following. Otherwise, what is the purpose of this example?
+
+	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
+
+	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
+
+	TODO: provide at least a mean value for the posterior of  $\hat{\ell}_1$ here?
 
 	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives FALSE.
+
+	TODO: I think you wanted to use $\hat{\ell}_1 - 0.01 \geq \ell_1$ and I corrected it above. Can you verify whether you want + or -?
 	\end{enumerate}
@@
 	\begin{enumerate}
-	\item As the diagnostic in previous iteration 1 was FALSE, update $\ell_2$ with $\hat{\ell}_1$.
-
-	\item Computing the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.
-
-	\item Fitting the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+	\item As the diagnostic in iteration 1 was FALSE, set $\ell_2 = \hat{\ell}_1$.
+
+	\item Compute the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.
+
+	TODO: why make this a single step here when it was two steps in Section 4.4.1?
+
+	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
 
@@
 	\begin{enumerate}
-	\item As the diagnostic in step 4 in previous iteration 3 was TRUE, increasing $m_4 = m_3 + 5$.
-
-	\item Computing the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Fitting the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.
+	\item As the diagnostic in step 4 in previous iteration 3 was TRUE, set $m_4 = m_3 + 5$.
+
+	\item Compute the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Fit the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.
 
@@
 
-	\item As RMSE, $R^2$ and ELPD are stable relative to previous iteration 4 the approximation must be sufficiently accurate.
+	\item As RMSE, $R^2$, and ELPD are stable relative to previous iteration 4, the HSGP approximation is likely sufficiently accurate and the procedure ends here.
 	\end{enumerate}
@@
 
-\paragraph*{Data set with the exact GP length-scale $\ell_{\text{GP}} = 1.4$:} 
+\paragraph*{Data set with the squared exponential kernel and true length-scale $\ell_{\text{GP}} = 1.4$:} 
 \begin{itemize}
@@
 	\begin{enumerate}
-	\item Making the initial guess that the length-scale $\ell_1=1$.
-
-	\item Computing the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Computing the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
-
-	\item Fitting the HSGP model and obtain the estimated $\hat{\ell}_1$.
+	\item Make the initial guess that the length-scale $\ell_1=1$.
+
+	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
+
+	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
 
@@
 	\begin{enumerate}
-	\item As the diagnostic in previous iteration 1 was TRUE, increasing $m_2 = m_1 + 5$.
-
-	\item Computing the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Fitting the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+	\item As the diagnostic in previous iteration 1 was TRUE, set $m_2 = m_1 + 5$.
+
+	\item Compute the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
 
@@
 
-	\item As RMSE, $R^2$ and ELPD are stable relative to previous iteration 2 the approximation must be sufficiently accurate.
+	\item As RMSE, $R^2$, and ELPD are stable relative to previous iteration 2, the HSGP approximation is likely sufficiently accurate and the procedure ends here.
 	\end{enumerate}
@@
 
-It is worth emphasizing that using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential in uni-dimensional cases with very large data sets, and even more in low-dimensional cases ($D \leq 4$; Notice from section~\ref{sec_method_multi} that for $D > 4$ it starts to be unpractical in any case) even with small data sets, since knowing the optimum values for $m$ and $c$ for each individual dimension can reduce significantly the computational requirements in these multi-dimensional cases, otherwise they would be unfeasible to be fitted.
+It is worth emphasizing that using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential in uni-dimensional cases with very large data sets.
+Using such optimal values will be even more essential in multi-dimensional cases ($D \leq 4$; notice from Section~\ref{sec_method_multi} that for $D > 4$ HSGPs start to be unpractical in any case) even with small data sets, since knowing the optimum values for $m$ and $c$ for each individual dimension can significantly reduce the computational requirements in these multi-dimensional cases; otherwise fitting HSGPs would be infeasible here.
 
@@
 %\multirow{5}{*}{ 0.05} 
-% & 1 & 0.50 & 1.60 & 6 & 0.127 & FALSE & 0.800 & 0.264 & -0.980 & 0.748 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_1.png}}\\
+% & 1 & 0.50 & 1.60 & 6 & 0.127 & FALSE & 0.800 & 0.264 & -0.980 & 0.748 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_1.png}}\\
 % & 2 & 0.127 & 1.20 & 17 & 0.061 & FALSE & 0.427 & 0.779 & -0.396 & 0.330 & \\
@@
 \multirow{5}{*}{ 0.08} 
- & 1 & 0.50 & 1.60 & 6 & 0.169 & FALSE & 0.447 & 0.610 & -0.396 & 0.360 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_2.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.169 & FALSE & 0.447 & 0.610 & -0.396 & 0.360 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_2.png}}\\
  & 2 & 0.169 & 1.20 & 13 & 0.068 & FALSE & 0.357 & 0.738 & -0.176 & 0.242 & \\
@@
 \multirow{3}{*}{ 0.13} 
- & 1 & 0.50 & 1.60 & 6 & 0.250 & FALSE & 0.461 & 0.625 & -0.482 & 0.330 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_3.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.250 & FALSE & 0.461 & 0.625 & -0.482 & 0.330 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_3.png}}\\
  & 2 & 0.250 & 1.20 & 9 & 0.148 & FALSE & 0.352 & 0.780 & -0.135 & 0.153 & \\
@@
 \multirow{3}{*}{ 0.25} 
- & 1 & 0.50 & 1.60 & 6 & 0.237 & FALSE & 0.368 & 0.399 & -0.146 & 0.214 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_4.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.237 & FALSE & 0.368 & 0.399 & -0.146 & 0.214 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_4.png}}\\
  & 2 & 0.237 & 1.20 & 9 & 0.263 & TRUE & 0.294 & 0.603 & -0.085 & 0.024 & \\
@@
 \multirow{3}{*}{ 0.54} 
- & 1 & 0.50 & 1.60 & 6 & 0.376 & FALSE & 0.272 & 0.858 & 0.034 & 0.013 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_5.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.376 & FALSE & 0.272 & 0.858 & 0.034 & 0.013 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_5.png}}\\
  & 2 & 0.376 & 1.20 & 6 & 0.469 & TRUE & 0.277 & 0.855 & 0.029 & 0.051 & \\
@@
 \multirow{3}{*}{ 0.80} 
- & 1 & 1.00 & 3.20 & 6 & 0.516 & FALSE & 0.286 & 0.923 & 0.133 & 0.025 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_6.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 0.516 & FALSE & 0.286 & 0.923 & 0.133 & 0.025 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_6.png}}\\
  & 2 & 0.516 & 1.65 & 6 & 0.975 & TRUE & 0.286 & 0.923 & 0.107 & 0.030 & \\
@@
 \multirow{2}{*}{ 1.40} 
- & 1 & 1.00 & 3.20 & 6 & 1.023 & TRUE & 0.283 & 0.561 & 0.139 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_7.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 1.023 & TRUE & 0.283 & 0.561 & 0.139 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_7.png}}\\
  & 2 & 0.521 & 3.27 & 11 & 1.233 & TRUE & 0.281 & 0.566 & 0.136 & 0.003 & \\
@@
 \multirow{2}{*}{ 2.94} 
- & 1 & 0.50 & 1.60 & 6 & 1.192 & TRUE & 0.343 & 0.449 & -0.104 & 0.058 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_8.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 1.192 & TRUE & 0.343 & 0.449 & -0.104 & 0.058 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_8.png}}\\
  & 2 & 0.607 & 3.81 & 11 & 2.577 & TRUE & 0.341 & 0.454 & -0.074 & 0.008 & \\
@@
 \end{tabular}
-\caption{Diagnosis on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
-  \label{fig9_diagnosis}
+\caption{diagnostic on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+  \label{fig9_diagnostic}
 \end{figure*}
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_diagnosis_lscale_comparison.png}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_diagnostic_lscale_comparison.png}}
 \caption{Comparison of the true and estimated length-scales in every iteration for the different data sets varying the length-scale.}
-  \label{fig10_diagnosis_lscale_comparison}
+  \label{fig10_diagnostic_lscale_comparison}
 \end{figure}
@@
 
-%In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.
-
-%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships}.
+%In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnostic of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.
+
+%The diagnostic consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships}.
 Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
 Next, we demonstrate how to apply the diagnostic for approximation quality as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency}.
-Figure \ref{fig13_diagnosis} depicts the diagnosis applied to case study.
-In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnosis}.
+Figure \ref{fig13_diagnostic} depicts the diagnostic applied to case study.
+In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnostic}.
 In the second iteration, $m$ is increased to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_7_relationships} is $\ell^*=0.19$.
@@
 \end{tabular}
-\caption{Iterative diagnosis of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.
+\caption{Iterative diagnostic of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.
 The iterative steps go from the left side to the right side of the figure.}
-	\label{fig13_diagnosis}
+	\label{fig13_diagnostic}
 \end{figure*}
@@
 About 16\% of cases were censored.
-Predictors are \textit{age} ($x_1$), \textit{sex} ($x_2$), \textit{white blood cell} (WBC) ($x_3$) count at diagnosis with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($x_4$) which is a measure of deprivation for district of residence.
+Predictors are \textit{age} ($x_1$), \textit{sex} ($x_2$), \textit{white blood cell} (WBC) ($x_3$) count at diagnostic with 1 unit = $50\times109/L$, and the \textit{Townsend deprivation index} (TDI) ($x_4$) which is a measure of deprivation for district of residence.
 We denote $\bm{x}_i=(x_{i1},x_{i2},x_{i3},x_{i4}) \in {\rm I\!R}^{4}$ as the vector of predictor values for observation $i$.
@@
 
-The main contribution of this paper is an in-depth analysis and diagnosis of the performance and accuracy of the approximation in relation to the key factors of the method, that is, the number of basis functions, the boundary condition of the Laplace eigenfunctions, and the non-linearity of the function to be learned.
+The main contribution of this paper is an in-depth analysis and diagnostic of the performance and accuracy of the approximation in relation to the key factors of the method, that is, the number of basis functions, the boundary condition of the Laplace eigenfunctions, and the non-linearity of the function to be learned.
 Recommendations for the values of these key factors based on the recognized relations among them have been provided along with illustrations of these relations.
-These illustrations will not only help users to improve performance and save computation time, but also serve as a powerful diagnosis tool whether the chosen values for the number of basis functions and the boundary condition are adequate to fit to the data at hand with sufficient accuracy.
+These illustrations will not only help users to improve performance and save computation time, but also serve as a powerful diagnostic tool whether the chosen values for the number of basis functions and the boundary condition are adequate to fit to the data at hand with sufficient accuracy.
 

### 2022-02-18 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
reduce digits in fig 9

--- 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 \multirow{5}{*}{ 0.08} 
- & 1 & 0.50 & 1.60 & 6 & 0.169 & FALSE & 0.447 & 0.610 & -0.396 & 0.360 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_2.png}}\\
- & 2 & 0.169 & 1.20 & 13 & 0.068 & FALSE & 0.357 & 0.738 & -0.176 & 0.242 & \\
- & 3 & 0.068 & 1.20 & 31 & 0.080 & TRUE & 0.242 & 0.871 & 0.204 & 0.013 & \\
- & 4 & 0.058 & 1.20 & 36 & 0.081 & TRUE & 0.241 & 0.871 & 0.207 & 0.009 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.17 & FALSE & 0.45 & 0.61 & -0.40 & 0.36 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_2.png}}\\
+ & 2 & 0.17 & 1.20 & 13 & 0.07 & FALSE & 0.36 & 0.74 & -0.18 & 0.24 & \\
+ & 3 & 0.07 & 1.20 & 31 & 0.08 & TRUE & 0.24 & 0.87 & 0.20 & 0.01 & \\
+ & 4 & 0.06 & 1.20 & 36 & 0.08 & TRUE & 0.24 & 0.87 & 0.21 & 0.01 & \\
  \\[1mm]
@@
 \multirow{3}{*}{ 0.13} 
- & 1 & 0.50 & 1.60 & 6 & 0.250 & FALSE & 0.461 & 0.625 & -0.482 & 0.330 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_3.png}}\\
- & 2 & 0.250 & 1.20 & 9 & 0.148 & FALSE & 0.352 & 0.780 & -0.135 & 0.153 & \\
- & 3 & 0.148 & 1.20 & 15 & 0.144 & TRUE & 0.321 & 0.807 & -0.074 & 0.084 & \\
- & 4 & 0.105 & 1.20 & 20 & 0.117 & TRUE & 0.297 & 0.830 & -0.006 & 0.013 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.25 & FALSE & 0.46 & 0.63 & -0.49 & 0.33 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_3.png}}\\
+ & 2 & 0.25 & 1.20 & 9 & 0.15 & FALSE & 0.35 & 0.78 & -0.14 & 0.15 & \\
+ & 3 & 0.15 & 1.20 & 15 & 0.15 & TRUE & 0.32 & 0.81 & -0.07 & 0.08 & \\
+ & 4 & 0.11 & 1.20 & 20 & 0.12 & TRUE & 0.30 & 0.83 & -0.01 & 0.01 & \\
  \\[1mm]
@@
 \multirow{3}{*}{ 0.25} 
- & 1 & 0.50 & 1.60 & 6 & 0.237 & FALSE & 0.368 & 0.399 & -0.146 & 0.214 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_4.png}}\\
- & 2 & 0.237 & 1.20 & 9 & 0.263 & TRUE & 0.294 & 0.603 & -0.085 & 0.024 & \\
- & 3 & 0.250 & 1.20 & 14 & 0.257 & TRUE & 0.289 & 0.615 & -0.050 & 0.009 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.24 & FALSE & 0.37 & 0.40 & -0.15 & 0.21 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnosis_4.png}}\\
+ & 2 & 0.24 & 1.20 & 9 & 0.26 & TRUE & 0.30 & 0.60 & -0.09 & 0.02 & \\
+ & 3 & 0.25 & 1.20 & 14 & 0.26 & TRUE & 0.29 & 0.62 & -0.05 & 0.01 & \\
  \\
@@
 \multirow{3}{*}{ 0.54} 
- & 1 & 0.50 & 1.60 & 6 & 0.376 & FALSE & 0.272 & 0.858 & 0.034 & 0.013 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_5.png}}\\
- & 2 & 0.376 & 1.20 & 6 & 0.469 & TRUE & 0.277 & 0.855 & 0.029 & 0.051 & \\
- & 3 & 0.239 & 1.50 & 11 & 0.525 & TRUE & 0.272 & 0.858 & 0.036 & 0.014 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.38 & FALSE & 0.27 & 0.86 & 0.03 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_5.png}}\\
+ & 2 & 0.38 & 1.20 & 6 & 0.47 & TRUE & 0.28 & 0.86 & 0.03 & 0.05 & \\
+ & 3 & 0.24 & 1.50 & 11 & 0.53 & TRUE & 0.27 & 0.86 & 0.04 & 0.01 & \\
  \\
@@
 \multirow{3}{*}{ 0.80} 
- & 1 & 1.00 & 3.20 & 6 & 0.516 & FALSE & 0.286 & 0.923 & 0.133 & 0.025 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_6.png}}\\
- & 2 & 0.516 & 1.65 & 6 & 0.975 & TRUE & 0.286 & 0.923 & 0.107 & 0.030 & \\
- & 3 & 0.496 & 3.12 & 11 & 0.772 & TRUE & 0.283 & 0.924 & 0.109 & 0.005 & \\
+ & 1 & 1.00 & 3.20 & 6 & 0.52 & FALSE & 0.29 & 0.92 & 0.13 & 0.03 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_6.png}}\\
+ & 2 & 0.52 & 1.65 & 6 & 0.98 & TRUE & 0.29 & 0.92 & 0.11 & 0.03 & \\
+ & 3 & 0.50 & 3.12 & 11 & 0.77 & TRUE & 0.28 & 0.92 & 0.11 & 0.01 & \\
  \\
@@
 \multirow{2}{*}{ 1.40} 
- & 1 & 1.00 & 3.20 & 6 & 1.023 & TRUE & 0.283 & 0.561 & 0.139 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_7.png}}\\
- & 2 & 0.521 & 3.27 & 11 & 1.233 & TRUE & 0.281 & 0.566 & 0.136 & 0.003 & \\
+ & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_7.png}}\\
+ & 2 & 0.52 & 3.27 & 11 & 1.23 & TRUE & 0.28 & 0.57 & 0.14 & 0.00 & \\
  \\
@@
 \multirow{2}{*}{ 2.94} 
- & 1 & 0.50 & 1.60 & 6 & 1.192 & TRUE & 0.343 & 0.449 & -0.104 & 0.058 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_8.png}}\\
- & 2 & 0.607 & 3.81 & 11 & 2.577 & TRUE & 0.341 & 0.454 & -0.074 & 0.008 & \\
+ & 1 & 0.50 & 1.60 & 6 & 1.19 & TRUE & 0.34 & 0.45 & -0.10 & 0.06 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnosis_8.png}}\\
+ & 2 & 0.61 & 3.81 & 11 & 2.58 & TRUE & 0.34 & 0.45 & -0.07 & 0.01 & \\
  \\

### 2022-02-18 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex

--- 2022-02-17 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 
-\subsubsection{Relationships between $J$ and $\ell$ for a periodic squared exponential covariance function} \label{subsec_empiric_relations_periodic}
-
-TODO: Is $J$ equal to $m$ or is this something else?
-
-In Appendix~\ref{sec_periodic}, we analyze the accuracy of the linear representation of a periodic squared exponential covariance function and derive the minimum number of terms $J$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
+\subsubsection{Relationships between $m$ and $\ell$ for a periodic squared exponential covariance function} \label{subsec_empiric_relations_periodic}
+
+In Appendix~\ref{sec_periodic}, we analyze the accuracy of the linear representation of a periodic squared exponential covariance function and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
 Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
-This relationship between $J$ and $\ell$ is gathered in Figure~\ref{figB1_m_lscale_periodic}.
+This relationship between $m$ and $\ell$ is gathered in Figure~\ref{figB1_m_lscale_periodic}.
 
@@
 
-The numerical equation governing the relationships between $J$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\ref{eq_J_l_periodic} which is depicted next:
+The numerical equation governing the relationships between $m$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\ref{eq_J_l_periodic} which is depicted next:
 %
 \begin{align*}
-&J \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{J}
+&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m}
 \end{align*}
@@
 Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison} left).
-Here it comes out a possible weaknesses of the approach which consists in using too many basis functions relative to the true length-scale causes overestimating the posteriors.
-
-TODO: Please clarify the last sentence above. I don't get that is ment here.
 
@@
 Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
-Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids overestimation.
-
-TODO: overestimation of what?
+Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length scale that would cause overestimating the underlying function and bias the posterior towards smaller length scale values.
+
+TODO: overestimation of what? $\to$ Gabi: Please, see if it's better now.
 
 TODO: Personally, I think the two examples of the procedure below should go to an appendix, as they take a lot of space and don't add so much additional value.
+$\to$ Gabi: It's fine for me moving them to an appendix.
+If no-one disagree I'll move them later.
 
@@
 \multirow{5}{*}{ 0.08} 
- & 1 & 0.50 & 1.60 & 6 & 0.169 & FALSE & 0.447 & 0.610 & -0.396 & 0.360 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_2.png}}\\
- & 2 & 0.169 & 1.20 & 13 & 0.068 & FALSE & 0.357 & 0.738 & -0.176 & 0.242 & \\
- & 3 & 0.068 & 1.20 & 31 & 0.080 & TRUE & 0.242 & 0.871 & 0.204 & 0.013 & \\
- & 4 & 0.058 & 1.20 & 36 & 0.081 & TRUE & 0.241 & 0.871 & 0.207 & 0.009 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.17 & FALSE & 0.45 & 0.61 & -0.40 & 0.36 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_2.png}}\\
+ & 2 & 0.17 & 1.20 & 13 & 0.07 & FALSE & 0.36 & 0.74 & -0.18 & 0.24 & \\
+ & 3 & 0.07 & 1.20 & 31 & 0.08 & TRUE & 0.24 & 0.87 & 0.20 & 0.01 & \\
+ & 4 & 0.06 & 1.20 & 36 & 0.08 & TRUE & 0.24 & 0.87 & 0.21 & 0.01 & \\
  \\[1mm]
@@
 \multirow{3}{*}{ 0.13} 
- & 1 & 0.50 & 1.60 & 6 & 0.250 & FALSE & 0.461 & 0.625 & -0.482 & 0.330 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_3.png}}\\
- & 2 & 0.250 & 1.20 & 9 & 0.148 & FALSE & 0.352 & 0.780 & -0.135 & 0.153 & \\
- & 3 & 0.148 & 1.20 & 15 & 0.144 & TRUE & 0.321 & 0.807 & -0.074 & 0.084 & \\
- & 4 & 0.105 & 1.20 & 20 & 0.117 & TRUE & 0.297 & 0.830 & -0.006 & 0.013 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.25 & FALSE & 0.46 & 0.63 & -0.49 & 0.33 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_3.png}}\\
+ & 2 & 0.25 & 1.20 & 9 & 0.15 & FALSE & 0.35 & 0.78 & -0.14 & 0.15 & \\
+ & 3 & 0.15 & 1.20 & 15 & 0.15 & TRUE & 0.32 & 0.81 & -0.07 & 0.08 & \\
+ & 4 & 0.11 & 1.20 & 20 & 0.12 & TRUE & 0.30 & 0.83 & -0.01 & 0.01 & \\
  \\[1mm]
@@
 \multirow{3}{*}{ 0.25} 
- & 1 & 0.50 & 1.60 & 6 & 0.237 & FALSE & 0.368 & 0.399 & -0.146 & 0.214 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_4.png}}\\
- & 2 & 0.237 & 1.20 & 9 & 0.263 & TRUE & 0.294 & 0.603 & -0.085 & 0.024 & \\
- & 3 & 0.250 & 1.20 & 14 & 0.257 & TRUE & 0.289 & 0.615 & -0.050 & 0.009 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.24 & FALSE & 0.37 & 0.40 & -0.15 & 0.21 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_4.png}}\\
+ & 2 & 0.24 & 1.20 & 9 & 0.26 & TRUE & 0.30 & 0.60 & -0.09 & 0.02 & \\
+ & 3 & 0.25 & 1.20 & 14 & 0.26 & TRUE & 0.29 & 0.62 & -0.05 & 0.01 & \\
  \\
@@
 \multirow{3}{*}{ 0.54} 
- & 1 & 0.50 & 1.60 & 6 & 0.376 & FALSE & 0.272 & 0.858 & 0.034 & 0.013 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_5.png}}\\
- & 2 & 0.376 & 1.20 & 6 & 0.469 & TRUE & 0.277 & 0.855 & 0.029 & 0.051 & \\
- & 3 & 0.239 & 1.50 & 11 & 0.525 & TRUE & 0.272 & 0.858 & 0.036 & 0.014 & \\
+ & 1 & 0.50 & 1.60 & 6 & 0.38 & FALSE & 0.27 & 0.86 & 0.03 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_5.png}}\\
+ & 2 & 0.38 & 1.20 & 6 & 0.47 & TRUE & 0.28 & 0.86 & 0.03 & 0.05 & \\
+ & 3 & 0.24 & 1.50 & 11 & 0.53 & TRUE & 0.27 & 0.86 & 0.04 & 0.01 & \\
  \\
@@
 \multirow{3}{*}{ 0.80} 
- & 1 & 1.00 & 3.20 & 6 & 0.516 & FALSE & 0.286 & 0.923 & 0.133 & 0.025 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_6.png}}\\
- & 2 & 0.516 & 1.65 & 6 & 0.975 & TRUE & 0.286 & 0.923 & 0.107 & 0.030 & \\
- & 3 & 0.496 & 3.12 & 11 & 0.772 & TRUE & 0.283 & 0.924 & 0.109 & 0.005 & \\
+ & 1 & 1.00 & 3.20 & 6 & 0.52 & FALSE & 0.29 & 0.92 & 0.13 & 0.03 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_6.png}}\\
+ & 2 & 0.52 & 1.65 & 6 & 0.98 & TRUE & 0.29 & 0.92 & 0.11 & 0.03 & \\
+ & 3 & 0.50 & 3.12 & 11 & 0.77 & TRUE & 0.28 & 0.92 & 0.11 & 0.01 & \\
  \\
@@
 \multirow{2}{*}{ 1.40} 
- & 1 & 1.00 & 3.20 & 6 & 1.023 & TRUE & 0.283 & 0.561 & 0.139 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_7.png}}\\
- & 2 & 0.521 & 3.27 & 11 & 1.233 & TRUE & 0.281 & 0.566 & 0.136 & 0.003 & \\
+ & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_7.png}}\\
+ & 2 & 0.52 & 3.27 & 11 & 1.23 & TRUE & 0.28 & 0.57 & 0.14 & 0.00 & \\
  \\
@@
 \multirow{2}{*}{ 2.94} 
- & 1 & 0.50 & 1.60 & 6 & 1.192 & TRUE & 0.343 & 0.449 & -0.104 & 0.058 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_8.png}}\\
- & 2 & 0.607 & 3.81 & 11 & 2.577 & TRUE & 0.341 & 0.454 & -0.074 & 0.008 & \\
+ & 1 & 0.50 & 1.60 & 6 & 1.19 & TRUE & 0.34 & 0.45 & -0.10 & 0.06 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_8.png}}\\
+ & 2 & 0.61 & 3.81 & 11 & 2.58 & TRUE & 0.34 & 0.45 & -0.07 & 0.01 & \\
  \\

### 2022-02-22 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Edits to Sec. 4. Mainly typos and minor errors. Unified wording across whole paper as well.

@@
 A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation.
-In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic square exponential covariance function.
+In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic squared exponential covariance function.
 
@@
 Ultimately, these recommendations are based on the relationships among the number of basis functions $m$, the boundary condition $L$, and the length-scale $\ell$, which depend on the particular choice of the kernel function.
-In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
+In this work we investigate these relationships for the squared exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
 For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
@@
 The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models.
-Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
-Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
-Figure \ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
-Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and~$c$.
+Figure~\ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
+Figure~\ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
+Looking at the RMSEs in Figure~\ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
 Further, the less conservative choice of $m = 10$ and $c = 1.5$ could also produce a sufficiently accurately approximation depending on the application.
@@
 From these results, some general conclusions may be drawn:
-
+%
 \begin{itemize}
@@
 
-\subsection{Near linear proportionality between $m$, $c$, and $l$} \label{subsec_theoretic_linearity}
+\subsection{Near linear proportionality between $m$, $c$, and $\ell$} \label{subsec_theoretic_linearity}
 
@@
 % There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
-In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
+In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the length-scale $l$.
 % In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
@@
 % spectral densities and when they go to zero. This is illustrated in the same
-% notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
+% notebook after that figure. E.g. with length-scale l=1,l=1/2, l=1/4, l=1/8 to
 % explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
@@
 
-\subsection{Empirical discovering of the functional form of the relationships between $m$, $c$ and $l$} \label{subsec_empiric_relations}
+\subsection{Empirical discovering of the functional form of the relationships between $m$, $c$ and $\ell$} \label{subsec_empiric_relations}
 
@@
 \subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
   \label{fig6_relationships}
@@
 Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $l$.
-Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
-(\ref{eq_diff_covs}).
-We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
-(\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
+Figure~\ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure~\ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.~\eqref{eq_diff_covs}.
+We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.~\eqref{eq_diff_covs}, is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
 %
@@
 We obtain the following main conclusions:
-
+%
 \begin{itemize}
@@
 
-As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near
-linear proportionality between $m$, $\frac{l}{S}$ (lengthscale $l$ relative to the half range of the input domain $S$) and $c$.
+As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\frac{l}{S}$ (length-scale $l$ relative to the half range of the input domain $S$) and $c$.
 To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the practically useful constant terms.
@@
 \begin{align}\label{eq_c_vs_l_QE}
-&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2      
+&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2 .
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat52}
-&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
+&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2 .
 \end{align}
@@
 \begin{align}\label{eq_c_vs_l_mat32}
-&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2        
+&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2 .       
 \end{align}
 
-These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$).
+These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \to \infty$).
 Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions.
@@
 
-Figure~\ref{fig6_relationships} and previous equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
+Figure~\ref{fig6_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
 Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$.
@@
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+However, as an approximation, we can use the unidimensional GP conclusions in Figure~\ref{fig6_relationships} or equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
 
@@
 
-The numerical equation governing the relationships between $m$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\ref{eq_J_l_periodic} which is depicted next:
+The numerical equation governing the relationships between $m$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
 %
@@
 Guidelines for updating $m$ and/or $c$ are:
-
+%
 \begin{itemize}
@@
 
-If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
+If the initial length-scale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
 The following iterative procedure works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed for repeated fitting (e.g., in cross-validation or simulation based calibration).
@@
 Based on the above proposed diagnostics, we obtain a simple, iterative step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
-The procedure is split into two phases, Phase A and B, which have to be completed consecutively.
+The procedure is split into two phases, Phase~A and B, which have to be completed consecutively.
 %In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.
@@
 
-\item[A1.] Make an initial guess on the lengthscale $l_1$.
-If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \in [0.5, 1]$.
+\item[A1.] Make an initial guess on the length-scale $\ell_1$.
+If there is no useful information available, we recommend to start with a large length-scale, about $\ell_1 \in [0.5, 1]$.
 Set the iteration index to $i=1$.
 
-\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
+\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
 
@@
 
-\item[A5.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
+\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
 
@@
 
-If the diagnostic is TRUE, the HSGP approximation is close to be sufficiently accurate. Continue with Phase B.
+If the diagnostic is TRUE, the HSGP approximation is close to be sufficiently accurate. Continue with Phase~B.
 
@@
 
-\item[B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD).
+\item[B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predictive density (ELPD).
 
@@
 
-\item[B4.] Obtain the mimimum valid lengthscale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
+\item[B4.] Obtain the minimum valid length-scale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
 
@@
 
-\item[B6.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
+\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
 
@@
 
-If all the stability checks succeeed, the HSGP approximation of the latest model should be sufficiently accurate and the procedure ends here.
+If all the stability checks succeed, the HSGP approximation of the latest model should be sufficiently accurate and the procedure ends here.
 
@@
 \subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
-\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP lengthscale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
+\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP length-scale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
   \label{fig7_lscale_comparison}
@@
 \eqref{eq:rel_total_distance}.
-Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
+Ensuring that the relative distance is bounded by a small constant for the relevant length-scale implies a high quality approximation.
 
@@
 In this section, we first give a brief presentation of the results by \citet{solin2014explicit}, who obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
-Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic square exponential covariance function.
+Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic squared exponential covariance function.
 

### 2022-02-22 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Edits to Sec. 4. Mainly typos and minor errors. Unified wording across whole paper as well.

@@
 \begin{align*}
-&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m}
+&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m} .
 \end{align*}
@@
 obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness.
-Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.
-(\ref{eq_c_vs_l_QE}).
-Figure \ref{fig7_lscale_comparison} right shows the RMSE of the HSGP models with the exact GP model as the reference.
+Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.~(\ref{eq_c_vs_l_QE}).
+Figure~\ref{fig7_lscale_comparison} (right) shows the RMSE of the HSGP models with the exact GP model as the reference.
 It can be seen that accurate estimates of the length-scale imply small RMSEs.
 
-Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison} left).
+Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure~\ref{fig7_lscale_comparison} left).
 
@@
 Figure~\ref{fig10_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
-Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
-Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length scale that would cause overestimating the underlying function and bias the posterior towards smaller length scale values.
+Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length-scale and the true length-scale, are sufficient to reach the optimal values of $m$ and $c$.
+Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length-scale that would cause overestimating the underlying function and bias the posterior towards smaller length-scale values.
 
@@
 If no-one disagree I'll move them later.
+$\to$ Arno agrees.
 
@@
 
-In Section \ref{sec_accuracy}, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
-We specifically focused on the Matérn and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
+In Section~\ref{sec_accuracy}, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
+We specifically focused on the Mat\'ern and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
 The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
-The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
-\eqref{eq:rel_total_distance}.
+The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure~\ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.~\eqref{eq:rel_total_distance}.
 Ensuring that the relative distance is bounded by a small constant for the relevant length-scale implies a high quality approximation.

### 2022-02-22 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Notation.

@@
 
-\subsection{Near linear proportionality between $m$, $c$, and $\ell$} \label{subsec_theoretic_linearity}
+\subsection{Near linear proportionality between $m$, $c$ and $\ell$} \label{subsec_theoretic_linearity}
 
@@
 
-As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\frac{l}{S}$ (length-scale $l$ relative to the half range of the input domain $S$) and $c$.
+As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\frac{\ell}{S}$ (length-scale $\ell$ relative to the half range of the input domain $S$) and $c$.
 To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the practically useful constant terms.

### 2022-02-23 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Minor
Tex stuff

@@
 
-\author{Gabriel Riutort-Mayol \and Paul-Christian Bürkner \and Michael R.\ Andersen \and Arno Solin \and Aki Vehtari
+\author{Gabriel Riutort-Mayol \and Paul-Christian Bürkner \and Michael R.\ Andersen \and Arno~Solin \and Aki Vehtari
 }
@@
 \end{equation} 
-for $j=1,2,\cdots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
+for $j=1,2,\ldots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
 This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required.

### 2022-02-23 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Tex stuff

@@
 \begin{align*}
-<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) \,\mathrm{d}(\bm{x})
+\langle f,g \rangle=\int_\Omega f(\bm{x}) g(\bm{x}) \,\mathrm{d}(\bm{x})
 \end{align*} 
@@
 \end{equation}
-where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
+where $j=1,2,\ldots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
 %
@@
 
-From the empirical observations, a numerical equation governing the relationships between $J$ and $\ell$ were estimated in equation \ref{eq_J_l_periodic}, which show linear proportionality between $J$ and $\ell$.
+From the empirical observations, a numerical equation governing the relationships between $J$ and $\ell$ were estimated in equation \ref{eq_J_l_periodic}, which show linear proportionality between $J$ and $\ell$:
 %
 \begin{align}\label{eq_J_l_periodic}
-&J \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{J},
+&J \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{J}.
 \end{align}
@@
 \begin{align} \label{eq_cov_periodic_taylor_approx_2}
-k(x,x')\approx \; & \alpha \left( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x') \right. \nonumber\\
-& \left. + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \right).
+k(x,x')\approx \; & \alpha \bigg( \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 x)  \cos(j\omega_0 x')  \nonumber\\
+& + \sum_{j=1}^{J} \tilde{q}_j^2 \sin(j\omega_0 x) \sin(j\omega_0 x') \bigg).
 \end{align}
@@
 \begin{align} \label{eq_f_period}
-f(x) \approx \; & \alpha^{1/2} \left( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j \right. \nonumber\\
-& \left.+  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \right),
+f(x) \approx \; & \alpha^{1/2} \bigg( \sum_{j=0}^J  \tilde{q}_j \cos(j\omega_0 x) \beta_j \nonumber\\
+& +  \sum_{j=1}^J \tilde{q}_j \sin(j\omega_0 x) \beta_{J+1+j} \bigg),
 \end{align}

### 2022-02-23 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Minor edits in sec 4

@@
 
-In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ will include the smallest interval that contains all the inputs points $x_i$.
+In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ contains all the inputs points $x_i$.
 Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.
@@
 
-As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) provide the minimum length-scale that can be closely accurately given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data.
+As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of estimating the latent function from the data.
 Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.
@@
 
-If the initial length-scale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
+If the initial length-scale is not far from the optimal, and larger than the minimal $c$ and $m$ used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
 The following iterative procedure works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed for repeated fitting (e.g., in cross-validation or simulation based calibration).

### 2022-02-24 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
updating sec 4.5, case sutdy I and II, and adding new secs 4.5.2 and 4.8

--- 2022-02-23 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 
-\author{Gabriel Riutort-Mayol \and Paul-Christian Bürkner \and Michael R.\ Andersen \and Arno~Solin \and Aki Vehtari
+\author{Gabriel Riutort-Mayol \and Paul-Christian Bürkner \and Michael R.\ Andersen \and Arno Solin \and Aki Vehtari
 }
@@
 
-A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation.
-In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic squared exponential covariance function.
+A GP model with a periodic covariance function does not fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation.
+In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic square exponential covariance function.
 
@@
 Ultimately, these recommendations are based on the relationships among the number of basis functions $m$, the boundary condition $L$, and the length-scale $\ell$, which depend on the particular choice of the kernel function.
-In this work we investigate these relationships for the squared exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
+In this work we investigate these relationships for the square exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
 For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
@@
 The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models.
-Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and~$c$.
-Figure~\ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
-Figure~\ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
-Looking at the RMSEs in Figure~\ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$.
+Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
+Figure \ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
+Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
 Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application.
@@
 From these results, some general conclusions may be drawn:
-%
+
 \begin{itemize}
@@
 
-\subsection{Near linear proportionality between $m$, $c$ and $\ell$} \label{subsec_theoretic_linearity}
+\subsection{Near linear proportionality between $m$, $c$, and $\ell$} \label{subsec_theoretic_linearity}
 
@@
 
-When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when when $c$ is larger than the minimal value recommendation, the number of $m$ first terms needed to explain almost 100\% of the variation has a near linear relationship with $\ell$ and $c$.
+When considering squared exponential and Mat{\'e}rn covariance functions, we can show with simple algebra that when $c$ is larger than the minimal value recommendation, the number of $m$ first terms needed to explain almost 100\% of the variation has a near linear relationship with $\ell$ and $c$.
 With decreasing $\ell$, the $m$ should grow near linearly with $1/\ell$, and with increasing $c$, the $m$ should grow near linearly with $c$.
@@
 % There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
-In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the length-scale $l$.
+In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the lengthscale $l$.
 % In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
@@
 % spectral densities and when they go to zero. This is illustrated in the same
-% notebook after that figure. E.g. with length-scale l=1,l=1/2, l=1/4, l=1/8 to
+% notebook after that figure. E.g. with lengthscale l=1,l=1/2, l=1/4, l=1/8 to
 % explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
@@
 \subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
   \label{fig6_relationships}
@@
 
-Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $l$.
-Figure~\ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure~\ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.~\eqref{eq_diff_covs}.
-We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.~\eqref{eq_diff_covs}, is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
+Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ relative to the half-range $S$ of the input domain).
+Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) with length-scale $\frac{\ell}{S}$ (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+(\ref{eq_diff_covs}).
+We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq.
+(\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
 %
@@
 We obtain the following main conclusions:
-%
+
 \begin{itemize}
@@
 \item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
-\item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$.
+\item For a given $\ell$ there exist a minimum $c$ under which an accurate approximation is never going to be achieved regardless of $m$.
 This fact can be seen in Figure~\ref{fig6_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
@@
 
-\subsubsection{Relationships between $m$ and $\ell$ for a periodic squared exponential covariance function} \label{subsec_empiric_relations_periodic}
-
-In Appendix~\ref{sec_periodic}, we analyze the accuracy of the linear representation of a periodic squared exponential covariance function and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
-Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
-This relationship between $m$ and $\ell$ is gathered in Figure~\ref{figB1_m_lscale_periodic}.
-
 \subsubsection{Numerical equations}
@@
 
-As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\frac{\ell}{S}$ (length-scale $\ell$ relative to the half range of the input domain $S$) and $c$.
+As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\ell$ and $c$.
 To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the practically useful constant terms.
@@
 %
-\begin{align}\label{eq_m_l_QE}
-&m= 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 1.75 \, \frac{c}{m},
+\begin{align}
+&m= 1.75 \, \frac{c}{\nicefrac{\ell}{S}} \;\; \Leftrightarrow \;\; \nicefrac{\ell}{S}= 1.75 \, \frac{c}{m},	\label{eq_m_l_QE} \\
+&\text{with} \nonumber \\
+&c \geq \, 3.2 \, \nicefrac{\ell}{S} \;\; \& \;\, c \geq \, 1.2	\label{eq_c_vs_l_QE}
 \end{align}
-%
-with
-%
-\begin{align}\label{eq_c_vs_l_QE}
-&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2 .
+
+\paragraph*{Mat\'ern ($\nu$=5/2)}:
+%
+\begin{align}
+&m= 2.65 \, \frac{c}{\nicefrac{\ell}{S}} \;\; \Leftrightarrow \;\; \nicefrac{\ell}{S}= 2.65 \, \frac{c}{m}, \label{eq_m_l_mat52} \\
+&\text{with} \nonumber \\
+&c \, \geq \, 4.1 \, \nicefrac{\ell}{S} \;\; \& \;\, c \geq \, 1.2  \label{eq_c_vs_l_mat52}
 \end{align}
 
-\paragraph*{Mat\'ern ($\nu$=5/2)}:
-%
-\begin{align}\label{eq_m_l_mat52}
-&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
+\paragraph*{Mat\'ern ($\nu$=3/2)}:
+%
+\begin{align}
+&m= 3.42 \, \frac{c}{\nicefrac{\ell}{S}} \;\; \Leftrightarrow \;\; \nicefrac{\ell}{S}= 3.42 \, \frac{c}{m}, \label{eq_m_l_mat32}\\
+&\text{with} \nonumber \\
+&c \, \geq \, 4.5 \, \nicefrac{\ell}{S} \;\; \& \;\, c \geq \, 1.2 \label{eq_c_vs_l_mat32}
 \end{align}
 %
-with
-%
-\begin{align}\label{eq_c_vs_l_mat52}
-&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2 .
-\end{align}
-
-\paragraph*{Mat\'ern ($\nu$=3/2)}:
-%
-\begin{align}\label{eq_m_l_mat32}
-&m= 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 3.42 \, \frac{c}{m},
-\end{align}
-%
-with
-%
-\begin{align}\label{eq_c_vs_l_mat32}
-&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2 .       
-\end{align}
-
-These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \to \infty$).
+These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$).
 Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions.
@@
 
-Figure~\ref{fig6_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
+Figure~\ref{fig6_relationships} and previous equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
 Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$.
@@
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figure~\ref{fig6_relationships} or equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
-
-\subsubsection{Numerical equation for a periodic squared exponential covariance function} \label{sec_num_equation_periodic}
-
-The numerical equation governing the relationships between $m$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
+However, as an approximation, we can use the unidimensional GP conclusions in Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+
+\subsection{Relationships between $m$ and $\ell$ for a periodic squared exponential covariance function} \label{subsec_empiric_relations_periodic}
+
+As commented in Section~\ref{sec_method_periodic}, in Appendix~\ref{sec_periodic} we present an approximate linear representation of a periodic squared exponential covariance function.
+In Appendix~\ref{sec_periodic}, we also analyze the accuracy of this linear representation and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
+Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
+This relationship between $m$ and $\ell$ for a periodic squared exponential covariance function is gathered in Figure~\ref{figB1_m_lscale_periodic} and the numerical equation was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
 %
 \begin{align*}
-&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m} .
+&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m}
 \end{align*}
@@
 
-As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be closely approximated given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 
-Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided by equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used).
-If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good.
+Having obtained an estimate $\hat{\ell}$ of a given $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided as a function of $m$ and $c$ by equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used).
+If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good, or at least close to.
 If, however, $\hat{\ell}$ does not exceed the recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated.
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also match those obtained from eq.
-(\ref{eq_m_l_QE}).
-
-Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to minimizing the computational cost.
-
-Guidelines for updating $m$ and/or $c$ are:
+If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also match those obtained from equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
+
+Equations in Section~\ref{sec_num_equations} to update $m$ and $c$ imply:
 % 
 \begin{itemize}
-	\item $c$ must be big enough for the given length-scale $l$ of the function to be learned and,
+	\item $c$ must be big enough for the given length-scale $\ell$ of the function to be learned and,
 	\item $m$ must be big enough for the given $\ell$ and $c$. 
 \end{itemize}
-
-We recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using the equations in Section~\ref{sec_num_equations}; and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$, update $\ell$ with $\hat{\ell}$ and repeat the process.
-
-If the initial length-scale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
-The following iterative procedure works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed for repeated fitting (e.g., in cross-validation or simulation based calibration).
-
-%A detailed guideline with the steps to perform diagnostic, ensuring that a good approximation is achieved with the minimum computational requirements (i.e.
-minimum number of basis functions $m$) is provided in  next Section~\ref{sec_user_guide}.
+%
+If larger than minimal $c$ and $m$ for the given length-scale $\ell$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
+However, as the given length-scale $\ell$ of the function to be learned is unknown, we recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using the equations in Section~\ref{sec_num_equations}; and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$, which indicates a potential problem with the approximation, update $\ell$ with $\hat{\ell}$ and repeat the process until both:
+%
+\begin{itemize}
+\item the estimated $\hat{\ell}$ is larger than the minimum lengthscale determined by $c$ and $m$, which indicates the approximation must be sufficiently accurate, and
+\item the measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD) reach a stabilization.
+\end{itemize}
+
+As commented in next Section~\ref{sec_diagnostic_criticism}, the estimated $\hat{\ell}$ being larger than the minimum lengthscale determined by $c$ and $m$ is not a definite certain that the approximate is sufficiently accurate, although it is for most of the practical cases.
+Therefore, we recommend to look at the residual error of the fit until it stabilizes.
+
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to minimizing the computational cost.
+
+%If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
+The iterative procedure, briefly commented above and detailed in next Section~\ref{sec_user_guide}, works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed.
+Using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential for repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional functions with very large data sets, and even more essential in multi-dimensional cases ($D \leq 4$; notice from Section~\ref{sec_method_multi} that for $D > 4$ HSGPs start to be unpractical in any case) where knowing the optimum values for $m$ and $c$ for each individual dimension can significantly reduce the computational requirements in these multi-dimensional cases; otherwise fitting HSGPs would be unfeasible here.
 
@@
 Based on the above proposed diagnostics, we obtain a simple, iterative step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation.
-The procedure is split into two phases, Phase~A and B, which have to be completed consecutively.
+The procedure is split into two phases, Phase A and B, which have to be completed consecutively.
 %In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.
@@
 
-\item[A1.] Make an initial guess on the length-scale $\ell_1$.
-If there is no useful information available, we recommend to start with a large length-scale, about $\ell_1 \in [0.5, 1]$.
+\item[A1.] Make an initial guess on the lengthscale $\ell_1$.
+If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \in [0.5, 1]$.
 Set the iteration index to $i=1$.
 
-\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
+\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
 
@@
 
-\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
+\item[A5.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
 
@@
 
-If the diagnostic is TRUE, the HSGP approximation is close to be sufficiently accurate. Continue with Phase~B.
+If the diagnostic is TRUE, the HSGP approximation is close to be sufficiently accurate. Continue with Phase B.
 
@@
 
-\item[B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predictive density (ELPD).
+\item[B1.] For the current HSGP model, compute measures of predictive accuracy, for example, RMSE, $R^2$, or ELPD.
 
@@
 
-\item[B4.] Obtain the minimum valid length-scale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
+\item[B4.] Obtain the mimimum valid lengthscale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
 
@@
 
-\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
+\item[B6.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
 
@@
 
-If all the stability checks succeed, the HSGP approximation of the latest model should be sufficiently accurate and the procedure ends here.
+If all the stability checks succeeed, the HSGP approximation of the latest model should be sufficiently accurate and the procedure ends here.
 
@@
 
+\subsubsection{"Criticism" of the diagnostic} \label{sec_diagnostic_criticism}
+
+The basis of our diagnostic tool are the following:
+% 
+\begin{enumerate}
+\item observing a length-scale estimate $\hat{\ell}$ smaller than recommended minimum length-scale determined by $m$ and $c$ indicates a potential problem with the approximation,
+\item observing a $\hat{\ell}$ larger than the recommended minimum length scale implies that the approximation is accurate.
+\end{enumerate}
+
+The later statement, however, cannot be assumed to be true in general, that is, cannot be generalized to all possible cases.
+({\color{red} TODO: A major revision from a reviewer was related to that, i.e.
+to be critical with the diagnostic and discuss about possible cases where it could fail.
+I don't know if we could discuss a bit about this.
+I remember Michael commented an example of a potential case where that diagnostic statement could not be true.})
+
+For that reason, to ultimately diagnose whether the approximation is accurate, in addition to observe whether $\hat{\ell}$ exceeds the recommended minimum length scale, we recommend ensuring the measures of predictive accuracy reach an stabilization too.
+
 \subsection{Performance analysis of the diagnostics}
@@
 \subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
-\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP length-scale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
+\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP lengthscale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
   \label{fig7_lscale_comparison}
@@
 obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness.
-Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.~(\ref{eq_c_vs_l_QE}).
-Figure~\ref{fig7_lscale_comparison} (right) shows the RMSE of the HSGP models with the exact GP model as the reference.
+Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.
+(\ref{eq_c_vs_l_QE}).
+Figure \ref{fig7_lscale_comparison} right shows the RMSE of the HSGP models with the exact GP model as the reference.
 It can be seen that accurate estimates of the length-scale imply small RMSEs.
 
-Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure~\ref{fig7_lscale_comparison} left).
+Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_overestimated_posterios} as an illustration of the overestimating effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison} left).
 
@@
 \centering
-\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_oversmoothed_posterios.png}
-\caption{Oversmoothing effect on posterior functions when using too many number of basis functions relative to the wigglyness of the function to be learned.}
-  \label{fig8_oversmoothed_posterios}
+\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_overestimated_posterios.png}
+\caption{Overestimating effect on posterior functions when using too many number of basis functions relative to the wigglyness of the function to be learned.
+The dashed red line represents the true function.}
+  \label{fig8_overestimated_posterios}
 \end{figure}
 
-Figure~\ref{fig9_diagnostic} shows the iterative steps of applying the diagnostic procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison} left.
+Table~\ref{tab1_diagnostic} shows the iterative steps of applying the diagnostic procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison} left.
 It is clearly visible that by following our recommendations, an optimum solution with minimum computational requirements is achieved in these cases.
 Figure~\ref{fig10_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
-Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length-scale and the true length-scale, are sufficient to reach the optimal values of $m$ and $c$.
-Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length-scale that would cause overestimating the underlying function and bias the posterior towards smaller length-scale values.
-
-TODO: overestimation of what? $\to$ Gabi: Please, see if it's better now.
-
-TODO: Personally, I think the two examples of the procedure below should go to an appendix, as they take a lot of space and don't add so much additional value.
-$\to$ Gabi: It's fine for me moving them to an appendix.
-If no-one disagree I'll move them later.
-$\to$ Arno agrees.
-
-As concrete examples, the iterative steps applied to perform diagnostic on two of the data sets in Figure~\ref{fig9_diagnostic} are described next:
-
-\paragraph*{Data set with the squared exponential kernel and true length-scale $\ell_{\text{GP}} = 0.08$:} 
-\begin{itemize}
-\item[] Iteration 1:
-	\begin{enumerate}
-	\item Make the initial guess that the length-scale $\ell_1=0.5$.
-
-	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	TODO: name the value of $c_1$ and all the other quantities in the following. Otherwise, what is the purpose of this example?
-
-	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
-
-	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
-
-	TODO: provide at least a mean value for the posterior of  $\hat{\ell}_1$ here?
-
-	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives FALSE.
-
-	TODO: I think you wanted to use $\hat{\ell}_1 - 0.01 \geq \ell_1$ and I corrected it above. Can you verify whether you want + or -?
-	\end{enumerate}
-
-\item[] Iteration 2:
-	\begin{enumerate}
-	\item As the diagnostic in iteration 1 was FALSE, set $\ell_2 = \hat{\ell}_1$.
-
-	\item Compute the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.
-
-	TODO: why make this a single step here when it was two steps in Section 4.4.1?
-
-	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
-
-	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives FALSE.
-	\end{enumerate}
-
-\item[] Iteration 3:
-	\begin{enumerate}
-	\item As the diagnostic in previous iteration 2 was FALSE, updating $\ell_3$ with $\hat{\ell}_2$ and repeat the process as in iteration 2.
-
-	\item The diagnostic $\hat{\ell}_3 + 0.01 \geq \ell_3$ gives TRUE.
-	\end{enumerate}
-
-\item[] Iteration 4:
-	\begin{enumerate}
-	\item As the diagnostic in step 4 in previous iteration 3 was TRUE, set $m_4 = m_3 + 5$.
-
-	\item Compute the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Fit the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.
-
-	\item The diagnostic $\hat{\ell}_4 + 0.01 \geq \ell_4$ gives TRUE.
-
-	\item As RMSE, $R^2$, and ELPD are stable relative to previous iteration 4, the HSGP approximation is likely sufficiently accurate and the procedure ends here.
-	\end{enumerate}
-
-\end{itemize}
-
-\paragraph*{Data set with the squared exponential kernel and true length-scale $\ell_{\text{GP}} = 1.4$:} 
-\begin{itemize}
-\item[] Iteration 1:
-	\begin{enumerate}
-	\item Make the initial guess that the length-scale $\ell_1=1$.
-
-	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
-
-	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
-
-	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives TRUE.
-	\end{enumerate}
-
-\item[] Iteration 2:
-	\begin{enumerate}
-	\item As the diagnostic in previous iteration 1 was TRUE, set $m_2 = m_1 + 5$.
-
-	\item Compute the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
-
-	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives TRUE.
-
-	\item As RMSE, $R^2$, and ELPD are stable relative to previous iteration 2, the HSGP approximation is likely sufficiently accurate and the procedure ends here.
-	\end{enumerate}
-
-\end{itemize}
-
-It is worth emphasizing that using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential in uni-dimensional cases with very large data sets.
-Using such optimal values will be even more essential in multi-dimensional cases ($D \leq 4$; notice from Section~\ref{sec_method_multi} that for $D > 4$ HSGPs start to be unpractical in any case) even with small data sets, since knowing the optimum values for $m$ and $c$ for each individual dimension can significantly reduce the computational requirements in these multi-dimensional cases; otherwise fitting HSGPs would be infeasible here.
-
-\begin{figure*}
+Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
+Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length scale that would cause overestimating the underlying function and bias the posterior towards smaller length scale values.
+
+As concrete examples, the iterative steps applied to perform diagnostic on two of the data sets in Table~\ref{tab1_diagnostic} are described in Appendix~\ref{sec_ex_iterative_steps}.
+
+\begin{table*}
 \centering
@@
 \multirow{2}{*}{ 1.40} 
- & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.010 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_7.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_7.png}}\\
  & 2 & 0.52 & 3.27 & 11 & 1.23 & TRUE & 0.28 & 0.57 & 0.14 & 0.00 & \\
@@
 \caption{diagnostic on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
-  \label{fig9_diagnostic}
-\end{figure*}
+  \label{tab1_diagnostic}
+\end{table*}
 
@@
 
-In Section~\ref{sec_accuracy}, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
-We specifically focused on the Mat\'ern and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
+Above, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
+We specifically focused on the Matérn and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
 The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
-The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure~\ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.~\eqref{eq:rel_total_distance}.
-Ensuring that the relative distance is bounded by a small constant for the relevant length-scale implies a high quality approximation.
+The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
+\eqref{eq:rel_total_distance}.
+Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
+Other possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation are actually explaining.
+
+To select $c$ for many other covariance functions, users can be guided by equations~\eqref{eq_c_vs_l_QE}, \eqref{eq_c_vs_l_mat52} and \eqref{eq_c_vs_l_mat32}, as pointed in Section~\ref{sec_num_equations}.
+
+\subsection{On computational requirements in multi-dimensional cases}
+
+%In the online supplemental material, additional case studies are presented. From those examples, it can be seen how 
+The computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
+Even though, the computation time increases significantly with $D$ in bivariate input spaces, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
+For moderate sized datasets, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
+However, the HSGP method will be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
+In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
+
+Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
 
@@
 
-In this section, we will present several simulated and real case studies in which we apply the developed HSGP models. More case studies are presented in the online supplemental materials.
+In this section, we will present several simulated and real case studies in which we apply the developed HSGP models and the recommended steps to fit them.
+More case studies are presented in the online supplemental materials.
 
@@
 In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
-The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process using the Mat{\'e}rn ($\nu$=3/2) covariance function with marginal variance $\alpha=1$ and length-scale $\ell=0.15$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
+The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process $f(x) \sim \GP(0, k(x, x', \theta))$ with a Mat{\'e}rn ($\nu$=3/2) covariance function $k$ with marginal variance $\alpha=1$ and length-scale $\ell=0.2$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
-We split the dataset into three parts: $155$ data points are used for fitting the model (training set), $45$ data ponts are used for the interpolation test set, and the remaining $50$ data points are used for the extrapolation test set.
-
-The exact GP model for fitting and predicting this simulated dataset $\bm{y}$ can be written as follows,
-%
-\begin{align*}\label{eq_latentgp_simudata1}
-\begin{split}
-\bm{y} &= \bm{f} + \bm{\epsilon} \\
-\bm{\epsilon} &\sim \Normal(\bm{0}, \sigma^2 \bm{I}) \\
-f(x) &\sim \GP(0, k(x, x', \theta)),
-\end{split}
-\end{align*}
-%
-where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $\bm{I}$ representing the identity matrix.
+%We split the dataset into three parts: $155$ data points are used for fitting the model (training set), $45$ data ponts are used for the interpolation test set, and the remaining $50$ data points are used for the extrapolation test set.
+%
+%The exact GP model for fitting and predicting this simulated dataset $\bm{y}$ can be written as follows,
+%%
+%\begin{align*}\label{eq_latentgp_simudata1}
+%\begin{split}
+%\bm{y} &= \bm{f} + \bm{\epsilon} \\
+%\bm{\epsilon} &\sim \Normal(\bm{0}, \sigma^2 \bm{I}) \\
+%f(x) &\sim \GP(0, k(x, x', \theta)),
+%\end{split}
+%\end{align*}
+%%
+%where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $\bm{I}$ representing the identity matrix.
 The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn ($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
@@
 Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,2,\dots,n$.
-
-A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using MCMC sampling methods, would be its marginalized form,
-%
-\begin{equation*}\label{eq_marginalizedgp_simudata1}
-\bm{y} \sim \Normal(\bm{0},\bm{K}+ \sigma^2 \bm{I} ),
-\end{equation*}
-%
-where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
+% 
+%A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using MCMC sampling methods, would be its marginalized form,
+%%
+%\begin{equation*}\label{eq_marginalizedgp_simudata1}
+%\bm{y} \sim \Normal(\bm{0},\bm{K}+ \sigma^2 \bm{I} ),
+%\end{equation*}
+%%
+%where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
 
 In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn ($\nu$=3/2) spectral density $S$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
-In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+%In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
 
-The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.\break A $\mathrm{Gamma}(1,1)$ prior distribution has been used for both observation noise $\sigma$ and covariance function marginal variance $\alpha$, and a $\mathrm{Gamma}(3.75,25)$ prior distribution for length-scale $\ell$.
-We use the same prior distributions for the exact GP model as for the HSGP models.
-
-Figure~\ref{fig11_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=60$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~(\ref{eq_boundary})), and the spline model with 80 knots.
-The true data-generating function and the noisy observations are also plotted.
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
+We use the same prior distributions to fit the exact GP model as the HSGP models.
+
+The HSGP model is fitted following the recommended iterative steps as in Section~\ref{sec_user_guide}.
+A initial value for the minimum lengthscale $\ell$ to use at the first iteration is guessed to be 0.5.
+While diagnostic $\hat{\ell} + 0.01 \geq \ell$ is \textit{false}, $c$ and $m$ are updated by equations~\eqref{eq_c_vs_l_mat32} and \eqref{eq_m_l_mat32}, respectively, and the minimum $\ell$ is updated with the estimated $\hat{\ell}$.
+After the diagnostic generated the first \textit{true}, $m$ is updated by increasing the $m$ of the previous iteration by 5 additional basis functions, $c$ is updated by eq.~\eqref{eq_c_vs_l_mat32} as a function of the estimated $\hat{\ell}$ at previous iteration, and the minimum $\ell$ is set by eq.~\eqref{eq_m_l_mat32} as function of $c$ and $m$.
+Table~\ref{tab_caseI} contains the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ and the RMSE compute with both the data and the GP as the reference for every iterative steps of the fitting process.
+
+\begin{table}
+\centering
+\begin{tabular}{cccccccc}
+\arrayrulecolor{gray}\hline \\[-3mm]
+iter. & $\ell$ & $c$ & $m$ & $\hat{\ell}$ & $\hat{\ell} + 0.01 \geq \ell$ & RMSE & RMSE* \\ 
+\arrayrulecolor{lightgray}\hline \\[-1mm]
+ 1 & 0.50 & 2.25 & 16 & 0.12 & FALSE & 0.26 & 0.17 \\
+ 2 & 0.12 & 1.20 & 35 & 0.32 & TRUE & 0.18 & 0.01 \\
+ 3 & 0.02 & 1.20 & 40 & 0.28 & TRUE & 0.18 & 0.01 \\[1mm]
+\arrayrulecolor{gray}\hline
+\end{tabular}
+\caption{diagnostic on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+  \label{tab_caseI}
+\end{table}
+
+Figure~\ref{fig11_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
+In order to do model comparison a spline-based model with 60 knots is also fitted and plotted jointly with the GP and HSGP models.
+The spline-based model is fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
 The sample observations are plotted as circles and the out-of-sample or test data are plotted as crosses.
-The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation.
 The posteriors of the three models, exact GP, HSGP and spline, are similar in the interpolation regions of the input space.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.85\textwidth]{fig11_Posteriors_exI.pdf}
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{fig11_Posteriors_exI.png}
 \caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
   \label{fig11_Posteriors_exI}
-\end{figure*}
-
-In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted.
+\end{figure}
+
+%\textsc{In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted.
 Figure \ref{fig12_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots.
@@
 However, the spline model does not extrapolate data properly.
-Both models show roughly similar interpolating performance.
+Both models show roughly similar interpolating performance.}
 
@@
 Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
-Next, we demonstrate how to apply the diagnostic for approximation quality as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency}.
-Figure \ref{fig13_diagnostic} depicts the diagnostic applied to case study.
+%Next, we demonstrate how to apply the diagnostic for approximation quality as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency}.
+%Figure \ref{fig13_diagnostic} depicts the diagnostic applied to case study.
 In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnostic}.
@@
 
-\begin{figure*}[!t]
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_inter.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_extra.pdf}}
-\caption{Standardized root mean square error (SRMSE) of the different methods against the data-generating function.
+%\begin{figure*}[!t]
+%\centering
+%\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_inter.pdf}}
+%\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_extra.pdf}}
+%\caption{Standardized root mean square error (SRMSE) of the different methods against the data-generating function.
 SRMSE for interpolation (left) and SRMSE for extrapolation (right).
 The standard deviation of the mean of the SRMSE is plotted as dashed lines.}
-  \label{fig12_MSE_exI_inter}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\begin{tabular}{llll}
-{\Large 1}\;\;\; \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} &  {\Large 2}\;\;\; \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.12$} &  {\Large 3}\;\;\; \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} &  {\Large 4}\;\;\; \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
-	& \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} > \ell^*$\\[1mm]
-\hspace{-3mm} \includegraphics[scale=0.26, trim = 0mm 0mm 4mm 10mm, clip]{fig13_diag_iter1_exI.pdf} & \hspace{-3mm}
-\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter2_exI.pdf} & \hspace{-3mm}
-\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter3_exI.pdf} & \hspace{-4mm}
-\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter4_exI.pdf} \\[1mm]
-\end{tabular}
-\caption{Iterative diagnostic of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.
+%  \label{fig12_MSE_exI_inter}
+%\end{figure*}
+%
+%
+%\begin{figure*}
+%\centering
+%\begin{tabular}{llll}
+%{\Large 1}\;\;\; \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} &  {\Large 2}\;\;\; \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.12$} &  {\Large 3}\;\;\; \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} &  {\Large 4}\;\;\; \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
+%	& \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} > \ell^*$\\[1mm]
+%\hspace{-3mm} \includegraphics[scale=0.26, trim = 0mm 0mm 4mm 10mm, clip]{fig13_diag_iter1_exI.pdf} & \hspace{-3mm}
+%\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter2_exI.pdf} & \hspace{-3mm}
+%\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter3_exI.pdf} & \hspace{-4mm}
+%\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter4_exI.pdf} \\[1mm]
+%\end{tabular}
+%\caption{Iterative diagnostic of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.
 The iterative steps go from the left side to the right side of the figure.}
-	\label{fig13_diagnostic}
-\end{figure*}
-
-Figure~\ref{fig14_time_exI} shows computational times, in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+%	\label{fig13_diagnostic}
+%\end{figure*}
+
+Figure~\ref{fig14_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+Both the HSGP and splines models show roughly similar performance at sample data.
+The HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation.
+However, the spline model does not extrapolate data properly (Figure~\ref{fig11_Posteriors_exI}).
 The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 10mm 0mm, clip]{fig14_time_exI.pdf}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig14_rmse_time_exI.png}}
 \caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
@@
 
-In the online supplemental material, additional case studies are presented.
+%In the online supplemental material, additional case studies are presented.
 From those examples, it can be seen how the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
@@
 In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
-
-Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
+%
+%Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
 
@@
 
-The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
-%
-\begin{equation*}
-f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
-\end{equation*}
-%
-A horse-shoe prior allows for sparse distributed effects.
+%The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
+%%
+%\begin{equation*}
+%f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
+%\end{equation*}
+%%
+%A horse-shoe prior allows for sparse distributed effects.
 Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
@@
 
-The component $f_1(t)$ will be approximated using the HSGP model and the function values $f_1(t)$ are approximated as in eq.~(\ref{eq_approxf}), with the squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
-We use $m=30$ basis functions and a boundary factor $c=1.5$.
+The component $f_4(t)$ represents the special days effects, modeled as a t-student prior model with 1 degree of freedom and variance $\tau^2$:
+%
+\begin{equation*}
+f_4(t) \sim \text{t-student}(1,0,\tau^2).
+\end{equation*}
+%
+%A horse-shoe prior allows for sparse distributed effects.
+Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
+Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
+
+The component $f_1$ will be approximated using the HSGP model and the function values $f_1$ are approximated as in eq.~(\ref{eq_approxf}), with the squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+The year effects $f_2$ and week effects $f_3$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper.
+However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{sec_periodic}).
+Thus, the functions $f_2$ and $f_3$ are approximated as in eq.~(\ref{eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in eq.~(\ref{eq_q_2}).
+
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1 \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell, \ell,\ell\}$, respectively.
+A $\mathrm{Normal}(0,0.1)$ prior distribution has been used for the standard deviation $\tau$ of the t-student distribution with 1 degree of freedom used to model $f_4$ (i.e., the special days effects).
+
+The HSGP model is fitted following the recommended iterative steps as in Section~\ref{sec_user_guide}, where in each iteration the diagnosis is applied on $f_1$, $f_2$ and $f_3$, where for each one these functions the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated.
+For functions $f_2$ and $f_3$ there are not boundary factor $c$ as they use periodic covariance functions, and $m$ and minimum $\ell$ are updated by eq.~\eqref{eq_J_l_periodic}.
+A initial value for the minimum lengthscale $\ell_1$ of $f_1$ is guessed to correspond to around 3 years (i.e., $\ell_1=3$ years $= 3\cdot 365=1095$ days) of the input dimension.
+Initial values for the minimum lengthscales $\ell_2$ and $\ell_3$ of $f_2$ and $f_3$, respectively, are guessed to correspond to half of the period (i.e., $l_2=l_3=0.5$).
+
+\begin{table}
+\centering
+\begin{tabular}{ccccccccc}
+\arrayrulecolor{gray}\hline \\[-3mm]
+iter. & f & $\ell$ & $c$ & $m$ & $\hat{\ell}$ & $\hat{\ell} + 0.01 \geq \ell$ & RMSE & ELPD \\ 
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+   & $f_1$ & 0.52 & 1.20 & 7  & 0.53 & FALSE &      &  	   \\
+ 1 & $f_2$ & 0.50 & -    & 8  & 0.34 & FALSE & 0.32 & 0.05 \\
+   & $f_3$ & 0.50 & -    & 8  & 1.62 & TRUE  &      &  	   \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+   & $f_1$ & 0.50 & 1.20 & 8  & 0.54 & TRUE  &      &  	   \\
+ 2 & $f_2$ & 0.34 & -    & 11 & 0.29 & FALSE & 0.32 & 0.06 \\
+   & $f_3$ & 0.29 & -    & 13 & 1.53 & TRUE  &      &  	   \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+   & $f_1$ & 0.28 & 1.20 & 13 & 0.33 & TRUE  &      &  	   \\
+ 3 & $f_2$ & 0.29 & -    & 13 & 0.24 & FALSE & 0.30 & 0.12 \\
+   & $f_3$ & 0.29 & -    & 13 & 1.64 & TRUE  &      &  	   \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+   & $f_1$ & 0.28 & 1.20 & 18 & 0.23 & TRUE  &      &  	   \\
+ 4 & $f_2$ & 0.24 & -    & 16 & 0.23 & TRUE  & 0.30 & 0.15 \\
+   & $f_3$ & 0.29 & -    & 13 & 1.68 & TRUE  &      &  	   \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+   & $f_1$ & 0.16 & 1.20 & 23 & 0.20 & TRUE  &      &  	   \\
+ 5 & $f_2$ & 0.24 & -    & 16 & 0.23 & TRUE  & 0.29 & 0.18 \\
+   & $f_3$ & 0.29 & -    & 13 & 1.63 & TRUE  &      &  	   \\[1mm]
+\arrayrulecolor{gray}\hline
+\end{tabular}
+\caption{Iterative steps in HSGP model fitting of \textit{Birthday data} case study.
+The diagnosis procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three underlying functions $f_1$, $f_2$ and $f_3$.
+Notice that $f_2$ and $f_3$ do not have boundary factor $c$ as they use periodic covariance functions.
+RMSE and ELPD are evaluated for the model underlying function $\mu=f_1 + f_2 + f_3 + f_4$ and sample data.
+The full diagnosis process is applied until two \textit{trues} are achieved for each function.}
+  \label{tab_caseI}
+\end{table}
+
+%We use $m=30$ basis functions and a boundary factor $c=1.5$.
 The length-scale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum length-scale reported by Figure \ref{fig6_7_relationships} as a function of $m$ and $c$.
 This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
-
-The year effects $f_2(t)$ and week effects $f_3(t)$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper.
-However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{sec_periodic}).
-Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in eq.~(\ref{eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in eq.~(\ref{eq_q_2}).
-We use $J=10$ cosine terms.
+%
+%We use $J=10$ cosine terms.
 The length-scale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum length-scale reported by Figure \ref{figB1_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
 
-Figure~\ref{fig15_posteriors_birthday} shows the posterior means of the long-term trend $f_1(t)$ and yearly pattern $f_2(t)$ for the whole period, jointly with the observed data.
+Figure~\ref{fig15_posteriors_birthday} shows the posterior means of the long-term trend $f_1$ and yearly pattern $f_2$ for the whole period, jointly with the observed data.
 Figure \ref{fig16_posteriors_oneyear_birthday} shows the model for one year (1972) only.
-In this figure, the special days effects $f_4(t)$ in the year can be clearly represented.
-The posterior means of the the function $\mu(t)$ and the components $f_1(t)$ (long-term trend) and $f_2(t)$ (year pattern) are also plotted in this Figure \ref{fig16_posteriors_oneyear_birthday}.
-Figure \ref{fig17_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3(t)$ can be clearly represented.
-The mean of the the function $\mu(t)$ and components $f_1(t)$ (long-term trend), $f_2(t)$ (year pattern) and $f_4(t)$ (special-days effects) are also plotted in this Figure \ref{fig17_posteriors_onemonth_birthday}.
+In this figure, the special days effects $f_4$ in the year can be clearly represented.
+The posterior means of the the function $\mu$ and the components $f_1$ (long-term trend) and $f_2$ (year pattern) are also plotted in this Figure \ref{fig16_posteriors_oneyear_birthday}.
+Figure \ref{fig17_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3$ can be clearly represented.
+The mean of the the function $\mu$ and components $f_1$ (long-term trend), $f_2$ (year pattern) and $f_4$ (special-days effects) are also plotted in this Figure \ref{fig17_posteriors_onemonth_birthday}.
 
@@
 
-\begin{figure*}
-\centering
-\subfigure{\includegraphics[width=0.85\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig15_posteriors_birthday.pdf}}
-\caption{Posterior means of the long-term trend ($f_1(\cdot)$) and year effects pattern ($f_2(\cdot)$) for the whole series. }
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig15_posteriors_birthday.png}}
+\caption{Posterior means of the long-term trend $f_1$ and year effects pattern ($f_2$) for the whole series. }
   \label{fig15_posteriors_birthday}
-\end{figure*}
-
-\begin{figure*}[!t]
-\centering
-\subfigure{\includegraphics[width=0.85\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig16_posteriors_oneyear_birthday.pdf}}
-\caption{Posterior means of the function $\mu(\cdot)$ for the year 1972 of the series.
-The special days effects pattern ($f_4(\cdot)$) in the year is also represented, as well as the long-term trend ($f_1(\cdot)$) and year effects pattern ($f_2(\cdot)$).
+\end{figure}
+
+\begin{figure}%[!t]
+\centering
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig16_posteriors_oneyear_birthday.png}}
+\caption{Posterior means of the function $\mu$ for the year 1972 of the series.
+The special days effects pattern $f_4$ in the year is also represented, as well as the long-term trend $f_1$ and year effects pattern $f_2$.
 }
   \label{fig16_posteriors_oneyear_birthday}
-\end{figure*}
-
-\begin{figure*}[!h]
-\centering
-\subfigure{\includegraphics[width=0.85\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig17_posteriors_onemonth_birthday.pdf}}
-\caption{Posterior means of the function $\mu(\cdot)$ for the month of January of 1972.
-The week effects pattern ($f_3(\cdot)$) in the month is also represented, as well as the long-term trend ($f_1(\cdot)$), year effects pattern ($f_2(\cdot)$) and special days effects pattern ($f_4(\cdot)$).
+\end{figure}
+
+\begin{figure}%[!h]
+\centering
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig17_posteriors_onemonth_birthday.png}}
+\caption{Posterior means of the function $\mu$ for the month of January of 1972.
+The week effects pattern $f_3$ in the month is also represented, as well as the long-term trend $f_1$, year effects pattern $f_2$ and special days effects pattern $f_4$.
 }
   \label{fig17_posteriors_onemonth_birthday}
-\end{figure*}
+\end{figure}
 
@@
 In this section, we first give a brief presentation of the results by \citet{solin2014explicit}, who obtain an approximate linear representation of a periodic squared exponential covariance function based on expanding the periodic covariance function into a series of stochastic resonators.
-Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic squared exponential covariance function.
+Secondly, we analyze the accuracy of this approximation and, finally, we derive the GP model with this approximate periodic square exponential covariance function.
 
@@
 \end{equation} 
-for $j=1,2,\ldots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
+for $j=1,2,\cdots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind.
 This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required.
@@
 $\beta_j$, which makes posterior inference easier.
+
+\section{Description of the iterative steps applied to perform diagnostic on two of the data sets in Figure~\ref{tab1_diagnostic}}\label{sec_ex_iterative_steps}
+
+As concrete examples, the iterative steps applied to perform diagnostic on two of the data sets in Table~\ref{tab1_diagnostic} are described next:
+
+\paragraph*{Data set with the squared exponential kernel and true length-scale $\ell_{\text{GP}} = 0.08$:} 
+\begin{itemize}
+\item[] Iteration 1:
+	\begin{enumerate}
+	\item Make the initial guess that the length-scale $\ell_1=0.5$.
+
+	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	TODO: name the value of $c_1$ and all the other quantities in the following. Otherwise, what is the purpose of this example?
+
+	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
+
+	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
+
+	TODO: provide at least a mean value for the posterior of  $\hat{\ell}_1$ here?
+
+	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives FALSE.
+
+	TODO: I think you wanted to use $\hat{\ell}_1 - 0.01 \geq \ell_1$ and I corrected it above. Can you verify whether you want + or -?
+	\end{enumerate}
+
+\item[] Iteration 2:
+	\begin{enumerate}
+	\item As the diagnostic in iteration 1 was FALSE, set $\ell_2 = \hat{\ell}_1$.
+
+	\item Compute the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.
+
+	TODO: why make this a single step here when it was two steps in Section 4.4.1?
+
+	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+
+	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives FALSE.
+	\end{enumerate}
+
+\item[] Iteration 3:
+	\begin{enumerate}
+	\item As the diagnostic in previous iteration 2 was FALSE, updating $\ell_3$ with $\hat{\ell}_2$ and repeat the process as in iteration 2.
+
+	\item The diagnostic $\hat{\ell}_3 + 0.01 \geq \ell_3$ gives TRUE.
+	\end{enumerate}
+
+\item[] Iteration 4:
+	\begin{enumerate}
+	\item As the diagnostic in step 4 in previous iteration 3 was TRUE, set $m_4 = m_3 + 5$.
+
+	\item Compute the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Fit the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.
+
+	\item The diagnostic $\hat{\ell}_4 + 0.01 \geq \ell_4$ gives TRUE.
+
+	\item As RMSE, $R^2$, and ELPD are stable relative to previous iteration 4, the HSGP approximation is likely sufficiently accurate and the procedure ends here.
+	\end{enumerate}
+
+\end{itemize}
+
+\paragraph*{Data set with the squared exponential kernel and true length-scale $\ell_{\text{GP}} = 1.4$:} 
+\begin{itemize}
+\item[] Iteration 1:
+	\begin{enumerate}
+	\item Make the initial guess that the length-scale $\ell_1=1$.
+
+	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
+
+	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
+
+	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives TRUE.
+	\end{enumerate}
+
+\item[] Iteration 2:
+	\begin{enumerate}
+	\item As the diagnostic in previous iteration 1 was TRUE, set $m_2 = m_1 + 5$.
+
+	\item Compute the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).
+
+	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+
+	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives TRUE.
+
+	\item As RMSE, $R^2$, and ELPD are stable relative to previous iteration 2, the HSGP approximation is likely sufficiently accurate and the procedure ends here.
+	\end{enumerate}
+
+\end{itemize}
 

### 2022-02-24 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex

@@
 
-A GP model with a periodic covariance function does not fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation.
-In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic square exponential covariance function.
+A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation.
+In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic squared exponential covariance function.
 
@@
 %
-These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$).
+These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \to \infty$).
 Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions.
@@
 
-As stated above, equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be closely approximated given $m$ and $c$.
+As stated above, equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be closely accurately given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
-As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data.
+As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of estimating the latent function from the data.
 Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.
@@
 
-\item[A1.] Make an initial guess on the lengthscale $\ell_1$.
-If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \in [0.5, 1]$.
+\item[A1.] Make an initial guess on the length-scale $\ell_1$.
+If there is no useful information available, we recommend to start with a large length-scale, about $\ell_1 \in [0.5, 1]$.
 Set the iteration index to $i=1$.
 
-\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.
+\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
 
@@
 obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness.
-Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.
-(\ref{eq_c_vs_l_QE}).
-Figure \ref{fig7_lscale_comparison} right shows the RMSE of the HSGP models with the exact GP model as the reference.
+Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.~(\ref{eq_c_vs_l_QE}).
+Figure~\ref{fig7_lscale_comparison} (right) shows the RMSE of the HSGP models with the exact GP model as the reference.
 It can be seen that accurate estimates of the length-scale imply small RMSEs.

### 2022-02-25 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edited 4.5 and 4.8

@@
 
-As stated above, equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be closely accurately given $m$ and $c$.
+Equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be accurately inferred given $m$ and $c$.
 This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.
@@
 
-Having obtained an estimate $\hat{\ell}$ of a given $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided as a function of $m$ and $c$ by equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used).
-If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good, or at least close to.
-If, however, $\hat{\ell}$ does not exceed the recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated.
-If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also match those obtained from equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
+Having obtained an estimate $\hat{\ell}$ for a HSGP model with prespecified $m$ and $c$, we can check whether $\hat{\ell}$ exceeds the smallest length-scale, that can be accurately inferred, provided as a function of $m$ and $c$ by equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used).
+If $\hat{\ell}$ exceeds this value, the approximation is assumed to be good.
+If $\hat{\ell}$ does not exceed this value, the approximation may be inaccurate, and $m$ and/or $c$ need to be increased.
+In Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, $m = 10$ and $c = 1.5$ were sufficient for an accurate modeling of function with $\ell = 0.3$, which matches the diagnostic based on equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
 
@@
 \begin{itemize}
-	\item $c$ must be big enough for the given length-scale $\ell$ of the function to be learned and,
-	\item $m$ must be big enough for the given $\ell$ and $c$. 
+	\item $c$ must be big enough for a given $\ell$, and
+	\item $m$ must be big enough for given $\ell$ and $c$. 
 \end{itemize}
 %
-If larger than minimal $c$ and $m$ for the given length-scale $\ell$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
-However, as the given length-scale $\ell$ of the function to be learned is unknown, we recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using the equations in Section~\ref{sec_num_equations}; and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$, which indicates a potential problem with the approximation, update $\ell$ with $\hat{\ell}$ and repeat the process until both:
+If larger than minimal $c$ and $m$ (for a given $\ell$) are used in the initial HSGP model, it is likely that the results are already sufficiently accurate.
+As $\ell$ is initially unknown, we recommend using this diagnostic in an iterative procedure by starting with $c$ and $m$ based on some initial guess about $\ell$, and if the estimated $\hat{\ell}$ is below the diagnostic threshold, select new $c$ and $m$ using $\hat{\ell}$.
+This can be repeated until
 %
 \begin{itemize}
-\item the estimated $\hat{\ell}$ is larger than the minimum lengthscale determined by $c$ and $m$, which indicates the approximation must be sufficiently accurate, and
-\item the measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD) reach a stabilization.
+\item the estimated $\hat{\ell}$ is larger than the diagnostic threshold given $c$ and $m$, and
+\item the predictive accuracy measures, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD) don't improve.
 \end{itemize}
 
-As commented in next Section~\ref{sec_diagnostic_criticism}, the estimated $\hat{\ell}$ being larger than the minimum lengthscale determined by $c$ and $m$ is not a definite certain that the approximate is sufficiently accurate, although it is for most of the practical cases.
-Therefore, we recommend to look at the residual error of the fit until it stabilizes.
-
-Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to minimizing the computational cost.
-
+As commented in next Section~\ref{sec_diagnostic_criticism}, the estimated $\hat{\ell}$ being larger than diagnostic threshold doesn't guarantee that the approximate is sufficiently accurate, and thus we recommend to look at the predicitve accuracy measures, too.
+
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also provide the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that van be used to minimize the computational cost in repeated computations (e.g., in cross-validation and simulation based calibration)
 %If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
-The iterative procedure, briefly commented above and detailed in next Section~\ref{sec_user_guide}, works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed.
-Using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential for repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional functions with very large data sets, and even more essential in multi-dimensional cases ($D \leq 4$; notice from Section~\ref{sec_method_multi} that for $D > 4$ HSGPs start to be unpractical in any case) where knowing the optimum values for $m$ and $c$ for each individual dimension can significantly reduce the computational requirements in these multi-dimensional cases; otherwise fitting HSGPs would be unfeasible here.
+%The iterative procedure, briefly commented above and detailed in next Section~\ref{sec_user_guide}, works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed.
+%Using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential for repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional functions with very large data sets, and
+This even more useful in multi-dimensional cases ($2 \leq D \leq 4$), where knowing the smallest useful value of $m$ for each dimension has even bigger effect on the total computational cost.
 
@@
 
-\subsubsection{"Criticism" of the diagnostic} \label{sec_diagnostic_criticism}
-
-The basis of our diagnostic tool are the following:
-% 
-\begin{enumerate}
-\item observing a length-scale estimate $\hat{\ell}$ smaller than recommended minimum length-scale determined by $m$ and $c$ indicates a potential problem with the approximation,
-\item observing a $\hat{\ell}$ larger than the recommended minimum length scale implies that the approximation is accurate.
-\end{enumerate}
-
-The later statement, however, cannot be assumed to be true in general, that is, cannot be generalized to all possible cases.
+% \subsubsection{Criticism of the diagnostic} \label{sec_diagnostic_criticism}
+
+% The basis of our diagnostic tool are the following:
+% % 
+% \begin{enumerate}
+% \item observing a length-scale estimate $\hat{\ell}$ smaller than recommended minimum length-scale determined by $m$ and $c$ indicates a potential problem with the approximation,
+% \item observing a $\hat{\ell}$ larger than the recommended minimum length scale implies that the approximation is accurate.
+% \end{enumerate}
+
+% The later statement, however, cannot be assumed to be true in general, that is, cannot be generalized to all possible cases.
 ({\color{red} TODO: A major revision from a reviewer was related to that, i.e.
@@
 
-For that reason, to ultimately diagnose whether the approximation is accurate, in addition to observe whether $\hat{\ell}$ exceeds the recommended minimum length scale, we recommend ensuring the measures of predictive accuracy reach an stabilization too.
+% For that reason, to ultimately diagnose whether the approximation is accurate, in addition to observe whether $\hat{\ell}$ exceeds the recommended minimum length scale, we recommend ensuring the measures of predictive accuracy reach an stabilization too.
 
@@
 
-\subsection{On computational requirements in multi-dimensional cases}
+\subsection{On computational cost in multi-dimensional cases}
 
@@
 The computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-Even though, the computation time increases significantly with $D$ in bivariate input spaces, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
-For moderate sized datasets, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
-However, the HSGP method will be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
-In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
-
-Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
+In the experiments of this paper computation for HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
+For small and moderate sized datasets, HSGPs is likely to be slower than exact GPs already for $D=3$ and $D=4$ even with a relatively low number of basis functions ($m \gtrsim 5$ and $m \gtrsim 20$ respectively; see Figure~C.3 in the online material).
+The HSGP method can be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
+In all of the investigated cases, choosing the optimal booundary factor $c$ reduces the required number of basis functions $m$ noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore  reduces computational time drastically in particular in multidimensional cases.
+When HSGP or exact GP is faster depend also on the implementation details that can have big effect on the constants, and thus any more detailed recommendations can only be made in the context of the specific software implementation.
+
+Roughly similar or even worse behavior was found for thin plate regression splines used in the experiments, with serious challenges with computation time when doing inference for spline models for $D=3$ and with more than 10 basis functions, or even for $D=2$ and more than 40 basis functions (see Figures~B.3-right and C.3 in the online material).
 

### 2022-02-28 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
fix typos

@@
 Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
-Other possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation are actually explaining.
+Another possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation they are actually explaining.
 

### 2022-02-28 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
add reply_letter.tex

\documentclass[11pt]{report}
\usepackage{blindtext}
\usepackage{titlesec}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage{microtype}
\usepackage{lscape}

\usepackage[textwidth=17cm, textheight=21.2cm]{geometry}

\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}

\usepackage{epstopdf}

\usepackage[sectionbib,numbers,sort&compress]{natbib}

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
%\graphicspath{{./images/}}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{bm}

\usepackage{appendix}

\bibpunct[, ]{[}{]}{,}{}{,}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}

\setlength{\parskip}{0.5em}
\renewcommand{\baselinestretch}{1.1}

\renewcommand{\bibname}{References}

\usepackage{dashrule}

\usepackage{bibentry}
\bibliographystyle{apalike}

\makeatletter 
\renewcommand\BR@b@bibitem[2][]{\BR@bibitem[#1]{#2}\BR@c@bibitem{#2}}           
\makeatother

\nobibliography*

\begin{document}

We would like to thank the reviewers for their thorough comments and valuable feedback. First, a summary with the improvements, changes and new insights achieved from the review are outlined. Then, all comments are addressed and there is a separate reply to each reviewer.

Authors would like the summary of the updates and insights presented above to be sent to all of the reviewers.

\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

\subsection*{Summary of the updates and insights}

\begin{itemize}

\item We have improved the paper significantly, including new insights from new experiments and analysis. The readability of Sections 4 and 5 have also been improved.

\item A brief discussion about the theoretical evidence of the near linear proportionality between the setting parameters of the model has been added (Section 4.2).

\item In the first submission of the paper, the relationships between the setting parameters of the model were analyzed by simulation experiments and presented in form of graphs.

\item We have derived numerical equations that characterize the relationships between the setting parameters of the model, which are useful to automatize updating the parameters in the fitting process (Section 4.3.1).

\item The diagnostic rule that allowed to check whether the approximation was sufficiently accurate has been further discussed regarding its reliability and applicability (Sections 4.5).

\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1), which is based on:

\begin{enumerate}
\item Recommendations to set initial values for the parameters.
\item Safety diagnostic rules to check if the approximation is sufficiently accurate (Sections 4.5).
\item Numerical equations to update and find out the optimal values for the parameters (Sections 4.3.1).
\end{enumerate}

\item The proposed diagnostic procedure is a safety, general and extensive procedure that allows us to automatize the fitting process.

\item A thorough analysis of the performance of the diagnostic tool has been carried out (Section 4.6).

\item The diagnostic tool has been extended to add the analysis for the Mat\'ern-5/2 kernel. So the whole study finally provides the diagnostics for a squared exponential, Mat\'ern-3/2, Mat\'ern-5/2 (Sections 4.3 and 4.3.1), and periodic squared exponential (Sections 3.3, 4.4 and Appendix C) covariance functions.

\item By analyzing the constants characterizing the relationships for each covariance function, recommendations to perform diagnosis on any other covariance function different from those implemented in this study have been provided (Sections 4.2 and 4.7).

\item Explanations of how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).

\item In the first submission of the paper, a linear representation for a periodic squared exponential kernel and the corresponding diagnosis analysis were presented in Appendix B.

\item We have derived the numerical equations governing the relationships between the parameters of this linear representation of the periodic kernel. These equations have been included in the main text in Section 4.4.

\item We have applied the diagnostic tool on cases studies in Section 5. In that way, case studies in Section 5 have become more focused on the main goals of the paper, which are to facilitate the implementation of the method for practical users providing clear formulation of the method and provide a safety and general diagnostic tool to fit and diagnose the model.

\item We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set. We have applied the diagnostic procedure on it. Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing individually the different dimensions in each iteration. It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.

We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material.

{\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}

\item We have added Section 4.8 where a discussion of the performance of the model in multidimensional cases are presented.

\item We have emphasized in the paper that applying the diagnostic tool allows for minimum computational requirements that can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting such as cross-validation and simulation based calibration (Section 4.8).

\item A potential fitting overestimating effect when using too many basis functions relative to the wigglyness of the function to be learned has been commented. {\color{red}

(TODO: Currently we just have outlined but not discussed this. If we had time, it would be nice to discover why this overestimating effect is happening.)}

\item We expect this paper having a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios like uni-dimensional, multi(low)-dimensional functions and other problems that require repeated fitting.

\end{itemize}

\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

\subsection*{Reviewer \#1}

\subsubsection*{R\#1 Comment \#1}

Reviewer \#1 comments "The authors correctly identify that for practical use in probabilistic programming environments the method has to be reliable and automatic enough, justifying the more detailed analysis of its parameters to identify when the approximation can be trusted. In addition, they provide the required implementations for one PP environment (Stan), although this is done at the level of case studies rather than integrating the approximation more tightly as part of a probabilistic programming environment."

The Hilbert space approximation for GPs (HSGP) is presented as a general formulation for probabilistic modeling, so it can be implemented in different probabilistic programming frameworks, such as Stan, WinBugs, etc. The case study's models have also been formulated in general probabilistic language in the main manuscript, and Stan's implementations have also been provided as examples. Furthermore, HSGP model has been implemented in the \textit{brms} R-package that uses Stan to make model inference.

The diagnostic tool which consists in iterative steps to fit and diagnose the model has also been implemented in the \textit{brms} R-package. 
%This diagnostic iterative procedure is fundamentally based on a diagnostic rule to check whether the approximation is accurate and numerical equations allowing updating the key parameters of the method.

\subsubsection*{R\#1 Comment \#2}

Reviewer \#1 comments "The analysis is reasonable and provides tangible results on how the approximation scheme should be used, but does not go very deep. There are not new theoretical results but instead the analysis is purely empirical and in fact a standard study where each approximation parameter is varied at a time to identify the limits when the approximation starts to fail. The analysis is properly carried out and results in a practical diagnostic, but did not require notable scientific insight."

The performance of the HSGP model drastically depends on some setting parameters. The study pursues accurately measuring the sensitivity of the HSGP model to the values of these parameters, detecting when the approximation fails and when the parameter values are optimal in accuracy and computation cost. Ultimately, the aim of the study is to provide a diagnostic to check whether the approximation is sufficiently accurate, provide numerical equations to update the setting parameters if the approximation is not sufficiently accurate, and finally provide a safety and general procedure based on iterative steps to successfully fit the model, finding out optimal values for the setting parameters which allow for minimum computational requirements.

A theoretical approach to the diagnostic equations and rules is too complex that makes it not practical. The way to approach this is empirically. Making the empirical analysis we fully recognize the behavior of the approximation according to the parameters values, derive numerical equations gathering the relationships between the parameters that allow for updating the parameters and diagnosing the fitting.

Furthermore, by following our recommended diagnostic procedure, an optimum solution with minimum computational requirements is achieved, that can be essential in repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional cases with very large data sets, multi(low)-dimensional cases, even with small data sets, since knowing the optimum values for the setting parameters for each individual dimension can significantly reduce the computational requirements, otherwise fitting HSGPs would be unfeasible here.

\subsubsection*{R\#1 Comment \#3}

Reviewer \#1 comments "The results are intuitive, but the process of determining the parameters in practice remains somewhat cumbersome as one needs to inspect the curves in Figures 6 and 7 to make the decision and there are no definite guidelines on how to modify the parameters when observing insufficient accuracy (e.g. no hints on how much one should increase m)."

We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate. We have also elaborated a safety and general procedure based on iterative steps to fit the model with definite rules to update the parameters and make diagnosis.

\subsubsection*{R\#1 Comment \#4}

Reviewer \#1 comments "Furthermore, to use the approximation with any other kernel besides square exponential or Matérn one would essentially need to repeat the same analysis from scratch. This limits the value in probabilistic programming context. The authors indicate existing GP  libraries are limited and favour Stan for its generality, but since we now only have support for two specific kernels one could easily argue this approach is limited as well."

We have implemented the HSGP model and the diagnostic tool for four different kernels, the squared exponential, Matern-3/2, Matern-5/2, and periodic squared exponential kernels, which cover a broad range of learning functions in practical applications. Kernel combinations that satisfy the requirements can also be used. We have also given recommendations in Section 4.2 and 4.7 to update $m$ and $c$ and perform diagnosis on any other different stationary kernel from those implemented in our study.

The study deeply describes in Section 3 the formulation and implementation of the HSGP model, description that can be used to implement it for any other different stationary kernel. The only new thing that the user needs to compute is the Fourier transform of the kernel to obtain the spectral density of the kernel.

\subsubsection*{R\#1 Comment \#5}

Reviewer \#1 comments "it still feels like many educated practitioners would have found similar values for m and c by following a conventional validation procedure of applying the approximation with range of values for both on some example functions. While they would probably have ended up using slightly too large m or might spend a bit more computation while finding the best values, the practical difference might not be that big."

Speculation from the reviewers side; we do not really know the expertise of our users.

Rerunning takes computer and human time and we reduce that time by reducing the number of iterations and the computational requirements for each iteration.

Not all practitioners may be educated enough to come up with and adjust procedures themselves. So clear, justified guidelines (that we provide) will be helpful and even required for the method to be used for many users.

Apart from the number of basis function, the boundary factor $c$ plays a crucial role here and has to be set carefully as a function of the wigglyness (true length-scale) of the function to be learned. By following our diagnostic rules and equations, both $m$ and $c$ can be set and properly updated with minimal effort.

An optimum solution with minimum computational requirements can be achieved following our recommended diagnostic procedure, otherwise it would be difficult. As discussed in the paper and above in this letter, using optimal values for the key parameters that allow for minimum computational requirements can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting.

\subsubsection*{R\#1 Comment \#6}

Reviewer \#1 comments "The case studies are interesting and demonstrate the approximation scheme works in practice for simple statistical models while providing clear example codes, but they could be more focused towards the main goal of the paper."

The developed diagnostic tool and iterative steps to fit and perform diagnosis on the HSGP model have been made prominent in the case studies.

The HSGP model is a new and computationally efficient representation of a GP. GPs are not simple statistical models. Improving the computational efficiency of a GP favors its implementation as modular components in flexible and complex modeling. In this sense, case study II and IV are examples of the use of the model in these flexible and complex modeling scenarios.

\begin{itemize}
\item Case study II involves a function that is composed of three additive GP components with different kernels, squared exponential and periodic kernels. Additionally this case study has a large data set of more than 7000 observations which would be unfeasible to be fitted using the exact GP model.

\item Case study IV (Leukemia data set) is a 4D function with a multilevel GP structure. This case study is also a good example that by using our diagnostic tool you can easily get an optimal solution in 4D which makes it significantly faster than regular GPs. Otherwise, user would had probably been struggling searching for useful values for the parameters without any guide.
\end{itemize}

The only limitation of the method is the number of dimensions as these affect computation through the total number of the basis functions needed. But this does not necessarily means a statistical model being simple.
In fact due to this high dependency of the computational requirements of the method on the number of basis functions, the analysis and diagnosis tool developed in this paper arises to be essential for practical applications and in broad scenarios.

\begin{itemize}
\item Case study I involves a wiggly function with a Mat\'ern-3/2 kernel.

\item Case study III involves a two-dimensional function on which we use the diagnostic tool to fit the model. We show that, although the diagnostic tool was built for unidimensional covariance functions, it can be successfully used for multi-dimensional cases by diagnosing individually the different dimensions in each iteration. It allows gradually finding the optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.

Taking advantage of this case study we present additional insights about the performance of the method in 3D and 4D cases from other examples initially included in online material.

{\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}
\end{itemize}

\subsubsection*{R\#1 Comment \#7}

Reviewer \#1 comments "Section 5.1 takes approximately four full pages to present results for a simple artificial data. It describes one concrete example of setting the parameters (that remains somewhat vague -- "In the first iteration, we choose...", "In the second iteration, m is increased to..."), but the detailed comparisons against splines could have been much more brief in a paper focusing on how to make the approximation easier to use in practice."

We have simplified the section significantly but keeping references to equations in the main text of how to implement the approximation function. We have used the diagnosis tool and iterative steps to fit and perform diagnosis on the models. We have also improved the readability of the section.

\subsubsection*{R\#1 Comment \#8}

Reviewer \#1 comments "Section 5.2 is similarly a bit besides the point -- the example model is interesting but regarding the choice of the parameters you simply say "We use m=30 basis functions and a boundary factor c=1.5" followed by stating that the length-scale estimate is good. This is what we would expect to see in a paper that later uses the method, but for a paper that is introducing the diagnostic I would expect more details and e.g. explanations of what would happen if the parameters were poorly chosen. I understand fully that verbally describing an iterative process is difficult, but was still a bit let down by these case studies."

This case study involves an underlying function that is composed of three additive GP functions with different kernels, squared exponential and a periodic kernels. We have applied the diagnostic tool, describing the iterative steps to fit this model and performing diagnosis on the three additive GP components at the same time. This example also serves as an example of how useful our diagnosis can be in large 1D data sets.

\subsubsection*{R\#1 Comment \#9}

Reviewer \#1 comments "Section 5.3 presents a multivariate case and seems to mostly provide a negative result on applicability of HSGP for multivariate cases. This is fine as the method is valuable even if limited to low dimensions, but it would have been more interesting to see a more detailed analysis of the intermediate region, especially D=2 and D=3 that are frequent in spatial and spatiotemporal analysis problems -- a practitioner would appreciate clear guidelines on when to use HSGP and how to recognise when it is not a good solution."

It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown. In fact, for that reason the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration. Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.

We disagree with the reviewer that the HSGP performs poorer than the exact GP on this Leukemia data set. In fact using the diagnostic tool the user will discover easily the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP. Without the guide of the diagnostic tool the user probably would had made several trials until finding out an accurate and practical solution, consuming a lot of requirements in this demanding 4D case.

The statement "its applicability to multidimensional cases is mostly negative" is inaccurate, as it ultimately depends on the wigglyness of the function.

As commented across the paper and in particular in Section 4.8, the method is limited in practice to the number of dimensions and the wigglyness of the function across the different dimensions. We have observed that it can be useful for $D<5$ conditioned to the wigglyness of the function. However, for $D=2$ it is still useful and significantly faster than the exact GP even for very wiggly functions. In Section 4.8 we give brief details on the performance of the model in multi-D cases found from various case studies carried out by the authors and initially presented in the online material. In this sense, the use of our recommended diagnostic allows to find the optimal values for $c$ and $m$ which allows for minimum computational requirements which can be essential on these multidimensional cases.

We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set. We have applied the diagnostic procedure on it. Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing individually the different dimensions in each iteration. It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases. We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material. {\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}

\subsubsection*{R\#1 Comment \#10}

Reviewer \#1 comments "Overall, I believe the paper has clear value for practical use and is worth publishing in some form: It provides nice code examples for HSGP via case studies and an iterative procedure for diagnosing when the approximation accuracy is good enough. However, the scientific novelty of the work is fairly limited and perhaps insufficient for publication in a top venue, in particular because there are no claims or contributions that would extend beyond this specific approximation scheme. There are also no clear scientific insights in the process itself, since the analysis is purely empirical and follows a straightforward protocol. In summary, the manuscript is a borderline case even if addressing the remarks below."

This work is necessary to facilitate the implementation of the method for practical users, while Solin and Sarka (2020) fundamentally focus on the mathematical development of the model. We also analyze how the key factors of the model relate each other for accurate approximations. We derive numerical functions for these relationships that allow to update the parameters, perform diagnosis, and provide clear and useful guidelines for the whole procedure. So, we make it manageable for practical users, otherwise the practical applicability of the method would be highly limited.

We expect this paper to have a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios like uni-dimensional functions, multi-dimensional functions, and other problems that require repeated fitting such as cross-validation and simulation based calibration.

We do not see a problem in a purely simulative approach. It simply has different advantages and disadvantages over analytic approaches. In fact, as mentioned above, this study can only be approached from an empirical point of view. We would disagree on the scientific novelty (not all scientific novelty is mathematical in this regard). And we agree with the reviewer about the practical usefulness, which was our goal.

\subsubsection*{R\#1 Major revisions \#1}

\#1.1 - Reviewer \#1 comments: "The diagnostic involves comparing the estimated length-scale $\hat l$ against a minimum $l^*$ determined by m and c. I have trouble understanding the full validity of this process. I agree with your statement on page 8 about $\hat l$ being accurate if the approximation is accurate, but you do not state anything definite about the quality of $\hat l$ when the approximation is inaccurate. However, the process seems to implicitly assume we can still rely on $\hat l$ even then, with very relaxed justification ("If $\hat l$ does not exceed ... the approximation *may* be inaccurate"). Am I correct that your method would fail to diagnose a bad approximation if $\hat l$ was overestimated? I would like to see this discussion extended, at least by explicitly mentioning possible failure modes."

An inaccurate approximation implies that number of basis functions $m$ is not enough and/or the boundary factor $c$ is not adequate.

We can not state anything definite about the estimated lengthscale $\hat{\ell}$ when the approximation is inaccurate, basically because it might ultimately depend on the specific frequency patterns of the signal/function to be learned, and also because $c$ must have been properly set. Although, we can state that for most of the cases, if $c$ is large enough, $\hat{\ell}$ will be below that minimum supported length-scale ($\ell^*$) determined by $m$ and $c$. But, as said before, it can not be generalized to all case studies because it might depend on the particular frequency patterns of a signal. Future research will be focused on investigating potential signals with particular frequency patterns that could escape from this rule.

What we can definitely state, which is in fact what our diagnosis is based on, is the following:

\begin{enumerate}
\item  Observing an $\hat{\ell}$ smaller than $\ell^*$ (i.e. $\hat{\ell}$ is overestimated), and $c$ is large enough given $\ell^*$, indicates a potential problem with the approximation. This happens because in inaccurate approximation scenarios the true length-scale is smaller than $\ell^*$, so $\hat{\ell}$ will tend to be towards smaller values than $\ell^*$.

\item The other way around, that is, observing $\hat{\ell}$ larger than $\ell^*$ (i.e. $\hat{\ell}$ is underestimated), the approximation is assumed to be accurate because when the true length-scale is larger than $\ell^*$, $\hat{\ell}$ will tend to be larger than $\ell^*$. However, we do not still have full evidence that this can be generalized to all possible cases as explained above (but, we can state that this rule can be apply for most of the the practical cases). For that reason, in order to make the diagnostic fully safety we recommend re-running the model, updating parameters, until seeing predictive accuracy measures, such as RMSE, $R^2$ and ELPD, do not improve.

As largely demonstrated in the paper in Section 4.6 and Case studies, these diagnostic rules leads to a safety, useful and fast iterative procedure to fit the models with minimum computational requirements.

These arguments have been included in the paper in Section 4.5. 
\end{enumerate}

\noindent \#1.2 - Reviewer \#1 comments: "To me it seems Fig 5 (left) already hints that you tend to underestimate l when using bad approximation and the procedure may be valid in practice, but this is not transparent in the paper."

For all cases showing inaccurate approximation in Fig. 5 (left), $\hat{\ell}$ is below $\ell^*$, as long as $c$ is large enough given $\ell^*$. It can be checked by computing $\ell^*$ for each case in Fig. 5 (left) by using equation (18). Following we show a few of these inaccurate cases in Fig. 5 (left), showing diagnostic $\hat{\ell} < \ell^*$ can be trusted to diagnose a bad approximation, as long as $c$ is large enough given $\ell^*$:

\begin{table}[h]
\centering
\begin{tabular}{ c c c c c c }
$m$ & $c$ & $\ell^*$ & $\hat{\ell}$ & \\ 
\hline \\[-4mm]
7  & 4   & 1.00 & $0.46$ & $\to$ & $\hat{\ell} < \ell^*$ \\
15 & 4   & 0.46 & $0.25$ & $\to$ & $\hat{\ell} < \ell^*$ \\
5  & 1.5 & 0.53 & $0.45$ & $\to$ & $\hat{\ell} < \ell^*$ \\
7  & 1.5 & 0.38 & $0.27$ & $\to$ & $\hat{\ell} < \ell^*$ \\
10 & 2.5 & 0.44 & $0.25$ & $\to$ & $\hat{\ell} < \ell^*$ \\
7  & 2   & 0.50 & $0.28$ & $\to$ & $\hat{\ell} < \ell^*$ \\
15 & 3   & 0.35 & $0.28$ & $\to$ & $\hat{\ell} < \ell^*$ \\
7  & 1.2 & 0.30 & $0.25$ & $\to$ & $\hat{\ell} < \ell^*$ \\
\end{tabular}
\end{table}

However, when $c$ is not large enough given $\ell^*$, $\hat{\ell}$ can be larger than $\ell^*$ and the approximation being inaccurate, so the diagnosis will fail in this case, for example:

\begin{table}[h]
\centering
\begin{tabular}{ c c c c c c }
$m$ & $c$ & $\ell^*$ & $\hat{\ell}$ & \\ 
\hline \\[-4mm]
20 & 1.1 & 0.10 & $0.17$ & $\to$ & $\hat{\ell} > \ell^*$ \\
10 & 1.1 & 0.19 & $0.20$ & $\to$ & $\hat{\ell} > \ell^*$ \\
\end{tabular}
\end{table}

It shows that it is essential to follow our diagnosis tool to be sure we are using valid values in each iterative step.

\noindent \#1.3 Reviewer \#1 comments: "Do we have a reason for the error to be always towards understimating the lengthscale? Could this be proven?"

See response in \#1.1 above.

\subsubsection*{R\#1 Major revisions \#2}

Reviewer \#1 comments: "The process involves checking Figure 6 (or 7) for verifying whether the approximation is valid. This feels cumbersome in practice, and for better PP integration I would expect proper software support for this. It would be nice to at least outline how this could work, but even better to provide an actual tool. Both a visual tool plotting the current estimate as in Fig 13 and a textual feedback like "$\hat l$ is too small, you should either increase m or decrease c a bit -- we recommend the former" could work. One could also make this completely automatic, a loop that keeps on modifying the parameters until the threshold is reached. I would like you to discuss which of these options you would prefer and why."

We have developed and presented a diagnostic procedure that consists in iterative steps to fit and diagnose the model. Numerical equations are used to set initial values and update the key parameters in each iterative step. The recommended diagnostic procedure is a safety and extensive procedure that allows to automatize the applicability of the method. See details in Section 4.5 and examples in Section 4.6, Case studies and Appendix C.

\subsubsection*{R\#1 Major revisions \#3}

Reviewer \#1 comments: "The paper is in general well written, but the structure could be improved. I think the specific diagnostic is now lost in the middle of the story in Sections 4 and 5 and would benefit from clear presentation in dedicated section (that would also be a good place to discuss the software support; see comment 2 above). The writing of both Sections 4 and 5 is also quite verbose and text-heavy, which makes reading the paper somewhat tedious. Section 4.1 is 200+ lines of text with no structure and as mentioned above both 5.1 and 5.2 are somewhat long compared to their value. Since you anyway provide exact code and re-use models from previous papers, you could compress the general discussion of the models and focus more on the process of determining the approximation parameters."

Section 4 have been re-structured and updated by adding new subsections, such as a specific section on the theoretical evidence of near linear proportionality among the key parameters of the model, another section with the experimentation and resulting numerical diagnostic equations, another section supporting, detailing and discussing on the diagnostic rules, and another analyzing the performance of the diagnostic tool. Additionally, new subsections regarding how to proceed with other different kernels from those implemented in this study and some insights on how the method performs in multi-dimensional cases have also been added.

We have simplified and improved significantly the writing of Section 5. We have also focused the case studies more on the main goals of the paper:
\begin{enumerate}
\item Facilitate the implementation of the method for practical users providing clear formulation of the method.
\item Provide a safety and general diagnostic tool to fit and diagnose the model.
\end{enumerate}

\subsubsection*{R\#1 Major revisions \#4}

Reviewer \#1 comments: " It would be really interesting to see a bit more about how HSGP works for functions with 2-3 inputs, and in particular about how a practitioner can recognise whether it makes sense to use HSGP for their case or not. To me it seems clear HSGP is very often a reasonable method for D=1 and almost certainly breaks down for D>=5 or so unless working with very smooth functions (and for them almost any approximation works well), but clear diagnostics for quickly determining whether it could work for D=2 or D=3 would increase the contribution notably. Now the practical process would be to try it out and keep on increasing m until it starts working or one notices the computation is too slow, which feels wasteful -- as a practitioner I would rather just skip HSGP completely if finding out it does not help takes longer than solving the problem without the approximation."

Diagnostic has been build for GPs with unidimensional covariance functions. Future research will focus on building analytical models that provide the diagnostics in multi-dimensional cases. However, as an approximation, we can use the diagnostic for unidimensional functions to make diagnosis individually on the different dimensions of a multidimensional GP model.

See comment \#9 above for an extended answer.

\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

\subsection*{Reviewer \#2}

\subsubsection*{R\#2 Comment \#1}

Reviewer \#2 comments "The paper gives experimental results for the method of Solin and Sarkka  (2020). Overall the idea such experiments is interesting, but the actual experiments are not extensive enough given the pure applied nature of this work."

From the experiments, we have built numerical functions that characterize the relationships between the key parameters of the model which are the basis to perform diagnosis on the fitting. We have developed a diagnostic procedure to fit, diagnose and update parameters, based on iterative steps until an accurate solution is reached. Furthermore, this diagnostic tool provides the optimal values for the key parameters of the model that allow for minimizing the computational cost, which can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting such as cross-validation and simulated based calibration.

The diagnostic rules, the numerical functions that allow updating the parameters and the recommended iterative steps to fit and diagnose the model are extensive to the community and essential to make the model usable for practitioners.

All of this have been properly explained along Section 4 in the paper.

\subsubsection*{R\#2 Comment \#2}

Reviewer \#2 comments "Experiments are mainly low dimensional (all 1D except a single 4D example). Please discuss high dimensional cases ; indeed equation 8 suggests poor scaling in dimension. Can it be applied to high dimensional data?".

We have mention in the paper how dimensionality affects the expected number of basis functions needed and that the effect of dimensionality is only through this.

It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown. In fact, for that reason the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration. Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.

The Leukemia data set case study is a 4D data set. We have applied applied the diagnostic tool on it and we have been able to easily discover the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP. Without the guide of the diagnostic tool the user probably would had made several trials until finding out an accurate and practical solution, consuming a lot of requirements in this demanding 4D case.

We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set. We have applied the diagnostic procedure on it. Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing individually the different dimensions in each iteration. It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases. We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material. {\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}

As commented across the paper and in particular in Section 4.8, the method is limited in practice to the number of dimensions and the wigglyness of the function across the different dimensions. We have observed that it can be useful for $D<5$ conditioned to the wigglyness of the function. However, for $D=2$ it is still useful and significantly faster than the exact GP even for very wiggly functions. In Section 4.8 we give brief details on the performance of the model in multi-D cases found from various case studies carried out by the authors and initially presented in the online material. In this sense, the use of our recommended diagnostic allows to find the optimal values for $c$ and $m$ which allows for minimum computational requirements which can be essential on these multidimensional cases.

\subsubsection*{R\#2 Comment \#3}
Reviewer \#2 comments "There is no comparison with other techniques (except a spline method) ! Please compare with, e.g. "MCMC for Variationally Sparse Gaussian Processes" ; although this is SVI, it appears directly applicable for the problems considered here. Please also compare with non MCMC variational methods - while not directly applicable to generic MCMC methods, these are the sttandard and it would be nice to see how they compare in runtime."

Aki: R2 is not happy that we compare only to splines and suggests we should compare to SVI and other variational methods like splines and VI would belong to the same category. We can use Stan's ADVI to run variational inference results. We should include a model for which usual variational inducing point approaches are not directly applicable due to non-exponential family model or combination of different components so that we can argue that we we stick to comparing to black box VI. We have recently run ADVI and improved robust ADVI results for birthday and motorcycle examples (the latter has latent GP for both location and log sigma)
and VI is not that good (it's annoying how common belief it is that black box
VI would be great).

\subsubsection*{R\#2 Comment \#4}

Reviewer \#2 comments "On the eigenfunction expansion: 1) note that this is very similar to the Mercer expansion of e.g. Walder et al, "Fast Bayesian Intensity Estimation for the Permanental Process". 2) the Nystrom method also yields an approximate eigen expansion which has pros and cons to the method of the submission. For example, the Nystrom applies to any kernel (not just stationary) and yields a linear GP approximation, but requires some overhead for the setup. Given this submission is purely experimental, it would be nice to see comparisons.".

Aki: R2 asks comparisons to Nystrom, but that comparison was
already made by Solin and Särkkä so I don't think we should add that.

%\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

%\nobibliography{references}

\end{document}

### 2022-03-01 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
minor edits and comments

@@
 
-We would like to thank the reviewers for their thorough comments and valuable feedback.
-First, a summary with the improvements, changes and new insights achieved from the review are outlined.
-Then, all comments are addressed and there is a separate reply to each reviewer.
+We would like to thank the reviewers for their thorough comments and valuable feedback, which has helped us a lot in improving the paper.
+First, a summary of the improvements, changes, and new insights achieved during the revision are outlined.
+Then, all comments are addressed in a point-by-point manner, with separate replies to each reviewer.
+Throughout, comments or the reviewers are hightlighted in bold.
 
-Authors would like the summary of the updates and insights presented above to be sent to all of the reviewers.
+%Paul: all responses will be send to all reviewers anyway
+%Authors would like the summary of the updates and insights presented above to be sent to all of the reviewers.
 
@@
 
+%Paul: This list is way too long and detailed and appears to be redundant at times. I recommend focusing on between 4 and 7 or so points but more will be too much.
+
 \item We have improved the paper significantly, including new insights from new experiments and analysis. The readability of Sections 4 and 5 have also been improved.
 
-\item A brief discussion about the theoretical evidence of the near linear proportionality between the setting parameters of the model has been added (Section 4.2).
+\item A discussion about the theoretical evidence of the near linear proportionality between the setting parameters of the model has been added (Section 4.2).
 
@@
 
+%Paul: Don't forget the thank the reviewers for their comments to show some more appreciation of their comments and time they put into the review.
+
 \subsubsection*{R\#1 Comment \#1}
 
-Reviewer \#1 comments "The authors correctly identify that for practical use in probabilistic programming environments the method has to be reliable and automatic enough, justifying the more detailed analysis of its parameters to identify when the approximation can be trusted.
-In addition, they provide the required implementations for one PP environment (Stan), although this is done at the level of case studies rather than integrating the approximation more tightly as part of a probabilistic programming environment."
+\textbf{The authors correctly identify that for practical use in probabilistic programming environments the method has to be reliable and automatic enough, justifying the more detailed analysis of its parameters to identify when the approximation can be trusted.
+In addition, they provide the required implementations for one PP environment (Stan), although this is done at the level of case studies rather than integrating the approximation more tightly as part of a probabilistic programming environment.}
 
@@
 
-The diagnostic tool which consists in iterative steps to fit and diagnose the model has also been implemented in the \textit{brms} R-package. 
+%Paul: The statement below is incorrect. I have not implemented the diagnostic procedure in brms.
+%The diagnostic tool which consists in iterative steps to fit and diagnose the model has also been implemented in the \textit{brms} R-package. 
 %This diagnostic iterative procedure is fundamentally based on a diagnostic rule to check whether the approximation is accurate and numerical equations allowing updating the key parameters of the method.
@@
 
-Reviewer \#1 comments "The analysis is reasonable and provides tangible results on how the approximation scheme should be used, but does not go very deep.
+%Paul: I recommend highlighting either the reviewers comments or our responses in bold to clearly differentiate what the reviewers wrote and what we wrote. Example below.
+
+\textbf{The analysis is reasonable and provides tangible results on how the approximation scheme should be used, but does not go very deep.
 There are not new theoretical results but instead the analysis is purely empirical and in fact a standard study where each approximation parameter is varied at a time to identify the limits when the approximation starts to fail.
-The analysis is properly carried out and results in a practical diagnostic, but did not require notable scientific insight."
+The analysis is properly carried out and results in a practical diagnostic, but did not require notable scientific insight.}
 

### 2022-03-01 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edited abstract and added one ref to intro

@@
 Gaussian processes are powerful non-parametric probabilistic models for stochastic functions.
-However, they entail a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
+However, the direct implementation entails a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
 In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions.
-The main contribution of this paper is a detailed analysis of the performance and practical implementation of the method in relation to key factors such as the number of basis functions, domain of the prediction space, and smoothness of the latent function.
-We provide intuitive visualizations and recommendations for choosing the values of these factors, which make it easier for users to improve approximation accuracy and computational performance.
-We also propose diagnostics for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
-The proposed approach is simple and exhibits an attractive computational complexity due to its linear structure, and it is easy to implement in probabilistic programming frameworks.
+The main contribution of this paper is a detailed analysis of the performance, and practical recommendations for how to select the number of basis functions and the boundary factor.
+Intuitive visualizations and recommendations, make it easier for users to improve approximation accuracy and computational performance.
+We also propose diagnostics for checking that the number of basis functions and the boundary factor are adequate given the data.
+The approach is simple and exhibits an attractive computational complexity due to its linear structure, and it is easy to implement in probabilistic programming frameworks.
 Several illustrative examples of the performance and applicability of the method in the probabilistic programming language Stan are presented together with the underlying Stan model code.
@@
 Recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.,][]{grenander1981abstract} is presented by \citet{JSSv090i10}.
-
-{\color{blue} Aki: we could cite in the intro https://arxiv.org/abs/2111.01084 as a paper which
-reviews alternative methods}
+Yet another approach is to present Gaussian process using precision matrix, which is the inverse of the covariance matrix.
+If the precision matrix is sparse, computation taking the benefit of that sparsity can scale much better than $O(n^3)$.
+See, for example, review by \citet{Lindgren+Havard:2022:spde}.
 

### 2022-03-02 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
update response letter

@@
 
-%Paul: all responses will be send to all reviewers anyway
-%Authors would like the summary of the updates and insights presented above to be sent to all of the reviewers.
-
 \noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}
@@
 
+{\color{blue} Paul: This list is way too long and detailed and appears to be redundant at times. I recommend focusing on between 4 and 7 or so points but more will be too much.}
+
 \begin{itemize}
-
-%Paul: This list is way too long and detailed and appears to be redundant at times. I recommend focusing on between 4 and 7 or so points but more will be too much.
 
@@
 
-{\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}
+{\color{blue} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
 
@@
 
-\item A potential fitting overestimating effect when using too many basis functions relative to the wigglyness of the function to be learned has been commented. {\color{red}
+\item A potential fitting overestimating effect when using too many basis functions relative to the wigglyness of the function to be learned has been commented.
 
-(TODO: Currently we just have outlined but not discussed this. If we had time, it would be nice to discover why this overestimating effect is happening.)}
+{\color{blue} (Gabriel: Currently we just have outlined but not discussed this. If we had time, it would be nice to discover why this overestimating effect is happening.)}
 
@@
 
-%Paul: Don't forget the thank the reviewers for their comments to show some more appreciation of their comments and time they put into the review.
+Authors are really grateful to the reviewer for their valuable comments which have been very helpful to improve the paper.
 
@@
 
-%Paul: The statement below is incorrect. I have not implemented the diagnostic procedure in brms.
-%The diagnostic tool which consists in iterative steps to fit and diagnose the model has also been implemented in the \textit{brms} R-package. 
+We are planning to implement the diagnostic tool which consists in iterative steps to fit and diagnose the model in the \textit{brms} R-package.
+
 %This diagnostic iterative procedure is fundamentally based on a diagnostic rule to check whether the approximation is accurate and numerical equations allowing updating the key parameters of the method.
@@
 \subsubsection*{R\#1 Comment \#2}
-
-%Paul: I recommend highlighting either the reviewers comments or our responses in bold to clearly differentiate what the reviewers wrote and what we wrote. Example below.
 
@@
 
-Reviewer \#1 comments "The results are intuitive, but the process of determining the parameters in practice remains somewhat cumbersome as one needs to inspect the curves in Figures 6 and 7 to make the decision and there are no definite guidelines on how to modify the parameters when observing insufficient accuracy (e.g.
-no hints on how much one should increase m)."
+\textbf{The results are intuitive, but the process of determining the parameters in practice remains somewhat cumbersome as one needs to inspect the curves in Figures 6 and 7 to make the decision and there are no definite guidelines on how to modify the parameters when observing insufficient accuracy (e.g.
+no hints on how much one should increase m).}
 
@@
 
-Reviewer \#1 comments "Furthermore, to use the approximation with any other kernel besides square exponential or Matérn one would essentially need to repeat the same analysis from scratch.
+\textbf{Furthermore, to use the approximation with any other kernel besides square exponential or Matérn one would essentially need to repeat the same analysis from scratch.
 This limits the value in probabilistic programming context.
-The authors indicate existing GP  libraries are limited and favour Stan for its generality, but since we now only have support for two specific kernels one could easily argue this approach is limited as well."
+The authors indicate existing GP  libraries are limited and favour Stan for its generality, but since we now only have support for two specific kernels one could easily argue this approach is limited as well.}
 
@@
 
-Reviewer \#1 comments "it still feels like many educated practitioners would have found similar values for m and c by following a conventional validation procedure of applying the approximation with range of values for both on some example functions.
-While they would probably have ended up using slightly too large m or might spend a bit more computation while finding the best values, the practical difference might not be that big."
+\textbf{it still feels like many educated practitioners would have found similar values for m and c by following a conventional validation procedure of applying the approximation with range of values for both on some example functions.
+While they would probably have ended up using slightly too large m or might spend a bit more computation while finding the best values, the practical difference might not be that big.}
 
@@
 
-Reviewer \#1 comments "The case studies are interesting and demonstrate the approximation scheme works in practice for simple statistical models while providing clear example codes, but they could be more focused towards the main goal of the paper."
+\textbf{The case studies are interesting and demonstrate the approximation scheme works in practice for simple statistical models while providing clear example codes, but they could be more focused towards the main goal of the paper.}
 
@@
 
-{\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}
+{\color{red} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
 \end{itemize}
@@
 
-Reviewer \#1 comments "Section 5.1 takes approximately four full pages to present results for a simple artificial data.
-It describes one concrete example of setting the parameters (that remains somewhat vague -- "In the first iteration, we choose...", "In the second iteration, m is increased to..."), but the detailed comparisons against splines could have been much more brief in a paper focusing on how to make the approximation easier to use in practice."
+\textbf{Section 5.1 takes approximately four full pages to present results for a simple artificial data.
+It describes one concrete example of setting the parameters (that remains somewhat vague -- "In the first iteration, we choose...", "In the second iteration, m is increased to..."), but the detailed comparisons against splines could have been much more brief in a paper focusing on how to make the approximation easier to use in practice.}
 
@@
 
-Reviewer \#1 comments "Section 5.2 is similarly a bit besides the point -- the example model is interesting but regarding the choice of the parameters you simply say "We use m=30 basis functions and a boundary factor c=1.5" followed by stating that the length-scale estimate is good.
+\textbf{Section 5.2 is similarly a bit besides the point -- the example model is interesting but regarding the choice of the parameters you simply say "We use m=30 basis functions and a boundary factor c=1.5" followed by stating that the length-scale estimate is good.
 This is what we would expect to see in a paper that later uses the method, but for a paper that is introducing the diagnostic I would expect more details and e.g.
 explanations of what would happen if the parameters were poorly chosen.
-I understand fully that verbally describing an iterative process is difficult, but was still a bit let down by these case studies."
+I understand fully that verbally describing an iterative process is difficult, but was still a bit let down by these case studies.}
 
@@
 
-Reviewer \#1 comments "Section 5.3 presents a multivariate case and seems to mostly provide a negative result on applicability of HSGP for multivariate cases.
-This is fine as the method is valuable even if limited to low dimensions, but it would have been more interesting to see a more detailed analysis of the intermediate region, especially D=2 and D=3 that are frequent in spatial and spatiotemporal analysis problems -- a practitioner would appreciate clear guidelines on when to use HSGP and how to recognise when it is not a good solution."
+\textbf{Section 5.3 presents a multivariate case and seems to mostly provide a negative result on applicability of HSGP for multivariate cases.
+This is fine as the method is valuable even if limited to low dimensions, but it would have been more interesting to see a more detailed analysis of the intermediate region, especially D=2 and D=3 that are frequent in spatial and spatiotemporal analysis problems -- a practitioner would appreciate clear guidelines on when to use HSGP and how to recognise when it is not a good solution.}
 
@@
 We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material.
-{\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}
+{\color{red} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
 
@@
 
-Reviewer \#1 comments "Overall, I believe the paper has clear value for practical use and is worth publishing in some form: It provides nice code examples for HSGP via case studies and an iterative procedure for diagnosing when the approximation accuracy is good enough.
+\textbf{Overall, I believe the paper has clear value for practical use and is worth publishing in some form: It provides nice code examples for HSGP via case studies and an iterative procedure for diagnosing when the approximation accuracy is good enough.
 However, the scientific novelty of the work is fairly limited and perhaps insufficient for publication in a top venue, in particular because there are no claims or contributions that would extend beyond this specific approximation scheme.
 There are also no clear scientific insights in the process itself, since the analysis is purely empirical and follows a straightforward protocol.
-In summary, the manuscript is a borderline case even if addressing the remarks below."
+In summary, the manuscript is a borderline case even if addressing the remarks below.}
 
@@
 
-\#1.1 - Reviewer \#1 comments: "The diagnostic involves comparing the estimated length-scale $\hat l$ against a minimum $l^*$ determined by m and c.
+\textbf{\#1.1 - The diagnostic involves comparing the estimated length-scale $\hat l$ against a minimum $l^*$ determined by m and c.
 I have trouble understanding the full validity of this process.
@@
 Am I correct that your method would fail to diagnose a bad approximation if $\hat l$ was overestimated?
-I would like to see this discussion extended, at least by explicitly mentioning possible failure modes."
+I would like to see this discussion extended, at least by explicitly mentioning possible failure modes.}
 
@@
 
-\noindent \#1.2 - Reviewer \#1 comments: "To me it seems Fig 5 (left) already hints that you tend to underestimate l when using bad approximation and the procedure may be valid in practice, but this is not transparent in the paper."
+\textbf{\#1.2 - To me it seems Fig 5 (left) already hints that you tend to underestimate l when using bad approximation and the procedure may be valid in practice, but this is not transparent in the paper.}
 
@@
 
-\noindent \#1.3 Reviewer \#1 comments: "Do we have a reason for the error to be always towards understimating the lengthscale? Could this be proven?"
+\textbf{\#1.3 - Do we have a reason for the error to be always towards understimating the lengthscale? Could this be proven?}
 
@@
 
-Reviewer \#1 comments: "The process involves checking Figure 6 (or 7) for verifying whether the approximation is valid.
+\textbf{The process involves checking Figure 6 (or 7) for verifying whether the approximation is valid.
 This feels cumbersome in practice, and for better PP integration I would expect proper software support for this.
@@
 One could also make this completely automatic, a loop that keeps on modifying the parameters until the threshold is reached.
-I would like you to discuss which of these options you would prefer and why."
+I would like you to discuss which of these options you would prefer and why.}
 
@@
 
-Reviewer \#1 comments: "The paper is in general well written, but the structure could be improved.
+\textbf{The paper is in general well written, but the structure could be improved.
 I think the specific diagnostic is now lost in the middle of the story in Sections 4 and 5 and would benefit from clear presentation in dedicated section (that would also be a good place to discuss the software support; see comment 2 above).
@@
 Section 4.1 is 200+ lines of text with no structure and as mentioned above both 5.1 and 5.2 are somewhat long compared to their value.
-Since you anyway provide exact code and re-use models from previous papers, you could compress the general discussion of the models and focus more on the process of determining the approximation parameters."
+Since you anyway provide exact code and re-use models from previous papers, you could compress the general discussion of the models and focus more on the process of determining the approximation parameters.}
 
@@
 
-Reviewer \#1 comments: " It would be really interesting to see a bit more about how HSGP works for functions with 2-3 inputs, and in particular about how a practitioner can recognise whether it makes sense to use HSGP for their case or not.
+\textbf{It would be really interesting to see a bit more about how HSGP works for functions with 2-3 inputs, and in particular about how a practitioner can recognise whether it makes sense to use HSGP for their case or not.
 To me it seems clear HSGP is very often a reasonable method for D=1 and almost certainly breaks down for D>=5 or so unless working with very smooth functions (and for them almost any approximation works well), but clear diagnostics for quickly determining whether it could work for D=2 or D=3 would increase the contribution notably.
-Now the practical process would be to try it out and keep on increasing m until it starts working or one notices the computation is too slow, which feels wasteful -- as a practitioner I would rather just skip HSGP completely if finding out it does not help takes longer than solving the problem without the approximation."
+Now the practical process would be to try it out and keep on increasing m until it starts working or one notices the computation is too slow, which feels wasteful -- as a practitioner I would rather just skip HSGP completely if finding out it does not help takes longer than solving the problem without the approximation.}
 
@@
 
-Reviewer \#2 comments "The paper gives experimental results for the method of Solin and Sarkka  (2020).
-Overall the idea such experiments is interesting, but the actual experiments are not extensive enough given the pure applied nature of this work."
+\textbf{The paper gives experimental results for the method of Solin and Sarkka  (2020).
+Overall the idea such experiments is interesting, but the actual experiments are not extensive enough given the pure applied nature of this work.}
 
@@
 
-Reviewer \#2 comments "Experiments are mainly low dimensional (all 1D except a single 4D example).
+\textbf{Experiments are mainly low dimensional (all 1D except a single 4D example).
 Please discuss high dimensional cases ; indeed equation 8 suggests poor scaling in dimension.
-Can it be applied to high dimensional data?".
+Can it be applied to high dimensional data?.}
 
@@
 We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material.
-{\color{red} (TODO: I've not done this yet, but I could do it before the deadline.)}
+{\color{blue} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
 
@@
 \subsubsection*{R\#2 Comment \#3}
-Reviewer \#2 comments "There is no comparison with other techniques (except a spline method) !
+
+\textbf{There is no comparison with other techniques (except a spline method) !
 Please compare with, e.g.
 "MCMC for Variationally Sparse Gaussian Processes" ; although this is SVI, it appears directly applicable for the problems considered here.
-Please also compare with non MCMC variational methods - while not directly applicable to generic MCMC methods, these are the sttandard and it would be nice to see how they compare in runtime."
+Please also compare with non MCMC variational methods - while not directly applicable to generic MCMC methods, these are the sttandard and it would be nice to see how they compare in runtime.}
 
@@
 
-Reviewer \#2 comments "On the eigenfunction expansion: 1) note that this is very similar to the Mercer expansion of e.g.
+\textbf{On the eigenfunction expansion: 1) note that this is very similar to the Mercer expansion of e.g.
 Walder et al, "Fast Bayesian Intensity Estimation for the Permanental Process".
@@
 For example, the Nystrom applies to any kernel (not just stationary) and yields a linear GP approximation, but requires some overhead for the setup.
-Given this submission is purely experimental, it would be nice to see comparisons.".
+Given this submission is purely experimental, it would be nice to see comparisons.}
 

### 2022-03-05 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Fix table overflow
add paragrah about sampling issue due to large m

@@
 
-Furthermore, from Figure~\ref{fig7_lscale_comparison} left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e., true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_overestimated_posterios} as an illustration of the overestimating effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison} left).
+If much bigger $m$ than necessary is used (i.e., too much large $m$ relative to the wigglyness of the function to be learned), there can be computational challenges.
+In Figure~\ref{fig7_lscale_comparison}, we see that increasing $m$ much beyond what is needed, is leading to underestimating the lengthscale.
+Examining the convergence diagnostics revealed problems in exploring larger length-scales, which is likely due to the posterior having a difficult funnel shape \citep[see, e.g.,][]{Betancourt+Girolami:2019}.
+A largely underestimation of the length-scale leads to overestimating the underlying functions (see Figure~\ref{fig8_overestimated_posterios} as an illustration of the overestimating effect as a function of $m$ in a moderate smooth dataset).
+Increasing the boundary factor $c$ (more than that recommended in equations~\eqref{sec_num_equations}) reduces the funnellness of the posterior and the sampling problems in exploring larger length-scales mostly disappear and so underestimating the length-scale.
+
+Following our recommended diagnostic tool is a safety procedure to avoid using too much large $m$ that needed, as it allows us to select optimal values for $m$ and $c$ in each iteration.
 
@@
 \includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_overestimated_posterios.png}
-\caption{Overestimating effect on posterior functions when using too many number of basis functions relative to the wigglyness of the function to be learned.
-The dashed red line represents the true function.}
+\caption{Overestimation of the data points when using too many number of basis functions relative to the wigglyness of the data points. The dashed red line represents the true underlying function.}
   \label{fig8_overestimated_posterios}
@@
 \centering
+\setlength{\tabcolsep}{5pt}
 \begin{tabular}{ c c c c c c c c c c | c c}
@@
 \centering
+\setlength{\tabcolsep}{4pt}
 \begin{tabular}{ccccccccc}

### 2022-03-05 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
add new 2D case study

@@
 
+\subsection{Case study of simulated data for a bivariate function}
+
+This example consists of a synthetic dataset with $n=120$ ($i=1,\dots,n$) single draws from a GP prior as a function of input values $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$ in a bivariate input space ($D=2$).
+A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, for the first input dimension, and $\ell_2=0.35$, for the second input dimension, is used for the GP prior.
+Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^{120}$.
+
+In the HSGP model with $2$ input dimensions, the latent function $f(\bm{x})$ is approximated as in equation (13)% (\ref{eq_approxf_multi})
+, with the $2$D squared exponential spectral density $S$ as in eq.
+\eqref{eq_specdens_inf}, and the $2$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations \eqref{eq_eigenfunction_multi} and \eqref{eq_eigenvalue_multi}, respectively.
+
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
+We use the same prior distributions to fit the exact GP model as the HSGP models.
+
+The HSGP model is fitted following the recommended iterative steps as in Section~\ref{sec_user_guide}, where in each iteration the diagnosis is applied on both dimensions, where for each one the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated by equations in Section~\ref{sec_num_equations}.
+The initial values $0.5$ and $1$ are set for the minimum lengthscales $\ell_1$ and $\ell_2$, for first and second dimensions respectively.
+
+In order to do model comparison, in addition to the exact GP and HSGP models, a $2$D splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
+
+Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, exact GP, HSGP, and splines, fitted over the dataset.
+Sample observations are also plotted as circles.
+For the HSGP, $m_1=40$ and $m_2=40$ basis functions for each dimension, respectively, were used, which lead to a total of 1600 multivariate basis functions.
+A boundary factor for each dimension, $c_1=1.5$ and $c_2=1.5$, were used.
+For the spline-based model, 40 knots in each dimension were used.
+
+Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the data-generating function and the GP, HSGP and spline-based models, respectively.
+
+\begin{figure*}
+\begin{center}
+\begin{tabular}{ c c c c}
+\includegraphics[scale=0.40, trim = 0mm 18mm 35mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf} &
+\includegraphics[scale=0.40, trim = 19mm 18mm 35mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}  & 
+\includegraphics[scale=0.40, trim = 19mm 18mm 35mm 20mm, clip]{ch5_fig15_spfun_exII.pdf} & 
+\includegraphics[scale=0.35, trim = 150mm 10mm 0mm 0mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}
+\end{tabular}
+\end{center}
+\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline-based (d) models. Sample points are plotted as circles}
+  \label{ch5_fig15_Posteriors_exII}
+\end{figure*}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
+\caption{Error between the data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models. Sample points are plotted as circles.}
+  \label{ch5_fig16_errors_exII}
+\end{figure}
+
+In order to assess performance of the models as a function of the number of basis functions, boundary factor and knots, different models with different number of basis functions, for the HSGP, and different number of knots, for the spline-based model, have been fitted.
+In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
+Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the spline-based model.
+From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen an accurate approximation of the HSGP to the exact GP.
+However, the performance of the spline-based model is significantly worse.
+
+Figure \ref{ch5_fig17_RMSE_exII}-right shows the computation times of the different models as a function of the boundary factor, number of basis functions and knots.
+Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computation time.
+Even though in a bivariate input space the computational demands increase significantly with the number of dimensions or knots, either the HSGP or spline-based models work significantly better than exact GP, even for highly wiggly functions that require a high number of basis functions or knots for an accurate approximation.
+It is interesting to be noticed that just the computation of the input data for the spline model in a 2D input space and considering more than 50 knots per dimension is computationally very expensive.
+
+The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
+\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
+\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+  \label{ch5_fig17_RMSE_exII}
+\end{figure}
+
 \subsection{Leukemia data}\label{sec_bf_caseVI}

### 2022-03-05 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
remove text and fig 8 about sampling issue due to large m

@@
 
-If much bigger $m$ than necessary is used (i.e., too much large $m$ relative to the wigglyness of the function to be learned), there can be computational challenges.
+%If much bigger $m$ than necessary is used (i.e., too much large $m$ relative to the wigglyness of the function to be learned), there can be computational challenges.
 In Figure~\ref{fig7_lscale_comparison}, we see that increasing $m$ much beyond what is needed, is leading to underestimating the lengthscale.
@@
 Increasing the boundary factor $c$ (more than that recommended in equations~\eqref{sec_num_equations}) reduces the funnellness of the posterior and the sampling problems in exploring larger length-scales mostly disappear and so underestimating the length-scale.
-
-Following our recommended diagnostic tool is a safety procedure to avoid using too much large $m$ that needed, as it allows us to select optimal values for $m$ and $c$ in each iteration.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_overestimated_posterios.png}
-\caption{Overestimation of the data points when using too many number of basis functions relative to the wigglyness of the data points. The dashed red line represents the true underlying function.}
-  \label{fig8_overestimated_posterios}
-\end{figure}
+%
+%Following our recommended diagnostic tool is a safety procedure to avoid using too much large $m$ that needed, as it allows us to select optimal values for $m$ and $c$ in each iteration.
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_overestimated_posterios.png}
+%\caption{Overestimation of the data points when using too many number of basis functions relative to the wigglyness of the data points. The dashed red line represents the true underlying function.}
+%  \label{fig8_overestimated_posterios}
+%\end{figure}
 
@@
 Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
-Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length scale that would cause overestimating the underlying function and bias the posterior towards smaller length scale values.
+%Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length scale that would cause overestimating the underlying function and bias the posterior towards smaller length scale values.
 
@@
 \centering
+\setlength{\tabcolsep}{4pt}
 \begin{tabular}{cccccccc}
@@
 \begin{figure*}
-\begin{center}
-\begin{tabular}{ c c c c}
-\includegraphics[scale=0.40, trim = 0mm 18mm 35mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf} &
-\includegraphics[scale=0.40, trim = 19mm 18mm 35mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}  & 
-\includegraphics[scale=0.40, trim = 19mm 18mm 35mm 20mm, clip]{ch5_fig15_spfun_exII.pdf} & 
-\includegraphics[scale=0.35, trim = 150mm 10mm 0mm 0mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}
-\end{tabular}
-\end{center}
+\centering
+\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf}
+\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}
+\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_spfun_exII.pdf}
+\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}
 \caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline-based (d) models. Sample points are plotted as circles}
@@
   \label{ch5_fig16_errors_exII}
-\end{figure}
+\end{figure*}
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
-\subfigure{\includegraphics[scale=0.40, trim = 28mm 35mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
+\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
+\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
+\subfigure{\includegraphics[scale=0.35, trim = 28mm 70mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
 \caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}

### 2022-03-07 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
update response letter

@@
 Then, all comments are addressed in a point-by-point manner, with separate replies to each reviewer.
-Throughout, comments or the reviewers are hightlighted in bold.
+Throughout, comments of the reviewers are hightlighted in bold.
 
@@
 
-{\color{blue} Paul: This list is way too long and detailed and appears to be redundant at times. I recommend focusing on between 4 and 7 or so points but more will be too much.}
+\textit{General updates:}
 
@@
 
+\item A safety, general and extensive procedure that allows us to diagnose and automatize the fitting process has been proposed (Section 4).
+
+\item Section 5 has been focused on the main goals of the paper, which are to 1) facilitate the implementation of the method for practical users providing clear formulation of the method and 2) provide a safety and general diagnostic tool to fit and diagnose the model.
+
+\item The analysis has been extended to the Mat\'ern-5/2 kernel.
+So the whole study finally provides the diagnostics for a squared exponential, Mat\'ern-3/2, Mat\'ern-5/2 (Sections 4.3 and 4.3.1), and periodic squared exponential (Sections 3.3, 4.4 and Appendix C) covariance functions.
+
+\item A discussion of the performance of the model and the applicability of the diagnostic tool in multidimensional cases have been included in Section 4.8 and case study III.
+
+\item We expect this paper having a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios.
+
+\end{itemize}
+
+\noindent \textit{Specific updates:}
+
+\begin{itemize}
+
 \item A discussion about the theoretical evidence of the near linear proportionality between the setting parameters of the model has been added (Section 4.2).
 
-\item In the first submission of the paper, the relationships between the setting parameters of the model were analyzed by simulation experiments and presented in form of graphs.
+\item We have derived numerical equations that characterize the relationships between the setting parameters of the model for each particular kernel, which are useful to update and find out optimal values for the parameters (Section 4.3.1).
 
-\item We have derived numerical equations that characterize the relationships between the setting parameters of the model, which are useful to automatize updating the parameters in the fitting process (Section 4.3.1).
+\item The main diagnostic rule that indicates whether the approximation may be sufficiently accurate has been further discussed regarding its reliability and applicability (Section 4.5).
 
-\item The diagnostic rule that allowed to check whether the approximation was sufficiently accurate has been further discussed regarding its reliability and applicability (Sections 4.5).
+\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1 and Section 4.6), which is based on: 1) recommendations to set initial values for the parameters, 2) safety diagnostic rules to check if the approximation is sufficiently accurate (Sections 4.5), and 3) numerical equations to automatize updating the parameters in the fitting process (Sections 4.3.1).
 
-\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1), which is based on:
+\item Recommendations to perform diagnosis on any other covariance function different from those implemented in this study have been provided (Sections 4.2 and 4.7).
 
-\begin{enumerate}
-\item Recommendations to set initial values for the parameters.
-\item Safety diagnostic rules to check if the approximation is sufficiently accurate (Sections 4.5).
-\item Numerical equations to update and find out the optimal values for the parameters (Sections 4.3.1).
-\end{enumerate}
-
-\item The proposed diagnostic procedure is a safety, general and extensive procedure that allows us to automatize the fitting process.
-
-\item A thorough analysis of the performance of the diagnostic tool has been carried out (Section 4.6).
-
-\item The diagnostic tool has been extended to add the analysis for the Mat\'ern-5/2 kernel.
-So the whole study finally provides the diagnostics for a squared exponential, Mat\'ern-3/2, Mat\'ern-5/2 (Sections 4.3 and 4.3.1), and periodic squared exponential (Sections 3.3, 4.4 and Appendix C) covariance functions.
-
-\item By analyzing the constants characterizing the relationships for each covariance function, recommendations to perform diagnosis on any other covariance function different from those implemented in this study have been provided (Sections 4.2 and 4.7).
-
-\item Explanations of how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).
-
-\item In the first submission of the paper, a linear representation for a periodic squared exponential kernel and the corresponding diagnosis analysis were presented in Appendix B.
-
-\item We have derived the numerical equations governing the relationships between the parameters of this linear representation of the periodic kernel.
-These equations have been included in the main text in Section 4.4.
-
-\item We have applied the diagnostic tool on cases studies in Section 5.
-In that way, case studies in Section 5 have become more focused on the main goals of the paper, which are to facilitate the implementation of the method for practical users providing clear formulation of the method and provide a safety and general diagnostic tool to fit and diagnose the model.
-
-\item We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set.
-We have applied the diagnostic procedure on it.
-Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing individually the different dimensions in each iteration.
-It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
-
-We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material.
-
-{\color{blue} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
-
-\item We have added Section 4.8 where a discussion of the performance of the model in multidimensional cases are presented.
-
-\item We have emphasized in the paper that applying the diagnostic tool allows for minimum computational requirements that can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting such as cross-validation and simulation based calibration (Section 4.8).
-
-\item A potential fitting overestimating effect when using too many basis functions relative to the wigglyness of the function to be learned has been commented.
-
-{\color{blue} (Gabriel: Currently we just have outlined but not discussed this. If we had time, it would be nice to discover why this overestimating effect is happening.)}
-
-\item We expect this paper having a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios like uni-dimensional, multi(low)-dimensional functions and other problems that require repeated fitting.
+\item We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set, demonstrating the applicability and usefulness of the diagnostic tool in multidimensional cases.
+Additional results concerning computational cost in 2D, 3D and 4D have been presented and briefly discussed, emphasizing that applying the diagnostic tool allows for minimum computational requirements that can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting.
 
@@
 
-Authors are really grateful to the reviewer for their valuable comments which have been very helpful to improve the paper.
+Authors are really grateful to the reviewer for their valuable comments which have been very helpful to improve the paper, and for her/his dedicated time.
 
@@
 The case study's models have also been formulated in general probabilistic language in the main manuscript, and Stan's implementations have also been provided as examples.
-Furthermore, HSGP model has been implemented in the \textit{brms} R-package that uses Stan to make model inference.
-
-We are planning to implement the diagnostic tool which consists in iterative steps to fit and diagnose the model in the \textit{brms} R-package.
+Furthermore, HSGP model has been implemented in the \textit{brms} R-package that uses Stan to make model inference, and we are planning to implement the diagnostic tool which consists in iterative steps to fit and diagnose the model also in the \textit{brms} R-package.
+{\color{blue}(Gabriel: Paul, do we intend to implement the diagnostic procedure in brms?)}
 
@@
 The study pursues accurately measuring the sensitivity of the HSGP model to the values of these parameters, detecting when the approximation fails and when the parameter values are optimal in accuracy and computation cost.
-Ultimately, the aim of the study is to provide a diagnostic to check whether the approximation is sufficiently accurate, provide numerical equations to update the setting parameters if the approximation is not sufficiently accurate, and finally provide a safety and general procedure based on iterative steps to successfully fit the model, finding out optimal values for the setting parameters which allow for minimum computational requirements.
+Ultimately, the aim of the study is to provide: 1) a diagnostic to check whether the approximation is sufficiently accurate, 2) numerical equations to update the setting parameters if the approximation is not sufficiently accurate, and 3) a safety and general procedure based on iterative steps to successfully fit the model, finding out optimal values for the setting parameters which allow for minimum computational requirements.
 
@@
 
-Furthermore, by following our recommended diagnostic procedure, an optimum solution with minimum computational requirements is achieved, that can be essential in repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional cases with very large data sets, multi(low)-dimensional cases, even with small data sets, since knowing the optimum values for the setting parameters for each individual dimension can significantly reduce the computational requirements, otherwise fitting HSGPs would be unfeasible here.
+Furthermore, by following our recommended diagnostic procedure, an optimum solution with minimum computational requirements is achieved, that can be essential in repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional cases with very large data sets, and multi(low)-dimensional cases, even with small data sets, since knowing the optimum values for the setting parameters for each individual dimension can significantly reduce the computational requirements, otherwise fitting HSGPs would be unfeasible here.
 
@@
 
-We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate.
-We have also elaborated a safety and general procedure based on iterative steps to fit the model with definite rules to update the parameters and make diagnosis.
+We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate (Section 4.3.1).
+We have also elaborated a safety and general procedure based on iterative steps to fit the model with definite rules to update the parameters and make diagnosis (Sections 4.5 and 4.6).
 
@@
 
-We have implemented the HSGP model and the diagnostic tool for four different kernels, the squared exponential, Matern-3/2, Matern-5/2, and periodic squared exponential kernels, which cover a broad range of learning functions in practical applications.
+We have implemented the HSGP model and the diagnostic tool for four different kernels, the squared exponential, Matern-3/2, Matern-5/2, and periodic squared exponential kernels, which cover a broad range of learning functions in practical applications (Sections 4.3 and 4.4).
 Kernel combinations that satisfy the requirements can also be used.
-We have also given recommendations in Section 4.2 and 4.7 to update $m$ and $c$ and perform diagnosis on any other different stationary kernel from those implemented in our study.
+We have also given recommendations to update $m$ and $c$ and perform diagnosis on any other different stationary kernel from those implemented in our study, by analyzing the constants characterizing the relationships for each covariance function (Sections 4.2 and 4.7).
+Explanations of how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).
 
-The study deeply describes in Section 3 the formulation and implementation of the HSGP model, description that can be used to implement it for any other different stationary kernel.
-The only new thing that the user needs to compute is the Fourier transform of the kernel to obtain the spectral density of the kernel.
+The study deeply describes in Section 3 the formulation and implementation of the HSGP model, a description that can be used to implement it for any other different stationary kernel.
+The only new thing that the user needs to compute is the Fourier transform of the kernel to obtain the spectral density of the kernel (Section 3 and Section 4.7).
 
@@
 
-Speculation from the reviewers side; we do not really know the expertise of our users.
+We do not really know the expertise of our users.
 
@@
 \item Case study II involves a function that is composed of three additive GP components with different kernels, squared exponential and periodic kernels.
-Additionally this case study has a large data set of more than 7000 observations which would be unfeasible to be fitted using the exact GP model.
+Additionally this case study has a large data set of more than 7000 observations which would be unfeasible to be fitted using a regular formulation of the GP model.
 
@@
 
-\item Case study III involves a two-dimensional function on which we use the diagnostic tool to fit the model.
+\item Case study III involves 2D and 3D functions on which the diagnostic tool is used to fit and diagnose the model.
 We show that, although the diagnostic tool was built for unidimensional covariance functions, it can be successfully used for multi-dimensional cases by diagnosing individually the different dimensions in each iteration.
 It allows gradually finding the optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
-
-Taking advantage of this case study we present additional insights about the performance of the method in 3D and 4D cases from other examples initially included in online material.
-
-{\color{red} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
+We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D cases.
 \end{itemize}
@@
 
-It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
-In fact, for that reason the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
+The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already fast.
+
+It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
+For that reason, in fact, the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
+For D>5 even for very smooth functions seems to be impractical.
 
@@
 In fact using the diagnostic tool the user will discover easily the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP.
-Without the guide of the diagnostic tool the user probably would had made several trials until finding out an accurate and practical solution, consuming a lot of requirements in this demanding 4D case.
+Without the guide of the diagnostic tool the user probably would had made several trials until finding out an accurate and practical solution, consuming a lot of computation requirements in this demanding 4D case.
 
@@
 However, for $D=2$ it is still useful and significantly faster than the exact GP even for very wiggly functions.
-In Section 4.8 we give brief details on the performance of the model in multi-D cases found from various case studies carried out by the authors and initially presented in the online material.
+In Section 4.8 we give brief details on the performance of the model in multi-D cases.
 In this sense, the use of our recommended diagnostic allows to find the optimal values for $c$ and $m$ which allows for minimum computational requirements which can be essential on these multidimensional cases.
 
-We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set.
+We have added a new case study to Section 5 which consists in wiggly 2D and 3D simulated data sets.
 We have applied the diagnostic procedure on it.
@@
 It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
-We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material.
-{\color{red} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
+We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D.
 
@@
 
-We can not state anything definite about the estimated lengthscale $\hat{\ell}$ when the approximation is inaccurate, basically because it might ultimately depend on the specific frequency patterns of the signal/function to be learned, and also because $c$ must have been properly set.
+We can not state anything definite about the estimated lengthscale $\hat{\ell}$ when the approximation is inaccurate, basically because it might ultimately depend on the specific frequency patterns of the signal/function to be learned, and also because $c$ must be properly set.
 Although, we can state that for most of the cases, if $c$ is large enough, $\hat{\ell}$ will be below that minimum supported length-scale ($\ell^*$) determined by $m$ and $c$.
@@
 
-What we can definitely state, which is in fact what our diagnosis is based on, is the following:
+What we can definitely state, which is, in fact, what our diagnosis is based on, is the following:
 
@@
 
-We have simplified and improved significantly the writing of Section 5. We have also focused the case studies more on the main goals of the paper:
-\begin{enumerate}
-\item Facilitate the implementation of the method for practical users providing clear formulation of the method.
-\item Provide a safety and general diagnostic tool to fit and diagnose the model.
-\end{enumerate}
+We have simplified and improved significantly the writing of Section 5.
+We have also focused the case studies more on the main goals of the paper, which are: 1) to facilitate the implementation of the method for practical users providing clear formulation of the method, and 2) and to provide a safety and general diagnostic tool to fit and diagnose the model.
 
@@
 \subsection*{Reviewer \#2}
+
+Authors are really grateful to the reviewer for their valuable comments which have been very helpful to improve the paper, and for her/his time.
 
@@
 
-It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
-In fact, for that reason the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
+The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already fast.
+It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
+For that reason, in fact, the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
+For D>5 even for very smooth functions seems to be impractical.
 
@@
 
-We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set.
+We have added a new case study to Section 5 which consists in wiggly 2D and 3D simulated data sets.
 We have applied the diagnostic procedure on it.
@@
 It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
-We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D, from case studies initially included in the online material.
-{\color{blue} (Gabriel: I've not done this yet, but I could do it before the deadline.)}
+We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D.
 
@@
 However, for $D=2$ it is still useful and significantly faster than the exact GP even for very wiggly functions.
-In Section 4.8 we give brief details on the performance of the model in multi-D cases found from various case studies carried out by the authors and initially presented in the online material.
+In Section 4.8 we give brief details on the performance of the model in multi-D cases.
 In this sense, the use of our recommended diagnostic allows to find the optimal values for $c$ and $m$ which allows for minimum computational requirements which can be essential on these multidimensional cases.

### 2022-03-07 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
minor edits

@@
 However, the direct implementation entails a complexity that is computationally intractable when the number of observations is large, especially when estimated with fully Bayesian methods such as Markov chain Monte Carlo.
-In this paper, we focus on a novel approach for low-rank approximate Bayesian Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions.
+In this paper, we focus on a low-rank approximate Bayesian Gaussian processes, based on a basis function approximation via Laplace eigenfunctions for stationary covariance functions.
 The main contribution of this paper is a detailed analysis of the performance, and practical recommendations for how to select the number of basis functions and the boundary factor.
@@
 In this paper, the interactions among the key factors of the method such as the number of basis functions, domain of the prediction space, and properties of the true functional relationship between covariates and response variable, are investigated and analyzed in detail in relation to the computational performance and accuracy of the method.
-Practical recommendations are given for the values of the key factors based on intuitive graphical summaries that encode the recognized relationships.
+Practical recommendations are given for the values of the key factors based on simple diagnostic values and intuitive graphical summaries that encode the recognized relationships.
 Our recommendations will help users to choose valid and optimized values for these factors, improving computational performance without sacrificing modeling accuracy.
-We also propose  diagnostics to indicate whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
+The proposed diagnostic can be used to check whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
 
@@
 If $m$ is fixed to a sufficiently large value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation of the mean mainly near the boundaries, while the approximation of the standard deviation is affected across the whole domain.
-What is more, the approximation error tends to be bigger for the standard deviation than for the mean.
+The approximation error tends to be bigger for the standard deviation than for the mean.
 
@@
 
-The component $f_4(t)$ represents the special days effects, modeled as a t-student prior model with 1 degree of freedom and variance $\tau^2$:
+The component $f_4(t)$ represents the special days effects, modeled as a Student's $t$ prior model with 1 degree of freedom and variance $\tau^2$:
 %
 \begin{equation*}
-f_4(t) \sim \text{t-student}(1,0,\tau^2).
+f_4(t) \sim t(1,0,\tau^2).
 \end{equation*}
@@
 $\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1 \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell, \ell,\ell\}$, respectively.
-A $\mathrm{Normal}(0,0.1)$ prior distribution has been used for the standard deviation $\tau$ of the t-student distribution with 1 degree of freedom used to model $f_4$ (i.e., the special days effects).
+A $\mathrm{Normal}(0,0.1)$ prior distribution has been used for the standard deviation $\tau$ of the Student's $t$ distribution with 1 degree of freedom used to model $f_4$ (i.e., the special days effects).
 
@@
 
-Modeling unknown functions using exact GPs is computationally intractable for a lot of applications.
+Modeling unknown functions using exact GPs is computationally intractable in many applications.
 This problem becomes especially severe when performing full Bayesian inference using sampling-based methods.
-In this paper, a recent approach for a low-rank representation of stationary GPs, originally proposed by \citet{solin2018hilbert}, has been implemented and analyzed in detail.
+In this paper, a recent approach for a low-rank representation of stationary GPs, originally proposed by \citet{solin2018hilbert}, has been analyzed in detail.
 The method is based on a basis function approximation via Laplace eigenfunctions.
@@
 Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships}, is essential to avoid a excessive computational time especially in multivariate input spaces.
-However, in practice, input dimensionalities larger than three start to be quite computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
+However, in practice, input dimensionalities larger than three start to be computationally demanding even for moderately wiggly functions and few basis functions per input dimension.
 In these high dimensional cases, the proposed approximate GP methods may still be used for low-dimensional components in an additive modeling scheme but without modeling very high dimensional interactions, as complexity is linear with the number of additive components.
@@
 The obtained functional relationships between the key factors influencing the approximation not only help users to visually assess the accuracy of the method
-but can also serve an automatic diagnostic tool, if appropriately implemented.
+but can also serve an automatic diagnostic tool.
 In this paper, we primarily studied the functional relationships for univariate inputs.

### 2022-03-07 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
edited rebuttal

@@
 
-\item A safety, general and extensive procedure that allows us to diagnose and automatize the fitting process has been proposed (Section 4).
+\item A safe, general, and extensive procedure that allows us to diagnose and automatize the fitting process has been defined explicitly (Section 4).
 
-\item Section 5 has been focused on the main goals of the paper, which are to 1) facilitate the implementation of the method for practical users providing clear formulation of the method and 2) provide a safety and general diagnostic tool to fit and diagnose the model.
+\item Examples in Section 5 have been clarified.
 
 \item The analysis has been extended to the Mat\'ern-5/2 kernel.
-So the whole study finally provides the diagnostics for a squared exponential, Mat\'ern-3/2, Mat\'ern-5/2 (Sections 4.3 and 4.3.1), and periodic squared exponential (Sections 3.3, 4.4 and Appendix C) covariance functions.
+So the whole study provides the explicit diagnostic for a squared exponential, Mat\'ern-3/2, Mat\'ern-5/2 (Sections 4.3 and 4.3.1), and periodic squared exponential (Sections 3.3, 4.4 and Appendix C) covariance functions.
+In addition connection to other covariance functions is discussed.
 
@@
 
-\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1 and Section 4.6), which is based on: 1) recommendations to set initial values for the parameters, 2) safety diagnostic rules to check if the approximation is sufficiently accurate (Sections 4.5), and 3) numerical equations to automatize updating the parameters in the fitting process (Sections 4.3.1).
+\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1 and Section 4.6), which is based on: 1) recommendations to set initial values for the parameters, 2) diagnostic to check if the approximation is sufficiently accurate (Sections 4.5), and 3) numerical equations to automatize updating the parameters in the fitting process (Sections 4.3.1).
 
@@
 
-\item We have added a new case study to Section 5 which consists in a wiggly 2D simulated data set, demonstrating the applicability and usefulness of the diagnostic tool in multidimensional cases.
+\item We have added a new case study to Section 5 which consists of a wiggly 2D simulated data set, demonstrating the applicability and usefulness of the diagnostic tool in multidimensional cases.
 Additional results concerning computational cost in 2D, 3D and 4D have been presented and briefly discussed, emphasizing that applying the diagnostic tool allows for minimum computational requirements that can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting.
@@
 
-The Hilbert space approximation for GPs (HSGP) is presented as a general formulation for probabilistic modeling, so it can be implemented in different probabilistic programming frameworks, such as Stan, WinBugs, etc.
+The Hilbert space approximation for GPs (HSGP) is presented as a general formulation for probabilistic modeling, so it can be implemented in different probabilistic programming frameworks, such as Stan, PyMC, Pyro, Beanmachine, etc.
 The case study's models have also been formulated in general probabilistic language in the main manuscript, and Stan's implementations have also been provided as examples.
@@
 
-The only limitation of the method is the number of dimensions as these affect computation through the total number of the basis functions needed.
-But this does not necessarily means a statistical model being simple.
-In fact due to this high dependency of the computational requirements of the method on the number of basis functions, the analysis and diagnosis tool developed in this paper arises to be essential for practical applications and in broad scenarios.
+The only limitation of the method is the expoential scaling of the computation with increasing number of dimensions (assuming wigglines stays the same and the amount of data is sufficient to get sufficient informationa about that wiggliness).
+Due to the high dependency of the computational requirements of the method on the number of basis functions, the analysis and diagnosis tool developed in this paper arises to be essential for practical applications and in broad scenarios.
 
@@
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
-For D>5 even for very smooth functions seems to be impractical.
+For $D>5$ even for very smooth functions seems to be impractical.
 
@@
 
-This work is necessary to facilitate the implementation of the method for practical users, while Solin and Sarka (2020) fundamentally focus on the mathematical development of the model.
+Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, and this work is necessary to facilitate the easier use of the method for practical users.
 We also analyze how the key factors of the model relate each other for accurate approximations.
@@
 
-We expect this paper to have a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios like uni-dimensional functions, multi-dimensional functions, and other problems that require repeated fitting such as cross-validation and simulation based calibration.
+We expect this paper to have an impact on how widely the basis function GPs are used, as it is makes the method more usable in practice and in broad scenarios like uni-dimensional functions, multi-dimensional functions, and other problems that require repeated fitting such as cross-validation and simulation based calibration.
 
-We do not see a problem in a purely simulative approach.
+%We do not see a problem in a purely simulative approach.
 It simply has different advantages and disadvantages over analytic approaches.
@@
 \textbf{It would be really interesting to see a bit more about how HSGP works for functions with 2-3 inputs, and in particular about how a practitioner can recognise whether it makes sense to use HSGP for their case or not.
-To me it seems clear HSGP is very often a reasonable method for D=1 and almost certainly breaks down for D>=5 or so unless working with very smooth functions (and for them almost any approximation works well), but clear diagnostics for quickly determining whether it could work for D=2 or D=3 would increase the contribution notably.
+To me it seems clear HSGP is very often a reasonable method for D=1 and almost certainly breaks down for $D>=5$ or so unless working with very smooth functions (and for them almost any approximation works well), but clear diagnostics for quickly determining whether it could work for D=2 or D=3 would increase the contribution notably.
 Now the practical process would be to try it out and keep on increasing m until it starts working or one notices the computation is too slow, which feels wasteful -- as a practitioner I would rather just skip HSGP completely if finding out it does not help takes longer than solving the problem without the approximation.}
@@
 
-Aki: R2 is not happy that we compare only to splines and suggests we should compare to SVI and other variational methods like splines and VI would belong to the same category.
+The main comparison is the Hilbert space basis function approximation against the exact GP as a model.
+In all comparisons the inference conditional on the model is assumed to be accurate.
+We've used efficient dynamic HMC, the convergence diagnostics didn't reveal problems, and thus the accuracy of the inference is good.
+Comparisons of MCMC to variational inference approximations are out of scope of this paper.
+Some comparisons of MCMC and variational inference for basis function GPs can be found in
+
+Lu Zhang, Bob Carpenter, Andrew Gelman, and Aki Vehtari (2021). Pathfinder: Parallel quasi-Newton variational inference. arXiv preprint arXiv:2108.03782.
+
+Comparisons of the Hilbert space basis function approximations against other GP approximations are provided by Solin and Särkkä (2020), which is cited prominently in our paper.
+
+% Aki: R2 is not happy that we compare only to splines and suggests we should compare to SVI and other variational methods like splines and VI would belong to the same category.
 We can use Stan's ADVI to run variational inference results.
@@
 We have recently run ADVI and improved robust ADVI results for birthday and motorcycle examples (the latter has latent GP for both location and log sigma)
-and VI is not that good (it's annoying how common belief it is that black box
-VI would be great).
+% and VI is not that good (it's annoying how common belief it is that black box
+% VI would be great).
 
@@
 
-Aki: R2 asks comparisons to Nystrom, but that comparison was
-already made by Solin and Särkkä so I don't think we should add that.
+Comparisons of the Hilbert space basis function approximations against other GP approximations, including the Nyström method, are provided by Solin and Särkkä (2020), which is cited prominently in our paper.
+
+% Aki: R2 asks comparisons to Nystrom, but that comparison was
+% already made by Solin and Särkkä so I don't think we should add that.
 

### 2022-03-08 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
first round of edits to the response letter

@@
 
-\item A discussion of the performance of the model and the applicability of the diagnostic tool in multidimensional cases have been included in Section 4.8 and case study III.
+\item A discussion of the performance of the model and the applicability of the diagnostic tool in multidimensional cases have been included in Section 4.8 and Case study III.
 
-\item We expect this paper having a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios.
+%Paul: I don't see how this is an update of the paper
+%\item We expect this paper having a high impact on users as it is highly recommended to make the method usable in practice and in broad scenarios.
 
@@
 
-\item A discussion about the theoretical evidence of the near linear proportionality between the setting parameters of the model has been added (Section 4.2).
+\item A discussion about the theoretical evidence of the near linear proportionality between the hyperparameters of the model has been added (Section 4.2).
 
-\item We have derived numerical equations that characterize the relationships between the setting parameters of the model for each particular kernel, which are useful to update and find out optimal values for the parameters (Section 4.3.1).
+\item We have derived numerical equations that characterize the relationships between the hyperparameters of the model for each particular kernel, which are useful to update and find out optimal values for the hyper-parameters (Section 4.3.1).
 
-\item The main diagnostic rule that indicates whether the approximation may be sufficiently accurate has been further discussed regarding its reliability and applicability (Section 4.5).
+\item The main diagnostic rule that indicates whether the approximation may be sufficiently accurate has been further discussed regarding its reliability and practial applicability (Section 4.5).
 
-\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1 and Section 4.6), which is based on: 1) recommendations to set initial values for the parameters, 2) diagnostic to check if the approximation is sufficiently accurate (Sections 4.5), and 3) numerical equations to automatize updating the parameters in the fitting process (Sections 4.3.1).
+\item A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1 and Section 4.6), which is based on: 1) recommendations to set initial values for the hyperparameters, 2) diagnostic to check if the approximation is sufficiently accurate (Sections 4.5), and 3) numerical equations to automatize updating the parameters in the fitting process (Sections 4.3.1).
 
-\item Recommendations to perform diagnosis on any other covariance function different from those implemented in this study have been provided (Sections 4.2 and 4.7).
+\item Recommendations to perform diagnostics on any other covariance function different from those discussed in this study have been provided (Sections 4.2 and 4.7).
 
@@
 
-Authors are really grateful to the reviewer for their valuable comments which have been very helpful to improve the paper, and for her/his dedicated time.
+We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
 
@@
 The case study's models have also been formulated in general probabilistic language in the main manuscript, and Stan's implementations have also been provided as examples.
-Furthermore, HSGP model has been implemented in the \textit{brms} R-package that uses Stan to make model inference, and we are planning to implement the diagnostic tool which consists in iterative steps to fit and diagnose the model also in the \textit{brms} R-package.
-{\color{blue}(Gabriel: Paul, do we intend to implement the diagnostic procedure in brms?)}
+Furthermore, the HSGP model framework has been fully implemented in the \textit{brms} R-package that uses Stan as backend to make model inference.
+
+% and we are planning to implement the diagnostic tool which consists in iterative steps to fit and diagnose the model also in the \textit{brms} R-package.
+
+%Paul: Not in the near future; highly unlikely before the paper is accepted
+%{\color{blue}(Gabriel: Paul, do we intend to implement the diagnostic procedure in brms?)}
 
@@
 
-The performance of the HSGP model drastically depends on some setting parameters.
+The performance of the HSGP model drastically depends on some hyperparameters.
 The study pursues accurately measuring the sensitivity of the HSGP model to the values of these parameters, detecting when the approximation fails and when the parameter values are optimal in accuracy and computation cost.
-Ultimately, the aim of the study is to provide: 1) a diagnostic to check whether the approximation is sufficiently accurate, 2) numerical equations to update the setting parameters if the approximation is not sufficiently accurate, and 3) a safety and general procedure based on iterative steps to successfully fit the model, finding out optimal values for the setting parameters which allow for minimum computational requirements.
+Ultimately, the aim of the study is to provide: 1) a diagnostic to check whether the approximation is sufficiently accurate, 2) numerical equations to update the hyperparameters if the approximation is not sufficiently accurate, and 3) a safety and general procedure based on iterative steps to successfully fit the model, finding out optimal values for the hyperparameters which allow for minimum computational requirements.
 
-A theoretical approach to the diagnostic equations and rules is too complex that makes it not practical.
-The way to approach this is empirically.
-Making the empirical analysis we fully recognize the behavior of the approximation according to the parameters values, derive numerical equations gathering the relationships between the parameters that allow for updating the parameters and diagnosing the fitting.
+While we agree that a theoretical approach has a lot of merit, we do not see how such a theoretical approach to the diagnostic equations and rules would have been feasible.
+Thus we have approached these challenges empirically via numerical experiments, instead.
+By means of this empirical analysis, we got a thorough understanding of the behavior of the approximation as a function of the hyperparameters, derive numerical equations that capture the relationships between these parameters, which allow for an iterative procedure to ensure sufficient accuracy of the approximation.
 
-Furthermore, by following our recommended diagnostic procedure, an optimum solution with minimum computational requirements is achieved, that can be essential in repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional cases with very large data sets, and multi(low)-dimensional cases, even with small data sets, since knowing the optimum values for the setting parameters for each individual dimension can significantly reduce the computational requirements, otherwise fitting HSGPs would be unfeasible here.
+Furthermore, by following our recommended diagnostic procedure, an optimum solution with minimum computational requirements is achieved.
+This can be essential especially in cases where computation time is prohibitive, for example, when having to fit the model repeatedly (e.g., in cross-validation or simulation based calibration), uni-dimensional cases with very large data sets, or multi(low)-dimensional cases, even with small data sets.
+
+%since knowing the optimum values for the hyperparameters for each individual dimension can significantly reduce the computational requirements, otherwise fitting HSGPs would be unfeasible here.
 
@@
 We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate (Section 4.3.1).
-We have also elaborated a safety and general procedure based on iterative steps to fit the model with definite rules to update the parameters and make diagnosis (Sections 4.5 and 4.6).
+We have also added a general and robust iterative procedure to derive minimally sufficient hyperparameter values that ensure good GP approximations via our HSGP approach (Sections 4.5 and 4.6).
 
@@
 Kernel combinations that satisfy the requirements can also be used.
-We have also given recommendations to update $m$ and $c$ and perform diagnosis on any other different stationary kernel from those implemented in our study, by analyzing the constants characterizing the relationships for each covariance function (Sections 4.2 and 4.7).
-Explanations of how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).
+We have also given recommendations to update $m$ and $c$ and perform diagnostics on any other different stationary kernel from those implemented in our study, by analyzing the constants characterizing the relationships for each covariance function (Sections 4.2 and 4.7).
+Explanations on how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).
 
-The study deeply describes in Section 3 the formulation and implementation of the HSGP model, a description that can be used to implement it for any other different stationary kernel.
-The only new thing that the user needs to compute is the Fourier transform of the kernel to obtain the spectral density of the kernel (Section 3 and Section 4.7).
+In Section 3, we describe the formulation and implementation of the HSGP model in detail, a description that can be used to implement it for any other stationary kernel if required.
+The only kernel-specific function that the user needs to know is the Fourier transform of the kernel to obtain its spectral density (Section 3 and Section 4.7).
 
@@
 
-We do not really know the expertise of our users.
+Our experience tells us that the range of expertise in users of probabilistics programming languages such as Stan and their higher level interfaces such a brms is quite wide.
+We don't want to speculate on the percentage of users that would be able to come up with comparable solutions themselves.
+To study this empirically, one had to conduct extensive user studies and their are not in scope of this paper.
+We want to highlight though that rerunning of models and deciding about good updating rules takes a lot of human and compute time.
+We substantially reduce that time by reducing the number of model iterations, the computational requirements for each iteration, as well as eliminating the time an educated practioners would have had to spend to come up with some sensible rule themselves.
 
-Rerunning takes computer and human time and we reduce that time by reducing the number of iterations and the computational requirements for each iteration.
-
-Not all practitioners may be educated enough to come up with and adjust procedures themselves.
+%Not all practitioners may be educated enough to come up with and adjust procedures themselves.
 So clear, justified guidelines (that we provide) will be helpful and even required for the method to be used for many users.
 
-Apart from the number of basis function, the boundary factor $c$ plays a crucial role here and has to be set carefully as a function of the wigglyness (true length-scale) of the function to be learned.
+%Apart from the number of basis function, the boundary factor $c$ plays a crucial role here and has to be set carefully as a function of the wigglyness (true length-scale) of the function to be learned, which is unknown a prior.
 By following our diagnostic rules and equations, both $m$ and $c$ can be set and properly updated with minimal effort.
 
-An optimum solution with minimum computational requirements can be achieved following our recommended diagnostic procedure, otherwise it would be difficult.
+%An optimum solution with minimum computational requirements can be achieved following our recommended diagnostic procedure, otherwise it would be difficult.
 As discussed in the paper and above in this letter, using optimal values for the key parameters that allow for minimum computational requirements can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting.
@@
 
-The developed diagnostic tool and iterative steps to fit and perform diagnosis on the HSGP model have been made prominent in the case studies.
+The developed diagnostic tool and iterative steps to fit and perform diagnostics on the HSGP model have been made more prominent in the case studies, to better link them our main goals and the rest of the paper.
+In particular:
 
-The HSGP model is a new and computationally efficient representation of a GP.
-GPs are not simple statistical models.
+%The HSGP model is a new and computationally efficient representation of a GP.
 Improving the computational efficiency of a GP favors its implementation as modular components in flexible and complex modeling.
@@
 \begin{itemize}
+%\item Case study I involves a wiggly function with a Mat\'ern-3/2 kernel.
+
 \item Case study II involves a function that is composed of three additive GP components with different kernels, squared exponential and periodic kernels.
 Additionally this case study has a large data set of more than 7000 observations which would be unfeasible to be fitted using a regular formulation of the GP model.
-
-\item Case study IV (Leukemia data set) is a 4D function with a multilevel GP structure.
-This case study is also a good example that by using our diagnostic tool you can easily get an optimal solution in 4D which makes it significantly faster than regular GPs.
-Otherwise, user would had probably been struggling searching for useful values for the parameters without any guide.
-\end{itemize}
-
-The only limitation of the method is the expoential scaling of the computation with increasing number of dimensions (assuming wigglines stays the same and the amount of data is sufficient to get sufficient informationa about that wiggliness).
-Due to the high dependency of the computational requirements of the method on the number of basis functions, the analysis and diagnosis tool developed in this paper arises to be essential for practical applications and in broad scenarios.
-
-\begin{itemize}
-\item Case study I involves a wiggly function with a Mat\'ern-3/2 kernel.
 
@@
 We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D cases.
+
+\item Case study IV (Leukemia data set) is a 4D function with a multilevel GP structure.
+This case study is also a good example that by using our diagnostic tool you can easily get an optimal solution in 4D which makes it significantly faster than regular GPs.
+Otherwise, user would had probably been struggling searching for useful values for the hyperparameters without any guidance.
 \end{itemize}
+
+The only main limitation of the HSGP method is the expoential scaling of the computation with increasing number of input dimensions (assuming wigglines stays the same and the amount of data is sufficient to get sufficient information about that wiggliness).
+Due to the high dependency of the computational requirements of the method on the number of basis functions, the analysis and diagnostics tool developed in this paper arises to be essential for practical application in a wide range of scenarios.
 
@@
 We have simplified the section significantly but keeping references to equations in the main text of how to implement the approximation function.
-We have used the diagnosis tool and iterative steps to fit and perform diagnosis on the models.
+We have used the diagnostics tool and iterative steps to fit and perform diagnostics on the models.
 We have also improved the readability of the section.
@@
 This case study involves an underlying function that is composed of three additive GP functions with different kernels, squared exponential and a periodic kernels.
-We have applied the diagnostic tool, describing the iterative steps to fit this model and performing diagnosis on the three additive GP components at the same time.
-This example also serves as an example of how useful our diagnosis can be in large 1D data sets.
+We have applied the diagnostic tool, describing the iterative steps to fit this model and performing diagnostics on the three additive GP components at the same time.
+This example also serves as an example of how useful our diagnostics can be in large 1D data sets.
 
@@
 It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
-For that reason, in fact, the use of our diagnosis tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
+For that reason, in fact, the use of our diagnostics tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
@@
 We also analyze how the key factors of the model relate each other for accurate approximations.
-We derive numerical functions for these relationships that allow to update the parameters, perform diagnosis, and provide clear and useful guidelines for the whole procedure.
+We derive numerical functions for these relationships that allow to update the parameters, perform diagnostics, and provide clear and useful guidelines for the whole procedure.
 So, we make it manageable for practical users, otherwise the practical applicability of the method would be highly limited.
@@
 
-What we can definitely state, which is, in fact, what our diagnosis is based on, is the following:
+What we can definitely state, which is, in fact, what our diagnostics is based on, is the following:
 
@@
 
-However, when $c$ is not large enough given $\ell^*$, $\hat{\ell}$ can be larger than $\ell^*$ and the approximation being inaccurate, so the diagnosis will fail in this case, for example:
+However, when $c$ is not large enough given $\ell^*$, $\hat{\ell}$ can be larger than $\ell^*$ and the approximation being inaccurate, so the diagnostics will fail in this case, for example:
 
@@
 
-It shows that it is essential to follow our diagnosis tool to be sure we are using valid values in each iterative step.
+It shows that it is essential to follow our diagnostics tool to be sure we are using valid values in each iterative step.
 
@@
 Future research will focus on building analytical models that provide the diagnostics in multi-dimensional cases.
-However, as an approximation, we can use the diagnostic for unidimensional functions to make diagnosis individually on the different dimensions of a multidimensional GP model.
+However, as an approximation, we can use the diagnostic for unidimensional functions to make diagnostics individually on the different dimensions of a multidimensional GP model.
 
@@
 
-From the experiments, we have built numerical functions that characterize the relationships between the key parameters of the model which are the basis to perform diagnosis on the fitting.
+From the experiments, we have built numerical functions that characterize the relationships between the key parameters of the model which are the basis to perform diagnostics on the fitting.
 We have developed a diagnostic procedure to fit, diagnose and update parameters, based on iterative steps until an accurate solution is reached.

### 2022-03-08 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
finish editing response to reviewer 1

@@
 We have simplified the section significantly but keeping references to equations in the main text of how to implement the approximation function.
-We have used the diagnostics tool and iterative steps to fit and perform diagnostics on the models.
+We have used the diagnostic tools and iterative steps to fit and perform diagnostics on the models.
 We have also improved the readability of the section.
@@
 
-This case study involves an underlying function that is composed of three additive GP functions with different kernels, squared exponential and a periodic kernels.
-We have applied the diagnostic tool, describing the iterative steps to fit this model and performing diagnostics on the three additive GP components at the same time.
-This example also serves as an example of how useful our diagnostics can be in large 1D data sets.
+Following your suggestions, we improve the case study to be better integrated with the rest of the paper.
+This case study involves an underlying function that is composed of three additive GP functions with different kernels, two squared exponential kernels and a periodic kernel.
+We have now applied the diagnostic tool, describing the iterative steps to fit this model and perform diagnostics on the three additive GP components at the same time.
+This case study also serves as an example of how useful our diagnostics can be in large data sets with 1D HSGPs.
 
@@
 
-The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already fast.
+The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
 
 It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
-For that reason, in fact, the use of our diagnostics tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
+For that reason, in fact, the use of our diagnostics tool can be essential because it allows to iteratively fit the model using minimum computational requirements in each iteration.
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
-For $D>5$ even for very smooth functions seems to be impractical.
+For $D>5$, even for very smooth functions, HSGPS definitely seem to be impractical.
+For $D<3$ it is widely useful and significantly faster than the exact GP even for very wiggly functions.
 
@@
 In fact using the diagnostic tool the user will discover easily the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP.
-Without the guide of the diagnostic tool the user probably would had made several trials until finding out an accurate and practical solution, consuming a lot of computation requirements in this demanding 4D case.
+Without the guide of the diagnostic tool, the user would have likely made several trials until finding an accurate and practical solution, consuming a lot of computation requirements in this highly demanding 4D case.
+Insofar, the understanding that "its applicability to multidimensional cases is mostly negative" is not fully accurate, as it ultimately depends on the wigglyness of the function.
+We now better discuss this point in various places in the paper.
+Specifically, in Section 4.8, we provide details on the performance of the model in multi-dimensional cases.
 
-The statement "its applicability to multidimensional cases is mostly negative" is inaccurate, as it ultimately depends on the wigglyness of the function.
+%the method is limited in practice to the number of dimensions and the wigglyness of the function across the different dimensions.
+We have observed that it can be useful for $D<5$ depending on the wigglyness of the function.
+.
 
-As commented across the paper and in particular in Section 4.8, the method is limited in practice to the number of dimensions and the wigglyness of the function across the different dimensions.
-We have observed that it can be useful for $D<5$ conditioned to the wigglyness of the function.
-However, for $D=2$ it is still useful and significantly faster than the exact GP even for very wiggly functions.
-In Section 4.8 we give brief details on the performance of the model in multi-D cases.
-In this sense, the use of our recommended diagnostic allows to find the optimal values for $c$ and $m$ which allows for minimum computational requirements which can be essential on these multidimensional cases.
+%In this sense, the use of our recommended diagnostic allows to find the optimal values for $c$ and $m$ which allows for minimum computational requirements which can be essential on these multidimensional cases.
 
-We have added a new case study to Section 5 which consists in wiggly 2D and 3D simulated data sets.
-We have applied the diagnostic procedure on it.
-Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing individually the different dimensions in each iteration.
-It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
-We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D.
+We have also added a new case study to Section 5 which consists of wiggly 2D and 3D simulated data sets, where have applied the diagnostic procedure too.
+Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing the different dimensions in each iteration individually.
+%It allows us to find the optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
+In this new case study, we have also included a brief discussion and presentation of additional results concerning computational cost in 2D, 3D, and 4D.
 
@@
 
-Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, and this work is necessary to facilitate the easier use of the method for practical users.
+Thank you again for your thorough review.
+To us, also empirical analysis provides scientific insights, in particular as we can study much more complicated scenarios that we cannot feasibly approach my mathematical analysis alone.
+As such, we think our paper nicely complements existing research.
+
+Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, while our work is necessary to facilitate the easier use of the method for practical users.
 We also analyze how the key factors of the model relate each other for accurate approximations.
@@
 
-We expect this paper to have an impact on how widely the basis function GPs are used, as it is makes the method more usable in practice and in broad scenarios like uni-dimensional functions, multi-dimensional functions, and other problems that require repeated fitting such as cross-validation and simulation based calibration.
+We expect this paper to have a strong impact on how widely basis function GPs are used.
+Our contribution makes these approximate GPs more practically applicable for both uni-dimensional and multi-dimensional functions, even in combination with additional challenges that require repeated fitting such as cross-validation or simulation based calibration.
 
@@
 
-An inaccurate approximation implies that number of basis functions $m$ is not enough and/or the boundary factor $c$ is not adequate.
+We have substantially extended the discussion around our proposed diagnostic.
+In short, an inaccurate approximation implies that the number of basis functions $m$ is not enough and/or the boundary factor $c$ is too small.
+Our experiments yield consistently that, if $c$ is large enough, $\hat{\ell}$ will be below that minimum supported length-scale ($\ell^*$) determined by $m$ and $c$; and observation on which our diagnostic procedure rests.
+We cannot rule out with certainty that and inaccurate approximation might also yield an $\hat{\ell}$ above the minimum supported length-scale ($\ell^*$), too.
+However, we have not encountered this case in our experiments.
 
-We can not state anything definite about the estimated lengthscale $\hat{\ell}$ when the approximation is inaccurate, basically because it might ultimately depend on the specific frequency patterns of the signal/function to be learned, and also because $c$ must be properly set.
+As demonstrated in Section 4.6 and our case studies, our developed diagnostic rules leads to a reliable, useful, and fast iterative procedure to fit the models with minimum computational requirements.
+More arguments supporting this notion have been included in the paper in Section 4.5.
+
+%We cannot state anything definitive about the estimated lengthscale $\hat{\ell}$ when the approximation is inaccurate, basically because it might ultimately depend on the specific frequency patterns of the signal/function to be learned, and also because $c$ must be properly set.
 Although, we can state that for most of the cases, if $c$ is large enough, $\hat{\ell}$ will be below that minimum supported length-scale ($\ell^*$) determined by $m$ and $c$.
-But, as said before, it can not be generalized to all case studies because it might depend on the particular frequency patterns of a signal.
+But, as said before, it cannot be generalized to all case studies because it might depend on the particular frequency patterns of a signal.
 Future research will be focused on investigating potential signals with particular frequency patterns that could escape from this rule.
 
-What we can definitely state, which is, in fact, what our diagnostics is based on, is the following:
+%What we can say more definitely is the following:
 
-\begin{enumerate}
-\item  Observing an $\hat{\ell}$ smaller than $\ell^*$ (i.e.
-$\hat{\ell}$ is overestimated), and $c$ is large enough given $\ell^*$, indicates a potential problem with the approximation.
-This happens because in inaccurate approximation scenarios the true length-scale is smaller than $\ell^*$, so $\hat{\ell}$ will tend to be towards smaller values than $\ell^*$.
+%\begin{enumerate}
+%\item Observing an $\hat{\ell}$ smaller than $\ell^*$, for $c$ large enough given $\ell^*$, indicates a potential problem with the approximation.
+This happens because in such scenarios the true length-scale is smaller than $\ell^*$, so $\hat{\ell}$ will tend to take on values smaller than $\ell^*$.
 
-\item The other way around, that is, observing $\hat{\ell}$ larger than $\ell^*$ (i.e.
-$\hat{\ell}$ is underestimated), the approximation is assumed to be accurate because when the true length-scale is larger than $\ell^*$, $\hat{\ell}$ will tend to be larger than $\ell^*$.
+%\item The other way around, that is, observing $\hat{\ell}$ larger than $\ell^*$, the approximation is assumed to be accurate because when the true length-scale is larger than $\ell^*$, $\hat{\ell}$ will tend to be larger than $\ell^*$.
 However, we do not still have full evidence that this can be generalized to all possible cases as explained above (but, we can state that this rule can be apply for most of the the practical cases).
@@
 
-As largely demonstrated in the paper in Section 4.6 and Case studies, these diagnostic rules leads to a safety, useful and fast iterative procedure to fit the models with minimum computational requirements.
-
-These arguments have been included in the paper in Section 4.5. 
-\end{enumerate}
+%\end{enumerate}
 
 \textbf{\#1.2 - To me it seems Fig 5 (left) already hints that you tend to underestimate l when using bad approximation and the procedure may be valid in practice, but this is not transparent in the paper.}
+
+We agree and have made this observation and corresponding assumptions more explicit in the paper. Below, we detail some results corresponding to Figure 5 for your convenience.
 
@@
 5 (left) by using equation (18).
-Following we show a few of these inaccurate cases in Fig.
+
+%Following we show a few of these inaccurate cases in Fig.
 5 (left), showing diagnostic $\hat{\ell} < \ell^*$ can be trusted to diagnose a bad approximation, as long as $c$ is large enough given $\ell^*$:
@@
 
-It shows that it is essential to follow our diagnostics tool to be sure we are using valid values in each iterative step.
+Together, this demonstrates that it is essential to follow our diagnostic procedure to ensure using valid values in each iterative step.
 
@@
 
-See response in \#1.1 above.
+Yes. Please see our response to your comment \#1.1 above.
 
@@
 
-We have developed and presented a diagnostic procedure that consists in iterative steps to fit and diagnose the model.
+In response, we have developed and presented a diagnostic procedure that consists of iterative steps to fit and diagnose the model.
 Numerical equations are used to set initial values and update the key parameters in each iterative step.
-The recommended diagnostic procedure is a safety and extensive procedure that allows to automatize the applicability of the method.
-See details in Section 4.5 and examples in Section 4.6, Case studies and Appendix C.
+The recommended diagnostic procedure allows to automatize the application of the method.
+In the revised manuscript, please see Section 4.5 and examples in Section 4.6, Case studies, and Appendix C, for more details.
 
@@
 
-Section 4 have been re-structured and updated by adding new subsections, such as a specific section on the theoretical evidence of near linear proportionality among the key parameters of the model, another section with the experimentation and resulting numerical diagnostic equations, another section supporting, detailing and discussing on the diagnostic rules, and another analyzing the performance of the diagnostic tool.
-Additionally, new subsections regarding how to proceed with other different kernels from those implemented in this study and some insights on how the method performs in multi-dimensional cases have also been added.
+Section 4 has been re-structured and updated by adding new subsections.
+In particular this includes, a subsection on the theoretical evidence of near linear proportionality among the key hyperparameters of the model, a subsection with the experimentation and resulting numerical diagnostic equations, a subsection supporting, detailing, and discussing on the diagnostic rules, and a subsection analyzing the performance of the diagnostic tool.
+Additionally, new subsections regarding how to proceed with other different kernels from those implemented in this study and some insights on how the method performs in multi-dimensional cases have been added.
 
-We have simplified and improved significantly the writing of Section 5.
-We have also focused the case studies more on the main goals of the paper, which are: 1) to facilitate the implementation of the method for practical users providing clear formulation of the method, and 2) and to provide a safety and general diagnostic tool to fit and diagnose the model.
+We have simplified and significantly improved the writing in Section 5.
+We have also focused the case studies more on the main goals of the paper, which are: 1) to facilitate the implementation of the method for practical users providing clear a formulation of the method, and 2) and to provide an easy and general diagnostic tool to achieve good approximations via HSGPs.
 
@@
 
-Diagnostic has been build for GPs with unidimensional covariance functions.
-Future research will focus on building analytical models that provide the diagnostics in multi-dimensional cases.
-However, as an approximation, we can use the diagnostic for unidimensional functions to make diagnostics individually on the different dimensions of a multidimensional GP model.
-
-See comment \#9 above for an extended answer.
+Our diagnostic has been build for GPs with unidimensional covariance functions.
+Future research will focus on building analytical models that provide the diagnostic for multi-dimensional cases, as well.
+However, for now, we can use the diagnostic for unidimensional functions and apply it to each dimensionan of a multidimensional GP model individually, which seems to works quite well in our examples.
+Please see comment \#9 above for an extended answer on this topic.
 

### 2022-03-08 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
edits to the responses to reviewer 2

@@
 The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
-
 It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
@@
 
-Authors are really grateful to the reviewer for their valuable comments which have been very helpful to improve the paper, and for her/his time.
+We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
 
@@
 
-From the experiments, we have built numerical functions that characterize the relationships between the key parameters of the model which are the basis to perform diagnostics on the fitting.
-We have developed a diagnostic procedure to fit, diagnose and update parameters, based on iterative steps until an accurate solution is reached.
-Furthermore, this diagnostic tool provides the optimal values for the key parameters of the model that allow for minimizing the computational cost, which can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting such as cross-validation and simulated based calibration.
+In the revised version of the paper, we have substantially extended the experiments.
+On that basis, we have derived equations that characterize the relationships between the key hyperparameters of the HSGP model and  the goodness of the corresponding exact GP approximation.
+We have developed a procedure to diagnose the appropriateness of the hyperparameters and to update them based on iterative steps until an accurate approximation has been achieved.
+Furthermore, this diagnostic tool provides optimal values for those hyperparameters with respect to minimizing the computational cost subject to sufficient accuracy of the approximation.
+Minimizing computation cost is particularily important for very large data sets, multi-dimensional cases, or other problems that require repeated fitting, such as cross-validation and simulated based calibration.
+All of this new content is discussed in detail in Section 4 of the paper.
 
-The diagnostic rules, the numerical functions that allow updating the parameters and the recommended iterative steps to fit and diagnose the model are extensive to the community and essential to make the model usable for practitioners.
-
-All of this have been properly explained along Section 4 in the paper.
+%The diagnostic rules, the numerical functions that allow updating the parameters and the recommended iterative steps to fit and diagnose the model are extensive to the community and essential to make the model usable for practitioners.
 
@@
 
-We have mention in the paper how dimensionality affects the expected number of basis functions needed and that the effect of dimensionality is only through this.
+Indeed, the main limitation of the HSGP approach is its bad scaling with increasing number of input dimensions.
+In the revised version, we now make this more explicit and discuss the implications in detail, in particular in Section 4.8.
 
-The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already fast.
-It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
-For that reason, in fact, the use of our diagnostics tool can be essential because it allows iteratively to fit the model using minimum computational requirements in each iteration.
+The HSGP model is definitely highly useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast.
+It is difficult to give clear guidelines of when the application of the HSGP model becomes impractical in multidimensional cases, because it also depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
+For that reason, the use of our diagnostic tool is highly beneficial to minimize computational requirements in order to achieve a good approximation.
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
-For $D>5$ even for very smooth functions seems to be impractical.
+For $D>5$, even for very smooth functions, HSGPs indeed seem to be impractical and we now state this explicitely.
 
-The Leukemia data set case study is a 4D data set.
-We have applied applied the diagnostic tool on it and we have been able to easily discover the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP.
-Without the guide of the diagnostic tool the user probably would had made several trials until finding out an accurate and practical solution, consuming a lot of requirements in this demanding 4D case.
+In the Leukemia data set case study, we present a 4D HSGP model.
+We have applied the diagnostic tool and have been able to easily discover the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP.
+Without the guidance of the diagnostic tool, a user would have to go through several more trials until finding a sufficiently accurate solution, which would result in substantially increased human and computer time requirements.
 
-We have added a new case study to Section 5 which consists in wiggly 2D and 3D simulated data sets.
-We have applied the diagnostic procedure on it.
-Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing individually the different dimensions in each iteration.
-It allows finding the  optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
-We have included in this new case study a brief discussion and presentation of additional results concerning computational cost in 2D, 3D and 4D.
+We have also added a new case study to Section 5 which consists of wiggly 2D and 3D simulated data sets, where have applied the diagnostic procedure too.
+Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing the different dimensions in each iteration individually.
+%It allows us to find the optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
+In this new case study, we have also included a brief discussion and presentation of additional results concerning computational cost in 2D, 3D, and 4D.
 
-As commented across the paper and in particular in Section 4.8, the method is limited in practice to the number of dimensions and the wigglyness of the function across the different dimensions.
+%As commented across the paper and in particular in Section 4.8, the method is limited in practice to the number of dimensions and the wigglyness of the function across the different dimensions.
 We have observed that it can be useful for $D<5$ conditioned to the wigglyness of the function.
@@
 "MCMC for Variationally Sparse Gaussian Processes" ; although this is SVI, it appears directly applicable for the problems considered here.
-Please also compare with non MCMC variational methods - while not directly applicable to generic MCMC methods, these are the sttandard and it would be nice to see how they compare in runtime.}
+Please also compare with non MCMC variational methods - while not directly applicable to generic MCMC methods, these are the standard and it would be nice to see how they compare in runtime.}
 
-The main comparison is the Hilbert space basis function approximation against the exact GP as a model.
-In all comparisons the inference conditional on the model is assumed to be accurate.
-We've used efficient dynamic HMC, the convergence diagnostics didn't reveal problems, and thus the accuracy of the inference is good.
-Comparisons of MCMC to variational inference approximations are out of scope of this paper.
+The main comparison that we focus on is the Hilbert space basis function approximation (our HSGP method) against the exact GP model.
+In all comparisons, the inference conditional on the model is assumed to be accurate.
+In other words, we compare the exact to an approximate model, using asymptotically correct inference.
+For this purpose, we have used efficient dynamic HMC, the MCMC convergence diagnostics did not reveal problems, and thus the accuracy of the inference itself can be considered good.
+
+While we agree that other comparisons could interesting as well, for example MCMC vs.
+variational inference, they would blow up the already extensive scope of the paper.
 Some comparisons of MCMC and variational inference for basis function GPs can be found in
@@
 
-Comparisons of the Hilbert space basis function approximations against other GP approximations are provided by Solin and Särkkä (2020), which is cited prominently in our paper.
+Further comparisons of the Hilbert space basis function approximations against other GP approximations are provided by Solin and Särkkä (2020), which is cited prominently in our paper.
 

### 2022-03-09 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
Minor edits to first part of reply letter

@@
 So the whole study provides the explicit diagnostic for a squared exponential, Mat\'ern-3/2, Mat\'ern-5/2 (Sections 4.3 and 4.3.1), and periodic squared exponential (Sections 3.3, 4.4 and Appendix C) covariance functions.
-In addition connection to other covariance functions is discussed.
+In addition, the connection to other covariance functions is discussed.
 
@@
 
-\item We have derived numerical equations that characterize the relationships between the hyperparameters of the model for each particular kernel, which are useful to update and find out optimal values for the hyper-parameters (Section 4.3.1).
+\item We have derived numerical equations that characterize the relationships between the hyperparameters of the model for each particular kernel, which are useful to update and determine optimal values for the hyper-parameters (Section 4.3.1).
 
@@
 \item We have added a new case study to Section 5 which consists of a wiggly 2D simulated data set, demonstrating the applicability and usefulness of the diagnostic tool in multidimensional cases.
-Additional results concerning computational cost in 2D, 3D and 4D have been presented and briefly discussed, emphasizing that applying the diagnostic tool allows for minimum computational requirements that can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting.
+Additional results concerning the computational cost in 2D, 3D and 4D have been presented and briefly discussed, emphasizing that the diagnostic facilitates a reduction of the required computational ressources, which can be essential for very large data sets, multi-dimensional cases, or other problems that require repeated fitting.
 

### 2022-03-09 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
Minor edits to R1 comments in reply letter

@@
 We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate (Section 4.3.1).
-We have also added a general and robust iterative procedure to derive minimally sufficient hyperparameter values that ensure good GP approximations via our HSGP approach (Sections 4.5 and 4.6).
+We have also added a general and robust iterative procedure to derive minimally sufficient hyperparameter values that ensure accurate GP approximations via our HSGP approach (Sections 4.5 and 4.6).
 
@@
 We have also given recommendations to update $m$ and $c$ and perform diagnostics on any other different stationary kernel from those implemented in our study, by analyzing the constants characterizing the relationships for each covariance function (Sections 4.2 and 4.7).
-Explanations on how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).
 
 In Section 3, we describe the formulation and implementation of the HSGP model in detail, a description that can be used to implement it for any other stationary kernel if required.
-The only kernel-specific function that the user needs to know is the Fourier transform of the kernel to obtain its spectral density (Section 3 and Section 4.7).
+The only kernel-specific function that the user needs to know is the Fourier transform of the kernel to obtain its spectral density (Section 3).
+Moreover, we discuss how to estimate the number of basis functions needed for an accurate approximation for any other (stationary) covariance function have also been provided (Section 4.7).
 
@@
 To study this empirically, one had to conduct extensive user studies and their are not in scope of this paper.
-We want to highlight though that rerunning of models and deciding about good updating rules takes a lot of human and compute time.
+We want to highlight, though, that rerunning models iteratively for estimating reasonable hyperparameter values often takes a lot of human and compute time.
 We substantially reduce that time by reducing the number of model iterations, the computational requirements for each iteration, as well as eliminating the time an educated practioners would have had to spend to come up with some sensible rule themselves.
@@
 This case study is also a good example that by using our diagnostic tool you can easily get an optimal solution in 4D which makes it significantly faster than regular GPs.
-Otherwise, user would had probably been struggling searching for useful values for the hyperparameters without any guidance.
+% Otherwise, the user would had probably been struggling searching for useful values for the hyperparameters without any guidance.
 \end{itemize}
@@
 The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
-It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown.
+It is difficult to give specific guidelines for when the HSGP model will be impractical for multidimensional problems, because it ultimately depends on the wigglyness/ of the function to be learned which is, in most of the cases, unknown.
 For that reason, in fact, the use of our diagnostics tool can be essential because it allows to iteratively fit the model using minimum computational requirements in each iteration.
 Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied.
-For $D>5$, even for very smooth functions, HSGPS definitely seem to be impractical.
+For $D>5$, even for very smooth functions, HSGPs definitely seem to be impractical.
 For $D<3$ it is widely useful and significantly faster than the exact GP even for very wiggly functions.
 
-We disagree with the reviewer that the HSGP performs poorer than the exact GP on this Leukemia data set.
-In fact using the diagnostic tool the user will discover easily the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP.
-Without the guide of the diagnostic tool, the user would have likely made several trials until finding an accurate and practical solution, consuming a lot of computation requirements in this highly demanding 4D case.
+We disagree with the reviewer that the HSGP performs worse than the exact GP on this Leukemia data set.
+In fact, using the diagnostic tool the user will easily discover the optimal values for $c$ and $m$ that implies significantly faster computation than using the exact GP.
+Without the guide of the diagnostic tool, the user could easily end up using several trials until finding an accurate and practical solution, consuming a lot of computation requirements in this highly demanding 4D case.
 Insofar, the understanding that "its applicability to multidimensional cases is mostly negative" is not fully accurate, as it ultimately depends on the wigglyness of the function.
@@
 Thank you again for your thorough review.
-To us, also empirical analysis provides scientific insights, in particular as we can study much more complicated scenarios that we cannot feasibly approach my mathematical analysis alone.
+To us, empirical analysis also provides scientific insights, in particular as we can study much more complex scenarios that we cannot feasibly approach via mathematical analysis alone.
 As such, we think our paper nicely complements existing research.
 
-Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, while our work is necessary to facilitate the easier use of the method for practical users.
-We also analyze how the key factors of the model relate each other for accurate approximations.
+Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, while our work is necessary to decribe and facilitate applied use of the approximation.
+We also analyze how the key factors of the model relate to each other for accurate approximations.
 We derive numerical functions for these relationships that allow to update the parameters, perform diagnostics, and provide clear and useful guidelines for the whole procedure.

### 2022-03-09 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
minor edits to intro and conclusion

@@
 Another advantage is the reduced memory requirements of automatic differentiation methods used in modern probabilistic programming frameworks, such as Stan \citep{carpenter2017stan} and others.
-This is because the memory requirements of automatic differentiation scale with the size of the autodiff expression tree which in direct implementations is simpler for basis function than covariance matrix based approach.
+This is because the memory requirements of automatic differentiation scale with the size of the autodiff expression tree which in direct implementations is simpler for basis function than covariance matrix-based approaches.
 The basis function approach also provides an easy way to apply a non-centered parameterization of GPs, which reduces the posterior dependency between parameters representing the estimated function and the hyperparameters of the covariance function, which further improves MCMC efficiency.
@@
 Practical recommendations are given for the values of the key factors based on simple diagnostic values and intuitive graphical summaries that encode the recognized relationships.
-Our recommendations will help users to choose valid and optimized values for these factors, improving computational performance without sacrificing modeling accuracy.
+Our recommendations help users to choose valid and optimized values for these factors, improving computational performance without sacrificing modeling accuracy.
 The proposed diagnostic can be used to check whether the chosen values for the number of basis functions and the domain of the prediction space are adequate to model the data well.
+On that basis, we also develop an iterative procedure to achieve accuracte approximation performance with minimal computational costs.
 
@@
 In Section~\ref{sec_method}, the reduced-rank approximation to GPs proposed by \citet{solin2018hilbert} is described.
-In Section~\ref{sec_accuracy}, the accuracy of these approximations under several conditions using analytical and numerical methods is analyzed.
+In Section~\ref{sec_accuracy}, the accuracy of these approximations under several conditions is studied using both analytical and numerical methods.
+Practical diagnostics are developed there as well.
 Several case studies in which we fit exact and approximate GPs to real and simulated data are provided in Section~\ref{sec_cases}.
@@
 The method has an attractive computational cost as it effectively approximates GPs by linear models, which is also an attractive property in modular probabilistic programming programming frameworks.
-The dominating cost per log density evaluation (during sampling) is $O(nm+m)$, which is a big benefit in comparison to $O(n^3)$ of a exact GP model.
+The dominating cost per log density evaluation (during sampling) is $O(nm+m)$, which is a big benefit in comparison to $O(n^3)$ of an exact GP model.
 The obtained design matrix is independent of hyperparameters and therefore only needs to be constructed once, at cost $O(nm)$.
@@
 
-The main contribution of this paper is an in-depth analysis and diagnostic of the performance and accuracy of the approximation in relation to the key factors of the method, that is, the number of basis functions, the boundary condition of the Laplace eigenfunctions, and the non-linearity of the function to be learned.
-Recommendations for the values of these key factors based on the recognized relations among them have been provided along with illustrations of these relations.
+As one of the main contributions of this paper, we provided an in-depth analysis of the approximation's performance and accuracy in relation to the key factors of the method, that is, the number of basis functions, the boundary condition of the Laplace eigenfunctions, and the non-linearity of the function to be learned.
+On that basis, as our second main contribution, we developed practical diagnostics to assess the approximation's performance as well as an iterative procedure to obtain an accurate approximation with minimal computational costs.
+
+%We illustrated our findings in several case studies of real and simulated data. 
+%%Recommendations for the values of these key factors based on the recognized relations among them have been provided.
 These illustrations will not only help users to improve performance and save computation time, but also serve as a powerful diagnostic tool whether the chosen values for the number of basis functions and the boundary condition are adequate to fit to the data at hand with sufficient accuracy.
@@
 
-The obtained functional relationships between the key factors influencing the approximation not only help users to visually assess the accuracy of the method
-but can also serve an automatic diagnostic tool.
-In this paper, we primarily studied the functional relationships for univariate inputs.
+In this paper, the obtained functional relationships between the key factors influencing the approximation and corresponding diagnostics were studied primarily for univariate inputs.
 Accordingly, investigating the functional relationships more thoroughly for multivariate inputs remains a topic for future research.

### 2022-03-09 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
More edits to R1 responses in reply letter

@@
 
-In response, we have developed and presented a diagnostic procedure that consists of iterative steps to fit and diagnose the model.
+We completely agree and consequently, we have developed and described an automatic iterative diagnostic procedure for fitting and diagnosing the model.
 Numerical equations are used to set initial values and update the key parameters in each iterative step.
-The recommended diagnostic procedure allows to automatize the application of the method.
+% The recommended diagnostic procedure allows to automatize the application of the method. 
 In the revised manuscript, please see Section 4.5 and examples in Section 4.6, Case studies, and Appendix C, for more details.
@@
 
-Section 4 has been re-structured and updated by adding new subsections.
-In particular this includes, a subsection on the theoretical evidence of near linear proportionality among the key hyperparameters of the model, a subsection with the experimentation and resulting numerical diagnostic equations, a subsection supporting, detailing, and discussing on the diagnostic rules, and a subsection analyzing the performance of the diagnostic tool.
-Additionally, new subsections regarding how to proceed with other different kernels from those implemented in this study and some insights on how the method performs in multi-dimensional cases have been added.
+Section 4 has been re-structured and updated by adding new subsections:
+\begin{itemize}
+	\item a subsection on the theoretical evidence of near linear proportionality among the key hyperparameters of the model
+	\item a subsection with the experimentation and resulting numerical diagnostic equations
+	\item a subsection supporting, detailing, and discussing on the diagnostic rules
+	\item a subsection analyzing the performance of the diagnostic tool
+\end{itemize}
+
+%Additionally, new subsections regarding how to use other kernels from those implemented in this study and some insights on how the method performs in multi-dimensional cases have been added.
 
 We have simplified and significantly improved the writing in Section 5.
-We have also focused the case studies more on the main goals of the paper, which are: 1) to facilitate the implementation of the method for practical users providing clear a formulation of the method, and 2) and to provide an easy and general diagnostic tool to achieve good approximations via HSGPs.
+We have also adapted the case studies to be more aligned with the main goals of the paper, which are: 1) to facilitate the implementation of the method for practical users providing clear a formulation of the method, and 2) and to provide an easy and general diagnostic tool to achieve good approximations via HSGPs.
 

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
add case study III and update section 4.8

--- 2022-03-07 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
@@
 
-%In the online supplemental material, additional case studies are presented. From those examples, it can be seen how 
-The computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-In the experiments of this paper computation for HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
-For small and moderate sized datasets, HSGPs is likely to be slower than exact GPs already for $D=3$ and $D=4$ even with a relatively low number of basis functions ($m \gtrsim 5$ and $m \gtrsim 20$ respectively; see Figure~C.3 in the online material).
+The computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
 The HSGP method can be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
-In all of the investigated cases, choosing the optimal booundary factor $c$ reduces the required number of basis functions $m$ noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore  reduces computational time drastically in particular in multidimensional cases.
+
+The HSGP model is definitely useful in $1$D and $2$D, even for highly wiggly functions, unless in $2$D the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
+
+In the experiments of this paper for multivariate cases, mostly reported in Section~\ref{sec_computation_nD}, computation for HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderated-large sized ($n>1000$) $3$D datasets (even highly wiggly functions with large $m^* \approx 10000$); see Figure~\ref{fig_time_nD_log}).
+
+For small sized ($n<1000$) datasets, HSGPs is likely to be slower than exact GPs already for $D=3$ and $D=4$, with a relatively large number of multivariate basis functions ($m^* > 3000$) in $3$D, and with a relatively low number of multivariate basis functions ($m^* \gtrsim 3000$) in $4$D (see Figure~\ref{fig_time_nD_log}).
+
+%In all of the investigated cases, choosing the optimal booundary factor $c$ reduces the required number of basis functions $m$ noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore  reduces computational time drastically in particular in multidimensional cases.
+
+As it has been shown in case study III (Section~\ref{sec_caseIII}), the iterative diagnostic tool proposed in this study can be very useful for the multivariate cases as it allows to reduce $m^*$ to the minimum, reducing computational time drastically, and still getting an accurate approximation.
+Otherwise, user might be struggling in these multivariate cases.
+
 When HSGP or exact GP is faster depend also on the implementation details that can have big effect on the constants, and thus any more detailed recommendations can only be made in the context of the specific software implementation.
@@
 Initial values for the minimum lengthscales $\ell_2$ and $\ell_3$ of $f_2$ and $f_3$, respectively, are guessed to correspond to half of the period (i.e., $l_2=l_3=0.5$).
+After the diagnostic generated the first \textit{true}, $m$ is updated by increasing the $m$ of the previous iteration by 5 additional basis functions and $c$ and the minimum $\ell$ updated accordingly as explained in Section~\ref{sec_user_guide}.
+The full diagnosis process is applied until two \textit{trues} are achieved for each function.
+Table~\ref{tab_caseII} contains the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ for each function, and the RMSE compute with the data as the reference for every iterative steps of the fitting process.
 
@@
 The full diagnosis process is applied until two \textit{trues} are achieved for each function.}
-  \label{tab_caseI}
+  \label{tab_caseII}
 \end{table}
@@
 
-\subsection{Case study of simulated data for a bivariate function}
-
-This example consists of a synthetic dataset with $n=120$ ($i=1,\dots,n$) single draws from a GP prior as a function of input values $\bm{x}_i \in \{[-1,1],[-1,1]\} \subset {\rm I\!R}^2$ in a bivariate input space ($D=2$).
-A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, for the first input dimension, and $\ell_2=0.35$, for the second input dimension, is used for the GP prior.
-Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy set of observations $\bm{y} \in {\rm I\!R}^{120}$.
-
-In the HSGP model with $2$ input dimensions, the latent function $f(\bm{x})$ is approximated as in equation (13)% (\ref{eq_approxf_multi})
-, with the $2$D squared exponential spectral density $S$ as in eq.
-\eqref{eq_specdens_inf}, and the $2$-vector of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations \eqref{eq_eigenfunction_multi} and \eqref{eq_eigenvalue_multi}, respectively.
+\subsection{Case study of simulated data for $2$D and $3$D functions}\label{sec_caseIII}
+
+In this example we apply the diagnostic tool to fit and diagnose two different data sets, one data set simulated from a bivariate (D=$2$) function and another data set simulated from a $3$-dimensinal (D=$3$) function.
+Furthermore, we present results about the computational time required to fit the HSGP model in $2$D, $3$D and $4$D input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+
+The $2$D and $3$D synthetic datasets consist of $n=200$ and $n=1000$ ($i=1,\dots,n$) single draws from GP priors with input values $\bm{x}_i \in \{[-1,1],\dots,[-1,1]\} \subset {\rm I\!R}^D$, in a D-dimensional input space.
+A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, for the first input dimension, $\ell_2=0.3$, for the second input dimension, and $\ell_3=0.4$, for the third input dimension, is used for the GP priors.
+Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy sets of observations.
+
+In the HSGP models with $2$ and $3$ input dimensions, the underlying functions $f(\bm{x})$ to the data are approximated as in eq. (13)% (\ref{eq_approxf_multi})
+, with the $D$-dimensional squared exponential spectral densities $S$ as in eq.
+\eqref{eq_specdens_inf}, and the $D$-vectors of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations \eqref{eq_eigenfunction_multi} and \eqref{eq_eigenvalue_multi}, respectively.
 
 The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
-$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
-We use the same prior distributions to fit the exact GP model as the HSGP models.
-
-The HSGP model is fitted following the recommended iterative steps as in Section~\ref{sec_user_guide}, where in each iteration the diagnosis is applied on both dimensions, where for each one the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated by equations in Section~\ref{sec_num_equations}.
-The initial values $0.5$ and $1$ are set for the minimum lengthscales $\ell_1$ and $\ell_2$, for first and second dimensions respectively.
-
-In order to do model comparison, in addition to the exact GP and HSGP models, a $2$D splines-based model is also fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
-
-Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, exact GP, HSGP, and splines, fitted over the dataset.
+$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{InverseGamma}(2,0.5)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scales $\ell$, respectively.
+We use the same prior distributions to fit the exact GP model as the HSGP models on the synthetic data sets.
+
+The HSGP models are fitted following the recommended iterative steps as explained in Section~\ref{sec_user_guide}, where in each iteration the diagnosis is applied on every dimensions and, for each one of the dimensions, the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated by equations in Section~\ref{sec_num_equations}.
+The values $0.5$, $1$, and $1$ are set as initial values for the minimum lengthscales $\ell_1$, $\ell_2$ and $\ell_3$, for the first, second and third dimensions, respectively.
+In order to be as more efficient as possible in these multivariate cases, after the diagnostic generated the first \textit{true} for a certain dimension, its corresponding $m$ is updated by increasing the $m$ of the previous iteration by only 2 additional basis functions and, after the second \textit{true}, $m$ is no longer increased.
+The full diagnosis process is applied until two \textit{trues} are achieved for each dimension.
+
+Table~\ref{tab_caseIII_2D} and \ref{tab_caseIII_3D} contain the iterative steps to fit and diagnose the $2$D and $3$D data sets, respectively. 
+%These tables contain the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ for each dimension, and the RMSE compute with both the data and the GP as the reference for every iterative steps of the fitting process.
+The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{st}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{th}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately fit these notably wiggly data sets.
+Without using our recommended guidelines and diagnostics, it could have taken user several trials to get the optimal solution and to expend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
+
+\begin{figure*}[h]
+\centering
+\subfigure{\includegraphics[scale=0.315, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 151mm 14mm 2mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
+\hspace{1pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
+\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{ch5_fig17_legend_exII.pdf}}}
+\caption{Error between the $2$D data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models.
+Sample points are plotted as circles.
+The right side plot shows the root mean square error (RMSE) of the different methods, and plotted as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots (the same number of basis functions $m$ and knots per dimension are used, resulting in $m^D$ total number of multivariate basis functions).}
+  \label{ch5_fig16_errors_exII}
+\end{figure*}
+
+\begin{table}
+\centering
+\setlength{\tabcolsep}{3.6pt}
+\begin{tabular}{cccccccccc}
+\arrayrulecolor{gray}\hline \\[-3mm]
+ iter. & Dim. & $\ell$ & $c$ & $m$ & $\hat{\ell}$ & $\hat{\ell} + 0.01 \geq \ell$ & RMSE & RMSE* \\ 
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{2}{*}{ 1 } & $1$ & 0.50 & 1.6 & 6 & 0.19 & FALSE & \multirow{2}{*}{ 0.75 } & \multirow{2}{*}{ 0.72 }\\ 
+ & $2$ & 1.00 & 3.2 & 6 & 0.25 & FALSE &  &  \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{2}{*}{ 2 } & $1$ & 0.19 & 1.2 & 12 & 0.11 & FALSE  & \multirow{2}{*}{ 0.18 } & \multirow{2}{*}{ 0.12 }\\
+ & $2$ & 0.25 & 1.2 & 9 & 0.25 & TRUE &  &  \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{2}{*}{ 3 } & $1$ & 0.11 & 1.2 & 20 & 0.11 & TRUE  & \multirow{2}{*}{ 0.11 } & \multirow{2}{*}{ 0.03 }\\
+ & $2$ & 0.19 & 1.2 & 11 & 0.27 & TRUE &  &  \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{2}{*}{ 4 } & $1$ & 0.10 & 1.2 & 22 & 0.11 & TRUE  & \multirow{2}{*}{ 0.10 } & \multirow{2}{*}{ 0.02 }\\
+ & $2$ & 0.19 & 1.2 & 11 & 0.27 & TRUE  &  & \\[1mm] 
+\arrayrulecolor{lightgray}\hline
+ \multicolumn{8}{l}{ $\ell_{1_{GP}}=0.10$, \, $\ell_{2_{GP}}=0.29$ }
+\end{tabular}
+\caption{Iterative steps in HSGP model fitting on the $2$D data.
+The diagnosis procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the two dimensions $dim_1$ and $dim_2$.
+The full diagnosis process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
+RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+  \label{tab_caseIII_2D}
+\end{table}
+
+\begin{table}
+\centering
+\setlength{\tabcolsep}{3.6pt}
+\begin{tabular}{cccccccccc}
+\arrayrulecolor{gray}\hline \\[-3mm]
+ iter. & Dim. & $\ell$ & $c$ & $m$ & $\hat{\ell}$ & $\hat{\ell} + 0.01 \geq \ell$ & RMSE & RMSE* \\ 
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{3}{*}{ 1 } & $1$ & 0.50 & 1.6 & 6 & 0.14 & FALSE & \multirow{3}{*}{ 0.63 } & \multirow{3}{*}{ 0.58 }\\ 
+ & $2$ & 0.50 & 1.6 & 6 & 0.27 & FALSE &  &  \\
+ & $3$ & 1.00 & 3.2 & 6 & 0.23 & FALSE &  &  \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{3}{*}{ 2 } & $1$ & 0.13 & 1.2 & 16 & 0.11 & FALSE  & \multirow{3}{*}{ 0.13 } & \multirow{3}{*}{ 0.07 }\\
+ & $2$ & 0.27 & 1.2 & 8 & 0.29 & TRUE &  &  \\
+ & $3$ & 0.23 & 1.2 & 19 & 0.40 & TRUE &  &  \\
+\arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{3}{*}{ 3 } & $1$ & 0.11 & 1.2 & 19 & 0.12 & TRUE  & \multirow{3}{*}{ 0.11 } & \multirow{3}{*}{ 0.04 }\\
+ & $2$ & 0.21 & 1.2 & 10 & 0.29 & TRUE &  &  \\
+ & $3$ & 0.19 & 1.3 & 12 & 0.43 & TRUE &  &  \\
+ \arrayrulecolor{lightgray}\hline \\[-2mm]
+ \multirow{3}{*}{ 4 } & $1$ & 0.10 & 1.2 & 21 & 0.12 & TRUE  & \multirow{3}{*}{ 0.11 } & \multirow{3}{*}{ 0.03 }\\
+ & $2$ & 0.21 & 1.2 & 10 & 0.29 & TRUE &  &  \\
+ & $3$ & 0.19 & 1.3 & 12 & 0.43 & TRUE &  &  \\[1mm] 
+\arrayrulecolor{lightgray}\hline
+ \multicolumn{9}{l}{ $\ell_{1_{GP}}=0.11$, \,$\ell_{2_{GP}}=0.32$, \, $\ell_{3_{GP}}=0.43$ }
+\end{tabular}
+\caption{Iterative steps in HSGP model fitting on the $3$D data.
+The diagnosis procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three dimensions $dim_1$, $dim_2$ and $dim_3$.
+The full diagnosis process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
+RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+  \label{tab_caseIII_3D}
+\end{table}
+
+%\begin{figure*}
+%\centering
+%\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf}
+%\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}
+%\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_spfun_exII.pdf}
+%\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}
+%\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline-based (d) models. Sample points are plotted as circles}
+%  \label{ch5_fig15_Posteriors_exII}
+%\end{figure*}
+
+%\begin{figure}
+%\centering
+%\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
+%\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
+%\hspace{-1.8cm} \subfigure{\raisebox{0.5cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{ch5_fig17_legend_exII.pdf}}}
+%\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
+%  \label{ch5_fig17_RMSE_exII}
+%\end{figure}
+
+%Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, exact GP, HSGP, and splines, fitted over the dataset.
 Sample observations are also plotted as circles.
@@
 
-Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the data-generating function and the GP, HSGP and spline-based models, respectively.
+Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the $2$D data-generating function, from where the dataset was drawn, and the GP, HSGP and spline-based model solutions, respectively.
+The $2$D splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
+For the spline-based model, $40$ knots in each dimension were used.
+Figure \ref{ch5_fig16_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
+It can be seen an accurate approximation of the HSGP and exact GP to the $2$D data-generative function.
+However, the performance of the spline-based model is significantly worse.
+
+\subsubsection{Computation requirements in $2$D, $3$D and $4$D input spaces}\label{sec_computation_nD}
+
+Experiments have been conducted to empirically recognize time of computation of the HSGP model in multivariate cases and different number of sample data points.
+Dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} was used for the experiments.
+Implementation and sampling has been done only for the sample data.
+
+Figure~\ref{fig_time_nD_log} shows the computation time (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$ and $n=3000$, and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Both the time of computation and $m^*$ are plotted in the logarithm scale.
+
+If we look back at Figure~\ref{fig6_relationships} or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32} and \eqref{eq_m_l_mat52} we can find out that, for example for a squared exponential kernel, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted, after applying the iterative diagnostic steps, by only using approximately $10$ or $12$ basis functions as much, with the assumption that we have chosen the optimal value of the boundary factor $c$.
+And, for true lengthscales between $0.1$ and $0.3$, by only using between $10$ and $22$ basis functions as much, with the assumption that we have chosen the optimal value of the boundary factor $c$.
+
+So, for a very wiggly $2$D function (let's say with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed for an accurate approximation will be $m*=22\times 22= 484$, which results in an approximation significantly faster than the regular GP even in very small sized data sets (i.e., $n<300$) (see Figure~\ref{fig_time_nD_log}).
+Similarly, for a very wiggly $3$D function (let's say with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), we can expect an approximate minimum number of multivariate basis functions for an accurate approximation of $m*=22\times 22\times 22= 10648$, which, unless it is a small sized data set (i.e., $n<1000$), the method is significantly faster than the regular GP.
+Note the a $3$D function where each dimension has a lengthscale around $0.1$ is quite uncommon, so for most of the $3$D functions, unless it is a very small sized data set, the approximation will be significantly faster than the excat GP, if our iterative diagnostic rules are used as they allow to find out easily the solution with the minimum requirements.
+For $4$D data sets, the method can still be more efficient than the regular GP for moderate-large data sets (i.e., $n>1000$).
+
+Finally, for $D>5$ the method starts to be impractical even for smooth univariate functions and small $n$.
+However, in these cases the method may still be used in an additive modeling scheme with low-dimensional additive components.
+
+%Even though in a bivariate input space the computational demands increase significantly with the number of dimensions or knots, either the HSGP or spline-based models work significantly better than exact GP, even for highly wiggly functions that require a high number of basis functions or knots for an accurate approximation.
+
+%\begin{figure*}
+%\centering
+%\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_time_nD.png}
+%\caption{Computational time in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of multivariate basis functions $m^D$ and for different number of data points.}
+%  \label{fig_time_nD}
+%\end{figure*}
 
@@
 \centering
-\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf}
-\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}
-\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_spfun_exII.pdf}
-\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}
-\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline-based (d) models. Sample points are plotted as circles}
-  \label{ch5_fig15_Posteriors_exII}
+\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_time_nD_log.png}
+\caption{Computational time in seconds per iteration (iteration of the HMC sampling method) for different datasets with different dimensionalties and data points, and plotted as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$.}
+  \label{fig_time_nD_log}
 \end{figure*}
 
-\begin{figure*}
-\centering
-\subfigure{\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
-\caption{Error between the data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models. Sample points are plotted as circles.}
-  \label{ch5_fig16_errors_exII}
-\end{figure*}
-
-In order to assess performance of the models as a function of the number of basis functions, boundary factor and knots, different models with different number of basis functions, for the HSGP, and different number of knots, for the spline-based model, have been fitted.
+%In order to assess performance of the models as a function of the number of basis functions, boundary factor and knots, different models with different number of basis functions, for the HSGP, and different number of knots, for the spline-based model, have been fitted.
 In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
-Figure \ref{ch5_fig17_RMSE_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the spline-based model.
 From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen an accurate approximation of the HSGP to the exact GP.
 However, the performance of the spline-based model is significantly worse.
-
-Figure \ref{ch5_fig17_RMSE_exII}-right shows the computation times of the different models as a function of the boundary factor, number of basis functions and knots.
+%
+%Figure \ref{ch5_fig16_errors_exII}-right %\ref{ch5_fig17_RMSE_exII}-right 
+%shows the computation times of the different models as a function of the boundary factor, number of basis functions and knots.
 Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computation time.
@@
 The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
-
-\begin{figure}
-\centering
-\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
-\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
-\subfigure{\includegraphics[scale=0.35, trim = 28mm 70mm 112mm 0mm, clip]{ch5_fig17_legend_exII.pdf}}
-\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
-  \label{ch5_fig17_RMSE_exII}
-\end{figure}
 
@@
 Figure~\ref{fig19_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig19_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_7_relationships}.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig19_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_relationships}.
 Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig19_elpd_leukemia}-right).

### 2022-03-11 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex

@@
 We have applied the diagnostic tool and have been able to easily discover the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP.
-Without the guidance of the diagnostic tool, a user would have to go through several more trials until finding a sufficiently accurate solution, which would result in substantially increased human and computer time requirements.
+Without the guidance of the diagnostic tool, a user would have to go through several trials until finding a sufficiently accurate solution, which would result in substantially increased human and computer time requirements.
 
 We have also added a new case study to Section 5 which consists of wiggly 2D and 3D simulated data sets, where have applied the diagnostic procedure too.
-Even though the diagnostic was built for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing the different dimensions in each iteration individually.
+Even though the diagnostic was designed for unidimensional kernels, we demonstrate its applicability and usefulness in multidimensional cases by diagnosing the different dimensions in each iteration individually.
 %It allows us to find the optimal solution with the minimum requirements that can be essential to use the HSGP model in multidimensional cases.
@@
 
-The main comparison that we focus on is the Hilbert space basis function approximation (our HSGP method) against the exact GP model.
-In all comparisons, the inference conditional on the model is assumed to be accurate.
-In other words, we compare the exact to an approximate model, using asymptotically correct inference.
-For this purpose, we have used efficient dynamic HMC, the MCMC convergence diagnostics did not reveal problems, and thus the accuracy of the inference itself can be considered good.
+In the paper, we focus on the comparison between the Hilbert space basis function approximation (HSGP method) against an exact GP inference. 
+%In all comparisons, the inference conditional on the model is assumed to be accurate. In other words, we compare the exact to an approximate model, using asymptotically correct inference.
+For this purpose, we have used the efficient dynamic HMC algorithm with careful monitoring of the relevant MCMC convergence diagnostics to ensure that the Markov chains have mixed properly, and thus the accuracy of the inference itself can be considered good.
 
@@
 variational inference, they would blow up the already extensive scope of the paper.
-Some comparisons of MCMC and variational inference for basis function GPs can be found in
+However, we want to emphasize that several comparisons of MCMC and variational inference for basis function GPs can be found in
 
@@
 
-Further comparisons of the Hilbert space basis function approximations against other GP approximations are provided by Solin and Särkkä (2020), which is cited prominently in our paper.
+Furthermore, Solin and Särkkä (2020) compare the HSGP approximation to a range of difficult low-rank Gaussian process approximation methods based on inducing points.
+The results show that the HSGP approximation yields superior accuracy and therefore, our study is focused on the comparison against the exact model.
+
+ %of the Hilbert space basis function approximations against other GP approximations are provided by ,
 

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
quick look for undefined references

@@
 
-As commented in next Section~\ref{sec_diagnostic_criticism}, the estimated $\hat{\ell}$ being larger than diagnostic threshold doesn't guarantee that the approximate is sufficiently accurate, and thus we recommend to look at the predicitve accuracy measures, too.
+As commented above, the estimated $\hat{\ell}$ being larger than diagnostic threshold doesn't guarantee that the approximate is sufficiently accurate, and thus we recommend to look at the predicitve accuracy measures, too.
 
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships}, is essential to avoid a excessive computational time especially in multivariate input spaces.
+Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figure~\ref{fig6_relationships}, is essential to avoid a excessive computational time especially in multivariate input spaces.
 However, in practice, input dimensionalities larger than three start to be computationally demanding even for moderately wiggly functions and few basis functions per input dimension.

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
cite Stan as aksed by the Stan team

@@
 
-The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
 $\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
@@
 
-The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
 $\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1 \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell, \ell,\ell\}$, respectively.
@@
 
-The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual}.
+The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
 $\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{InverseGamma}(2,0.5)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scales $\ell$, respectively.
@@
 Experiments have been conducted to empirically recognize time of computation of the HSGP model in multivariate cases and different number of sample data points.
-Dynamic HMC algorithm implemented in Stan \citep{carpenter2017stan,betancourt2017conceptual} was used for the experiments.
+Dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} was used for the experiments.
 Implementation and sampling has been done only for the sample data.

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edits in 5.8

@@
 $\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
-We use the same prior distributions to fit the exact GP model as the HSGP models.
+We use the same prior distributions to fit the exact GP model.
 
@@
 $\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{InverseGamma}(2,0.5)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scales $\ell$, respectively.
-We use the same prior distributions to fit the exact GP model as the HSGP models on the synthetic data sets.
+We use the same prior distributions to fit the exact GP model.
 
@@
 The values $0.5$, $1$, and $1$ are set as initial values for the minimum lengthscales $\ell_1$, $\ell_2$ and $\ell_3$, for the first, second and third dimensions, respectively.
-In order to be as more efficient as possible in these multivariate cases, after the diagnostic generated the first \textit{true} for a certain dimension, its corresponding $m$ is updated by increasing the $m$ of the previous iteration by only 2 additional basis functions and, after the second \textit{true}, $m$ is no longer increased.
+In order to be as efficient as possible in these multivariate cases, after the diagnostic generated the first \textit{true} for a certain dimension, its corresponding $m$ is updated by increasing the $m$ of the previous iteration by only 2 additional basis functions and, after the second \textit{true}, $m$ is no longer increased.
 The full diagnosis process is applied until two \textit{trues} are achieved for each dimension.
@@
 The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{st}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{th}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately fit these notably wiggly data sets.
-Without using our recommended guidelines and diagnostics, it could have taken user several trials to get the optimal solution and to expend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
+Without using our recommended guidelines and diagnostics, it could have taken user several trials to get the optimal solution and to spend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
 
@@
 
-\subsubsection{Computation requirements in $2$D, $3$D and $4$D input spaces}\label{sec_computation_nD}
-
-Experiments have been conducted to empirically recognize time of computation of the HSGP model in multivariate cases and different number of sample data points.
+\subsubsection{Computation requirements in 2D, 3D and 4D input spaces}\label{sec_computation_nD}
+
+%Experiments have been conducted to empirically recognize time of computation of the HSGP model in multivariate cases and different number of sample data points.
 Dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} was used for the experiments.
@@
 
-Figure~\ref{fig_time_nD_log} shows the computation time (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$ and $n=3000$, and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Figure~\ref{fig_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$ and $n=3000$, and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
 Both the time of computation and $m^*$ are plotted in the logarithm scale.
 
-If we look back at Figure~\ref{fig6_relationships} or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32} and \eqref{eq_m_l_mat52} we can find out that, for example for a squared exponential kernel, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted, after applying the iterative diagnostic steps, by only using approximately $10$ or $12$ basis functions as much, with the assumption that we have chosen the optimal value of the boundary factor $c$.
-And, for true lengthscales between $0.1$ and $0.3$, by only using between $10$ and $22$ basis functions as much, with the assumption that we have chosen the optimal value of the boundary factor $c$.
-
-So, for a very wiggly $2$D function (let's say with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed for an accurate approximation will be $m*=22\times 22= 484$, which results in an approximation significantly faster than the regular GP even in very small sized data sets (i.e., $n<300$) (see Figure~\ref{fig_time_nD_log}).
-Similarly, for a very wiggly $3$D function (let's say with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), we can expect an approximate minimum number of multivariate basis functions for an accurate approximation of $m*=22\times 22\times 22= 10648$, which, unless it is a small sized data set (i.e., $n<1000$), the method is significantly faster than the regular GP.
-Note the a $3$D function where each dimension has a lengthscale around $0.1$ is quite uncommon, so for most of the $3$D functions, unless it is a very small sized data set, the approximation will be significantly faster than the excat GP, if our iterative diagnostic rules are used as they allow to find out easily the solution with the minimum requirements.
-For $4$D data sets, the method can still be more efficient than the regular GP for moderate-large data sets (i.e., $n>1000$).
+Looking back at Figure~\ref{fig6_relationships}, or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32} and \eqref{eq_m_l_mat52}, and assuming squared exponential covariance function, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted by only using $10--12$ basis functions.
+And, for lengthscales between $0.1$ and $0.3$, $10--22$ basis functions is sufficient.
+
+For a very wiggly $2$D function (let's say with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22= 484$, which results in significantly faster computation than the regular GP, even with small data sets (i.e., $n<300$) (see Figure~\ref{fig_time_nD_log}).
+For a very wiggly $3$D function (let's say with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n<1000$), the method is significantly faster than the regular GP.
+A $3$D function where each dimension has a lengthscale around $0.1$ is not that common in statistical analysis, and thus in many cases the approximation will be significantly faster than excat GP.
+For $4$D data sets, the method can still be more efficient than the exact GP for moderate-large data sets (i.e., $n>1000$).
 
 Finally, for $D>5$ the method starts to be impractical even for smooth univariate functions and small $n$.
-However, in these cases the method may still be used in an additive modeling scheme with low-dimensional additive components.
+However, in these cases the method may still be used for lower dimensional componenets in an additive modeling scheme.
 

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Edits to sec 4.8

@@
 
-\subsection{On computational cost in multi-dimensional cases}
-
-The computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-The HSGP method can be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
-
-The HSGP model is definitely useful in $1$D and $2$D, even for highly wiggly functions, unless in $2$D the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
-
-In the experiments of this paper for multivariate cases, mostly reported in Section~\ref{sec_computation_nD}, computation for HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderated-large sized ($n>1000$) $3$D datasets (even highly wiggly functions with large $m^* \approx 10000$); see Figure~\ref{fig_time_nD_log}).
+\subsection{The computational cost in the multi-dimensional setting}
+
+The HSGP model is computationally superior $1$D and $2$D, even for highly wiggly functions, except when the number of data points so small ($n \lesssim 300$) that regular formulation of the GP is already reasonably fast itself.
+However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
+Yet, the HSGP method can still be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
+
+In our experiments with for multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderated-large sized ($n>1000$) $3$D datasets (even highly wiggly functions with large $m^* \approx 10000$); see Figure~\ref{fig_time_nD_log}).
 
@@
 
-As it has been shown in case study III (Section~\ref{sec_caseIII}), the iterative diagnostic tool proposed in this study can be very useful for the multivariate cases as it allows to reduce $m^*$ to the minimum, reducing computational time drastically, and still getting an accurate approximation.
-Otherwise, user might be struggling in these multivariate cases.
-
-When HSGP or exact GP is faster depend also on the implementation details that can have big effect on the constants, and thus any more detailed recommendations can only be made in the context of the specific software implementation.
+As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum, reducing computational time drastically, and still getting an accurate approximation.% Otherwise, the user might be struggling in these multivariate cases.
+
+Whether HSGP or exact GP is faster will also depend on the specific implementation details, which can have big effect on the constants, and thus further detailed recommendations may depend on the specific software implementation.
 

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Edits to sec 5.3

@@
 
-Roughly similar or even worse behavior was found for thin plate regression splines used in the experiments, with serious challenges with computation time when doing inference for spline models for $D=3$ and with more than 10 basis functions, or even for $D=2$ and more than 40 basis functions (see Figures~B.3-right and C.3 in the online material).
+We observed roughly similar or even worse behavior for the thin plate regression splines used in the experiments, with serious challenges with computation time when doing inference for spline models for $D=3$ and with more than 10 basis functions, or even for $D=2$ and more than 40 basis functions (see Figures~B.3-right and C.3 in the online material).
 
@@
 
-\subsection{Case study of simulated data for $2$D and $3$D functions}\label{sec_caseIII}
-
-In this example we apply the diagnostic tool to fit and diagnose two different data sets, one data set simulated from a bivariate (D=$2$) function and another data set simulated from a $3$-dimensinal (D=$3$) function.
-Furthermore, we present results about the computational time required to fit the HSGP model in $2$D, $3$D and $4$D input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+\subsection{Case study: Simulated data for $2$D and $3$D functions}\label{sec_caseIII}
+
+In this case study, we apply the diagnostic tool to fit and diagnose two different data sets, one data set simulated from a bivariate (D=$2$) function and another data set simulated from a $3$-dimensinal (D=$3$) function.
+Furthermore, we present results or the computational time required to fit the HSGP model in $2$D, $3$D and $4$D input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
 
 The $2$D and $3$D synthetic datasets consist of $n=200$ and $n=1000$ ($i=1,\dots,n$) single draws from GP priors with input values $\bm{x}_i \in \{[-1,1],\dots,[-1,1]\} \subset {\rm I\!R}^D$, in a D-dimensional input space.
-A squared exponential covariance function, with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, for the first input dimension, $\ell_2=0.3$, for the second input dimension, and $\ell_3=0.4$, for the third input dimension, is used for the GP priors.
-Gaussian noise standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy sets of observations.
-
-In the HSGP models with $2$ and $3$ input dimensions, the underlying functions $f(\bm{x})$ to the data are approximated as in eq. (13)% (\ref{eq_approxf_multi})
+We use a GP prior with a squared exponential covariance function with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension.
+I.I.D.
+Gaussian noise with standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy sets of observations.
+
+In the HSGP models with $2$ and $3$ input dimensions, the underlying functions $f(\bm{x})$ are approximated as in eq. (13)% (\ref{eq_approxf_multi})
 , with the $D$-dimensional squared exponential spectral densities $S$ as in eq.
@@
 
-The HSGP models are fitted following the recommended iterative steps as explained in Section~\ref{sec_user_guide}, where in each iteration the diagnosis is applied on every dimensions and, for each one of the dimensions, the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated by equations in Section~\ref{sec_num_equations}.
-The values $0.5$, $1$, and $1$ are set as initial values for the minimum lengthscales $\ell_1$, $\ell_2$ and $\ell_3$, for the first, second and third dimensions, respectively.
-In order to be as efficient as possible in these multivariate cases, after the diagnostic generated the first \textit{true} for a certain dimension, its corresponding $m$ is updated by increasing the $m$ of the previous iteration by only 2 additional basis functions and, after the second \textit{true}, $m$ is no longer increased.
+The HSGP models are fitted following the recommended iterative steps as explained in Section~\ref{sec_user_guide}, where the diagnosis is applied on every dimensions.
+For each dimension, the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated using the equations in Section~\ref{sec_num_equations}.
+The values $0.5$, $1$, and $1$ are set as initial values for the minimum lengthscales $\ell_1$, $\ell_2$ and $\ell_3$, respectively.
+In order to be as efficient as possible, after the diagnostic generated the first \textit{true} for a certain dimension, its corresponding $m$ is updated by increasing the $m$ of the previous iteration by only 2 additional basis functions and, after the second \textit{true}, $m$ is no longer increased.
 The full diagnosis process is applied until two \textit{trues} are achieved for each dimension.
@@
 %These tables contain the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ for each dimension, and the RMSE compute with both the data and the GP as the reference for every iterative steps of the fitting process.
-The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{st}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{th}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately fit these notably wiggly data sets.
+The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{\text{st}}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{\text{th}}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately fit these notably wiggly data sets.
 Without using our recommended guidelines and diagnostics, it could have taken user several trials to get the optimal solution and to spend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
@@
 
-Figure \ref{ch5_fig16_errors_exII} shows the difference functions between the $2$D data-generating function, from where the dataset was drawn, and the GP, HSGP and spline-based model solutions, respectively.
-The $2$D splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+Figure \ref{ch5_fig16_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions, respectively.
+The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
@@
 Figure \ref{ch5_fig16_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
-It can be seen an accurate approximation of the HSGP and exact GP to the $2$D data-generative function.
-However, the performance of the spline-based model is significantly worse.
+It can be seen that the posterior mean for HSGP and exact GP accurately represents the true data-generating function.
+In contrast, the accuracy of the spline-based model is significantly worse.
 

### 2022-03-12 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
more edits to Section 4.8 and 5.3

@@
 
-The HSGP model is computationally superior $1$D and $2$D, even for highly wiggly functions, except when the number of data points so small ($n \lesssim 300$) that regular formulation of the GP is already reasonably fast itself.
+The HSGP model is computationally superior in $1$D and $2$D even for highly wiggly functions, except when the number of data points so small ($n \lesssim 300$) that exact GPs are already reasonably fast themselves.
 However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-Yet, the HSGP method can still be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
-
-In our experiments with for multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderated-large sized ($n>1000$) $3$D datasets (even highly wiggly functions with large $m^* \approx 10000$); see Figure~\ref{fig_time_nD_log}).
-
-For small sized ($n<1000$) datasets, HSGPs is likely to be slower than exact GPs already for $D=3$ and $D=4$, with a relatively large number of multivariate basis functions ($m^* > 3000$) in $3$D, and with a relatively low number of multivariate basis functions ($m^* \gtrsim 3000$) in $4$D (see Figure~\ref{fig_time_nD_log}).
+Yet, the HSGP method can still be computationally faster than the exact GP for larger datasets due the latter's cubic scaling in $n$.
+
+In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3$D datasets ($n>1000$), even for highly wiggly functions with large $m^* \approx 10000$ (see Figure~\ref{fig_time_nD_log}).
+
+For small sized datasets ($n<1000$), HSGPs are likely to be slower than exact GPs already for $D=3$ and $D=4$, with a relatively large number of multivariate basis functions ($m^* > 3000$) in $3$D, and with a relatively low number of multivariate basis functions ($m^* \gtrsim 3000$) in $4$D (see Figure~\ref{fig_time_nD_log}).
 
@@
 
-As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum, reducing computational time drastically, and still getting an accurate approximation.% Otherwise, the user might be struggling in these multivariate cases.
-
-Whether HSGP or exact GP is faster will also depend on the specific implementation details, which can have big effect on the constants, and thus further detailed recommendations may depend on the specific software implementation.
-
-We observed roughly similar or even worse behavior for the thin plate regression splines used in the experiments, with serious challenges with computation time when doing inference for spline models for $D=3$ and with more than 10 basis functions, or even for $D=2$ and more than 40 basis functions (see Figures~B.3-right and C.3 in the online material).
+As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum sufficient value, reducing computational time drastically, and still getting an accurate approximation.% Otherwise, the user might be struggling in these multivariate cases.
+
+Whether HSGP or exact GP is faster will also depend on the specific implementation details, which can have big effects on the scaling constants.
+Thus, more detailed recommendations would depend on the specific software implementation.
+
+In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions, or even for $D=2$ and more than 40 basis functions (see Figures~B.3-right and C.3 in the online supplement).
 
@@
 
-The $2$D and $3$D synthetic datasets consist of $n=200$ and $n=1000$ ($i=1,\dots,n$) single draws from GP priors with input values $\bm{x}_i \in \{[-1,1],\dots,[-1,1]\} \subset {\rm I\!R}^D$, in a D-dimensional input space.
-We use a GP prior with a squared exponential covariance function with hyperparameters marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension.
-I.I.D.
-Gaussian noise with standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy sets of observations.
+The $2$D and $3$D synthetic datasets consist of either $n=200$ and $n=1000$ ($i=1,\dots,n$) individual draws from GP priors with input values $\bm{x}_i \in \{[-1,1],\dots,[-1,1]\} \subset {\rm I\!R}^D$, in a D-dimensional input space.
+We use a GP prior with a squared exponential covariance function with  marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension.
+Independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy sets of observations.
 
@@
 
-The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
-$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{InverseGamma}(2,0.5)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scales $\ell$, respectively.
-We use the same prior distributions to fit the exact GP model.
-
-The HSGP models are fitted following the recommended iterative steps as explained in Section~\ref{sec_user_guide}, where the diagnosis is applied on every dimensions.
-For each dimension, the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated using the equations in Section~\ref{sec_num_equations}.
+The joint posterior parameter distributions are estimated via the dynamic HMC sampling algorithm implemented in Stan \citep{StanTeam:2021}.
+$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$, and $\mathrm{InverseGamma}(2,0.5)$ priors were used for the observation noise $\sigma$, marginal variance $\alpha$, and length-scales $\ell$, respectively.
+We used the same priors to fit the exact GP model serving as benchmark.
+
+The HSGP models are fitted following the recommended iterative procedure detailed in Section~\ref{sec_user_guide}, where the diagnostic is applied on every dimension, separately.
+For each dimension, the parameters $c$, $m$, minimum $\ell$, estimated $\hat{\ell}$, and the diagnostic $\hat{\ell} + 0.01 \geq \ell$ are updated using the equations in Section~\ref{sec_num_equations}.
 The values $0.5$, $1$, and $1$ are set as initial values for the minimum lengthscales $\ell_1$, $\ell_2$ and $\ell_3$, respectively.
 In order to be as efficient as possible, after the diagnostic generated the first \textit{true} for a certain dimension, its corresponding $m$ is updated by increasing the $m$ of the previous iteration by only 2 additional basis functions and, after the second \textit{true}, $m$ is no longer increased.
-The full diagnosis process is applied until two \textit{trues} are achieved for each dimension.
+The full diagnostic process is applied until two \textit{trues} are achieved for each dimension.
 
@@
 %These tables contain the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ for each dimension, and the RMSE compute with both the data and the GP as the reference for every iterative steps of the fitting process.
-The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{\text{st}}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{\text{th}}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately fit these notably wiggly data sets.
-Without using our recommended guidelines and diagnostics, it could have taken user several trials to get the optimal solution and to spend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
+The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{\text{st}}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{\text{th}}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately approximate these notably wiggly functions.
+Without using our recommended guidelines and diagnostics, it would likely have taken the user several more trials to get the optimal solution and to spend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
 
@@
 \arrayrulecolor{lightgray}\hline
- \multicolumn{8}{l}{ $\ell_{1_{GP}}=0.10$, \, $\ell_{2_{GP}}=0.29$ }
+ \multicolumn{8}{l}{True lengthscales: $\ell_{1_{GP}}=0.10$, \, $\ell_{2_{GP}}=0.29$ }
 \end{tabular}
 \caption{Iterative steps in HSGP model fitting on the $2$D data.
-The diagnosis procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the two dimensions $dim_1$ and $dim_2$.
-The full diagnosis process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
-RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+The diagnostic procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$, and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the two input dimensions.
+The full diagnostic process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
+RMSE and RMSE* refer to the root mean squared error of the HSGP computed against the true function and the exact GP, respectively.}
   \label{tab_caseIII_2D}
@@
 \arrayrulecolor{lightgray}\hline
- \multicolumn{9}{l}{ $\ell_{1_{GP}}=0.11$, \,$\ell_{2_{GP}}=0.32$, \, $\ell_{3_{GP}}=0.43$ }
+ \multicolumn{9}{l}{True lengthscales: $\ell_{1_{GP}}=0.11$, \,$\ell_{2_{GP}}=0.32$, \, $\ell_{3_{GP}}=0.43$ }
 \end{tabular}
 \caption{Iterative steps in HSGP model fitting on the $3$D data.
-The diagnosis procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three dimensions $dim_1$, $dim_2$ and $dim_3$.
-The full diagnosis process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
-RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+The diagnostic procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$m and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three input dimensions.
+The full diagnostic process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
+RMSE and RMSE* refer to the root mean squared error of the HSGP computed against the true function and the exact GP, respectively.}
   \label{tab_caseIII_3D}
@@
 
-\subsubsection{Computation requirements in 2D, 3D and 4D input spaces}\label{sec_computation_nD}
+\subsubsection{Computation requirements in 2D, 3D, and 4D input spaces}\label{sec_computation_nD}
 
@@
 
-Figure~\ref{fig_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$ and $n=3000$, and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
-Both the time of computation and $m^*$ are plotted in the logarithm scale.
-
-Looking back at Figure~\ref{fig6_relationships}, or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32} and \eqref{eq_m_l_mat52}, and assuming squared exponential covariance function, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted by only using $10--12$ basis functions.
-And, for lengthscales between $0.1$ and $0.3$, $10--22$ basis functions is sufficient.
-
-For a very wiggly $2$D function (let's say with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22= 484$, which results in significantly faster computation than the regular GP, even with small data sets (i.e., $n<300$) (see Figure~\ref{fig_time_nD_log}).
-For a very wiggly $3$D function (let's say with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n<1000$), the method is significantly faster than the regular GP.
+Figure~\ref{fig_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$Dm and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Both the time of computation and $m^*$ are plotted in the logarithmic scale.
+
+Looking back at Figure~\ref{fig6_relationships}, or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32}, and \eqref{eq_m_l_mat52}, and assuming squared exponential covariance function, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted using only $10--12$ basis functions.
+For lengthscales between $0.1$ and $0.3$, $10--22$ basis functions are sufficient.
+
+For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22= 484$, which results in significantly faster computation than the regular GP, even with small data sets (i.e., $n<300$) (see Figure~\ref{fig_time_nD_log}).
+For a very wiggly $3$D function (say, with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n<1000$), the method is still significantly faster than the regular GP.
 A $3$D function where each dimension has a lengthscale around $0.1$ is not that common in statistical analysis, and thus in many cases the approximation will be significantly faster than excat GP.
-For $4$D data sets, the method can still be more efficient than the exact GP for moderate-large data sets (i.e., $n>1000$).
-
-Finally, for $D>5$ the method starts to be impractical even for smooth univariate functions and small $n$.
-However, in these cases the method may still be used for lower dimensional componenets in an additive modeling scheme.
+For $4$D data sets, the method can still be more efficient than the exact GP for moderate-to-large data sets (i.e., $n>1000$).
+
+Finally, for $D>5$ the method starts to be impractical even for smooth univariate functions.
+However, in these cases, the method may still be used for lower dimensional componenets in an additive modeling scheme.
 
@@
 
-The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
+The Stan model codes for the exact GP, the HSGPm and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
 

### 2022-03-12 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
last edits

@@
 \item the estimated $\hat{\ell}$ is larger than the diagnostic threshold given $c$ and $m$, and
-\item the predictive accuracy measures, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD) don't improve.
+\item the predictive accuracy measures, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD) do not improve.
 \end{itemize}
 
-As commented above, the estimated $\hat{\ell}$ being larger than diagnostic threshold doesn't guarantee that the approximate is sufficiently accurate, and thus we recommend to look at the predicitve accuracy measures, too.
+As commented above, the estimated $\hat{\ell}$ being larger than diagnostic threshold does not guarantee that the approximate is sufficiently accurate, and thus we recommend to look at the predicitve accuracy measures, too.
 
@@
 %Using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential for repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional functions with very large data sets, and
-This even more useful in multi-dimensional cases ($2 \leq D \leq 4$), where knowing the smallest useful value of $m$ for each dimension has even bigger effect on the total computational cost.
+This is even more useful in multi-dimensional cases ($2 \leq D \leq 4$), where knowing the smallest useful value of $m$ for each dimension has even bigger effect on the total computational cost.
 
@@
 
-The HSGP model is computationally superior in $1$D and $2$D even for highly wiggly functions, except when the number of data points so small ($n \lesssim 300$) that exact GPs are already reasonably fast themselves.
+The HSGP model is computationally superior in $1$D and $2$D even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$) that exact GPs are already reasonably fast themselves.
 However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
@@
 
-In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3$D datasets ($n>1000$), even for highly wiggly functions with large $m^* \approx 10000$ (see Figure~\ref{fig_time_nD_log}).
-
-For small sized datasets ($n<1000$), HSGPs are likely to be slower than exact GPs already for $D=3$ and $D=4$, with a relatively large number of multivariate basis functions ($m^* > 3000$) in $3$D, and with a relatively low number of multivariate basis functions ($m^* \gtrsim 3000$) in $4$D (see Figure~\ref{fig_time_nD_log}).
+In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3$D datasets ($n \gtrsim 1000$), even for highly wiggly 3D functions (e.g., $\ell_1, \ell_2, \ell_3 \approx 0.1$).
+
+{\color{blue} For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for highly to moderated wiggly $3$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3 \gtrsim 0.3$) and for overall smooth $4$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3, \ell_4 \gtrsim 0.4$).}
 
@@
 
-As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum sufficient value, reducing computational time drastically, and still getting an accurate approximation.% Otherwise, the user might be struggling in these multivariate cases.
+As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum sufficient value, reducing computational time drastically, and still getting an accurate approximation.
+{\color{blue} For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\ell \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \ell \gtrsim 0.1$; and $m \gtrsim 20$ for $\ell \lesssim 0.1$), which, from results presented in Figure~\ref{fig_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3$D functions and smooth $4$D functions.}
 
@@
 
-In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions, or even for $D=2$ and more than 40 basis functions (see Figures~B.3-right and C.3 in the online supplement).
+In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions per dimension, or even for $D=2$ and more than 40 basis functions per dimension (see Figures~B.3-right and C.3 in the online supplement).
 
@@
 The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
-$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
+$\mathrm{Normal}(0,1)$, \linebreak $\mathrm{Normal}(0,3)$ and $\mathrm{Gamma}(1.2,0.2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variance $\alpha$, and length-scale $\ell$, respectively.
 We use the same prior distributions to fit the exact GP model.
@@
 
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
+The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/}} \linebreak {\small \url{ basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
 
@@
 The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
-$\mathrm{Normal}(0,1)$, $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1 \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell, \ell,\ell\}$, respectively.
+$\mathrm{Normal}(0,1)$, \linebreak $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1 \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell_1, \ell_2, \ell_3\}$, respectively.
 A $\mathrm{Normal}(0,0.1)$ prior distribution has been used for the standard deviation $\tau$ of the Student's $t$ distribution with 1 degree of freedom used to model $f_4$ (i.e., the special days effects).
@@
 
-The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}\,.
+The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/}} \linebreak {\small \url{master/Paper/Case-study_Birthday-data}}\,.
 
@@
 In this case study, we apply the diagnostic tool to fit and diagnose two different data sets, one data set simulated from a bivariate (D=$2$) function and another data set simulated from a $3$-dimensinal (D=$3$) function.
-Furthermore, we present results or the computational time required to fit the HSGP model in $2$D, $3$D and $4$D input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
-
-The $2$D and $3$D synthetic datasets consist of either $n=200$ and $n=1000$ ($i=1,\dots,n$) individual draws from GP priors with input values $\bm{x}_i \in \{[-1,1],\dots,[-1,1]\} \subset {\rm I\!R}^D$, in a D-dimensional input space.
-We use a GP prior with a squared exponential covariance function with  marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension.
-Independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the GP draws to form the final noisy sets of observations.
+Furthermore, in Section~\ref{sec_computation_nD} we present results of the computational time required to fit the HSGP model in $2$D, $3$D and $4$D input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+
+{\color{blue} $2$D and $3$D synthetic functions were drawn from $2$D and $3$D GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively.
+Squared exponential covariance functions with marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension, were used in the GP priors.
+$200$ and $1000$ data points were sampled from the $2$D and $3$D drawn functions, respectively, and independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the data points to form the final noisy sets of observations.}
 
@@
 \arrayrulecolor{lightgray}\hline
- \multicolumn{8}{l}{True lengthscales: $\ell_{1_{GP}}=0.10$, \, $\ell_{2_{GP}}=0.29$ }
+ \multicolumn{8}{l}{{\small Exact GP length-scales:} $\ell_{1_{GP}}=0.10$, \, $\ell_{2_{GP}}=0.29$ }
 \end{tabular}
@@
 \arrayrulecolor{lightgray}\hline
- \multicolumn{9}{l}{True lengthscales: $\ell_{1_{GP}}=0.11$, \,$\ell_{2_{GP}}=0.32$, \, $\ell_{3_{GP}}=0.43$ }
+ \multicolumn{9}{l}{{\small Exact GP length-scales:} $\ell_{1_{GP}}=0.11$, \,$\ell_{2_{GP}}=0.32$, \, $\ell_{3_{GP}}=0.43$ }
 \end{tabular}
@@
 
-Figure \ref{ch5_fig16_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions, respectively.
-The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+{\color{blue} Figure \ref{ch5_fig16_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2$D data set.
+For the $3$D data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3$D input space.} The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
@@
 
-Figure~\ref{fig_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$Dm and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Figure~\ref{fig_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
 Both the time of computation and $m^*$ are plotted in the logarithmic scale.
@@
 
-For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22= 484$, which results in significantly faster computation than the regular GP, even with small data sets (i.e., $n<300$) (see Figure~\ref{fig_time_nD_log}).
-For a very wiggly $3$D function (say, with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n<1000$), the method is still significantly faster than the regular GP.
+For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig_time_nD_log}).
+For a very wiggly $3$D function (say, with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n \lesssim 1000$), the method is still significantly faster than the regular GP.
 A $3$D function where each dimension has a lengthscale around $0.1$ is not that common in statistical analysis, and thus in many cases the approximation will be significantly faster than excat GP.
-For $4$D data sets, the method can still be more efficient than the exact GP for moderate-to-large data sets (i.e., $n>1000$).
+For $4$D data sets, the method can still be more efficient than the exact GP for moderate-to-large data sets (i.e., $n \lesssim 1000$).
 
@@
 
-The Stan model codes for the exact GP, the HSGPm and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
+The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/}} \linebreak {\small \url{basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
 

### 2022-03-12 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Tiny edit

@@
 
-{\color{blue} For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for highly to moderated wiggly $3$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3 \gtrsim 0.3$) and for overall smooth $4$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3, \ell_4 \gtrsim 0.4$).}
+{\color{blue} For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for moderate to highly wiggly $3$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3 \gtrsim 0.3$) and for overall smooth $4$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3, \ell_4 \gtrsim 0.4$).}
 

### 2022-03-14 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
latest/submitted version of revised munuscript and rebuttal

@@
 
-% There is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the length-scale $\ell$ of the approximated function.
-In this section we discuss about the existing theoretical evidence that the minimum number of basis functions $m$ needed for an accurate approximation is linearly proportional to the the boundary factor $c$ and inversely proportional to the length-scale $l$.
-% In the next Section \ref{subsec_empiric_relations} we conduct an empirical analysis through simulations to prove this expected linear proportionality between $m$, $c$ and $l$ and compute the specific coefficients of proportionality for every kernel.
-
-% {\color{blue}Aki: We can also look at the mean zero-level upcrossings, which for QE is
-% proportional to 1/l which gives as that part, and the effect of c should also
-% be quite linear when c is not very close to 1 and m is not very small as it
-% linearly makes the interval shorter so that the there are less zero-level
-% upcrossings available in the basis functions. You can visualize this using
-% {\small \url{https://avehtari.github.io/casestudies/Motorcycle/motorcycle_gpcourse.html}} and
-% look how the plot of 6 basis functions change when you set c=1, c=1.5, c=2.25
-% (ie $1.5^2$). Instead of looking at mean zero-level crossings, we can look at the
-% spectral densities and when they go to zero. This is illustrated in the same
-% notebook after that figure. E.g. with length-scale l=1,l=1/2, l=1/4, l=1/8 to
-% explain 99\% of the prior variance we need 4, 8, 16, and 32 basis functions
-% (linear in l) and to explain 80\% of the prior variance we need 2, 4, 8, 16
-% basis functions. These spectral densities don't depend on c. When c increaes
-% the basis functions are truncated and for example going from c=1.5 to c=2.25,
-% basis function 9 has the flexibility basis function 6 had, so 1.5 more basis
-% functions are needed. So we have proportionality c/l. }
-
-% {\color{red} Gabi: The spectral densities do depend on $c$, see equations (\ref{eq_approxf}) and (\ref{eq_eigenvalue}).}
-
 \subsection{Empirical discovering of the functional form of the relationships between $m$, $c$ and $\ell$} \label{subsec_empiric_relations}
@@
 
-% From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved.
-Following we show the numerical functions governing these relationships for every kernel.
-
 \paragraph*{Squared exponential}:
@@
 
-Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also provide the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that van be used to minimize the computational cost in repeated computations (e.g., in cross-validation and simulation based calibration)
-%If the initial lengthscale is not far from the optimal, and larger than minimal $c$ and $m$ are used in the initial HSGP model, it is likely that the initial results are already sufficiently accurate.
-%The iterative procedure, briefly commented above and detailed in next Section~\ref{sec_user_guide}, works as safety approach and also allows finding a smaller sufficient $m$ which can be useful if the user wants to optimize the computational speed.
-%Using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential for repeated fitting (e.g., in cross-validation or simulation based calibration), uni-dimensional functions with very large data sets, and
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also provide the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that van be used to minimize the computational cost in repeated computations (e.g., in cross-validation and simulation based calibration).
 This is even more useful in multi-dimensional cases ($2 \leq D \leq 4$), where knowing the smallest useful value of $m$ for each dimension has even bigger effect on the total computational cost.
@@
 The procedure is split into two phases, Phase~A and B, which have to be completed consecutively.
-%In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.
 
@@
 \end{enumerate}
-
-% \subsubsection{Criticism of the diagnostic} \label{sec_diagnostic_criticism}
-
-% The basis of our diagnostic tool are the following:
-% % 
-% \begin{enumerate}
-% \item observing a length-scale estimate $\hat{\ell}$ smaller than recommended minimum length-scale determined by $m$ and $c$ indicates a potential problem with the approximation,
-% \item observing a $\hat{\ell}$ larger than the recommended minimum length scale implies that the approximation is accurate.
-% \end{enumerate}
-
-% The later statement, however, cannot be assumed to be true in general, that is, cannot be generalized to all possible cases.
-({\color{red} TODO: A major revision from a reviewer was related to that, i.e.
-to be critical with the diagnostic and discuss about possible cases where it could fail.
-I don't know if we could discuss a bit about this.
-I remember Michael commented an example of a potential case where that diagnostic statement could not be true.})
-
-% For that reason, to ultimately diagnose whether the approximation is accurate, in addition to observe whether $\hat{\ell}$ exceeds the recommended minimum length scale, we recommend ensuring the measures of predictive accuracy reach an stabilization too.
 
@@
 
-%If much bigger $m$ than necessary is used (i.e., too much large $m$ relative to the wigglyness of the function to be learned), there can be computational challenges.
-In Figure~\ref{fig7_lscale_comparison}, we see that increasing $m$ much beyond what is needed, is leading to underestimating the lengthscale.
-Examining the convergence diagnostics revealed problems in exploring larger length-scales, which is likely due to the posterior having a difficult funnel shape \citep[see, e.g.,][]{Betancourt+Girolami:2019}.
-A largely underestimation of the length-scale leads to overestimating the underlying functions (see Figure~\ref{fig8_overestimated_posterios} as an illustration of the overestimating effect as a function of $m$ in a moderate smooth dataset).
-Increasing the boundary factor $c$ (more than that recommended in equations~\eqref{sec_num_equations}) reduces the funnellness of the posterior and the sampling problems in exploring larger length-scales mostly disappear and so underestimating the length-scale.
-%
-%Following our recommended diagnostic tool is a safety procedure to avoid using too much large $m$ that needed, as it allows us to select optimal values for $m$ and $c$ in each iteration.
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_overestimated_posterios.png}
-%\caption{Overestimation of the data points when using too many number of basis functions relative to the wigglyness of the data points. The dashed red line represents the true underlying function.}
-%  \label{fig8_overestimated_posterios}
-%\end{figure}
-
 Table~\ref{tab1_diagnostic} shows the iterative steps of applying the diagnostic procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison} left.
 It is clearly visible that by following our recommendations, an optimum solution with minimum computational requirements is achieved in these cases.
-Figure~\ref{fig10_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
+Figure~\ref{fig8_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
 Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
-%Furthermore, as can be seen in Figure~\ref{fig10_diagnostic_lscale_comparison}, following our diagnostic recommendations avoids the use of a too large number of basis functions relative to the true length scale that would cause overestimating the underlying function and bias the posterior towards smaller length scale values.
 
@@
 $\ell_{\text{GP}}$ & iter. & $\ell$ & $c$ & $m$ & $\hat{\ell}$ & $\hat{\ell} + 0.01 \geq \ell$ & RMSE & $R^2$ & ELPD & RMSE* & \\ 
-%\arrayrulecolor{lightgray}\hline \\[-1mm]
-%\multirow{5}{*}{ 0.05} 
-% & 1 & 0.50 & 1.60 & 6 & 0.127 & FALSE & 0.800 & 0.264 & -0.980 & 0.748 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_1.png}}\\
-% & 2 & 0.127 & 1.20 & 17 & 0.061 & FALSE & 0.427 & 0.779 & -0.396 & 0.330 & \\
-% & 3 & 0.061 & 1.20 & 35 & 0.059 & TRUE & 0.264 & 0.908 & 0.008 & 0.063 & \\
-% & 4 & 0.053 & 1.20 & 40 & 0.059 & TRUE & 0.254 & 0.915 & 0.088 & 0.029 & \\
-% \\[1mm]
 \arrayrulecolor{lightgray}\hline \\[-1mm]
 \multirow{5}{*}{ 0.08} 
- & 1 & 0.50 & 1.60 & 6 & 0.17 & FALSE & 0.45 & 0.61 & -0.40 & 0.36 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_2.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.17 & FALSE & 0.45 & 0.61 & -0.40 & 0.36 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_2.png}}\\
  & 2 & 0.17 & 1.20 & 13 & 0.07 & FALSE & 0.36 & 0.74 & -0.18 & 0.24 & \\
@@
 \multirow{3}{*}{ 0.13} 
- & 1 & 0.50 & 1.60 & 6 & 0.25 & FALSE & 0.46 & 0.63 & -0.49 & 0.33 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_3.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.25 & FALSE & 0.46 & 0.63 & -0.49 & 0.33 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_3.png}}\\
  & 2 & 0.25 & 1.20 & 9 & 0.15 & FALSE & 0.35 & 0.78 & -0.14 & 0.15 & \\
@@
 \multirow{3}{*}{ 0.25} 
- & 1 & 0.50 & 1.60 & 6 & 0.24 & FALSE & 0.37 & 0.40 & -0.15 & 0.21 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{fig9_diagnostic_4.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.24 & FALSE & 0.37 & 0.40 & -0.15 & 0.21 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_4.png}}\\
  & 2 & 0.24 & 1.20 & 9 & 0.26 & TRUE & 0.30 & 0.60 & -0.09 & 0.02 & \\
@@
 \multirow{3}{*}{ 0.54} 
- & 1 & 0.50 & 1.60 & 6 & 0.38 & FALSE & 0.27 & 0.86 & 0.03 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_5.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.38 & FALSE & 0.27 & 0.86 & 0.03 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_5.png}}\\
  & 2 & 0.38 & 1.20 & 6 & 0.47 & TRUE & 0.28 & 0.86 & 0.03 & 0.05 & \\
@@
 \multirow{3}{*}{ 0.80} 
- & 1 & 1.00 & 3.20 & 6 & 0.52 & FALSE & 0.29 & 0.92 & 0.13 & 0.03 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_6.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 0.52 & FALSE & 0.29 & 0.92 & 0.13 & 0.03 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_6.png}}\\
  & 2 & 0.52 & 1.65 & 6 & 0.98 & TRUE & 0.29 & 0.92 & 0.11 & 0.03 & \\
@@
 \multirow{2}{*}{ 1.40} 
- & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_7.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_7.png}}\\
  & 2 & 0.52 & 3.27 & 11 & 1.23 & TRUE & 0.28 & 0.57 & 0.14 & 0.00 & \\
@@
 \multirow{2}{*}{ 2.94} 
- & 1 & 0.50 & 1.60 & 6 & 1.19 & TRUE & 0.34 & 0.45 & -0.10 & 0.06 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{fig9_diagnostic_8.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 1.19 & TRUE & 0.34 & 0.45 & -0.10 & 0.06 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_8.png}}\\
  & 2 & 0.61 & 3.81 & 11 & 2.58 & TRUE & 0.34 & 0.45 & -0.07 & 0.01 & \\
@@
 \end{tabular}
-\caption{diagnostic on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+\caption{Diagnostic on various data sets with varying smoothness.
+$\ell_{GP}$ refers to the exact GP length-scale estimate.
+$\ell$ refers to the minimum valid length-scale that can be accurately estimated determined by $m$ and $c$ as per Section \ref{sec_num_equations}, except for the first iteration which is the initial guess of the length-scale.
+$\hat{\ell}$ refers to the HSGP length-scale estimate.
+RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
   \label{tab1_diagnostic}
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_diagnostic_lscale_comparison.png}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_diagnostic_lscale_comparison.png}}
 \caption{Comparison of the true and estimated length-scales in every iteration for the different data sets varying the length-scale.}
-  \label{fig10_diagnostic_lscale_comparison}
+  \label{fig8_diagnostic_lscale_comparison}
 \end{figure}
@@
 
-The HSGP model is computationally superior in $1$D and $2$D even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$) that exact GPs are already reasonably fast themselves.
+The HSGP model is computationally superior in $1$D and $2$D even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$, i.e., $n$ smaller than some value around 300) that exact GPs are already reasonably fast themselves.
 However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
@@
 
-In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3$D datasets ($n \gtrsim 1000$), even for highly wiggly 3D functions (e.g., $\ell_1, \ell_2, \ell_3 \approx 0.1$).
-
-{\color{blue} For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for moderate to highly wiggly $3$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3 \gtrsim 0.3$) and for overall smooth $4$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3, \ell_4 \gtrsim 0.4$).}
-
-%In all of the investigated cases, choosing the optimal booundary factor $c$ reduces the required number of basis functions $m$ noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore  reduces computational time drastically in particular in multidimensional cases.
+In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3$D datasets ($n \gtrsim 1000$, i.e., $n$ greater than some value around 1000), even for highly wiggly 3D functions (e.g., $\ell_1, \ell_2, \ell_3 \approx 0.1$).
+
+For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for highly to moderated wiggly $3$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3 \gtrsim 0.3$) and for overall smooth $4$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3, \ell_4 \gtrsim 0.4$).
 
 As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum sufficient value, reducing computational time drastically, and still getting an accurate approximation.
-{\color{blue} For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\ell \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \ell \gtrsim 0.1$; and $m \gtrsim 20$ for $\ell \lesssim 0.1$), which, from results presented in Figure~\ref{fig_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3$D functions and smooth $4$D functions.}
+For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\ell \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \ell \gtrsim 0.1$; and $m \gtrsim 20$ for $\ell \lesssim 0.1$), which, from results presented in Figure~\ref{fig15_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3$D functions and smooth $4$D functions.
 
@@
 
-In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions per dimension, or even for $D=2$ and more than 40 basis functions per dimension (see Figures~B.3-right and C.3 in the online supplement).
+In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions per dimension, or even for $D=2$ and more than 40 basis functions per dimension (see Figures~B.3 right and C.3 in the online supplement).
 
@@
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
-%We split the dataset into three parts: $155$ data points are used for fitting the model (training set), $45$ data ponts are used for the interpolation test set, and the remaining $50$ data points are used for the extrapolation test set.
-%
-%The exact GP model for fitting and predicting this simulated dataset $\bm{y}$ can be written as follows,
-%%
-%\begin{align*}\label{eq_latentgp_simudata1}
-%\begin{split}
-%\bm{y} &= \bm{f} + \bm{\epsilon} \\
-%\bm{\epsilon} &\sim \Normal(\bm{0}, \sigma^2 \bm{I}) \\
-%f(x) &\sim \GP(0, k(x, x', \theta)),
-%\end{split}
-%\end{align*}
-%%
-%where $\bm{f}=\{f(x_i)\}_{i=1}^n$ represents the underlying function at the input values $x_i$, and $\bm{\epsilon}$ is the Gaussian noise term with variance $\sigma^2$, with $\bm{I}$ representing the identity matrix.
-The function $f:{\rm I\!R} \to {\rm I\!R}$ is a GP prior with the Mat{\'e}rn ($\nu$=3/2) covariance function $k$, which depends on the inputs $\bm{x}$ and hyperparameters $\theta=\{\alpha,\ell\}$.
-The hyperparameters $\alpha$ and $\ell$ represent the marginal variance and length-scale, respectively, of the GP process.
-Saying that the function $f(\cdot)$ follows a GP model is equivalent to say that $\bm{f}$ is multivariate Gaussian distributed with covariance matrix $\bm{K}$, where $K_{ij}=k(x_i,x_j,\theta)$, with $i,j=1,2,\dots,n$.
-% 
-%A more computationally efficient formulation of a GP model with Gaussian likelihood, and for probabilistic inference using MCMC sampling methods, would be its marginalized form,
-%%
-%\begin{equation*}\label{eq_marginalizedgp_simudata1}
-%\bm{y} \sim \Normal(\bm{0},\bm{K}+ \sigma^2 \bm{I} ),
-%\end{equation*}
-%%
-%where the function values $\bm{f}$ have been integrated out, yielding a lower-dimensional parameter space over which to do inference, reducing the time of computation and improving the sampling and the effective number of samples.
 
 In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn ($\nu$=3/2) spectral density $S$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
-%In order to do model comparison, in addition to the exact GP model and HSGP model, a spline-based model is also fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
 
@@
 \end{tabular}
-\caption{diagnostic on various data sets with varying smoothness. RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
+\caption{Diagnostic on various data sets with varying smoothness.
+$\ell$ refers to the minimum valid length-scale that can be accurately estimated determined by $m$ and $c$, except for the first iteration which is the initial guess of the length-scale.
+RMSE* refers to the root mean squared error computed with the exact GP model as the reference.}
   \label{tab_caseI}
@@
 
-Figure~\ref{fig11_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
+Figure~\ref{fig9_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
 In order to do model comparison a spline-based model with 60 knots is also fitted and plotted jointly with the GP and HSGP models.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig11_Posteriors_exI.png}
+\includegraphics[width=0.5\textwidth]{fig9_Posteriors_exI.png}
 \caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
-  \label{fig11_Posteriors_exI}
+  \label{fig9_Posteriors_exI}
 \end{figure}
 
-%\textsc{In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted.
-Figure \ref{fig12_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots.
-The SRMSE is computed against the data-generating model.
-From Figures \ref{fig11_Posteriors_exI} and \ref{fig12_MSE_exI_inter}, it is seen that the HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation.
-However, the spline model does not extrapolate data properly.
-Both models show roughly similar interpolating performance.}
-
-%In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnostic of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.
-
-%The diagnostic consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships}.
-Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$.
-%Next, we demonstrate how to apply the diagnostic for approximation quality as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency}.
-%Figure \ref{fig13_diagnostic} depicts the diagnostic applied to case study.
-In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnostic}.
-In the second iteration, $m$ is increased to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig6_7_relationships} is $\ell^*=0.19$.
-As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{fig12_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP).
-In the third iteration, $m$ is increased to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships suggests a minimum length-scale of $\ell^*=0.15$.
-As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{fig12_MSE_exI_inter}-left).
-In the fourth iteration, $m$ is increased up $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum length-scale generated by the model of relationships is $\ell^*=0.11$, and the approximation is finally diagnosed to be accurate as $\hat{\ell} > \ell^*$.
-
-%\begin{figure*}[!t]
-%\centering
-%\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_inter.pdf}}
-%\subfigure{\includegraphics[scale=0.38, trim = 0mm 5mm 3mm 0mm, clip]{fig12_MSE_exI_extra.pdf}}
-%\caption{Standardized root mean square error (SRMSE) of the different methods against the data-generating function.
-SRMSE for interpolation (left) and SRMSE for extrapolation (right).
-The standard deviation of the mean of the SRMSE is plotted as dashed lines.}
-%  \label{fig12_MSE_exI_inter}
-%\end{figure*}
-%
-%
-%\begin{figure*}
-%\centering
-%\begin{tabular}{llll}
-%{\Large 1}\;\;\; \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} &  {\Large 2}\;\;\; \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.12$} &  {\Large 3}\;\;\; \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} &  {\Large 4}\;\;\; \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\;\; \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\;\; {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
-%	& \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} < \ell^*$ & \hspace{1.6cm} $\hat{\ell} > \ell^*$\\[1mm]
-%\hspace{-3mm} \includegraphics[scale=0.26, trim = 0mm 0mm 4mm 10mm, clip]{fig13_diag_iter1_exI.pdf} & \hspace{-3mm}
-%\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter2_exI.pdf} & \hspace{-3mm}
-%\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter3_exI.pdf} & \hspace{-4mm}
-%\includegraphics[scale=0.26, trim = 14mm 0mm 4mm 10mm, clip]{fig13_diag_iter4_exI.pdf} \\[1mm]
-%\end{tabular}
-%\caption{Iterative diagnostic of the approximation as a function of the number of basis functions $m$ and boundary factor $c$.
-The iterative steps go from the left side to the right side of the figure.}
-%	\label{fig13_diagnostic}
-%\end{figure*}
-
-Figure~\ref{fig14_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+Figure~\ref{fig10_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
 Both the HSGP and splines models show roughly similar performance at sample data.
 The HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation.
-However, the spline model does not extrapolate data properly (Figure~\ref{fig11_Posteriors_exI}).
+However, the spline model does not extrapolate data properly (Figure~\ref{fig9_Posteriors_exI}).
 The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig14_rmse_time_exI.png}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_rmse_time_exI.png}}
 \caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
@@
 The standard deviation of the computational time is plotted as dashed lines.}
-  \label{fig14_time_exI}
+  \label{fig10_time_exI}
 \end{figure}
@@
 The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/}} \linebreak {\small \url{ basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
-
-%In the online supplemental material, additional case studies are presented.
-From those examples, it can be seen how the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of basis functions in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
-Even though, for a bivariate input space, the computation time increases significantly with $D$, the HSGP model works significantly faster than the exact GP for most of the non-linear $2D$ functions (even highly wiggly functions; see Figures B.3-right and C.3 in the online material).
-For moderate sized datasets, HSGPs tend to be slower than exact GPs for $D>3$ with a relatively low number of basis functions ($m \gtrsim 5$), as well as even for $D=3$ with a moderate high number of basis functions ($m \gtrsim 20$; see Figure~C.3 in the online material).
-However, the HSGP method will be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.
-In all of the investigated cases, choosing the optimal boundary factor in the HSGP approximation reduces the number of required basis functions noticeably (see Figures~A.3, B.3-left and C.2 in the online material) and therefore also reduces computational time drastically in particular in multivariate input spaces.
-%
-%Roughly similar or even worse behavior was found for splines where serious difficulties with computation time were encountered in building spline models for $D=3$ and with more than 10 knots, or even for $D=2$ and more than 40 knots (see Figures~B.3-right and C.3 in the online material).
 
@@
 
-%The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
-%%
-%\begin{equation*}
-%f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
-%\end{equation*}
-%%
-%A horse-shoe prior allows for sparse distributed effects.
-Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
-Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
-
 The component $f_4(t)$ represents the special days effects, modeled as a Student's $t$ prior model with 1 degree of freedom and variance $\tau^2$:
@@
 %
-%A horse-shoe prior allows for sparse distributed effects.
-Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage.
-Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.
-
 The component $f_1$ will be approximated using the HSGP model and the function values $f_1$ are approximated as in eq.~(\ref{eq_approxf}), with the squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
@@
 The joint posterior parameter distributions are estimated by sampling using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021}.
-$\mathrm{Normal}(0,1)$, \linebreak $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1 \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell_1, \ell_2, \ell_3\}$, respectively.
+$\mathrm{Normal}(0,1)$, \linebreak $\mathrm{Normal}(0,10)$ and $\mathrm{Normal}(0,2)$ prior distributions has been used for the observation noise $\sigma$, covariance function marginal variances $\bm{\alpha}=\{\alpha_1, \alpha_2, \alpha_3\}$, and length-scales $\bm{\ell}=\{\ell_1, \ell_2, \ell_3\}$, respectively.
 A $\mathrm{Normal}(0,0.1)$ prior distribution has been used for the standard deviation $\tau$ of the Student's $t$ distribution with 1 degree of freedom used to model $f_4$ (i.e., the special days effects).
@@
 
-%We use $m=30$ basis functions and a boundary factor $c=1.5$.
-The length-scale estimate $\hat{\ell}_1$, for this component, normalized by half of the range of the input $x_1$, is bigger than the minimum length-scale reported by Figure \ref{fig6_7_relationships} as a function of $m$ and $c$.
-This means that the chosen number of basis functions and the boundary factor are suitable values for modeling the input effects sufficiently accurate.
-%
-%We use $J=10$ cosine terms.
-The length-scale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum length-scale reported by Figure \ref{figB1_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
-
-Figure~\ref{fig15_posteriors_birthday} shows the posterior means of the long-term trend $f_1$ and yearly pattern $f_2$ for the whole period, jointly with the observed data.
-Figure \ref{fig16_posteriors_oneyear_birthday} shows the model for one year (1972) only.
+Figure~\ref{fig11_posteriors_birthday} shows the posterior means of the long-term trend $f_1$ and yearly pattern $f_2$ for the whole period, jointly with the observed data.
+Figure \ref{fig12_posteriors_oneyear_birthday} shows the model for one year (1972) only.
 In this figure, the special days effects $f_4$ in the year can be clearly represented.
-The posterior means of the the function $\mu$ and the components $f_1$ (long-term trend) and $f_2$ (year pattern) are also plotted in this Figure \ref{fig16_posteriors_oneyear_birthday}.
-Figure \ref{fig17_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3$ can be clearly represented.
-The mean of the the function $\mu$ and components $f_1$ (long-term trend), $f_2$ (year pattern) and $f_4$ (special-days effects) are also plotted in this Figure \ref{fig17_posteriors_onemonth_birthday}.
+The posterior means of the the function $\mu$ and the components $f_1$ (long-term trend) and $f_2$ (year pattern) are also plotted in this Figure \ref{fig12_posteriors_oneyear_birthday}.
+Figure \ref{fig13_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3$ can be clearly represented.
+The mean of the the function $\mu$ and components $f_1$ (long-term trend), $f_2$ (year pattern) and $f_4$ (special-days effects) are also plotted in this Figure \ref{fig13_posteriors_onemonth_birthday}.
 
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig15_posteriors_birthday.png}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig11_posteriors_birthday.png}}
 \caption{Posterior means of the long-term trend $f_1$ and year effects pattern ($f_2$) for the whole series. }
-  \label{fig15_posteriors_birthday}
+  \label{fig11_posteriors_birthday}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig16_posteriors_oneyear_birthday.png}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig12_posteriors_oneyear_birthday.png}}
 \caption{Posterior means of the function $\mu$ for the year 1972 of the series.
@@
 }
-  \label{fig16_posteriors_oneyear_birthday}
+  \label{fig12_posteriors_oneyear_birthday}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig17_posteriors_onemonth_birthday.png}}
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig13_posteriors_onemonth_birthday.png}}
 \caption{Posterior means of the function $\mu$ for the month of January of 1972.
@@
 }
-  \label{fig17_posteriors_onemonth_birthday}
+  \label{fig13_posteriors_onemonth_birthday}
 \end{figure}
@@
 
-{\color{blue} $2$D and $3$D synthetic functions were drawn from $2$D and $3$D GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively.
+$2$D and $3$D synthetic functions were drawn from $2$D and $3$D GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively.
 Squared exponential covariance functions with marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension, were used in the GP priors.
-$200$ and $1000$ data points were sampled from the $2$D and $3$D drawn functions, respectively, and independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the data points to form the final noisy sets of observations.}
+$200$ and $1000$ data points were sampled from the $2$D and $3$D drawn functions, respectively, and independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the data points to form the final noisy sets of observations.
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.315, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig16_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 151mm 14mm 2mm 20mm, clip]{ch5_fig16_error_colorbar_exII.pdf}}
-\hspace{1pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
-\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{ch5_fig17_legend_exII.pdf}}}
+\subfigure{\includegraphics[scale=0.315, trim = 1mm 18mm 39.2mm 20mm, clip]{fig14_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig14_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig14_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 151mm 14mm 2mm 20mm, clip]{fig14_error_colorbar_exII.pdf}}
+\hspace{1pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{fig14_RMSE_exII.pdf}}
+\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{fig14_legend_exII.pdf}}}
 \caption{Error between the $2$D data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models.
@@
 The right side plot shows the root mean square error (RMSE) of the different methods, and plotted as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots (the same number of basis functions $m$ and knots per dimension are used, resulting in $m^D$ total number of multivariate basis functions).}
-  \label{ch5_fig16_errors_exII}
+  \label{fig14_errors_exII}
 \end{figure*}
@@
 \caption{Iterative steps in HSGP model fitting on the $3$D data.
-The diagnostic procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$m and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three input dimensions.
+The diagnostic procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three input dimensions.
 The full diagnostic process is applied until two consecutive \textit{trues} in the diagnostic are achieved for each dimension.
@@
 
-%\begin{figure*}
-%\centering
-%\includegraphics[scale=0.36, trim = 1mm 18mm 39.2mm 20mm, clip]{ch5_fig15_gpfun_exII.pdf}
-%\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_bffun_exII.pdf}
-%\includegraphics[scale=0.36, trim = 19mm 18mm 39.2mm 20mm, clip]{ch5_fig15_spfun_exII.pdf}
-%\includegraphics[scale=0.30, trim = 150mm 5mm 0mm 20mm, clip]{ch5_fig15_colorbar_vertical_exII.pdf}
-%\caption{Data-generating function (a) and posterior predictive mean functions of the GP (b), HSGP (c) and spline-based (d) models. Sample points are plotted as circles}
-%  \label{ch5_fig15_Posteriors_exII}
-%\end{figure*}
-
-%\begin{figure}
-%\centering
-%\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 5mm 0mm, clip]{ch5_fig17_RMSE_exII.pdf}}
-%\subfigure{\includegraphics[scale=0.2, trim = 0mm 0mm 10mm 0mm, clip]{ch5_fig17_time_exII.pdf}}
-%\hspace{-1.8cm} \subfigure{\raisebox{0.5cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{ch5_fig17_legend_exII.pdf}}}
-%\caption{Root mean square error (RMSE) (left) and computational time (right) in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
-%  \label{ch5_fig17_RMSE_exII}
-%\end{figure}
-
-%Figure \ref{ch5_fig15_Posteriors_exII} shows the data-generating GP function, from where the dataset was drawn, and the mean posterior predictive functions of the three models, exact GP, HSGP, and splines, fitted over the dataset.
-Sample observations are also plotted as circles.
-For the HSGP, $m_1=40$ and $m_2=40$ basis functions for each dimension, respectively, were used, which lead to a total of 1600 multivariate basis functions.
-A boundary factor for each dimension, $c_1=1.5$ and $c_2=1.5$, were used.
-For the spline-based model, 40 knots in each dimension were used.
-
-{\color{blue} Figure \ref{ch5_fig16_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2$D data set.
-For the $3$D data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3$D input space.} The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
+Figure \ref{fig14_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2$D data set.
+For the $3$D data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3$D input space.
+The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
 A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
 For the spline-based model, $40$ knots in each dimension were used.
-Figure \ref{ch5_fig16_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
+Figure \ref{fig14_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
 It can be seen that the posterior mean for HSGP and exact GP accurately represents the true data-generating function.
@@
 
-%Experiments have been conducted to empirically recognize time of computation of the HSGP model in multivariate cases and different number of sample data points.
-Dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} was used for the experiments.
-Implementation and sampling has been done only for the sample data.
-
-Figure~\ref{fig_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Figure~\ref{fig15_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
 Both the time of computation and $m^*$ are plotted in the logarithmic scale.
@@
 
-For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig_time_nD_log}).
+For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig15_time_nD_log}).
 For a very wiggly $3$D function (say, with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n \lesssim 1000$), the method is still significantly faster than the regular GP.
@@
 
-%Even though in a bivariate input space the computational demands increase significantly with the number of dimensions or knots, either the HSGP or spline-based models work significantly better than exact GP, even for highly wiggly functions that require a high number of basis functions or knots for an accurate approximation.
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_time_nD.png}
-%\caption{Computational time in seconds per iteration (iteration of the HMC sampling method) of the different methods, and plotted as a function of the boundary factor $c$, number of multivariate basis functions $m^D$ and for different number of data points.}
-%  \label{fig_time_nD}
-%\end{figure*}
-
 \begin{figure*}
 \centering
-\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_time_nD_log.png}
+\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig15_time_nD_log.png}
 \caption{Computational time in seconds per iteration (iteration of the HMC sampling method) for different datasets with different dimensionalties and data points, and plotted as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$.}
-  \label{fig_time_nD_log}
+  \label{fig15_time_nD_log}
 \end{figure*}
-
-%In order to assess performance of the models as a function of the number of basis functions, boundary factor and knots, different models with different number of basis functions, for the HSGP, and different number of knots, for the spline-based model, have been fitted.
-In each of these models, the same boundary factor, number of basis functions and knots per dimension were used.
-From Figures \ref{ch5_fig16_errors_exII} and \ref{ch5_fig17_RMSE_exII}-left, it can be seen an accurate approximation of the HSGP to the exact GP.
-However, the performance of the spline-based model is significantly worse.
-%
-%Figure \ref{ch5_fig16_errors_exII}-right %\ref{ch5_fig17_RMSE_exII}-right 
-%shows the computation times of the different models as a function of the boundary factor, number of basis functions and knots.
-Figure \ref{ch5_fig17_RMSE_exII} reveals that choosing the optimal boundary factor allows for less number of basis functions and less computation time.
-Even though in a bivariate input space the computational demands increase significantly with the number of dimensions or knots, either the HSGP or spline-based models work significantly better than exact GP, even for highly wiggly functions that require a high number of basis functions or knots for an accurate approximation.
-It is interesting to be noticed that just the computation of the input data for the spline model in a 2D input space and considering more than 50 knots per dimension is computationally very expensive.
 
@@
 \centering
-\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{fig18_posteriors_leukemia.pdf}
+\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{fig16_posteriors_leukemia.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
 The thick line in each graph is the posterior mean estimated using a HSGP model, and the thin lines represent pointwise 95\% credible intervals.}
-  \label{fig18_posteriors_leukemia}
+  \label{fig16_posteriors_leukemia}
 \end{figure*}
@@
 
-Figure \ref{fig18_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
+Figure \ref{fig16_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
 These posterior estimates correspond to the HSGP model with $m=10$ basis functions and $c=3$ boundary factor.
 There are clear  non-linear patterns and the right bottom subplot also shows that the conditional function associated with WBC has an interaction with TDI.
-Figure~\ref{fig19_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig19_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_relationships}.
-Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig19_elpd_leukemia}-right).
+Figure~\ref{fig17_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig17_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_relationships}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig17_elpd_leukemia}-right).
 With a moderate number of univariate basis functions ($m=15$), the HSGP model becomes slower than the exact GP model, in this specific application with $3$ input variables, as the total number of multivariate basis functions becomes $15^3 = 3375$ and is therefore quite high.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig19_elpd_leukemia.pdf}} 
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig19_time_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig17_elpd_leukemia.pdf}} 
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig17_time_leukemia.pdf}}
 \caption{Expected log predictive density (ELPD; left) and time of computation in seconds per iteration (iteration of the HMC sampling method; right) as a function of the number of basis functions $m$ and boundary factor $c$.}
-  \label{fig19_elpd_leukemia}
+  \label{fig17_elpd_leukemia}
 \end{figure*}
@@
 On that basis, as our second main contribution, we developed practical diagnostics to assess the approximation's performance as well as an iterative procedure to obtain an accurate approximation with minimal computational costs.
-
-%We illustrated our findings in several case studies of real and simulated data. 
-%%Recommendations for the values of these key factors based on the recognized relations among them have been provided.
-These illustrations will not only help users to improve performance and save computation time, but also serve as a powerful diagnostic tool whether the chosen values for the number of basis functions and the boundary condition are adequate to fit to the data at hand with sufficient accuracy.
 

### 2022-05-14 Paper/Case-study_2D&3D-Simulated-data/Rnb_case2D_diagnosis.Rmd
add Case-study_2D&3D-Simulated-data

---
title: '2D simulated case study: Diagnosis of the HSGP approximation'
author: 'Gabriel Riutort-Mayol'
date: "`r format(Sys.Date())`."
output:
  bookdown::html_document2:
    theme: readable
    toc: yes
    toc_depth: 4
    number_sections: true
    toc_float: 
      collapsed: true
    code_download: yes
    df_print: kable
---

<style type="text/css">
body, td{ font-size: 14px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
h1.title {
  font-size: 24px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
}
h3 { /* Header 3 */
    font-size: 14px;
  color: Black;
}
h4 { /* Header 4 */
    font-size: 14px;
  color: Black;
}
#TOC {
  color: DarkBlue;
  font-size: 16px;
}
</style>

This notebook implements the HSGP model fit and diagnosis on the 2D dataset in case study 5.3 '*Simulated data for 2D and 3D functions*' in [Riutort-Mayol, G., Bürkner, P. C., Andersen, M. R., Solin, A., & Vehtari, A. (2020; updated in 2022). Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming. arXiv preprint arXiv:2004.11408.](https://arxiv.org/pdf/2004.11408.pdf)

```{r include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

# Load packages

```{r message=FALSE, warn=FALSE, eval=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(RColorBrewer)
library(ggplot2)
library(geoR)
```

# Synthetic dataset

A synthetic dataset $\boldsymbol{y}$ with a true data generating function $f(\boldsymbol{x})$, with $\boldsymbol{x}_i \in {\rm I\!R}^2$, where $f(\boldsymbol{x})$ is a Gaussian process (GP) with a squared exponential kernel. Gaussian errors $\boldsymbol{\epsilon}$ are added to $f(\boldsymbol{x})$ to form the final synthetic noisy dataset $\boldsymbol{y}$.

\begin{align*}
\begin{split}
\boldsymbol{y} &= \boldsymbol{f} + \boldsymbol{\epsilon} \\
\boldsymbol{\epsilon} &\sim \text{Normal}(\boldsymbol{0}, \sigma^2 \boldsymbol{I}) \\
f(\boldsymbol{x}) &\sim \text{GP}(0, k(\boldsymbol{x}, \boldsymbol{x}', \theta)),
\end{split}
\end{align*}

where $k(\boldsymbol{x}, \boldsymbol{x}', \theta)$ is a squared exponential kernel as a function of input values $\boldsymbol{x}$ and parameters $\theta=(\alpha,\ell)$, where $\alpha$ is the marginal variance of the GP and $\ell$ the lengthscale of the kernel.

## Draw the true data generative function $f(x)$ from a GP

### Input predictive space

Matrix of 2D-predictive input points `x_pred`, length of `x_pred` (`N_pred`)

```{r eval=TRUE}
x_pred <- pred_grid(coords= c(-1,1), y.coords= c(-1,1), by= 0.04)  
N_pred <- dim(x_pred)[1]
N_pred
```

Number of dimensions `D`

```{r eval=TRUE}
D <- 2 
D
```

Half range of the input domain

```{r eval=TRUE, message=FALSE, warn=FALSE}
S <- abs((max(x_pred) - min(x_pred))/2)
S
```

### Draw $f(x)$ using Stan

Marginal variance `gpscale` ($\alpha$) and lengthscale `lscale` ($\ell$)

```{r eval=TRUE}
gpscale <- 1.1
lscale <- vector()
lscale[1] <- 0.1
lscale[2] <- 0.3
```

Random variable `eta` to simulate from the GP

```{r eval=TRUE}
seed <- 88646
set.seed(seed)
eta <- rnorm(N_pred, 0, 1)
```

Data to be passed to Stan

```{r eval=TRUE}
standata <- list(D= D, 
                 N_pred= N_pred, 
                 x_pred= x_pred,
                 gpscale= gpscale, 
                 lscale= lscale, 
                 eta= eta)
```

Model sampling

```{r eval=TRUE, results=FALSE}
simuGP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_SIMU_2D.stan")
fit_SIMU <- simuGP_mod$sample(data= standata,
                                iter_warmup=1,
                                iter_sampling=1,
                                chains=1, thin=1, init=0.5,
                                fixed_param=TRUE)
fit_SIMU <- read_stan_csv(fit_SIMU$output_files())
```

Drawn $f(x)$

```{r eval=TRUE, echo=FALSE}
f_true <- extract(fit_SIMU, pars="f_true", permuted = TRUE, inc_warmup = FALSE, include = TRUE)$f_true[1,]
xx <- unique(x_pred[,1])

col1 <- brewer.pal(11,"BrBG")[11:1] 
brk <- seq(-4.00,max(f_true), length.out= 12)

filled.contour(xx, xx, matrix(f_true,length(xx),length(xx)), levels = brk, col = brewer.pal(11,"BrBG")[11:1], plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); abline(h= seq(-1,1,0.5), col=grey(0.7)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.7), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03))
text(-0.3,0.9,"(a)", font=2, cex=1.5);
text( par("usr")[1] - 0.30, mean(c(-1,1)), labels = expression(x[2]), srt = 0, xpd = TRUE, cex=2, font=1, family="serif")
```

## Adding noise to $f(x)$

```{r eval=TRUE}
set.seed(1234)
y_pred <- f_true + rnorm(N_pred, 0, 0.2)
```

## Sampling data set

```{r eval=TRUE}
nobs <- 150 
set.seed(4567)
vv_sample <- sample(1:N_pred, nobs)
N_sample <- length(vv_sample)
```

# GP model fitting

## Data

Data to Stan

```{r eval=TRUE}
standata_GP <- list(D= D, 
                    N_sample= N_sample,
                    vv_sample= vv_sample,
                    N_pred= N_pred, 
                    x_pred= x_pred, 
                    y_pred= y_pred) 
str(standata_GP)
```

## Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
GP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_GP_marginalized_2D.stan")
fit_GP <- GP_mod$sample(data= standata_GP,
                          iter_warmup=50,
                          iter_sampling=50,
                          chains=4, thin=1, init=2,
                          save_warmup=FALSE)
fit_GP <- read_stan_csv(fit_GP$output_files())
```

Summaries and simulation chains for the variables after warmup

```{r fig.height=5, fig.width=10, eval=TRUE}
param = c("lscale","gpscale[1]","noise")
print(summary(fit_GP, pars = param, probs = c(0.025, 0.5, 0.975))$summary)
traceplot(fit_GP, pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

## Posterior predictive distributions

```{r eval=TRUE, include=TRUE, echo=FALSE}
gp12_f_pred <- summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
gp12_elpd <- summary(fit_GP, pars= c("elpd"), probs = c(0.025, 0.5, 0.975), digits_summary= 4)$summary[,c(1,4,6)]
gp12_time <- sum(get_elapsed_time(fit_GP)[2])/(fit_GP@sim$iter - fit_GP@sim$warmup)
gp12_mse <- sqrt(sum((f_true - gp12_f_pred[,1])^2)/N_pred)

col1 <- brewer.pal(11,"BrBG")[11:1] 
# brk <- seq(min(gp12_f_pred[,1]),max(gp12_f_pred[,1]), length.out= 12)

filled.contour(xx, xx, matrix(gp12_f_pred[,1],length(xx),length(xx)), levels = brk, col = brewer.pal(11,"BrBG")[11:1], plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); points(x_pred[vv_sample,], cex=1, lwd=1, pch=1);  abline(h= seq(-1,1,0.5), col=grey(0.7)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.7), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03)) #; grid(col=grey(0.90), lty=1)
text(-0.3,0.9,"(b)   GP", font=2, cex=1.5);
text( par("usr")[1] - 0.30, mean(c(-1,1)), labels = expression(x[2]), srt = 0, xpd = TRUE, cex=2, font=1, family="serif")
```

## Error plot

```{r eval=TRUE, include=TRUE, echo=FALSE}
error <- f_true - gp12_f_pred[,1]

col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(6), colorRampPalette(c("white", "red"), bias=10)(6)[2:6])#brewer.pal(11,"RdBu")[1:11]
brk_error <- seq(min(error),max(error), length.out= 12)

filled.contour(xx, xx, matrix(error,length(xx),length(xx)), levels= brk_error, col= col3, plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); points(x_pred[vv_sample,], cex=1, lwd=1, pch=1);  abline(h= seq(-1,1,0.5), col=grey(0.6)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.6), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03))
text(-0.3,1.12,"(a)", font=2, cex=1.5);
text( par("usr")[1] - 0.30, mean(c(-1,1)), labels = expression(x[2]), srt = 0, xpd = TRUE, cex=2, font=1, family="serif")
```

# Functional relationships among $m$, $\ell$ and $c$ parameters in the HSGP approach

Relationship among the number of basis functions $m$, the lengthscale $\ell$ and the boundary factor $c$ for a **squared exponential kernel** in the HSGP approach:

$$\begin{align*}
m= 1.75 \frac{c}{\frac{\ell}{S}} \;\; \Leftrightarrow \;\; \frac{\ell}{S}= 1.75 \, \frac{c}{m}  (\#eq:m-l-QE)
\end{align*}$$

```{r eval=TRUE, include=TRUE}
m_QE <- function(c,l,S) ceiling(1.75 * c / (l/S))
l_QE <- function(c,m,S) round(S * 1.75 * c / m, 3)   
```

with

$$\begin{align*}
c \, \geq \, 3.2 \, \frac{\ell}{S}\;\; \& \;\;c \, \geq \frac{1}{2}  (\#eq:c-QE)
\end{align*}$$

as a function of $\ell$ and $S$, where $S$ is the half range of the input domain,

```{r eval=TRUE, include=TRUE}
c_vs_l_QE <- function(l,S){
    c =  3.2*l/S
    if(c < 1.2)
      c = 1.2
  c
}
```

# Diagnostic function over the lengthscale of the HSGP model

The lengthscale diagnosis function is defined as:

$$\begin{align*}
\hat{\ell} + 0.01 \geq \ell  (\#eq:diagnostic)
\end{align*}$$

```{r eval=TRUE, include=TRUE}
diagnostic <- function(l,l_hat) l_hat + 0.01 >= l
```

# HSGP fit and diagnosis

Initializing programming objects

```{r eval=TRUE}
standata <- list()
fit<- list()
diagnosis <- list()

l1 <- vector()  # lengthscale for dimension 1
l2 <- vector()  # lengthscale for dimension 2
c1 <- vector()  # boundary factor for dimension 1
c2 <- vector()  # boundary factor for dimension 2
m1 <- vector()  # number of basis functions for dimension 1
m2 <- vector()  # number of basis functions for dimension 2
l_hat1 <- vector()  # HSGP lengthscale estimate for dimension 1
l_hat2 <- vector()  # HSGP lengthscale estimate for dimension 2
check1 <- vector()  # Diagnostic for dimension 1
check2 <- vector()  # Diagnostic for dimension 2
rmse_gp <- vector()  # root mean square error of GP model
rmse <- vector()  # root mean square error of HSGP model
eR <- vector()  # coefficient of determination of HSGP model
elpd <- vector()  # expected log predictive density of HSGP model
```

## Iteration 1

```{r eval=TRUE, include=TRUE, results='hold'}
i <- 1
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If iteration $i=1$:

1. Making a first guess for lengthscale $\ell_i$

2. The boundary factor $c_i$ has to fulfill equation \@ref(eq:c-QE)

3. The number of basis functions $m_i$ has to fulfill equation \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE}
if(i==1){
  l1[i] <- 0.5
  c1[i] <- c_vs_l_QE(l=l1[i], S)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S)

  l2[i] <- 1
  c2[i] <- c_vs_l_QE(l=l2[i], S)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
		mm = mm+1
		indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE, message=FALSE, warn=FALSE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * S, 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE, include=TRUE}
f <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against data
residual_gp <- gp12_f_pred[vv_sample] - f[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell}_i$ in the HSGP approximation

```{r eval=TRUE, include=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell}_i + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE, include=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

```{r eval=TRUE, include=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat > l", "rmse_gp", "rmse", "R2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 2

```{r eval=TRUE, include=TRUE, results='hold'}
i <- 2
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i=2$:

1. Update $m_i = m_{i-1} + 2$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE, results='hold'}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i==2){
  m1[i] <- m1[i-1] + 2
  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S)
  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S)

  m2[i] <- m2[i-1] + 2
  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S)
  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE, results='hold'}
if(!diagnostic(l1[i-1], l_hat1[i-1])){
  l1[i] <- l_hat1[i-1]
  c1[i] <- c_vs_l_QE(l=l1[i], S=S)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S)

  l2[i] <- l_hat2[i-1]
  c2[i] <- c_vs_l_QE(l=l2[i], S=S)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
			mm = mm+1
			indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * S, 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE, include=TRUE}
f <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against data
residual_gp <- gp12_f_pred[vv_sample] - f[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell_i}$ in the HSGP approximation

```{r eval=TRUE, include=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 3)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 3)
```

Check the diagnostic of whether $\, \hat{\ell_i} + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE, include=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

```{r eval=TRUE, include=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat > l", "rmse_gp", "rmse", "R2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 3

```{r eval=TRUE, include=TRUE, results='hold'}
i <- 3
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} + 0.01 \geq \ell_{i-2}$ \@ref(eq:diagnostic) is FALSE:

1. Update $m_i = m_{i-1} + 2$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE, results='hold'}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & !diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1] + 2
  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S)
  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S)

  m2[i] <- m2[i-1] + 2
  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S)
  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} + 0.01 \geq \ell_{i-2}$ \@ref(eq:diagnostic) is TRUE:

1. Update $m_i = m_{i-1}$

2. Update $c_i = c_{i-1}$

3. Update $\ell_i = \ell_{i-1}$

```{r eval=TRUE, include=TRUE, results='hold'}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1]
  c1[i] <- c1[i-1]
  l1[i] <- l1[i-1]

  m2[i] <- m2[i-1]
  c2[i] <- c2[i-1]
  l2[i] <- l2[i-1]
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE, results='hold'}
if(!diagnostic(l1[i-1], l_hat1[i-1])){
  l1[i] <- l_hat1[i-1]
  c1[i] <- c_vs_l_QE(l=l1[i], S=S)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S)

  l2[i] <- l_hat2[i-1]
  c2[i] <- c_vs_l_QE(l=l2[i], S=S)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
			mm = mm+1
			indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * S, 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE, include=TRUE}
f <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against data
residual_gp <- gp12_f_pred[vv_sample] - f[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell_i}$ in the HSGP approximation

```{r eval=TRUE, include=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell_i} + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE, include=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

```{r eval=TRUE, include=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat > l", "rmse_gp", "rmse", "R2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 4

```{r eval=TRUE, include=TRUE, results='hold'}
i <- 4
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} \geq \ell_{i-2}$ \@ref(eq:diagnostic) is FALSE:

1. Update $m_i = m_{i-1} + 2$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE, results='hold'}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & !diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1] + 2
  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S)
  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S)

  m2[i] <- m2[i-1] + 2
  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S)
  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} + 0.01 \geq \ell_{i-2}$ \@ref(eq:diagnostic) is TRUE:

1. Update $m_i = m_{i-1}$

2. Update $c_i = c_{i-1}$

3. Update $\ell_i = \ell_{i-1}$

```{r eval=TRUE, include=TRUE, results='hold'}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1]
  c1[i] <- c1[i-1]
  l1[i] <- l1[i-1]

  m2[i] <- m2[i-1]
  c2[i] <- c2[i-1]
  l2[i] <- l2[i-1]
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-QE)

```{r eval=TRUE, include=TRUE, results='hold'}
if(!diagnostic(l1[i-1], l_hat1[i-1])){
  l1[i] <- l_hat1[i-1]
  c1[i] <- c_vs_l_QE(l=l1[i], S=S)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S)

  l2[i] <- l_hat2[i-1]
  c2[i] <- c_vs_l_QE(l=l2[i], S=S)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
			mm = mm+1
			indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * S, 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE, include=TRUE}
f <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against data
residual_gp <- gp12_f_pred[vv_sample] - f[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell_i}$ in the HSGP approximation

```{r eval=TRUE, include=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell_i} + 0.01 \geq \ell_i$

```{r eval=TRUE, include=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

```{r eval=TRUE, include=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat > l", "rmse_gp", "rmse", "R2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

### 2022-05-15 Paper/Case-study_2D&3D-Simulated-data/2Ddataset/Rnb_case2D_diagnosis.Rmd
reorganize files

--- 2022-05-14 Paper/Case-study_2D&3D-Simulated-data/Rnb_case2D_diagnosis.Rmd
@@
 library(ggplot2)
-library(geoR)
 ```
@@
 
-where $k(\boldsymbol{x}, \boldsymbol{x}', \theta)$ is a squared exponential kernel as a function of input values $\boldsymbol{x}$ and parameters $\theta=(\alpha,\ell)$, where $\alpha$ is the marginal variance of the GP and $\ell$ the lengthscale of the kernel.
+where $k(\boldsymbol{x}, \boldsymbol{x}', \theta)$ is a squared exponential kernel as a function of input values $\boldsymbol{x}$ and parameters $\theta=(\alpha,\boldsymbol{\ell})$, where $\alpha$ is the marginal variance of the GP function and $\boldsymbol{\ell}=(\ell_1,\ell_2)$ the lengthscale of the kernel for dimensions 1 and 2, respectively.
 
@@
 
-Matrix of 2D-predictive input points `x_pred`, length of `x_pred` (`N_pred`)
-
-```{r eval=TRUE}
-x_pred <- pred_grid(coords= c(-1,1), y.coords= c(-1,1), by= 0.04)  
+Matrix of 2D-predictive input points `x_pred`
+
+```{r eval=TRUE}
+x_pred <- expand.grid(x1= seq(-1,1,by=0.04), x2= seq(-1,1,by=0.04))  
+str(x_pred)
+```
+
+Length of `x_pred` (`N_pred`)
+
+```{r eval=TRUE}
 N_pred <- dim(x_pred)[1]
@@
 
-Half range of the input domain
-
-```{r eval=TRUE, message=FALSE, warn=FALSE}
-S <- abs((max(x_pred) - min(x_pred))/2)
-S
+Half range `S` of every input dimension
+
+```{r eval=TRUE}
+S1 <- abs((max(x_pred$x1) - min(x_pred$x1))/2)
+S1
+S2 <- abs((max(x_pred$x2) - min(x_pred$x2))/2)
+S2
 ```
@@
 
-Marginal variance `gpscale` ($\alpha$) and lengthscale `lscale` ($\ell$)
+Marginal variance `gpscale` ($\alpha$) and multivariate lengthscale `lscale` ($\boldsymbol{\ell}$)
 
@@
 
-```{r eval=TRUE, results=FALSE}
+```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
 simuGP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_SIMU_2D.stan")
@@
 
+```{r eval=TRUE}
+f_true <- extract(fit_SIMU, pars="f_true", permuted = TRUE, inc_warmup = FALSE, include = TRUE)$f_true[1,]
+```
+
 ```{r eval=TRUE, echo=FALSE}
-f_true <- extract(fit_SIMU, pars="f_true", permuted = TRUE, inc_warmup = FALSE, include = TRUE)$f_true[1,]
 xx <- unique(x_pred[,1])
@@
 
-## Posterior predictive distributions
-
-```{r eval=TRUE, include=TRUE, echo=FALSE}
-gp12_f_pred <- summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
-gp12_elpd <- summary(fit_GP, pars= c("elpd"), probs = c(0.025, 0.5, 0.975), digits_summary= 4)$summary[,c(1,4,6)]
-gp12_time <- sum(get_elapsed_time(fit_GP)[2])/(fit_GP@sim$iter - fit_GP@sim$warmup)
-gp12_mse <- sqrt(sum((f_true - gp12_f_pred[,1])^2)/N_pred)
-
+## Posterior predictive distribution
+
+```{r eval=TRUE}
+f_gp <- summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
+```
+
+```{r eval=TRUE, echo=FALSE}
 col1 <- brewer.pal(11,"BrBG")[11:1] 
-# brk <- seq(min(gp12_f_pred[,1]),max(gp12_f_pred[,1]), length.out= 12)
-
-filled.contour(xx, xx, matrix(gp12_f_pred[,1],length(xx),length(xx)), levels = brk, col = brewer.pal(11,"BrBG")[11:1], plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); points(x_pred[vv_sample,], cex=1, lwd=1, pch=1);  abline(h= seq(-1,1,0.5), col=grey(0.7)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.7), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03)) #; grid(col=grey(0.90), lty=1)
+# brk <- seq(min(f_gp[,1]),max(f_gp[,1]), length.out= 12)
+
+filled.contour(xx, xx, matrix(f_gp[,1],length(xx),length(xx)), levels = brk, col = brewer.pal(11,"BrBG")[11:1], plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); points(x_pred[vv_sample,], cex=1, lwd=1, pch=1);  abline(h= seq(-1,1,0.5), col=grey(0.7)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.7), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03)) #; grid(col=grey(0.90), lty=1)
 text(-0.3,0.9,"(b)   GP", font=2, cex=1.5);
@@
 
-```{r eval=TRUE, include=TRUE, echo=FALSE}
-error <- f_true - gp12_f_pred[,1]
-
+```{r eval=TRUE}
+error <- f_true - f_gp[,1]
+```
+
+```{r eval=TRUE, echo=FALSE}
 col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(6), colorRampPalette(c("white", "red"), bias=10)(6)[2:6])#brewer.pal(11,"RdBu")[1:11]
@@
 
-```{r eval=TRUE, include=TRUE}
+```{r eval=TRUE}
 m_QE <- function(c,l,S) ceiling(1.75 * c / (l/S))
@@
 
-as a function of $\ell$ and $S$, where $S$ is the half range of the input domain,
-
-```{r eval=TRUE, include=TRUE}
+as a function of $\ell$ and $S$, where $S$ is the half range of the input domain
+
+```{r eval=TRUE}
 c_vs_l_QE <- function(l,S){
@@
 
-# Diagnostic function over the lengthscale of the HSGP model
-
-The lengthscale diagnosis function is defined as:
+# Lengthscale diagnostic
+
+The HSGP lengthscale diagnostic is defined as:
 
@@
 
-```{r eval=TRUE, include=TRUE}
+```{r eval=TRUE}
 diagnostic <- function(l,l_hat) l_hat + 0.01 >= l
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 i <- 1
@@
 
-```{r eval=TRUE, include=TRUE}
+```{r eval=TRUE}
 if(i==1){
   l1[i] <- 0.5
-  c1[i] <- c_vs_l_QE(l=l1[i], S)
-  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S)
+  c1[i] <- c_vs_l_QE(l=l1[i], S=S1)
+  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S1)
 
   l2[i] <- 1
-  c2[i] <- c_vs_l_QE(l=l2[i], S)
-  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S)
+  c2[i] <- c_vs_l_QE(l=l2[i], S=S2)
+  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S2)
 }
@@
 
-```{r eval=TRUE, message=FALSE, warn=FALSE}
+```{r eval=TRUE}
 standata[[i]] <- list(D= D, 
@@
                       indices= indices,
-                      L= c(c1[i], c2[i]) * S, 
+                      L= c(c1[i], c2[i]) * c(S1, S2), 
                       x_pred= x_pred, 
@@
 
-```{r eval=TRUE, include=TRUE}
-f <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
+```{r eval=TRUE}
+f_hsgp <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
 
 # rmse against data
-residual <- standata[[i]]$y_pred[vv_sample] - f[,1]
+residual <- standata[[i]]$y_pred[vv_sample] - f_hsgp[,1]
 rmse[i] <- round(sqrt(mean(residual^2)), 2)
 
-# rmse against data
-residual_gp <- gp12_f_pred[vv_sample] - f[,1]
+# rmse against gp
+residual_gp <- f_gp[vv_sample] - f_hsgp[,1]
 rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)
@@
 
-```{r eval=TRUE, include=TRUE}
+```{r eval=TRUE}
 l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
@@
 
-```{r eval=TRUE, include=TRUE}
+```{r eval=TRUE}
 check1[i] <- diagnostic(l1[i], l_hat1[i])
@@
 
-```{r eval=TRUE, include=TRUE}
+  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.
+
+  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.
+
+```{r eval=TRUE}
 diagnosis[[i]] <- data.frame(iter= c(i, i),
@@
                               elpd = c(NA, elpd[i]))
-names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat > l", "rmse_gp", "rmse", "R2", "elpd")
+names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")
 
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 i <- 2
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 if(diagnostic(l1[i-1], l_hat1[i-1]) & i==2){
   m1[i] <- m1[i-1] + 2
-  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S)
-  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S)
-
+  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S1)
+  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S1)
+}
+
+if(diagnostic(l2[i-1], l_hat2[i-1]) & i==2){
   m2[i] <- m2[i-1] + 2
-  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S)
-  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S)
+  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S2)
+  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S2)
 }
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 if(!diagnostic(l1[i-1], l_hat1[i-1])){
   l1[i] <- l_hat1[i-1]
-  c1[i] <- c_vs_l_QE(l=l1[i], S=S)
-  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S)
-
+  c1[i] <- c_vs_l_QE(l=l1[i], S=S1)
+  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S1)
+}
+
+if(!diagnostic(l2[i-1], l_hat2[i-1])){
   l2[i] <- l_hat2[i-1]
-  c2[i] <- c_vs_l_QE(l=l2[i], S=S)
-  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S)
+  c2[i] <- c_vs_l_QE(l=l2[i], S=S2)
+  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S2)
 }
@@
 
-```{r eval=TRUE, include=TRUE}
+```{r eval=TRUE}
 l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 3)
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 i <- 3
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & !diagnostic(l1[i-2], l_hat1[i-2])){
   m1[i] <- m1[i-1] + 2
-  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S)
-  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S)
-
+  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S1)
+  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S1)
+}
+
+if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & !diagnostic(l2[i-2], l_hat2[i-2])){  
   m2[i] <- m2[i-1] + 2
-  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S)
-  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S)
+  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S2)
+  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S2)
 }
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & diagnostic(l1[i-2], l_hat1[i-2])){
@@
   l1[i] <- l1[i-1]
-
+}
+
+if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & diagnostic(l2[i-2], l_hat2[i-2])){  
   m2[i] <- m2[i-1]
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 i <- 4
@@
 
-```{r eval=TRUE, include=TRUE, results='hold'}
+```{r eval=TRUE}
 if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & !diagnostic(l1[i-2], l_hat1[i-2])){
   m1[i] <- m1[i-1] + 2
-  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S)
-  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S)
-
+  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S1)
+  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S1)
+}
+
+if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & !diagnostic(l2[i-2], l_hat2[i-2])){ 
   m2[i] <- m2[i-1] + 2
-  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S)
-  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S)
+  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S2)
+  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S2)
 }
@@
   l1[i] <- l1[i-1]
-
+}
+
+if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & diagnostic(l2[i-2], l_hat2[i-2])){ 
   m2[i] <- m2[i-1]

### 2022-05-17 Paper/Case-study_1D-Simulated-data/updated_version/Diagnosis_case_study/Rnb_case1D_diagnosis.Rmd
update case study I

---
title: '1D simulated case study: Diagnosis of the HSGP approximation'
author: 'Gabriel Riutort-Mayol'
date: "`r format(Sys.Date())`."
output:
  bookdown::html_document2:
    theme: readable
    toc: yes
    toc_depth: 4
    number_sections: true
    toc_float: 
      collapsed: true
    code_download: yes
    df_print: kable
---

<style type="text/css">
body, td{ font-size: 14px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
h1.title {
  font-size: 24px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
}
h3 { /* Header 3 */
    font-size: 14px;
  color: Black;
}
h4 { /* Header 4 */
    font-size: 14px;
  color: Black;
}
#TOC {
  color: DarkBlue;
  font-size: 16px;
}
</style>

This notebook implements the HSGP model fit and diagnosis on the 1D dataset in case study 5.1 '*Simulated data for a 1D function*' in [Riutort-Mayol, G., Bürkner, P. C., Andersen, M. R., Solin, A., & Vehtari, A. (2020; updated in 2022). Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming. arXiv preprint arXiv:2004.11408.](https://arxiv.org/pdf/2004.11408.pdf)

```{r include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

# Load packages

```{r message=FALSE, warn=FALSE, eval=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(RColorBrewer)
library(ggplot2)
```

# Synthetic dataset

A synthetic dataset $\boldsymbol{y}$ with a true data generating function $f(\boldsymbol{x})$, with $x_i \in {\rm I\!R}$, where $f(\boldsymbol{x})$ is a Gaussian process (GP) with a squared exponential kernel. Gaussian errors $\boldsymbol{\epsilon}$ are added to $f(\boldsymbol{x})$ to form the final synthetic noisy dataset $\boldsymbol{y}$.

\begin{align*}
\begin{split}
\boldsymbol{y} &= \boldsymbol{f} + \boldsymbol{\epsilon} \\
\boldsymbol{\epsilon} &\sim \text{Normal}(\boldsymbol{0}, \sigma^2 \boldsymbol{I}) \\
f(x) &\sim \text{GP}(0, k(x, x', \theta)),
\end{split}
\end{align*}

where $k(x, x', \theta)$ is a squared exponential kernel as a function of input values $x$ and parameters $\theta=(\alpha,\ell)$, where $\alpha$ is the marginal variance of the GP function and $\ell$ the lengthscale of the kernel.

## Draw the true data generative function $f(x)$ from a GP

### Input predictive space

Matrix of 1D-predictive input points `x_pred`

```{r eval=TRUE}
x_pred <- seq(-1, 1, 0.002)  
str(x_pred)
```

Vector of indices of `x_pred` (`vv_pred`)

```{r eval=TRUE}
vv_pred <- 1:length(x_pred)
```

Length of `x_pred` (`N_pred`)

```{r eval=TRUE}
N_pred <- length(x_pred)
N_pred
```

Number of dimensions `D`

```{r eval=TRUE}
D <- 2 
D
```

Half range `S` of the input domain

```{r eval=TRUE}
S <- abs((max(x_pred) - min(x_pred))/2)
S
```

### Draw $f(x)$ using Stan

`param_v` identifies the order $\nu$ of the kernel

```{r eval=TRUE}
param_v <- 3
```

Marginal variance `gpscale` ($\alpha$) and lengthscale `lscale` ($\ell$)

```{r eval=TRUE}
gpscale <- 1
lscale <- 0.2
```

Random variable `eta` to simulate from the GP

```{r eval=TRUE}
seed <- 88646
set.seed(seed)
eta <- rnorm(N_pred, 0, 1)
```

Data to be passed to Stan

```{r eval=TRUE}
standata <- list(x= x_pred, 
                  N= N_pred, 
                  param_v= param_v, 
                  gpscale= gpscale, 
                  lscale= lscale, 
                  eta= eta)
```

Model sampling

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
simuGP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_SIMU_1D.stan")
fit_SIMU <- simuGP_mod$sample(data= standata,
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=1, thin=1, init=0.5,
                                fixed_param=TRUE)
fit_SIMU <- read_stan_csv(fit_SIMU$output_files())
```

Drawn $f(x)$

```{r eval=TRUE, fig.height=2, fig.width=5}
f_true <- summary(fit_SIMU, pars = "f", probs = c(0.025, 0.5, 0.975))$summary[,1]
ggplot(data.frame(f=f_true, x=x_pred), aes(x=x, y=f))  +
  geom_line(color = 'black', size=1) +
  theme_classic()
```

### Adding noise to $f(x)$

Five different noisy datasets $\boldsymbol{y}$ over true data generative function $f(x)$

```{r eval=TRUE, fig.height=2, fig.width=5}
seed <- 10453
set.seed(seed)
sd <- 0.2
y <- f_true + rnorm(N_pred, 0, sd)
ggplot(data.frame(f=f_true, x=x_pred), aes(x=x, y=f))  +
    geom_line(color = 'black', size=1) +
    geom_point(y= y, size=0.7, color='black') +
    theme_classic()
```

## Data

A random sample `x_sample` of size n=250 from the input predicting points `x_pred`

```{r eval=TRUE, fig.height=2, fig.width=5}
set.seed(seed)
vv_sample <- sort(sample(vv_pred, 250))
x_sample <- x_pred[vv_sample]
N_sample <- length(x_sample)
ggplot(data.frame(f=f_true[vv_sample], x=x_sample), aes(x=x, y=f))  +
    geom_line(color = 'black', size=1) +
    geom_point(y= y[vv_sample], size=0.9, color='black') +
    theme_classic()
```

### Interpolation predicting data

Select samples from `x_sample` for interpolation

```{r eval=TRUE}
set.seed(seed)
vv_inter <- vv_sample[which((x_sample>(-0.50)&x_sample<(-0.35))|(x_sample>(0.13)&x_sample<0.30))]
N_inter <- length(vv_inter)
x_inter <- x_pred[vv_inter]
```

### Extrapolation predicting data

Select samples from `x_sample` for extrapolation

```{r eval=TRUE}
vv_extra <- vv_sample[which(x_sample<(-0.8)|x_sample>0.82)]
N_extra <- length(vv_extra)
x_extra <- x_pred[vv_extra]
```

### Training data

Select samples from `x_sample` for training

```{r eval=TRUE}
vv_train <- setdiff(vv_sample, c(vv_extra, vv_inter))
N_train <- length(vv_train)
x_train <- x_pred[vv_train]
```

### Plot of training (black), interpolation (red) and extrapolation (red) points

```{r eval=TRUE, fig.height=2, fig.width=5, echo=FALSE}
ggplot(data.frame(f=f_true[vv_sample], x=x_sample), aes(x=x, y=f), size=0.7, color='black')  +
  geom_line() +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= as.vector(y[vv_train]), x=x_train), size=0.7, color='black') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= as.vector(y[vv_inter]), x=x_inter), size=0.7, color='red') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= as.vector(y[vv_extra]), x=x_extra), size=0.7, color='red') +
  theme_classic()
```

# GP model fitting

Initialization of programming objects

```{r eval=TRUE, eval=TRUE}
fit_GP <- list()
```

## Data

Data to Stan

```{r eval=TRUE}
standata_GP <- list(x_pred= x_pred, 
                     y= y, 
                     vv_train= vv_train, 
                     N_pred= N_pred, 
                     N_train= N_train, 
                     param_v= param_v,
                     f_true= f_true) 
```

## Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
GP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_GP_marginalized_1D.stan")
fit_GP <- GP_mod$sample(data= standata_GP,
                          iter_warmup=50,
                          iter_sampling=50,
                          chains=4, thin=1, init=0.5)
fit_GP <- read_stan_csv(fit_GP$output_files())
```

Summaries and simulation chains for the variable estimates after warmup

```{r eval=TRUE, fig.height=3, fig.width=10}
param = c("lscale","gpscale","sigma")
print(summary(fit_GP, pars = param, probs = c(0.025, 0.5, 0.975))$summary)
traceplot(fit_GP, pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

## Posterior predictive distributions

```{r eval=TRUE, include=TRUE, echo=TRUE, echo=FALSE}
f_pred_gp <- summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]
time_GP <- sum(get_elapsed_time(fit_GP)[2])/(fit_GP@sim$iter - fit_GP@sim$warmup)
```

```{r eval=TRUE, fig.height=4, fig.width=10}
df_pred_gp <- data.frame(as.data.frame(f_pred_gp), x=x_pred)
ggplot(df_pred_gp, aes(x=x, y=mean))  +
  geom_line(size=1.3, color='red') +
  geom_ribbon(aes(ymin=X2.5., ymax=X97.5., color=NA), fill='red', size=0.7, linetype=2, alpha=0.1) +
  # geom_line(mapping=aes(x=x, y=f), data=data.frame(f=f_true[vv_sample], x=x_sample), size=0.7, color='black') +  //true function
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red') +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red') +
  scale_color_brewer(palette="Set1") +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12))
```

# Functional relationships among $m$, $\ell$ and $c$ parameters in the HSGP approach

Relationship among the number of basis functions $m$, the lengthscale $\ell$ and the boundary factor $c$ for a **Mattern-3/2 kernel** in the HSGP approach:

$$\begin{align*}
m= 3.42 \, \frac{c}{\frac{\ell}{S}} \;\; \Leftrightarrow \;\; \frac{\ell}{S}= 3.42 \, \frac{c}{m}        (\#eq:m-l-Mattern)
\end{align*}$$

```{r eval=TRUE}
m_Mattern32 <- function(c,l,S) ceiling(3.42 * c / (l/S))  
l_Mattern32 <- function(c,m,S) round(S * 3.42 * c / m, 3)  
```

with

$$\begin{align*}
c \, \geq \, 4.5 \, \frac{\ell}{S}\;\; \& \;\; c \, \geq \frac{1}{2}  (\#eq:c-Mattern)
\end{align*}$$

as a function of $\ell$ and $S$, where $S$ is the half range of the input domain

```{r eval=TRUE}
c_vs_l_Mattern32 <- function(l,S){
    c =  4.5*l/S
    if(c < 1.2)
      c = 1.2
  c
}
```

# Lengthscale diagnostic

The HSGP lengthscale diagnostic is defined as:

$$\begin{align*}
\hat{\ell} + 0.01 \geq \ell  (\#eq:diagnostic)
\end{align*}$$

where $\hat{\ell}$ is the lengthscale estimate in the HSGP approximation and $\ell$ is the minimum lengthscale determined by the $m$ and $c$ used in the HSGP approximation by using equation \@ref(eq:m-l-Mattern)

```{r eval=TRUE}
diagnostic <- function(l,l_hat) l_hat + 0.01 >= l
```

# HSGP model fit and diagnosis

Initializing programming objects.

```{r eval=TRUE}
standata <- list()
fit <- list()
diagnosis <- list()

l <- vector()
c <- vector()
m <- vector()
l_hat <- vector()
l_hat_min <- vector()
l_hat_max <- vector()
sd_hat <- vector()
check <- vector()
rmse <- vector()
rmse_gp <- vector()
eR <- vector()
elpd <- vector()
f <- vector()
```

## Iteration 1

```{r eval=TRUE}
i <- 1
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If iteration $i=1$:

1. Making a first guess for lengthscale $\ell_i$

2. The boundary factor $c_i$ has to fulfill equation \@ref(eq:c-Mattern)

3. The number of basis functions $m_i$ has to fulfill equation \@ref(eq:m-l-Mattern)

```{r eval=TRUE}
if(i==1){
  l[i] <- 0.5
  c[i] <- c_vs_l_Mattern32(l=l[i], S=S)
  m[i] <- m_Mattern32(c=c[i], l=l[i], S=S)
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(y= y, 
                       f_true= f_true, 
                       x_pred= x_pred, 
                       vv_train= vv_train, 
                       N_pred= N_pred, 
                       N_train= N_train, 
                       param_v= param_v, 
                       c= c[i], 
                       L= c[i]*S, 
                       M= m[i])
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_1D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE, fig.height=3, fig.width=10}
param = c("lscale","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior mean functions.

```{r eval=TRUE, fig.height=4, fig.width=10}
f_pred_hsgp <- summary(fit[[i]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]

df_f_GP <- data.frame(as.data.frame(f_pred_gp), x=x_pred, model="GP")
df_f_HSGP <- data.frame(as.data.frame(f_pred_hsgp), x=x_pred, model="HSGP")
df_f_HSGP_GP <- rbind(df_f_GP, df_f_HSGP)

ggplot(df_f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=model, color=model))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"))$summary[,1]

# rmse against data
residual <- standata[[i]]$y[vv_train] - f_hsgp
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_pred_gp[vv_train] -f_hsgp
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell}_i$ in the HSGP approximation

```{r eval=TRUE}
l_hat[i] <- round(summary(fit[[i]], pars = "lscale", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell}_i + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE}
check[i] <- diagnostic(l[i], l_hat[i])
```

Summary diagnosis table

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= i,
                              l= l[i],
                              c= c[i],
                              m= m[i],
                              l_hat= l_hat[i],
                              'l_hat > l' = check[i],
                              rsme_gp = rmse_gp[i],
                              rsme = rmse[i],
                              R2 = eR[i],
                              elpd = elpd[i])
names(diagnosis[[i]]) <- c("iter", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 2

```{r eval=TRUE}
i <- 2
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE:

1. Update $m_i = m_{i-1} + 5$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-Mattern) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-Mattern)

```{r eval=TRUE}
if(diagnostic(l[i-1], l_hat[i-1])){
  m[i] <- m[i-1] + 5
  c[i] <- c_vs_l_Mattern32(l=l_hat[i-1], S=S)
  l[i] <- l_Mattern32(c=c[i], m=m[i], S=S)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-Mattern)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-Mattern)

```{r eval=TRUE}
if(!diagnostic(l[i-1], l_hat[i-1])){
  l[i] <- l_hat[i-1]
  c[i] <- c_vs_l_Mattern32(l=l[i], S=S)
  m[i] <- m_Mattern32(c=c[i], l=l[i], S=S)
}
```

### Data to be passed to Stan

```{r eval=TRUE, message=FALSE, warn=FALSE}
standata[[i]] <- list(y= y, 
                       f_true= f_true, 
                       x_pred= x_pred, 
                       vv_train= vv_train, 
                       N_pred= N_pred, 
                       N_train= N_train, 
                       param_v= param_v, 
                       c= c[i], 
                       L= c[i]*S, 
                       M= m[i])
str(standata[[i]])
```

### Model fitting

Model fitting.

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_1D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE, fig.height=3, fig.width=10}
param = c("lscale","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior mean functions.

```{r eval=TRUE, fig.height=4, fig.width=10}
f_pred_hsgp <- summary(fit[[i]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]

df_f_GP <- data.frame(as.data.frame(f_pred_gp), x=x_pred, model="GP")
df_f_HSGP <- data.frame(as.data.frame(f_pred_hsgp), x=x_pred, model="HSGP")
df_f_HSGP_GP <- rbind(df_f_GP, df_f_HSGP)

ggplot(df_f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=model, color=model))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"))$summary[,1]

# rmse against data
residual <- standata[[i]]$y[vv_train] - f_hsgp
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_pred_gp[vv_train] -f_hsgp
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell}_i$ in the HSGP approximation

```{r eval=TRUE}
l_hat[i] <- round(summary(fit[[i]], pars = "lscale", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell}_i + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE}
check[i] <- diagnostic(l[i], l_hat[i])
```

Summary diagnosis table

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= i,
                              l= l[i],
                              c= c[i],
                              m= m[i],
                              l_hat= l_hat[i],
                              'l_hat > l' = check[i],
                              rsme_gp = rmse_gp[i],
                              rsme = rmse[i],
                              R2 = eR[i],
                              elpd = elpd[i])
names(diagnosis[[i]]) <- c("iter", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 3

```{r eval=TRUE}
i <- 3
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE:

1. Update $m_i = m_{i-1} + 5$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-Mattern) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-Mattern)

```{r eval=TRUE}
if(diagnostic(l[i-1], l_hat[i-1])){
  m[i] <- m[i-1] + 5
  c[i] <- c_vs_l_Mattern32(l=l_hat[i-1], S=S)
  l[i] <- l_Mattern32(c=c[i], m=m[i], S=S)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-Mattern)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-Mattern)

```{r eval=TRUE}
if(!diagnostic(l[i-1], l_hat[i-1])){
  l[i] <- l_hat[i-1]
  c[i] <- c_vs_l_Mattern32(l=l[i], S=S)
  m[i] <- m_Mattern32(c=c[i], l=l[i], S=S)
}
```

### Data to be passed to Stan

```{r eval=TRUE, message=FALSE, warn=FALSE}
standata[[i]] <- list(y= y, 
                       f_true= f_true, 
                       x_pred= x_pred, 
                       vv_train= vv_train, 
                       N_pred= N_pred, 
                       N_train= N_train, 
                       param_v= param_v, 
                       c= c[i], 
                       L= c[i]*S, 
                       M= m[i])
str(standata[[i]])
```

### Model fitting

Model fitting.

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_1D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE, fig.height=3, fig.width=10}
param = c("lscale","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Plot of the posterior mean functions.

```{r eval=TRUE, fig.height=4, fig.width=10}
f_pred_hsgp <- summary(fit[[i]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]

df_f_GP <- data.frame(as.data.frame(f_pred_gp), x=x_pred, model="GP")
df_f_HSGP <- data.frame(as.data.frame(f_pred_hsgp), x=x_pred, model="HSGP")
df_f_HSGP_GP <- rbind(df_f_GP, df_f_HSGP)

ggplot(df_f_HSGP_GP, aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=model, color=model))  +
  geom_line(size=1.3) +
  geom_ribbon(alpha=0.5, color=NA) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_train], x=x_train), size=1, color='black', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_inter], x=x_inter), size=1, color='red', inherit.aes=FALSE) +
  geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y[vv_extra], x=x_extra), size=1, color='red', inherit.aes=FALSE) +
  labs(y="y", x="x") +
  theme_classic() +
  theme(axis.title.y=element_text(angle=0, vjust=0.5, size=12)) 
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"))$summary[,1]

# rmse against data
residual <- standata[[i]]$y[vv_train] - f_hsgp
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_pred_gp[vv_train] -f_hsgp
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell}_i$ in the HSGP approximation

```{r eval=TRUE}
l_hat[i] <- round(summary(fit[[i]], pars = "lscale", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell}_i + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE}
check[i] <- diagnostic(l[i], l_hat[i])
```

Summary diagnosis table

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= i,
                              l= l[i],
                              c= c[i],
                              m= m[i],
                              l_hat= l_hat[i],
                              'l_hat > l' = check[i],
                              rsme_gp = rmse_gp[i],
                              rsme = rmse[i],
                              R2 = eR[i],
                              elpd = elpd[i])
names(diagnosis[[i]]) <- c("iter", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

# Plot of GP and HSGP posterior mean functions.

```{r fig.height=4, fig.width=10, eval=TRUE, message=FALSE, warning=FALSE}
df_f_HSGP <- list()
df_f_HSGP_GP <- list()

for(iter in 1:i){
  df_f_GP_diag <- data.frame(as.data.frame(summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]), x=x_pred, type="GP", iteration=iter)
  df_f_HSGP[[iter]] <- data.frame(as.data.frame(summary(fit[[iter]], pars = c("f_pred"), probs = c(0.025, 0.5, 0.975))$summary[,c(1,4,6)]), x=x_pred, type="HSGP", iteration=iter)
  df_f_HSGP_GP[[iter]] <- rbind(df_f_GP_diag, df_f_HSGP[[iter]])
  df_f_HSGP_GP[[iter]]$iteration <- as.factor(df_f_HSGP_GP[[iter]]$iteration)
}
for(s in 1:i) 
  if(s==1){ 
    df_f_HSGP_GP[[s]] <- df_f_HSGP_GP[[s]]
  }else{
    df_f_HSGP_GP[[s]] <- rbind(df_f_HSGP_GP[[s-1]], df_f_HSGP_GP[[s]])}

ggplot(df_f_HSGP_GP[[i]][df_f_HSGP_GP[[i]]$type=="HSGP",], aes(x=x, y=mean, ymin=X2.5., ymax=X97.5., fill=iteration, color=iteration))  +
  geom_line(size=2) +
  geom_ribbon(alpha=0.1, color=NA) +
  geom_line(mapping=aes(x=x, y=mean), data=df_f_GP_diag, size=2, colour="#e41a1c", linetype=2.5, inherit.aes=FALSE) +
  # geom_ribbon(mapping=aes(x=x, y=mean, ymin=X2.5, ymax=X97.5), data=df_f_GP_diag, alpha=0.2, color=NA, fill="#e41a1c", inherit.aes=FALSE) +
  # geom_point(mapping=aes(x=x, y=y), data=data.frame(y= y, x=x_pred), size=2, color='black', inherit.aes=FALSE) +
  labs(y="", x="x") +
  scale_fill_manual(values = RColorBrewer::brewer.pal(i, "Blues")) +
  scale_color_manual(values = RColorBrewer::brewer.pal(i, "Blues")) +
  theme_classic() +
  theme(legend.position="right", axis.title.y=element_text(angle=0, vjust=0.5), axis.title=element_text(size=20), axis.text=element_text(size=18), legend.title = element_text(size=20), legend.text = element_text(size=18), legend.key.size = unit(1, 'cm'), legend.key.height = unit(0.8, 'cm'), legend.key.width = unit(1, 'cm'))
```

### 2022-07-05 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
revised manuscript

@@
 \usepackage{soul}
-\usepackage{xcolor}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{amssymb}
@@
 \institute{Gabriel Riutort-Mayol \at
-              Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain\\
-              \email{gabriuma@gmail.com}           %  \\
+              Foundation for the Promotion of Health and Biomedical Research of Valencia Region (FISABIO), Spain\\
+              \emph{Corresponding Author} \\
+              \email{gabriuma@gmail.com; riutort\_gab@gva.es}       %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
            Aki Vehtari \at
-              Department of Computer Science, Aalto University, Finland
+              Department of Computer Science, Aalto University, Finland\\
+              \email{aki.vehtari@aalto.fi}
 }
@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the degree of non-linearity of the function to be estimated, which is in turn characterized by the length-scale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity ({\color{blue}wigglyness/smoothness}) of the function to be estimated, which is in turn characterized by the length-scale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 In this work we investigate these relationships for the squared exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
-For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the covariance function.
+For other kernels, the relationships will be slightly different depending on the {\color{blue}smoothness or wigglyness of the non-linear effects generated from the covariance function}.
 
@@
 
-\begin{figure*} %[!t]
-\centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Post_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig1_Sigma_J.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig1_Cov_J.pdf}}
+\begin{figure}
+%\centering
+\includegraphics[width=0.485\textwidth, trim = 2mm 0mm 0mm 0mm, clip]{fig1_Post_J_L.png}
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
-  \label{fig1_Post_J}
-\end{figure*}
+\caption{Mean posterior predictive functions (top), posterior standard deviations (center), and covariance functions (bottom) of both the exact GP model (red line) and the HSGP model for 1) different number of basis functions $m$, with the boundary factor fixed to a large enough value (left column) and 2) different values of the boundary factor $c$, with a large enough fixed number of basis functions (right column).}
+  \label{fig1_Post_J_L}
+\end{figure}
 
@@
 \centering
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Post_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 3mm 0mm, clip]{fig2_Sigma_L.pdf}}
-\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{fig2_Cov_L.pdf}}
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig2_Post.png}}
 \vspace{-1mm}
-\caption{Mean posterior predictive functions (left), posterior standard deviations (center), and covariance functions (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
-  \label{fig2_Post_L}
+\caption{Posterior mean predictive functions (top), posterior standard deviations (center) and  covariance functions (bottom) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$ (columns).}
+  \label{fig2_Post}
 \end{figure*}
@@
 In this example, the length-scale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model.
-Figures~\ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
-For a sufficiently large fixed value of $c$, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation for both the posterior mean or uncertainty.
+Figures~\ref{fig1_Post_J_L} and \ref{fig1_Post_J_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictive mean and standard deviation of the estimated function as well as on the covariance function itself.
+For a sufficiently large fixed value of $c$, Figure \ref{fig1_Post_J_L} shows clearly how $m$ affects the accuracy on the approximation for both the posterior mean or uncertainty.
 It is seen that if the number of basis functions $m$ is too small, the estimated function tend to be overly smooth because the necessary high frequency components are missing.
 In general, the higher the degree of wigglyness of the function to be estimated, the larger number of  basis functions will be required.
-If $m$ is fixed to a sufficiently large value, Figure~\ref{fig2_Post_L} shows that $c$ affects the approximation of the mean mainly near the boundaries, while the approximation of the standard deviation is affected across the whole domain.
+If $m$ is fixed to a sufficiently large value, Figure~\ref{fig1_Post_J_L} shows that $c$ affects the approximation of the mean mainly near the boundaries, while the approximation of the standard deviation is affected across the whole domain.
 The approximation error tends to be bigger for the standard deviation than for the mean.
 
-\begin{figure*}
-\centering
-\begin{tabular}{ c c c }
-\arrayrulecolor{darkgray}\hline
-c = 1.05 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part1.pdf}
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part1.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part1.pdf} & 
-\multirow{43}{*}{ \includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{fig3_legend.pdf}}\\ 
-\arrayrulecolor{lightgray}\cline{1-2}
-c = 1.1 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part2.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part2.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part2.pdf} &\\
-\cline{1-2}
-c = 1.2 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part3.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part3.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part3.pdf} &\\
-\cline{1-2}
-c = 1.5 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part4.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part4.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part4.pdf} & \\
-\cline{1-2}
-c = 2 &
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Post_part5.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{fig3_Sigma_part5.pdf} 
-\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{fig3_Cov_part5.pdf} & \\
-\cline{1-2}
-c = 2.5 &
-\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Post_part6.pdf}
-\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{fig3_Sigma_part6.pdf}
-\includegraphics[scale=0.215, trim = 0mm 4mm 5mm 14mm, clip]{fig3_Cov_part6.pdf} &\\
-\arrayrulecolor{darkgray}\hline
-\end{tabular}
-\caption{Posterior mean predictive functions (left), posterior standard deviations (center) and  covariance functions (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
-  \label{fig3_Post_part1}
-\end{figure*}
-
-\begin{figure*} %[!t]
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig4_MSE_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig4_MSE_vs_c.pdf}}
+Next, we analyze how the interaction between $m$ and $c$ affects the quality of the approximation.
+The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models.
+Figure~\ref{fig2_Post} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and~$c$.
+Figure~\ref{fig3_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
+Figure~\ref{fig4_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
+Looking at the RMSEs in Figure~\ref{fig3_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
+Further, the less conservative choice of $m = 10$ and $c = 1.5$ could also produce a sufficiently accurately approximation depending on the application.
+We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig2_Post}.
+From these results, some general conclusions may be drawn:
+%
+\begin{itemize}
+\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
+\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
+\end{itemize}
+
+\begin{figure} %[!t]
+\centering
+\subfigure{\includegraphics[width=0.49\textwidth, trim = 2mm 0mm 3mm 0mm, clip]{fig3_MSE_vs_J.png}}
 \caption{Root mean square error (RMSE) of the proposed HSGP models computed against the exact GP model.
@@
 }
-  \label{fig4_MSE_vs_J}
-\end{figure*}
-
-\begin{figure*} %[!h]
-\centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig5_lscale_vs_J.pdf}}
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig5_magnitud_vs_J.pdf}}
+\label{fig3_MSE_vs_J}
+\end{figure}
+
+\begin{figure} %[!t]
+\centering
+\subfigure{\includegraphics[width=0.49\textwidth, trim = 2mm 0mm 2mm 0mm, clip]{fig4_lscale_vs_J.png}}
 \caption{Estimated length-scale (left) and marginal variance (right) parameters of both exact GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
-  \label{fig5_lscale_vs_J}
-\end{figure*}
-
-Next, we analyze how the interaction between $m$ and $c$ affects the quality of the approximation.
-The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models.
-Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and~$c$.
-Figure~\ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model.
-Figure~\ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models.
-Looking at the RMSEs in Figure~\ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$.
-Further, the less conservative choice of $m = 10$ and $c = 1.5$ could also produce a sufficiently accurately approximation depending on the application.
-We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig3_Post_part1}.
-From these results, some general conclusions may be drawn:
-%
-\begin{itemize}
-\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
-\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
-\end{itemize}
+\label{fig4_lscale_vs_J}
+\end{figure}
 
@@
 %
-When $c$ is below our recommendations given $\ell$, the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ are changed.
+When $c$ is below our recommendations given $\ell$ ({\color{blue}recommendations that we give later throughout the paper and specifically in the next Sections \ref{subsec_empiric_relations} and \ref{sec_num_equations}}), the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ are changed.
 
@@
 \centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
+\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig5_relationships.png}}
 \caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
-  \label{fig6_relationships}
+  \label{fig5_relationships}
 \end{figure*}
@@
 Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ relative to the half-range $S$ of the input domain).
-Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) with length-scale $\frac{\ell}{S}$ (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+Figure \ref{fig5_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) with {\color{blue}normalized} length-scale $\frac{\ell}{S}$ (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
@@
 \begin{itemize}
-\item As $\ell$ increases, $m$ required for an accurate approximation decrease.
+\item As $\ell$ increases, $m$ required for an accurate approximation decreases.
+({\color{blue}Notice that the larger $\ell$ the smoother the functions generated from a covariance function, and vice versa; see covariance functions equations in Section \ref{ch4_sec_cov}.})
 \item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
 \item For a given $\ell$ there exist a minimum $c$ under which an accurate approximation is never going to be achieved regardless of $m$.
-This fact can be seen in Figure~\ref{fig6_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
+This fact can be seen in Figure~\ref{fig5_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
 As $\ell$ increases, the minimum valid $c$ also increases.
@@
 As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\ell$ and $c$.
-To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically derived the practically useful constant terms.
+To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically ({\color{blue}checked this linear relationship and} derived the practically useful constant terms.
 We require a lower bound for $c$ of $c \geq 1.2$ such that the equations below are precise enough for practical application.
@@
 
-Figure~\ref{fig6_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
+{\color{blue}Note that in the above equations we have made explicit the use of the normalized length-scale by the half-range of the input space ($\frac{\ell}{S}$) (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$).
+By using the normalized length-scale, the equations, figures, and conclusions made throughout the paper can easily be referred to any specific scale of the input space.}
+
+Figure~\ref{fig5_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
 Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$.
@@
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
-However, as an approximation, we can use the unidimensional GP conclusions in Figure~\ref{fig6_relationships} or equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
+However, as an approximation, we can use the unidimensional GP conclusions in Figure~\ref{fig5_relationships} or equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.
 
@@
 In Appendix~\ref{sec_periodic}, we also analyze the accuracy of this linear representation and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
-Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition.
+Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition{\color{blue}, nor is there a need for a normalized length scale, since the length-scale refers to the period of a sinusoidal function}.
 This relationship between $m$ and $\ell$ for a periodic squared exponential covariance function is gathered in Figure~\ref{figB1_m_lscale_periodic} and the numerical equation was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
@@
 If $\hat{\ell}$ does not exceed this value, the approximation may be inaccurate, and $m$ and/or $c$ need to be increased.
-In Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, $m = 10$ and $c = 1.5$ were sufficient for an accurate modeling of function with $\ell = 0.3$, which matches the diagnostic based on equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
+In Figures~\ref{fig3_MSE_vs_J} and \ref{fig4_lscale_vs_J}, $m = 10$ and $c = 1.5$ were sufficient for an accurate modeling of function with $\ell = 0.3$, which matches the diagnostic based on equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
 
@@
 
-Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also provide the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that van be used to minimize the computational cost in repeated computations (e.g., in cross-validation and simulation based calibration).
+Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, the equations in the previous Section~\ref{sec_num_equations} also provide the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that can be used to minimize the computational cost in repeated computations (e.g., in cross-validation and simulation based calibration).
 This is even more useful in multi-dimensional cases ($2 \leq D \leq 4$), where knowing the smallest useful value of $m$ for each dimension has even bigger effect on the total computational cost.
@@
 
-\item[A1.] Make an initial guess on the length-scale $\ell_1$.
-If there is no useful information available, we recommend to start with a large length-scale, about $\ell_1 \in [0.5, 1]$.
-Set the iteration index to $i=1$.
-
-\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
-
-\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
-
-If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell_{i+1} = \hat{\ell}_{i}$, increase the iteration index $i = i + 1$, and go back to A2.
+\item[A1.] Set the iteration index to $i=1$.
+Make an initial guess on the length-scale $\ell^{(i)}$.
+If there is no useful information available, we recommend to start with a large length-scale, about $\ell^{(i)} \in [0.5, 1]$.
+({\color{blue}Remember that $\ell$ refers to the normalized length-scale $\frac{\ell}{S}$.})
+
+\item[A2.] Obtain the minimum valid boundary factor $c^{(i)}$ determined by $\ell = \ell^{(i)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m^{(i)}$ determined by $\ell = \ell^{(i)}$ and $c = c^{(i)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[A4.] Fit an HSGP model using $m^{(i)}$ and $c^{(i)}$ and ensure convergence of the MCMC chains.
+
+\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(i)} - 0.01 \geq \ell^{(i)}$.
+
+If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell^{(i+1)} = \hat{\ell}^{(i)}$, increase the iteration index $i = i + 1$, and go back to A2.
 
@@
 
-\item[B2.] Set $m_{i+1} = m_{i} + 5$ and increase the iteration index $i = i + 1$.
-
-\item[B3.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \hat{\ell}_{i-1}$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[B4.] Obtain the minimum valid length-scale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[B5.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
-
-\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}_i - 0.01 \geq \ell_i$.
-
-Check the stability of both $\hat{\ell}_i$ and the measures of predictive accuracy relative to the previous iteration.
+\item[B2.] Set $m^{(i+1)} = m^{(i)} + 5$ and increase the iteration index $i = i + 1$.
+
+\item[B3.] Obtain the minimum valid boundary factor $c^{(i)}$ determined by $\ell = \hat{\ell}^{(i-1)}$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[B4.] Obtain the minimum valid length-scale $\ell^{(i)}$ that can be accurately estimated by $m = m^{(i)}$ and $c = c^{(i)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[B5.] Fit an HSGP model using $m^{(i)}$ and $c^{(i)}$ and ensure convergence of the MCMC chains.
+
+\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(i)} - 0.01 \geq \ell^{(i)}$.
+
+Check the stability of both $\hat{\ell}^{(i)}$ and the measures of predictive accuracy relative to the previous iteration.
 
@@
 
+({\color{blue}In Section \ref{sec_requirements_diagnosis} readers can find a brief remark on the iterations requirements in the diagnosis procedure.})
+
 \subsection{Performance analysis of the diagnostics}
 
-\begin{figure*}
-\centering
-\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
-\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP length-scale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
-  \label{fig7_lscale_comparison}
-\end{figure*}
-
 In this section, we first illustrate that accurate estimates of the length-scale implies accurate approximations via HSGPs.
-Figure~\ref{fig7_lscale_comparison} left shows a comparison of the length-scale estimates
+Figure~\ref{fig6_lscale_comparison} left shows a comparison of the length-scale estimates
 obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness.
 Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.~(\ref{eq_c_vs_l_QE}).
-Figure~\ref{fig7_lscale_comparison} (right) shows the RMSE of the HSGP models with the exact GP model as the reference.
+Figure~\ref{fig6_lscale_comparison} (right) shows the RMSE of the HSGP models with the exact GP model as the reference.
 It can be seen that accurate estimates of the length-scale imply small RMSEs.
 
-Table~\ref{tab1_diagnostic} shows the iterative steps of applying the diagnostic procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison} left.
+Table~\ref{tab1_diagnostic} shows the iterative steps of applying the diagnostic procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig6_lscale_comparison}.
 It is clearly visible that by following our recommendations, an optimum solution with minimum computational requirements is achieved in these cases.
-Figure~\ref{fig8_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
+Figure~\ref{fig7_diagnostic_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set.
 Between two and four iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$.
@@
 
+\begin{figure}[t]
+\centering
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_lscale_comparison_2.png}}
+\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP length-scale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
+  \label{fig6_lscale_comparison}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_diagnostic_lscale_comparison.png}}
+\caption{Comparison of the true and estimated length-scales in every iteration for the different data sets varying the length-scale.}
+  \label{fig7_diagnostic_lscale_comparison}
+\end{figure}
+
+\subsection{Other covariance functions}
+
+Above, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
+We specifically focused on the Matérn and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
+The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
+The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig5_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
+\eqref{eq:rel_total_distance}.
+Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
+Another possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation they are actually explaining{\color{blue}, as already mentioned in Section \ref{subsec_theoretic_linearity}}.
+
+To select $c$ for many other covariance functions, users can be guided by equations~\eqref{eq_c_vs_l_QE}, \eqref{eq_c_vs_l_mat52} and \eqref{eq_c_vs_l_mat32}, as pointed in Section~\ref{sec_num_equations}.
+
 \begin{table*}
 \centering
-\setlength{\tabcolsep}{5pt}
+\setlength{\tabcolsep}{4pt}
 \begin{tabular}{ c c c c c c c c c c | c c}
@@
 \arrayrulecolor{lightgray}\hline \\[-1mm]
-\multirow{5}{*}{ 0.08} 
- & 1 & 0.50 & 1.60 & 6 & 0.17 & FALSE & 0.45 & 0.61 & -0.40 & 0.36 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_2.png}}\\
+\multirow{4}{*}{ 0.08} 
+ & 1 & 0.50 & 1.60 & 6 & 0.17 & FALSE & 0.45 & 0.61 & -0.40 & 0.36 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_2.png}}\\
  & 2 & 0.17 & 1.20 & 13 & 0.07 & FALSE & 0.36 & 0.74 & -0.18 & 0.24 & \\
@@
  & 4 & 0.06 & 1.20 & 36 & 0.08 & TRUE & 0.24 & 0.87 & 0.21 & 0.01 & \\
+ \\
  \\[1mm]
  \arrayrulecolor{lightgray}\hline \\[-1mm]
-\multirow{3}{*}{ 0.13} 
- & 1 & 0.50 & 1.60 & 6 & 0.25 & FALSE & 0.46 & 0.63 & -0.49 & 0.33 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_3.png}}\\
+\multirow{4}{*}{ 0.13} 
+ & 1 & 0.50 & 1.60 & 6 & 0.25 & FALSE & 0.46 & 0.63 & -0.49 & 0.33 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_3.png}}\\
  & 2 & 0.25 & 1.20 & 9 & 0.15 & FALSE & 0.35 & 0.78 & -0.14 & 0.15 & \\
@@
  & 4 & 0.11 & 1.20 & 20 & 0.12 & TRUE & 0.30 & 0.83 & -0.01 & 0.01 & \\
+ \\
  \\[1mm]
@@
 \multirow{3}{*}{ 0.25} 
- & 1 & 0.50 & 1.60 & 6 & 0.24 & FALSE & 0.37 & 0.40 & -0.15 & 0.21 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_4.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.24 & FALSE & 0.37 & 0.40 & -0.15 & 0.21 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 5mm, clip]{tab1_diagnostic_4.png}}\\
  & 2 & 0.24 & 1.20 & 9 & 0.26 & TRUE & 0.30 & 0.60 & -0.09 & 0.02 & \\
  & 3 & 0.25 & 1.20 & 14 & 0.26 & TRUE & 0.29 & 0.62 & -0.05 & 0.01 & \\
+ \\
  \\
@@
 \multirow{3}{*}{ 0.54} 
- & 1 & 0.50 & 1.60 & 6 & 0.38 & FALSE & 0.27 & 0.86 & 0.03 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_5.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 0.38 & FALSE & 0.27 & 0.86 & 0.03 & 0.01 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_5.png}}\\
  & 2 & 0.38 & 1.20 & 6 & 0.47 & TRUE & 0.28 & 0.86 & 0.03 & 0.05 & \\
  & 3 & 0.24 & 1.50 & 11 & 0.53 & TRUE & 0.27 & 0.86 & 0.04 & 0.01 & \\
+ \\
  \\
@@
 \multirow{3}{*}{ 0.80} 
- & 1 & 1.00 & 3.20 & 6 & 0.52 & FALSE & 0.29 & 0.92 & 0.13 & 0.03 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_6.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 0.52 & FALSE & 0.29 & 0.92 & 0.13 & 0.03 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_6.png}}\\
  & 2 & 0.52 & 1.65 & 6 & 0.98 & TRUE & 0.29 & 0.92 & 0.11 & 0.03 & \\
  & 3 & 0.50 & 3.12 & 11 & 0.77 & TRUE & 0.28 & 0.92 & 0.11 & 0.01 & \\
+ \\
  \\
@@
 \multirow{2}{*}{ 1.40} 
- & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.01 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_7.png}}\\
+ & 1 & 1.00 & 3.20 & 6 & 1.02 & TRUE & 0.28 & 0.56 & 0.14 & 0.01 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_7.png}}\\
  & 2 & 0.52 & 3.27 & 11 & 1.23 & TRUE & 0.28 & 0.57 & 0.14 & 0.00 & \\
+ \\
  \\
@@
 \multirow{2}{*}{ 2.94} 
- & 1 & 0.50 & 1.60 & 6 & 1.19 & TRUE & 0.34 & 0.45 & -0.10 & 0.06 & \multirow{5}{*}{ \includegraphics[scale=0.20, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_8.png}}\\
+ & 1 & 0.50 & 1.60 & 6 & 1.19 & TRUE & 0.34 & 0.45 & -0.10 & 0.06 & \multirow{6}{*}{ \includegraphics[scale=0.25, trim = 0mm 10mm 0mm 3mm, clip]{tab1_diagnostic_8.png}}\\
  & 2 & 0.61 & 3.81 & 11 & 2.58 & TRUE & 0.34 & 0.45 & -0.07 & 0.01 & \\
+ \\
  \\
@@
 
-\begin{figure}
-\centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_diagnostic_lscale_comparison.png}}
-\caption{Comparison of the true and estimated length-scales in every iteration for the different data sets varying the length-scale.}
-  \label{fig8_diagnostic_lscale_comparison}
-\end{figure}
-
-\subsection{Other covariance functions}
-
-Above, we thoroughly studied the relationship between the number of basis functions in the approximation and the approximation accuracy across different configurations.
-We specifically focused on the Matérn and squared exponential families of covariance functions, yet there exists other families of stationary covariance functions.
-The basis function approximation can easily be implemented for any stationary covariance function, where the spectral density function is available.
-The assess the accuracy of a basis function approximation for a kernel, where the diagnostic plots like Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) are not available, we suggest to use the relative total variational distance between the true covariance function and the approximate covariance function as given in eq.
-\eqref{eq:rel_total_distance}.
-Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
-Another possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation they are actually explaining.
-
-To select $c$ for many other covariance functions, users can be guided by equations~\eqref{eq_c_vs_l_QE}, \eqref{eq_c_vs_l_mat52} and \eqref{eq_c_vs_l_mat32}, as pointed in Section~\ref{sec_num_equations}.
-
-\subsection{The computational cost in the multi-dimensional setting}
-
-The HSGP model is computationally superior in $1$D and $2$D even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$, i.e., $n$ smaller than some value around 300) that exact GPs are already reasonably fast themselves.
+\subsection{The computational cost in the multi-dimensional setting}\label{sec_cost_multi}
+
+The HSGP model is computationally superior {\color{blue} to the exact GP} in $1$D and $2$D even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$, i.e., $n$ smaller than some value around 300) that exact GPs are already reasonably fast themselves.
 However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
@@
 As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum sufficient value, reducing computational time drastically, and still getting an accurate approximation.
-For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\ell \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \ell \gtrsim 0.1$; and $m \gtrsim 20$ for $\ell \lesssim 0.1$), which, from results presented in Figure~\ref{fig15_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3$D functions and smooth $4$D functions.
+For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\ell \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \ell \gtrsim 0.1$; and $m \gtrsim 20$ for $\ell \lesssim 0.1$), which, from results presented in Figure~\ref{fig14_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3$D functions and smooth $4$D functions.
 
@@
 In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions per dimension, or even for $D=2$ and more than 40 basis functions per dimension (see Figures~B.3 right and C.3 in the online supplement).
+
+\subsubsection{Remark on the iterations requirements in the diagnosis procedure}\label{sec_requirements_diagnosis}
+
+{\color{blue}The limitation of the approach that can cause too long computation time does not come from the iterative diagnosis procedure, but from the computational cost of the HMC sampling method to fit the HSGP model at each diagnosis iteration.
+As said extensively throughout the paper, the computational cost of a single model fit depends ultimately on the non-linearity (wigglyness) of the function to be estimated and the number of dimensions, which affects the HMC sampling of a model fit, but not the number of iterations required to complete the diagnosis procedure.
+Then, the user has to care about the computational cost of fitting a model, but not about the number of iterations needed to complete the diagnosis procedure, which are few in any case.
+Few iterations in the diagnosis procedure are sufficient to obtain an accurate model approximation even in multi(low)-dimensional cases (that need a large number of basis functions), as can be seen in Table 1 and Tables 2-4 in the following sections.}
 
@@
 
-Figure~\ref{fig9_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
+Figure~\ref{fig8_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
 In order to do model comparison a spline-based model with 60 knots is also fitted and plotted jointly with the GP and HSGP models.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig9_Posteriors_exI.png}
+\includegraphics[width=0.485\textwidth]{fig8_Posteriors_exI.png}
 \caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
-  \label{fig9_Posteriors_exI}
+  \label{fig8_Posteriors_exI}
 \end{figure}
 
-Figure~\ref{fig10_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+Figure~\ref{fig9_rmse_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
 Both the HSGP and splines models show roughly similar performance at sample data.
 The HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation.
-However, the spline model does not extrapolate data properly (Figure~\ref{fig9_Posteriors_exI}).
+However, the spline model does not extrapolate data properly (Figure~\ref{fig8_Posteriors_exI}).
 The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_rmse_time_exI.png}}
+\subfigure{\includegraphics[width=0.485\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig9_rmse_time_exI.png}}
 \caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
@@
 The standard deviation of the computational time is plotted as dashed lines.}
-  \label{fig10_time_exI}
+  \label{fig9_rmse_time_exI}
 \end{figure}
 
-The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/}} \linebreak {\small \url{ basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
+The Stan model code for the exact GP model, the HSGP model, the spline model, and R-code to reproduce this case study can be found online at {\small \url{ https://github.com/gabriuma/}} \linebreak {\small \url{ basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
 
@@
 
-Figure~\ref{fig11_posteriors_birthday} shows the posterior means of the long-term trend $f_1$ and yearly pattern $f_2$ for the whole period, jointly with the observed data.
-Figure \ref{fig12_posteriors_oneyear_birthday} shows the model for one year (1972) only.
+Figure~\ref{fig10_posteriors_birthday} shows the posterior means of the long-term trend $f_1$ and yearly pattern $f_2$ for the whole period, jointly with the observed data.
+Figure \ref{fig11_posteriors_oneyear_birthday} shows the model for one year (1972) only.
 In this figure, the special days effects $f_4$ in the year can be clearly represented.
-The posterior means of the the function $\mu$ and the components $f_1$ (long-term trend) and $f_2$ (year pattern) are also plotted in this Figure \ref{fig12_posteriors_oneyear_birthday}.
-Figure \ref{fig13_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3$ can be clearly represented.
-The mean of the the function $\mu$ and components $f_1$ (long-term trend), $f_2$ (year pattern) and $f_4$ (special-days effects) are also plotted in this Figure \ref{fig13_posteriors_onemonth_birthday}.
-
-The Stan model code for the approximate GP model of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/}} \linebreak {\small \url{master/Paper/Case-study_Birthday-data}}\,.
+The posterior means of the the function $\mu$ and the components $f_1$ (long-term trend) and $f_2$ (year pattern) are also plotted in this Figure \ref{fig11_posteriors_oneyear_birthday}.
+Figure \ref{fig12_posteriors_onemonth_birthday} show the process in the month of January of 1972 only, where the week pattern $f_3$ can be clearly represented.
+The mean of the the function $\mu$ and components $f_1$ (long-term trend), $f_2$ (year pattern) and $f_4$ (special-days effects) are also plotted in this Figure \ref{fig12_posteriors_onemonth_birthday}.
+
+The Stan model code for the HSGP model and R-code to reproduce this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Birthday-data}}\,.
 
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig11_posteriors_birthday.png}}
+\subfigure{\includegraphics[width=0.49\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig10_posteriors_birthday.png}}
 \caption{Posterior means of the long-term trend $f_1$ and year effects pattern ($f_2$) for the whole series. }
-  \label{fig11_posteriors_birthday}
+  \label{fig10_posteriors_birthday}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig12_posteriors_oneyear_birthday.png}}
+\subfigure{\includegraphics[width=0.49\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig11_posteriors_oneyear_birthday.png}}
 \caption{Posterior means of the function $\mu$ for the year 1972 of the series.
@@
 }
-  \label{fig12_posteriors_oneyear_birthday}
+  \label{fig11_posteriors_oneyear_birthday}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig13_posteriors_onemonth_birthday.png}}
+\subfigure{\includegraphics[width=0.49\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig12_posteriors_onemonth_birthday.png}}
 \caption{Posterior means of the function $\mu$ for the month of January of 1972.
@@
 }
-  \label{fig13_posteriors_onemonth_birthday}
+  \label{fig12_posteriors_onemonth_birthday}
 \end{figure}
@@
 \centering
-\subfigure{\includegraphics[scale=0.315, trim = 1mm 18mm 39.2mm 20mm, clip]{fig14_errorGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig14_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig14_errorSP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.30, trim = 151mm 14mm 2mm 20mm, clip]{fig14_error_colorbar_exII.pdf}}
-\hspace{1pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{fig14_RMSE_exII.pdf}}
-\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{fig14_legend_exII.pdf}}}
+\subfigure{\includegraphics[scale=0.315, trim = 1mm 18mm 39.2mm 20mm, clip]{fig13_errorGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig13_errorHSGP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig13_errorSP_exII.pdf}}
+\subfigure{\includegraphics[scale=0.30, trim = 151mm 14mm 2mm 20mm, clip]{fig13_error_colorbar_exII.pdf}}
+\hspace{1pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{fig13_RMSE_exII.pdf}}
+\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{fig13_legend_exII.pdf}}}
 \caption{Error between the $2$D data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models.
@@
 The right side plot shows the root mean square error (RMSE) of the different methods, and plotted as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots (the same number of basis functions $m$ and knots per dimension are used, resulting in $m^D$ total number of multivariate basis functions).}
-  \label{fig14_errors_exII}
+  \label{fig13_errors_exII}
 \end{figure*}
@@
 
-Figure \ref{fig14_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2$D data set.
+Figure \ref{fig13_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2$D data set.
 For the $3$D data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3$D input space.
@@
 For the spline-based model, $40$ knots in each dimension were used.
-Figure \ref{fig14_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
+Figure \ref{fig13_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
 It can be seen that the posterior mean for HSGP and exact GP accurately represents the true data-generating function.
@@
 
+The Stan model codes for the exact GP model, the HSGP model, and the spline-based model, and R-code to reproduce this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D&3D-Simulated-data}}\,.
+
 \subsubsection{Computation requirements in 2D, 3D, and 4D input spaces}\label{sec_computation_nD}
 
-Figure~\ref{fig15_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Figure~\ref{fig14_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
 Both the time of computation and $m^*$ are plotted in the logarithmic scale.
 
-Looking back at Figure~\ref{fig6_relationships}, or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32}, and \eqref{eq_m_l_mat52}, and assuming squared exponential covariance function, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted using only $10--12$ basis functions.
-For lengthscales between $0.1$ and $0.3$, $10--22$ basis functions are sufficient.
-
-For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig15_time_nD_log}).
+Looking back at Figure~\ref{fig5_relationships}, or equations~\eqref{eq_m_l_QE}, \eqref{eq_m_l_mat32}, and \eqref{eq_m_l_mat52}, and assuming squared exponential covariance function, any univariate function (or single dimensions of a multivariate function) with true lengthscales bigger than $0.3$ can be accurately fitted using only $10-12$ basis functions.
+For lengthscales between $0.1$ and $0.3$, $10-22$ basis functions are sufficient.
+
+For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig14_time_nD_log}).
 For a very wiggly $3$D function (say, with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n \lesssim 1000$), the method is still significantly faster than the regular GP.
@@
 Finally, for $D>5$ the method starts to be impractical even for smooth univariate functions.
-However, in these cases, the method may still be used for lower dimensional componenets in an additive modeling scheme.
+However, in these cases, the method may still be used for lower dimensional components in an additive modeling scheme.
+
+The Stan model codes for the exact GP model and the HSGP model for these 2D, 3D, and 4D cases can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Stan_code_2D_3D_4D}}\,.
 
@@
 \centering
-\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig15_time_nD_log.png}
+\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig14_time_nD_log.png}
 \caption{Computational time in seconds per iteration (iteration of the HMC sampling method) for different datasets with different dimensionalties and data points, and plotted as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$.}
-  \label{fig15_time_nD_log}
+  \label{fig14_time_nD_log}
 \end{figure*}
-
-The Stan model codes for the exact GP, the HSGP and the spline-based models of this case study can be found online at {\small \url{https://github.com/gabriuma/}} \linebreak {\small \url{basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D-Simulated-data}}\,.
 
@@
 \centering
-\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{fig16_posteriors_leukemia.pdf}
+\includegraphics[scale=0.70, trim = 0mm 5mm 0mm 0mm, clip]{fig15_posteriors_leukemia.pdf}
 \caption{Expected lifetime conditional comparison for each predictor with other predictors fixed to their mean values.
 The thick line in each graph is the posterior mean estimated using a HSGP model, and the thin lines represent pointwise 95\% credible intervals.}
-  \label{fig16_posteriors_leukemia}
+  \label{fig15_posteriors_leukemia}
 \end{figure*}
@@
 
-Figure \ref{fig16_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
+Figure \ref{fig15_posteriors_leukemia} shows estimated conditional functions of each predictor with all others fixed to their mean values.
 These posterior estimates correspond to the HSGP model with $m=10$ basis functions and $c=3$ boundary factor.
 There are clear  non-linear patterns and the right bottom subplot also shows that the conditional function associated with WBC has an interaction with TDI.
-Figure~\ref{fig17_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
-As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig17_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig6_relationships}.
-Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig17_elpd_leukemia}-right).
+Figure~\ref{fig16_elpd_leukemia} shows the expected log predictive density \citep[ELPD; ][]{vehtari_2012,Vehtari+Gelman+Gabry:2017_practical} and time of computation as function of the number of univariate basis functions $m$ ($m^{\ast}=m^D$ in eq.~(\ref{eq_approxf_multi})) and boundary factor $c$.
+As the functions are smooth, a few number of basis functions and a large boundary factor are required to obtain a good approximation (Figure \ref{fig16_elpd_leukemia}-left); Small boundary factors are not appropriate for models for large length-scales, as can be seen in Figure \ref{fig5_relationships}.
+Increasing the boundary factor also significantly increases the time of computation (Figure \ref{fig16_elpd_leukemia}-right).
 With a moderate number of univariate basis functions ($m=15$), the HSGP model becomes slower than the exact GP model, in this specific application with $3$ input variables, as the total number of multivariate basis functions becomes $15^3 = 3375$ and is therefore quite high.
@@
 \centering
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig17_elpd_leukemia.pdf}} 
-\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig17_time_leukemia.pdf}}
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig16_elpd_leukemia.pdf}} 
+\subfigure{\includegraphics[scale=0.38, trim = 0mm 8mm 0mm 0mm, clip]{fig16_time_leukemia.pdf}}
 \caption{Expected log predictive density (ELPD; left) and time of computation in seconds per iteration (iteration of the HMC sampling method; right) as a function of the number of basis functions $m$ and boundary factor $c$.}
-  \label{fig17_elpd_leukemia}
+  \label{fig16_elpd_leukemia}
 \end{figure*}
 
-The Stan model code for the exact GP and the approximate GP models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
+The Stan model code for the exact GP and the HSGP models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
 %
@@
 The main drawback of the approach is that its computational complexity scales exponentially with the number of input dimensions.
-Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figure~\ref{fig6_relationships}, is essential to avoid a excessive computational time especially in multivariate input spaces.
+Hence, choosing optimal values for the number of basis functions and the boundary factor, using the recommendations and diagnostics provided in Figure~\ref{fig5_relationships}, is essential to avoid a excessive computational time especially in multivariate input spaces.
 However, in practice, input dimensionalities larger than three start to be computationally demanding even for moderately wiggly functions and few basis functions per input dimension.

### 2022-07-05 Paper/Manuscript_Stats&Comp/reply_letter_2nd_STCO.tex
response letter to 2nd revision

\documentclass[11pt]{report}
\usepackage{blindtext}
\usepackage{titlesec}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage{microtype}
\usepackage{lscape}

\usepackage[textwidth=17cm, textheight=21.2cm]{geometry}

\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}

\usepackage{epstopdf}

\usepackage[sectionbib,numbers,sort&compress]{natbib}

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}
\usepackage[dvipsnames]{xcolor}
%\graphicspath{{./images/}}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{bm}

\usepackage{appendix}

\bibpunct[, ]{[}{]}{,}{}{,}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}

\setlength{\parskip}{0.5em}
\renewcommand{\baselinestretch}{1.1}

\renewcommand{\bibname}{References}

\usepackage{dashrule}

\usepackage{bibentry}
\bibliographystyle{apalike}

\makeatletter 
\renewcommand\BR@b@bibitem[2][]{\BR@bibitem[#1]{#2}\BR@c@bibitem{#2}}           
\makeatother

\nobibliography*

\begin{document}

We would like to thank the reviewers for their thorough comments and valuable feedback, which has helped us a lot in improving the paper. First, the updates to the manuscript that come from the authors' own initiative are outlined. Then, all reviewers' comments are addressed in a point-by-point manner, with separate replies to each reviewer. 
Throughout, comments of the reviewers are hightlighted in bold.

\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

\subsection*{Authors' own updates}

\begin{itemize}

\item We have updated the appearance of Figures 1, 2, 3, 4 and 5 (old numbering) in order to make them more compact and thus the overall appearance of the paper more pleasing. The new numbering corresponding to those figures is Fig. 1, 2, 3 and 4.

\item We have revised the entire manuscript making some minor edits, such as rewriting a few sentences and adding minor little additional information.

\end{itemize}

\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

\subsection*{Reviewer \#1}

We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.

%\subsubsection*{R\#1 General comment and feedback for the authors}
%
%\textbf{Overall, the paper has clearly improved from the original submission and the revision addresses all of my major revision requests adequately. I still think the paper is a bit borderline in terms of the technical contribution itself, but the revised version reads better, is more clear on the limitations, clarifies a few open issues, and the experimental section has been improved. In particular, the experimental section is now very extensive and provides rich information about the practical use and limitations of the approach. The paper certainly has potential for impact as a practical guide and tools for using HSGP and now is clearly more suitable for publication. I do not have any major comments for the paper and would be happy to see it accepted in current form.}
%
%\textbf{Quick feedback on (some of) the author responses:}
%
%\textbf{1. Apologies for skipping the statement on brms integration. I agree it is valuable as it makes the contribution more accessible for a broad audience that might not be interested in the technical details.}
%
%\textbf{3. This resolves by remark well. Good addition that will help practitioners and can be implemented directly for the software as default criteria.}
%
%\textbf{4. I appreciate the effort of extending the treatment. I still have a bit of an issue of how this needs to be done separately for each possible kernel and the implications that has on general probabilistic programming environments where part of the value of the tool is in providing extreme flexibility, but I do agree that for practical uses e.g. with brms the set of kernels you now provided is useful and likely to cover most user needs.}
%
%\textbf{5. Fair enough; I certainly agree some (if not many) practitioners would fail to do this properly.}
%
%\textbf{6-8. I think the experimental section indeed reads better now and is overall very extensive and detailed.}
%
%\textbf{9. By mostly negative results I meant that the value of the approach is not obvious and hence a practitioner would get the impression that investigating the value of HSGP for higher-dimensional cases would not be worthwhile since the gain would likely be small compared to the effort. Apologies for the slightly confusing statement, since I agree you do show some advantage on the Leukemia data. The new experiments on simulated data make this aspect more clear.}
%
%\textbf{Finally, a quick comment related to R\#2 Comment \#3: I agree here with the authors that empirical comparison against alternative inference techniques is not necessary and would not really make the paper stronger either.}

We would also like to thank you for the additional feedback on some of our responses.

\subsubsection*{R\#1 Minor comment \#1}

\textbf{The new Figure 6 is, for some reason, very blurry. The same holds for the small images in Table 1 (btw, I like this presentation style where the numbers are put next to the figures in a clear manner) and some other figures as well. Maybe check all of them for resolution/vectorization.}

Figure 6 may look a bit fuzzy, but it is actually a visual effect because the curves were constructed for discrete values of the number of basis functions $m$ (Y-axis), which makes them look like stair-shaped curves when plotted, and its appearance can be confused with a blurred image.

We have increased the size of the figures in Table 1.

The authors think that the rest of the figures look good in their current form.

\subsubsection*{R\#1 Minor comment \#2}

\textbf{The algorithms in Section 4.5.1 do not have clear stopping conditions for failure cases.  For problematic cases (e.g. clearly too high dimensionality) the result would be a loop of  extremely many iterations, right? In Section 4.6 you explain how a few iterations were enough in your example cases, but maybe you could already in 4.5.1 give some suggestion on when the user should give up, with something like "in case you do not find a satisfactory solution in N iterations, you might want to reconsider using HSGP or perhaps need to change X or Y".}

The limitation of the approach that can cause too long computation time does not come from the iterative diagnosis procedure, but from the computational cost of the HMC sampling method to fit the HSGP model at each diagnosis iteration, that is, the more input dimensions ($D$) the more multivariate basis functions ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit, but not the number of iterations required to complete the diagnosis procedure.

In fact, the diagnosis procedure converges quickly (with few iterations) to an accurate approximation even in multi(low)-dimensional cases (that need a large number of basis functions), as can be seen in Tables 1-5. It is true that more input dimensions may lead to the need for a few more iterations in the diagnosis procedure due to the fact that the diagnostic check is performed individually on each dimension and an accurate approximation of the model will be reached when the diagnostic check passes for all dimensions together. But this fact should only lead to the need for only a couple of additional iterations, since the approach is only suitable for dimensionalities $<= 5$ (Remember that, as stated throughout the paper, the HSGP approach is only suitable for $D<=3$, with (very) wiggly functions, and for $D<=5$, with (very) smooth functions). For larger dimensionalities, the approach can still be used in an additive modeling framework.

As an example, Table 5 summarizes the iterative diagnosis for a 3D function formed by very/moderate wiggly functions in each of its dimensions. For the 4D and 5D cases we do not expect a significant increase in the number of diagnosis iterations, even more so since for the 4D and 5D cases the HSGP model can only fit smooth (or very smooth in 5D) functions.

Therefore, the user should be concerned about the computational cost of fitting a model in a single iteration of the diagnosis procedure, but not about the number of iterations required to complete the diagnosis procedure, which should be few in any case. The computational cost of fitting a single model ultimately depends on the wigglyness of the function and the number of dimensions. We have provided extensive information throughout the paper on the computational requirements for model fitting as a function of wigglyness and number of dimensions, information that is also summarized in Figure 15.

Although that has been said extensively throughout the paper, we have added a remark in Section 4.8.1 and a note at the end of Section 4.5.1, as suggested by the reviewer, which we also think can help to make it more clear.

\subsection*{Reviewer \#2}

We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.

\subsubsection*{R\#2 Comment \#1}

\textbf{the suggestion that, as a purely empirical work, more alternative methods should be compared with was rejected by the authors because it would "would blow up the already extensive scope of the paper". however, much of the paper includes extensive material from earlier work which need not be duplicated. also, the author comment that "focus on the comparison between the Hilbert space basis function approximation (HSGP method) against an exact GP inference", but this is misleading as they actually include a specific baseline of spline regression, which seems like a rather arbitrary and weak one (the authors had a chance to change my mind here, but did not attempt to do so).}

We included the solution of a spline model (a model related to the GP/HSGP model) only in case study I, with no other intention than to show, beside the GP and HSGP solutions, the solution of a spline model, which has similar (of the same order) computational cost and limitations in multidimensional cases. It can be seen that a basic formulation of a spline model causes significantly worse extrapolation accuracy.

\subsubsection*{R\#2 Comment \#2}

\textbf{a major item of new content is "A diagnostic procedure based on iterative steps to fit and diagnose the approximation has been proposed (Section 4.5.1 and Section 4.6)". however, the very first step A1 in section 4.5.1 looks odd. the choice of $\ell_1$ surely depends on the scale of the input data. if the data scale is large, won't the method fail ? it could be that i misunderstood this, e.g. that there is some preprocessing step required, but given that there is a confusing collision of notation with the $\ell_1$ of line 50, i expect other readers to interpret it as i did (or worse, to assume they are referring to the $\ell_1$ of line 50).}

As said in the paper, we work with a normalized lengthscale, i.e., a lengthscale $\ell$ relative to the half-range ($S$) of the input domain ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain. We have added an additional paragraph to make it more clear in the second-to-last paragraph in Section 4.3.1.

We agree with the reviewer that the $\ell_i$ notation in Section 4.5.1, where $i$ is the subscript for the diagnosis iteration, may conflict with the $\ell_i$ notation used in the article where $i$ is the subscript for dimensions. Therefore, we have changed the notation in Section 4.5.1 from $\ell_i$ to $\ell^{(i)}$ to differentiate it from the notation in the rest of the article where subscripts refer to dimensions.

%\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}

%\nobibliography{references}

\end{document}

### 2022-07-12 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
edit paper and response letter 2

@@
 Figure \ref{fig5_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) with {\color{blue}normalized} length-scale $\frac{\ell}{S}$ (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+More precisely, for a given GP model (with a particular covariance function) with {\color{blue}normalized} length-scale $\frac{\ell}{S}$ (from now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
@@
 
-\item[A1.] Set the iteration index to $i=1$.
-Make an initial guess on the length-scale $\ell^{(i)}$.
-If there is no useful information available, we recommend to start with a large length-scale, about $\ell^{(i)} \in [0.5, 1]$.
-({\color{blue}Remember that $\ell$ refers to the normalized length-scale $\frac{\ell}{S}$.})
-
-\item[A2.] Obtain the minimum valid boundary factor $c^{(i)}$ determined by $\ell = \ell^{(i)}$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[A3.] Obtain the mimimum valid number of basis functions $m^{(i)}$ determined by $\ell = \ell^{(i)}$ and $c = c^{(i)}$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[A4.] Fit an HSGP model using $m^{(i)}$ and $c^{(i)}$ and ensure convergence of the MCMC chains.
-
-\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(i)} - 0.01 \geq \ell^{(i)}$.
-
-If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell^{(i+1)} = \hat{\ell}^{(i)}$, increase the iteration index $i = i + 1$, and go back to A2.
+\item[A1.] Set the iteration index to $k=1$.
+Make an initial guess on the length-scale $\ell^{(k)}$.
+If there is no useful information available, we recommend to start with a large length-scale, about $\ell^{(k)} \in [0.5, 1]$.
+({\color{blue}Remember that $\ell$ here refers to the normalized length-scale $\frac{\ell}{S}$.})
+
+\item[A2.] Obtain the minimum valid boundary factor $c^{(k)}$ determined by $\ell = \ell^{(k)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m^{(k)}$ determined by $\ell = \ell^{(k)}$ and $c = c^{(k)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[A4.] Fit an HSGP model using $m^{(k)}$ and $c^{(k)}$ and ensure convergence of the MCMC chains.
+
+\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(k)} - 0.01 \geq \ell^{(k)}$.
+
+If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell^{(k+1)} = \hat{\ell}^{(k)}$, increase the iteration index $k = k + 1$, and go back to A2.
 
@@
 
-\item[B2.] Set $m^{(i+1)} = m^{(i)} + 5$ and increase the iteration index $i = i + 1$.
-
-\item[B3.] Obtain the minimum valid boundary factor $c^{(i)}$ determined by $\ell = \hat{\ell}^{(i-1)}$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[B4.] Obtain the minimum valid length-scale $\ell^{(i)}$ that can be accurately estimated by $m = m^{(i)}$ and $c = c^{(i)}$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[B5.] Fit an HSGP model using $m^{(i)}$ and $c^{(i)}$ and ensure convergence of the MCMC chains.
-
-\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(i)} - 0.01 \geq \ell^{(i)}$.
-
-Check the stability of both $\hat{\ell}^{(i)}$ and the measures of predictive accuracy relative to the previous iteration.
+\item[B2.] Set $m^{(k+1)} = m^{(k)} + 5$ and increase the iteration index $k = k + 1$.
+
+\item[B3.] Obtain the minimum valid boundary factor $c^{(k)}$ determined by $\ell = \hat{\ell}^{(k-1)}$ for the given kernel as per Section \ref{sec_num_equations}.
+
+\item[B4.] Obtain the minimum valid length-scale $\ell^{(k)}$ that can be accurately estimated by $m = m^{(k)}$ and $c = c^{(k)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+
+\item[B5.] Fit an HSGP model using $m^{(k)}$ and $c^{(k)}$ and ensure convergence of the MCMC chains.
+
+\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(k)} - 0.01 \geq \ell^{(k)}$.
+
+Check the stability of both $\hat{\ell}^{(k)}$ and the measures of predictive accuracy relative to the previous iteration.
 
@@
 
-({\color{blue}In Section \ref{sec_requirements_diagnosis} readers can find a brief remark on the iterations requirements in the diagnosis procedure.})
+{\color{blue} We have seen this procedure do converge quickly after only few iterations in all the investigated cases (see Section~\ref{sec_cases}).
+That said, we cannot rule out that there may be cases where convergence is much slower (in terms of number of required diagnostic iterations) to a degree where HSGPs become impractical and an early stopping of the procedure would be advisable.
+Importantly though, this scenario should not be confused with the scaling of computational cost due to higher dimensional input spaces, an issue discussed in detail in Section~\ref{sec_cost_multi}.
+In a nutshell, according to our experiments, increasing the dimensionality of the input space does not increase the number of required diagnostic iterations to a relevant degree but only the computational cost per iteration.}
 
@@
 
-\subsubsection{Remark on the iterations requirements in the diagnosis procedure}\label{sec_requirements_diagnosis}
-
-{\color{blue}The limitation of the approach that can cause too long computation time does not come from the iterative diagnosis procedure, but from the computational cost of the HMC sampling method to fit the HSGP model at each diagnosis iteration.
-As said extensively throughout the paper, the computational cost of a single model fit depends ultimately on the non-linearity (wigglyness) of the function to be estimated and the number of dimensions, which affects the HMC sampling of a model fit, but not the number of iterations required to complete the diagnosis procedure.
-Then, the user has to care about the computational cost of fitting a model, but not about the number of iterations needed to complete the diagnosis procedure, which are few in any case.
-Few iterations in the diagnosis procedure are sufficient to obtain an accurate model approximation even in multi(low)-dimensional cases (that need a large number of basis functions), as can be seen in Table 1 and Tables 2-4 in the following sections.}
-
 \section{Case studies}\label{sec_cases}

### 2022-07-12 Paper/Manuscript_Stats&Comp/reply_letter_2nd_STCO.tex
edit paper and response letter 2

@@
 
-\item We have updated the appearance of Figures 1, 2, 3, 4 and 5 (old numbering) in order to make them more compact and thus the overall appearance of the paper more pleasing.
-The new numbering corresponding to those figures is Fig.
-1, 2, 3 and 4.
+\item We have updated the appearance of Figures 1, 2, 3, 4, and 5 (old numbering) in order to make them more compact and thus the overall appearance of the paper more pleasing.
+The new numbering corresponding to those figures is now Fig.
+1, 2, 3, and 4.
 
@@
 
-We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
+Thank you again for your valuable comments and suggestions.
+
+%We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
+We would also like to thank you for the additional feedback on some of our responses.
 
@@
 
-We would also like to thank you for the additional feedback on some of our responses.
-
 \subsubsection*{R\#1 Minor comment \#1}
@@
 Figure 6 may look a bit fuzzy, but it is actually a visual effect because the curves were constructed for discrete values of the number of basis functions $m$ (Y-axis), which makes them look like stair-shaped curves when plotted, and its appearance can be confused with a blurred image.
+We think the resolution of the figure itself is sufficient.
+We have increased the size of the figures in Table 1 and have checked and updated the other figures as we saw fit.
 
-We have increased the size of the figures in Table 1.
-
-The authors think that the rest of the figures look good in their current form.
+%The authors think that the rest of the figures look good in their current form.
 
@@
 
-The limitation of the approach that can cause too long computation time does not come from the iterative diagnosis procedure, but from the computational cost of the HMC sampling method to fit the HSGP model at each diagnosis iteration, that is, the more input dimensions ($D$) the more multivariate basis functions ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit, but not the number of iterations required to complete the diagnosis procedure.
+Based on the evidence that we have at this point, we can unfortunately not give evidence-based suggestions as to when to stop pre-convergence and try something else because we have seen only very few iterations to be required in all the investigated cases.
+With regard to high-dimensional cases, the main problem would not be more iterations (as far as we have seen) but the computational cost of each iteration.
+More concretely, as we discuss in detail in the paper, the more input dimensions ($D$) we have, the more multivariate basis functions we get ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit (per iteration), but not (in a relevant manner) the number of iterations required to complete the diagnostic procedure.
+We have now clarified these points at the end of Section 4.5.1.
 
-In fact, the diagnosis procedure converges quickly (with few iterations) to an accurate approximation even in multi(low)-dimensional cases (that need a large number of basis functions), as can be seen in Tables 1-5.
-It is true that more input dimensions may lead to the need for a few more iterations in the diagnosis procedure due to the fact that the diagnostic check is performed individually on each dimension and an accurate approximation of the model will be reached when the diagnostic check passes for all dimensions together.
+%The limitation of the approach that can cause too long computation time does not come from the iterative diagnostic procedure, but from the computational cost of the HMC sampling method to fit the HSGP model at each diagnostic iteration, that is, the more input dimensions ($D$) the more multivariate basis functions ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit, but not the number of iterations required to complete the diagnostic procedure.
+
+%In fact, the diagnostic procedure converges quickly (with few iterations) to an accurate approximation even in multi(low)-dimensional cases (that need a large number of basis functions), as can be seen in Tables 1-5.
+It is true that more input dimensions may lead to the need for a few more iterations in the diagnostic procedure due to the fact that the diagnostic check is performed individually on each dimension and an accurate approximation of the model will be reached when the diagnostic check passes for all dimensions together.
 But this fact should only lead to the need for only a couple of additional iterations, since the approach is only suitable for dimensionalities $<= 5$ (Remember that, as stated throughout the paper, the HSGP approach is only suitable for $D<=3$, with (very) wiggly functions, and for $D<=5$, with (very) smooth functions).
@@
 
-As an example, Table 5 summarizes the iterative diagnosis for a 3D function formed by very/moderate wiggly functions in each of its dimensions.
-For the 4D and 5D cases we do not expect a significant increase in the number of diagnosis iterations, even more so since for the 4D and 5D cases the HSGP model can only fit smooth (or very smooth in 5D) functions.
+%As an example, Table 5 summarizes the iterative diagnostic for a 3D function formed by very/moderate wiggly functions in each of its dimensions.
+For the 4D and 5D cases we do not expect a significant increase in the number of diagnostic iterations, even more so since for the 4D and 5D cases the HSGP model can only fit smooth (or very smooth in 5D) functions.
 
-Therefore, the user should be concerned about the computational cost of fitting a model in a single iteration of the diagnosis procedure, but not about the number of iterations required to complete the diagnosis procedure, which should be few in any case.
+%Therefore, the user should be concerned about the computational cost of fitting a model in a single iteration of the diagnostic procedure, but not about the number of iterations required to complete the diagnostic procedure, which should be few in any case.
 The computational cost of fitting a single model ultimately depends on the wigglyness of the function and the number of dimensions.
@@
 
-Although that has been said extensively throughout the paper, we have added a remark in Section 4.8.1 and a note at the end of Section 4.5.1, as suggested by the reviewer, which we also think can help to make it more clear.
+%Although that has been said extensively throughout the paper, we have added a remark in Section 4.8.1 and a note at the end of Section 4.5.1, as suggested by the reviewer, which we also think can help to make it more clear.
 
@@
 
-We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
+%We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
+
+Thank you again for your valuable comments and suggestions.
 
@@
 
-We included the solution of a spline model (a model related to the GP/HSGP model) only in case study I, with no other intention than to show, beside the GP and HSGP solutions, the solution of a spline model, which has similar (of the same order) computational cost and limitations in multidimensional cases.
+We apologize for not making a stronger case in terms of performed (and not performed) comparisons earlier.
+We are aware that this paper could be extended in various different valuable directions, as is probably the case with most papers.
+One of such directions is of course to add more comparisons with other methods.
+
+In this paper, we have focussed on the empirical accuracy of the HSGP approximation of exact GPs as you are aware.
+Including the spline in one of the case studies was to show that excessively applied methods that are similar in some ways (here: reduced-rank thin-plate splines, a method whose corresponding papers have been cited thousands of times already), does not perform better out of the box than our HSGPs.
+But we understand and agree that including this particular spline here in this particular case study is not a strong case in terms of comparison with other methods.
+Accordingly, we decided to remove it to not invoke a false impression of our focus or claims.
+
+We are not sure whether we can convince you of the sensibility and sufficiency of the focus of our paper that we have chosen.
+But what we can say is that based on the informal feedback from the community we have recieved so far, the other reviewers' and also the editor's comments, as well as over 20 citation of the corresponding preprint already, a lot of people see value in our work and its existing focus.
+There is certainly a lot of value in comparing HSGPs with competing methods other than exact GPs, but we do not see a way to make this part of the current paper, not only in terms of paper space, but also in terms of time, personnel, and monetary resources we currently have available to work on this topic.
+
+%We included the solution of a spline model (a model related to the GP/HSGP model) only in case study I, with no other intention than to show, beside the GP and HSGP solutions, the solution of a spline model, which has similar (of the same order) computational cost and limitations in multidimensional cases.
 It can be seen that a basic formulation of a spline model causes significantly worse extrapolation accuracy.
@@
 
-As said in the paper, we work with a normalized lengthscale, i.e., a lengthscale $\ell$ relative to the half-range ($S$) of the input domain ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
-We have added an additional paragraph to make it more clear in the second-to-last paragraph in Section 4.3.1.
+We apologize for creating confusion with our notation here.
+We work with a normalized lengthscale, i.e., a lengthscale $\ell$ relative to the half-range ($S$) of the input domain ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
+We have now clarified the notation by using different symbols for lengthscale and normalized lengthscale throughout the paper.
 
-We agree with the reviewer that the $\ell_i$ notation in Section 4.5.1, where $i$ is the subscript for the diagnosis iteration, may conflict with the $\ell_i$ notation used in the article where $i$ is the subscript for dimensions.
-Therefore, we have changed the notation in Section 4.5.1 from $\ell_i$ to $\ell^{(i)}$ to differentiate it from the notation in the rest of the article where subscripts refer to dimensions.
+%added an additional paragraph to make it more clear in the second-to-last paragraph in Section 4.3.1.
+
+Thank you for also your comment about the $\ell_i$ notation in Section 4.5.1, where $i$ is the subscript for the diagnostic iteration, conflicting with the $\ell_i$ notation used in the article where $i$ is the subscript for dimensions.
+Therefore, we have changed the notation in Section 4.5.1 from $\ell_i$ to $\ell^{(k)}$ to clearly differentiate it from the notation in the rest of the article where subscripts refer to dimensions.
 

### 2022-07-22 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
More edits to the paper and response letter 2

@@
 \usepackage{nicefrac}
+%\usepackage{ulem}
 
@@
 The accuracy and speed of the HSGP model depends on several interrelated factors, most notably on the number of basis functions and on the boundary condition of the Laplace eigenfunctions.
-Furthermore, appropriate values for these factors will depend on the degree of non-linearity ({\color{blue}wigglyness/smoothness}) of the function to be estimated, which is in turn characterized by the length-scale of the covariance function.
+Furthermore, appropriate values for these factors will depend on the degree of non-linearity (wigglyness/smoothness) of the function to be estimated, which is in turn characterized by the length-scale of the covariance function.
 In this section, we analyze the effects of the number of basis functions and the boundary condition on the approximation accuracy.
@@
 In this work we investigate these relationships for the squared exponential and the Mat{\'e}rn ($\nu=3/2$ and $\nu=5/2$) covariance functions in the present section, and for the periodic squared exponential covariance function in Appendix~\ref{sec_periodic}.
-For other kernels, the relationships will be slightly different depending on the {\color{blue}smoothness or wigglyness of the non-linear effects generated from the covariance function}.
+For other kernels, the relationships will be slightly different depending on the smoothness or wigglyness of the non-linear effects generated from the covariance function.
 
@@
 
-In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ contains all the inputs points $x_i$.
-Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.
-Let $S = \max_{i}|x_i|$, then it follows that $x_i \in \left[-S, S\right]$ for all $i$.
-We now define $L$ as
-\begin{equation}\label{eq_boundary}
+\setstcolor{blue}
+
+In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ contains all the inputs points $x_i$
+\st{. Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.}
+{\color{blue}, and the set of input points $\{x_i\}_{i=1}^n$ has some kind of centering around zero.}
+Let 
+%
+\begin{align}\label{eq_boundary_S}
+S=\max_{i}|x_i|, 
+\end{align}
+%
+then it follows that $x_i \in \left[-S, S\right]$ for all $i$. We now define $L$ as
+%
+\begin{align}\label{eq_boundary}
 L=c \cdot S,
-\end{equation} 
-%
-where $S > 0$ represents the half-range of the input space, and $c \geq 1$ is the proportional extension factor.
+\end{align} 
+%
+where $S > 0$ represents  
+%the half-range of the input space
+%{\color{blue}the boundary of the zero-mean minimum symmetric domain $\Psi=[-S,S]$ containing the input values $\bm{x}$}
+{\color{blue}the maximum absolute value of the input space}, 
+and $c \geq 1$ is the proportional extension factor.
 In the following, we will refer to $c$ as the boundary factor of the approximation.
-The boundary factor can also be regarded as the boundary $L$ normalized by the half-range $S$ of the input space.
+The boundary factor can also be regarded as the boundary $L$ normalized by
+%the half-range $S$ of the input space.
+{\color{blue}the boundary $S$}. 
+{\color{blue}Notice that $\bm{x}$ need not be symmetric around zero (although it is recommended that $\bm{x}$ fits exactly the interval $\left[-S, S\right]$ to optimize computation and the iterative diagnostic presented in Section~\ref{subsec_diagnostics}), but contained in $\left[-S, S\right]$}.
 
@@
 We start by illustrating how the number of basis functions $m$ and boundary factor $c$ influence the accuracy of the HSGP approximations individually.
-For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of length-scale $\ell=0.3$ and marginal variance $\alpha=1$, using input values from the zero-mean input domain with half-range $S=1$.
+For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of length-scale $\ell=0.3$ and marginal variance $\alpha=1$, using input values
+%from the zero-mean input domain with half-range $S=1$
+{\color{blue}within the interval $[-1,1]$ that leads to a boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}.
+Notice that in this case the domain of input points $x_i$ matches exactly the interval $\left[-S, S\right]$}.
 Several HSGP models with varying $m$ and $c$ are fitted to this data.
@@
 %
-When $c$ is below our recommendations given $\ell$ ({\color{blue}recommendations that we give later throughout the paper and specifically in the next Sections \ref{subsec_empiric_relations} and \ref{sec_num_equations}}), the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ are changed.
+When $c$ is below our recommendations given $\ell$ (recommendations that we give later throughout the paper and specifically in the next Sections \ref{subsec_empiric_relations} and \ref{sec_num_equations}), the effect of $\ell$ and $c$ to the recommended $m$ is more non-linear, but as long as we stay in the recommended range the linearity assumption is useful thumb rule for how to change $m$, if $\ell$ or $c$ are changed.
 
@@
 \subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig5_relationships.png}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale $\ell$ normalized by {\color{blue}the input space boundary $S$} ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
   \label{fig5_relationships}
@@
 
-Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ relative to the half-range $S$ of the input domain).
-Figure \ref{fig5_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) with {\color{blue}normalized} length-scale $\frac{\ell}{S}$ (from now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\ell$.
+Figure \ref{fig5_relationships} depicts how $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ normalized {\color{blue}by the input space boundary $S$}) interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) {\color{blue}with length-scale $\ell$
+%(from now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$)
+ and given a input space boundary $S$ and a boundary factor $c$}, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
@@
 where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions.
-Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$.
-Of course, we may also read it as providing the minimum $\ell$ that can be approximated with high accuracy given $m$ and $c$.
+Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\frac{\ell}{S}$ and $m$.
+Of course, we may also read it as providing the minimum $\frac{\ell}{S}$ that can be approximated with high accuracy given $m$ and $c$.
 We obtain the following main conclusions:
@@
 \item As $\ell$ increases, $m$ required for an accurate approximation decreases.
-({\color{blue}Notice that the larger $\ell$ the smoother the functions generated from a covariance function, and vice versa; see covariance functions equations in Section \ref{ch4_sec_cov}.})
+(Notice that the larger $\ell$ the smoother the functions generated from a covariance function, and vice versa; see covariance functions equations in Section \ref{ch4_sec_cov}.)
 \item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
@@
 As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\ell$ and $c$.
-To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically ({\color{blue}checked this linear relationship and} derived the practically useful constant terms.
+To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically {\color{blue}checked this linear relationship and} derived the practically useful constant terms.
 We require a lower bound for $c$ of $c \geq 1.2$ such that the equations below are precise enough for practical application.
@@
 
-{\color{blue}Note that in the above equations we have made explicit the use of the normalized length-scale by the half-range of the input space ($\frac{\ell}{S}$) (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$).
-By using the normalized length-scale, the equations, figures, and conclusions made throughout the paper can easily be referred to any specific scale of the input space.}
-
-Figure~\ref{fig5_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$.
-Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$.
+{\color{blue}
+%By using the normalized length-scale $\frac{\ell}{S}$, the equations, figures, and conclusions made throughout the paper regarding the diagnostic of the method can easily be referred to any specific scale of the input space.
+}
+
+Figure~\ref{fig5_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\frac{\ell}{S}$.
+Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\frac{\ell_1}{S}$ and $\frac{\ell_2}{S}$.
 More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions.
-For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$.
+For instance, in the two-dimensional case, it would depend on the ratio between $\frac{\ell_1}{S}$ and $\frac{\ell_2}{S}$.
 Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases.
@@
 In Appendix~\ref{sec_periodic}, we also analyze the accuracy of this linear representation and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel.
-Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition{\color{blue}, nor is there a need for a normalized length scale, since the length-scale refers to the period of a sinusoidal function}.
+Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition, nor is there a need for a normalized length-scale, since the length-scale refers to the period of a sinusoidal function.
 This relationship between $m$ and $\ell$ for a periodic squared exponential covariance function is gathered in Figure~\ref{figB1_m_lscale_periodic} and the numerical equation was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
@@
 
-Having obtained an estimate $\hat{\ell}$ for a HSGP model with prespecified $m$ and $c$, we can check whether $\hat{\ell}$ exceeds the smallest length-scale, that can be accurately inferred, provided as a function of $m$ and $c$ by equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used).
+Having obtained an estimate $\hat{\ell}$ for a HSGP model with prespecified $m$ and $c$, we can check whether $\hat{\ell}$ exceeds the smallest length-scale that can be accurately inferred, provided as a function of $m$ and $c$ by equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used).
 If $\hat{\ell}$ exceeds this value, the approximation is assumed to be good.
 If $\hat{\ell}$ does not exceed this value, the approximation may be inaccurate, and $m$ and/or $c$ need to be increased.
-In Figures~\ref{fig3_MSE_vs_J} and \ref{fig4_lscale_vs_J}, $m = 10$ and $c = 1.5$ were sufficient for an accurate modeling of function with $\ell = 0.3$, which matches the diagnostic based on equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
+In Figures~\ref{fig3_MSE_vs_J} and \ref{fig4_lscale_vs_J}, $m = 10$ and $c = 1.5$ were sufficient for an accurate modeling of function with $\ell = 0.3$ and $S=1$, which matches the diagnostic based on equations \eqref{eq_m_l_QE} and \eqref{eq_c_vs_l_QE}.
 
@@
 
+\item[A0.] {\color{blue}Compute the boundary $S$ of the input values by using eq.~\eqref{eq_boundary_S}}.
+
 \item[A1.] Set the iteration index to $k=1$.
 Make an initial guess on the length-scale $\ell^{(k)}$.
-If there is no useful information available, we recommend to start with a large length-scale, about $\ell^{(k)} \in [0.5, 1]$.
-({\color{blue}Remember that $\ell$ here refers to the normalized length-scale $\frac{\ell}{S}$.})
-
-\item[A2.] Obtain the minimum valid boundary factor $c^{(k)}$ determined by $\ell = \ell^{(k)}$ for the given kernel as per Section~\ref{sec_num_equations}.
-
-\item[A3.] Obtain the mimimum valid number of basis functions $m^{(k)}$ determined by $\ell = \ell^{(k)}$ and $c = c^{(k)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+If there is no useful information available, we recommend to start with a large length-scale, {\color{blue}such as the normalized length-scale being about $\frac{\ell^{(k)}}{S} \in [0.5, 1]$}.
+%({\color{blue}Remember that $\ell$ here refers to the normalized length-scale $\frac{\ell}{S}$.})
+
+\item[A2.] Obtain the minimum valid boundary factor $c^{(k)}$ determined by $\ell=\ell^{(k)}$ for the given kernel and boundary $S$ as per Section~\ref{sec_num_equations}.
+
+\item[A3.] Obtain the mimimum valid number of basis functions $m^{(k)}$ determined by $\ell=\ell^{(k)}$ and $c = c^{(k)}$ for the given kernel and boundary $S$ as per Section~\ref{sec_num_equations}.
 
@@
 
-\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(k)} - 0.01 \geq \ell^{(k)}$.
+\item[A5.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(k)} + 0.01 \geq \ell^{(k)}$.
 
@@
 
-\item[B3.] Obtain the minimum valid boundary factor $c^{(k)}$ determined by $\ell = \hat{\ell}^{(k-1)}$ for the given kernel as per Section \ref{sec_num_equations}.
-
-\item[B4.] Obtain the minimum valid length-scale $\ell^{(k)}$ that can be accurately estimated by $m = m^{(k)}$ and $c = c^{(k)}$ for the given kernel as per Section~\ref{sec_num_equations}.
+\item[B3.] Obtain the minimum valid boundary factor $c^{(k)}$ determined by $\ell = \hat{\ell}^{(k-1)}$ for the given kernel and boundary $S$ as per Section \ref{sec_num_equations}.
+
+\item[B4.] Obtain the minimum valid length-scale $\ell^{(k)}$ that can be accurately estimated by $m = m^{(k)}$ and $c = c^{(k)}$ for the given kernel and boundary $S$ as per Section~\ref{sec_num_equations}.
 
@@
 
-\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(k)} - 0.01 \geq \ell^{(k)}$.
+\item[B6.] Perform the length-scale diagnostic by checking if $\hat{\ell}^{(k)} + 0.01 \geq \ell^{(k)}$.
 
@@
 
-{\color{blue} We have seen this procedure do converge quickly after only few iterations in all the investigated cases (see Section~\ref{sec_cases}).
+{\color{blue} We have seen this procedure do converge quickly after only few iterations in all the investigated cases (see Figure~\ref{fig7_diagnostic_lscale_comparison}, Table~\ref{tab1_diagnostic}, and Section~\ref{sec_cases}).
 That said, we cannot rule out that there may be cases where convergence is much slower (in terms of number of required diagnostic iterations) to a degree where HSGPs become impractical and an early stopping of the procedure would be advisable.
@@
 Ensuring that the relative distance is bounded by a small constant for the relevant lengthscale implies a high quality approximation.
-Another possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation they are actually explaining{\color{blue}, as already mentioned in Section \ref{subsec_theoretic_linearity}}.
+Another possibility to asses the accuracy of the approximation is to look at the cumulative sum of the spectral densities terms used in the series expansion and find out how much of the total variation they are actually explaining, as already mentioned in Section \ref{subsec_theoretic_linearity}.
 
@@
 
-The HSGP model is computationally superior {\color{blue} to the exact GP} in $1$D and $2$D even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$, i.e., $n$ smaller than some value around 300) that exact GPs are already reasonably fast themselves.
+The HSGP model is computationally superior to the exact GP in $1D$ and $2D$ even for highly wiggly functions, except when the number of data points is so small ($n \lesssim 300$, i.e., $n$ smaller than some value around 300) that exact GPs are already reasonably fast themselves.
 However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})).
@@
 
-In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3$D datasets ($n \gtrsim 1000$, i.e., $n$ greater than some value around 1000), even for highly wiggly 3D functions (e.g., $\ell_1, \ell_2, \ell_3 \approx 0.1$).
-
-For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for highly to moderated wiggly $3$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3 \gtrsim 0.3$) and for overall smooth $4$D functions (e.g., $\ell_1 \lesssim 0.1$, and $\ell_2, \ell_3, \ell_4 \gtrsim 0.4$).
+In our experiments of multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderate-to-large sized $3D$ datasets ($n \gtrsim 1000$, i.e., $n$ greater than some value around 1000), even for highly wiggly $3D$ functions (e.g., $\frac{\ell_1}{S}, \frac{\ell_2}{S}, \frac{\ell_3}{S} \approx 0.1$).
+
+For small sized datasets ($n\lesssim 1000$), HSGPs are likely to be slower than exact GPs already for highly to moderated wiggly $3D$ functions (e.g., $\frac{\ell_1}{S} \lesssim 0.1$, and $\frac{\ell_2}{S}, \frac{\ell_3}{S} \gtrsim 0.3$) and for overall smooth $4D$ functions (e.g., $\frac{\ell_1}{S} \lesssim 0.1$, and $\frac{\ell_2}{S}, \frac{\ell_3}{S}, \frac{\ell_4}{S} \gtrsim 0.4$).
 
 As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum sufficient value, reducing computational time drastically, and still getting an accurate approximation.
-For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\ell \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \ell \gtrsim 0.1$; and $m \gtrsim 20$ for $\ell \lesssim 0.1$), which, from results presented in Figure~\ref{fig14_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3$D functions and smooth $4$D functions.
+For example, assuming a squared exponential covariance function, choosing the optimal value for $c$ allows one to use few basis functions in every single dimension (i.e., $m \lesssim 10$ for $\frac{\ell}{S} \gtrsim 0.3$; $20 \gtrsim m \gtrsim 10$ for $0.3 \gtrsim \frac{\ell}{S} \gtrsim 0.1$; and $m \gtrsim 20$ for $\frac{\ell}{S} \lesssim 0.1$), which, from results presented in Figure~\ref{fig14_time_nD_log}, implies that the HSGP model can be, in general terms, useful for highly wiggly $3D$ functions and smooth $4D$ functions.
 
@@
 In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
-The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process $f(x) \sim \GP(0, k(x, x', \theta))$ with a Mat{\'e}rn ($\nu$=3/2) covariance function $k$ with marginal variance $\alpha=1$ and length-scale $\ell=0.2$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$.
+The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process $f(x) \sim \GP(0, k(x, x', \theta))$ with a Mat{\'e}rn ($\nu$=3/2) covariance function $k$ with marginal variance $\alpha=1$ and length-scale $\ell=0.2$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$ {\color{blue} that lead to an input space boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}}.
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
@@
 
-Figure~\ref{fig8_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot 1= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
+Figure~\ref{fig8_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot S= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
 In order to do model comparison a spline-based model with 60 knots is also fitted and plotted jointly with the GP and HSGP models.
@@
 Thus, the functions $f_2$ and $f_3$ are approximated as in eq.~(\ref{eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in eq.~(\ref{eq_q_2}).
+{\color{blue}The input variable $\bm{t}=\{1,2,\dots,7305\}$ is previously standardized to have zero mean and unit variance $\left(\frac{\bm{t}-\text{mean}(\bm{t})}{\text{sd}(\bm{t})}\right)$, and then it follows a input space boundary $S=1.732$}.
 
@@
 For functions $f_2$ and $f_3$ there are not boundary factor $c$ as they use periodic covariance functions, and $m$ and minimum $\ell$ are updated by eq.~\eqref{eq_J_l_periodic}.
-A initial value for the minimum lengthscale $\ell_1$ of $f_1$ is guessed to correspond to around 3 years (i.e., $\ell_1=3$ years $= 3\cdot 365=1095$ days) of the input dimension.
-Initial values for the minimum lengthscales $\ell_2$ and $\ell_3$ of $f_2$ and $f_3$, respectively, are guessed to correspond to half of the period (i.e., $l_2=l_3=0.5$).
+A initial value for the minimum lengthscale $\ell_1$ of $f_1$ is guessed to correspond to around 3 years (i.e., $\ell_1=3$ years $= 3\cdot 365=1095$ days) {\color{blue}in the original scale or around $0.52$ (i.e., $\ell_1=\nicefrac{1095}{\text{sd}(\bm{t})}=0.52$) in the standardized scale used as input in the model}.
+Initial values for the minimum lengthscales $\ell_2$ and $\ell_3$ of $f_2$ and $f_3$, respectively, are guessed to correspond to half of the period (i.e., $\ell_2=\ell_3=0.5$).
 After the diagnostic generated the first \textit{true}, $m$ is updated by increasing the $m$ of the previous iteration by 5 additional basis functions and $c$ and the minimum $\ell$ updated accordingly as explained in Section~\ref{sec_user_guide}.
@@
 
-In this case study, we apply the diagnostic tool to fit and diagnose two different data sets, one data set simulated from a bivariate (D=$2$) function and another data set simulated from a $3$-dimensinal (D=$3$) function.
-Furthermore, in Section~\ref{sec_computation_nD} we present results of the computational time required to fit the HSGP model in $2$D, $3$D and $4$D input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
-
-$2$D and $3$D synthetic functions were drawn from $2$D and $3$D GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively.
+In this case study, we apply the diagnostic tool to fit and diagnose two different data sets, one data set simulated from a bivariate ($D=2$) function and another data set simulated from a $3$-dimensinal ($D=3$) function.
+Furthermore, in Section~\ref{sec_computation_nD} we present results of the computational time required to fit the HSGP model in $2D$, $3D$ and $4D$ input spaces, with different sized data sets and as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+
+$2D$ and $3D$ synthetic functions were drawn from $2D$ and $3D$ GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively, {\color{blue}and then it follows a input space boundary $S=1$ for both the $2D$ and $3D$ functions}.
 Squared exponential covariance functions with marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension, were used in the GP priors.
-$200$ and $1000$ data points were sampled from the $2$D and $3$D drawn functions, respectively, and independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the data points to form the final noisy sets of observations.
+$200$ and $1000$ data points were sampled from the $2D$ and $3D$ drawn functions, respectively, and independent Gaussian noise with standard deviation $\sigma=0.2$ was added to the data points to form the final noisy sets of observations.
 
@@
 
-Table~\ref{tab_caseIII_2D} and \ref{tab_caseIII_3D} contain the iterative steps to fit and diagnose the $2$D and $3$D data sets, respectively. 
+Table~\ref{tab_caseIII_2D} and \ref{tab_caseIII_3D} contain the iterative steps to fit and diagnose the $2D$ and $3D$ data sets, respectively. 
 %These tables contain the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ for each dimension, and the RMSE compute with both the data and the GP as the reference for every iterative steps of the fitting process.
-The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{\text{st}}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2$D and $3$D data sets, respectively) to the $4^{\text{th}}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2$D and $3$D data sets, respectively) required to accurately approximate these notably wiggly functions.
+The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{\text{st}}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2D$ and $3D$ data sets, respectively) to the $4^{\text{th}}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2D$ and $3D$ data sets, respectively) required to accurately approximate these notably wiggly functions.
 Without using our recommended guidelines and diagnostics, it would likely have taken the user several more trials to get the optimal solution and to spend a significant amount of time, as using a larger number of basis functions in multivariate cases increases time of computation drastically.
@@
 \hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{fig13_legend_exII.pdf}}}
-\caption{Error between the $2$D data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models.
+\caption{Error between the $2D$ data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models.
 Sample points are plotted as circles.
@@
 \end{tabular}
-\caption{Iterative steps in HSGP model fitting on the $2$D data.
+\caption{Iterative steps in HSGP model fitting on the $2D$ data.
 The diagnostic procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$, and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the two input dimensions.
@@
 \end{tabular}
-\caption{Iterative steps in HSGP model fitting on the $3$D data.
+\caption{Iterative steps in HSGP model fitting on the $3D$ data.
 The diagnostic procedure updating the parameters minimum $\ell$, $c$, $m$, estimated $\hat{\ell}$ and diagnostic $\hat{\ell} + 0.01 \geq \ell$ is applied on the three input dimensions.
@@
 
-Figure \ref{fig13_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2$D data set.
-For the $3$D data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3$D input space.
+Figure \ref{fig13_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2D$ data set.
+For the $3D$ data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3D$ input space.
 The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
@@
 For the spline-based model, $40$ knots in each dimension were used.
-Figure \ref{fig13_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2$D spline-based model.
+Figure \ref{fig13_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2D$ spline-based model.
 It can be seen that the posterior mean for HSGP and exact GP accurately represents the true data-generating function.
@@
 
-Figure~\ref{fig14_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2$D, $3$D and $4$D input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
+Figure~\ref{fig14_time_nD_log} shows the computation time for dynamic HMC (in seconds per iteration) in $2D$, $3D$ and $4D$ input spaces and different sized data sets, $n=300, n=500, n=1000$, and $n=3000$, as a function of the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ used in the approximation.
 Both the time of computation and $m^*$ are plotted in the logarithmic scale.
@@
 
-For a very wiggly $2$D function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig14_time_nD_log}).
+For a very wiggly $2D$ function (say, with $\ell_1=0.1$ and $\ell_2=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22= 484$, which results in significantly faster computation than the exact GP, even with small data sets (i.e., $n \lesssim 300$) (see Figure~\ref{fig14_time_nD_log}).
 For a very wiggly $3$D function (say, with $\ell_1=0.1$, $\ell_2=0.1$ and $\ell_3=0.1$), the approximate number of multivariate basis functions needed is $m^*=22\times 22\times 22= 10648$, which, unless for small data sets (i.e., $n \lesssim 1000$), the method is still significantly faster than the regular GP.
-A $3$D function where each dimension has a lengthscale around $0.1$ is not that common in statistical analysis, and thus in many cases the approximation will be significantly faster than excat GP.
-For $4$D data sets, the method can still be more efficient than the exact GP for moderate-to-large data sets (i.e., $n \lesssim 1000$).
+A $3D$ function where each dimension has a lengthscale around $0.1$ is not that common in statistical analysis, and thus in many cases the approximation will be significantly faster than excat GP.
+For $4D$ data sets, the method can still be more efficient than the exact GP for moderate-to-large data sets (i.e., $n \lesssim 1000$).
 
@@
 
-The Stan model codes for the exact GP model and the HSGP model for these 2D, 3D, and 4D cases can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Stan_code_2D_3D_4D}}\,.
+The Stan model codes for the exact GP model and the HSGP model for these $2D$, $3D$, and $4D$ cases can be found online at {\small \url{https://github.com/gabriuma/basis_functions_}} \linebreak {\small \url{approach_to_GP/tree/master/Paper/Stan_code_2D_}} \linebreak {\small \url{3D_4D}}\,.
 
@@
 	\begin{enumerate}
-	\item Make the initial guess that the length-scale $\ell_1=0.5$.
-
-	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	TODO: name the value of $c_1$ and all the other quantities in the following. Otherwise, what is the purpose of this example?
-
-	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
-
-	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
-
-	TODO: provide at least a mean value for the posterior of  $\hat{\ell}_1$ here?
+	\item Compute the boundary $S$ of the input values by using eq.~\eqref{eq_boundary_S} $\to$ $S=1$.
+	\item Make an initial guess for the length-scale $\ell$ $\to$ $\ell_1=0.5$.
+
+	\item Compute the optimal $c_1$ as a function of $\ell_1$ and $S$ by using eq.~(\ref{eq_c_vs_l_QE}) $\to$ $c_1=1.60$.
+
+	\item Compute the minimum $m_1$ as a function of $c_1$, $\ell_1$, and $S$ by using eq.~(\ref{eq_m_l_QE}) $\to$ $m_1=6$.
+
+	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$  $\to$ $\text{mean}(\hat{\ell}_1)=0.17$.
 
@@
 
-	TODO: I think you wanted to use $\hat{\ell}_1 - 0.01 \geq \ell_1$ and I corrected it above. Can you verify whether you want + or -?
 	\end{enumerate}
@@
 
-	\item Compute the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.
-
-	TODO: why make this a single step here when it was two steps in Section 4.4.1?
-
-	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+	\item Compute the optimal $c_2$ as a function of $\ell_2$ and $S$ by using eq.~\eqref{eq_c_vs_l_QE} $\to$ $c_2=1.20$.
+
+	\item Compute the minimum $m_2$ as a function of $c_2$, $\ell_2$, and $S$ by using eq.~\eqref{eq_m_l_QE} $\to$ $m_2=13$.
+
+	\item Fit the HSGP model using $c_2$ and $m_2$ and obtain the estimated $\hat{\ell}_2$ $\to$ $\hat{\ell}_2=0.07$.
 
@@
 
-	\item Compute the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Fit the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.
+	\item Compute the optimal $c_4$ as a function of $\ell_3$ and $S$ by using eq.~(\ref{eq_c_vs_l_QE}) $to$ $c_4=1.20$.
+
+	\item Fit the HSGP model using $c_4$ and $m_4$ and obtain the estimated $\hat{\ell}_4$ $to$ $\hat{\ell}_4=0.08$.
 
@@
 	\begin{enumerate}
-	\item Make the initial guess that the length-scale $\ell_1=1$.
-
-	\item Compute the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Compute the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).
-
-	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$.
+	\item Compute the boundary $S$ of the input values by using eq.~\eqref{eq_boundary_S} $\to$ $S=1$.
+	\item Make an initial guess for the length-scale $\ell$ $\to$ $\ell_1=1$.
+
+	\item Compute the optimal $c_1$ as a function of $\ell_1$ and $S$ by using eq.~(\ref{eq_c_vs_l_QE}) $\to$ $c_1=3.20$.
+
+	\item Compute the minimum $m_1$ as a function of $c_1$, $\ell_1$, and $S$ by using eq.~(\ref{eq_m_l_QE}) $\to$ $m_1=6$.
+
+	\item Fit the HSGP model and obtain the estimated $\hat{\ell}_1$ $\to$ $\hat{\ell}_1=1.02$.
 
@@
 
-	\item Compute the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).
-
-	\item Fit the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.
+	\item Compute the optimal $c_2$ as a function of $\ell_2$ and $S$ by using eq.~(\ref{eq_c_vs_l_QE}) $\to$ $c_2=3.27$.
+
+	\item Fit the HSGP model using $c_2$ and $m_2$ and obtain the estimated $\hat{\ell}_2$ $\to$ $\hat{\ell}_2=1.23$.
 

### 2022-07-22 Paper/Manuscript_Stats&Comp/reply_letter_2nd_STCO.tex
More edits to the paper and response letter 2

@@
 We apologize for creating confusion with our notation here.
-We work with a normalized lengthscale, i.e., a lengthscale $\ell$ relative to the half-range ($S$) of the input domain ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
+{\color{blue}The diagnostic Figure 5 and equations (19-24) work with a normalized lengthscale, i.e., a lengthscale $\ell$ normalized by the zero-mean, symmetric interval $S$ which contains the inpus values $x_i$ ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
+In concrete in step A.1 in Section 4.5.1 we recommend values for a first guess of the lengthscale based on the normalized lengthscale.
+We have now clarified the role and use of the normalized lengthscale relative to $S$ throughout the paper.}
+
+%We work with a normalized lengthscale, i.e., a lengthscale $\ell$ relative to the half-range ($S$) of the input domain ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
 We have now clarified the notation by using different symbols for lengthscale and normalized lengthscale throughout the paper.
-
-%added an additional paragraph to make it more clear in the second-to-last paragraph in Section 4.3.1.
 

### 2022-07-27 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
some edits of blue texts

@@
 {\color{blue}the boundary $S$}. 
-{\color{blue}Notice that $\bm{x}$ need not be symmetric around zero (although it is recommended that $\bm{x}$ fits exactly the interval $\left[-S, S\right]$ to optimize computation and the iterative diagnostic presented in Section~\ref{subsec_diagnostics}), but contained in $\left[-S, S\right]$}.
+{\color{blue}Notice that $\bm{x}$ need not be symmetric around zero, but need to be contained in $\left[-S, S\right]$} (although it is recommended that $\bm{x}$ fits tightly the interval $\left[-S, S\right]$ to optimize computation and the iterative diagnostic presented in Section~\ref{subsec_diagnostics}).
 
@@
 {\color{blue}within the interval $[-1,1]$ that leads to a boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}.
-Notice that in this case the domain of input points $x_i$ matches exactly the interval $\left[-S, S\right]$}.
+In this case the domain of input points $x_i$ matches exactly the interval $\left[-S, S\right]$}.
 Several HSGP models with varying $m$ and $c$ are fitted to this data.
@@
 
-{\color{blue} We have seen this procedure do converge quickly after only few iterations in all the investigated cases (see Figure~\ref{fig7_diagnostic_lscale_comparison}, Table~\ref{tab1_diagnostic}, and Section~\ref{sec_cases}).
+{\color{blue} In our experiments this procedure converges quickly after only few iterations in all the cases (see Figure~\ref{fig7_diagnostic_lscale_comparison}, Table~\ref{tab1_diagnostic}, and Section~\ref{sec_cases}).
 That said, we cannot rule out that there may be cases where convergence is much slower (in terms of number of required diagnostic iterations) to a degree where HSGPs become impractical and an early stopping of the procedure would be advisable.

### 2022-08-05 Paper/Manuscript_Stats&Comp/reply_letter_2nd_STCO.tex
Minor edits to reply letter

@@
 
-\item We have revised the entire manuscript making some minor edits, such as rewriting a few sentences and adding minor little additional information.
+\item We have revised the entire manuscript making several minor edits, such as rewriting a few sentences and adding minor additional information.
 
@@
 With regard to high-dimensional cases, the main problem would not be more iterations (as far as we have seen) but the computational cost of each iteration.
-More concretely, as we discuss in detail in the paper, the more input dimensions ($D$) we have, the more multivariate basis functions we get ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit (per iteration), but not (in a relevant manner) the number of iterations required to complete the diagnostic procedure.
+More concretely, as we discuss in detail in the paper, the more input dimensions ($D$) we have, the number of required multivariate basis functions increases ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit (per iteration), but not (in a relevant manner) the number of iterations required to complete the diagnostic procedure.
 We have now clarified these points at the end of Section 4.5.1.

### 2022-08-05 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Fix typo in leukemia experiment
Change color of add text to blue
A few more edits

@@
 \st{. Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.}
-{\color{blue}, and the set of input points $\{x_i\}_{i=1}^n$ has some kind of centering around zero.}
+{\color{blue}, and the set of input points $\{x_i\}_{i=1}^n$ is centered around zero.}
 Let 
@@
 Make an initial guess on the length-scale $\ell^{(k)}$.
-If there is no useful information available, we recommend to start with a large length-scale, {\color{blue}such as the normalized length-scale being about $\frac{\ell^{(k)}}{S} \in [0.5, 1]$}.
+If there is no useful information available, we recommend to start with a large length-scale, {\color{blue}such as the normalized length-scale being in the interval $\frac{\ell^{(k)}}{S} \in [0.5, 1]$}.
 %({\color{blue}Remember that $\ell$ here refers to the normalized length-scale $\frac{\ell}{S}$.})
@@
 
-$2D$ and $3D$ synthetic functions were drawn from $2D$ and $3D$ GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively, {\color{blue}and then it follows a input space boundary $S=1$ for both the $2D$ and $3D$ functions}.
+$2D$ and $3D$ synthetic functions were drawn from $2D$ and $3D$ GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively, {\color{blue}and the input space boundary is $S=1$ for both the $2D$ and $3D$ functions}.
 Squared exponential covariance functions with marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension, were used in the GP priors.
@@
 p(y_i > t_i \mid f) &= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i \\
-&=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
+&=  1 - \Phi\! \left( \frac{\log(t_i)-f(\bm{x}_i)}{\sigma} \right),
 \end{align*}
 %
-where $y_i>t_i$ denotes the unobserved survival time.
+where $y_i>t_i$ denotes the unobserved survival time {\color{blue} and $\Phi$ is the standard normal cumulative distribution function}.
 The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$.

### 2022-08-05 Paper/Manuscript_Stats&Comp/reply_letter_2nd_STCO.tex
Small edits in reply letter

@@
 
+\section*{\normalsize Reply to reviewer comments for \\ \Large`Practical Hilbert space approximate Bayesian Gaussian processes \\ for probabilistic programming'}
+
 We would like to thank the reviewers for their thorough comments and valuable feedback, which has helped us a lot in improving the paper.
@@
 
-\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}
+\noindent \hdashrule{\textwidth}{0.2pt}{2mm 1pt}
 
@@
 
-\item We have updated the appearance of Figures 1, 2, 3, 4, and 5 (old numbering) in order to make them more compact and thus the overall appearance of the paper more pleasing.
-The new numbering corresponding to those figures is now Fig.
-1, 2, 3, and 4.
+\item We have updated the appearance of Figures~1, 2, 3, 4, and 5 (old numbering) in order to make them more compact and thus the overall appearance of the paper more pleasing.
+The new numbering corresponding to those figures is now Fig.~1, 2, 3, and 4.
 
@@
 
-Figure 6 may look a bit fuzzy, but it is actually a visual effect because the curves were constructed for discrete values of the number of basis functions $m$ (Y-axis), which makes them look like stair-shaped curves when plotted, and its appearance can be confused with a blurred image.
+Figure 6 may look a bit fuzzy, but it is actually a visual effect because the curves were constructed for discrete values of the number of basis functions $m$ ($y$-axis), which makes them look like stair-shaped curves when plotted, and its appearance can be confused with a blurred image.
 We think the resolution of the figure itself is sufficient.
@@
 With regard to high-dimensional cases, the main problem would not be more iterations (as far as we have seen) but the computational cost of each iteration.
-More concretely, as we discuss in detail in the paper, the more input dimensions ($D$) we have, the number of required multivariate basis functions increases ($m^*=nm^D$, with $n$=number of points and $m$=number of univariate basis functions), which affects the HMC sampling of a model fit (per iteration), but not (in a relevant manner) the number of iterations required to complete the diagnostic procedure.
-We have now clarified these points at the end of Section 4.5.1.
+More concretely, as we discuss in detail in the paper, the more input dimensions ($D$) we have, the number of required multivariate basis functions increases ($m^*=nm^D$, with $n$ = number of points and $m$ = number of univariate basis functions), which affects the HMC sampling of a model fit (per iteration), but not (in a relevant manner) the number of iterations required to complete the diagnostic procedure.
+We have now clarified these points at the end of Section~4.5.1.
 
@@
 We are not sure whether we can convince you of the sensibility and sufficiency of the focus of our paper that we have chosen.
-But what we can say is that based on the informal feedback from the community we have recieved so far, the other reviewers' and also the editor's comments, as well as over 20 citation of the corresponding preprint already, a lot of people see value in our work and its existing focus.
+But what we can say is that based on the informal feedback from the community we have received so far, the other reviewers' and also the editor's comments, as well as over 20 citation of the corresponding preprint already, a lot of people see value in our work and its existing focus.
 There is certainly a lot of value in comparing HSGPs with competing methods other than exact GPs, but we do not see a way to make this part of the current paper, not only in terms of paper space, but also in terms of time, personnel, and monetary resources we currently have available to work on this topic.
@@
 We apologize for creating confusion with our notation here.
-{\color{blue}The diagnostic Figure 5 and equations (19-24) work with a normalized lengthscale, i.e., a lengthscale $\ell$ normalized by the zero-mean, symmetric interval $S$ which contains the inpus values $x_i$ ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
+{\color{blue}The diagnostic Figure~5 and equations (19--24) work with a normalized lengthscale, i.e., a lengthscale $\ell$ normalized by the zero-mean, symmetric interval $S$ which contains the inpus values $x_i$ ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
 In concrete in step A.1 in Section 4.5.1 we recommend values for a first guess of the lengthscale based on the normalized lengthscale.
@@
 
-Thank you for also your comment about the $\ell_i$ notation in Section 4.5.1, where $i$ is the subscript for the diagnostic iteration, conflicting with the $\ell_i$ notation used in the article where $i$ is the subscript for dimensions.
-Therefore, we have changed the notation in Section 4.5.1 from $\ell_i$ to $\ell^{(k)}$ to clearly differentiate it from the notation in the rest of the article where subscripts refer to dimensions.
+Thank you for also your comment about the $\ell_i$ notation in Section~4.5.1, where $i$ is the subscript for the diagnostic iteration, conflicting with the $\ell_i$ notation used in the article where $i$ is the subscript for dimensions.
+Therefore, we have changed the notation in Section~4.5.1 from $\ell_i$ to $\ell^{(k)}$ to clearly differentiate it from the notation in the rest of the article where subscripts refer to dimensions.
 

### 2022-08-15 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
updated manuscript submitted

@@
               \emph{Corresponding Author} \\
-              \email{gabriuma@gmail.com; riutort\_gab@gva.es}       %  \\
+              \email{riutort\_gab@gva.es; gabriuma@gmail.com}       %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
 The former is commonly known as the squared exponential or exponentiated quadratic covariance function.
-As an example, assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernel $K_{\nu}$ written above takes the form
+As an example, assuming the Euclidean distance between observations, $r=|\bm{x}-\bm{x}'|_{L2}=\sqrt{\sum_{i=1}^{D}(x_i-x_i')^2}$, the kernel $k_{\infty}$ written above takes the form
 %
@@
 \begin{align*}
-S_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^\intercal \bm{\omega} \right)^{-(\nu+D/2)}
+s_{\nu}(\bm{\omega})&= \alpha \, \frac{2^D\pi^{D/2}\Gamma(\nu+D/2)(2\nu)^{\nu}}{\Gamma(\nu)\, \ell^{2\nu}}\left(\frac{2\nu}{\ell^2}+4\pi^2\bm{\omega}^\intercal \bm{\omega} \right)^{-(\nu+D/2)}
 \end{align*}
@@
 \begin{align}
-S_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\frac{1}{2}\ell^2 \bm{\omega}^\intercal \bm{\omega}), \label{eq_specdens_inf}  \\
-%S_{\frac{1}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+1}{2}} \label{eq_specdens_12},     \\
-S_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+3}{2}} \label{eq_specdens_32},     \\
-S_{\frac{5}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+5}{2}){5}^{5/2}}{\frac{3}{4}\sqrt{\pi}\ell^5}\left(\frac{5}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+5}{2}}.
+s_{\infty}(\bm{\omega})&= \alpha\,(\sqrt{2\pi})^D  \ell^D \exp(-\frac{1}{2}\ell^2 \bm{\omega}^\intercal \bm{\omega}), \label{eq_specdens_inf}  \\
+%s_{\frac{1}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+1}{2})}{\sqrt{\pi}\ell}\left(\frac{1}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+1}{2}} \label{eq_specdens_12},     \\
+s_{\frac{3}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+3}{2}){3}^{3/2}}{\frac{1}{2}\sqrt{\pi}\ell^3}\left(\frac{3}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+3}{2}} \label{eq_specdens_32},     \\
+s_{\frac{5}{2}}(\bm{\omega})&= \alpha\,\frac{2^D\pi^{D/2}\Gamma(\frac{D+5}{2}){5}^{5/2}}{\frac{3}{4}\sqrt{\pi}\ell^5}\left(\frac{5}{\ell^2}+\bm{\omega}^\intercal \bm{\omega} \right)^{-\frac{D+5}{2}}.
 \label{eq_specdens_52}
@@
 \begin{align*}
-S_{\infty}(\bm{\omega})&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 \omega_i^2 \right),   \\
-%S_{\frac{1}{2}}(\bm{\omega})&= \alpha \, 8\,\pi \prod_{i=1}^{3}\ell_i\left(1+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-2},    \\
-S_{\frac{3}{2}}(\bm{\omega})&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-3},   \\
-S_{\frac{5}{2}}(\bm{\omega})&= \alpha \, \frac{64}{3}\,\pi {5}^{5/2}\prod_{i=1}^{3}\ell_i\left(5+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-4}.
+s_{\infty}(\bm{\omega})&= \alpha \, (2\pi)^{3/2}  \prod_{i=1}^{3} \ell_i  \exp\!\left(-\frac{1}{2} \sum_{i=1}^{3} \ell_i^2 \omega_i^2 \right),   \\
+%s_{\frac{1}{2}}(\bm{\omega})&= \alpha \, 8\,\pi \prod_{i=1}^{3}\ell_i\left(1+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-2},    \\
+s_{\frac{3}{2}}(\bm{\omega})&= \alpha \, 32\,\pi {3}^{3/2}\prod_{i=1}^{3}\ell_i\left(3+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-3},   \\
+s_{\frac{5}{2}}(\bm{\omega})&= \alpha \, \frac{64}{3}\,\pi {5}^{5/2}\prod_{i=1}^{3}\ell_i\left(5+\sum_{i=1}^{3}\ell_i^2 \omega_i^2 \right)^{-4}.
 \end{align*}
@@
 \begin{equation}\label{eq_approxcov}
-k(x,x') = \sum_{j=1}^\infty S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
-\end{equation} 
-%
-where $S_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ is the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
+k(x,x') = \sum_{j=1}^\infty s_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'),
+\end{equation} 
+%
+where $s_{\theta}$ is the spectral density of the stationary covariance function $k$ (see Section \ref{ch4_sec_cov}) and $\theta$ is the set of hyperparameters of $k$ \citep{rasmussen2006gaussian}.
 The terms $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(x)\}_{j=1}^{\infty}$ are the sets of eigenvalues and eigenvectors, respectively, of the Laplacian operator in the given domain $\Omega$.
@@
 \begin{equation}
-k(x,x') \approx \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \bm{\phi}(x'), \nonumber
-\end{equation}
-%
-where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is a diagonal matrix of the spectral density evaluated at the square root of the eigenvalues, that is, $S_{\theta}(\sqrt{\lambda_j})$,
+k(x,x') \approx \sum_{j=1}^m s_{\theta}(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x') = \bm{\phi}(x)^\intercal \Delta \bm{\phi}(x'), \nonumber
+\end{equation}
+%
+where $\bm{\phi}(x)=\{\phi_j(x)\}_{j=1}^{m} \in {\rm I\!R}^{m}$ is the column vector of basis functions, and $\Delta  \in {\rm I\!R}^{m\times m}$ is a diagonal matrix of the spectral density evaluated at the square root of the eigenvalues, that is, $s_{\theta}(\sqrt{\lambda_j})$,
 %
@@
 \Delta =  \begin{bmatrix}
-    S_{\theta}(\sqrt{\lambda_1}) & & \\
+    s_{\theta}(\sqrt{\lambda_1}) & & \\
     & \ddots & \nonumber \\
-    & & S_{\theta}(\sqrt{\lambda_m}) \\
+    & & s_{\theta}(\sqrt{\lambda_m}) \\
   \end{bmatrix}.
@@
 \begin{equation}\label{eq_approxf}
-f(x) \approx \sum_{j=1}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
+f(x) \approx \sum_{j=1}^m \left( s_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
 \end{equation}
@@
 A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the hyperparameters of the covariance function $\theta$.
-Instead, the only dependence of the model on $\theta$ is through the spectral density $S_{\theta}$.
-The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions.
+Instead, the only dependence of the model on $\theta$ is through the spectral density $s_{\theta}$.
+The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $s_{\theta}$ goes rapidly to zero for bounded covariance functions.
 Therefore, eq.~\eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$.
@@
 Then, the number of multivariate eigenfunctions and eigenvalues is $m^{\ast} = m_{1} \cdot m_{2} \cdot m_{3} = 12$ and the matrix $\mathbb{S}\in {\rm I\!N}^{12 \times 3}$ is given by
-%
-%\begin{align}\small
-%\mathbb{S}=
-%\left[ {\begin{array}{ccc}
-%1 & 1 & 1 \nonumber \\
-%1 & 1 & 2 \\
-%1 & 1 & 3 \\
-%1 & 2 & 1 \\
-%1 & 2 & 2 \\
-%1 & 2 & 3 \\
-%2 & 1 & 1 \\
-%2 & 1 & 2 \\
-%2 & 1 & 3 \\
-%2 & 2 & 1 \\
-%2 & 2 & 2 \\
-%2 & 2 & 3 
-%\end{array} } \right].
-%\end{align} 
 %
@@
 k(\bm{x},\bm{x}') \approx \sum_{j=1}^{m^{\ast}} 
-S^{\ast}_{\theta}\left(\sqrt{\bm{\lambda}^{\ast}_j}\right)
+s^{\ast}_{\theta}\left(\sqrt{\bm{\lambda}^{\ast}_j}\right)
 \phi^{\ast}_j(\bm{x}) \phi^{\ast}_j(\bm{x}'),
@@
 %
-where $S^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}) as a function of $\sqrt{\bm{\lambda}^{\ast}_j}$ that denotes the element-wise square root of the vector $\bm{\lambda}^{\ast}_j$.
+where $s^{\ast}_{\theta}$ is the spectral density of the $D$-dimensional covariance function (see Section \ref{ch4_sec_cov}) as a function of $\sqrt{\bm{\lambda}^{\ast}_j}$ that denotes the element-wise square root of the vector $\bm{\lambda}^{\ast}_j$.
 We can now write the approximate series expansion of the multivariate function $f$ as
@@
 f(\bm{x}) \approx \sum_{j=1}^{m^{\ast}} 
-\left( S^{\ast}_{\theta} \left(\sqrt{\bm{\lambda}^{\ast}_j} \right)\right)^{\! \frac{1}{2}} \phi^{\ast}_j(\bm{x}) \beta_j, 
+\left( s^{\ast}_{\theta} \left(\sqrt{\bm{\lambda}^{\ast}_j} \right)\right)^{\! \frac{1}{2}} \phi^{\ast}_j(\bm{x}) \beta_j, 
 \end{equation}
@@
 \begin{equation*}
-k(\tau) = \sum_{j=1}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
+k(\tau) = \sum_{j=1}^{\infty} s_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
 \end{equation*}
@@
 \begin{equation}\label{eq_diff_covs}
- \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \,\mathrm{d}\tau  < \varepsilon.
+ \int | k(\tau) - \sum_{j=1}^m s_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \,\mathrm{d}\tau  < \varepsilon.
 \end{equation}
@@
 
-In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ contains all the inputs points $x_i$
-\st{. Without loss of generality, we can assume all data points are contained in a symmetric interval around zero.}
-{\color{blue}, and the set of input points $\{x_i\}_{i=1}^n$ is centered around zero.}
+In this work, we choose $L$ such that the domain $\Omega = \left[-L, L\right]$ contains all the inputs points $x_i$, and the set $\{x_i\}_{i=1}^n$ of input points is centered around zero.
 Let 
@@
 %
-where $S > 0$ represents  
-%the half-range of the input space
-%{\color{blue}the boundary of the zero-mean minimum symmetric domain $\Psi=[-S,S]$ containing the input values $\bm{x}$}
-{\color{blue}the maximum absolute value of the input space}, 
+where $S > 0$ represents the maximum absolute value of the input space, 
 and $c \geq 1$ is the proportional extension factor.
 In the following, we will refer to $c$ as the boundary factor of the approximation.
-The boundary factor can also be regarded as the boundary $L$ normalized by
-%the half-range $S$ of the input space.
-{\color{blue}the boundary $S$}. 
-{\color{blue}Notice that $\bm{x}$ need not be symmetric around zero, but need to be contained in $\left[-S, S\right]$} (although it is recommended that $\bm{x}$ fits tightly the interval $\left[-S, S\right]$ to optimize computation and the iterative diagnostic presented in Section~\ref{subsec_diagnostics}).
+The boundary factor can also be regarded as the boundary $L$ normalized by the boundary $S$.
+Notice that $\bm{x}$ need not be symmetric around zero, but need to be contained in $\left[-S, S\right]$ (although it is recommended that $\bm{x}$ fits tightly the interval $\left[-S, S\right]$ to optimize computation and the iterative diagnostic presented in Section~\ref{subsec_diagnostics}).
 
 \begin{figure}
-%\centering
+\centering
 \includegraphics[width=0.485\textwidth, trim = 2mm 0mm 0mm 0mm, clip]{fig1_Post_J_L.png}
-\vspace{-1mm}
 \caption{Mean posterior predictive functions (top), posterior standard deviations (center), and covariance functions (bottom) of both the exact GP model (red line) and the HSGP model for 1) different number of basis functions $m$, with the boundary factor fixed to a large enough value (left column) and 2) different values of the boundary factor $c$, with a large enough fixed number of basis functions (right column).}
@@
 We start by illustrating how the number of basis functions $m$ and boundary factor $c$ influence the accuracy of the HSGP approximations individually.
-For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of length-scale $\ell=0.3$ and marginal variance $\alpha=1$, using input values
-%from the zero-mean input domain with half-range $S=1$
-{\color{blue}within the interval $[-1,1]$ that leads to a boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}.
-In this case the domain of input points $x_i$ matches exactly the interval $\left[-S, S\right]$}.
+For this purpose, a set of noisy observations are drawn from an exact GP model with a squared exponential covariance function of length-scale $\ell=0.3$ and marginal variance $\alpha=1$, using input values within the interval $[-1,1]$ that leads to a boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}.
+In this case the domain of input points $x_i$ matches exactly the interval $\left[-S, S\right]$.
 Several HSGP models with varying $m$ and $c$ are fitted to this data.
@@
 
-\begin{figure} %[!t]
+\begin{figure}
 \centering
@@
 \subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig5_relationships.png}}
-\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale $\ell$ normalized by {\color{blue}the input space boundary $S$} ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
+\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale $\ell$ normalized by the input space boundary $S$ ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
   \label{fig5_relationships}
@@
 Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\ell$.
-Figure \ref{fig5_relationships} depicts how $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ normalized {\color{blue}by the input space boundary $S$}) interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
-More precisely, for a given GP model (with a particular covariance function) {\color{blue}with length-scale $\ell$
-%(from now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$)
- and given a input space boundary $S$ and a boundary factor $c$}, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
+Figure \ref{fig5_relationships} depicts how $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ normalized by the input space boundary $S$) interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension.
+More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a input space boundary $S$ and a boundary factor $c$, Figure \ref{fig5_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq.
 (\ref{eq_diff_covs}).
@@
 As explained in Section~\ref{subsec_theoretic_linearity}, when $c$ is large enough, there is a near linear proportionality between $m$, $\ell$ and $c$.
-To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically {\color{blue}checked this linear relationship and} derived the practically useful constant terms.
+To obtain practical numerical functions that can be used to guide the selection of these parameters, we have empirically checked this linear relationship and derived the practically useful constant terms.
 We require a lower bound for $c$ of $c \geq 1.2$ such that the equations below are precise enough for practical application.
@@
 If the aim is to find minimal $m$ to speedup the computation, a further refined formula can be obtained for new covariance functions.
-
-{\color{blue}
-%By using the normalized length-scale $\frac{\ell}{S}$, the equations, figures, and conclusions made throughout the paper regarding the diagnostic of the method can easily be referred to any specific scale of the input space.
-}
 
@@
 
-\item[A0.] {\color{blue}Compute the boundary $S$ of the input values by using eq.~\eqref{eq_boundary_S}}.
+\item[A0.] Compute the boundary $S$ of the input values by using eq.~\eqref{eq_boundary_S}.
 
@@
 Make an initial guess on the length-scale $\ell^{(k)}$.
-If there is no useful information available, we recommend to start with a large length-scale, {\color{blue}such as the normalized length-scale being in the interval $\frac{\ell^{(k)}}{S} \in [0.5, 1]$}.
-%({\color{blue}Remember that $\ell$ here refers to the normalized length-scale $\frac{\ell}{S}$.})
+If there is no useful information available, we recommend to start with a large length-scale, such as the normalized length-scale being in the interval $\frac{\ell^{(k)}}{S} \in [0.5, 1]$.
 
@@
 
-{\color{blue} In our experiments this procedure converges quickly after only few iterations in all the cases (see Figure~\ref{fig7_diagnostic_lscale_comparison}, Table~\ref{tab1_diagnostic}, and Section~\ref{sec_cases}).
+In our experiments this procedure converges quickly after only few iterations in all the cases (see Figure~\ref{fig7_diagnostic_lscale_comparison}, Table~\ref{tab1_diagnostic}, and Section~\ref{sec_cases}).
 That said, we cannot rule out that there may be cases where convergence is much slower (in terms of number of required diagnostic iterations) to a degree where HSGPs become impractical and an early stopping of the procedure would be advisable.
 Importantly though, this scenario should not be confused with the scaling of computational cost due to higher dimensional input spaces, an issue discussed in detail in Section~\ref{sec_cost_multi}.
-In a nutshell, according to our experiments, increasing the dimensionality of the input space does not increase the number of required diagnostic iterations to a relevant degree but only the computational cost per iteration.}
+In a nutshell, according to our experiments, increasing the dimensionality of the input space does not increase the number of required diagnostic iterations to a relevant degree but only the computational cost per iteration.
 
@@
 
-In our experiments, we observed roughly similar or even worse scaling behavior for the thin plate splines in comparison to HSGPs, with serious challenges with computation time when doing inference with spline models for $D=3$ and more than 10 basis functions per dimension, or even for $D=2$ and more than 40 basis functions per dimension (see Figures~B.3 right and C.3 in the online supplement).
-
 \section{Case studies}\label{sec_cases}
@@
 In this experiment, we analyze a synthetic dataset with $n = 250$ observations, where the true data generating process is a Gaussian process with additive noise.
-The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process $f(x) \sim \GP(0, k(x, x', \theta))$ with a Mat{\'e}rn ($\nu$=3/2) covariance function $k$ with marginal variance $\alpha=1$ and length-scale $\ell=0.2$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$ {\color{blue} that lead to an input space boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}}.
+The data points are simulated from the model $y_i = f(x_i) + \epsilon_i$, where $f$ is a sample from a Gaussian process $f(x) \sim \GP(0, k(x, x', \theta))$ with a Mat{\'e}rn ($\nu$=3/2) covariance function $k$ with marginal variance $\alpha=1$ and length-scale $\ell=0.2$ at inputs values $\bm{x}=(x_1,x_2,\dots,x_n)$ with $x_i \in [-1,1]$ that lead to an input space boundary $S$ of $S=1$ as per eq.~\eqref{eq_boundary_S}.
 $\epsilon_i$ is additive Gaussian noise with standard deviation $\sigma=0.2$.
 
-In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn ($\nu$=3/2) spectral density $S$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
+In the HSGP model, the latent function values $f(x)$ are approximated as in eq.~(\ref{eq_approxf}), with the Mat{\'e}rn ($\nu$=3/2) spectral density $s$ as in eq.~(\ref{eq_specdens_32}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}), respectively.
 
@@
 Figure~\ref{fig8_Posteriors_exI} shows the posteriors predictive distributions of the exact GP and the HSGP models, the later using the parameter values as the third iterative step, $c=1.2$ ($L=c\cdot S= 1.2$; see eq.~\eqref{eq_boundary}) and $m=40$ basis functions.
-In order to do model comparison a spline-based model with 60 knots is also fitted and plotted jointly with the GP and HSGP models.
-The spline-based model is fitted using the thin plate regression spline approach by \citet{wood2003thin} and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this spline model using the R-package \textit{brms} \citep{burkner2017brms}.
-The sample observations are plotted as circles and the out-of-sample or test data are plotted as crosses.
-The posteriors of the three models, exact GP, HSGP and spline, are similar in the interpolation regions of the input space.
-However, when extrapolating the spline model solution clearly differs from the exact GP and HSGP models as well as the actual observations.
 
@@
 \includegraphics[width=0.485\textwidth]{fig8_Posteriors_exI.png}
-\caption{Posterior predictive means of the proposed HSGP model, the exact GP model, and the spline model. 95\% credible intervals are plotted as dashed lines.}
+\caption{Posterior predictive means of the proposed HSGP model and the exact GP model. 95\% credible intervals are plotted as dashed lines.}
   \label{fig8_Posteriors_exI}
@@
 
-Figure~\ref{fig9_rmse_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
-Both the HSGP and splines models show roughly similar performance at sample data.
-The HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation.
-However, the spline model does not extrapolate data properly (Figure~\ref{fig8_Posteriors_exI}).
-The HSGP model is on average roughly 400 times faster than the exact GP and 10 times faster than the spline model for this particular model and data.
-Also, it is seen that the computation time increases slowly as a function of the number of basis functions.
+Figure~\ref{fig9_rmse_time_exI} shows both the standardized root mean squared error (SRMSE) of the models for the sample data and the computational times in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$.
+The HSGP model is on average roughly 400 times faster than the exact GP for this particular model and data.
+Also, it is seen that the computation time increases slowly as a function of $m$.
 
@@
 \subfigure{\includegraphics[width=0.485\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig9_rmse_time_exI.png}}
-\caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$, for the HSGP model, and knots, for the spline model.
+\caption{Computational time (y-axis), in seconds per iteration (iteration of the HMC sampling method), as a function of the number of basis functions $m$.
 The y-axis is on a logarithmic scale.
@@
 
-The Stan model code for the exact GP model, the HSGP model, the spline model, and R-code to reproduce this case study can be found online at {\small \url{ https://github.com/gabriuma/}} \linebreak {\small \url{ basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
+The Stan model code for the exact GP model and the HSGP model, and R-code to reproduce this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_}} \linebreak {\small \url{functions_approach_to_GP/tree/master/Paper/}} \linebreak {\small \url{Case-study_1D-Simulated-data}}\,.
 
@@
 %
-The component $f_1$ will be approximated using the HSGP model and the function values $f_1$ are approximated as in eq.~(\ref{eq_approxf}), with the squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
+The component $f_1$ will be approximated using the HSGP model and the function values $f_1$ are approximated as in eq.~(\ref{eq_approxf}), with the squared exponential spectral density $s$ as in eq.~(\ref{eq_specdens_inf}), and eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ as in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}).
 The year effects $f_2$ and week effects $f_3$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper.
@@
 Thus, the functions $f_2$ and $f_3$ are approximated as in eq.~(\ref{eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in eq.~(\ref{eq_q_2}).
-{\color{blue}The input variable $\bm{t}=\{1,2,\dots,7305\}$ is previously standardized to have zero mean and unit variance $\left(\frac{\bm{t}-\text{mean}(\bm{t})}{\text{sd}(\bm{t})}\right)$, and then it follows a input space boundary $S=1.732$}.
+The input variable $\bm{t}=\{1,2,\dots,7305\}$ is previously standardized to have zero mean and unit variance $\left(\frac{\bm{t}-\text{mean}(\bm{t})}{\text{sd}(\bm{t})}\right)$, and then it follows a input space boundary $S=1.732$.
 
@@
 For functions $f_2$ and $f_3$ there are not boundary factor $c$ as they use periodic covariance functions, and $m$ and minimum $\ell$ are updated by eq.~\eqref{eq_J_l_periodic}.
-A initial value for the minimum lengthscale $\ell_1$ of $f_1$ is guessed to correspond to around 3 years (i.e., $\ell_1=3$ years $= 3\cdot 365=1095$ days) {\color{blue}in the original scale or around $0.52$ (i.e., $\ell_1=\nicefrac{1095}{\text{sd}(\bm{t})}=0.52$) in the standardized scale used as input in the model}.
+A initial value for the minimum lengthscale $\ell_1$ of $f_1$ is guessed to correspond to around 3 years (i.e., $\ell_1=3$ years $= 3\cdot 365=1095$ days) in the original scale or around $0.52$ (i.e., $\ell_1=\nicefrac{1095}{\text{sd}(\bm{t})}=0.52$) in the standardized scale used as input in the model.
 Initial values for the minimum lengthscales $\ell_2$ and $\ell_3$ of $f_2$ and $f_3$, respectively, are guessed to correspond to half of the period (i.e., $\ell_2=\ell_3=0.5$).
@@
 
-\begin{figure}%[!t]
+\begin{figure}
 \centering
@@
 
-\begin{figure}%[!h]
+\begin{figure}
 \centering
@@
 
-$2D$ and $3D$ synthetic functions were drawn from $2D$ and $3D$ GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively, {\color{blue}and the input space boundary is $S=1$ for both the $2D$ and $3D$ functions}.
+$2D$ and $3D$ synthetic functions were drawn from $2D$ and $3D$ GP priors, with input values $\bm{x}_i \in [-1,1]^2$ and $\bm{x}_i \in [-1,1]^3$, respectively, and the input space boundary is $S=1$ for both the $2D$ and $3D$ functions.
 Squared exponential covariance functions with marginal variance $\alpha=1$ and length-scales $\ell_1=0.10$, $\ell_2=0.3$, and $\ell_3=0.4$, where $\ell_i$ is the length scale for the $i$'th dimension, were used in the GP priors.
@@
 
-In the HSGP models with $2$ and $3$ input dimensions, the underlying functions $f(\bm{x})$ are approximated as in eq. (13)% (\ref{eq_approxf_multi})
-, with the $D$-dimensional squared exponential spectral densities $S$ as in eq.
+In the HSGP models with $2$ and $3$ input dimensions, the underlying functions $f(\bm{x})$ are approximated as in eq.
+(13), with the $D$-dimensional squared exponential spectral densities $s$ as in eq.
 \eqref{eq_specdens_inf}, and the $D$-vectors of eigenvalues $\bm{\lambda}_j$ and the multivariate eigenfunctions $\phi_j$ as in equations \eqref{eq_eigenfunction_multi} and \eqref{eq_eigenvalue_multi}, respectively.
@@
 Table~\ref{tab_caseIII_2D} and \ref{tab_caseIII_3D} contain the iterative steps to fit and diagnose the $2D$ and $3D$ data sets, respectively. 
-%These tables contain the values for the parameters $\ell$, $c$, $m$, the estimated $\hat{\ell}$, the diagnostic $\hat{\ell} + 0.01 \geq \ell$ for each dimension, and the RMSE compute with both the data and the GP as the reference for every iterative steps of the fitting process.
 The minimum requirements to fit the models were easily achieved by performing 4 iterations: from a $1^{\text{st}}$ iteration that uses few multivariate basis functions $m^*$ ($m^*=m_1(=6) \times m_2(=6)=36$ and $m^*=m_1(=6) \times m_2(=6) \times m_3(=6)=216$ for the $2D$ and $3D$ data sets, respectively) to the $4^{\text{th}}$ iteration that uses the minimum multivariate basis functions ($m^*=m_1(=22) \times m_2(=11)=242$ and $m^*=m_1(=21) \times m_2(=10) \times m_3(=12)=2520$ for the $2D$ and $3D$ data sets, respectively) required to accurately approximate these notably wiggly functions.
@@
 \subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig13_errorHSGP_exII.pdf}}
-\subfigure{\includegraphics[scale=0.315, trim = 19mm 18mm 39.2mm 20mm, clip]{fig13_errorSP_exII.pdf}}
 \subfigure{\includegraphics[scale=0.30, trim = 151mm 14mm 2mm 20mm, clip]{fig13_error_colorbar_exII.pdf}}
-\hspace{1pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{fig13_RMSE_exII.pdf}}
-\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{fig13_legend_exII.pdf}}}
-\caption{Error between the $2D$ data-generating function and posterior mean functions of the GP (a), HSGP (b) and spline-based (c) models.
+\hspace{20pt}\subfigure{\includegraphics[scale=0.27, trim =0mm 5mm 10mm 0mm, clip]{fig13_RMSE_exII_nueva.pdf}}
+\hspace{-1.4cm} \subfigure{\raisebox{1.65cm}{\includegraphics[scale=0.30, trim = 28mm 70mm 112mm 20mm, clip]{fig13_legend_exII_nueva.pdf}}}
+\caption{Error between the $2D$ data-generating function and posterior mean functions of the GP (a) and HSGP (b) models.
 Sample points are plotted as circles.
-The right side plot shows the root mean square error (RMSE) of the different methods, and plotted as a function of the boundary factor $c$, number of univariate basis functions $m$ and knots (the same number of basis functions $m$ and knots per dimension are used, resulting in $m^D$ total number of multivariate basis functions).}
+The right side plot shows the root mean square error (RMSE) of the different methods, and plotted as a function of the boundary factor $c$ and number of univariate basis functions $m$ (the same $m$ is used per dimension, resulting in a total number of $m^D$ multivariate basis functions).}
   \label{fig13_errors_exII}
@@
 
-Figure \ref{fig13_errors_exII} shows the difference  between the true underlying data-generating function and the posterior mean for GP, HSGP and spline-based model solutions for the $2D$ data set.
-For the $3D$ data set plots with the posterior functions are not shown because it is difficult to plot functions in a $3D$ input space.
-The splines-based model was fitted using a cubic spline basis, penalized by the conventional integrated square second derivative cubic spline penalty \citep{wood2017generalized}, and implemented in the R-package \textit{mgcv} \citep{wood2011mgcv}.
-A Bayesian approach is used to fit this spline-based model using the R-package \textit{brms} \citep{burkner2017brms}.
-For the spline-based model, $40$ knots in each dimension were used.
-Figure \ref{fig13_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$, for the HSGP, and knots, for the $2D$ spline-based model.
-It can be seen that the posterior mean for HSGP and exact GP accurately represents the true data-generating function.
-In contrast, the accuracy of the spline-based model is significantly worse.
-
-The Stan model codes for the exact GP model, the HSGP model, and the spline-based model, and R-code to reproduce this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_2D&3D-Simulated-data}}\,.
+Figure~\ref{fig13_errors_exII} shows the difference between the true underlying data-generating function and the posterior means of GP and HSGP models for the $2D$ data set.
+For the $3D$ data set, posterior functions are not plotted because it is difficult to plot functions in a $3D$ input space.
+Figure \ref{fig13_errors_exII}-left shows the root mean squared error (RMSE), computed against the data-generating function, as a function of the boundary factor $c$ and number of univariate basis functions $m$.
+
+The Stan model codes for the exact GP model and the HSGP model, and R-code to reproduce this case study can be found online at {\small \url{https://github.com/gabriuma/basis_} \linebreak \url{functions_approach_to_GP/tree/master/Paper/} \linebreak \url{Case-study_2D&3D-Simulated-data}}\,.
 
@@
 %
-where $y_i>t_i$ denotes the unobserved survival time {\color{blue} and $\Phi$ is the standard normal cumulative distribution function}.
+where $y_i>t_i$ denotes the unobserved survival time and $\Phi$ is the standard normal cumulative distribution function.
 The latent function $f(\cdot)$ is modeled as a Gaussian process, centered around a linear model of the predictors $\bm{x}$, and with a squared exponential covariance function $k$.
@@
 
-Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in eq.~(\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar length-scale) squared exponential spectral density $S$ as in eq.~(\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
+Using the HSGP approximation, the functions $h(\bm{x})$ and $g(\bm{x})$ are approximated as in eq.~(\ref{eq_approxf_multi}), with the $D$-dimensional (with a scalar length-scale) squared exponential spectral density $s$ as in eq.~(\ref{eq_specdens_inf}), and the multivariate eigenfunctions $\phi_j$ and the $D$-vector of eigenvalues $\bm{\lambda}_j$ as in equations (\ref{eq_eigenfunction_multi}) and  (\ref{eq_eigenvalue_multi}), respectively.
 
@@
 
-The Stan model code for the exact GP and the HSGP models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_Leukemia-data}}\,.
-%
+The Stan model code for the exact GP and the HSGP models of this case study can be found online at {\small \url{https://github.com/gabriuma/basis_functions_approach_to_GP/} \linebreak \url{tree/master/Paper/Case-study_Leukemia-data}}\,.
 
@@
 \begin{equation*}
-S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \,\mathrm{d}\bm{\tau}, \nonumber
+s(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \,\mathrm{d}\bm{\tau}, \nonumber
 \end{equation*}
@@
 This allows for a Fourier representation of the operator $\mathcal{K}$ as a transfer function which is the spectral density of the Gaussian process.
-Thus, the spectral density $S(\bm{w})$ also gives the approximate eigenvalues of the operator $\mathcal{K}$.
-
-In the isotropic case $S(\bm{w}) = S(||\bm{w}||)$ and assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
-%
-\begin{equation}\label{eq_S}
-S(||\bm{w}||)=a_0+a_1||\bm{w}||^2+a_2(||\bm{w}||^2)^2+a_3(||\bm{w}||^2)^3+\cdots.
-\end{equation}
-The Fourier transform of the Laplace operator $\nabla^2$ is $-||\bm{w}||$, thus the Fourier transform of $S(||\bm{w}||)$ is
+Thus, the spectral density $s(\bm{w})$ also gives the approximate eigenvalues of the operator $\mathcal{K}$.
+
+In the isotropic case $s(\bm{w}) = s(||\bm{w}||)$ and assuming that the spectral density function $s(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
+%
+\begin{equation}\label{eq_s}
+s(||\bm{w}||)=a_0+a_1||\bm{w}||^2+a_2(||\bm{w}||^2)^2+a_3(||\bm{w}||^2)^3+\cdots.
+\end{equation}
+The Fourier transform of the Laplace operator $\nabla^2$ is $-||\bm{w}||$, thus the Fourier transform of $s(||\bm{w}||)$ is
 %
@@
 \end{equation} 
-By letting $||\bm{w}||^2=\lambda_j$ the spectral density in eq.~(\ref{eq_S}) becomes
+By letting $||\bm{w}||^2=\lambda_j$ the spectral density in eq.~(\ref{eq_s}) becomes
 %
 \begin{equation*}
-S(\sqrt{\lambda_j})=a_0+a_1\lambda_j+a_2\lambda_j^2+a_3\lambda_j^3+\cdots,
+s(\sqrt{\lambda_j})=a_0+a_1\lambda_j+a_2\lambda_j^2+a_3\lambda_j^3+\cdots,
 \end{equation*}
@@
 \begin{equation}\label{eq_k_2}
-k(\bm{x},\bm{x}')= \sum_j S(\sqrt{\lambda_j}) \phi_j(\bm{x}) \phi_j(\bm{x}'),
-\end{equation} 
-where $S(\cdot)$ is the spectral density of the covariance function, $\lambda_j$ is the $j$th eigenvalue and $\phi_j(\cdot)$ the eigenfunction of the Laplace operator in a given domain.
+k(\bm{x},\bm{x}')= \sum_j s(\sqrt{\lambda_j}) \phi_j(\bm{x}) \phi_j(\bm{x}'),
+\end{equation} 
+where $s(\cdot)$ is the spectral density of the covariance function, $\lambda_j$ is the $j$th eigenvalue and $\phi_j(\cdot)$ the eigenfunction of the Laplace operator in a given domain.
 
@@
 \begin{acknowledgements}
-We thank Academy of Finland (grants 298742, 308640, and 313122), Finnish Center for Artificial Intelligence, and Technology Industries of Finland Centennial Foundation (grant 70007503; Artificial Intelligence for Research and Development) for partial support of this research.
+We thank Academy of Finland (grants 298742, 308640, and 313122), Instituto de Salud Carlos III, Spain (grant CD21/00186 - Sara Borrell Postdoctoral Fellowship), Finnish Center for Artificial Intelligence, and Technology Industries of Finland Centennial Foundation (grant 70007503; Artificial Intelligence for Research and Development), and Conselleria de Innovación, Universidades, Ciencia y Sociedad Digital, Generalitat Valenciana, Spain (grant AICO/2020/285) for partial support of this research.
 We also acknowledge the computational resources provided by the Aalto Science-IT project.

### 2022-08-15 Paper/Manuscript_Stats&Comp/reply_letter_2nd_STCO.tex
updated response letter submitted

@@
 
-%We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
-We would also like to thank you for the additional feedback on some of our responses.
-
-%\subsubsection*{R\#1 General comment and feedback for the authors}
-%
-%\textbf{Overall, the paper has clearly improved from the original submission and the revision addresses all of my major revision requests adequately.
-I still think the paper is a bit borderline in terms of the technical contribution itself, but the revised version reads better, is more clear on the limitations, clarifies a few open issues, and the experimental section has been improved.
-In particular, the experimental section is now very extensive and provides rich information about the practical use and limitations of the approach.
-The paper certainly has potential for impact as a practical guide and tools for using HSGP and now is clearly more suitable for publication.
-I do not have any major comments for the paper and would be happy to see it accepted in current form.}
-%
-%\textbf{Quick feedback on (some of) the author responses:}
-%
-%\textbf{1.
-Apologies for skipping the statement on brms integration.
-I agree it is valuable as it makes the contribution more accessible for a broad audience that might not be interested in the technical details.}
-%
-%\textbf{3. This resolves by remark well. Good addition that will help practitioners and can be implemented directly for the software as default criteria.}
-%
-%\textbf{4.
-I appreciate the effort of extending the treatment.
-I still have a bit of an issue of how this needs to be done separately for each possible kernel and the implications that has on general probabilistic programming environments where part of the value of the tool is in providing extreme flexibility, but I do agree that for practical uses e.g.
-with brms the set of kernels you now provided is useful and likely to cover most user needs.}
-%
-%\textbf{5. Fair enough; I certainly agree some (if not many) practitioners would fail to do this properly.}
-%
-%\textbf{6-8. I think the experimental section indeed reads better now and is overall very extensive and detailed.}
-%
-%\textbf{9.
-By mostly negative results I meant that the value of the approach is not obvious and hence a practitioner would get the impression that investigating the value of HSGP for higher-dimensional cases would not be worthwhile since the gain would likely be small compared to the effort.
-Apologies for the slightly confusing statement, since I agree you do show some advantage on the Leukemia data.
-The new experiments on simulated data make this aspect more clear.}
-%
-%\textbf{Finally, a quick comment related to R\#2 Comment \#3: I agree here with the authors that empirical comparison against alternative inference techniques is not necessary and would not really make the paper stronger either.}
-
 \subsubsection*{R\#1 Minor comment \#1}
@@
 We have increased the size of the figures in Table 1 and have checked and updated the other figures as we saw fit.
-
-%The authors think that the rest of the figures look good in their current form.
 
@@
 
-%Although that has been said extensively throughout the paper, we have added a remark in Section 4.8.1 and a note at the end of Section 4.5.1, as suggested by the reviewer, which we also think can help to make it more clear.
-
 \subsection*{Reviewer \#2}
-
-%We are grateful for your valuable comments and suggestions, which have been very helpful to improve the paper, and more generally for your time you have dedicated to reviewing it.
 
@@
 
-%We included the solution of a spline model (a model related to the GP/HSGP model) only in case study I, with no other intention than to show, beside the GP and HSGP solutions, the solution of a spline model, which has similar (of the same order) computational cost and limitations in multidimensional cases.
-It can be seen that a basic formulation of a spline model causes significantly worse extrapolation accuracy.
-
 \subsubsection*{R\#2 Comment \#2}
@@
 We apologize for creating confusion with our notation here.
-{\color{blue}The diagnostic Figure~5 and equations (19--24) work with a normalized lengthscale, i.e., a lengthscale $\ell$ normalized by the zero-mean, symmetric interval $S$ which contains the inpus values $x_i$ ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
+The diagnostic Figure~5 and equations (19--24) work with a normalized lengthscale, i.e., a lengthscale $\ell$ normalized by the zero-mean, symmetric interval $S$ which contains the inpus values $x_i$ ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
 In concrete in step A.1 in Section 4.5.1 we recommend values for a first guess of the lengthscale based on the normalized lengthscale.
-We have now clarified the role and use of the normalized lengthscale relative to $S$ throughout the paper.}
-
-%We work with a normalized lengthscale, i.e., a lengthscale $\ell$ relative to the half-range ($S$) of the input domain ($\ell/S$), so the conclusions can be easily referred to any specific scale of the input domain.
-We have now clarified the notation by using different symbols for lengthscale and normalized lengthscale throughout the paper.
+We have now clarified the role and use of the normalized lengthscale relative to $S$ throughout the paper.
 
@@
 
-%\noindent \hdashrule{12.5cm}{0.2pt}{2mm 1pt}
-
-%\nobibliography{references}
-
 \end{document}
